### 2014-06-02 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
First draft for classical hamiltonian derivation

%\documentclass[aps,prd,twocolumn,floatfix]{revtex4}   % style for Physical Review B and AJP are similar
\documentclass[twocolumn,floatfix,nofootinbib]{revtex4}   % style for Physical Review B and AJP are similar
%\documentclass[12pt,aps,prb,preprint]{revtex4}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}

\theoremstyle{theorem}
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{sublem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{ax}[thm]{Axiom}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{clm}[thm]{Claim}
\newtheorem{lemdef}[thm]{Lemma-Definition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{rem}[thm]{Remark}
\newtheorem{ques}[thm]{Question}
\newtheorem{cons}[thm]{\rm\bfseries{Construction}}
\newtheorem{exm}[thm]{Example}
\newtheorem{conds}[thm]{Condition}
\newtheorem{proper}[thm]{Property}
\newtheorem{defnlem}[thm]{Definition-Lemma}
\newtheorem{assump}[thm]{Assumption}
\newtheorem{warn}[thm]{Warning}
\newtheorem{situ}[thm]{Situation}

\begin{document}

\title{Deriving Hamiltonian mechanics from physics, without passing passing through Lagrangians.}
\author{Gabriele Carcassi}
\affiliation{Brookhaven National Laboratory, Upton, NY 11973}
\email{carcassi@bnl.gov}
\date{August 25, 2013}

\begin{abstract}
We derive the Hamiltonian formulation of classical mechanics directly, without reference to Lagrangian mechanics.
We start from the definition of states and the labels used to identify them, and show how simple properties required
on those labels lead to the Hamiltonian formulation.\end{abstract}

\maketitle

\section{Introduction}
In physics Hamiltonian mechanics is usually presented as a reformulation of Lagrangian mechanics, which is how it was originally developed. In this work we aim to derive it on its own, starting from simple physical assumptions, without any reference to Lagrangian mechanics. We believe that provides more direct insight in the theory and a direct physical interpretation of its geometric features.

Hamiltonian is either math or physics from Lagrangian. Lagrangian are not well motivated anyway.

We start from scratch and find Hamiltonian. $dt \wedge dE$

\section{States}
\textbf{State definition}. Suppose we have a physical system that can be found in different configurations. Each unique configuration, or configuration state, is identified by a label. The label could correspond to a measurement, a way to prepare the system or any other mechanism that can be used to distinguish between configuration states. For our purposes, it does not matter what it is, just that it exists.

\begin{defn}\label{statedef}
Let $I$ be a set of labels. Let $\mathbbm{C}_{I}$ be the set of all possible configuration states of our system identified by the labels. $\forall i \in I, \exists \mathbbm{c}_{i} \in \mathbbm{C}_{I}$.
\end{defn}

The following definitions allow us to use multiple labels.

\begin{defn}\label{combine label}
Let $I_1$ and $I_2$ be two sets of labels. We call $I = \langle I_1, I_2 \rangle$ the combined set of labels if an injective map exists such that $\forall i \in I \rightarrowtail i_1 \in I_1, i_2 \in I_2$.
\end{defn}

\begin{defn}\label{orth}
The two label sets $I_1$ and $I_2$ are said to be orthogonal iff $I = \langle I_1, I_2 \rangle = I_1 \times I_2$.
\end{defn}

The following assumptions and definitions allow us to describe a configuration state by its parts.

\begin{assump}\label{classical}
(\emph{Classical assumption}) The system is infinitesimally reducible: it can be thought as composed by two or more similar but smaller systems, each in its own configuration state, which also can be thought as composed by two or more, ad infinitum.
\end{assump}

\begin{defn}\label{statedef}
Let $\mathbbm{S}_I$ be the set of all possible configuration states of the infinitesimal subdivision. We call this set phase space. We call each element $\mathbbm{s}_i$ a state.
\end{defn}

\begin{cor}\label{statedistr}
Each configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}_i$. In physics term, the distribution can be visualized as a histogram of discrete data.
\end{cor}

Under the classical assumption, we can then limit ourselves to the study of the labels of states and their properties without losing generality.

\textbf{State mapping}. We will now concentrate ourselves on deterministic and reversible evolution, defined as:
\begin{defn}\label{determ}
Under deterministic evolution, the future state (i.e. label) is fully identified by the present state: $i'=f(i)$. Under reversible evolution, the present state is fully identified by the future state: $i=g(i')$.
\end{defn}

\begin{cor}\label{detAndRev}
Under deterministic and reversible evolution $i'=f(i)$ and $i=f^{-1}(i')$
\end{cor}

\begin{cor}\label{discreteEv}
The deterministic and reversible evolution of a reducible configuration state $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$ is given by $\mathbbm{c}_{j'}=\sum D'(i) \mathbbm{s}_{i}=\sum D(f^{-1}(i)) \mathbbm{s}_{i}$.
\end{cor}

The value of the new distribution at a label is simply the value of the old distribution at the previous label. The bars of the discrete histogram move place, but keep the same height.

\begin{defn}\label{labelsIndep}
Let $I_1$ and $I_2$ be two orthogonal sets of labels and $I = I_1 \times I_2$. Let $f:i \in I \leftrightarrow i' \in I'$ be a bijective map. The labels are said independent iff we can define two sets of orthogonal labels $I_1'$ and $I_2'$ such that we can define separate mappings:
\begin{enumerate}
\item $I_1' \times I_2' = I$
\item there exists a bijective map $f_1: i_1 \in I_1 \leftrightarrow i_1' \in I_1'$
\item there exists a bijective map $f_2: i_2 \in I_2 \leftrightarrow i_2' \in I_2'$
\item $f(i) = \{f_1(i_1), f_2(i_2)\}$
\end{enumerate}
\end{defn}

\textbf{State counting}. Given that deterministic and reversible evolution defines a bijective map between past and future states, we have the following.

\begin{cor}\label{labelsCount}
Let $\hat{I} \subset I$ of finite size $n(\hat{I})$. Let $f(\hat{I})$ and $f^{-1}(\hat{I})$ be respectively the forward and backward image of a deterministic and reversible mapping. We then have $n(\hat{I})=n(f(\hat{I}))=n(f^{-1}(\hat{I}))$, the number of states is conserved.
\end{cor}

\begin{cor}\label{labelsMultiCount}
Let $\hat{I_1} \subset I_1$ and $\hat{I_2} \subset I_2$ of finite size $n(\hat{I_1})$ and $n(\hat{I_2})$. Let $f(\hat{I_1})$, $f(\hat{I_2})$, $f^{-1}(\hat{I_1})$ and $f^{-1}(\hat{I_2})$ be the corresponding forward and backward images. We then have $n(\hat{I_1}\times\hat{I_2})=n(\hat{I_1})n(\hat{I_2})=n(f(\hat{I_1})) n(f(\hat{I_2}))=n(f^{-1}(\hat{I_1})) n(f^{-1}(\hat{I_2}))$, the total number of states is conserved and remains the product of the number of labels in each set.
\end{cor}

These statements, properly generalized for the continuous case, will allow us to define a metric. Conserving that metric will lead to Hamiltonian mechanics.

\section{Numeric labels}

Labels often will be numbers. For integers, we simply use $\mathbb{Z}$ as our set of labels. For real number, we need to update our definitions.

\textbf{Discrete labels over $\mathbb{R}$}.

\begin{defn}\label{disclabelsoverr}
Consider a continuous numeric range. We divide the full range into contiguous cells. Let $I$ be the set of cells. For each cell we have a center value $x: I \mapsto \mathbb{R}$ and a width $w: I \mapsto \mathbb{R}$. $I$ is a set of discrete labels over a continuous range.
\end{defn}

\begin{cor}\label{disclabelsoverrdist}
Each configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i=\sum \rho(i) w(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ is defined as before, and $\rho(i)\equiv D(i) / w(i)$ is the density of the distribution for the cell. In physics term, the distribution can be visualized as a histogram where $w(i)$, $\rho(i)$ and $D(i)$ are respectively the width, height and area of the each bin.
\end{cor}

\begin{cor}\label{discreteEv}
Under deterministic and reversible evolution $i' = f(i)$, where $i', i \in I$, we have $\mathbbm{c}_j=\sum D'(i') \mathbbm{s}_i'=\sum \rho'(i') w(i') \mathbbm{s}_i' = \sum D(i) \mathbbm{s}_i' = \sum rho(i) w(i) \mathbbm{s}_i'$. $\rho'(i') = \rho(i) w(i) / w(i')$.
\end{cor}

The area moves from one cell of the histogram to the other. The height need to be adjusted if the cell is of a different size.

\begin{defn}\label{discreteHomogeneous}
A set $I$ of discrete labels over a continuous range is said to be homogeneous if $w(i)=k$: the bins are of equal width.
\end{defn}

With homogeneous labels no adjusting is needed, and the range can be used as a measure of the number of labels.

\textbf{Continuous labels over $\mathbb{R}$}.

\begin{defn}\label{continuousLabels}
A state variable $X$ is the continuous limit of a set $I$ of discrete labels over a continuous range.
\end{defn}

To prepare for the limit we define $m(i)=w(i)/\Delta w$, where $\Delta w$ represents the average width of the cells. We increase the number of the cells and reduce their width such that $max(m(i))$ is bounded. In the limit we'll have a cell for each value, so we can use $x(i)$ (or simply x) instead of $i$ for the label. $\rho$ and $m$ will converge to functions defined over $X$. The corresponding configuration state will become $\mathbbm{c}_j=\int \rho(x) m(x) \mathbbm{s}_x ds$.

\begin{prop}\label{continuousMapping}
Let $f: X \mapsto X$ be a deterministic and reversible mapping on a state variable. The mapping must be continuous.
\end{prop}
%http://en.wikipedia.org/wiki/Limiting_density_of_discrete_points
Assume mapping is discontinuous at point $x$. Consider the cell at $x$ of width $m(x)dx$. The cell would be split into two, so it would not be mapped to one and only one other cell.

\begin{prop}\label{widthMapping}
Let $f: X \mapsto X$ be a deterministic and reversible mapping on a state variable. Then $m(x') dx' \equiv m(f(x)) df(x)/dx dx = m(x) dx$. If $X$ is homogeneous, then $dx' = dx$ and the range gives us a measure of the number of states.
\end{prop}
The mapping must be done so that the width of the cells is mapped as well, not just the center value.

\section{Single degree of freedom}

\begin{defn}\label{sdof}
A degree of freedom is a label set $X$ given by a pair of homogeneous and orthogonal state variables $P$ and $Q$. We call these conjugate variables. A state variable is said conjugate of the first iff they together form a pair of conjugate variables.
\end{defn}

\begin{prop}\label{sdofMap}
Let $f: X \mapsto X$ be a deterministic and reversible mapping on a degree of freedom. Then $dx' \wedge dp' = dx \wedge dp$.
\end{prop}

This is the generalization of \ref{widthMapping} for two state variable. The density $\rho(p,q)$ will be defined on cells of area proportional to $dx \wedge dp$. When mapping one cell to another, it is the area that will need to remain the same and that can be used as a measure of number of labels. We can re-express area conservation as the invariance of the vector product, which defines a metric in terms of labels along a degree of freedom.

\begin{prop}\label{sdofInvariance}
Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by two conjugate variables. Let
\begin{align*}
\omega_{\alpha, \beta} = \left[
  \begin{array}{cc}
    0 & 1 \\
    -1 & 0 \\
  \end{array}
\right] \\
\end{align*}
then $v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}=v^{\alpha} \omega_{\alpha, \beta} w^{\beta}$ under detrev transformation.
\end{prop}

\begin{lem}\label{genAntisim}
Let $v$ and $w$ be two vectors. Let $v^{\alpha} \omega_{\alpha, \beta} w^{\alpha}$ be an antisymmetric metric product conserved under a continuous transformation parameterized by $t$. We can then define a generating function $H$ such that:
\begin{enumerate}
\item $S^{\alpha} \equiv d_{t}x^{\alpha}$, vector field that tells us how the state variables change
\item $S_{\beta} \equiv S^{\alpha} \omega_{\alpha, \beta}$
\item $S_{\alpha} = \partial_{\alpha}H$
\end{enumerate}
\end{lem}

Simply applying the vector transformation rules under continuous transformation we have:
\begin{align*}
V^{\alpha} \Omega_{\alpha, \beta} W^{\beta} &= V'^{\alpha} \Omega_{\alpha, \beta} W'^{\beta}  \\
&= (V^{\alpha} + \partial_{\gamma} S^{\alpha} dt V^{\gamma}) \Omega_{\alpha, \beta} ( W^{\beta} + \partial_{\delta} S^{\beta} W^{\delta} dt) \\
&= V^{\alpha} \Omega_{\alpha, \beta} W^{\beta} + (\partial_{\gamma} S^{\alpha} V^{\gamma} \Omega_{\alpha, \beta} W^{\beta} \\
 &+ V^{\alpha} \Omega_{\alpha, \beta} \partial_{\delta} S^{\beta} W^{\delta}) dt + O(dt^2)
\end{align*}
\begin{align*}
V^{\gamma} W^{\beta} \partial_{\gamma} S_{\beta} - V^{\alpha} W^{\delta} \partial_{\delta} S_{\alpha} = 0
\end{align*}
\begin{align*}
\partial_{\alpha} S_{\beta} - \partial_{\beta} S_{\alpha} &= curl(S_{\alpha}) = 0 \\
S_{\alpha} &= \partial_{\alpha}H
\end{align*}

\begin{prop}\label{sdofHam}
The evolution for a single degree of freedom is given by:
\begin{align*}
d_{t}q &= \partial_{p} H \\
d_{t}p &= - \partial_{q} H
\end{align*}
\end{prop}

This is just expanding \ref{genAntisim} with the metric defined in \ref{sdofInvariance}. We recognise Hamilton's equations for one degree of freedom.

\section{Multiple degrees of freedom}

\begin{defn}\label{mdof}
Two degrees of freedom are said independent if the corresponding label set $X^1$ and $X^2$ are independent.
\end{defn}

\begin{prop}\label{mdofInvariance}
Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by two independent degrees of freedom. Let
\begin{align*}
\omega_{\alpha, \beta} = \left[
  \begin{array}{cccc}
    0 & 1 & 0 & 0 \\
    -1 & 0 & 0 & 0 \\
    0 & 0 & 0 & 1 \\
    0 & 0 & -1 & 0 \\
  \end{array}
\right] \\
\end{align*}
then $v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}=v^{\alpha} \omega_{\alpha, \beta} w^{\beta}$ under detrev transformation.
\end{prop}

The orthogonality of the labels corresponds to orthogonality in the manifold. Consider a hypercube oriented in such a way that some sides align to the first degree of freedom and some sides align to the second. Let $V$ its hypervolume and $V'$ the hypervolume of its map. and Let $A_1$ and $A_2$ the area of two regions of each degree of freedom. Let $A_1'$ and $A_2'$ be the corresponding maps. From \ref{labelsCount} and \ref{labelsMultiCount} we have $V=V'=A_1A_2=A_1'A_2'$, which can hold only if the two degrees of freedom are orthogonal and remain orthogonal. This means we need to require the conservation of the scalar product across independent degrees of freedoms, while still requiring conservation of the vector product within. Our metric will be given by:
\begin{align*}
\omega_{\alpha, \beta} =  \left[
  \begin{array}{cc}
    1 & 0 \\
    0 & 1 \\
  \end{array}
\right] \otimes \left[
  \begin{array}{cc}
    0 & 1 \\
    -1 & 0 \\
  \end{array}
\right] \\
\end{align*}
The metric corresponds to a measure of the labels defined on the area given by two arbitrary vectors. For an infinitesimal region, this corresponds to $dx^1 \wedge dp^1 + dx^2 \wedge dp^2$, the sum of the projections on the independent planes.

\begin{prop}\label{mdofHam}
The evolution for a multiple degrees of freedom is given by:
\begin{align*}
d_{t}q^i &= \partial_{p^i} H \\
d_{t}p^i &= - \partial_{q^i} H
\end{align*}
\end{prop}

Expand \ref{genAntisim} with the metric defined in \ref{mdofInvariance}. We recognise Hamilton equations for multiple degrees of freedom.

\section{Time dependence}
So far we have assumed that both state labeling and mapping do not change in time. To be general, then we have to introduce an extra degree of freedom.

\begin{defn}\label{tdof}
The temporal degree of freedom is a label set $X$ given by the pair of conjugate variables $T$ and $E$. We call extended phase space the outer product between phase space and the temporal degree of freedom.
\end{defn}

\begin{prop}\label{tdofMonotonic}
Let $s$ be the parameter of a trajectory in the extended phase space of a deterministic and reversible system. The trajectory must be continuous. There must exist a strictly monotonic function $t(s)$. States connected by a trajectory where $dt/ds>0$ are called particle states, while the ones for which $dt/ds<0$ are called anti-particle states.
\end{prop}

The trajectory has to be continuous in both spatial and temporal variables because of \ref{continuousMapping}. The mapping still needs to give one and only one value in time, so in principle we should be able to use $t$ as a parameter, we much have an invertible mapping between $t$ and $s$, which means we must have a strictly monotonic $t(s)$. Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call particle and anti-particle states is also conventional. What is physical, instead, is that $dt/ds$ along a trajectory cannot change sign, so particle and anti-particle states cannot be connected by deterministic and reversible evolution.

\begin{prop}\label{tdofInvariant}
Let $v$ and $w$ be two vectors defined on the tangent space of the extended phase space for one degree of freedom. Let $\alpha$ and $\beta$ be indexes for $t, E, x, p$. Let
\begin{align*}
\omega_{\alpha, \beta} = \omega_{\alpha, \beta} =  \left[
  \begin{array}{cc}
    -1 & 0 \\
    0 & 1 \\
  \end{array}
\right] \otimes \left[
  \begin{array}{cc}
    0 & 1 \\
    -1 & 0 \\
  \end{array}
\right]
= \left[
  \begin{array}{cccc}
    0 & -1 & 0 & 0 \\
    1 & 0 & 0 & 0 \\
    0 & 0 & 0 & 1 \\
    0 & 0 & -1 & 0 \\
  \end{array}
\right] \\
\end{align*}
then $v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}=v^{\alpha} \omega_{\alpha, \beta} w^{\beta}$ under detrev transformation.
\end{prop}

$\langle T, E \rangle$ are not independent labels from $\langle Q, P \rangle$ as they do not define new states. They are not necessarily orthogonal directions in the extended phase space. Looking back at \ref{disclabelsoverr}, cells need to be defined on the plane where $\langle Q, P \rangle$ changes: they are not defined on the plane of constant $\langle E, T \rangle$ (they are not orthogonal), but on the plane perpendicular to $\langle Q, P \rangle$. Areas defined on that plane are then going to be invariant.

Like before, we have three planes of conjugate variables forming a right triangle-like relationship with one side invariant. In this case, the right angle is between the invariant and the plane of constant $\langle Q, P \rangle$ (on which $dt$ and $dE$ are defined). So we must have $dt \wedge dE + k = dq \wedge dp$. Which corresponds to the Minkowski product across d.o.f. and the vector product within. The metric, with a space-like convention, gives us a measure for states.

\begin{prop}\label{tdofConstrain}
The evolution is constrained by $H_{e}=k$, the generating function defined in \ref{genAntisim} generalized for the extended phase space.
\end{prop}

Given N degrees of freedom, phase space is $\mathbb{R}^{2*N}$. Extended phase space is $\mathbb{R}^{2*N + 2}$. Phase space evolved in time is $\mathbb{R}^{2*N + 1}$. Since $H_{e}$ is constant through the evolution, it can serve both as the generating function and as the evolution constrain. By convention, we set $H_{e}=0$.

\begin{prop}\label{tdofHam}
The evolution for time varying multiple degrees of freedom is given by:
\begin{align*}
d_{s}t &= - \partial_{E} H_{e} \\
d_{s}E &= \partial_{t} H_{e} \\
d_{s}q^i &= \partial_{p^i} H_{e} \\
d_{s}p^i &= - \partial_{q^i} H_{e}
\end{align*}
\end{prop}

Expand \ref{genAntisim} with the metric defined in \ref{tdofInvariant}. We recognise Hamilton equations in the extended phase space.

It should not be a surprise that the equations do not mention $c$. In fact, nothing says that all $q^i$ represent space, and with the same unit, and that the laws of motion are invariant in all inertial frames. The only requirement we have is that the areas of each degrees of freedom represent the same measure for labels.

We wrap up with an example. Let $H_{E}= mc^2 + ((p^i)^2 - E^2/c^2) / 2m$. This represents the rest energy of a free particle, the conjugate of which is proper time. Therefore the parametrization of a particle-state is $s=\tau$ while the parametrization for an anti-particle state is $s=-\tau$. If we apply \ref{tdofHam} we have:
\begin{align*}
d_{s}t &= E / mc^2 \\
d_{s}E &= 0 \\
d_{s}q^i &= p^i / m \\
d_{s}p^i &= 0
\end{align*}
For particles, we have:
\begin{align*}
E / c &= m c d_{\tau}t = m U^0 = P^0 \\
p^i &= m d_{\tau}q^i = m U^i = P^i \\
\end{align*}
so we recognize the four-momentum $P^\alpha = {E/c, p^i}$. For anti-particles:
\begin{align*}
- E / c &= m c d_{\tau}t = m U^0 = P^0 \\
- p^i &= m d_{\tau}q^i = m U^i = P^i \\
\end{align*}
we end up with a minus sign between the four-momentum and the conjugate variables $P^\alpha = {-E/c, -p^i}$.

\section{Conclusion}
I hope this work convinced you that we can derive Hamiltonian mechanics from simple definitions of label, states, determinism and reversibility. I believe this approach helps give direct physical meaning to phase space and its geometric properties, and it shows that Hamiltonian mechanics can stand on its own, without Lagrangians. I find it uncanny how so much can be derived with so little.

While no mathematical breakthrough is revealed (we have just put together in a novel way the known pieces of the puzzle), I am not aware of any work that focus on this aspect. We already have ideas on how to extend this work to Quantum Mechanics, and we hope to gain insight on how to extend to General Relativity.

\section{Thermodynamics}
\begin{align*}
\oint \delta E &= \oint (\frac{dE}{dt} \delta t + \frac{dx_i}{dt} \delta p_i - \frac{dp_i}{dt} \delta x_i) \\
 &= \oint (\frac{dE}{d\tau} \frac{d\tau}{dt} \delta t + \frac{dx_i}{d\tau}\frac{d\tau}{dt} \delta p_i - \frac{dp_i}{d\tau}\frac{d\tau}{dt} \delta x_i) \\
\oint \frac{dt}{d\tau} \delta E &= \oint (\frac{dE}{d\tau} \delta t + \frac{dx_i}{d\tau} \delta p_i - \frac{dp_i}{d\tau} \delta x_i) \\
\oint (\frac{dt}{d\tau} \delta E &- \frac{dE}{d\tau} \delta t - \frac{dx_i}{d\tau} \delta p_i + \frac{dp_i}{d\tau} \delta x_i) = 0 \\
\end{align*}

\section{Math notes}

Define a product of vectors defined in phase space (space tangent at each point in phase space).

\begin{align*}
\vec{V} \ast \vec{W} &= V^{\alpha} \Omega_{\alpha, \beta} W^{\beta} \\
\Omega_{\alpha, \beta} &= \left(
                      \begin{array}{cccccccc}
                        0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
                        -1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
                        0 & 0 & 0 & -1 & 0 & 0 & 0 & 0 \\
                        0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
                        0 & 0 & 0 & 0 & 0 & -1 & 0 & 0 \\
                        0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
                        0 & 0 & 0 & 0 & 0 & 0 & 0 & -1 \\
                        0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
                      \end{array}
                    \right) \\
\end{align*}

\begin{align*}
x'^{\alpha}&=x^{\alpha} + S^{\alpha} d\tau \\
S^{\alpha} &= d_{\tau}x^{\alpha}  \\
S_{\beta} &\equiv S^{\alpha} \Omega_{\alpha, \beta} \\
S_{\alpha} &= - S^{\beta} \Omega_{\alpha, \beta}
\end{align*}

\begin{align*}
V^{\alpha} \Omega_{\alpha, \beta} W^{\beta} &= V'^{\alpha} \Omega_{\alpha, \beta} W'^{\beta}  \\
&= (V^{\alpha} + \partial_{\gamma} S^{\alpha} dt V^{\gamma}) \Omega_{\alpha, \beta} ( W^{\beta} + \partial_{\delta} S^{\beta} W^{\delta} dt) \\
&= V^{\alpha} \Omega_{\alpha, \beta} W^{\beta} + (\partial_{\gamma} S^{\alpha} V^{\gamma} \Omega_{\alpha, \beta} W^{\beta} \\
 &+ V^{\alpha} \Omega_{\alpha, \beta} \partial_{\delta} S^{\beta} W^{\delta}) dt + O(dt^2) \\
\end{align*}

\begin{align*}
V^{\gamma} W^{\beta} \partial_{\gamma} S_{\beta} - V^{\alpha} W^{\delta} \partial_{\delta} S_{\alpha} = 0
\end{align*}

\begin{align*}
\partial_{\alpha} S_{\beta} - \partial_{\beta} S_{\alpha} &= 0 \\
S_{\alpha} &= \partial_{\alpha}H \\
S^{\alpha} &= d_{\tau}x^{\alpha}
\end{align*}

\begin{align*}
d_{\tau}(ct) &= S^{ct} = S_{E/c} = \partial_{E/c} H \\
d_{\tau}t &= \partial_{E} H \\
\end{align*}

\begin{align*}
d_{\tau}(E/c) &= S^{E/c} = - S_{ct} = - \partial_{ct} H \\
d_{\tau}E &= - \partial_{t} H \\
\end{align*}

\begin{align*}
d_{\tau}(x) &= S^{x} = - S_{p_x} = - \partial_{p_x} H \\
d_{\tau}x &= - \partial_{p_x} H \\
d_{\tau}y &= - \partial_{p_y} H \\
d_{\tau}z &= - \partial_{p_z} H \\
\end{align*}

\begin{align*}
d_{\tau}(p_x) &= S^{p_x} = S_{x} = \partial_{x} H \\
d_{\tau}p_x &= \partial_{x} H \\
d_{\tau}p_y &= \partial_{y} H \\
d_{\tau}p_z &= \partial_{z} H \\
\end{align*}

All 8 equations:
\begin{align*}
d_{\tau}t &= \partial_{E} H \\
d_{\tau}x &= - \partial_{p_x} H \\
d_{\tau}y &= - \partial_{p_y} H \\
d_{\tau}z &= - \partial_{p_z} H \\
d_{\tau}E &= - \partial_{t} H \\
d_{\tau}p_x &= \partial_{x} H \\
d_{\tau}p_y &= \partial_{y} H \\
d_{\tau}p_z &= \partial_{z} H \\
\end{align*}

Free massive particle:
\begin{align*}
H &= (E^2/c^2 - p_x^2 - p_y^2 - p_z^2)/2m \\
d_{\tau}t &= E/mc^2 \\
d_{\tau}x &= p_x/m \\
d_{\tau}y &= p_y/m \\
d_{\tau}z &= p_z/m \\
d_{\tau}E &= 0 \\
d_{\tau}p_x &= 0 \\
d_{\tau}p_y &= 0 \\
d_{\tau}p_z &= 0 \\
\end{align*}

\begin{thebibliography}{5}

\bibitem{extend} No bib right now.

\end{thebibliography}

\end{document}

### 2014-06-02 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Removing old math notes

@@
 
-\section{Thermodynamics}
-\begin{align*}
-\oint \delta E &= \oint (\frac{dE}{dt} \delta t + \frac{dx_i}{dt} \delta p_i - \frac{dp_i}{dt} \delta x_i) \\
- &= \oint (\frac{dE}{d\tau} \frac{d\tau}{dt} \delta t + \frac{dx_i}{d\tau}\frac{d\tau}{dt} \delta p_i - \frac{dp_i}{d\tau}\frac{d\tau}{dt} \delta x_i) \\
-\oint \frac{dt}{d\tau} \delta E &= \oint (\frac{dE}{d\tau} \delta t + \frac{dx_i}{d\tau} \delta p_i - \frac{dp_i}{d\tau} \delta x_i) \\
-\oint (\frac{dt}{d\tau} \delta E &- \frac{dE}{d\tau} \delta t - \frac{dx_i}{d\tau} \delta p_i + \frac{dp_i}{d\tau} \delta x_i) = 0 \\
-\end{align*}
-
-\section{Math notes}
-
-Define a product of vectors defined in phase space (space tangent at each point in phase space).
-
-\begin{align*}
-\vec{V} \ast \vec{W} &= V^{\alpha} \Omega_{\alpha, \beta} W^{\beta} \\
-\Omega_{\alpha, \beta} &= \left(
-                      \begin{array}{cccccccc}
-                        0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
-                        -1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-                        0 & 0 & 0 & -1 & 0 & 0 & 0 & 0 \\
-                        0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
-                        0 & 0 & 0 & 0 & 0 & -1 & 0 & 0 \\
-                        0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
-                        0 & 0 & 0 & 0 & 0 & 0 & 0 & -1 \\
-                        0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
-                      \end{array}
-                    \right) \\
-\end{align*}
-
-\begin{align*}
-x'^{\alpha}&=x^{\alpha} + S^{\alpha} d\tau \\
-S^{\alpha} &= d_{\tau}x^{\alpha}  \\
-S_{\beta} &\equiv S^{\alpha} \Omega_{\alpha, \beta} \\
-S_{\alpha} &= - S^{\beta} \Omega_{\alpha, \beta}
-\end{align*}
-
-\begin{align*}
-V^{\alpha} \Omega_{\alpha, \beta} W^{\beta} &= V'^{\alpha} \Omega_{\alpha, \beta} W'^{\beta}  \\
-&= (V^{\alpha} + \partial_{\gamma} S^{\alpha} dt V^{\gamma}) \Omega_{\alpha, \beta} ( W^{\beta} + \partial_{\delta} S^{\beta} W^{\delta} dt) \\
-&= V^{\alpha} \Omega_{\alpha, \beta} W^{\beta} + (\partial_{\gamma} S^{\alpha} V^{\gamma} \Omega_{\alpha, \beta} W^{\beta} \\
- &+ V^{\alpha} \Omega_{\alpha, \beta} \partial_{\delta} S^{\beta} W^{\delta}) dt + O(dt^2) \\
-\end{align*}
-
-\begin{align*}
-V^{\gamma} W^{\beta} \partial_{\gamma} S_{\beta} - V^{\alpha} W^{\delta} \partial_{\delta} S_{\alpha} = 0
-\end{align*}
-
-\begin{align*}
-\partial_{\alpha} S_{\beta} - \partial_{\beta} S_{\alpha} &= 0 \\
-S_{\alpha} &= \partial_{\alpha}H \\
-S^{\alpha} &= d_{\tau}x^{\alpha}
-\end{align*}
-
-\begin{align*}
-d_{\tau}(ct) &= S^{ct} = S_{E/c} = \partial_{E/c} H \\
-d_{\tau}t &= \partial_{E} H \\
-\end{align*}
-
-\begin{align*}
-d_{\tau}(E/c) &= S^{E/c} = - S_{ct} = - \partial_{ct} H \\
-d_{\tau}E &= - \partial_{t} H \\
-\end{align*}
-
-\begin{align*}
-d_{\tau}(x) &= S^{x} = - S_{p_x} = - \partial_{p_x} H \\
-d_{\tau}x &= - \partial_{p_x} H \\
-d_{\tau}y &= - \partial_{p_y} H \\
-d_{\tau}z &= - \partial_{p_z} H \\
-\end{align*}
-
-\begin{align*}
-d_{\tau}(p_x) &= S^{p_x} = S_{x} = \partial_{x} H \\
-d_{\tau}p_x &= \partial_{x} H \\
-d_{\tau}p_y &= \partial_{y} H \\
-d_{\tau}p_z &= \partial_{z} H \\
-\end{align*}
-
-All 8 equations:
-\begin{align*}
-d_{\tau}t &= \partial_{E} H \\
-d_{\tau}x &= - \partial_{p_x} H \\
-d_{\tau}y &= - \partial_{p_y} H \\
-d_{\tau}z &= - \partial_{p_z} H \\
-d_{\tau}E &= - \partial_{t} H \\
-d_{\tau}p_x &= \partial_{x} H \\
-d_{\tau}p_y &= \partial_{y} H \\
-d_{\tau}p_z &= \partial_{z} H \\
-\end{align*}
-
-Free massive particle:
-\begin{align*}
-H &= (E^2/c^2 - p_x^2 - p_y^2 - p_z^2)/2m \\
-d_{\tau}t &= E/mc^2 \\
-d_{\tau}x &= p_x/m \\
-d_{\tau}y &= p_y/m \\
-d_{\tau}z &= p_z/m \\
-d_{\tau}E &= 0 \\
-d_{\tau}p_x &= 0 \\
-d_{\tau}p_y &= 0 \\
-d_{\tau}p_z &= 0 \\
-\end{align*}
-
 \begin{thebibliography}{5}

### 2014-06-14 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Fixed title
First review
Concentrate on bijective maps, detrev is one such.
Various improvement to the text.

@@
 
-\title{Deriving Hamiltonian mechanics from physics, without passing passing through Lagrangians.}
+\title{Deriving Hamiltonian mechanics from physics, without passing through Lagrangians.}
 \author{Gabriele Carcassi}
@@
 \textbf{State definition}.
-Suppose we have a physical system that can be found in different configurations.
+Let's assume  we have a physical system that can be found in different configurations.
 Each unique configuration, or configuration state, is identified by a label.
-The label could correspond to a measurement, a way to prepare the system or any other mechanism that can be used to distinguish between configuration states.
-For our purposes, it does not matter what it is, just that it exists.
+For our purposes, it does not matter what it is (a measurement, a way to prepare the system, future behavior, ...), just that it exists.
 
@@
 Let $I_1$ and $I_2$ be two sets of labels.
-We call $I = \langle I_1, I_2 \rangle$ the combined set of labels if an injective map exists such that $\forall i \in I \rightarrowtail i_1 \in I_1, i_2 \in I_2$.
+We call $I = \langle I_1, I_2 \rangle$ a combined set of labels if an injective map exists such that $\forall i \in I \rightarrowtail i_1 \in I_1, i_2 \in I_2$.
 \end{defn}
@@
 
-Under the classical assumption, we can then limit ourselves to the study of the labels of states and their properties without losing generality.
+Under the classical assumption, we can then limit ourselves to study the labels of states and their properties without losing generality.
 
-\textbf{State mapping}. We will now concentrate ourselves on deterministic and reversible evolution, defined as:
-\begin{defn}\label{determ}
-Under deterministic evolution, the future state (i.e.
-label) is fully identified by the present state: $i'=f(i)$.
-Under reversible evolution, the present state is fully identified by the future state: $i=g(i')$.
-\end{defn}
+\textbf{State mapping}. We will now concentrate ourselves on deterministic and reversible evolution.
+\begin{assump}\label{detAss}
+The system undergoes deterministic (future state identified by the present state) and reversible (past state identified by the present state) evolution.
+\end{assump}
 
-\begin{cor}\label{detAndRev}
-Under deterministic and reversible evolution $i'=f(i)$ and $i=f^{-1}(i')$
+\begin{prop}\label{detMap}
+Let $I$ be a set of labels of a system that undergoes deterministic and reversible evolution. There exists a bijective map $f:i \in I \leftrightarrow i' \in I$.
+\end{prop}
+
+\begin{cor}\label{discreteEv}
+The evolution of a reducible configuration state $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$ under a bijective map is given by $\mathbbm{c}_{j'}=\sum D'(i) \mathbbm{s}_{i}=\sum D(f^{-1}(i)) \mathbbm{s}_{i}$.
 \end{cor}
 
-\begin{cor}\label{discreteEv}
-The deterministic and reversible evolution of a reducible configuration state $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$ is given by $\mathbbm{c}_{j'}=\sum D'(i) \mathbbm{s}_{i}=\sum D(f^{-1}(i)) \mathbbm{s}_{i}$.
-\end{cor}
-
-The value of the new distribution at a label is simply the value of the old distribution at the previous label. The bars of the discrete histogram move place, but keep the same height.
+Mathematically, studying \ref{detAss} means studying bijective maps.
+The evolution of a distribution simply moves the elements around: the bars of the discrete histogram move place, but keep the same height.
 
@@
 Let $I_1$ and $I_2$ be two orthogonal sets of labels and $I = I_1 \times I_2$.
-Let $f:i \in I \leftrightarrow i' \in I'$ be a bijective map.
+Let $f:i \in I \leftrightarrow i' \in I$ be a bijective map.
 The labels are said independent iff we can define two sets of orthogonal labels $I_1'$ and $I_2'$ such that we can define separate mappings:
@@
 
-\textbf{State counting}. Given that deterministic and reversible evolution defines a bijective map between past and future states, we have the following.
+\textbf{State counting}. We now look at how the cardinality of states evolves under a bijective map.
 
@@
 Let $\hat{I} \subset I$ of finite size $n(\hat{I})$.
-Let $f(\hat{I})$ and $f^{-1}(\hat{I})$ be respectively the forward and backward image of a deterministic and reversible mapping.
+Let $f(\hat{I})$ and $f^{-1}(\hat{I})$ be respectively the forward and backward image of a bijective map.
 We then have $n(\hat{I})=n(f(\hat{I}))=n(f^{-1}(\hat{I}))$, the number of states is conserved.
@@
 Let $f(\hat{I_1})$, $f(\hat{I_2})$, $f^{-1}(\hat{I_1})$ and $f^{-1}(\hat{I_2})$ be the corresponding forward and backward images.
-We then have $n(\hat{I_1}\times\hat{I_2})=n(\hat{I_1})n(\hat{I_2})=n(f(\hat{I_1})) n(f(\hat{I_2}))=n(f^{-1}(\hat{I_1})) n(f^{-1}(\hat{I_2}))$, the total number of states is conserved and remains the product of the number of labels in each set.
+We then have $n(\hat{I_1}\times\hat{I_2})=n(\hat{I_1})n(\hat{I_2})=n(f(\hat{I_1})) n(f(\hat{I_2}))=n(f^{-1}(\hat{I_1})) n(f^{-1}(\hat{I_2}))=n(\hat{f(I_1)}\times\hat{f(I_2)})=n(f^{-1}(\hat{I_1})\times f^{-1}(\hat{I_2}))$, the total number of states is conserved and remains the product of the number of labels in each set.
 \end{cor}
@@
 
-Labels often will be numbers. For integers, we simply use $\mathbb{Z}$ as our set of labels. For real number, we need to update our definitions.
-
-\textbf{Discrete labels over $\mathbb{R}$}.
+\textbf{Discrete labels over $\mathbb{R}$}. Labels often will be numbers. For integers, we simply use $\mathbb{Z}$ as our set of labels. For real number, we need to update our definitions.
 
@@
 \begin{cor}\label{discreteEv}
-Under deterministic and reversible evolution $i' = f(i)$, where $i', i \in I$, we have $\mathbbm{c}_j=\sum D'(i') \mathbbm{s}_i'=\sum \rho'(i') w(i') \mathbbm{s}_i' = \sum D(i) \mathbbm{s}_i' = \sum rho(i) w(i) \mathbbm{s}_i'$.
+Let $f: i \in I \leftrightarrow i'$ be a bijective map, we have $\mathbbm{c}_j=\sum D'(i') \mathbbm{s}_i'=\sum \rho'(i') w(i') \mathbbm{s}_i' = \sum D(i) \mathbbm{s}_i' = \sum \rho(i) w(i) \mathbbm{s}_i'$.
 $\rho'(i') = \rho(i) w(i) / w(i')$.
@@
 
-\textbf{Continuous labels over $\mathbb{R}$}.
+\textbf{Continuous labels over $\mathbb{R}$}. We now make the bin width arbitrarily small.
 
@@
 To prepare for the limit we define $m(i)=w(i)/\Delta w$, where $\Delta w$ represents the average width of the cells.
-We increase the number of the cells and reduce their width such that $max(m(i))$ is bounded.
-In the limit we'll have a cell for each value, so we can use $x(i)$ (or simply x) instead of $i$ for the label.
+We increase the number of the cells and reduce $\Delta w$ such that $max(m(i))$ is bounded.
+In the limit we'll have a cell for each value, so we can use $x(i)$ (or simply $x$) instead of $i$ for the label.
 $\rho$ and $m$ will converge to functions defined over $X$.
@@
 \begin{prop}\label{continuousMapping}
-Let $f: X \mapsto X$ be a deterministic and reversible mapping on a state variable. The mapping must be continuous.
+Let $f: X \mapsto X$ be a bijective map on a state variable. The mapping must be continuous.
 \end{prop}
 %http://en.wikipedia.org/wiki/Limiting_density_of_discrete_points
-Assume mapping is discontinuous at point $x$. Consider the cell at $x$ of width $m(x)dx$. The cell would be split into two, so it would not be mapped to one and only one other cell.
+Assume mapping is discontinuous at point $x$. Consider the cell at $x$ of width $m(x)dx$. The cell would be split into two, so it would not be mapped to one and only one cell.
 
@@
 Let $f: X \mapsto X$ be a deterministic and reversible mapping on a state variable.
-Then $m(x') dx' \equiv m(f(x)) df(x)/dx dx = m(x) dx$.
+Then $m(x') dx' \equiv m(f(x)) \frac{\partial f(x)}{\partial x} dx = m(x) dx$.
 If $X$ is homogeneous, then $dx' = dx$ and the range gives us a measure of the number of states.
@@
 We call these conjugate variables.
-A state variable is said conjugate of the first iff they together form a pair of conjugate variables.
+A state variable is said conjugate of the first iff together they form a pair of conjugate variables.
 \end{defn}
@@
 \begin{prop}\label{sdofMap}
-Let $f: X \mapsto X$ be a deterministic and reversible mapping on a degree of freedom. Then $dx' \wedge dp' = dx \wedge dp$.
+Let $f: X \mapsto X$ be a bijective map on a degree of freedom. Then $dx' \wedge dp' = dx \wedge dp$.
 \end{prop}
@@
 This is the generalization of \ref{widthMapping} for two state variable.
-The density $\rho(p,q)$ will be defined on cells of area proportional to $dx \wedge dp$.
-When mapping one cell to another, it is the area that will need to remain the same and that can be used as a measure of number of labels.
-We can re-express area conservation as the invariance of the vector product, which defines a metric in terms of labels along a degree of freedom.
+The density $\rho(p,q)$ will be defined on cells of infinitesimal area proportional to $dx \wedge dp$.
+When mapping one cell to another, the infinitesimal area will remain the same.
+The area in phase space of one degree of freedom can then be used as a measure for the cardinality of the labels.
+\ref{labelsCount} becomes area conservation for conjugate variables.
+Area conservation is equivalent to requiring the invariance of the vector product.
 
@@
 \end{align*}
-then $v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}=v^{\alpha} \omega_{\alpha, \beta} w^{\beta}$ under detrev transformation.
+then $v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}=v^{\alpha} \omega_{\alpha, \beta} w^{\beta}$ under a bijective map.
 \end{prop}
@@
 Let $v$ and $w$ be two vectors.
-Let $v^{\alpha} \omega_{\alpha, \beta} w^{\alpha}$ be an antisymmetric metric product conserved under a continuous transformation parameterized by $t$.
-We can then define a generating function $H$ such that:
-\begin{enumerate}
-\item $S^{\alpha} \equiv d_{t}x^{\alpha}$, vector field that tells us how the state variables change
-\item $S_{\beta} \equiv S^{\alpha} \omega_{\alpha, \beta}$
-\item $S_{\alpha} = \partial_{\alpha}H$
-\end{enumerate}
+Let $v^{\alpha} \omega_{\alpha, \beta} w^{\alpha}$ be an antisymmetric product conserved under a continuous transformation parameterized by $t$.
+We can then define a function $H$ such that given $S^{\alpha} \equiv d_{t}x^{\alpha}$, the vector field that tells us how the state variables change, $S_{\beta} \equiv S^{\alpha} \omega_{\alpha, \beta}$, we have $S_{\alpha} = \partial_{\alpha}H$.
 \end{lem}
@@
 \omega_{\alpha, \beta} =  \left[
-  \begin{array}{cccc}
-    0 & 1 & 0 & 0 \\
-    -1 & 0 & 0 & 0 \\
-    0 & 0 & 0 & 1 \\
-    0 & 0 & -1 & 0 \\
-  \end{array}
-\right] \\
-\end{align*}
-then $v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}=v^{\alpha} \omega_{\alpha, \beta} w^{\beta}$ under detrev transformation.
-\end{prop}
-
-The orthogonality of the labels corresponds to orthogonality in the manifold.
-Consider a hypercube oriented in such a way that some sides align to the first degree of freedom and some sides align to the second.
-Let $V$ its hypervolume and $V'$ the hypervolume of its map.
-and Let $A_1$ and $A_2$ the area of two regions of each degree of freedom.
-Let $A_1'$ and $A_2'$ be the corresponding maps.
-From \ref{labelsCount} and \ref{labelsMultiCount} we have $V=V'=A_1A_2=A_1'A_2'$, which can hold only if the two degrees of freedom are orthogonal and remain orthogonal.
-This means we need to require the conservation of the scalar product across independent degrees of freedoms, while still requiring conservation of the vector product within.
-Our metric will be given by:
-\begin{align*}
-\omega_{\alpha, \beta} =  \left[
   \begin{array}{cc}
@@
   \end{array}
+\right] =
+\left[
+  \begin{array}{cccc}
+    0 & 1 & 0 & 0 \\
+    -1 & 0 & 0 & 0 \\
+    0 & 0 & 0 & 1 \\
+    0 & 0 & -1 & 0 \\
+  \end{array}
 \right] \\
 \end{align*}
-The metric corresponds to a measure of the labels defined on the area given by two arbitrary vectors.
+then $v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}=v^{\alpha} \omega_{\alpha, \beta} w^{\beta}$ under a bijective map.
+\end{prop}
+
+The orthogonality of the labels corresponds to orthogonality in phase space, which the mapping needs to preserve across independent degrees of freedom.
+The cardinality of labels within each degree of freedom corresponds to the area in phase space\footnote{We assume we are using the same unit across d.o.f.}, which the mapping also needs to preserve.
+This is equivalent to require the conservation of the scalar product across independent degrees of freedoms, while still requiring conservation of the vector product within.
+That leads us to the metric defined by \ref{mdofInvariance}.
+The metric generalizes \ref{sdofInvariance} to give us the cardinality of labels defined on the area given by two arbitrary vectors.
 For an infinitesimal region, this corresponds to $dx^1 \wedge dp^1 + dx^2 \wedge dp^2$, the sum of the projections on the independent planes.
@@
 \section{Time dependence}
-So far we have assumed that both state labeling and mapping do not change in time. To be general, then we have to introduce an extra degree of freedom.
+So far we have assumed that both state labeling and mapping do not change in time. To be able to use time as a label, we now introduce an extra degree of freedom.
 
@@
 There must exist a strictly monotonic function $t(s)$.
-States connected by a trajectory where $dt/ds>0$ are called particle states, while the ones for which $dt/ds<0$ are called anti-particle states.
+States are connected by a trajectory where either $dt/ds>0$ (particle states) of $dt/ds<0$ (anti-particle states).
 \end{prop}
@@
 The trajectory has to be continuous in both spatial and temporal variables because of \ref{continuousMapping}.
-The mapping still needs to give one and only one value in time, so in principle we should be able to use $t$ as a parameter, we much have an invertible mapping between $t$ and $s$, which means we must have a strictly monotonic $t(s)$.
+The mapping still needs to give one and only one value in time, so we must have an invertible mapping between $t$ and $s$, which means we must have a strictly monotonic $t(s)$.
 Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call particle and anti-particle states is also conventional.

### 2014-06-15 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Reviewed explanation for time d.o.f.

@@
 \begin{prop}\label{sdofMap}
-Let $f: X \mapsto X$ be a bijective map on a degree of freedom. Then $dx' \wedge dp' = dx \wedge dp$.
+Let $f: X \mapsto X$ be a bijective map on a degree of freedom. Then $dq' \wedge dp' = dq \wedge dp$.
 \end{prop}
@@
 This is the generalization of \ref{widthMapping} for two state variable.
-The density $\rho(p,q)$ will be defined on cells of infinitesimal area proportional to $dx \wedge dp$.
+The density $\rho(p,q)$ will be defined on cells of infinitesimal area proportional to $dq \wedge dp$.
 When mapping one cell to another, the infinitesimal area will remain the same.
@@
 The metric generalizes \ref{sdofInvariance} to give us the cardinality of labels defined on the area given by two arbitrary vectors.
-For an infinitesimal region, this corresponds to $dx^1 \wedge dp^1 + dx^2 \wedge dp^2$, the sum of the projections on the independent planes.
+For an infinitesimal region, this corresponds to $dq^1 \wedge dp^1 + dq^2 \wedge dp^2$, the sum of the projections on the independent planes.
 
@@
 \begin{prop}\label{tdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of the extended phase space for one degree of freedom. Let $\alpha$ and $\beta$ be indexes for $t, E, x, p$. Let
+Let $v$ and $w$ be two vectors defined on the tangent space of the extended phase space for one degree of freedom. Let $\alpha$ and $\beta$ be indexes for $t, E, q, p$. Let
 \begin{align*}
-\omega_{\alpha, \beta} = \omega_{\alpha, \beta} =  \left[
+\omega_{\alpha, \beta} =  \left[
   \begin{array}{cc}
@@
 $\langle T, E \rangle$ are not independent labels from $\langle Q, P \rangle$ as they do not define new states.
-They are not necessarily orthogonal directions in the extended phase space.
-Looking back at \ref{disclabelsoverr}, cells need to be defined on the plane where $\langle Q, P \rangle$ changes: they are not defined on the plane of constant $\langle E, T \rangle$ (they are not orthogonal), but on the plane perpendicular to $\langle Q, P \rangle$.
-Areas defined on that plane are then going to be invariant.
+So they are not necessarily orthogonal in the extended phase space.
+Looking back at \ref{disclabelsoverr}, cells need to be defined on the plane where $\langle Q, P \rangle$ (maximally) changes: this is not defined on the plane of constant $\langle E, T \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but on the plane perpendicular to constant $\langle Q, P \rangle$ where $dt \wedge dE$ is defined.
+On that plane we can properly count states and define our invariant.
 
-Like before, we have three planes of conjugate variables forming a right triangle-like relationship with one side invariant.
-In this case, the right angle is between the invariant and the plane of constant $\langle Q, P \rangle$ (on which $dt$ and $dE$ are defined).
-So we must have $dt \wedge dE + k = dq \wedge dp$.
+As in the case of independent d.o.f., we have three planes of conjugate variables forming a right triangle-like relationship with one side invariant.
+\begin{align*}
+dq^1 \wedge dp^1 + dq^2 \wedge dp^2 = k \\
+dt \wedge dE + k = dq \wedge dp
+\end{align*}
+In the previous case, the right angle was between the two independent degree of freedom.
+In this case, the right angle is between the invariant and the plane of constant $\langle Q, P \rangle$ where $dt \wedge dE$ is defined.
+We rewrite it as $dq \wedge dp - dt \wedge dE = k$.
 Which corresponds to the Minkowski product across d.o.f.
 and the vector product within.
-The metric, with a space-like convention, gives us a measure for states.
+The metric, with a space-like convention, still gives us the cardinality of labels within a degree of freedom.
 

### 2014-06-15 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
More notation and language cleanup

@@
 \begin{prop}\label{detMap}
-Let $I$ be a set of labels of a system that undergoes deterministic and reversible evolution. There exists a bijective map $f:i \in I \leftrightarrow i' \in I$.
+Let $I$ be a set of labels of a system that undergoes deterministic and reversible evolution. There exists a bijective map $f:I \leftrightarrow I$.
 \end{prop}
@@
 Let $I_1$ and $I_2$ be two orthogonal sets of labels and $I = I_1 \times I_2$.
-Let $f:i \in I \leftrightarrow i' \in I$ be a bijective map.
+Let $f:I \leftrightarrow I$ be a bijective map.
 The labels are said independent iff we can define two sets of orthogonal labels $I_1'$ and $I_2'$ such that we can define separate mappings:
@@
 \item $I_1' \times I_2' = I$
-\item there exists a bijective map $f_1: i_1 \in I_1 \leftrightarrow i_1' \in I_1'$
-\item there exists a bijective map $f_2: i_2 \in I_2 \leftrightarrow i_2' \in I_2'$
+\item there exists a bijective map $f_1: I_1 \leftrightarrow I_1'$
+\item there exists a bijective map $f_2: I_2 \leftrightarrow I_2'$
 \item $f(i) = \{f_1(i_1), f_2(i_2)\}$
@@
 \begin{cor}\label{discreteEv}
-Let $f: i \in I \leftrightarrow i'$ be a bijective map, we have $\mathbbm{c}_j=\sum D'(i') \mathbbm{s}_i'=\sum \rho'(i') w(i') \mathbbm{s}_i' = \sum D(i) \mathbbm{s}_i' = \sum \rho(i) w(i) \mathbbm{s}_i'$.
-$\rho'(i') = \rho(i) w(i) / w(i')$.
+Let $f: I \leftrightarrow I$ be a bijective map, we have $\mathbbm{c}_j=\sum D'(i) \mathbbm{s}_i=\sum \rho'(i) w(i) \mathbbm{s}_i = \sum D(f^{-1}(i)) \mathbbm{s}_i = \sum \rho(f^{-1}(i)) w(f^{-1}(i)) \mathbbm{s}_i$.
+$\rho'(i) = \rho(f^{-1}(i)) w(f^{-1}(i)) / w(i)$.
 \end{cor}
@@
 
-With homogeneous labels no adjusting is needed, and the range can be used as a measure of the number of labels.
+With homogeneous labels no adjustment is needed, and the range can be used as a measure of the number of labels.
 
@@
 \begin{prop}\label{continuousMapping}
-Let $f: X \mapsto X$ be a bijective map on a state variable. The mapping must be continuous.
+Let $f: X \leftrightarrow X$ be a bijective map on a state variable. The mapping must be continuous.
 \end{prop}
@@
 \begin{prop}\label{widthMapping}
-Let $f: X \mapsto X$ be a deterministic and reversible mapping on a state variable.
+Let $f: X \leftrightarrow X$ be a bijective map on a state variable.
 Then $m(x') dx' \equiv m(f(x)) \frac{\partial f(x)}{\partial x} dx = m(x) dx$.
@@
 \begin{prop}\label{sdofMap}
-Let $f: X \mapsto X$ be a bijective map on a degree of freedom. Then $dq' \wedge dp' = dq \wedge dp$.
+Let $f: X \leftrightarrow X$ be a bijective map on a degree of freedom. Then $dq' \wedge dp' = dq \wedge dp$.
 \end{prop}
@@
 \begin{align*}
-V^{\alpha} \Omega_{\alpha, \beta} W^{\beta} &= V'^{\alpha} \Omega_{\alpha, \beta} W'^{\beta}  \\
-&= (V^{\alpha} + \partial_{\gamma} S^{\alpha} dt V^{\gamma}) \Omega_{\alpha, \beta} ( W^{\beta} + \partial_{\delta} S^{\beta} W^{\delta} dt) \\
-&= V^{\alpha} \Omega_{\alpha, \beta} W^{\beta} + (\partial_{\gamma} S^{\alpha} V^{\gamma} \Omega_{\alpha, \beta} W^{\beta} \\
- &+ V^{\alpha} \Omega_{\alpha, \beta} \partial_{\delta} S^{\beta} W^{\delta}) dt + O(dt^2)
+v^{\alpha} \omega_{\alpha, \beta} w^{\beta} &= v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}  \\
+&= (v^{\alpha} + \partial_{\gamma} S^{\alpha} dt v^{\gamma}) \omega_{\alpha, \beta} ( w^{\beta} + \partial_{\delta} S^{\beta} w^{\delta} dt) \\
+&= v^{\alpha} \omega_{\alpha, \beta} w^{\beta} + (\partial_{\gamma} S^{\alpha} v^{\gamma} \omega_{\alpha, \beta} w^{\beta} \\
+ &+ v^{\alpha} \omega_{\alpha, \beta} \partial_{\delta} S^{\beta} w^{\delta}) dt + O(dt^2)
 \end{align*}
 \begin{align*}
-V^{\gamma} W^{\beta} \partial_{\gamma} S_{\beta} - V^{\alpha} W^{\delta} \partial_{\delta} S_{\alpha} = 0
+v^{\gamma} w^{\beta} \partial_{\gamma} S_{\beta} - v^{\alpha} w^{\delta} \partial_{\delta} S_{\alpha} = 0
 \end{align*}
@@
 The trajectory has to be continuous in both spatial and temporal variables because of \ref{continuousMapping}.
-The mapping still needs to give one and only one value in time, so we must have an invertible mapping between $t$ and $s$, which means we must have a strictly monotonic $t(s)$.
+Since determinism and reversibility are defined in time, the trajectory must traverse all times once and only once: we must have an invertible mapping between $t$ and $s$, which means we must have a strictly monotonic $t(s)$.
+$dt/ds$ along a trajectory cannot change sign, so we have the division between particle and anti-particle states.
 Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call particle and anti-particle states is also conventional.
-What is physical, instead, is that $dt/ds$ along a trajectory cannot change sign, so particle and anti-particle states cannot be connected by deterministic and reversible evolution.
+What is physical and not conventional, though, is that particle and anti-particle states cannot be connected by deterministic and reversible evolution.
 

### 2014-06-15 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Settled on a notation for energy and extended hamiltonian

@@
 \begin{prop}\label{mdofInvariance}
-Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by two independent degrees of freedom. Let
+Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by two independent degrees of freedom.
+Let $\alpha$ and $\beta$ be indexes for the state variables $q^1, p^1, q^2, p^2$.
+Let
 \begin{align*}
@@
 \begin{prop}\label{tdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of the extended phase space for one degree of freedom. Let $\alpha$ and $\beta$ be indexes for $t, E, q, p$. Let
+Let $v$ and $w$ be two vectors defined on the tangent space of the extended phase space for one degree of freedom. Let $\alpha$ and $\beta$ be indexes for the state variables $t, e, q, p$. Let
 \begin{align*}
@@
 So they are not necessarily orthogonal in the extended phase space.
-Looking back at \ref{disclabelsoverr}, cells need to be defined on the plane where $\langle Q, P \rangle$ (maximally) changes: this is not defined on the plane of constant $\langle E, T \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but on the plane perpendicular to constant $\langle Q, P \rangle$ where $dt \wedge dE$ is defined.
+Looking back at \ref{disclabelsoverr}, cells need to be defined on the plane where $\langle Q, P \rangle$ (maximally) changes: this is not the plane of constant $\langle E, T \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle Q, P \rangle$ where $dt \wedge de$ is defined.
 On that plane we can properly count states and define our invariant.
@@
 dq^1 \wedge dp^1 + dq^2 \wedge dp^2 = k \\
-dt \wedge dE + k = dq \wedge dp
+dt \wedge de + k = dq \wedge dp
 \end{align*}
 In the previous case, the right angle was between the two independent degree of freedom.
-In this case, the right angle is between the invariant and the plane of constant $\langle Q, P \rangle$ where $dt \wedge dE$ is defined.
-We rewrite it as $dq \wedge dp - dt \wedge dE = k$.
+In this case, the right angle is between the invariant and the plane of constant $\langle Q, P \rangle$ where $dt \wedge de$ is defined.
+We rewrite it as $dq \wedge dp - dt \wedge de = k$.
 Which corresponds to the Minkowski product across d.o.f.
@@
 
+\begin{prop}\label{tdofHam}
+The evolution for time varying multiple degrees of freedom is given by:
+\begin{align*}
+d_{s}t &= - \partial_{e} \mathcal{H} \\
+d_{s}e &= \partial_{t} \mathcal{H} \\
+d_{s}q^i &= \partial_{p^i} \mathcal{H} \\
+d_{s}p^i &= - \partial_{q^i} \mathcal{H}
+\end{align*}
+\end{prop}
+
+Expand \ref{genAntisim} with the metric defined in \ref{tdofInvariant}, calling the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$.
+We recognise Hamilton equations in the extended phase space.
+
+It should not be a surprise that the equations do not mention $c$.
+In fact, nothing says that all $q^i$ represent space, and with the same unit, and that the laws of motion are invariant in all inertial frames.
+The only requirement we have is that the areas of each degrees of freedom represent the same cardinality for labels.
+
 \begin{prop}\label{tdofConstrain}
-The evolution is constrained by $H_{e}=k$, the generating function defined in \ref{genAntisim} generalized for the extended phase space.
+The evolution is constrained by $\mathcal{H}=k$, the generating function defined in \ref{genAntisim} generalized for the extended phase space.
 \end{prop}
@@
 Phase space evolved in time is $\mathbb{R}^{2*N + 1}$.
-Since $H_{e}$ is constant through the evolution, it can serve both as the generating function and as the evolution constrain.
-By convention, we set $H_{e}=0$.
+Since $\mathcal{H}$ is constant through the evolution, it can serve both as the generating function and as the evolution constrain.
+By convention, we set $\mathcal{H}=0$.
 
-\begin{prop}\label{tdofHam}
-The evolution for time varying multiple degrees of freedom is given by:
-\begin{align*}
-d_{s}t &= - \partial_{E} H_{e} \\
-d_{s}E &= \partial_{t} H_{e} \\
-d_{s}q^i &= \partial_{p^i} H_{e} \\
-d_{s}p^i &= - \partial_{q^i} H_{e}
-\end{align*}
-\end{prop}
-
-Expand \ref{genAntisim} with the metric defined in \ref{tdofInvariant}. We recognise Hamilton equations in the extended phase space.
-
-It should not be a surprise that the equations do not mention $c$.
-In fact, nothing says that all $q^i$ represent space, and with the same unit, and that the laws of motion are invariant in all inertial frames.
-The only requirement we have is that the areas of each degrees of freedom represent the same measure for labels.
-
+\textbf{Example}.
 We wrap up with an example.
-Let $H_{E}= mc^2 + ((p^i)^2 - E^2/c^2) / 2m$.
-This represents the rest energy of a free particle, the conjugate of which is proper time.
-Therefore the parametrization of a particle-state is $s=\tau$ while the parametrization for an anti-particle state is $s=-\tau$.
+Let $\mathcal{H}= mc^2 + ((p^i)^2 - E^2/c^2) / 2m$.
+The non-constant term represents the rest energy of a free particle, the conjugate of which is proper time.
+Therefore the parametrization of a particle state is $s=\tau$ while the parametrization for an anti-particle state is $s=-\tau$.
 If we apply \ref{tdofHam} we have:

### 2014-06-17 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Another pass at language and math details.

@@
 Each configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}_i$.
-In physics term, the distribution can be visualized as a histogram of discrete data.
+In physics term, the distribution can be visualized as a histogram over the discrete labels.
 \end{cor}
@@
 \item there exists a bijective map $f_2: I_2 \leftrightarrow I_2'$
-\item $f(i) = \{f_1(i_1), f_2(i_2)\}$
+\item $f(i) = \langle f_1(i_1), f_2(i_2)\rangle$
 \end{enumerate}
@@
 
-\begin{cor}\label{disclabelsoverrdist}
+\begin{prop}\label{disclabelsoverrdist}
 Each configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i=\sum \rho(i) w(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ is defined as before, and $\rho(i)\equiv D(i) / w(i)$ is the density of the distribution for the cell.
 In physics term, the distribution can be visualized as a histogram where $w(i)$, $\rho(i)$ and $D(i)$ are respectively the width, height and area of the each bin.
-\end{cor}
+\end{prop}
 
@@
 In the limit we'll have a cell for each value, so we can use $x(i)$ (or simply $x$) instead of $i$ for the label.
-$\rho$ and $m$ will converge to functions defined over $X$.
-The corresponding configuration state will become $\mathbbm{c}_j=\int \rho(x) m(x) \mathbbm{s}_x ds$.
+$\rho$ and $m$ will converge to functions defined over $x$.
+The corresponding configuration state will become $\mathbbm{c}_j=\int \rho(x) m(x) \mathbbm{s}_x dx$.
 
@@
 \begin{prop}\label{widthMapping}
-Let $f: X \leftrightarrow X$ be a bijective map on a state variable.
-Then $m(x') dx' \equiv m(f(x)) \frac{\partial f(x)}{\partial x} dx = m(x) dx$.
-If $X$ is homogeneous, then $dx' = dx$ and the range gives us a measure of the number of states.
+Let $f: X \leftrightarrow X$ be a bijective map on a state variable, and $x'=f(x)$.
+Then $\frac{dx'}{dx} = \frac{m(x')}{m(x)} $.
+If $X$ is homogeneous, then $dx' = dx$ and the range gives us a measure of the cardinality of labels.
 \end{prop}
 The mapping must be done so that the width of the cells is mapped as well, not just the center value.
+The width of the transported cell $m(x)dx \rightarrow m(x) \frac{dx'}{dx} dx$ must be equal to the width of the target cell $m(x')dx$.
 
@@
 
-This is the generalization of \ref{widthMapping} for two state variable.
+This is the equivalent of \ref{widthMapping} for two state variable.
 The density $\rho(p,q)$ will be defined on cells of infinitesimal area proportional to $dq \wedge dp$.
@@
 
-This is just expanding \ref{genAntisim} with the metric defined in \ref{sdofInvariance}. We recognise Hamilton's equations for one degree of freedom.
+Simply expand \ref{genAntisim} with the metric defined in \ref{sdofInvariance}. We recognise Hamilton's equations for one degree of freedom.
 
@@
 \section{Time dependence}
-So far we have assumed that both state labeling and mapping do not change in time. To be able to use time as a label, we now introduce an extra degree of freedom.
+So far we have assumed that both state labeling and mapping do not change in time. In that case, we need to to use time as a label. We introduce an extra degree of freedom.
 
@@
 There must exist a strictly monotonic function $t(s)$.
-States are connected by a trajectory where either $dt/ds>0$ (particle states) of $dt/ds<0$ (anti-particle states).
+States are connected by a trajectory where either $dt/ds>0$ (particle states) or $dt/ds<0$ (anti-particle states).
 \end{prop}
@@
 \begin{align*}
-dq^1 \wedge dp^1 + dq^2 \wedge dp^2 = k \\
-dt \wedge de + k = dq \wedge dp
+&dq^1 \wedge dp^1 + dq^2 \wedge dp^2 = k \\
+&dt \wedge de + k = dq \wedge dp
 \end{align*}
@@
 
-Expand \ref{genAntisim} with the metric defined in \ref{tdofInvariant}, calling the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$.
+Expand \ref{genAntisim} with the metric defined in \ref{tdofInvariant}, with the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$.
 We recognise Hamilton equations in the extended phase space.
@@
 Since $\mathcal{H}$ is constant through the evolution, it can serve both as the generating function and as the evolution constrain.
-By convention, we set $\mathcal{H}=0$.
+By convention, we can set $\mathcal{H}=0$ without loss of generality as changing $\mathcal{H}$ by a constant does not change the equation of motion.
 
@@
 Let $\mathcal{H}= mc^2 + ((p^i)^2 - E^2/c^2) / 2m$.
-The non-constant term represents the rest energy of a free particle, the conjugate of which is proper time.
-Therefore the parametrization of a particle state is $s=\tau$ while the parametrization for an anti-particle state is $s=-\tau$.
+The constant term represents the rest energy of a free particle, the conjugate of which is proper time.
+The parametrization of a particle state is $s=\tau$ while the parametrization for an anti-particle state is $s=-\tau$.
 If we apply \ref{tdofHam} we have:
@@
 \section{Conclusion}
-I hope this work convinced you that we can derive Hamiltonian mechanics from simple definitions of label, states, determinism and reversibility.
-I believe this approach helps give direct physical meaning to phase space and its geometric properties, and it shows that Hamiltonian mechanics can stand on its own, without Lagrangians.
-I find it uncanny how so much can be derived with so little.
+By deriving Hamiltonian mechanics from simple definitions of labels, states, determinism and reversibility we have given more direct physical meaning to phase space and its geometric properties, and shown that a good part of the Hamiltonian framework can stand on its own, without Lagrangians.
+No mathematical breakthrough is revealed here, yet I am not aware of any work that brings all the pieces of the puzzle in quite this way: so much derived from so little.
 
-While no mathematical breakthrough is revealed (we have just put together in a novel way the known pieces of the puzzle), I am not aware of any work that focus on this aspect.
-We already have ideas on how to extend this work to Quantum Mechanics, and we hope to gain insight on how to extend to General Relativity.
+The hope is that, by continuing in this approach, we can shed more light on why nature has the laws it has; and show that they are not some arbitrary rules, but indeed necessary given few simple assumptions.
 

### 2014-06-22 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Fixed introduction. Better introduction and meaning for the metric. Other improvements.

@@
 In this work we aim to derive it on its own, starting from simple physical assumptions, without any reference to Lagrangian mechanics.
-We believe that provides more direct insight in the theory and a direct physical interpretation of its geometric features.
 
-Hamiltonian is either math or physics from Lagrangian. Lagrangian are not well motivated anyway.
+We'll give simple definitions of states in terms of labels and determinism/reversibility in terms of bijective maps; define a metric based on the idea that bijective maps preserve the cardinality of labels; obtain Hamilton's equations from the invariance of that metric.
 
-We start from scratch and find Hamiltonian. $dt \wedge dE$
+No mathematical breakthrough should be expected: the goal, after all, is to derive the \emph{known} framework from a set of \emph{simple} definitions in the most \emph{obvious} way possible.
+All proofs are no longer than a couple of paragraphs, so the word theorem is avoided in favor of proposition and corollary.
 
@@
 \textbf{State definition}.
-Let's assume  we have a physical system that can be found in different configurations.
-Each unique configuration, or configuration state, is identified by a label.
-For our purposes, it does not matter what it is (a measurement, a way to prepare the system, future behavior, ...), just that it exists.
+We assume we have a physical system that can be found in different configurations.
+We call each unique configuration \emph{configuration state} and assign it a \emph{label}.
+For our purposes, it does not matter what the label is (a measurement, a way to prepare the system, future behavior, ...), just that it exists.
 
@@
 \begin{cor}\label{statedistr}
-Each configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}_i$.
+Each classical configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}_i$.
 In physics term, the distribution can be visualized as a histogram over the discrete labels.
@@
 \begin{assump}\label{detAss}
-The system undergoes deterministic (future state identified by the present state) and reversible (past state identified by the present state) evolution.
+(\emph{Determinism and Reversibility}) The system undergoes deterministic (future state identified by the present state) and reversible (past state identified by the present state) evolution.
 \end{assump}
@@
 \begin{cor}\label{discreteEv}
-The evolution of a reducible configuration state $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$ under a bijective map is given by $\mathbbm{c}_{j'}=\sum D'(i) \mathbbm{s}_{i}=\sum D(f^{-1}(i)) \mathbbm{s}_{i}$.
+The evolution of a classical configuration state $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$ under a bijective map is given by $\mathbbm{c}_{j'}=\sum D'(i) \mathbbm{s}_{i}=\sum D(f^{-1}(i)) \mathbbm{s}_{i}$.
 \end{cor}
 
-Mathematically, studying \ref{detAss} means studying bijective maps.
-The evolution of a distribution simply moves the elements around: the bars of the discrete histogram move place, but keep the same height.
+Mathematically, studying \ref{detAss} means studying bijective maps. The evolution of a distribution simply moves the elements around: the bars of the histogram move place, but keep the same height.
 
@@
 Let $f:I \leftrightarrow I$ be a bijective map.
-The labels are said independent iff we can define two sets of orthogonal labels $I_1'$ and $I_2'$ such that we can define separate mappings:
+The labels are said independent iff we can define two sets of orthogonal labels $I_1'$ and $I_2'$ such that we can define separate maps:
 \begin{enumerate}
@@
 
-\begin{cor}\label{labelsCount}
+\begin{prop}\label{labelsCount}
 Let $\hat{I} \subset I$ of finite size $n(\hat{I})$.
@@
 We then have $n(\hat{I})=n(f(\hat{I}))=n(f^{-1}(\hat{I}))$, the number of states is conserved.
-\end{cor}
+\end{prop}
 
-\begin{cor}\label{labelsMultiCount}
+\begin{prop}\label{labelsMultiCount}
 Let $\hat{I_1} \subset I_1$ and $\hat{I_2} \subset I_2$ of finite size $n(\hat{I_1})$ and $n(\hat{I_2})$.
@@
 We then have $n(\hat{I_1}\times\hat{I_2})=n(\hat{I_1})n(\hat{I_2})=n(f(\hat{I_1})) n(f(\hat{I_2}))=n(f^{-1}(\hat{I_1})) n(f^{-1}(\hat{I_2}))=n(\hat{f(I_1)}\times\hat{f(I_2)})=n(f^{-1}(\hat{I_1})\times f^{-1}(\hat{I_2}))$, the total number of states is conserved and remains the product of the number of labels in each set.
-\end{cor}
+\end{prop}
 
-These statements, properly generalized for the continuous case, will allow us to define a metric. Conserving that metric will lead to Hamiltonian mechanics.
+Bijective maps preserve the number of states as they provide one-to-one association between future and past labels.
 
@@
 \begin{prop}\label{disclabelsoverrdist}
-Each configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i=\sum \rho(i) w(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ is defined as before, and $\rho(i)\equiv D(i) / w(i)$ is the density of the distribution for the cell.
+Each classical configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i=\sum \rho(i) w(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ is defined as before, and $\rho(i)\equiv D(i) / w(i)$ is the density of the distribution for the cell.
 In physics term, the distribution can be visualized as a histogram where $w(i)$, $\rho(i)$ and $D(i)$ are respectively the width, height and area of the each bin.
@@
 
-To prepare for the limit we define $m(i)=w(i)/\Delta w$, where $\Delta w$ represents the average width of the cells.
-We increase the number of the cells and reduce $\Delta w$ such that $max(m(i))$ is bounded.
+To prepare for the limit we define $m(i)=w(i)/\bar{w}$, where $\bar{w}$ represents the average width of the cells.
+We increase the number of the cells and reduce $\bar{w}$ while keeping $m(i)$ finite.
 In the limit we'll have a cell for each value, so we can use $x(i)$ (or simply $x$) instead of $i$ for the label.
@@
 Let $f: X \leftrightarrow X$ be a bijective map on a state variable, and $x'=f(x)$.
-Then $\frac{dx'}{dx} = \frac{m(x')}{m(x)} $.
-If $X$ is homogeneous, then $dx' = dx$ and the range gives us a measure of the cardinality of labels.
+Then $dx' = \frac{m(x')}{m(x)} dx$.
+If $X$ is homogeneous, then $dx' = dx$, the range gives us a measure of the cardinality of labels and must be conserved.
 \end{prop}
 The mapping must be done so that the width of the cells is mapped as well, not just the center value.
-The width of the transported cell $m(x)dx \rightarrow m(x) \frac{dx'}{dx} dx$ must be equal to the width of the target cell $m(x')dx$.
+The width of the transported cell $m(x)dx \rightarrow m(x) dx'$ must be equal to the width of the target cell $m(x')dx$, which gives us the first part.
+If $X$ is homogeneous, $m(x)=m(x')=1$, which gives us the second part.
+
+Let $\Delta x$ be a finite range of $X$. Before the limit, we have:
+\begin{align*}
+\Delta x = \frac{\Delta x}{1} = \frac{n(\Delta x) k}{n(1) k} = \frac{n(\Delta x)}{n(1)}
+\end{align*}
+where $n(\Delta x)$ and $n(1)$ are the number of labels/cells in the $\Delta x$ and unit range, while k their width.
+The range can be seen as the ration between the number of labels in the range and the number of labels in the unit range.
+That ratio must be conserved by the bijective map and remains well defined during the limit.
+In this sense, the range can be used as a measure of the cardinality of the labels.
 
@@
 \ref{labelsCount} becomes area conservation for conjugate variables.
-Area conservation is equivalent to requiring the invariance of the vector product.
 
@@
 \end{prop}
+
+Area conservation is equivalent to requiring the invariance of the vector product, which is what $\omega_{\alpha, \beta}$ represents.
 
@@
 So they are not necessarily orthogonal in the extended phase space.
-Looking back at \ref{disclabelsoverr}, cells need to be defined on the plane where $\langle Q, P \rangle$ (maximally) changes: this is not the plane of constant $\langle E, T \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle Q, P \rangle$ where $dt \wedge de$ is defined.
+Looking back at \ref{disclabelsoverr}, cells need to be defined on the plane where $\langle Q, P \rangle$ (maximally) change: this is not the plane of constant $\langle E, T \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle Q, P \rangle$ where $dt \wedge de$ is defined.
 On that plane we can properly count states and define our invariant.
@@
 and the vector product within.
-The metric, with a space-like convention, still gives us the cardinality of labels within a degree of freedom.
+The metric, with a space-like convention, still gives us the cardinality of labels within a degree of freedom.\footnote{Adjusted to avoid double counting}
 
@@
 \begin{prop}\label{tdofConstrain}
-The evolution is constrained by $\mathcal{H}=k$, the generating function defined in \ref{genAntisim} generalized for the extended phase space.
+The evolution is constrained by $\mathcal{H}=k$.
 \end{prop}
 
-Given N degrees of freedom, phase space is $\mathbb{R}^{2*N}$.
-Extended phase space is $\mathbb{R}^{2*N + 2}$.
-Phase space evolved in time is $\mathbb{R}^{2*N + 1}$.
 Since $\mathcal{H}$ is constant through the evolution, it can serve both as the generating function and as the evolution constrain.
-By convention, we can set $\mathcal{H}=0$ without loss of generality as changing $\mathcal{H}$ by a constant does not change the equation of motion.
+By convention, we could set $\mathcal{H}=0$ without loss of generality as changing $\mathcal{H}$ by a constant does not change the equation of motion.
+This reduces extended phase space to $\mathbb{R}^{2*N + 1}$, the state labels plus time.
 
-\textbf{Example}.
-We wrap up with an example.
-Let $\mathcal{H}= mc^2 + ((p^i)^2 - E^2/c^2) / 2m$.
-The constant term represents the rest energy of a free particle, the conjugate of which is proper time.
-The parametrization of a particle state is $s=\tau$ while the parametrization for an anti-particle state is $s=-\tau$.
-If we apply \ref{tdofHam} we have:
+\textbf{Example}. We wrap up with an example. Let $\mathcal{H} = mc^2 + ((p^i)^2 - E^2/c^2) / 2m$.  If we apply \ref{tdofHam} we have:
 \begin{align*}
@@
 \end{align*}
-For particles, we have:
+For particles, let $s=\tau$:
 \begin{align*}
@@
 \end{align*}
-so we recognize the four-momentum $P^\alpha = {E/c, p^i}$. For anti-particles:
+so we recognize the four-momentum $P^\alpha = {E/c, p^i}$ and $\tau$ proper time. For anti-particles, let $s=-\tau$:
 \begin{align*}
@@
 \end{align*}
-we end up with a minus sign between the four-momentum and the conjugate variables $P^\alpha = {-E/c, -p^i}$.
+we end up with a minus sign between the four-momentum and the conjugate variables $P^\alpha = {-E/c, -p^i}$.\footnote{This is consistent with a free Dirac field ...} $\mathcal{H}$ relates to the rest energy, the parametrization to proper time (its conjugate).
+Parametrization and time are aligned for particles and anti-aligned for anti-particles.
 

### 2014-06-23 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Highlight terms in their definitions. Few other minor improvements.

@@
 
-\title{Deriving Hamiltonian mechanics from physics, without passing through Lagrangians.}
+\title{Deriving Hamiltonian mechanics from physics, without passing through Lagrangian.}
 \author{Gabriele Carcassi}
@@
 \email{carcassi@bnl.gov}
-\date{August 25, 2013}
+\date{May 23, 2014}
 
@@
 We derive the Hamiltonian formulation of classical mechanics directly, without reference to Lagrangian mechanics.
-We start from the definition of states and the labels used to identify them, and show how simple properties required
-on those labels lead to the Hamiltonian formulation.\end{abstract}
+We start from the definition of states in terms of labels used to identify them, and show how, under a deterministic and
+reversible process, the conservation of the cardinality of the labels leads to Hamilton's equations.\end{abstract}
 
@@
 Let $I_1$ and $I_2$ be two sets of labels.
-We call $I = \langle I_1, I_2 \rangle$ a combined set of labels if an injective map exists such that $\forall i \in I \rightarrowtail i_1 \in I_1, i_2 \in I_2$.
+We call $I = \langle I_1, I_2 \rangle$ a \emph{combined set of labels} if an injective map exists such that $\forall i \in I \rightarrowtail i_1 \in I_1, i_2 \in I_2$.
 \end{defn}
@@
 \begin{defn}\label{orth}
-The two label sets $I_1$ and $I_2$ are said to be orthogonal iff $I = \langle I_1, I_2 \rangle = I_1 \times I_2$.
+The two label sets $I_1$ and $I_2$ are said to be \emph{orthogonal} iff $I = \langle I_1, I_2 \rangle = I_1 \times I_2$.
 \end{defn}
@@
 \begin{defn}\label{statedef}
-Let $\mathbbm{S}_I$ be the set of all possible configuration states of the infinitesimal subdivision. We call this set phase space. We call each element $\mathbbm{s}_i$ a state.
+Let $\mathbbm{S}_I$ be the set of all possible configuration states of the infinitesimal subdivision. We call this set \emph{phase space}. We call each element $\mathbbm{s}_i$ a \emph{state}.
 \end{defn}
@@
 
-\textbf{State mapping}. We will now concentrate ourselves on deterministic and reversible evolution.
+\textbf{State mapping}. We will now concentrate on deterministic and reversible evolution.
 \begin{assump}\label{detAss}
@@
 Let $f:I \leftrightarrow I$ be a bijective map.
-The labels are said independent iff we can define two sets of orthogonal labels $I_1'$ and $I_2'$ such that we can define separate maps:
+The labels are said \emph{independent} iff we can define two sets of orthogonal labels $I_1'$ and $I_2'$ such that we can define separate maps:
 \begin{enumerate}
@@
 For each cell we have a center value $x: I \mapsto \mathbb{R}$ and a width $w: I \mapsto \mathbb{R}$.
-$I$ is a set of discrete labels over a continuous range.
+$I$ is a set of \emph{discrete labels over a continuous range}.
 \end{defn}
@@
 \begin{defn}\label{discreteHomogeneous}
-A set $I$ of discrete labels over a continuous range is said to be homogeneous if $w(i)=k$: the bins are of equal width.
+A set $I$ of discrete labels over a continuous range is said to be \emph{homogeneous} if $w(i)=k$: the bins are of equal width.
 \end{defn}
@@
 \begin{defn}\label{continuousLabels}
-A state variable $X$ is the continuous limit of a set $I$ of discrete labels over a continuous range.
+A \emph{state variable} $X$ is the continuous limit of a set $I$ of discrete labels over a continuous range.
 \end{defn}
@@
 \begin{defn}\label{sdof}
-A degree of freedom is a label set $X$ given by a pair of homogeneous and orthogonal state variables $P$ and $Q$.
-We call these conjugate variables.
+A \emph{degree of freedom} is a label set $X$ given by a pair of homogeneous and orthogonal state variables $P$ and $Q$.
+We call these \emph{conjugate variables}.
 A state variable is said conjugate of the first iff together they form a pair of conjugate variables.
@@
 
-This is the equivalent of \ref{widthMapping} for two state variable.
+This is the equivalent of \ref{widthMapping} for two state variables.
 The density $\rho(p,q)$ will be defined on cells of infinitesimal area proportional to $dq \wedge dp$.
@@
 
-\begin{prop}\label{sdofInvariance}
+\begin{prop}\label{sdofInvariant}
 Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by two conjugate variables. Let
@@
 
-Simply expand \ref{genAntisim} with the metric defined in \ref{sdofInvariance}. We recognise Hamilton's equations for one degree of freedom.
+Simply expand \ref{genAntisim} with the metric defined in \ref{sdofInvariant}. We recognise Hamilton's equations for one degree of freedom.
 
@@
 \begin{defn}\label{mdof}
-Two degrees of freedom are said independent if the corresponding label set $X^1$ and $X^2$ are independent.
+Two degrees of freedom are said \emph{independent} if the corresponding label set $X^1$ and $X^2$ are independent.
 \end{defn}
 
-\begin{prop}\label{mdofInvariance}
+\begin{prop}\label{mdofInvariant}
 Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by two independent degrees of freedom.
@@
 This is equivalent to require the conservation of the scalar product across independent degrees of freedoms, while still requiring conservation of the vector product within.
-That leads us to the metric defined by \ref{mdofInvariance}.
-The metric generalizes \ref{sdofInvariance} to give us the cardinality of labels defined on the area given by two arbitrary vectors.
+That leads us to the metric defined by \ref{mdofInvariant}.
+The metric generalizes \ref{sdofInvariant} to give us the cardinality of labels defined on the area given by two arbitrary vectors.
 For an infinitesimal region, this corresponds to $dq^1 \wedge dp^1 + dq^2 \wedge dp^2$, the sum of the projections on the independent planes.
@@
 
-Expand \ref{genAntisim} with the metric defined in \ref{mdofInvariance}. We recognise Hamilton equations for multiple degrees of freedom.
+Expand \ref{genAntisim} with the metric defined in \ref{mdofInvariant}. We recognise Hamilton equations for multiple degrees of freedom.
 
@@
 \begin{defn}\label{tdof}
-The temporal degree of freedom is a label set $X$ given by the pair of conjugate variables $T$ and $E$.
+The \emph{temporal degree of freedom} is a label set $X$ given by the pair of conjugate variables $T$ and $E$.
 We call extended phase space the outer product between phase space and the temporal degree of freedom.
@@
 
-As in the case of independent d.o.f., we have three planes of conjugate variables forming a right triangle-like relationship with one side invariant.
+As in the case of independent d.o.f., we have right triangle-like relationship between the plane where the invariant is defined and its projections on the planes defined by each d.o.f.:
 \begin{align*}
@@
 \end{align*}
-In the previous case, the right angle was between the two independent degree of freedom.
+In the previous case, the right angle was between the two independent d.o.f..
 In this case, the right angle is between the invariant and the plane of constant $\langle Q, P \rangle$ where $dt \wedge de$ is defined.
@@
 
-Expand \ref{genAntisim} with the metric defined in \ref{tdofInvariant}, with the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$.
+Take the metric from \ref{tdofInvariant}, add multiple independent d.o.f as in \ref{mdofInvariant}, use \ref{genAntisim} with the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$.
 We recognise Hamilton equations in the extended phase space.
 
-It should not be a surprise that the equations do not mention $c$.
+It should not be a surprise that the equations do not mention the speed of light $c$.
 In fact, nothing says that all $q^i$ represent space, and with the same unit, and that the laws of motion are invariant in all inertial frames.
@@
 By convention, we could set $\mathcal{H}=0$ without loss of generality as changing $\mathcal{H}$ by a constant does not change the equation of motion.
-This reduces extended phase space to $\mathbb{R}^{2*N + 1}$, the state labels plus time.
+This reduces extended phase space to $\mathbb{R}^{2*N + 1}$, the state variables plus time.
 
@@
 \end{align*}
-so we recognize the four-momentum $P^\alpha = {E/c, p^i}$ and $\tau$ proper time. For anti-particles, let $s=-\tau$:
+so we recognize the four-momentum $P^\alpha = [E/c, p^i]$ and $\tau$ proper time. For anti-particles, let $s=-\tau$:
 \begin{align*}
@@
 \end{align*}
-we end up with a minus sign between the four-momentum and the conjugate variables $P^\alpha = {-E/c, -p^i}$.\footnote{This is consistent with a free Dirac field ...} $\mathcal{H}$ relates to the rest energy, the parametrization to proper time (its conjugate).
+we end up with a minus sign between the four-momentum and the conjugate variables $P^\alpha = [-E/c, -p^i]$.\footnote{This is consistent with the Dirac field where we have $[i\hbar\partial_t, -i\hbar\partial_{x^i}]$ for generators and $\psi^\dagger\gamma^0[i\hbar\partial_t, -i\hbar\partial_{x^i}]\psi$ for observables, and $\gamma^0$ is $1$ for particles and $-1$ for anti-particles.} $\mathcal{H}$ relates to the rest energy, the parametrization to proper time (its conjugate).
 Parametrization and time are aligned for particles and anti-aligned for anti-particles.
@@
 
-The hope is that, by continuing in this approach, we can shed more light on why nature has the laws it has; and show that they are not some arbitrary rules, but indeed necessary given few simple assumptions.
+The hope is that, by continuing in this approach, we can shed more light on why the laws of physics are what they are; and show that they are not arbitrary rules, but necessary given few simple assumptions.
 

### 2014-06-27 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Including Christine's comment and some bibliography

@@
 \newtheorem{defnlem}[thm]{Definition-Lemma}
-\newtheorem{assump}[thm]{Assumption}
+\newtheorem*{assump1}{Classical assumption}
+\newtheorem*{assump2}{Determinism and Reversibility assumption}
 \newtheorem{warn}[thm]{Warning}
@@
 
-\title{Deriving Hamiltonian mechanics from physics, without passing through Lagrangian.}
+\title{From physical principles to Hamiltonian mechanics}
 \author{Gabriele Carcassi}
@@
 
-We'll give simple definitions of states in terms of labels and determinism/reversibility in terms of bijective maps; define a metric based on the idea that bijective maps preserve the cardinality of labels; obtain Hamilton's equations from the invariance of that metric.
+We'll give simple definitions of states in terms of labels and of determinism/reversibility in terms of bijective maps; define a metric based on the idea that bijective maps preserve the cardinality of labels; obtain Hamilton's equations from the invariance of that metric.
 
 No mathematical breakthrough should be expected: the goal, after all, is to derive the \emph{known} framework from a set of \emph{simple} definitions in the most \emph{obvious} way possible.
-All proofs are no longer than a couple of paragraphs, so the word theorem is avoided in favor of proposition and corollary.
+No proof is longer than a couple of paragraphs, so the word \emph{theorem} is avoided in favor of \emph{proposition} and \emph{corollary}.
 
@@
 Let $I_1$ and $I_2$ be two sets of labels.
-We call $I = \langle I_1, I_2 \rangle$ a \emph{combined set of labels} if an injective map exists such that $\forall i \in I \rightarrowtail i_1 \in I_1, i_2 \in I_2$.
+We call $I = \langle I_1, I_2 \rangle$ a \emph{combined set of labels} if an injective map $f:I \rightarrowtail I_1, I_2$ exists that maps an element of $I$ to a pair of elements of $I_1$ and $I_2$.
 \end{defn}
@@
 
-\begin{assump}\label{classical}
-(\emph{Classical assumption}) The system is infinitesimally reducible: it can be thought as composed by two or more similar but smaller systems, each in its own configuration state, which also can be thought as composed by two or more, ad infinitum.
-\end{assump}
+\begin{assump1}\label{classical}
+The system is infinitely reducible: it can be thought as composed by two or more similar but smaller systems, each in its own configuration state, which also can be thought as composed by two or more, ad infinitum.
+\end{assump1}
 
@@
 
-Under the classical assumption, we can then limit ourselves to study the labels of states and their properties without losing generality.
+Under the classical assumption, we can then limit ourselves to study the labels of states and their properties without loss of generality.
 
 \textbf{State mapping}. We will now concentrate on deterministic and reversible evolution.
-\begin{assump}\label{detAss}
-(\emph{Determinism and Reversibility}) The system undergoes deterministic (future state identified by the present state) and reversible (past state identified by the present state) evolution.
-\end{assump}
+
+\begin{assump2}
+The system undergoes deterministic (future state identified by the present state) and reversible (past state identified by the present state) evolution.
+\end{assump2}
 
@@
 
-Mathematically, studying \ref{detAss} means studying bijective maps. The evolution of a distribution simply moves the elements around: the bars of the histogram move place, but keep the same height.
+Mathematically, assuming determinism and reversibility means studying bijective maps.
+The evolution of a distribution simply moves the elements around: the bars of the histogram move place, but keep the same height.
 
@@
 Let $f:I \leftrightarrow I$ be a bijective map.
-The labels are said \emph{independent} iff we can define two sets of orthogonal labels $I_1'$ and $I_2'$ such that we can define separate maps:
+We call the labels \emph{independent} iff we can define two sets of orthogonal labels $I_1'$ and $I_2'$ such that we can define separate maps:
 \begin{enumerate}
@@
 
-\textbf{Discrete labels over $\mathbb{R}$}. Labels often will be numbers. For integers, we simply use $\mathbb{Z}$ as our set of labels. For real number, we need to update our definitions.
+\textbf{Discrete labels over $\mathbb{R}$}. Labels often will be numbers. For integers, we simply use $\mathbb{Z}$ as our set of labels. For real numbers, we need to update our definitions.
 
@@
 Each classical configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i=\sum \rho(i) w(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ is defined as before, and $\rho(i)\equiv D(i) / w(i)$ is the density of the distribution for the cell.
-In physics term, the distribution can be visualized as a histogram where $w(i)$, $\rho(i)$ and $D(i)$ are respectively the width, height and area of the each bin.
+In physics terms, the distribution can be visualized as a histogram where $w(i)$, $\rho(i)$ and $D(i)$ are respectively the width, height and area of each bin.
 \end{prop}
@@
 \begin{cor}\label{discreteEv}
-Let $f: I \leftrightarrow I$ be a bijective map, we have $\mathbbm{c}_j=\sum D'(i) \mathbbm{s}_i=\sum \rho'(i) w(i) \mathbbm{s}_i = \sum D(f^{-1}(i)) \mathbbm{s}_i = \sum \rho(f^{-1}(i)) w(f^{-1}(i)) \mathbbm{s}_i$.
+Let $f: I \leftrightarrow I$ be a bijective map.
+We have $\mathbbm{c}_{j'}=\sum D'(i) \mathbbm{s}_i=\sum \rho'(i) w(i) \mathbbm{s}_i = \sum D(f^{-1}(i)) \mathbbm{s}_i = \sum \rho(f^{-1}(i)) w(f^{-1}(i)) \mathbbm{s}_i$.
 $\rho'(i) = \rho(f^{-1}(i)) w(f^{-1}(i)) / w(i)$.
@@
 
-The area moves from one cell of the histogram to the other. The height need to be adjusted if the cell is of a different size.
+The area moves from one cell of the histogram to the other. The height needs to be adjusted if the cell is of a different size.
 
@@
 
-To prepare for the limit we define $m(i)=w(i)/\bar{w}$, where $\bar{w}$ represents the average width of the cells.
+To prepare for the limit we define $m(i)=w(i)/\bar{w}$\footnote{This $m(x)$ solves the same problems addressed by the invariant measure $m(x)$ introduced by Jaynes\cite{Jaynes} in the limiting density of discrete points, which adjusts Shannon's differential information entropy definition to make it invariant.\cite{Shannon}}, where $\bar{w}$ represents the average width of the cells.
 We increase the number of the cells and reduce $\bar{w}$ while keeping $m(i)$ finite.
@@
 \end{prop}
-%http://en.wikipedia.org/wiki/Limiting_density_of_discrete_points
-Assume mapping is discontinuous at point $x$. Consider the cell at $x$ of width $m(x)dx$. The cell would be split into two, so it would not be mapped to one and only one cell.
+Assume mapping is discontinuous at point $x$.
+Consider the cell at $x$ of width $m(x)dx$.
+The cell would be split into two, so it would not be mapped to one and only one cell.
+Therefore the mapping must be continuous.
 
@@
 Then $dx' = \frac{m(x')}{m(x)} dx$.
-If $X$ is homogeneous, then $dx' = dx$, the range gives us a measure of the cardinality of labels and must be conserved.
+If $X$ is homogeneous, then $dx' = dx$; the range gives us a measure of the cardinality of labels and must be conserved.
 \end{prop}
@@
 \end{align*}
-where $n(\Delta x)$ and $n(1)$ are the number of labels/cells in the $\Delta x$ and unit range, while k their width.
-The range can be seen as the ration between the number of labels in the range and the number of labels in the unit range.
+where $n(\Delta x)$ and $n(1)$ are the number of labels/cells in the $\Delta x$ and unit range, while k is their width.
+The range can be seen as the ratio between the number of labels in the range and the number of labels in the unit range.
 That ratio must be conserved by the bijective map and remains well defined during the limit.
@@
 \begin{defn}\label{sdof}
-A \emph{degree of freedom} is a label set $X$ given by a pair of homogeneous and orthogonal state variables $P$ and $Q$.
-We call these \emph{conjugate variables}.
-A state variable is said conjugate of the first iff together they form a pair of conjugate variables.
+A \emph{degree of freedom} is a label set $X$ given by a pair of homogeneous and orthogonal state variables $P$ and $Q$. We call these \emph{conjugate variables}.
 \end{defn}
@@
 Let $v^{\alpha} \omega_{\alpha, \beta} w^{\alpha}$ be an antisymmetric product conserved under a continuous transformation parameterized by $t$.
-We can then define a function $H$ such that given $S^{\alpha} \equiv d_{t}x^{\alpha}$, the vector field that tells us how the state variables change, $S_{\beta} \equiv S^{\alpha} \omega_{\alpha, \beta}$, we have $S_{\alpha} = \partial_{\alpha}H$.
+We can then define a function $H$ such that given $S^{\alpha} \equiv d_{t}x^{\alpha}$ and $S_{\beta} \equiv S^{\alpha} \omega_{\alpha, \beta}$, we have $S_{\alpha} = \partial_{\alpha}H$.
 \end{lem}
 
-Simply applying the vector transformation rules under continuous transformation we have:
+$S^{\alpha}$ is the vector field that represents how the state variables change. Simply applying the vector transformation rules under continuous transformation we have:
 \begin{align*}
@@
 
-Simply expand \ref{genAntisim} with the metric defined in \ref{sdofInvariant}. We recognise Hamilton's equations for one degree of freedom.
+Simply expand \ref{genAntisim} with the metric defined in \ref{sdofInvariant}. We recognize Hamilton's equations for one degree of freedom.
 
@@
 \begin{defn}\label{mdof}
-Two degrees of freedom are said \emph{independent} if the corresponding label set $X^1$ and $X^2$ are independent.
+Two degrees of freedom are said to be \emph{independent} if the corresponding label sets $X^1$ and $X^2$ are independent.
 \end{defn}
@@
 The orthogonality of the labels corresponds to orthogonality in phase space, which the mapping needs to preserve across independent degrees of freedom.
-The cardinality of labels within each degree of freedom corresponds to the area in phase space\footnote{We assume we are using the same unit across d.o.f.}, which the mapping also needs to preserve.
-This is equivalent to require the conservation of the scalar product across independent degrees of freedoms, while still requiring conservation of the vector product within.
+The cardinality of labels within each degree of freedom corresponds to the area in phase space\footnote{We assume we are using the same unit across d.o.f.}, which the mapping also needs to preserve.\footnote{These statements provide a direct physical interpretation for Gromov's non-squeezing theorem.\cite{Gromov}\cite{deGosson}\cite{Stewart}} This is equivalent to requiring the conservation of the scalar product across independent degrees of freedom, while still requiring conservation of the vector product within.
 That leads us to the metric defined by \ref{mdofInvariant}.
@@
 For an infinitesimal region, this corresponds to $dq^1 \wedge dp^1 + dq^2 \wedge dp^2$, the sum of the projections on the independent planes.
+Moreover, because of \ref{labelsMultiCount}, the product of the areas of each d.o.f.
+is also conserved; the volume of phase space corresponds to the cardinality of states, which a bijective map conserves: this is the Liouville's theorem.
 
 \begin{prop}\label{mdofHam}
-The evolution for a multiple degrees of freedom is given by:
+The evolution for multiple degrees of freedom is given by:
 \begin{align*}
@@
 
-Expand \ref{genAntisim} with the metric defined in \ref{mdofInvariant}. We recognise Hamilton equations for multiple degrees of freedom.
+Expand \ref{genAntisim} with the metric defined in \ref{mdofInvariant}. We recognize Hamilton's equations for multiple degrees of freedom.
 
 \section{Time dependence}
-So far we have assumed that both state labeling and mapping do not change in time. In that case, we need to to use time as a label. We introduce an extra degree of freedom.
+So far we have assumed that neither state labeling nor mapping change in time. If they do, we need to to use time as a label and therefore introduce an extra degree of freedom.
 
@@
 The \emph{temporal degree of freedom} is a label set $X$ given by the pair of conjugate variables $T$ and $E$.
-We call extended phase space the outer product between phase space and the temporal degree of freedom.
+We call \emph{extended phase space} the outer product between phase space and the temporal degree of freedom.
 \end{defn}
@@
 \begin{prop}\label{tdofMonotonic}
-Let $s$ be the parameter of a trajectory in the extended phase space of a deterministic and reversible system.
-The trajectory must be continuous.
-There must exist a strictly monotonic function $t(s)$.
-States are connected by a trajectory where either $dt/ds>0$ (particle states) or $dt/ds<0$ (anti-particle states).
+Let $s$ be the parameter of a trajectory in the extended phase space of a deterministic and reversible system. The trajectory must be continuous. There must exist a strictly monotonic function $t(s)$.
 \end{prop}
@@
 Since determinism and reversibility are defined in time, the trajectory must traverse all times once and only once: we must have an invertible mapping between $t$ and $s$, which means we must have a strictly monotonic $t(s)$.
-$dt/ds$ along a trajectory cannot change sign, so we have the division between particle and anti-particle states.
+
+\begin{defn}\label{tdofAntistates}
+We call \emph{standard states} those connected by a trajectory where $dt/ds>0$. We call \emph{anti-states} those connected by a trajectory where $dt/ds<0$.
+\end{defn}
+
+Since $t(s)$ is strictly monotonic, $dt/ds$ along a trajectory cannot change sign, so we have the division between particle and anti-particle states.
 Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call particle and anti-particle states is also conventional.
@@
 \begin{prop}\label{tdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of the extended phase space for one degree of freedom. Let $\alpha$ and $\beta$ be indexes for the state variables $t, e, q, p$. Let
+Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by the temporal degree of freedom and one standard degree of freedom.
+Let $\alpha$ and $\beta$ be indexes for the state variables $t, e, q, p$.
+Let
 \begin{align*}
@@
 \end{align*}
-then $v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}=v^{\alpha} \omega_{\alpha, \beta} w^{\beta}$ under detrev transformation.
+then $v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}=v^{\alpha} \omega_{\alpha, \beta} w^{\beta}$ under deterministic and reversible evolution.
 \end{prop}
@@
 
-As in the case of independent d.o.f., we have right triangle-like relationship between the plane where the invariant is defined and its projections on the planes defined by each d.o.f.:
+We have right triangle-like relationship between the plane where the invariant is defined and its projections on the planes defined by each d.o.f., similar to the multiple d.o.f.:
 \begin{align*}
-&dq^1 \wedge dp^1 + dq^2 \wedge dp^2 = k \\
-&dt \wedge de + k = dq \wedge dp
+m.d.o.f &dq^1 \wedge dp^1 + dq^2 \wedge dp^2 = k \\
+t.d.o.t &dt \wedge de + k = dq \wedge dp
 \end{align*}
-In the previous case, the right angle was between the two independent d.o.f..
+But in the previous case, the right angle was between the two independent d.o.f..
 In this case, the right angle is between the invariant and the plane of constant $\langle Q, P \rangle$ where $dt \wedge de$ is defined.
 We rewrite it as $dq \wedge dp - dt \wedge de = k$.
-Which corresponds to the Minkowski product across d.o.f.
+This corresponds to the Minkowski product across d.o.f.
 and the vector product within.
-The metric, with a space-like convention, still gives us the cardinality of labels within a degree of freedom.\footnote{Adjusted to avoid double counting}
+The metric, with a space-like convention, still gives us the cardinality of labels within a degree of freedom.\footnote{Adjusted to avoid double counting.}
 
@@
 Take the metric from \ref{tdofInvariant}, add multiple independent d.o.f as in \ref{mdofInvariant}, use \ref{genAntisim} with the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$.
-We recognise Hamilton equations in the extended phase space.
+If we set $q^{n+1}=t$ and $p^{n+1}=-e$, we recognise Hamilton equations in the extended phase space.\cite{Synge}\cite{Lanczos}\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.
+In fact, it may not need be constant within and across trajectories.}
 
 It should not be a surprise that the equations do not mention the speed of light $c$.
-In fact, nothing says that all $q^i$ represent space, and with the same unit, and that the laws of motion are invariant in all inertial frames.
+In fact, nothing says that all $q^i$ represent space or that the laws of motion are invariant in all inertial frames.
 The only requirement we have is that the areas of each degrees of freedom represent the same cardinality for labels.
@@
 
-Since $\mathcal{H}$ is constant through the evolution, it can serve both as the generating function and as the evolution constrain.
+Since $\mathcal{H}$ is constant through the evolution, it can serve both as the generating function and as the evolution constraint.
 By convention, we could set $\mathcal{H}=0$ without loss of generality as changing $\mathcal{H}$ by a constant does not change the equation of motion.
@@
 
-\textbf{Example}. We wrap up with an example. Let $\mathcal{H} = mc^2 + ((p^i)^2 - E^2/c^2) / 2m$.  If we apply \ref{tdofHam} we have:
+\textbf{Example}. We wrap up with an example. Let $\mathcal{H} = mc^2 + ((p^i)^2 - E^2/c^2) / 2m$ = 0.  If we apply \ref{tdofHam} we have:
 \begin{align*}
@@
 \end{align*}
-For particles, let $s=\tau$:
+For standard states, let $s=\tau$:
 \begin{align*}
@@
 \end{align*}
-so we recognize the four-momentum $P^\alpha = [E/c, p^i]$ and $\tau$ proper time. For anti-particles, let $s=-\tau$:
+so we recognize the four-momentum $P^\alpha = [E/c, p^i]$ and $\tau$ proper time. For anti-states, let $s=-\tau$:
 \begin{align*}
@@
 we end up with a minus sign between the four-momentum and the conjugate variables $P^\alpha = [-E/c, -p^i]$.\footnote{This is consistent with the Dirac field where we have $[i\hbar\partial_t, -i\hbar\partial_{x^i}]$ for generators and $\psi^\dagger\gamma^0[i\hbar\partial_t, -i\hbar\partial_{x^i}]\psi$ for observables, and $\gamma^0$ is $1$ for particles and $-1$ for anti-particles.} $\mathcal{H}$ relates to the rest energy, the parametrization to proper time (its conjugate).
-Parametrization and time are aligned for particles and anti-aligned for anti-particles.
+Parametrization and time are aligned for standard states and anti-aligned for anti-states.
 
@@
 By deriving Hamiltonian mechanics from simple definitions of labels, states, determinism and reversibility we have given more direct physical meaning to phase space and its geometric properties, and shown that a good part of the Hamiltonian framework can stand on its own, without Lagrangians.
-No mathematical breakthrough is revealed here, yet I am not aware of any work that brings all the pieces of the puzzle in quite this way: so much derived from so little.
+The further emergence of the Minkowsky metric and the distinction between standard and anti-states by the inclusion of the temporal d.o.g.
+is also noteworthy.
+No mathematical breakthrough is revealed here, yet I am not aware of any work that brings all the pieces of the puzzle together in quite this way: so much derived from so little.
 
@@
 
-\bibitem{extend} No bib right now.
+\bibitem{Jaynes} Jaynes, E. T.: "Information theory and statistical mechanics", (1963)
+\bibitem{Shannon} Shannon, C. E.: "A mathematical theory of communications", The Bell System Technical Journal, Vol. 27, pp. 379–423, 623–656, (1948)
+\bibitem{Gromov} Gromov, M. L.: "Pseudo holomorphic curves in symplectic manifolds". Inventiones Mathematicae 82: 307–347, (1985)
+\bibitem{deGosson} de Gosson, M. A.: "The symplectic camel and the uncertainty principle: the tip of an iceberg?", Foundation of Physics 39, pp. 194–214, (2009)
+\bibitem{Stewart} Stewart, I.: "The symplectic camel", Nature 329(6134), 17–18 (1987)
+\bibitem{Struckmeier} Struckmeier, J.: "Hamiltonian dynamics on the symplectic extended phase space for autonomous and non-autonomous systems", J. Phys. A: Math. Gen 38, 1257-1278, (2005)
+\bibitem{Synge} Synge, J. L.: Encyclopedia of Physics Vol 3/1, Springer (1960)
+\bibitem{Lanczos} Lanczos, C.: "The variational principles of mechanics", University of Toronto Press (1949)
 

### 2014-06-28 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Reviewing label names

@@
 
-\begin{defn}\label{combine label}
+\begin{defn}\label{labelsCombine}
 Let $I_1$ and $I_2$ be two sets of labels.
@@
 
-\begin{defn}\label{orth}
+\begin{defn}\label{labelsOrth}
 The two label sets $I_1$ and $I_2$ are said to be \emph{orthogonal} iff $I = \langle I_1, I_2 \rangle = I_1 \times I_2$.
@@
 
-\begin{defn}\label{statedef}
+\begin{defn}\label{classicalPhaseSpace}
 Let $\mathbbm{S}_I$ be the set of all possible configuration states of the infinitesimal subdivision. We call this set \emph{phase space}. We call each element $\mathbbm{s}_i$ a \emph{state}.
@@
 
-\begin{cor}\label{statedistr}
+\begin{cor}\label{classicalDistribution}
 Each classical configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}_i$.
@@
 
-\begin{prop}\label{detMap}
+\begin{prop}\label{detrevMap}
 Let $I$ be a set of labels of a system that undergoes deterministic and reversible evolution. There exists a bijective map $f:I \leftrightarrow I$.
@@
 
-\begin{cor}\label{discreteEv}
+\begin{cor}\label{detrevDist}
 The evolution of a classical configuration state $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$ under a bijective map is given by $\mathbbm{c}_{j'}=\sum D'(i) \mathbbm{s}_{i}=\sum D(f^{-1}(i)) \mathbbm{s}_{i}$.
@@
 
-\begin{defn}\label{disclabelsoverr}
+\begin{defn}\label{discreteLabelDef}
 Consider a continuous numeric range.
@@
 
-\begin{prop}\label{disclabelsoverrdist}
+\begin{prop}\label{discreteLabelDist}
 Each classical configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i=\sum \rho(i) w(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ is defined as before, and $\rho(i)\equiv D(i) / w(i)$ is the density of the distribution for the cell.
@@
 
-\begin{cor}\label{discreteEv}
+\begin{cor}\label{discreteLabelEv}
 Let $f: I \leftrightarrow I$ be a bijective map.
@@
 
-\begin{defn}\label{discreteHomogeneous}
+\begin{defn}\label{discreteLabelHomogeneous}
 A set $I$ of discrete labels over a continuous range is said to be \emph{homogeneous} if $w(i)=k$: the bins are of equal width.
@@
 So they are not necessarily orthogonal in the extended phase space.
-Looking back at \ref{disclabelsoverr}, cells need to be defined on the plane where $\langle Q, P \rangle$ (maximally) change: this is not the plane of constant $\langle E, T \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle Q, P \rangle$ where $dt \wedge de$ is defined.
+Looking back at \ref{discreteLabelDef}, cells need to be defined on the plane where $\langle Q, P \rangle$ (maximally) change: this is not the plane of constant $\langle E, T \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle Q, P \rangle$ where $dt \wedge de$ is defined.
 On that plane we can properly count states and define our invariant.
@@
 Take the metric from \ref{tdofInvariant}, add multiple independent d.o.f as in \ref{mdofInvariant}, use \ref{genAntisim} with the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$.
-If we set $q^{n+1}=t$ and $p^{n+1}=-e$, we recognise Hamilton equations in the extended phase space.\cite{Synge}\cite{Lanczos}\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.
-In fact, it may not need be constant within and across trajectories.}
+If we set $q^{n+1}=t$ and $p^{n+1}=-e$, we recognise Hamilton equations in the extended phase space.\cite{Synge}\cite{Lanczos}\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}
 

### 2014-06-28 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Fixing e capitalization and spelling of Louisville

@@
 Moreover, because of \ref{labelsMultiCount}, the product of the areas of each d.o.f.
-is also conserved; the volume of phase space corresponds to the cardinality of states, which a bijective map conserves: this is the Liouville's theorem.
+is also conserved; the volume of phase space corresponds to the cardinality of states, which a bijective map conserves: this is the Louisville's theorem.
 
@@
 Take the metric from \ref{tdofInvariant}, add multiple independent d.o.f as in \ref{mdofInvariant}, use \ref{genAntisim} with the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$.
-If we set $q^{n+1}=t$ and $p^{n+1}=-e$, we recognise Hamilton equations in the extended phase space.\cite{Synge}\cite{Lanczos}\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}
+If we set\footnote{We avoided using $p^{n+1}$ as it hides the minus sign from the metric, making it seem that the temporal d.o.f is just another independent d.o.f.} $q^{n+1}=t$ and $p^{n+1}=-e$, we recognise Hamilton equations in the extended phase space.\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}\cite{Synge}\cite{Lanczos}
 
@@
 
-\textbf{Example}. We wrap up with an example. Let $\mathcal{H} = mc^2 + ((p^i)^2 - E^2/c^2) / 2m$ = 0.  If we apply \ref{tdofHam} we have:
+\textbf{Example}. We wrap up with an example. Let $\mathcal{H} = mc^2 + ((p^i)^2 - e^2/c^2) / 2m$ = 0.  If we apply \ref{tdofHam} we have:
 \begin{align*}
-d_{s}t &= E / mc^2 \\
-d_{s}E &= 0 \\
+d_{s}t &= e / mc^2 \\
+d_{s}e &= 0 \\
 d_{s}q^i &= p^i / m \\
@@
 \end{align*}
-For standard states, let $s=\tau$:
+For standard states (positive $e$), let $s=\tau$:
 \begin{align*}
-E / c &= m c d_{\tau}t = m U^0 = P^0 \\
+e / c &= m c d_{\tau}t = m U^0 = P^0 \\
 p^i &= m d_{\tau}q^i = m U^i = P^i \\
 \end{align*}
-so we recognize the four-momentum $P^\alpha = [E/c, p^i]$ and $\tau$ proper time. For anti-states, let $s=-\tau$:
+so we recognize the four-momentum $P^\alpha = [e/c, p^i]$ and $\tau$ proper time. For anti-states (negative $e$), let $s=-\tau$:
 \begin{align*}
-- E / c &= m c d_{\tau}t = m U^0 = P^0 \\
+- e / c &= m c d_{\tau}t = m U^0 = P^0 \\
 - p^i &= m d_{\tau}q^i = m U^i = P^i \\
 \end{align*}
-we end up with a minus sign between the four-momentum and the conjugate variables $P^\alpha = [-E/c, -p^i]$.\footnote{This is consistent with the Dirac field where we have $[i\hbar\partial_t, -i\hbar\partial_{x^i}]$ for generators and $\psi^\dagger\gamma^0[i\hbar\partial_t, -i\hbar\partial_{x^i}]\psi$ for observables, and $\gamma^0$ is $1$ for particles and $-1$ for anti-particles.} $\mathcal{H}$ relates to the rest energy, the parametrization to proper time (its conjugate).
+we end up with a minus sign between the four-momentum and the conjugate variables $P^\alpha = [-e/c, -p^i]$.\footnote{This is consistent with the Dirac field where we have $[i\hbar\partial_t, -i\hbar\partial_{x^i}]$ for generators and $\psi^\dagger\gamma^0[i\hbar\partial_t, -i\hbar\partial_{x^i}]\psi$ for observables, and $\gamma^0$ is $1$ for particles and $-1$ for anti-particles.} $\mathcal{H}$ relates to the rest energy, the parametrization to proper time (its conjugate).
 Parametrization and time are aligned for standard states and anti-aligned for anti-states.

### 2014-06-30 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Fixing cite format and aligning equations.

@@
 
-To prepare for the limit we define $m(i)=w(i)/\bar{w}$\footnote{This $m(x)$ solves the same problems addressed by the invariant measure $m(x)$ introduced by Jaynes\cite{Jaynes} in the limiting density of discrete points, which adjusts Shannon's differential information entropy definition to make it invariant.\cite{Shannon}}, where $\bar{w}$ represents the average width of the cells.
+To prepare for the limit we define $m(i)=w(i)/\bar{w}$\footnote{This $m(x)$ solves the same problems addressed by the invariant measure $m(x)$ introduced by Jaynes\cite{Jaynes}, which modifies Shannon's differential information entropy definition to be invariant under coordinate transformation\cite{Shannon}.}, where $\bar{w}$ represents the average width of the cells.
 We increase the number of the cells and reduce $\bar{w}$ while keeping $m(i)$ finite.
@@
 The orthogonality of the labels corresponds to orthogonality in phase space, which the mapping needs to preserve across independent degrees of freedom.
-The cardinality of labels within each degree of freedom corresponds to the area in phase space\footnote{We assume we are using the same unit across d.o.f.}, which the mapping also needs to preserve.\footnote{These statements provide a direct physical interpretation for Gromov's non-squeezing theorem.\cite{Gromov}\cite{deGosson}\cite{Stewart}} This is equivalent to requiring the conservation of the scalar product across independent degrees of freedom, while still requiring conservation of the vector product within.
+The cardinality of labels within each degree of freedom corresponds to the area in phase space\footnote{We assume we are using the same unit across d.o.f.}, which the mapping also needs to preserve.\footnote{These statements provide a direct physical interpretation for Gromov's non-squeezing theorem\cite{Gromov,deGosson,Stewart}.} This is equivalent to requiring the conservation of the scalar product across independent degrees of freedom, while still requiring conservation of the vector product within.
 That leads us to the metric defined by \ref{mdofInvariant}.
@@
 \begin{align*}
-m.d.o.f &dq^1 \wedge dp^1 + dq^2 \wedge dp^2 = k \\
-t.d.o.t &dt \wedge de + k = dq \wedge dp
+m.d.o.f \;\;\; &dq^1 \wedge dp^1 + dq^2 \wedge dp^2 = k \\
+t.d.o.f \;\;\; &dt \wedge de + k = dq \wedge dp \\
 \end{align*}
@@
 Take the metric from \ref{tdofInvariant}, add multiple independent d.o.f as in \ref{mdofInvariant}, use \ref{genAntisim} with the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$.
-If we set\footnote{We avoided using $p^{n+1}$ as it hides the minus sign from the metric, making it seem that the temporal d.o.f is just another independent d.o.f.} $q^{n+1}=t$ and $p^{n+1}=-e$, we recognise Hamilton equations in the extended phase space.\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}\cite{Synge}\cite{Lanczos}
+If we set\footnote{We avoided using $p^{n+1}$ as it hides the minus sign from the metric, making it seem that the temporal d.o.f is just another independent d.o.f.} $q^{n+1}=t$ and $p^{n+1}=-e$, we recognise Hamilton equations in the extended phase space\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}\cite{Synge,Lanczos}.
 

### 2014-06-30 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Fixing quotes in bibliography

@@
 
-\bibitem{Jaynes} Jaynes, E. T.: "Information theory and statistical mechanics", (1963)
-\bibitem{Shannon} Shannon, C. E.: "A mathematical theory of communications", The Bell System Technical Journal, Vol. 27, pp. 379–423, 623–656, (1948)
-\bibitem{Gromov} Gromov, M. L.: "Pseudo holomorphic curves in symplectic manifolds". Inventiones Mathematicae 82: 307–347, (1985)
-\bibitem{deGosson} de Gosson, M. A.: "The symplectic camel and the uncertainty principle: the tip of an iceberg?", Foundation of Physics 39, pp. 194–214, (2009)
-\bibitem{Stewart} Stewart, I.: "The symplectic camel", Nature 329(6134), 17–18 (1987)
-\bibitem{Struckmeier} Struckmeier, J.: "Hamiltonian dynamics on the symplectic extended phase space for autonomous and non-autonomous systems", J. Phys. A: Math. Gen 38, 1257-1278, (2005)
+\bibitem{Jaynes} Jaynes, E. T.: ``Information theory and statistical mechanics'', (1963)
+\bibitem{Shannon} Shannon, C. E.: ``A mathematical theory of communications'', The Bell System Technical Journal, Vol. 27, pp. 379–423, 623–656, (1948)
+\bibitem{Gromov} Gromov, M. L.: ``Pseudo holomorphic curves in symplectic manifolds''. Inventiones Mathematicae 82: 307–347, (1985)
+\bibitem{deGosson} de Gosson, M. A.: ``The symplectic camel and the uncertainty principle: the tip of an iceberg?'', Foundation of Physics 39, pp. 194–214, (2009)
+\bibitem{Stewart} Stewart, I.: ``The symplectic camel'', Nature 329(6134), 17–18 (1987)
+\bibitem{Lanczos} Lanczos, C.: ``The variational principles of mechanics'', University of Toronto Press (1949)
 \bibitem{Synge} Synge, J. L.: Encyclopedia of Physics Vol 3/1, Springer (1960)
-\bibitem{Lanczos} Lanczos, C.: "The variational principles of mechanics", University of Toronto Press (1949)
+\bibitem{Struckmeier} Struckmeier, J.: ``Hamiltonian dynamics on the symplectic extended phase space for autonomous and non-autonomous systems'', J. Phys. A: Math. Gen 38, 1257-1278, (2005)
 

### 2014-06-30 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Changed position of citation.

@@
 
-To prepare for the limit we define $m(i)=w(i)/\bar{w}$\footnote{This $m(x)$ solves the same problems addressed by the invariant measure $m(x)$ introduced by Jaynes\cite{Jaynes}, which modifies Shannon's differential information entropy definition to be invariant under coordinate transformation\cite{Shannon}.}, where $\bar{w}$ represents the average width of the cells.
+To prepare for the limit we define $m(i)=w(i)/\bar{w}$\footnote{This $m(x)$ solves the same problems addressed by the invariant measure $m(x)$ introduced by Jaynes\cite{Jaynes}, which modifies Shannon's differential information entropy definition\cite{Shannon} to be invariant under coordinate transformation.}, where $\bar{w}$ represents the average width of the cells.
 We increase the number of the cells and reduce $\bar{w}$ while keeping $m(i)$ finite.

### 2014-06-30 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Adding reference for Hamilton's equation.

@@
 
-Simply expand \ref{genAntisim} with the metric defined in \ref{sdofInvariant}. We recognize Hamilton's equations for one degree of freedom.
+Simply expand \ref{genAntisim} with the metric defined in \ref{sdofInvariant}. We recognize Hamilton's equations for one degree of freedom\cite{classical_dynamics}.
 
@@
 
-Expand \ref{genAntisim} with the metric defined in \ref{mdofInvariant}. We recognize Hamilton's equations for multiple degrees of freedom.
+Expand \ref{genAntisim} with the metric defined in \ref{mdofInvariant}. We recognize Hamilton's equations for multiple degrees of freedom\cite{classical_dynamics}.
 
@@
 \bibitem{Shannon} Shannon, C. E.: ``A mathematical theory of communications'', The Bell System Technical Journal, Vol. 27, pp. 379–423, 623–656, (1948)
+\bibitem{classical_dynamics} J. V. Jose', E. J. Saletan: ``Classical Dynamics'', Cambridge University Press, (1998)
 \bibitem{Gromov} Gromov, M. L.: ``Pseudo holomorphic curves in symplectic manifolds''. Inventiones Mathematicae 82: 307–347, (1985)

### 2014-06-30 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Minor improvements after another pass

@@
 
-To prepare for the limit we define $m(i)=w(i)/\bar{w}$\footnote{This $m(x)$ solves the same problems addressed by the invariant measure $m(x)$ introduced by Jaynes\cite{Jaynes}, which modifies Shannon's differential information entropy definition\cite{Shannon} to be invariant under coordinate transformation.}, where $\bar{w}$ represents the average width of the cells.
+To prepare for the limit we define\footnote{This $m(x)$ solves the same problems addressed by the invariant measure $m(x)$ introduced by Jaynes\cite{Jaynes}, which modifies Shannon's differential information entropy definition\cite{Shannon} to be invariant under coordinate transformation.} $m(i)=w(i)/\bar{w}$, where $\bar{w}$ represents the average width of the cells.
 We increase the number of the cells and reduce $\bar{w}$ while keeping $m(i)$ finite.
@@
 
-The trajectory has to be continuous in both spatial and temporal variables because of \ref{continuousMapping}.
+The trajectory has to be continuous in both standard and temporal variables because of \ref{continuousMapping}.
 Since determinism and reversibility are defined in time, the trajectory must traverse all times once and only once: we must have an invertible mapping between $t$ and $s$, which means we must have a strictly monotonic $t(s)$.
@@
 \begin{defn}\label{tdofAntistates}
-We call \emph{standard states} those connected by a trajectory where $dt/ds>0$. We call \emph{anti-states} those connected by a trajectory where $dt/ds<0$.
+We call \emph{standard states} those connected by a trajectory where $d_{s}t>0$. We call \emph{anti-states} those connected by a trajectory where $d_{s}t<0$.
 \end{defn}
 
-Since $t(s)$ is strictly monotonic, $dt/ds$ along a trajectory cannot change sign, so we have the division between particle and anti-particle states.
+Since $t(s)$ is strictly monotonic, $d_{s}t$ along a trajectory cannot change sign, so we have the division between particle and anti-particle states.
 Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call particle and anti-particle states is also conventional.
@@
 
-We have right triangle-like relationship between the plane where the invariant is defined and its projections on the planes defined by each d.o.f., similar to the multiple d.o.f.:
+We have a right triangle-like relationship between the plane where the invariant is defined and its projections on the planes defined by each d.o.f., similar to the multiple d.o.f.:
 \begin{align*}
@@
 Take the metric from \ref{tdofInvariant}, add multiple independent d.o.f as in \ref{mdofInvariant}, use \ref{genAntisim} with the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$.
+
 If we set\footnote{We avoided using $p^{n+1}$ as it hides the minus sign from the metric, making it seem that the temporal d.o.f is just another independent d.o.f.} $q^{n+1}=t$ and $p^{n+1}=-e$, we recognise Hamilton equations in the extended phase space\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}\cite{Synge,Lanczos}.
@@
 By deriving Hamiltonian mechanics from simple definitions of labels, states, determinism and reversibility we have given more direct physical meaning to phase space and its geometric properties, and shown that a good part of the Hamiltonian framework can stand on its own, without Lagrangians.
-The further emergence of the Minkowsky metric and the distinction between standard and anti-states by the inclusion of the temporal d.o.g.
+The further emergence of the Minkowsky metric and the distinction between standard and anti-states by the inclusion of the temporal d.o.f.
 is also noteworthy.

### 2014-07-17 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Fixed warnings
Switching to revtex 4.1 fixes the footnotes on last page!
Removing unneeded theorem definitions
Changing affiliation and e-mail address
Update title and contact info

@@
-%\documentclass[aps,prd,twocolumn,floatfix]{revtex4}   % style for Physical Review B and AJP are similar
-\documentclass[twocolumn,floatfix,nofootinbib]{revtex4}   % style for Physical Review B and AJP are similar
-%\documentclass[12pt,aps,prb,preprint]{revtex4}
+\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
 
@@
 
-\theoremstyle{theorem}
 \newtheorem{thm}{Theorem}[section]
@@
 \newtheorem{lem}[thm]{Lemma}
-\newtheorem{sublem}[thm]{Lemma}
 \newtheorem{prop}[thm]{Proposition}
-\newtheorem{ax}[thm]{Axiom}
-\newtheorem{conj}[thm]{Conjecture}
-\newtheorem{clm}[thm]{Claim}
-\newtheorem{lemdef}[thm]{Lemma-Definition}
 
@@
 \newtheorem{defn}[thm]{Definition}
-\newtheorem{rem}[thm]{Remark}
-\newtheorem{ques}[thm]{Question}
-\newtheorem{cons}[thm]{\rm\bfseries{Construction}}
-\newtheorem{exm}[thm]{Example}
-\newtheorem{conds}[thm]{Condition}
-\newtheorem{proper}[thm]{Property}
-\newtheorem{defnlem}[thm]{Definition-Lemma}
 \newtheorem*{assump1}{Classical assumption}
 \newtheorem*{assump2}{Determinism and Reversibility assumption}
-\newtheorem{warn}[thm]{Warning}
-\newtheorem{situ}[thm]{Situation}
 
@@
 
-\title{From physical principles to Hamiltonian mechanics}
+\title{From physical principles to classical Hamiltonian mechanics}
 \author{Gabriele Carcassi}
-\affiliation{Brookhaven National Laboratory, Upton, NY 11973}
-\email{carcassi@bnl.gov}
-\date{May 23, 2014}
+\affiliation{University of Michigan, Ann Arbor, MI 48109}
+\email{carcassi@umich.edu}
+\date{June 16, 2014}
 
@@
 
-\begin{thebibliography}{5}
+\begin{thebibliography}{0}
 

### 2014-09-13 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
First pass in using set notation for labels

@@
 
-\section{States}
-\textbf{State definition}.
-We assume we have a physical system that can be found in different configurations.
-We call each unique configuration \emph{configuration state} and assign it a \emph{label}.
-For our purposes, it does not matter what the label is (a measurement, a way to prepare the system, future behavior, ...), just that it exists.
+\section{States, Labels and Maps}
 
 \begin{defn}\label{statedef}
-Let $I$ be a set of labels.
-Let $\mathbbm{C}_{I}$ be the set of all possible configuration states of our system identified by the labels.
-$\forall i \in I, \exists \mathbbm{c}_{i} \in \mathbbm{C}_{I}$.
+Fix a physical system for study. We define the set $\mathbbm{C}$ of all physically distinguishable configurations for that system. Each element $\mathbbm{c}$ we call \emph{configuration state}.
 \end{defn}
-
-The following definitions allow us to use multiple labels.
-
-\begin{defn}\label{labelsCombine}
-Let $I_1$ and $I_2$ be two sets of labels.
-We call $I = \langle I_1, I_2 \rangle$ a \emph{combined set of labels} if an injective map $f:I \rightarrowtail I_1, I_2$ exists that maps an element of $I$ to a pair of elements of $I_1$ and $I_2$.
-\end{defn}
-
-\begin{defn}\label{labelsOrth}
-The two label sets $I_1$ and $I_2$ are said to be \emph{orthogonal} iff $I = \langle I_1, I_2 \rangle = I_1 \times I_2$.
-\end{defn}
-
-The following assumptions and definitions allow us to describe a configuration state by its parts.
 
@@
 \begin{defn}\label{classicalPhaseSpace}
-Let $\mathbbm{S}_I$ be the set of all possible configuration states of the infinitesimal subdivision. We call this set \emph{phase space}. We call each element $\mathbbm{s}_i$ a \emph{state}.
+Let $\mathbbm{S}$ be the set of all possible configuration states of the infinitesimal subdivision. We call this set \emph{phase space}. We call each element $\mathbbm{s}$ a \emph{state}.
 \end{defn}
@@
 \begin{cor}\label{classicalDistribution}
-Each classical configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}_i$.
-In physics term, the distribution can be visualized as a histogram over the discrete labels.
+Each classical configuration state $\mathbbm{c} \in \mathbbm{C}$ is a distribution over states: $\mathbbm{c}=\sum\limits_{\mathbbm{s} \in \mathbbm{S}} D(\mathbbm{s}) \mathbbm{s}$, where $D:I\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}_i$.
+The distribution can be visualized as a histogram over the states in phase space.
 \end{cor}
 
-Under the classical assumption, we can then limit ourselves to study the labels of states and their properties without loss of generality.
+Under the classical assumption, we can then limit ourselves to the study of the infinitesimal elements, their states and their properties without loss of generality.
+
+\begin{defn}\label{label}
+To help identify states, we introduce the following: we call a \emph{label} a set of states $i\subset\mathbbm{S}$; a \emph{set of labels} a collection of disjoint labels $I | \forall i_1,i_2\in I, i_1\bigcap i_2 = \emptyset$; a \emph{state variable} a set of labels $X$ that covers all of phase space: $\bigcup\limits_{i \in X}i=\mathbbm{S}$.
+Therefore a state is in one and only one label of a state variable.
+\end{defn}
+
+\begin{defn}\label{labelsCombine}
+Let $I_1$ and $I_2$ be two sets of labels.
+We can define the \emph{combined set}, $\langle I_1, I_2 \rangle$, whose labels consist of all the non-empty intersections of one label of $I_1$ and one of $I_2$.
+If all intersections are non-empty, $I_1$ and $I_2$ are said to be \emph{orthogonal}, and the total number of labels of $\langle I_1, I_2 \rangle$ is given by the product of the number of labels in $I_1$ and $I_2$.
+\end{defn}
 
@@
 \begin{prop}\label{detrevMap}
-Let $I$ be a set of labels of a system that undergoes deterministic and reversible evolution. There exists a bijective map $f:I \leftrightarrow I$.
+Let $S$ be the set of configuration states of a system that undergoes deterministic and reversible evolution. There exists a bijective map $f:S \leftrightarrow S$.
 \end{prop}
@@
 \begin{cor}\label{detrevDist}
-The evolution of a classical configuration state $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$ under a bijective map is given by $\mathbbm{c}_{j'}=\sum D'(i) \mathbbm{s}_{i}=\sum D(f^{-1}(i)) \mathbbm{s}_{i}$.
+The evolution of a classical configuration state $\mathbbm{c}=\sum D(s) \mathbbm{s}$ under a bijective map is given by $\mathbbm{c'}=\sum D'(s) \mathbbm{s}=\sum D(f^{-1}(s)) \mathbbm{s}$.
 \end{cor}
@@
 
+\begin{cor}\label{detrevDist}
+Given a label $i$, the image $f(i)$ is also a label containing the same number of states.
+Given a set of labels $I$, the image $f(I)$ is also a set of labels containing the same number of labels.
+Given a state variable $X$, the image $f(X)$ is also a state variable.
+\end{cor}
+
 \begin{defn}\label{labelsIndep}
-Let $I_1$ and $I_2$ be two orthogonal sets of labels and $I = I_1 \times I_2$.
-Let $f:I \leftrightarrow I$ be a bijective map.
-We call the labels \emph{independent} iff we can define two sets of orthogonal labels $I_1'$ and $I_2'$ such that we can define separate maps:
-\begin{enumerate}
-\item $I_1' \times I_2' = I$
-\item there exists a bijective map $f_1: I_1 \leftrightarrow I_1'$
-\item there exists a bijective map $f_2: I_2 \leftrightarrow I_2'$
-\item $f(i) = \langle f_1(i_1), f_2(i_2)\rangle$
-\end{enumerate}
+Let $I_1$ and $I_2$ be two orthogonal sets of labels.
+They are said to be \emph{independent} if and only if the image of the combined set $f(\langle I_1, I_2 \rangle)$ is equal to the combined set of the images $\langle f(I_1), f(I_2) \rangle$.
+It follows that $f(I_1)$ and $f(I_2)$ are orthogonal.
 \end{defn}

### 2014-09-19 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Reviewed all article with new terminology for label/set/state variable

@@
 Under the classical assumption, we can then limit ourselves to the study of the infinitesimal elements, their states and their properties without loss of generality.
+To help identify states, we introduce the following definitions.
 
 \begin{defn}\label{label}
-To help identify states, we introduce the following: we call a \emph{label} a set of states $i\subset\mathbbm{S}$; a \emph{set of labels} a collection of disjoint labels $I | \forall i_1,i_2\in I, i_1\bigcap i_2 = \emptyset$; a \emph{state variable} a set of labels $X$ that covers all of phase space: $\bigcup\limits_{i \in X}i=\mathbbm{S}$.
-Therefore a state is in one and only one label of a state variable.
+We call a \emph{label} a set of states $i\subset\mathbbm{S}$; a \emph{set of labels} a collection of disjoint labels $I | \forall i_1,i_2\in I, i_1\bigcap i_2 = \emptyset$; a \emph{state variable} a set of labels $X$ that covers all of phase space: $\bigcup\limits_{i \in X}i=\mathbbm{S}$.
+Therefore a state belongs to one and only one label of a state variable.
 \end{defn}
@@
 We can define the \emph{combined set}, $\langle I_1, I_2 \rangle$, whose labels consist of all the non-empty intersections of one label of $I_1$ and one of $I_2$.
-If all intersections are non-empty, $I_1$ and $I_2$ are said to be \emph{orthogonal}, and the total number of labels of $\langle I_1, I_2 \rangle$ is given by the product of the number of labels in $I_1$ and $I_2$.
+If all intersections are non-empty, $I_1$ and $I_2$ are said to be \emph{independent}, and we have $n(\langle I_1, I_2 \rangle)=n(I_1)*n(I_2)$ where $n: I \rightarrow \mathbbm{R}$ gives the number of labels of each set.
 \end{defn}
@@
 
-\begin{cor}\label{detrevDist}
+\begin{cor}\label{labelsCount}
 Given a label $i$, the image $f(i)$ is also a label containing the same number of states.
@@
 Given a state variable $X$, the image $f(X)$ is also a state variable.
+Given two independent sets of labels $I_1$ and $I_2$, the images $f(I_1)$ and $f(I_2)$ are also independent.
 \end{cor}
 
-\begin{defn}\label{labelsIndep}
-Let $I_1$ and $I_2$ be two orthogonal sets of labels.
-They are said to be \emph{independent} if and only if the image of the combined set $f(\langle I_1, I_2 \rangle)$ is equal to the combined set of the images $\langle f(I_1), f(I_2) \rangle$.
-It follows that $f(I_1)$ and $f(I_2)$ are orthogonal.
-\end{defn}
-
-\textbf{State counting}. We now look at how the cardinality of states evolves under a bijective map.
-
-\begin{prop}\label{labelsCount}
-Let $\hat{I} \subset I$ of finite size $n(\hat{I})$.
-Let $f(\hat{I})$ and $f^{-1}(\hat{I})$ be respectively the forward and backward image of a bijective map.
-We then have $n(\hat{I})=n(f(\hat{I}))=n(f^{-1}(\hat{I}))$, the number of states is conserved.
-\end{prop}
-
-\begin{prop}\label{labelsMultiCount}
-Let $\hat{I_1} \subset I_1$ and $\hat{I_2} \subset I_2$ of finite size $n(\hat{I_1})$ and $n(\hat{I_2})$.
-Let $f(\hat{I_1})$, $f(\hat{I_2})$, $f^{-1}(\hat{I_1})$ and $f^{-1}(\hat{I_2})$ be the corresponding forward and backward images.
-We then have $n(\hat{I_1}\times\hat{I_2})=n(\hat{I_1})n(\hat{I_2})=n(f(\hat{I_1})) n(f(\hat{I_2}))=n(f^{-1}(\hat{I_1})) n(f^{-1}(\hat{I_2}))=n(\hat{f(I_1)}\times\hat{f(I_2)})=n(f^{-1}(\hat{I_1})\times f^{-1}(\hat{I_2}))$, the total number of states is conserved and remains the product of the number of labels in each set.
-\end{prop}
-
-Bijective maps preserve the number of states as they provide one-to-one association between future and past labels.
+Bijective maps preserve the number of labels as they provide one-to-one association between future and past.
+And they do so for each independent set of labels.
+These simple results, properly generalized to the continuous case, will give us Hamiltonian flow.
 
@@
 
-\textbf{Discrete labels over $\mathbb{R}$}. Labels often will be numbers. For integers, we simply use $\mathbb{Z}$ as our set of labels. For real numbers, we need to update our definitions.
+\textbf{Discrete labels over $\mathbb{R}$}.
+Labels often will be numbers.
+For integers, there will be a one-to-one map between $z\in\mathbb{Z}$ and $i\in I$.
+Therefore we can use $I$ to both refer to the set of labels and the set of numbers they correspond to.
+For real numbers, each label will correspond to an interval.
 
@@
 
-With homogeneous labels no adjustment is needed, and the range can be used as a measure of the number of labels.
+With homogeneous labels no adjustment is needed, and the range can be used as a measure of the number of labels.\footnote{Note that homogeneous labels are not arbitrary: they depend on the map.
+Any non-linear transformation will transform homogeneous labels in non-homogeneous labels.
+Therefore either we require specific labels, and the map will tell us whether they are homogeneous, or require homogeneous labels, and the map will determine what they are, up to a linear transformation.}
 
@@
 \begin{defn}\label{continuousLabels}
-A \emph{state variable} $X$ is the continuous limit of a set $I$ of discrete labels over a continuous range.
+A \emph{continuous state variable} $X$ is the continuous limit of a set $I$ of discrete labels over all $\mathbb{R}$.
 \end{defn}
@@
 \begin{prop}\label{continuousMapping}
-Let $f: X \leftrightarrow X$ be a bijective map on a state variable. The mapping must be continuous.
+Let $f: X \leftrightarrow X$ be a bijective map on a continuous state variable. The mapping must be continuous.
 \end{prop}
@@
 \begin{defn}\label{sdof}
-A \emph{degree of freedom} is a label set $X$ given by a pair of homogeneous and orthogonal state variables $P$ and $Q$. We call these \emph{conjugate variables}.
+A \emph{degree of freedom} is a label set $X$ given by a pair\footnote{Having exactly two variables for each d.o.f.
+is linked to the laws of physics being second order.
+It's justification will be part of a later work.} of homogeneous and orthogonal state variables $P$ and $Q$.
+We call these \emph{conjugate variables}.
 \end{defn}
@@
 The area in phase space of one degree of freedom can then be used as a measure for the cardinality of the labels.
-\ref{labelsCount} becomes area conservation for conjugate variables.
+\ref{labelsCount} becomes area conservation for conjugate variables.\footnote{P' and Q' remain homogeneous but not, in general, orthogonal.}
 
@@
 
-The orthogonality of the labels corresponds to orthogonality in phase space, which the mapping needs to preserve across independent degrees of freedom.
-The cardinality of labels within each degree of freedom corresponds to the area in phase space\footnote{We assume we are using the same unit across d.o.f.}, which the mapping also needs to preserve.\footnote{These statements provide a direct physical interpretation for Gromov's non-squeezing theorem\cite{Gromov,deGosson,Stewart}.} This is equivalent to requiring the conservation of the scalar product across independent degrees of freedom, while still requiring conservation of the vector product within.
+The independence of the homogeneous state variables corresponds to orthogonality in phase space: from \ref{labelsCombine} the product between the number of labels on each d.o.f.
+(i.e.
+the area), must be equal to the number of combines labels (i.e.
+the hyper-volume), which is true only if the d.o.f are orthogonal in phase space.
+From \ref{labelsCount}, the mapping will preserve the cardinality of labels, the area\footnote{We assume we are using the same unit across d.o.f.} on each d.o.f, and the independence, orthogonality across d.o.f.\footnote{These statements provide a direct physical interpretation for Gromov's non-squeezing theorem\cite{Gromov,deGosson,Stewart}.} This is equivalent to requiring the conservation of the scalar product across independent degrees of freedom, while still requiring conservation of the vector product within.
 That leads us to the metric defined by \ref{mdofInvariant}.
@@
 For an infinitesimal region, this corresponds to $dq^1 \wedge dp^1 + dq^2 \wedge dp^2$, the sum of the projections on the independent planes.
-Moreover, because of \ref{labelsMultiCount}, the product of the areas of each d.o.f.
-is also conserved; the volume of phase space corresponds to the cardinality of states, which a bijective map conserves: this is the Louisville's theorem.
+Volume in phase space corresponds to the cardinality of the combined labels, and is therefore conserved: this is the Louisville's theorem.
 
@@
 
-Since $t(s)$ is strictly monotonic, $d_{s}t$ along a trajectory cannot change sign, so we have the division between particle and anti-particle states.
+Since $t(s)$ is strictly monotonic, $d_{s}t$ along a trajectory cannot change sign, so we have the division between standard and anti states.
 Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call particle and anti-particle states is also conventional.
-What is physical and not conventional, though, is that particle and anti-particle states cannot be connected by deterministic and reversible evolution.
+What is physical and not conventional, though, is that standard and anti states cannot be connected by deterministic and reversible evolution.
 
@@
 
-$\langle T, E \rangle$ are not independent labels from $\langle Q, P \rangle$ as they do not define new states.
+$\langle T, E \rangle$ are not independent from $\langle Q, P \rangle$ as they do not define new states.
 So they are not necessarily orthogonal in the extended phase space.

### 2014-09-22 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Another pass at making the language consistent

@@
 No proof is longer than a couple of paragraphs, so the word \emph{theorem} is avoided in favor of \emph{proposition} and \emph{corollary}.
+What is surprising how so much can be derived from very few definitions.
 
@@
 \begin{defn}\label{statedef}
-Fix a physical system for study. We define the set $\mathbbm{C}$ of all physically distinguishable configurations for that system. Each element $\mathbbm{c}$ we call \emph{configuration state}.
+Suppose we have a physical system to study.
+We define the set $\mathbbm{C}$ of all physically distinguishable configurations for that system.
+Each element $\mathbbm{c}$ we call \emph{configuration state}.
 \end{defn}
@@
 \begin{cor}\label{classicalDistribution}
-Each classical configuration state $\mathbbm{c} \in \mathbbm{C}$ is a distribution over states: $\mathbbm{c}=\sum\limits_{\mathbbm{s} \in \mathbbm{S}} D(\mathbbm{s}) \mathbbm{s}$, where $D:I\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}_i$.
+Each classical configuration state $\mathbbm{c} \in \mathbbm{C}$ is a distribution over states: $\mathbbm{c}=\sum\limits_{\mathbbm{s} \in \mathbbm{S}} D(\mathbbm{s}) \mathbbm{s}$, where $D:\mathbbm{S}\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}$.
 The distribution can be visualized as a histogram over the states in phase space.
@@
 
-\textbf{State mapping}. We will now concentrate on deterministic and reversible evolution.
+We now want to study how states and labels evolve in time, under the following assumption.
 
@@
 \begin{prop}\label{detrevMap}
-Let $S$ be the set of configuration states of a system that undergoes deterministic and reversible evolution. There exists a bijective map $f:S \leftrightarrow S$.
+Let $\mathbbm{S}$ be the phase space of a system that undergoes deterministic and reversible evolution.
+There exists a bijective map $f:\mathbbm{S} \leftrightarrow \mathbbm{S}$ between past and future states.
 \end{prop}
@@
 The evolution of a classical configuration state $\mathbbm{c}=\sum D(s) \mathbbm{s}$ under a bijective map is given by $\mathbbm{c'}=\sum D'(s) \mathbbm{s}=\sum D(f^{-1}(s)) \mathbbm{s}$.
+The evolution of fraction of the system in a label $D(i)=\sum\limits_{\mathbbm{S} \in i} D(\mathbbm{s})$ is given by $D'(i)=D(f^{-1}(i))$.
 \end{cor}
@@
 \textbf{Discrete labels over $\mathbb{R}$}.
-Labels often will be numbers.
+Labels are often numbers.
 For integers, there will be a one-to-one map between $z\in\mathbb{Z}$ and $i\in I$.
 Therefore we can use $I$ to both refer to the set of labels and the set of numbers they correspond to.
-For real numbers, each label will correspond to an interval.
+Moreover, from now on we will assume that each numeric label $i$ is enough to identify one and only one state $\mathbb{s}_i$.
+This way we will be able to define the map on the numeric labels directly.
+
+For real numbers, each label corresponds to an interval, which forces us to update our definitions.
 
@@
 \begin{prop}\label{discreteLabelDist}
-Each classical configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i=\sum \rho(i) w(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ is defined as before, and $\rho(i)\equiv D(i) / w(i)$ is the density of the distribution for the cell.
+The configuration state over a discrete label becomes $\mathbbm{c}=\sum D(i) \mathbbm{s}_i dx$.
+The fraction of the system in a label is given by $D(i)=\rho(i) w(i)$ where $\rho(i)\equiv D(i) / w(i)$ is the density of the distribution for the cell.
 In physics terms, the distribution can be visualized as a histogram where $w(i)$, $\rho(i)$ and $D(i)$ are respectively the width, height and area of each bin.
@@
 \begin{cor}\label{discreteLabelEv}
-Let $f: I \leftrightarrow I$ be a bijective map.
-We have $\mathbbm{c}_{j'}=\sum D'(i) \mathbbm{s}_i=\sum \rho'(i) w(i) \mathbbm{s}_i = \sum D(f^{-1}(i)) \mathbbm{s}_i = \sum \rho(f^{-1}(i)) w(f^{-1}(i)) \mathbbm{s}_i$.
-$\rho'(i) = \rho(f^{-1}(i)) w(f^{-1}(i)) / w(i)$.
+Let $f: I \leftrightarrow I$ be a bijective map.\footnote{} We have $D'(i) = \rho'(i) w(i) = D(f^{-1}(i)) = \rho(f^{-1}(i)) w(f^{-1}(i))$. $\rho'(i) = \rho(f^{-1}(i)) w(f^{-1}(i)) / w(i)$.
 \end{cor}
@@
 
-With homogeneous labels no adjustment is needed, and the range can be used as a measure of the number of labels.\footnote{Note that homogeneous labels are not arbitrary: they depend on the map.
-Any non-linear transformation will transform homogeneous labels in non-homogeneous labels.
-Therefore either we require specific labels, and the map will tell us whether they are homogeneous, or require homogeneous labels, and the map will determine what they are, up to a linear transformation.}
+With homogeneous labels no adjustment is needed, and the range can be used as a measure of the number of labels.\footnote{Since the boundaries between intervals are part of the label, a non-linear transformation will typically transform state variables and its labels from homogeneous to non-homogeneous.
+Therefore, there is a class of state variables (i.e.
+phase space basis) that are "special" to the system, not for their physical significance per se, but for their relationship to the state definition.}
 
@@
 $\rho$ and $m$ will converge to functions defined over $x$.
-The corresponding configuration state will become $\mathbbm{c}_j=\int \rho(x) m(x) \mathbbm{s}_x dx$.
+The corresponding configuration state will become $\mathbbm{c}=\int \rho(x) m(x) \mathbbm{s}_x dx$.
 
@@
 \begin{defn}\label{sdof}
-A \emph{degree of freedom} is a label set $X$ given by a pair\footnote{Having exactly two variables for each d.o.f.
+A \emph{degree of freedom} is a homogeneous two dimensional state variable $X$ given by the combined pair\footnote{Having exactly two variables for each d.o.f.
 is linked to the laws of physics being second order.
-It's justification will be part of a later work.} of homogeneous and orthogonal state variables $P$ and $Q$.
+Its justification outside of the scope, will be part of a later work.} of two orthogonal state variables $Q$ and $P$.
 We call these \emph{conjugate variables}.
@@
 \begin{prop}\label{sdofMap}
-Let $f: X \leftrightarrow X$ be a bijective map on a degree of freedom. Then $dq' \wedge dp' = dq \wedge dp$.
+Let $f: X \leftrightarrow X$ be a bijective map on a degree of freedom. Let $Q'=f(Q)$ $P'=f(P)$. Then $dq' \wedge dp' = dq \wedge dp$.
 \end{prop}
 
-This is the equivalent of \ref{widthMapping} for two state variables.
+This is the equivalent of \ref{widthMapping} for a two dimensional state variables.
 The density $\rho(p,q)$ will be defined on cells of infinitesimal area proportional to $dq \wedge dp$.
@@
 
-\begin{defn}\label{mdof}
-Two degrees of freedom are said to be \emph{independent} if the corresponding label sets $X^1$ and $X^2$ are independent.
-\end{defn}
-
 \begin{prop}\label{mdofInvariant}
@@
 For an infinitesimal region, this corresponds to $dq^1 \wedge dp^1 + dq^2 \wedge dp^2$, the sum of the projections on the independent planes.
-Volume in phase space corresponds to the cardinality of the combined labels, and is therefore conserved: this is the Louisville's theorem.
+Moreover, volume in phase space corresponds to the cardinality of the combined labels (i.e.
+the states), and is therefore conserved: this is the Louisville's theorem.
 
@@
 \begin{defn}\label{tdof}
-The \emph{temporal degree of freedom} is a label set $X$ given by the pair of conjugate variables $T$ and $E$.
+The \emph{temporal degree of freedom} is homogeneous two dimensional state variable $X$ given by the pair of conjugate variables $T$ and $E$.
 We call \emph{extended phase space} the outer product between phase space and the temporal degree of freedom.
@@
 
-Since $t(s)$ is strictly monotonic, $d_{s}t$ along a trajectory cannot change sign, so we have the division between standard and anti states.
+Since $t(s)$ is strictly monotonic, $d_{s}t$ along a trajectory cannot change sign, so we have the division between standard and anti-states.
 Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call particle and anti-particle states is also conventional.
-What is physical and not conventional, though, is that standard and anti states cannot be connected by deterministic and reversible evolution.
+What is physical and not conventional, though, is that standard and anti-states cannot be connected by deterministic and reversible evolution.
 

### 2014-09-23 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Final pass for language change

@@
 We'll give simple definitions of states in terms of labels and of determinism/reversibility in terms of bijective maps; define a metric based on the idea that bijective maps preserve the cardinality of labels; obtain Hamilton's equations from the invariance of that metric.
+In so doing we will gain better insight in various aspects, such as the geometry of phase space, continuity under evolution, conjugate variable and more.
 
@@
 Under the classical assumption, we can then limit ourselves to the study of the infinitesimal elements, their states and their properties without loss of generality.
-To help identify states, we introduce the following definitions.
+To help identify states, we introduce the following concepts.
 
@@
 We can define the \emph{combined set}, $\langle I_1, I_2 \rangle$, whose labels consist of all the non-empty intersections of one label of $I_1$ and one of $I_2$.
-If all intersections are non-empty, $I_1$ and $I_2$ are said to be \emph{independent}, and we have $n(\langle I_1, I_2 \rangle)=n(I_1)*n(I_2)$ where $n: I \rightarrow \mathbbm{R}$ gives the number of labels of each set.
+If all intersections are non-empty, $I_1$ and $I_2$ are said to be \emph{independent}, and we have $n(\langle I_1, I_2 \rangle)=n(I_1)n(I_2)$ where $n: I \rightarrow \mathbbm{R}$ gives the number of labels of each set.
 \end{defn}
@@
 The evolution of a classical configuration state $\mathbbm{c}=\sum D(s) \mathbbm{s}$ under a bijective map is given by $\mathbbm{c'}=\sum D'(s) \mathbbm{s}=\sum D(f^{-1}(s)) \mathbbm{s}$.
-The evolution of fraction of the system in a label $D(i)=\sum\limits_{\mathbbm{S} \in i} D(\mathbbm{s})$ is given by $D'(i)=D(f^{-1}(i))$.
+The evolution of the fraction of the system in a label $D(i)=\sum\limits_{\mathbbm{S} \in i} D(\mathbbm{s})$ is given by $D'(i)=D(f^{-1}(i))$.
 \end{cor}
@@
 Given two independent sets of labels $I_1$ and $I_2$, the images $f(I_1)$ and $f(I_2)$ are also independent.
+Therefore $n(f(\langle I_1, I_2 \rangle))=n(f(I_1))n(f(I_2))=n(I_1)n(I_2)=n(\langle I_1, I_2 \rangle)$
 \end{cor}
@@
 \textbf{Discrete labels over $\mathbb{R}$}.
-Labels are often numbers.
-For integers, there will be a one-to-one map between $z\in\mathbb{Z}$ and $i\in I$.
+Numeric labels are often used in physics.
+For integers, there will be a one-to-one map between $z\in\mathbbm{Z}$ and $i\in I$.
 Therefore we can use $I$ to both refer to the set of labels and the set of numbers they correspond to.
-Moreover, from now on we will assume that each numeric label $i$ is enough to identify one and only one state $\mathbb{s}_i$.
-This way we will be able to define the map on the numeric labels directly.
+Moreover, from now on we will assume that each numeric label $i$ is enough to identify one and only one state $\mathbbm{s}_i$.
+This allows us to define the map on the numeric labels directly.
 
-For real numbers, each label corresponds to an interval, which forces us to update our definitions.
+For real numbers, each label corresponds to an interval according to the following:
 
@@
 \begin{cor}\label{discreteLabelEv}
-Let $f: I \leftrightarrow I$ be a bijective map.\footnote{} We have $D'(i) = \rho'(i) w(i) = D(f^{-1}(i)) = \rho(f^{-1}(i)) w(f^{-1}(i))$. $\rho'(i) = \rho(f^{-1}(i)) w(f^{-1}(i)) / w(i)$.
+Let $f: I \leftrightarrow I$ be a bijective map. We have $D'(i) = \rho'(i) w(i) = D(f^{-1}(i)) = \rho(f^{-1}(i)) w(f^{-1}(i))$. $\rho'(i) = \rho(f^{-1}(i)) w(f^{-1}(i)) / w(i)$.
 \end{cor}
@@
 Therefore, there is a class of state variables (i.e.
-phase space basis) that are "special" to the system, not for their physical significance per se, but for their relationship to the state definition.}
+a phase space basis) that hold "special status", not for what they represent physically per se, but for their relationship to state definition.}
 
@@
 We increase the number of the cells and reduce $\bar{w}$ while keeping $m(i)$ finite.
-In the limit we'll have a cell for each value, so we can use $x(i)$ (or simply $x$) instead of $i$ for the label.
+In the limit we'll have a cell for each value, so we can use $x(i)$ (or simply $x$) instead of $i$ to identify the label.
 $\rho$ and $m$ will converge to functions defined over $x$.
@@
 is linked to the laws of physics being second order.
-Its justification outside of the scope, will be part of a later work.} of two orthogonal state variables $Q$ and $P$.
+Its justification outside of the scope, will be part of a later work.} of orthogonal state variables $Q$ and $P$.
 We call these \emph{conjugate variables}.
@@
 The area in phase space of one degree of freedom can then be used as a measure for the cardinality of the labels.
-\ref{labelsCount} becomes area conservation for conjugate variables.\footnote{P' and Q' remain homogeneous but not, in general, orthogonal.}
+\ref{labelsCount} becomes area conservation\footnote{Under a generic pair of state variables, the conserved quantity related to the cardinality of labels would be $\int\limits_\mathcal{S}\frac{d\hat{q}\wedge d\hat{p}}{m(\hat{q},\hat{p})} $, not the area.
+Conjugate variables, therefore, simplify the equations by eliminating "apparent" dynamic that would just be the result of the difference in density of non-homogeneous labels.} for conjugate variables.\footnote{P' and Q' remain homogeneous but not, in general, orthogonal.}
 
@@
 \section{Time dependence}
-So far we have assumed that neither state labeling nor mapping change in time. If they do, we need to to use time as a label and therefore introduce an extra degree of freedom.
+So far we have assumed that neither state labeling nor mapping change in time. If they do, we also need to to use time as a label and therefore introduce an extra degree of freedom.
 
@@
 Since $t(s)$ is strictly monotonic, $d_{s}t$ along a trajectory cannot change sign, so we have the division between standard and anti-states.
-Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call particle and anti-particle states is also conventional.
+Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call standard and anti-states is also conventional.
 What is physical and not conventional, though, is that standard and anti-states cannot be connected by deterministic and reversible evolution.
@@
 Since $\mathcal{H}$ is constant through the evolution, it can serve both as the generating function and as the evolution constraint.
-By convention, we could set $\mathcal{H}=0$ without loss of generality as changing $\mathcal{H}$ by a constant does not change the equation of motion.
+By convention, we can set $\mathcal{H}=0$ without loss of generality as changing $\mathcal{H}$ by a constant does not change the equation of motion.
 This reduces extended phase space to $\mathbb{R}^{2*N + 1}$, the state variables plus time.

### 2014-09-29 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Another pass with minor corrections.

@@
 We'll give simple definitions of states in terms of labels and of determinism/reversibility in terms of bijective maps; define a metric based on the idea that bijective maps preserve the cardinality of labels; obtain Hamilton's equations from the invariance of that metric.
-In so doing we will gain better insight in various aspects, such as the geometry of phase space, continuity under evolution, conjugate variable and more.
+In so doing we will be able to answer basic questions such as: what does the geometry of phase space represents?
+why evolution is continuous?
+why are conjugate variable special?
 
@@
 \begin{prop}\label{discreteLabelDist}
-The configuration state over a discrete label becomes $\mathbbm{c}=\sum D(i) \mathbbm{s}_i dx$.
+The configuration state over a set of discrete labels becomes $\mathbbm{c}=\sum D(i) \mathbbm{s}_i$.
 The fraction of the system in a label is given by $D(i)=\rho(i) w(i)$ where $\rho(i)\equiv D(i) / w(i)$ is the density of the distribution for the cell.
@@
 With homogeneous labels no adjustment is needed, and the range can be used as a measure of the number of labels.\footnote{Since the boundaries between intervals are part of the label, a non-linear transformation will typically transform state variables and its labels from homogeneous to non-homogeneous.
-Therefore, there is a class of state variables (i.e.
+Therefore, homogeneity identifies a class of state variables (i.e.
 a phase space basis) that hold "special status", not for what they represent physically per se, but for their relationship to state definition.}
@@
 is linked to the laws of physics being second order.
-Its justification outside of the scope, will be part of a later work.} of orthogonal state variables $Q$ and $P$.
+Its justification is outside of the scope, and will be part of a later work.} of orthogonal state variables $Q$ and $P$.
 We call these \emph{conjugate variables}.
@@
 \ref{labelsCount} becomes area conservation\footnote{Under a generic pair of state variables, the conserved quantity related to the cardinality of labels would be $\int\limits_\mathcal{S}\frac{d\hat{q}\wedge d\hat{p}}{m(\hat{q},\hat{p})} $, not the area.
-Conjugate variables, therefore, simplify the equations by eliminating "apparent" dynamic that would just be the result of the difference in density of non-homogeneous labels.} for conjugate variables.\footnote{P' and Q' remain homogeneous but not, in general, orthogonal.}
+Conjugate variables, therefore, simplify the equations by eliminating "apparent" dynamic that would just be the result of the difference in density of non-homogeneous labels.} for conjugate variables.\footnote{P' and Q' remain a homogeneous pair but not, in general, orthogonal.}
 

### 2014-10-01 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Last pass after Christine's comments

@@
 We'll give simple definitions of states in terms of labels and of determinism/reversibility in terms of bijective maps; define a metric based on the idea that bijective maps preserve the cardinality of labels; obtain Hamilton's equations from the invariance of that metric.
-In so doing we will be able to answer basic questions such as: what does the geometry of phase space represents?
-why evolution is continuous?
-why are conjugate variable special?
+In so doing we will be able to answer basic questions such as: what does the geometry of phase space represent?
+why is evolution continuous?
+why are conjugate variables special?
 
@@
 \begin{cor}\label{detrevDist}
-The evolution of a classical configuration state $\mathbbm{c}=\sum D(s) \mathbbm{s}$ under a bijective map is given by $\mathbbm{c'}=\sum D'(s) \mathbbm{s}=\sum D(f^{-1}(s)) \mathbbm{s}$.
+The evolution of a classical configuration state $\mathbbm{c}=\sum D(\mathbbm{s}) \mathbbm{s}$ under a bijective map is given by $\mathbbm{c'}=\sum D'(\mathbbm{s}) \mathbbm{s}=\sum D(f^{-1}(s)) \mathbbm{s}$.
 The evolution of the fraction of the system in a label $D(i)=\sum\limits_{\mathbbm{S} \in i} D(\mathbbm{s})$ is given by $D'(i)=D(f^{-1}(i))$.
@@
 
-With homogeneous labels no adjustment is needed, and the range can be used as a measure of the number of labels.\footnote{Since the boundaries between intervals are part of the label, a non-linear transformation will typically transform state variables and its labels from homogeneous to non-homogeneous.
+With homogeneous labels no adjustment is needed, and the range can be used as a measure of the number of labels.\footnote{Since the boundaries between intervals are part of the label, a non-linear transformation will typically transform state variables and their labels from homogeneous to non-homogeneous.
 Therefore, homogeneity identifies a class of state variables (i.e.
@@
 The width of the transported cell $m(x)dx \rightarrow m(x) dx'$ must be equal to the width of the target cell $m(x')dx$, which gives us the first part.
-If $X$ is homogeneous, $m(x)=m(x')=1$, which gives us the second part.
+If $X$ is homogeneous, $m(x)=m(x')$, which gives us the second part.
 
@@
 \begin{defn}\label{sdof}
-A \emph{degree of freedom} is a homogeneous two dimensional state variable $X$ given by the combined pair\footnote{Having exactly two variables for each d.o.f.
+A \emph{degree of freedom} is a homogeneous two dimensional combined state variable $X=\langle Q, P \rangle$ given by the combined pair\footnote{Having exactly two variables for each d.o.f.
 is linked to the laws of physics being second order.
@@
 \begin{prop}\label{sdofMap}
-Let $f: X \leftrightarrow X$ be a bijective map on a degree of freedom. Let $Q'=f(Q)$ $P'=f(P)$. Then $dq' \wedge dp' = dq \wedge dp$.
+Let $f: X \leftrightarrow X$ be a bijective map on a degree of freedom. Let $Q'=f(Q)$ and $P'=f(P)$. Then $dq' \wedge dp' = dq \wedge dp$.
 \end{prop}
 
-This is the equivalent of \ref{widthMapping} for a two dimensional state variables.
+This is the equivalent of \ref{widthMapping} for a two dimensional state variable.
 The density $\rho(p,q)$ will be defined on cells of infinitesimal area proportional to $dq \wedge dp$.
@@
 \ref{labelsCount} becomes area conservation\footnote{Under a generic pair of state variables, the conserved quantity related to the cardinality of labels would be $\int\limits_\mathcal{S}\frac{d\hat{q}\wedge d\hat{p}}{m(\hat{q},\hat{p})} $, not the area.
-Conjugate variables, therefore, simplify the equations by eliminating "apparent" dynamic that would just be the result of the difference in density of non-homogeneous labels.} for conjugate variables.\footnote{P' and Q' remain a homogeneous pair but not, in general, orthogonal.}
+Conjugate variables, therefore, simplify the equations by eliminating "apparent" dynamics that would just be the result of the difference in density of non-homogeneous labels.} for conjugate variables.\footnote{P' and Q' remain a homogeneous pair but not, in general, orthogonal.}
 
@@
 Moreover, volume in phase space corresponds to the cardinality of the combined labels (i.e.
-the states), and is therefore conserved: this is the Louisville's theorem.
+the states), and is therefore conserved: this is Liouville's theorem.
 
@@
 \begin{defn}\label{tdof}
-The \emph{temporal degree of freedom} is homogeneous two dimensional state variable $X$ given by the pair of conjugate variables $T$ and $E$.
+The \emph{temporal degree of freedom} is a homogeneous two dimensional state variable $X=\langle T, E \rangle$ given by the pair of conjugate variables $T$ and $E$.
 We call \emph{extended phase space} the outer product between phase space and the temporal degree of freedom.
@@
 So they are not necessarily orthogonal in the extended phase space.
-Looking back at \ref{discreteLabelDef}, cells need to be defined on the plane where $\langle Q, P \rangle$ (maximally) change: this is not the plane of constant $\langle E, T \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle Q, P \rangle$ where $dt \wedge de$ is defined.
+Looking back at \ref{discreteLabelDef}, cells need to be defined on the plane where $\langle Q, P \rangle$ (maximally) change: this is not the plane of constant $\langle T, E \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle Q, P \rangle$ where $dt \wedge de$ is defined.
 On that plane we can properly count states and define our invariant.
@@
 In fact, nothing says that all $q^i$ represent space or that the laws of motion are invariant in all inertial frames.
-The only requirement we have is that the areas of each degrees of freedom represent the same cardinality for labels.
+The only requirement we have is that the areas of each degree of freedom represent the same cardinality for labels.
 
@@
 By deriving Hamiltonian mechanics from simple definitions of labels, states, determinism and reversibility we have given more direct physical meaning to phase space and its geometric properties, and shown that a good part of the Hamiltonian framework can stand on its own, without Lagrangians.
-The further emergence of the Minkowsky metric and the distinction between standard and anti-states by the inclusion of the temporal d.o.f.
+The further emergence of the Minkowski metric and the distinction between standard and anti-states by the inclusion of the temporal d.o.f.
 is also noteworthy.

### 2014-10-01 2014ClassicalLagrangianDerivation/classicalHamiltonianDerivation.tex
Starting to work on kinematic (classical Lagrangian)

--- 2014-09-23 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
@@
 
-\title{From physical principles to classical Hamiltonian mechanics}
+\title{From physical principles to classical Lagrangian mechanics}
 \author{Gabriele Carcassi}
@@
 \begin{abstract}
-We derive the Hamiltonian formulation of classical mechanics directly, without reference to Lagrangian mechanics.
-We start from the definition of states in terms of labels used to identify them, and show how, under a deterministic and
-reversible process, the conservation of the cardinality of the labels leads to Hamilton's equations.\end{abstract}
+We derive the Langrangian formalism from the assumption that trajectories in space-time are the output of a deterministic and reversible process. \end{abstract}
 
@@
 \section{Introduction}
-In physics Hamiltonian mechanics is usually presented as a reformulation of Lagrangian mechanics, which is how it was originally developed.
-In this work we aim to derive it on its own, starting from simple physical assumptions, without any reference to Lagrangian mechanics.
 
-We'll give simple definitions of states in terms of labels and of determinism/reversibility in terms of bijective maps; define a metric based on the idea that bijective maps preserve the cardinality of labels; obtain Hamilton's equations from the invariance of that metric.
-In so doing we will gain better insight in various aspects, such as the geometry of phase space, continuity under evolution, conjugate variable and more.
+\section{Hamiltonian review}
 
-No mathematical breakthrough should be expected: the goal, after all, is to derive the \emph{known} framework from a set of \emph{simple} definitions in the most \emph{obvious} way possible.
-No proof is longer than a couple of paragraphs, so the word \emph{theorem} is avoided in favor of \emph{proposition} and \emph{corollary}.
-What is surprising how so much can be derived from very few definitions.
+\section{Kinematics}
+Assume trajectory are deterministic and reversible. Have state.
 
-\section{States, Labels and Maps}
+How much state? Find that is two per degree of freedom. Position and momentum are only state variables.
 
-\begin{defn}\label{statedef}
-Suppose we have a physical system to study.
-We define the set $\mathbbm{C}$ of all physically distinguishable configurations for that system.
-Each element $\mathbbm{c}$ we call \emph{configuration state}.
-\end{defn}
+Inertial frame
 
-\begin{assump1}\label{classical}
-The system is infinitely reducible: it can be thought as composed by two or more similar but smaller systems, each in its own configuration state, which also can be thought as composed by two or more, ad infinitum.
-\end{assump1}
+\section{Connect to Hamiltonian}
+Let q=x
 
-\begin{defn}\label{classicalPhaseSpace}
-Let $\mathbbm{S}$ be the set of all possible configuration states of the infinitesimal subdivision. We call this set \emph{phase space}. We call each element $\mathbbm{s}$ a \emph{state}.
-\end{defn}
+Find that p must be contra-variant. Must be monotonic. And a linear transformation of v. Introduce mass. Introduce gauge.
 
-\begin{cor}\label{classicalDistribution}
-Each classical configuration state $\mathbbm{c} \in \mathbbm{C}$ is a distribution over states: $\mathbbm{c}=\sum\limits_{\mathbbm{s} \in \mathbbm{S}} D(\mathbbm{s}) \mathbbm{s}$, where $D:\mathbbm{S}\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}$.
-The distribution can be visualized as a histogram over the states in phase space.
-\end{cor}
+\section{Lagrangian}
 
-Under the classical assumption, we can then limit ourselves to the study of the infinitesimal elements, their states and their properties without loss of generality.
-To help identify states, we introduce the following concepts.
-
-\begin{defn}\label{label}
-We call a \emph{label} a set of states $i\subset\mathbbm{S}$; a \emph{set of labels} a collection of disjoint labels $I | \forall i_1,i_2\in I, i_1\bigcap i_2 = \emptyset$; a \emph{state variable} a set of labels $X$ that covers all of phase space: $\bigcup\limits_{i \in X}i=\mathbbm{S}$.
-Therefore a state belongs to one and only one label of a state variable.
-\end{defn}
-
-\begin{defn}\label{labelsCombine}
-Let $I_1$ and $I_2$ be two sets of labels.
-We can define the \emph{combined set}, $\langle I_1, I_2 \rangle$, whose labels consist of all the non-empty intersections of one label of $I_1$ and one of $I_2$.
-If all intersections are non-empty, $I_1$ and $I_2$ are said to be \emph{independent}, and we have $n(\langle I_1, I_2 \rangle)=n(I_1)n(I_2)$ where $n: I \rightarrow \mathbbm{R}$ gives the number of labels of each set.
-\end{defn}
-
-We now want to study how states and labels evolve in time, under the following assumption.
-
-\begin{assump2}
-The system undergoes deterministic (future state identified by the present state) and reversible (past state identified by the present state) evolution.
-\end{assump2}
-
-\begin{prop}\label{detrevMap}
-Let $\mathbbm{S}$ be the phase space of a system that undergoes deterministic and reversible evolution.
-There exists a bijective map $f:\mathbbm{S} \leftrightarrow \mathbbm{S}$ between past and future states.
-\end{prop}
-
-\begin{cor}\label{detrevDist}
-The evolution of a classical configuration state $\mathbbm{c}=\sum D(s) \mathbbm{s}$ under a bijective map is given by $\mathbbm{c'}=\sum D'(s) \mathbbm{s}=\sum D(f^{-1}(s)) \mathbbm{s}$.
-The evolution of the fraction of the system in a label $D(i)=\sum\limits_{\mathbbm{S} \in i} D(\mathbbm{s})$ is given by $D'(i)=D(f^{-1}(i))$.
-\end{cor}
-
-Mathematically, assuming determinism and reversibility means studying bijective maps.
-The evolution of a distribution simply moves the elements around: the bars of the histogram move place, but keep the same height.
-
-\begin{cor}\label{labelsCount}
-Given a label $i$, the image $f(i)$ is also a label containing the same number of states.
-Given a set of labels $I$, the image $f(I)$ is also a set of labels containing the same number of labels.
-Given a state variable $X$, the image $f(X)$ is also a state variable.
-Given two independent sets of labels $I_1$ and $I_2$, the images $f(I_1)$ and $f(I_2)$ are also independent.
-Therefore $n(f(\langle I_1, I_2 \rangle))=n(f(I_1))n(f(I_2))=n(I_1)n(I_2)=n(\langle I_1, I_2 \rangle)$
-\end{cor}
-
-Bijective maps preserve the number of labels as they provide one-to-one association between future and past.
-And they do so for each independent set of labels.
-These simple results, properly generalized to the continuous case, will give us Hamiltonian flow.
-
-\section{Numeric labels}
-
-\textbf{Discrete labels over $\mathbb{R}$}.
-Numeric labels are often used in physics.
-For integers, there will be a one-to-one map between $z\in\mathbbm{Z}$ and $i\in I$.
-Therefore we can use $I$ to both refer to the set of labels and the set of numbers they correspond to.
-Moreover, from now on we will assume that each numeric label $i$ is enough to identify one and only one state $\mathbbm{s}_i$.
-This allows us to define the map on the numeric labels directly.
-
-For real numbers, each label corresponds to an interval according to the following:
-
-\begin{defn}\label{discreteLabelDef}
-Consider a continuous numeric range.
-We divide the full range into contiguous cells.
-Let $I$ be the set of cells.
-For each cell we have a center value $x: I \mapsto \mathbb{R}$ and a width $w: I \mapsto \mathbb{R}$.
-$I$ is a set of \emph{discrete labels over a continuous range}.
-\end{defn}
-
-\begin{prop}\label{discreteLabelDist}
-The configuration state over a discrete label becomes $\mathbbm{c}=\sum D(i) \mathbbm{s}_i dx$.
-The fraction of the system in a label is given by $D(i)=\rho(i) w(i)$ where $\rho(i)\equiv D(i) / w(i)$ is the density of the distribution for the cell.
-In physics terms, the distribution can be visualized as a histogram where $w(i)$, $\rho(i)$ and $D(i)$ are respectively the width, height and area of each bin.
-\end{prop}
-
-\begin{cor}\label{discreteLabelEv}
-Let $f: I \leftrightarrow I$ be a bijective map. We have $D'(i) = \rho'(i) w(i) = D(f^{-1}(i)) = \rho(f^{-1}(i)) w(f^{-1}(i))$. $\rho'(i) = \rho(f^{-1}(i)) w(f^{-1}(i)) / w(i)$.
-\end{cor}
-
-The area moves from one cell of the histogram to the other. The height needs to be adjusted if the cell is of a different size.
-
-\begin{defn}\label{discreteLabelHomogeneous}
-A set $I$ of discrete labels over a continuous range is said to be \emph{homogeneous} if $w(i)=k$: the bins are of equal width.
-\end{defn}
-
-With homogeneous labels no adjustment is needed, and the range can be used as a measure of the number of labels.\footnote{Since the boundaries between intervals are part of the label, a non-linear transformation will typically transform state variables and its labels from homogeneous to non-homogeneous.
-Therefore, there is a class of state variables (i.e.
-a phase space basis) that hold "special status", not for what they represent physically per se, but for their relationship to state definition.}
-
-\textbf{Continuous labels over $\mathbb{R}$}. We now make the bin width arbitrarily small.
-
-\begin{defn}\label{continuousLabels}
-A \emph{continuous state variable} $X$ is the continuous limit of a set $I$ of discrete labels over all $\mathbb{R}$.
-\end{defn}
-
-To prepare for the limit we define\footnote{This $m(x)$ solves the same problems addressed by the invariant measure $m(x)$ introduced by Jaynes\cite{Jaynes}, which modifies Shannon's differential information entropy definition\cite{Shannon} to be invariant under coordinate transformation.} $m(i)=w(i)/\bar{w}$, where $\bar{w}$ represents the average width of the cells.
-We increase the number of the cells and reduce $\bar{w}$ while keeping $m(i)$ finite.
-In the limit we'll have a cell for each value, so we can use $x(i)$ (or simply $x$) instead of $i$ to identify the label.
-$\rho$ and $m$ will converge to functions defined over $x$.
-The corresponding configuration state will become $\mathbbm{c}=\int \rho(x) m(x) \mathbbm{s}_x dx$.
-
-\begin{prop}\label{continuousMapping}
-Let $f: X \leftrightarrow X$ be a bijective map on a continuous state variable. The mapping must be continuous.
-\end{prop}
-Assume mapping is discontinuous at point $x$.
-Consider the cell at $x$ of width $m(x)dx$.
-The cell would be split into two, so it would not be mapped to one and only one cell.
-Therefore the mapping must be continuous.
-
-\begin{prop}\label{widthMapping}
-Let $f: X \leftrightarrow X$ be a bijective map on a state variable, and $x'=f(x)$.
-Then $dx' = \frac{m(x')}{m(x)} dx$.
-If $X$ is homogeneous, then $dx' = dx$; the range gives us a measure of the cardinality of labels and must be conserved.
-\end{prop}
-The mapping must be done so that the width of the cells is mapped as well, not just the center value.
-The width of the transported cell $m(x)dx \rightarrow m(x) dx'$ must be equal to the width of the target cell $m(x')dx$, which gives us the first part.
-If $X$ is homogeneous, $m(x)=m(x')=1$, which gives us the second part.
-
-Let $\Delta x$ be a finite range of $X$. Before the limit, we have:
-\begin{align*}
-\Delta x = \frac{\Delta x}{1} = \frac{n(\Delta x) k}{n(1) k} = \frac{n(\Delta x)}{n(1)}
-\end{align*}
-where $n(\Delta x)$ and $n(1)$ are the number of labels/cells in the $\Delta x$ and unit range, while k is their width.
-The range can be seen as the ratio between the number of labels in the range and the number of labels in the unit range.
-That ratio must be conserved by the bijective map and remains well defined during the limit.
-In this sense, the range can be used as a measure of the cardinality of the labels.
-
-\section{Single degree of freedom}
-
-\begin{defn}\label{sdof}
-A \emph{degree of freedom} is a homogeneous two dimensional state variable $X$ given by the combined pair\footnote{Having exactly two variables for each d.o.f.
-is linked to the laws of physics being second order.
-Its justification outside of the scope, will be part of a later work.} of orthogonal state variables $Q$ and $P$.
-We call these \emph{conjugate variables}.
-\end{defn}
-
-\begin{prop}\label{sdofMap}
-Let $f: X \leftrightarrow X$ be a bijective map on a degree of freedom. Let $Q'=f(Q)$ $P'=f(P)$. Then $dq' \wedge dp' = dq \wedge dp$.
-\end{prop}
-
-This is the equivalent of \ref{widthMapping} for a two dimensional state variables.
-The density $\rho(p,q)$ will be defined on cells of infinitesimal area proportional to $dq \wedge dp$.
-When mapping one cell to another, the infinitesimal area will remain the same.
-The area in phase space of one degree of freedom can then be used as a measure for the cardinality of the labels.
-\ref{labelsCount} becomes area conservation\footnote{Under a generic pair of state variables, the conserved quantity related to the cardinality of labels would be $\int\limits_\mathcal{S}\frac{d\hat{q}\wedge d\hat{p}}{m(\hat{q},\hat{p})} $, not the area.
-Conjugate variables, therefore, simplify the equations by eliminating "apparent" dynamic that would just be the result of the difference in density of non-homogeneous labels.} for conjugate variables.\footnote{P' and Q' remain homogeneous but not, in general, orthogonal.}
-
-\begin{prop}\label{sdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by two conjugate variables. Let
-\begin{align*}
-\omega_{\alpha, \beta} = \left[
-  \begin{array}{cc}
-    0 & 1 \\
-    -1 & 0 \\
-  \end{array}
-\right] \\
-\end{align*}
-then $v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}=v^{\alpha} \omega_{\alpha, \beta} w^{\beta}$ under a bijective map.
-\end{prop}
-
-Area conservation is equivalent to requiring the invariance of the vector product, which is what $\omega_{\alpha, \beta}$ represents.
-
-\begin{lem}\label{genAntisim}
-Let $v$ and $w$ be two vectors.
-Let $v^{\alpha} \omega_{\alpha, \beta} w^{\alpha}$ be an antisymmetric product conserved under a continuous transformation parameterized by $t$.
-We can then define a function $H$ such that given $S^{\alpha} \equiv d_{t}x^{\alpha}$ and $S_{\beta} \equiv S^{\alpha} \omega_{\alpha, \beta}$, we have $S_{\alpha} = \partial_{\alpha}H$.
-\end{lem}
-
-$S^{\alpha}$ is the vector field that represents how the state variables change. Simply applying the vector transformation rules under continuous transformation we have:
-\begin{align*}
-v^{\alpha} \omega_{\alpha, \beta} w^{\beta} &= v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}  \\
-&= (v^{\alpha} + \partial_{\gamma} S^{\alpha} dt v^{\gamma}) \omega_{\alpha, \beta} ( w^{\beta} + \partial_{\delta} S^{\beta} w^{\delta} dt) \\
-&= v^{\alpha} \omega_{\alpha, \beta} w^{\beta} + (\partial_{\gamma} S^{\alpha} v^{\gamma} \omega_{\alpha, \beta} w^{\beta} \\
- &+ v^{\alpha} \omega_{\alpha, \beta} \partial_{\delta} S^{\beta} w^{\delta}) dt + O(dt^2)
-\end{align*}
-\begin{align*}
-v^{\gamma} w^{\beta} \partial_{\gamma} S_{\beta} - v^{\alpha} w^{\delta} \partial_{\delta} S_{\alpha} = 0
-\end{align*}
-\begin{align*}
-\partial_{\alpha} S_{\beta} - \partial_{\beta} S_{\alpha} &= curl(S_{\alpha}) = 0 \\
-S_{\alpha} &= \partial_{\alpha}H
-\end{align*}
-
-\begin{prop}\label{sdofHam}
-The evolution for a single degree of freedom is given by:
-\begin{align*}
-d_{t}q &= \partial_{p} H \\
-d_{t}p &= - \partial_{q} H
-\end{align*}
-\end{prop}
-
-Simply expand \ref{genAntisim} with the metric defined in \ref{sdofInvariant}. We recognize Hamilton's equations for one degree of freedom\cite{classical_dynamics}.
-
-\section{Multiple degrees of freedom}
-
-\begin{prop}\label{mdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by two independent degrees of freedom.
-Let $\alpha$ and $\beta$ be indexes for the state variables $q^1, p^1, q^2, p^2$.
-Let
-\begin{align*}
-\omega_{\alpha, \beta} =  \left[
-  \begin{array}{cc}
-    1 & 0 \\
-    0 & 1 \\
-  \end{array}
-\right] \otimes \left[
-  \begin{array}{cc}
-    0 & 1 \\
-    -1 & 0 \\
-  \end{array}
-\right] =
-\left[
-  \begin{array}{cccc}
-    0 & 1 & 0 & 0 \\
-    -1 & 0 & 0 & 0 \\
-    0 & 0 & 0 & 1 \\
-    0 & 0 & -1 & 0 \\
-  \end{array}
-\right] \\
-\end{align*}
-then $v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}=v^{\alpha} \omega_{\alpha, \beta} w^{\beta}$ under a bijective map.
-\end{prop}
-
-The independence of the homogeneous state variables corresponds to orthogonality in phase space: from \ref{labelsCombine} the product between the number of labels on each d.o.f.
-(i.e.
-the area), must be equal to the number of combines labels (i.e.
-the hyper-volume), which is true only if the d.o.f are orthogonal in phase space.
-From \ref{labelsCount}, the mapping will preserve the cardinality of labels, the area\footnote{We assume we are using the same unit across d.o.f.} on each d.o.f, and the independence, orthogonality across d.o.f.\footnote{These statements provide a direct physical interpretation for Gromov's non-squeezing theorem\cite{Gromov,deGosson,Stewart}.} This is equivalent to requiring the conservation of the scalar product across independent degrees of freedom, while still requiring conservation of the vector product within.
-That leads us to the metric defined by \ref{mdofInvariant}.
-The metric generalizes \ref{sdofInvariant} to give us the cardinality of labels defined on the area given by two arbitrary vectors.
-For an infinitesimal region, this corresponds to $dq^1 \wedge dp^1 + dq^2 \wedge dp^2$, the sum of the projections on the independent planes.
-Moreover, volume in phase space corresponds to the cardinality of the combined labels (i.e.
-the states), and is therefore conserved: this is the Louisville's theorem.
-
-\begin{prop}\label{mdofHam}
-The evolution for multiple degrees of freedom is given by:
-\begin{align*}
-d_{t}q^i &= \partial_{p^i} H \\
-d_{t}p^i &= - \partial_{q^i} H
-\end{align*}
-\end{prop}
-
-Expand \ref{genAntisim} with the metric defined in \ref{mdofInvariant}. We recognize Hamilton's equations for multiple degrees of freedom\cite{classical_dynamics}.
-
-\section{Time dependence}
-So far we have assumed that neither state labeling nor mapping change in time. If they do, we also need to to use time as a label and therefore introduce an extra degree of freedom.
-
-\begin{defn}\label{tdof}
-The \emph{temporal degree of freedom} is homogeneous two dimensional state variable $X$ given by the pair of conjugate variables $T$ and $E$.
-We call \emph{extended phase space} the outer product between phase space and the temporal degree of freedom.
-\end{defn}
-
-\begin{prop}\label{tdofMonotonic}
-Let $s$ be the parameter of a trajectory in the extended phase space of a deterministic and reversible system. The trajectory must be continuous. There must exist a strictly monotonic function $t(s)$.
-\end{prop}
-
-The trajectory has to be continuous in both standard and temporal variables because of \ref{continuousMapping}.
-Since determinism and reversibility are defined in time, the trajectory must traverse all times once and only once: we must have an invertible mapping between $t$ and $s$, which means we must have a strictly monotonic $t(s)$.
-
-\begin{defn}\label{tdofAntistates}
-We call \emph{standard states} those connected by a trajectory where $d_{s}t>0$. We call \emph{anti-states} those connected by a trajectory where $d_{s}t<0$.
-\end{defn}
-
-Since $t(s)$ is strictly monotonic, $d_{s}t$ along a trajectory cannot change sign, so we have the division between standard and anti-states.
-Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call standard and anti-states is also conventional.
-What is physical and not conventional, though, is that standard and anti-states cannot be connected by deterministic and reversible evolution.
-
-\begin{prop}\label{tdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by the temporal degree of freedom and one standard degree of freedom.
-Let $\alpha$ and $\beta$ be indexes for the state variables $t, e, q, p$.
-Let
-\begin{align*}
-\omega_{\alpha, \beta} =  \left[
-  \begin{array}{cc}
-    -1 & 0 \\
-    0 & 1 \\
-  \end{array}
-\right] \otimes \left[
-  \begin{array}{cc}
-    0 & 1 \\
-    -1 & 0 \\
-  \end{array}
-\right]
-= \left[
-  \begin{array}{cccc}
-    0 & -1 & 0 & 0 \\
-    1 & 0 & 0 & 0 \\
-    0 & 0 & 0 & 1 \\
-    0 & 0 & -1 & 0 \\
-  \end{array}
-\right] \\
-\end{align*}
-then $v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}=v^{\alpha} \omega_{\alpha, \beta} w^{\beta}$ under deterministic and reversible evolution.
-\end{prop}
-
-$\langle T, E \rangle$ are not independent from $\langle Q, P \rangle$ as they do not define new states.
-So they are not necessarily orthogonal in the extended phase space.
-Looking back at \ref{discreteLabelDef}, cells need to be defined on the plane where $\langle Q, P \rangle$ (maximally) change: this is not the plane of constant $\langle E, T \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle Q, P \rangle$ where $dt \wedge de$ is defined.
-On that plane we can properly count states and define our invariant.
-
-We have a right triangle-like relationship between the plane where the invariant is defined and its projections on the planes defined by each d.o.f., similar to the multiple d.o.f.:
-\begin{align*}
-m.d.o.f \;\;\; &dq^1 \wedge dp^1 + dq^2 \wedge dp^2 = k \\
-t.d.o.f \;\;\; &dt \wedge de + k = dq \wedge dp \\
-\end{align*}
-But in the previous case, the right angle was between the two independent d.o.f..
-In this case, the right angle is between the invariant and the plane of constant $\langle Q, P \rangle$ where $dt \wedge de$ is defined.
-We rewrite it as $dq \wedge dp - dt \wedge de = k$.
-This corresponds to the Minkowski product across d.o.f.
-and the vector product within.
-The metric, with a space-like convention, still gives us the cardinality of labels within a degree of freedom.\footnote{Adjusted to avoid double counting.}
-
-\begin{prop}\label{tdofHam}
-The evolution for time varying multiple degrees of freedom is given by:
-\begin{align*}
-d_{s}t &= - \partial_{e} \mathcal{H} \\
-d_{s}e &= \partial_{t} \mathcal{H} \\
-d_{s}q^i &= \partial_{p^i} \mathcal{H} \\
-d_{s}p^i &= - \partial_{q^i} \mathcal{H}
-\end{align*}
-\end{prop}
-
-Take the metric from \ref{tdofInvariant}, add multiple independent d.o.f as in \ref{mdofInvariant}, use \ref{genAntisim} with the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$.
-
-If we set\footnote{We avoided using $p^{n+1}$ as it hides the minus sign from the metric, making it seem that the temporal d.o.f is just another independent d.o.f.} $q^{n+1}=t$ and $p^{n+1}=-e$, we recognise Hamilton equations in the extended phase space\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}\cite{Synge,Lanczos}.
-
-It should not be a surprise that the equations do not mention the speed of light $c$.
-In fact, nothing says that all $q^i$ represent space or that the laws of motion are invariant in all inertial frames.
-The only requirement we have is that the areas of each degrees of freedom represent the same cardinality for labels.
-
-\begin{prop}\label{tdofConstrain}
-The evolution is constrained by $\mathcal{H}=k$.
-\end{prop}
-
-Since $\mathcal{H}$ is constant through the evolution, it can serve both as the generating function and as the evolution constraint.
-By convention, we can set $\mathcal{H}=0$ without loss of generality as changing $\mathcal{H}$ by a constant does not change the equation of motion.
-This reduces extended phase space to $\mathbb{R}^{2*N + 1}$, the state variables plus time.
-
-\textbf{Example}. We wrap up with an example. Let $\mathcal{H} = mc^2 + ((p^i)^2 - e^2/c^2) / 2m$ = 0.  If we apply \ref{tdofHam} we have:
-\begin{align*}
-d_{s}t &= e / mc^2 \\
-d_{s}e &= 0 \\
-d_{s}q^i &= p^i / m \\
-d_{s}p^i &= 0
-\end{align*}
-For standard states (positive $e$), let $s=\tau$:
-\begin{align*}
-e / c &= m c d_{\tau}t = m U^0 = P^0 \\
-p^i &= m d_{\tau}q^i = m U^i = P^i \\
-\end{align*}
-so we recognize the four-momentum $P^\alpha = [e/c, p^i]$ and $\tau$ proper time. For anti-states (negative $e$), let $s=-\tau$:
-\begin{align*}
-- e / c &= m c d_{\tau}t = m U^0 = P^0 \\
-- p^i &= m d_{\tau}q^i = m U^i = P^i \\
-\end{align*}
-we end up with a minus sign between the four-momentum and the conjugate variables $P^\alpha = [-e/c, -p^i]$.\footnote{This is consistent with the Dirac field where we have $[i\hbar\partial_t, -i\hbar\partial_{x^i}]$ for generators and $\psi^\dagger\gamma^0[i\hbar\partial_t, -i\hbar\partial_{x^i}]\psi$ for observables, and $\gamma^0$ is $1$ for particles and $-1$ for anti-particles.} $\mathcal{H}$ relates to the rest energy, the parametrization to proper time (its conjugate).
-Parametrization and time are aligned for standard states and anti-aligned for anti-states.
+From Hamiltonian conservation to Lagrangian. p monotonic means convex Hamiltonian: can use Legendre transform.
 
 \section{Conclusion}
-By deriving Hamiltonian mechanics from simple definitions of labels, states, determinism and reversibility we have given more direct physical meaning to phase space and its geometric properties, and shown that a good part of the Hamiltonian framework can stand on its own, without Lagrangians.
-The further emergence of the Minkowsky metric and the distinction between standard and anti-states by the inclusion of the temporal d.o.f.
-is also noteworthy.
-No mathematical breakthrough is revealed here, yet I am not aware of any work that brings all the pieces of the puzzle together in quite this way: so much derived from so little.
-
-The hope is that, by continuing in this approach, we can shed more light on why the laws of physics are what they are; and show that they are not arbitrary rules, but necessary given few simple assumptions.
 

### 2014-10-17 2014ClassicalLagrangianDerivation/classicalHamiltonianDerivation.tex
Initial ideas from kinematic assumption to possible initial conditions

@@
 \newtheorem*{assump2}{Determinism and Reversibility assumption}
+\newtheorem*{assump3}{Kinematic assumption}
 
@@
 \section{Kinematics}
+\begin{assump3}\label{kinematicAssumption}
+The study of the motion of a body is equivalent to study its state under deterministic and reversible evolution.
+\end{assump3}
+
+\begin{cor}\label{}
+Consider all states at $t=t_0$ and all possible trajectories, to each state corresponds one and only one trajectory; to each trajectory corresponds one and only one state.
+\end{cor}
+
+If studying the motion and state evolution are equivalent, given a state we must be able to reconstruct the trajectory, and given the trajectory we must be able to reconstruct the state.
+Given that the choice of state coordinates is arbitrary, we can use position and time as a set of continuous state variables: $q^\alpha = x^\alpha$.
+With that in mind:
+
+\begin{prop}\label{}
+A metric $g_{\alpha\beta}$ must be defined on the manifold identified by $x^\alpha$. All trajectories must be continuous.
+\end{prop}
+
+As each label corresponds to an infinitesimal cell, we need distances properly defined.
+This requires a metric defined on the manifold.
+And since evolution is continuous on state variables, it will be continuous in $x^\alpha$ as well.
+
+\begin{defn}\label{continuousLabels}
+An \emph{inertial frame} is one for which each direction of space represents a homogeneous state variable of an independent degree of freedom.\footnote{While a local inertial frame always exists, the existence of a global one is an added requirement.
+We add it because it makes the discussion easier and the proofs more obvious.
+Most of the results, though, will hold without that assumption.
+This will be important in future works that will try to extend this framework to general relativity.}
+\end{defn}
+
+\begin{cor}\label{continuousLabels}
+The metric associated with an inertial frame is
+\begin{align*}
+g_{\alpha, \beta} =  \left[
+  \begin{array}{cccc}
+    -1 & 0 & 0 & 0 \\
+    0 & 1 & 0 & 0 \\
+    0 & 0 & 1 & 0 \\
+    0 & 0 & 0 & 1 \\
+  \end{array}
+\right] \\
+\end{align*}
+\end{cor}
+
+The relationships between coordinates is the same as in phase space: $X^i$ must be orthogonal with each other, as the corresponding $Q^i$ are orthogonal in phase space; $X^i$ and $X^0$ are not in general orthogonal and it is the length perpendicular to time that gets conserved.
+Given that $X^\alpha$ are homogeneous, the metric must be invariant under translation, does not change in space-time, it is constant.
+With a suitable choice of units, we can set $|g_{\alpha\alpha}|=1$.
+All these combines gives us the Minkowski metric.
+We also introduce a constant $c$\footnote{Here $c$ is not defined as a a speed: it's the constant that allows to convert the cardinality of labels between space and time coordinates.} to convert time intervals to space intervals with the same number of labels, so we have $x^0=ct$.
+
+It's important to note that the geometry of extended phase space and the geometry of space-time are linked: we cannot define cells in one without also defining cells in the other.
+This is true for all reference frames, not just inertial ones.
+
+\begin{prop}\label{continuousLabels}
+The position $x^\alpha$ and the velocity $dx^i/dt$ are necessary and sufficient initial conditions to determine the state of the system and its whole trajectory.
+\end{prop}
+
+The initial conditions will be given by the position and its first $n$ derivatives: $x^i_k=d^kx^i(t_0)/dt^k$ where $k=0..n$.
+This means that there is a function that, given the initial conditions, gives us the trajectory: $x(t)=f(t_0, x^i_k)$.
+We note that the initial conditions can be changed by an active transformation, therefore counting the number of possible initial conditions is equivalent to counting the possible number of active transformations.
+Under such change, though, the metric cannot change: it needs to be a change in initial conditions in the same reference frame, or we would double count each case.
+Defining the transformation for the position also defines how all its derivatives change, and by requiring the invariance of the metric, our options are actually very limited.
+Assuming we start from an inertial frame, only a linear transformation in all $x^\alpha$ can preserve the metric.
+A linear transformation can only change position and velocity, so the initial condition can only be limited to those.
+We also note that the active transformation can always change the velocity of $x(t)$, so $f$ must depend on the velocity or it would not be able to reach all possible trajectories defined by an active transformation in the same reference frame.
+Therefore position and velocity are both necessary and sufficient initial condition.
+
+\begin{prop}
+Principle of relativity. The laws of motion have the same form for all inertial observers.
+\end{prop}
+
+We saw that an active transformation that does not change the metric is a change in the initial conditions.
+Such a change should not affect the form of the function that given the initial conditions gives us the whole trajectory: the observer is the same, the law of motion is the same, they just shift the value of the arguments.
+But if the form of the laws of motion is preserved under such active transformations, it will also not change under passive transformation that preserve the metric.
+Given that all inertial observers can be reached by a linear transformation, which preserves the metric, then the laws of motion will have the same form.
+
+%x(t)=f(x_0, x_1, ... x_n)
+
+%What is the space size for a particular frame?
+
+%We can reach any initial condition by coordinate changes
+%But they also change f
+%Free parameters are going to be the ones we can change without changing f
+%How many coordinate changes are available to use without changing f?
+%The ones that preserve the metric
+
+%l_\alpha^\gamma l_\beta^\delta g_{\gamma\delta} = g_{\alpha\beta}
+
+%Suppose different coordinate frame with same metric
+%x'(t)=f(x'_0, x'_1, ... x'_n)
+
 Assume trajectory are deterministic and reversible. Have state.
@@
 \section{Connect to Hamiltonian}
-Let q=x
+As we have seen, $X^\alpha$ and $U^\alpha$ are suitable state variables to fully identify our degrees of freedom. Nothing tells us, though, that
+they are conjugate variables.
+But, since for every set of initial condition there must be one and only one state associated with it, there must be a bijective transformation such that $q^\alpha=q^\alpha(x^\alpha,u^\alpha)$ and $p^\alpha=p^\alpha(x^\alpha,u^\alpha)$.
+These function must be monotonic in both variable to make sure they are invertible.
+
+Given the degree of arbitrariness in choice, we can set $q^\alpha=x^\alpha$.
+This creates a direct link between how space is measured ($g_{\alpha\beta}(x^\alpha)$) and the width of our cells in phase space ($m(q)$)
 

### 2014-10-29 2014ClassicalLagrangianDerivation/classicalHamiltonianDerivation.tex
Worked out the math from linear relationships between momentums to the end

@@
 \section{Connect to Hamiltonian}
-As we have seen, $X^\alpha$ and $U^\alpha$ are suitable state variables to fully identify our degrees of freedom. Nothing tells us, though, that
-they are conjugate variables.
-But, since for every set of initial condition there must be one and only one state associated with it, there must be a bijective transformation such that $q^\alpha=q^\alpha(x^\alpha,u^\alpha)$ and $p^\alpha=p^\alpha(x^\alpha,u^\alpha)$.
-These function must be monotonic in both variable to make sure they are invertible.
+\begin{prop}\label{continuousLabels}
+Under the kinematic assumption, there must exist a bijective transformation $q^\alpha=q^\alpha(x^\alpha,u^\alpha)$ and $p^\alpha=p^\alpha(x^\alpha,u^\alpha)$ between initial conditions and state variables.
+As such, they are monotonic in both variables.
+\end{prop}
 
-Given the degree of arbitrariness in choice, we can set $q^\alpha=x^\alpha$.
+As we have seen, $X^\alpha$ and $U^\alpha$ are state variables that fully identify our degrees of freedom. Nothing tells us, though, that
+they are conjugate variables. But, since for every set of initial condition there must be one and only one state associated with it. This means the transformation must be invertible, monotonic.
+
+\begin{prop}\label{continuousLabels}
+Under the kinematic assumption, the extended phase space is defined on the conjugate variables $q^\alpha=x^\alpha$ and $p^\alpha=g_{\alpha\beta}(mu^\alpha+p_0^\alpha(x)$.
+\end{prop}
+
+Given the degree of arbitrariness in choice of transformation, we can set $q^\alpha=x^\alpha$.
 This creates a direct link between how space is measured ($g_{\alpha\beta}(x^\alpha)$) and the width of our cells in phase space ($m(q)$)
+
+$p_\alpha=m\frac{dx_\alpha}{ds}+\hat{p}_\alpha(x)$
+
+$\frac{dx^\alpha}{ds}=\frac{\partial H}{\partial p_\alpha}=\frac{1}{m}(p^\alpha-\hat{p}^\alpha(x))$
+
+$H=\frac{1}{2m}(p_\alpha-\hat{p}_\alpha(x))(p^\alpha-\hat{p}^\alpha(x))+\hat{H}(x)$
+
+$\frac{dp_\alpha}{ds}=-\frac{\partial H}{\partial x^\alpha}=$
+
+$\frac{1}{2m}[\frac{\partial\hat{p}_\beta(x)}{\partial x^\alpha} (p^\beta -\hat{p}^\beta(x)) + (p_\beta -\hat{p}_\beta(x))\frac{\partial\hat{p}^\beta(x)}{\partial x^\alpha} ]-\frac{\partial \hat{H}(x)}{\partial x^\alpha}=$
+
+$\frac{1}{2}[\frac{\partial\hat{p}_\beta(x)}{\partial x^\alpha} (\frac{dx^\beta}{ds}) + (\frac{dx_\beta}{ds})\frac{\partial\hat{p}^\beta(x)}{\partial x^\alpha} ]-\frac{\partial \hat{H}(x)}{\partial x^\alpha}=$
+
+$m\frac{d^2x_\alpha}{ds^2}+\frac{d\hat{p}_\alpha(x)}{ds}$
+
+$m\frac{d^2x_\alpha}{ds^2}=\frac{1}{2}[\frac{\partial\hat{p}_\beta(x)}{\partial x^\alpha} (\frac{dx^\beta}{ds}) + (\frac{dx_\beta}{ds})\frac{\partial\hat{p}^\beta(x)}{\partial x^\alpha} ]-\frac{\partial\hat{p}_\alpha(x)}{\partial x^\beta}\frac{dx^\beta}{ds}
+-\frac{\partial \hat{H}(x)}{\partial x^\alpha}$
+
+$m\frac{d^2x_\alpha}{ds^2}=(\frac{\partial\hat{p}_\beta(x)}{\partial x^\alpha} - \frac{\partial\hat{p}_\alpha(x)}{\partial x^\beta} ) \frac{dx\beta}{ds}
+-\frac{\partial \hat{H}(x)}{\partial x^\alpha}$
 

### 2015-01-16 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Lagrangian derivation now becomes complete classical. First pass at the introduction

--- 2014-10-29 2014ClassicalLagrangianDerivation/classicalHamiltonianDerivation.tex
@@
 
-\title{From physical principles to classical Lagrangian mechanics}
+\title{From physical principles to relativistic classical Hamiltonian and Lagrangian particle mechanics}
 \author{Gabriele Carcassi}
@@
 \begin{abstract}
-We derive the Langrangian formalism from the assumption that trajectories in space-time are the output of a deterministic and reversible process. \end{abstract}
-
+We introduce three fundamental principles (infinite reducibility, deterministic and reversible evolution, ) and show that classical mechanics can be derived them.
+The core idea is that deterministic and reversible systems preserve the number of labels we use to identify states, which puts considerable constraints and leads to the Hamiltonian and Lagrangian frameworks, space-time being locally Minkowskian, the general form of the Hamiltonian and equations of motion equivalent to those for relativistic charged particles.
+Many familiar fundamental concepts and constants (such as mass, the speed of light, the Plank constant, the vector potential) are introduced and defined in a novel way, thus providing different insight into them.
+The derivation strives to use definition and mathematical concepts that would allow a future extension to quantum mechanics.\end{abstract}
 \maketitle
@@
 \section{Introduction}
+Classical mechanics is usually founded on Newton's laws.
+These, though, are insufficient to derive the Lagrangian and Hamiltonian formalism, and usually other ad-hoc assumptions (e.g.
+conservative forces) are introduced.
+Special relativity is based on two principles (invariance of the speed of light and the principle of relativity), which lead to the Minkowskian nature of space-time but not to the equation of motion, which are a consistent reformulation of the non-relativistic ones.
+What we'll do in this work is see if we can re-organize all the known elements and equations in a more consistent and comprehensive way.
+The hope is that this gives us better insight on why the fundamental laws are what they are.
+Also, given that classical and quantum mechanics are formally equivalent, there is a chance that such work could be extended to quantum mechanics.
+
+We'll first mathematically define states and the labels (e.g.
+position, momentum) we use to identify them.
+Using the principle of infinite reducibility (or classical assumption), the idea that each state is always divisible into the states of its sub-systems, we show that it is sufficient to describe the evolution of infinitesimal subsystem within their state space (or phase space).
+
+Under the assumption of deterministic and reversible evolution, to each state corresponds one and only one future (or past) state, we show that the cardinality of labels (and states) is conserved.
+This allows us to define a metric on phase space, and the conservation of that metric leads to the Hamiltonian framework.
+That is: we can derive the Hamiltonian framework on its own merit and show that it is equivalent to deterministic and reversible motion.
+
+We then introduce the principle of kinematic equivalence, the idea that studying trajectories is equivalent to studying states, and show that under this assumption we can define a transformation between state and kinematic variables.
+Noting that a passive coordinate change in space-time is equivalent to a re-labelling of states, and as such should conserve the phase space metric (i.e.
+the cardinality of states), the metric of space-time is constrained to be locally Minkowskian.
+Under the kinematic equivalence, the relationship between velocity and conjugate momentum has to be monotonic, leading to a concave Hamiltonian which allows a Legendre transformation leading to the Lagrangian.
+We can constrain the Hamiltonian further, and show that the most general form of motion under the three condition is the one of a geodesics modified by the force given by a vector potential, such as the one of a relativistic charged particle.
+
+The above outline should help the reader maintain the big picture, as each section will go through the many details and definitions.
+While we try to keep a precise language, no mathematical breakthrough should be expected: the goal, after all, is to derive the \emph{known} framework from a set of \emph{simple} definitions in the most \emph{obvious} way possible.
+No proof is longer than a couple of paragraphs, so the word \emph{theorem} is avoided in favor of \emph{proposition} and \emph{corollary}.
+What is surprising how so much can be derived from very few definitions.
 
@@
 \begin{prop}\label{continuousLabels}
-Under the kinematic assumption, the extended phase space is defined on the conjugate variables $q^\alpha=x^\alpha$ and $p^\alpha=g_{\alpha\beta}(mu^\alpha+p_0^\alpha(x)$.
+Under the kinematic assumption, the extended phase space is defined on the conjugate variables $q^\alpha=x^\alpha$ and $p_\alpha=m\frac{dx_\alpha}{ds}+\hat{p}_\alpha(x)$.
 \end{prop}
@@
 
+$\partial_\alpha \delta^\beta_\gamma = 0 = \partial_\alpha g^{\beta\delta} g_{\delta\gamma} + g^{\beta\delta} \partial_\alpha g_{\delta\gamma}$
+
+$\partial_\alpha g^{\beta\delta} g_{\delta\gamma} = - g^{\beta\delta} \partial_\alpha g_{\delta\gamma}$
+
+$\partial_\alpha g^{\beta\delta} g_{\delta\gamma} g^{\gamma\epsilon} = - g^{\beta\delta} \partial_\alpha g_{\delta\gamma} g^{\gamma\epsilon}$
+
+$\partial_\alpha g^{\beta\epsilon} = - g^{\beta\delta} g^{\gamma\epsilon} \partial_\alpha g_{\delta\gamma} $
+
 $p_\alpha=m\frac{dx_\alpha}{ds}+\hat{p}_\alpha(x)$
@@
 
-$H=\frac{1}{2m}(p_\alpha-\hat{p}_\alpha(x))(p^\alpha-\hat{p}^\alpha(x))+\hat{H}(x)$
+$H=\frac{1}{2m}(p_\alpha-\hat{p}_\alpha)g^{\alpha\beta}(p_\beta-\hat{p}_\beta(x))+\hat{H}(x)$
 
-$\frac{dp_\alpha}{ds}=-\frac{\partial H}{\partial x^\alpha}=$
+$\frac{dp_\alpha}{ds}=-\frac{\partial H}{\partial q^\alpha}=$
 
-$\frac{1}{2m}[\frac{\partial\hat{p}_\beta(x)}{\partial x^\alpha} (p^\beta -\hat{p}^\beta(x)) + (p_\beta -\hat{p}_\beta(x))\frac{\partial\hat{p}^\beta(x)}{\partial x^\alpha} ]-\frac{\partial \hat{H}(x)}{\partial x^\alpha}=$
+$\frac{1}{2m}[\partial_\alpha \hat{p}_\beta g^{\beta \gamma} (p_\gamma -\hat{p}_\gamma)
+ - (p_\beta -\hat{p}_\beta) \partial_\alpha g^{\beta \gamma} (p_\gamma -\hat{p}_\gamma)
+ + (p_\beta -\hat{p}_\beta) g^{\beta \gamma} \partial_\alpha \hat{p}_\gamma ]- \partial_\alpha \hat{H}(x)=$
 
-$\frac{1}{2}[\frac{\partial\hat{p}_\beta(x)}{\partial x^\alpha} (\frac{dx^\beta}{ds}) + (\frac{dx_\beta}{ds})\frac{\partial\hat{p}^\beta(x)}{\partial x^\alpha} ]-\frac{\partial \hat{H}(x)}{\partial x^\alpha}=$
+$\frac{1}{2}[\partial_\alpha \hat{p}_\beta u^\beta
+- m u^\delta g_{\delta\beta} \partial_\alpha g^{\beta \gamma} u^\epsilon g_{\epsilon\gamma}
++ u^\gamma \partial_\alpha \hat{p}_\gamma ]- \partial_\alpha \hat{H}(x)=$
+
+$\partial_\alpha \hat{p}_\beta u^\beta + \frac{m}{2}u^\beta \partial_\alpha g_{\beta \gamma} u^\gamma
+- \partial_\alpha \hat{H}(x)=$
 

### 2015-01-16 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Reviewed state definitions

@@
 We'll first mathematically define states and the labels (e.g.
-position, momentum) we use to identify them.
+position, momentum, temperature, pressure) we use to identify them.
 Using the principle of infinite reducibility (or classical assumption), the idea that each state is always divisible into the states of its sub-systems, we show that it is sufficient to describe the evolution of infinitesimal subsystem within their state space (or phase space).
@@
 The above outline should help the reader maintain the big picture, as each section will go through the many details and definitions.
+We'll keep names and notation as consistent as possible to current use, but this may lead to some non sequitur as it will not be immediately clear why the different definition will be at the end equivalent to the standard one.
+This is still preferable than introducing a whole new set of names one has to become familiar.
+
 While we try to keep a precise language, no mathematical breakthrough should be expected: the goal, after all, is to derive the \emph{known} framework from a set of \emph{simple} definitions in the most \emph{obvious} way possible.
 No proof is longer than a couple of paragraphs, so the word \emph{theorem} is avoided in favor of \emph{proposition} and \emph{corollary}.
-What is surprising how so much can be derived from very few definitions.
+The novel, and surprising, result is how so much can be derived from so little.
+
+\section{States, Labels and Maps}
+
+The first two sections are dedicated to properly introduce states and the labels we use to identify them (such as position or temperature).
+The two issues that needs to be clarified are the nature of the system described (whole system vs infinitesimal component) and the cardinality of the states (discrete vs continuous variables).
+When talking about infinitesimal components and continuous labels, in fact, one must always remember that they are the result of a limit.
+Failure to do so would result in improper characterization of the system or counting of labels.
+We start with the whole system and discrete labels, then study the two limits separately.\footnote{Whole system vs infinitesimal component is also source of confusion when comparing classical and quantum states.
+Quantum states are always whole systems and are always distributions as there is no state uniquely attributed to its infinitesimal components.}
+
+\begin{defn}\label{statedef}
+Suppose we have a physical system to study.
+We define the set $\mathbbm{C}$ of all physically distinguishable configurations for that system.
+Each element $\mathbbm{c}$ we call \emph{configuration state}.
+\end{defn}
+
+\begin{assump1}\label{classical}
+The system is infinitely reducible: it can be thought as composed by two or more similar but smaller systems, each in its own configuration state, which also can be thought as composed by two or more, ad infinitum.
+\end{assump1}
+
+\begin{defn}\label{classicalPhaseSpace}
+Let $\mathbbm{S}$ be the set of all possible configuration states of the infinitesimal subdivision. We call this set \emph{phase space}. We call each element $\mathbbm{s}$ a \emph{state}.
+\end{defn}
+
+\begin{cor}\label{classicalDistribution}
+Each classical configuration state $\mathbbm{c} \in \mathbbm{C}$ is a distribution over states: $\mathbbm{c}=\sum\limits_{\mathbbm{s} \in \mathbbm{S}} D(\mathbbm{s}) \mathbbm{s}$, where $D:\mathbbm{S}\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}$.
+The distribution can be visualized as a histogram over the states in phase space.
+\end{cor}
+
+Under the classical assumption, we can then limit ourselves to the study of the infinitesimal elements, their states and their properties without loss of generality.
+To help identify states, we introduce the following concepts.
+
+\begin{defn}\label{label}
+We call a \emph{label} a set of states $i\subset\mathbbm{S}$; a \emph{set of labels} a collection of disjoint labels $I | \forall i_1,i_2\in I, i_1\bigcap i_2 = \emptyset$; a \emph{state variable} a set of labels $X$ that covers all of phase space: $\bigcup\limits_{i \in X}i=\mathbbm{S}$.
+Therefore a state belongs to one and only one label of a state variable.
+\end{defn}
+
+\begin{defn}\label{labelsCombine}
+Let $I_1$ and $I_2$ be two sets of labels.
+We can define the \emph{combined set}, $\langle I_1, I_2 \rangle$, whose labels consist of all the non-empty intersections of one label of $I_1$ and one of $I_2$.
+If all intersections are non-empty, $I_1$ and $I_2$ are said to be \emph{independent}, and we have $n(\langle I_1, I_2 \rangle)=n(I_1)n(I_2)$ where $n: I \rightarrow \mathbbm{R}$ gives the number of labels of each set.
+\end{defn}
+
+We now want to study how states and labels evolve in time, under the following assumption.
+
+\begin{assump2}
+The system undergoes deterministic (future state identified by the present state) and reversible (past state identified by the present state) evolution.
+\end{assump2}
+
+\begin{prop}\label{detrevMap}
+Let $\mathbbm{S}$ be the phase space of a system that undergoes deterministic and reversible evolution.
+There exists a bijective map $f:\mathbbm{S} \leftrightarrow \mathbbm{S}$ between past and future states.
+\end{prop}
+
+\begin{cor}\label{detrevDist}
+The evolution of a classical configuration state $\mathbbm{c}=\sum D(\mathbbm{s}) \mathbbm{s}$ under a bijective map is given by $\mathbbm{c'}=\sum D'(\mathbbm{s}) \mathbbm{s}=\sum D(f^{-1}(s)) \mathbbm{s}$.
+The evolution of the fraction of the system in a label $D(i)=\sum\limits_{\mathbbm{S} \in i} D(\mathbbm{s})$ is given by $D'(i)=D(f^{-1}(i))$.
+\end{cor}
+
+Mathematically, assuming determinism and reversibility means studying bijective maps.
+The evolution of a distribution simply moves the elements around: the bars of the histogram move place, but keep the same height.
+
+\begin{cor}\label{labelsCount}
+Given a label $i$, the image $f(i)$ is also a label containing the same number of states.
+Given a set of labels $I$, the image $f(I)$ is also a set of labels containing the same number of labels.
+Given a state variable $X$, the image $f(X)$ is also a state variable.
+Given two independent sets of labels $I_1$ and $I_2$, the images $f(I_1)$ and $f(I_2)$ are also independent.
+Therefore $n(f(\langle I_1, I_2 \rangle))=n(f(I_1))n(f(I_2))=n(I_1)n(I_2)=n(\langle I_1, I_2 \rangle)$
+\end{cor}
+
+Bijective maps preserve the number of labels as they provide one-to-one association between future and past.
+And they do so for each independent set of labels.
+These simple results using discrete labels, properly generalized to the continuous case, will give us Hamiltonian flow.
 

### 2015-01-16 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Adding continuous labels

@@
 These simple results using discrete labels, properly generalized to the continuous case, will give us Hamiltonian flow.
+
+\section{Continuous labels}
+
+Numeric labels are prevalent in physics.
+For the discrete case, we can use integers and expect a one-to-one map between $z\in\mathbbm{Z}$ and $i\in I$.
+The deterministic map than becomes $f:\mathbbm{Z} \leftrightarrow \mathbbm{Z}$.
+For the continuous case, one may simply expect to replace $\mathbbm{Z}$ with $\mathbbm{R}$, but this presents a problem.
+In the continuous limit, the distribution of our infinitesimal states will become a density which is defined over infinitesimal intervals.
+Since the distribution and the states must be defined on the same object, the label is really an interval over $\mathbbm{R}$, not a point.
+This also makes physical sense, as we never deal with points per se, but widths that can be made arbitrarily small.
+With this in mind, we have the following definitions.
+
+\begin{defn}\label{sdof}
+A \emph{degree of freedom} is a state variable identified by an infinitesimal cell of $\mathbbm{R}$.
+\end{defn}
+
+\begin{defn}\label{sdof}
+We define the \emph{generalized coordinate} $q$ as the center of each cell.
+We define the \emph{cell number}\footnote{This represents the classical analogue of the wave number.} $k$ such that $k \, dq$ represents the width of each cell.
+\end{defn}
+
+\begin{cor}\label{continuousLabels}
+Each degree of freedom is identified by the pair of labels $\langle q,k \rangle$. As $k\,dq$ is invariant under coordinate changes, $dk\wedge dq$ is invariant and $k$ is contravariant.
+\end{cor}
+
+\begin{prop}\label{continuousLabels}
+The cardinality of labels on a degree of freedom is given by $\hbar \, dq \wedge dk$, where $\hbar$ is a constant that defines the unit for labels.
+\end{prop}
+
+Since $q$ and $k$ are independent labels, the number of labels defined over a degree of freedom is the product of the labels defined on $q$ and $k$ (\ref{labelsCombine}).
+The number of labels for an interval of $q$ or $k$ will be proportional to said interval, therefore we have $\hbar \, dq \wedge dk$, where the constant is given by the choice of units for labels.
+It is customary to combine $\hbar$ with $k$.
+
+\begin{defn}\label{sdof}
+We define the \emph{conjugate momentum} $p=\hbar k$.
+\end{defn}
+
+\begin{cor}\label{continuousLabels}
+Each degree of freedom is identified by the pair of labels $\langle q,p \rangle$. $p dq$ and $dp \wedge dq$ are invariant, $p$ is contravariant.
+\end{cor}
+
+One can recognize the one-form $\theta = p dq$ and the two-form $\omega = dq \wedge dp$.
+They are intimately linked to the way states are defined ($\theta / \hbar$ is the cell width) and counted ($\omega$ is the cardinality of the labels).
+As such, they are truly fundamental objects.
 

### 2015-01-18 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Single degree of freedom evolution

@@
 \begin{prop}\label{continuousLabels}
-The cardinality of labels on a degree of freedom is given by $\hbar \, dq \wedge dk$, where $\hbar$ is a constant that defines the unit for labels.
+The cardinality of labels on a degree of freedom is given by $\omega = \hbar \, dq \wedge dk$, where $\hbar$ is a constant that defines the unit for labels.
 \end{prop}
@@
 \begin{cor}\label{continuousLabels}
-Each degree of freedom is identified by the pair of labels $\langle q,p \rangle$. $p dq$ and $dp \wedge dq$ are invariant, $p$ is contravariant.
+Each degree of freedom is identified by the pair of labels $\langle q,p \rangle$. $\theta_0 = p dq$ and $\omega = dq \wedge dp$ are invariant, $p$ is contravariant.
 \end{cor}
 
-One can recognize the one-form $\theta = p dq$ and the two-form $\omega = dq \wedge dp$.
-They are intimately linked to the way states are defined ($\theta / \hbar$ is the cell width) and counted ($\omega$ is the cardinality of the labels).
-As such, they are truly fundamental objects.
+\begin{prop}\label{continuousLabelDist}
+The configuration state over a degree of freedom is $\mathbbm{c}=\int D(q,p) \mathbbm{s}_{qp} = \int \rho(q,p) dq \wedge dp \mathbbm{s}_{qp}$, where $\rho(q,p)\equiv D(q,p) / \omega$ is the distribution density for each label.
+\end{prop}
+
+This is simply the limit of \ref{classicalDistribution} for the continuous case. Note that, since $\mathbbm{c}$ and $\omega$ are invariant under coordinate transformations, so is $\rho$.
+
+In the language of differential geometry, we recognise the function $\rho$, the one-form $\theta_0$ and the two-form $\omega$.
+They are truly fundamental objects as they are intimately linked to the way states are defined ($\rho$ is the distribution, $\theta_0 / \hbar$ is the cell width) and counted ($\omega$ is the cardinality of the labels) and they don't depend on coordinate choice.
+
+\section{Single degree of freedom}
+
+We can now derive the equation of motions for the state variables under deterministic and reversible evolution.
+
+\begin{defn}\label{canonical}
+We define \emph{canonical transformation} a bijective map on one (or more) degree of freedoms.
+\end{defn}
+
+\begin{prop}\label{continuousMapping}
+A canonical transformation must be continuous and preserve $\omega$.
+\end{prop}
+
+A bijective map conserves the cardinality of labels, see \ref{labelsCount}, therefore $\omega$ must be conserved.
+The map must be continuous in q, or it would split some cells into two parts: a cell would not be mapped to one and only one other cell, the mapping would not be bijective.
+The reverse mapping must be continuous in q as well, or the inverse would not map to one and only one cell.
+\begin{align*}
+dq' &= \frac{\partial q'}{\partial q} dq + \frac{\partial q'}{\partial p} dp \\
+dq &= \frac{\partial q}{\partial q'} dq' + \frac{\partial q}{\partial p'} dp'
+\end{align*}
+We can re-express $dp'$ in terms of $dq$ and $dp$, as the conservation of $\omega$ means the map is non-degenerate, find that all partial derivatives are well defined, and therefore the mapping is continuous in p as well.
+
+\begin{cor}\label{sdofInvariant}
+Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by two conjugate variables. Let
+\begin{align*}
+\omega_{a, b} = \left[
+  \begin{array}{cc}
+    0 & 1 \\
+    -1 & 0 \\
+  \end{array}
+\right] \\
+\end{align*}
+then $v'^{a} \omega_{a, b} w'^{b}=v^{a} \omega_{a, b} w^{b}$ under a canonical transformation, where $a$ and $b$ represent component along coordinates $\xi^{a} \equiv \{q,p\}$.
+\end{cor}
+
+Here we are simply expressing $\omega$, $v$ and $w$ with their respective components and underlying the fact that $\omega$ defines the metric conserved under canonical transformations.
+
+\begin{lem}\label{genAntisim}
+Let $v$ and $w$ be two vectors.
+Let $v^{a} \omega_{a, b} w^{b}$ be an antisymmetric product conserved under a continuous transformation parameterized by $t$.
+We can then define a function $H$ such that given $S^{a} \equiv d_{t}\xi^{a}$ and $S_{b} \equiv S^{a} \omega_{a, b}$, we have $S_{a} = \partial_{a}H$.
+\end{lem}
+
+$S^{a}$ is the vector field that represents how the state variables change. Simply applying the vector transformation rules under continuous transformation we have:
+\begin{align*}
+v^{a} \omega_{a, b} w^{b} &= v'^{a} \omega_{a, c} w'^{b}  \\
+&= (v^{a} + \partial_{c} S^{a} dt v^{c}) \omega_{a, b} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
+&= v^{a} \omega_{a, b} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{a, b} w^{b} \\
+ &+ v^{a} \omega_{a, b} \partial_{d} S^{b} w^{d}) dt + O(dt^2)
+\end{align*}
+\begin{align*}
+v^{c} w^{b} \partial_{c} S_{b} - v^{a} w^{d} \partial_{d} S_{a} = 0
+\end{align*}
+\begin{align*}
+\partial_{a} S_{b} - \partial_{b} S_{a} &= curl(S_{a}) = 0 \\
+S_{a} &= \partial_{a}H
+\end{align*}
+
+\begin{prop}\label{sdofHam}
+The time evolution for a single degree of freedom is given by:
+\begin{align*}
+d_{t}q &= \partial_{p} H \\
+d_{t}p &= - \partial_{q} H
+\end{align*}
+\end{prop}
+
+Simply expand \ref{genAntisim} with the metric defined in \ref{sdofInvariant}. We recognize Hamilton's equations for one degree of freedom\cite{classical_dynamics}.
 

### 2015-01-18 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Adding multiple and time d.o.f.

@@
 
-\section{Hamiltonian review}
+\section{Multiple degrees of freedom}
+
+\begin{prop}\label{mdofInvariant}
+Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by two independent degrees of freedom.
+Let $a$ and $b$ be indexes for the state variables $\xi^a\equiv \{q^i, p_i\}$.
+Let
+\begin{align*}
+\omega_{\alpha, \beta} =  \left[
+  \begin{array}{cc}
+    0 & 1 \\
+    -1 & 0 \\
+  \end{array}
+\right] \otimes \left[
+  \begin{array}{cc}
+    1 & 0 \\
+    0 & 1 \\
+  \end{array}
+\right] =
+\left[
+  \begin{array}{cccc}
+    0 & 0 & 1 & 0 \\
+    0 & 0 & 0 & 1 \\
+    -1 & 0 & 0 & 0 \\
+    0 & -1 & 0 & 0 \\
+  \end{array}
+\right] \\
+\end{align*}
+then $v'^{a} \omega_{a, b} w'^{b}=v^{a} \omega_{a, b} w^{a}$ under a canonical transformation.
+\end{prop}
+
+The independence between degrees of freedoms corresponds to orthogonality in phase space: from \ref{labelsCombine} the product between the number of labels on each d.o.f.
+(i.e.
+the area), must be equal to the number of combines labels (i.e.
+the hyper-volume), which is true only if the d.o.f are orthogonal in phase space.
+From \ref{labelsCount}, the mapping will preserve the cardinality of labels, the area\footnote{We assume we are using the same unit across d.o.f.} on each d.o.f, and the independence, orthogonality across d.o.f.\footnote{These statements provide a direct physical interpretation for Gromov's non-squeezing theorem\cite{Gromov,deGosson,Stewart}.} This is equivalent to requiring the conservation of the scalar product across independent degrees of freedom, while still requiring conservation of the vector product within.
+That leads us to the metric defined by \ref{mdofInvariant}.
+The metric generalizes \ref{sdofInvariant} to give us the cardinality of labels defined on the area given by two arbitrary directions in phase space.
+For an infinitesimal region, this corresponds to $dq^1 \wedge dp^1 + dq^2 \wedge dp^2$, the sum of the projections on the independent planes.
+Moreover, volume in phase space corresponds to the cardinality of the combined labels (i.e.
+the states), and is therefore conserved: this is Liouville's theorem.
+
+\begin{prop}\label{mdofHam}
+The evolution for multiple degrees of freedom is given by:
+\begin{align*}
+d_{t}q^i &= \partial_{p_i} H \\
+d_{t}p_i &= - \partial_{q^i} H
+\end{align*}
+\end{prop}
+
+Expand \ref{genAntisim} with the metric defined in \ref{mdofInvariant}. We recognize Hamilton's equations for multiple degrees of freedom\cite{classical_dynamics}.
+
+\section{Time dependence}
+
+So far we have assumed that neither state labeling nor mapping change in time. If they do, we also need to to use time as a label and therefore introduce an extra degree of freedom.
+
+\begin{defn}\label{tdof}
+The \emph{temporal degree of freedom} is a state variable identified by temporal cells.
+The center of each cell is identified by $t$, the width by $\omega dt$, the conjugate variable $E\equiv\hbar\omega$.
+We call \emph{extended phase space} the outer product between phase space and the temporal degree of freedom.
+\end{defn}
+
+\begin{prop}\label{tdofMonotonic}
+Let $s$ be the parameter of a trajectory in the extended phase space of a deterministic and reversible system. The trajectory must be continuous. There must exist a strictly monotonic function $t(s)$.
+\end{prop}
+
+The trajectory has to be continuous in both standard and temporal variables because of \ref{continuousMapping}.
+Since determinism and reversibility are defined in time, the trajectory must traverse all times once and only once: we must have an invertible mapping between $t$ and $s$, which means we must have a strictly monotonic $t(s)$.
+
+\begin{defn}\label{tdofAntistates}
+We call \emph{standard states} those connected by a trajectory where $d_{s}t>0$. We call \emph{anti-states} those connected by a trajectory where $d_{s}t<0$.
+\end{defn}
+
+Since $t(s)$ is strictly monotonic, $d_{s}t$ along a trajectory cannot change sign, so we have the division between standard and anti-states.
+Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call standard and anti-states is also conventional.
+What is physical and not conventional, though, is that standard and anti-states cannot be connected by deterministic and reversible evolution.
+
+\begin{prop}\label{tdofInvariant}
+Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by the temporal degree of freedom and one standard degree of freedom.
+Let $a$ and $b$ be indexes for the state variables $\xi^a\equiv\{t, E, q, p\}$.
+Let
+\begin{align*}
+\omega_{a, b} =  \left[
+  \begin{array}{cc}
+    0 & 1 \\
+    -1 & 0 \\
+  \end{array}
+\right] \otimes \left[
+  \begin{array}{cc}
+    -1 & 0 \\
+    0 & 1 \\
+  \end{array}
+\right]
+= \left[
+  \begin{array}{cccc}
+    0 & 0 & -1 & 0 \\
+    0 & 0 & 0 & 1 \\
+    1 & 0 & 0 & 0 \\
+    0 & -1 & 0 & 0 \\
+  \end{array}
+\right] \\
+\end{align*}
+then $v'^{a} \omega_{a, b} w'^{b}=v^{a} \omega_{a, b} w^{b}$ under deterministic and reversible evolution.
+\end{prop}
+
+$\langle t, E \rangle$ are not independent from $\langle q, p \rangle$ as they do not define new states.
+So they are not necessarily orthogonal in the extended phase space.
+Looking back at \ref{discreteLabelDef}, cells need to be defined on the plane where $\langle q, p \rangle$ (maximally) change: this is not the plane of constant $\langle t, E \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle q, p \rangle$ where $dt \wedge dE$ is defined.
+On that plane we can properly count states and define our invariant.
+
+We have a right triangle-like relationship between the plane where the invariant is defined and its projections on the planes defined by each d.o.f., similar to the multiple d.o.f.:
+\begin{align*}
+m.d.o.f \;\;\; &dq^1 \wedge dp_1 + dq^2 \wedge dp_2 = k \\
+t.d.o.f \;\;\; &dt \wedge dE + k = dq \wedge dp \\
+\end{align*}
+But in the previous case, the right angle was between the two independent d.o.f..
+In this case, the right angle is between the invariant and the plane of constant $\langle q, p \rangle$ where $dt \wedge dE$ is defined.
+We rewrite it as $dq \wedge dp - dt \wedge dE = k$.
+This corresponds to the Minkowski product across d.o.f.
+and the vector product within.
+The metric, with a space-like convention, still gives us the cardinality of labels within a degree of freedom.\footnote{Adjusted to avoid double counting.}
+
+\begin{prop}\label{tdofHam}
+The evolution for time varying multiple degrees of freedom is given by:
+\begin{align*}
+d_{s}t &= - \partial_{E} \mathcal{H} \\
+d_{s}E &= \partial_{t} \mathcal{H} \\
+d_{s}q^i &= \partial_{p_i} \mathcal{H} \\
+d_{s}p_i &= - \partial_{q^i} \mathcal{H}
+\end{align*}
+\end{prop}
+
+Take the metric from \ref{tdofInvariant}, add multiple independent d.o.f as in \ref{mdofInvariant}, use \ref{genAntisim} with the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$.
+
+If we set\footnote{We avoided using $p^{n+1}$ as it hides the minus sign from the metric, making it seem that the temporal d.o.f is just another independent d.o.f.} $q^{n+1}=t$ and $p^{n+1}=-e$, we recognise Hamilton equations in the extended phase space\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}\cite{Synge,Lanczos}.
+
+It should not be a surprise that the equations do not mention the speed of light $c$.
+In fact, nothing says that all $q^i$ represent space or that the laws of motion are invariant in all inertial frames.
+The only requirement we have is that the areas of each degree of freedom represent the same cardinality for labels.
+
+\begin{prop}\label{tdofConstrain}
+The evolution is constrained by $\mathcal{H}=k$.
+\end{prop}
+
+Since $\mathcal{H}$ is constant through the evolution, it can serve both as the generating function and as the evolution constraint.
+By convention, we can set $\mathcal{H}=0$ without loss of generality as changing $\mathcal{H}$ by a constant does not change the equation of motion.
+This reduces extended phase space to $\mathbb{R}^{2*N + 1}$, the state variables plus time.
 

### 2015-01-21 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Kinematics up to relativity

@@
 \newtheorem*{assump2}{Determinism and Reversibility assumption}
-\newtheorem*{assump3}{Kinematic assumption}
+\newtheorem*{assump3}{Kinematic equivalence assumption}
 
@@
 \section{Kinematics}
+So far we talked about states without reference to what they physically represent.
+
 \begin{assump3}\label{kinematicAssumption}
-The study of the motion of a body is equivalent to study its state under deterministic and reversible evolution.
+The study of the trajectory (kinematic) of a body is equivalent to study its state (dynamic) under deterministic and reversible evolution.
 \end{assump3}
@@
 \begin{cor}\label{}
-Consider all states at $t=t_0$ and all possible trajectories, to each state corresponds one and only one trajectory; to each trajectory corresponds one and only one state.
+Given all possible space-time trajectory $x^\alpha(s)$ and all possible phase space trajectory $\xi^a(s)$, there exist a bijective function $f: x^\alpha(s) \leftrightarrow \xi^a(s)$ that links each space-time trajectory with one and only one phase space trajectory.
 \end{cor}
 
-If studying the motion and state evolution are equivalent, given a state we must be able to reconstruct the trajectory, and given the trajectory we must be able to reconstruct the state.
-Given that the choice of state coordinates is arbitrary, we can use position and time as a set of continuous state variables: $q^\alpha = x^\alpha$.
-With that in mind:
+If studying the motion and state evolution are equivalent, then we must be able to go back and forth between the two pictures.
+Without losing generality, we can chose $x^\alpha=f(q^i,t)$ to be a function of only $q^i$ and $t$.
 
-\begin{prop}\label{}
-A metric $g_{\alpha\beta}$ must be defined on the manifold identified by $x^\alpha$. All trajectories must be continuous.
+\begin{prop}\label{locallyMinkowski}
+Space-time is locally Minkowskian Rimmanian manifold.
+That is, there exists a metric $g_ij$ that at any point can be expressed, with a suitable choice of coordinate, as $dx^\alpha g_{\alpha \beta}|_P dx^\beta=dx^\alpha\eta_{\alpha \beta}dx^\beta=(dx^i)^2 - (dx^0)^2=(dx^i)^2 - c^2dt^2$, where $c=n(dq)/n(dt)$ is the ration between the density of states in space and time.
 \end{prop}
 
-As each label corresponds to an infinitesimal cell, we need distances properly defined.
-This requires a metric defined on the manifold.
-And since evolution is continuous on state variables, it will be continuous in $x^\alpha$ as well.
+As states are defined on intervals, a metric $g$ must be defined on $Q$.
+Such a metric must be consistent with $\omega$, as both must be invariant under coordinate transformations.
+The idea is that each $(dx^\alpha)^2$ can be made to both represent a length square in space-time and an area in phase space, linking the two metrics.
+For each spatial d.o.f.
+fix $dp_i=\lambda dq^i$, we have $dq^i \wedge dp_i = \lambda (dx^i)^2$ where $\lambda$ converts from units of length squared to cardinality of labels.
+For the temporal d.o.f fix $dE = \lambda c^2 dt$, we have $dt \wedge dE = \lambda c^2 dt^2$.
+That is: the area in time squared is converted to an area in length squared that have the same density of states (as per definition of $c$) and then to cardinality of labels.
+The phase space invariant is $\omega = \lambda [(dq^i)^2 - c^2 dt^2]$.
 
-\begin{defn}\label{continuousLabels}
-An \emph{inertial frame} is one for which each direction of space represents a homogeneous state variable of an independent degree of freedom.\footnote{While a local inertial frame always exists, the existence of a global one is an added requirement.
-We add it because it makes the discussion easier and the proofs more obvious.
-Most of the results, though, will hold without that assumption.
-This will be important in future works that will try to extend this framework to general relativity.}
-\end{defn}
+Assume the choice of coordinates $\x^alpha$ locally diagonalizes $g(x^\alpha)$, each diagonal element being either $\pm 1$ (such coordinate system always exists).
+Set $x^i=q^i$ and $x^0=ct$.
+We have $\omega = \lambda ((dx^i)^2 - (dx^0)^2)$ and $dx^\alpha g_{\alpha \beta} dx^\beta=dx^\alpha g_{\alpha \alpha}dx^\alpha$.
+Both are invariant under coordinate transformation for any $dx^\alpha$.
+This can only be if $g_{ii}(x^\alpha)=1$ and $g_{00}(x^\alpha)=-1$.
 
-\begin{cor}\label{continuousLabels}
-The metric associated with an inertial frame is
+It is fitting that deterministic and reversible evolution requires space-time to be locally Minkowskian, as this clearly defines past and future events.
+To make us understand better the role of $c$, as we defined it, we prove the following.
+
+\begin{prop}\label{locallyMinkowski}
+The speed of a body cannot exceed $c$ under the kinematic equivalence assumption.
+\end{prop}
+
+Consider a movement $ds$ along any trajectory.
+This will go through $n(dt)$ labels in time and $n(dq)$ labels in space.
+As the motion is continuous, the number of labels in space cannot exceed the number of labels in time, we can't skip them, therefore $n(dq)\leq c n(dt)$.
+So we have:
 \begin{align*}
-g_{\alpha, \beta} =  \left[
-  \begin{array}{cccc}
-    -1 & 0 & 0 & 0 \\
-    0 & 1 & 0 & 0 \\
-    0 & 0 & 1 & 0 \\
-    0 & 0 & 0 & 1 \\
-  \end{array}
-\right] \\
+\frac{n(dq)}{n(dt)}= \frac{dq}{dt} \leq c
 \end{align*}
-\end{cor}
-
-The relationships between coordinates is the same as in phase space: $X^i$ must be orthogonal with each other, as the corresponding $Q^i$ are orthogonal in phase space; $X^i$ and $X^0$ are not in general orthogonal and it is the length perpendicular to time that gets conserved.
-Given that $X^\alpha$ are homogeneous, the metric must be invariant under translation, does not change in space-time, it is constant.
-With a suitable choice of units, we can set $|g_{\alpha\alpha}|=1$.
-All these combines gives us the Minkowski metric.
-We also introduce a constant $c$\footnote{Here $c$ is not defined as a a speed: it's the constant that allows to convert the cardinality of labels between space and time coordinates.} to convert time intervals to space intervals with the same number of labels, so we have $x^0=ct$.
-
-It's important to note that the geometry of extended phase space and the geometry of space-time are linked: we cannot define cells in one without also defining cells in the other.
-This is true for all reference frames, not just inertial ones.
 

### 2015-01-22 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Up to expression for Hamiltonian

@@
 
-Assume the choice of coordinates $\x^alpha$ locally diagonalizes $g(x^\alpha)$, each diagonal element being either $\pm 1$ (such coordinate system always exists).
+Assume the choice of coordinates $x^\alpha$ locally diagonalizes $g(x^\alpha)$, each diagonal element being either $\pm 1$ (such coordinate system always exists).
 Set $x^i=q^i$ and $x^0=ct$.
@@
 \begin{align*}
-\frac{n(dq)}{n(dt)}= \frac{dq}{dt} \leq c
+\frac{n(dq)}{n(dt)}= \frac{dq}{dt} \leq c \\
 \end{align*}
 
-\begin{prop}\label{continuousLabels}
-The position $x^\alpha$ and the velocity $dx^i/dt$ are necessary and sufficient initial conditions to determine the state of the system and its whole trajectory.
+\begin{prop}\label{initialConditions}
+Let $u^\alpha = d_s x^\alpha$ be the four-velocity, where the parametrization $s$ is chosen, by convention, to be proper time.
+The position $x^\alpha$ and velocity $u^\alpha$ are necessary and sufficient initial conditions to determine the state of the system and its whole trajectory.
 \end{prop}
 
-The initial conditions will be given by the position and its first $n$ derivatives: $x^i_k=d^kx^i(t_0)/dt^k$ where $k=0..n$.
-This means that there is a function that, given the initial conditions, gives us the trajectory: $x(t)=f(t_0, x^i_k)$.
-We note that the initial conditions can be changed by an active transformation, therefore counting the number of possible initial conditions is equivalent to counting the possible number of active transformations.
-Under such change, though, the metric cannot change: it needs to be a change in initial conditions in the same reference frame, or we would double count each case.
-Defining the transformation for the position also defines how all its derivatives change, and by requiring the invariance of the metric, our options are actually very limited.
-Assuming we start from an inertial frame, only a linear transformation in all $x^\alpha$ can preserve the metric.
-A linear transformation can only change position and velocity, so the initial condition can only be limited to those.
-We also note that the active transformation can always change the velocity of $x(t)$, so $f$ must depend on the velocity or it would not be able to reach all possible trajectories defined by an active transformation in the same reference frame.
-Therefore position and velocity are both necessary and sufficient initial condition.
+The equation of motion \ref{tdofHam} can be at most second order in $q^i$, they can at most be second order in $x^\alpha=\{ct, q^i\}$ so only position and velocity can be candidates for the initial conditions.
+Fixing time, phase space is $\mathbbm{R}^{2n}$, too big to be covered by position only, but just right to be covered by both.
+
+\begin{prop}\label{kineticMomentum}
+Let $x^\alpha=\{ct, q^i\}$ and $p_\alpha=\{-E/c, p_i\}$.
+Then $p_\alpha= m g_{\alpha \beta}u^\beta + \hat{p}_\alpha(x^\gamma)$, where $m=n(dp_\alpha)/n(d(g_{\alpha \beta}u^\beta))$ is the ration between the density of states in conjugate momentum and in contravariant velocity, and is called \emph{inertial mass}.
+\end{prop}
+
+For \ref{kinematicAssumption} and \ref{kineticMomentum}, there must exist $p_\alpha=p_\alpha(x^\beta , u^\gamma)$. We express it in terms of $u_\alpha\equiv g_{\alpha \beta} u^\beta$. We have:
+\begin{align*}
+\omega &= q^i\wedge p_i - ct \wedge E/c \\
+&=x^\alpha \wedge p_\alpha \\
+&=dx^\alpha \wedge \frac{\partial p_\alpha}{\partial u_\beta}du_\beta + dx^\alpha \wedge \frac{\partial p_\alpha}{\partial x^\gamma}dx^\gamma \\
+&=\frac{\partial p_\alpha}{\partial u_\beta}du_\beta dx^\alpha \\
+\end{align*}
+Consider the expression $m du_\alpha dx^\alpha$: this gives us the density of labels in position and velocity converted to the conjugate variables.
+In other world, we should have $\omega=m du_\alpha dx^\alpha$.
+This means:
+\begin{align*}
+\frac{\partial p_\alpha}{\partial u_\beta} &= m \delta^\beta_\alpha \\
+p_\alpha &= m g_{\alpha \beta}u^\beta + \hat{p}_\alpha(x^\gamma)
+\end{align*}
+where $\hat{p}$ is an arbitrary function.
+
+To convince ourselves that $m$ is indeed the inertial mass, consider applying a force.
+We are changing the state through the velocity, meaning changing the conjugate momentum.
+The higher the mass, the more states we'll have to go through to reach the same velocity.
+The higher the mass, the more change is required, the more force needs to be applied.
+
+\begin{prop}\label{kineticHamiltonian}
+The extended Hamiltonian is $\mathcal{H}=\frac{1}{2m}(p_\alpha-\hat{p}_\alpha(x))g^{\alpha\beta}(p_\beta-\hat{p}_\beta(x))+\hat{\mathcal{H}}(x)$
+\end{prop}
+We have:
+\begin{align*}
+\frac{dq^\alpha}{ds} &= u^\alpha \\
+&= \frac{1}{m}g^{\alpha\beta}(p_\beta-\hat{p}_\beta) \\
+&= \frac{\partial \mathcal{H}}{dp_\alpha} \\
+\end{align*}
+Integrating we have the expression for the Hamiltonian, where $\hat{\mathcal{H}}$ is an arbitrary function.
 

### 2015-01-23 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Up to the relativistic equation of motion

@@
 
-\begin{prop}
-Principle of relativity. The laws of motion have the same form for all inertial observers.
+\begin{prop}\label{kineticLaw}
+Let $\hat{p}_\alpha = q A_\alpha$, $F_{\alpha \beta} \equiv \partial_\alpha A_\beta - \partial_\beta A_\alpha$ and $\hat{H} = 0$, the equations of motion are $m \nabla_s u^\alpha = g^{\alpha\beta} F_{\beta \gamma} q u^\gamma$.
 \end{prop}
 
-We saw that an active transformation that does not change the metric is a change in the initial conditions.
-Such a change should not affect the form of the function that given the initial conditions gives us the whole trajectory: the observer is the same, the law of motion is the same, they just shift the value of the arguments.
-But if the form of the laws of motion is preserved under such active transformations, it will also not change under passive transformation that preserve the metric.
-Given that all inertial observers can be reached by a linear transformation, which preserves the metric, then the laws of motion will have the same form.
+We first derive the following relationship for later use:
+\begin{align*}
+\partial_\alpha \delta^\beta_\gamma &= 0 = \partial_\alpha g^{\beta\delta} g_{\delta\gamma} + g^{\beta\delta} \partial_\alpha g_{\delta\gamma}\\
+\partial_\alpha g^{\beta\epsilon} &= - g^{\beta\delta} g^{\gamma\epsilon} \partial_\alpha g_{\delta\gamma} 
+\end{align*}
 
-%x(t)=f(x_0, x_1, ... x_n)
+We expand \ref{tdofHam} using \ref{kineticHamiltonian}:
+\begin{align*}
+u^\alpha &= \frac{dq^\alpha}{ds} = \frac{\partial \mathcal{H}}{dp_\alpha} \\
+&= \frac{1}{m}g^{\alpha\beta}(p_\beta-\hat{p}_\beta) \\
+d_s p_\alpha &= - \frac{\partial \mathcal{H}}{\partial q^\alpha} \\
+&=\frac{1}{2m}[\partial_\alpha \hat{p}_\beta g^{\beta \gamma} (p_\gamma -\hat{p}_\gamma) \\
+ &- (p_\beta -\hat{p}_\beta) \partial_\alpha g^{\beta \gamma} (p_\gamma -\hat{p}_\gamma) \\
+ &+ (p_\beta -\hat{p}_\beta) g^{\beta \gamma} \partial_\alpha \hat{p}_\gamma ]- \partial_\alpha \hat{H} \\
+&=\frac{1}{2}[\partial_\alpha \hat{p}_\beta u^\beta
+- m u^\delta g_{\delta\beta} \partial_\alpha g^{\beta \gamma} u^\epsilon g_{\epsilon\gamma}
++ u^\gamma \partial_\alpha \hat{p}_\gamma ]- \partial_\alpha \hat{H}
+\end{align*}
 
-%What is the space size for a particular frame?
+We then calculate the four-force:
+\begin{align*}
+m d_s u^\alpha &= d_s g^{\alpha\beta}(p_\beta-\hat{p}_\beta) + g^{\alpha\beta} d_s (p_\beta-\hat{p}_\beta) \\
+&= \partial_\gamma g^{\alpha\beta} d_s x^\gamma m g_{\beta \delta} u^\delta + g^{\alpha\beta} (d_s p_\beta - \partial_\gamma \hat{p}_\beta d_s x^\gamma) \\
+&= \partial_\gamma g^{\alpha\beta} u^\gamma m g_{\beta \delta} u^\delta + g^{\alpha\beta} \frac{1}{2} [\partial_\beta \hat{p}_\gamma u^\gamma
+- m u^\epsilon g_{\epsilon\gamma} \partial_\beta g^{\gamma \delta} u^\zeta g_{\zeta\delta} \\
+&+ u^\delta \partial_\beta \hat{p}_\delta ]- g^{\alpha\beta} \partial_\beta \hat{H} - g^{\alpha\beta} \partial_\gamma \hat{p}_\beta u^\gamma \\
+&= m  g_{\beta \delta} \partial_\gamma g^{\alpha\beta} u^\gamma u^\delta - \frac{1}{2} m g^{\alpha\beta} g_{\zeta\delta} g_{\epsilon\gamma} \partial_\beta g^{\gamma \delta} u^\epsilon u^\zeta  \\
+&+ g^{\alpha\beta} \partial_\beta \hat{p}_\gamma u^\gamma - g^{\alpha\beta} \partial_\gamma \hat{p}_\beta u^\gamma
+- g^{\alpha\beta} \partial_\beta \hat{H}\\
+&= - m  g^{\alpha \beta} \partial_\gamma g_{\beta\delta} u^\gamma u^\delta + \frac{1}{2} m g^{\alpha\beta} \partial_\beta g_{\gamma \delta} u^\gamma u^\delta  \\
+&+ g^{\alpha\beta} (\partial_\beta \hat{p}_\gamma - \partial_\gamma \hat{p}_\beta ) u^\gamma
+- g^{\alpha\beta} \partial_\beta \hat{H}\\
+&= - m \frac{1}{2} g^{\alpha \beta} ( \partial_\gamma g_{\beta\delta} + \partial_\delta g_{\beta\gamma} - \partial_\beta g_{\gamma \delta} ) u^\gamma u^\delta  \\
+&+ g^{\alpha\beta} (\partial_\beta \hat{p}_\gamma - \partial_\gamma \hat{p}_\beta ) u^\gamma
+- g^{\alpha\beta} \partial_\beta \hat{H}\\
+&= - m \Gamma ^\alpha_{\ \gamma \delta} u^\gamma u^\delta + g^{\alpha\beta} (\partial_\beta \hat{p}_\gamma - \partial_\gamma \hat{p}_\beta ) u^\gamma
+- g^{\alpha\beta} \partial_\beta \hat{H}
+\end{align*}
+\begin{align*}
+m \nabla_{u} u^\alpha &= m (d_s u^\alpha + \Gamma ^\alpha_{\ \gamma \delta} u^\gamma u^\delta)  \\
+&= g^{\alpha\beta} (\partial_\beta \hat{p}_\gamma - \partial_\gamma \hat{p}_\beta ) u^\gamma - g^{\alpha\beta} \partial_\beta \hat{H}\\
+&= g^{\alpha\beta} (\nabla_\beta \hat{p}_\gamma - \nabla_\gamma \hat{p}_\beta ) u^\gamma - g^{\alpha\beta} \nabla_\beta \hat{H}
+\end{align*}
 
-%We can reach any initial condition by coordinate changes
-%But they also change f
-%Free parameters are going to be the ones we can change without changing f
-%How many coordinate changes are available to use without changing f?
-%The ones that preserve the metric
+The equation is manifestly covariant.
+If $\hat{p}$ and $\hat{H}$ are zero, absence of forces, we recognize the geodesic equation.
+We substitute $\hat{p}_\alpha = q A_\alpha$ and $\hat{H} = 0$ and have:
 
-%l_\alpha^\gamma l_\beta^\delta g_{\gamma\delta} = g_{\alpha\beta}
+\begin{align*}
+m \nabla_{u} u^\alpha &= g^{\alpha\beta} (\partial_\beta A_\gamma - \partial_\gamma A_\beta ) q u^\gamma \\
+&= g^{\alpha\beta} F_{\beta \gamma} q u^\gamma\\
+\end{align*}
 
-%Suppose different coordinate frame with same metric
-%x'(t)=f(x'_0, x'_1, ... x'_n)
-
-Assume trajectory are deterministic and reversible. Have state.
-
-How much state? Find that is two per degree of freedom. Position and momentum are only state variables.
-
-Inertial frame
-
-\section{Connect to Hamiltonian}
-\begin{prop}\label{continuousLabels}
-Under the kinematic assumption, there must exist a bijective transformation $q^\alpha=q^\alpha(x^\alpha,u^\alpha)$ and $p^\alpha=p^\alpha(x^\alpha,u^\alpha)$ between initial conditions and state variables.
-As such, they are monotonic in both variables.
-\end{prop}
-
-As we have seen, $X^\alpha$ and $U^\alpha$ are state variables that fully identify our degrees of freedom. Nothing tells us, though, that
-they are conjugate variables. But, since for every set of initial condition there must be one and only one state associated with it. This means the transformation must be invertible, monotonic.
-
-\begin{prop}\label{continuousLabels}
-Under the kinematic assumption, the extended phase space is defined on the conjugate variables $q^\alpha=x^\alpha$ and $p_\alpha=m\frac{dx_\alpha}{ds}+\hat{p}_\alpha(x)$.
-\end{prop}
-
-Given the degree of arbitrariness in choice of transformation, we can set $q^\alpha=x^\alpha$.
-This creates a direct link between how space is measured ($g_{\alpha\beta}(x^\alpha)$) and the width of our cells in phase space ($m(q)$)
-
-$\partial_\alpha \delta^\beta_\gamma = 0 = \partial_\alpha g^{\beta\delta} g_{\delta\gamma} + g^{\beta\delta} \partial_\alpha g_{\delta\gamma}$
-
-$\partial_\alpha g^{\beta\delta} g_{\delta\gamma} = - g^{\beta\delta} \partial_\alpha g_{\delta\gamma}$
-
-$\partial_\alpha g^{\beta\delta} g_{\delta\gamma} g^{\gamma\epsilon} = - g^{\beta\delta} \partial_\alpha g_{\delta\gamma} g^{\gamma\epsilon}$
-
-$\partial_\alpha g^{\beta\epsilon} = - g^{\beta\delta} g^{\gamma\epsilon} \partial_\alpha g_{\delta\gamma} $
-
-$p_\alpha=m\frac{dx_\alpha}{ds}+\hat{p}_\alpha(x)$
-
-$\frac{dx^\alpha}{ds}=\frac{\partial H}{\partial p_\alpha}=\frac{1}{m}(p^\alpha-\hat{p}^\alpha(x))$
-
-$H=\frac{1}{2m}(p_\alpha-\hat{p}_\alpha)g^{\alpha\beta}(p_\beta-\hat{p}_\beta(x))+\hat{H}(x)$
-
-$\frac{dp_\alpha}{ds}=-\frac{\partial H}{\partial q^\alpha}=$
-
-$\frac{1}{2m}[\partial_\alpha \hat{p}_\beta g^{\beta \gamma} (p_\gamma -\hat{p}_\gamma)
- - (p_\beta -\hat{p}_\beta) \partial_\alpha g^{\beta \gamma} (p_\gamma -\hat{p}_\gamma)
- + (p_\beta -\hat{p}_\beta) g^{\beta \gamma} \partial_\alpha \hat{p}_\gamma ]- \partial_\alpha \hat{H}(x)=$
-
-$\frac{1}{2}[\partial_\alpha \hat{p}_\beta u^\beta
-- m u^\delta g_{\delta\beta} \partial_\alpha g^{\beta \gamma} u^\epsilon g_{\epsilon\gamma}
-+ u^\gamma \partial_\alpha \hat{p}_\gamma ]- \partial_\alpha \hat{H}(x)=$
-
-$\partial_\alpha \hat{p}_\beta u^\beta + \frac{m}{2}u^\beta \partial_\alpha g_{\beta \gamma} u^\gamma
-- \partial_\alpha \hat{H}(x)=$
-
-$m\frac{d^2x_\alpha}{ds^2}+\frac{d\hat{p}_\alpha(x)}{ds}$
-
-$m\frac{d^2x_\alpha}{ds^2}=\frac{1}{2}[\frac{\partial\hat{p}_\beta(x)}{\partial x^\alpha} (\frac{dx^\beta}{ds}) + (\frac{dx_\beta}{ds})\frac{\partial\hat{p}^\beta(x)}{\partial x^\alpha} ]-\frac{\partial\hat{p}_\alpha(x)}{\partial x^\beta}\frac{dx^\beta}{ds}
--\frac{\partial \hat{H}(x)}{\partial x^\alpha}$
-
-$m\frac{d^2x_\alpha}{ds^2}=(\frac{\partial\hat{p}_\beta(x)}{\partial x^\alpha} - \frac{\partial\hat{p}_\alpha(x)}{\partial x^\beta} ) \frac{dx\beta}{ds}
--\frac{\partial \hat{H}(x)}{\partial x^\alpha}$
-
-Find that p must be contra-variant. Must be monotonic. And a linear transformation of v. Introduce mass. Introduce gauge.
+This is the relativistic equation for a charged particle.
 

### 2015-01-27 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Reviewed outline and first section

@@
 \newtheorem{defn}[thm]{Definition}
-\newtheorem*{assump1}{Classical assumption}
+\newtheorem*{assump1}{Infinite reducibility (or classical) assumption}
 \newtheorem*{assump2}{Determinism and Reversibility assumption}
@@
 \begin{abstract}
-We introduce three fundamental principles (infinite reducibility, deterministic and reversible evolution, ) and show that classical mechanics can be derived them.
-The core idea is that deterministic and reversible systems preserve the number of labels we use to identify states, which puts considerable constraints and leads to the Hamiltonian and Lagrangian frameworks, space-time being locally Minkowskian, the general form of the Hamiltonian and equations of motion equivalent to those for relativistic charged particles.
+We show that classical mechanics, including Hamiltonian, Lagrangian, non-relativistic Newtonian gravitation and relativistic electromagnetism, can be derived from three fundamental assumptions: infinite reducibility, deterministic and reversible evolution, and kinematic equivalence.
+The core idea is that deterministic and reversible systems preserve the cardinality of a set of states, which puts considerable constraints on the equations of motion.
 Many familiar fundamental concepts and constants (such as mass, the speed of light, the Plank constant, the vector potential) are introduced and defined in a novel way, thus providing different insight into them.
@@
 \section{Introduction}
+
 Classical mechanics is usually founded on Newton's laws.
@@
 Also, given that classical and quantum mechanics are formally equivalent, there is a chance that such work could be extended to quantum mechanics.
-
-We'll first mathematically define states and the labels (e.g.
-position, momentum, temperature, pressure) we use to identify them.
-Using the principle of infinite reducibility (or classical assumption), the idea that each state is always divisible into the states of its sub-systems, we show that it is sufficient to describe the evolution of infinitesimal subsystem within their state space (or phase space).
-
-Under the assumption of deterministic and reversible evolution, to each state corresponds one and only one future (or past) state, we show that the cardinality of labels (and states) is conserved.
-This allows us to define a metric on phase space, and the conservation of that metric leads to the Hamiltonian framework.
-That is: we can derive the Hamiltonian framework on its own merit and show that it is equivalent to deterministic and reversible motion.
-
-We then introduce the principle of kinematic equivalence, the idea that studying trajectories is equivalent to studying states, and show that under this assumption we can define a transformation between state and kinematic variables.
-Noting that a passive coordinate change in space-time is equivalent to a re-labelling of states, and as such should conserve the phase space metric (i.e.
-the cardinality of states), the metric of space-time is constrained to be locally Minkowskian.
-Under the kinematic equivalence, the relationship between velocity and conjugate momentum has to be monotonic, leading to a concave Hamiltonian which allows a Legendre transformation leading to the Lagrangian.
-We can constrain the Hamiltonian further, and show that the most general form of motion under the three condition is the one of a geodesics modified by the force given by a vector potential, such as the one of a relativistic charged particle.
 
@@
 
+\section{Outline}
+
+We give here a brief a general conceptual overview, hoping it will help guide the reader through mathematical details.
+
+We'll first mathematically define states and the labels we use to identify them (e.g.
+position, momentum, temperature, pressure).
+We introduce the infinite reducibility assumption (or classical assumption): each state is divisible into the states of its sub-systems.
+It is then sufficient to describe the evolution of infinitesimal subsystem within their state space (corresponding to point particles in phase space).
+
+We introduce the deterministic and reversible evolution assumption: to each state corresponds one and only one future (or past) state.
+The cardinality of a set of states is therefore conserved under evolution.
+This allows us to define a metric $\omega$ on phase space, and the conservation of that metric leads to the Hamiltonian framework.
+That is: we can derive the Hamiltonian framework on its own merit and show that it is equivalent to deterministic and reversible motion.
+
+We introduce the assumption of kinematic equivalence: studying trajectories is equivalent to studying states.
+We show that there must be link between the space-time metric $g$ and the phase-space metric $\omega$, as both quantities must be conserved by passive coordinate transformations.
+That link constrains the space-time metric to be locally Minkowskian.
+A transformation between state variable $(q, p)$ and kinematic variables $(x, \dot{x})$.
+To be invertible, the relationship between velocity and conjugate momentum has to be monotonic, leading to a concave Hamiltonian which allows a Legendre transformation leading to the Lagrangian.
+We can constrain the Hamiltonian further, and show that the most general equation of motion under the three condition is the one of a geodesic modified by the force given by a vector potential, such as the one of a relativistic charged particle, and a scalar potential, such as the one of newtonian gravitation.
+
 \section{States, Labels and Maps}
 
-The first two sections are dedicated to properly introduce states and the labels we use to identify them (such as position or temperature).
-The two issues that needs to be clarified are the nature of the system described (whole system vs infinitesimal component) and the cardinality of the states (discrete vs continuous variables).
-When talking about infinitesimal components and continuous labels, in fact, one must always remember that they are the result of a limit.
-Failure to do so would result in improper characterization of the system or counting of labels.
-We start with the whole system and discrete labels, then study the two limits separately.\footnote{Whole system vs infinitesimal component is also source of confusion when comparing classical and quantum states.
+The first two sections are dedicated to properly define states.
+In particular, we will need to introduce more precise terminology to be able to make two crucial distinctions.
+The first is between the state of the whole system and one of its parts.\footnote{Whole system vs infinitesimal component is also source of confusion when comparing classical and quantum states.
 Quantum states are always whole systems and are always distributions as there is no state uniquely attributed to its infinitesimal components.}
+\begin{center}
+    \begin{tabular}{ | p{2.5cm} | p{5.5cm} | }
+    \hline
+    Configuration state & The state of the whole system being described. \\ \hline
+    Elemental state (or just state) & The state of an infinitesimal part of the system. \\ \hline
+    \end{tabular}
+\end{center}
+The second is, when referring to a physical quantity, between the space of all possible values, a particular value or a set of possible values.
+\begin{center}
+    \begin{tabular}{ | p{2.5cm} | p{5.5cm} | }
+    \hline
+    State variable & A quantity that must be specified to identify a state (e.g. position). \\ \hline
+    Label & A particular value for a state variable (e.g. position = $5m$). \\ \hline
+    Label set & A set of possible values for a state variable (e.g. position = $[4.5m, 5.5m]$). \\
+    \hline
+    \end{tabular}
+\end{center}
+When talking about infinitesimal components and continuous state variables, we must always remember that they are the result of a limit.
+Therefore we will start with discrete definitions and do the limit, and see how some properties will extend from discrete case to continuous.
+These are practically the only new terms introduced by this work, but they are fundamental as most of the later derivation will be based on \emph{counting the labels} (i.e.
+counting the possible values of a physical quantity) and making sure that such number is conserved (i.e.
+the number of cases is the same).
+
+This section will mathematically define the above concepts, limited to discrete state variables, and derive some basic properties that hold after applying bijective maps.
 
@@
 \begin{defn}\label{classicalPhaseSpace}
-Let $\mathbbm{S}$ be the set of all possible configuration states of the infinitesimal subdivision. We call this set \emph{phase space}. We call each element $\mathbbm{s}$ a \emph{state}.
+Let $\mathbbm{S}$ be the set of all possible configuration states of the infinitesimal subdivision.
+We call this set \emph{phase space}.
+We call each $\mathbbm{s} \in \mathbbm{S}$ an \emph{elemental state}, or simply \emph{state}.
 \end{defn}
@@
 \end{defn}
+
+TODO: NOTATION maybe $X$ should be $\mathcal{I}$?
 

### 2015-02-06 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Reviewed second section
Limit definition of states and QT.

@@
 
-Numeric labels are prevalent in physics.
-For the discrete case, we can use integers and expect a one-to-one map between $z\in\mathbbm{Z}$ and $i\in I$.
-The deterministic map than becomes $f:\mathbbm{Z} \leftrightarrow \mathbbm{Z}$.
-For the continuous case, one may simply expect to replace $\mathbbm{Z}$ with $\mathbbm{R}$, but this presents a problem.
-In the continuous limit, the distribution of our infinitesimal states will become a density which is defined over infinitesimal intervals.
-Since the distribution and the states must be defined on the same object, the label is really an interval over $\mathbbm{R}$, not a point.
+The above definitions readily apply for labels identified by integers.\footnote{For each label $i$ there is one and only one integer $z$.} Let $z \in \mathbbm{Z}$, we have $\mathbbm{c}=\sum\limits_{z \in \mathbbm{Z}} D(z) \mathbbm{s}(z)$ and the deterministic map becomes $f:\mathbbm{Z} \leftrightarrow \mathbbm{Z}$.
+For the continuous case, one may simply expect to replace $z \in \mathbbm{Z}$ with $r \in \mathbbm{R}$, but this does not work.
+In the continuous limit, we would have $\mathbbm{c}=\int\limits_{r \in \mathbbm{R}} \rho(r) dr \mathbbm{s}(r)$, where $\rho(r) = D(r) / dr$.
+The continuous distribution $\rho$ is a density, defined over interval $dr$.
+That is: it's really $\rho(r, dr(r))$, function of both the center and the width of the interval.
+A bijective map on just $r$ is not sufficient, $dr$ must be mapped as well.
+On one side we claim the state fully identified by $r$, on the other $dr$ (an different label) is required for the density and the bijective map.
+We can't have it both ways.
+
+In the continuous case, then, the appropriate label corresponds to a cell over $\mathbbm{R}$, not a point.
 This also makes physical sense, as we never deal with points per se, but widths that can be made arbitrarily small.
+An cell corresponds to $\mathbbm{R}^2$: the center and the width.
 With this in mind, we have the following definitions.
@@
 \begin{defn}\label{sdof}
-A \emph{degree of freedom} is a state variable identified by an infinitesimal cell of $\mathbbm{R}$.
+A \emph{degree of freedom} is a state variable identified by an infinitesimal cell of $\mathbbm{R}$.\footnote{For each label $i$ there is one and only infinitesimal cell.}
 \end{defn}
@@
 \begin{cor}\label{continuousLabels}
-Each degree of freedom is identified by the pair of labels $\langle q,k \rangle$. As $k\,dq$ is invariant under coordinate changes, $dk\wedge dq$ is invariant and $k$ is contravariant.
+Each degree of freedom is identified by the pair of labels $\langle q,k \rangle \in QT^*$. $k\,dq$ is invariant under coordinate changes, $dq\wedge dk$ is invariant and $k$ is contravariant.
 \end{cor}
 
+Let's start with the discrete case.
+Let $Q$ be a finite region of $\mathbbm{R}$.
+Divide the region in $N$ equal intervals of center $q$ and length $\Delta q$.\footnote{The definition could be mathematically more general.
+We use equally spaced intervals as it makes the derivation less cumbersome.} Let $K$ be a finite region of $\mathbbm{R}$.
+Divide the region in $M$ equal intervals of center $k$ and length $\Delta k$.
+Let $I$ be the label set identified by the cells in $Q$ with center $q(i)$ and width $k(i) \Delta q$.
+The configuration state will be given by $\mathbbm{c}=\sum\limits_{i \in \mathbbm{I}} \rho(q(i), k(i)) \Delta q \Delta k \mathbbm{s}(q(i), k(i))$, where $\rho(q(i), k(i)) = D(q(i), k(i)) / (\Delta q \Delta k)$.
+
+As we increase $N$ and $M$, the cardinality of the label set $I$ increases, $\Delta q$ and $\Delta k$ decrease and so does the width of the cells.
+Our definitions, though, do not change: for each $\langle q,k \rangle$ we have one and only one state.
+In the limit, we will cover every possible center $q$ and evert possible cell number $k$.
+Our configuration state becomes $\mathbbm{c}=\int\limits_{q \in Q \; k\in K} \rho(q, k) dq \wedge dk \mathbbm{s}(q(i), k(i))$.
+We can then arbitrarily increase the size of $Q$, and note that none of the definition change.
+
+If we apply a coordinate change $q'=q'(q)$, we change the labels but the state must remain the same, defined at the same point with the same width.
+$k' dq'$ must be then equal to $k dq$; $k'= k dq / dq'$ is contravariant, $dq' \wedge dk'= dq'/dq dq \wedge dk dq/dq' = dq \wedge dk$ is invariant.
+This means that $\rho$ is also invariant, which makes sense: the density depends only on the state (the cell), and not the coordinate chosen to represent them.
+As $k$ is contravariant, in the limit $K$ becomes the cotangent space of $Q$.
+
 \begin{prop}\label{continuousLabels}
-The cardinality of labels on a degree of freedom is given by $\omega = \hbar \, dq \wedge dk$, where $\hbar$ is a constant that defines the unit for labels.
+The cardinality of a label set $I$, subset of a degree of freedom, is given by $n(I)=\int \omega$, where $\omega = \hbar \, dq \wedge dk$ and $\hbar$ is a constant that defines the unit of state cardinality.
 \end{prop}
 
-Since $q$ and $k$ are independent labels, the number of labels defined over a degree of freedom is the product of the labels defined on $q$ and $k$ (\ref{labelsCombine}).
-The number of labels for an interval of $q$ or $k$ will be proportional to said interval, therefore we have $\hbar \, dq \wedge dk$, where the constant is given by the choice of units for labels.
-It is customary to combine $\hbar$ with $k$.
+As we deal with continuous labels, a label set is uncountable: given any range $\delta q$ and $\delta k$ there are an infinite number of possible labels $I$.
+Let $I_0$ be another set, defined on $\delta q_0$ and $\delta k_0$.
+In the discrete case, we have:
 
-\begin{defn}\label{sdof}
-We define the \emph{conjugate momentum} $p=\hbar k$.
+\begin{align*}
+\frac{n(I)}{n(I_0)} = \frac{\delta q \delta k / \Delta q \Delta k}{\delta q_0 \delta k_0 / \Delta q \Delta k} = \frac{\delta q \delta k}{\delta q_0 \delta k_0}
+\end{align*}
+
+That is: while the cardinality of each set diverges, the ratio between the two remains finite.\footnote{Such treatment is exactly equivalent to what one does in information theory to extend Shannon's entropy to the continuous case.} We can redefine our cardinality as the ratio of our set and our reference set $I_0$.
+That is $n(I)=\hbar \delta q \delta k$ where $\hbar$ is chosen such that $n(I_0)=\hbar \delta q_0 \delta k_0=1$.\footnote{The choice of angular momentum for $\hbar$ derives from the relationship between $p$ and $dq/dt$, which will be derived later.
+The scale in classical mechanics is arbitrary: we have no physical reason to choose a reference label set over another.
+In quantum mechanics, instead, we will have a preference: the set that corresponds to a single quantized system.} In general, the label set is not rectangular in $\langle q,k \rangle$ and we have $n(I)=\int \hbar dq \wedge dk$
+
+\begin{defn}\label{conjugateMomentum}
+We define the \emph{conjugate momentum} as $p=\hbar k$.
 \end{defn}
 
-\begin{cor}\label{continuousLabels}
-Each degree of freedom is identified by the pair of labels $\langle q,p \rangle$. $\theta_0 = p dq$ and $\omega = dq \wedge dp$ are invariant, $p$ is contravariant.
+As $\hbar dq \wedge dk$ is truly fundamental, it is customary to use $dq \wedge (\hbar dk) = dq \wedge dp$.
+This way the area formed by generalized coordinate and conjugate momentum corresponds to the cardinality of states defined on such area.
+
+\begin{cor}\label{continuousConjugateRelationships}
+Each degree of freedom is identified by the pair of labels $\langle q,p \rangle \in QT^*$.
+$\theta_0 = p dq$ and $\omega = dq \wedge dp$ are invariant, $p$ is contravariant.
+The configuration state over a degree of freedom is $\mathbbm{c}=\int \rho(q,p) dq \wedge dp \mathbbm{s}_{qp}$, where $\rho(q,p)\equiv D(q,p) / \omega$ is the distribution density for each label.
 \end{cor}
 
-\begin{prop}\label{continuousLabelDist}
-The configuration state over a degree of freedom is $\mathbbm{c}=\int D(q,p) \mathbbm{s}_{qp} = \int \rho(q,p) dq \wedge dp \mathbbm{s}_{qp}$, where $\rho(q,p)\equiv D(q,p) / \omega$ is the distribution density for each label.
-\end{prop}
+This restates all the previous findings in terms of conjugate momentum.
 
-This is simply the limit of \ref{classicalDistribution} for the continuous case. Note that, since $\mathbbm{c}$ and $\omega$ are invariant under coordinate transformations, so is $\rho$.
-
-In the language of differential geometry, we recognise the function $\rho$, the one-form $\theta_0$ and the two-form $\omega$.
+In the language of differential geometry, we recognise phase space as $Q$ extended by its cotangent space $T^*$, the function $\rho$, the one-form $\theta_0$ and the two-form $\omega$.
 They are truly fundamental objects as they are intimately linked to the way states are defined ($\rho$ is the distribution, $\theta_0 / \hbar$ is the cell width) and counted ($\omega$ is the cardinality of the labels) and they don't depend on coordinate choice.

### 2015-02-11 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Another language review

@@
 Special relativity is based on two principles (invariance of the speed of light and the principle of relativity), which lead to the Minkowskian nature of space-time but not to the equation of motion, which are a consistent reformulation of the non-relativistic ones.
-What we'll do in this work is see if we can re-organize all the known elements and equations in a more consistent and comprehensive way.
-The hope is that this gives us better insight on why the fundamental laws are what they are.
-Also, given that classical and quantum mechanics are formally equivalent, there is a chance that such work could be extended to quantum mechanics.
+What we'll do in this work is re-organize all the known elements and equations in a more consistent and comprehensive way, leading to better insight on why the fundamental concepts and laws are what they are.
+The derivation will also include classical analogues of strictly quantum concepts to facilitate a future extension to quantum mechanics.
 
@@
     Configuration state & The state of the whole system being described. \\ \hline
-    Elemental state (or just state) & The state of an infinitesimal part of the system. \\ \hline
+    Particle state (or simply state) & The state of an infinitesimal part (i.e. particle) of the system. \\ \hline
     \end{tabular}
@@
 The system is infinitely reducible: it can be thought as composed by two or more similar but smaller systems, each in its own configuration state, which also can be thought as composed by two or more, ad infinitum.
+We call \emph{particle} such an infinitesimal part.
 \end{assump1}
@@
 \begin{defn}\label{classicalPhaseSpace}
-Let $\mathbbm{S}$ be the set of all possible configuration states of the infinitesimal subdivision.
+Let $\mathbbm{S}$ be the set of all possible configuration states for a particle.
 We call this set \emph{phase space}.
-We call each $\mathbbm{s} \in \mathbbm{S}$ an \emph{elemental state}, or simply \emph{state}.
+We call each $\mathbbm{s} \in \mathbbm{S}$ an \emph{particle state}, or simply \emph{state}.
 \end{defn}
@@
 \begin{cor}\label{classicalDistribution}
-Each classical configuration state $\mathbbm{c} \in \mathbbm{C}$ is a distribution over states: $\mathbbm{c}=\sum\limits_{\mathbbm{s} \in \mathbbm{S}} D(\mathbbm{s}) \mathbbm{s}$, where $D:\mathbbm{S}\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}$.
+Each classical configuration state $\mathbbm{c} \in \mathbbm{C}$ is a distribution over particle states: $\mathbbm{c}=\sum\limits_{\mathbbm{s} \in \mathbbm{S}} D(\mathbbm{s}) \mathbbm{s}$, where $D:\mathbbm{S}\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}$.
 The distribution can be visualized as a histogram over the states in phase space.
@@
 
-Under the classical assumption, we can then limit ourselves to the study of the infinitesimal elements, their states and their properties without loss of generality.
+Under the classical assumption, we can then limit ourselves to study the particles of the system, their states and their properties without loss of generality.
 To help identify states, we introduce the following concepts.
@@
 \begin{defn}\label{label}
-We call a \emph{label} a set of states $i\subset\mathbbm{S}$; a \emph{set of labels} a collection of disjoint labels $I | \forall i_1,i_2\in I, i_1\bigcap i_2 = \emptyset$; a \emph{state variable} a set of labels $X$ that covers all of phase space: $\bigcup\limits_{i \in X}i=\mathbbm{S}$.
+We call a \emph{label} a set of states $i\subset\mathbbm{S}$; a \emph{set of labels} a collection of disjoint labels $I | \forall i_1,i_2\in I, i_1\bigcap i_2 = \emptyset$; a \emph{state variable} a set of labels $\mathbbm{I}$ that covers all of phase space: $\bigcup\limits_{i \in \mathbbm{I}}i=\mathbbm{S}$.
 Therefore a state belongs to one and only one label of a state variable.
 \end{defn}
-
-TODO: NOTATION maybe $X$ should be $\mathcal{I}$?
 
@@
 This also makes physical sense, as we never deal with points per se, but widths that can be made arbitrarily small.
-An cell corresponds to $\mathbbm{R}^2$: the center and the width.
-With this in mind, we have the following definitions.
+An cell corresponds to $\mathbbm{R}^2$: the center and the width.\footnote{In other words: the particles of the system are not point-like, but infinitesimal cell-like.
+Not only this leads to a more direct understanding of phase-space, it is also more consistent with general relativity (the mass is spread across a small region, not forming a singularity) and quantum mechanics (position is really the central value of a small distribution).} With this in mind, we have the following definitions.
 
 \begin{defn}\label{sdof}
-A \emph{degree of freedom} is a state variable identified by an infinitesimal cell of $\mathbbm{R}$.\footnote{For each label $i$ there is one and only infinitesimal cell.}
+A \emph{degree of freedom} is a state variable identified by an infinitesimal cell of a manifold $\mathbbm{Q}$.\footnote{For each label $i$ there is one and only infinitesimal cell.}
 \end{defn}
@@
 \begin{cor}\label{continuousLabels}
-Each degree of freedom is identified by the pair of labels $\langle q,k \rangle \in QT^*$. $k\,dq$ is invariant under coordinate changes, $dq\wedge dk$ is invariant and $k$ is contravariant.
+Each degree of freedom is identified by the pair of labels $\langle q,k \rangle \in \mathbf{T}^*\mathbbm{Q}$.
+$k\,dq$ is invariant under coordinate changes, $dq\wedge dk$ is invariant and $k$ is contravariant.
 \end{cor}
@@
 Let's start with the discrete case.
-Let $Q$ be a finite region of $\mathbbm{R}$.
+Let $Q$ be a finite region of $\mathbbm{Q}$.
 Divide the region in $N$ equal intervals of center $q$ and length $\Delta q$.\footnote{The definition could be mathematically more general.
@@
 Our configuration state becomes $\mathbbm{c}=\int\limits_{q \in Q \; k\in K} \rho(q, k) dq \wedge dk \mathbbm{s}(q(i), k(i))$.
-We can then arbitrarily increase the size of $Q$, and note that none of the definition change.
+We can repeat the process increasing or changing the region covered by $Q$ and $K$, until we cover all of $\mathbbm{Q}$ and $\mathbbm{R}$.
 
@@
 As $k$ is contravariant, in the limit $K$ becomes the cotangent space of $Q$.
+Phase space is the cotangent bundle $\mathbf{T}^*\mathbbm{Q}$.
 
@@
 \begin{cor}\label{continuousConjugateRelationships}
-Each degree of freedom is identified by the pair of labels $\langle q,p \rangle \in QT^*$.
+Each degree of freedom is identified by the pair of labels $\langle q,p \rangle \in \mathbf{T}^*\mathbbm{Q}$.
 $\theta_0 = p dq$ and $\omega = dq \wedge dp$ are invariant, $p$ is contravariant.
@@
 
-In the language of differential geometry, we recognise phase space as $Q$ extended by its cotangent space $T^*$, the function $\rho$, the one-form $\theta_0$ and the two-form $\omega$.
+In the language of differential geometry, we recognise phase space as the cotangent bundle $\mathbf{T}^*\mathbbm{Q}$, the function $\rho$, the one-form $\theta_0$ and the two-form $\omega$.
 They are truly fundamental objects as they are intimately linked to the way states are defined ($\rho$ is the distribution, $\theta_0 / \hbar$ is the cell width) and counted ($\omega$ is the cardinality of the labels) and they don't depend on coordinate choice.
@@
 
-We can now derive the equation of motions for the state variables under deterministic and reversible evolution.
+Now that we have properly defined continuous labels and the cardinality of their sets, we will extend bijective maps for the continuous case.
+As these must preserve the cardinality of label sets, they are constrained in their form, which coincides with Hamilton's equation.
+In this section we'll study a single degree of freedom, then extend in the next two section to multiple degrees of freedom and to the time dependent case.
 
@@
 A bijective map conserves the cardinality of labels, see \ref{labelsCount}, therefore $\omega$ must be conserved.
-The map must be continuous in q, or it would split some cells into two parts: a cell would not be mapped to one and only one other cell, the mapping would not be bijective.
+The map must be continuous in q: suppose it isn't, it would split some cells into two parts, a cell would not be mapped to one and only one other cell, the mapping would not be bijective.
 The reverse mapping must be continuous in q as well, or the inverse would not map to one and only one cell.

### 2015-02-11 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Another pass at the language

@@
 \newtheorem*{assump1}{Infinite reducibility (or classical) assumption}
-\newtheorem*{assump2}{Determinism and Reversibility assumption}
-\newtheorem*{assump3}{Kinematic equivalence assumption}
+\newtheorem*{assump2}{Determinism and Reversibility (or Hamiltonian) assumption}
+\newtheorem*{assump3}{Kinematic equivalence (or Lagrangian) assumption}
 
@@
 
-The above outline should help the reader maintain the big picture, as each section will go through the many details and definitions.
+We will use concepts from different disciplines, such as set theory, differential geometry, relativity, Hamiltonian and Lagrangian mechanics.
+This means we will find an interesting connection between them, but one may require familiarity with their basics to appreciate it.
 We'll keep names and notation as consistent as possible to current use, but this may lead to some non sequitur as it will not be immediately clear why the different definition will be at the end equivalent to the standard one.
@@
 
-We give here a brief a general conceptual overview, hoping it will help guide the reader through mathematical details.
+We give here a brief a general conceptual overview, hoping it will help guide the reader through the mathematical details.
 
@@
 position, momentum, temperature, pressure).
-We introduce the infinite reducibility assumption (or classical assumption): each state is divisible into the states of its sub-systems.
-It is then sufficient to describe the evolution of infinitesimal subsystem within their state space (corresponding to point particles in phase space).
+We introduce the infinite reducibility assumption (or classical assumption): each state is divisible into the states of its parts.
+It is then sufficient to describe the evolution of infinitesimal parts (i.e.
+particles) within their state space (i.e.
+phase space).
 
-We introduce the deterministic and reversible evolution assumption: to each state corresponds one and only one future (or past) state.
+We introduce the deterministic and reversible evolution assumption (or Hamiltonian assumption): to each state corresponds one and only one future (or past) state.
 The cardinality of a set of states is therefore conserved under evolution.
@@
 
-We introduce the assumption of kinematic equivalence: studying trajectories is equivalent to studying states.
+We introduce the assumption of kinematic equivalence (or Lagrangian assumption): studying trajectories is equivalent to studying states.
 We show that there must be link between the space-time metric $g$ and the phase-space metric $\omega$, as both quantities must be conserved by passive coordinate transformations.
@@
 
-The first two sections are dedicated to properly define states.
+This and the next section are dedicated to properly define states.
 In particular, we will need to introduce more precise terminology to be able to make two crucial distinctions.
@@
 \end{center}
+When talking about particles of a system we must always remember that they are the result of a limit.
+Therefore we will start with discrete definitions and do the limit, and see how some properties will extend from discrete case to continuous.
+
 The second is, when referring to a physical quantity, between the space of all possible values, a particular value or a set of possible values.
@@
 \end{center}
-When talking about infinitesimal components and continuous state variables, we must always remember that they are the result of a limit.
-Therefore we will start with discrete definitions and do the limit, and see how some properties will extend from discrete case to continuous.
+We will use set theory to define labels (e.g. "position = $5m$" will correspond to the set of all states located at the given position).
 These are practically the only new terms introduced by this work, but they are fundamental as most of the later derivation will be based on \emph{counting the labels} (i.e.
@@
 \begin{defn}\label{statedef}
-Suppose we have a physical system to study.
-We define the set $\mathbbm{C}$ of all physically distinguishable configurations for that system.
-Each element $\mathbbm{c}$ we call \emph{configuration state}.
+Fix a physical system to study. We define the set $\mathbbm{C}$ of all physically distinguishable configurations for that system. Each element $\mathbbm{c}$ we call \emph{configuration state}.
 \end{defn}
@@
 We call this set \emph{phase space}.
-We call each $\mathbbm{s} \in \mathbbm{S}$ an \emph{particle state}, or simply \emph{state}.
+We call each $\mathbbm{s} \in \mathbbm{S}$ a \emph{particle state}, or simply \emph{state}.
 \end{defn}
@@
 
+\begin{defn}\label{discreteCardinality}
+Let $I$ be a finite, countable set of labels. We define the \emph{cardinality} $n: I \rightarrow \mathbbm{N}$ as the number of labels in the set.
+\end{defn}
+
 \begin{defn}\label{labelsCombine}
@@
 We can define the \emph{combined set}, $\langle I_1, I_2 \rangle$, whose labels consist of all the non-empty intersections of one label of $I_1$ and one of $I_2$.
-If all intersections are non-empty, $I_1$ and $I_2$ are said to be \emph{independent}, and we have $n(\langle I_1, I_2 \rangle)=n(I_1)n(I_2)$ where $n: I \rightarrow \mathbbm{R}$ gives the number of labels of each set.
+If all intersections are non-empty, $I_1$ and $I_2$ are said to be \emph{independent}, and we have $n(\langle I_1, I_2 \rangle)=n(I_1)n(I_2)$.
 \end{defn}
@@
 \begin{cor}\label{detrevDist}
-The evolution of a classical configuration state $\mathbbm{c}=\sum D(\mathbbm{s}) \mathbbm{s}$ under a bijective map is given by $\mathbbm{c'}=\sum D'(\mathbbm{s}) \mathbbm{s}=\sum D(f^{-1}(s)) \mathbbm{s}$.
+The evolution of a classical configuration state $\mathbbm{c}=\sum D(\mathbbm{s}) \mathbbm{s}$ under a bijective map is given by $\mathbbm{c'}=\sum D'(\mathbbm{s}) \mathbbm{s}=\sum D(f^{-1}(\mathbbm{s})) \mathbbm{s}$.
 The evolution of the fraction of the system in a label $D(i)=\sum\limits_{\mathbbm{S} \in i} D(\mathbbm{s})$ is given by $D'(i)=D(f^{-1}(i))$.
@@
 Given a set of labels $I$, the image $f(I)$ is also a set of labels containing the same number of labels.
-Given a state variable $X$, the image $f(X)$ is also a state variable.
+Given a state variable $\mathbbm{I}$, the image $f(\mathbbm{I})$ is also a state variable.
 Given two independent sets of labels $I_1$ and $I_2$, the images $f(I_1)$ and $f(I_2)$ are also independent.
@@
 
-\section{Continuous labels}
+\section{Numeric labels}
 
-The above definitions readily apply for labels identified by integers.\footnote{For each label $i$ there is one and only one integer $z$.} Let $z \in \mathbbm{Z}$, we have $\mathbbm{c}=\sum\limits_{z \in \mathbbm{Z}} D(z) \mathbbm{s}(z)$ and the deterministic map becomes $f:\mathbbm{Z} \leftrightarrow \mathbbm{Z}$.
+We now focus on labels that can be identified by numbers, where we have a bijective map between the label and a number in a set.
+The above definitions readily apply for labels identified by integers.
+Let $z \in \mathbbm{Z}$, we have $\mathbbm{c}=\sum\limits_{z \in \mathbbm{Z}} D(z) \mathbbm{s}(z)$ and the deterministic map becomes $f:\mathbbm{Z} \leftrightarrow \mathbbm{Z}$.
+
 For the continuous case, one may simply expect to replace $z \in \mathbbm{Z}$ with $r \in \mathbbm{R}$, but this does not work.
@@
 A bijective map on just $r$ is not sufficient, $dr$ must be mapped as well.
-On one side we claim the state fully identified by $r$, on the other $dr$ (an different label) is required for the density and the bijective map.
+On one side we claim the state fully identified by $r$, on the other $dr$ (a different label) is required for the density and the bijective map.
 We can't have it both ways.
 
-In the continuous case, then, the appropriate label corresponds to a cell over $\mathbbm{R}$, not a point.
+In the continuous case, then, the appropriate label corresponds to a cell, not a point.
 This also makes physical sense, as we never deal with points per se, but widths that can be made arbitrarily small.
-An cell corresponds to $\mathbbm{R}^2$: the center and the width.\footnote{In other words: the particles of the system are not point-like, but infinitesimal cell-like.
+An cell of $\mathbbm{R}$ corresponds to $\mathbbm{R}^2$: the center and the width.\footnote{In other words: the particles of the system are not point-like, but infinitesimal cell-like.
 Not only this leads to a more direct understanding of phase-space, it is also more consistent with general relativity (the mass is spread across a small region, not forming a singularity) and quantum mechanics (position is really the central value of a small distribution).} With this in mind, we have the following definitions.
@@
 \begin{defn}\label{sdof}
-A \emph{degree of freedom} is a state variable identified by an infinitesimal cell of a manifold $\mathbbm{Q}$.\footnote{For each label $i$ there is one and only infinitesimal cell.}
+A \emph{degree of freedom} is a state variable identified by an infinitesimal cell of a one dimensional manifold $\mathbbm{Q}$.\footnote{For each label $i$ there is one and only infinitesimal cell.}
 \end{defn}
@@
 Let $I$ be the label set identified by the cells in $Q$ with center $q(i)$ and width $k(i) \Delta q$.
-The configuration state will be given by $\mathbbm{c}=\sum\limits_{i \in \mathbbm{I}} \rho(q(i), k(i)) \Delta q \Delta k \mathbbm{s}(q(i), k(i))$, where $\rho(q(i), k(i)) = D(q(i), k(i)) / (\Delta q \Delta k)$.
+The configuration state will be given by $\mathbbm{c}=\sum\limits_{i \in I} \rho(q(i), k(i)) \Delta q \Delta k \, \mathbbm{s}(q(i), k(i))$, where $\rho(q(i), k(i)) = D(q(i), k(i)) / (\Delta q \Delta k)$.
 
 As we increase $N$ and $M$, the cardinality of the label set $I$ increases, $\Delta q$ and $\Delta k$ decrease and so does the width of the cells.
-Our definitions, though, do not change: for each $\langle q,k \rangle$ we have one and only one state.
-In the limit, we will cover every possible center $q$ and evert possible cell number $k$.
-Our configuration state becomes $\mathbbm{c}=\int\limits_{q \in Q \; k\in K} \rho(q, k) dq \wedge dk \mathbbm{s}(q(i), k(i))$.
+Our definitions, though, do not change: for each $\langle q,k \rangle$ we have one and only one cell, one state.
+In the limit, we will cover every possible center $q$ and every possible cell number $k$.
+Our configuration state becomes $\mathbbm{c}=\int\limits_{q \in Q \; k\in K} \rho(q, k) dq \wedge dk \, \mathbbm{s}(q(i), k(i))$.
 We can repeat the process increasing or changing the region covered by $Q$ and $K$, until we cover all of $\mathbbm{Q}$ and $\mathbbm{R}$.
@@
 If we apply a coordinate change $q'=q'(q)$, we change the labels but the state must remain the same, defined at the same point with the same width.
-$k' dq'$ must be then equal to $k dq$; $k'= k dq / dq'$ is contravariant, $dq' \wedge dk'= dq'/dq dq \wedge dk dq/dq' = dq \wedge dk$ is invariant.
+$k' dq'$ must be then equal to $k dq$; $k'= k dq / dq'$ is contravariant, $dq' \wedge dk'= dq'/dq \, dq \wedge dk \, dq/dq' = dq \wedge dk$ is invariant.
 This means that $\rho$ is also invariant, which makes sense: the density depends only on the state (the cell), and not the coordinate chosen to represent them.
@@
 
-\begin{prop}\label{continuousLabels}
-The cardinality of a label set $I$, subset of a degree of freedom, is given by $n(I)=\int \omega$, where $\omega = \hbar \, dq \wedge dk$ and $\hbar$ is a constant that defines the unit of state cardinality.
+\begin{prop}\label{relativeCardinality}
+Let $I$ be a closed dense set of labels, subset of a degree of freedom.
+We define \emph{relative cardinality} $n: I \rightarrow \mathbbm{R}$ the ratio between the number of labels in $I$ and the ones of a reference set $I_0$.
+We find $n(I)=\int \omega$, where $\omega = \hbar \, dq \wedge dk$ and $\hbar$ is the constant that determines the unit (i.e.
+chosen so $n(I_0)=1$).
 \end{prop}
@@
 
-That is: while the cardinality of each set diverges, the ratio between the two remains finite.\footnote{Such treatment is exactly equivalent to what one does in information theory to extend Shannon's entropy to the continuous case.} We can redefine our cardinality as the ratio of our set and our reference set $I_0$.
+That is: while the cardinality of each set diverges, the ratio between the two remains finite.\footnote{Such treatment is exactly equivalent to what one does in information theory to extend Shannon's entropy to the continuous case.} For continuous state variables we can define the relative cardinality as the ratio of a label set and our reference set $I_0$.
 That is $n(I)=\hbar \delta q \delta k$ where $\hbar$ is chosen such that $n(I_0)=\hbar \delta q_0 \delta k_0=1$.\footnote{The choice of angular momentum for $\hbar$ derives from the relationship between $p$ and $dq/dt$, which will be derived later.
-The scale in classical mechanics is arbitrary: we have no physical reason to choose a reference label set over another.
-In quantum mechanics, instead, we will have a preference: the set that corresponds to a single quantized system.} In general, the label set is not rectangular in $\langle q,k \rangle$ and we have $n(I)=\int \hbar dq \wedge dk$
+The value for $\hbar$ in classical mechanics is arbitrary: we have no physical reason to choose a reference label set over another.
+In quantum mechanics, instead, we will have a preference: the set that corresponds to a single quantized system.
+We use this arbitrariness to set $\hbar$ to the value to the known constant.} In general, the label set is not rectangular in $\langle q,k \rangle$ and we have $n(I)=\int \hbar dq \wedge dk$
 
@@
 
-As $\hbar dq \wedge dk$ is truly fundamental, it is customary to use $dq \wedge (\hbar dk) = dq \wedge dp$.
+As $\hbar dq \wedge dk$ is truly fundamental, it is customary to group $\hbar$ with $k$: $dq \wedge (\hbar dk) = dq \wedge dp$.
 This way the area formed by generalized coordinate and conjugate momentum corresponds to the cardinality of states defined on such area.

### 2015-02-11 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Language review up to Kinematics

@@
 
-Now that we have properly defined continuous labels and the cardinality of their sets, we will extend bijective maps for the continuous case.
-As these must preserve the cardinality of label sets, they are constrained in their form, which coincides with Hamilton's equation.
+Now that we have properly defined continuous labels and the relative cardinality of their sets, we will extend bijective maps to the continuous case.
+As these must preserve the cardinality of label sets, they are constrained in their form.
+The use of a bijective map for infinitesimal time evolution will lead us to Hamilton's equation.
 In this section we'll study a single degree of freedom, then extend in the next two section to multiple degrees of freedom and to the time dependent case.
@@
 \begin{defn}\label{canonical}
-We define \emph{canonical transformation} a bijective map on one (or more) degree of freedoms.
+We define \emph{canonical transformation} a bijective map on one (or more) degree of freedom.
 \end{defn}
@@
 
-A bijective map conserves the cardinality of labels, see \ref{labelsCount}, therefore $\omega$ must be conserved.
+As we saw \ref{labelsCount}, a bijective map conserves the cardinality of labels.
+In the continuous case, it will conserve relative cardinality and therefore $\omega$.
 The map must be continuous in q: suppose it isn't, it would split some cells into two parts, a cell would not be mapped to one and only one other cell, the mapping would not be bijective.
@@
 \end{align*}
-We can re-express $dp'$ in terms of $dq$ and $dp$, as the conservation of $\omega$ means the map is non-degenerate, find that all partial derivatives are well defined, and therefore the mapping is continuous in p as well.
+We can re-express $dp'$ in terms of $dq$ and $dp$, as the conservation of $\omega$ means the map is non-degenerate.
+All partial derivatives are well defined, and therefore the mapping is continuous in p as well.
 
 \begin{cor}\label{sdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by two conjugate variables. Let
+Let $v$ and $w$ be two vectors defined on the tangent space of the phase space for one degree of freedom. Let
 \begin{align*}
@@
 \begin{prop}\label{mdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by two independent degrees of freedom.
+Let $v$ and $w$ be two vectors defined on the tangent space of the phase space for two independent degrees of freedom.
 Let $a$ and $b$ be indexes for the state variables $\xi^a\equiv \{q^i, p_i\}$.
@@
 The metric generalizes \ref{sdofInvariant} to give us the cardinality of labels defined on the area given by two arbitrary directions in phase space.
-For an infinitesimal region, this corresponds to $dq^1 \wedge dp^1 + dq^2 \wedge dp^2$, the sum of the projections on the independent planes.
+For an infinitesimal region, this corresponds to $dq^1 \wedge dp_1 + dq^2 \wedge dp_2$, the sum of the projections on the independent planes.
 Moreover, volume in phase space corresponds to the cardinality of the combined labels (i.e.
-the states), and is therefore conserved: this is Liouville's theorem.
+the states), and is therefore conserved: this is Liouville's theorem for Hamiltonian mechanics.
 
@@
 Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call standard and anti-states is also conventional.
-What is physical and not conventional, though, is that standard and anti-states cannot be connected by deterministic and reversible evolution.
+What is physical and not conventional, though, is that standard and anti-states cannot be connected by deterministic and reversible evolution.\footnote{This represents a classical analogue for quantum anti-particle states.}
 
@@
 Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by the temporal degree of freedom and one standard degree of freedom.
-Let $a$ and $b$ be indexes for the state variables $\xi^a\equiv\{t, E, q, p\}$.
+Let $a$ and $b$ be indexes for the state variables $\xi^a\equiv\{t, q, E, p\}$.
 Let
@@
 So they are not necessarily orthogonal in the extended phase space.
-Looking back at \ref{discreteLabelDef}, cells need to be defined on the plane where $\langle q, p \rangle$ (maximally) change: this is not the plane of constant $\langle t, E \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle q, p \rangle$ where $dt \wedge dE$ is defined.
+States are defined on the plane where $\langle q, p \rangle$ (maximally) change: this is not the plane of constant $\langle t, E \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle q, p \rangle$ where $dt \wedge dE$ is defined.
 On that plane we can properly count states and define our invariant.
@@
 and the vector product within.
-The metric, with a space-like convention, still gives us the cardinality of labels within a degree of freedom.\footnote{Adjusted to avoid double counting.}
+The metric, with a space-like convention, still gives us the cardinality of labels within a degree of freedom, adjusting \ref{sdofInvariant} to avoid "double counting".
 
@@
 
-It should not be a surprise that the equations do not mention the speed of light $c$.
-In fact, nothing says that all $q^i$ represent space or that the laws of motion are invariant in all inertial frames.
-The only requirement we have is that the areas of each degree of freedom represent the same cardinality for labels.
-
 \begin{prop}\label{tdofConstrain}
@@
 By convention, we can set $\mathcal{H}=0$ without loss of generality as changing $\mathcal{H}$ by a constant does not change the equation of motion.
-This reduces extended phase space to $\mathbb{R}^{2*N + 1}$, the state variables plus time.
+This reduces extended phase space to $2N + 1$ components, the state variables plus time.
 

### 2015-02-12 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Finished language review

@@
 \newtheorem*{assump2}{Determinism and Reversibility (or Hamiltonian) assumption}
-\newtheorem*{assump3}{Kinematic equivalence (or Lagrangian) assumption}
+\newtheorem*{assump3}{Kinematic equivalence assumption}
 
@@
 
-We introduce the assumption of kinematic equivalence (or Lagrangian assumption): studying trajectories is equivalent to studying states.
+We introduce the assumption of kinematic equivalence: studying trajectories is equivalent to studying states.
 We show that there must be link between the space-time metric $g$ and the phase-space metric $\omega$, as both quantities must be conserved by passive coordinate transformations.
@@
 Space-time is locally Minkowskian Rimmanian manifold.
-That is, there exists a metric $g_ij$ that at any point can be expressed, with a suitable choice of coordinate, as $dx^\alpha g_{\alpha \beta}|_P dx^\beta=dx^\alpha\eta_{\alpha \beta}dx^\beta=(dx^i)^2 - (dx^0)^2=(dx^i)^2 - c^2dt^2$, where $c=n(dq)/n(dt)$ is the ration between the density of states in space and time.
+That is, there exists a metric $g$ that at any point can be expressed, with a suitable choice of coordinate, as $dx^\alpha g_{\alpha \beta}(P) dx^\beta=dx^\alpha\eta_{\alpha \beta}dx^\beta=(dx^i)^2 - (dx^0)^2=(dx^i)^2 - c^2dt^2$, where $c=n(dq)/n(dt)$ is the ratio between the density of states in space and time.
 \end{prop}
 
-As states are defined on intervals, a metric $g$ must be defined on $Q$.
+As states are defined on intervals, a metric $g$ must be defined on $\mathcal{M}$.
 Such a metric must be consistent with $\omega$, as both must be invariant under coordinate transformations.
@@
 For each spatial d.o.f.
-fix $dp_i=\lambda dq^i$, we have $dq^i \wedge dp_i = \lambda (dx^i)^2$ where $\lambda$ converts from units of length squared to cardinality of labels.
+fix $dp_i=\lambda dq^i$, we have $dq^i \wedge dp_i = \lambda (dx^i)^2$ where $\lambda$ converts from length squared to the label cardinality contained in the area.
 For the temporal d.o.f fix $dE = \lambda c^2 dt$, we have $dt \wedge dE = \lambda c^2 dt^2$.
-That is: the area in time squared is converted to an area in length squared that have the same density of states (as per definition of $c$) and then to cardinality of labels.
+That is: the area in time squared is converted to an area in length squared that have the same density of states (as per definition of $c$) and then to label cardinality.
 The phase space invariant is $\omega = \lambda [(dq^i)^2 - c^2 dt^2]$.
@@
 
-The equation of motion \ref{tdofHam} can be at most second order in $q^i$, they can at most be second order in $x^\alpha=\{ct, q^i\}$ so only position and velocity can be candidates for the initial conditions.
+As the equation of motion \ref{tdofHam} can be at most second order in $\{t, q^i\}$, they can at most be second order in $x^\alpha=\{ct, q^i\}$ so only position and velocity can be candidates for the initial conditions.
 Fixing time, phase space is $\mathbbm{R}^{2n}$, too big to be covered by position only, but just right to be covered by both.
+
+\begin{defn}\label{inertialMass}
+Let $dp_\alpha$ an infinitesimal interval of conjugate momentum and $d(g_{\alpha \beta}u^\beta)$ an infinitesimal interval of the covariant components of the four velocity.
+The ratio $m=n(dp_\alpha)/n(d(g_{\alpha \beta}u^\beta))$ between their relative label cardinality is called \emph{inertial mass}.
+\end{defn}
 
@@
 Let $x^\alpha=\{ct, q^i\}$ and $p_\alpha=\{-E/c, p_i\}$.
-Then $p_\alpha= m g_{\alpha \beta}u^\beta + \hat{p}_\alpha(x^\gamma)$, where $m=n(dp_\alpha)/n(d(g_{\alpha \beta}u^\beta))$ is the ration between the density of states in conjugate momentum and in contravariant velocity, and is called \emph{inertial mass}.
+Then $p_\alpha= m g_{\alpha \beta}u^\beta + \hat{p}_\alpha(x^\gamma)$, where $\hat{p}_\alpha : \mathcal{M} \rightarrow \mathbbm{R}$ is a function defined on the extended configuration manifold.
 \end{prop}
@@
 Consider the expression $m du_\alpha dx^\alpha$: this gives us the density of labels in position and velocity converted to the conjugate variables.
-In other world, we should have $\omega=m du_\alpha dx^\alpha$.
-This means:
+That is: $\omega=m du_\alpha dx^\alpha$.
+Combining the two:
 \begin{align*}
@@
 \begin{prop}\label{kineticHamiltonian}
-The extended Hamiltonian is $\mathcal{H}=\frac{1}{2m}(p_\alpha-\hat{p}_\alpha(x))g^{\alpha\beta}(p_\beta-\hat{p}_\beta(x))+\hat{\mathcal{H}}(x)$
+The extended Hamiltonian is $\mathcal{H}=\frac{1}{2m}(p_\alpha-\hat{p}_\alpha(x))g^{\alpha\beta}(p_\beta-\hat{p}_\beta(x))+\hat{\mathcal{H}}(x)$, where $\hat{\mathcal{H}}_\alpha : \mathcal{M} \rightarrow \mathbbm{R}$ is a function defined on the extended configuration manifold.
 \end{prop}
@@
 
-\begin{prop}\label{kineticLaw}
-Let $\hat{p}_\alpha = q A_\alpha$, $F_{\alpha \beta} \equiv \partial_\alpha A_\beta - \partial_\beta A_\alpha$ and $\hat{H} = 0$, the equations of motion are $m \nabla_s u^\alpha = g^{\alpha\beta} F_{\beta \gamma} q u^\gamma$.
+Now that we have found the general form of the Hamiltonian, we show that this is compatible with the established fundamental classical theories.
+
+\begin{prop}\label{relativisticEM}
+Let $\hat{p}_\alpha = q A_\alpha$, $F_{\alpha \beta} \equiv \partial_\alpha A_\beta - \partial_\beta A_\alpha$ and $\hat{H} = 0$, the equations of motion are $m \nabla_u u^\alpha = g^{\alpha\beta} F_{\beta \gamma} q u^\gamma$.
+These are the relativistic equations for a charged particle.
 \end{prop}
@@
 
-This is the relativistic equation for a charged particle.
+This is indeed compatible with general relativity and classical electro-magnetism.
+
+\begin{prop}\label{newtonianGravitation}
+Assume $\mathcal{M}$ flat and time independent motion, the Hamiltonian simplifies to $H=\frac{1}{2m}(p_i-\hat{p}_\alpha(x))^2)+\hat{H}(x)$.
+Let $\hat{p}_i = 0$ and $\hat{H} = mV$, the equations of motion are $m d_t v^i = - m \partial_i V$, where $v^i=dx^1/dt$.
+These are equations for a particle under a Netwonian gravitational potential.
+\end{prop}
+
+Given flat space-time and time independent motion, we can use $t$ as a parameter for the motion in $\mathbbm{Q}$.
+Repeating \ref{kineticMomentum} and \ref{initialConditions} for the non-relativistic case leads to the above Hamiltonian.
+Similar to \ref{relativisticEM}, use \ref{mdofHam} with the newly found Hamiltonian and find the equation of motion.
+
+\begin{defn}\label{lagrangian}
+Let $\mathcal{H}$ be an extended hamiltonian defined on $\mathbf{T}^*\mathcal{M}$.
+Under \ref{kinematicAssumption}, we can define the Legendre transform $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$ which we call \emph{extended Lagrangian}.
+For the time invariant case, we define the \emph{Lagrangian} $L=u^\alpha p_\alpha - H$.
+\end{defn}
+
+The Legendre transform can be defined only if $\mathcal{H}$ is convex in $p_\alpha$.
+The Hamiltonian found at \ref{kineticHamiltonian} is convex, so the Lagrangian can always be defined.
+While the kinetic equivalence assumption is sufficient, it's not necessary.
+What is necessary is \ref{initialConditions}: as long as position and velocity are enough to determine the state, $u^\alpha=f(q^\alpha, p_\alpha)$ must be monotonic in $p_\alpha$, which means $\mathcal{H}$ is convex in $p_\alpha$.
+What happens in that case is that the space-time trajectories may become more or less dense, and $n(du_\alpha)$ will be a function of $x^\alpha$ as well.
+So while Hamiltonian mechanics coincides with deterministic and reversible evolution, Lagrangian mechanics coincides with initial conditions given by position and velocity.
 

### 2015-02-20 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Another pass and conclusions

@@
 \newtheorem*{assump1}{Infinite reducibility (or classical) assumption}
-\newtheorem*{assump2}{Determinism and Reversibility (or Hamiltonian) assumption}
+\newtheorem*{assump2}{Determinism and Reversibility assumption}
 \newtheorem*{assump3}{Kinematic equivalence assumption}
@@
 The core idea is that deterministic and reversible systems preserve the cardinality of a set of states, which puts considerable constraints on the equations of motion.
-Many familiar fundamental concepts and constants (such as mass, the speed of light, the Plank constant, the vector potential) are introduced and defined in a novel way, thus providing different insight into them.
+This perspective links different concepts from different branches of math and physics (cardinality of a set, cotangent bundle for phase space, locally Minkowskian space-time manifold, inertial mass) providing new insights.
 The derivation strives to use definition and mathematical concepts that would allow a future extension to quantum mechanics.\end{abstract}
@@
 
-We will use concepts from different disciplines, such as set theory, differential geometry, relativity, Hamiltonian and Lagrangian mechanics.
-This means we will find an interesting connection between them, but one may require familiarity with their basics to appreciate it.
-We'll keep names and notation as consistent as possible to current use, but this may lead to some non sequitur as it will not be immediately clear why the different definition will be at the end equivalent to the standard one.
+We will use concepts from different disciplines, such as set theory, differential geometry, relativity, Hamiltonian and Lagrangian mechanics, and will find interesting connections between them.
+We'll keep names and notation as consistent as possible to current use across the different disciplines, but this may lead to some non sequitur as it will not be immediately clear why the different definition will be at the end equivalent to the standard one.
 This is still preferable than introducing a whole new set of names one has to become familiar.
 
-While we try to keep a precise language, no mathematical breakthrough should be expected: the goal, after all, is to derive the \emph{known} framework from a set of \emph{simple} definitions in the most \emph{obvious} way possible.
+No mathematical breakthrough should be expected: the goal, after all, is to derive the \emph{known} framework from a set of \emph{simple} definitions in the most \emph{obvious} way possible.
 No proof is longer than a couple of paragraphs, so the word \emph{theorem} is avoided in favor of \emph{proposition} and \emph{corollary}.
@@
 
-We introduce the deterministic and reversible evolution assumption (or Hamiltonian assumption): to each state corresponds one and only one future (or past) state.
+We introduce the deterministic and reversible evolution assumption: to each state corresponds one and only one future (or past) state.
 The cardinality of a set of states is therefore conserved under evolution.
@@
 We introduce the assumption of kinematic equivalence: studying trajectories is equivalent to studying states.
-We show that there must be link between the space-time metric $g$ and the phase-space metric $\omega$, as both quantities must be conserved by passive coordinate transformations.
+We show that there must be link between the space-time metric $g$ and the phase-space metric $\omega$, as both quantities must be conserved under passive coordinate transformations.
 That link constrains the space-time metric to be locally Minkowskian.
-A transformation between state variable $(q, p)$ and kinematic variables $(x, \dot{x})$.
+A transformation between state variables $(q, p)$ and kinematic variables $(x, \dot{x})$ must exist.
 To be invertible, the relationship between velocity and conjugate momentum has to be monotonic, leading to a concave Hamiltonian which allows a Legendre transformation leading to the Lagrangian.
-We can constrain the Hamiltonian further, and show that the most general equation of motion under the three condition is the one of a geodesic modified by the force given by a vector potential, such as the one of a relativistic charged particle, and a scalar potential, such as the one of newtonian gravitation.
+We can constrain the Hamiltonian further, and show that the most general equation of motion under the three assumptions is the one of a geodesic modified by the force given by a vector potential (such as the one of a relativistic charged particle) and a scalar potential (such as the one of newtonian gravitation).
 
@@
 When talking about particles of a system we must always remember that they are the result of a limit.
-Therefore we will start with discrete definitions and do the limit, and see how some properties will extend from discrete case to continuous.
+Therefore we will start with discrete definitions, and see how some properties will extend from discrete case to continuous.
 
-The second is, when referring to a physical quantity, between the space of all possible values, a particular value or a set of possible values.
+The second distinctions is, when referring to a physical quantity, between the space of all possible values, a particular value or a set of possible values.
 \begin{center}
@@
 \end{center}
-We will use set theory to define labels (e.g. "position = $5m$" will correspond to the set of all states located at the given position).
-These are practically the only new terms introduced by this work, but they are fundamental as most of the later derivation will be based on \emph{counting the labels} (i.e.
+
+We will first define discrete labels using set theory and in the next section generalize to the continuous case using differential geometry.
+These new terms are fundamental as most of the later derivation will be based on \emph{counting the labels} (i.e.
 counting the possible values of a physical quantity) and making sure that such number is conserved (i.e.
 the number of cases is the same).
-
-This section will mathematically define the above concepts, limited to discrete state variables, and derive some basic properties that hold after applying bijective maps.
 
@@
 Given a label $i$, the image $f(i)$ is also a label containing the same number of states.
-Given a set of labels $I$, the image $f(I)$ is also a set of labels containing the same number of labels.
+Given a set of labels $I$, the image $f(I)$ is also a set of labels containing the same number of labels $n(I) = n(f(I))$.
 Given a state variable $\mathbbm{I}$, the image $f(\mathbbm{I})$ is also a state variable.
@@
 We now focus on labels that can be identified by numbers, where we have a bijective map between the label and a number in a set.
-The above definitions readily apply for labels identified by integers.
+The definitions in the previous section readily apply for labels identified by integers.
 Let $z \in \mathbbm{Z}$, we have $\mathbbm{c}=\sum\limits_{z \in \mathbbm{Z}} D(z) \mathbbm{s}(z)$ and the deterministic map becomes $f:\mathbbm{Z} \leftrightarrow \mathbbm{Z}$.
@@
 This also makes physical sense, as we never deal with points per se, but widths that can be made arbitrarily small.
-An cell of $\mathbbm{R}$ corresponds to $\mathbbm{R}^2$: the center and the width.\footnote{In other words: the particles of the system are not point-like, but infinitesimal cell-like.
+A cell of $\mathbbm{R}$ corresponds to $\mathbbm{R}^2$: a center and a width.\footnote{In other words: the particles of the system are not point-like, but infinitesimal cell-like.
 Not only this leads to a more direct understanding of phase-space, it is also more consistent with general relativity (the mass is spread across a small region, not forming a singularity) and quantum mechanics (position is really the central value of a small distribution).} With this in mind, we have the following definitions.
@@
 Each degree of freedom is identified by the pair of labels $\langle q,k \rangle \in \mathbf{T}^*\mathbbm{Q}$.
-$k\,dq$ is invariant under coordinate changes, $dq\wedge dk$ is invariant and $k$ is contravariant.
+Under coordinate changes, $k\,dq$ and $dq\wedge dk$ are invariant and $k$ is contravariant.
 \end{cor}
@@
 
-That is: while the cardinality of each set diverges, the ratio between the two remains finite.\footnote{Such treatment is exactly equivalent to what one does in information theory to extend Shannon's entropy to the continuous case.} For continuous state variables we can define the relative cardinality as the ratio of a label set and our reference set $I_0$.
+That is: while the cardinality of each set diverges, the ratio between the two remains finite.\footnote{Such treatment is exactly equivalent to what one does in information theory to extend Shannon's entropy to the continuous case.
+There is a link between label cardinality and informational entropy, whichwe do not explore here for brevity.} For continuous state variables we can define the relative cardinality as the ratio of a label set and our reference set $I_0$.
 That is $n(I)=\hbar \delta q \delta k$ where $\hbar$ is chosen such that $n(I_0)=\hbar \delta q_0 \delta k_0=1$.\footnote{The choice of angular momentum for $\hbar$ derives from the relationship between $p$ and $dq/dt$, which will be derived later.
@@
 $\theta_0 = p dq$ and $\omega = dq \wedge dp$ are invariant, $p$ is contravariant.
-The configuration state over a degree of freedom is $\mathbbm{c}=\int \rho(q,p) dq \wedge dp \mathbbm{s}_{qp}$, where $\rho(q,p)\equiv D(q,p) / \omega$ is the distribution density for each label.
+The configuration state over a degree of freedom is $\mathbbm{c}=\int \rho(q,p) dq \wedge dp \, \mathbbm{s}_{qp}$, where $\rho(q,p)\equiv D(q,p) / \omega$ is the distribution density for each label.
 \end{cor}
@@
 \begin{defn}\label{canonical}
-We define \emph{canonical transformation} a bijective map on one (or more) degree of freedom.
+We call \emph{canonical transformation} a bijective map on $\mathbf{T}^*\mathbbm{Q}$.
 \end{defn}
@@
 \begin{cor}\label{sdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of the phase space for one degree of freedom. Let
+Let $v$ and $w$ be two vectors defined on the tangent space of $\mathbf{T}^*\mathbbm{Q}$ for one degree of freedom. Let
 \begin{align*}
@@
 
-\section{Multiple degrees of freedom}
+\section{Multiple independent degrees of freedom}
 
 \begin{prop}\label{mdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of the phase space for two independent degrees of freedom.
+Let $v$ and $w$ be two vectors defined on the tangent space of the phase space $\mathbf{T}^*\mathbbm{Q}$ for two independent degrees of freedom.
 Let $a$ and $b$ be indexes for the state variables $\xi^a\equiv \{q^i, p_i\}$.
@@
 \begin{prop}\label{tdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of the manifold identified by the temporal degree of freedom and one standard degree of freedom.
+Let $\mathcal{M}\equiv\mathbbm{T}\times\mathbbm{Q}$ the configuration manifold extended by time.
+Let $v$ and $w$ be two vectors defined on the tangent space of extended phase space $\mathbf{T}^*\mathcal{M}\equiv\mathbf{T}^*\mathbbm{T}\times\mathbf{T}^*\mathbbm{Q}$ for the temporal degree of freedom and one standard degree of freedom.
 Let $a$ and $b$ be indexes for the state variables $\xi^a\equiv\{t, q, E, p\}$.
@@
 \section{Kinematics}
-So far we talked about states without reference to what they physically represent.
+It's now time to turn our attention to the configuration manifold $mathbbm{Q}$, as what we are really interested in studying is the motion of bodies in physical space.
+As the trajectories go through time as well, we will directly study the time dependent case on the extended configuration manifold $mathcal{M}$.
+
+We will introduce another assumption: that the trajectories are enough to fully describe the system.
+We can expect this to hold true if the system is elemental (it has no internal structure) and is sufficiently isolated.
+At that point, if the motion is the result of a deterministic and reversible process, no two trajectories can be attributed to the same state as there is nothing else that could affect them.
 
@@
 \begin{cor}\label{}
-Given all possible space-time trajectory $x^\alpha(s)$ and all possible phase space trajectory $\xi^a(s)$, there exist a bijective function $f: x^\alpha(s) \leftrightarrow \xi^a(s)$ that links each space-time trajectory with one and only one phase space trajectory.
+Given all possible trajectories $x^\alpha(s)$ in the extended configuration manifold $\mathcal{M}$ and all possible trajectory $\xi^a(s)$ in the extended phase space $\mathbf{T}^*\mathcal{M}$, there exist a bijective function $f: x^\alpha(s) \leftrightarrow \xi^a(s)$ that links each space-time trajectory with one and only one phase space trajectory.
 \end{cor}
@@
 \begin{prop}\label{locallyMinkowski}
-Space-time is locally Minkowskian Rimmanian manifold.
+The space-time extended configuration manifold $\mathcal{M}$ is a locally Minkowskian Riemannian manifold.
 That is, there exists a metric $g$ that at any point can be expressed, with a suitable choice of coordinate, as $dx^\alpha g_{\alpha \beta}(P) dx^\beta=dx^\alpha\eta_{\alpha \beta}dx^\beta=(dx^i)^2 - (dx^0)^2=(dx^i)^2 - c^2dt^2$, where $c=n(dq)/n(dt)$ is the ratio between the density of states in space and time.
@@
 As states are defined on intervals, a metric $g$ must be defined on $\mathcal{M}$.
-Such a metric must be consistent with $\omega$, as both must be invariant under coordinate transformations.
+Such a metric must be consistent with $\omega$ as defined on $\mathbf{T}^*\mathcal{M}$, as both must be invariant under coordinate transformations.
 The idea is that each $(dx^\alpha)^2$ can be made to both represent a length square in space-time and an area in phase space, linking the two metrics.
 For each spatial d.o.f.
-fix $dp_i=\lambda dq^i$, we have $dq^i \wedge dp_i = \lambda (dx^i)^2$ where $\lambda$ converts from length squared to the label cardinality contained in the area.
+fix $dp_i=\lambda dq^i$, we have $dq^i \wedge dp_i = \lambda (dq^i)^2$ where $\lambda$ converts from length squared to the label cardinality contained in the area.
 For the temporal d.o.f fix $dE = \lambda c^2 dt$, we have $dt \wedge dE = \lambda c^2 dt^2$.
@@
 \begin{prop}\label{initialConditions}
-Let $u^\alpha = d_s x^\alpha$ be the four-velocity, where the parametrization $s$ is chosen, by convention, to be proper time.
+Let $x^\alpha={ct, q^i}$ and $u^\alpha = d_s x^\alpha$ be the four-velocity, where the parametrization $s$ is chosen, by convention, to be proper time.
 The position $x^\alpha$ and velocity $u^\alpha$ are necessary and sufficient initial conditions to determine the state of the system and its whole trajectory.
@@
 
-For \ref{kinematicAssumption} and \ref{kineticMomentum}, there must exist $p_\alpha=p_\alpha(x^\beta , u^\gamma)$. We express it in terms of $u_\alpha\equiv g_{\alpha \beta} u^\beta$. We have:
+Given \ref{initialConditions}, there must exist $p_\alpha=p_\alpha(x^\beta , u^\gamma)$. We express it in terms of $u_\alpha\equiv g_{\alpha \beta} u^\beta$. We have:
 \begin{align*}
@@
 \end{align*}
-Consider the expression $m du_\alpha dx^\alpha$: this gives us the density of labels in position and velocity converted to the conjugate variables.
+Consider the expression $m du_\alpha dx^\alpha$: this invariant gives us the density of labels in position and velocity converted to the conjugate variables.
 That is: $\omega=m du_\alpha dx^\alpha$.
@@
 Let $\mathcal{H}$ be an extended hamiltonian defined on $\mathbf{T}^*\mathcal{M}$.
-Under \ref{kinematicAssumption}, we can define the Legendre transform $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$ which we call \emph{extended Lagrangian}.
+Under the kinetic assumption, we can define the Legendre transform $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$ which we call \emph{extended Lagrangian}.
 For the time invariant case, we define the \emph{Lagrangian} $L=u^\alpha p_\alpha - H$.
@@
 What is necessary is \ref{initialConditions}: as long as position and velocity are enough to determine the state, $u^\alpha=f(q^\alpha, p_\alpha)$ must be monotonic in $p_\alpha$, which means $\mathcal{H}$ is convex in $p_\alpha$.
-What happens in that case is that the space-time trajectories may become more or less dense, and $n(du_\alpha)$ will be a function of $x^\alpha$ as well.
-So while Hamiltonian mechanics coincides with deterministic and reversible evolution, Lagrangian mechanics coincides with initial conditions given by position and velocity.
-
-\section{Lagrangian}
-
-From Hamiltonian conservation to Lagrangian. p monotonic means convex Hamiltonian: can use Legendre transform.
+What happens in that case is that the space-time trajectories may become more or less dense, $du_\alpha$ and $dx^\alpha$ would not be enough to define the label cardinality and therefore the state density $\rho$.
+In short: deterministic and reversible evolution gives us Hamiltonian mechanics, position and velocity as initial conditions Lagrangian mechanics, and the kinetic assumption (which links the differentials of state variables and initial conditions) gives us the narrowed Hamiltonian form.
 
 \section{Conclusion}
+
+As we have touched many areas, we could only scratch the surface.
+There are many more details that can be expanded upon.
+Yet, the intent is just to give the overall picture, which could be summarized in the following points.
+
+\begin{itemize}
+  \item Classical states are those that describe every infinitesimal part.
+  \item Classical particles are infinitesimal cells. Phase space is the cotangent bundle $\mathbf{T}^*\mathbbm{Q}$ because that's the space of infinitesimal cells.
+  \item Hamiltonian mechanics coincides with deterministic and reversible evolution.
+  \item Hamiltonian flow is the conservation of number of labels for each independent degree of freedom.
+  \item The conservation of number of states is what ultimately gives space-time its locally Minkowskian nature.
+  \item Lagrangian mechanics coincides with position and velocity being necessary and sufficient initial conditions.
+  \item The motion of an isolated elementary system, for which the kinetic assumption is valid, is restricted to a Hamiltonian that can describe fundamental classical forces.
+\end{itemize}
+
+This helps clarify and understand much of the classical framework in a more cohesive way. The notion of cardinality of labels and states is what's at the heart.
+
+While this is limited to classical particle mechanics, it should be obvious how, at least in principle, this work could be extended.
+For field theories, the \emph{kinematic assumption} should be substituted by a \emph{kinematic field assumption}: what we are studying are not trajectories $x^\alpha(s)$ but fields $\psi(x^\alpha)$.
+The field values at each point become a set of independent state variables, where their conjugates $\pi(x^\alpha)$ will be linked to $d_s\psi(x^\alpha)$.
+For quantum mechanics, the \emph{infinite reducibility assumption} has to leave way to an \emph{irreducibility assumption}: the state of parts of a quantum system cannot be known.
+The configuration state as a whole undergoes deterministic and reversible evolution, while the motion of its parts does not: it is non deterministic.
+
+The hope is that, by continuing in this approach, we can shed more light on why the laws of physics are what they are; and show that they are not arbitrary rules, but necessary given few simple assumptions.
 

### 2015-02-23 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Another full pass to the language and notation.

@@
 \begin{abstract}
-We show that classical mechanics, including Hamiltonian, Lagrangian, non-relativistic Newtonian gravitation and relativistic electromagnetism, can be derived from three fundamental assumptions: infinite reducibility, deterministic and reversible evolution, and kinematic equivalence.
+We show that classical mechanics (Hamiltonian and Lagrangian consistent with relativistic electromagnetism) can be derived from three fundamental assumptions: infinite reducibility, deterministic and reversible evolution, and kinematic equivalence.
 The core idea is that deterministic and reversible systems preserve the cardinality of a set of states, which puts considerable constraints on the equations of motion.
-This perspective links different concepts from different branches of math and physics (cardinality of a set, cotangent bundle for phase space, locally Minkowskian space-time manifold, inertial mass) providing new insights.
-The derivation strives to use definition and mathematical concepts that would allow a future extension to quantum mechanics.\end{abstract}
+This perspective links different concepts from different branches of math and physics (e.g.
+cardinality of a set, cotangent bundle for phase space, Hamiltonian flow, locally Minkowskian space-time manifold) providing new insights.
+The derivation strives to use definition and mathematical concepts compatible with a future extensions to field theories and quantum mechanics.\end{abstract}
 \maketitle
@@
 
-Classical mechanics is usually founded on Newton's laws.
-These, though, are insufficient to derive the Lagrangian and Hamiltonian formalism, and usually other ad-hoc assumptions (e.g.
+Classical particle mechanics is usually founded on Newton's laws.
+These, though, are insufficient to derive the full Lagrangian and Hamiltonian formalism, and usually other ad-hoc assumptions (e.g.
 conservative forces) are introduced.
@@
 What we'll do in this work is re-organize all the known elements and equations in a more consistent and comprehensive way, leading to better insight on why the fundamental concepts and laws are what they are.
-The derivation will also include classical analogues of strictly quantum concepts to facilitate a future extension to quantum mechanics.
+The derivation will also include classical analogues of strictly quantum notions to facilitate a future extension to quantum mechanics.
 
-We will use concepts from different disciplines, such as set theory, differential geometry, relativity, Hamiltonian and Lagrangian mechanics, and will find interesting connections between them.
-We'll keep names and notation as consistent as possible to current use across the different disciplines, but this may lead to some non sequitur as it will not be immediately clear why the different definition will be at the end equivalent to the standard one.
-This is still preferable than introducing a whole new set of names one has to become familiar.
+We'll use concepts from different disciplines, such as set theory, differential geometry, relativity, Hamiltonian and Lagrangian mechanics, and we'll find interesting connections between them.
+We'll keep names and notation as consistent as possible to current use across the different disciplines.
+This may sometime lead to some non sequitur as it will not be immediately clear why the new definitions are equivalent to the standard ones.
+These are typically resolved by subsequent derivation of the expected properties.
 
@@
 
-We give here a brief a general conceptual overview, hoping it will help guide the reader through the mathematical details.
+We give here a brief general conceptual overview, hoping it will help guide the reader through the mathematical details.
 
@@
 We introduce the assumption of kinematic equivalence: studying trajectories is equivalent to studying states.
-We show that there must be link between the space-time metric $g$ and the phase-space metric $\omega$, as both quantities must be conserved under passive coordinate transformations.
-That link constrains the space-time metric to be locally Minkowskian.
+We show that there must be a link between the space-time metric $g$ and the phase-space metric $\omega$, as both quantities must be conserved under passive coordinate transformations.
+That link will constrain the space-time metric to be locally Minkowskian.
 A transformation between state variables $(q, p)$ and kinematic variables $(x, \dot{x})$ must exist.
@@
 The first is between the state of the whole system and one of its parts.\footnote{Whole system vs infinitesimal component is also source of confusion when comparing classical and quantum states.
-Quantum states are always whole systems and are always distributions as there is no state uniquely attributed to its infinitesimal components.}
+Quantum states are always whole and are always distributions as there is no state uniquely attributed to parts of a quantum system.}
 \begin{center}
@@
 When talking about particles of a system we must always remember that they are the result of a limit.
-Therefore we will start with discrete definitions, and see how some properties will extend from discrete case to continuous.
+Therefore we will start with discrete definitions, and see how some properties will extend to the continuous case.
 
@@
 \begin{defn}\label{discreteCardinality}
-Let $I$ be a finite, countable set of labels. We define the \emph{cardinality} $n: I \rightarrow \mathbbm{N}$ as the number of labels in the set.
+Let $I$ be a finite, countable set of labels. We define the \emph{cardinality} $n(I) \rightarrow \mathbbm{N}$ as the number of labels in the set.
 \end{defn}
@@
 A cell of $\mathbbm{R}$ corresponds to $\mathbbm{R}^2$: a center and a width.\footnote{In other words: the particles of the system are not point-like, but infinitesimal cell-like.
-Not only this leads to a more direct understanding of phase-space, it is also more consistent with general relativity (the mass is spread across a small region, not forming a singularity) and quantum mechanics (position is really the central value of a small distribution).} With this in mind, we have the following definitions.
+Not only this leads to a more direct understanding of classical phase-space, it is also more consistent with general relativity (the mass is spread across a small region, not forming a singularity) and quantum mechanics (position is really the central value of a small distribution).} With this in mind, we have the following definitions.
 
@@
 In the limit, we will cover every possible center $q$ and every possible cell number $k$.
-Our configuration state becomes $\mathbbm{c}=\int\limits_{q \in Q \; k\in K} \rho(q, k) dq \wedge dk \, \mathbbm{s}(q(i), k(i))$.
+Our configuration state becomes $\mathbbm{c}=\int\limits_{q \in Q \; k\in K} \rho(q, k) dq \wedge dk \, \mathbbm{s}(q, k)$.
 We can repeat the process increasing or changing the region covered by $Q$ and $K$, until we cover all of $\mathbbm{Q}$ and $\mathbbm{R}$.
@@
 \begin{prop}\label{relativeCardinality}
-Let $I$ be a closed dense set of labels, subset of a degree of freedom.
-We define \emph{relative cardinality} $n: I \rightarrow \mathbbm{R}$ the ratio between the number of labels in $I$ and the ones of a reference set $I_0$.
+Let $I \subset \mathbf{T}^*\mathbbm{Q}$ be a closed dense set of labels, subset of a degree of freedom (i.e.
+a region of phase space).
+We define \emph{relative cardinality} $n(I) \rightarrow \mathbbm{R}$ the ratio between the number of labels in $I$ and the ones of a reference set $I_0$.
 We find $n(I)=\int \omega$, where $\omega = \hbar \, dq \wedge dk$ and $\hbar$ is the constant that determines the unit (i.e.
-chosen so $n(I_0)=1$).
+fixed so that $n(I_0) = 1$).
 \end{prop}
@@
 There is a link between label cardinality and informational entropy, whichwe do not explore here for brevity.} For continuous state variables we can define the relative cardinality as the ratio of a label set and our reference set $I_0$.
-That is $n(I)=\hbar \delta q \delta k$ where $\hbar$ is chosen such that $n(I_0)=\hbar \delta q_0 \delta k_0=1$.\footnote{The choice of angular momentum for $\hbar$ derives from the relationship between $p$ and $dq/dt$, which will be derived later.
+That is $n(I)=\hbar \delta q \delta k$ where $\hbar=n(I_0)/(\delta q_0 \delta k_0)$ and $n(I_0)\equiv1$ by definition.\footnote{The choice of angular momentum for $\hbar$ derives from the relationship between $p$ and $dq/dt$, which will be derived later.
 The value for $\hbar$ in classical mechanics is arbitrary: we have no physical reason to choose a reference label set over another.
 In quantum mechanics, instead, we will have a preference: the set that corresponds to a single quantized system.
-We use this arbitrariness to set $\hbar$ to the value to the known constant.} In general, the label set is not rectangular in $\langle q,k \rangle$ and we have $n(I)=\int \hbar dq \wedge dk$
+We use this arbitrariness to set the value of our classical analogue of $\hbar$ to the known constant.} In general, the label set is not rectangular in $\langle q,k \rangle$ and we have $n(I)=\int \hbar dq \wedge dk$
 
@@
 $\theta_0 = p dq$ and $\omega = dq \wedge dp$ are invariant, $p$ is contravariant.
-The configuration state over a degree of freedom is $\mathbbm{c}=\int \rho(q,p) dq \wedge dp \, \mathbbm{s}_{qp}$, where $\rho(q,p)\equiv D(q,p) / \omega$ is the distribution density for each label.
+The configuration state is $\mathbbm{c}=\int \rho(q,p) dq \wedge dp \, \mathbbm{s}(q, p)$, where $\rho(q,p)\equiv D(q,p) / \omega$ is the distribution density for each label.
 \end{cor}
@@
 \begin{cor}\label{sdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of $\mathbf{T}^*\mathbbm{Q}$ for one degree of freedom. Let
+Let $v$ and $w$ be two vectors defined on the tangent space of the phase space $\mathbf{T}^*\mathbbm{Q}$ for one degree of freedom. Let
 \begin{align*}
@@
 \begin{defn}\label{tdof}
-The \emph{temporal degree of freedom} is a state variable identified by temporal cells.
-The center of each cell is identified by $t$, the width by $\omega dt$, the conjugate variable $E\equiv\hbar\omega$.
-We call \emph{extended phase space} the outer product between phase space and the temporal degree of freedom.
+We call \emph{extended phase space} the cotangent bundle $\mathbf{T}^*\mathcal{M}$, where the manifold $\mathcal{M}$ identifies the possible center values for all infinitesimal cells at all times.
+We call \emph{temporal degree of freedom} the state variable identified by temporal cells.
+The center of each cell is identified by $t$, the width by $\hat{\omega} dt$, the conjugate variable $E\equiv\hbar\hat{\omega}$.\footnote{$\hat{\omega}$ is the classical analogue of the wave frequency.
+We use $\hat{\omega}$ to distinguish from the phase space metric $\omega$.}
 \end{defn}
@@
 \begin{prop}\label{tdofInvariant}
-Let $\mathcal{M}\equiv\mathbbm{T}\times\mathbbm{Q}$ the configuration manifold extended by time.
-Let $v$ and $w$ be two vectors defined on the tangent space of extended phase space $\mathbf{T}^*\mathcal{M}\equiv\mathbf{T}^*\mathbbm{T}\times\mathbf{T}^*\mathbbm{Q}$ for the temporal degree of freedom and one standard degree of freedom.
+Let $v$ and $w$ be two vectors defined on the tangent space of extended phase space $\mathbf{T}^*\mathcal{M}$ for the temporal degree of freedom and one standard degree of freedom.
 Let $a$ and $b$ be indexes for the state variables $\xi^a\equiv\{t, q, E, p\}$.
@@
 \section{Kinematics}
-It's now time to turn our attention to the configuration manifold $mathbbm{Q}$, as what we are really interested in studying is the motion of bodies in physical space.
-As the trajectories go through time as well, we will directly study the time dependent case on the extended configuration manifold $mathcal{M}$.
-
-We will introduce another assumption: that the trajectories are enough to fully describe the system.
-We can expect this to hold true if the system is elemental (it has no internal structure) and is sufficiently isolated.
+In the previous section we made no constraint on what our state variables $q^i$ actually represent.
+In this section we turn our attention to the study of the motion of a body.
+In particular, we will introduce another assumption: that the trajectories are enough to fully describe the system.
+We can expect this to hold true if the system is elemental (it has no relevant internal structure) and is sufficiently isolated.
 At that point, if the motion is the result of a deterministic and reversible process, no two trajectories can be attributed to the same state as there is nothing else that could affect them.
+As position and time will be state variables, the space-time manifold will coincide with $\mathcal{M}$ and the extended phase space with $\mathbf{T}^*\mathcal{M}$.
 
@@
 
-\begin{cor}\label{}
-Given all possible trajectories $x^\alpha(s)$ in the extended configuration manifold $\mathcal{M}$ and all possible trajectory $\xi^a(s)$ in the extended phase space $\mathbf{T}^*\mathcal{M}$, there exist a bijective function $f: x^\alpha(s) \leftrightarrow \xi^a(s)$ that links each space-time trajectory with one and only one phase space trajectory.
+\begin{cor}\label{singleTrajectory}
+Fix a system under deterministic and reversible evolution.
+Given all its possible trajectories $x^\alpha(s)$ in the space-time manifold $\mathcal{M}$ and all its possible trajectory $\xi^a(s)$ in the extended phase space $\mathbf{T}^*\mathcal{M}$, there exist a bijective function $f: x^\alpha(s) \leftrightarrow \xi^a(s)$ that links each space-time trajectory with one and only one phase space trajectory.
 \end{cor}
@@
 If studying the motion and state evolution are equivalent, then we must be able to go back and forth between the two pictures.
-Without losing generality, we can chose $x^\alpha=f(q^i,t)$ to be a function of only $q^i$ and $t$.
+Without losing generality, we can chose $x^\alpha=f(q^i,t)$ to be a linear function of only $q^i$ and $t$.
 
 \begin{prop}\label{locallyMinkowski}
-The space-time extended configuration manifold $\mathcal{M}$ is a locally Minkowskian Riemannian manifold.
-That is, there exists a metric $g$ that at any point can be expressed, with a suitable choice of coordinate, as $dx^\alpha g_{\alpha \beta}(P) dx^\beta=dx^\alpha\eta_{\alpha \beta}dx^\beta=(dx^i)^2 - (dx^0)^2=(dx^i)^2 - c^2dt^2$, where $c=n(dq)/n(dt)$ is the ratio between the density of states in space and time.
+The space-time manifold $\mathcal{M}$ is a locally Minkowskian Riemannian manifold.
+That is, there exists a metric $g$ that at any point can be expressed, with a suitable choice of coordinate, as $g=dx^\alpha g_{\alpha \beta}(P) dx^\beta=dx^\alpha\eta_{\alpha \beta}dx^\beta=(dx^i)^2 - (dx^0)^2=(dx^i)^2 - c^2dt^2$, where $c=n(dq)/n(dt)$ is the ratio between the density of states in space and time.
 \end{prop}
@@
 Set $x^i=q^i$ and $x^0=ct$.
-We have $\omega = \lambda ((dx^i)^2 - (dx^0)^2)$ and $dx^\alpha g_{\alpha \beta} dx^\beta=dx^\alpha g_{\alpha \alpha}dx^\alpha$.
+We have $\omega = \lambda ((dx^i)^2 - (dx^0)^2)$ and $g=dx^\alpha g_{\alpha \beta} dx^\beta=dx^\alpha g_{\alpha \alpha}dx^\alpha$.
 Both are invariant under coordinate transformation for any $dx^\alpha$.
@@
 \begin{defn}\label{inertialMass}
-Let $dp_\alpha$ an infinitesimal interval of conjugate momentum and $d(g_{\alpha \beta}u^\beta)$ an infinitesimal interval of the covariant components of the four velocity.
-The ratio $m=n(dp_\alpha)/n(d(g_{\alpha \beta}u^\beta))$ between their relative label cardinality is called \emph{inertial mass}.
+Let $dp_\alpha$ an infinitesimal interval of conjugate momentum and $du_\alpha\equiv d(g_{\alpha \beta}u^\beta)$ an infinitesimal interval of the covariant components of the four velocity.
+The ratio $m=n(dp_\alpha)/n(du_\alpha)$ between their relative label cardinality is called \emph{inertial mass}.
 \end{defn}
@@
 Let $x^\alpha=\{ct, q^i\}$ and $p_\alpha=\{-E/c, p_i\}$.
-Then $p_\alpha= m g_{\alpha \beta}u^\beta + \hat{p}_\alpha(x^\gamma)$, where $\hat{p}_\alpha : \mathcal{M} \rightarrow \mathbbm{R}$ is a function defined on the extended configuration manifold.
+Then $p_\alpha= m g_{\alpha \beta}u^\beta + \hat{p}_\alpha(x^\gamma)$, where $\hat{p}_\alpha : \mathcal{M} \rightarrow \mathbbm{R}$ is a function defined on the space-time manifold.
 \end{prop}
@@
 \begin{prop}\label{kineticHamiltonian}
-The extended Hamiltonian is $\mathcal{H}=\frac{1}{2m}(p_\alpha-\hat{p}_\alpha(x))g^{\alpha\beta}(p_\beta-\hat{p}_\beta(x))+\hat{\mathcal{H}}(x)$, where $\hat{\mathcal{H}}_\alpha : \mathcal{M} \rightarrow \mathbbm{R}$ is a function defined on the extended configuration manifold.
+The extended Hamiltonian is $\mathcal{H}=\frac{1}{2m}(p_\alpha-\hat{p}_\alpha(x))g^{\alpha\beta}(p_\beta-\hat{p}_\beta(x))+\hat{\mathcal{H}}(x)$, where $\hat{\mathcal{H}} : \mathcal{M} \rightarrow \mathbbm{R}$ is a function defined on the space-time manifold.
 \end{prop}
@@
 \begin{align*}
-\frac{dq^\alpha}{ds} &= u^\alpha \\
+\frac{dx^\alpha}{ds} &= u^\alpha \\
 &= \frac{1}{m}g^{\alpha\beta}(p_\beta-\hat{p}_\beta) \\
-&= \frac{\partial \mathcal{H}}{dp_\alpha} \\
+&= \frac{\partial \mathcal{H}}{\partial p_\alpha} \\
 \end{align*}
@@
 
-Now that we have found the general form of the Hamiltonian, we show that this is compatible with the established fundamental classical theories.
+Now that we have found the constrained form of the Hamiltonian, we show that this is compatible with the established fundamental classical theories.
 
 \begin{prop}\label{relativisticEM}
-Let $\hat{p}_\alpha = q A_\alpha$, $F_{\alpha \beta} \equiv \partial_\alpha A_\beta - \partial_\beta A_\alpha$ and $\hat{H} = 0$, the equations of motion are $m \nabla_u u^\alpha = g^{\alpha\beta} F_{\beta \gamma} q u^\gamma$.
+Let $\hat{p}_\alpha = q A_\alpha$, $F_{\alpha \beta} \equiv \partial_\alpha A_\beta - \partial_\beta A_\alpha$ and $\hat{\mathcal{H}} = 0$, the equations of motion are $m \nabla_u u^\alpha = g^{\alpha\beta} F_{\beta \gamma} q u^\gamma$.
 These are the relativistic equations for a charged particle.
@@
 \begin{align*}
-u^\alpha &= \frac{dq^\alpha}{ds} = \frac{\partial \mathcal{H}}{dp_\alpha} \\
+u^\alpha &= \frac{dx^\alpha}{ds} = \frac{\partial \mathcal{H}}{dp_\alpha} \\
 &= \frac{1}{m}g^{\alpha\beta}(p_\beta-\hat{p}_\beta) \\
@@
  &- (p_\beta -\hat{p}_\beta) \partial_\alpha g^{\beta \gamma} (p_\gamma -\hat{p}_\gamma) \\
- &+ (p_\beta -\hat{p}_\beta) g^{\beta \gamma} \partial_\alpha \hat{p}_\gamma ]- \partial_\alpha \hat{H} \\
+ &+ (p_\beta -\hat{p}_\beta) g^{\beta \gamma} \partial_\alpha \hat{p}_\gamma ]- \partial_\alpha \hat{\mathcal{H}} \\
 &=\frac{1}{2}[\partial_\alpha \hat{p}_\beta u^\beta
 - m u^\delta g_{\delta\beta} \partial_\alpha g^{\beta \gamma} u^\epsilon g_{\epsilon\gamma}
-+ u^\gamma \partial_\alpha \hat{p}_\gamma ]- \partial_\alpha \hat{H}
++ u^\gamma \partial_\alpha \hat{p}_\gamma ]- \partial_\alpha \hat{\mathcal{H}}
 \end{align*}
@@
 &= \partial_\gamma g^{\alpha\beta} d_s x^\gamma m g_{\beta \delta} u^\delta + g^{\alpha\beta} (d_s p_\beta - \partial_\gamma \hat{p}_\beta d_s x^\gamma) \\
-&= \partial_\gamma g^{\alpha\beta} u^\gamma m g_{\beta \delta} u^\delta + g^{\alpha\beta} \frac{1}{2} [\partial_\beta \hat{p}_\gamma u^\gamma
-- m u^\epsilon g_{\epsilon\gamma} \partial_\beta g^{\gamma \delta} u^\zeta g_{\zeta\delta} \\
-&+ u^\delta \partial_\beta \hat{p}_\delta ]- g^{\alpha\beta} \partial_\beta \hat{H} - g^{\alpha\beta} \partial_\gamma \hat{p}_\beta u^\gamma \\
+&= \partial_\gamma g^{\alpha\beta} u^\gamma m g_{\beta \delta} u^\delta + g^{\alpha\beta} \frac{1}{2} [\partial_\beta \hat{p}_\gamma u^\gamma \\
+&- m u^\epsilon g_{\epsilon\gamma} \partial_\beta g^{\gamma \delta} u^\zeta g_{\zeta\delta} \\
+&+ u^\delta \partial_\beta \hat{p}_\delta ]- g^{\alpha\beta} \partial_\beta \hat{\mathcal{H}} - g^{\alpha\beta} \partial_\gamma \hat{p}_\beta u^\gamma \\
 &= m  g_{\beta \delta} \partial_\gamma g^{\alpha\beta} u^\gamma u^\delta - \frac{1}{2} m g^{\alpha\beta} g_{\zeta\delta} g_{\epsilon\gamma} \partial_\beta g^{\gamma \delta} u^\epsilon u^\zeta  \\
 &+ g^{\alpha\beta} \partial_\beta \hat{p}_\gamma u^\gamma - g^{\alpha\beta} \partial_\gamma \hat{p}_\beta u^\gamma
-- g^{\alpha\beta} \partial_\beta \hat{H}\\
+- g^{\alpha\beta} \partial_\beta \hat{\mathcal{H}}\\
 &= - m  g^{\alpha \beta} \partial_\gamma g_{\beta\delta} u^\gamma u^\delta + \frac{1}{2} m g^{\alpha\beta} \partial_\beta g_{\gamma \delta} u^\gamma u^\delta  \\
 &+ g^{\alpha\beta} (\partial_\beta \hat{p}_\gamma - \partial_\gamma \hat{p}_\beta ) u^\gamma
-- g^{\alpha\beta} \partial_\beta \hat{H}\\
+- g^{\alpha\beta} \partial_\beta \hat{\mathcal{H}}\\
 &= - m \frac{1}{2} g^{\alpha \beta} ( \partial_\gamma g_{\beta\delta} + \partial_\delta g_{\beta\gamma} - \partial_\beta g_{\gamma \delta} ) u^\gamma u^\delta  \\
 &+ g^{\alpha\beta} (\partial_\beta \hat{p}_\gamma - \partial_\gamma \hat{p}_\beta ) u^\gamma
-- g^{\alpha\beta} \partial_\beta \hat{H}\\
+- g^{\alpha\beta} \partial_\beta \hat{\mathcal{H}}\\
 &= - m \Gamma ^\alpha_{\ \gamma \delta} u^\gamma u^\delta + g^{\alpha\beta} (\partial_\beta \hat{p}_\gamma - \partial_\gamma \hat{p}_\beta ) u^\gamma
-- g^{\alpha\beta} \partial_\beta \hat{H}
+- g^{\alpha\beta} \partial_\beta \hat{\mathcal{H}}
 \end{align*}
@@
 m \nabla_{u} u^\alpha &= m (d_s u^\alpha + \Gamma ^\alpha_{\ \gamma \delta} u^\gamma u^\delta)  \\
-&= g^{\alpha\beta} (\partial_\beta \hat{p}_\gamma - \partial_\gamma \hat{p}_\beta ) u^\gamma - g^{\alpha\beta} \partial_\beta \hat{H}\\
-&= g^{\alpha\beta} (\nabla_\beta \hat{p}_\gamma - \nabla_\gamma \hat{p}_\beta ) u^\gamma - g^{\alpha\beta} \nabla_\beta \hat{H}
+&= g^{\alpha\beta} (\partial_\beta \hat{p}_\gamma - \partial_\gamma \hat{p}_\beta ) u^\gamma - g^{\alpha\beta} \partial_\beta \hat{\mathcal{H}}\\
+&= g^{\alpha\beta} (\nabla_\beta \hat{p}_\gamma - \nabla_\gamma \hat{p}_\beta ) u^\gamma - g^{\alpha\beta} \nabla_\beta \hat{\mathcal{H}}
 \end{align*}
@@
 The equation is manifestly covariant.
-If $\hat{p}$ and $\hat{H}$ are zero, absence of forces, we recognize the geodesic equation.
-We substitute $\hat{p}_\alpha = q A_\alpha$ and $\hat{H} = 0$ and have:
+If $\hat{p}$ and $\hat{\mathcal{H}}$ are zero, absence of forces, we recognize the geodesic equation.
+We substitute $\hat{p}_\alpha = q A_\alpha$ and $\hat{\mathcal{H}} = 0$ and have:
 
@@
 Assume $\mathcal{M}$ flat and time independent motion, the Hamiltonian simplifies to $H=\frac{1}{2m}(p_i-\hat{p}_\alpha(x))^2)+\hat{H}(x)$.
-Let $\hat{p}_i = 0$ and $\hat{H} = mV$, the equations of motion are $m d_t v^i = - m \partial_i V$, where $v^i=dx^1/dt$.
+Let $\hat{p}_i = 0$ and $\hat{H} = mV$, the equations of motion are $m d_t v^i = - m \partial_i V$, where $v^i=dx^i/dt$.
 These are equations for a particle under a Netwonian gravitational potential.
@@
 Under the kinetic assumption, we can define the Legendre transform $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$ which we call \emph{extended Lagrangian}.
-For the time invariant case, we define the \emph{Lagrangian} $L=u^\alpha p_\alpha - H$.
+For the time invariant case, we define the \emph{Lagrangian} $L=v^i p_i - H$.
 \end{defn}
@@
 The Hamiltonian found at \ref{kineticHamiltonian} is convex, so the Lagrangian can always be defined.
-While the kinetic equivalence assumption is sufficient, it's not necessary.
+While the kinetic equivalence assumption is sufficient, it's not necessary for the existence of a Lagrangian.
 What is necessary is \ref{initialConditions}: as long as position and velocity are enough to determine the state, $u^\alpha=f(q^\alpha, p_\alpha)$ must be monotonic in $p_\alpha$, which means $\mathcal{H}$ is convex in $p_\alpha$.
 What happens in that case is that the space-time trajectories may become more or less dense, $du_\alpha$ and $dx^\alpha$ would not be enough to define the label cardinality and therefore the state density $\rho$.
-In short: deterministic and reversible evolution gives us Hamiltonian mechanics, position and velocity as initial conditions Lagrangian mechanics, and the kinetic assumption (which links the differentials of state variables and initial conditions) gives us the narrowed Hamiltonian form.
+In short: deterministic and reversible evolution gives us Hamiltonian mechanics, position and velocity as initial conditions Lagrangian mechanics, and the kinetic assumption (which links the differentials of the state variables and the initial conditions) gives us the narrowed Hamiltonian form.
 
@@
 
-As we have touched many areas, we could only scratch the surface.
-There are many more details that can be expanded upon.
-Yet, the intent is just to give the overall picture, which could be summarized in the following points.
+While there are many more details that we could have expanded up, as we have touched many areas we could only scratch the surface, the intent is to give the overall picture, which could be summarized in the following points.
 
@@
   \item Hamiltonian flow is the conservation of number of labels for each independent degree of freedom.
-  \item The conservation of number of states is what ultimately gives space-time its locally Minkowskian nature.
+  \item Deterministic and reversible evolution is what ultimately gives space-time its locally Minkowskian nature.
   \item Lagrangian mechanics coincides with position and velocity being necessary and sufficient initial conditions.
@@
 
-This helps clarify and understand much of the classical framework in a more cohesive way. The notion of cardinality of labels and states is what's at the heart.
+This helps clarify and understand much of the classical framework in a more cohesive way and the notion of cardinality of labels and states is at the heart.
 
@@
 For field theories, the \emph{kinematic assumption} should be substituted by a \emph{kinematic field assumption}: what we are studying are not trajectories $x^\alpha(s)$ but fields $\psi(x^\alpha)$.
-The field values at each point become a set of independent state variables, where their conjugates $\pi(x^\alpha)$ will be linked to $d_s\psi(x^\alpha)$.
+The field values at each point become a set of independent state variables and their conjugates $\pi(x^\alpha)$ will be linked to $d_s\psi(x^\alpha)$.
 For quantum mechanics, the \emph{infinite reducibility assumption} has to leave way to an \emph{irreducibility assumption}: the state of parts of a quantum system cannot be known.
-The configuration state as a whole undergoes deterministic and reversible evolution, while the motion of its parts does not: it is non deterministic.
+The configuration state as a whole undergoes deterministic and reversible evolution, while the motion of its parts does not.
 

### 2015-02-24 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Minor edits

@@
 \begin{abstract}
-We show that classical mechanics (Hamiltonian and Lagrangian consistent with relativistic electromagnetism) can be derived from three fundamental assumptions: infinite reducibility, deterministic and reversible evolution, and kinematic equivalence.
+We show that classical particle mechanics (Hamiltonian and Lagrangian consistent with relativistic electromagnetism) can be derived from three fundamental assumptions: infinite reducibility, deterministic and reversible evolution, and kinematic equivalence.
 The core idea is that deterministic and reversible systems preserve the cardinality of a set of states, which puts considerable constraints on the equations of motion.
@@
 To be invertible, the relationship between velocity and conjugate momentum has to be monotonic, leading to a concave Hamiltonian which allows a Legendre transformation leading to the Lagrangian.
-We can constrain the Hamiltonian further, and show that the most general equation of motion under the three assumptions is the one of a geodesic modified by the force given by a vector potential (such as the one of a relativistic charged particle) and a scalar potential (such as the one of newtonian gravitation).
+We can constrain the Hamiltonian further, and show that the most general equation of motion under the three assumptions is the one of a geodesic modified by the force given by a vector potential (such as the one of a relativistic charged particle) and a scalar potential (such as the one of newtonian gravitation in the non-relativistic case).
 
@@
 The first is between the state of the whole system and one of its parts.\footnote{Whole system vs infinitesimal component is also source of confusion when comparing classical and quantum states.
-Quantum states are always whole and are always distributions as there is no state uniquely attributed to parts of a quantum system.}
+Quantum states are always whole and are always distributions as there is no state attributable to parts of a quantum system.}
 \begin{center}
@@
 
-The second distinctions is, when referring to a physical quantity, between the space of all possible values, a particular value or a set of possible values.
+The second distinction is, when referring to a physical quantity, between the space of all possible values, a particular value or a set of possible values.
 \begin{center}
@@
 The continuous distribution $\rho$ is a density, defined over interval $dr$.
-That is: it's really $\rho(r, dr(r))$, function of both the center and the width of the interval.
+That is: it's really $\rho(r, dr)$, function of both the center and the width of the interval.
 A bijective map on just $r$ is not sufficient, $dr$ must be mapped as well.
@@
 A cell of $\mathbbm{R}$ corresponds to $\mathbbm{R}^2$: a center and a width.\footnote{In other words: the particles of the system are not point-like, but infinitesimal cell-like.
-Not only this leads to a more direct understanding of classical phase-space, it is also more consistent with general relativity (the mass is spread across a small region, not forming a singularity) and quantum mechanics (position is really the central value of a small distribution).} With this in mind, we have the following definitions.
+Not only this leads to a more direct understanding of classical phase-space, it is also more consistent with general relativity (the mass is spread across a small region, not forming a singularity) and quantum mechanics (position is always the central value of a small distribution).} With this in mind, we have the following definitions.
 
@@
 
-If we apply a coordinate change $q'=q'(q)$, we change the labels but the state must remain the same, defined at the same point with the same width.
-$k' dq'$ must be then equal to $k dq$; $k'= k dq / dq'$ is contravariant, $dq' \wedge dk'= dq'/dq \, dq \wedge dk \, dq/dq' = dq \wedge dk$ is invariant.
+If we apply a coordinate change $q'=q'(q)$, we change the labels but the state must remain the same, defined at the same point with the same width.\footnote{The notion that states are what they are no matter how we label them already contains the principle of relativity.} $k' dq'$ must be then equal to $k dq$; $k'= k dq / dq'$ is contravariant, $dq' \wedge dk'= dq'/dq \, dq \wedge dk \, dq/dq' = dq \wedge dk$ is invariant.
 This means that $\rho$ is also invariant, which makes sense: the density depends only on the state (the cell), and not the coordinate chosen to represent them.
@@
 
-That is: while the cardinality of each set diverges, the ratio between the two remains finite.\footnote{Such treatment is exactly equivalent to what one does in information theory to extend Shannon's entropy to the continuous case.
-There is a link between label cardinality and informational entropy, whichwe do not explore here for brevity.} For continuous state variables we can define the relative cardinality as the ratio of a label set and our reference set $I_0$.
+That is: while the cardinality of each set diverges, the ratio between the two remains finite.\footnote{Such treatment is equivalent to what one does in information theory to extend Shannon's entropy\cite{Shannon} to the continuous case.
+There is a link between label cardinality and informational entropy, including Jaynes\cite{Jaynes} invariant formulation, which we do not explore here for brevity.} For continuous state variables we can define the relative cardinality as the ratio of a label set and our reference set $I_0$.
 That is $n(I)=\hbar \delta q \delta k$ where $\hbar=n(I_0)/(\delta q_0 \delta k_0)$ and $n(I_0)\equiv1$ by definition.\footnote{The choice of angular momentum for $\hbar$ derives from the relationship between $p$ and $dq/dt$, which will be derived later.
@@
 In quantum mechanics, instead, we will have a preference: the set that corresponds to a single quantized system.
-We use this arbitrariness to set the value of our classical analogue of $\hbar$ to the known constant.} In general, the label set is not rectangular in $\langle q,k \rangle$ and we have $n(I)=\int \hbar dq \wedge dk$
+We use this arbitrariness to set the value of our classical analogue of $\hbar$ to the known constant.} In general, the label set is not rectangular in $\langle q,k \rangle$ and we have $n(I)=\int \hbar dq \wedge dk$.\footnote{It should be stressed that no notion of uncertainty is used here.
+Though there is a link between label cardinality and uncertainty, which we do not explore here for brevity, label cardinality is the more fundamental as we can still count the number of possible states even in an ideal case of no uncertainty.}
 

### 2015-02-24 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Minor edits

@@
 \begin{defn}\label{canonical}
-We call \emph{canonical transformation} a bijective map on $\mathbf{T}^*\mathbbm{Q}$.
+We call \emph{canonical transformation} the continuous limit of a bijective map, as defined by \ref{detrevMap}, on $\mathbf{T}^*\mathbbm{Q}$.
 \end{defn}
@@
 
-If we set\footnote{We avoided using $p^{n+1}$ as it hides the minus sign from the metric, making it seem that the temporal d.o.f is just another independent d.o.f.} $q^{n+1}=t$ and $p^{n+1}=-e$, we recognise Hamilton equations in the extended phase space\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}\cite{Synge,Lanczos}.
+If we set\footnote{We avoided using $p_{n+1}$ as it hides the minus sign from the metric, making it seem that the temporal d.o.f is just another independent d.o.f.} $q^{n+1}=t$ and $p_{n+1}=-E$, we recognise Hamilton equations in the extended phase space\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}\cite{Synge,Lanczos}.
 
@@
 In particular, we will introduce another assumption: that the trajectories are enough to fully describe the system.
-We can expect this to hold true if the system is elemental (it has no relevant internal structure) and is sufficiently isolated.
+We can expect this to hold true if the system is elementary (it has no relevant internal structure) and is sufficiently isolated.
 At that point, if the motion is the result of a deterministic and reversible process, no two trajectories can be attributed to the same state as there is nothing else that could affect them.
@@
 \begin{prop}\label{initialConditions}
-Let $x^\alpha={ct, q^i}$ and $u^\alpha = d_s x^\alpha$ be the four-velocity, where the parametrization $s$ is chosen, by convention, to be proper time.
+Let $x^\alpha=\{ct, q^i\}$ and $u^\alpha = d_s x^\alpha$ be the four-velocity, where the parametrization $s$ is chosen, by convention, to be proper time.
 The position $x^\alpha$ and velocity $u^\alpha$ are necessary and sufficient initial conditions to determine the state of the system and its whole trajectory.
@@
 As the equation of motion \ref{tdofHam} can be at most second order in $\{t, q^i\}$, they can at most be second order in $x^\alpha=\{ct, q^i\}$ so only position and velocity can be candidates for the initial conditions.
-Fixing time, phase space is $\mathbbm{R}^{2n}$, too big to be covered by position only, but just right to be covered by both.
+Fixing time, phase space has $2N$ components, too big to be covered by position only, but just right to be covered by both.
 

### 2015-03-01 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Adding the notion of cardinality conversion constants

@@
 This restates all the previous findings in terms of conjugate momentum.
-
 In the language of differential geometry, we recognise phase space as the cotangent bundle $\mathbf{T}^*\mathbbm{Q}$, the function $\rho$, the one-form $\theta_0$ and the two-form $\omega$.
 They are truly fundamental objects as they are intimately linked to the way states are defined ($\rho$ is the distribution, $\theta_0 / \hbar$ is the cell width) and counted ($\omega$ is the cardinality of the labels) and they don't depend on coordinate choice.
+
+As we'll need to work with different physical units, which may change $q$ or $p$ independently, it is useful to introduce the following notation and concepts.
+
+\begin{defn}\label{differentialCardinality}
+Let $dq$ be a finite contiguous range for a continuous state variable, we define $n_q(dq)$ as the cardinality of the set $I_{dq}$ that comprise all the labels in the range.
+For contravariant (conjugate) variables, we reverse the indexes in the notation: $n^p(dp)$.
+\end{defn}
+
+\begin{cor}\label{linearCardinality}
+Given $dq = a dq^a + b da^b$, then $n_q(dq) = a n_q(dq^a) + b n_q(da^b)$. That is, the relative cardinality is a linear operator.
+\end{cor}
+
+The linearity should be obvious: as we combine or increase the range, we combine and increase the number of possible labels.\footnote{Since the relative cardinality is linear, invariant under coordinate transformation and conserved by deterministic and reversible motion, it should come to no surprise that linearity will play a significant role.
+This is true for both classical (e.g.
+forces combine linearly) and quantum (e.g.
+linear inner product) concepts that are tightly linked to this fundamental operator.}
+
+\begin{defn}\label{cardinalityConversionConstant}
+Let $q^a$ and $q^b$ be two continuous state variables measured in different units, and $n_a(dq^a)$ and $n_b(dq^b)$ the relative cardinality defined on intervals of the respective units.
+We define the \emph{cardinality conversion constant} $\mathfrak{C}^b_a$ that converts a range of $q^a$ to a range of $q^b$ with the same cardinality.
+That is: $dq^b = \mathfrak{C}^b_a dq^a$ and $n_b(dq^b)=n_a(dq^a)$.
+For contravariant (conjugate) variables, we reverse the indexes in the notation: $dp_b = \mathfrak{C}^a_b dp_a$.
+\end{defn}
+
+As the relative cardinality $n_q(dq)$ is invariant, the function $n_q$ cannot be since $dq$ is covariant.
+This means that, when changing units, we have to be careful.\footnote{This is only a problem when considering the number of labels of a single state variable $n_q(dq)$, instead of a whole degree of freedom $n_{\langle q, p \rangle}(\langle dq, dp \rangle)$: $\langle dq, dp \rangle$ is invariant and so is $n_{\langle q, p \rangle}$.}
+Fortunately, since the $n_q$ is linear, we can always find a coefficient that allows us to convert.
+For example: $\hbar \equiv \mathfrak{C}^k_p$ is the cardinality conversion constant between cell number and conjugate moment.
+In the future, we will have to convert intervals of time into space, or intervals of velocity into conjugate momentum, so the idea of cardinality conversion constants will reappear.
 
@@
 
+\begin{defn}\label{invariantSpeed}
+Let $x$ be a space variable and $t$ the time variable.
+We define the \emph{invariant speed} as the cardinality conversion constant $c \equiv \mathfrak{C}^x_t$ between $x$ and $t$.
+That is, if $dx=cdt$ then $n_x(dx)=n_t(dt)$.
+\end{defn}
+
 \begin{prop}\label{locallyMinkowski}
 The space-time manifold $\mathcal{M}$ is a locally Minkowskian Riemannian manifold.
-That is, there exists a metric $g$ that at any point can be expressed, with a suitable choice of coordinate, as $g=dx^\alpha g_{\alpha \beta}(P) dx^\beta=dx^\alpha\eta_{\alpha \beta}dx^\beta=(dx^i)^2 - (dx^0)^2=(dx^i)^2 - c^2dt^2$, where $c=n(dq)/n(dt)$ is the ratio between the density of states in space and time.
+That is, there exists a metric $g$ that at any point can be expressed, with a suitable choice of coordinate, as $g=dx^\alpha g_{\alpha \beta}(P) dx^\beta=dx^\alpha\eta_{\alpha \beta}dx^\beta=(dx^i)^2 - (dx^0)^2=(dx^i)^2 - c^2dt^2$.
 \end{prop}
@@
 
-\begin{prop}\label{locallyMinkowski}
-The speed of a body cannot exceed $c$ under the kinematic equivalence assumption.
+\begin{prop}\label{speedLimit}
+The speed of a body cannot exceed the invariant speed $c$ under the kinematic equivalence assumption.
 \end{prop}
@@
 Consider a movement $ds$ along any trajectory.
-This will go through $n(dt)$ labels in time and $n(dq)$ labels in space.
-As the motion is continuous, the number of labels in space cannot exceed the number of labels in time, we can't skip them, therefore $n(dq)\leq c n(dt)$.
-So we have:
+This will go through $n_t(dt)$ labels in time and $n_x(dx)$ labels in space.
+As the motion is continuous, the number of labels in space cannot exceed the number of labels in time, we can't skip them:
 \begin{align*}
-\frac{n(dq)}{n(dt)}= \frac{dq}{dt} \leq c \\
+\frac{n_x(dx)}{n_t(dt)} &= \frac{n_x(dx)}{n_x(c dt)} = \frac{dx}{c dt} \leq 1 \\
+\frac{dx}{dt} & \leq c \\
 \end{align*}
+
+Thus we find that $c$ is the well known relativistic constant.\footnote{Deterministic and reversible motion at speeds greater than $c$ (i.e.
+tachyon) is not allowed.
+Non-deterministic and non-reversible motion motion at speeds greater than $c$ (i.e.
+correlations at a distant) is allowed.}
 
@@
 The ratio $m=n(dp_\alpha)/n(du_\alpha)$ between their relative label cardinality is called \emph{inertial mass}.
+\end{defn}
+
+\begin{defn}\label{invariantSpeed}
+Let $p_\alpha$ be a component of conjugate momentum and $u_\alpha \equiv g_{\alpha \beta}u^\beta$ be a contravariant component of the four velocity.
+We define the \emph{inertial mass} as the cardinality conversion constant $m \equiv \mathfrak{C}^p_u$ between $p$ and $tu$.
+That is, if $dp=mdu$ then $n^p(dp)=n^u(du)$.
 \end{defn}
@@
   \item Deterministic and reversible evolution is what ultimately gives space-time its locally Minkowskian nature.
+  \item $\hbar$, $c$ and $m$ can be seen as conversion constants that preserve cardinality
   \item Lagrangian mechanics coincides with position and velocity being necessary and sufficient initial conditions.

### 2015-03-12 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Removed duplicated paragraph

@@
 \begin{defn}\label{inertialMass}
-Let $dp_\alpha$ an infinitesimal interval of conjugate momentum and $du_\alpha\equiv d(g_{\alpha \beta}u^\beta)$ an infinitesimal interval of the covariant components of the four velocity.
-The ratio $m=n(dp_\alpha)/n(du_\alpha)$ between their relative label cardinality is called \emph{inertial mass}.
-\end{defn}
-
-\begin{defn}\label{invariantSpeed}
 Let $p_\alpha$ be a component of conjugate momentum and $u_\alpha \equiv g_{\alpha \beta}u^\beta$ be a contravariant component of the four velocity.
-We define the \emph{inertial mass} as the cardinality conversion constant $m \equiv \mathfrak{C}^p_u$ between $p$ and $tu$.
-That is, if $dp=mdu$ then $n^p(dp)=n^u(du)$.
+We define the \emph{inertial mass} as the cardinality conversion constant $m \equiv \mathfrak{C}^p_u$ between $p$ and $u$.
+That is, if $dp_\alpha=mdu_\alpha$ then $n^p(dp_\alpha)=n^u(du_\alpha)$.
 \end{defn}

### 2015-03-12 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
After Christine comments.

@@
 This perspective links different concepts from different branches of math and physics (e.g.
-cardinality of a set, cotangent bundle for phase space, Hamiltonian flow, locally Minkowskian space-time manifold) providing new insights.
-The derivation strives to use definition and mathematical concepts compatible with a future extensions to field theories and quantum mechanics.\end{abstract}
+cardinality of a set, cotangent bundle for phase space, Hamiltonian flow, locally Minkowskian space-time manifold), providing new insights.
+The derivation strives to use definitions and mathematical concepts compatible with future extensions to field theories and quantum mechanics.\end{abstract}
 \maketitle
@@
 conservative forces) are introduced.
-Special relativity is based on two principles (invariance of the speed of light and the principle of relativity), which lead to the Minkowskian nature of space-time but not to the equation of motion, which are a consistent reformulation of the non-relativistic ones.
+Special relativity is based on two principles (invariance of the speed of light and the principle of relativity), which lead to the Minkowskian nature of space-time but not to the equations of motion, which are a consistent reformulation of the non-relativistic ones.
 What we'll do in this work is re-organize all the known elements and equations in a more consistent and comprehensive way, leading to better insight on why the fundamental concepts and laws are what they are.
@@
 
-We'll use concepts from different disciplines, such as set theory, differential geometry, relativity, Hamiltonian and Lagrangian mechanics, and we'll find interesting connections between them.
+We'll use concepts from different disciplines, such as set theory, differential geometry, relativity, Hamiltonian and Lagrangian mechanics, and we'll find interesting connections among them.
 We'll keep names and notation as consistent as possible to current use across the different disciplines.
-This may sometime lead to some non sequitur as it will not be immediately clear why the new definitions are equivalent to the standard ones.
+This may sometimes lead to some non sequitur as it will not be immediately clear why the new definitions are equivalent to the standard ones.
 These are typically resolved by subsequent derivation of the expected properties.
@@
 To be invertible, the relationship between velocity and conjugate momentum has to be monotonic, leading to a concave Hamiltonian which allows a Legendre transformation leading to the Lagrangian.
-We can constrain the Hamiltonian further, and show that the most general equation of motion under the three assumptions is the one of a geodesic modified by the force given by a vector potential (such as the one of a relativistic charged particle) and a scalar potential (such as the one of newtonian gravitation in the non-relativistic case).
+We can constrain the Hamiltonian further, and show that the most general equation of motion under the three assumptions is the one of a geodesic modified by the force given by a vector potential (such as the one of a relativistic charged particle) and a scalar potential (such as the one of Newtonian gravitation in the non-relativistic case).
 
@@
 
-This and the next section are dedicated to properly define states.
+This and the next section are dedicated to properly defining states.
 In particular, we will need to introduce more precise terminology to be able to make two crucial distinctions.
@@
 
-The second distinction is, when referring to a physical quantity, between the space of all possible values, a particular value or a set of possible values.
+The second distinction comes up when dealing with physical quantities. We need to be able to distinguish among a particular value, a set of possible values, and the space of all possible values.
 \begin{center}
@@
 \begin{assump1}\label{classical}
-The system is infinitely reducible: it can be thought as composed by two or more similar but smaller systems, each in its own configuration state, which also can be thought as composed by two or more, ad infinitum.
+The system is infinitely reducible: it can be thought of as composed of two or more similar but smaller systems, each in its own configuration state, which in turn can be thought of as composed of two or more, ad infinitum.
 We call \emph{particle} such an infinitesimal part.
@@
 A cell of $\mathbbm{R}$ corresponds to $\mathbbm{R}^2$: a center and a width.\footnote{In other words: the particles of the system are not point-like, but infinitesimal cell-like.
-Not only this leads to a more direct understanding of classical phase-space, it is also more consistent with general relativity (the mass is spread across a small region, not forming a singularity) and quantum mechanics (position is always the central value of a small distribution).} With this in mind, we have the following definitions.
+Not only does this lead to a more direct understanding of classical phase-space, it is also more consistent with general relativity (the mass is spread across a small region, not forming a singularity) and quantum mechanics (position is always the central value of a small distribution).} With this in mind, we have the following definitions.
 
 \begin{defn}\label{sdof}
-A \emph{degree of freedom} is a state variable identified by an infinitesimal cell of a one dimensional manifold $\mathbbm{Q}$.\footnote{For each label $i$ there is one and only infinitesimal cell.}
+A \emph{degree of freedom} is a state variable identified by an infinitesimal cell of a one dimensional manifold $\mathbbm{Q}$.\footnote{For each label $i$ there is one and only one infinitesimal cell.}
 \end{defn}
 
-\begin{defn}\label{sdof}
+\begin{defn}\label{generalizedCoordinates}
 We define the \emph{generalized coordinate} $q$ as the center of each cell.
@@
 \begin{cor}\label{continuousLabels}
-Each degree of freedom is identified by the pair of labels $\langle q,k \rangle \in \mathbf{T}^*\mathbbm{Q}$.
+Each degree of freedom is identified by the pair of labels $\langle q,k \rangle \in \mathbf{T}^*\mathbbm{Q}$ where $\mathbf{T}^*\mathbbm{Q}$ is the cotangent bundle of $\mathbbm{Q}$.
 Under coordinate changes, $k\,dq$ and $dq\wedge dk$ are invariant and $k$ is contravariant.
@@
 Let $Q$ be a finite region of $\mathbbm{Q}$.
-Divide the region in $N$ equal intervals of center $q$ and length $\Delta q$.\footnote{The definition could be mathematically more general.
+Divide the region into $N$ equal intervals of center $q$ and length $\Delta q$.\footnote{The definition could be mathematically more general.
 We use equally spaced intervals as it makes the derivation less cumbersome.} Let $K$ be a finite region of $\mathbbm{R}$.
@@
 
-If we apply a coordinate change $q'=q'(q)$, we change the labels but the state must remain the same, defined at the same point with the same width.\footnote{The notion that states are what they are no matter how we label them already contains the principle of relativity.} $k' dq'$ must be then equal to $k dq$; $k'= k dq / dq'$ is contravariant, $dq' \wedge dk'= dq'/dq \, dq \wedge dk \, dq/dq' = dq \wedge dk$ is invariant.
+If we apply a coordinate change $q'=q'(q)$, we change the labels but the state must remain the same, defined at the same point with the same width.\footnote{The notion that states are what they are no matter how we label them already contains the principle of relativity.} $k' dq'$ must be then equal to $k dq$; $k'= k dq / dq'$ is contravariant, $dq' \wedge dk'= (dq'/dq) \, (dq \wedge dk) \, (dq/dq') = dq \wedge dk$ is invariant.
 This means that $\rho$ is also invariant, which makes sense: the density depends only on the state (the cell), and not the coordinate chosen to represent them.
@@
 
-\begin{prop}\label{relativeCardinality}
+\begin{defn}\label{relativeCardinality}
 Let $I \subset \mathbf{T}^*\mathbbm{Q}$ be a closed dense set of labels, subset of a degree of freedom (i.e.
 a region of phase space).
-We define \emph{relative cardinality} $n(I) \rightarrow \mathbbm{R}$ the ratio between the number of labels in $I$ and the ones of a reference set $I_0$.
+We define \emph{relative cardinality} $n(I) \rightarrow \mathbbm{R}$ as the ratio between the number of labels in $I$ and the ones of a reference set $I_0$.
 We find $n(I)=\int \omega$, where $\omega = \hbar \, dq \wedge dk$ and $\hbar$ is the constant that determines the unit (i.e.
 fixed so that $n(I_0) = 1$).
-\end{prop}
+\end{defn}
 
@@
 We use this arbitrariness to set the value of our classical analogue of $\hbar$ to the known constant.} In general, the label set is not rectangular in $\langle q,k \rangle$ and we have $n(I)=\int \hbar dq \wedge dk$.\footnote{It should be stressed that no notion of uncertainty is used here.
-Though there is a link between label cardinality and uncertainty, which we do not explore here for brevity, label cardinality is the more fundamental as we can still count the number of possible states even in an ideal case of no uncertainty.}
+Though there is a link between label cardinality and uncertainty, which we do not explore here for brevity, label cardinality is the more fundamental quantity as we can still count the number of possible states even in an ideal case of no uncertainty.}
 
@@
 This restates all the previous findings in terms of conjugate momentum.
-In the language of differential geometry, we recognise phase space as the cotangent bundle $\mathbf{T}^*\mathbbm{Q}$, the function $\rho$, the one-form $\theta_0$ and the two-form $\omega$.
+In the language of differential geometry, we recognize phase space as the cotangent bundle $\mathbf{T}^*\mathbbm{Q}$, the function $\rho$, the one-form $\theta_0$ and the two-form $\omega$.
 They are truly fundamental objects as they are intimately linked to the way states are defined ($\rho$ is the distribution, $\theta_0 / \hbar$ is the cell width) and counted ($\omega$ is the cardinality of the labels) and they don't depend on coordinate choice.
@@
 \begin{defn}\label{differentialCardinality}
-Let $dq$ be a finite contiguous range for a continuous state variable, we define $n_q(dq)$ as the cardinality of the set $I_{dq}$ that comprise all the labels in the range.
-For contravariant (conjugate) variables, we reverse the indexes in the notation: $n^p(dp)$.
+Let $dq$ be a finite contiguous range for a continuous state variable, we define $n_q(dq)$ as the cardinality of the set $I_{dq}$ that comprises all the labels in the range.
+For contravariant (conjugate) variables, we reverse the indices in the notation: $n^p(dp)$.
 \end{defn}
@@
 \begin{cor}\label{linearCardinality}
-Given $dq = a dq^a + b da^b$, then $n_q(dq) = a n_q(dq^a) + b n_q(da^b)$. That is, the relative cardinality is a linear operator.
+Given $dq = a dq^a + b dq^b$, then $n_q(dq) = a n_q(dq^a) + b n_q(da^b)$. That is, the relative cardinality is a linear operator.
 \end{cor}
 
-The linearity should be obvious: as we combine or increase the range, we combine and increase the number of possible labels.\footnote{Since the relative cardinality is linear, invariant under coordinate transformation and conserved by deterministic and reversible motion, it should come to no surprise that linearity will play a significant role.
+The linearity should be obvious: as we combine or increase the range, we combine or increase the number of possible labels.\footnote{Since the relative cardinality is linear, invariant under coordinate transformation and conserved by deterministic and reversible motion, it should come as no surprise that linearity will play a significant role.
 This is true for both classical (e.g.
@@
 That is: $dq^b = \mathfrak{C}^b_a dq^a$ and $n_b(dq^b)=n_a(dq^a)$.
-For contravariant (conjugate) variables, we reverse the indexes in the notation: $dp_b = \mathfrak{C}^a_b dp_a$.
+For contravariant (conjugate) variables, we reverse the indices in the notation: $dp_b = \mathfrak{C}^a_b dp_a$.
 \end{defn}
@@
 Fortunately, since the $n_q$ is linear, we can always find a coefficient that allows us to convert.
-For example: $\hbar \equiv \mathfrak{C}^k_p$ is the cardinality conversion constant between cell number and conjugate moment.
+For example: $\hbar \equiv \mathfrak{C}^k_p$ is the cardinality conversion constant between cell number and conjugate momentum.
 In the future, we will have to convert intervals of time into space, or intervals of velocity into conjugate momentum, so the idea of cardinality conversion constants will reappear.
@@
 As these must preserve the cardinality of label sets, they are constrained in their form.
-The use of a bijective map for infinitesimal time evolution will lead us to Hamilton's equation.
-In this section we'll study a single degree of freedom, then extend in the next two section to multiple degrees of freedom and to the time dependent case.
+The use of a bijective map for infinitesimal time evolution will lead us to Hamilton's equations.
+In this section we'll study a single degree of freedom, then extend in the next two sections to multiple degrees of freedom and to the time dependent case.
 
@@
 
-As we saw \ref{labelsCount}, a bijective map conserves the cardinality of labels.
+As we saw in \ref{labelsCount}, a bijective map conserves the cardinality of labels.
 In the continuous case, it will conserve relative cardinality and therefore $\omega$.
-The map must be continuous in q: suppose it isn't, it would split some cells into two parts, a cell would not be mapped to one and only one other cell, the mapping would not be bijective.
-The reverse mapping must be continuous in q as well, or the inverse would not map to one and only one cell.
+The map must be continuous in $q$: suppose it isn't, it would split some cells into two parts, a cell would not be mapped to one and only one other cell, the mapping would not be bijective.
+The reverse mapping must be continuous in $q$ as well, or the inverse would not map to one and only one cell.
 \begin{align*}
@@
 We can re-express $dp'$ in terms of $dq$ and $dp$, as the conservation of $\omega$ means the map is non-degenerate.
-All partial derivatives are well defined, and therefore the mapping is continuous in p as well.
+All partial derivatives are well defined, and therefore the mapping is continuous in $p$ as well.
 
@@
 \begin{align*}
-\omega_{a, b} = \left[
+\omega_{ab} = \left[
   \begin{array}{cc}
@@
 \end{align*}
-then $v'^{a} \omega_{a, b} w'^{b}=v^{a} \omega_{a, b} w^{b}$ under a canonical transformation, where $a$ and $b$ represent component along coordinates $\xi^{a} \equiv \{q,p\}$.
+then $v'^{a} \omega_{ab} w'^{b}=v^{a} \omega_{ab} w^{b}$ under a canonical transformation, where $a$ and $b$ represent components along coordinates $\xi^{a} \equiv \{q,p\}$.
 \end{cor}
 
-Here we are simply expressing $\omega$, $v$ and $w$ with their respective components and underlying the fact that $\omega$ defines the metric conserved under canonical transformations.
+Here we are simply expressing $\omega$, $v$ and $w$ with their respective components and underlining the fact that $\omega$ defines the metric conserved under canonical transformations.
 
@@
 Let $v$ and $w$ be two vectors.
-Let $v^{a} \omega_{a, b} w^{b}$ be an antisymmetric product conserved under a continuous transformation parameterized by $t$.
-We can then define a function $H$ such that given $S^{a} \equiv d_{t}\xi^{a}$ and $S_{b} \equiv S^{a} \omega_{a, b}$, we have $S_{a} = \partial_{a}H$.
+Let $v^{a} \omega_{ab} w^{b}$ be an antisymmetric product conserved under a continuous transformation parameterized by $t$.
+We can then define a function $H$ such that given $S^{a} \equiv d_{t}\xi^{a}$ and $S_{b} \equiv S^{a} \omega_{ab}$, we have $S_{a} = \partial_{a}H$.
 \end{lem}
@@
 \begin{align*}
-v^{a} \omega_{a, b} w^{b} &= v'^{a} \omega_{a, c} w'^{b}  \\
-&= (v^{a} + \partial_{c} S^{a} dt v^{c}) \omega_{a, b} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
-&= v^{a} \omega_{a, b} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{a, b} w^{b} \\
- &+ v^{a} \omega_{a, b} \partial_{d} S^{b} w^{d}) dt + O(dt^2)
+v^{a} \omega_{ab} w^{b} &= v'^{a} \omega_{ab} w'^{b}  \\
+&= (v^{a} + \partial_{c} S^{a} dt v^{c}) \omega_{ab} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
+&= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} \\
+ &+ v^{a} \omega_{ab} \partial_{d} S^{b} w^{d}) dt + O(dt^2)
 \end{align*}
@@
 Let $v$ and $w$ be two vectors defined on the tangent space of the phase space $\mathbf{T}^*\mathbbm{Q}$ for two independent degrees of freedom.
-Let $a$ and $b$ be indexes for the state variables $\xi^a\equiv \{q^i, p_i\}$.
+Let $a$ and $b$ be indices for the state variables $\xi^a\equiv \{q^i, p_i\}$.
 Let
 \begin{align*}
-\omega_{\alpha, \beta} =  \left[
+\omega_{ab} =  \left[
   \begin{array}{cc}
@@
 \end{align*}
-then $v'^{a} \omega_{a, b} w'^{b}=v^{a} \omega_{a, b} w^{a}$ under a canonical transformation.
+then $v'^{a} \omega_{a b} w'^{b}=v^{a} \omega_{ab} w^{b}$ under a canonical transformation.
 \end{prop}
@@
 (i.e.
-the area), must be equal to the number of combines labels (i.e.
+the area), must be equal to the number of combined labels (i.e.
 the hyper-volume), which is true only if the d.o.f are orthogonal in phase space.
@@
 
-So far we have assumed that neither state labeling nor mapping change in time. If they do, we also need to to use time as a label and therefore introduce an extra degree of freedom.
+So far we have assumed that neither state labeling nor mapping changes in time. If they do, we also need to to use time as a label and therefore introduce an extra degree of freedom.
 
@@
 Let $v$ and $w$ be two vectors defined on the tangent space of extended phase space $\mathbf{T}^*\mathcal{M}$ for the temporal degree of freedom and one standard degree of freedom.
-Let $a$ and $b$ be indexes for the state variables $\xi^a\equiv\{t, q, E, p\}$.
+Let $a$ and $b$ be indices for the state variables $\xi^a\equiv\{t, q, E, p\}$.
 Let
 \begin{align*}
-\omega_{a, b} =  \left[
+\omega_{ab} =  \left[
   \begin{array}{cc}
@@
 \end{align*}
-then $v'^{a} \omega_{a, b} w'^{b}=v^{a} \omega_{a, b} w^{b}$ under deterministic and reversible evolution.
+then $v'^{a} \omega_{ab} w'^{b}=v^{a} \omega_{ab} w^{b}$ under deterministic and reversible evolution.
 \end{prop}
@@
 States are defined on the plane where $\langle q, p \rangle$ (maximally) change: this is not the plane of constant $\langle t, E \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle q, p \rangle$ where $dt \wedge dE$ is defined.
-On that plane we can properly count states and define our invariant.
+On that plane we can properly count states and define our metric.
 
-We have a right triangle-like relationship between the plane where the invariant is defined and its projections on the planes defined by each d.o.f., similar to the multiple d.o.f.:
+We have a right triangle-like relationship between the plane where the metric is defined and its projections on the planes defined by each d.o.f., similar to the multiple d.o.f.:
 \begin{align*}
-m.d.o.f \;\;\; &dq^1 \wedge dp_1 + dq^2 \wedge dp_2 = k \\
-t.d.o.f \;\;\; &dt \wedge dE + k = dq \wedge dp \\
+m.d.o.f \;\;\; &dq^1 \wedge dp_1 + dq^2 \wedge dp_2 = \omega \\
+t.d.o.f \;\;\; &dt \wedge dE + \omega = dq \wedge dp \\
 \end{align*}
 But in the previous case, the right angle was between the two independent d.o.f..
-In this case, the right angle is between the invariant and the plane of constant $\langle q, p \rangle$ where $dt \wedge dE$ is defined.
-We rewrite it as $dq \wedge dp - dt \wedge dE = k$.
+In this case, the right angle is between the metric and the plane of constant $\langle q, p \rangle$ where $dt \wedge dE$ is defined.
+We rewrite it as $dq \wedge dp - dt \wedge dE = \omega$.
 This corresponds to the Minkowski product across d.o.f.
 and the vector product within.
-The metric, with a space-like convention, still gives us the cardinality of labels within a degree of freedom, adjusting \ref{sdofInvariant} to avoid "double counting".
+The metric \ref{tdofInvariant}, with a space-like convention, still gives us the cardinality of labels within a degree of freedom, adjusting \ref{sdofInvariant} to avoid "double counting".
 
@@
 
-If we set\footnote{We avoided using $p_{n+1}$ as it hides the minus sign from the metric, making it seem that the temporal d.o.f is just another independent d.o.f.} $q^{n+1}=t$ and $p_{n+1}=-E$, we recognise Hamilton equations in the extended phase space\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}\cite{Synge,Lanczos}.
+If we set\footnote{We avoided using $p_{n+1}$ as it hides the minus sign from the metric, making it seem that the temporal d.o.f is just another independent d.o.f.} $q^{n+1}=t$ and $p_{n+1}=-E$, we recognize Hamilton equations in the extended phase space\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}\cite{Synge,Lanczos}.
 
@@
 \begin{assump3}\label{kinematicAssumption}
-The study of the trajectory (kinematic) of a body is equivalent to study its state (dynamic) under deterministic and reversible evolution.
+The study of the trajectory (kinematics) of a body is equivalent to study of its state (dynamics) under deterministic and reversible evolution.
 \end{assump3}
@@
 Fix a system under deterministic and reversible evolution.
-Given all its possible trajectories $x^\alpha(s)$ in the space-time manifold $\mathcal{M}$ and all its possible trajectory $\xi^a(s)$ in the extended phase space $\mathbf{T}^*\mathcal{M}$, there exist a bijective function $f: x^\alpha(s) \leftrightarrow \xi^a(s)$ that links each space-time trajectory with one and only one phase space trajectory.
+Given all its possible trajectories $x^\alpha(s)$ in the space-time manifold $\mathcal{M}$ and all its possible trajectories $\xi^a(s)$ in the extended phase space $\mathbf{T}^*\mathcal{M}$, there exists a bijective function $f: x^\alpha(s) \leftrightarrow \xi^a(s)$ that links each space-time trajectory with one and only one phase space trajectory.
 \end{cor}
@@
 If studying the motion and state evolution are equivalent, then we must be able to go back and forth between the two pictures.
-Without losing generality, we can chose $x^\alpha=f(q^i,t)$ to be a linear function of only $q^i$ and $t$.
+Without losing generality, we can choose $x^\alpha=f(q^i,t)$ to be a linear function of only $q^i$ and $t$.
 
@@
 The space-time manifold $\mathcal{M}$ is a locally Minkowskian Riemannian manifold.
-That is, there exists a metric $g$ that at any point can be expressed, with a suitable choice of coordinate, as $g=dx^\alpha g_{\alpha \beta}(P) dx^\beta=dx^\alpha\eta_{\alpha \beta}dx^\beta=(dx^i)^2 - (dx^0)^2=(dx^i)^2 - c^2dt^2$.
+That is, there exists a metric $g$ that at any point P can be expressed, with a suitable choice of coordinate, as $g=dx^\alpha g_{\alpha \beta}(P) dx^\beta=dx^\alpha\eta_{\alpha \beta}dx^\beta=(dx^i)^2 - (dx^0)^2=(dx^i)^2 - c^2dt^2$, where $\eta_{\alpha \beta}$ is the Minkowski metric.
 \end{prop}
@@
 Such a metric must be consistent with $\omega$ as defined on $\mathbf{T}^*\mathcal{M}$, as both must be invariant under coordinate transformations.
-The idea is that each $(dx^\alpha)^2$ can be made to both represent a length square in space-time and an area in phase space, linking the two metrics.
+The idea is that each $(dx^\alpha)^2$ can be made to represent both a length squared in space-time and an area in phase space, linking the two metrics.
 For each spatial d.o.f.
@@
 For the temporal d.o.f fix $dE = \lambda c^2 dt$, we have $dt \wedge dE = \lambda c^2 dt^2$.
-That is: the area in time squared is converted to an area in length squared that have the same density of states (as per definition of $c$) and then to label cardinality.
-The phase space invariant is $\omega = \lambda [(dq^i)^2 - c^2 dt^2]$.
+That is: the area in time squared is converted to an area in length squared that has the same density of states (as per definition of $c$) and then converted again to its label cardinality.
+The phase space metric is $\omega = \lambda [(dq^i)^2 - c^2 dt^2]$.
 
@@
 Consider a movement $ds$ along any trajectory.
-This will go through $n_t(dt)$ labels in time and $n_x(dx)$ labels in space.
-As the motion is continuous, the number of labels in space cannot exceed the number of labels in time, we can't skip them:
+This will go through $|n_t(dt)|$ labels in time and $|n_x(dx)|$ labels in space.
+As we go through labels in time, we may go through fewer labels in space (e.g.
+the particle remains still) but not more: we cannot skip space labels or the motion would not be continuous.
+So we have:
 \begin{align*}
-\frac{n_x(dx)}{n_t(dt)} &= \frac{n_x(dx)}{n_x(c dt)} = \frac{dx}{c dt} \leq 1 \\
-\frac{dx}{dt} & \leq c \\
+\frac{|n_x(dx)|}{|n_t(dt)|} &= \frac{|n_x(dx)|}{|n_x(c dt|)} = \frac{|dx|}{|c dt|} \leq 1 \\
+\frac{|dx|}{|dt|} & \leq c \\
 \end{align*}
@@
 tachyon) is not allowed.
-Non-deterministic and non-reversible motion motion at speeds greater than $c$ (i.e.
-correlations at a distant) is allowed.}
+Non-deterministic and non-reversible motion at speeds greater than $c$ (e.g.
+correlations at a distance) is allowed.}
 
@@
 
-As the equation of motion \ref{tdofHam} can be at most second order in $\{t, q^i\}$, they can at most be second order in $x^\alpha=\{ct, q^i\}$ so only position and velocity can be candidates for the initial conditions.
+As the equations of motion \ref{tdofHam} can be at most second order in $\{t, q^i\}$, they can at most be second order in $x^\alpha=\{ct, q^i\}$ so only position and velocity can be candidates for the initial conditions.
 Fixing time, phase space has $2N$ components, too big to be covered by position only, but just right to be covered by both.
@@
 We define the \emph{inertial mass} as the cardinality conversion constant $m \equiv \mathfrak{C}^p_u$ between $p$ and $u$.
-That is, if $dp_\alpha=mdu_\alpha$ then $n^p(dp_\alpha)=n^u(du_\alpha)$.
+That is, if $dp_\alpha=mdu_\alpha$ then $n^p(dp_\alpha)=n^u(du_\alpha)$.\footnote{Note that \ref{kinematicAssumption} rules out massless particles: they can have the same position/velocity with different momentum, therefore the same trajectory for different states.
+When generalizing to field theory, the derivative of the position will be replaced by the derivative of the field (i.e.
+its oscillation in space-time), which holds for massive and massless fields.}
 \end{defn}
@@
 Let $x^\alpha=\{ct, q^i\}$ and $p_\alpha=\{-E/c, p_i\}$.
-Then $p_\alpha= m g_{\alpha \beta}u^\beta + \hat{p}_\alpha(x^\gamma)$, where $\hat{p}_\alpha : \mathcal{M} \rightarrow \mathbbm{R}$ is a function defined on the space-time manifold.
+Then $p_\alpha= m g_{\alpha \beta}u^\beta + \hat{p}_\alpha(x)$, where $\hat{p}_\alpha : \mathcal{M} \rightarrow \mathbbm{R}$ is a function defined on the space-time manifold.
 \end{prop}
@@
 \end{align*}
-where $\hat{p}$ is an arbitrary function.
+where $\hat{p}$ is an arbitrary function.\footnote{The conjugate momentum $p$ and the kinetic momentum $mu$ are therefore different in general: only their differentials are linked.
+Moreover: $p$ is unique only once $\hat{p}$ is fixed, but $\hat{p}$ is arbitrary.
+Given that only $u$ is linked to the trajectory, any transformation in $\{p, \hat{p} \}$ that does not change $u$ describes the same motion.
+We call \emph{gauge transformation} such transformation that redefines conjugate momentum while keeping kinetic momentum unvaried.}
 
@@
 \begin{prop}\label{relativisticEM}
-Let $\hat{p}_\alpha = q A_\alpha$, $F_{\alpha \beta} \equiv \partial_\alpha A_\beta - \partial_\beta A_\alpha$ and $\hat{\mathcal{H}} = 0$, the equations of motion are $m \nabla_u u^\alpha = g^{\alpha\beta} F_{\beta \gamma} q u^\gamma$.
+Let $\hat{p}_\alpha = q A_\alpha$, $F_{\alpha \beta} \equiv \partial_\alpha A_\beta - \partial_\beta A_\alpha$ and $\hat{\mathcal{H}} = 0$.
+Then the equations of motion are $m \nabla_u u^\alpha = g^{\alpha\beta} F_{\beta \gamma} q u^\gamma$.
 These are the relativistic equations for a charged particle.
@@
 \end{align*}
+where $\nabla$ is the covariant derivative and $\Gamma$ the Christoffel symbols.
 
 The equation is manifestly covariant.
-If $\hat{p}$ and $\hat{\mathcal{H}}$ are zero, absence of forces, we recognize the geodesic equation.
+If $\hat{p}$ and $\hat{\mathcal{H}}$ are zero (i.e.
+in the absence of forces) we recognize the geodesic equation.
 We substitute $\hat{p}_\alpha = q A_\alpha$ and $\hat{\mathcal{H}} = 0$ and have:
@@
 
-This is indeed compatible with general relativity and classical electro-magnetism.
+This is indeed compatible with general relativity and classical electromagnetism.
 
 \begin{prop}\label{newtonianGravitation}
-Assume $\mathcal{M}$ flat and time independent motion, the Hamiltonian simplifies to $H=\frac{1}{2m}(p_i-\hat{p}_\alpha(x))^2)+\hat{H}(x)$.
+Assume $\mathcal{M}$ flat and time independent motion.
+Then the Hamiltonian simplifies to $H=\frac{1}{2m}(p_i-\hat{p}_\alpha(x))^2+\hat{H}(x)$.
 Let $\hat{p}_i = 0$ and $\hat{H} = mV$, the equations of motion are $m d_t v^i = - m \partial_i V$, where $v^i=dx^i/dt$.
@@
 \begin{defn}\label{lagrangian}
-Let $\mathcal{H}$ be an extended hamiltonian defined on $\mathbf{T}^*\mathcal{M}$.
-Under the kinetic assumption, we can define the Legendre transform $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$ which we call \emph{extended Lagrangian}.
+Let $\mathcal{H}$ be an extended Hamiltonian defined on $\mathbf{T}^*\mathcal{M}$.
+Under the kinetic assumption, we can define the Legendre transform $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$ which we call the \emph{extended Lagrangian}.
 For the time invariant case, we define the \emph{Lagrangian} $L=v^i p_i - H$.
@@
 What happens in that case is that the space-time trajectories may become more or less dense, $du_\alpha$ and $dx^\alpha$ would not be enough to define the label cardinality and therefore the state density $\rho$.
-In short: deterministic and reversible evolution gives us Hamiltonian mechanics, position and velocity as initial conditions Lagrangian mechanics, and the kinetic assumption (which links the differentials of the state variables and the initial conditions) gives us the narrowed Hamiltonian form.
+In short: deterministic and reversible evolution gives us Hamiltonian mechanics, position and velocity as initial conditions gives us Lagrangian mechanics, and the kinetic assumption (which links the differentials of the state variables and the initial conditions) gives us the narrowed Hamiltonian form.
 
@@
 
-While there are many more details that we could have expanded up, as we have touched many areas we could only scratch the surface, the intent is to give the overall picture, which could be summarized in the following points.
+While there are many more details that we could have expanded upon, as we have touched many areas we could only scratch the surface.
+The intent is to give the overall picture, which could be summarized in the following points.
 
@@
 While this is limited to classical particle mechanics, it should be obvious how, at least in principle, this work could be extended.
-For field theories, the \emph{kinematic assumption} should be substituted by a \emph{kinematic field assumption}: what we are studying are not trajectories $x^\alpha(s)$ but fields $\psi(x^\alpha)$.
-The field values at each point become a set of independent state variables and their conjugates $\pi(x^\alpha)$ will be linked to $d_s\psi(x^\alpha)$.
-For quantum mechanics, the \emph{infinite reducibility assumption} has to leave way to an \emph{irreducibility assumption}: the state of parts of a quantum system cannot be known.
+For field theories, the \emph{kinematic assumption} should be substituted by a \emph{kinematic field assumption}: what we are studying are not trajectories $x^\alpha(s)$ but fields $\psi(x)$.
+The field values at each point become a set of independent state variables, and their conjugates $\pi(x^\alpha)$ will be linked to $d\partial_s\psi(x^\alpha)$.
+For quantum mechanics, the \emph{infinite reducibility assumption} has to give way to an \emph{irreducibility assumption}: the state of parts of a quantum system cannot be known.
 The configuration state as a whole undergoes deterministic and reversible evolution, while the motion of its parts does not.
@@
 
+\bibitem{Shannon} Shannon, C. E.: ``A mathematical theory of communications'', The Bell System Technical Journal, Vol. 27, pp. 379–423, 623–656, (1948)
 \bibitem{Jaynes} Jaynes, E. T.: ``Information theory and statistical mechanics'', (1963)
-\bibitem{Shannon} Shannon, C. E.: ``A mathematical theory of communications'', The Bell System Technical Journal, Vol. 27, pp. 379–423, 623–656, (1948)
 \bibitem{classical_dynamics} J. V. Jose', E. J. Saletan: ``Classical Dynamics'', Cambridge University Press, (1998)

### 2015-03-15 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Acknowledgements, bibliography and other minor edits

@@
 
-Here we are simply expressing $\omega$, $v$ and $w$ with their respective components and underlining the fact that $\omega$ defines the metric conserved under canonical transformations.
+Here we are simply expressing $\omega$, $v$ and $w$ in terms of their respective components and underlining the fact that $\omega$ defines the metric conserved under canonical transformations.
 
@@
 We define the \emph{inertial mass} as the cardinality conversion constant $m \equiv \mathfrak{C}^p_u$ between $p$ and $u$.
-That is, if $dp_\alpha=mdu_\alpha$ then $n^p(dp_\alpha)=n^u(du_\alpha)$.\footnote{Note that \ref{kinematicAssumption} rules out massless particles: they can have the same position/velocity with different momentum, therefore the same trajectory for different states.
+That is, if $dp_\alpha=mdu_\alpha$ then $n^p(dp_\alpha)=n^u(du_\alpha)$.\footnote{Note that the kinematic assumption rules out massless particles: they can have the same position and velocity with different momentum, therefore the same trajectory for different states.
 When generalizing to field theory, the derivative of the position will be replaced by the derivative of the field (i.e.
@@
 \frac{\partial p_\alpha}{\partial u_\beta} &= m \delta^\beta_\alpha \\
-p_\alpha &= m g_{\alpha \beta}u^\beta + \hat{p}_\alpha(x^\gamma)
+p_\alpha &= m g_{\alpha \beta}u^\beta + \hat{p}_\alpha(x)
 \end{align*}
-where $\hat{p}$ is an arbitrary function.\footnote{The conjugate momentum $p$ and the kinetic momentum $mu$ are therefore different in general: only their differentials are linked.
-Moreover: $p$ is unique only once $\hat{p}$ is fixed, but $\hat{p}$ is arbitrary.
-Given that only $u$ is linked to the trajectory, any transformation in $\{p, \hat{p} \}$ that does not change $u$ describes the same motion.
-We call \emph{gauge transformation} such transformation that redefines conjugate momentum while keeping kinetic momentum unvaried.}
+where $\hat{p}$ is an arbitrary function.\footnote{The conjugate momentum $p$ and the kinetic momentum $mu$ are therefore different in general: they are linked only through their differentials.
+Moreover, $p$ is unique only once $\hat{p}$ is fixed, but $\hat{p}$ is arbitrary.
+Given that $u$ determines the trajectory, any transformation in $\{p, \hat{p} \}$ that does not change $u$ describes the same motion.
+We call \emph{gauge transformation} such a transformation that redefines conjugate momentum while keeping kinetic momentum unvaried.
+This is the particle mechanics equivalent of gauge transformations in field theory.}
 
@@
 
+\section{Acknowledgements}
+
+It is my pleasure to thank Christine A. Aidala, Lydia Bieri, Robert Geroch and Steven J. Miller for helpful discussions and support.
+
 \begin{thebibliography}{0}
 
-\bibitem{Shannon} Shannon, C. E.: ``A mathematical theory of communications'', The Bell System Technical Journal, Vol. 27, pp. 379–423, 623–656, (1948)
-\bibitem{Jaynes} Jaynes, E. T.: ``Information theory and statistical mechanics'', (1963)
-\bibitem{classical_dynamics} J. V. Jose', E. J. Saletan: ``Classical Dynamics'', Cambridge University Press, (1998)
-\bibitem{Gromov} Gromov, M. L.: ``Pseudo holomorphic curves in symplectic manifolds''. Inventiones Mathematicae 82: 307–347, (1985)
-\bibitem{deGosson} de Gosson, M. A.: ``The symplectic camel and the uncertainty principle: the tip of an iceberg?'', Foundation of Physics 39, pp. 194–214, (2009)
-\bibitem{Stewart} Stewart, I.: ``The symplectic camel'', Nature 329(6134), 17–18 (1987)
-\bibitem{Lanczos} Lanczos, C.: ``The variational principles of mechanics'', University of Toronto Press (1949)
-\bibitem{Synge} Synge, J. L.: Encyclopedia of Physics Vol 3/1, Springer (1960)
-\bibitem{Struckmeier} Struckmeier, J.: ``Hamiltonian dynamics on the symplectic extended phase space for autonomous and non-autonomous systems'', J. Phys. A: Math. Gen 38, 1257-1278, (2005)
+\bibitem{Shannon} Shannon, C. E., ``A mathematical theory of communication'', The Bell System Technical Journal, Vol. 27, pp. 379--423, 623--656, (1948).
+\bibitem{Jaynes} Jaynes, E. T., ``Information theory and statistical mechanics'', Statistical Physics 3, pp. 181--218, (1963).
+\bibitem{classical_dynamics} J. V. Jos\'{e}, E. J. Saletan, ``Classical Dynamics'', Cambridge University Press, (1998).
+\bibitem{Gromov} Gromov, M. L., ``Pseudo holomorphic curves in symplectic manifolds'', Inventiones Mathematicae 82, pp. 307--347, (1985).
+\bibitem{deGosson} de Gosson, M. A., ``The symplectic camel and the uncertainty principle: the tip of an iceberg?'', Foundations of Physics 39, pp. 194--214, (2009).
+\bibitem{Stewart} Stewart, I., ``The symplectic camel'', Nature 329, pp. 17--18, (1987).
+\bibitem{Lanczos} Lanczos, C., ``The variational principles of mechanics'', University of Toronto Press, (1949).
+\bibitem{Synge} Synge, J. L., ``Classical dynamics'', Encyclopedia of Physics Vol 3/1, Springer (1960).
+\bibitem{Struckmeier} Struckmeier, J., ``Hamiltonian dynamics on the symplectic extended phase space for autonomous and non-autonomous systems'', J. Phys. A: Math. Gen. 38, pp. 1257--1278, (2005).
 

### 2015-03-15 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
Last review before submission

@@
 \email{carcassi@umich.edu}
-\date{June 16, 2014}
+\date{March 15, 2015}
 
@@
 In particular, we will need to introduce more precise terminology to be able to make two crucial distinctions.
-The first is between the state of the whole system and one of its parts.\footnote{Whole system vs infinitesimal component is also source of confusion when comparing classical and quantum states.
+The first is between the state of the whole system and one of its parts.\footnote{Whole system vs.
+infinitesimal component is also a source of confusion when comparing classical and quantum states.
 Quantum states are always whole and are always distributions as there is no state attributable to parts of a quantum system.}
@@
     Configuration state & The state of the whole system being described. \\ \hline
-    Particle state (or simply state) & The state of an infinitesimal part (i.e. particle) of the system. \\ \hline
+    \raggedright{Particle state (or simply state)} & The state of an infinitesimal part (i.e. particle) of the system. \\ \hline
     \end{tabular}
@@
 A cell of $\mathbbm{R}$ corresponds to $\mathbbm{R}^2$: a center and a width.\footnote{In other words: the particles of the system are not point-like, but infinitesimal cell-like.
-Not only does this lead to a more direct understanding of classical phase-space, it is also more consistent with general relativity (the mass is spread across a small region, not forming a singularity) and quantum mechanics (position is always the central value of a small distribution).} With this in mind, we have the following definitions.
+Not only does this lead to a more direct understanding of classical phase space, it is also more consistent with general relativity (the mass is spread across a small region, not forming a singularity) and quantum mechanics (position is always the central value of a small distribution).} With this in mind, we have the following definitions.
 
@@
 We use equally spaced intervals as it makes the derivation less cumbersome.} Let $K$ be a finite region of $\mathbbm{R}$.
-Divide the region in $M$ equal intervals of center $k$ and length $\Delta k$.
+Divide the region into $M$ equal intervals of center $k$ and length $\Delta k$.
 Let $I$ be the label set identified by the cells in $Q$ with center $q(i)$ and width $k(i) \Delta q$.
@@
 If we apply a coordinate change $q'=q'(q)$, we change the labels but the state must remain the same, defined at the same point with the same width.\footnote{The notion that states are what they are no matter how we label them already contains the principle of relativity.} $k' dq'$ must be then equal to $k dq$; $k'= k dq / dq'$ is contravariant, $dq' \wedge dk'= (dq'/dq) \, (dq \wedge dk) \, (dq/dq') = dq \wedge dk$ is invariant.
-This means that $\rho$ is also invariant, which makes sense: the density depends only on the state (the cell), and not the coordinate chosen to represent them.
+This means that $\rho$ is also invariant, which makes sense: the density depends only on the state (the cell), and not the coordinate chosen to represent it.
 As $k$ is contravariant, in the limit $K$ becomes the cotangent space of $Q$.
@@
 \begin{defn}\label{relativeCardinality}
-Let $I \subset \mathbf{T}^*\mathbbm{Q}$ be a closed dense set of labels, subset of a degree of freedom (i.e.
+Let $I \subset \mathbf{T}^*\mathbbm{Q}$ be a closed dense set of labels, a subset of a degree of freedom (i.e.
 a region of phase space).
@@
 That is: while the cardinality of each set diverges, the ratio between the two remains finite.\footnote{Such treatment is equivalent to what one does in information theory to extend Shannon's entropy\cite{Shannon} to the continuous case.
-There is a link between label cardinality and informational entropy, including Jaynes\cite{Jaynes} invariant formulation, which we do not explore here for brevity.} For continuous state variables we can define the relative cardinality as the ratio of a label set and our reference set $I_0$.
+There is a link between label cardinality and informational entropy, including Jaynes' invariant formulation\cite{Jaynes}, which we do not explore here for brevity.} For continuous state variables we can define the relative cardinality as the ratio of a label set and our reference set $I_0$.
 That is $n(I)=\hbar \delta q \delta k$ where $\hbar=n(I_0)/(\delta q_0 \delta k_0)$ and $n(I_0)\equiv1$ by definition.\footnote{The choice of angular momentum for $\hbar$ derives from the relationship between $p$ and $dq/dt$, which will be derived later.
@@
 Each degree of freedom is identified by the pair of labels $\langle q,p \rangle \in \mathbf{T}^*\mathbbm{Q}$.
-$\theta_0 = p dq$ and $\omega = dq \wedge dp$ are invariant, $p$ is contravariant.
+$\theta_0 = p dq$ and $\omega = dq \wedge dp$ are invariant; $p$ is contravariant.
 The configuration state is $\mathbbm{c}=\int \rho(q,p) dq \wedge dp \, \mathbbm{s}(q, p)$, where $\rho(q,p)\equiv D(q,p) / \omega$ is the distribution density for each label.
@@
 \begin{cor}\label{linearCardinality}
-Given $dq = a dq^a + b dq^b$, then $n_q(dq) = a n_q(dq^a) + b n_q(da^b)$. That is, the relative cardinality is a linear operator.
+Given $dq = a dq^a + b dq^b$, then $n_q(dq) = a n_q(dq^a) + b n_q(dq^b)$. That is, the relative cardinality is a linear operator.
 \end{cor}
@@
 v^{a} \omega_{ab} w^{b} &= v'^{a} \omega_{ab} w'^{b}  \\
-&= (v^{a} + \partial_{c} S^{a} dt v^{c}) \omega_{ab} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
+&= (v^{a} + \partial_{c} S^{a} v^{c} dt) \omega_{ab} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
 &= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} \\
@@
 
-The independence between degrees of freedoms corresponds to orthogonality in phase space: from \ref{labelsCombine} the product between the number of labels on each d.o.f.
+The independence between degrees of freedom corresponds to orthogonality in phase space: from \ref{labelsCombine} the product between the number of labels on each d.o.f.
 (i.e.
@@
 The center of each cell is identified by $t$, the width by $\hat{\omega} dt$, the conjugate variable $E\equiv\hbar\hat{\omega}$.\footnote{$\hat{\omega}$ is the classical analogue of the wave frequency.
-We use $\hat{\omega}$ to distinguish from the phase space metric $\omega$.}
+We use $\hat{\omega}$ to distinguish from the phase-space metric $\omega$.}
 \end{defn}
@@
 and the vector product within.
-The metric \ref{tdofInvariant}, with a space-like convention, still gives us the cardinality of labels within a degree of freedom, adjusting \ref{sdofInvariant} to avoid "double counting".
+The metric \ref{tdofInvariant}, with a space-like convention, still gives us the cardinality of labels within a degree of freedom, adjusting \ref{sdofInvariant} to avoid ``double counting''.
 
@@
 
-If we set\footnote{We avoided using $p_{n+1}$ as it hides the minus sign from the metric, making it seem that the temporal d.o.f is just another independent d.o.f.} $q^{n+1}=t$ and $p_{n+1}=-E$, we recognize Hamilton equations in the extended phase space\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}\cite{Synge,Lanczos}.
+If we set\footnote{We avoided using $p_{n+1}$ as it hides the minus sign from the metric, making it seem that the temporal d.o.f is just another independent d.o.f.} $q^{n+1}=t$ and $p_{n+1}=-E$, we recognize Hamilton's equations in the extended phase space\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}\cite{Synge,Lanczos}.
 
@@
 Fix a system under deterministic and reversible evolution.
-Given all its possible trajectories $x^\alpha(s)$ in the space-time manifold $\mathcal{M}$ and all its possible trajectories $\xi^a(s)$ in the extended phase space $\mathbf{T}^*\mathcal{M}$, there exists a bijective function $f: x^\alpha(s) \leftrightarrow \xi^a(s)$ that links each space-time trajectory with one and only one phase space trajectory.
+Given all its possible trajectories $x^\alpha(s)$ in the space-time manifold $\mathcal{M}$ and all its possible trajectories $\xi^a(s)$ in the extended phase space $\mathbf{T}^*\mathcal{M}$, there exists a bijective function $f: x^\alpha(s) \leftrightarrow \xi^a(s)$ that links each space-time trajectory with one and only one phase-space trajectory.
 \end{cor}
@@
 If studying the motion and state evolution are equivalent, then we must be able to go back and forth between the two pictures.
-Without losing generality, we can choose $x^\alpha=f(q^i,t)$ to be a linear function of only $q^i$ and $t$.
+Without losing generality, we can choose $x^\alpha=f(t,q^i)$ to be a linear function of only $q^i$ and $t$.
 
@@
 For the temporal d.o.f fix $dE = \lambda c^2 dt$, we have $dt \wedge dE = \lambda c^2 dt^2$.
-That is: the area in time squared is converted to an area in length squared that has the same density of states (as per definition of $c$) and then converted again to its label cardinality.
-The phase space metric is $\omega = \lambda [(dq^i)^2 - c^2 dt^2]$.
+That is: the area in time squared is converted to an area in length squared that has the same density of states (as per definition of $c$) and then is converted again to its label cardinality.
+The phase-space metric is $\omega = \lambda [(dq^i)^2 - c^2 dt^2]$.
 
@@
 \begin{align*}
-\frac{|n_x(dx)|}{|n_t(dt)|} &= \frac{|n_x(dx)|}{|n_x(c dt|)} = \frac{|dx|}{|c dt|} \leq 1 \\
+\frac{|n_x(dx)|}{|n_t(dt)|} &= \frac{|n_x(dx)|}{|n_x(c dt)|} = \frac{|dx|}{|c dt|} \leq 1 \\
 \frac{|dx|}{|dt|} & \leq c \\
@@
 Thus we find that $c$ is the well known relativistic constant.\footnote{Deterministic and reversible motion at speeds greater than $c$ (i.e.
-tachyon) is not allowed.
+a tachyon) is not allowed.
 Non-deterministic and non-reversible motion at speeds greater than $c$ (e.g.
@@
 \begin{prop}\label{newtonianGravitation}
-Assume $\mathcal{M}$ flat and time independent motion.
+Assume $\mathcal{M}$ flat and time-independent motion.
 Then the Hamiltonian simplifies to $H=\frac{1}{2m}(p_i-\hat{p}_\alpha(x))^2+\hat{H}(x)$.
 Let $\hat{p}_i = 0$ and $\hat{H} = mV$, the equations of motion are $m d_t v^i = - m \partial_i V$, where $v^i=dx^i/dt$.
-These are equations for a particle under a Netwonian gravitational potential.
+These are equations for a particle under a Newtonian gravitational potential.
 \end{prop}
 
-Given flat space-time and time independent motion, we can use $t$ as a parameter for the motion in $\mathbbm{Q}$.
+Given flat space-time and time-independent motion, we can use $t$ as a parameter for the motion in $\mathbbm{Q}$.
 Repeating \ref{kineticMomentum} and \ref{initialConditions} for the non-relativistic case leads to the above Hamiltonian.
-Similar to \ref{relativisticEM}, use \ref{mdofHam} with the newly found Hamiltonian and find the equation of motion.
+Similar to \ref{relativisticEM}, use \ref{mdofHam} with the newly found Hamiltonian and find the equations of motion.
 
@@
 Let $\mathcal{H}$ be an extended Hamiltonian defined on $\mathbf{T}^*\mathcal{M}$.
-Under the kinetic assumption, we can define the Legendre transform $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$ which we call the \emph{extended Lagrangian}.
+Under the kinematic assumption, we can define the Legendre transform $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$ which we call the \emph{extended Lagrangian}.
 For the time invariant case, we define the \emph{Lagrangian} $L=v^i p_i - H$.
@@
 The Hamiltonian found at \ref{kineticHamiltonian} is convex, so the Lagrangian can always be defined.
-While the kinetic equivalence assumption is sufficient, it's not necessary for the existence of a Lagrangian.
+While the kinematic equivalence assumption is sufficient, it's not necessary for the existence of a Lagrangian.
 What is necessary is \ref{initialConditions}: as long as position and velocity are enough to determine the state, $u^\alpha=f(q^\alpha, p_\alpha)$ must be monotonic in $p_\alpha$, which means $\mathcal{H}$ is convex in $p_\alpha$.
 What happens in that case is that the space-time trajectories may become more or less dense, $du_\alpha$ and $dx^\alpha$ would not be enough to define the label cardinality and therefore the state density $\rho$.
-In short: deterministic and reversible evolution gives us Hamiltonian mechanics, position and velocity as initial conditions gives us Lagrangian mechanics, and the kinetic assumption (which links the differentials of the state variables and the initial conditions) gives us the narrowed Hamiltonian form.
+In short: deterministic and reversible evolution gives us Hamiltonian mechanics, position and velocity as initial conditions gives us Lagrangian mechanics, and the kinematic assumption (which links the differentials of the state variables and the initial conditions) gives us the narrowed Hamiltonian form.
 
@@
   \item Deterministic and reversible evolution is what ultimately gives space-time its locally Minkowskian nature.
-  \item $\hbar$, $c$ and $m$ can be seen as conversion constants that preserve cardinality
+  \item $\hbar$, $c$ and $m$ can be seen as conversion constants that preserve cardinality.
   \item Lagrangian mechanics coincides with position and velocity being necessary and sufficient initial conditions.
-  \item The motion of an isolated elementary system, for which the kinetic assumption is valid, is restricted to a Hamiltonian that can describe fundamental classical forces.
+  \item The motion of an isolated elementary system, for which the kinematic assumption is valid, is restricted to a Hamiltonian that can describe fundamental classical forces.
 \end{itemize}

### 2015-07-15 ReferenceCards/ComplexInnerProduct.tex
Adding reference card for inner product

\documentclass[10pt]{article}
\usepackage{fixltx2e}
\usepackage[orthodox,l2tabu,abort]{nag}

% Page layout
\usepackage[landscape,margin=0.5in]{geometry}
\usepackage{multicol}
\usepackage{xcolor}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}

\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{arrows.meta}

% Title area
\usepackage{titling} % Allows for use of date, author, etc. after \maketitle
% Ref: http://tex.stackexchange.com/questions/3988/titlesec-versus-titling-mangling-thetitle
\let\oldtitle\title
\renewcommand{\title}[1]{\oldtitle{#1}\newcommand{\mythetitle}{#1}}
\renewcommand{\maketitle}{%
{\begin{center}\Large \mythetitle\end{center}}
}

% Document divisions
\usepackage{titlesec}
\setcounter{secnumdepth}{0}
\titlespacing{\section}{0pt}{0pt}{0pt}
\titlespacing{\subsection}{0pt}{0pt}{0pt}
\usepackage{nopageno} % To keep \section from resetting page style

\setlength{\parindent}{0pt} % disabling indentation by default

% Lists
\usepackage{enumitem} % for consistent formatting of lists
\newlist{ttdesc}{description}{1}
\setlist[ttdesc]{font=\ttfamily,noitemsep}
\usepackage{calc} % for \widthof

% Code
\usepackage{listings}
\lstset{language=[LaTeX]TeX,%
  basicstyle=\itshape,%
  keywordstyle=\normalfont\ttfamily,%
  morekeywords={part,chapter,subsection,subsubsection,paragraph,subparagraph}%
  }

\usepackage{lipsum}

\title{Complex Inner Product}
\author{Gabriele Carcassi}
\date{2015}

\begin{document}
\begin{multicols}{3}
\maketitle

\section{Inner product}

\begin{tikzpicture}[scale=1]
    \tikzstyle{help line}=[semithick,color=gray]
    \tikzstyle{vector}=[-Stealth,semithick,color=black]
    \tikzstyle{guideline}=[semithick,color=gray]
    \tikzstyle{guideline text down}=[black,below=2pt]
    \tikzstyle{guideline text right}=[black,right=1pt]
    \tikzstyle{guideline text up}=[black,above=2pt]
    \tikzstyle{guideline text left}=[black,left=1pt]
    %\draw[style=help line,step=2cm] (-2.5,-2.5) grid (2.5,2.5);
    % Draw a1 + b1 i
	\draw[style=vector] (0,0) -- (4,1);
	\draw[style=guideline] (0,0) -- node[style=guideline text down] {$a_1$} (4,0);
	\draw[style=guideline] (4,0) -- node[style=guideline text right] {$b_1$} (4,1);

    % Draw a2 + b2 i
	\draw[style=vector] (0,0) -- (1.5,3);
	\draw[style=guideline] (0,0) -- node[style=guideline text left] {$b_2$} (0,3);
	\draw[style=guideline] (0,3) -- node[style=guideline text up] {$a_2$} (1.5,3);
\end{tikzpicture}

\begin{align*}
\overline{c_2} c_1 &= (a_2 - b_2 i) (a_1 + b_1 i) \\
&= a_1 a_2 + b_1 b_2 + \frac {a_1 b_2 - a_2 b_1}{i} \\
\end{align*}

\section{Imaginary part}

\begin{tikzpicture}[scale=1]
    \tikzstyle{axis}=[->,semithick,color=black]
    \tikzstyle{vector}=[-Stealth,semithick,color=black]
    \tikzstyle{parallel}=[semithick,color=black]
    \tikzstyle{guideline}=[semithick,color=gray]
    \tikzstyle{guideline text down}=[black,below=2pt]
    \tikzstyle{guideline text right}=[black,right=1pt]
    \tikzstyle{guideline text up}=[black,above=2pt]
    \tikzstyle{guideline text left}=[black,left=1pt]

	\coordinate (a_1) at (5,0);
	\coordinate (b_1) at (0,1);
	\coordinate (v_1) at ($(a_1)+(b_1)$);

	\coordinate (a_2) at (1.5,0);
	\coordinate (b_2) at (0,4);
	\coordinate (v_2) at ($(a_2)+(b_2)$);

    % Axis
	\draw[style=axis] (-0.5,0) -- ($(a_1)+(a_2)+(0.5,0)$) node[style=guideline text right] {$Re$};
	\draw[style=axis] (0,-0.5) -- ($(b_1)+(b_2)+(0,0.5)$) node[style=guideline text up] {$Im$};

    \filldraw[gray,fill=red,opacity=0.5](0,0)--(a_1)--(v_1)--cycle;
    \filldraw[gray,fill=red,opacity=0.5](v_2)--($(b_1)+(v_2)$)--($(v_1)+(v_2)$)--cycle;
    \filldraw[gray,fill=orange,opacity=0.5](0,0)--(b_2)--(v_2)--cycle;
    \filldraw[gray,fill=orange,opacity=0.5](v_1)--($(v_1)+(a_2)$)--($(v_1)+(v_2)$)--cycle;
    \filldraw[gray,fill=brown,opacity=0.5](b_2)--(v_2)--($(b_1)+(v_2)$)--($(b_1)+(b_2)$)--cycle;
    \filldraw[gray,fill=brown,opacity=0.5](a_1)--(v_1)--($(a_2)+(v_1)$)--($(a_1)+(a_2)$)--cycle;

    % Bottom
	\draw[style=guideline] (0,0) -- node[style=guideline text down] {$a_1$} (a_1);
	\draw[style=guideline] (a_1) -- node[style=guideline text down] {$a_2$} ($(a_1)+(a_2)$);
	\draw[style=guideline] (v_1) -- ($(v_1)+(a_2)$);

    % Left
	\draw[style=guideline] (0,0) -- node[style=guideline text left] {$b_2$} (b_2);
	\draw[style=guideline] (b_2) -- node[style=guideline text left] {$b_1$} ($(b_2)+(b_1)$);
	\draw[style=guideline] (v_2) -- ($(v_2)+(b_1)$);

    % Top
	\draw[style=guideline] ($(b_2)+(b_1)$) -- node[style=guideline text up] {$a_2$} ($(v_2)+(b_1)$);
	\draw[style=guideline] ($(v_2)+(b_1)$) -- node[style=guideline text up] {$a_1$} ($(v_1)+(v_2)$);
	\draw[style=guideline] (b_2) -- (v_2);

    % Right
	\draw[style=guideline] ($(a_1)+(a_2)$) -- node[style=guideline text right] {$b_1$} ($(v_1)+(a_2)$);
	\draw[style=guideline] ($(v_1)+(a_2)$) -- node[style=guideline text right] {$b_2$} ($(v_1)+(v_2)$);
	\draw[style=guideline] (a_1) -- (v_1);

    % Draw a1 + b1 i
	\draw[style=vector] (0,0) -- node[style=guideline text up] {$c_1$} (v_1);
	\draw[style=vector] (0,0) -- node[style=guideline text right] {$c_2$} (v_2);
	\draw[style=parallel] (v_1) -- ($(v_1)+(v_2)$);
	\draw[style=parallel] (v_2) -- ($(v_1)+(v_2)$);
	\draw[align=center] ($0.5*(v_1)+0.5*(v_2)$) node {$a_1 b_2 - a_2 b_1$};

\end{tikzpicture}

\begin{align*}
Area&=(a_1+a_2)(b_1+b_2) - \textcolor{red}{a_1 b_1} - \textcolor{orange}{a_2 b_2}
- \textcolor{brown}{2 a_2 b_1}\\
&= a_1 b_2 - a_2 b_1 \\
\end{align*}

\begin{tikzpicture}
	% general shift to north east
	\coordinate (O) at (0.5,0.5);
	\draw[semithick] (0,0) -- (4,1);% bottom line in front
	\draw[dashed,color=gray] (O) -- ($(4,1)+(O)$);% bottom line in the back
	\draw[semithick] (0,3) -- (4,2);% top line in front
	\draw[semithick] ($(0,3)+(O)$) -- ($(4,2)+(O)$);% top line in the back
	\draw[semithick] (0,3) -- ($(0,3)+(O)$);% line to the back, top left
	\draw[semithick] (4,2) -- ($(4,2)+(O)$);% line to the back, top right
	\draw[semithick] (4,1) -- ($(4,1)+(O)$);% line to the back, bottom right
	\draw[dashed,color=gray] (0,0) -- (O);% line to the back, bottom left
	% the first angle is 180°+atan(0.25)
	% the second angle is 180°-atan(0.25)
	% the radius is sqrt(6^2+1.5^2)
	\draw[semithick] (0,0) arc (194.036:165.964:6.185);% left arc in front
	\draw[dashed,color=gray] (O) arc (194.036:165.964:6.185);% left arc in the back
	% the first angle is 180°+atan(0.25)
	% the second angle is 180°-atan(0.25)
	% the radius is 1/3*sqrt(6^2+1.5^2)
	\draw[semithick] (4,1) arc (194.036:165.964:2.062);% right arc in front
	\draw[semithick] ($(4,1)+(O)$) arc (194.036:165.964:2.062);% right arc in the back
	\draw (-0.5,1.7) node {$b_1$};
	\draw (3.6,1.7) node {$b_2$};
	\draw (0,3.5) node {$s$};
	\draw[|-,semithick] (0,-0.5) -- (4,-0.5);
	\draw[|->,semithick] (4,-0.5) -- (4.5,-0.5);
	\draw (0,-1) node {$x=0$};
	\draw (4,-1) node {$x=l$};
\end{tikzpicture}

\noindent Copyright \textcopyright{} \thedate{} \theauthor{}

\end{multicols}
\end{document}

### 2015-07-15 2015CompleteDerivation/completeDerivation.tex
Starting to think about a new version which includes quantum

--- 2015-03-15 2014CompleteClassicalMechanics/completeClassicalMechanics.tex
@@
 The novel, and surprising, result is how so much can be derived from so little.
+
+\section{About}
+
+The work is organized into:
+\begin{description}
+  \item[Assumptions] these characterize the physical system we are studying and constitute the premise of our discussion.
+A \textbf{rationale} follows each assumption, which uses physical and sometimes philosophical arguments to motivate why (or why not) such an assumption makes sense (in a particular case).
+  \item[Definitions] these encode into mathematical language the physical concepts we are studying.
+A \textbf{justification} follows each definition, which uses physical and mathematical arguments to explain why such a definition and its properties are necessary to describe a particular physical concept.
+  \item[Propositions] these are statements that mathematically follow from definitions or other propositions.
+A mathematical \textbf{proof} follows each proposition.
+No mathematical breakthrough should be expected as we mostly use well known results from different fields.
+The goal is to show how those results make sense in light of the physical assumptions we start with.
+\end{description}
+This should allow you, the reader, to focus on the parts you are most interested in and skim the rest.
+The philosophically inclined may focus on the assumptions, their rationale and the definitions.
+The mathematically inclined may focus on the definitions, the propositions and their proofs.
+The scientist on the assumptions, the definitions and their justifications, and the propositions.
+
+\section{New}
+
+%TODO: it would be nice to have "examples" always refer to the same cases, and that they pertain to different type of physics.
+E.g.
+trajectory, gravity, thermodynamics, fluid dynamics, biology, electromagnetism...
+
+We start by fixing a physical system we want to study, that is something we can interact with and perform measurements on (e.g.
+a cat, a planet, water).
+We set what particular aspect we want to study (e.g.
+the trajectory when falling, the motion around a star, its flow in a pipe).
+We call state a particular configuration in time of the portion under study.
+We call internal dynamics the remainder of the system.
+
+\begin{assump2}
+The physical system undergoes deterministic (future state identified by the present state) and reversible (past state identified by the present state) evolution.
+\end{assump2}
+
+\textbf{Rationale}.
+While the choice of the object of our study (i.e.
+the system and the aspect we are interested in) is arbitrary, this assumption is at the heart of being able to write physics laws, to define physical states and to physically study a system.
+If we are to study a physical system through reliable experimentation and observation, we need deterministic and reversible processes to manipulate its state.
+We can provide different arguments that point in the same direction.
+
+First, to be able to identify the system, we must first be able to isolate it from the environment.
+Intuitively, we can distinguish between two chairs because we can move the first to another room and sit on it without somehow having to drag the second as well.
+More precisely, we can manipulate the state of the first system without affecting the second, and vice-versa.
+If we extend this idea, to be able to properly call something a system we need to show that it is isolated: if I affect its state, the state of the other systems is not affected and vice-versa.
+In other words: the system future and past states are only determined by its own state.
+
+Second, the aim of physics is to write laws that can be used to make prediction that can be validated experimentally.
+If I drop an anvil from a tower, it will accelerate at $9.81 m/s^2$; if I want the anvil to reach the ground at $x m/s$ I have to drop it from $y m$.
+To the extent that we want to make predictions in time, we need to have a one-to-one correspondence between initial and final states
+
+Third, operationally we must reliably prepare and measure states.
+That is, we need a process for which the input settings of our preparing device determine the outgoing state of the system; and a process for which the incoming state of the system can be reconstructed by the output of the measuring device.
+That is, our system must participate in a deterministic and reversible process with the preparing and measuring device.
+How else could we calibrate our experimental apparatus?
+
+This link between state definition and deterministic processes should not be too surprising as the state, in the context of thermodynamics and system theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system}.
+One source of confusion, though, may come from statistical process: those are not deterministic!
+The confusion is that the object of our study stops at the distribution (the ensemble), the part that can be indeed calculated, measured and prepared.
+Another source of confusion is that what is considered state changes depending on the context: the position of a molecule is part of the state for a single molecule system but not for a volume of gas.
+And again because in the latter context is not part of the deterministic motion of the system.
+
+So, if it seems we need to require deterministic and reversible motion, why we call this an assumption?
+Why not call it requirement?
+Because it's a \emph{simplifying} assumption: it can never be totally achieved.
+A system can be prepared or measured up to a certain level of precision.
+Perfect isolation of a system is impossible both practically (e.g.
+black-body radiation, gravity, ...) and conceptually (e.g.
+if the system is perfectly isolated, we can't prepare it, we can't measure it, we can't study it: it's not a physical system).
+If the system is always interacting with the environment, then how can we get deterministic and reversible evolution?
+We can if the environment and the internal dynamics of the system interact in such a way that they affect and are little affected by the aspect we are studying.
+For example: the trajectory of a falling cat does not greatly affect and is not greatly affected by its digestive system or external temperature.
+
+\begin{assump1}\label{classical}
+The system is infinitely reducible: it can be thought of as composed of two or more similar but smaller systems, each in its own configuration state, which in turn can be thought of as composed of two or more, ad infinitum.
+We call \emph{particle} such an infinitesimal part.
+\end{assump1}
+
+\textbf{Rationale}.
+This makes a clear demarcation of what a classical object is (and conversely what a quantum object is not): it is something that can be fully described up to its infinitesimal constituents.
+This is clearly a \emph{simplifying} assumption, and it is instructive to understand when and why it cannot be used.
+
+The first problem is methodological.
+As we saw before, we need access to a deterministic and reversible process to be able to study a system and define a state.
+For two balls, we can imagine studying pieces of the balls in isolation, and then describe the collision between the two by describing what happens at each piece.
+The classical assumption holds.
+For an electron and a photon, we cannot take pieces of the electron or the photon, study them in isolation and then describe how each part moves during Compton scattering.
+The classical assumption does not hold: we do not have suitable physical processes at our disposal.
+
+The second problem is more conceptual.
+As we saw before, a system cannot be fully isolated from the environment.
+The smaller the part we study, the more it will be susceptible to fluctuations from the environment, the less we can justify the deterministic assumption, the less we are in a position to define its state.
+While the overall position of a ball can be considered fairly unaffected by its surrounding, the position of each small piece is going to be constantly affected by air and light scattering on its surface.
+
+If the scale of the problem
 

### 2015-07-16 ReferenceCards/ComplexInnerProduct.tex
Adding angle
Added real part diagram

@@
 \begin{document}
-\begin{multicols}{3}
 \maketitle
+\begin{multicols}{2}
 
@@
 \begin{tikzpicture}[scale=1]
-    \tikzstyle{help line}=[semithick,color=gray]
+    \tikzstyle{axis}=[->,semithick,color=black]
     \tikzstyle{vector}=[-Stealth,semithick,color=black]
+    \tikzstyle{parallel}=[semithick,color=black]
     \tikzstyle{guideline}=[semithick,color=gray]
@@
     \tikzstyle{guideline text left}=[black,left=1pt]
+
+	\coordinate (O) at (0,0);
+
+	\coordinate (a_1) at (4,0);
+	\coordinate (b_1) at (0,1);
+	\coordinate (v_1) at ($(a_1)+(b_1)$);
+
+	\coordinate (a_2) at (1.5,0);
+	\coordinate (b_2) at (0,3);
+	\coordinate (v_2) at ($(a_2)+(b_2)$);
+
+    % Axis
+	\draw[style=axis] (-0.5,0) -- ($(a_1)+(0.5,0)$) node[style=guideline text right] {$Re$};
+	\draw[style=axis] (0,-0.5) -- ($(b_2)+(0,0.5)$) node[style=guideline text up] {$Im$};
+
     %\draw[style=help line,step=2cm] (-2.5,-2.5) grid (2.5,2.5);
     % Draw a1 + b1 i
-	\draw[style=vector] (0,0) -- (4,1);
-	\draw[style=guideline] (0,0) -- node[style=guideline text down] {$a_1$} (4,0);
-	\draw[style=guideline] (4,0) -- node[style=guideline text right] {$b_1$} (4,1);
+	\draw[style=guideline] (O) -- node[style=guideline text down] {$a_1$} (a_1)
+                               -- node[style=guideline text right] {$b_1$} (v_1);
+	\draw[style=vector] (O) -- node[style=guideline text up] {$c_1$} (v_1);
 
     % Draw a2 + b2 i
-	\draw[style=vector] (0,0) -- (1.5,3);
-	\draw[style=guideline] (0,0) -- node[style=guideline text left] {$b_2$} (0,3);
-	\draw[style=guideline] (0,3) -- node[style=guideline text up] {$a_2$} (1.5,3);
+	\draw[style=guideline] (O) -- node[style=guideline text left] {$b_2$} (b_2)
+	                           -- node[style=guideline text up] {$a_2$} (v_2);
+	\draw[style=vector] (O) -- node[style=guideline text right] {$c_2$} (v_2);
+
+    \draw let \p1 = (O),    % Center
+              \p2 = (v_1),    % First vector
+              \p3 = (v_2),  % Second vector
+              \n1 = {0.2},    % Position of arc on first vector
+              \p4 = ($(\p1)!\n1!(\p2)$),    % Point where the arc starts
+              \n2 = {atan2(\y2,\x2)},       % First angle
+              \n3 = {atan2(\y3,\x3)},       % Second angle
+              \n4 = {veclen(\y4,\x4)},       % Arc radius
+              \n5 = {\n2 + (\n3 - \n2) / 2},       % Mid angle
+              \n6 = {\n4 + 7}       % Label radius
+               in (\p4) -- ++(\p1) arc   (\n2:\n3:\n4) node at(\n5:\n6) {$\theta$} ;
+
 \end{tikzpicture}
@@
 &= a_1 a_2 + b_1 b_2 + \frac {a_1 b_2 - a_2 b_1}{i} \\
+&= r_2 e^{-i\theta_2} r_1 e^{i\theta_1} \\
+&= r_1 r_2 e^{-i\theta} \\
+&= r_1 r_2 cos(\theta) + \frac {r_1 r_2 sin(\theta)}{i} \\
+&= c_1 \cdot c_2 + \frac {c_1 \wedge c_2}{i} \\
+\end{align*}
+
+\vfill
+\columnbreak
+
+\section{Real part}
+\begin{center}
+\begin{tikzpicture}[scale=1]
+    \tikzstyle{axis}=[->,semithick,color=black]
+    \tikzstyle{vector}=[-Stealth,semithick,color=black]
+    \tikzstyle{parallel}=[semithick,color=black]
+    \tikzstyle{guideline}=[semithick,color=gray]
+    \tikzstyle{guideline text down}=[black,below=2pt]
+    \tikzstyle{guideline text right}=[black,right=1pt]
+    \tikzstyle{guideline text up}=[black,above=2pt]
+    \tikzstyle{guideline text left}=[black,left=1pt]
+
+	\coordinate (a_1) at (5,0);
+	\coordinate (b_1) at (0,1);
+	\coordinate (v_1) at ($(a_1)+(b_1)$);
+
+	\coordinate (a_2) at (1.5,0);
+	\coordinate (b_2) at (0,4);
+	\coordinate (v_2) at ($(a_2)+(b_2)$);
+
+	\coordinate (a_2r) at (0,1.5);
+	\coordinate (b_2r) at (-4,0);
+	\coordinate (v_2r) at ($(a_2r)+(b_2r)$);
+
+    % Axis
+	\draw[style=axis] ($(b_2r)-(0.5,0)$) -- ($(a_1)+(0.5,0)$) node[style=guideline text right] {$Re$};
+	\draw[style=axis] (0,-0.5) -- ($(b_1)+(a_2r)+(0,0.5)$) node[style=guideline text up] {$Im$};
+
+    \filldraw[gray,fill=red,opacity=0.5](0,0)--(a_1)--(v_1)--cycle;
+    \filldraw[gray,fill=red,opacity=0.5](v_2r)--($(b_1)+(v_2r)$)--($(v_1)+(v_2r)$)--cycle;
+    \filldraw[gray,fill=orange,opacity=0.5](0,0)--(b_2r)--(v_2r)--cycle;
+    \filldraw[gray,fill=orange,opacity=0.5](v_1)--($(v_1)+(a_2r)$)--($(v_1)+(v_2r)$)--cycle;
+
+    % Bottom triangle
+	\draw[style=guideline] (0,0) -- node[style=guideline text down] {$a_1$} (a_1);
+	\draw[style=guideline] (0,0) -- node[style=guideline text down] {$b_2$} (b_2r);
+
+    % Left
+	\draw[style=guideline] (b_2r) -- node[style=guideline text left] {$a_2$} (v_2r)
+	                              -- node[style=guideline text left] {$b_1$} ($(v_2r)+(b_1)$);
+
+    % Top
+	\draw[style=guideline] ($(v_2r)+(b_1)$) -- node[style=guideline text up] {$a_1$} ($(v_2r)+(v_1)$)
+                                            -- node[style=guideline text up] {$b_2$} ($(v_1)+(a_2r)$);
+
+    % Right
+	\draw[style=guideline] ($(v_1)+(a_2r)$) -- node[style=guideline text right] {$a_2$} (v_1)
+	                                        -- node[style=guideline text right] {$b_1$} (a_1);
+
+    % Draw a1 + b1 i
+	\draw[style=vector] (0,0) -- node[style=guideline text up] {$c_1$} (v_1);
+	\draw[style=vector, gray] (0,0) -- (v_2);
+	\draw[style=vector] (0,0) -- node[style=guideline text up] {$c_2^{-90^\circ}$} (v_2r);
+	\draw[style=parallel] (v_1) -- ($(v_1)+(v_2r)$);
+	\draw[style=parallel] (v_2r) -- ($(v_1)+(v_2r)$);
+	\draw[align=center] ($0.5*(v_1)+0.5*(v_2r)$) node {$a_1 a_2 + b_1 b_2$ \\ $c_1 \cdot c_2$};
+
+\end{tikzpicture}
+\end{center}
+
+\begin{align*}
+Area&=(a_1+b_2)(a_2+b_1) - \textcolor{red}{a_1 b_1} - \textcolor{orange}{a_2 b_2} = a_1 a_2 + b_1 b_2 \\
 \end{align*}
@@
 \section{Imaginary part}
-
+\begin{center}
 \begin{tikzpicture}[scale=1]
@@
 	\draw[style=parallel] (v_2) -- ($(v_1)+(v_2)$);
-	\draw[align=center] ($0.5*(v_1)+0.5*(v_2)$) node {$a_1 b_2 - a_2 b_1$};
+	\draw[align=center] ($0.5*(v_1)+0.5*(v_2)$) node {$a_1 b_2 - a_2 b_1$ \\ $c_1 \wedge c_2$};
 
 \end{tikzpicture}
+\end{center}
 
@@
 Area&=(a_1+a_2)(b_1+b_2) - \textcolor{red}{a_1 b_1} - \textcolor{orange}{a_2 b_2}
-- \textcolor{brown}{2 a_2 b_1}\\
-&= a_1 b_2 - a_2 b_1 \\
+- \textcolor{brown}{2 a_2 b_1} = a_1 b_2 - a_2 b_1 \\
 \end{align*}
 
-\begin{tikzpicture}
-	% general shift to north east
-	\coordinate (O) at (0.5,0.5);
-	\draw[semithick] (0,0) -- (4,1);% bottom line in front
-	\draw[dashed,color=gray] (O) -- ($(4,1)+(O)$);% bottom line in the back
-	\draw[semithick] (0,3) -- (4,2);% top line in front
-	\draw[semithick] ($(0,3)+(O)$) -- ($(4,2)+(O)$);% top line in the back
-	\draw[semithick] (0,3) -- ($(0,3)+(O)$);% line to the back, top left
-	\draw[semithick] (4,2) -- ($(4,2)+(O)$);% line to the back, top right
-	\draw[semithick] (4,1) -- ($(4,1)+(O)$);% line to the back, bottom right
-	\draw[dashed,color=gray] (0,0) -- (O);% line to the back, bottom left
-	% the first angle is 180°+atan(0.25)
-	% the second angle is 180°-atan(0.25)
-	% the radius is sqrt(6^2+1.5^2)
-	\draw[semithick] (0,0) arc (194.036:165.964:6.185);% left arc in front
-	\draw[dashed,color=gray] (O) arc (194.036:165.964:6.185);% left arc in the back
-	% the first angle is 180°+atan(0.25)
-	% the second angle is 180°-atan(0.25)
-	% the radius is 1/3*sqrt(6^2+1.5^2)
-	\draw[semithick] (4,1) arc (194.036:165.964:2.062);% right arc in front
-	\draw[semithick] ($(4,1)+(O)$) arc (194.036:165.964:2.062);% right arc in the back
-	\draw (-0.5,1.7) node {$b_1$};
-	\draw (3.6,1.7) node {$b_2$};
-	\draw (0,3.5) node {$s$};
-	\draw[|-,semithick] (0,-0.5) -- (4,-0.5);
-	\draw[|->,semithick] (4,-0.5) -- (4.5,-0.5);
-	\draw (0,-1) node {$x=0$};
-	\draw (4,-1) node {$x=l$};
-\end{tikzpicture}
+\end{multicols}
 
 \noindent Copyright \textcopyright{} \thedate{} \theauthor{}
-
-\end{multicols}
 \end{document}

### 2015-10-15 2015CompleteDerivation/completeDerivation.tex
Added an outline and discussion on the fundamental model

@@
 
+\section{Outline}
+
+Deterministic/reversible assumption.
+Describe fundamental model of physics.
+System is always interacting with environment, both deterministically and non-deterministically.
+Deterministic means only depends on the state of the system.
+Set of states needs to be invariant under non-deterministic interaction: the non-deterministic interaction affects what are the states of the system.
+Deterministic/reversible evolution is necessary to even be able to talk about a system (trivial deterministic process).
+
+Define: set of states, deterministic evolution as permutation, labels/state variables, cardinality of state variable, state id function (=1 for a particular state).
+Can we do for continuous variables as well?
+Can we find divergence free evolution for det/rev and independent variables?
+
+Infinite reducibility assumption.
+Decomposable systems.
+Rule of composition.
+Inverse and null state to study changes.
+Set of transformations to increase and decrease the quantity in each state.
+Define magnitude of the system (probably needed to guarantee things converge in math?).
+
+Define: set of states as a vector space, inner product.
+
+Continuous quantities. Need for densities to be a function of the state (not state variable). Defined on infinitesimal interval. Define: degree of freedom, state cardinality, metric (T*Q).
+
 \section{New}
@@
 
-We start by fixing a physical system we want to study, that is something we can interact with and perform measurements on (e.g.
+We start by fixing a \emph{physical system} we want to study, that is something we can interact with and perform measurements on (e.g.
 a cat, a planet, water).
+We call \emph{environment} everything else.
 We set what particular aspect we want to study (e.g.
 the trajectory when falling, the motion around a star, its flow in a pipe).
-We call state a particular configuration in time of the portion under study.
-We call internal dynamics the remainder of the system.
+We call \emph{state} a particular configuration in time of the portion under study.
+Since the state does not, in general, exhaust the description of the system, a part remains \emph{unstated}, and as such we'll call it, for lack of a better word.
+With this in mind, we introduce the following assumption:
 
 \begin{assump2}
-The physical system undergoes deterministic (future state identified by the present state) and reversible (past state identified by the present state) evolution.
+The state of the physical system under study undergoes deterministic and reversible evolution.
 \end{assump2}
+
+By deterministic (and reversible) we mean that future (and past) states are uniquely identified by the present state.
+We can then think of the state of the system as the part that interacts deterministically and reversibly with the environment (and itself) while the unstated part as interacting non-deterministically with the environment (and itself).
+We call this setup the \emph{fundamental model of physics}, as it is the simplest model that still captures the crucial aspects of studying and writing laws for a physical system.
+
+Much of the focus later will be given to the state under deterministic and reversible evolution.
+However, the non-deterministic part plays a fundamental part as well.
+In fact: it is this part of the evolution that determines what states are available and their description.
+Suppose we study the motion of a small particle, its state under gravitational and inertial forces will be properly described by the position and momentum of the center of mass.
+Suppose we study the same small particle as suspended in a fluid, as it undergoes Brownian motion: its state will be a probability distribution for position and momentum of the center of mass.
+Gravitational and inertial forces have not changed, yet the set of states have.
+The issue is that the set of states has to invariant under the non-deterministic evolution as well.
+That is: if we apply the non-deterministic evolution to any state, we still need to end up in one of the possible states.
+If the Brownian motion is strong enough, the state gets knocked out of the set of states determined by a simple position/momentum pair.
+
+A similar more drastic effect: consider a book and its motion under gravitation and inertial forces, its state being the position and momentum of the center of mass.
+As we increase the temperature of the air around the book, its motion is unaffected until, at some point, the book burns.
+Clearly, the non-deterministic evolution has brought one of the states outside the set.
+
+As we have seen, sometimes the state is identified by a distribution (either statistical or actual).
+Even in this case, the state can be deterministic and reversible.
+That is: given the distribution at one time we can determine the distribution at future times.
+The shape and the parameters of the distribution can be deterministic, even if the individual trajectories are not (those fall within the unstated part).
+In fact: we cannot assume trajectory and states are defined at all for the unstated part.
+Consider a muon and its decay products: it is clear that the three outgoing particles have a state and trajectory of their own, it is clear that the mass and energy they have came from the muon, yet we cannot assign some internal state or trajectory for those parts of mass and energy before the muon decayed.
+What happens in the muon stays in the muon, so to speak.
+
+Now that we have clarified what we mean by deterministic and reversible evolution (even on a statistical ensemble), and how also the non-deterministic part is key to define states (as the set as a whole needs to be invariant under it), we can ask ourselves: when is this assumption valid?
 

### 2015-10-19 2015CompleteDerivation/completeDerivation.tex
Added detrev justification and first few math definitions

@@
 Since the state does not, in general, exhaust the description of the system, a part remains \emph{unstated}, and as such we'll call it, for lack of a better word.
-With this in mind, we introduce the following assumption:
+With this in mind, we introduce the following:
 
@@
 Much of the focus later will be given to the state under deterministic and reversible evolution.
-However, the non-deterministic part plays a fundamental part as well.
-In fact: it is this part of the evolution that determines what states are available and their description.
+However, the non-deterministic part plays a fundamental role as well.
+In fact: it is this aspect of the evolution that determines what states are available and their description.
 Suppose we study the motion of a small particle, its state under gravitational and inertial forces will be properly described by the position and momentum of the center of mass.
@@
 Gravitational and inertial forces have not changed, yet the set of states have.
-The issue is that the set of states has to invariant under the non-deterministic evolution as well.
+The issue is that the set of states must be closed under the non-deterministic evolution as well.
 That is: if we apply the non-deterministic evolution to any state, we still need to end up in one of the possible states.
@@
 A similar more drastic effect: consider a book and its motion under gravitation and inertial forces, its state being the position and momentum of the center of mass.
-As we increase the temperature of the air around the book, its motion is unaffected until, at some point, the book burns.
-Clearly, the non-deterministic evolution has brought one of the states outside the set.
+As we increase the temperature of the air around the book, its motion remains unaffected until, at some point, the book burns.
+Clearly, the non-deterministic evolution has brought one of the states outside the set, to the point that the system is no longer recognizable.
 
@@
 In fact: we cannot assume trajectory and states are defined at all for the unstated part.
-Consider a muon and its decay products: it is clear that the three outgoing particles have a state and trajectory of their own, it is clear that the mass and energy they have came from the muon, yet we cannot assign some internal state or trajectory for those parts of mass and energy before the muon decayed.
-What happens in the muon stays in the muon, so to speak.
+Consider a muon and its decay products: it is clear that the three outgoing particles have a state and trajectory of their own, it is clear that the resulting total mass and energy came from the muon.
+Yet, before the decay, we cannot ascribe an internal state and trajectory to those parts.
+The state of a muon is not some combination of the state of an electron and two neutrinos.
 
@@
 \textbf{Rationale}.
-While the choice of the object of our study (i.e.
-the system and the aspect we are interested in) is arbitrary, this assumption is at the heart of being able to write physics laws, to define physical states and to physically study a system.
-If we are to study a physical system through reliable experimentation and observation, we need deterministic and reversible processes to manipulate its state.
+The claim is that deterministic and reversible evolution is necessary to be able to study a physical system.
+That is: the only part of a system that can be studied through reliable and reproducible experimentation or observation is the part that undergoes deterministic and reversible evolution.
 We can provide different arguments that point in the same direction.
 
-First, to be able to identify the system, we must first be able to isolate it from the environment.
-Intuitively, we can distinguish between two chairs because we can move the first to another room and sit on it without somehow having to drag the second as well.
-More precisely, we can manipulate the state of the first system without affecting the second, and vice-versa.
-If we extend this idea, to be able to properly call something a system we need to show that it is isolated: if I affect its state, the state of the other systems is not affected and vice-versa.
-In other words: the system future and past states are only determined by its own state.
+First, to be able to identify the system, we must first be able to tell it apart from the environment.
+Intuitively, we can distinguish between two chairs because we can move the first to another room and sit on it without having touched the second.
+We can manipulate the state of the first system without affecting the second, and vice-versa.
+So, to tell a system apart from its environment, there has to to be a certain amount of isolation between the two: if I affect its state, the state of the other systems is not affected and vice-versa.
+This means that the system future and past states are only determined by its own state and the state undergoes deterministic and reversible evolution.
 
@@
 If I drop an anvil from a tower, it will accelerate at $9.81 m/s^2$; if I want the anvil to reach the ground at $x m/s$ I have to drop it from $y m$.
-To the extent that we want to make predictions in time, we need to have a one-to-one correspondence between initial and final states
+To the extent that we want to make predictions in time, we need to have a one-to-one correspondence between initial and final states.
 
@@
 This link between state definition and deterministic processes should not be too surprising as the state, in the context of thermodynamics and system theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system}.
-One source of confusion, though, may come from statistical process: those are not deterministic!
-The confusion is that the object of our study stops at the distribution (the ensemble), the part that can be indeed calculated, measured and prepared.
-Another source of confusion is that what is considered state changes depending on the context: the position of a molecule is part of the state for a single molecule system but not for a volume of gas.
-And again because in the latter context is not part of the deterministic motion of the system.
+As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can be indeed calculated, measured and prepared, not each individual element.
 
 So, if it seems we need to require deterministic and reversible motion, why we call this an assumption?
-Why not call it requirement?
-Because it's a \emph{simplifying} assumption: it can never be totally achieved.
+First of all, because it's an idealization: it can never be completely achieved in practice.
 A system can be prepared or measured up to a certain level of precision.
@@
 black-body radiation, gravity, ...) and conceptually (e.g.
-if the system is perfectly isolated, we can't prepare it, we can't measure it, we can't study it: it's not a physical system).
-If the system is always interacting with the environment, then how can we get deterministic and reversible evolution?
-We can if the environment and the internal dynamics of the system interact in such a way that they affect and are little affected by the aspect we are studying.
-For example: the trajectory of a falling cat does not greatly affect and is not greatly affected by its digestive system or external temperature.
+if the system is perfectly isolated, we cannot interact with it: how can it studied through experimentation?).
+It's a simplifying assumption that can only be taken if the environment and the internal dynamics of the system interact in such a way that they affect and are little affected by the aspect we are studying.
+As we saw before, for example, assuming that the states consists of the position and momentum of the center of mass means assuming that the Brownian motion of the body is negligible.
+
+The second reason is that technically the assumption comes first.
+If we start looking for a law, we are implicitly assuming that there exists one to be found.
+The opposite assumption, there is no law to be found, can't really be experimentally proven exhaustively (i.e.
+you can only show that you have not found it among a certain class).
+Very often the hard part in science is finding for what system and in what conditions such assumption hold.
+Finding the law after that is a much simpler endeavour, and is something that is routinely done by students in their intro labs.
+
+We are now ready to capture the elements of our discussion and their properties through mathematical definitions.
+
+\begin{defn}\label{statedef}
+Fix a physical system for study.
+There exists a set $\mathfrak{S}$, which we call \emph{state space}, consisting of all the physically distinguishable configurations $\mathfrak{s}$, which we call \emph{states}.
+There exists a bijective map $f_{\Delta t}:\mathfrak{S} \leftrightarrow \mathfrak{S}$, which we call \emph{time evolution operator}, that links initial and final states.
+\end{defn}
+
+\textbf{Justification}. The collection of all possible states from a set. The deterministic and reversible evolution maps one element in the set to one and only one element of the set.$\Box$
+
+In practice, we use physical quantities to identify states.
+For example, the state for an ideal gas is given by pressure, volume and temperature.
+To clarify whether we are referring to a particular value of such quantity, or the quantity itself, we introduce the following terminology:
+\begin{center}
+    \begin{tabular}{ | p{2.5cm} | p{5.5cm} | }
+    \hline
+    State variable & A quantity (discrete or continuous) that must be specified to identify a state (e.g. position). \\ \hline
+    Label & A particular value for a state variable (e.g. position = $5m$). \\ \hline
+    Label range & A set of possible values for a state variable (e.g. position = $[4.5m, 5.5m]$). \\
+    \hline
+    \end{tabular}
+\end{center}
+
+\begin{defn}\label{label}
+Let $\mathbbm{I}$ be a set.
+Let $f_\mathbbm{I} : \mathfrak{S} \rightarrow \mathbbm{I}$ be a function that maps each state to an element of the set.
+We call the pair $(\mathbbm{I}, f_\mathbbm{I})$ a \emph{state variable}.
+We call the pair $(i, f_\mathbbm{I}^{-1}(i))$ a \emph{label}, where $i \in \mathbbm{I}$ and $f_\mathbbm{I}^{-1}(i) = \{\mathfrak{s} \in \mathfrak{S} | f_\mathbbm{I}(\mathfrak{s}) = i\}$ is the reverse image of $i$.
+We call the pair $(I, f_\mathbbm{I}^{-1}(I))$ a \emph{label range}, where $I \subset \mathbbm{I}$.
+\end{defn}
+
+\textbf{Justification}.
+A state variable associates each state to one of its values.
+A label, one of the possible values, can also be thought as representing all states that are identified by its value.
+A label range, a set of possible values, can also be thought as representing all states that are identified by those values.
+Abusing the notation, $\mathbbm{I}$ will represent both the set of possible values and the function, and $i$ both the possible value and the set of states.$\Box$
+
+labels/state variables, cardinality of state variable, state id function (=1 for a particular state).
+Can we do for continuous variables as well?
+Can we find divergence free evolution for det/rev and independent variables?
 

### 2015-10-19 2015CompleteDerivation/completeDerivation.tex
Removed copied part from previous article

@@
 
-\title{From physical principles to relativistic classical Hamiltonian and Lagrangian particle mechanics}
+\title{From physical principles to ???}
 \author{Gabriele Carcassi}
@@
 \begin{abstract}
-We show that classical particle mechanics (Hamiltonian and Lagrangian consistent with relativistic electromagnetism) can be derived from three fundamental assumptions: infinite reducibility, deterministic and reversible evolution, and kinematic equivalence.
-The core idea is that deterministic and reversible systems preserve the cardinality of a set of states, which puts considerable constraints on the equations of motion.
-This perspective links different concepts from different branches of math and physics (e.g.
-cardinality of a set, cotangent bundle for phase space, Hamiltonian flow, locally Minkowskian space-time manifold), providing new insights.
-The derivation strives to use definitions and mathematical concepts compatible with future extensions to field theories and quantum mechanics.\end{abstract}
+Deterministic/reversible assumption.
+Describe fundamental model of physics.
+System is always interacting with environment, both deterministically and non-deterministically.
+Deterministic means only depends on the state of the system.
+Set of states needs to be invariant under non-deterministic interaction: the non-deterministic interaction affects what are the states of the system.
+Deterministic/reversible evolution is necessary to even be able to talk about a system (trivial deterministic process).
+
+Define: set of states, deterministic evolution as permutation, labels/state variables, cardinality of state variable, state id function (=1 for a particular state).
+Can we do for continuous variables as well?
+Can we find divergence free evolution for det/rev and independent variables?
+
+Infinite reducibility assumption.
+Decomposable systems.
+Rule of composition.
+Inverse and null state to study changes.
+Set of transformations to increase and decrease the quantity in each state.
+Define magnitude of the system (probably needed to guarantee things converge in math?).
+
+Define: set of states as a vector space, inner product.
+
+Continuous quantities. Need for densities to be a function of the state (not state variable). Defined on infinitesimal interval. Define: degree of freedom, state cardinality, metric (T*Q).
+\end{abstract}
 \maketitle
@@
 Special relativity is based on two principles (invariance of the speed of light and the principle of relativity), which lead to the Minkowskian nature of space-time but not to the equations of motion, which are a consistent reformulation of the non-relativistic ones.
-What we'll do in this work is re-organize all the known elements and equations in a more consistent and comprehensive way, leading to better insight on why the fundamental concepts and laws are what they are.
-The derivation will also include classical analogues of strictly quantum notions to facilitate a future extension to quantum mechanics.
+Quantum mechanics is simply founded on its mathematical formulation, leaving its physical significance open to debate.
+One can't come to the conclusion that, unfortunately, the landscape of fundamental physics is a piece-meal of concepts and idea, sometime apparently inconsistent with each other.
+Yet: the physical world is one, and our physical theories are simply models of such reality.
+If we clarify what those models represent, and on what assumptions are based, shouldn't we be able to have a coherent and more satisfying picture?
 
-We'll use concepts from different disciplines, such as set theory, differential geometry, relativity, Hamiltonian and Lagrangian mechanics, and we'll find interesting connections among them.
+This work aims to re-organize the known elements and equations in a more consistent and comprehensive way, leading to better insight on why the fundamental concepts and laws are what they are.
+We will start with physical assumptions, that clarify the models we impose on the physical world, and give arguments on when those assumptions can be considered valid.
+We will justify our mathematical definitions, based on the formal properties of the objects of our discussion.
+And will, as a consequence, re-derive known results and theories.
+We will strive to do this in a way that is mathematically meaningful, philosophically consistent and mathematically precise.
+
+TODO: high level view of the derivation
+
+%We'll use concepts from different disciplines, such as set theory, differential geometry, relativity, Hamiltonian and Lagrangian mechanics, and we'll find interesting connections among them.
 We'll keep names and notation as consistent as possible to current use across the different disciplines.
@@
 
-No mathematical breakthrough should be expected: the goal, after all, is to derive the \emph{known} framework from a set of \emph{simple} definitions in the most \emph{obvious} way possible.
+%No mathematical breakthrough should be expected: the goal, after all, is to derive the \emph{known} framework from a set of \emph{simple} definitions in the most \emph{obvious} way possible.
 No proof is longer than a couple of paragraphs, so the word \emph{theorem} is avoided in favor of \emph{proposition} and \emph{corollary}.
@@
 
-\section{Outline}
-
-Deterministic/reversible assumption.
-Describe fundamental model of physics.
-System is always interacting with environment, both deterministically and non-deterministically.
-Deterministic means only depends on the state of the system.
-Set of states needs to be invariant under non-deterministic interaction: the non-deterministic interaction affects what are the states of the system.
-Deterministic/reversible evolution is necessary to even be able to talk about a system (trivial deterministic process).
-
-Define: set of states, deterministic evolution as permutation, labels/state variables, cardinality of state variable, state id function (=1 for a particular state).
-Can we do for continuous variables as well?
-Can we find divergence free evolution for det/rev and independent variables?
-
-Infinite reducibility assumption.
-Decomposable systems.
-Rule of composition.
-Inverse and null state to study changes.
-Set of transformations to increase and decrease the quantity in each state.
-Define magnitude of the system (probably needed to guarantee things converge in math?).
-
-Define: set of states as a vector space, inner product.
-
-Continuous quantities. Need for densities to be a function of the state (not state variable). Defined on infinitesimal interval. Define: degree of freedom, state cardinality, metric (T*Q).
-
-\section{New}
+\section{Determinism and reversibility}
 
@@
 
-\section{Outline}
-
-We give here a brief general conceptual overview, hoping it will help guide the reader through the mathematical details.
-
-We'll first mathematically define states and the labels we use to identify them (e.g.
-position, momentum, temperature, pressure).
-We introduce the infinite reducibility assumption (or classical assumption): each state is divisible into the states of its parts.
-It is then sufficient to describe the evolution of infinitesimal parts (i.e.
-particles) within their state space (i.e.
-phase space).
-
-We introduce the deterministic and reversible evolution assumption: to each state corresponds one and only one future (or past) state.
-The cardinality of a set of states is therefore conserved under evolution.
-This allows us to define a metric $\omega$ on phase space, and the conservation of that metric leads to the Hamiltonian framework.
-That is: we can derive the Hamiltonian framework on its own merit and show that it is equivalent to deterministic and reversible motion.
-
-We introduce the assumption of kinematic equivalence: studying trajectories is equivalent to studying states.
-We show that there must be a link between the space-time metric $g$ and the phase-space metric $\omega$, as both quantities must be conserved under passive coordinate transformations.
-That link will constrain the space-time metric to be locally Minkowskian.
-A transformation between state variables $(q, p)$ and kinematic variables $(x, \dot{x})$ must exist.
-To be invertible, the relationship between velocity and conjugate momentum has to be monotonic, leading to a concave Hamiltonian which allows a Legendre transformation leading to the Lagrangian.
-We can constrain the Hamiltonian further, and show that the most general equation of motion under the three assumptions is the one of a geodesic modified by the force given by a vector potential (such as the one of a relativistic charged particle) and a scalar potential (such as the one of Newtonian gravitation in the non-relativistic case).
-
-\section{States, Labels and Maps}
-
-This and the next section are dedicated to properly defining states.
-In particular, we will need to introduce more precise terminology to be able to make two crucial distinctions.
-The first is between the state of the whole system and one of its parts.\footnote{Whole system vs.
-infinitesimal component is also a source of confusion when comparing classical and quantum states.
-Quantum states are always whole and are always distributions as there is no state attributable to parts of a quantum system.}
-\begin{center}
-    \begin{tabular}{ | p{2.5cm} | p{5.5cm} | }
-    \hline
-    Configuration state & The state of the whole system being described. \\ \hline
-    \raggedright{Particle state (or simply state)} & The state of an infinitesimal part (i.e. particle) of the system. \\ \hline
-    \end{tabular}
-\end{center}
-When talking about particles of a system we must always remember that they are the result of a limit.
-Therefore we will start with discrete definitions, and see how some properties will extend to the continuous case.
-
-The second distinction comes up when dealing with physical quantities. We need to be able to distinguish among a particular value, a set of possible values, and the space of all possible values.
-\begin{center}
-    \begin{tabular}{ | p{2.5cm} | p{5.5cm} | }
-    \hline
-    State variable & A quantity that must be specified to identify a state (e.g. position). \\ \hline
-    Label & A particular value for a state variable (e.g. position = $5m$). \\ \hline
-    Label set & A set of possible values for a state variable (e.g. position = $[4.5m, 5.5m]$). \\
-    \hline
-    \end{tabular}
-\end{center}
-
-We will first define discrete labels using set theory and in the next section generalize to the continuous case using differential geometry.
-These new terms are fundamental as most of the later derivation will be based on \emph{counting the labels} (i.e.
-counting the possible values of a physical quantity) and making sure that such number is conserved (i.e.
-the number of cases is the same).
-
-\begin{defn}\label{statedef}
-Fix a physical system to study. We define the set $\mathbbm{C}$ of all physically distinguishable configurations for that system. Each element $\mathbbm{c}$ we call \emph{configuration state}.
-\end{defn}
-
-\begin{assump1}\label{classical}
-The system is infinitely reducible: it can be thought of as composed of two or more similar but smaller systems, each in its own configuration state, which in turn can be thought of as composed of two or more, ad infinitum.
-We call \emph{particle} such an infinitesimal part.
-\end{assump1}
-
-\begin{defn}\label{classicalPhaseSpace}
-Let $\mathbbm{S}$ be the set of all possible configuration states for a particle.
-We call this set \emph{phase space}.
-We call each $\mathbbm{s} \in \mathbbm{S}$ a \emph{particle state}, or simply \emph{state}.
-\end{defn}
-
-\begin{cor}\label{classicalDistribution}
-Each classical configuration state $\mathbbm{c} \in \mathbbm{C}$ is a distribution over particle states: $\mathbbm{c}=\sum\limits_{\mathbbm{s} \in \mathbbm{S}} D(\mathbbm{s}) \mathbbm{s}$, where $D:\mathbbm{S}\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}$.
-The distribution can be visualized as a histogram over the states in phase space.
-\end{cor}
-
-Under the classical assumption, we can then limit ourselves to study the particles of the system, their states and their properties without loss of generality.
-To help identify states, we introduce the following concepts.
-
-\begin{defn}\label{label}
-We call a \emph{label} a set of states $i\subset\mathbbm{S}$; a \emph{set of labels} a collection of disjoint labels $I | \forall i_1,i_2\in I, i_1\bigcap i_2 = \emptyset$; a \emph{state variable} a set of labels $\mathbbm{I}$ that covers all of phase space: $\bigcup\limits_{i \in \mathbbm{I}}i=\mathbbm{S}$.
-Therefore a state belongs to one and only one label of a state variable.
-\end{defn}
-
-\begin{defn}\label{discreteCardinality}
-Let $I$ be a finite, countable set of labels. We define the \emph{cardinality} $n(I) \rightarrow \mathbbm{N}$ as the number of labels in the set.
-\end{defn}
-
-\begin{defn}\label{labelsCombine}
-Let $I_1$ and $I_2$ be two sets of labels.
-We can define the \emph{combined set}, $\langle I_1, I_2 \rangle$, whose labels consist of all the non-empty intersections of one label of $I_1$ and one of $I_2$.
-If all intersections are non-empty, $I_1$ and $I_2$ are said to be \emph{independent}, and we have $n(\langle I_1, I_2 \rangle)=n(I_1)n(I_2)$.
-\end{defn}
-
-We now want to study how states and labels evolve in time, under the following assumption.
-
-\begin{assump2}
-The system undergoes deterministic (future state identified by the present state) and reversible (past state identified by the present state) evolution.
-\end{assump2}
-
-\begin{prop}\label{detrevMap}
-Let $\mathbbm{S}$ be the phase space of a system that undergoes deterministic and reversible evolution.
-There exists a bijective map $f:\mathbbm{S} \leftrightarrow \mathbbm{S}$ between past and future states.
-\end{prop}
-
-\begin{cor}\label{detrevDist}
-The evolution of a classical configuration state $\mathbbm{c}=\sum D(\mathbbm{s}) \mathbbm{s}$ under a bijective map is given by $\mathbbm{c'}=\sum D'(\mathbbm{s}) \mathbbm{s}=\sum D(f^{-1}(\mathbbm{s})) \mathbbm{s}$.
-The evolution of the fraction of the system in a label $D(i)=\sum\limits_{\mathbbm{S} \in i} D(\mathbbm{s})$ is given by $D'(i)=D(f^{-1}(i))$.
-\end{cor}
-
-Mathematically, assuming determinism and reversibility means studying bijective maps.
-The evolution of a distribution simply moves the elements around: the bars of the histogram move place, but keep the same height.
-
-\begin{cor}\label{labelsCount}
-Given a label $i$, the image $f(i)$ is also a label containing the same number of states.
-Given a set of labels $I$, the image $f(I)$ is also a set of labels containing the same number of labels $n(I) = n(f(I))$.
-Given a state variable $\mathbbm{I}$, the image $f(\mathbbm{I})$ is also a state variable.
-Given two independent sets of labels $I_1$ and $I_2$, the images $f(I_1)$ and $f(I_2)$ are also independent.
-Therefore $n(f(\langle I_1, I_2 \rangle))=n(f(I_1))n(f(I_2))=n(I_1)n(I_2)=n(\langle I_1, I_2 \rangle)$
-\end{cor}
-
-Bijective maps preserve the number of labels as they provide one-to-one association between future and past.
-And they do so for each independent set of labels.
-These simple results using discrete labels, properly generalized to the continuous case, will give us Hamiltonian flow.
-
-\section{Numeric labels}
-
-We now focus on labels that can be identified by numbers, where we have a bijective map between the label and a number in a set.
-The definitions in the previous section readily apply for labels identified by integers.
-Let $z \in \mathbbm{Z}$, we have $\mathbbm{c}=\sum\limits_{z \in \mathbbm{Z}} D(z) \mathbbm{s}(z)$ and the deterministic map becomes $f:\mathbbm{Z} \leftrightarrow \mathbbm{Z}$.
-
-For the continuous case, one may simply expect to replace $z \in \mathbbm{Z}$ with $r \in \mathbbm{R}$, but this does not work.
-In the continuous limit, we would have $\mathbbm{c}=\int\limits_{r \in \mathbbm{R}} \rho(r) dr \mathbbm{s}(r)$, where $\rho(r) = D(r) / dr$.
-The continuous distribution $\rho$ is a density, defined over interval $dr$.
-That is: it's really $\rho(r, dr)$, function of both the center and the width of the interval.
-A bijective map on just $r$ is not sufficient, $dr$ must be mapped as well.
-On one side we claim the state fully identified by $r$, on the other $dr$ (a different label) is required for the density and the bijective map.
-We can't have it both ways.
-
-In the continuous case, then, the appropriate label corresponds to a cell, not a point.
-This also makes physical sense, as we never deal with points per se, but widths that can be made arbitrarily small.
-A cell of $\mathbbm{R}$ corresponds to $\mathbbm{R}^2$: a center and a width.\footnote{In other words: the particles of the system are not point-like, but infinitesimal cell-like.
-Not only does this lead to a more direct understanding of classical phase space, it is also more consistent with general relativity (the mass is spread across a small region, not forming a singularity) and quantum mechanics (position is always the central value of a small distribution).} With this in mind, we have the following definitions.
-
-\begin{defn}\label{sdof}
-A \emph{degree of freedom} is a state variable identified by an infinitesimal cell of a one dimensional manifold $\mathbbm{Q}$.\footnote{For each label $i$ there is one and only one infinitesimal cell.}
-\end{defn}
-
-\begin{defn}\label{generalizedCoordinates}
-We define the \emph{generalized coordinate} $q$ as the center of each cell.
-We define the \emph{cell number}\footnote{This represents the classical analogue of the wave number.} $k$ such that $k \, dq$ represents the width of each cell.
-\end{defn}
-
-\begin{cor}\label{continuousLabels}
-Each degree of freedom is identified by the pair of labels $\langle q,k \rangle \in \mathbf{T}^*\mathbbm{Q}$ where $\mathbf{T}^*\mathbbm{Q}$ is the cotangent bundle of $\mathbbm{Q}$.
-Under coordinate changes, $k\,dq$ and $dq\wedge dk$ are invariant and $k$ is contravariant.
-\end{cor}
-
-Let's start with the discrete case.
-Let $Q$ be a finite region of $\mathbbm{Q}$.
-Divide the region into $N$ equal intervals of center $q$ and length $\Delta q$.\footnote{The definition could be mathematically more general.
-We use equally spaced intervals as it makes the derivation less cumbersome.} Let $K$ be a finite region of $\mathbbm{R}$.
-Divide the region into $M$ equal intervals of center $k$ and length $\Delta k$.
-Let $I$ be the label set identified by the cells in $Q$ with center $q(i)$ and width $k(i) \Delta q$.
-The configuration state will be given by $\mathbbm{c}=\sum\limits_{i \in I} \rho(q(i), k(i)) \Delta q \Delta k \, \mathbbm{s}(q(i), k(i))$, where $\rho(q(i), k(i)) = D(q(i), k(i)) / (\Delta q \Delta k)$.
-
-As we increase $N$ and $M$, the cardinality of the label set $I$ increases, $\Delta q$ and $\Delta k$ decrease and so does the width of the cells.
-Our definitions, though, do not change: for each $\langle q,k \rangle$ we have one and only one cell, one state.
-In the limit, we will cover every possible center $q$ and every possible cell number $k$.
-Our configuration state becomes $\mathbbm{c}=\int\limits_{q \in Q \; k\in K} \rho(q, k) dq \wedge dk \, \mathbbm{s}(q, k)$.
-We can repeat the process increasing or changing the region covered by $Q$ and $K$, until we cover all of $\mathbbm{Q}$ and $\mathbbm{R}$.
-
-If we apply a coordinate change $q'=q'(q)$, we change the labels but the state must remain the same, defined at the same point with the same width.\footnote{The notion that states are what they are no matter how we label them already contains the principle of relativity.} $k' dq'$ must be then equal to $k dq$; $k'= k dq / dq'$ is contravariant, $dq' \wedge dk'= (dq'/dq) \, (dq \wedge dk) \, (dq/dq') = dq \wedge dk$ is invariant.
-This means that $\rho$ is also invariant, which makes sense: the density depends only on the state (the cell), and not the coordinate chosen to represent it.
-As $k$ is contravariant, in the limit $K$ becomes the cotangent space of $Q$.
-Phase space is the cotangent bundle $\mathbf{T}^*\mathbbm{Q}$.
-
-\begin{defn}\label{relativeCardinality}
-Let $I \subset \mathbf{T}^*\mathbbm{Q}$ be a closed dense set of labels, a subset of a degree of freedom (i.e.
-a region of phase space).
-We define \emph{relative cardinality} $n(I) \rightarrow \mathbbm{R}$ as the ratio between the number of labels in $I$ and the ones of a reference set $I_0$.
-We find $n(I)=\int \omega$, where $\omega = \hbar \, dq \wedge dk$ and $\hbar$ is the constant that determines the unit (i.e.
-fixed so that $n(I_0) = 1$).
-\end{defn}
-
-As we deal with continuous labels, a label set is uncountable: given any range $\delta q$ and $\delta k$ there are an infinite number of possible labels $I$.
-Let $I_0$ be another set, defined on $\delta q_0$ and $\delta k_0$.
-In the discrete case, we have:
-
-\begin{align*}
-\frac{n(I)}{n(I_0)} = \frac{\delta q \delta k / \Delta q \Delta k}{\delta q_0 \delta k_0 / \Delta q \Delta k} = \frac{\delta q \delta k}{\delta q_0 \delta k_0}
-\end{align*}
-
-That is: while the cardinality of each set diverges, the ratio between the two remains finite.\footnote{Such treatment is equivalent to what one does in information theory to extend Shannon's entropy\cite{Shannon} to the continuous case.
-There is a link between label cardinality and informational entropy, including Jaynes' invariant formulation\cite{Jaynes}, which we do not explore here for brevity.} For continuous state variables we can define the relative cardinality as the ratio of a label set and our reference set $I_0$.
-That is $n(I)=\hbar \delta q \delta k$ where $\hbar=n(I_0)/(\delta q_0 \delta k_0)$ and $n(I_0)\equiv1$ by definition.\footnote{The choice of angular momentum for $\hbar$ derives from the relationship between $p$ and $dq/dt$, which will be derived later.
-The value for $\hbar$ in classical mechanics is arbitrary: we have no physical reason to choose a reference label set over another.
-In quantum mechanics, instead, we will have a preference: the set that corresponds to a single quantized system.
-We use this arbitrariness to set the value of our classical analogue of $\hbar$ to the known constant.} In general, the label set is not rectangular in $\langle q,k \rangle$ and we have $n(I)=\int \hbar dq \wedge dk$.\footnote{It should be stressed that no notion of uncertainty is used here.
-Though there is a link between label cardinality and uncertainty, which we do not explore here for brevity, label cardinality is the more fundamental quantity as we can still count the number of possible states even in an ideal case of no uncertainty.}
-
-\begin{defn}\label{conjugateMomentum}
-We define the \emph{conjugate momentum} as $p=\hbar k$.
-\end{defn}
-
-As $\hbar dq \wedge dk$ is truly fundamental, it is customary to group $\hbar$ with $k$: $dq \wedge (\hbar dk) = dq \wedge dp$.
-This way the area formed by generalized coordinate and conjugate momentum corresponds to the cardinality of states defined on such area.
-
-\begin{cor}\label{continuousConjugateRelationships}
-Each degree of freedom is identified by the pair of labels $\langle q,p \rangle \in \mathbf{T}^*\mathbbm{Q}$.
-$\theta_0 = p dq$ and $\omega = dq \wedge dp$ are invariant; $p$ is contravariant.
-The configuration state is $\mathbbm{c}=\int \rho(q,p) dq \wedge dp \, \mathbbm{s}(q, p)$, where $\rho(q,p)\equiv D(q,p) / \omega$ is the distribution density for each label.
-\end{cor}
-
-This restates all the previous findings in terms of conjugate momentum.
-In the language of differential geometry, we recognize phase space as the cotangent bundle $\mathbf{T}^*\mathbbm{Q}$, the function $\rho$, the one-form $\theta_0$ and the two-form $\omega$.
-They are truly fundamental objects as they are intimately linked to the way states are defined ($\rho$ is the distribution, $\theta_0 / \hbar$ is the cell width) and counted ($\omega$ is the cardinality of the labels) and they don't depend on coordinate choice.
-
-As we'll need to work with different physical units, which may change $q$ or $p$ independently, it is useful to introduce the following notation and concepts.
-
-\begin{defn}\label{differentialCardinality}
-Let $dq$ be a finite contiguous range for a continuous state variable, we define $n_q(dq)$ as the cardinality of the set $I_{dq}$ that comprises all the labels in the range.
-For contravariant (conjugate) variables, we reverse the indices in the notation: $n^p(dp)$.
-\end{defn}
-
-\begin{cor}\label{linearCardinality}
-Given $dq = a dq^a + b dq^b$, then $n_q(dq) = a n_q(dq^a) + b n_q(dq^b)$. That is, the relative cardinality is a linear operator.
-\end{cor}
-
-The linearity should be obvious: as we combine or increase the range, we combine or increase the number of possible labels.\footnote{Since the relative cardinality is linear, invariant under coordinate transformation and conserved by deterministic and reversible motion, it should come as no surprise that linearity will play a significant role.
-This is true for both classical (e.g.
-forces combine linearly) and quantum (e.g.
-linear inner product) concepts that are tightly linked to this fundamental operator.}
-
-\begin{defn}\label{cardinalityConversionConstant}
-Let $q^a$ and $q^b$ be two continuous state variables measured in different units, and $n_a(dq^a)$ and $n_b(dq^b)$ the relative cardinality defined on intervals of the respective units.
-We define the \emph{cardinality conversion constant} $\mathfrak{C}^b_a$ that converts a range of $q^a$ to a range of $q^b$ with the same cardinality.
-That is: $dq^b = \mathfrak{C}^b_a dq^a$ and $n_b(dq^b)=n_a(dq^a)$.
-For contravariant (conjugate) variables, we reverse the indices in the notation: $dp_b = \mathfrak{C}^a_b dp_a$.
-\end{defn}
-
-As the relative cardinality $n_q(dq)$ is invariant, the function $n_q$ cannot be since $dq$ is covariant.
-This means that, when changing units, we have to be careful.\footnote{This is only a problem when considering the number of labels of a single state variable $n_q(dq)$, instead of a whole degree of freedom $n_{\langle q, p \rangle}(\langle dq, dp \rangle)$: $\langle dq, dp \rangle$ is invariant and so is $n_{\langle q, p \rangle}$.}
-Fortunately, since the $n_q$ is linear, we can always find a coefficient that allows us to convert.
-For example: $\hbar \equiv \mathfrak{C}^k_p$ is the cardinality conversion constant between cell number and conjugate momentum.
-In the future, we will have to convert intervals of time into space, or intervals of velocity into conjugate momentum, so the idea of cardinality conversion constants will reappear.
-
-\section{Single degree of freedom}
-
-Now that we have properly defined continuous labels and the relative cardinality of their sets, we will extend bijective maps to the continuous case.
-As these must preserve the cardinality of label sets, they are constrained in their form.
-The use of a bijective map for infinitesimal time evolution will lead us to Hamilton's equations.
-In this section we'll study a single degree of freedom, then extend in the next two sections to multiple degrees of freedom and to the time dependent case.
-
-\begin{defn}\label{canonical}
-We call \emph{canonical transformation} the continuous limit of a bijective map, as defined by \ref{detrevMap}, on $\mathbf{T}^*\mathbbm{Q}$.
-\end{defn}
-
-\begin{prop}\label{continuousMapping}
-A canonical transformation must be continuous and preserve $\omega$.
-\end{prop}
-
-As we saw in \ref{labelsCount}, a bijective map conserves the cardinality of labels.
-In the continuous case, it will conserve relative cardinality and therefore $\omega$.
-The map must be continuous in $q$: suppose it isn't, it would split some cells into two parts, a cell would not be mapped to one and only one other cell, the mapping would not be bijective.
-The reverse mapping must be continuous in $q$ as well, or the inverse would not map to one and only one cell.
-\begin{align*}
-dq' &= \frac{\partial q'}{\partial q} dq + \frac{\partial q'}{\partial p} dp \\
-dq &= \frac{\partial q}{\partial q'} dq' + \frac{\partial q}{\partial p'} dp'
-\end{align*}
-We can re-express $dp'$ in terms of $dq$ and $dp$, as the conservation of $\omega$ means the map is non-degenerate.
-All partial derivatives are well defined, and therefore the mapping is continuous in $p$ as well.
-
-\begin{cor}\label{sdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of the phase space $\mathbf{T}^*\mathbbm{Q}$ for one degree of freedom. Let
-\begin{align*}
-\omega_{ab} = \left[
-  \begin{array}{cc}
-    0 & 1 \\
-    -1 & 0 \\
-  \end{array}
-\right] \\
-\end{align*}
-then $v'^{a} \omega_{ab} w'^{b}=v^{a} \omega_{ab} w^{b}$ under a canonical transformation, where $a$ and $b$ represent components along coordinates $\xi^{a} \equiv \{q,p\}$.
-\end{cor}
-
-Here we are simply expressing $\omega$, $v$ and $w$ in terms of their respective components and underlining the fact that $\omega$ defines the metric conserved under canonical transformations.
-
-\begin{lem}\label{genAntisim}
-Let $v$ and $w$ be two vectors.
-Let $v^{a} \omega_{ab} w^{b}$ be an antisymmetric product conserved under a continuous transformation parameterized by $t$.
-We can then define a function $H$ such that given $S^{a} \equiv d_{t}\xi^{a}$ and $S_{b} \equiv S^{a} \omega_{ab}$, we have $S_{a} = \partial_{a}H$.
-\end{lem}
-
-$S^{a}$ is the vector field that represents how the state variables change. Simply applying the vector transformation rules under continuous transformation we have:
-\begin{align*}
-v^{a} \omega_{ab} w^{b} &= v'^{a} \omega_{ab} w'^{b}  \\
-&= (v^{a} + \partial_{c} S^{a} v^{c} dt) \omega_{ab} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
-&= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} \\
- &+ v^{a} \omega_{ab} \partial_{d} S^{b} w^{d}) dt + O(dt^2)
-\end{align*}
-\begin{align*}
-v^{c} w^{b} \partial_{c} S_{b} - v^{a} w^{d} \partial_{d} S_{a} = 0
-\end{align*}
-\begin{align*}
-\partial_{a} S_{b} - \partial_{b} S_{a} &= curl(S_{a}) = 0 \\
-S_{a} &= \partial_{a}H
-\end{align*}
-
-\begin{prop}\label{sdofHam}
-The time evolution for a single degree of freedom is given by:
-\begin{align*}
-d_{t}q &= \partial_{p} H \\
-d_{t}p &= - \partial_{q} H
-\end{align*}
-\end{prop}
-
-Simply expand \ref{genAntisim} with the metric defined in \ref{sdofInvariant}. We recognize Hamilton's equations for one degree of freedom\cite{classical_dynamics}.
-
-\section{Multiple independent degrees of freedom}
-
-\begin{prop}\label{mdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of the phase space $\mathbf{T}^*\mathbbm{Q}$ for two independent degrees of freedom.
-Let $a$ and $b$ be indices for the state variables $\xi^a\equiv \{q^i, p_i\}$.
-Let
-\begin{align*}
-\omega_{ab} =  \left[
-  \begin{array}{cc}
-    0 & 1 \\
-    -1 & 0 \\
-  \end{array}
-\right] \otimes \left[
-  \begin{array}{cc}
-    1 & 0 \\
-    0 & 1 \\
-  \end{array}
-\right] =
-\left[
-  \begin{array}{cccc}
-    0 & 0 & 1 & 0 \\
-    0 & 0 & 0 & 1 \\
-    -1 & 0 & 0 & 0 \\
-    0 & -1 & 0 & 0 \\
-  \end{array}
-\right] \\
-\end{align*}
-then $v'^{a} \omega_{a b} w'^{b}=v^{a} \omega_{ab} w^{b}$ under a canonical transformation.
-\end{prop}
-
-The independence between degrees of freedom corresponds to orthogonality in phase space: from \ref{labelsCombine} the product between the number of labels on each d.o.f.
-(i.e.
-the area), must be equal to the number of combined labels (i.e.
-the hyper-volume), which is true only if the d.o.f are orthogonal in phase space.
-From \ref{labelsCount}, the mapping will preserve the cardinality of labels, the area\footnote{We assume we are using the same unit across d.o.f.} on each d.o.f, and the independence, orthogonality across d.o.f.\footnote{These statements provide a direct physical interpretation for Gromov's non-squeezing theorem\cite{Gromov,deGosson,Stewart}.} This is equivalent to requiring the conservation of the scalar product across independent degrees of freedom, while still requiring conservation of the vector product within.
-That leads us to the metric defined by \ref{mdofInvariant}.
-The metric generalizes \ref{sdofInvariant} to give us the cardinality of labels defined on the area given by two arbitrary directions in phase space.
-For an infinitesimal region, this corresponds to $dq^1 \wedge dp_1 + dq^2 \wedge dp_2$, the sum of the projections on the independent planes.
-Moreover, volume in phase space corresponds to the cardinality of the combined labels (i.e.
-the states), and is therefore conserved: this is Liouville's theorem for Hamiltonian mechanics.
-
-\begin{prop}\label{mdofHam}
-The evolution for multiple degrees of freedom is given by:
-\begin{align*}
-d_{t}q^i &= \partial_{p_i} H \\
-d_{t}p_i &= - \partial_{q^i} H
-\end{align*}
-\end{prop}
-
-Expand \ref{genAntisim} with the metric defined in \ref{mdofInvariant}. We recognize Hamilton's equations for multiple degrees of freedom\cite{classical_dynamics}.
-
-\section{Time dependence}
-
-So far we have assumed that neither state labeling nor mapping changes in time. If they do, we also need to to use time as a label and therefore introduce an extra degree of freedom.
-
-\begin{defn}\label{tdof}
-We call \emph{extended phase space} the cotangent bundle $\mathbf{T}^*\mathcal{M}$, where the manifold $\mathcal{M}$ identifies the possible center values for all infinitesimal cells at all times.
-We call \emph{temporal degree of freedom} the state variable identified by temporal cells.
-The center of each cell is identified by $t$, the width by $\hat{\omega} dt$, the conjugate variable $E\equiv\hbar\hat{\omega}$.\footnote{$\hat{\omega}$ is the classical analogue of the wave frequency.
-We use $\hat{\omega}$ to distinguish from the phase-space metric $\omega$.}
-\end{defn}
-
-\begin{prop}\label{tdofMonotonic}
-Let $s$ be the parameter of a trajectory in the extended phase space of a deterministic and reversible system. The trajectory must be continuous. There must exist a strictly monotonic function $t(s)$.
-\end{prop}
-
-The trajectory has to be continuous in both standard and temporal variables because of \ref{continuousMapping}.
-Since determinism and reversibility are defined in time, the trajectory must traverse all times once and only once: we must have an invertible mapping between $t$ and $s$, which means we must have a strictly monotonic $t(s)$.
-
-\begin{defn}\label{tdofAntistates}
-We call \emph{standard states} those connected by a trajectory where $d_{s}t>0$. We call \emph{anti-states} those connected by a trajectory where $d_{s}t<0$.
-\end{defn}
-
-Since $t(s)$ is strictly monotonic, $d_{s}t$ along a trajectory cannot change sign, so we have the division between standard and anti-states.
-Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call standard and anti-states is also conventional.
-What is physical and not conventional, though, is that standard and anti-states cannot be connected by deterministic and reversible evolution.\footnote{This represents a classical analogue for quantum anti-particle states.}
-
-\begin{prop}\label{tdofInvariant}
-Let $v$ and $w$ be two vectors defined on the tangent space of extended phase space $\mathbf{T}^*\mathcal{M}$ for the temporal degree of freedom and one standard degree of freedom.
-Let $a$ and $b$ be indices for the state variables $\xi^a\equiv\{t, q, E, p\}$.
-Let
-\begin{align*}
-\omega_{ab} =  \left[
-  \begin{array}{cc}
-    0 & 1 \\
-    -1 & 0 \\
-  \end{array}
-\right] \otimes \left[
-  \begin{array}{cc}
-    -1 & 0 \\
-    0 & 1 \\
-  \end{array}
-\right]
-= \left[
-  \begin{array}{cccc}
-    0 & 0 & -1 & 0 \\
-    0 & 0 & 0 & 1 \\
-    1 & 0 & 0 & 0 \\
-    0 & -1 & 0 & 0 \\
-  \end{array}
-\right] \\
-\end{align*}
-then $v'^{a} \omega_{ab} w'^{b}=v^{a} \omega_{ab} w^{b}$ under deterministic and reversible evolution.
-\end{prop}
-
-$\langle t, E \rangle$ are not independent from $\langle q, p \rangle$ as they do not define new states.
-So they are not necessarily orthogonal in the extended phase space.
-States are defined on the plane where $\langle q, p \rangle$ (maximally) change: this is not the plane of constant $\langle t, E \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle q, p \rangle$ where $dt \wedge dE$ is defined.
-On that plane we can properly count states and define our metric.
-
-We have a right triangle-like relationship between the plane where the metric is defined and its projections on the planes defined by each d.o.f., similar to the multiple d.o.f.:
-\begin{align*}
-m.d.o.f \;\;\; &dq^1 \wedge dp_1 + dq^2 \wedge dp_2 = \omega \\
-t.d.o.f \;\;\; &dt \wedge dE + \omega = dq \wedge dp \\
-\end{align*}
-But in the previous case, the right angle was between the two independent d.o.f..
-In this case, the right angle is between the metric and the plane of constant $\langle q, p \rangle$ where $dt \wedge dE$ is defined.
-We rewrite it as $dq \wedge dp - dt \wedge dE = \omega$.
-This corresponds to the Minkowski product across d.o.f.
-and the vector product within.
-The metric \ref{tdofInvariant}, with a space-like convention, still gives us the cardinality of labels within a degree of freedom, adjusting \ref{sdofInvariant} to avoid ``double counting''.
-
-\begin{prop}\label{tdofHam}
-The evolution for time varying multiple degrees of freedom is given by:
-\begin{align*}
-d_{s}t &= - \partial_{E} \mathcal{H} \\
-d_{s}E &= \partial_{t} \mathcal{H} \\
-d_{s}q^i &= \partial_{p_i} \mathcal{H} \\
-d_{s}p_i &= - \partial_{q^i} \mathcal{H}
-\end{align*}
-\end{prop}
-
-Take the metric from \ref{tdofInvariant}, add multiple independent d.o.f as in \ref{mdofInvariant}, use \ref{genAntisim} with the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$.
-
-If we set\footnote{We avoided using $p_{n+1}$ as it hides the minus sign from the metric, making it seem that the temporal d.o.f is just another independent d.o.f.} $q^{n+1}=t$ and $p_{n+1}=-E$, we recognize Hamilton's equations in the extended phase space\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}\cite{Synge,Lanczos}.
-
-\begin{prop}\label{tdofConstrain}
-The evolution is constrained by $\mathcal{H}=k$.
-\end{prop}
-
-Since $\mathcal{H}$ is constant through the evolution, it can serve both as the generating function and as the evolution constraint.
-By convention, we can set $\mathcal{H}=0$ without loss of generality as changing $\mathcal{H}$ by a constant does not change the equation of motion.
-This reduces extended phase space to $2N + 1$ components, the state variables plus time.
-
-\section{Kinematics}
-In the previous section we made no constraint on what our state variables $q^i$ actually represent.
-In this section we turn our attention to the study of the motion of a body.
-In particular, we will introduce another assumption: that the trajectories are enough to fully describe the system.
-We can expect this to hold true if the system is elementary (it has no relevant internal structure) and is sufficiently isolated.
-At that point, if the motion is the result of a deterministic and reversible process, no two trajectories can be attributed to the same state as there is nothing else that could affect them.
-As position and time will be state variables, the space-time manifold will coincide with $\mathcal{M}$ and the extended phase space with $\mathbf{T}^*\mathcal{M}$.
-
-\begin{assump3}\label{kinematicAssumption}
-The study of the trajectory (kinematics) of a body is equivalent to study of its state (dynamics) under deterministic and reversible evolution.
-\end{assump3}
-
-\begin{cor}\label{singleTrajectory}
-Fix a system under deterministic and reversible evolution.
-Given all its possible trajectories $x^\alpha(s)$ in the space-time manifold $\mathcal{M}$ and all its possible trajectories $\xi^a(s)$ in the extended phase space $\mathbf{T}^*\mathcal{M}$, there exists a bijective function $f: x^\alpha(s) \leftrightarrow \xi^a(s)$ that links each space-time trajectory with one and only one phase-space trajectory.
-\end{cor}
-
-If studying the motion and state evolution are equivalent, then we must be able to go back and forth between the two pictures.
-Without losing generality, we can choose $x^\alpha=f(t,q^i)$ to be a linear function of only $q^i$ and $t$.
-
-\begin{defn}\label{invariantSpeed}
-Let $x$ be a space variable and $t$ the time variable.
-We define the \emph{invariant speed} as the cardinality conversion constant $c \equiv \mathfrak{C}^x_t$ between $x$ and $t$.
-That is, if $dx=cdt$ then $n_x(dx)=n_t(dt)$.
-\end{defn}
-
-\begin{prop}\label{locallyMinkowski}
-The space-time manifold $\mathcal{M}$ is a locally Minkowskian Riemannian manifold.
-That is, there exists a metric $g$ that at any point P can be expressed, with a suitable choice of coordinate, as $g=dx^\alpha g_{\alpha \beta}(P) dx^\beta=dx^\alpha\eta_{\alpha \beta}dx^\beta=(dx^i)^2 - (dx^0)^2=(dx^i)^2 - c^2dt^2$, where $\eta_{\alpha \beta}$ is the Minkowski metric.
-\end{prop}
-
-As states are defined on intervals, a metric $g$ must be defined on $\mathcal{M}$.
-Such a metric must be consistent with $\omega$ as defined on $\mathbf{T}^*\mathcal{M}$, as both must be invariant under coordinate transformations.
-The idea is that each $(dx^\alpha)^2$ can be made to represent both a length squared in space-time and an area in phase space, linking the two metrics.
-For each spatial d.o.f.
-fix $dp_i=\lambda dq^i$, we have $dq^i \wedge dp_i = \lambda (dq^i)^2$ where $\lambda$ converts from length squared to the label cardinality contained in the area.
-For the temporal d.o.f fix $dE = \lambda c^2 dt$, we have $dt \wedge dE = \lambda c^2 dt^2$.
-That is: the area in time squared is converted to an area in length squared that has the same density of states (as per definition of $c$) and then is converted again to its label cardinality.
-The phase-space metric is $\omega = \lambda [(dq^i)^2 - c^2 dt^2]$.
-
-Assume the choice of coordinates $x^\alpha$ locally diagonalizes $g(x^\alpha)$, each diagonal element being either $\pm 1$ (such coordinate system always exists).
-Set $x^i=q^i$ and $x^0=ct$.
-We have $\omega = \lambda ((dx^i)^2 - (dx^0)^2)$ and $g=dx^\alpha g_{\alpha \beta} dx^\beta=dx^\alpha g_{\alpha \alpha}dx^\alpha$.
-Both are invariant under coordinate transformation for any $dx^\alpha$.
-This can only be if $g_{ii}(x^\alpha)=1$ and $g_{00}(x^\alpha)=-1$.
-
-It is fitting that deterministic and reversible evolution requires space-time to be locally Minkowskian, as this clearly defines past and future events.
-To make us understand better the role of $c$, as we defined it, we prove the following.
-
-\begin{prop}\label{speedLimit}
-The speed of a body cannot exceed the invariant speed $c$ under the kinematic equivalence assumption.
-\end{prop}
-
-Consider a movement $ds$ along any trajectory.
-This will go through $|n_t(dt)|$ labels in time and $|n_x(dx)|$ labels in space.
-As we go through labels in time, we may go through fewer labels in space (e.g.
-the particle remains still) but not more: we cannot skip space labels or the motion would not be continuous.
-So we have:
-\begin{align*}
-\frac{|n_x(dx)|}{|n_t(dt)|} &= \frac{|n_x(dx)|}{|n_x(c dt)|} = \frac{|dx|}{|c dt|} \leq 1 \\
-\frac{|dx|}{|dt|} & \leq c \\
-\end{align*}
-
-Thus we find that $c$ is the well known relativistic constant.\footnote{Deterministic and reversible motion at speeds greater than $c$ (i.e.
-a tachyon) is not allowed.
-Non-deterministic and non-reversible motion at speeds greater than $c$ (e.g.
-correlations at a distance) is allowed.}
-
-\begin{prop}\label{initialConditions}
-Let $x^\alpha=\{ct, q^i\}$ and $u^\alpha = d_s x^\alpha$ be the four-velocity, where the parametrization $s$ is chosen, by convention, to be proper time.
-The position $x^\alpha$ and velocity $u^\alpha$ are necessary and sufficient initial conditions to determine the state of the system and its whole trajectory.
-\end{prop}
-
-As the equations of motion \ref{tdofHam} can be at most second order in $\{t, q^i\}$, they can at most be second order in $x^\alpha=\{ct, q^i\}$ so only position and velocity can be candidates for the initial conditions.
-Fixing time, phase space has $2N$ components, too big to be covered by position only, but just right to be covered by both.
-
-\begin{defn}\label{inertialMass}
-Let $p_\alpha$ be a component of conjugate momentum and $u_\alpha \equiv g_{\alpha \beta}u^\beta$ be a contravariant component of the four velocity.
-We define the \emph{inertial mass} as the cardinality conversion constant $m \equiv \mathfrak{C}^p_u$ between $p$ and $u$.
-That is, if $dp_\alpha=mdu_\alpha$ then $n^p(dp_\alpha)=n^u(du_\alpha)$.\footnote{Note that the kinematic assumption rules out massless particles: they can have the same position and velocity with different momentum, therefore the same trajectory for different states.
-When generalizing to field theory, the derivative of the position will be replaced by the derivative of the field (i.e.
-its oscillation in space-time), which holds for massive and massless fields.}
-\end{defn}
-
-\begin{prop}\label{kineticMomentum}
-Let $x^\alpha=\{ct, q^i\}$ and $p_\alpha=\{-E/c, p_i\}$.
-Then $p_\alpha= m g_{\alpha \beta}u^\beta + \hat{p}_\alpha(x)$, where $\hat{p}_\alpha : \mathcal{M} \rightarrow \mathbbm{R}$ is a function defined on the space-time manifold.
-\end{prop}
-
-Given \ref{initialConditions}, there must exist $p_\alpha=p_\alpha(x^\beta , u^\gamma)$. We express it in terms of $u_\alpha\equiv g_{\alpha \beta} u^\beta$. We have:
-\begin{align*}
-\omega &= q^i\wedge p_i - ct \wedge E/c \\
-&=x^\alpha \wedge p_\alpha \\
-&=dx^\alpha \wedge \frac{\partial p_\alpha}{\partial u_\beta}du_\beta + dx^\alpha \wedge \frac{\partial p_\alpha}{\partial x^\gamma}dx^\gamma \\
-&=\frac{\partial p_\alpha}{\partial u_\beta}du_\beta dx^\alpha \\
-\end{align*}
-Consider the expression $m du_\alpha dx^\alpha$: this invariant gives us the density of labels in position and velocity converted to the conjugate variables.
-That is: $\omega=m du_\alpha dx^\alpha$.
-Combining the two:
-\begin{align*}
-\frac{\partial p_\alpha}{\partial u_\beta} &= m \delta^\beta_\alpha \\
-p_\alpha &= m g_{\alpha \beta}u^\beta + \hat{p}_\alpha(x)
-\end{align*}
-where $\hat{p}$ is an arbitrary function.\footnote{The conjugate momentum $p$ and the kinetic momentum $mu$ are therefore different in general: they are linked only through their differentials.
-Moreover, $p$ is unique only once $\hat{p}$ is fixed, but $\hat{p}$ is arbitrary.
-Given that $u$ determines the trajectory, any transformation in $\{p, \hat{p} \}$ that does not change $u$ describes the same motion.
-We call \emph{gauge transformation} such a transformation that redefines conjugate momentum while keeping kinetic momentum unvaried.
-This is the particle mechanics equivalent of gauge transformations in field theory.}
-
-To convince ourselves that $m$ is indeed the inertial mass, consider applying a force.
-We are changing the state through the velocity, meaning changing the conjugate momentum.
-The higher the mass, the more states we'll have to go through to reach the same velocity.
-The higher the mass, the more change is required, the more force needs to be applied.
-
-\begin{prop}\label{kineticHamiltonian}
-The extended Hamiltonian is $\mathcal{H}=\frac{1}{2m}(p_\alpha-\hat{p}_\alpha(x))g^{\alpha\beta}(p_\beta-\hat{p}_\beta(x))+\hat{\mathcal{H}}(x)$, where $\hat{\mathcal{H}} : \mathcal{M} \rightarrow \mathbbm{R}$ is a function defined on the space-time manifold.
-\end{prop}
-We have:
-\begin{align*}
-\frac{dx^\alpha}{ds} &= u^\alpha \\
-&= \frac{1}{m}g^{\alpha\beta}(p_\beta-\hat{p}_\beta) \\
-&= \frac{\partial \mathcal{H}}{\partial p_\alpha} \\
-\end{align*}
-Integrating we have the expression for the Hamiltonian, where $\hat{\mathcal{H}}$ is an arbitrary function.
-
-Now that we have found the constrained form of the Hamiltonian, we show that this is compatible with the established fundamental classical theories.
-
-\begin{prop}\label{relativisticEM}
-Let $\hat{p}_\alpha = q A_\alpha$, $F_{\alpha \beta} \equiv \partial_\alpha A_\beta - \partial_\beta A_\alpha$ and $\hat{\mathcal{H}} = 0$.
-Then the equations of motion are $m \nabla_u u^\alpha = g^{\alpha\beta} F_{\beta \gamma} q u^\gamma$.
-These are the relativistic equations for a charged particle.
-\end{prop}
-
-We first derive the following relationship for later use:
-\begin{align*}
-\partial_\alpha \delta^\beta_\gamma &= 0 = \partial_\alpha g^{\beta\delta} g_{\delta\gamma} + g^{\beta\delta} \partial_\alpha g_{\delta\gamma}\\
-\partial_\alpha g^{\beta\epsilon} &= - g^{\beta\delta} g^{\gamma\epsilon} \partial_\alpha g_{\delta\gamma}
-\end{align*}
-
-We expand \ref{tdofHam} using \ref{kineticHamiltonian}:
-\begin{align*}
-u^\alpha &= \frac{dx^\alpha}{ds} = \frac{\partial \mathcal{H}}{dp_\alpha} \\
-&= \frac{1}{m}g^{\alpha\beta}(p_\beta-\hat{p}_\beta) \\
-d_s p_\alpha &= - \frac{\partial \mathcal{H}}{\partial q^\alpha} \\
-&=\frac{1}{2m}[\partial_\alpha \hat{p}_\beta g^{\beta \gamma} (p_\gamma -\hat{p}_\gamma) \\
- &- (p_\beta -\hat{p}_\beta) \partial_\alpha g^{\beta \gamma} (p_\gamma -\hat{p}_\gamma) \\
- &+ (p_\beta -\hat{p}_\beta) g^{\beta \gamma} \partial_\alpha \hat{p}_\gamma ]- \partial_\alpha \hat{\mathcal{H}} \\
-&=\frac{1}{2}[\partial_\alpha \hat{p}_\beta u^\beta
-- m u^\delta g_{\delta\beta} \partial_\alpha g^{\beta \gamma} u^\epsilon g_{\epsilon\gamma}
-+ u^\gamma \partial_\alpha \hat{p}_\gamma ]- \partial_\alpha \hat{\mathcal{H}}
-\end{align*}
-
-We then calculate the four-force:
-\begin{align*}
-m d_s u^\alpha &= d_s g^{\alpha\beta}(p_\beta-\hat{p}_\beta) + g^{\alpha\beta} d_s (p_\beta-\hat{p}_\beta) \\
-&= \partial_\gamma g^{\alpha\beta} d_s x^\gamma m g_{\beta \delta} u^\delta + g^{\alpha\beta} (d_s p_\beta - \partial_\gamma \hat{p}_\beta d_s x^\gamma) \\
-&= \partial_\gamma g^{\alpha\beta} u^\gamma m g_{\beta \delta} u^\delta + g^{\alpha\beta} \frac{1}{2} [\partial_\beta \hat{p}_\gamma u^\gamma \\
-&- m u^\epsilon g_{\epsilon\gamma} \partial_\beta g^{\gamma \delta} u^\zeta g_{\zeta\delta} \\
-&+ u^\delta \partial_\beta \hat{p}_\delta ]- g^{\alpha\beta} \partial_\beta \hat{\mathcal{H}} - g^{\alpha\beta} \partial_\gamma \hat{p}_\beta u^\gamma \\
-&= m  g_{\beta \delta} \partial_\gamma g^{\alpha\beta} u^\gamma u^\delta - \frac{1}{2} m g^{\alpha\beta} g_{\zeta\delta} g_{\epsilon\gamma} \partial_\beta g^{\gamma \delta} u^\epsilon u^\zeta  \\
-&+ g^{\alpha\beta} \partial_\beta \hat{p}_\gamma u^\gamma - g^{\alpha\beta} \partial_\gamma \hat{p}_\beta u^\gamma
-- g^{\alpha\beta} \partial_\beta \hat{\mathcal{H}}\\
-&= - m  g^{\alpha \beta} \partial_\gamma g_{\beta\delta} u^\gamma u^\delta + \frac{1}{2} m g^{\alpha\beta} \partial_\beta g_{\gamma \delta} u^\gamma u^\delta  \\
-&+ g^{\alpha\beta} (\partial_\beta \hat{p}_\gamma - \partial_\gamma \hat{p}_\beta ) u^\gamma
-- g^{\alpha\beta} \partial_\beta \hat{\mathcal{H}}\\
-&= - m \frac{1}{2} g^{\alpha \beta} ( \partial_\gamma g_{\beta\delta} + \partial_\delta g_{\beta\gamma} - \partial_\beta g_{\gamma \delta} ) u^\gamma u^\delta  \\
-&+ g^{\alpha\beta} (\partial_\beta \hat{p}_\gamma - \partial_\gamma \hat{p}_\beta ) u^\gamma
-- g^{\alpha\beta} \partial_\beta \hat{\mathcal{H}}\\
-&= - m \Gamma ^\alpha_{\ \gamma \delta} u^\gamma u^\delta + g^{\alpha\beta} (\partial_\beta \hat{p}_\gamma - \partial_\gamma \hat{p}_\beta ) u^\gamma
-- g^{\alpha\beta} \partial_\beta \hat{\mathcal{H}}
-\end{align*}
-\begin{align*}
-m \nabla_{u} u^\alpha &= m (d_s u^\alpha + \Gamma ^\alpha_{\ \gamma \delta} u^\gamma u^\delta)  \\
-&= g^{\alpha\beta} (\partial_\beta \hat{p}_\gamma - \partial_\gamma \hat{p}_\beta ) u^\gamma - g^{\alpha\beta} \partial_\beta \hat{\mathcal{H}}\\
-&= g^{\alpha\beta} (\nabla_\beta \hat{p}_\gamma - \nabla_\gamma \hat{p}_\beta ) u^\gamma - g^{\alpha\beta} \nabla_\beta \hat{\mathcal{H}}
-\end{align*}
-where $\nabla$ is the covariant derivative and $\Gamma$ the Christoffel symbols.
-
-The equation is manifestly covariant.
-If $\hat{p}$ and $\hat{\mathcal{H}}$ are zero (i.e.
-in the absence of forces) we recognize the geodesic equation.
-We substitute $\hat{p}_\alpha = q A_\alpha$ and $\hat{\mathcal{H}} = 0$ and have:
-
-\begin{align*}
-m \nabla_{u} u^\alpha &= g^{\alpha\beta} (\partial_\beta A_\gamma - \partial_\gamma A_\beta ) q u^\gamma \\
-&= g^{\alpha\beta} F_{\beta \gamma} q u^\gamma\\
-\end{align*}
-
-This is indeed compatible with general relativity and classical electromagnetism.
-
-\begin{prop}\label{newtonianGravitation}
-Assume $\mathcal{M}$ flat and time-independent motion.
-Then the Hamiltonian simplifies to $H=\frac{1}{2m}(p_i-\hat{p}_\alpha(x))^2+\hat{H}(x)$.
-Let $\hat{p}_i = 0$ and $\hat{H} = mV$, the equations of motion are $m d_t v^i = - m \partial_i V$, where $v^i=dx^i/dt$.
-These are equations for a particle under a Newtonian gravitational potential.
-\end{prop}
-
-Given flat space-time and time-independent motion, we can use $t$ as a parameter for the motion in $\mathbbm{Q}$.
-Repeating \ref{kineticMomentum} and \ref{initialConditions} for the non-relativistic case leads to the above Hamiltonian.
-Similar to \ref{relativisticEM}, use \ref{mdofHam} with the newly found Hamiltonian and find the equations of motion.
-
-\begin{defn}\label{lagrangian}
-Let $\mathcal{H}$ be an extended Hamiltonian defined on $\mathbf{T}^*\mathcal{M}$.
-Under the kinematic assumption, we can define the Legendre transform $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$ which we call the \emph{extended Lagrangian}.
-For the time invariant case, we define the \emph{Lagrangian} $L=v^i p_i - H$.
-\end{defn}
-
-The Legendre transform can be defined only if $\mathcal{H}$ is convex in $p_\alpha$.
-The Hamiltonian found at \ref{kineticHamiltonian} is convex, so the Lagrangian can always be defined.
-While the kinematic equivalence assumption is sufficient, it's not necessary for the existence of a Lagrangian.
-What is necessary is \ref{initialConditions}: as long as position and velocity are enough to determine the state, $u^\alpha=f(q^\alpha, p_\alpha)$ must be monotonic in $p_\alpha$, which means $\mathcal{H}$ is convex in $p_\alpha$.
-What happens in that case is that the space-time trajectories may become more or less dense, $du_\alpha$ and $dx^\alpha$ would not be enough to define the label cardinality and therefore the state density $\rho$.
-In short: deterministic and reversible evolution gives us Hamiltonian mechanics, position and velocity as initial conditions gives us Lagrangian mechanics, and the kinematic assumption (which links the differentials of the state variables and the initial conditions) gives us the narrowed Hamiltonian form.
-
-\section{Conclusion}
-
-While there are many more details that we could have expanded upon, as we have touched many areas we could only scratch the surface.
-The intent is to give the overall picture, which could be summarized in the following points.
-
-\begin{itemize}
-  \item Classical states are those that describe every infinitesimal part.
-  \item Classical particles are infinitesimal cells. Phase space is the cotangent bundle $\mathbf{T}^*\mathbbm{Q}$ because that's the space of infinitesimal cells.
-  \item Hamiltonian mechanics coincides with deterministic and reversible evolution.
-  \item Hamiltonian flow is the conservation of number of labels for each independent degree of freedom.
-  \item Deterministic and reversible evolution is what ultimately gives space-time its locally Minkowskian nature.
-  \item $\hbar$, $c$ and $m$ can be seen as conversion constants that preserve cardinality.
-  \item Lagrangian mechanics coincides with position and velocity being necessary and sufficient initial conditions.
-  \item The motion of an isolated elementary system, for which the kinematic assumption is valid, is restricted to a Hamiltonian that can describe fundamental classical forces.
-\end{itemize}
-
-This helps clarify and understand much of the classical framework in a more cohesive way and the notion of cardinality of labels and states is at the heart.
-
-While this is limited to classical particle mechanics, it should be obvious how, at least in principle, this work could be extended.
-For field theories, the \emph{kinematic assumption} should be substituted by a \emph{kinematic field assumption}: what we are studying are not trajectories $x^\alpha(s)$ but fields $\psi(x)$.
-The field values at each point become a set of independent state variables, and their conjugates $\pi(x^\alpha)$ will be linked to $d\partial_s\psi(x^\alpha)$.
-For quantum mechanics, the \emph{infinite reducibility assumption} has to give way to an \emph{irreducibility assumption}: the state of parts of a quantum system cannot be known.
-The configuration state as a whole undergoes deterministic and reversible evolution, while the motion of its parts does not.
-
-The hope is that, by continuing in this approach, we can shed more light on why the laws of physics are what they are; and show that they are not arbitrary rules, but necessary given few simple assumptions.
-
-\section{Acknowledgements}
-
-It is my pleasure to thank Christine A. Aidala, Lydia Bieri, Robert Geroch and Steven J. Miller for helpful discussions and support.
-
 \begin{thebibliography}{0}

### 2015-10-22 2015CompleteDerivation/completeDerivation.tex
Adding vector space definition and justification

@@
 
-labels/state variables, cardinality of state variable, state id function (=1 for a particular state).
+TODO: labels/state variables, cardinality of state variable, state id function (=1 for a particular state).
 Can we do for continuous variables as well?
 Can we find divergence free evolution for det/rev and independent variables?
+
+\section{Composite systems and reducibility}
+
+We know introduce the idea that a system is decomposable, that is it is made of smaller parts, and we characterize the relationship between the whole and the parts with the following assumption:
 
@@
 
-If the scale of the problem
+TODO:
+
+We are now ready to capture the elements of our discussion and their properties through mathematical definitions.
+
+\begin{defn}\label{vector space}
+The state space $\mathfrak{C}$ for a infinitesimally reducible system is a vector space.
+Such vector is free, and admits a basis $\mathfrak{S}$ that represent the state space for the infinitesimal parts.
+\end{defn}
+
+\textbf{Justification}.
+We first claim that the state space $\mathfrak{C}$ is an additive monoid: there exist a law of composition $+ : \mathfrak{C} \times \mathfrak{C} \rightarrow \mathfrak{C}$ that takes two states and returns one that is the physical composition of the two.
+The domain and codomain match because the system is homogeneous.
+The law is commutative $\mathfrak{c}_1 +\mathfrak{c}_2 = \mathfrak{c}_2+\mathfrak{c}_1$ and associative $(\mathfrak{c}_1 + \mathfrak{c}_2) + \mathfrak{c}_3 = \mathfrak{c}_1 + (\mathfrak{c}_2 + \mathfrak{c}_3)$, as it does not matter in what order we compose the parts.
+There exist a zero element $\mathfrak{c} + 0 = \mathfrak{c}$ and it represent the state with no system.
+
+We then claim it is an additive group.
+As we want to describe changes during the evolution, i.e.
+$\Delta \mathfrak{c}$, we introduce an inverse $\mathfrak{c}_2$ for each element $\mathfrak{c}_1$ such that $\mathfrak{c}_1 + \mathfrak{c}_2 = 0$.
+Such inverse may not represent a true physical state, yet we'll still call $\mathfrak{C}$ the state space committing an abuse of terminology.
+Having an inverse, $\mathfrak{C}$ is an additive (i.e.
+abelian) group.
+
+We now consider the set of transformations $A$ that increase or decrease the size of the system by a constant.
+This forms a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics sense (a physical quantity with a value for each point in space).} that is isomorphic to $\mathbbm{R}$.
+In fact, let $a: \mathbbm{R} \rightarrow A$ be a mapping between a number and the transformation that increases or decreases the size of the system by that number; define on $A$ an addition $+: A \times A \rightarrow A$ and a multiplication $*: A \times A \rightarrow A$ such that $a(x) + a(y) = a(x+y)$ and $a(x) * a(y) = a(x*y)$, so that the sum and product of the transformation is equal to the sum and product of their respective factors.
+$A$ is a field, as its sum and multiplication have all properties of $\mathbbm{R}$ which is a field.
+$a$ is an homomorphism by definition and is invertible by the very definition of $A$, each element fully identified by its constant.
+Therefore $A$ and $\mathbbm{R}$ are isomorphic.
+
+We now claim that the state space $\mathfrak{C}$ is a vector space over $\mathbbm{R}$.
+The abelian group $\mathfrak{C}$ can be extended with the operations defined by $A$, as each element $a \in A$ is a map $a : \mathfrak{C} \rightarrow \mathfrak{C}$.
+The map has the following properties: $(a_1 + a_2) \mathfrak{c} = a_1 \mathfrak{c} + a_2 \mathfrak{c}$, increasing the size of the system by the sum of two constant is the same as combining the separate increases, and $a (\mathfrak{c}_1 + \mathfrak{c}_2) = a \mathfrak{c}_1 + a \mathfrak{c}_2$, increasing the size of the total system is the same as the combination of the increased parts.
+$\mathfrak{C}$ is a module over $A$, which is a field and isomorphic to $\mathbbm{R}$.
+We can then think of $\mathfrak{C}$ as a vector space over $\mathbbm{R}$.
 

### 2015-10-22 2015CompleteDerivation/completeDerivation.tex
Creating latex env for rationale/justification/proof

@@
 \newtheorem*{assump3}{Kinematic equivalence assumption}
+
+\newenvironment{rationale}{\textbf{Rationale}.}{$\Box$}
+\newenvironment{justification}{\textbf{Justification}.}{$\Box$}
+\renewenvironment{proof}{\textbf{Proof}.}{$\Box$}
 
@@
 
-\textbf{Rationale}.
+\begin{rationale}
 The claim is that deterministic and reversible evolution is necessary to be able to study a physical system.
@@
 Finding the law after that is a much simpler endeavour, and is something that is routinely done by students in their intro labs.
+\end{rationale}
 
@@
 
-\textbf{Justification}. The collection of all possible states from a set. The deterministic and reversible evolution maps one element in the set to one and only one element of the set.$\Box$
+\begin{justification}
+The collection of all possible states from a set. The deterministic and reversible evolution maps one element in the set to one and only one element of the set.
+\end{justification}
 
@@
 
-\textbf{Justification}.
+\begin{justification}
 A state variable associates each state to one of its values.
@@
 A label range, a set of possible values, can also be thought as representing all states that are identified by those values.
-Abusing the notation, $\mathbbm{I}$ will represent both the set of possible values and the function, and $i$ both the possible value and the set of states.$\Box$
+Abusing the notation, $\mathbbm{I}$ will represent both the set of possible values and the function, and $i$ both the possible value and the set of states.
+\end{justification}
 
@@
 
-\textbf{Rationale}.
+\begin{rationale}
 This makes a clear demarcation of what a classical object is (and conversely what a quantum object is not): it is something that can be fully described up to its infinitesimal constituents.
@@
 While the overall position of a ball can be considered fairly unaffected by its surrounding, the position of each small piece is going to be constantly affected by air and light scattering on its surface.
+\end{rationale}
 
@@
 \begin{defn}\label{vector space}
-The state space $\mathfrak{C}$ for a infinitesimally reducible system is a vector space.
-Such vector is free, and admits a basis $\mathfrak{S}$ that represent the state space for the infinitesimal parts.
+The state space $\mathfrak{C}$ for a infinitesimally reducible system is a vector space. It admits a basis $\mathfrak{S}$ that represents the state space for the infinitesimal parts.
 \end{defn}
 
-\textbf{Justification}.
+\begin{justification}
 We first claim that the state space $\mathfrak{C}$ is an additive monoid: there exist a law of composition $+ : \mathfrak{C} \times \mathfrak{C} \rightarrow \mathfrak{C}$ that takes two states and returns one that is the physical composition of the two.
@@
 We can then think of $\mathfrak{C}$ as a vector space over $\mathbbm{R}$.
+\end{justification}
 

### 2015-10-22 2015CompleteDerivation/completeDerivation.tex
Reviewed counter and styles for assump/defn/prop

@@
 
-\newtheorem{thm}{Theorem}[section]
-\newtheorem{cor}[thm]{Corollary}
-\newtheorem{lem}[thm]{Lemma}
-\newtheorem{prop}[thm]{Proposition}
+\newtheorem{assump}{Assumption}
+\renewcommand*{\theassump}{\Roman{assump}}
+\newtheorem{prop}{Proposition}[section]
 
 \theoremstyle{definition}
-\newtheorem{defn}[thm]{Definition}
-\newtheorem*{assump1}{Infinite reducibility (or classical) assumption}
-\newtheorem*{assump2}{Determinism and Reversibility assumption}
-\newtheorem*{assump3}{Kinematic equivalence assumption}
+\newtheorem{defn}[prop]{Definition}
 
-\newenvironment{rationale}{\textbf{Rationale}.}{$\Box$}
-\newenvironment{justification}{\textbf{Justification}.}{$\Box$}
-\renewenvironment{proof}{\textbf{Proof}.}{$\Box$}
+\newenvironment{rationale}{\emph{Rationale}.}{$\Box$}
+\newenvironment{justification}{\emph{Justification}.}{$\Box$}
+\renewenvironment{proof}{\emph{Proof}.}{$\Box$}
 
@@
 
-\begin{assump2}
+\begin{assump}[Determinism and reversability]
 The state of the physical system under study undergoes deterministic and reversible evolution.
-\end{assump2}
+\end{assump}
 
@@
 
-\begin{assump1}\label{classical}
+\begin{assump}[Infinite reducibility or Classical]\label{classical}
 The system is infinitely reducible: it can be thought of as composed of two or more similar but smaller systems, each in its own configuration state, which in turn can be thought of as composed of two or more, ad infinitum.
 We call \emph{particle} such an infinitesimal part.
-\end{assump1}
+\end{assump}
 

### 2015-10-24 2015CompleteDerivation/completeDerivation.tex
Using qed symbol instead of box
More work on the assumption rationale

@@
 
-\newenvironment{rationale}{\emph{Rationale}.}{$\Box$}
-\newenvironment{justification}{\emph{Justification}.}{$\Box$}
-\renewenvironment{proof}{\emph{Proof}.}{$\Box$}
+\newenvironment{rationale}{\emph{Rationale}.}{\qed}
+\newenvironment{justification}{\emph{Justification}.}{\qed}
+\renewenvironment{proof}{\emph{Proof}.}{\qed}
 
@@
 In fact: it is this aspect of the evolution that determines what states are available and their description.
-Suppose we study the motion of a small particle, its state under gravitational and inertial forces will be properly described by the position and momentum of the center of mass.
-Suppose we study the same small particle as suspended in a fluid, as it undergoes Brownian motion: its state will be a probability distribution for position and momentum of the center of mass.
+Suppose we study the motion of a cannonball, its state under gravitational and inertial forces will be properly described by the position and momentum of the center of mass.
+While light and air molecules scatter off its surface randomly, its trajectory is not greatly affect as it is a massive rigid body.
+Suppose we study the motion of a small particle, small enough that the random scattering does influence the trajectory, and it undergoes Brownian motion: its state will be a probability distribution for position and momentum of the center of mass.
 Gravitational and inertial forces have not changed, yet the set of states have.
@@
 The shape and the parameters of the distribution can be deterministic, even if the individual trajectories are not (those fall within the unstated part).
-In fact: we cannot assume trajectory and states are defined at all for the unstated part.
-Consider a muon and its decay products: it is clear that the three outgoing particles have a state and trajectory of their own, it is clear that the resulting total mass and energy came from the muon.
+In fact: we cannot assume trajectories and states are always defined at all for the unstated part.
+Consider a muon and its decay into an electron and two neutrinos: it is clear that the three outgoing particles have a state and trajectory of their own, it is clear that the resulting total mass and energy came from the muon.
 Yet, before the decay, we cannot ascribe an internal state and trajectory to those parts.
 The state of a muon is not some combination of the state of an electron and two neutrinos.
+That is: the unstated part is not just microstates.
 
@@
 We can manipulate the state of the first system without affecting the second, and vice-versa.
-So, to tell a system apart from its environment, there has to to be a certain amount of isolation between the two: if I affect its state, the state of the other systems is not affected and vice-versa.
+So, to tell a system apart from its environment, there has to to be a certain amount of isolation between the two: if I affect its state, the state of other systems is not affected and vice-versa.
 This means that the system future and past states are only determined by its own state and the state undergoes deterministic and reversible evolution.
@@
 if the system is perfectly isolated, we cannot interact with it: how can it studied through experimentation?).
-It's a simplifying assumption that can only be taken if the environment and the internal dynamics of the system interact in such a way that they affect and are little affected by the aspect we are studying.
+It's a simplifying assumption that can only be taken if the environment and the internal dynamics of the system interact in such a way that they little affect and are little affected by the aspect we are studying.
 As we saw before, for example, assuming that the states consists of the position and momentum of the center of mass means assuming that the Brownian motion of the body is negligible.
@@
 \begin{justification}
-The collection of all possible states from a set. The deterministic and reversible evolution maps one element in the set to one and only one element of the set.
+The collection of all possible states form a set. The deterministic and reversible evolution maps one element in the set to one and only one element of the set.
 \end{justification}
@@
 We call the pair $(\mathbbm{I}, f_\mathbbm{I})$ a \emph{state variable}.
-We call the pair $(i, f_\mathbbm{I}^{-1}(i))$ a \emph{label}, where $i \in \mathbbm{I}$ and $f_\mathbbm{I}^{-1}(i) = \{\mathfrak{s} \in \mathfrak{S} | f_\mathbbm{I}(\mathfrak{s}) = i\}$ is the reverse image of $i$.
+We call the pair $(i, f_\mathbbm{I}^{-1}(i))$ a \emph{label}, where $i \in \mathbbm{I}$ and $f_\mathbbm{I}^{-1}(i) = \{\mathfrak{s} \in \mathfrak{S} | f_\mathbbm{I}(\mathfrak{s}) = i\}$ is the inverse image of $i$.
 We call the pair $(I, f_\mathbbm{I}^{-1}(I))$ a \emph{label range}, where $I \subset \mathbbm{I}$.
@@
 
-We know introduce the idea that a system is decomposable, that is it is made of smaller parts, and we characterize the relationship between the whole and the parts with the following assumption:
+We now introduce the idea that a system is decomposable, that is it is made of smaller parts.
+To simplify our discussion, we will always assume that our system is homogeneous: all parts can be characterized in the same way, that they are under the same external and internal forces, that the set of states they can occupy is the same.
+For a non-homogeneous system, we would divide it in homogeneous components and consider the state of all of them.
+Ultimately, what homogenous means depends on context (e.g.
+air can be thought as homogeneous if the mixture of gasses does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is conserved by time evolution.
+
+We further characterize the relationship between the whole and the parts with the following assumption:
 
 \begin{assump}[Infinite reducibility or Classical]\label{classical}
-The system is infinitely reducible: it can be thought of as composed of two or more similar but smaller systems, each in its own configuration state, which in turn can be thought of as composed of two or more, ad infinitum.
+The system is infinitely reducible: it can be thought of as composed of two or more similar but smaller systems, each in its own state, which in turn can be thought of as composed of two or more, ad infinitum.
 We call \emph{particle} such an infinitesimal part.
+We call \emph{classical} a system for which such an assumption is valid.
 \end{assump}
 
+In other words: if we know the state and the evolution for the whole system, we also know the evolution for all its parts; if we know the state and evolution for all the parts, we know the state and the evolution for the whole system.
+By smaller we do not mean spatial extent, but of smaller amount (e.g.
+less mass): even at the same spatial location we can imagine that there are infinitely many components (e.g.
+the mass, as a number, is continuously divisible(.
+
 \begin{rationale}
-This makes a clear demarcation of what a classical object is (and conversely what a quantum object is not): it is something that can be fully described up to its infinitesimal constituents.
-This is clearly a \emph{simplifying} assumption, and it is instructive to understand when and why it cannot be used.
+This is clearly a \emph{simplifying} assumption, and it is instructive to understand when it breaks down.
 
@@
 As we saw before, we need access to a deterministic and reversible process to be able to study a system and define a state.
-For two balls, we can imagine studying pieces of the balls in isolation, and then describe the collision between the two by describing what happens at each piece.
+For two balls, we can imagine to isolate some pieces (small enough to be considered infinitesimal in respect to the ball but big enough to contain enough molecules to be considered homogenous) and then describe the collision between the two by describing what happens at each piece.
 The classical assumption holds.
@@
 As we saw before, a system cannot be fully isolated from the environment.
-The smaller the part we study, the more it will be susceptible to fluctuations from the environment, the less we can justify the deterministic assumption, the less we are in a position to define its state.
-While the overall position of a ball can be considered fairly unaffected by its surrounding, the position of each small piece is going to be constantly affected by air and light scattering on its surface.
+A certain amount of unstated part needs to remain so that the state of the system can be insulated from the external non-deterministic interaction.
+In practice, the state of each piece of our material is still given by the average of a collection of a large amount of molecules, their respective motion depending not just on the state of the system but on the external environment.
+The classical assumption can therefore hold only if we assume that the description of each pieces is, again, not complete.
+Assuming it at a fundamental level would mean that the motion of each and every molecule of the system is only influenced by the state of the system, no external factor counts.
+As such, it would be a contradiction with it being a physical system: one we can study experimentally.
+
+While ultimately flawed, the classical assumption can be considered valid for a great number of macroscopic system, and that is why is very useful.
+One should be cautioned, though, that nowhere in the assumption the spatial extent of the system is mentioned.
+We may as well have a macroscopic system where a clear independent state cannot be assigned to each part, and the assumption would not hold.
 \end{rationale}
-
-TODO:
 

### 2015-12-10 2015CompleteDerivation/completeDerivation.tex
Dividing decomposability, reducibility and classical evolution

@@
 
+Given that scientific reductionism (i.e.
+the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics, we should explore how to characterize a system in terms of its components.
+In general, this is quite a complicated thing to do, that requires intimate knowledge of the system at hand.
+We will study an idealized limit case: one where the components are homogeneous and infinitesimal.
+What we'll find is that under the assumption that the system is reducible (i.e.
+its state is equivalent to the state of the parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian particle mechanics.
+
+\subsection{Composite systems}
+As mentioned before, we will limit ourselves to the case where our system is homogeneous: all parts can be characterized in the same way, they are under the same external and internal forces, the set of states they can occupy is the same.
+For a non-homogeneous system, we would divide it in homogeneous components and consider the state of all of them.
+Ultimately, what homogenous means depends on context (e.g.
+air can be thought as homogeneous if the mixture of gasses does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is conserved by time evolution.
+
+\begin{defn}\label{reducible_state_space}
+The state space $\mathfrak{C}$ for a homogeneous decomposable system is an additive abelian (i.e. commutative) group.
+\end{defn}
+
+\begin{justification}
+We first claim that the state space $\mathfrak{C}$ is an additive monoid: there exist a law of composition $+ : \mathfrak{C} \times \mathfrak{C} \rightarrow \mathfrak{C}$ that takes two states and returns one that is the physical composition of the two.
+The domain and codomain match because the system is homogeneous.
+The law is commutative $\mathfrak{c}_1 +\mathfrak{c}_2 = \mathfrak{c}_2+\mathfrak{c}_1$ and associative $(\mathfrak{c}_1 + \mathfrak{c}_2) + \mathfrak{c}_3 = \mathfrak{c}_1 + (\mathfrak{c}_2 + \mathfrak{c}_3)$, as it does not matter in what order we compose the parts.
+There exist a zero element $\mathfrak{c} + 0 = \mathfrak{c}$ and it represent the state with no system.
+
+We then claim it is an additive group.
+As we want to describe changes during the evolution, i.e.
+$\Delta \mathfrak{c}$, we introduce an inverse $\mathfrak{c}_2$ for each element $\mathfrak{c}_1$ such that $\mathfrak{c}_1 + \mathfrak{c}_2 = 0$.
+Such inverse may not represent a true physical state, yet we'll still call $\mathfrak{C}$ the state space committing an abuse of terminology.
+Having an inverse, $\mathfrak{C}$ is an additive (i.e.
+abelian) group.
+\end{justification}
+
+\subsection{Classical composite systems}
+We also want to be able to express the state of the composite system in terms of the state of the parts.
+For example, we'll want to relate the state variables of the composite system to the state variables of the parts.
+This in general will depend on how much the state of the composite system "knows" about its part.
+For example, the position and orientation of an ideal rigid body is enough to define where all its constituents are, while the volume/pressure/temperature of an ideal gas is not enough to determine the position and momentum of all its molecules.
+We need to characterize the system further.
+
+We will call a classical system one that is infinitesimally reducible. That is: it is made of arbitrarily small parts, which we call particles, and the state of the whole system is equivalent to the
+
+We also need to be able to characterize the composite state in terms of the
+
+% Infinitesimally decomposable and reducible
+% We have structure for pieces (their state space).
+We can use labels to identify pieces.
+We just need to consider state transformations that do not change the state variable of the pieces and are independent of state variable of the pieces.
+Increase number of pieces is the only thing.
+% State space is vector space over R
+
+% Det/rev for infinitesimal part
+% Inconsistent: 1.
+X=AxB A'=F(A) B'=F(B) Isolated sub-systems.
+Does not work for measuring.
+2.
+A'=F(A) B'=F(A,B) Ideal non-interfering measurement: deterministic but not reversible.
+3.
+A'=F(A,B) B'=F(A,B) X det/rev whole, but A not deterministic itself.
+But if you consider X a sub-system of X and something else, you can make the same argument.
+Whole universe deterministic, but none of its parts.
+4.
+A=As x Au ; B=Bs x Bu ; As' = F(As) Au=F(X) Bu=F(X) Bs = F(As, Au) ; more realistic picture.
+Still non-reversible (if As' = F(As), then nothing else can be F(As) and have bijective map).
+
+% Continuous state variable: Q is measured at constant t
+
 We now introduce the idea that a system is decomposable, that is it is made of smaller parts.
@@
 \begin{justification}
-We first claim that the state space $\mathfrak{C}$ is an additive monoid: there exist a law of composition $+ : \mathfrak{C} \times \mathfrak{C} \rightarrow \mathfrak{C}$ that takes two states and returns one that is the physical composition of the two.
-The domain and codomain match because the system is homogeneous.
-The law is commutative $\mathfrak{c}_1 +\mathfrak{c}_2 = \mathfrak{c}_2+\mathfrak{c}_1$ and associative $(\mathfrak{c}_1 + \mathfrak{c}_2) + \mathfrak{c}_3 = \mathfrak{c}_1 + (\mathfrak{c}_2 + \mathfrak{c}_3)$, as it does not matter in what order we compose the parts.
-There exist a zero element $\mathfrak{c} + 0 = \mathfrak{c}$ and it represent the state with no system.
-
-We then claim it is an additive group.
-As we want to describe changes during the evolution, i.e.
-$\Delta \mathfrak{c}$, we introduce an inverse $\mathfrak{c}_2$ for each element $\mathfrak{c}_1$ such that $\mathfrak{c}_1 + \mathfrak{c}_2 = 0$.
-Such inverse may not represent a true physical state, yet we'll still call $\mathfrak{C}$ the state space committing an abuse of terminology.
-Having an inverse, $\mathfrak{C}$ is an additive (i.e.
-abelian) group.
+Since the system is decomposable, the state space is an abelian group.
 
@@
 \end{justification}
+
+\section{Quantum systems}
+%Note on composite systems and distinguishability.
+When putting together two classical system, just sum distributions.
+When putting together quantum system we have to clarify: combine to create one quantum system (sum in vector space) or two separate quantum system (symmetrized tensor product).
+In classical mechanics, 2a + 2b = (a+b) + (a+b).
+In quantum note |a> x |b> = |a>|b> + |b>|a> but (|a> + |b>) x (|a> + |b>) = |a>|a> + ...
+That is two quantum states in two different states is not the same as two quantum states spread equally over those two states.
+In classical they are.)
 

### 2015-12-11 2015CompleteDerivation/completeDerivation.tex
Justification for classical state space

@@
 
-We will call a classical system one that is infinitesimally reducible. That is: it is made of arbitrarily small parts, which we call particles, and the state of the whole system is equivalent to the
+We will call a classical system one that is infinitesimally reducible.
+That is: it is made of arbitrarily small parts, which we call particles, and the state of the whole system is equivalent to the sum of the states of the parts.
+Given the state space $\mathfrak{S}$ for the parts, the state $\mathfrak{c}$ of the composite system is going to be the sum of all particle states, weighted by the amount of particles that are in that space.
+Since the system is infinitesimally reducible, such amount, no matter the units we choose, will be a real number.
+In other words: the classical state is a real valued distribution over $\mathfrak{S}$.
+The state space $\mathfrak{C}$ of the composite system is a real vector space.
 
-We also need to be able to characterize the composite state in terms of the
+A different way to think about it, one that we can later generalize to irreducible (i.e.
+quantum) systems is the following.
+First we consider all that the composite system can tell us about each part.
+In the classical case, everything: its full state $\mathfrak{s}$.
+This part can be described by an appropriate set of state variables.
+Any transformation that changes such information can be described by how the state variables are affected.
+There remains the set of transformations that leave unchanged what the composite system can tell about each part.
+In the classical case, since the everything about each part is in the state of the composite system, the only transformation left is the increasing or decreasing the amount of particles by a scalar multiple.
+
+\begin{defn}\label{vector space}
+The state space $\mathfrak{C}$ for a classical (i.e.
+infinitesimally reducible) system is a vector space over $\mathbbm{R}$.
+It admits a basis isomorphic to the state space $\mathfrak{S}$ of its particles (i.e.
+infinitesimal parts).
+\end{defn}
+
+\begin{justification}
+Since the system is decomposable, the state space is an abelian group.
+
+We now consider the set of transformations $A$ that increase or decrease the amount of parts of the system by a constant.
+We claim $A$ is a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics sense (a physical quantity with a value for each point in space).} isomorphic to $\mathbbm{R}$.
+Consider $a: \mathbbm{R} \rightarrow A$ the mapping between a number and the transformation that increases or decreases the size of the system by that number; define on $A$ an addition $+: A \times A \rightarrow A$ and a multiplication $*: A \times A \rightarrow A$ such that $a(x) + a(y) = a(x+y)$ and $a(x) * a(y) = a(x*y)$, so that the sum and product of the transformation is equal to the sum and product of their respective factors.
+$a$ is an isomorphism between $A$ and $\mathbbm{R}$.
+$A$ is isomorphic to $\mathbbm{R}$ as a field.
+
+We now claim that the state space $\mathfrak{C}$ is a vector space over $\mathbbm{R}$.
+The abelian group $\mathfrak{C}$ can be extended with the operations defined by $A$, as each element $a \in A$ is a map $a : \mathfrak{C} \rightarrow \mathfrak{C}$.
+The map has the following properties: $(a_1 + a_2) \mathfrak{c} = a_1 \mathfrak{c} + a_2 \mathfrak{c}$, increasing the size of the system by the sum of two constant is the same as combining the separate increases, and $a (\mathfrak{c}_1 + \mathfrak{c}_2) = a \mathfrak{c}_1 + a \mathfrak{c}_2$, increasing the size of the total system is the same as the combination of the increased parts.
+$\mathfrak{C}$ is a module over $A$, which is a field and isomorphic to $\mathbbm{R}$.
+
+Consider now the injection map $e: \mathfrak{S} \hookrightarrow \mathfrak{C}$ that for each particle state returns a composite state constituted only by a fixed amount of particles are in such state.
+The image of such injection is a basis for $\mathfrak{C}$.
+We first claim that all elements are linearly independent.
+Consider $\sum a_i e(\mathfrak{s}_i)$, with $a_i \in \mathbbm{R}$ and $\mathfrak{s}_i \in \mathfrak{S}$.
+This corresponds to a physical state with $a_i$ particles in state $\mathfrak{s}_i$.
+This is equal to the empty state if and only if $a_i=0 \forall i$.
+This is true even if $\mathfrak{S}$ is infinite dimensional.
+We now claim they span.
+Consider $\mathfrak{c} + \sum a_i e(\mathfrak{s}_i)$.
+This can always be made equal to the empty state by setting $a_i$ to be the opposite of the amount of particles of $\mathfrak{c}$ that are in state $\mathfrak{s}_i$.
+\end{justification}
 
@@
 
-\begin{defn}\label{vector space}
-The state space $\mathfrak{C}$ for a infinitesimally reducible system is a vector space. It admits a basis $\mathfrak{S}$ that represents the state space for the infinitesimal parts.
-\end{defn}
-
-\begin{justification}
-Since the system is decomposable, the state space is an abelian group.
-
-We now consider the set of transformations $A$ that increase or decrease the size of the system by a constant.
-This forms a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics sense (a physical quantity with a value for each point in space).} that is isomorphic to $\mathbbm{R}$.
-In fact, let $a: \mathbbm{R} \rightarrow A$ be a mapping between a number and the transformation that increases or decreases the size of the system by that number; define on $A$ an addition $+: A \times A \rightarrow A$ and a multiplication $*: A \times A \rightarrow A$ such that $a(x) + a(y) = a(x+y)$ and $a(x) * a(y) = a(x*y)$, so that the sum and product of the transformation is equal to the sum and product of their respective factors.
-$A$ is a field, as its sum and multiplication have all properties of $\mathbbm{R}$ which is a field.
-$a$ is an homomorphism by definition and is invertible by the very definition of $A$, each element fully identified by its constant.
-Therefore $A$ and $\mathbbm{R}$ are isomorphic.
-
-We now claim that the state space $\mathfrak{C}$ is a vector space over $\mathbbm{R}$.
-The abelian group $\mathfrak{C}$ can be extended with the operations defined by $A$, as each element $a \in A$ is a map $a : \mathfrak{C} \rightarrow \mathfrak{C}$.
-The map has the following properties: $(a_1 + a_2) \mathfrak{c} = a_1 \mathfrak{c} + a_2 \mathfrak{c}$, increasing the size of the system by the sum of two constant is the same as combining the separate increases, and $a (\mathfrak{c}_1 + \mathfrak{c}_2) = a \mathfrak{c}_1 + a \mathfrak{c}_2$, increasing the size of the total system is the same as the combination of the increased parts.
-$\mathfrak{C}$ is a module over $A$, which is a field and isomorphic to $\mathbbm{R}$.
-We can then think of $\mathfrak{C}$ as a vector space over $\mathbbm{R}$.
-\end{justification}
-
 \section{Quantum systems}

### 2015-12-14 2015CompleteDerivation/completeDerivation.tex
Refined composite/classical state definitions

--- 2015-12-10 2015CompleteDerivation/completeDerivation.tex
@@
 In general, this is quite a complicated thing to do, that requires intimate knowledge of the system at hand.
-We will study an idealized limit case: one where the components are homogeneous and infinitesimal.
-What we'll find is that under the assumption that the system is reducible (i.e.
+So we simplify our problem and study an idealized case: one where the components are homogeneous and infinitesimal.
+What we'll find is that under the additional assumption that the system is reducible (i.e.
 its state is equivalent to the state of the parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian particle mechanics.
@@
 \subsection{Composite systems}
-As mentioned before, we will limit ourselves to the case where our system is homogeneous: all parts can be characterized in the same way, they are under the same external and internal forces, the set of states they can occupy is the same.
+As mentioned before, we will limit ourselves to the case where our system is homogeneous: the parts are indistinguishable, the set of states they can occupy is the same, they are under the same external and internal forces.
 For a non-homogeneous system, we would divide it in homogeneous components and consider the state of all of them.
-Ultimately, what homogenous means depends on context (e.g.
+What homogenous means depends on context (e.g.
 air can be thought as homogeneous if the mixture of gasses does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is conserved by time evolution.
+Nonetheless, we can claim the following:
 
 \begin{defn}\label{reducible_state_space}
-The state space $\mathfrak{C}$ for a homogeneous decomposable system is an additive abelian (i.e. commutative) group.
+The state space $\mathbbm{C}$ for a homogeneous decomposable system is an additive abelian (i.e. commutative) group.
 \end{defn}
@@
 \begin{justification}
-We first claim that the state space $\mathfrak{C}$ is an additive monoid: there exist a law of composition $+ : \mathfrak{C} \times \mathfrak{C} \rightarrow \mathfrak{C}$ that takes two states and returns one that is the physical composition of the two.
+We claim $\mathbbm{C}$ is an additive monoid.
+There exist a law of composition $+ : \mathbbm{C} \times \mathbbm{C} \rightarrow \mathbbm{C}$ that takes two states and returns one that is the physical composition of the two.
 The domain and codomain match because the system is homogeneous.
-The law is commutative $\mathfrak{c}_1 +\mathfrak{c}_2 = \mathfrak{c}_2+\mathfrak{c}_1$ and associative $(\mathfrak{c}_1 + \mathfrak{c}_2) + \mathfrak{c}_3 = \mathfrak{c}_1 + (\mathfrak{c}_2 + \mathfrak{c}_3)$, as it does not matter in what order we compose the parts.
-There exist a zero element $\mathfrak{c} + 0 = \mathfrak{c}$ and it represent the state with no system.
+The law is commutative $\mathbbm{c}_1 +\mathbbm{c}_2 = \mathbbm{c}_2+\mathbbm{c}_1$ and associative $(\mathbbm{c}_1 + \mathbbm{c}_2) + \mathbbm{c}_3 = \mathbbm{c}_1 + (\mathbbm{c}_2 + \mathbbm{c}_3)$, as it does not matter in what order we physically compose the parts.
+There exist a unique zero element $\mathbbm{c} + 0 = \mathbbm{c}$ and it represent the physically empty state (i.e.
+no system).
 
-We then claim it is an additive group.
-As we want to describe changes during the evolution, i.e.
-$\Delta \mathfrak{c}$, we introduce an inverse $\mathfrak{c}_2$ for each element $\mathfrak{c}_1$ such that $\mathfrak{c}_1 + \mathfrak{c}_2 = 0$.
-Such inverse may not represent a true physical state, yet we'll still call $\mathfrak{C}$ the state space committing an abuse of terminology.
-Having an inverse, $\mathfrak{C}$ is an additive (i.e.
-abelian) group.
+We claim $\mathbbm{C}$ is an additive group.
+As we want to describe changes during the evolution (i.e.
+$\hat{\mathbbm{c}} = \mathbbm{c} +\Delta \mathbbm{c}$) we introduce an inverse $\mathbbm{c}_2$ for each element $\mathbbm{c}_1$ such that $\mathbbm{c}_1 + \mathbbm{c}_2 = 0$.
+The inverse may not represent a proper physical state, yet we'll still call $\mathbbm{C}$ the state space committing an abuse of terminology.
+The inverse $- : \mathbbm{C} \rightarrow \mathbbm{C}$ is physically unique and well defined.
 \end{justification}
@@
 For example, we'll want to relate the state variables of the composite system to the state variables of the parts.
-This in general will depend on how much the state of the composite system "knows" about its part.
+This in general will depend on how much the state of the composite system "knows" about its parts.
 For example, the position and orientation of an ideal rigid body is enough to define where all its constituents are, while the volume/pressure/temperature of an ideal gas is not enough to determine the position and momentum of all its molecules.
-We need to characterize the system further.
+Therefore we need to characterize the system further.
 
-We will call a classical system one that is infinitesimally reducible. That is: it is made of arbitrarily small parts, which we call particles, and the state of the whole system is equivalent to the
+We will call a classical system one that is infinitesimally reducible.
+That is: it is made of arbitrarily small parts, which we call particles, and the state of the whole system is equivalent to the sum of the states of the parts.
+Given the state space $\mathbbm{S}$ for the parts, giving a composite state $\mathbbm{c}$ is equivalent to give the amount of particles that are present in each state.
+Since the system is infinitesimally reducible, such amount can be increased or decreased infinitesimally and will be a real number.
+In other words: the classical state is a real valued distribution over $\mathbbm{S}$.
+This intuitive picture is unfortunately not suitable to be generalized to different contexts, so we develop another.
 
-We also need to be able to characterize the composite state in terms of the
+Consider the following expression:
+\begin{align*}
+\mathbbm{c} = \sum\limits_{i \in \mathbbm{I}} a_i \mathbbm{e}(i)
+\end{align*}
+$\mathbbm{I}$ is best description of an infinitesimal part that a composite system can give.
+$\mathbbm{e}(i)$ represents a composite system made of a unit amount of particles, all prepared in the state determined by $i$.
+$a_i$ represents a transformation that changes the whole state, without affecting the value $i$ for any infinitesimal part.
+For a classical system, $i$ represents the full information about the infinitesimal part, and therefore includes the value of all state variables; the only transformation left is increasing/decreasing the amount of particles by a scalar multiple.
+For a different system, where $i$ does not represent the full state of the parts, $\mathbbm{e}(i)$ will be an ensemble and $a_i$ will include internal transformations.
+This is intuitively how we decouple the composite state into the part $i$ that describes the infinitesimal parts, and the part $a_i$ that describes the rest.
+Giving a full description of the state space of a decomposable system means fully characterizing these components.
+And we can see how this fits for the classical case as well: the state variables gives us the description for an infinitesimal part, but the amount of particles for each state is a description of the composite state.
 
-% Infinitesimally decomposable and reducible
-% We have structure for pieces (their state space).
-We can use labels to identify pieces.
-We just need to consider state transformations that do not change the state variable of the pieces and are independent of state variable of the pieces.
-Increase number of pieces is the only thing.
-% State space is vector space over R
+\begin{defn}\label{classical_vector space}
+The state space $\mathbbm{C}$ for a homogeneous classical (i.e.
+infinitesimally reducible) system is a vector space over $\mathbbm{R}$.
+It admits a basis isomorphic to the state space $\mathbbm{S}$ of its particles (i.e.
+infinitesimal parts).
+\end{defn}
+
+\begin{justification}
+The state space is an abelian group as the system is homogeneous and decomposable and \ref{reducible_state_space}.
+
+We now consider the set of transformations $A$ that increase or decrease the amount of parts of the system by a constant.
+We claim $A$ is a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics sense (a physical quantity with a value for each point in space).} isomorphic to $\mathbbm{R}$.
+Consider $a: \mathbbm{R} \rightarrow A$ the mapping between a number and the transformation that increases or decreases the size of the system by that number.
+This transformation exists: the system is infinitesimally decomposable and the amount can be changed continuously.
+Define on $A$ an addition $+: A \times A \rightarrow A$ and a multiplication $*: A \times A \rightarrow A$ such that $a(x) + a(y) = a(x+y)$ and $a(x) * a(y) = a(x*y)$, so that the sum and product of the transformation is equal to the sum and product of their respective factors.
+$a$ is an isomorphism between $A$ and $\mathbbm{R}$.
+$A$ is isomorphic to $\mathbbm{R}$ as a field.
+
+We now claim that the state space $\mathbbm{C}$ is a vector space over $\mathbbm{R}$.
+The abelian group $\mathbbm{C}$ can be extended with the operations defined by $A$, as each element $a \in A$ is a map $a : \mathbbm{C} \rightarrow \mathbbm{C}$.
+The map has the following properties: $(a_1 + a_2) \mathbbm{c} = a_1 \mathbbm{c} + a_2 \mathbbm{c}$, increasing the size of the system by the sum of two constant is the same as combining the separate increases, and $a (\mathbbm{c}_1 + \mathbbm{c}_2) = a \mathbbm{c}_1 + a \mathbbm{c}_2$, increasing the size of the total system is the same as the combination of the increased parts.
+$\mathbbm{C}$ is a module over $A$, which is a field and isomorphic to $\mathbbm{R}$.
+
+Consider now the injection map $\mathbbm{e}: \mathbbm{S} \hookrightarrow \mathbbm{C}$ that for each particle state returns a composite state constituted only by a predetermined amount of particles in such state.
+The image of such injection is a basis for $\mathbbm{C}$.
+We first claim that all elements are linearly independent.
+Consider $\sum\limits_{\mathbbm{s}_i \in \mathbbm{S}} a_i \mathbbm{e}(\mathbbm{s}_i)$, with $a_i \in \mathbbm{R}$ and .
+This corresponds to a physical state with $a_i$ particles in each state $\mathbbm{s}_i$.
+This is equal to the empty state if and only if $a_i=0 \forall i$.
+This holds even if $\mathbbm{S}$ is not finite.
+We now claim they span.
+Consider $\mathbbm{c} + \sum\limits_{\mathbbm{s}_i \in \mathbbm{S}} a_i \mathbbm{e}(\mathbbm{s}_i)$, $a_i \in \mathbbm{R}$.
+This can always be made equal to the empty state by setting $a_i$ to be the opposite of the amount of particles of $\mathbbm{c}$ that are in state $\mathbbm{s}_i$.
+\end{justification}
+
+\subsection{Classical assumption}
 
@@
 % Continuous state variable: Q is measured at constant t
-
-We now introduce the idea that a system is decomposable, that is it is made of smaller parts.
-To simplify our discussion, we will always assume that our system is homogeneous: all parts can be characterized in the same way, that they are under the same external and internal forces, that the set of states they can occupy is the same.
-For a non-homogeneous system, we would divide it in homogeneous components and consider the state of all of them.
-Ultimately, what homogenous means depends on context (e.g.
-air can be thought as homogeneous if the mixture of gasses does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is conserved by time evolution.
 

### 2015-12-14 2015CompleteDerivation/completeDerivation.tex
Using mathcal for state spaces

@@
 \usepackage{amsthm}
+\usepackage{dutchcal}
 
@@
 \begin{defn}\label{reducible_state_space}
-The state space $\mathbbm{C}$ for a homogeneous decomposable system is an additive abelian (i.e. commutative) group.
+The state space $\mathcal{C}$ for a homogeneous decomposable system is an additive abelian (i.e. commutative) group.
 \end{defn}
@@
 \begin{justification}
-We claim $\mathbbm{C}$ is an additive monoid.
-There exist a law of composition $+ : \mathbbm{C} \times \mathbbm{C} \rightarrow \mathbbm{C}$ that takes two states and returns one that is the physical composition of the two.
+We claim $\mathcal{C}$ is an additive monoid.
+There exist a law of composition $+ : \mathcal{C} \times \mathcal{C} \rightarrow \mathcal{C}$ that takes two states and returns one that is the physical composition of the two.
 The domain and codomain match because the system is homogeneous.
-The law is commutative $\mathbbm{c}_1 +\mathbbm{c}_2 = \mathbbm{c}_2+\mathbbm{c}_1$ and associative $(\mathbbm{c}_1 + \mathbbm{c}_2) + \mathbbm{c}_3 = \mathbbm{c}_1 + (\mathbbm{c}_2 + \mathbbm{c}_3)$, as it does not matter in what order we physically compose the parts.
-There exist a unique zero element $\mathbbm{c} + 0 = \mathbbm{c}$ and it represent the physically empty state (i.e.
+The law is commutative $\mathcal{c}_1 +\mathcal{c}_2 = \mathcal{c}_2+\mathcal{c}_1$ and associative $(\mathcal{c}_1 + \mathcal{c}_2) + \mathcal{c}_3 = \mathcal{c}_1 + (\mathcal{c}_2 + \mathcal{c}_3)$, as it does not matter in what order we physically compose the parts.
+There exist a unique zero element $\mathcal{c} + 0 = \mathcal{c}$ and it represent the physically empty state (i.e.
 no system).
 
-We claim $\mathbbm{C}$ is an additive group.
+We claim $\mathcal{C}$ is an additive group.
 As we want to describe changes during the evolution (i.e.
-$\hat{\mathbbm{c}} = \mathbbm{c} +\Delta \mathbbm{c}$) we introduce an inverse $\mathbbm{c}_2$ for each element $\mathbbm{c}_1$ such that $\mathbbm{c}_1 + \mathbbm{c}_2 = 0$.
-The inverse may not represent a proper physical state, yet we'll still call $\mathbbm{C}$ the state space committing an abuse of terminology.
-The inverse $- : \mathbbm{C} \rightarrow \mathbbm{C}$ is physically unique and well defined.
+$\hat{\mathcal{c}} = \mathcal{c} +\Delta \mathcal{c}$) we introduce an inverse $\mathcal{c}_2$ for each element $\mathcal{c}_1$ such that $\mathcal{c}_1 + \mathcal{c}_2 = 0$.
+The inverse may not represent a proper physical state, yet we'll still call $\mathcal{C}$ the state space committing an abuse of terminology.
+The inverse $- : \mathcal{C} \rightarrow \mathcal{C}$ is physically unique and well defined.
 \end{justification}
@@
 That is: it is made of arbitrarily small parts, which we call particles, and the state of the whole system is equivalent to the sum of the states of the parts.
-Given the state space $\mathbbm{S}$ for the parts, giving a composite state $\mathbbm{c}$ is equivalent to give the amount of particles that are present in each state.
+Given the state space $\mathcal{S}$ for the parts, giving a composite state $\mathcal{c}$ is equivalent to give the amount of particles that are present in each state.
 Since the system is infinitesimally reducible, such amount can be increased or decreased infinitesimally and will be a real number.
-In other words: the classical state is a real valued distribution over $\mathbbm{S}$.
+In other words: the classical state is a real valued distribution over $\mathcal{S}$.
 This intuitive picture is unfortunately not suitable to be generalized to different contexts, so we develop another.
@@
 \begin{align*}
-\mathbbm{c} = \sum\limits_{i \in \mathbbm{I}} a_i \mathbbm{e}(i)
+\mathcal{c} = \sum\limits_{i \in \mathbbm{I}} a_i \mathcal{e}(i)
 \end{align*}
 $\mathbbm{I}$ is best description of an infinitesimal part that a composite system can give.
-$\mathbbm{e}(i)$ represents a composite system made of a unit amount of particles, all prepared in the state determined by $i$.
+$\mathcal{e}(i)$ represents a composite system made of a unit amount of particles, all prepared in the state determined by $i$.
 $a_i$ represents a transformation that changes the whole state, without affecting the value $i$ for any infinitesimal part.
 For a classical system, $i$ represents the full information about the infinitesimal part, and therefore includes the value of all state variables; the only transformation left is increasing/decreasing the amount of particles by a scalar multiple.
-For a different system, where $i$ does not represent the full state of the parts, $\mathbbm{e}(i)$ will be an ensemble and $a_i$ will include internal transformations.
+For a different system, where $i$ does not represent the full state of the parts, $\mathcal{e}(i)$ will be an ensemble and $a_i$ will include internal transformations.
 This is intuitively how we decouple the composite state into the part $i$ that describes the infinitesimal parts, and the part $a_i$ that describes the rest.
@@
 \begin{defn}\label{classical_vector space}
-The state space $\mathbbm{C}$ for a homogeneous classical (i.e.
+The state space $\mathcal{C}$ for a homogeneous classical (i.e.
 infinitesimally reducible) system is a vector space over $\mathbbm{R}$.
-It admits a basis isomorphic to the state space $\mathbbm{S}$ of its particles (i.e.
+It admits a basis isomorphic to the state space $\mathcal{S}$ of its particles (i.e.
 infinitesimal parts).
@@
 
-We now claim that the state space $\mathbbm{C}$ is a vector space over $\mathbbm{R}$.
-The abelian group $\mathbbm{C}$ can be extended with the operations defined by $A$, as each element $a \in A$ is a map $a : \mathbbm{C} \rightarrow \mathbbm{C}$.
-The map has the following properties: $(a_1 + a_2) \mathbbm{c} = a_1 \mathbbm{c} + a_2 \mathbbm{c}$, increasing the size of the system by the sum of two constant is the same as combining the separate increases, and $a (\mathbbm{c}_1 + \mathbbm{c}_2) = a \mathbbm{c}_1 + a \mathbbm{c}_2$, increasing the size of the total system is the same as the combination of the increased parts.
-$\mathbbm{C}$ is a module over $A$, which is a field and isomorphic to $\mathbbm{R}$.
+We now claim that the state space $\mathcal{C}$ is a vector space over $\mathbbm{R}$.
+The abelian group $\mathcal{C}$ can be extended with the operations defined by $A$, as each element $a \in A$ is a map $a : \mathcal{C} \rightarrow \mathcal{C}$.
+The map has the following properties: $(a_1 + a_2) \mathcal{c} = a_1 \mathcal{c} + a_2 \mathcal{c}$, increasing the size of the system by the sum of two constant is the same as combining the separate increases, and $a (\mathcal{c}_1 + \mathcal{c}_2) = a \mathcal{c}_1 + a \mathcal{c}_2$, increasing the size of the total system is the same as the combination of the increased parts.
+$\mathcal{C}$ is a module over $A$, which is a field and isomorphic to $\mathbbm{R}$.
 
-Consider now the injection map $\mathbbm{e}: \mathbbm{S} \hookrightarrow \mathbbm{C}$ that for each particle state returns a composite state constituted only by a predetermined amount of particles in such state.
-The image of such injection is a basis for $\mathbbm{C}$.
+Consider now the injection map $\mathcal{e}: \mathcal{S} \hookrightarrow \mathcal{C}$ that for each particle state returns a composite state constituted only by a predetermined amount of particles in such state.
+The image of such injection is a basis for $\mathcal{C}$.
 We first claim that all elements are linearly independent.
-Consider $\sum\limits_{\mathbbm{s}_i \in \mathbbm{S}} a_i \mathbbm{e}(\mathbbm{s}_i)$, with $a_i \in \mathbbm{R}$ and .
-This corresponds to a physical state with $a_i$ particles in each state $\mathbbm{s}_i$.
+Consider $\sum\limits_{\mathcal{s}_i \in \mathcal{S}} a_i \mathcal{e}(\mathcal{s}_i)$, with $a_i \in \mathbbm{R}$ and .
+This corresponds to a physical state with $a_i$ particles in each state $\mathcal{s}_i$.
 This is equal to the empty state if and only if $a_i=0 \forall i$.
-This holds even if $\mathbbm{S}$ is not finite.
+This holds even if $\mathcal{S}$ is not finite.
 We now claim they span.
-Consider $\mathbbm{c} + \sum\limits_{\mathbbm{s}_i \in \mathbbm{S}} a_i \mathbbm{e}(\mathbbm{s}_i)$, $a_i \in \mathbbm{R}$.
-This can always be made equal to the empty state by setting $a_i$ to be the opposite of the amount of particles of $\mathbbm{c}$ that are in state $\mathbbm{s}_i$.
+Consider $\mathcal{c} + \sum\limits_{\mathcal{s}_i \in \mathcal{S}} a_i \mathcal{e}(\mathcal{s}_i)$, $a_i \in \mathbbm{R}$.
+This can always be made equal to the empty state by setting $a_i$ to be the opposite of the amount of particles of $\mathcal{c}$ that are in state $\mathcal{s}_i$.
 \end{justification}

### 2015-12-17 2015CompleteDerivation/completeDerivation.tex
Better state variable definitions. Other minor corrections.

@@
 \begin{defn}\label{statedef}
-Fix a physical system for study.
-There exists a set $\mathfrak{S}$, which we call \emph{state space}, consisting of all the physically distinguishable configurations $\mathfrak{s}$, which we call \emph{states}.
-There exists a bijective map $f_{\Delta t}:\mathfrak{S} \leftrightarrow \mathfrak{S}$, which we call \emph{time evolution operator}, that links initial and final states.
+The state space $\mathcal{S}$ of a physical system is a set.
 \end{defn}
@@
 \begin{justification}
-The collection of all possible states form a set. The deterministic and reversible evolution maps one element in the set to one and only one element of the set.
+The collection of all possible states forms a set.
 \end{justification}
 
-In practice, we use physical quantities to identify states.
+\begin{defn}\label{detrevmap}
+A deterministic and reversible evolution is a bijective map $f_{\Delta t}:\mathcal{S} \leftrightarrow \mathcal{S}$ on the state space.
+\end{defn}
+
+\begin{justification}
+The system is deterministic: there exists a map $f:\mathcal{S} \rightarrow \mathcal{S}$ that returns the final state given the initial state.
+The system is reversible: there exists a map $g:\mathcal{S} \rightarrow \mathcal{S}$ that returns the initial state given the final state.
+$g \circ f = id$ as mapping forward and then backward must return the original element.
+\end{justification}
+
+%In practice, we use physical quantities to identify states.
 For example, the state for an ideal gas is given by pressure, volume and temperature.
 To clarify whether we are referring to a particular value of such quantity, or the quantity itself, we introduce the following terminology:
-\begin{center}
-    \begin{tabular}{ | p{2.5cm} | p{5.5cm} | }
-    \hline
-    State variable & A quantity (discrete or continuous) that must be specified to identify a state (e.g. position). \\ \hline
-    Label & A particular value for a state variable (e.g. position = $5m$). \\ \hline
-    Label range & A set of possible values for a state variable (e.g. position = $[4.5m, 5.5m]$). \\
-    \hline
-    \end{tabular}
-\end{center}
+%\begin{center}
+%    \begin{tabular}{ | p{2.5cm} | p{5.5cm} | }
+%    \hline
+%    State variable & A quantity (discrete or continuous) that must be specified to identify a state (e.g. position). \\ \hline
+%    Label & A particular value for a state variable (e.g. position = $5m$). \\ \hline
+%    Label range & A set of possible values for a state variable (e.g. position = $[4.5m, 5.5m]$). \\
+%    \hline
+%    \end{tabular}
+%\end{center}
 
@@
 Let $\mathbbm{I}$ be a set.
-Let $f_\mathbbm{I} : \mathfrak{S} \rightarrow \mathbbm{I}$ be a function that maps each state to an element of the set.
-We call the pair $(\mathbbm{I}, f_\mathbbm{I})$ a \emph{state variable}.
-We call the pair $(i, f_\mathbbm{I}^{-1}(i))$ a \emph{label}, where $i \in \mathbbm{I}$ and $f_\mathbbm{I}^{-1}(i) = \{\mathfrak{s} \in \mathfrak{S} | f_\mathbbm{I}(\mathfrak{s}) = i\}$ is the inverse image of $i$.
-We call the pair $(I, f_\mathbbm{I}^{-1}(I))$ a \emph{label range}, where $I \subset \mathbbm{I}$.
+We call $q : \mathcal{S} \rightarrow \mathbbm{I}$ a \emph{state variable} and a \emph{possible value} (or \emph{possibility}) of $q$ an element $i \in \mathbbm{I}$.
+Also $\mathcal{s}_1 \neq \mathcal{s}_2 \Leftrightarrow \exists q : \mathcal{S} \rightarrow \mathbbm{I} | q(\mathcal{s}_1)\neq q(\mathcal{s}_2) \forall \mathcal{s}_1, \mathcal{s}_2 \in \mathcal{S}$.
 \end{defn}
@@
 \begin{justification}
-A state variable associates each state to one of its values.
-A label, one of the possible values, can also be thought as representing all states that are identified by its value.
-A label range, a set of possible values, can also be thought as representing all states that are identified by those values.
-Abusing the notation, $\mathbbm{I}$ will represent both the set of possible values and the function, and $i$ both the possible value and the set of states.
+Consider a physical process that can distinguish between different initial conditions.
+Let $\mathbbm{I}$ be the set of all possible outcomes.
+To each possible state $\mathcal{s}$ will correspond one and only one outcome.
+Let $q : \mathcal{S} \rightarrow \mathbbm{I}$ be the map between state and outcome.
+Such object is physically well defined.
+
+Let $\mathcal{s}_1, \mathcal{s}_2 \in \mathcal{S}$.
+Suppose $\mathcal{s}_1 \neq \mathcal{s}_2$.
+$\mathcal{s}_1$ and $\mathcal{s}_2$ are physically distinguishable.
+There exists a physical process with different outcomes.
+Let $q : \mathcal{S} \rightarrow \mathbbm{I}$ be the state variable associated to that process.
+Then $q(\mathcal{s}_1)\neq q(\mathcal{s}_2)$.
+Now suppose $\nexists q : \mathcal{S} \rightarrow \mathbbm{I} | q(\mathcal{s}_1) \neq q(\mathcal{s}_2)$.
+Then no process can distinguish between the two, they are physically indistinguishable therefore $\mathcal{s}_1 = \mathcal{s}_2$.
 \end{justification}
@@
 air can be thought as homogeneous if the mixture of gasses does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is conserved by time evolution.
-Nonetheless, we can claim the following:
+Nonetheless, we can generally claim the following:
 
@@
 \end{align*}
-$\mathbbm{I}$ is best description of an infinitesimal part that a composite system can give.
+$\mathbbm{I}$ is a state variable representing the best description of an infinitesimal part that a composite system can give.
 $\mathcal{e}(i)$ represents a composite system made of a unit amount of particles, all prepared in the state determined by $i$.
@@
 For a classical system, $i$ represents the full information about the infinitesimal part, and therefore includes the value of all state variables; the only transformation left is increasing/decreasing the amount of particles by a scalar multiple.
-For a different system, where $i$ does not represent the full state of the parts, $\mathcal{e}(i)$ will be an ensemble and $a_i$ will include internal transformations.
+For a different system, where $i$ does not represent the full state of the parts, $\mathcal{e}(i)$ will be an ensemble and $a_i$ also represents an internal transformation of such ensable.
 This is intuitively how we decouple the composite state into the part $i$ that describes the infinitesimal parts, and the part $a_i$ that describes the rest.
-Giving a full description of the state space of a decomposable system means fully characterizing these components.
+Giving a full description of the state space of a decomposable system means fully characterizing these components: the state variable $\mathbbm{I}$ and the set of transformations $A$.
 And we can see how this fits for the classical case as well: the state variables gives us the description for an infinitesimal part, but the amount of particles for each state is a description of the composite state.
@@
 This transformation exists: the system is infinitesimally decomposable and the amount can be changed continuously.
-Define on $A$ an addition $+: A \times A \rightarrow A$ and a multiplication $*: A \times A \rightarrow A$ such that $a(x) + a(y) = a(x+y)$ and $a(x) * a(y) = a(x*y)$, so that the sum and product of the transformation is equal to the sum and product of their respective factors.
-$a$ is an isomorphism between $A$ and $\mathbbm{R}$.
-$A$ is isomorphic to $\mathbbm{R}$ as a field.
+Define on $A$ an addition $+: A \times A \rightarrow A$ and a multiplication $*: A \times A \rightarrow A$ such that $a(x) + a(y) = a(x+y)$ and $a(x) * a(y) = a(x*y)$, $x,y \in \mathbbm{R}$, so that the sum and product of the transformation is equal to the sum and product of their respective factors.
+$a$ is an isomorphism between $A$ and $\mathbbm{R}$ as fields.
 
@@
 We first claim that all elements are linearly independent.
-Consider $\sum\limits_{\mathcal{s}_i \in \mathcal{S}} a_i \mathcal{e}(\mathcal{s}_i)$, with $a_i \in \mathbbm{R}$ and .
+Consider $\sum\limits_{\mathcal{s}_i \in \mathcal{S}} a_i \mathcal{e}(\mathcal{s}_i)$, $a_i \in \mathbbm{R}$.
 This corresponds to a physical state with $a_i$ particles in each state $\mathcal{s}_i$.

### 2015-12-18 2015CompleteDerivation/completeDerivation.tex
Adding more on state variable. Other improvements.

@@
 
+\section{Style objective}
+
+The writing style should have the following goals:
+\begin{description}
+  \item[Skip math] one should be able to skip the mathematical definitions/propositions and still get a coherent narrative.
+  \item[Well-defined physical objects] each mathematical objects need to be physically well-defined. It must shown to exist and be unique.
+\end{description}
+Non-goals:
+\begin{description}
+  \item[Can't skip physics in justifications] justifications of mathematical definitions will inherently be rooted in the preceding physics discussion.
+\end{description}
+Notation:
+\begin{description}
+  \item[state variable and space] state space (cont/discr), state variable (cont/discr), possibility (cont/discr), range of possibility (cont/discr), set of possibilities (cont/discr)
+\end{description}
+
 \section{Determinism and reversibility}
@@
 
-\begin{assump}[Determinism and reversability]
+\begin{assump}[Determinism and reversibility]
 The state of the physical system under study undergoes deterministic and reversible evolution.
@@
 
-\begin{defn}\label{label}
+\begin{defn}\label{state_variable}
 Let $\mathbbm{I}$ be a set.
@@
 
+\begin{defn}\label{independent_state_variables}
+Let $q_1 : \mathcal{S} \rightarrow \mathbbm{I}_1$ and $q_2 : \mathcal{S} \rightarrow \mathbbm{I}_2$ be two state variables.
+They are said \emph{independent} if $q_1^{-1}(i_1)\cap q_2^{-1}(i_2) \neq 0 \forall i_1 \in \mathbbm{I}_1, i_2 \in \mathbbm{I}_2$.
+\end{defn}
+
+\begin{justification}
+If $q_1$ are $q_2$ independent state variables we can modify a state such that only one variable is affected. TODO finish
+\end{justification}
+
+\begin{prop}\label{discrete_state_space}
+The space state $\mathcal{S}$ for a system fully identified by a set of $n$ independent discrete state variables $q^i : \mathcal{S} \rightarrow \mathbbm{I}^i$ is isomorphic to $\prod\limits_{i=1}^n\mathbbm{I}^i$.
+Let $I^i \subseteq \mathbbm{I}^i$ be finite sets of possibilities for each variable.
+$\#(\bigcap\limits_{i=1}^{n}{q^i}^{-1}(I^i))=\prod\limits_{i=1}^{n}\#(I^i)$.
+\end{prop}
+
+\begin{proof}
+TODO
+\end{proof}
+
+\begin{defn}\label{continuous_state_space}
+The space state $\mathcal{Q}$ for a system fully identified by a set of $n$ independent continuous state variables is a manifold of dimension $n$.
+\end{defn}
+
+\begin{justification}
+TODO: set of state variables locally define a chart; inherit topology from charts (small change in state makes small change in state variable, makes small changes in outcomes?)
+\end{justification}
+
 TODO: labels/state variables, cardinality of state variable, state id function (=1 for a particular state).
@@
 
-\subsection{Composite systems}
-As mentioned before, we will limit ourselves to the case where our system is homogeneous: the parts are indistinguishable, the set of states they can occupy is the same, they are under the same external and internal forces.
-For a non-homogeneous system, we would divide it in homogeneous components and consider the state of all of them.
-What homogenous means depends on context (e.g.
+\subsection{Homogeneous systems}
+As mentioned before, we will limit ourselves to the case where our system is homogeneous: the parts are indistinguishable, the set of states they can occupy is the same, they are under the same external and internal forces.\footnote{Non-homogeneous systems can in principle be decomposed in homogeneous components.
+Therefore their study can be reduced to the study of homogeneous systems.} What homogenous means depends on context (e.g.
 air can be thought as homogeneous if the mixture of gasses does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is conserved by time evolution.
-Nonetheless, we can generally claim the following:
+Nonetheless, the state space of such system is richer in the sense that it comes equipped with the notion of composition and its rules.
+These are suitable captured mathematically by an additive abelian group.
 
@@
 As we want to describe changes during the evolution (i.e.
-$\hat{\mathcal{c}} = \mathcal{c} +\Delta \mathcal{c}$) we introduce an inverse $\mathcal{c}_2$ for each element $\mathcal{c}_1$ such that $\mathcal{c}_1 + \mathcal{c}_2 = 0$.
-The inverse may not represent a proper physical state, yet we'll still call $\mathcal{C}$ the state space committing an abuse of terminology.
-The inverse $- : \mathcal{C} \rightarrow \mathcal{C}$ is physically unique and well defined.
+$\hat{\mathcal{c}} = \mathcal{c} +\Delta \mathcal{c}$) we introduce an inverse $- : \mathcal{C} \rightarrow \mathcal{C}$ such that $\mathcal{c} + ( - \mathcal{c}) = 0 \forall \mathcal{c} \in \mathcal{C}$.
+Such inverse may introduce objects that do not properly represent a physical state, but a state change.
+A change of a physically distinguishable object is itself physically distinguishable therefore is physically well defined.
+We'll still call $\mathcal{C}$ the state space committing an abuse of terminology.
 \end{justification}
 
-\subsection{Classical composite systems}
+\subsection{Classical homogeneous systems}
 We also want to be able to express the state of the composite system in terms of the state of the parts.
-For example, we'll want to relate the state variables of the composite system to the state variables of the parts.
+For example, given the state of a fluid we'll want to know how its parts are distributed.
 This in general will depend on how much the state of the composite system "knows" about its parts.
@@
 That is: it is made of arbitrarily small parts, which we call particles, and the state of the whole system is equivalent to the sum of the states of the parts.
-Given the state space $\mathcal{S}$ for the parts, giving a composite state $\mathcal{c}$ is equivalent to give the amount of particles that are present in each state.
+Given the state space $\mathcal{S}$ for the parts, identifying a composite state $\mathcal{c}$ is equivalent to quantify the amount of particles that are present in each state.
 Since the system is infinitesimally reducible, such amount can be increased or decreased infinitesimally and will be a real number.
@@
 \end{align*}
-$\mathbbm{I}$ is a state variable representing the best description of an infinitesimal part that a composite system can give.
+$\mathbbm{I}$ represents all the possible values of a state variable representing the best description of an infinitesimal part that a composite system can give.
 $\mathcal{e}(i)$ represents a composite system made of a unit amount of particles, all prepared in the state determined by $i$.
@@
 For a different system, where $i$ does not represent the full state of the parts, $\mathcal{e}(i)$ will be an ensemble and $a_i$ also represents an internal transformation of such ensable.
-This is intuitively how we decouple the composite state into the part $i$ that describes the infinitesimal parts, and the part $a_i$ that describes the rest.
+This is intuitively how we decouple the composite state into the part $i$ that describes the infinitesimal parts, and the part $a_i$ that describes how the parts are composed.
 Giving a full description of the state space of a decomposable system means fully characterizing these components: the state variable $\mathbbm{I}$ and the set of transformations $A$.
-And we can see how this fits for the classical case as well: the state variables gives us the description for an infinitesimal part, but the amount of particles for each state is a description of the composite state.
+In the classical case $\mathcal{e}(i)$ form a basis of a real vector space, where $i$ is the set of state variable to identify the state of an an infinitesimal part, and $a_i$ a real valued coefficient that represents the amount of particles in each state, which describes how the parts are combined.
 
@@
 The map has the following properties: $(a_1 + a_2) \mathcal{c} = a_1 \mathcal{c} + a_2 \mathcal{c}$, increasing the size of the system by the sum of two constant is the same as combining the separate increases, and $a (\mathcal{c}_1 + \mathcal{c}_2) = a \mathcal{c}_1 + a \mathcal{c}_2$, increasing the size of the total system is the same as the combination of the increased parts.
-$\mathcal{C}$ is a module over $A$, which is a field and isomorphic to $\mathbbm{R}$.
+$\mathcal{C}$ is a vector space over $A$, which is a field and isomorphic to $\mathbbm{R}$.
 
@@
 
-\subsection{Classical assumption}
+\subsection{Classical hamiltonian assumption}
 
@@
 
-\begin{assump}[Infinite reducibility or Classical]\label{classical}
-The system is infinitely reducible: it can be thought of as composed of two or more similar but smaller systems, each in its own state, which in turn can be thought of as composed of two or more, ad infinitum.
-We call \emph{particle} such an infinitesimal part.
-We call \emph{classical} a system for which such an assumption is valid.
+\begin{assump}[Classical hamiltonian assumption]\label{classical}
+The evolution of the classical homogeneous system under study is deterministic, reversible and reducible.
 \end{assump}

### 2016-01-13 2015CompleteDerivation/completeDerivation.tex
Introduction and detrev reviewed

@@
 
-Classical particle mechanics is usually founded on Newton's laws.
-These, though, are insufficient to derive the full Lagrangian and Hamiltonian formalism, and usually other ad-hoc assumptions (e.g.
-conservative forces) are introduced.
-Special relativity is based on two principles (invariance of the speed of light and the principle of relativity), which lead to the Minkowskian nature of space-time but not to the equations of motion, which are a consistent reformulation of the non-relativistic ones.
-Quantum mechanics is simply founded on its mathematical formulation, leaving its physical significance open to debate.
-One can't come to the conclusion that, unfortunately, the landscape of fundamental physics is a piece-meal of concepts and idea, sometime apparently inconsistent with each other.
-Yet: the physical world is one, and our physical theories are simply models of such reality.
-If we clarify what those models represent, and on what assumptions are based, shouldn't we be able to have a coherent and more satisfying picture?
+While some branches of fundamental physics (e.g.
+Newtonian mechanics and special relativity) are founded on physical laws or principles others (e.g.
+Lagrangian/Hamiltonian/Quantum mechanics) start by postulating mathematical frameworks or relationships.
+The latter approach presents a significant drawback: the mathematical structure is not enough to define the physical system it describes.
+We are left asking: under what physical conditions a particular system is described by Lagrangian (or Quantum) mechanics?
+Naturally one \emph{can} define a Lagrangian system (or a Quantum system) as one that obeys those rules, but that begs the question.
+Another issue is that each element of the mathematical structure may or may not correspond to some physical concept.
+Suppose the earning (or losses) of two companies are governed by Hamiltonian's equations, which variable is on the cotangent space of the other?
+In other words: if we want to fully appreciate what the fundamental theories of physics are describing, what each mathematical concept corresponds to in the physical world, we need to start by characterizing the physical system we are describing in a precise enough way that it can be encoded in mathematical language.
+Naturally, the physical insight we gain will be relevant for only that type of system (i.e.
+other different systems may \emph{happen} to use the same equations), but what we'll see is that the assumptions we make are minimal and can apply to a variety of systems, especially the ones that are considered fundamental.
+It is indeed surprising how so much can be derived by imposing so little.
 
-This work aims to re-organize the known elements and equations in a more consistent and comprehensive way, leading to better insight on why the fundamental concepts and laws are what they are.
+TODO: high level view of the derivation
+
+%This work aims to re-organize the known elements and equations in a more consistent and comprehensive way, leading to better insight on why the fundamental concepts and laws are what they are.
 We will start with physical assumptions, that clarify the models we impose on the physical world, and give arguments on when those assumptions can be considered valid.
@@
 We will strive to do this in a way that is mathematically meaningful, philosophically consistent and mathematically precise.
-
-TODO: high level view of the derivation
 
@@
 
+As the terms used have different meaning in different contexts, let's first clarify them.
 By deterministic (and reversible) we mean that future (and past) states are uniquely identified by the present state.
@@
 While light and air molecules scatter off its surface randomly, its trajectory is not greatly affect as it is a massive rigid body.
-Suppose we study the motion of a small particle, small enough that the random scattering does influence the trajectory, and it undergoes Brownian motion: its state will be a probability distribution for position and momentum of the center of mass.
+Suppose we study the motion of a small particle, small enough that the random scattering does influence the trajectory and it undergoes Brownian motion: its state will be a probability distribution for position and momentum of the center of mass.
 Gravitational and inertial forces have not changed, yet the set of states have.
@@
 That is: given the distribution at one time we can determine the distribution at future times.
-The shape and the parameters of the distribution can be deterministic, even if the individual trajectories are not (those fall within the unstated part).
-In fact: we cannot assume trajectories and states are always defined at all for the unstated part.
+The shape and the parameters of the distribution can be deterministic, even if the trajectories of the parts are not (those fall within the unstated part).
+In fact: we \emph{cannot} assume trajectories and states are always defined at all for the unstated part.
 Consider a muon and its decay into an electron and two neutrinos: it is clear that the three outgoing particles have a state and trajectory of their own, it is clear that the resulting total mass and energy came from the muon.
-Yet, before the decay, we cannot ascribe an internal state and trajectory to those parts.
+Yet, before the decay, we cannot ascribe an internal independent state and evolution to each part.
 The state of a muon is not some combination of the state of an electron and two neutrinos.
@@
 This link between state definition and deterministic processes should not be too surprising as the state, in the context of thermodynamics and system theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system}.
-As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can be indeed calculated, measured and prepared, not each individual element.
+As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can be indeed calculated, measured and prepared.
+We can also describe the evolution of each element of the distribution provided that: we have a way to isolate it and study it under deterministic and reversible motion (so that we can define its states); the non-deterministic motion does not alter the system (the set of states remains the same).
+For example, consider a wooden ball described by position and momentum of the center of mass: if randomly scattered by other balls retain the set of states, if thrown into a fire it does not.
 
@@
 It's a simplifying assumption that can only be taken if the environment and the internal dynamics of the system interact in such a way that they little affect and are little affected by the aspect we are studying.
-As we saw before, for example, assuming that the states consists of the position and momentum of the center of mass means assuming that the Brownian motion of the body is negligible.
+As we saw before, for example, assuming that the states consists of the position and momentum of the center of mass requires assuming that the Brownian motion of the body is negligible.
 
@@
 \begin{justification}
-The collection of all possible states forms a set.
+The collection of all possible states forms a set as we can physically distinguish between them.
 \end{justification}

### 2016-01-14 2015CompleteDerivation/completeDerivation.tex
Adding case with no env. Intro to state var

@@
 We start by fixing a \emph{physical system} we want to study, that is something we can interact with and perform measurements on (e.g.
-a cat, a planet, water).
+a planet, water).
 We call \emph{environment} everything else.
 We set what particular aspect we want to study (e.g.
-the trajectory when falling, the motion around a star, its flow in a pipe).
-We call \emph{state} a particular configuration in time of the portion under study.
+the motion around a star, the flow in a pipe).
+We call \emph{state} a physically distinguishable configuration of the portion under study at a particular time.
 Since the state does not, in general, exhaust the description of the system, a part remains \emph{unstated}, and as such we'll call it, for lack of a better word.
-With this in mind, we introduce the following:
+Note that the environment plays an essential role here as it's what allows us to define two states as physically distinguishable: we can find an external process (i.e.
+a different environment) whose outcome changes depending on the different state.
+As such processes are what we can use to perform measurements, we consider the experimental apparatus (and us performing measurements) independent of the system, part of the environment.
+This is true even in the case of general relativity, where we can imagine multiple researchers on small spaceships collecting data without greatly influencing the motion of stars and planets.
+We call this triad of state, unstated part and environment the \emph{fundamental model of physics}, as it is the simplest model that still captures the crucial aspects of studying and writing laws for a physical system.\footnote{The case where the physical system is the whole universe and there is no environment presents practical problems (e.g.
+What physical device can we use to store and process the state of the whole universe to make predictions and compare?) and conceptual challenges (e.g.
+How do we define physically distinguishable?
+Can the physical laws predict which measurements are we going to make?
+Does that limit what is actually distinguishable?) that the current physical theories do not seem to be equipped to address, and therefore will be absent from our discussion.} With this in mind, we introduce the following:
 
@@
 We can then think of the state of the system as the part that interacts deterministically and reversibly with the environment (and itself) while the unstated part as interacting non-deterministically with the environment (and itself).
-We call this setup the \emph{fundamental model of physics}, as it is the simplest model that still captures the crucial aspects of studying and writing laws for a physical system.
 
@@
 
+To identify and name states, we will typically use a set of quantities, typically numbers.
+For example, the orbital of an electron in a hydrogen item is identified by the quantum numbers $n$, $l$, $m$ and $s$.
+We call each of these quantities \emph{state variables} and \emph{state vector} a set of state variables that can identify one and only one state.
+We avoid the term coordinates, even though that's what they are mathematically, as it will create confusion with spatial coordinates.
+We also avoid the term observable or measurable, as not all state variables map directly to the outcome of an experiment (e.g.
+conjugate momentum as it is not gauge invariant).
+It's clear, though, that even in those cases there must exist, at least in principle, a way to distinguish between the possible values of the state variable (e.g once we fix the vector potential, conjugate momentum can be measured via the kinetic momentum.
+Therefore we can assume that all state variables are in the end physically distinguishable, yet we do not assume all state variables have equal standing in physical meaning.
+Their purpose is just to label states, and the chosen name is a good reminder.
+
 \begin{defn}\label{state_variable}
@@
 \end{justification}
+
+\begin{prop}\label{state_vector}
+TODO: State vector always exists
+\end{prop}
+
+When assembling multiple state variable into a state vector, it is important to understand how they relate to each other.
+Consider the orbital of an electron in a hydrogen item, which is identified by the quantum numbers $n$, $l$, $m$ and $s$.
+For each combination of $n$, $l$ and $m$, the spin $s$ can have to value.
+The choice of spin is independent from the rest.
+The choice of $l$, though, depends on the choice of $n$: for $n=1$ only $l=s$ is available; for $n=2$ we can choose $l=s$ or $l=p$.
+The choices are not independent.
+That is: two or more variables are independent if there always exist a state for any possible combination.
 

### 2016-01-14 2015CompleteDerivation/completeDerivation.tex
More on state variables

@@
 That is, our system must participate in a deterministic and reversible process with the preparing and measuring device.
-How else could we calibrate our experimental apparatus?
+Without it we wouldn't be able to calibrate our experimental apparatus.
 
-This link between state definition and deterministic processes should not be too surprising as the state, in the context of thermodynamics and system theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system}.
+This link between state definition and deterministic processes should not be too surprising as the state, in the context of thermodynamics and systems theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system}.
 As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can be indeed calculated, measured and prepared.
@@
 For example, the orbital of an electron in a hydrogen item is identified by the quantum numbers $n$, $l$, $m$ and $s$.
-We call each of these quantities \emph{state variables} and \emph{state vector} a set of state variables that can identify one and only one state.
-We avoid the term coordinates, even though that's what they are mathematically, as it will create confusion with spatial coordinates.
+We call each of these quantities \emph{state variable} and \emph{state vector} a set of state variables that can identify one and only one state.
+We avoid the term coordinates, even though that's what they mathematically are, as it will create confusion with space coordinates.
 We also avoid the term observable or measurable, as not all state variables map directly to the outcome of an experiment (e.g.
 conjugate momentum as it is not gauge invariant).
-It's clear, though, that even in those cases there must exist, at least in principle, a way to distinguish between the possible values of the state variable (e.g once we fix the vector potential, conjugate momentum can be measured via the kinetic momentum.
+It's clear, though, that even in those cases there must exist, at least in principle, a way to distinguish between the possible values of the state variable (e.g once we fix the vector potential, conjugate momentum can be measured via the kinetic momentum).
 Therefore we can assume that all state variables are in the end physically distinguishable, yet we do not assume all state variables have equal standing in physical meaning.
-Their purpose is just to label states, and the chosen name is a good reminder.
+Conversely, some measurable quantities are not directly state variables: they may not be defined, as states are, at constant time (e.g.
+velocity) or may depend on the process at hand (e.g.
+acceleration).
+Yet, once a process is fully defined (e.g.
+a Hamiltonian is given), we may have a relationship between those observables and state variables.
+The purpose of state variables is just to label states, which are physically distinguishable and defined at a particular time.
+Therefore the state variable will be physically distinguishable and defined at constant time.
+But there is nothing telling us that they must be directly physically meaningful or that their relationship to measurable quantities is the same if we change physical processes.
 
@@
 
+To keep track of the state evolution, we'll need to understand how state variables change under continuous and reversible evolution.
+The first approach is to study the change in the state variables values from $q^i(\mathcal{s})$ to $q^i(\hat{\mathcal{s}})$.
+This is especially useful when state variables correspond to physically meaningful quantities.
+For example, we track how the temperature or the pressure of an ideal gas has changed.
+However, what we'll want to do is to find relationships that are common for all deterministic and reversible processes, which is hard to do in this picture.
+The second approach is to evolve the state variables as well.
+The idea is to keep the connection to the initial state by labelling the future state, instead by the future values of $q^i$, by the original one.
+So we introduce a new set of variables for which $\hat{q}^i(\hat{\mathcal{s}})=q^i(\mathcal{s})$.
+We can always do this because we have bijective correspondence between initial and final states.
+This makes it easier to study relationships that are invariant during the evolution: once we verify them for $q^i$ on the initial states, they will automatically hold for the $\hat{q}^i$ at later (or previous) times.
+Naturally, as $\hat{q}^i$ are new variables that will not hold the same physical meaning.
+For example, each will be a combination of pressure and temperature that depends on the particular process.
+We'll typically use the first picture to describe equations of motions and the second to describe invariants.
+Note, though, that the two picture are tightly linked.
+In fact:
+\begin{align*}
+\Delta q^i &= q^i(\hat{\mathcal{s}}) - q^i(\mathcal{s}) \\
+\hat{q}^i(\hat{\mathcal{s}}) &= q^i(\hat{\mathcal{s}}) - \Delta q^i
+\end{align*}
+
+\begin{prop}\label{evolved variable}
+TODO: Define and prove existence
+\end{prop}
+
 When assembling multiple state variable into a state vector, it is important to understand how they relate to each other.
@@
 The choices are not independent.
-That is: two or more variables are independent if there always exist a state for any possible combination.
+That is: two or more variables are independent if there always exist a state for any possible combination, if the total number of states is the cartesian product of the possibilities of each variable.
+For discrete space variables,
 

### 2016-01-20 2015CompleteDerivation/completeDerivation.tex
More on state variable evolution

@@
 
-To keep track of the state evolution, we'll need to understand how state variables change under continuous and reversible evolution.
-The first approach is to study the change in the state variables values from $q^i(\mathcal{s})$ to $q^i(\hat{\mathcal{s}})$.
-This is especially useful when state variables correspond to physically meaningful quantities.
-For example, we track how the temperature or the pressure of an ideal gas has changed.
-However, what we'll want to do is to find relationships that are common for all deterministic and reversible processes, which is hard to do in this picture.
-The second approach is to evolve the state variables as well.
+To study time evolution, we need to describe how state variables change under continuous and reversible evolution.
+There are two ways to do it, and we'll need both.
+The first approach is to \emph{evolve} the state variables from the initial value $q^i(\mathcal{s})$ to the final value $q^i(\hat{\mathcal{s}})$.
+The result is a trajectory $q^i(t)$ which is especially useful when state variables correspond to physically meaningful quantities.
+For example, we track how the temperature or the pressure of an ideal gas changes.
+Evolved state variables, however, make it hard to find relationships that are invariant and common to all deterministic and reversible processes.
+
+The second approach is to transport the state variables.
 The idea is to keep the connection to the initial state by labelling the future state, instead by the future values of $q^i$, by the original one.
-So we introduce a new set of variables for which $\hat{q}^i(\hat{\mathcal{s}})=q^i(\mathcal{s})$.
-We can always do this because we have bijective correspondence between initial and final states.
-This makes it easier to study relationships that are invariant during the evolution: once we verify them for $q^i$ on the initial states, they will automatically hold for the $\hat{q}^i$ at later (or previous) times.
-Naturally, as $\hat{q}^i$ are new variables that will not hold the same physical meaning.
-For example, each will be a combination of pressure and temperature that depends on the particular process.
-We'll typically use the first picture to describe equations of motions and the second to describe invariants.
-Note, though, that the two picture are tightly linked.
+So we introduce a new set of variables for which $\hat{q}^i(\hat{\mathcal{s}})=q^i(\mathcal{s})$, which we can always do as the evolution is deterministic and reversible.
+The level sets of $\hat{q}^i$ are the evolved level sets of $q^i$, which allows us to study how groups of states evolve in time.
+Moreover, once we identify an invariant for $q^i$ this is automatically satisfied by $\hat{q}^i$.
+Given that the value of the transported state variables does not change during evolution, and that it is unique for each initial state, transported state variable provide a way to label the trajectories themselves.
+While the numeric value of the transported state variables remains the same, the state variable (and its physical meaning) changes.
+For example, pressure and temperature will be combined depending on the particular process.
+
+Evolved state variables are useful to write equations of motions, study how physical quantities change, form a physical picture of what happens.
+Transported state variables are useful to write invariants, study state space trajectories, from a geometric picture for the state space.
+The two, though, are tightly linked.
 In fact:
@@
 \end{align*}
+Intuitively, the transported variables change in the oppose way as the evolution, so that the value of the variables remain the same.\footnote{One should not confuse evolved/transported state variable with active/passive transformations or with Schroedinger/Heisenberg pictures.
+In those cases, the choice is whether to change the state or the coordinates/observables.
+In our case, the state is always changing.
+The choice is whether to change the state variables or their value.}
 
@@
 
-When assembling multiple state variable into a state vector, it is important to understand how they relate to each other.
+\begin{prop}\label{transported variable}
+TODO: Define and prove existence
+\end{prop}
+
+When assembling multiple state variables into a state vector, it is important to understand how they relate to each other.
 Consider the orbital of an electron in a hydrogen item, which is identified by the quantum numbers $n$, $l$, $m$ and $s$.
-For each combination of $n$, $l$ and $m$, the spin $s$ can have to value.
+For each combination of $n$, $l$ and $m$, the spin $s$ can have two values.
 The choice of spin is independent from the rest.
@@
 The choices are not independent.
-That is: two or more variables are independent if there always exist a state for any possible combination, if the total number of states is the cartesian product of the possibilities of each variable.
-For discrete space variables,
+That is: two or more variables are independent if there always exists a state for any possible combination, if the total number of states is the cartesian product of the possibilities of each variable.
+
+For discrete state variables, the total number of states is the product of the possibilities of each independent state variable.
+That is: if $\Delta q^1$ and $Delta q^2$ are range of possibilities for two independent variables, the total number of possibilities $\#(\Delta q^1) \#(\Delta q^2)$.
+These relationships are maintained during evolution by the transported variables.
+Since the value remains unchanged, the range of possibilities for each value remain the same: $\#(\Delta \hat{q}^i) = \#(\Delta q^i)$.
+The variables remain independent as well, all possible combination remain, and we have: $\#(\Delta \hat{q}^1) \#(\Delta \hat{q}^2) = \#(\Delta q^1) \#(\Delta q^2)$.
+These relationship will later be extended in the continuous case to independent degrees of freedom for a Hamiltonian system, and will justify the symplectic form.
+Though this can't be done in general and requires further assumptions.
+
+In the case of continuous state, if each state can be identified by n independent state variables then, at least locally, we can chart the state space on $\mathcal{R}$: the state space is a manifold.
+Studying in general how ranges of possibilities evolve, though, becomes problematic.
+Time evolution will map points, while range of continuous possibilities will be length.
+We would need a metric, invariant during the evolution, telling us how ranges of possibilities map from one variable to another, possibly describing different physical quantities.
+We will not study this case, but we do note that, if such a metric exist, it depends on the system and on the choice of state variables.
 
@@
 
-TODO: labels/state variables, cardinality of state variable, state id function (=1 for a particular state).
-Can we do for continuous variables as well?
-Can we find divergence free evolution for det/rev and independent variables?
+We have seen that physical systems undergoing deterministic and reversible evolution are properly described by a dynamical system.
+Systems described by a discrete set of states have nice properties for counting possibilities of independent variables, which are maintained during the evolution.
+System described by a continuous set of states, instead, do not posses in general such properties.
 

### 2016-01-25 2015CompleteDerivation/completeDerivation.tex
Minor changes for decomposable systems

@@
 $a_i$ represents a transformation that changes the whole state, without affecting the value $i$ for any infinitesimal part.
-For a classical system, $i$ represents the full information about the infinitesimal part, and therefore includes the value of all state variables; the only transformation left is increasing/decreasing the amount of particles by a scalar multiple.
-For a different system, where $i$ does not represent the full state of the parts, $\mathcal{e}(i)$ will be an ensemble and $a_i$ also represents an internal transformation of such ensable.
+For a classical system, $i$ represents the full information about the infinitesimal part, and therefore includes the value of all state variables (i.e.
+the state vector); the only transformation left is increasing/decreasing the amount of particles by a scalar multiple.
+For a different system, where $i$ does not represent the full state of the parts, $\mathcal{e}(i)$ will be an ensemble and $a_i$ also represents an internal transformation of such ensemble.
 This is intuitively how we decouple the composite state into the part $i$ that describes the infinitesimal parts, and the part $a_i$ that describes how the parts are composed.
 Giving a full description of the state space of a decomposable system means fully characterizing these components: the state variable $\mathbbm{I}$ and the set of transformations $A$.
-In the classical case $\mathcal{e}(i)$ form a basis of a real vector space, where $i$ is the set of state variable to identify the state of an an infinitesimal part, and $a_i$ a real valued coefficient that represents the amount of particles in each state, which describes how the parts are combined.
+In the classical case $\mathcal{e}(i)$ form a basis of a real vector space, where $i$ is the state vector for an infinitesimal part, and $a_i$ a real valued coefficient that represents the amount of particles in each state, which describes how the parts are combined.
 
@@
 The map has the following properties: $(a_1 + a_2) \mathcal{c} = a_1 \mathcal{c} + a_2 \mathcal{c}$, increasing the size of the system by the sum of two constant is the same as combining the separate increases, and $a (\mathcal{c}_1 + \mathcal{c}_2) = a \mathcal{c}_1 + a \mathcal{c}_2$, increasing the size of the total system is the same as the combination of the increased parts.
-$\mathcal{C}$ is a vector space over $A$, which is a field and isomorphic to $\mathbbm{R}$.
+$\mathcal{C}$ is a module over $A$, which is a field and isomorphic to $\mathbbm{R}$.
 

### 2016-02-04 2015CompleteDerivation/completeDerivation.tex
Rewrite of the fundamental model

@@
 
-We start by fixing a \emph{physical system} we want to study, that is something we can interact with and perform measurements on (e.g.
-a planet, water).
+We start by fixing a \emph{physical system}, meaning something we can interact with and perform measurements on (e.g.
+a planet, a fluid, ...).
 We call \emph{environment} everything else.
 We set what particular aspect we want to study (e.g.
-the motion around a star, the flow in a pipe).
-We call \emph{state} a physically distinguishable configuration of the portion under study at a particular time.
-Since the state does not, in general, exhaust the description of the system, a part remains \emph{unstated}, and as such we'll call it, for lack of a better word.
+the motion around a star, the flow in a pipe, ...).
+We call \emph{state} a physically distinguishable configuration of the aspect under study at a particular time (e.g.
+position/momentum of center of mass, velocity field, ...).
+Since the state does not, in general, exhaust the description of the system, a part remains \emph{unstated}, and as such we'll call it, for lack of a better word (e.g.
+the chemical composition, the motion of each of its molecules, ...).
 Note that the environment plays an essential role here as it's what allows us to define two states as physically distinguishable: we can find an external process (i.e.
-a different environment) whose outcome changes depending on the different state.
-As such processes are what we can use to perform measurements, we consider the experimental apparatus (and us performing measurements) independent of the system, part of the environment.
-This is true even in the case of general relativity, where we can imagine multiple researchers on small spaceships collecting data without greatly influencing the motion of stars and planets.
-We call this triad of state, unstated part and environment the \emph{fundamental model of physics}, as it is the simplest model that still captures the crucial aspects of studying and writing laws for a physical system.\footnote{The case where the physical system is the whole universe and there is no environment presents practical problems (e.g.
-What physical device can we use to store and process the state of the whole universe to make predictions and compare?) and conceptual challenges (e.g.
+part of the environment) whose outcome changes depending on the different state.
+As such processes are what we can use to perform measurements, we consider the experimental apparatus (and us performing measurements) independent of the system, part of the environment.\footnote{This is true even in the case of general relativity, where we can imagine multiple researchers on small spaceships collecting data without greatly influencing the motion of stars and planets.
+The case where the physical system is the whole universe and there is no environment presents practical problems  and conceptual challenges that the current physical theories do not seem to be equipped to address, and therefore will be absent from our discussion.
+For example, what physical device can we use to store and process the state of the whole universe to make predictions and compare?
 How do we define physically distinguishable?
-Can the physical laws predict which measurements are we going to make?
-Does that limit what is actually distinguishable?) that the current physical theories do not seem to be equipped to address, and therefore will be absent from our discussion.} With this in mind, we introduce the following:
+Do the physical laws determine which measurements we are going to make and does that limit what is actually distinguishable?}
 
-\begin{assump}[Determinism and reversibility]
+In this context, we define a process \emph{deterministic} if the state at a given time uniquely identifies states at future times, and \emph{reversible} if it identifies states at past times.
+This means that, in general, the evolution of the unstated part is non-deterministic and/or non-reversible in the sense that the state of the system is not enough to determine its evolution.
+For example, the overall motion of a fluid in a pipe can be considered deterministic even if the motion of each molecule is not.
+The motion is deterministic and/or reversible only as far as the state is concerned.
+With this in mind, we introduce the following:
+
+\begin{assump}[Determinism and reversibility]\label{detrevass}
 The state of the physical system under study undergoes deterministic and reversible evolution.
@@
 
-As the terms used have different meaning in different contexts, let's first clarify them.
-By deterministic (and reversible) we mean that future (and past) states are uniquely identified by the present state.
-We can then think of the state of the system as the part that interacts deterministically and reversibly with the environment (and itself) while the unstated part as interacting non-deterministically with the environment (and itself).
+\begin{rationale}
+As it is an assumption, we first need to discuss when it is valid.
+More specifically, we need to understand that the non-deterministic/non-reversible evolution of the unstated part plays as much as a fundamental role as the deterministic/reversible evolution of the state.
+In fact: the non-deterministic part contributes in determining what states are available to the system.
 
-Much of the focus later will be given to the state under deterministic and reversible evolution.
-However, the non-deterministic part plays a fundamental role as well.
-In fact: it is this aspect of the evolution that determines what states are available and their description.
 Suppose we study the motion of a cannonball, its state under gravitational and inertial forces will be properly described by the position and momentum of the center of mass.
@@
 Gravitational and inertial forces have not changed, yet the set of states have.
-The issue is that the set of states must be closed under the non-deterministic evolution as well.
+The issue is that the non-deterministic evolution must not change the set of states itself.
 That is: if we apply the non-deterministic evolution to any state, we still need to end up in one of the possible states.
-If the Brownian motion is strong enough, the state gets knocked out of the set of states determined by a simple position/momentum pair.
+If the Brownian motion is strong enough, a simple position/momentum pair is not sufficient.
 
@@
 
-As we have seen, sometimes the state is identified by a distribution (either statistical or actual).
+As we have hinted, sometimes the state is identified by a distribution (either statistical or actual).
 Even in this case, the state can be deterministic and reversible.
@@
 
-Now that we have clarified what we mean by deterministic and reversible evolution (even on a statistical ensemble), and how also the non-deterministic part is key to define states (as the set as a whole needs to be invariant under it), we can ask ourselves: when is this assumption valid?
+It should be also clear that what constitute state and unstated part does not depend only on the system under study, but also on the processes we are considering.
+In some circumstances, the chemical composition of a fluid may be relevant, in others it may not.
+The choices of environment, state and unstated part are not independent from each other.
+By choosing a particular set of states, we are not only saying that the state evolution is well approximated by a deterministic/reversible map from initial to final state, but we are also saying that the non-deterministic/non-reversible evolution of the unstated part does not change the nature of the system, and processes that do not satisfy these conditions are not under consideration.
 
-\begin{rationale}
-The claim is that deterministic and reversible evolution is necessary to be able to study a physical system.
-That is: the only part of a system that can be studied through reliable and reproducible experimentation or observation is the part that undergoes deterministic and reversible evolution.
+As with all assumptions, we should also ask whether it is necessary.
+That is: could we define a set of physically distinguishable states and yet have no deterministic and reversible processes defined on them?
+The claim is that this assumption is indeed needed, as without it we cannot properly define states or write useful physics laws.
 We can provide different arguments that point in the same direction.
 
-First, to be able to identify the system, we must first be able to tell it apart from the environment.
+First, to be able to identify the system, we must be able to tell it apart from anything else.
 Intuitively, we can distinguish between two chairs because we can move the first to another room and sit on it without having touched the second.
 We can manipulate the state of the first system without affecting the second, and vice-versa.
-So, to tell a system apart from its environment, there has to to be a certain amount of isolation between the two: if I affect its state, the state of other systems is not affected and vice-versa.
-This means that the system future and past states are only determined by its own state and the state undergoes deterministic and reversible evolution.
+So, to identify a system it has to be sufficiently isolated from everything else.
+This means that the system future and past states are with good approximation determined only by its own state: the state undergoes deterministic and reversible evolution.
 
-Second, the aim of physics is to write laws that can be used to make prediction that can be validated experimentally.
-If I drop an anvil from a tower, it will accelerate at $9.81 m/s^2$; if I want the anvil to reach the ground at $x m/s$ I have to drop it from $y m$.
+Second, the aim of physics is to write laws that can be used to make predictions that can be validated experimentally.
+If I drop an anvil from a tower, it will accelerate at $9.81$~m/s$^2$; if I want the anvil to reach the ground at $x$~m/s I have to drop it from $y$~m.
 To the extent that we want to make predictions in time, we need to have a one-to-one correspondence between initial and final states.
@@
 
+Fourth, to be able to ascribe a property to a system we need to claim that, at least for a finite period of time, the system either held or did not hold such property.
+That is: there is a deterministic and reversible process for that finite period of time for such property.
+
 This link between state definition and deterministic processes should not be too surprising as the state, in the context of thermodynamics and systems theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system}.
 As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can be indeed calculated, measured and prepared.
-We can also describe the evolution of each element of the distribution provided that: we have a way to isolate it and study it under deterministic and reversible motion (so that we can define its states); the non-deterministic motion does not alter the system (the set of states remains the same).
-For example, consider a wooden ball described by position and momentum of the center of mass: if randomly scattered by other balls retain the set of states, if thrown into a fire it does not.
+We can also describe the evolution of each element of the distribution provided that: we have a way to isolate it and study it under deterministic and reversible motion (so that we can define microstates); the non-deterministic motion does not alter the system (the set of microstates is preserved by the evolution).
 
-So, if it seems we need to require deterministic and reversible motion, why we call this an assumption?
-First of all, because it's an idealization: it can never be completely achieved in practice.
+As with all assumptions, we should stress that it's an idealization: it can never be completely achieved in practice.
 A system can be prepared or measured up to a certain level of precision.
-Perfect isolation of a system is impossible both practically (e.g.
+Perfect determinism and isolation of a system is impossible both practically (e.g.
 black-body radiation, gravity, ...) and conceptually (e.g.
-if the system is perfectly isolated, we cannot interact with it: how can it studied through experimentation?).
+if the system is perfectly isolated, we cannot interact with it: how can it be physically distinguished?).
 It's a simplifying assumption that can only be taken if the environment and the internal dynamics of the system interact in such a way that they little affect and are little affected by the aspect we are studying.
-As we saw before, for example, assuming that the states consists of the position and momentum of the center of mass requires assuming that the Brownian motion of the body is negligible.
+As we saw before, for example, assuming that the state consists of the position and momentum of the center of mass requires assuming that the Brownian motion of the body is negligible.
 
-The second reason is that technically the assumption comes first.
-If we start looking for a law, we are implicitly assuming that there exists one to be found.
-The opposite assumption, there is no law to be found, can't really be experimentally proven exhaustively (i.e.
-you can only show that you have not found it among a certain class).
-Very often the hard part in science is finding for what system and in what conditions such assumption hold.
-Finding the law after that is a much simpler endeavour, and is something that is routinely done by students in their intro labs.
+Yet, this is a fundamental assumption in the sense that it is needed.
+If a particular set of states does not satisfy deterministic and reversible evolution under certain conditions, what we do is to keep at it until we find a set that does.
+That is: we work to restore the assumption.
+Finding new sets of states with new laws of evolution is, in fact, what leads to new physics.
+We will therefore call \emph{fundamental model of physics} the triad of state, unstated part and environment, with the assumption that the state may undergo deterministic and reversible evolution, and the unstated part undergoes non-deterministic non-reversible evolution that does not alter the set of states.
 \end{rationale}
+
+\section{States}
 

### 2016-02-07 2015CompleteDerivation/completeDerivation.tex
Adding physical distinguishability and topology part

@@
 
-\section{Determinism and reversibility}
+\section{On studying a physical system}
 
@@
 
-We are now ready to capture the elements of our discussion and their properties through mathematical definitions.
+Before we are able to define how states should be captured mathematically, we need to better specify what we mean by physically distinguishable.
+Therefore we need to characterise what we mean by measurement, particularly because it has become such an overloaded term.
+
+Consider the motion of a cannonball under inertial and gravitational forces.
+Light will be scattering off of it; as it lands the ground will be deformed and its temperature will rise.
+Those external physical process, which happen no matter what we do, can be used to distinguish the motion of the cannonball as their outcome are correlated.
+Therefore we can learn the position by looking at the reflected light, and learn the final kinetic momentum by looking at the deformation of the ground.
+In principle, any external process that has a correlation with the states under study can be used to perform a measurement, and any measurement is based on such a process.
+That is: for us a measurement is simply a physical process that we can use to distinguish states.
+There is no particular role played by the observer, except maybe create the device and forcing the system under study to interact with it.
+
+This external process may be quite complicated: when a charged particle enters a calorimeter, a shower of particles is produced, photons are captured and are directed to photomultipliers, a current is read out, the current is then digitized, and so on.
+Some processes interfere with the system, they affect its dynamics, and other are destructive, the system no longer can be described by the original set of states.
+A bubble chamber is an example of the first (e.g.
+the magnetic field curves the motion of a charged particle); burning a substance to determine its caloric content is an example of the latter.
+In these cases, special care needs to make sure to link the outcomes with the original state, unaffected state.
+But in the end, the conceptual description remains the same.
+
+We have a physical process that provides an outcome.
+This process has a set of potential outcomes.
+To each potential outcome we can associate a set of states.
+For example, if the electron follows this path, then its state is among the ones that has spin $1/2$.
+Since all measurements are finite precision, each outcome can be labelled by a finite set of digits, so the set of outcomes is countable.
+Note that different outcomes for the same process can overlap.
+For example, $4.12 \pm 0.05$ cm and $4.13 \pm 0.05$ cm are both legitimate possible outputs of the same measurement device.
+But since all states must be distinguishable, given two arbitrary states there must be a process precise enough to tell them apart.
+That is: the potential outcomes associated with the two states do not overlap.
+For example, $4.12 \pm 0.0005$ cm and $4.13 \pm 0.0005$ cm do not overlap anymore.
+
+We can also combine two outcome of two different processes.
+For example, if one tells us spin is $+1/2$ and another that position is $4.13 \pm 0.05$ cm, then can say the state has both.
+That is, the state is inside both of the sets of states associated with the two outcomes: the intersection of two sets associated with an outcome is also a set associated with an outcome.
+We can make similar arguments about the union as we can always reduce the precision by combining different outcome of the same process.
+
+This model maps very naturally to a topological space.
+The states are the elements of a set and the physical outcome provides a topology on that set.
+The fact that two elements of the set can be distinguished requires the space to be Hausdorff.
+The fact that the potential outcome of a process are countable requires the space to be second countable.
 
 \begin{defn}\label{statedef}
-The state space $\mathcal{S}$ of a physical system is a set.
+The state space $\mathcal{S}$ of a physical system is a Hausdorff and second countable topological space.
 \end{defn}

### 2016-02-08 2015CompleteDerivation/completeDerivation.tex
Updating status of different sections

@@
 
-%TODO: it would be nice to have "examples" always refer to the same cases, and that they pertain to different type of physics.
-E.g.
-trajectory, gravity, thermodynamics, fluid dynamics, biology, electromagnetism...
+% Status: Conceptual work done. Text ready for review.
+
+%TODO: incorporate feedback from Dave.
+Determinism definition as "relative" to choice of state and external processes may need to be reinforced (the single sentence pointing that out is read, forgotten, and then confusion later follows).
+May want to completely avoid QM at this point.
+See if can mention without raising further distracting questions.
 
@@
 
+% Status: concepts still need to be finished. Started to work on linking topology to physically distinguishable states.
+
+%Open questions: smoothness (obviously required to define densities later, is it required in general? What does it means that physical processes/measurements don't give a smooth topology?)
+
+%Things to add: discrete sets evolved continuous = identity map. Subdivide finite interval, same permutation at each time step, exclude odd perms, third-odd perms, ...
+
 Before we are able to define how states should be captured mathematically, we need to better specify what we mean by physically distinguishable.
@@
 \begin{justification}
-The collection of all possible states forms a set as we can physically distinguish between them.
+TODO: all possible states form a set.
+All possible set of states associated with the outcome of the process form a topology.
+Verify requirements on entire set, empty set, intersections and unions.
+Distinguishability of two elements in terms of process outcome leads to Hausdorff.
+Processes have countable possible outcome means base for topology must be countable, or we would have elements that are not distinguishable, therefore second countability.
 \end{justification}

### 2016-02-11 2015CompleteDerivation/completeDerivation.tex
Status and todos update
More on processes and outcomes

@@
 Before we are able to define how states should be captured mathematically, we need to better specify what we mean by physically distinguishable.
-Therefore we need to characterise what we mean by measurement, particularly because it has become such an overloaded term.
+Therefore we need to characterize what we mean by measurement, particularly because it has become such an overloaded term.
 
@@
 Light will be scattering off of it; as it lands the ground will be deformed and its temperature will rise.
-Those external physical process, which happen no matter what we do, can be used to distinguish the motion of the cannonball as their outcome are correlated.
+Those external physical processes, which happen no matter what we do, can be used to distinguish the motion of the cannonball as their outcomes are correlated.
 Therefore we can learn the position by looking at the reflected light, and learn the final kinetic momentum by looking at the deformation of the ground.
@@
 
-This external process may be quite complicated: when a charged particle enters a calorimeter, a shower of particles is produced, photons are captured and are directed to photomultipliers, a current is read out, the current is then digitized, and so on.
+This external process may be quite complicated: when a particle enters a calorimeter, a shower of particles is produced, photons are captured and are directed to photomultipliers, a current is read out, the current is then digitized, and so on.
 Some processes interfere with the system, they affect its dynamics, and other are destructive, the system no longer can be described by the original set of states.
-A bubble chamber is an example of the first (e.g.
-the magnetic field curves the motion of a charged particle); burning a substance to determine its caloric content is an example of the latter.
-In these cases, special care needs to make sure to link the outcomes with the original state, unaffected state.
+A tracking chamber is an example of the first (the magnetic field curves the motion of a charged particle); burning a substance to determine its caloric content is an example of the latter.
+In these cases, special care is needed to make sure to link the outcomes with the original state, unaffected state.
 But in the end, the conceptual description remains the same.
+
+%TODO: It's not about precision! Discrete variables are known to infinite precision! Is about information.
 
@@
 To each potential outcome we can associate a set of states.
-For example, if the electron follows this path, then its state is among the ones that has spin $1/2$.
-Since all measurements are finite precision, each outcome can be labelled by a finite set of digits, so the set of outcomes is countable.
+For example, if the electron follows a certain path, then its state is among the ones that has spin $+1/2$.
+Since all measurements are finite precision, each outcome can be labeled by a finite set of digits, so the set of outcomes is countable.
 Note that different outcomes for the same process can overlap.
@@
 
-We can also combine two outcome of two different processes.
-For example, if one tells us spin is $+1/2$ and another that position is $4.13 \pm 0.05$ cm, then can say the state has both.
+%TODO If we start "combining processes" then is it really _a_ process?
+
+We can also combines two outcome of two different processes.
+For example, if one tells us spin is $+1/2$ and another that position is $4.13 \pm 0.05$ cm, then we can say the state has both.
 That is, the state is inside both of the sets of states associated with the two outcomes: the intersection of two sets associated with an outcome is also a set associated with an outcome.
@@
 
+With our state space defined, deterministic and reversible evolution corresponds to a bijective map between initial and final states.
+
 \begin{defn}\label{detrevmap}
@@
 
-%In practice, we use physical quantities to identify states.
-For example, the state for an ideal gas is given by pressure, volume and temperature.
-To clarify whether we are referring to a particular value of such quantity, or the quantity itself, we introduce the following terminology:
-%\begin{center}
-%    \begin{tabular}{ | p{2.5cm} | p{5.5cm} | }
-%    \hline
-%    State variable & A quantity (discrete or continuous) that must be specified to identify a state (e.g. position). \\ \hline
-%    Label & A particular value for a state variable (e.g. position = $5m$). \\ \hline
-%    Label range & A set of possible values for a state variable (e.g. position = $[4.5m, 5.5m]$). \\
-%    \hline
-%    \end{tabular}
-%\end{center}
-
-To identify and name states, we will typically use a set of quantities, typically numbers.
+To identify and name states we will use a set of quantities, typically numbers.
 For example, the orbital of an electron in a hydrogen item is identified by the quantum numbers $n$, $l$, $m$ and $s$.
@@
 The second approach is to transport the state variables.
-The idea is to keep the connection to the initial state by labelling the future state, instead by the future values of $q^i$, by the original one.
+The idea is to keep the connection to the initial state by labeling the future state, instead by the future values of $q^i$, by the original one.
 So we introduce a new set of variables for which $\hat{q}^i(\hat{\mathcal{s}})=q^i(\mathcal{s})$, which we can always do as the evolution is deterministic and reversible.
@@
 \begin{justification}
-TODO: set of state variables locally define a chart; inherit topology from charts (small change in state makes small change in state variable, makes small changes in outcomes?)
+TODO: set of state variables locally define a chart, isomorphic to $\mathbbm{R}$.
 \end{justification}
@@
 \section{Composite systems and reducibility}
+
+%Status: conceptual work is done, need to write down
+
+%TODO: need more examples within explanation to make more intuitive
 
@@
 This intuitive picture is unfortunately not suitable to be generalized to different contexts, so we develop another.
+
+%TODO: move second picture to QM
 

### 2016-02-12 2015CompleteDerivation/completeDerivation.tex
Added the basic points for non-relativistic hamiltonian

@@
 This model maps very naturally to a topological space.
-The states are the elements of a set and the physical outcome provides a topology on that set.
+The states are elements of a set and the physical outcome provides a topology on that set.
 The fact that two elements of the set can be distinguished requires the space to be Hausdorff.
@@
 
+Show countable basis. $\mathcal{c}= \sum\limits_{\mathcal{s} \in \mathcal{S}} a(\mathcal{s}) \mathcal{e}(\mathcal{s})=\sum\limits_{q \in \mathbbm{Q}} a(q) \mathcal{e}(q)$
+
+% Continuous state variable: Q is measured at constant t
+
+Uncountable basis does not work. $\mathcal{c}= \int\limits_{q \in \mathbbm{Q}} \rho(q) \mathcal{e}(q)dq$ but $\rho(q')=|\partial _{q}q'|\rho(q)\neq\rho(q)$
+
+% TODO: need to understand a valid notation for composite state as infinite sum. The densities is the coefficient, the basis are are states, what is the differential?
+
+Find smallest object that can have an invariant density. $\mathcal{c}= \int\limits_{T^*\mathbbm{Q}} \rho(q,k) \mathcal{e}(q,k)dq\wedge dk$
+
+\begin{defn}\label{classical_phase_space}
+The state space $\mathcal{S}$ for the particles of a homogeneous classical  system is a cotangent bundle $T^*\mathbbm{Q}$.
+\end{defn}
+
+\begin{justification}
+	TODO
+\end{justification}
+
+Density allows to count particle states. Example: two uniform distribution for the same amount of particles.
+
+Area in each d.o.f.
+is number of possibilities.
+Total number of states is product of number of possibilities for independent d.o.f.
+Independent d.o.f.
+are orthogonal.
+Leads to omega metric.
+As in discrete case, conserved during evolution.
+
+Also: we can search for omega looking from constraints. Must be linear. Must be skew-symmetric (applied to the same vector twice must be zero). Must be invariant by coord transformation.
+
+\begin{defn}\label{symplectic_manifold}
+	The state space $\mathcal{S}$ for the particles of a homogeneous classical  system is a symplectic manifold with metric $\omega$, the canonical two-form.
+\end{defn}
+
+\begin{justification}
+	TODO
+\end{justification}
+
 \subsection{Classical hamiltonian assumption}
@@
 Still non-reversible (if As' = F(As), then nothing else can be F(As) and have bijective map).
-
-% Continuous state variable: Q is measured at constant t
 
@@
 
+\begin{defn}\label{canonical_transformation}
+	A deterministic and reversible process is a diffeomorphism that preserves $\omega$, that is a canonical transformation.
+\end{defn}
+
+\begin{justification}
+	TODO
+\end{justification}
+
+\begin{defn}\label{hamiltonian}
+	A continuous deterministic and reversible process admits a potential $H$, and the laws of evolution are of the form (Hamilton equations).
+\end{defn}
+
+\begin{justification}
+	TODO
+\end{justification}
+
 \section{Quantum systems}

### 2016-02-12 2015CompleteDerivation/completeDerivation.tex
Incorporated Dave's comments on first section

@@
 
-% Status: Conceptual work done. Text ready for review.
-
-%TODO: incorporate feedback from Dave.
-Determinism definition as "relative" to choice of state and external processes may need to be reinforced (the single sentence pointing that out is read, forgotten, and then confusion later follows).
-May want to completely avoid QM at this point.
-See if can mention without raising further distracting questions.
+% Status: Conceptual work done. Text ready for review. Incorporated Dave feedback
 
@@
 In this context, we define a process \emph{deterministic} if the state at a given time uniquely identifies states at future times, and \emph{reversible} if it identifies states at past times.
-This means that, in general, the evolution of the unstated part is non-deterministic and/or non-reversible in the sense that the state of the system is not enough to determine its evolution.
-For example, the overall motion of a fluid in a pipe can be considered deterministic even if the motion of each molecule is not.
-The motion is deterministic and/or reversible only as far as the state is concerned.
+While this is a common enough definition, we need to be clear how this applies to the unstated part.
+That is: the evolution of the unstated part is \emph{always} non-deterministic (and non-reversible) in the sense that the state of the system does not determine its evolution.
+For example, suppose we define the state as position and momentum of the center of mass of a cannonball.
+Suppose that the process is deterministic on that state.
+What does it tell us about the cannonball temperature, or about the motion of each of its atoms?
+Nothing.
+In this sense, the unstated part is non-deterministic and non-reversible.
+Could we extend the state and the laws of evolution to account for temperature?
+Yes, but that would be a different process defined on a different state.
+Does it mean the unstated part is always evolving chaotically?
+Not at all: the temperature may remain constant throughout the motion of the cannonball.
+Yet we wouldn't know, since we are not studying it: the motion is deterministic and/or reversible only as far as the state is concerned.
 With this in mind, we introduce the following:
@@
 The shape and the parameters of the distribution can be deterministic, even if the trajectories of the parts are not (those fall within the unstated part).
-In fact: we \emph{cannot} assume trajectories and states are always defined at all for the unstated part.
-Consider a muon and its decay into an electron and two neutrinos: it is clear that the three outgoing particles have a state and trajectory of their own, it is clear that the resulting total mass and energy came from the muon.
-Yet, before the decay, we cannot ascribe an internal independent state and evolution to each part.
-The state of a muon is not some combination of the state of an electron and two neutrinos.
-That is: the unstated part is not just microstates.
+In fact: we \emph{cannot} assume trajectories and states are always defined at all for the unstated part, as this includes also the unknown unknowns.
+Will return to this aspect when discussing quantum systems.
 
@@
 
+% Muon example.
+Consider a muon and its decay into an electron and two neutrinos: it is clear that the three outgoing particles have a state and trajectory of their own, it is clear that the resulting total mass and energy came from the muon.
+Yet, before the decay, we cannot ascribe an internal independent state and evolution to each part.
+The state of a muon is not some combination of the state of an electron and two neutrinos.
+That is: the unstated part is not just microstates.
+
 \begin{thebibliography}{0}

### 2016-02-16 2015CompleteDerivation/completeDerivation.tex
Fix typo

@@
 For discrete state variables, the total number of states is the product of the possibilities of each independent state variable.
-That is: if $\Delta q^1$ and $Delta q^2$ are range of possibilities for two independent variables, the total number of possibilities $\#(\Delta q^1) \#(\Delta q^2)$.
+That is: if $\Delta q^1$ and $\Delta q^2$ are range of possibilities for two independent variables, the total number of possibilities $\#(\Delta q^1) \#(\Delta q^2)$.
 These relationships are maintained during evolution by the transported variables.

### 2016-02-16 2015CompleteDerivation/completeDerivation.tex
Finished annotating state section with all open issues

@@
 
+TODO: det/rev map from manifold to manifold.
+Must be physically distinguishable.
+Can't be the whole class of functions: refining domain does not refine codomain.
+(mathematically, cardinality greater than continuoum, i.e.
+not second countable) We need that arbitrarily refining domain refines the codomain.
+Equivalent to continuous function(?).
+OR Measurement/process outcomes from initial states corresponds to outcomes on final states.
+That is: we map open sets to open sets.
+
 \begin{defn}\label{detrevmap}
-A deterministic and reversible evolution is a bijective map $f_{\Delta t}:\mathcal{S} \leftrightarrow \mathcal{S}$ on the state space.
+A deterministic and reversible evolution is a self-homeomorphism on the state space $\mathcal{S}$.
 \end{defn}
@@
 $g \circ f = id$ as mapping forward and then backward must return the original element.
+
+TODO: show it maps open sets to open sets.
 \end{justification}
@@
 But there is nothing telling us that they must be directly physically meaningful or that their relationship to measurable quantities is the same if we change physical processes.
+
+TODO: maybe best to define state vector directly, and state variables as part of state vector?
+Define discrete state variable when co-domain is integers and define continuous state variable when co-domain is real?
+Can we prove a state vector always exists, of real variables?
+For discrete topology, we should always be able to do it.
+In general, state vector defines charts, so we should not have a single chart for the whole state space.
+What does it mean physically?
+Does it matter?
 
@@
 
-\begin{prop}\label{evolved variable}
-TODO: Define and prove existence
-\end{prop}
+\begin{defn}\label{evolved_transported_variable}
+TODO: Define evolved and transported variables in terms of pushback/pushforward
+\end{defn}
 
-\begin{prop}\label{transported variable}
-TODO: Define and prove existence
-\end{prop}
+\begin{justification}
+TODO: prove existence
+\end{justification}
 
@@
 \end{proof}
+
+TODO: in continuous, not necessarily the Cartesian product.
+But we can always create a chart with state variables.
+State space is a manifold of dimension n.
+In principle, we could say more.
+For example, a deterministic/reversible map would not map a set to its subset or superset.
+We could define a metric, this would give us an ergotic system.
+Not going to do that.
 

### 2016-02-18 2015CompleteDerivation/completeDerivation.tex
More notes on classical systems

@@
 
+TODO: Still not clear what's the best way to introduce things.
+We do want to show that state variable are maps from states to "labels".
+We want to call discrete variables those that map to integer and continuous variables those that map to real numbers.
+Then we want to concentrate on state space that can be described by a finite number of either discrete variables or continuous variables.
+State vector may not be valid in general: e.g.
+we can't assume we have a single chart for the whole space.
+
 \begin{prop}\label{state_vector}
@@
 
+TODO: should we note already that the state space is a Lie group?
+
 \subsection{Classical homogeneous systems}
@@
 
+Need to define inner product.
+Introduce $Q:\mathcal{C} \to \mathbbm{R}$ amount of "stuff" of the system.
+$Q:\mathcal{S}\times\mathcal{C} \to \mathbbm{R}$ amount of stuff of the composite system in that particular state.
+We can use those those to construct inner product, and to show that states are bounded, but: norm induced by the inner product is not physical.
+
+\begin{defn}\label{classical_inner_product}
+	TODO: there exist an inner product.
+\end{defn}
+
+\begin{justification}
+	TODO
+\end{justification}
+
 Show countable basis. $\mathcal{c}= \sum\limits_{\mathcal{s} \in \mathcal{S}} a(\mathcal{s}) \mathcal{e}(\mathcal{s})=\sum\limits_{q \in \mathbbm{Q}} a(q) \mathcal{e}(q)$
@@
 
-% TODO: need to understand a valid notation for composite state as infinite sum. The densities is the coefficient, the basis are are states, what is the differential?
+TODO: need to understand a valid notation for composite state as infinite sum. The densities is the coefficient, the basis are are states, what is the differential?
 
-Find smallest object that can have an invariant density. $\mathcal{c}= \int\limits_{T^*\mathbbm{Q}} \rho(q,k) \mathcal{e}(q,k)dq\wedge dk$
+Find smallest object that can have an invariant density (inner product). $\mathcal{c}= \int\limits_{T^*\mathbbm{Q}} \rho(q,k) \mathcal{e}(q,k)dq\wedge dk$
 
@@
 \end{justification}
+
+TODO: generalization of inner product gives us $\int \rho_1(q, p) \rho_2(q, p) dq\wedge wp$.
+Space of continuous functions is not a complete metric space.
+Extend to square integrable.
+Get a Hilbert space.
+But topology is not a second countable: more than aleph 1.
+No physical justification, in fact we introduce unphysical functions (i.e.
+0 rational, 1 irrational).
+Discontinuous functions not physical states: they are not distinguishable.
 

### 2016-03-03 2015CompleteDerivation/completeDerivation.tex
Incorporated Isaac feedback

@@
 
-% Status: Conceptual work done. Text ready for review. Incorporated Dave feedback
+% Status: Conceptual work done. Text ready for review. Incorporated Dave and Isaac feedback
 
@@
 Gravitational and inertial forces have not changed, yet the set of states have.
-The issue is that the non-deterministic evolution must not change the set of states itself.
-That is: if we apply the non-deterministic evolution to any state, we still need to end up in one of the possible states.
-If the Brownian motion is strong enough, a simple position/momentum pair is not sufficient.
+In other words: the set of states must also be closed under non-deterministic.
+If the Brownian motion is not negligible, we do not end in a well defined position/momentum pair, even if we start from one.
+%Reworded this for Isaac
 
@@
 
-Fourth, to be able to ascribe a property to a system we need to claim that, at least for a finite period of time, the system either held or did not hold such property.
+Fourth, to be able to ascribe a property to a system we need to claim that, at least for a finite interval of time, the system either held or did not hold such property.
 That is: there is a deterministic and reversible process for that finite period of time for such property.

### 2016-03-03 2015CompleteDerivation/completeDerivation.tex
More work on  classical state space

@@
 
-\subsection{Homogeneous systems}
-As mentioned before, we will limit ourselves to the case where our system is homogeneous: the parts are indistinguishable, the set of states they can occupy is the same, they are under the same external and internal forces.\footnote{Non-homogeneous systems can in principle be decomposed in homogeneous components.
-Therefore their study can be reduced to the study of homogeneous systems.} What homogenous means depends on context (e.g.
-air can be thought as homogeneous if the mixture of gasses does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is conserved by time evolution.
-Nonetheless, the state space of such system is richer in the sense that it comes equipped with the notion of composition and its rules.
-These are suitable captured mathematically by an additive abelian group.
+\subsection{Homogeneous decomposable systems}
+The notion that the system is decomposable means the states is equipped with a rule of composition that allows to write $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$: the composite system is the sum of its part.
+For example, the state of a ball is equal to the state of its left and right parts.
+
+The notion that the system is homogeneous means that the states of the composite and of the parts are not unrelated: they are made of the same material.\footnote{Whether a system is homogeneous depends on context (e.g.
+air can be thought as homogeneous if the mixture of gases does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is maintained by by time evolution.} In fact, the state space $\mathcal{C}$ of all systems composed of such material will contain the state of system and all its part (i.e $\mathcal{c}, \mathcal{c}_1, \mathcal{c}_2 \in \mathcal{C}$).
+Since combining any two systems made of a homogeneous material will always give us a system made of the same homogeneous material (e.g.
+combining elements made of water gives us another element made of water), the state space $\mathcal{C}$ is closed under composition.
+
+As we study the system under evolution, we'll also want to study state changes.
+For example, if a stable mixture of gas expands, we'll want to know the difference between the initial and final distributions.
+That is: $\Delta\mathcal{c}=\hat{\mathcal{c}}-\mathcal{c}$.
+Note that such state difference may not describe a physical state: it may remove some material from one location to add it somewhere else.
+Yet, these state changes are still physically distinguishable objects that provide a configuration for the same homogeneous material, therefore we extend the state space $\mathcal{C}$ to include state differences.
+All combined, this gives us the structure of an abelian group.
 
@@
 
-TODO: should we note already that the state space is a Lie group?
+Note that, due to \ref{statedef}, $\mathcal{C}$ is also a topological space.
 
@@
 That is: it is made of arbitrarily small parts, which we call particles, and the state of the whole system is equivalent to the sum of the states of the parts.
-Given the state space $\mathcal{S}$ for the parts, identifying a composite state $\mathcal{c}$ is equivalent to quantify the amount of particles that are present in each state.
-Since the system is infinitesimally reducible, such amount can be increased or decreased infinitesimally and will be a real number.
-In other words: the classical state is a real valued distribution over $\mathcal{S}$.
-This intuitive picture is unfortunately not suitable to be generalized to different contexts, so we develop another.
+Given the state space $\mathcal{S}$ for the infinitesimal parts, a composite state $\mathcal{c}$ will tell us the amount of material for each possible particle state.
+That is, such states comes with a function $Q: \mathcal{S}\times\mathcal{C} \to \mathbbm{R}$ that returns the amount of material in the composite state $\mathcal{c}$ that is prepared in state $\mathcal{c}$.
+For example, given a certain configuration of gas, we can tell the amount that is at a specific point in place with a specific value of momentum.
+The notion that the parts are infinitesimal requires the co-domain of $Q$ to be continuous, as opposed to discrete.
+Note that the state space $\mathcal{S}$, instead, can be finite.
+For example, for a system composed of different tanks connected by pipes, the state of the composite system could be the overall distribution of water among the tanks.
+The amount in each tank is continuous (as we assume the water to be infinitesimally divisible) yet there are only a finite number of tanks the water can be placed.
 
-%TODO: move second picture to QM
-
-Consider the following expression:
-\begin{align*}
-\mathcal{c} = \sum\limits_{i \in \mathbbm{I}} a_i \mathcal{e}(i)
-\end{align*}
-$\mathbbm{I}$ represents all the possible values of a state variable representing the best description of an infinitesimal part that a composite system can give.
-$\mathcal{e}(i)$ represents a composite system made of a unit amount of particles, all prepared in the state determined by $i$.
-$a_i$ represents a transformation that changes the whole state, without affecting the value $i$ for any infinitesimal part.
-For a classical system, $i$ represents the full information about the infinitesimal part, and therefore includes the value of all state variables (i.e.
-the state vector); the only transformation left is increasing/decreasing the amount of particles by a scalar multiple.
-For a different system, where $i$ does not represent the full state of the parts, $\mathcal{e}(i)$ will be an ensemble and $a_i$ also represents an internal transformation of such ensemble.
-This is intuitively how we decouple the composite state into the part $i$ that describes the infinitesimal parts, and the part $a_i$ that describes how the parts are composed.
-Giving a full description of the state space of a decomposable system means fully characterizing these components: the state variable $\mathbbm{I}$ and the set of transformations $A$.
-In the classical case $\mathcal{e}(i)$ form a basis of a real vector space, where $i$ is the state vector for an infinitesimal part, and $a_i$ a real valued coefficient that represents the amount of particles in each state, which describes how the parts are combined.
+As we combine the state of finite parts, the distributions over particle states sum (i.e. they combine linearly). This gives us the structure of a real vector space.
 
@@
 
+%TODO: moved from classical
+%This intuitive picture is unfortunately not suitable to be generalized to different contexts, so we develop another.
+
+%Consider the following expression:
+%\begin{align*}
+%\mathcal{c} = \sum\limits_{i \in \mathbbm{I}} a_i %\mathcal{e}(i)
+%\end{align*}
+%$\mathbbm{I}$ represents all the possible values of a state variable representing the best description of an infinitesimal part that a composite system can give.
+$\mathcal{e}(i)$ represents a composite system made of a unit amount of particles, all prepared in the state determined by $i$.
+$a_i$ represents a transformation that changes the whole state, without affecting the value $i$ for any infinitesimal part.
+For a classical system, $i$ represents the full information about the infinitesimal part, and therefore includes the value of all state variables (i.e.
+the state vector); the only transformation left is increasing/decreasing the amount of particles by a scalar multiple.
+For a different system, where $i$ does not represent the full state of the parts, $\mathcal{e}(i)$ will be an ensemble and $a_i$ also represents an internal transformation of such ensemble.
+This is intuitively how we decouple the composite state into the part $i$ that describes the infinitesimal parts, and the part $a_i$ that describes how the parts are composed.
+Giving a full description of the state space of a decomposable system means fully characterizing these components: the state variable $\mathbbm{I}$ and the set of transformations $A$.
+In the classical case $\mathcal{e}(i)$ form a basis of a real vector space, where $i$ is the state vector for an infinitesimal part, and $a_i$ a real valued coefficient that represents the amount of particles in each state, which describes how the parts are combined.
+
 \begin{thebibliography}{0}

### 2016-03-09 2015CompleteDerivation/completeDerivation.tex
More work on composite state definition

@@
 Gravitational and inertial forces have not changed, yet the set of states have.
-In other words: the set of states must also be closed under non-deterministic.
+In other words: the set of states must also be closed under non-deterministic evolution.
 If the Brownian motion is not negligible, we do not end in a well defined position/momentum pair, even if we start from one.
@@
 
-%TODO: need more examples within explanation to make more intuitive
-
 Given that scientific reductionism (i.e.
@@
 \subsection{Homogeneous decomposable systems}
-The notion that the system is decomposable means the states is equipped with a rule of composition that allows to write $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$: the composite system is the sum of its part.
+The notion that a system is decomposable means the states is equipped with a rule of composition that allows to write $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$: the composite system is the sum of its part.
 For example, the state of a ball is equal to the state of its left and right parts.
@@
 \begin{defn}\label{reducible_state_space}
-The state space $\mathcal{C}$ for a homogeneous decomposable system is an additive abelian (i.e. commutative) group.
+The state space $\mathcal{C}$ for decomposable systems made of a homogeneous material is an additive abelian (i.e. commutative) group.
 \end{defn}
@@
 There exist a law of composition $+ : \mathcal{C} \times \mathcal{C} \rightarrow \mathcal{C}$ that takes two states and returns one that is the physical composition of the two.
-The domain and codomain match because the system is homogeneous.
+The domain and codomain match because the material is homogeneous.
 The law is commutative $\mathcal{c}_1 +\mathcal{c}_2 = \mathcal{c}_2+\mathcal{c}_1$ and associative $(\mathcal{c}_1 + \mathcal{c}_2) + \mathcal{c}_3 = \mathcal{c}_1 + (\mathcal{c}_2 + \mathcal{c}_3)$, as it does not matter in what order we physically compose the parts.
 There exist a unique zero element $\mathcal{c} + 0 = \mathcal{c}$ and it represent the physically empty state (i.e.
-no system).
+no amount of material).
 
@@
 Such inverse may introduce objects that do not properly represent a physical state, but a state change.
-A change of a physically distinguishable object is itself physically distinguishable therefore is physically well defined.
-We'll still call $\mathcal{C}$ the state space committing an abuse of terminology.
+A change of a physically distinguishable object is itself physically distinguishable therefore $\mathcal{C}$ is still a Hausdorff and second countable topological manifold as per \ref{statedef}.
+We'll still call $\mathcal{C}$ state space committing an abuse of terminology.
 \end{justification}
-
-Note that, due to \ref{statedef}, $\mathcal{C}$ is also a topological space.
 
@@
 Given the state space $\mathcal{S}$ for the infinitesimal parts, a composite state $\mathcal{c}$ will tell us the amount of material for each possible particle state.
-That is, such states comes with a function $Q: \mathcal{S}\times\mathcal{C} \to \mathbbm{R}$ that returns the amount of material in the composite state $\mathcal{c}$ that is prepared in state $\mathcal{c}$.
-For example, given a certain configuration of gas, we can tell the amount that is at a specific point in place with a specific value of momentum.
-The notion that the parts are infinitesimal requires the co-domain of $Q$ to be continuous, as opposed to discrete.
-Note that the state space $\mathcal{S}$, instead, can be finite.
+That is, each state is fully identified by a function $a: \mathcal{S} \to \mathbbm{R}$ that returns the amount of material in the composite state $\mathcal{c}$ that is prepared in state $\mathcal{s}$.
+For example, given a certain configuration of gas, we can tell the amount that is at a specific point in space with a specific value of momentum.
+The notion that the parts are infinitesimal requires the co-domain of $a$ to be continuous, as opposed to discrete.
+The state space $\mathcal{S}$, instead, can be finite.
 For example, for a system composed of different tanks connected by pipes, the state of the composite system could be the overall distribution of water among the tanks.
-The amount in each tank is continuous (as we assume the water to be infinitesimally divisible) yet there are only a finite number of tanks the water can be placed.
+The amount in each tank is continuous (as we assume the water to be infinitesimally divisible) yet there are only a finite number of tanks the water can be placed into.
 
-As we combine the state of finite parts, the distributions over particle states sum (i.e. they combine linearly). This gives us the structure of a real vector space.
+As we combine the state of finite parts, we sum the distributions over particle states $a(\mathcal{s})=a_1(\mathcal{s})+a_2(\mathcal{s})$: the amount of material in the composite state is the sum of the amount of material of the parts.
+This gives us the structure of a real vector space.
 
@@
 The state space $\mathcal{C}$ for a homogeneous classical (i.e.
-infinitesimally reducible) system is a vector space over $\mathbbm{R}$.
+infinitesimally reducible) material is a vector space over $\mathbbm{R}$.
 It admits a basis isomorphic to the state space $\mathcal{S}$ of its particles (i.e.
@@
 \begin{justification}
-The state space is an abelian group as the system is homogeneous and decomposable and \ref{reducible_state_space}.
+The state space is an abelian group as the material is homogeneous and decomposable and \ref{reducible_state_space}.
 
-We now consider the set of transformations $A$ that increase or decrease the amount of parts of the system by a constant.
-We claim $A$ is a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics sense (a physical quantity with a value for each point in space).} isomorphic to $\mathbbm{R}$.
-Consider $a: \mathbbm{R} \rightarrow A$ the mapping between a number and the transformation that increases or decreases the size of the system by that number.
+We now consider the set of transformations $T$ that increase or decrease the amount of material in the system by a constant.
+We claim $T$ is a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics sense (a physical quantity with a value for each point in space).} isomorphic to $\mathbbm{R}$.
+Consider $\tau: \mathbbm{R} \rightarrow T$ the mapping between a number and the transformation that increases or decreases the amount of material by that number.
 This transformation exists: the system is infinitesimally decomposable and the amount can be changed continuously.
-Define on $A$ an addition $+: A \times A \rightarrow A$ and a multiplication $*: A \times A \rightarrow A$ such that $a(x) + a(y) = a(x+y)$ and $a(x) * a(y) = a(x*y)$, $x,y \in \mathbbm{R}$, so that the sum and product of the transformation is equal to the sum and product of their respective factors.
-$a$ is an isomorphism between $A$ and $\mathbbm{R}$ as fields.
+Define on $T$ an addition $+: T \times T \rightarrow T$ and a multiplication $*: T \times T \rightarrow T$ such that $\tau(a) + \tau(b) = \tau(a+b)$ and $\tau(a) * \tau(b) = \tau(a*b)$, $a,b \in \mathbbm{R}$, so that the sum and product of the transformation is equal to the sum and product of their respective factors.
+$\tau$ is an isomorphism between $T$ and $\mathbbm{R}$ as fields.
 
 We now claim that the state space $\mathcal{C}$ is a vector space over $\mathbbm{R}$.
-The abelian group $\mathcal{C}$ can be extended with the operations defined by $A$, as each element $a \in A$ is a map $a : \mathcal{C} \rightarrow \mathcal{C}$.
-The map has the following properties: $(a_1 + a_2) \mathcal{c} = a_1 \mathcal{c} + a_2 \mathcal{c}$, increasing the size of the system by the sum of two constant is the same as combining the separate increases, and $a (\mathcal{c}_1 + \mathcal{c}_2) = a \mathcal{c}_1 + a \mathcal{c}_2$, increasing the size of the total system is the same as the combination of the increased parts.
-$\mathcal{C}$ is a module over $A$, which is a field and isomorphic to $\mathbbm{R}$.
+The abelian group $\mathcal{C}$ can be extended with the operations defined by $T$, as each element $\tau \in T$ is a map $\tau : \mathcal{C} \rightarrow \mathcal{C}$.
+The map has the following properties: $(\tau_1 + \tau_2) \mathcal{c} = \tau_1 \mathcal{c} + \tau_2 \mathcal{c}$, increasing the amount of material by the sum of two constant is the same as combining the separate increases, and $\tau (\mathcal{c}_1 + \mathcal{c}_2) = \tau \mathcal{c}_1 + a \mathcal{c}_2$, increasing the amount of the total system is the same as the combination of the increased parts.
+$\mathcal{C}$ is a module over $T$, which is a field and isomorphic to $\mathbbm{R}$.
+$\mathcal{C}$ is (isomorphic to) a real vector space.
 
-Consider now the injection map $\mathcal{e}: \mathcal{S} \hookrightarrow \mathcal{C}$ that for each particle state returns a composite state constituted only by a predetermined amount of particles in such state.
+Consider now the injection map $\mathcal{e}: \mathcal{S} \hookrightarrow \mathcal{C}$ that for each particle state returns a composite state constituted only by a predetermined unit amount of material in such state.
 The image of such injection is a basis for $\mathcal{C}$.
 We first claim that all elements are linearly independent.
-Consider $\sum\limits_{\mathcal{s}_i \in \mathcal{S}} a_i \mathcal{e}(\mathcal{s}_i)$, $a_i \in \mathbbm{R}$.
-This corresponds to a physical state with $a_i$ particles in each state $\mathcal{s}_i$.
-This is equal to the empty state if and only if $a_i=0 \forall i$.
-This holds even if $\mathcal{S}$ is not finite.
+Consider $\sum\limits_{\mathcal{s} \in \mathcal{S}} \tau(a(\mathcal{s})) \mathcal{e}(\mathcal{s})$, $a: \mathcal{S} \to \mathbbm{R}$.
+This corresponds to a physical state with $a(\mathcal{s})$ material in each state $\mathcal{s}$.
+This is equal to the empty state if and only if $a(\mathcal{s})=0 \forall \mathcal{s} \in \mathcal{S}$.
 We now claim they span.
-Consider $\mathcal{c} + \sum\limits_{\mathcal{s}_i \in \mathcal{S}} a_i \mathcal{e}(\mathcal{s}_i)$, $a_i \in \mathbbm{R}$.
-This can always be made equal to the empty state by setting $a_i$ to be the opposite of the amount of particles of $\mathcal{c}$ that are in state $\mathcal{s}_i$.
+Consider $\mathcal{c} + \sum\limits_{\mathcal{s} \in \mathcal{S}} \tau(a(\mathcal{s})) \mathcal{e}(\mathcal{s})$.
+This can always be made equal to the empty state by setting $a(\mathcal{s})$ to be the opposite of the amount of material of $\mathcal{c}$ found in state $\mathcal{s}$.
+Both arguments hold even if $\mathcal{S}$ is an infinite or uncountable set, as they hold for each finite subset of $\mathcal{S}$ independently and therefore $\mathcal{C}=\prod\limits_{\mathcal{s} \in \mathcal{S}} span({\mathcal{e}(\mathcal{s})})$.
 \end{justification}

### 2016-03-09 2015CompleteDerivation/completeDerivation.tex
Reviewed state definitions

@@
 
-\section{States}
+\section{States and state space}
 
@@
 
-Before we are able to define how states should be captured mathematically, we need to better specify what we mean by physically distinguishable.
-Therefore we need to characterize what we mean by measurement, particularly because it has become such an overloaded term.
+We now proceed to characterize states and their being physically distinguishable in more precise terms so that we can capture their description mathematically.
+What we'll see is that the set of all physical processes that can be used to gain information about the system, that can be used to perform a measurement, induce a topology on the set of states that is Hausdorff and second countable.
+That is: physical distinguishability  is mathematically captured by topological distinguishability.
+Deterministic and reversible evolution will then preserve the topology, and they will be mathematically captured by self-homeomorphisms.
+
+We'll focus on state spaces that can be described by a set of independent state variables, either discrete or continuous, and see what can be said in general on the evolution of state variables.
+
+\subsection{States and topology}
+
+As the term measurement has become a particularly loaded, let's first characterize what we mean by physical distinguishability in our context.
 
 Consider the motion of a cannonball under inertial and gravitational forces.
-Light will be scattering off of it; as it lands the ground will be deformed and its temperature will rise.
-Those external physical processes, which happen no matter what we do, can be used to distinguish the motion of the cannonball as their outcomes are correlated.
-Therefore we can learn the position by looking at the reflected light, and learn the final kinetic momentum by looking at the deformation of the ground.
+Light will scatter off of it; as it lands the ground will be deformed and the impact will make the temperature rise.
+Those external physical processes, which happen no matter what we do, can be used to distinguish the motion of the cannonball as their outcomes are correlated.\footnote{Note that by requiring a correlation we are implying that measurements are repeatable.
+That is, for similarly prepared states we get similar outcomes.} Therefore we can learn the position by looking at the reflected light, and learn the final kinetic energy by looking at the deformation of the ground.
 In principle, any external process that has a correlation with the states under study can be used to perform a measurement, and any measurement is based on such a process.
 That is: for us a measurement is simply a physical process that we can use to distinguish states.
-There is no particular role played by the observer, except maybe create the device and forcing the system under study to interact with it.
+There is no particular role played by the observer, except maybe setting up such a process and forcing the system under study to interact with it.
 
@@
 A tracking chamber is an example of the first (the magnetic field curves the motion of a charged particle); burning a substance to determine its caloric content is an example of the latter.
-In these cases, special care is needed to make sure to link the outcomes with the original state, unaffected state.
-But in the end, the conceptual description remains the same.
+In these cases, special care is needed to make sure to link the outcomes with the original state.
+But in the end, the conceptual description remains the same: each possible outcome of the process will be associated with a set of states consistent with that outcome (e.g.
+if the cannonball deformed the ground by this amount than the final state its kinetic energy at impact was within this range).
 
-%TODO: It's not about precision! Discrete variables are known to infinite precision! Is about information.
-
-We have a physical process that provides an outcome.
-This process has a set of potential outcomes.
-To each potential outcome we can associate a set of states.
+Each physical process, then, can be thought as having a set of potential outcomes.
+To each potential outcome we can associate a set of states that are consistent with it.
 For example, if the electron follows a certain path, then its state is among the ones that has spin $+1/2$.
-Since all measurements are finite precision, each outcome can be labeled by a finite set of digits, so the set of outcomes is countable.
+However precise our measurements are, each outcome is expressed by a finite set of digits, therefore the set of outcomes is countable.
 Note that different outcomes for the same process can overlap.
@@
 
-%TODO If we start "combining processes" then is it really _a_ process?
+We can also conceptually combine two different processes into a single one.
+That is: having a way to measure quantity $x$ and a way to measure quantity $y$ gives us a way to measure the combination $(x,y)$.
+Formally, the outcomes of the combined process will be the intersections of each pair of outcomes of the original processes.
 
-We can also combines two outcome of two different processes.
-For example, if one tells us spin is $+1/2$ and another that position is $4.13 \pm 0.05$ cm, then we can say the state has both.
-That is, the state is inside both of the sets of states associated with the two outcomes: the intersection of two sets associated with an outcome is also a set associated with an outcome.
-We can make similar arguments about the union as we can always reduce the precision by combining different outcome of the same process.
+We can also coarsen a single process.
+That is: having a way to measure $(x,y)$ gives us a way to measure $x$ alone (or any $f(x,y)$).
+Formally, the outcomes of the coarsened process are given by performing the union of some outcomes of the original process.
 
@@
 With our state space defined, deterministic and reversible evolution corresponds to a bijective map between initial and final states.
-
-TODO: det/rev map from manifold to manifold.
-Must be physically distinguishable.
-Can't be the whole class of functions: refining domain does not refine codomain.
-(mathematically, cardinality greater than continuoum, i.e.
-not second countable) We need that arbitrarily refining domain refines the codomain.
-Equivalent to continuous function(?).
-OR Measurement/process outcomes from initial states corresponds to outcomes on final states.
-That is: we map open sets to open sets.
+Moreover, a physical process that can distinguish initial states is also a physical process that can distinguish final states.
+That is: the topology is mapped and preserved by the deterministic and reversible evolution (i.e.
+initial and final states are equally distinguishable).
+The evolution is then a self-homeomorphism on the state space.
 
 \begin{defn}\label{detrevmap}
-A deterministic and reversible evolution is a self-homeomorphism on the state space $\mathcal{S}$.
+A deterministic and reversible evolution map is a self-homeomorphism on the state space $f:\mathcal{S} \rightarrow \mathcal{S}$.
 \end{defn}
@@
 \begin{justification}
-The system is deterministic: there exists a map $f:\mathcal{S} \rightarrow \mathcal{S}$ that returns the final state given the initial state.
+We claim that $f$ exists. The system is deterministic: given an initial state $\mathcal{s} \in \mathcal{S}$ there exists a well defined final state $f(\mathcal{s})=\hat{\mathcal{s}} \in \mathcal{S}$.
+
+We claim that $f$ is a bijection.
 The system is reversible: there exists a map $g:\mathcal{S} \rightarrow \mathcal{S}$ that returns the initial state given the final state.
-$g \circ f = id$ as mapping forward and then backward must return the original element.
+$g \circ f = f \circ g = id$ as mapping forward and then backward or backwards and then forward must return the original element.
+$f$ admits $g$ as an inverse.
+$f$ is a bijection
 
-TODO: show it maps open sets to open sets.
+We claim $f$ is an open map.
+Let $U \subseteq \mathcal{S}$ represent an outcome of a process.
+$U$ is an open set in the state space topology by definition.
+Given the deterministic and reversible evolution $f$, an outcome on the initial states can be used to distinguish final states.
+The same outcome associated with $U$ as initial states will also be associated with $f(U)$ as final states.
+$f(U)$ is associated with an outcome.
+$f(U)$ is an open set in the state space topology by definition.
 \end{justification}
+
+%TODO: do continuous function form a manifold Hausdorf and second countable?
+
+\subsection{Labeling states}
 

### 2016-03-11 2015CompleteDerivation/completeDerivation.tex
Starting to shape up state labelling

@@
 For example, the orbital of an electron in a hydrogen item is identified by the quantum numbers $n$, $l$, $m$ and $s$.
-We call each of these quantities \emph{state variable} and \emph{state vector} a set of state variables that can identify one and only one state.
-We avoid the term coordinates, even though that's what they mathematically are, as it will create confusion with space coordinates.
-We also avoid the term observable or measurable, as not all state variables map directly to the outcome of an experiment (e.g.
-conjugate momentum as it is not gauge invariant).
-It's clear, though, that even in those cases there must exist, at least in principle, a way to distinguish between the possible values of the state variable (e.g once we fix the vector potential, conjugate momentum can be measured via the kinetic momentum).
-Therefore we can assume that all state variables are in the end physically distinguishable, yet we do not assume all state variables have equal standing in physical meaning.
-Conversely, some measurable quantities are not directly state variables: they may not be defined, as states are, at constant time (e.g.
-velocity) or may depend on the process at hand (e.g.
-acceleration).
-Yet, once a process is fully defined (e.g.
-a Hamiltonian is given), we may have a relationship between those observables and state variables.
-The purpose of state variables is just to label states, which are physically distinguishable and defined at a particular time.
-Therefore the state variable will be physically distinguishable and defined at constant time.
-But there is nothing telling us that they must be directly physically meaningful or that their relationship to measurable quantities is the same if we change physical processes.
+We call each of these quantities \emph{state variables} and \emph{state vector} a set of state variables that can identify one and only one state.
+We call a \emph{possibility} a possible value that can be taken by a state variable.
+We call a state variable \emph{discrete} or \emph{continuous} if the possibilities are integer or real numbers respectively.
+
+We purposely use the term state variables instead of coordinates (even though that's what they are mathematically) as it will create confusion with space-time coordinates.
+We also avoid the term observable or measurable, as not all state variables may be directly physically meaningful (e.g.
+conjugate momentum in a gauge theory).
+The only requirements for state variables is that they identify states.
+This means the possible values for state variables are physically distinguishable and are defined at equal time (since states are defined and mapped at a particular time).
+\footnote{This is the main reason that a quantity like velocity is not a suitable state variable, as it is  defined over an interval of time.
+Therefore velocity is always physically meaningful but is not a state variable in general, while conjugate momentum is always a state variable but is physically meaningful in general.
+As we'll see later, only once a link is established between them they can both be physically meaningful state variables.}
+
+\begin{defn}\label{state_variable}
+	A \emph{state variable} is a homeomphism $q : U \rightarrow \mathbbm{L}$ where $U \subseteq \mathcal{S}$ and $\mathbbm{L}$ is the topological space for the possible values.
+If $\mathbbm{L}\cong \mathbbm{Z}$ the variable is said \emph{discrete}.
+If $\mathbbm{L}\cong \mathbbm{R}$ the variable is said \emph{continuous}.
+	%We also have $\mathcal{s}_1 \neq \mathcal{s}_2 \Leftrightarrow \exists q : U \rightarrow \mathbbm{L} | q(\mathcal{s}_1)\neq q(\mathcal{s}_2) \forall \mathcal{s}_1, \mathcal{s}_2 \in U$.
+\end{defn}
+
+\begin{justification}
+	We claim $\mathbbm{L}$ is a topological space.
+The collection of all possible values form a set.
+Such set is physically distinguishable.
+The physical outcomes induce a Hausdorff and second countable topology, as for \ref{statedef}.
+
+	We claim $q$ is a homeomorphism.
+Elements of $\mathcal{S}$ and $\mathbbm{L}$ are both physically distinguishable therefore a map between them is well defined.
+Let $V \subseteq U$ the set of states associated with a physical outcome.
+$V$ is an open set in the state topology.
+Given the map $q$ an outcome on states can be used to distinguish possibilities of the state variables.
+The same outcome is associated with $g(V)$.
+$g(V)$ is an open set in the topology of possible values.
+
+	%Let $\mathcal{s}_1, \mathcal{s}_2 \in \mathcal{S}$.
+Suppose $\mathcal{s}_1 \neq \mathcal{s}_2$.
+$\mathcal{s}_1$ and $\mathcal{s}_2$ are physically distinguishable.
+There exists a physical process with different outcomes.
+Let $q : \mathcal{S} \rightarrow \mathbbm{I}$ be the state variable associated to that process.
+Then $q(\mathcal{s}_1)\neq q(\mathcal{s}_2)$.
+Now suppose $\nexists q : \mathcal{S} \rightarrow \mathbbm{I} | q(\mathcal{s}_1) \neq q(\mathcal{s}_2)$.
+Then no process can distinguish between the two, they are physically indistinguishable therefore $\mathcal{s}_1 = \mathcal{s}_2$.
+\end{justification}
+
+When assembling multiple state variables into a state vector, it is important to understand how they relate to each other.
+Consider the orbital of an electron in a hydrogen item, which is identified by the quantum numbers $n$, $l$, $m$ and $s$.
+For each combination of $n$, $l$ and $m$, the spin $s$ can have two values.
+The choice of spin is independent from the rest.
+The choice of $l$, though, depends on the choice of $n$: for $n=1$ only $l=s$ is available; for $n=2$ we can choose $l=s$ or $l=p$.
+The choices are not independent.
+That is: two or more variables are independent if there always exists a state for any possible combination, if the total number of states is the cartesian product of the possibilities of each variable.
+
+\begin{defn}\label{independent_state_variables}
+	Two state variables $q_1 : U \rightarrow \mathbbm{L}_1$ and $q_2 : U \rightarrow \mathbbm{L}_2$ are said \emph{independent} if $\exists \mathcal{s} | q_1(\mathcal{s})=l_1, q_2(\mathcal{s})=l_2 \forall l_1 \in q_1(U), l_2 \in q_2(U)$.
+\end{defn}
+
+\begin{justification}
+	TODO What is there to justify?
+\end{justification}
+
+In general, the entire state space may not be identified by a predetermined set of independent state variables.
+Consider the state of a pool table, determined by the number of balls together with position and momentum of the center of mass: the number of state variables is not fixed as it depends on the number of balls.
+But deterministic and reversible evolution cannot take us from a different number of continuous independent state variables (i.e.
+there is no homeomorphism between $\mathbbm{R}^n$ and $\mathbbm{R}^m$).
+That is the number of balls cannot change under deterministic and reversible evolution.
+Therefore we can restrict ourselves to the case where the number of continuous independent state variables is constant without loss of generality.
+This means that, at least locally, the state space is always homeomorphic to $\mathbbm{R}^n$, and is therefore a manifold.
+
+The discrete variable do not present such problems, as any number of them can be flattened out in a single one (i.e.
+$\mathbbm{Z}^n$ is homeomorphic to $\mathbbm{Z}$).
+They will determine the number of connected components of the manifold.
+Once we introduce a continuous parameter for time evolution, though, these also become irrelevant as continuous time evolution cannot move states across disconnected components.
+
+\begin{defn}\label{continuous_state_space}
+	The space state $\mathcal{Q}$ for a system fully identified by a set of $n$ independent continuous state variables is a manifold of dimension $n$.
+\end{defn}
+
+\begin{justification}
+	TODO: set of state variables locally define a chart, isomorphic to $\mathbbm{R}$.
+\end{justification}
 
@@
 Does it matter?
-
-\begin{defn}\label{state_variable}
-Let $\mathbbm{I}$ be a set.
-We call $q : \mathcal{S} \rightarrow \mathbbm{I}$ a \emph{state variable} and a \emph{possible value} (or \emph{possibility}) of $q$ an element $i \in \mathbbm{I}$.
-Also $\mathcal{s}_1 \neq \mathcal{s}_2 \Leftrightarrow \exists q : \mathcal{S} \rightarrow \mathbbm{I} | q(\mathcal{s}_1)\neq q(\mathcal{s}_2) \forall \mathcal{s}_1, \mathcal{s}_2 \in \mathcal{S}$.
-\end{defn}
-
-\begin{justification}
-Consider a physical process that can distinguish between different initial conditions.
-Let $\mathbbm{I}$ be the set of all possible outcomes.
-To each possible state $\mathcal{s}$ will correspond one and only one outcome.
-Let $q : \mathcal{S} \rightarrow \mathbbm{I}$ be the map between state and outcome.
-Such object is physically well defined.
-
-Let $\mathcal{s}_1, \mathcal{s}_2 \in \mathcal{S}$.
-Suppose $\mathcal{s}_1 \neq \mathcal{s}_2$.
-$\mathcal{s}_1$ and $\mathcal{s}_2$ are physically distinguishable.
-There exists a physical process with different outcomes.
-Let $q : \mathcal{S} \rightarrow \mathbbm{I}$ be the state variable associated to that process.
-Then $q(\mathcal{s}_1)\neq q(\mathcal{s}_2)$.
-Now suppose $\nexists q : \mathcal{S} \rightarrow \mathbbm{I} | q(\mathcal{s}_1) \neq q(\mathcal{s}_2)$.
-Then no process can distinguish between the two, they are physically indistinguishable therefore $\mathcal{s}_1 = \mathcal{s}_2$.
-\end{justification}
 
@@
 
-When assembling multiple state variables into a state vector, it is important to understand how they relate to each other.
-Consider the orbital of an electron in a hydrogen item, which is identified by the quantum numbers $n$, $l$, $m$ and $s$.
-For each combination of $n$, $l$ and $m$, the spin $s$ can have two values.
-The choice of spin is independent from the rest.
-The choice of $l$, though, depends on the choice of $n$: for $n=1$ only $l=s$ is available; for $n=2$ we can choose $l=s$ or $l=p$.
-The choices are not independent.
-That is: two or more variables are independent if there always exists a state for any possible combination, if the total number of states is the cartesian product of the possibilities of each variable.
-
 For discrete state variables, the total number of states is the product of the possibilities of each independent state variable.
@@
 We will not study this case, but we do note that, if such a metric exist, it depends on the system and on the choice of state variables.
-
-\begin{defn}\label{independent_state_variables}
-Let $q_1 : \mathcal{S} \rightarrow \mathbbm{I}_1$ and $q_2 : \mathcal{S} \rightarrow \mathbbm{I}_2$ be two state variables.
-They are said \emph{independent} if $q_1^{-1}(i_1)\cap q_2^{-1}(i_2) \neq 0 \forall i_1 \in \mathbbm{I}_1, i_2 \in \mathbbm{I}_2$.
-\end{defn}
-
-\begin{justification}
-If $q_1$ are $q_2$ independent state variables we can modify a state such that only one variable is affected. TODO finish
-\end{justification}
 
@@
 Not going to do that.
-
-\begin{defn}\label{continuous_state_space}
-The space state $\mathcal{Q}$ for a system fully identified by a set of $n$ independent continuous state variables is a manifold of dimension $n$.
-\end{defn}
-
-\begin{justification}
-TODO: set of state variables locally define a chart, isomorphic to $\mathbbm{R}$.
-\end{justification}
 

### 2016-03-16 2015CompleteDerivation/completeDerivation.tex
Adding justification for topological space

@@
 \begin{justification}
-TODO: all possible states form a set.
-All possible set of states associated with the outcome of the process form a topology.
-Verify requirements on entire set, empty set, intersections and unions.
-Distinguishability of two elements in terms of process outcome leads to Hausdorff.
-Processes have countable possible outcome means base for topology must be countable, or we would have elements that are not distinguishable, therefore second countability.
+We claim $\mathcal{S}$ is a set. Each state is well defined as it is physically distinguishable. The collection of all possible states form a set.
+
+We claim $\mathcal{S}$ has a topology $\mathcal{T}$.
+Consider the set of all possible physical outcomes associated with all physical processes.
+Each possible outcome is associated with a set of states that are compatible with that outcome.
+Let $\mathcal{T}$ be the set of all sets associated with all physical outcomes.
+$\mathcal{S} \in \mathcal{T}$ and is associated with unavoidable outcomes.
+$\varnothing \in \mathcal{T}$ and is associated with impossible outcomes.
+Let $V_1, V_2 \in \mathcal{T}$.
+Then, by definition, there exist a process $P_1$ that admits $V_1$ as an outcome and a process $P_2$ that admits $V_2$ as an outcome.
+Consider the process $P$ that combines the outcomes of $P_1$ and $P_2$.
+This always exist physically as we can prepare the same state multiple times and let it interact with each process separately.
+$P$ will have a possible outcome $V$ corresponding to the case where $P_1$ gave outcome $V_1$ and $P_2$ gave outcome $V_2$.
+The states compatible with $V$ must be in both $V_1$ and $V_2$, that is $V = V_1 \cap V_2$.
+Therefore $\mathcal{T}$ is closed under intersection.
+Let $V_1, V_2 \in \mathcal{T}$.
+If they are physically distinguishable, then there exists a physical process $P$ that admits both as outcomes.
+Given $P$, we can always construct the process $P_0$ that combines $V_1, V_2$ into a single outcome $V$ by "forgetting" which of the two was given.
+The states compatible with $V$ must be in either $V_1$ or $V_2$, that is $V = V_1 \cup V_2$.
+Therefore $\mathcal{T}$ is closed under union.
+$\mathcal{T}$ is a topology by definition.
+
+We claim that $\mathcal{S}$ is Hausdorff.
+Let $\mathcal{s_1}, \mathcal{s_2} \in \mathcal{S}$.
+As states are physically distinguishable, there must exist a physical process with two outcomes $V_1, V_2 \in \mathcal{T}$ for which $s_1 \in V_1, s_2 \in V_2, V_1 \cap V_2 = \varnothing$.
+$\mathcal{S}$ is Hausdorff by definition.
+
+We claim that $\mathcal{S}$ is second countable.
+Consider the set of outcomes of a process.
+As each is identified by a finite number of digits.
+The set of outcomes is countable.
+This remains true when combining processes: the number of process we can combine is countable as so it is the number of times we can prepare the same state and the intersections of countable many sets is still countable.
+Consider now the set, not necessarily countable, of all processes $\mathcal{P}$ that cover the whole space (i.e.
+union of outcomes is the whole space).
+Consider now the super process $P_s$ that is the combination of all elements in $\mathcal{P}$.
+This represents the best possible identification of each state, as no other process can further refine the set of possibilities.
+As states are physically distinguishable, their best possible identification must be reachable by a physical process.
+$P_s$ is itself a physical process.
+The set of outcomes represent a basis for the topology because given the outcome of this super process we can construct the outcome of all other less refined processes.
+The outcomes of $P_s$ process are countable.
+The topology has a countable basis.
+$\mathcal{S}$ is second countable by definition.
+
 \end{justification}

### 2016-03-18 2015CompleteDerivation/completeDerivation.tex
Incorporated Christine feedback. Final pass up to state labels (excluded)

@@
 
-% Status: Conceptual work done. Text ready for review. Incorporated Dave and Isaac feedback
+% Status: Conceptual work done. Text ready. Incorporated Dave, Isaac and Christine feedback
+
+Our first task is to extend the standard picture of system plus environment by differentiating between the state of the system (i.e.
+the aspects under study) from the unstated part of the system (i.e.
+the aspects missing from our description).
+
+We will assume that the state evolves according to a deterministic and reversible law (i.e.
+for each present state there is one and only one future state).
+While the unstated part does not influence the state evolution, it constrains what states are available and what type of description can or cannot be given to the system.
+This understanding is what will allow us to treat classical, statistical and, later, quantum states within the same framework.
+
+\subsection{States and their evolution}
 
@@
 
-In this context, we define a process \emph{deterministic} if the state at a given time uniquely identifies states at future times, and \emph{reversible} if it identifies states at past times.
+In this context, we define the evolution of the system \emph{deterministic} if the state at a given time uniquely identifies states at future times, and \emph{reversible} if it uniquely identifies states at past times.
 While this is a common enough definition, we need to be clear how this applies to the unstated part.
-That is: the evolution of the unstated part is \emph{always} non-deterministic (and non-reversible) in the sense that the state of the system does not determine its evolution.
+That is, the evolution of the unstated part under the same process is \emph{always} non-deterministic (and non-reversible) in the sense that the state of the system does not determine its evolution.
 For example, suppose we define the state as position and momentum of the center of mass of a cannonball.
-Suppose that the process is deterministic on that state.
+Suppose that the evolution is deterministic on that state.
 What does it tell us about the cannonball temperature, or about the motion of each of its atoms?
@@
 Does it mean the unstated part is always evolving chaotically?
-Not at all: the temperature may remain constant throughout the motion of the cannonball.
+Not at all.
+The temperature may remain constant throughout the motion of the cannonball.
 Yet we wouldn't know, since we are not studying it: the motion is deterministic and/or reversible only as far as the state is concerned.
@@
 As it is an assumption, we first need to discuss when it is valid.
-More specifically, we need to understand that the non-deterministic/non-reversible evolution of the unstated part plays as much as a fundamental role as the deterministic/reversible evolution of the state.
-In fact: the non-deterministic part contributes in determining what states are available to the system.
+More specifically, we need to understand that the non-deterministic/non-reversible evolution of the unstated part plays as much of a fundamental role as the deterministic/reversible evolution of the state.
+In fact, the non-deterministic part contributes in determining what states are available to the system.
 
-Suppose we study the motion of a cannonball, its state under gravitational and inertial forces will be properly described by the position and momentum of the center of mass.
-While light and air molecules scatter off its surface randomly, its trajectory is not greatly affect as it is a massive rigid body.
+Suppose we study the motion of a cannonball; its state under gravitational and inertial forces will be properly described by the position and momentum of the center of mass.
+While light and air molecules may scatter off its surface unpredictably, its trajectory is not greatly affected as it is a massive rigid body.
 Suppose we study the motion of a small particle, small enough that the random scattering does influence the trajectory and it undergoes Brownian motion: its state will be a probability distribution for position and momentum of the center of mass.
-Gravitational and inertial forces have not changed, yet the set of states have.
-In other words: the set of states must also be closed under non-deterministic evolution.
+Gravitational and inertial forces have not changed, yet the set of states has.
+In other words, the set of states must also be closed under non-deterministic evolution.
 If the Brownian motion is not negligible, we do not end in a well defined position/momentum pair, even if we start from one.
-%Reworded this for Isaac
 
@@
 As we increase the temperature of the air around the book, its motion remains unaffected until, at some point, the book burns.
-Clearly, the non-deterministic evolution has brought one of the states outside the set, to the point that the system is no longer recognizable.
+Clearly, the non-deterministic evolution has pushed one of the states outside the set, to the point that the system is no longer recognizable.
 
@@
 Even in this case, the state can be deterministic and reversible.
-That is: given the distribution at one time we can determine the distribution at future times.
+That is, given the distribution at one time we can determine the distribution at future times.
 The shape and the parameters of the distribution can be deterministic, even if the trajectories of the parts are not (those fall within the unstated part).
-In fact: we \emph{cannot} assume trajectories and states are always defined at all for the unstated part, as this includes also the unknown unknowns.
-Will return to this aspect when discussing quantum systems.
+In fact, we \emph{cannot} assume trajectories and states are always defined at all for the unstated part, as this includes also the unknown unknowns.
+We will return to this aspect when discussing quantum systems.
 
-It should be also clear that what constitute state and unstated part does not depend only on the system under study, but also on the processes we are considering.
+It should also be clear that what constitutes state and unstated part does not depend only on the system under study, but also on the processes we are considering.
 In some circumstances, the chemical composition of a fluid may be relevant, in others it may not.
@@
 As with all assumptions, we should also ask whether it is necessary.
-That is: could we define a set of physically distinguishable states and yet have no deterministic and reversible processes defined on them?
+That is, could we define a set of physically distinguishable states and yet have no deterministic and reversible processes defined on them?
 The claim is that this assumption is indeed needed, as without it we cannot properly define states or write useful physics laws.
@@
 Fourth, to be able to ascribe a property to a system we need to claim that, at least for a finite interval of time, the system either held or did not hold such property.
-That is: there is a deterministic and reversible process for that finite period of time for such property.
+That is, there is a deterministic and reversible process for that finite period of time for such property.
 
 This link between state definition and deterministic processes should not be too surprising as the state, in the context of thermodynamics and systems theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system}.
-As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can be indeed calculated, measured and prepared.
+As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can indeed be calculated, measured and prepared.
 We can also describe the evolution of each element of the distribution provided that: we have a way to isolate it and study it under deterministic and reversible motion (so that we can define microstates); the non-deterministic motion does not alter the system (the set of microstates is preserved by the evolution).
 
-As with all assumptions, we should stress that it's an idealization: it can never be completely achieved in practice.
+As with many assumptions, we should stress that it's an idealization: it can never be completely achieved in practice.
 A system can be prepared or measured up to a certain level of precision.
@@
 If a particular set of states does not satisfy deterministic and reversible evolution under certain conditions, what we do is to keep at it until we find a set that does.
-That is: we work to restore the assumption.
+That is, we work to restore the assumption.
 Finding new sets of states with new laws of evolution is, in fact, what leads to new physics.
@@
 
-% Status: concepts still need to be finished. Started to work on linking topology to physically distinguishable states.
+% Status: First subsection. Conceptual work done. Text ready. Incorporated Christine feedback
+% Second subsection. Conceptual work almost done. Need to finish the text
 
@@
 
-%Things to add: discrete sets evolved continuous = identity map. Subdivide finite interval, same permutation at each time step, exclude odd perms, third-odd perms, ...
-
 We now proceed to characterize states and their being physically distinguishable in more precise terms so that we can capture their description mathematically.
 What we'll see is that the set of all physical processes that can be used to gain information about the system, that can be used to perform a measurement, induce a topology on the set of states that is Hausdorff and second countable.
-That is: physical distinguishability  is mathematically captured by topological distinguishability.
+That is, physical distinguishability is mathematically captured by topological distinguishability.
 Deterministic and reversible evolution will then preserve the topology, and they will be mathematically captured by self-homeomorphisms.
@@
 
-As the term measurement has become a particularly loaded, let's first characterize what we mean by physical distinguishability in our context.
+As the term "measurement" has become particularly loaded, let's first characterize what we mean by physical distinguishability in our context.
 
@@
 In principle, any external process that has a correlation with the states under study can be used to perform a measurement, and any measurement is based on such a process.
-That is: for us a measurement is simply a physical process that we can use to distinguish states.
-There is no particular role played by the observer, except maybe setting up such a process and forcing the system under study to interact with it.
+That is, for us a measurement is simply a physical process that we can use to distinguish states.
+Setting up an experiment means choosing a particular process with desired outcomes and forcing the system under study to interact with it.
+After that, there is no special role played by the "observer" in making outcomes come about.
 
 This external process may be quite complicated: when a particle enters a calorimeter, a shower of particles is produced, photons are captured and are directed to photomultipliers, a current is read out, the current is then digitized, and so on.
-Some processes interfere with the system, they affect its dynamics, and other are destructive, the system no longer can be described by the original set of states.
+Some processes interfere with the system, they affect its dynamics, and others are destructive, the system no longer can be described by the original set of states.
 A tracking chamber is an example of the first (the magnetic field curves the motion of a charged particle); burning a substance to determine its caloric content is an example of the latter.
@@
 But in the end, the conceptual description remains the same: each possible outcome of the process will be associated with a set of states consistent with that outcome (e.g.
-if the cannonball deformed the ground by this amount than the final state its kinetic energy at impact was within this range).
+if the cannonball deformed the ground by this amount then its kinetic energy at impact was within this range).
 
@@
 For example, if the electron follows a certain path, then its state is among the ones that has spin $+1/2$.
-However precise our measurements are, each outcome is expressed by a finite set of digits, therefore the set of outcomes is countable.
-Note that different outcomes for the same process can overlap.
+However precise our measurements are, each outcome is expressed by a finite set of digits; therefore, the set of outcomes is countable.\footnote{The information provided by the process as measured by Shannon's entropy is finite.} Note that different outcomes for the same process can overlap.
 For example, $4.12 \pm 0.05$ cm and $4.13 \pm 0.05$ cm are both legitimate possible outputs of the same measurement device.
 But since all states must be distinguishable, given two arbitrary states there must be a process precise enough to tell them apart.
-That is: the potential outcomes associated with the two states do not overlap.
+That is, the potential outcomes associated with the two states do not overlap.
 For example, $4.12 \pm 0.0005$ cm and $4.13 \pm 0.0005$ cm do not overlap anymore.
@@
 We can also conceptually combine two different processes into a single one.
-That is: having a way to measure quantity $x$ and a way to measure quantity $y$ gives us a way to measure the combination $(x,y)$.
-Formally, the outcomes of the combined process will be the intersections of each pair of outcomes of the original processes.
+That is, having a way to measure quantity $x$ and a way to measure quantity $y$ gives us a way to measure the combination $(x,y)$.
+For ensembles, if we can measure the marginal distribution $\rho_x(x)$ and the marginal distribution $\rho_y(x)$, we know the joint distribution $\rho(x,y)$ has to be compatible with both.
+Note, though, that this does not provide a way to fully measure $\rho(x,y)$ as we know nothing about the correlation between the two variables.\footnote{The quantum case is similar: we measure marginal distributions and rule out states that are incompatible with them.} Formally, the states compatible with the outcomes of the combined process will be the intersections of the states compatible with each pair of outcomes of the original processes.
 
 We can also coarsen a single process.
-That is: having a way to measure $(x,y)$ gives us a way to measure $x$ alone (or any $f(x,y)$).
+That is, having a way to measure $(x,y)$ gives us a way to measure $x$ alone (or any $f(x,y)$).
 Formally, the outcomes of the coarsened process are given by performing the union of some outcomes of the original process.
@@
 This model maps very naturally to a topological space.
-The states are elements of a set and the physical outcome provides a topology on that set.
+The states are elements of a set and the physical outcomes provide a topology on that set.
 The fact that two elements of the set can be distinguished requires the space to be Hausdorff.
-The fact that the potential outcome of a process are countable requires the space to be second countable.
+The fact that the potential outcomes of a process are countable requires the space to be second countable.
 
@@
 The set of outcomes is countable.
-This remains true when combining processes: the number of process we can combine is countable as so it is the number of times we can prepare the same state and the intersections of countable many sets is still countable.
-Consider now the set, not necessarily countable, of all processes $\mathcal{P}$ that cover the whole space (i.e.
+This remains true when combining processes: the number of processes we can combine is finite as so it is the number of times we can prepare the same state and the intersections of countable many sets is still countable.
+Consider now the set, not necessarily finite, of all processes $\mathcal{P}$ that cover the whole space (i.e.
 union of outcomes is the whole space).
@@
 As states are physically distinguishable, their best possible identification must be reachable by a physical process.
-$P_s$ is itself a physical process.
+The outcomes of $P_s$ are countable.
 The set of outcomes represent a basis for the topology because given the outcome of this super process we can construct the outcome of all other less refined processes.
-The outcomes of $P_s$ process are countable.
 The topology has a countable basis.
@@
 
+\end{justification}
+
+Note that the arguments that lead to the topological space had nothing to do with states per se, just that they are physically distinguishable.
+States, though, are not the only objects with that property.
+In fact: any element of a set of physical object (e.g.
+forces, physical properties such as mass or charge, time) needs to be physically distinguishable to be well defined.
+We can generalize the above justification: any set of physically distinguishable elements is a topological space.
+
+\begin{defn}\label{topologically_distinguishable}
+	Any set $\mathcal{S}$ of physically distinguishable elements is a Hausdorff and second countable topological space.
+\end{defn}
+
+\begin{justification}
+	Same justification as in \ref{statedef} with "state" replaced by "element" of the set $\mathcal{S}$.
 \end{justification}
@@
 $f(U)$ is an open set in the state space topology by definition.
+\end{justification}
+
+Again, we note that the arguments that lead to an open map had nothing to do with states per se, just that the relationship is between physically distinguishable objects.
+We can then generalize the above justification.
+
+\begin{defn}\label{openmap}
+	A map $f:\mathcal{S_1} \rightarrow \mathcal{S_2}$ between two sets of physically distinguishable elements $\mathcal{S_1}$ and $\mathcal{S_2}$ is an open map.
+\end{defn}
+
+\begin{justification}
+	Same justification as in \ref{detrevmap} with "initial states" and "final states" replaced by "elements" of $\mathcal{S_1}$ and $\mathcal{S_2}$ respectively.
 \end{justification}
@@
 So we simplify our problem and study an idealized case: one where the components are homogeneous and infinitesimal.
-What we'll find is that under the additional assumption that the system is reducible (i.e.
-its state is equivalent to the state of the parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian particle mechanics.
+What we'll find is that under the additional assumption that the system is infinitesimally reducible (i.e.
+its state is equivalent to the state of its parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian particle mechanics.
 
@@
 The notion that the system is homogeneous means that the states of the composite and of the parts are not unrelated: they are made of the same material.\footnote{Whether a system is homogeneous depends on context (e.g.
-air can be thought as homogeneous if the mixture of gases does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is maintained by by time evolution.} In fact, the state space $\mathcal{C}$ of all systems composed of such material will contain the state of system and all its part (i.e $\mathcal{c}, \mathcal{c}_1, \mathcal{c}_2 \in \mathcal{C}$).
+air can be thought as homogeneous if the mixture of gases does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is maintained by by time evolution.} In fact, the state space $\mathcal{C}$ of all systems composed of such material will the state of the system as well as the states of all its part (i.e $\mathcal{c}, \mathcal{c}_1, \mathcal{c}_2 \in \mathcal{C}$).
 Since combining any two systems made of a homogeneous material will always give us a system made of the same homogeneous material (e.g.

### 2016-03-20 2015CompleteDerivation/completeDerivation.tex
Starting to prove manifold for continuous time evolution

@@
 They will determine the number of connected components of the manifold.
-Once we introduce a continuous parameter for time evolution, though, these also become irrelevant as continuous time evolution cannot move states across disconnected components.
+Once we introduce a continuous parameter for time evolution, though, these also become irrelevant as continuous time evolution requires continuous trajectories that cannot move states across disconnected components.
+This justifies the special interest in path connected manifolds, as these where trajectories for deterministic and reversible continuous evolution live.
 
-\begin{defn}\label{continuous_state_space}
-	The space state $\mathcal{Q}$ for a system fully identified by a set of $n$ independent continuous state variables is a manifold of dimension $n$.
-\end{defn}
+\begin{prop}\label{continuous_state_space}
+	Let $\mathcal{s} \in \mathcal{S}$ a state within a state space.
+The set of states $\mathcal{S}'\subseteq\mathcal{S}$ potentially reachable from $\mathcal{s}$ by deterministic and reversible continuous evolution is a path connected manifold of dimension equal to the number of independent continuous state variables necessary to identify it.
+\end{prop}
 
-\begin{justification}
-	TODO: set of state variables locally define a chart, isomorphic to $\mathbbm{R}$.
-\end{justification}
+\begin{proof}
+	We claim $\mathcal{S}'$ is a manifold.
+Let $\mathcal{s} \in \mathcal{S}$.
+Let $n$ be the number of independent continuous variables needed to identify $\mathcal{s}$.
+There exists a neighborhood $U$ around $\mathcal{s}$ where we have $(q_1,...,q_n):U\rightarrow \mathbbm{R}^n$.
+This map is a bijection as $s$ is identified by those variables.
+$\mathcal{S}$ is homeomorphic to $\mathbbm{R}^n$ around  $\mathcal{s}$.
+Let $\mathcal{S}'$ be the set of all states potentially reachable by deterministic and reversible evolution from $\mathcal{s}$.
+Deterministic and reversible evolution is a homeomorphism between initial and final states.
+$\forall \hat{\mathcal{s}} \in \mathcal{S}'$ the evolution of $U$ will contain $\mathcal{s}$ and will be isomorphic to $\mathbbm{R}^n$.
+$\mathcal{S}'$ equipped with the subspace topology is a Hausdorff, second countable topological space everywhere homeomorphic to $\mathbbm{R}^n$.
+$\mathcal{S}'$ is a manifold of dimension $n$.
+	TODO: proof path-connectedness
+\end{proof}
 

### 2016-03-20 2015CompleteDerivation/completeDerivation.tex
Maps are continuous not open

@@
 With our state space defined, deterministic and reversible evolution corresponds to a bijective map between initial and final states.
-Moreover, a physical process that can distinguish initial states is also a physical process that can distinguish final states.
+Moreover, a physical process that can distinguish final states is also a physical process that can distinguish initial states.
 That is: the topology is mapped and preserved by the deterministic and reversible evolution (i.e.
@@
 
+We claim $f$ is continuous.
+Let $U \subseteq \mathcal{S}$ represent an outcome of a process $P$ on the final states.
+$U$ is an open set in the (final) state space topology by definition.
+Consider the process $P_0$ that evolves the initial states and then distinguishes the final state with $P$.
+$P_0$ is a process that distinguishes initial states.
+The set of initial states compatible with $U$ are $f^{-1}(U)$.
+$f^{-1}(U)$ is an open set in the (initial) state space topology by definition.
+$f$ is a continuous map.
+
 We claim that $f$ is a bijection.
@@
 
-We claim $f$ is an open map.
-Let $U \subseteq \mathcal{S}$ represent an outcome of a process.
-$U$ is an open set in the state space topology by definition.
-Given the deterministic and reversible evolution $f$, an outcome on the initial states can be used to distinguish final states.
-The same outcome associated with $U$ as initial states will also be associated with $f(U)$ as final states.
-$f(U)$ is associated with an outcome.
-$f(U)$ is an open set in the state space topology by definition.
+We claim that $f$ is a self-homeomorphism as it is a continuous bijection.
 \end{justification}
 
-Again, we note that the arguments that lead to an open map had nothing to do with states per se, just that the relationship is between physically distinguishable objects.
+Again, we note that the arguments that lead to the existence of a continuous map had nothing to do with states per se, just that the relationship is between physically distinguishable objects.
 We can then generalize the above justification.
@@
 \begin{defn}\label{openmap}
-	A map $f:\mathcal{S_1} \rightarrow \mathcal{S_2}$ between two sets of physically distinguishable elements $\mathcal{S_1}$ and $\mathcal{S_2}$ is an open map.
+	A map $f:\mathcal{S_1} \rightarrow \mathcal{S_2}$ between two sets of physically distinguishable elements $\mathcal{S_1}$ and $\mathcal{S_2}$ is a continuous map.
 \end{defn}
@@
 \begin{justification}
-	Same justification as in \ref{detrevmap} with "initial states" and "final states" replaced by "elements" of $\mathcal{S_1}$ and $\mathcal{S_2}$ respectively.
+	Same justification for existence and continuity as in \ref{detrevmap} with "initial states" and "final states" replaced by "elements" of $\mathcal{S_1}$ and $\mathcal{S_2}$ respectively.
 \end{justification}
+
+The generality of this result explains why in physics one always assumes functions to be well behaved.
+This is not a matter of practical convenience or usefulness, but it's because continuous functions represent actual physically distinguishable objects.
 

### 2016-03-25 2015CompleteDerivation/completeDerivation.tex
Adding mention of statistical and systematic uncertainty on physical processes

@@
 Light will scatter off of it; as it lands the ground will be deformed and the impact will make the temperature rise.
-Those external physical processes, which happen no matter what we do, can be used to distinguish the motion of the cannonball as their outcomes are correlated.\footnote{Note that by requiring a correlation we are implying that measurements are repeatable.
-That is, for similarly prepared states we get similar outcomes.} Therefore we can learn the position by looking at the reflected light, and learn the final kinetic energy by looking at the deformation of the ground.
+Those external physical processes, which happen no matter what we do, can be used to distinguish the motion of the cannonball as their outcomes are correlated.
+Therefore we can learn the position by looking at the reflected light, and learn the final kinetic energy by looking at the deformation of the ground.
 In principle, any external process that has a correlation with the states under study can be used to perform a measurement, and any measurement is based on such a process.
@@
 A tracking chamber is an example of the first (the magnetic field curves the motion of a charged particle); burning a substance to determine its caloric content is an example of the latter.
-In these cases, special care is needed to make sure to link the outcomes with the original state.
-But in the end, the conceptual description remains the same: each possible outcome of the process will be associated with a set of states consistent with that outcome (e.g.
-if the cannonball deformed the ground by this amount then its kinetic energy at impact was within this range).
+Therefore intimate knowledge of the process is always needed to ensure that one makes the proper link between outcomes and the \emph{original} states, and properly accounts for systematic uncertainties that would skew that link.
 
-Each physical process, then, can be thought as having a set of potential outcomes.
-To each potential outcome we can associate a set of states that are consistent with it.
-For example, if the electron follows a certain path, then its state is among the ones that has spin $+1/2$.
-However precise our measurements are, each outcome is expressed by a finite set of digits; therefore, the set of outcomes is countable.\footnote{The information provided by the process as measured by Shannon's entropy is finite.} Note that different outcomes for the same process can overlap.
+Repeatability is also fundamental.
+First, to make sure the process is indeed correlated to the states.
+Second, because "a single take does not a measurement make".
+One has to gather enough statistics.
+Note that the number of takes influences the outcomes: with greater statistics the precision and number of distinguishable cases increases.
+Therefore the processes, as we define them, may require repeated interactions with similarly prepared states.
+They may even be a combination of different kinds of interactions that, taken all together, create a set of distinguishable outcomes.
+But in the end, however complicated it is, the conceptual model remains the same: each process has a set of possible outcomes, and each possible outcome will be associated with a set of states consistent with that outcome.
+For example, if the cannonball deformed the ground by this amount then its kinetic energy at impact was within this range; if the electron follows a certain path, then its state is among the ones that has spin $+1/2$.
+However precise our measurements are, we can only gather a finite amount of statistics and each outcome is expressed by a finite set of digits; therefore, the set of outcomes is countable.\footnote{The information provided by the process as measured by Shannon's entropy is finite.} Note that different outcomes for the same process can overlap.
 For example, $4.12 \pm 0.05$ cm and $4.13 \pm 0.05$ cm are both legitimate possible outputs of the same measurement device.

### 2016-03-25 2015CompleteDerivation/completeDerivation.tex
Different argument for second countable

@@
 
-We claim $\mathcal{S}$ has a topology $\mathcal{T}$.
+We claim $\mathcal{S}$ has a topology $\mathsf{T}$.
 Consider the set of all possible physical outcomes associated with all physical processes.
 Each possible outcome is associated with a set of states that are compatible with that outcome.
-Let $\mathcal{T}$ be the set of all sets associated with all physical outcomes.
-$\mathcal{S} \in \mathcal{T}$ and is associated with unavoidable outcomes.
-$\varnothing \in \mathcal{T}$ and is associated with impossible outcomes.
-Let $V_1, V_2 \in \mathcal{T}$.
+Let $\mathsf{T}$ be the set of all sets associated with all physical outcomes.
+$\mathcal{S} \in \mathsf{T}$ and is associated with unavoidable outcomes.
+$\varnothing \in \mathsf{T}$ and is associated with impossible outcomes.
+Let $V_1, V_2 \in \mathsf{T}$.
 Then, by definition, there exist a process $P_1$ that admits $V_1$ as an outcome and a process $P_2$ that admits $V_2$ as an outcome.
@@
 The states compatible with $V$ must be in both $V_1$ and $V_2$, that is $V = V_1 \cap V_2$.
-Therefore $\mathcal{T}$ is closed under intersection.
-Let $V_1, V_2 \in \mathcal{T}$.
+Therefore $\mathsf{T}$ is closed under intersection.
+Let $V_1, V_2 \in \mathsf{T}$.
 If they are physically distinguishable, then there exists a physical process $P$ that admits both as outcomes.
@@
 The states compatible with $V$ must be in either $V_1$ or $V_2$, that is $V = V_1 \cup V_2$.
-Therefore $\mathcal{T}$ is closed under union.
-$\mathcal{T}$ is a topology by definition.
+Therefore $\mathsf{T}$ is closed under union.
+$\mathsf{T}$ is a topology by definition.
 
@@
 Let $\mathcal{s_1}, \mathcal{s_2} \in \mathcal{S}$.
-As states are physically distinguishable, there must exist a physical process with two outcomes $V_1, V_2 \in \mathcal{T}$ for which $s_1 \in V_1, s_2 \in V_2, V_1 \cap V_2 = \varnothing$.
+As states are physically distinguishable, there must exist a physical process with two outcomes $V_1, V_2 \in \mathsf{T}$ for which $s_1 \in V_1, s_2 \in V_2, V_1 \cap V_2 = \varnothing$.
 $\mathcal{S}$ is Hausdorff by definition.
@@
 This remains true when combining processes: the number of processes we can combine is finite as so it is the number of times we can prepare the same state and the intersections of countable many sets is still countable.
-Consider now the set, not necessarily finite, of all processes $\mathcal{P}$ that cover the whole space (i.e.
-union of outcomes is the whole space).
-Consider now the super process $P_s$ that is the combination of all elements in $\mathcal{P}$.
-This represents the best possible identification of each state, as no other process can further refine the set of possibilities.
-As states are physically distinguishable, their best possible identification must be reachable by a physical process.
-The outcomes of $P_s$ are countable.
-The set of outcomes represent a basis for the topology because given the outcome of this super process we can construct the outcome of all other less refined processes.
-The topology has a countable basis.
-$\mathcal{S}$ is second countable by definition.
+Consider a bases for the topology.
+Suppose it is of cardinality greater than countability.
+Then no physical process could ever distinguish between the elements of the basis.
+The basis, and therefore the outcomes, would not be physically distinguishable.
+The basis for the topology cannot be of cardinality greater than countability.
+By contradiction, $\mathcal{S}$ must be second countable.
 

### 2016-03-25 2015CompleteDerivation/completeDerivation.tex
Minor corrections and changes in notation

@@
 \begin{defn}\label{detrevmap}
-A deterministic and reversible evolution map is a self-homeomorphism on the state space $f:\mathcal{S} \rightarrow \mathcal{S}$.
+A deterministic and reversible evolution map is a self-homeomorphism on the state space $\mathcal{T}_{\Delta t}:\mathcal{S} \rightarrow \mathcal{S}$.
 \end{defn}
@@
 \begin{justification}
-We claim that $f$ exists. The system is deterministic: given an initial state $\mathcal{s} \in \mathcal{S}$ there exists a well defined final state $f(\mathcal{s})=\hat{\mathcal{s}} \in \mathcal{S}$.
+We claim that $\mathcal{T}_{\Delta t}$ exists.
+The system is deterministic: given an initial state $\mathcal{s} \in \mathcal{S}$ there exists a well defined final state $\mathcal{T}_{\Delta t}(\mathcal{s})=\hat{\mathcal{s}} \in \mathcal{S}$.
 
-We claim $f$ is continuous.
+We claim $\mathcal{T}_{\Delta t}$ is continuous.
 Let $U \subseteq \mathcal{S}$ represent an outcome of a process $P$ on the final states.
@@
 $P_0$ is a process that distinguishes initial states.
-The set of initial states compatible with $U$ are $f^{-1}(U)$.
-$f^{-1}(U)$ is an open set in the (initial) state space topology by definition.
-$f$ is a continuous map.
+The set of initial states compatible with $U$ are $\mathcal{T}_{\Delta t}^{-1}(U)$.
+$\mathcal{T}_{\Delta t}^{-1}(U)$ is an open set in the (initial) state space topology by definition.
+$\mathcal{T}_{\Delta t}$ is a continuous map.
 
-We claim that $f$ is a bijection.
-The system is reversible: there exists a map $g:\mathcal{S} \rightarrow \mathcal{S}$ that returns the initial state given the final state.
-$g \circ f = f \circ g = id$ as mapping forward and then backward or backwards and then forward must return the original element.
-$f$ admits $g$ as an inverse.
-$f$ is a bijection
+We claim that $\mathcal{T}_{\Delta t}$ is a bijection.
+The system is reversible: there exists a map $\mathcal{T}_{-\Delta t}:\mathcal{S} \rightarrow \mathcal{S}$ that returns the initial state given the final state.
+$\mathcal{T}_{-\Delta t} \circ \mathcal{T}_{\Delta t} = \mathcal{T}_{\Delta t} \circ \mathcal{T}_{-\Delta t} = id$ as mapping forward and then backward or backwards and then forward must return the original element.
+$\mathcal{T}_{\Delta t}$ admits $\mathcal{T}_{-\Delta t}$ as an inverse.
+$\mathcal{T}_{\Delta t}$ is a bijection
 
-We claim that $f$ is a self-homeomorphism as it is a continuous bijection.
+We claim that $\mathcal{T}_{\Delta t}$ is a self-homeomorphism as it is a continuous bijection.
 \end{justification}
 
-Again, we note that the arguments that lead to the existence of a continuous map had nothing to do with states per se, just that the relationship is between physically distinguishable objects.
+Again, we note that the arguments that lead to continuity had nothing to do with states per se, just that the relationship is between physically distinguishable objects.
 We can then generalize the above justification.
 
-\begin{defn}\label{openmap}
+\begin{defn}\label{continuous_map}
 	A map $f:\mathcal{S_1} \rightarrow \mathcal{S_2}$ between two sets of physically distinguishable elements $\mathcal{S_1}$ and $\mathcal{S_2}$ is a continuous map.
@@
 \begin{justification}
-	Same justification for existence and continuity as in \ref{detrevmap} with "initial states" and "final states" replaced by "elements" of $\mathcal{S_1}$ and $\mathcal{S_2}$ respectively.
+	Same justification for continuity as in \ref{detrevmap} with "initial states" and "final states" replaced by "elements" of $\mathcal{S_1}$ and $\mathcal{S_2}$ respectively.
 \end{justification}

### 2016-03-25 2015CompleteDerivation/completeDerivation.tex
Discussion on discontinuities

@@
 The generality of this result explains why in physics one always assumes functions to be well behaved.
-This is not a matter of practical convenience or usefulness, but it's because continuous functions represent actual physically distinguishable objects.
+As the result was derived from our notion of physical distinguishability, this is not a matter of practical convenience or usefulness.
+Suppose we were able to prepare a force field that was zero everywhere in space except at a single point.
+This gives us a way to tag a specific point.
+A finite precision measurement of the force provides us infinite precision of space, which we ruled out.\footnote{This assumes that we are able to position the probe perfectly, which we could do if we able to manipulate forces at that precision.} This is what the math is telling us, that if we said before that outcomes can't distinguish isolated points (i.e.
+the topology is second countable) then neither can maps.
+It is physical consistency that limits us to continuous functions.
 
-%TODO: do continuous function form a manifold Hausdorf and second countable?
+While functions with few discontinuities are useful and used in physics and engineering, they are employed for idealized cases  (e.g.
+a signal change is fast enough, a charge distribution is small enough) that are usually treated differently (i.e.
+propagation in material discontinuities).
+While this may be obvious and intuitive to the physicist, it may be troubling to the mathematician as the proper use of many mathematical techniques requires the inclusion of discontinuous functions.
+This is less of a problem than it would seem at first.
+Once we made sure that the objects and their relationships are physically meaningful, we can extend our mathematical spaces for the purpose of math computations.
+Our physically meaningful continuous function can be expanded into a sum of discontinuous functions.
+One just has to be mindful of the extension and be wary that mathematical results that depend on such extension may or may not be physically meaningful.
 

### 2016-03-25 2015CompleteDerivation/completeDerivation.tex
Reviewed labeling states section

@@
 For example, the orbital of an electron in a hydrogen item is identified by the quantum numbers $n$, $l$, $m$ and $s$.
-We call each of these quantities \emph{state variables} and \emph{state vector} a set of state variables that can identify one and only one state.
+We call each of these quantities \emph{state variables}.
 We call a \emph{possibility} a possible value that can be taken by a state variable.
 We call a state variable \emph{discrete} or \emph{continuous} if the possibilities are integer or real numbers respectively.
+From now on, we are going to consider state spaces whose states can be identified by a finite set of discrete and continuous state variables (i.e.
+the state space is locally isomorphic to $\bigcup\limits_{1 \leq i \leq n} \mathbbm{R}^{n_i}$).\footnote{TODO: Is there a topological space that is Hausdorff and second countable that is not like that?}
 
 We purposely use the term state variables instead of coordinates (even though that's what they are mathematically) as it will create confusion with space-time coordinates.
-We also avoid the term observable or measurable, as not all state variables may be directly physically meaningful (e.g.
+We also avoid the term observable or measurable, as not all state variables may be directly physically tangible (e.g.
 conjugate momentum in a gauge theory).
@@
 \footnote{This is the main reason that a quantity like velocity is not a suitable state variable, as it is  defined over an interval of time.
-Therefore velocity is always physically meaningful but is not a state variable in general, while conjugate momentum is always a state variable but is physically meaningful in general.
-As we'll see later, only once a link is established between them they can both be physically meaningful state variables.}
+Therefore velocity is always physically well defined but is not a state variable in general, while conjugate momentum is always a state variable but is not physically well defined by itself (it requires the vector potential to be specified as well).
+As we'll see later, only once a link is established between them they can both be physically well defined state variables.}
 
 \begin{defn}\label{state_variable}
-	A \emph{state variable} is a homeomphism $q : U \rightarrow \mathbbm{L}$ where $U \subseteq \mathcal{S}$ and $\mathbbm{L}$ is the topological space for the possible values.
+	A \emph{state variable} is a continuous map $q : U \rightarrow \mathbbm{L}$ where $U \subseteq \mathcal{S}$ and $\mathbbm{L}$ is the topological space for the possible values.
 If $\mathbbm{L}\cong \mathbbm{Z}$ the variable is said \emph{discrete}.
 If $\mathbbm{L}\cong \mathbbm{R}$ the variable is said \emph{continuous}.
-	%We also have $\mathcal{s}_1 \neq \mathcal{s}_2 \Leftrightarrow \exists q : U \rightarrow \mathbbm{L} | q(\mathcal{s}_1)\neq q(\mathcal{s}_2) \forall \mathcal{s}_1, \mathcal{s}_2 \in U$.
 \end{defn}
@@
 \begin{justification}
-	We claim $\mathbbm{L}$ is a topological space.
-The collection of all possible values form a set.
-Such set is physically distinguishable.
-The physical outcomes induce a Hausdorff and second countable topology, as for \ref{statedef}.
+	We claim $\mathbbm{L}$ is a topological space. $\mathbbm{L}$ is a set of physically distinguishable possibilities. $\mathbbm{L}$ is a topological space because of \ref{topologically_distinguishable}.
 
-	We claim $q$ is a homeomorphism.
-Elements of $\mathcal{S}$ and $\mathbbm{L}$ are both physically distinguishable therefore a map between them is well defined.
-Let $V \subseteq U$ the set of states associated with a physical outcome.
-$V$ is an open set in the state topology.
-Given the map $q$ an outcome on states can be used to distinguish possibilities of the state variables.
-The same outcome is associated with $g(V)$.
-$g(V)$ is an open set in the topology of possible values.
-
-	%Let $\mathcal{s}_1, \mathcal{s}_2 \in \mathcal{S}$.
-Suppose $\mathcal{s}_1 \neq \mathcal{s}_2$.
-$\mathcal{s}_1$ and $\mathcal{s}_2$ are physically distinguishable.
-There exists a physical process with different outcomes.
-Let $q : \mathcal{S} \rightarrow \mathbbm{I}$ be the state variable associated to that process.
-Then $q(\mathcal{s}_1)\neq q(\mathcal{s}_2)$.
-Now suppose $\nexists q : \mathcal{S} \rightarrow \mathbbm{I} | q(\mathcal{s}_1) \neq q(\mathcal{s}_2)$.
-Then no process can distinguish between the two, they are physically indistinguishable therefore $\mathcal{s}_1 = \mathcal{s}_2$.
+	We claim $q$ is a continuous map. $q$ is a map between two sets of physically distinguishable elements. $q$ is continuous because of \ref{continuous_map}.
 \end{justification}
 
-When assembling multiple state variables into a state vector, it is important to understand how they relate to each other.
+When combining multiple state variables, it is important to understand how they relate to each other.
 Consider the orbital of an electron in a hydrogen item, which is identified by the quantum numbers $n$, $l$, $m$ and $s$.
@@
 That is: two or more variables are independent if there always exists a state for any possible combination, if the total number of states is the cartesian product of the possibilities of each variable.
+We call \emph{state vector} a collection of independent state variables that fully identify a state.
 
@@
 \end{defn}
-
-\begin{justification}
-	TODO What is there to justify?
-\end{justification}
 
@@
 Once we introduce a continuous parameter for time evolution, though, these also become irrelevant as continuous time evolution requires continuous trajectories that cannot move states across disconnected components.
-This justifies the special interest in path connected manifolds, as these where trajectories for deterministic and reversible continuous evolution live.
+This justifies the special interest in path connected manifolds, as this is where trajectories for deterministic and reversible continuous evolution live.
 
@@
 Deterministic and reversible evolution is a homeomorphism between initial and final states.
-$\forall \hat{\mathcal{s}} \in \mathcal{S}'$ the evolution of $U$ will contain $\mathcal{s}$ and will be isomorphic to $\mathbbm{R}^n$.
+$\forall \hat{\mathcal{s}} \in \mathcal{S}'$ the evolution of $U$ will contain $\hat{\mathcal{s}}$ and will be isomorphic to $\mathbbm{R}^n$.
 $\mathcal{S}'$ equipped with the subspace topology is a Hausdorff, second countable topological space everywhere homeomorphic to $\mathbbm{R}^n$.
 $\mathcal{S}'$ is a manifold of dimension $n$.
-	TODO: proof path-connectedness
+
+	We claim $\mathcal{S}'$ is path connected.
+Let $\hat{\mathcal{s}} \in \mathcal{S}'$.
+There exist map $\lambda : [t_0,t_1] \rightarrow \mathcal{S}'$ where $\lambda(t_0)=\mathcal{s}$, $\lambda(t_1)=\hat{\mathcal{s}}$ and $t_0$ and $t_1$ are the initial and final time respectively.
+$\lambda$ is a map from two physically distinguishable quantities, time and states.
+$\lambda$ is continuous by \ref{continuous_map}.
+All $\hat{\mathcal{s}} \in \mathcal{S}'$ are path connected to $\mathcal{s}$.
+$\mathcal{S}'$ is path connected.
 \end{proof}
+
+\subsection{Evolution of state variables}
 

### 2016-03-30 2015CompleteDerivation/completeDerivation.tex
State variable evolution is now ready for review

@@
 For example, if the cannonball deformed the ground by this amount then its kinetic energy at impact was within this range; if the electron follows a certain path, then its state is among the ones that has spin $+1/2$.
+
 However precise our measurements are, we can only gather a finite amount of statistics and each outcome is expressed by a finite set of digits; therefore, the set of outcomes is countable.\footnote{The information provided by the process as measured by Shannon's entropy is finite.} Note that different outcomes for the same process can overlap.
@@
 
-TODO: maybe best to define state vector directly, and state variables as part of state vector?
-Define discrete state variable when co-domain is integers and define continuous state variable when co-domain is real?
-Can we prove a state vector always exists, of real variables?
-For discrete topology, we should always be able to do it.
-In general, state vector defines charts, so we should not have a single chart for the whole state space.
-What does it mean physically?
-Does it matter?
-
-TODO: Still not clear what's the best way to introduce things.
-We do want to show that state variable are maps from states to "labels".
-We want to call discrete variables those that map to integer and continuous variables those that map to real numbers.
-Then we want to concentrate on state space that can be described by a finite number of either discrete variables or continuous variables.
-State vector may not be valid in general: e.g.
-we can't assume we have a single chart for the whole space.
-
-\begin{prop}\label{state_vector}
-TODO: State vector always exists
-\end{prop}
-
-To study time evolution, we need to describe how state variables change under continuous and reversible evolution.
+To study time evolution, we need to describe how state variables change under deterministic and reversible evolution.
 There are two ways to do it, and we'll need both.
 The first approach is to \emph{evolve} the state variables from the initial value $q^i(\mathcal{s})$ to the final value $q^i(\hat{\mathcal{s}})$.
-The result is a trajectory $q^i(t)$ which is especially useful when state variables correspond to physically meaningful quantities.
+The result is a trajectory $q^i(t)=q^i(\lambda(t))$ which is especially useful when state variables correspond to physically meaningful quantities.
 For example, we track how the temperature or the pressure of an ideal gas changes.
@@
 The idea is to keep the connection to the initial state by labeling the future state, instead by the future values of $q^i$, by the original one.
+For example, the evolved state variable will not tell us the current value of pressure, but the one for the initial conditions.
 So we introduce a new set of variables for which $\hat{q}^i(\hat{\mathcal{s}})=q^i(\mathcal{s})$, which we can always do as the evolution is deterministic and reversible.
-The level sets of $\hat{q}^i$ are the evolved level sets of $q^i$, which allows us to study how groups of states evolve in time.
-Moreover, once we identify an invariant for $q^i$ this is automatically satisfied by $\hat{q}^i$.
-Given that the value of the transported state variables does not change during evolution, and that it is unique for each initial state, transported state variable provide a way to label the trajectories themselves.
-While the numeric value of the transported state variables remains the same, the state variable (and its physical meaning) changes.
-For example, pressure and temperature will be combined depending on the particular process.
+The level sets of $\hat{q}^i$ are the evolved level sets of $q^i$ (e.g.
+all the states that started with a particular value for pressure) which allows us to study how groups of states evolve in time.
+Given that the value of the transported state variables does not change during evolution, and that it is unique for each initial state, transported state variables also provide a way to label the trajectories themselves.
 
 Evolved state variables are useful to write equations of motions, study how physical quantities change, form a physical picture of what happens.
-Transported state variables are useful to write invariants, study state space trajectories, from a geometric picture for the state space.
-The two, though, are tightly linked.
-In fact:
-\begin{align*}
-\Delta q^i &= q^i(\hat{\mathcal{s}}) - q^i(\mathcal{s}) \\
-\hat{q}^i(\hat{\mathcal{s}}) &= q^i(\hat{\mathcal{s}}) - \Delta q^i
-\end{align*}
-Intuitively, the transported variables change in the oppose way as the evolution, so that the value of the variables remain the same.\footnote{One should not confuse evolved/transported state variable with active/passive transformations or with Schroedinger/Heisenberg pictures.
-In those cases, the choice is whether to change the state or the coordinates/observables.
+Transported state variables are useful to write invariants, study state space trajectories, form a geometric picture for the state space.\footnote{One should not confuse evolved/transported state variable with active/passive transformations or with Schroedinger/Heisenberg pictures.
+In those cases, the choice is between changing the state or the coordinates/observables.
 In our case, the state is always changing.
-The choice is whether to change the state variables or their value.}
+The choice is between tracking the change using different state variables or different values.}
 
 \begin{defn}\label{evolved_transported_variable}
-TODO: Define evolved and transported variables in terms of pushback/pushforward
+Let $q^i$ be a set of state variables and $\mathcal{T}_{\Delta t}$ a deterministic and reversible evolution map.
+The evolved state variables are given by $q^i \circ \mathcal{T}_{\Delta t}$.
+The transported state variables are given by $q^i \circ \mathcal{T}_{-\Delta t}$.
 \end{defn}
 
-\begin{justification}
-TODO: prove existence
-\end{justification}
+To show the usefulness of transported state variables, we can prove the following:
 
-For discrete state variables, the total number of states is the product of the possibilities of each independent state variable.
-That is: if $\Delta q^1$ and $\Delta q^2$ are range of possibilities for two independent variables, the total number of possibilities $\#(\Delta q^1) \#(\Delta q^2)$.
-These relationships are maintained during evolution by the transported variables.
-Since the value remains unchanged, the range of possibilities for each value remain the same: $\#(\Delta \hat{q}^i) = \#(\Delta q^i)$.
-The variables remain independent as well, all possible combination remain, and we have: $\#(\Delta \hat{q}^1) \#(\Delta \hat{q}^2) = \#(\Delta q^1) \#(\Delta q^2)$.
-These relationship will later be extended in the continuous case to independent degrees of freedom for a Hamiltonian system, and will justify the symplectic form.
-Though this can't be done in general and requires further assumptions.
-
-In the case of continuous state, if each state can be identified by n independent state variables then, at least locally, we can chart the state space on $\mathcal{R}$: the state space is a manifold.
-Studying in general how ranges of possibilities evolve, though, becomes problematic.
-Time evolution will map points, while range of continuous possibilities will be length.
-We would need a metric, invariant during the evolution, telling us how ranges of possibilities map from one variable to another, possibly describing different physical quantities.
-We will not study this case, but we do note that, if such a metric exist, it depends on the system and on the choice of state variables.
-
-\begin{prop}\label{discrete_state_space}
-The space state $\mathcal{S}$ for a system fully identified by a set of $n$ independent discrete state variables $q^i : \mathcal{S} \rightarrow \mathbbm{I}^i$ is isomorphic to $\prod\limits_{i=1}^n\mathbbm{I}^i$.
-Let $I^i \subseteq \mathbbm{I}^i$ be finite sets of possibilities for each variable.
-$\#(\bigcap\limits_{i=1}^{n}{q^i}^{-1}(I^i))=\prod\limits_{i=1}^{n}\#(I^i)$.
+\begin{prop}\label{discrete_state_metric}
+	Let $U \subseteq \mathcal{S}$ a set of states fully identified by a set of $n$ independent discrete state variables $q^i$.
+Let $\Delta q^i = (q^i)^{-1}(U)$ the range of possibilities of each variables.
+Then $\#(\Delta q^i)=\#(\Delta \hat{q}^i) \; \forall i$ and $\#(U)=\prod\limits_{i=1}^{n}\#(\Delta q^i)=\prod\limits_{i=1}^{n}\#(\Delta \hat{q}^i)=\#(\hat{U})$ where $\hat{U}=\mathcal{T}_{\Delta t}(U)$.
 \end{prop}
@@
 \begin{proof}
-TODO
+	We claim $\#(\Delta q^i)=\#(\Delta \hat{q}^i)$.
+For each $\mathcal{s} \in U$, let $\hat{\mathcal{s}}=\mathcal{T}_{\Delta t}(\mathcal{s})$.
+We have $\hat{q}^i(\hat{\mathcal{s}}) = q^i(\mathcal{s})$.
+Therefore $\Delta q^i = (q^i)^{-1}(U) = (\hat{q}^i)^{-1}(\hat{U})=\Delta \hat{q}^i$.
+$\#(\Delta q^i)=\#(\Delta \hat{q}^i)$.
+
+	We claim $\#(U)=\prod\limits_{i=1}^{n}\#(\Delta q^i)$. As $q^i$ are independent variables, the states are the Cartesian product of the possibilities of each variable.
 \end{proof}
 
-TODO: in continuous, not necessarily the Cartesian product.
-But we can always create a chart with state variables.
-State space is a manifold of dimension n.
-In principle, we could say more.
-For example, a deterministic/reversible map would not map a set to its subset or superset.
-We could define a metric, this would give us an ergotic system.
-Not going to do that.
+What happens is that, because the variables are independent, the total number of states is the product of the number of possibilities for each variables.
+If $\Delta q^1$ and $\Delta q^2$ are ranges of possibilities for two independent variables, the total number of possibilities $\#(\Delta q^1) \#(\Delta q^2)$.
+By construction, the set of possibilities for each independent variables $\Delta q^i$ is the same as the set of possibilities for the transported variable $\Delta \hat{q}^i$.
+Also, the transported variables remain independent.
+Therefore the relationship $\#(\Delta \hat{q}^1) \#(\Delta \hat{q}^2) = \#(\Delta q^1) \#(\Delta q^2)$ is valid throughout the evolution.
 
-We have seen that physical systems undergoing deterministic and reversible evolution are properly described by a dynamical system.
-Systems described by a discrete set of states have nice properties for counting possibilities of independent variables, which are maintained during the evolution.
-System described by a continuous set of states, instead, do not posses in general such properties.
+We'll see that very similar relationships are what define Hamiltonian and Lagrangian mechanics.
+But recovering them in the continuous case is trickier, and in fact will be part of the challenge in the following sections.
+Consider the map $q'=aq$ with $0<a<1$.
+At first glance, it's a bijective continuous map so we may think it can represent a deterministic and reversible evolution.
+Yet, $\Delta q = [-b, b] \supset \Delta q'$: a set of states is mapped to a proper subset (i.e.
+to fewer states) which does not make sense for a reversible process.
+In the limit where we apply the map an infinite amount of times, any value of $q$ will be brought infinitely close to $0$, which also does not sound like a reversible process.
+The issue is that while the evolution map will map points to point, not all point to point map can be considered deterministic and reversible.
+Discrete sets already come with a way to measure and compare the number of elements which is preserved by any bijective map.
+With continuous sets, we need to do extra work to properly define a measure (or metric) that allows to count states and possibilities.
 
@@
 the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics, we should explore how to characterize a system in terms of its components.
+That is, we want to study the relationship between the state space $\mathcal{C}$ of the composite system and the state space $\mathcal{S}$ of its parts.
 In general, this is quite a complicated thing to do, that requires intimate knowledge of the system at hand.

### 2016-03-30 2015CompleteDerivation/completeDerivation.tex
Minor improvements in the first 2 main sections

@@
 
-Our first task is to extend the standard picture of system plus environment by differentiating between the state of the system (i.e.
+Our first task is to develop a conceptual model that applies to all realms of physics we'll be considering: classical, statistical and, later, quantum mechanics.
+We will take the standard picture of system plus environment and extend it to differentiate between the state of the system (i.e.
 the aspects under study) from the unstated part of the system (i.e.
@@
 for each present state there is one and only one future state).
-While the unstated part does not influence the state evolution, it constrains what states are available and what type of description can or cannot be given to the system.
-This understanding is what will allow us to treat classical, statistical and, later, quantum states within the same framework.
+While the unstated part does not influence the state evolution, we'll see how it constrains what states are available and what type of description can or cannot be given to the system.
 
@@
 
-In this context, we define the evolution of the system \emph{deterministic} if the state at a given time uniquely identifies states at future times, and \emph{reversible} if it uniquely identifies states at past times.
+In this context, we call the evolution of the system \emph{deterministic} if the state at a given time uniquely identifies states at future times, and \emph{reversible} if it uniquely identifies states at past times.
 While this is a common enough definition, we need to be clear how this applies to the unstated part.
@@
 That is, given the distribution at one time we can determine the distribution at future times.
-The shape and the parameters of the distribution can be deterministic, even if the trajectories of the parts are not (those fall within the unstated part).
-In fact, we \emph{cannot} assume trajectories and states are always defined at all for the unstated part, as this includes also the unknown unknowns.
+The shape and the parameters of the distribution can be deterministic, even if the trajectories of the parts are not as they fall within the unstated part.
+Note that we \emph{cannot} assume trajectories and states are always defined for the unstated part, as this includes also the unknown unknowns.
 We will return to this aspect when discussing quantum systems.
@@
 
-We now proceed to characterize states and their being physically distinguishable in more precise terms so that we can capture their description mathematically.
-What we'll see is that the set of all physical processes that can be used to gain information about the system, that can be used to perform a measurement, induce a topology on the set of states that is Hausdorff and second countable.
+We now proceed to characterize states and physically distinguishability in more precise terms so that we can capture their description mathematically.
+What we'll see is that the outcomes of all physical processes that can be used to gain information about the system, that can be used to perform a measurement, induce a topology on the set of states that is Hausdorff and second countable.
 That is, physical distinguishability is mathematically captured by topological distinguishability.
@@
 Consider the motion of a cannonball under inertial and gravitational forces.
-Light will scatter off of it; as it lands the ground will be deformed and the impact will make the temperature rise.
+Light will scatter off of it; as it lands the ground will be deformed and the impact will make the temperature slightly rise.
 Those external physical processes, which happen no matter what we do, can be used to distinguish the motion of the cannonball as their outcomes are correlated.
@@
 That is, for us a measurement is simply a physical process that we can use to distinguish states.
-Setting up an experiment means choosing a particular process with desired outcomes and forcing the system under study to interact with it.
+Setting up an experiment means choosing a particular process with desired outcomes and forcing the system under study to interact with it one or more times.
 After that, there is no special role played by the "observer" in making outcomes come about.
@@
 Note that the number of takes influences the outcomes: with greater statistics the precision and number of distinguishable cases increases.
-Therefore the processes, as we define them, may require repeated interactions with similarly prepared states.
+Therefore the processes, as we defined them, may require repeated interactions with similarly prepared states.
 They may even be a combination of different kinds of interactions that, taken all together, create a set of distinguishable outcomes.
@@
 Let $\mathsf{T}$ be the set of all sets associated with all physical outcomes.
+$\varnothing \in \mathsf{T}$ and is associated with impossible outcomes.
 $\mathcal{S} \in \mathsf{T}$ and is associated with unavoidable outcomes.
-$\varnothing \in \mathsf{T}$ and is associated with impossible outcomes.
 Let $V_1, V_2 \in \mathsf{T}$.
@@
 States, though, are not the only objects with that property.
-In fact: any element of a set of physical object (e.g.
+In fact: any element of a set of physical objects (e.g.
 forces, physical properties such as mass or charge, time) needs to be physically distinguishable to be well defined.
@@
 The generality of this result explains why in physics one always assumes functions to be well behaved.
-As the result was derived from our notion of physical distinguishability, this is not a matter of practical convenience or usefulness.
+As the result was derived from our notion of physical distinguishability, this is not a matter of practical convenience.
 Suppose we were able to prepare a force field that was zero everywhere in space except at a single point.
 This gives us a way to tag a specific point.
-A finite precision measurement of the force provides us infinite precision of space, which we ruled out.\footnote{This assumes that we are able to position the probe perfectly, which we could do if we able to manipulate forces at that precision.} This is what the math is telling us, that if we said before that outcomes can't distinguish isolated points (i.e.
+A finite precision measurement of the force provides us infinite precision of space, which we ruled out.\footnote{This assumes that we are able to position the probe perfectly, which we could do if we were able to manipulate forces at that precision.} This is what the math is telling us, that if we said before that outcomes can't distinguish isolated points (i.e.
 the topology is second countable) then neither can maps.
@@
 While functions with few discontinuities are useful and used in physics and engineering, they are employed for idealized cases  (e.g.
-a signal change is fast enough, a charge distribution is small enough) that are usually treated differently (i.e.
+a signal change is fast enough, a charge distribution is small enough) that are often treated as a special case (i.e.
 propagation in material discontinuities).
@@
 
-\subsection{Labeling states}
+\subsection{Labeling states and manifolds}
 
-To identify and name states we will use a set of quantities, typically numbers.
+To identify and name states one uses a set of quantities, typically numbers.
 For example, the orbital of an electron in a hydrogen item is identified by the quantum numbers $n$, $l$, $m$ and $s$.
@@
 We call a state variable \emph{discrete} or \emph{continuous} if the possibilities are integer or real numbers respectively.
-From now on, we are going to consider state spaces whose states can be identified by a finite set of discrete and continuous state variables (i.e.
-the state space is locally isomorphic to $\bigcup\limits_{1 \leq i \leq n} \mathbbm{R}^{n_i}$).\footnote{TODO: Is there a topological space that is Hausdorff and second countable that is not like that?}
+From now on, we are going to consider state spaces whose states can be identified, at least within a region, by a finite set of discrete and continuous state variables (i.e.
+the state space is locally isomorphic to $\bigcup\limits_{1 \leq i \leq n} \mathbbm{R}^{n_i}$).\footnote{TODO: Is there a more physically meaningful requirement?
+Could we have a set of states that are not locally described by a set of quantities?}
 
-We purposely use the term state variables instead of coordinates (even though that's what they are mathematically) as it will create confusion with space-time coordinates.
+We purposely use the term state variables instead of coordinates (even though that's what they are mathematically) as it would create confusion with space-time coordinates.
 We also avoid the term observable or measurable, as not all state variables may be directly physically tangible (e.g.
@@
 \begin{defn}\label{state_variable}
-	A \emph{state variable} is a continuous map $q : U \rightarrow \mathbbm{L}$ where $U \subseteq \mathcal{S}$ and $\mathbbm{L}$ is the topological space for the possible values.
+	A \emph{state variable} is a continuous map $q : U \rightarrow \mathbbm{L}$ where $U \subseteq \mathcal{S}$ and $\mathbbm{L}$ is the space for the possible values.
 If $\mathbbm{L}\cong \mathbbm{Z}$ the variable is said \emph{discrete}.
@@
 When combining multiple state variables, it is important to understand how they relate to each other.
-Consider the orbital of an electron in a hydrogen item, which is identified by the quantum numbers $n$, $l$, $m$ and $s$.
+Consider the orbital of an electron in a hydrogen atom, which is identified by the quantum numbers $n$, $l$, $m$ and $s$.
 For each combination of $n$, $l$ and $m$, the spin $s$ can have two values.
@@
 
-The discrete variable do not present such problems, as any number of them can be flattened out in a single one (i.e.
+Discrete variables do not present such problems, as any number of them can be flattened out in a single one (i.e.
 $\mathbbm{Z}^n$ is homeomorphic to $\mathbbm{Z}$).
 They will determine the number of connected components of the manifold.
-Once we introduce a continuous parameter for time evolution, though, these also become irrelevant as continuous time evolution requires continuous trajectories that cannot move states across disconnected components.
+Once we introduce a continuous parameter for time evolution, though, these become irrelevant as continuous time evolution requires continuous trajectories that cannot move states across disconnected components.
 This justifies the special interest in path connected manifolds, as this is where trajectories for deterministic and reversible continuous evolution live.
@@
 Given that scientific reductionism (i.e.
-the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics, we should explore how to characterize a system in terms of its components.
+the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics, we now explore how to characterize a system in terms of its components.
 That is, we want to study the relationship between the state space $\mathcal{C}$ of the composite system and the state space $\mathcal{S}$ of its parts.

### 2016-03-31 2015CompleteDerivation/completeDerivation.tex
Minor typos correction from Christine

@@
 But in the end, however complicated it is, the conceptual model remains the same: each process has a set of possible outcomes, and each possible outcome will be associated with a set of states consistent with that outcome.
-For example, if the cannonball deformed the ground by this amount then its kinetic energy at impact was within this range; if the electron follows a certain path, then its state is among the ones that has spin $+1/2$.
+For example, if the cannonball deformed the ground by this amount then its kinetic energy at impact was within this range; if the electron follows a certain path, then its state is among the ones that have spin $+1/2$.
 
@@
 Consider the set of outcomes of a process.
-As each is identified by a finite number of digits.
-The set of outcomes is countable.
-This remains true when combining processes: the number of processes we can combine is finite as so it is the number of times we can prepare the same state and the intersections of countable many sets is still countable.
-Consider a bases for the topology.
-Suppose it is of cardinality greater than countability.
+As each is identified by a finite number of digits, the set of outcomes is countable.
+This remains true when combining processes: the number of processes we can combine is finite as is the number of times we can prepare the same state and the intersections of countably many sets is still countable.
+Consider a basis for the topology.
+Suppose it is of cardinality greater than countable.
 Then no physical process could ever distinguish between the elements of the basis.
 The basis, and therefore the outcomes, would not be physically distinguishable.
-The basis for the topology cannot be of cardinality greater than countability.
+The basis for the topology cannot be of cardinality greater than countable.
 By contradiction, $\mathcal{S}$ must be second countable.
@@
 $\mathcal{T}_{\Delta t}$ admits $\mathcal{T}_{-\Delta t}$ as an inverse.
-$\mathcal{T}_{\Delta t}$ is a bijection
+$\mathcal{T}_{\Delta t}$ is a bijection.
 
@@
 This gives us a way to tag a specific point.
+But it also allows us to create an outcome compatible with only a single point.
 A finite precision measurement of the force provides us infinite precision of space, which we ruled out.\footnote{This assumes that we are able to position the probe perfectly, which we could do if we were able to manipulate forces at that precision.} This is what the math is telling us, that if we said before that outcomes can't distinguish isolated points (i.e.
@@
 
-While functions with few discontinuities are useful and used in physics and engineering, they are employed for idealized cases  (e.g.
-a signal change is fast enough, a charge distribution is small enough) that are often treated as a special case (i.e.
-propagation in material discontinuities).
+While functions with few discontinuities are useful and used in physics and engineering, they are employed for idealized cases  (e.g.~a signal change is fast enough, a charge distribution is small enough) that are often treated as a special case (e.g.~propagation in material discontinuities).
 While this may be obvious and intuitive to the physicist, it may be troubling to the mathematician as the proper use of many mathematical techniques requires the inclusion of discontinuous functions.
@@
 
-\subsection{Labeling states and manifolds}
+\subsection{Manifolds and labeling states}
 
 To identify and name states one uses a set of quantities, typically numbers.
-For example, the orbital of an electron in a hydrogen item is identified by the quantum numbers $n$, $l$, $m$ and $s$.
+For example, the orbital of an electron in a hydrogen atom is identified by the quantum numbers $n$, $l$, $m$ and $s$.
 We call each of these quantities \emph{state variables}.
@@
 From now on, we are going to consider state spaces whose states can be identified, at least within a region, by a finite set of discrete and continuous state variables (i.e.
-the state space is locally isomorphic to $\bigcup\limits_{1 \leq i \leq n} \mathbbm{R}^{n_i}$).\footnote{TODO: Is there a more physically meaningful requirement?
+the state space is locally isomorphic to $\bigcup\limits_{1 \leq i \leq n} \mathbbm{R}^{m_i}$).\footnote{TODO: Is there a more physically meaningful requirement?
 Could we have a set of states that are not locally described by a set of quantities?}
@@
 Therefore velocity is always physically well defined but is not a state variable in general, while conjugate momentum is always a state variable but is not physically well defined by itself (it requires the vector potential to be specified as well).
-As we'll see later, only once a link is established between them they can both be physically well defined state variables.}
+As we'll see later, if there exists a one to one map between velocity and conjugate momentum they can both be physically well defined state variables.}
 
@@
 $\mathbbm{Z}^n$ is homeomorphic to $\mathbbm{Z}$).
-They will determine the number of connected components of the manifold.
+They will determine the number of connected components of the state space.
 Once we introduce a continuous parameter for time evolution, though, these become irrelevant as continuous time evolution requires continuous trajectories that cannot move states across disconnected components.
@@
 There exists a neighborhood $U$ around $\mathcal{s}$ where we have $(q_1,...,q_n):U\rightarrow \mathbbm{R}^n$.
-This map is a bijection as $s$ is identified by those variables.
+This map is a bijection as $\mathcal{s}$ is identified by those variables.
 $\mathcal{S}$ is homeomorphic to $\mathbbm{R}^n$ around  $\mathcal{s}$.
@@
 Deterministic and reversible evolution is a homeomorphism between initial and final states.
-$\forall \hat{\mathcal{s}} \in \mathcal{S}'$ the evolution of $U$ will contain $\hat{\mathcal{s}}$ and will be isomorphic to $\mathbbm{R}^n$.
+$\forall \hat{\mathcal{s}} \in \mathcal{S}'$ there exist a map $\mathcal{T}_{\Delta t}:\mathcal{S} \rightarrow \mathcal{S}$ such that $\hat{\mathcal{s}} =\mathcal{T}_{\Delta t}(\mathcal{s})$.
+$\hat{\mathcal{s}} \in \mathcal{T}_{\Delta t}(U)$ and $\mathcal{T}_{\Delta t}(U)$ is isomorphic to $\mathbbm{R}^n$.
 $\mathcal{S}'$ equipped with the subspace topology is a Hausdorff, second countable topological space everywhere homeomorphic to $\mathbbm{R}^n$.
@@
 There exist map $\lambda : [t_0,t_1] \rightarrow \mathcal{S}'$ where $\lambda(t_0)=\mathcal{s}$, $\lambda(t_1)=\hat{\mathcal{s}}$ and $t_0$ and $t_1$ are the initial and final time respectively.
-$\lambda$ is a map from two physically distinguishable quantities, time and states.
+$\lambda$ is a map between two physically distinguishable quantities, time and states.
 $\lambda$ is continuous by \ref{continuous_map}.
@@
 
-What happens is that, because the variables are independent, the total number of states is the product of the number of possibilities for each variables.
-If $\Delta q^1$ and $\Delta q^2$ are ranges of possibilities for two independent variables, the total number of possibilities $\#(\Delta q^1) \#(\Delta q^2)$.
+What happens is that, because the variables are independent, the total number of states is the product of the number of possibilities for each variable.
+If $\Delta q^1$ and $\Delta q^2$ are ranges of possibilities for two independent variables, the total number of possibilities is $\#(\Delta q^1) \#(\Delta q^2)$.
 By construction, the set of possibilities for each independent variables $\Delta q^i$ is the same as the set of possibilities for the transported variable $\Delta \hat{q}^i$.
@@
 In the limit where we apply the map an infinite amount of times, any value of $q$ will be brought infinitely close to $0$, which also does not sound like a reversible process.
-The issue is that while the evolution map will map points to point, not all point to point map can be considered deterministic and reversible.
+The issue is that while deterministic evolution will map points to points, not all point to point maps can be considered deterministic and reversible.
 Discrete sets already come with a way to measure and compare the number of elements which is preserved by any bijective map.

### 2016-04-01 2015CompleteDerivation/completeDerivation.tex
First pass at classical homogeneous system

@@
 The notion that the system is homogeneous means that the states of the composite and of the parts are not unrelated: they are made of the same material.\footnote{Whether a system is homogeneous depends on context (e.g.
-air can be thought as homogeneous if the mixture of gases does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is maintained by by time evolution.} In fact, the state space $\mathcal{C}$ of all systems composed of such material will the state of the system as well as the states of all its part (i.e $\mathcal{c}, \mathcal{c}_1, \mathcal{c}_2 \in \mathcal{C}$).
+air can be thought as homogeneous if the mixture of gases does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is maintained by time evolution.} In fact, the state space $\mathcal{C}$ of all systems composed of such material will include the state of the system as well as the states of all its part (i.e $\mathcal{c}, \mathcal{c}_1, \mathcal{c}_2 \in \mathcal{C}$).
 Since combining any two systems made of a homogeneous material will always give us a system made of the same homogeneous material (e.g.
@@
 We also want to be able to express the state of the composite system in terms of the state of the parts.
-For example, given the state of a fluid we'll want to know how its parts are distributed.
+For example, given the state of a fluid we'll want to know how the material is distributed in space.
 This in general will depend on how much the state of the composite system "knows" about its parts.
@@
 We will call a classical system one that is infinitesimally reducible.
-That is: it is made of arbitrarily small parts, which we call particles, and the state of the whole system is equivalent to the sum of the states of the parts.
+That is: it is made of arbitrarily small parts, which we call particles, and the state of the whole system coincides with the sum of the states of the parts.
 Given the state space $\mathcal{S}$ for the infinitesimal parts, a composite state $\mathcal{c}$ will tell us the amount of material for each possible particle state.
-That is, each state is fully identified by a function $a: \mathcal{S} \to \mathbbm{R}$ that returns the amount of material in the composite state $\mathcal{c}$ that is prepared in state $\mathcal{s}$.
-For example, given a certain configuration of gas, we can tell the amount that is at a specific point in space with a specific value of momentum.
-The notion that the parts are infinitesimal requires the co-domain of $a$ to be continuous, as opposed to discrete.
+That is, each state is fully identified by a function $\rho: \mathcal{S} \to \mathbbm{R}$ that returns the amount of material in the composite state $\mathcal{c}$ that is prepared in the particle state $\mathcal{s}$.
+For example, given a certain configuration of gas, we can tell the amount of material that is at a specific point in space with a specific value of momentum.
+
+The notion that the parts are infinitesimal requires the co-domain of $\rho$ to be continuous, as opposed to discrete.
 The state space $\mathcal{S}$, instead, can be finite.
@@
 The amount in each tank is continuous (as we assume the water to be infinitesimally divisible) yet there are only a finite number of tanks the water can be placed into.
+As it links two physically distinguishable objects, $\rho$ is a continuous functions for reasoned discussed before.
 
-As we combine the state of finite parts, we sum the distributions over particle states $a(\mathcal{s})=a_1(\mathcal{s})+a_2(\mathcal{s})$: the amount of material in the composite state is the sum of the amount of material of the parts.
-This gives us the structure of a real vector space.
+As we combine the state of finite parts, we sum the distributions over particle states $\rho(\mathcal{s})=\rho_1(\mathcal{s})+\rho_2(\mathcal{s})$: the amount of material in the composite state is the sum of the amount of material of the parts.
+We can also increase or decrease the amount of material by a constant factor $\rho_1(\mathcal{s})=a\rho_2(\mathcal{s})$.
+This gives us the structure of a real vector space that is isomorphic to the space of continuous functions.
 
@@
 The state space $\mathcal{C}$ for a homogeneous classical (i.e.
-infinitesimally reducible) material is a vector space over $\mathbbm{R}$.
-It admits a basis isomorphic to the state space $\mathcal{S}$ of its particles (i.e.
-infinitesimal parts).
+infinitesimally reducible) material is a vector space over $\mathbbm{R}$ isomorphic to a subspace of $C(\mathcal{S}) = \{\rho:\mathcal{S} \rightarrow \mathbbm{R} | \rho$ is continuous $\}$, where $\mathcal{S}$ is the state space of an infinitesimal amount of material.
 \end{defn}
@@
 \begin{justification}
-The state space is an abelian group as the material is homogeneous and decomposable and \ref{reducible_state_space}.
+We claim $\mathcal{C}$ is an abelian group. $\mathcal{C}$ is a decomposable homogeneous system. $\mathcal{C}$ is an abelian group  by \ref{reducible_state_space}.
 
-We now consider the set of transformations $T$ that increase or decrease the amount of material in the system by a constant.
-We claim $T$ is a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics sense (a physical quantity with a value for each point in space).} isomorphic to $\mathbbm{R}$.
+Consider the set of transformations $T$ that increase or decrease the amount of material in the system by a constant.
+We claim $T$ is a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics or differential geometry sense (a physical quantity/tensor with a value for each point in space).} isomorphic to $\mathbbm{R}$.
 Consider $\tau: \mathbbm{R} \rightarrow T$ the mapping between a number and the transformation that increases or decreases the amount of material by that number.
@@
 
-We now claim that the state space $\mathcal{C}$ is a vector space over $\mathbbm{R}$.
+We claim $\mathcal{C}$ is a vector space over $\mathbbm{R}$.
 The abelian group $\mathcal{C}$ can be extended with the operations defined by $T$, as each element $\tau \in T$ is a map $\tau : \mathcal{C} \rightarrow \mathcal{C}$.
-The map has the following properties: $(\tau_1 + \tau_2) \mathcal{c} = \tau_1 \mathcal{c} + \tau_2 \mathcal{c}$, increasing the amount of material by the sum of two constant is the same as combining the separate increases, and $\tau (\mathcal{c}_1 + \mathcal{c}_2) = \tau \mathcal{c}_1 + a \mathcal{c}_2$, increasing the amount of the total system is the same as the combination of the increased parts.
+The map has the following properties: $(\tau_1 + \tau_2) \mathcal{c} = \tau_1 \mathcal{c} + \tau_2 \mathcal{c}$, increasing the amount of material by the sum of two constant is the same as combining the separate increases, and $\tau (\mathcal{c}_1 + \mathcal{c}_2) = \tau \mathcal{c}_1 + \tau \mathcal{c}_2$, increasing the amount of the total system is the same as the combination of the increased parts.
 $\mathcal{C}$ is a module over $T$, which is a field and isomorphic to $\mathbbm{R}$.
@@
 
-Consider now the injection map $\mathcal{e}: \mathcal{S} \hookrightarrow \mathcal{C}$ that for each particle state returns a composite state constituted only by a predetermined unit amount of material in such state.
-The image of such injection is a basis for $\mathcal{C}$.
-We first claim that all elements are linearly independent.
-Consider $\sum\limits_{\mathcal{s} \in \mathcal{S}} \tau(a(\mathcal{s})) \mathcal{e}(\mathcal{s})$, $a: \mathcal{S} \to \mathbbm{R}$.
-This corresponds to a physical state with $a(\mathcal{s})$ material in each state $\mathcal{s}$.
-This is equal to the empty state if and only if $a(\mathcal{s})=0 \forall \mathcal{s} \in \mathcal{S}$.
-We now claim they span.
-Consider $\mathcal{c} + \sum\limits_{\mathcal{s} \in \mathcal{S}} \tau(a(\mathcal{s})) \mathcal{e}(\mathcal{s})$.
-This can always be made equal to the empty state by setting $a(\mathcal{s})$ to be the opposite of the amount of material of $\mathcal{c}$ found in state $\mathcal{s}$.
-Both arguments hold even if $\mathcal{S}$ is an infinite or uncountable set, as they hold for each finite subset of $\mathcal{S}$ independently and therefore $\mathcal{C}=\prod\limits_{\mathcal{s} \in \mathcal{S}} span({\mathcal{e}(\mathcal{s})})$.
+We claim $\mathcal{C}$ is isomorphic to a subspace of $C(\mathcal{S})$ as a vector space over $\mathbbm{R}$.
+Let $\rho:\mathcal{S} \rightarrow \mathbbm{R}$ be a map returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$.
+$\rho \in C(\mathcal{S})$ by \ref{continuous_map}.
+Let $G \subset C(\mathcal{S})$ be the functions that correspond to physically meaningful configurations.
+To each $\mathcal{c} \in \mathcal{C}$ corresponds one and only one $\rho_\mathcal{c} \in G$ as the system is infinitesimally reducible.
+Let $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$, then $\rho_\mathcal{c}=\rho_{\mathcal{c}_1}+\rho_{\mathcal{c}_2}$ as the amount of material of the composition of two states is the sum of the individual amounts.
+Let $\mathcal{c}_1=\tau(a)\mathcal{c}_2$, then $\rho_{\mathcal{c}_1}=a\rho_{\mathcal{c}_2}$ as $\tau(a)$ increases the amount of material by the factor $a$.
+$\mathcal{C}\cong G$ as vector spaces.
+$G$ is a subspace of $C(\mathcal{S})$.
+
 \end{justification}
 
-Need to define inner product.
-Introduce $Q:\mathcal{C} \to \mathbbm{R}$ amount of "stuff" of the system.
-$Q:\mathcal{S}\times\mathcal{C} \to \mathbbm{R}$ amount of stuff of the composite system in that particular state.
-We can use those those to construct inner product, and to show that states are bounded, but: norm induced by the inner product is not physical.
+Note that the isomorphism is to a subspace of continuous functions as we may have to exclude, for example, those that are not bounded or those that do not vanish at infinity.
+These may, in some cases, not represent physically meaningful system.
 
-\begin{defn}\label{classical_inner_product}
-	TODO: there exist an inner product.
+\subsection{Coordinate invariant densities}
+
+Math: need linear functional from $\rho$ to total amount of material.
+For Riesz representation theorem, there can be only one.
+Density must be invariant under coordinate transformation.
+Density live on $T^*\mathbbm{Q}$.
+Must be able to define marginal distributions on submanifold.
+Metric is two form.
+Metric must be form invariant under coord transformation.
+Metric is canonical two form.
+
+Physics, 1D, show that densities do not work on Q.
+Show they work on T*Q.
+Define degree of freedom.
+Independent degree of freedom are orthogonal.
+Area in each d.o.f.
+is number of possibilities.
+Total number of states is product of number of possibilities for independent d.o.f.
+Independent d.o.f.
+are orthogonal.
+Leads to omega metric.
+Density allows to count particle states.
+Example: two uniform distribution for the same amount of particles.
+
+Also: we can search for omega looking from constraints. Must be linear. Must be skew-symmetric (applied to the same vector twice must be zero). Must be invariant by coord transformation.
+
+\begin{defn}\label{classical_phase_space}
+The state space $\mathcal{S}$ for the particles of a homogeneous classical  system is a cotangent bundle $T^*\mathbbm{Q}$ with symplectic metric $\omega$.
 \end{defn}
@@
 
-Show countable basis. $\mathcal{c}= \sum\limits_{\mathcal{s} \in \mathcal{S}} a(\mathcal{s}) \mathcal{e}(\mathcal{s})=\sum\limits_{q \in \mathbbm{Q}} a(q) \mathcal{e}(q)$
-
-% Continuous state variable: Q is measured at constant t
-
-Uncountable basis does not work. $\mathcal{c}= \int\limits_{q \in \mathbbm{Q}} \rho(q) \mathcal{e}(q)dq$ but $\rho(q')=|\partial _{q}q'|\rho(q)\neq\rho(q)$
-
-TODO: need to understand a valid notation for composite state as infinite sum. The densities is the coefficient, the basis are are states, what is the differential?
-
-Find smallest object that can have an invariant density (inner product). $\mathcal{c}= \int\limits_{T^*\mathbbm{Q}} \rho(q,k) \mathcal{e}(q,k)dq\wedge dk$
-
-\begin{defn}\label{classical_phase_space}
-The state space $\mathcal{S}$ for the particles of a homogeneous classical  system is a cotangent bundle $T^*\mathbbm{Q}$.
-\end{defn}
-
-\begin{justification}
-	TODO
-\end{justification}
-
-TODO: generalization of inner product gives us $\int \rho_1(q, p) \rho_2(q, p) dq\wedge wp$.
-Space of continuous functions is not a complete metric space.
-Extend to square integrable.
-Get a Hilbert space.
-But topology is not a second countable: more than aleph 1.
-No physical justification, in fact we introduce unphysical functions (i.e.
-0 rational, 1 irrational).
-Discontinuous functions not physical states: they are not distinguishable.
-
-Density allows to count particle states. Example: two uniform distribution for the same amount of particles.
-
-Area in each d.o.f.
-is number of possibilities.
-Total number of states is product of number of possibilities for independent d.o.f.
-Independent d.o.f.
-are orthogonal.
-Leads to omega metric.
-As in discrete case, conserved during evolution.
-
-Also: we can search for omega looking from constraints. Must be linear. Must be skew-symmetric (applied to the same vector twice must be zero). Must be invariant by coord transformation.
-
-\begin{defn}\label{symplectic_manifold}
-	The state space $\mathcal{S}$ for the particles of a homogeneous classical  system is a symplectic manifold with metric $\omega$, the canonical two-form.
-\end{defn}
-
-\begin{justification}
-	TODO
-\end{justification}
-
-\subsection{Classical hamiltonian assumption}
+\subsection{Classical Hamiltonian assumption}
 

### 2016-04-15 2015CompleteDerivation/completeDerivation.tex
Starting to tackle classical phase space

@@
 Could we extend the state and the laws of evolution to account for temperature?
-Yes, but that would be a different process defined on a different state.
+Yes, but that would be a different evolution defined on a different state.
 Does it mean the unstated part is always evolving chaotically?
@@
 The temperature may remain constant throughout the motion of the cannonball.
-Yet we wouldn't know, since we are not studying it: the motion is deterministic and/or reversible only as far as the state is concerned.
+Yet we wouldn't know, since we are not studying it: the evolution is deterministic and/or reversible only as far as the state is concerned.
 With this in mind, we introduce the following:
@@
 Suppose we study the motion of a small particle, small enough that the random scattering does influence the trajectory and it undergoes Brownian motion: its state will be a probability distribution for position and momentum of the center of mass.
-Gravitational and inertial forces have not changed, yet the set of states has.
-In other words, the set of states must also be closed under non-deterministic evolution.
+Gravitational and inertial forces have not changed, yet the states have changed from ``pure" to statistical ensembles.
+In other words, the set of states must be closed under both the deterministic evolution of the state and the non-deterministic evolution of the unstated part.
 If the Brownian motion is not negligible, we do not end in a well defined position/momentum pair, even if we start from one.
@@
 That is, given the distribution at one time we can determine the distribution at future times.
-The shape and the parameters of the distribution can be deterministic, even if the trajectories of the parts are not as they fall within the unstated part.
+The shape and the parameters of the distribution can be deterministic, even if the evolution of the parts are not as they fall within the unstated part.
 Note that we \emph{cannot} assume trajectories and states are always defined for the unstated part, as this includes also the unknown unknowns.
@@
 
-We now proceed to characterize states and physically distinguishability in more precise terms so that we can capture their description mathematically.
+We now proceed to characterize states and physical distinguishability in more precise terms so that we can capture their description mathematically.
 What we'll see is that the outcomes of all physical processes that can be used to gain information about the system, that can be used to perform a measurement, induce a topology on the set of states that is Hausdorff and second countable.
@@
 That is, having a way to measure quantity $x$ and a way to measure quantity $y$ gives us a way to measure the combination $(x,y)$.
-For ensembles, if we can measure the marginal distribution $\rho_x(x)$ and the marginal distribution $\rho_y(x)$, we know the joint distribution $\rho(x,y)$ has to be compatible with both.
+For ensembles, if we can measure the marginal distribution $\rho_x(x)$ and the marginal distribution $\rho_y(y)$, we know the joint distribution $\rho(x,y)$ has to be compatible with both.
 Note, though, that this does not provide a way to fully measure $\rho(x,y)$ as we know nothing about the correlation between the two variables.\footnote{The quantum case is similar: we measure marginal distributions and rule out states that are incompatible with them.} Formally, the states compatible with the outcomes of the combined process will be the intersections of the states compatible with each pair of outcomes of the original processes.
@@
 \begin{justification}
-We claim $\mathcal{S}$ is a set. Each state is well defined as it is physically distinguishable. The collection of all possible states form a set.
+We claim $\mathcal{S}$ is a set. Each state is well defined as it is physically distinguishable. The collection of all possible states forms a set.
 
@@
 Let $V_1, V_2 \in \mathsf{T}$.
-Then, by definition, there exist a process $P_1$ that admits $V_1$ as an outcome and a process $P_2$ that admits $V_2$ as an outcome.
+Then, by definition, there exists a process $P_1$ that admits $V_1$ as an outcome and a process $P_2$ that admits $V_2$ as an outcome.
 Consider the process $P$ that combines the outcomes of $P_1$ and $P_2$.
-This always exist physically as we can prepare the same state multiple times and let it interact with each process separately.
+This always exists physically as we can prepare the same state multiple times and let it interact with each process separately.
 $P$ will have a possible outcome $V$ corresponding to the case where $P_1$ gave outcome $V_1$ and $P_2$ gave outcome $V_2$.
@@
 As each is identified by a finite number of digits, the set of outcomes is countable.
-This remains true when combining processes: the number of processes we can combine is finite as is the number of times we can prepare the same state and the intersections of countably many sets is still countable.
+This remains true when combining processes: the number of processes we can combine is finite as is the number of times we can prepare the same state and the set of intersections obtained from the countable outcomes of a finite number of processes is still countable.
 Consider a basis for the topology.
 Suppose it is of cardinality greater than countable.
-Then no physical process could ever distinguish between the elements of the basis.
-The basis, and therefore the outcomes, would not be physically distinguishable.
+Then the basis would be able to distinguish among states in a way that no physical process ever could.
+The topology could distinguish among states that are not physically distinguishable.
 The basis for the topology cannot be of cardinality greater than countable.
@@
 
-Note that the arguments that lead to the topological space had nothing to do with states per se, just that they are physically distinguishable.
+Note that the arguments that led to the topological space had nothing to do with states per se, just that they are physically distinguishable.
 States, though, are not the only objects with that property.
@@
 
-Note that the isomorphism is to a subspace of continuous functions as we may have to exclude, for example, those that are not bounded or those that do not vanish at infinity.
-These may, in some cases, not represent physically meaningful system.
+Note that not all continuous functions correspond to physically meaningful states: we must make sure they describe a finite amount of material.
 
@@
 
-Math: need linear functional from $\rho$ to total amount of material.
-For Riesz representation theorem, there can be only one.
-Density must be invariant under coordinate transformation.
-Density live on $T^*\mathbbm{Q}$.
-Must be able to define marginal distributions on submanifold.
-Metric is two form.
-Metric must be form invariant under coord transformation.
-Metric is canonical two form.
+For each classical composite state will also tell us the total amount of material, once we fixed unit (e.g.
+mole, kg, ...).
+That is, there will be a functional $\Lambda : \mathcal{C} \rightarrow \mathbbm{R}$ that tells us how much gas we have for a particular gas configuration.
+As we combine parts, the total amount of material will be the sum of their parts.
+As we increase or decrease by a constant factor the amount of material will increase or decrease by the same factor.
+That is: $\Lambda(a_1 \mathcal{c}_1 + a_2 \mathcal{c}_2) = a_1 \Lambda(\mathcal{c}_1) + a_2 \Lambda(\mathcal{c}_2)$.
+For a proper state, one that is not the difference of two states, we will expect a positive amount of material.
+So, if a density $\rho(\mathcal{S})$ is positive everywhere, then the total amount is also positive.
+This makes $\Lambda$ a positive linear functional.
 
-Physics, 1D, show that densities do not work on Q.
-Show they work on T*Q.
-Define degree of freedom.
-Independent degree of freedom are orthogonal.
-Area in each d.o.f.
-is number of possibilities.
-Total number of states is product of number of possibilities for independent d.o.f.
-Independent d.o.f.
-are orthogonal.
-Leads to omega metric.
-Density allows to count particle states.
-Example: two uniform distribution for the same amount of particles.
+Being able to tell the total amount is not enough.
+Experimentally, what we do is measure the amount of material for a region of phase space.
+For example, the amount of gas within a particular region of space for a particular range of momentum.
+In general, given an outcome of our physical process (i.e.
+an open set in the topology for $\mathcal{S}$), each composite state should tell us the amount of material associated with it (i.e.
+contained in the open set of states).
+For each $U \subseteq \mathcal{S}$ we'll have a positive linear functional $\Lambda_U : \mathcal{C} \rightarrow \mathbbm{R}$ whose value is always bounded.
 
@@
 
-\begin{defn}\label{classical_phase_space}
-The state space $\mathcal{S}$ for the particles of a homogeneous classical  system is a cotangent bundle $T^*\mathbbm{Q}$ with symplectic metric $\omega$.
+In mathematical terms, we can use the Riesz representation theorem for linear functionals to give $\mathcal{S}$ the structure of a measure space, with a suitable Borel $\sigma$-algebra and measure $\mu$.
+
+\begin{defn}\label{integration}
+	The state space $\mathcal{S}$ for the particles of a homogeneous classical material is endowed with a measure $\mu$.
+The state space $\mathcal{C}$ is isomorphic to the space of continuous functions that are integrable over any subset of $\mathcal{S}$.
+That is: $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$ where $L^1_{ew}(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq \mathcal{S}\}$.
 \end{defn}
@@
 \end{justification}
+
+Before we can properly identify $d\mu$, it turns out that we need to better characterize $\mathcal{S}$.
+
+On a state space with discrete topology and a single state variable we simply have $\Lambda_U (\mathcal{c}) : \sum \limits_{q=a}^b D(q)$.
+For one dimensional continuous states, we would simply expect $\Lambda_U (\mathcal{c}) : \int_a^b \rho (q) dq$.
+By a simple change of state variable, we see that this does not work: $\rho(q')= \rho(q) dq/dq'$.
+This would make $\rho$ not a function of just the state $\mathcal{s}$, but also of the state variables we are using.
+This is inconsistent with our previous definition.
+Moreover if the transformation is not differential, the density is not even well-defined.
+This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows coordinate independent densities.
+
+To do that $\mathcal{S}$ has to be, first of all, a differentiable manifold. This way, at least, the densities is well defined for all state variables.
+
+If $\rho$ is to remain unchanged when one state variable changes, then one or more other state variables have to change as well so that the Jacobian of the transformation is unitary.
+This means that we cannot change physical units to all the variables as we please: they are part of a unit system.
+For example, changing units of position will also change units of momentum.
+Let $\mathcal{Q}$ be the manifold charted by a the minimum set of state variables that define such system of units.
+An element of $\mathcal{S}$ will be a geometrical object in such space.
+As these are states for infinitesimal parts, it will be a local object in $\mathcal{Q}$.
+As we saw before, it cannot be a point as the density $\rho(q')= \rho(q) dq/dq'$ is not invariant.
+It cannot be a vector $v(q) e_q$, as the density $\rho(q', v')= \rho(q, v) (dq/dq')^2$ is not invariant.
+A covector $k(q) dq$ is fine as $\rho(q', k')= \rho(q, k) (dq/dq') (dq'/dq) = \rho(q, k)$.
+In other words: a covector is the simplest local geometrical object that allows coordinate independent densities.
+We can identify $\mathcal{S}$ as the cotangent bundle $T^*\mathcal{Q}$, the space of all possible covectors.
+
+If $\mathcal{Q}$ is one dimensional, then the state is fully identified by a pair $(q, k)$.
+$d\mu$ is proportional to $dq \wedge dk$.
+We set $d\mu = \hbar dq \wedge dk$, where $\hbar$ is the proportionality constant that will depend on the unit chosen to count the possibilities of the pair $(q, k)$.
+The cardinality for states in a finite (i.e.
+compact) region $U \subseteq \mathcal{S}$ will be given by $\int_U \hbar dq \wedge dk$.
+
+If $\mathcal{Q}$ is $n$-dimensional, $d\mu$ is proportional to $dq^n \wedge dk_n$.
+But it's not just the volume that is preserved as we change state variables.
+Each pair $(q^i, k_i)$ represents an independent set of state variables, with the cardinality of possibilities given by $\hbar dq^i \wedge dk_i$.
+We call such a pair a degree of freedom.
+As they are independent the total number of states is the product of the possibilities of each d.o.f.: $\hbar^n dq^n \wedge dk_n = \prod \limits_{i=1}^n \hbar dq^i \wedge dk_i$.
+This is equivalent to saying that independent d.o.f.
+are orthogonal surfaces in $\mathcal{S}$.
+
+The cardinality of possibilities for each d.o.f.
+(i.e.
+the wedge product within d.o.f) and the orthogonality of different d.o.f.
+(i.e.
+the scalar product across d.o.f.) must be the same regardless of the choice of state variables.
+We can express both requirements mathematically in a compact way.
+We first define conjugate momentum as $p_i=\hbar k_i$ and unified state variables as $\xi^a\equiv \{q^i, p_i\}$.
+Then we consider the canonical two-form $\omega$ given by the following components:
+\begin{align*}
+\omega_{ab} =  \left[
+\begin{array}{cc}
+0 & 1 \\
+-1 & 0 \\
+\end{array}
+\right] \otimes I_n =
+\left[
+\begin{array}{cc}
+0 & I_n \\
+-I_n & 0 \\
+\end{array}
+\right] \\
+\end{align*}
+It returns the wedge product within a d.o.f.
+and the scalar product across.
+Requiring the invariance of this metric under state variables assures us the cardinalities of states and possibilities is well defined.
+
+The state space $\mathcal{S}$ is the symplectic manifold $(T^*\mathcal{Q}, \omega)$.
+This is the simplest manifold that allows us to define coordinate invariant densities.
+The metric allows us to measure the cardinality of possibilities within and across independent d.o.f.
+
+\begin{defn}\label{classical_phase_space}
+The state space $\mathcal{S}$ for the particles of a homogeneous classical  system is a symplectic manifold form by a cotangent bundle $T^*\mathcal{Q}$ equipped with the canonical two-form $\omega$.
+\end{defn}
+
+\begin{justification}
+	TODO: we can search for omega looking from constraints. Must be linear. Must be skew-symmetric (applied to the same vector twice must be zero). Must be invariant by coord transformation.
+\end{justification}
+
+We should note that we are starting to be in a position similar to the one discussed in \ref{discrete_state_metric} for discrete states.
+We have both a topological space and a metric that allows to count states and possibilities.
+The only thing we started from was the idea of an infinitesimally reducible system.
+That lead to a state space $\mathcal{S}$ for the infinitesimal parts, on which we must be able to define state variable independent distributions.
+Which in turn gave us degrees of freedom made by pair of conjugate variables and the symplectic metric typical of classical phase space.
+
+Before discussing time evolution, we should investigate what assuming our state is infinitesimally reducible means in terms of the fundamental model discussed in Assumption \ref{detrevass}.
 

### 2016-04-15 2015CompleteDerivation/completeDerivation.tex
Starting to tackle classical phase space

@@
 
-Also: we can search for omega looking from constraints. Must be linear. Must be skew-symmetric (applied to the same vector twice must be zero). Must be invariant by coord transformation.
+The ability to associate finite amount of material with sets of states allows us to define finite regions of $\mathcal{S}$ and compare them.
+Consider a region of position and momentum in phase space.
+If we are able to spread a finite amount of material into a uniform distribution, then we know we have a finite region.
+And the density will give us an indication of how big the region is: if we double the region, the density will half.
+In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique way to measure the size of sets of $\mathcal{S}$, a metric $d\mu$ and our positive linear functionals become integrals: $\Lambda_U (\mathcal{c}) : \int_U \rho_{\mathcal{c}} d \mu$.
 

### 2016-04-20 2015CompleteDerivation/completeDerivation.tex
Another pass at classical Hamiltonian section

@@
 the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics, we now explore how to characterize a system in terms of its components.
-That is, we want to study the relationship between the state space $\mathcal{C}$ of the composite system and the state space $\mathcal{S}$ of its parts.
+That is, we want to study the relationship between the state space $\mathcal{C}$ of a composite system and the state space $\mathcal{S}$ of its parts.
 In general, this is quite a complicated thing to do, that requires intimate knowledge of the system at hand.
-So we simplify our problem and study an idealized case: one where the components are homogeneous and infinitesimal.
+So we simplify our problem and study the system is made of infinitesimal homogeneous parts.
 What we'll find is that under the additional assumption that the system is infinitesimally reducible (i.e.
-its state is equivalent to the state of its parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian particle mechanics.
+its state is equivalent to the state of its parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian mechanics.
 
@@
 For example, if a stable mixture of gas expands, we'll want to know the difference between the initial and final distributions.
-That is: $\Delta\mathcal{c}=\hat{\mathcal{c}}-\mathcal{c}$.
+That is: $\delta\mathcal{c}=\hat{\mathcal{c}}-\mathcal{c}$.
 Note that such state difference may not describe a physical state: it may remove some material from one location to add it somewhere else.
@@
 As we want to describe changes during the evolution (i.e.
-$\hat{\mathcal{c}} = \mathcal{c} +\Delta \mathcal{c}$) we introduce an inverse $- : \mathcal{C} \rightarrow \mathcal{C}$ such that $\mathcal{c} + ( - \mathcal{c}) = 0 \forall \mathcal{c} \in \mathcal{C}$.
+$\hat{\mathcal{c}} = \mathcal{c} +\delta \mathcal{c}$) we introduce an inverse $- : \mathcal{C} \rightarrow \mathcal{C}$ such that $\mathcal{c} + ( - \mathcal{c}) = 0 \forall \mathcal{c} \in \mathcal{C}$.
 Such inverse may introduce objects that do not properly represent a physical state, but a state change.
@@
 We can also increase or decrease the amount of material by a constant factor $\rho_1(\mathcal{s})=a\rho_2(\mathcal{s})$.
-This gives us the structure of a real vector space that is isomorphic to the space of continuous functions.
+This gives us the structure of a real vector space that is isomorphic to a subspace of continuous functions.
 
@@
 We claim $\mathcal{C}$ is isomorphic to a subspace of $C(\mathcal{S})$ as a vector space over $\mathbbm{R}$.
-Let $\rho:\mathcal{S} \rightarrow \mathbbm{R}$ be a map returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$.
+$\forall \mathcal{c} \in \mathcal{C} \exists !
+\rho_{\mathcal{c}}:\mathcal{S} \rightarrow \mathbbm{R}$ returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$.
 $\rho \in C(\mathcal{S})$ by \ref{continuous_map}.
-Let $G \subset C(\mathcal{S})$ be the functions that correspond to physically meaningful configurations.
-To each $\mathcal{c} \in \mathcal{C}$ corresponds one and only one $\rho_\mathcal{c} \in G$ as the system is infinitesimally reducible.
-Let $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$, then $\rho_\mathcal{c}=\rho_{\mathcal{c}_1}+\rho_{\mathcal{c}_2}$ as the amount of material of the composition of two states is the sum of the individual amounts.
-Let $\mathcal{c}_1=\tau(a)\mathcal{c}_2$, then $\rho_{\mathcal{c}_1}=a\rho_{\mathcal{c}_2}$ as $\tau(a)$ increases the amount of material by the factor $a$.
-$\mathcal{C}\cong G$ as vector spaces.
-$G$ is a subspace of $C(\mathcal{S})$.
-
+Let $\varrho : \mathcal{C} \rightarrow C(\mathcal{S})$ such that $\varrho(\mathcal{c}) \mapsto \rho_\mathcal{c}$.
+As the system is reducible, two distinct composite state must represent different distributions.
+$\forall \mathcal{c_1}, \mathcal{c_2} \in \mathcal{C}, \mathcal{c_1} \neq \mathcal{c_2} \implies \varrho(\mathcal{c_1}) \neq \varrho(\mathcal{c_1})$.
+$\varrho$ is injective.
+$\varrho$ is a bijection between $\mathcal{C}$ and $\varrho(\mathcal{C})$.
+Let $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$, then $\varrho(\mathcal{c})=\varrho(\mathcal{c}_1)+\varrho(\mathcal{c}_2)$ as the amount of material of the composition of two states is the sum of the individual amounts.
+Let $\mathcal{c}_1=\tau(a)\mathcal{c}_2$, then $\varrho(\mathcal{c}_1)=a \varrho(\mathcal{c}_2)$ as $\tau(a)$ increases the amount of material by the factor $a$.
+$\varrho$ is a homomorphism.
+$\mathcal{C}\cong \varrho(\mathcal{C}) \subseteq C(\mathcal{S})$ as vector spaces.
 \end{justification}
 
-Note that not all continuous functions correspond to physically meaningful states: we must make sure they describe a finite amount of material.
+\subsection{Invariant densities}
 
-\subsection{Coordinate invariant densities}
+In the case where the topology of $\mathcal{S}$ is not discrete, the distribution $\rho_\mathcal{c}$ is a density which is not something we directly measure.
+What we do measure experimentally are finite amounts of material.
+For example, the amount of fluid in a particular volume within a particular range of momentum specified in some units (e.g.
+mass, kg, ...).
+This means that given a composite state $\mathcal{c}$ and a set $U \subseteq \mathcal{S}$ of particle states compatible with an outcome of a process, we must be able to tell the amount of material that we will find associated with that outcome.
+That is for each open set $U \subseteq \mathcal{S}$ there will be a functional $\Lambda_U : \mathcal{C} \rightarrow \mathbbm{R}$.
 
-For each classical composite state will also tell us the total amount of material, once we fixed unit (e.g.
-mole, kg, ...).
-That is, there will be a functional $\Lambda : \mathcal{C} \rightarrow \mathbbm{R}$ that tells us how much gas we have for a particular gas configuration.
 As we combine parts, the total amount of material will be the sum of their parts.
 As we increase or decrease by a constant factor the amount of material will increase or decrease by the same factor.
-That is: $\Lambda(a_1 \mathcal{c}_1 + a_2 \mathcal{c}_2) = a_1 \Lambda(\mathcal{c}_1) + a_2 \Lambda(\mathcal{c}_2)$.
+That is: $\Lambda_U(a_1 \mathcal{c}_1 + a_2 \mathcal{c}_2) = a_1 \Lambda_U(\mathcal{c}_1) + a_2 \Lambda_U(\mathcal{c}_2)$.
 For a proper state, one that is not the difference of two states, we will expect a positive amount of material.
-So, if a density $\rho(\mathcal{S})$ is positive everywhere, then the total amount is also positive.
-This makes $\Lambda$ a positive linear functional.
-
-Being able to tell the total amount is not enough.
-Experimentally, what we do is measure the amount of material for a region of phase space.
-For example, the amount of gas within a particular region of space for a particular range of momentum.
-In general, given an outcome of our physical process (i.e.
-an open set in the topology for $\mathcal{S}$), each composite state should tell us the amount of material associated with it (i.e.
-contained in the open set of states).
-For each $U \subseteq \mathcal{S}$ we'll have a positive linear functional $\Lambda_U : \mathcal{C} \rightarrow \mathbbm{R}$ whose value is always bounded.
+So, if a density $\rho_\mathcal{c}(\mathcal{S})$ is positive everywhere, then the total amount is also positive.
+This makes $\Lambda_U$ a positive linear functional.
+As the amount of material we will measure is always finite, the functional applied to any composite state $\mathcal{c}$ over any set $U$ will be finite.
 
@@
 And the density will give us an indication of how big the region is: if we double the region, the density will half.
-In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique way to measure the size of sets of $\mathcal{S}$, a metric $d\mu$ and our positive linear functionals become integrals: $\Lambda_U (\mathcal{c}) : \int_U \rho_{\mathcal{c}} d \mu$.
+In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique measure $\mu$ for the size of sets of $\mathcal{S}$, a metric $d\mu$ and our positive linear functionals become integrals: $\Lambda_U (\mathcal{c}) : \int_U \rho_{\mathcal{c}} d \mu$.
 
-In mathematical terms, we can use the Riesz representation theorem for linear functionals to give $\mathcal{S}$ the structure of a measure space, with a suitable Borel $\sigma$-algebra and measure $\mu$.
+This intuitive picture is formalized mathematically by the Riesz representation theorem for linear functionals, which gives $\mathcal{S}$ the structure of a measure space, with a suitable Borel $\sigma$-algebra and measure $\mu$.
 
@@
 
-\begin{assump}[Classical hamiltonian assumption]\label{classical}
-	The evolution of the classical homogeneous system under study is deterministic, reversible and reducible.
+\begin{assump}[Classical Hamiltonian assumption]\label{classical}
+	The system under study is composed of an infinitesimally reducible homogeneous material and each part undergoes deterministic and reversible evolution.
 \end{assump}

### 2016-04-21 2015CompleteDerivation/completeDerivation.tex
Incorporated Josh feedback

@@
 Yet we wouldn't know, since we are not studying it: the evolution is deterministic and/or reversible only as far as the state is concerned.
-With this in mind, we introduce the following:
+With this in mind, we will restrict ourselves to the cases where the following is valid:
 
@@
 As we increase the temperature of the air around the book, its motion remains unaffected until, at some point, the book burns.
-Clearly, the non-deterministic evolution has pushed one of the states outside the set, to the point that the system is no longer recognizable.
+Clearly, the non-deterministic evolution has pushed one of the states outside the set of states, to the point that the system is no longer recognizable.
 
@@
 If I drop an anvil from a tower, it will accelerate at $9.81$~m/s$^2$; if I want the anvil to reach the ground at $x$~m/s I have to drop it from $y$~m.
-To the extent that we want to make predictions in time, we need to have a one-to-one correspondence between initial and final states.
+To the extent that we want to make predictions in time, we need to have a correspondence between initial and final states.
 
@@
 That is, we need a process for which the input settings of our preparing device determine the outgoing state of the system; and a process for which the incoming state of the system can be reconstructed by the output of the measuring device.
-That is, our system must participate in a deterministic and reversible process with the preparing and measuring device.
+That is, our system must, at least in some cases, be able to participate in a deterministic and reversible process with the preparing and measuring device.
 Without it we wouldn't be able to calibrate our experimental apparatus.
@@
 The second approach is to transport the state variables.
-The idea is to keep the connection to the initial state by labeling the future state, instead by the future values of $q^i$, by the original one.
+The idea is to keep the connection to the initial state by labeling the future state by the original value of $q^i$, instead of by the future values.
 For example, the evolved state variable will not tell us the current value of pressure, but the one for the initial conditions.
@@
 \begin{prop}\label{discrete_state_metric}
-	Let $U \subseteq \mathcal{S}$ a set of states fully identified by a set of $n$ independent discrete state variables $q^i$.
-Let $\Delta q^i = (q^i)^{-1}(U)$ the range of possibilities of each variables.
-Then $\#(\Delta q^i)=\#(\Delta \hat{q}^i) \; \forall i$ and $\#(U)=\prod\limits_{i=1}^{n}\#(\Delta q^i)=\prod\limits_{i=1}^{n}\#(\Delta \hat{q}^i)=\#(\hat{U})$ where $\hat{U}=\mathcal{T}_{\Delta t}(U)$.
 \end{prop}
@@
 We have $\hat{q}^i(\hat{\mathcal{s}}) = q^i(\mathcal{s})$.
-Therefore $\Delta q^i = (q^i)^{-1}(U) = (\hat{q}^i)^{-1}(\hat{U})=\Delta \hat{q}^i$.
+Therefore $\Delta q^i = q^i(U) = \hat{q}^i(\hat{U})=\Delta \hat{q}^i$.
 $\#(\Delta q^i)=\#(\Delta \hat{q}^i)$.
@@
 If $\Delta q^1$ and $\Delta q^2$ are ranges of possibilities for two independent variables, the total number of possibilities is $\#(\Delta q^1) \#(\Delta q^2)$.
-By construction, the set of possibilities for each independent variables $\Delta q^i$ is the same as the set of possibilities for the transported variable $\Delta \hat{q}^i$.
+By construction, the set $\Delta q^i$ of possibilities for each independent variable is the same as the set $\Delta \hat{q}^i$ of possibilities for the transported variable.
 Also, the transported variables remain independent.
@@
 At first glance, it's a bijective continuous map so we may think it can represent a deterministic and reversible evolution.
-Yet, $\Delta q = [-b, b] \supset \Delta q'$: a set of states is mapped to a proper subset (i.e.
+Yet, $\Delta q \equiv [-b, b] \supset \Delta q'$: a set of states is mapped to a proper subset (i.e.
 to fewer states) which does not make sense for a reversible process.

### 2016-04-21 2015CompleteDerivation/completeDerivation.tex
Incorporated Josh feedback

@@
 \begin{prop}\label{discrete_state_metric}
+	Let $U \subseteq \mathcal{S}$ a set of states fully identified by a set of $n$ independent discrete state variables $q^i$.
+Let $\Delta q^i = q^i(U)$ the range of possibilities of each variable.
+Then $\#(\Delta q^i)=\#(\Delta \hat{q}^i) \; \forall i$ and $\#(U)=\prod\limits_{i=1}^{n}\#(\Delta q^i)=\prod\limits_{i=1}^{n}\#(\Delta \hat{q}^i)=\#(\hat{U})$ where $\hat{U}=\mathcal{T}_{\Delta t}(U)$ and $\#$ denotes the cardinality of the given set.
 \end{prop}

### 2016-04-21 2015CompleteDerivation/completeDerivation.tex
More work on classical Hamiltonian

@@
 That is, we want to study the relationship between the state space $\mathcal{C}$ of a composite system and the state space $\mathcal{S}$ of its parts.
-In general, this is quite a complicated thing to do, that requires intimate knowledge of the system at hand.
-So we simplify our problem and study the system is made of infinitesimal homogeneous parts.
+In general, this is quite a complicated thing to do, which requires intimate knowledge of the system at hand.
+So we simplify our problem and study a system made of infinitesimal homogeneous parts.
 What we'll find is that under the additional assumption that the system is infinitesimally reducible (i.e.
@@
 \subsection{Homogeneous decomposable systems}
-The notion that a system is decomposable means the states is equipped with a rule of composition that allows to write $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$: the composite system is the sum of its part.
+The notion that a system is decomposable means the states are equipped with a rule of composition that allows one to write $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$: the composite system is the sum of its parts.
 For example, the state of a ball is equal to the state of its left and right parts.
@@
 The notion that the system is homogeneous means that the states of the composite and of the parts are not unrelated: they are made of the same material.\footnote{Whether a system is homogeneous depends on context (e.g.
-air can be thought as homogeneous if the mixture of gases does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is maintained by time evolution.} In fact, the state space $\mathcal{C}$ of all systems composed of such material will include the state of the system as well as the states of all its part (i.e $\mathcal{c}, \mathcal{c}_1, \mathcal{c}_2 \in \mathcal{C}$).
+air can be thought as homogeneous if the mixture of gases does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is maintained by time evolution.} In fact, the state space $\mathcal{C}$ of all systems composed of such material will include the state of the system as well as the states of all its parts (i.e.
+$\mathcal{c}, \mathcal{c}_1, \mathcal{c}_2 \in \mathcal{C}$).
 Since combining any two systems made of a homogeneous material will always give us a system made of the same homogeneous material (e.g.
@@
 We claim $\mathcal{C}$ is an additive monoid.
-There exist a law of composition $+ : \mathcal{C} \times \mathcal{C} \rightarrow \mathcal{C}$ that takes two states and returns one that is the physical composition of the two.
+There exists a law of composition $+ : \mathcal{C} \times \mathcal{C} \rightarrow \mathcal{C}$ that takes two states and returns one that is the physical composition of the two.
 The domain and codomain match because the material is homogeneous.
 The law is commutative $\mathcal{c}_1 +\mathcal{c}_2 = \mathcal{c}_2+\mathcal{c}_1$ and associative $(\mathcal{c}_1 + \mathcal{c}_2) + \mathcal{c}_3 = \mathcal{c}_1 + (\mathcal{c}_2 + \mathcal{c}_3)$, as it does not matter in what order we physically compose the parts.
-There exist a unique zero element $\mathcal{c} + 0 = \mathcal{c}$ and it represent the physically empty state (i.e.
+There exists a unique zero element $\mathcal{c} + 0 = \mathcal{c}$ and it represents the physically empty state (i.e.
 no amount of material).
@@
 Such inverse may introduce objects that do not properly represent a physical state, but a state change.
-A change of a physically distinguishable object is itself physically distinguishable therefore $\mathcal{C}$ is still a Hausdorff and second countable topological manifold as per \ref{statedef}.
+A change of a physically distinguishable object is itself physically distinguishable therefore $\mathcal{C}$ is still a Hausdorff and second countable topological manifold as per \ref{topologically_distinguishable}.
 We'll still call $\mathcal{C}$ state space committing an abuse of terminology.
@@
 That is, each state is fully identified by a function $\rho: \mathcal{S} \to \mathbbm{R}$ that returns the amount of material in the composite state $\mathcal{c}$ that is prepared in the particle state $\mathcal{s}$.
-For example, given a certain configuration of gas, we can tell the amount of material that is at a specific point in space with a specific value of momentum.
+For example, given a certain configuration of gas, we can tell the density of material that is at a specific point in space with a specific value of momentum.
 
 The notion that the parts are infinitesimal requires the co-domain of $\rho$ to be continuous, as opposed to discrete.
-The state space $\mathcal{S}$, instead, can be finite.
+The state space $\mathcal{S}$, instead, can have a discrete topology.
 For example, for a system composed of different tanks connected by pipes, the state of the composite system could be the overall distribution of water among the tanks.
 The amount in each tank is continuous (as we assume the water to be infinitesimally divisible) yet there are only a finite number of tanks the water can be placed into.
-As it links two physically distinguishable objects, $\rho$ is a continuous functions for reasoned discussed before.
+As it links two physically distinguishable objects, $\rho$ is a continuous function for reasons discussed before.
 
-As we combine the state of finite parts, we sum the distributions over particle states $\rho(\mathcal{s})=\rho_1(\mathcal{s})+\rho_2(\mathcal{s})$: the amount of material in the composite state is the sum of the amount of material of the parts.
+As we combine the states of different parts, we sum the distributions over particle states $\rho(\mathcal{s})=\rho_1(\mathcal{s})+\rho_2(\mathcal{s})$: the amount of material in the composite state is the sum of the amount of material of the parts.
 We can also increase or decrease the amount of material by a constant factor $\rho_1(\mathcal{s})=a\rho_2(\mathcal{s})$.
@@
 The state space $\mathcal{C}$ for a homogeneous classical (i.e.
-infinitesimally reducible) material is a vector space over $\mathbbm{R}$ isomorphic to a subspace of $C(\mathcal{S}) = \{\rho:\mathcal{S} \rightarrow \mathbbm{R} | \rho$ is continuous $\}$, where $\mathcal{S}$ is the state space of an infinitesimal amount of material.
+infinitesimally reducible) material is a vector space over $\mathbbm{R}$ isomorphic to a subspace of $C(\mathcal{S}) = \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is continuous$\}$, where $\mathcal{S}$ is the state space of an infinitesimal amount of material.
 \end{defn}
@@
 
-Consider the set of transformations $T$ that increase or decrease the amount of material in the system by a constant.
+Consider the set of transformations $T$ that increase or decrease the amount of material in the system by a constant factor.
 We claim $T$ is a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics or differential geometry sense (a physical quantity/tensor with a value for each point in space).} isomorphic to $\mathbbm{R}$.
@@
 The abelian group $\mathcal{C}$ can be extended with the operations defined by $T$, as each element $\tau \in T$ is a map $\tau : \mathcal{C} \rightarrow \mathcal{C}$.
-The map has the following properties: $(\tau_1 + \tau_2) \mathcal{c} = \tau_1 \mathcal{c} + \tau_2 \mathcal{c}$, increasing the amount of material by the sum of two constant is the same as combining the separate increases, and $\tau (\mathcal{c}_1 + \mathcal{c}_2) = \tau \mathcal{c}_1 + \tau \mathcal{c}_2$, increasing the amount of the total system is the same as the combination of the increased parts.
+The map has the following properties: $(\tau_1 + \tau_2) \mathcal{c} = \tau_1 \mathcal{c} + \tau_2 \mathcal{c}$, increasing the amount of material by the sum of two constant factors is the same as combining the separate increases, and $\tau (\mathcal{c}_1 + \mathcal{c}_2) = \tau \mathcal{c}_1 + \tau \mathcal{c}_2$, increasing the amount of the total system is the same as the combination of the increased parts.
 $\mathcal{C}$ is a module over $T$, which is a field and isomorphic to $\mathbbm{R}$.
@@
 Let $\varrho : \mathcal{C} \rightarrow C(\mathcal{S})$ such that $\varrho(\mathcal{c}) \mapsto \rho_\mathcal{c}$.
-As the system is reducible, two distinct composite state must represent different distributions.
+As the system is reducible, two distinct composite states must represent different distributions.
 $\forall \mathcal{c_1}, \mathcal{c_2} \in \mathcal{C}, \mathcal{c_1} \neq \mathcal{c_2} \implies \varrho(\mathcal{c_1}) \neq \varrho(\mathcal{c_1})$.
@@
 $\varrho$ is a homomorphism.
-$\mathcal{C}\cong \varrho(\mathcal{C}) \subseteq C(\mathcal{S})$ as vector spaces.
+$\mathcal{C}$ is isomorphic to $\varrho(\mathcal{C}) \subseteq C(\mathcal{S})$ as vector spaces.
 \end{justification}
@@
 
-In the case where the topology of $\mathcal{S}$ is not discrete, the distribution $\rho_\mathcal{c}$ is a density which is not something we directly measure.
+In the case where the topology of $\mathcal{S}$ is not discrete, the distribution $\rho_\mathcal{c}$ is a density, which is not something we directly measure.
 What we do measure experimentally are finite amounts of material.
 For example, the amount of fluid in a particular volume within a particular range of momentum specified in some units (e.g.
-mass, kg, ...).
+moles, kg, ...).
 This means that given a composite state $\mathcal{c}$ and a set $U \subseteq \mathcal{S}$ of particle states compatible with an outcome of a process, we must be able to tell the amount of material that we will find associated with that outcome.
@@
 
-As we combine parts, the total amount of material will be the sum of their parts.
-As we increase or decrease by a constant factor the amount of material will increase or decrease by the same factor.
+As we combine parts, the total amount of material will be the sum of the material of the parts.
+As we increase or decrease by a constant factor the total amount of material in a state, we will find the total amount of material of that state increased or decreased by that factor.
 That is: $\Lambda_U(a_1 \mathcal{c}_1 + a_2 \mathcal{c}_2) = a_1 \Lambda_U(\mathcal{c}_1) + a_2 \Lambda_U(\mathcal{c}_2)$.
 For a proper state, one that is not the difference of two states, we will expect a positive amount of material.
-So, if a density $\rho_\mathcal{c}(\mathcal{S})$ is positive everywhere, then the total amount is also positive.
+So, if a density $\rho_\mathcal{c}$ is positive everywhere, then the total amount is also positive.
 This makes $\Lambda_U$ a positive linear functional.
@@
 Consider a region of position and momentum in phase space.
-If we are able to spread a finite amount of material into a uniform distribution, then we know we have a finite region.
-And the density will give us an indication of how big the region is: if we double the region, the density will half.
-In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique measure $\mu$ for the size of sets of $\mathcal{S}$, a metric $d\mu$ and our positive linear functionals become integrals: $\Lambda_U (\mathcal{c}) : \int_U \rho_{\mathcal{c}} d \mu$.
+If we are able to spread a finite amount of material into a non-infinitesimal uniform distribution, then we know we have a finite region.
+And the density will give us an indication of how big the region is: if we double the region, the density will halve.
+In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique measure $\mu$ for the size of sets of $\mathcal{S}$, a metric $d\mu$ and our positive linear functionals become integrals: $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
 
@@
 The state space $\mathcal{C}$ is isomorphic to the space of continuous functions that are integrable over any subset of $\mathcal{S}$.
-That is: $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$ where $L^1_{ew}(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq \mathcal{S}\}$.
+That is: $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$ where $L^1_{ew}(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq \mathcal{S}\}$.
 \end{defn}
@@
 
-On a state space with discrete topology and a single state variable we simply have $\Lambda_U (\mathcal{c}) : \sum \limits_{q=a}^b D(q)$.
-For one dimensional continuous states, we would simply expect $\Lambda_U (\mathcal{c}) : \int_a^b \rho (q) dq$.
-By a simple change of state variable, we see that this does not work: $\rho(q')= \rho(q) dq/dq'$.
-This would make $\rho$ not a function of just the state $\mathcal{s}$, but also of the state variables we are using.
+On a state space with discrete topology and a single state variable we simply have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho(q)$.
+For states identified by a single continuous state variable we would expect $\Lambda_U (\mathcal{c}) = \int_a^b \rho (q) dq$.
+By a simple change of state variable, we see that this does not work in general: $\rho(q')= \rho(q) dq/dq'$.
+This would make $\rho$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using.
 This is inconsistent with our previous definition.
 Moreover if the transformation is not differential, the density is not even well-defined.
-This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows coordinate independent densities.
+This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows state-variable-independent densities.
+Under arbitrary change of state variables, the density has to remain the same, which means that the Jacobian of the transformation must exist and must be unitary.
 
-To do that $\mathcal{S}$ has to be, first of all, a differentiable manifold. This way, at least, the densities is well defined for all state variables.
+For the Jacobian to exist $\mathcal{S}$ has to be a differentiable manifold; that is the transformation between state variables is always differentiable.
+This way, at least, the density is well defined for any choice of state variables.
 
-If $\rho$ is to remain unchanged when one state variable changes, then one or more other state variables have to change as well so that the Jacobian of the transformation is unitary.
-This means that we cannot change physical units to all the variables as we please: they are part of a unit system.
+If $\rho$ is to remain unchanged and the Jacobian to be unitary we cannot simply change one state variable as we please.
+If we change one at least another has to change in some coordinated way such that the Jacobian of the total transformation is unitary.
+This means that we cannot change physical units of all the variables as we please: they are part of a unit system.
 For example, changing units of position will also change units of momentum.
-Let $\mathcal{Q}$ be the manifold charted by a the minimum set of state variables that define such system of units.
+Let $\mathcal{Q}$ be the manifold charted by the minimum set of state variables that define such system of units.
 An element of $\mathcal{S}$ will be a geometrical object in such space.
@@
 It cannot be a vector $v(q) e_q$, as the density $\rho(q', v')= \rho(q, v) (dq/dq')^2$ is not invariant.
-A covector $k(q) dq$ is fine as $\rho(q', k')= \rho(q, k) (dq/dq') (dq'/dq) = \rho(q, k)$.
+A covector $k(q) dq$ is instead fine as $\rho(q', k')= \rho(q, k) (dq/dq') (dq'/dq) = \rho(q, k)$.
 In other words: a covector is the simplest local geometrical object that allows coordinate independent densities.
@@
 If $\mathcal{Q}$ is one dimensional, then the state is fully identified by a pair $(q, k)$.
-$d\mu$ is proportional to $dq \wedge dk$.
+These variables are independent, so the total number of possibilities in a small region of $\mathcal{S}$ is proportional to $dq \wedge dk$.
+That is if we double the range of $dq$ or $dk$ then the number of possibilities doubles.
 We set $d\mu = \hbar dq \wedge dk$, where $\hbar$ is the proportionality constant that will depend on the unit chosen to count the possibilities of the pair $(q, k)$.
-The cardinality for states in a finite (i.e.
+That is, a unitary range of possibilities for $q$ and $k$ will give us $\hbar$ possibilities.\footnote{The actual value and physical dimensions of $\hbar$ are determined by the system of units, and should not be taken to describe some intrinsic physical property.
+Only dimensionless relationships to other physical constants, such as the fine structure constant $\alpha = e^2/\hbar c$, can possess that trait.
+That is why one can choose "natural units" for which $\hbar=1$.
+In SI units the relationship $\hbar dk = dp = m du$ between kinetic and conjugate momentum, derived later, sets the relationship between units.} The cardinality for states in a finite (i.e.
 compact) region $U \subseteq \mathcal{S}$ will be given by $\int_U \hbar dq \wedge dk$.
@@
 (i.e.
-the wedge product within d.o.f) and the orthogonality of different d.o.f.
+the wedge product within a d.o.f) and the orthogonality of different d.o.f.
 (i.e.
@@
 and the scalar product across.
-Requiring the invariance of this metric under state variables assures us the cardinalities of states and possibilities is well defined.
+Requiring the invariance of this metric under state variable changes assures us the cardinality of states and possibilities is well defined.
 
-The state space $\mathcal{S}$ is the symplectic manifold $(T^*\mathcal{Q}, \omega)$.
-This is the simplest manifold that allows us to define coordinate invariant densities.
+This gives us insight on the physical meaning of the geometrical structure $T^*\mathcal{Q}$.
+The canonical one-form $\theta=k(q) dq$ represents the geometrical object we associate with each particle state, which is not a point in $\mathcal{Q}$.
+The canonical two-form $\omega=\Sigma \, dq^i \wedge dk_i$ is the metric that quantifies the possibilities described by two given state variables.
+Note, instead, that the relationship $\omega = - \hbar d\theta$, while mathematically true, has no clear physical meaning as we have not defined what the exterior derivative on a state actually represents.
+
+We can capture the above discussion by stating that the state space $\mathcal{S}$ is the symplectic manifold $(T^*\mathcal{Q}, \omega)$.
+This is the simplest manifold that allows us to define state-variable-invariant densities.
 The metric allows us to measure the cardinality of possibilities within and across independent d.o.f.
@@
 \begin{defn}\label{classical_phase_space}
-The state space $\mathcal{S}$ for the particles of a homogeneous classical  system is a symplectic manifold form by a cotangent bundle $T^*\mathcal{Q}$ equipped with the canonical two-form $\omega$.
+The state space $\mathcal{S}$ for the particles of a homogeneous classical  system is a symplectic manifold formed by a cotangent bundle $T^*\mathcal{Q}$ equipped with the canonical two-form $\omega$.
 \end{defn}
@@
 We should note that we are starting to be in a position similar to the one discussed in \ref{discrete_state_metric} for discrete states.
-We have both a topological space and a metric that allows to count states and possibilities.
-The only thing we started from was the idea of an infinitesimally reducible system.
-That lead to a state space $\mathcal{S}$ for the infinitesimal parts, on which we must be able to define state variable independent distributions.
-Which in turn gave us degrees of freedom made by pair of conjugate variables and the symplectic metric typical of classical phase space.
+We have both a topological space and a metric that allows us to count states and possibilities.
+We recovered the metric starting from the idea of an infinitesimally reducible system.
+That led to a state space $\mathcal{S}$ for the infinitesimal parts, on which we must be able to define state-variable-independent distributions, which in turn gave us degrees of freedom made by pairs of conjugate variables and the symplectic metric typical of classical phase space.
+In short: being able to measure the amount of material is what allows us to measure the cardinality of states.
 

### 2016-04-22 2015CompleteDerivation/completeDerivation.tex
Rationale on classical assumption

@@
 \begin{rationale}
+
+	% 1. We need a process that can interact with parts
+	% 2. State cannot be full description. Cannot stop interacting with the rest of the universe, we need unstated part to keep interacting (no measurement)
+	% 3. Complete isolation means no single universe. Cannot have meaningful determinism at different level of scale.
+	% 4. Evolution that exchange information means systems cannot always be isolated
+	% 5.
+It's an approximation.
+But the approximation breaks down the more the assumption is taken literally.
+The two cases that can make the assumption hold break down in the case of infinitesimal pieces.
+The set of states of the described object is "big" compared to the set of states of the probe.
+If both infinitesimal homogeneous pieces, that cannot hold.
+The laws are not symmetric so that the probe is more affecte by the system under sudy.
+If both infinitesimal homogeneous pieces, the law is symmetric.
+	% 6. A detrev system can't tell time.
+
+	% x = v_0 t + x_0 
+	% v = v_0
+
+	% Problems. Due to parts that are independent
+	%           Due to parts that are infinitesimal
+	%           Due to everything is state
+
+	% 1. Need a process for parts, not always have it. (1)
+	% 2. Isolation (and detrev) becomes untenable as parts are infinitesimal (5)
+	% 3. If it were tenable, we would have separate universes (2,3,4)
+	%    Detrev individual parts are not part of same universe (3)
+	%    If parts interact deterministically, still no good: can (2) decompose locally in d.o.f. And infinitesimal parts are local
+	%    Completely detrev system cannot be interacted with
+	% => Tension between two desires: system is isolated so that we can define independent state, system is not isolated so that we can interact with it.
+	% 4. A system that is detrev at all levels cannot tell time (6)
+
 	This is clearly a \emph{simplifying} assumption, and it is instructive to understand when it breaks down.
@@
 	The first problem is methodological.
-As we saw before, we need access to a deterministic and reversible process to be able to study a system and define a state.
-For two balls, we can imagine to isolate some pieces (small enough to be considered infinitesimal in respect to the ball but big enough to contain enough molecules to be considered homogenous) and then describe the collision between the two by describing what happens at each piece.
-The classical assumption holds.
-For an electron and a photon, we cannot take pieces of the electron or the photon, study them in isolation and then describe how each part moves during Compton scattering.
-The classical assumption does not hold: we do not have suitable physical processes at our disposal.
+To be able to talk about the state of a part we need a physical process that is able to distinguish them.
+For a pool ball we can imagine marking one spot with a different color.
+This allows us to track it as the ball moves or collides with other balls.
+For an electron, instead, we do not have a way to mark a piece.
+In fact when two electron scatter we can't even tell which is which, let alone what part went where.
+The classical assumption does not hold because we do not have suitable physical processes at our disposal.
 
-	The second problem is more conceptual.
-As we saw before, a system cannot be fully isolated from the environment.
-A certain amount of unstated part needs to remain so that the state of the system can be insulated from the external non-deterministic interaction.
-In practice, the state of each piece of our material is still given by the average of a collection of a large amount of molecules, their respective motion depending not just on the state of the system but on the external environment.
-The classical assumption can therefore hold only if we assume that the description of each pieces is, again, not complete.
-Assuming it at a fundamental level would mean that the motion of each and every molecule of the system is only influenced by the state of the system, no external factor counts.
-As such, it would be a contradiction with it being a physical system: one we can study experimentally.
+	The second, more conceptual, problem is that the idealization that allow us to hold Assumption \ref{detrevass} break down for infinitesimal parts.
+One case where a system is sufficiently isolated to define a state is when it is "big enough".
+For example, the cannonball is big compared to the photons that scatter on its surface.
+Clearly this breaks down for infinitesimal parts: as we keep dividing the material, we will find an amount that will be affected by the interaction with the environment, and that is not going to be sufficiently isolated.
+The other case where Assumption \ref{detrevass} holds is where the interaction is not symmetrical.
+For example, the electric charge of the system under study are much larger than the charges in the environment.
+Therefore the electromagnetic field generated can be considered sufficiently isolated.
+This also breaks down: as the parts are homogeneous they have will have the same charge density and the interaction needs to be symmetrical.
+
+	In other words: the definition of a state requires some kind of asymmetry between system under study and environment.
+If we say that the system is really just infinitesimal and homogeneous pieces, then the claim to that asymmetry is lost.
+Therefore our justification for holding the assumption is that the pieces are small compared to some scale but large compared to another.
+For example, water is a homogeneous material if we consider it made up of parts sufficiently small for the scale of the problem at hand but big enough to contain a large number of molecules.
+
+	Other conceptual problems stem from two competing requirements.
+One one side we want the system under study to be isolated so that we can define independent states.
+On the other side we do not want the system isolated so that we can define physical process on it.
+For example, if we were to take the state of an infinitesimal part to really be its full description, with no unstated part, then each piece would evolve independently, oblivious to the other parts.
+Each part would essentially live in its own separate universe.
+
+	If we assume the whole is deterministic and reversible but each part is not, then their evolution can depend on the states of the other parts.
+But this does not solve the problem.
+Let's ignore the question of how we were able to define the states of the parts.
+Let's also ignore that the type of forces one would use (i.e.
+the ones with scalar and vector potentials) are precisely the ones we derive by assuming the pieces are not interacting with each other.
+We simply need to note that we can always locally separate the evolution into independent degrees of freedom.
+For example, the position and momentum of two particle may effect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way too see that is in terms of the transported variables: they retain the original value, they clearly evolve independently.
+And since the pieces are infinitesimal, a local decomposition is all that is needed.
+While such description may be cumbersome to achieve in practice, conceptually is still possible, so one still ends up with a set of degrees of freedom that evolve independently, oblivious to each other.
+
+	The same type of problem arises on the whole system when we consider its interaction to our probe or other parts of the environment.
+Say $a \in A$ is the state of our system and $b \in B$ the state of our probe, $\hat{a}=f(a)$ because the system is deterministic, $\hat{b}=g(a,b)$ because the probe needs to be affected by the system, then $(a, b) \mapsto (\hat{a}, \hat{b})$ is not a reversible map.
+In other words: we can't expect to have deterministic and reversible evolution at all levels of decomposable systems where the parts interact in any physically meaningful way.
+
+	One final conceptual problem is with time itself: a deterministic and reversible system cannot tell time.
+If it did, we would be able to find a quantity that changes through time but is invariant under state variable changes.
+The problem is that deterministic and reversible time evolution is equivalent to a state variable change.
+So, all quantities that are invariant under state variable changes are also invariant under time evolution.
 
 	While ultimately flawed, the classical assumption can be considered valid for a great number of macroscopic system, and that is why is very useful.
-One should be cautioned, though, that nowhere in the assumption the spatial extent of the system is mentioned.
-We may as well have a macroscopic system where a clear independent state cannot be assigned to each part, and the assumption would not hold.
+Understanding its shortcoming will help us later to see how the quantum assumption solves, at least partially, some of these issues.
+
+	As a final note: we caution in automatically assuming that the classical assumption is valid for macroscopic systems.
+Nowhere in the assumption the spatial extent of the system is mentioned.
+It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, and the assumption would not hold.
 \end{rationale}

### 2016-04-22 2015CompleteDerivation/completeDerivation.tex
Cleaned up leading up to rationale for classical assumption

@@
 
-Before discussing time evolution, we should investigate what assuming our state is infinitesimally reducible means in terms of the fundamental model discussed in Assumption \ref{detrevass}.
-
 \subsection{Classical Hamiltonian assumption}
 
-% Det/rev for infinitesimal part
-% Inconsistent: 1.
-X=AxB A'=F(A) B'=F(B) Isolated sub-systems.
-Does not work for measuring.
-2.
-A'=F(A) B'=F(A,B) Ideal non-interfering measurement: deterministic but not reversible.
-3.
-A'=F(A,B) B'=F(A,B) X det/rev whole, but A not deterministic itself.
-But if you consider X a sub-system of X and something else, you can make the same argument.
-Whole universe deterministic, but none of its parts.
-4.
-A=As x Au ; B=Bs x Bu ; As' = F(As) Au=F(X) Bu=F(X) Bs = F(As, Au) ; more realistic picture.
-Still non-reversible (if As' = F(As), then nothing else can be F(As) and have bijective map).
-
-We further characterize the relationship between the whole and the parts with the following assumption:
+Now that we have fully characterized what we mean by a classical system, we can now stipulate the following:
 
@@
 
-In other words: if we know the state and the evolution for the whole system, we also know the evolution for all its parts; if we know the state and evolution for all the parts, we know the state and the evolution for the whole system.
-By smaller we do not mean spatial extent, but of smaller amount (e.g.
-less mass): even at the same spatial location we can imagine that there are infinitely many components (e.g.
-the mass, as a number, is continuously divisible(.
+\begin{rationale}
+	The idea is that time evolution specifies a map for the state space $\mathcal{S}$ of the infinitesimal parts.
+And knowing how the parts evolve, we also know how to evolve the whole, that elements in $\mathcal{C}$.
 
-\begin{rationale}
-
-	% 1. We need a process that can interact with parts
-	% 2. State cannot be full description. Cannot stop interacting with the rest of the universe, we need unstated part to keep interacting (no measurement)
-	% 3. Complete isolation means no single universe. Cannot have meaningful determinism at different level of scale.
-	% 4. Evolution that exchange information means systems cannot always be isolated
-	% 5.
-It's an approximation.
-But the approximation breaks down the more the assumption is taken literally.
-The two cases that can make the assumption hold break down in the case of infinitesimal pieces.
-The set of states of the described object is "big" compared to the set of states of the probe.
-If both infinitesimal homogeneous pieces, that cannot hold.
-The laws are not symmetric so that the probe is more affecte by the system under sudy.
-If both infinitesimal homogeneous pieces, the law is symmetric.
-	% 6. A detrev system can't tell time.
-
-	% x = v_0 t + x_0 
-	% v = v_0
-
-	% Problems. Due to parts that are independent
-	%           Due to parts that are infinitesimal
-	%           Due to everything is state
-
-	% 1. Need a process for parts, not always have it. (1)
-	% 2. Isolation (and detrev) becomes untenable as parts are infinitesimal (5)
-	% 3. If it were tenable, we would have separate universes (2,3,4)
-	%    Detrev individual parts are not part of same universe (3)
-	%    If parts interact deterministically, still no good: can (2) decompose locally in d.o.f. And infinitesimal parts are local
-	%    Completely detrev system cannot be interacted with
-	% => Tension between two desires: system is isolated so that we can define independent state, system is not isolated so that we can interact with it.
-	% 4. A system that is detrev at all levels cannot tell time (6)
-
-	This is clearly a \emph{simplifying} assumption, and it is instructive to understand when it breaks down.
+	Consistently with what we said in Assumption \ref{detrevass}, if we defined a state for the infinitesimal parts, then one such deterministic and reversible evolution should exist.
+Yet, the idea that we can assign states to infinitesimal parts should be considered only a \emph{simplifying} assumption, and it is instructive to understand when it breaks down.
 

### 2016-04-25 2015CompleteDerivation/completeDerivation.tex
Reviewed rationale for classical assumption

@@
 \begin{rationale}
-	The idea is that time evolution specifies a map for the state space $\mathcal{S}$ of the infinitesimal parts.
-And knowing how the parts evolve, we also know how to evolve the whole, that elements in $\mathcal{C}$.
+	The idea is that time evolution specifies a map for the state space $\mathcal{S}$ of each infinitesimal part.
+Knowing how the parts evolve tells us how a composite state $\mathcal{c}$ evolves as well.
 
-	Consistently with what we said in Assumption \ref{detrevass}, if we defined a state for the infinitesimal parts, then one such deterministic and reversible evolution should exist.
-Yet, the idea that we can assign states to infinitesimal parts should be considered only a \emph{simplifying} assumption, and it is instructive to understand when it breaks down.
+	Consistently with what we said in Assumption \ref{detrevass}, if we defined a state for each particle, then a deterministic and reversible evolution on such state must exist.
+Yet, the idea that we can assign states to infinitesimal parts should be considered only a \emph{simplifying} assumption, and it is instructive to understand when and how it breaks down.
 
 	The first problem is methodological.
-To be able to talk about the state of a part we need a physical process that is able to distinguish them.
+To be able to talk about the states of a part we need a physical process that is able to distinguish between them.
 For a pool ball we can imagine marking one spot with a different color.
@@
 For an electron, instead, we do not have a way to mark a piece.
-In fact when two electron scatter we can't even tell which is which, let alone what part went where.
-The classical assumption does not hold because we do not have suitable physical processes at our disposal.
+In fact when two electrons scatter we can't even tell which is which, let alone what part went where.
+The classical assumption may not hold because we do not have suitable physical processes at our disposal.
 
-	The second, more conceptual, problem is that the idealization that allow us to hold Assumption \ref{detrevass} break down for infinitesimal parts.
+	The second, more conceptual, problem is that the idealization that allow us to hold Assumption \ref{detrevass} break down as the parts get smaller.
 One case where a system is sufficiently isolated to define a state is when it is "big enough".
-For example, the cannonball is big compared to the photons that scatter on its surface.
-Clearly this breaks down for infinitesimal parts: as we keep dividing the material, we will find an amount that will be affected by the interaction with the environment, and that is not going to be sufficiently isolated.
+Recall the cannonball whose motion is sufficiently unaffected by the photons that scatter off its surface.
+Clearly this breaks down for infinitesimal parts: as we keep dividing the material, we will find an amount that will be affected by the interaction with the environment, and therefore not sufficiently isolated.
 The other case where Assumption \ref{detrevass} holds is where the interaction is not symmetrical.
-For example, the electric charge of the system under study are much larger than the charges in the environment.
-Therefore the electromagnetic field generated can be considered sufficiently isolated.
-This also breaks down: as the parts are homogeneous they have will have the same charge density and the interaction needs to be symmetrical.
+For example, suppose the electric charge of the system under study is much larger than the charges in the environment.
+The electromagnetic field generated can be considered determined only by our system with good approximation.
+But this also breaks down: if part of the environment is made of the same homogeneous material the charge density will be the same and the interaction between particles will be symmetrical.
 
 	In other words: the definition of a state requires some kind of asymmetry between system under study and environment.
-If we say that the system is really just infinitesimal and homogeneous pieces, then the claim to that asymmetry is lost.
-Therefore our justification for holding the assumption is that the pieces are small compared to some scale but large compared to another.
+If we try to assume that the whole universe is really made of infinitesimal homogeneous parts, then the claim to that asymmetry is lost.
+Therefore our justification for holding the assumption has to be that the pieces are small compared to some scale but large compared to another.
 For example, water is a homogeneous material if we consider it made up of parts sufficiently small for the scale of the problem at hand but big enough to contain a large number of molecules.
 
-	Other conceptual problems stem from two competing requirements.
-One one side we want the system under study to be isolated so that we can define independent states.
-On the other side we do not want the system isolated so that we can define physical process on it.
+	Other conceptual problems stem from the following two competing requirements: on one side we want the system under study to be isolated so that we can define independent states; on the other side we do not want the system isolated so that we can define physical process on it.
 For example, if we were to take the state of an infinitesimal part to really be its full description, with no unstated part, then each piece would evolve independently, oblivious to the other parts.
-Each part would essentially live in its own separate universe.
+Each particle would essentially live in its own separate universe.
 
-	If we assume the whole is deterministic and reversible but each part is not, then their evolution can depend on the states of the other parts.
-But this does not solve the problem.
+	If we assume the whole system is deterministic and reversible but each part is not, then their evolution can depend on the states of the other parts.
+But this does not solve the problem entirely.
 Let's ignore the question of how we were able to define the states of the parts.
-Let's also ignore that the type of forces one would use (i.e.
-the ones with scalar and vector potentials) are precisely the ones we derive by assuming the pieces are not interacting with each other.
-We simply need to note that we can always locally separate the evolution into independent degrees of freedom.
-For example, the position and momentum of two particle may effect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way too see that is in terms of the transported variables: they retain the original value, they clearly evolve independently.
+Let's also ignore that the type of forces one would typically use (i.e.
+ones that conserve energy or ones with scalar and vector potentials) are precisely the ones we derive by assuming the pieces are not interacting with each other.
+The problem is that we can always locally separate the evolution into independent degrees of freedom.
+For example, the position and momentum of two particles may effect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way too see that is in terms of the transported variables: they retain the original value, they clearly evolve independently.
 And since the pieces are infinitesimal, a local decomposition is all that is needed.
-While such description may be cumbersome to achieve in practice, conceptually is still possible, so one still ends up with a set of degrees of freedom that evolve independently, oblivious to each other.
+While such description may be cumbersome to achieve in practice, conceptually is still possible.
+So, even if the particles are in principle interacting, the system is described by degrees of freedom that evolve independently, oblivious to each other.
 
 	The same type of problem arises on the whole system when we consider its interaction to our probe or other parts of the environment.
-Say $a \in A$ is the state of our system and $b \in B$ the state of our probe, $\hat{a}=f(a)$ because the system is deterministic, $\hat{b}=g(a,b)$ because the probe needs to be affected by the system, then $(a, b) \mapsto (\hat{a}, \hat{b})$ is not a reversible map.
+Say $a \in A$ is the state of our system and $b \in B$ the state of our probe, the future state $\hat{a}=f(a)$ because the system is deterministic, but $\hat{b}=g(a,b)$ because the probe needs to be affected by the system.
+This means $(a, b) \mapsto (\hat{a}, \hat{b})$ is not a reversible map.
 In other words: we can't expect to have deterministic and reversible evolution at all levels of decomposable systems where the parts interact in any physically meaningful way.
@@
 The problem is that deterministic and reversible time evolution is equivalent to a state variable change.
+In fact, at a fixed time, we can choose to describe the system with the transported state variables of any past or future times.
 So, all quantities that are invariant under state variable changes are also invariant under time evolution.
 
-	While ultimately flawed, the classical assumption can be considered valid for a great number of macroscopic system, and that is why is very useful.
+	In the light of what we discussed, we cannot take the classical assumption to be fundamental, in the sense that we cannot take it to apply to all of physics.
+While ultimately flawed, the  assumption can be considered valid for a great number of macroscopic system, and that is why is very useful.
 Understanding its shortcoming will help us later to see how the quantum assumption solves, at least partially, some of these issues.
 
-	As a final note: we caution in automatically assuming that the classical assumption is valid for macroscopic systems.
-Nowhere in the assumption the spatial extent of the system is mentioned.
-It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, and the assumption would not hold.
+	As a final note: we caution in automatically thinking that the classical assumption is valid for all macroscopic systems.
+It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, in which case the assumption would not hold.
 \end{rationale}
+
+\subsection{Hamiltonian mechanics}
 

### 2016-04-25 2015CompleteDerivation/completeDerivation.tex
Added section on Hamilton equations

@@
 
-We are now ready to capture the elements of our discussion and their properties through mathematical definitions.
+We are finally ready to write the equations of motions. As per \ref{detrevmap} our evolution is at least a self-homeomorphism $f:\mathcal{S} \leftrightarrow \mathcal{S}$. But this is not enough.
+
+The evolution must map the distribution point-wise so that the value associated at the initial state is the same as the value.
+All of the system that starts in $\mathcal{s}$ has to end up in $\hat{\mathcal{s}}$.
+In math terms $\rho_{\hat{\mathcal{c}}} (\hat{\mathcal{s}}) = \rho_\mathcal{c}(\mathcal{s})$.
+
+In the same way, if $U \in \mathcal{S}$ is a set of initial particle state and $\Lambda_U(\mathcal{c})$ is the amount of material associated to that set, then we expect it to be equal to the amount of material $\Lambda_{\hat{U}}(\hat{\mathcal{c}})$ associated with the set of final states $\hat{U} \in \mathcal{S}$.
+
+But probably the best way to look at it is that initial and final set of states have to contain the same cardinality of states and possibilities, which where defined by our metric $\omega$.
+Therefore the area in each degree of freedom, which represents the number of possibilities in said d.o.f., needs to be mapped to an equal area in the transported d.o.f.
+(i.e.
+the d.o.f.
+defined by the transported state variables).
+Independent d.o.f.
+must remain independent, and therefore transported orthogonal d.o.f.
+remain orthogonal.
+This means that the product of possibilities of independent d.o.f.
+is also conserved.
+These statements are the physical justification of the Gromov's non-squeezing theorem and Liouville's theorem.
+And they give intuitive insight on the geometry of Hamiltonian systems.
+
+Mathematically, under the classical Hamilotnian assumption,  deterministic and reversible evolution is an self-isometry. That is, it does not just preserve the topology but also the metric.
 
 \begin{defn}\label{canonical_transformation}
-	A deterministic and reversible process is a diffeomorphism that preserves $\omega$, that is a canonical transformation.
+	A deterministic and reversible evolution map for a homogeneous classical material is a canonical transformation and self-isometry, $\mathcal{T}_{\Delta t}:(T^*\mathcal{Q}, \omega) \rightarrow (T^*\mathcal{Q}, \omega)$.
 \end{defn}
@@
 \begin{justification}
-	TODO
+	We claim $\mathcal{T}_{\Delta t}$ is a self-homeomorphism. $\mathcal{T}_{\Delta t}$ is a deterministic and reversible evolution map and \ref{detrevmap}.
+	We claim $\mathcal{T}_{\Delta t}$ is an self-isometry on $(T^*\mathcal{Q}, \omega)$.
+The Jacobian for $\mathcal{T}_{\Delta t}$ has to exist and be different from zero in order for the distribution on final states to be defined.
+$\mathcal{T}_{\Delta t}$ is a diffeomorphism.
+$\omega$ is the metric for the cardinality of possibilities.
+A deterministic and reversible process conserves the number of states and possibilities.
+$\omega$ is invariant under a deterministic and reversible evolution map.
 \end{justification}
 
-\begin{defn}\label{hamiltonian}
-	A continuous deterministic and reversible process admits a potential $H$, and the laws of evolution are of the form (Hamilton equations).
-\end{defn}
+In the special case of continuous time evolution we can prove the following:
 
-\begin{justification}
-	TODO
-\end{justification}
+\begin{prop}\label{hamiltonian}
+	A continuous deterministic and reversible process for a homogeneous classical material admits a Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ that allows to write the laws of evolution as
+\begin{align*}
+d_{t}q^i &= \partial_{p_i} H \\
+d_{t}p_i &= - \partial_{q^i} H
+\end{align*}
+\end{prop}
+
+\begin{proof}
+We claim the vector field $S \in T\mathcal{Q}$ for the infinitesimal displacement $S = d\mathcal{s}/dt$ admits a potential $H$ such that, in component notation, $S^{a} \omega_{ab} = \partial_{b}H$.
+Let $v, w \in T_P\mathcal{Q}$ be two vectors defined at a point $P \in \mathcal{Q}$.
+Let $v^a, w^b$ their components in terms of unified state variables.
+Let $v', w' \in T_{\hat{P}}\mathcal{Q}$ the vector transported by the evolution map.
+For \ref{canonical_transformation} we have $v^{a} \omega_{ab} w^{b} = v'^{a} \omega_{ab} w'^{b}$.
+The evolution is a diffeomorphism, so we have $\xi(t+dt)^a = \xi^a + S^a dt$.
+The vector components change according to $v'^a = \partial_b \xi(t+dt)^a v^b = (\delta^a_b + \partial_b S^a dt) v^b$.
+We have:
+\begin{align*}
+v^{a} \omega_{ab} w^{b} &= v'^{a} \omega_{ab} w'^{b}  \\
+&= (v^{a} + \partial_{c} S^{a} v^{c} dt) \omega_{ab} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
+&= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} \\
+&+ v^{a} \omega_{ab} \partial_{d} S^{b} w^{d}) dt + O(dt^2)
+\end{align*}
+We define $S_{b} \equiv S^{a} \omega_{ab}$.
+The metric is invariant iff $v^{c} w^{b} \partial_{c} S_{b} - v^{a} w^{d} \partial_{d} S_{a} = 0$.
+The relationship must true for and pair of vector, therefore $\partial_{a} S_{b} - \partial_{b} S_{a} = curl(S_{a}) = 0$.
+$S$ is a curl free vector field and admits a potential $H$ such that $S_{a} = \partial_{a}H$.
+
+We claim the state variables evolve according to Hamilton's equations.
+$S^{a} \omega_{ab} = d_t\xi^a \omega_{ab} = \partial_{b}H$.
+For $b=\{1,...,n\}$ we have $d_tp_i (-1) = \partial_{q_i} H$.
+For $a=\{n+1,...,2n\}$ we have $d_tq^i (+1) = \partial_{p_i} H$.
+\end{proof}
+
+We recognize the familiar set of Hamilton's equations.
+They are the set of equation to describe the deterministic and reversible motion of the infinitesimal parts of an infinitesimally reducible homogeneous material.
+The challenge in their derivation mainly lies in the necessary use of different branches of mathematics (e.g.
+topology, measure theory, differential geometry, symplectic geometry).
+The conceptual meaning, on the other hand, is hopefully straight forward.
+
+\section{Time dependent evolution}
+
+\section{Kinematic equivalence}
 
 \section{Quantum systems}
+
 %Note on composite systems and distinguishability.

### 2016-04-29 2015CompleteDerivation/completeDerivation.tex
Another clarification after discussing with Josh

@@
 While this is a common enough definition, we need to be clear how this applies to the unstated part.
-That is, the evolution of the unstated part under the same process is \emph{always} non-deterministic (and non-reversible) in the sense that the state of the system does not determine its evolution.
+Note that the concept of determinism outlined here is context dependent because the state itself represents only the part of the system that we choose to (or can) describe.
+In this sense, the unstated part is \emph{always} non-deterministic (and non-reversible) in the sense that the state of the system does not determine its evolution.
 For example, suppose we define the state as position and momentum of the center of mass of a cannonball.

### 2016-04-29 2015CompleteDerivation/completeDerivation.tex
Reviewed classical Hamiltonian assumption with Christine

@@
 \begin{defn}\label{classical_phase_space}
-The state space $\mathcal{S}$ for the particles of a homogeneous classical  system is a symplectic manifold formed by a cotangent bundle $T^*\mathcal{Q}$ equipped with the canonical two-form $\omega$.
+The state space $\mathcal{S}$ for the particles of a homogeneous classical material is a symplectic manifold formed by a cotangent bundle $T^*\mathcal{Q}$ equipped with the canonical two-form $\omega$.
 \end{defn}
@@
 
-Now that we have fully characterized what we mean by a classical system, we can now stipulate the following:
+Now that we have fully characterized what we mean by a classical system, we can stipulate the following:
 
@@
 
-	The second, more conceptual, problem is that the idealization that allow us to hold Assumption \ref{detrevass} break down as the parts get smaller.
-One case where a system is sufficiently isolated to define a state is when it is "big enough".
+	The second, more conceptual, problem is that the idealizations that allow us to hold Assumption \ref{detrevass} break down as the parts get smaller.
+One such idealization is when the system under study is "big enough".
 Recall the cannonball whose motion is sufficiently unaffected by the photons that scatter off its surface.
 Clearly this breaks down for infinitesimal parts: as we keep dividing the material, we will find an amount that will be affected by the interaction with the environment, and therefore not sufficiently isolated.
-The other case where Assumption \ref{detrevass} holds is where the interaction is not symmetrical.
+The other idealization is when the interaction between system and environment is not symmetrical.
 For example, suppose the electric charge of the system under study is much larger than the charges in the environment.
@@
 	In other words: the definition of a state requires some kind of asymmetry between system under study and environment.
-If we try to assume that the whole universe is really made of infinitesimal homogeneous parts, then the claim to that asymmetry is lost.
+If we try to assume that both the system and environment are really made of infinitesimal homogeneous parts, then the claim to that asymmetry is lost.
 Therefore our justification for holding the assumption has to be that the pieces are small compared to some scale but large compared to another.
@@
 
-	Other conceptual problems stem from the following two competing requirements: on one side we want the system under study to be isolated so that we can define independent states; on the other side we do not want the system isolated so that we can define physical process on it.
+	Other conceptual problems stem from the following two competing requirements: on one side we want the system under study to be isolated so that we can define independent states; on the other side we do not want the system isolated so that we can define physical processes on it.
 For example, if we were to take the state of an infinitesimal part to really be its full description, with no unstated part, then each piece would evolve independently, oblivious to the other parts.
-Each particle would essentially live in its own separate universe.
+Each particle would essentially reside in its own separate physical universe.
 
@@
 The problem is that we can always locally separate the evolution into independent degrees of freedom.
-For example, the position and momentum of two particles may effect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way too see that is in terms of the transported variables: they retain the original value, they clearly evolve independently.
+For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value, they clearly evolve independently.
 And since the pieces are infinitesimal, a local decomposition is all that is needed.
-While such description may be cumbersome to achieve in practice, conceptually is still possible.
+While such description may be cumbersome to achieve in practice, conceptually it is still possible.
 So, even if the particles are in principle interacting, the system is described by degrees of freedom that evolve independently, oblivious to each other.
 
-	The same type of problem arises on the whole system when we consider its interaction to our probe or other parts of the environment.
+	The same type of problem arises on the whole system when we consider its interaction with our probe or other parts of the environment.
 Say $a \in A$ is the state of our system and $b \in B$ the state of our probe, the future state $\hat{a}=f(a)$ because the system is deterministic, but $\hat{b}=g(a,b)$ because the probe needs to be affected by the system.
@@
 	In the light of what we discussed, we cannot take the classical assumption to be fundamental, in the sense that we cannot take it to apply to all of physics.
-While ultimately flawed, the  assumption can be considered valid for a great number of macroscopic system, and that is why is very useful.
-Understanding its shortcoming will help us later to see how the quantum assumption solves, at least partially, some of these issues.
-
-	As a final note: we caution in automatically thinking that the classical assumption is valid for all macroscopic systems.
+While ultimately flawed, the  assumption can be considered valid for a great number of macroscopic systems, and that is why it is very useful.
+Understanding its shortcomings will help us later to see how the quantum assumption solves, at least partially, some of these issues.
+
+	As a final note: we caution against automatically thinking that the classical assumption is valid for all macroscopic systems.
 It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, in which case the assumption would not hold.

### 2016-04-29 2015CompleteDerivation/completeDerivation.tex
Another pass at equations of motion

@@
 For an electron, instead, we do not have a way to mark a piece.
-In fact when two electrons scatter we can't even tell which is which, let alone what part went where.
+In fact when two electrons scatter we can't even tell which is which, let alone what portion went where.
 The classical assumption may not hold because we do not have suitable physical processes at our disposal.
@@
 
-We are finally ready to write the equations of motions. As per \ref{detrevmap} our evolution is at least a self-homeomorphism $f:\mathcal{S} \leftrightarrow \mathcal{S}$. But this is not enough.
-
-The evolution must map the distribution point-wise so that the value associated at the initial state is the same as the value.
+We are finally ready to write the equations of motion. As per \ref{detrevmap} our evolution is at least a self-homeomorphism $f:\mathcal{S} \leftrightarrow \mathcal{S}$. But this is not enough.
+
+The evolution must map the distribution point-wise so that the value associated at the initial state is the same as the value at the final state.
 All of the system that starts in $\mathcal{s}$ has to end up in $\hat{\mathcal{s}}$.
@@
 
-In the same way, if $U \in \mathcal{S}$ is a set of initial particle state and $\Lambda_U(\mathcal{c})$ is the amount of material associated to that set, then we expect it to be equal to the amount of material $\Lambda_{\hat{U}}(\hat{\mathcal{c}})$ associated with the set of final states $\hat{U} \in \mathcal{S}$.
-
-But probably the best way to look at it is that initial and final set of states have to contain the same cardinality of states and possibilities, which where defined by our metric $\omega$.
-Therefore the area in each degree of freedom, which represents the number of possibilities in said d.o.f., needs to be mapped to an equal area in the transported d.o.f.
+In the same way we expect the total amount of material to be conserved.
+If $U \in \mathcal{S}$ is a set of initial particle states and $\Lambda_U(\mathcal{c})$ is the amount of material associated with that set, then we expect it to be equal to the amount of material $\Lambda_{\hat{U}}(\hat{\mathcal{c}})$ associated with the set of final states $\hat{U} \in \mathcal{S}$.
+
+But probably the best way to look at it is that initial and final sets of states have to possess the same cardinality of states and possibilities, which were defined by our metric $\omega$.
+Therefore the area within a degree of freedom, which represents the number of possibilities in said d.o.f., needs to be mapped to an equal area within the transported d.o.f.
 (i.e.
@@
 is also conserved.
-These statements are the physical justification of the Gromov's non-squeezing theorem and Liouville's theorem.
+These statements are the physical justification of Gromov's non-squeezing theorem and Liouville's theorem.
 And they give intuitive insight on the geometry of Hamiltonian systems.
 
-Mathematically, under the classical Hamilotnian assumption,  deterministic and reversible evolution is an self-isometry. That is, it does not just preserve the topology but also the metric.
+Mathematically, under the classical Hamiltonian assumption,  deterministic and reversible evolution is a self-symplectomorphism (or self-isometry or canonical transformation depending on your math training).
+That is, it does not just preserve the topology but also the symplectic metric $\omega$.
 
 \begin{defn}\label{canonical_transformation}
-	A deterministic and reversible evolution map for a homogeneous classical material is a canonical transformation and self-isometry, $\mathcal{T}_{\Delta t}:(T^*\mathcal{Q}, \omega) \rightarrow (T^*\mathcal{Q}, \omega)$.
+	A deterministic and reversible evolution map for a homogeneous classical material is self-symplectomorphism.
+That is: $\mathcal{T}_{\Delta t}: T^*\mathcal{Q} \rightarrow T^*\mathcal{Q}$ and $\mathcal{T}_{\Delta t}^*\omega = \omega$ where $\mathcal{T}_{\Delta t}^*$ is the pullback of $\mathcal{T}_{\Delta t}$.
 \end{defn}
@@
 \begin{justification}
-	We claim $\mathcal{T}_{\Delta t}$ is a self-homeomorphism. $\mathcal{T}_{\Delta t}$ is a deterministic and reversible evolution map and \ref{detrevmap}.
-	We claim $\mathcal{T}_{\Delta t}$ is an self-isometry on $(T^*\mathcal{Q}, \omega)$.
-The Jacobian for $\mathcal{T}_{\Delta t}$ has to exist and be different from zero in order for the distribution on final states to be defined.
+	We claim $\mathcal{T}_{\Delta t}$ is a self-homeomorphism on $T^*\mathcal{Q}$.
+The state space for particles of a homogeneous classical material is $T^*\mathcal{Q}$ by \ref{classical_phase_space} $\mathcal{T}_{\Delta t}$ is a deterministic and reversible evolution map and \ref{detrevmap}.
+
+	We claim $\mathcal{T}_{\Delta t}$ is a symplectomorphism on $(T^*\mathcal{Q}, \omega)$.
+The distribution on final states must be defined.
+The Jacobian for $\mathcal{T}_{\Delta t}$ exists and is non-zero.
 $\mathcal{T}_{\Delta t}$ is a diffeomorphism.
+A deterministic and reversible process conserves the number of states and possibilities.
 $\omega$ is the metric for the cardinality of possibilities.
-A deterministic and reversible process conserves the number of states and possibilities.
 $\omega$ is invariant under a deterministic and reversible evolution map.
+$\mathcal{T}_{\Delta t}$ is a symplectomorphism by definition.
 \end{justification}
@@
 
+It's important to realize that, during the derivation, multiple mathematical features (e.g.
+invariant densities, cotangent bundle for phase space, symplectomorphism) where justified by the same physical assumption.
+As such, the math itself does not give us a conceptually unified picture, it gets lost.
+This is one of the reason we believe that starting from the physical description is objectively better if we are to come to a better understanding of our physical theories.
+
+Note that we could have taken different approaches.
+For example, we could have appealed to information theory and asked our invariant distributions to preserve Shannon's information entropy, as no information should be gained or lost during a deterministic and reversible process.
+Or we could have appealed to statistical mechanics and required that the determinant of the covariance matrix be conserved, as deterministic and reversible process should be defined at the same level of precision.
+With suitable treatment of independent d.o.f.
+both these approaches recover Hamiltonian mechanics as well.
+While the full treatment goes outside the scope of this work, we want to underline that the definitions used in this work go a long way to building bridges at the core of different mathematical and scientific disciplines.
+
 \section{Time dependent evolution}

### 2016-04-29 2015CompleteDerivation/completeDerivation.tex
Starting to outline time dependent evolution

@@
 
+We now generalize our discussion to include time dependent dynamics.
+This is needed when the evolution map itself changes in time or when a change of state variable is time dependent (e.g.
+changing to a moving frame).
+
+To do this we will extend phase space to include the temporal degree of freedom.
+We'll find that the dynamics is described by the relativistic Hamiltonian mechanics.
+We'll also find that particle states divide into standard and anti-states depending whether time is aligned or anti-aligned with the affine parameter.
+
+\subsection{Time changes}
+
+The discussion so far has been limited to the case where both the set of states and the evolution map are defined at equal time and never change throughout the evolution.
+This is too restrictive as there are very reasonable situation in which this does not hold.
+
+The first obvious case is that is that the map may be time dependent.
+Note that this does not affect our definition of determinism and reversibility: we still can tell final state from initial state and vice-versa.
+We just need to specify the time as well.
+At this point we have no way to do that.
+For example, the Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ is just a function of the state.
+
+The second case is when we perform a transformation for the time parameter $t'=t'(t)$.
+The equations of motion transform and we have $d_{t}t' d_{t'}q^i = \partial_{p_i} H$.
+As both $q^i$ and $H$ are independent of time, they cannot be redefined to include $d_{t}t'$.
+Physically the transformation is introducing fictitious forces that are not conservative, so they cannot be expressed by a Hamiltonian.
+But this means that we have an ill defined mathematical framework as the notion of deterministic and reversibility is not defined in a way that is independent of time transformations.
+We need a framework that is capable of handling the fictitious forces as well.
+
+The third case is when we perform a state variable transformation $q'=q'(q,t)$ that is time dependent (e.g.
+$x'=x+vt$).
+Our composite state distribution $\rho(q)$ becomes $\rho(q',t)$: it is no longer defined at equal time.
+This means that both our measure $\mu$ and our metric $d\mu$ need to be modified.
+
+The point is that we can't simply stick time in the Hamiltonian and expect everything to works out.
+But we can't ignore the problem either as the three cases outlined are pretty useful situation to study.
+So we need to go back to our definitions and amend them properly.
+
+\subsection{Extended phase space}
+
+% Extend Q to include time -> M. Extended phase space is a hypersurface T*M.
+
+% Call s (affine parameter) the parameter of the evolution.
+Distribution defined on a n-dimensional surface in T*M.
+Density does not change under time transformation.
+Use t'=t + \Delta t(q^i, p_i) so that t' is constant.
+Now we are at equal t': the metric is \Sigma dq^i \wedge dp_i.
+Metric is the projection on the space orthogonal to t and E.
+That is \Sigma dq^i \wedge dp_i - dt \wege dE.
+
+% definition, simplectic metric on phase space
+
+% justification.
+Need to be able to define integration for everywhere and space and on compact time.
+Metric exists.
+Must be a two-form.
+Two-form is skewsymmetric.
+All skew symmetric forms are decomposable in hyperplanes.
+T*Q is a subspace of T*M.
+Only one hyperplane left.
+Metric must be \Sigma dq^i \wedge dp_i \pm dt \wedge dE.
+Can't be positive, as ()t, E) is not an independent d.o.f.
+
+\subsection{Relativistic Hamiltonian mechanics}
+
+% find equation of motion
+
+% find constraint
+
+\subsection{anti-states}
+
+% relationship between s and t must be monotonic. Hamiltonian partitions extended phase space. Distribution must be always zero at the boundary.
+
 \section{Kinematic equivalence}

### 2016-04-29 2015CompleteDerivation/completeDerivation.tex
Comment on conservative forces and detrev
First stab at the measure justification

@@
 
-As we combine parts, the total amount of material will be the sum of the material of the parts.
-As we increase or decrease by a constant factor the total amount of material in a state, we will find the total amount of material of that state increased or decreased by that factor.
+As we combine parts, or increase the amount of material in each part, the total amount of material found will have to be consistent with those operations.
 That is: $\Lambda_U(a_1 \mathcal{c}_1 + a_2 \mathcal{c}_2) = a_1 \Lambda_U(\mathcal{c}_1) + a_2 \Lambda_U(\mathcal{c}_2)$.
@@
 \begin{justification}
-	TODO
+	We claim there exists a positive linear functional $\Lambda_U : \mathcal{C} \rightarrow \mathbbm{R}$ for each open set $U \subseteq \mathcal{S}$.
+Let $\Lambda_{U} : \mathcal{C} \rightarrow \mathbbm{R}$ the functional that returns the amount of material in $\mathcal{c}$ compatible with the outcome associated with the open set $U \subseteq \mathcal{S}$.
+Let $U_1, U_2$ two disjoint sets, then $\Lambda_U = \Lambda_{U_1} + \Lambda_{U_2}$ as the material found in $U$ must be either in $U_1$ or $U_2$.
+$\Lambda_U$ is linear.
+$\Lambda_U(a_1 \mathcal{c}_1 + a_2 \mathcal{c}_2) = a_1 \Lambda_U(\mathcal{c}_1) + a_2 \Lambda_U(\mathcal{c}_2)$ as it has to be consistent with the operations of composing states and increasing/decrease the amount of material by a factor.
+$\Lambda_U$ is positive.
+If the value of distribution at each particle state is positive then the total amount of material is positive.
+
+	We claim that $\mathcal{S}$ is endowed with a Borel measure $\mu$.
+$\mathcal{S}$ is locally compact as it is a manifold.
+At each point there exists an compact open a neighborhood $U$.
+$U$ with the subspace topology is a compact Hausdorff space on which is defined a positive linear function $\Lambda_U$ on the linear vector space $\mathcal{C}$ isomorphic to the space of continuous compactly supported real valued functions defined on $U$.
+For the Riesz representation theorem for linear functionals, there exists a unique Borel measure $\mu_U$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu_U$.
+$\Lambda_U$ is the sum for disjoint subsets, therefore $\Lambda_V (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu_U + \int_{V\backslash U} \rho_{\mathcal{c}} d \mu_V$ for any open set $V \supseteq U$.
+The measure is unique in the family of all functionals $\Lambda_U$.
+$\Lambda_U(\mathcal{c})$ is finite for any open, therefore $\Lambda_{\mathcal{S}} (\mathcal{c}) = \int_{\mathcal{S}} \rho_{\mathcal{c}} d \mu$.
+The measure is unique on the whole space.
 \end{justification}
@@
 They are the set of equation to describe the deterministic and reversible motion of the infinitesimal parts of an infinitesimally reducible homogeneous material.
+In other words: \textbf{the forces that conserve energy (i.e.
+the value of a sutiable Hamiltonian) are exactly the ones that provide deterministic and reversible motion}.
 The challenge in their derivation mainly lies in the necessary use of different branches of mathematics (e.g.

### 2016-05-09 2015CompleteDerivation/completeDerivation.tex
First full pass at justification for integration

@@
 \begin{defn}\label{integration}
-	The state space $\mathcal{S}$ for the particles of a homogeneous classical material is endowed with a measure $\mu$.
+	The state space $\mathcal{S}$ for the particles of a homogeneous classical material is endowed with a Borel measure $\mu$.
 The state space $\mathcal{C}$ is isomorphic to the space of continuous functions that are integrable over any subset of $\mathcal{S}$.
@@
 Let $\Lambda_{U} : \mathcal{C} \rightarrow \mathbbm{R}$ the functional that returns the amount of material in $\mathcal{c}$ compatible with the outcome associated with the open set $U \subseteq \mathcal{S}$.
-Let $U_1, U_2$ two disjoint sets, then $\Lambda_U = \Lambda_{U_1} + \Lambda_{U_2}$ as the material found in $U$ must be either in $U_1$ or $U_2$.
 $\Lambda_U$ is linear.
@@
 $\Lambda_U$ is positive.
-If the value of distribution at each particle state is positive then the total amount of material is positive.
-
-	We claim that $\mathcal{S}$ is endowed with a Borel measure $\mu$.
+If the value of the distribution for each particle state is positive then the total amount of material is positive.
+
+	We claim that $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$.
+Let $U \subseteq \mathcal{S}$ be the open set of particle states associated with an outcome.
+Let $\mathcal{c} \in \mathcal{S}$ a composite state.
+As physically we always work with finite quantities, the amount of material of $\mathcal{c}$ associated with the outcome $U$ must be finite.
+$|\Lambda_{U}(\mathcal{c})| < \infty$.
+
+	We claim that $\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2}$ $\forall $ open $U_1, U_2 \in \mathcal{S}$.
+Let $U_1, U_2 \in \mathcal{S}$ be the open sets of particles states associated with two outcomes.
+Suppose $U_1 \cap U_2 = 0$, $\Lambda_{U_1 \cup U_2} = \Lambda_{U_1} + \Lambda_{U_2}$ as the material found in $U_1 \cup U_2$ must be either in $U_1$ or $U_2$.
+Suppose $U_1 \cap U_2 \neq 0$, $\Lambda_{U_1 \cup U_2} = \Lambda_{U_1} + \Lambda_{U_2} - \Lambda_{U_1 \cap U_2}$ as the sum of the material associated with each outcome will double count the intersection.
+
+	We claim that each $U \subseteq \mathcal{S}$ is endowed with a Borel measure $\mu_U$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu_U$.
 $\mathcal{S}$ is locally compact as it is a manifold.
-At each point there exists an compact open a neighborhood $U$.
+At each point there exists a compact open a neighborhood $U$.
 $U$ with the subspace topology is a compact Hausdorff space on which is defined a positive linear function $\Lambda_U$ on the linear vector space $\mathcal{C}$ isomorphic to the space of continuous compactly supported real valued functions defined on $U$.
 For the Riesz representation theorem for linear functionals, there exists a unique Borel measure $\mu_U$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu_U$.
-$\Lambda_U$ is the sum for disjoint subsets, therefore $\Lambda_V (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu_U + \int_{V\backslash U} \rho_{\mathcal{c}} d \mu_V$ for any open set $V \supseteq U$.
-The measure is unique in the family of all functionals $\Lambda_U$.
-$\Lambda_U(\mathcal{c})$ is finite for any open, therefore $\Lambda_{\mathcal{S}} (\mathcal{c}) = \int_{\mathcal{S}} \rho_{\mathcal{c}} d \mu$.
+
+	We claim that $\mathcal{S}$ is endowed with a unique Borel measure $\mu$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
+Let $V \subseteq \mathcal{S}$ an open set.
+Let $U \subseteq V$ not necessarily open.
+As $\rho_{\mathcal{c}}$ is continuous, $\int_{U} \rho_{\mathcal{c}} d \mu_V = \int_{\mathrm{Int} U} \rho_{\mathcal{c}} d \mu_V$.
+We extend $\Lambda_U \equiv \Lambda_{\mathrm{Int}U}$ to all sets.
+We have $\Lambda_V (\mathcal{c}) = \int_V \rho_{\mathcal{c}} d \mu_V = \int_U \rho_{\mathcal{c}} d \mu_V + \int_{V\backslash U} \rho_{\mathcal{c}} d \mu_V$.
+We also have $\Lambda_V (\mathcal{c}) = \Lambda_U (\mathcal{c}) + \Lambda_{V\backslash U} (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu_U + \int_{\mathrm{Int} V \backslash U} \rho_{\mathcal{c}} d \mu_{\mathrm{Int} V \backslash U} $.
+As the two have to be equal for all $U, V, \rho_{\mathcal{c}}$, we must have $\int_{U} \rho_{\mathcal{c}} d \mu_U = \int_{U} \rho_{\mathcal{c}} d \mu_V$.
+Therefore $d\mu$ is unique for each open set in the family of all functionals.
+As $\mathcal{S}$ is locally compact, we can find a cover of compact sets.
+$\Lambda_{\mathcal{S}}(\mathcal{c})$ will be the sum of contributions of the elements of the cover and their intersections.
+As $\Lambda_U$ is always finite, $\Lambda_{\mathcal{S}}$ is a finite sum of finite contributions.
+Therefore we have  $\Lambda_{\mathcal{S}} (\mathcal{c}) = \int_{\mathcal{S}} \rho_{\mathcal{c}} d \mu$.
 The measure is unique on the whole space.
+
+	We claim $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$ where $L^1_{ew}(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq \mathcal{S}\}$.
+As per \ref{classical_vector space} $\mathcal{C}$, is isomorphic to a subspace $G \subseteq C(\mathcal{S})$ as a vector space over $\mathbbm{R}$.
+As $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$, then $G \subseteq L^1_{ew}(\mathcal{S}, \mu)$.
+TODO: can we prove mathematically that we have to span the whole space to be closed under linear combinations?
 \end{justification}

### 2016-05-10 2015CompleteDerivation/completeDerivation.tex
Review integration justification

@@
 	Let $U \subseteq \mathcal{S}$ a set of states fully identified by a set of $n$ independent discrete state variables $q^i$.
-Let $\Delta q^i = q^i(U)$ the range of possibilities of each variable.
+Let $\Delta q^i \equiv q^i(U)$ the range of possibilities of each variable.
 Then $\#(\Delta q^i)=\#(\Delta \hat{q}^i) \; \forall i$ and $\#(U)=\prod\limits_{i=1}^{n}\#(\Delta q^i)=\prod\limits_{i=1}^{n}\#(\Delta \hat{q}^i)=\#(\hat{U})$ where $\hat{U}=\mathcal{T}_{\Delta t}(U)$ and $\#$ denotes the cardinality of the given set.
@@
 That is: it is made of arbitrarily small parts, which we call particles, and the state of the whole system coincides with the sum of the states of the parts.
-Given the state space $\mathcal{S}$ for the infinitesimal parts, a composite state $\mathcal{c}$ will tell us the amount of material for each possible particle state.
+Given the state space $\mathcal{S}$ for the infinitesimal parts, which we assume it to be a manifold consistently with \ref{continuous_state_space}, a composite state $\mathcal{c}$ will tell us the amount of material for each possible particle state.
 That is, each state is fully identified by a function $\rho: \mathcal{S} \to \mathbbm{R}$ that returns the amount of material in the composite state $\mathcal{c}$ that is prepared in the particle state $\mathcal{s}$.
@@
 
-The notion that the parts are infinitesimal requires the co-domain of $\rho$ to be continuous, as opposed to discrete.
+The notion that the parts are infinitesimal requires the co-domain of $\rho$ to be a real number, as opposed to an integer.
 The state space $\mathcal{S}$, instead, can have a discrete topology.
@@
 The amount in each tank is continuous (as we assume the water to be infinitesimally divisible) yet there are only a finite number of tanks the water can be placed into.
-As it links two physically distinguishable objects, $\rho$ is a continuous function for reasons discussed before.
+As it links two physically distinguishable objects, $\rho$ is a continuous function as discussed before.
 
@@
 \begin{justification}
-We claim $\mathcal{C}$ is an abelian group. $\mathcal{C}$ is a decomposable homogeneous system. $\mathcal{C}$ is an abelian group  by \ref{reducible_state_space}.
+We claim $\mathcal{C}$ is an abelian group. $\mathcal{C}$ is the state space for a decomposable homogeneous system. $\mathcal{C}$ is an abelian group  by \ref{reducible_state_space}.
 
@@
 \rho_{\mathcal{c}}:\mathcal{S} \rightarrow \mathbbm{R}$ returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$.
-$\rho \in C(\mathcal{S})$ by \ref{continuous_map}.
+$\rho_{\mathcal{c}} \in C(\mathcal{S})$ by \ref{continuous_map}.
 Let $\varrho : \mathcal{C} \rightarrow C(\mathcal{S})$ such that $\varrho(\mathcal{c}) \mapsto \rho_\mathcal{c}$.
@@
 And the density will give us an indication of how big the region is: if we double the region, the density will halve.
-In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique measure $\mu$ for the size of sets of $\mathcal{S}$, a metric $d\mu$ and our positive linear functionals become integrals: $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
+In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique measure $\mu$ for the size of sets of $\mathcal{S}$.
+From the measure we get a metric $d\mu$ and our positive linear functionals become integrals: $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
 
@@
 \begin{defn}\label{integration}
-	The state space $\mathcal{S}$ for the particles of a homogeneous classical material is endowed with a Borel measure $\mu$.
+	Let $\mathcal{S}$ be the state space for the particles of a homogeneous classical material.
+Let each state $\mathcal{s} \in \mathcal{S}$ be fully identified by a finite number of independent state variables.
+Then $\mathcal{S}$ is endowed with a Borel measure $\mu$.
 The state space $\mathcal{C}$ is isomorphic to the space of continuous functions that are integrable over any subset of $\mathcal{S}$.
@@
 \begin{justification}
-	We claim there exists a positive linear functional $\Lambda_U : \mathcal{C} \rightarrow \mathbbm{R}$ for each open set $U \subseteq \mathcal{S}$.
-Let $\Lambda_{U} : \mathcal{C} \rightarrow \mathbbm{R}$ the functional that returns the amount of material in $\mathcal{c}$ compatible with the outcome associated with the open set $U \subseteq \mathcal{S}$.
+	We claim there exists a positive linear functional $\Lambda_U : \mathcal{C} \rightarrow \mathbbm{R}$ for each $U \subseteq \mathcal{S}$ for which $\Lambda_U = \Lambda_{\mathrm{int}(U)}$.
+Let $\Lambda_{U} : \mathcal{C} \rightarrow \mathbbm{R}$ be the functional that returns the amount of material in $\mathcal{c}$ compatible with the outcome associated with the open set $U \subseteq \mathcal{S}$.
 $\Lambda_U$ is linear.
@@
 If the value of the distribution for each particle state is positive then the total amount of material is positive.
+Let $U \subseteq \mathcal{S}$ not necessarily open.
+Define $\Lambda_U$ as $\Lambda_{\mathrm{int}(U)}$.
 
 	We claim that $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$.
-Let $U \subseteq \mathcal{S}$ be the open set of particle states associated with an outcome.
+Let $U \subseteq \mathcal{S}$ be an open set of particle states associated with an outcome.
 Let $\mathcal{c} \in \mathcal{S}$ a composite state.
-As physically we always work with finite quantities, the amount of material of $\mathcal{c}$ associated with the outcome $U$ must be finite.
+The amount of material of $\mathcal{c}$ associated with the outcome $U$ must be finite, as physically we always work with finite quantities.
 $|\Lambda_{U}(\mathcal{c})| < \infty$.
-
-	We claim that $\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2}$ $\forall $ open $U_1, U_2 \in \mathcal{S}$.
-Let $U_1, U_2 \in \mathcal{S}$ be the open sets of particles states associated with two outcomes.
+Let $U \subseteq \mathcal{S}$ not necessarily open.
+The same argument applies to $\mathrm{int}(U)$.
+
+	We claim that $\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2}$ $\forall $ $U_1, U_2 \in \mathcal{S}$.
+Let $U_1, U_2 \in \mathcal{S}$ be open sets of particle states associated with two outcomes.
 Suppose $U_1 \cap U_2 = 0$, $\Lambda_{U_1 \cup U_2} = \Lambda_{U_1} + \Lambda_{U_2}$ as the material found in $U_1 \cup U_2$ must be either in $U_1$ or $U_2$.
 Suppose $U_1 \cap U_2 \neq 0$, $\Lambda_{U_1 \cup U_2} = \Lambda_{U_1} + \Lambda_{U_2} - \Lambda_{U_1 \cap U_2}$ as the sum of the material associated with each outcome will double count the intersection.
-
-	We claim that each $U \subseteq \mathcal{S}$ is endowed with a Borel measure $\mu_U$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu_U$.
-$\mathcal{S}$ is locally compact as it is a manifold.
-At each point there exists a compact open a neighborhood $U$.
-$U$ with the subspace topology is a compact Hausdorff space on which is defined a positive linear function $\Lambda_U$ on the linear vector space $\mathcal{C}$ isomorphic to the space of continuous compactly supported real valued functions defined on $U$.
+Let $U_1, U_2 \in \mathcal{S}$ not necessarily open.
+The same argument applies to their interiors.
+
+	We claim that at each $\mathcal{s} \in \mathcal{S}$ there exists a neighborhood $U \subseteq \mathcal{S}$ endowed with a Borel measure $\mu_U$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu_U$.
+Each $\mathcal{s} \in \mathcal{S}$ is identified by a finite number of independent state variables.
+$\mathcal{S}$ is a manifold.
+$\mathcal{S}$ is locally compact.
+At each point there exists a compact neighborhood $U$.
+$U$ with the subspace topology is a compact Hausdorff topological space on which is defined a positive linear functional $\Lambda_U$ on the linear vector space $\mathcal{C}$ isomorphic to the space of continuous real valued functions defined on $U$.
 For the Riesz representation theorem for linear functionals, there exists a unique Borel measure $\mu_U$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu_U$.
+As $\rho_{\mathcal{c}}$ is continuous $\int_{U} \rho_{\mathcal{c}} d \mu_U = \int_{\mathrm{int}(U)} \rho_{\mathcal{c}} d \mu_U$ which is consistent with $\Lambda_U = \Lambda_{\mathrm{int}(U)}$.
 
 	We claim that $\mathcal{S}$ is endowed with a unique Borel measure $\mu$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
-Let $V \subseteq \mathcal{S}$ an open set.
-Let $U \subseteq V$ not necessarily open.
-As $\rho_{\mathcal{c}}$ is continuous, $\int_{U} \rho_{\mathcal{c}} d \mu_V = \int_{\mathrm{Int} U} \rho_{\mathcal{c}} d \mu_V$.
-We extend $\Lambda_U \equiv \Lambda_{\mathrm{Int}U}$ to all sets.
+Let $U \subseteq V \subseteq \mathcal{S}$ two compact subsets.
 We have $\Lambda_V (\mathcal{c}) = \int_V \rho_{\mathcal{c}} d \mu_V = \int_U \rho_{\mathcal{c}} d \mu_V + \int_{V\backslash U} \rho_{\mathcal{c}} d \mu_V$.
-We also have $\Lambda_V (\mathcal{c}) = \Lambda_U (\mathcal{c}) + \Lambda_{V\backslash U} (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu_U + \int_{\mathrm{Int} V \backslash U} \rho_{\mathcal{c}} d \mu_{\mathrm{Int} V \backslash U} $.
+We also have $\Lambda_V (\mathcal{c}) = \Lambda_U (\mathcal{c}) + \Lambda_{V\backslash U} (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu_U + \int_{V \backslash U} \rho_{\mathcal{c}} d \mu_{\overline{V \backslash U}} $.
 As the two have to be equal for all $U, V, \rho_{\mathcal{c}}$, we must have $\int_{U} \rho_{\mathcal{c}} d \mu_U = \int_{U} \rho_{\mathcal{c}} d \mu_V$.
-Therefore $d\mu$ is unique for each open set in the family of all functionals.
+Therefore $d\mu$ is unique and does not depend on the choice of neighborhood.
 As $\mathcal{S}$ is locally compact, we can find a cover of compact sets.
@@
 As $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$, then $G \subseteq L^1_{ew}(\mathcal{S}, \mu)$.
-TODO: can we prove mathematically that we have to span the whole space to be closed under linear combinations?
+$G \subseteq C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$.
+Let $\rho_\mathcal{c}$ be the distribution associated with the state $\mathcal{c}$.
+Let $\rho \in C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$ be a continuous function everywhere integral.
+If $\rho_\mathcal{c} \neq \rho$ then $\exists U \in \mathcal{S} | \int_{U} \rho_\mathcal{c} d \mu \neq \int_{U} \rho d \mu$.
+There exist an outcome in $\mathcal{S}$ that can physically distinguish the two distributions.
+There must be a state $\mathcal{c}_1 \in \mathcal{C} \; | \; \varrho_(\mathcal{c}_1)=\rho$.
+$\varrho : \mathcal{C} \rightarrow C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$ is surjective.
+$\varrho$ is an isomorphism between $\mathcal{C}$ and $C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$ as vector spaces.
 \end{justification}

### 2016-05-11 2015CompleteDerivation/completeDerivation.tex
Adding a section for purely mathematical demonstrations

@@
 \newtheorem{prop}{Proposition}[section]
+\newtheorem{thrm}{Theorem}[section]
 
@@
 Therefore $d\mu$ is unique and does not depend on the choice of neighborhood.
-As $\mathcal{S}$ is locally compact, we can find a cover of compact sets.
-$\Lambda_{\mathcal{S}}(\mathcal{c})$ will be the sum of contributions of the elements of the cover and their intersections.
-As $\Lambda_U$ is always finite, $\Lambda_{\mathcal{S}}$ is a finite sum of finite contributions.
-Therefore we have  $\Lambda_{\mathcal{S}} (\mathcal{c}) = \int_{\mathcal{S}} \rho_{\mathcal{c}} d \mu$.
-The measure is unique on the whole space.
+Let $\{U_\alpha\}_{\alpha \in A}$ a cover of $\mathcal{S}$ where each $U_\alpha$ is compact.
+This is possible as $\mathcal{S}$ is locally compact.
+We have
+\begin{align*}
+	\Lambda_{\mathcal{S}}(\mathcal{c}) &= \sum \limits_{\alpha \in A} \Lambda_{U_\alpha}(\mathcal{c}) - \sum \limits_{\alpha \in A} \sum \limits_{\beta \in A}^{\beta \neq\alpha} \Lambda_{U_\alpha \cap U_\beta}(\mathcal{c}) \\
+	&= \sum \limits_{\alpha \in A} \int_{U_\alpha} \rho_{\mathcal{c}} d\mu - \sum \limits_{\alpha \in A} \sum \limits_{\beta \in A}^{\beta \neq\alpha} \int_{U_\alpha \cap U_\beta}(\mathcal{c}) \rho_{\mathcal{c}} d\mu \\
+	 &= \int_{\mathcal{S}} \rho_{\mathcal{c}} d\mu
+\end{align*}
+	as $\Lambda_U$ is always finite. The measure is unique on the whole space.
 
@@
 \end{justification}
+
+As we identified $C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$ as the set of distributions that are physically meaningful, it's useful to show how commonly used functions spaces do not fit the bill.
+We do this briefly by finding counterexample.
+The set of continuous functions $C(\mathcal{S})$ would include functions that diverge at infinity.
+The set of continuous functions with compact support $C_C(\mathcal{S})$ would exclude Gaussians.
+The set of continuous functions that vanish at infinity $C_0(\mathcal{S})$ includes function whose integral is infinite.
+Schwartz space $S(\mathcal{S})$ requires functions to be infinitely smooth and we do not see a physical requirement for that.
 
@@
 
+TODO
+
 % Extend Q to include time -> M. Extended phase space is a hypersurface T*M.
@@
 
+TODO
+
 % find equation of motion
@@
 \section{Kinematic equivalence}
+TODO
 
 \section{Quantum systems}
+
+TODO
 
@@
 
+\section{Appendix: mathematical support}
+
+In this section we include mathematical demonstration in support of the physical arguments of the paper.
+
+\begin{thrm}\label{everywhere_integrable_is_lebesgue_integrable}
+	Let $\mathcal{L}(S,\mu) \equiv \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\}$.
+$\mathcal{L}(S,\mu)=L^1(S,\mu)$ where $L^1(S,\mu) = \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; \int_{S} |\rho| d\mu < \infty \}$
+\end{thrm}
+
+\begin{proof}
+	We claim $\mathcal{L}(S,\mu) \supseteq L^1(S,\mu)$.
+Let $\rho \in L^1(S,\mu)$.
+Let $U \subseteq S$.
+$\lVert \rho \rVert_U \equiv \int_{U} |\rho| d\mu < \int_{\mathcal{S}} |\rho| d\mu < \infty$.
+$\lVert \rho \rVert_U = \lVert \rho^+ \rVert_U + \lVert \rho^- \rVert_U$.
+$\lVert \rho^+ \rVert_U < \infty$ and $\lVert \rho^- \rVert_U < \infty$.
+$|\int_{U} \rho d\mu| = |\int_{U} \rho^+ d\mu - \int_{U} \rho^- d\mu| = |\int_{U} \rho^+ d\mu| - |\int_{U} \rho^- d\mu| = \lVert \rho^+ \rVert_U - \lVert \rho^- \rVert_U < \infty$.
+$\rho \in\mathcal{L}(S,\mu)$.
+
+	We claim $L^1(S,\mu) \supseteq \mathcal{L}(S,\mu)$.
+Let $\rho \in \mathcal{L}(S,\mu)$.
+Let $S^+ \equiv  \{ s \in S \; | \; \rho(s) > 0\}$.
+Let $S^- \equiv  \{ s \in S \; | \; \rho(s) < 0\}$.
+$|\int_{S^+} \rho d\mu| = |\int_{S} \rho^+ d\mu| = \lVert \rho^+ \rVert < \infty$ and $|\int_{S^-} \rho d\mu| = |\int_{S} \rho^- d\mu| = \lVert \rho^- \rVert < \infty$.
+$\lVert \rho \rVert = \lVert \rho^+ \rVert_U + \lVert \rho^- \rVert_U < \infty$.
+$\rho \in L^1(S,\mu)$.
+
+	We claim $L^1(S,\mu) = \mathcal{L}(S,\mu)$. $\mathcal{L}(S,\mu) \supseteq L^1(S,\mu)$ and $L^1(S,\mu) \supseteq \mathcal{L}(S,\mu)$.	
+\end{proof}
+
+\begin{thrm}[Modified Riesz theorem]\label{modified_riesz_theorem}
+	Let $\mathcal{L}(S,\mu) \equiv \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\}$.
+$\mathcal{L}(S,\mu)=L^1(S,\mu)$ where $L^1(S,\mu) = \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; \int_{S} |\rho| d\mu < \infty \}$
+\end{thrm}
+
 \begin{thebibliography}{0}

### 2016-05-11 2015CompleteDerivation/completeDerivation.tex
Added Riesz theorem extension in math appendix

@@
 
+\begin{thrm}[Extended Riesz theorem]\label{extended_riesz_theorem}
+	Let $(S, \tau)$ a locally compact Hausdorff space.
+Let $\Lambda = \{\Lambda_U : C(S) \rightarrow \mathbbm{R}\}_{U \subseteq S}$ a family of positive linear functionals such that $\forall U \subseteq S \; \Lambda_U = \Lambda_{\mathrm{int}(U)}$ and $\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2} \; \forall U_1, U_2 \subseteq S$.
+Then there exists a unique Borel measure $\mu$ such that $\Lambda_U (\rho) = \int_{U} \rho d\mu \; \forall \rho \in C(S)$.
+\end{thrm}
+
+\begin{proof}
+	We claim that at each $s \in S$ there exists a compact neighborhood $U \subseteq S$ endowed with a Borel measure $\mu_U$ such that $\Lambda_U (\rho) = \int_U \rho d \mu_U \; \forall \rho \in C(S)$.
+$S$ is locally compact.
+$\forall s \in S \; \exists U \subseteq S$ such that $U$ is compact.
+$U$ with the subspace topology is a compact Hausdorff topological space on which is defined a positive linear functional $\Lambda_U : C(S) \rightarrow \mathbbm{R}$.
+For the Riesz representation theorem for linear functionals, there exists a unique Borel measure $\mu_U$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho d \mu_U \; \forall \rho \in C(S)$.
+
+	We claim that $S$ is endowed with a unique Borel measure $\mu$ such that $\Lambda_U (\rho) = \int_U \rho d \mu$.
+Let $U \subseteq V \subseteq \mathcal{S}$ two compact subsets.
+Let $\rho \in C(S)$.
+$\Lambda_V (\rho) = \int_V \rho d \mu_V = \int_U \rho d \mu_V + \int_{V\backslash U} \rho d \mu_V$.
+Also $\Lambda_V (\rho) = \Lambda_U (\rho) + \Lambda_{V\backslash U} (\rho) = \int_U \rho d \mu_U + \int_{V \backslash U} \rho d \mu_{\overline{V \backslash U}} $.
+Therefore $\int_{U} \rho_{\mathcal{c}} d \mu_U = \int_{U} \rho_{\mathcal{c}} d \mu_V$.
+$d\mu$ is unique and does not depend on the choice of neighborhood.
+$S$ is locally compact.
+It admits a cover $\{U_\alpha\}_{\alpha \in A}$ where each $U_\alpha$ is compact.
+	\begin{align*}
+	\Lambda_{\mathcal{S}}(\mathcal{c}) &= \sum \limits_{\alpha \in A} \Lambda_{U_\alpha}(\mathcal{c}) - \sum \limits_{\alpha \in A} \sum \limits_{\beta \in A}^{\beta \neq\alpha} \Lambda_{U_\alpha \cap U_\beta}(\mathcal{c}) \\
+	&= \sum \limits_{\alpha \in A} \int_{U_\alpha} \rho_{\mathcal{c}} d\mu - \sum \limits_{\alpha \in A} \sum \limits_{\beta \in A}^{\beta \neq\alpha} \int_{U_\alpha \cap U_\beta}(\mathcal{c}) \rho_{\mathcal{c}} d\mu \\
+	&= \int_{\mathcal{S}} \rho_{\mathcal{c}} d\mu
+	\end{align*}
+	The measure is unique on the whole space.
+\end{proof}
+
 \begin{thrm}\label{everywhere_integrable_is_lebesgue_integrable}
@@
 
-\begin{thrm}[Modified Riesz theorem]\label{modified_riesz_theorem}
-	Let $\mathcal{L}(S,\mu) \equiv \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\}$.
-$\mathcal{L}(S,\mu)=L^1(S,\mu)$ where $L^1(S,\mu) = \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; \int_{S} |\rho| d\mu < \infty \}$
-\end{thrm}
-
 \begin{thebibliography}{0}

### 2016-05-11 2015CompleteDerivation/completeDerivation.tex
Reviewed integration that now uses the math theorems in appendix

@@
 The state space $\mathcal{C}$ is isomorphic to the space of continuous functions that are integrable over any subset of $\mathcal{S}$.
-That is: $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$ where $L^1_{ew}(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq \mathcal{S}\}$.
+That is: $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ where $L^1(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \; \int_{\mathcal{S}} |\rho| d\mu < \infty \}$.
 \end{defn}
@@
 
-	We claim that at each $\mathcal{s} \in \mathcal{S}$ there exists a neighborhood $U \subseteq \mathcal{S}$ endowed with a Borel measure $\mu_U$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu_U$.
+	We claim that $\mathcal{S}$ is endowed with a unique Borel measure $\mu$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
 Each $\mathcal{s} \in \mathcal{S}$ is identified by a finite number of independent state variables.
@@
 $\mathcal{S}$ is locally compact.
-At each point there exists a compact neighborhood $U$.
-$U$ with the subspace topology is a compact Hausdorff topological space on which is defined a positive linear functional $\Lambda_U$ on the linear vector space $\mathcal{C}$ isomorphic to the space of continuous real valued functions defined on $U$.
-For the Riesz representation theorem for linear functionals, there exists a unique Borel measure $\mu_U$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu_U$.
-As $\rho_{\mathcal{c}}$ is continuous $\int_{U} \rho_{\mathcal{c}} d \mu_U = \int_{\mathrm{int}(U)} \rho_{\mathcal{c}} d \mu_U$ which is consistent with $\Lambda_U = \Lambda_{\mathrm{int}(U)}$.
-
-	We claim that $\mathcal{S}$ is endowed with a unique Borel measure $\mu$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
-Let $U \subseteq V \subseteq \mathcal{S}$ two compact subsets.
-We have $\Lambda_V (\mathcal{c}) = \int_V \rho_{\mathcal{c}} d \mu_V = \int_U \rho_{\mathcal{c}} d \mu_V + \int_{V\backslash U} \rho_{\mathcal{c}} d \mu_V$.
-We also have $\Lambda_V (\mathcal{c}) = \Lambda_U (\mathcal{c}) + \Lambda_{V\backslash U} (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu_U + \int_{V \backslash U} \rho_{\mathcal{c}} d \mu_{\overline{V \backslash U}} $.
-As the two have to be equal for all $U, V, \rho_{\mathcal{c}}$, we must have $\int_{U} \rho_{\mathcal{c}} d \mu_U = \int_{U} \rho_{\mathcal{c}} d \mu_V$.
-Therefore $d\mu$ is unique and does not depend on the choice of neighborhood.
-Let $\{U_\alpha\}_{\alpha \in A}$ a cover of $\mathcal{S}$ where each $U_\alpha$ is compact.
-This is possible as $\mathcal{S}$ is locally compact.
-We have
-\begin{align*}
-	\Lambda_{\mathcal{S}}(\mathcal{c}) &= \sum \limits_{\alpha \in A} \Lambda_{U_\alpha}(\mathcal{c}) - \sum \limits_{\alpha \in A} \sum \limits_{\beta \in A}^{\beta \neq\alpha} \Lambda_{U_\alpha \cap U_\beta}(\mathcal{c}) \\
-	&= \sum \limits_{\alpha \in A} \int_{U_\alpha} \rho_{\mathcal{c}} d\mu - \sum \limits_{\alpha \in A} \sum \limits_{\beta \in A}^{\beta \neq\alpha} \int_{U_\alpha \cap U_\beta}(\mathcal{c}) \rho_{\mathcal{c}} d\mu \\
-	 &= \int_{\mathcal{S}} \rho_{\mathcal{c}} d\mu
-\end{align*}
-	as $\Lambda_U$ is always finite. The measure is unique on the whole space.
-
-	We claim $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$ where $L^1_{ew}(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq \mathcal{S}\}$.
+$\mathcal{C} \cong G \subseteq C(\mathcal{S})$ therefore $\Lambda_U(\mathcal{c}) \cong \Lambda_U(\rho_\mathcal{c})$.
+$\Lambda = \{\Lambda_U : C(\mathcal{S}) \rightarrow \mathbbm{R}\}_{U \subseteq \mathcal{S}}$ is a family of positive linear functionals such that $\forall U \subseteq \mathcal{S} \; \Lambda_U = \Lambda_{\mathrm{int}(U)}$ and $\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2} \; \forall U_1, U_2 \subseteq \mathcal{S}$.
+For the extension \ref{extended_riesz_theorem} of the Riesz representation theorem for linear functionals
+	there exists a unique Borel measure $\mu$ such that $\Lambda_U (\mathcal{c}) = \int_{U} \rho_\mathcal{c} d\mu$.
+
+	We claim $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$.
 As per \ref{classical_vector space} $\mathcal{C}$, is isomorphic to a subspace $G \subseteq C(\mathcal{S})$ as a vector space over $\mathbbm{R}$.
-As $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$, then $G \subseteq L^1_{ew}(\mathcal{S}, \mu)$.
-$G \subseteq C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$.
+As $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$, then $G \subseteq \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\} = L^1(\mathcal{S}, \mu)$ for \ref{everywhere_integrable_is_lebesgue_integrable}.
+$G \subseteq C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$.
 Let $\rho_\mathcal{c}$ be the distribution associated with the state $\mathcal{c}$.
-Let $\rho \in C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$ be a continuous function everywhere integral.
+Let $\rho \in C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ be a continuous function everywhere integral.
 If $\rho_\mathcal{c} \neq \rho$ then $\exists U \in \mathcal{S} | \int_{U} \rho_\mathcal{c} d \mu \neq \int_{U} \rho d \mu$.
@@
 There must be a state $\mathcal{c}_1 \in \mathcal{C} \; | \; \varrho_(\mathcal{c}_1)=\rho$.
-$\varrho : \mathcal{C} \rightarrow C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$ is surjective.
-$\varrho$ is an isomorphism between $\mathcal{C}$ and $C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$ as vector spaces.
+$\varrho : \mathcal{C} \rightarrow C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ is surjective.
+$\varrho$ is an isomorphism between $\mathcal{C}$ and $C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as vector spaces.
 \end{justification}
 
-As we identified $C(\mathcal{S}) \cap  L^1_{ew}(\mathcal{S}, \mu)$ as the set of distributions that are physically meaningful, it's useful to show how commonly used functions spaces do not fit the bill.
-We do this briefly by finding counterexample.
+As we identified $C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as the set of distributions that are physically meaningful, it's useful to briefly show how other commonly used functions spaces do not fit the bill.
 The set of continuous functions $C(\mathcal{S})$ would include functions that diverge at infinity.
@@
 The set of continuous functions that vanish at infinity $C_0(\mathcal{S})$ includes function whose integral is infinite.
-Schwartz space $S(\mathcal{S})$ requires functions to be infinitely smooth and we do not see a physical requirement for that.
-
-Before we can properly identify $d\mu$, it turns out that we need to better characterize $\mathcal{S}$.
+Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a physical argument for.
+The set of Lebesgue integrable functions $L^1(\mathcal{S})$ includes discontinuous functions that for \ref{continuous_map} are unphysical.
+
+Another point is that while the norm associated to $L^1(\mathcal{S})$ is $\int_{\mathcal{S}} |\rho| d\mu$, the pseudo-norm $\int_{\mathcal{S}} \rho d\mu$ is perhaps more physically meaningful.
+For proper states, the two are the same and they represent the total amount of material.
+For state changes they differ.
+The second represents the amount of material added (or taken away if negative) by the state change.
+For example, if $\rho_{\delta \mathcal{c}}$ is the change due to deterministic and reversible evolution, the amount of material does not change and therefore $\int_{\mathcal{S}} \rho_{\delta \mathcal{c}} d\mu = 0$.
+The first norm would represent the total amount of material that is changing (i.e.
+being added and being removed), therefore $\int_{\mathcal{S}} |\rho_{\delta \mathcal{c}}| d\mu = 0$ means no material is moving, no change is occurring.
+
+We now turn our attention to the metric $d\mu$ which we need to properly identify. It turns out that before we can do it, we need to better characterize $\mathcal{S}$.
 

### 2016-05-12 2015CompleteDerivation/completeDerivation.tex
Including Christine comments on classical systems

@@
 Discrete sets already come with a way to measure and compare the number of elements which is preserved by any bijective map.
-With continuous sets, we need to do extra work to properly define a measure (or metric) that allows to count states and possibilities.
+With continuous sets, we need to do extra work to properly define a measure (or metric) that allows one to count states and possibilities.
 
@@
 The notion that a system is decomposable means the states are equipped with a rule of composition that allows one to write $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$: the composite system is the sum of its parts.
-For example, the state of a ball is equal to the state of its left and right parts.
+For example, the state of a ball is equal to the state of its top and bottom parts.
 
 The notion that the system is homogeneous means that the states of the composite and of the parts are not unrelated: they are made of the same material.\footnote{Whether a system is homogeneous depends on context (e.g.
-air can be thought as homogeneous if the mixture of gases does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is maintained by time evolution.} In fact, the state space $\mathcal{C}$ of all systems composed of such material will include the state of the system as well as the states of all its parts (i.e.
+air can be thought as homogeneous if the mixture of gases does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is maintained by time evolution.} In fact, the state space $\mathcal{C}$ of all systems composed of such material will include the state of the system as well as the states of  its parts (i.e.
 $\mathcal{c}, \mathcal{c}_1, \mathcal{c}_2 \in \mathcal{C}$).
@@
 
-\subsection{Classical homogeneous systems}
-We also want to be able to express the state of the composite system in terms of the state of the parts.
+\subsection{Classical homogeneous decomposable systems}
+We also want to be able to express the state of the composite system in terms of the state of each part.
 For example, given the state of a fluid we'll want to know how the material is distributed in space.
@@
 We will call a classical system one that is infinitesimally reducible.
-That is: it is made of arbitrarily small parts, which we call particles, and the state of the whole system coincides with the sum of the states of the parts.
-Given the state space $\mathcal{S}$ for the infinitesimal parts, which we assume it to be a manifold consistently with \ref{continuous_state_space}, a composite state $\mathcal{c}$ will tell us the amount of material for each possible particle state.
-That is, each state is fully identified by a function $\rho: \mathcal{S} \to \mathbbm{R}$ that returns the amount of material in the composite state $\mathcal{c}$ that is prepared in the particle state $\mathcal{s}$.
+That is, it is made of arbitrarily small parts, which we call particles, and giving the state of the whole system is equivalent to giving the states of all parts.
+Given the state space $\mathcal{S}$ for the infinitesimal parts, which we assume to be a manifold consistently with \ref{continuous_state_space}, a composite state $\mathcal{c} \in \mathcal{C}$ will tell us the amount of material for each possible particle state.
+That is, each state is fully identified by a function $\rho_\mathcal{c}: \mathcal{S} \to \mathbbm{R}$ that returns the amount of material in the composite state $\mathcal{c}$ that is prepared in the particle state $\mathcal{s}$.
 For example, given a certain configuration of gas, we can tell the density of material that is at a specific point in space with a specific value of momentum.
 
-The notion that the parts are infinitesimal requires the co-domain of $\rho$ to be a real number, as opposed to an integer.
+The notion that the parts are infinitesimal requires the co-domain of $\rho_\mathcal{c}$ to be a real number, as opposed to an integer.
 The state space $\mathcal{S}$, instead, can have a discrete topology.
@@
 The amount in each tank is continuous (as we assume the water to be infinitesimally divisible) yet there are only a finite number of tanks the water can be placed into.
-As it links two physically distinguishable objects, $\rho$ is a continuous function as discussed before.
-
-As we combine the states of different parts, we sum the distributions over particle states $\rho(\mathcal{s})=\rho_1(\mathcal{s})+\rho_2(\mathcal{s})$: the amount of material in the composite state is the sum of the amount of material of the parts.
-We can also increase or decrease the amount of material by a constant factor $\rho_1(\mathcal{s})=a\rho_2(\mathcal{s})$.
+As it links two physically distinguishable objects, $\rho_\mathcal{c}$ is a continuous function as discussed before.
+
+As we combine the states of different parts, we sum the distributions over particle states $\rho_{\mathcal{c}}(\mathcal{s})=\rho_{\mathcal{c}_1}(\mathcal{s})+\rho_{\mathcal{c}_2}(\mathcal{s})$: the amount of material in the composite state is the sum of the amount of material of the parts.
+We can also increase or decrease the amount of material by a constant factor $\rho_{\mathcal{c}_1}(\mathcal{s})=a\rho_{\mathcal{c}_2}(\mathcal{s})$.
 This gives us the structure of a real vector space that is isomorphic to a subspace of continuous functions.
@@
 We claim $T$ is a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics or differential geometry sense (a physical quantity/tensor with a value for each point in space).} isomorphic to $\mathbbm{R}$.
-Consider $\tau: \mathbbm{R} \rightarrow T$ the mapping between a number and the transformation that increases or decreases the amount of material by that number.
+Consider $\tau: \mathbbm{R} \rightarrow T$ the mapping between a number and the transformation that increases or decreases the amount of material by that factor.
 This transformation exists: the system is infinitesimally decomposable and the amount can be changed continuously.
@@
 As the system is reducible, two distinct composite states must represent different distributions.
-$\forall \mathcal{c_1}, \mathcal{c_2} \in \mathcal{C}, \mathcal{c_1} \neq \mathcal{c_2} \implies \varrho(\mathcal{c_1}) \neq \varrho(\mathcal{c_1})$.
+$\forall \mathcal{c_1}, \mathcal{c_2} \in \mathcal{C}, \mathcal{c_1} \neq \mathcal{c_2} \implies \varrho(\mathcal{c_1}) \neq \varrho(\mathcal{c_2})$.
 $\varrho$ is injective.
@@
 $\varrho$ is a homomorphism.
-$\mathcal{C}$ is isomorphic to $\varrho(\mathcal{C}) \subseteq C(\mathcal{S})$ as vector spaces.
+$\mathcal{C}$ is isomorphic to $\varrho(\mathcal{C}) \subseteq C(\mathcal{S})$ as a vector space.
 \end{justification}
@@
 In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique measure $\mu$ for the size of sets of $\mathcal{S}$.
-From the measure we get a metric $d\mu$ and our positive linear functionals become integrals: $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
+From the measure we get a metric $d\mu$ and our positive linear functionals become integrals, $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
 
@@
 \begin{justification}
-	We claim there exists a positive linear functional $\Lambda_U : \mathcal{C} \rightarrow \mathbbm{R}$ for each $U \subseteq \mathcal{S}$ for which $\Lambda_U = \Lambda_{\mathrm{int}(U)}$.
+	We claim there exists a positive linear functional $\Lambda_U : \mathcal{C} \rightarrow \mathbbm{R}$ for each $U \subseteq \mathcal{S}$ such that $\Lambda_U = \Lambda_{\mathrm{int}(U)}$.
 Let $\Lambda_{U} : \mathcal{C} \rightarrow \mathbbm{R}$ be the functional that returns the amount of material in $\mathcal{c}$ compatible with the outcome associated with the open set $U \subseteq \mathcal{S}$.
 $\Lambda_U$ is linear.
-$\Lambda_U(a_1 \mathcal{c}_1 + a_2 \mathcal{c}_2) = a_1 \Lambda_U(\mathcal{c}_1) + a_2 \Lambda_U(\mathcal{c}_2)$ as it has to be consistent with the operations of composing states and increasing/decrease the amount of material by a factor.
+$\Lambda_U(a_1 \mathcal{c}_1 + a_2 \mathcal{c}_2) = a_1 \Lambda_U(\mathcal{c}_1) + a_2 \Lambda_U(\mathcal{c}_2)$ as it has to be consistent with the operations of composing states and increasing/decreasing the amount of material by a factor.
 $\Lambda_U$ is positive.
@@
 Let $U \subseteq \mathcal{S}$ be an open set of particle states associated with an outcome.
-Let $\mathcal{c} \in \mathcal{S}$ a composite state.
+Let $\mathcal{c} \in \mathcal{C}$ a composite state.
 The amount of material of $\mathcal{c}$ associated with the outcome $U$ must be finite, as physically we always work with finite quantities.
@@
 $\Lambda = \{\Lambda_U : C(\mathcal{S}) \rightarrow \mathbbm{R}\}_{U \subseteq \mathcal{S}}$ is a family of positive linear functionals such that $\forall U \subseteq \mathcal{S} \; \Lambda_U = \Lambda_{\mathrm{int}(U)}$ and $\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2} \; \forall U_1, U_2 \subseteq \mathcal{S}$.
-For the extension \ref{extended_riesz_theorem} of the Riesz representation theorem for linear functionals
+By the extension \ref{extended_riesz_theorem} of the Riesz representation theorem for linear functionals
 	there exists a unique Borel measure $\mu$ such that $\Lambda_U (\mathcal{c}) = \int_{U} \rho_\mathcal{c} d\mu$.
@@
 	We claim $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$.
-As per \ref{classical_vector space} $\mathcal{C}$, is isomorphic to a subspace $G \subseteq C(\mathcal{S})$ as a vector space over $\mathbbm{R}$.
-As $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$, then $G \subseteq \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\} = L^1(\mathcal{S}, \mu)$ for \ref{everywhere_integrable_is_lebesgue_integrable}.
+As per \ref{classical_vector space}, $\mathcal{C}$ is isomorphic to a subspace $G \subseteq C(\mathcal{S})$ as a vector space over $\mathbbm{R}$.
+As $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$, then $G \subseteq \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\} = L^1(\mathcal{S}, \mu)$ by \ref{everywhere_integrable_is_lebesgue_integrable}.
 $G \subseteq C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$.
 Let $\rho_\mathcal{c}$ be the distribution associated with the state $\mathcal{c}$.
-Let $\rho \in C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ be a continuous function everywhere integral.
-If $\rho_\mathcal{c} \neq \rho$ then $\exists U \in \mathcal{S} | \int_{U} \rho_\mathcal{c} d \mu \neq \int_{U} \rho d \mu$.
-There exist an outcome in $\mathcal{S}$ that can physically distinguish the two distributions.
-There must be a state $\mathcal{c}_1 \in \mathcal{C} \; | \; \varrho_(\mathcal{c}_1)=\rho$.
+Let $\rho \in C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ be a Lebesgue integrable continuous function.
+If $\rho_\mathcal{c} \neq \rho$ then $\exists U \in \mathcal{S} \; | \; \int_{U} \rho_\mathcal{c} d \mu \neq \int_{U} \rho d \mu$.
+There exists an outcome in $\mathcal{S}$ that can physically distinguish the two distributions.
+There must be a state $\mathcal{c}_1 \in \mathcal{C} \; | \; \varrho(\mathcal{c}_1)=\rho$.
 $\varrho : \mathcal{C} \rightarrow C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ is surjective.
@@
 
-As we identified $C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as the set of distributions that are physically meaningful, it's useful to briefly show how other commonly used functions spaces do not fit the bill.
+As we identified the set of Lebesque integrable continuous functions $C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as the set of distributions that are physically meaningful, it's useful to briefly show how other commonly used function spaces do not fit the bill.
 The set of continuous functions $C(\mathcal{S})$ would include functions that diverge at infinity.
 The set of continuous functions with compact support $C_C(\mathcal{S})$ would exclude Gaussians.
-The set of continuous functions that vanish at infinity $C_0(\mathcal{S})$ includes function whose integral is infinite.
+The set of continuous functions that vanish at infinity $C_0(\mathcal{S})$ includes functions whose integral is infinite.
 Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a physical argument for.
@@
 
-We now turn our attention to the metric $d\mu$ which we need to properly identify. It turns out that before we can do it, we need to better characterize $\mathcal{S}$.
+We now turn our attention to the metric $d\mu$ which we need to properly identify. It turns out that before we can do so, we need to better characterize $\mathcal{S}$.
 

### 2016-05-13 2015CompleteDerivation/completeDerivation.tex
More review on classical hamiltonian systems

@@
 
-On a state space with discrete topology and a single state variable we simply have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho(q)$.
-For states identified by a single continuous state variable we would expect $\Lambda_U (\mathcal{c}) = \int_a^b \rho (q) dq$.
-By a simple change of state variable, we see that this does not work in general: $\rho(q')= \rho(q) dq/dq'$.
-This would make $\rho$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using.
+On a state space with discrete topology and a single state variable we simply have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho_\mathcal{c}(q)$.
+For states identified by a single continuous state variable we would expect $\Lambda_U (\mathcal{c}) = \int_a^b \rho_\mathcal{c} (q) dq$.
+By a simple change of state variable, we see that this does not work in general: $\rho_\mathcal{c}(q')= \rho(q)_\mathcal{c} dq/dq'$.
+This would make $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using.
 This is inconsistent with our previous definition.
@@
 
-For the Jacobian to exist $\mathcal{S}$ has to be a differentiable manifold; that is the transformation between state variables is always differentiable.
-This way, at least, the density is well defined for any choice of state variables.
-
-If $\rho$ is to remain unchanged and the Jacobian to be unitary we cannot simply change one state variable as we please.
+For the Jacobian to exist the transformation between state variables has to be differentiable.
+Therefore a manifold that guarantees densities to be well defined is one that guarantees that state variable transformations are differentiable: a differentiable manifold.
+$\mathcal{S}$ has a differentiable structure in the sense that it is on that structure that densities are defined.
+In the same way that discontinuous changes of state variable can be detected because they are incompatible with the topology, non differential changes of state variable can be detected because they are incompatible with state-variable-independent densities.
+
+If $\rho_\mathcal{c}$ is to remain unchanged and the Jacobian to be unitary we cannot simply change one state variable as we please.
 If we change one at least another has to change in some coordinated way such that the Jacobian of the total transformation is unitary.
@@
 An element of $\mathcal{S}$ will be a geometrical object in such space.
-As these are states for infinitesimal parts, it will be a local object in $\mathcal{Q}$.
-As we saw before, it cannot be a point as the density $\rho(q')= \rho(q) dq/dq'$ is not invariant.
-It cannot be a vector $v(q) e_q$, as the density $\rho(q', v')= \rho(q, v) (dq/dq')^2$ is not invariant.
-A covector $k(q) dq$ is instead fine as $\rho(q', k')= \rho(q, k) (dq/dq') (dq'/dq) = \rho(q, k)$.
+For example, it will be a point, a distance, two points, the ratio between two distances, the inverse of a distance: something that is expressible in the unit defined.
+As the particle state represent configurations of infinitesimal parts, it will be a local object in $\mathcal{Q}$, defined around a point.
+As we saw before, it cannot be a point itself as the density $\rho(q')_\mathcal{c}= \rho(q)_\mathcal{c} dq/dq'$ is not invariant.
+It cannot be a vector $v(q) e_q$, as the density $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$ is not invariant.
+A covector $k(q) dq$ is instead fine as $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$.
 In other words: a covector is the simplest local geometrical object that allows coordinate independent densities.
@@
 
-	If we assume the whole system is deterministic and reversible but each part is not, then their evolution can depend on the states of the other parts.
+	If we just assume the whole system is deterministic and reversible, then each part can evolve depending the states of the other parts.
 But this does not solve the problem entirely.
@@
 This means $(a, b) \mapsto (\hat{a}, \hat{b})$ is not a reversible map.
-In other words: we can't expect to have deterministic and reversible evolution at all levels of decomposable systems where the parts interact in any physically meaningful way.
+In other words: we can't expect to have deterministic and reversible evolution at all levels of decomposable systems if the parts interact in any physically meaningful way.
 

### 2016-05-13 2015CompleteDerivation/completeDerivation.tex
General reorganization
Now we have Assumptions/Propositions/Theorems. Reviewed initial blurb to
put it in a state that is reasonable for public consumption.

@@
 \renewcommand*{\theassump}{\Roman{assump}}
+\newtheorem{thrm}{Theorem}[section]
+
+\theoremstyle{definition}
 \newtheorem{prop}{Proposition}[section]
-\newtheorem{thrm}{Theorem}[section]
-
-\theoremstyle{definition}
-\newtheorem{defn}[prop]{Definition}
 
@@
 
-\title{From physical principles to ???}
+\title{DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics}
 \author{Gabriele Carcassi}
@@
 \email{carcassi@umich.edu}
-\date{March 15, 2015}
+\date{\today}
 
 \begin{abstract}
-Deterministic/reversible assumption.
-Describe fundamental model of physics.
-System is always interacting with environment, both deterministically and non-deterministically.
-Deterministic means only depends on the state of the system.
-Set of states needs to be invariant under non-deterministic interaction: the non-deterministic interaction affects what are the states of the system.
-Deterministic/reversible evolution is necessary to even be able to talk about a system (trivial deterministic process).
-
-Define: set of states, deterministic evolution as permutation, labels/state variables, cardinality of state variable, state id function (=1 for a particular state).
-Can we do for continuous variables as well?
-Can we find divergence free evolution for det/rev and independent variables?
-
-Infinite reducibility assumption.
-Decomposable systems.
-Rule of composition.
-Inverse and null state to study changes.
-Set of transformations to increase and decrease the quantity in each state.
-Define magnitude of the system (probably needed to guarantee things converge in math?).
-
-Define: set of states as a vector space, inner product.
-
-Continuous quantities. Need for densities to be a function of the state (not state variable). Defined on infinitesimal interval. Define: degree of freedom, state cardinality, metric (T*Q).
+\textbf{This is manuscript is a work in progress.} Ideas are constantly reshaped to find more precise and elegant arguments. It is provided as is to stimulate discussion.
+
+The aim of this work is to show that particle mechanics, both classical and quantum, Hamiltonian and Lagrangian, can be derived from few simple physical assumptions.
+
+Assuming deterministic and reversible time evolution will give us a dynamical system, whose set of states form a topological space and whose law of evolution is a self-homeomorphism.
+Assuming the system is infinitesimally reducible, giving the state and the dynamics of the whole system is equivalent to giving the state and the dynamics of its infinitesimal parts, will give us a classical Hamiltonian system.
+Assuming the system is irreducible, giving the state and the dynamics of the whole system tells us nothing about the state and the dynamics of its substructure, will give us a quantum Hamiltonian system.
+Assuming kinematic equivalence, studying trajectories is equivalent to studying state evolution, will give us Lagrangian mechanics together and limit the form of the Hamiltonian/Lagrangian to the one with scalar and vector potential forces.
+
+We work strives as much as possible to be philosophically consistent, physically meaningful and mathematically precise.
 \end{abstract}
@@
 
-While some branches of fundamental physics (e.g.
+TODO: Point out most modern physics starts by assuming some mathematical framework. Pros and cons of that.
+
+TODO: High level view of the derivation
+
+%While some branches of fundamental physics (e.g.
 Newtonian mechanics and special relativity) are founded on physical laws or principles others (e.g.
@@
 
-TODO: high level view of the derivation
-
 %This work aims to re-organize the known elements and equations in a more consistent and comprehensive way, leading to better insight on why the fundamental concepts and laws are what they are.
@@
 
-\section{About}
+\section{Organization and style}
 
@@
 A \textbf{rationale} follows each assumption, which uses physical and sometimes philosophical arguments to motivate why (or why not) such an assumption makes sense (in a particular case).
-  \item[Definitions] these encode into mathematical language the physical concepts we are studying.
-A \textbf{justification} follows each definition, which uses physical and mathematical arguments to explain why such a definition and its properties are necessary to describe a particular physical concept.
-  \item[Propositions] these are statements that mathematically follow from definitions or other propositions.
+  \item[Propositions] these characterize the mathematical properties of the physical concepts we are studying.
+A \textbf{justification} follows each proposition to show why such characterization follows from the physical properties of the system.
+While the style of a justification is formal, it is not a pure mathematical proof as it necessarily has to mix physical and mathematical arguments.
+
+  Before or after each proposition there is a \textbf{discussion} that conceptually explains the motivation and the result in a less formal, more approachable way.
+
+  \item[Theorems] these are pure mathematical statements that are used to simplify propositions.
 A mathematical \textbf{proof} follows each proposition.
 No mathematical breakthrough should be expected as we mostly use well known results from different fields.
-The goal is to show how those results make sense in light of the physical assumptions we start with.
+Theorems are grouped in the appendix.
 \end{description}
-This should allow you, the reader, to focus on the parts you are most interested in and skim the rest.
-The philosophically inclined may focus on the assumptions, their rationale and the definitions.
-The mathematically inclined may focus on the definitions, the propositions and their proofs.
-The scientist on the assumptions, the definitions and their justifications, and the propositions.
-
-\section{Style objective}
-
-The writing style should have the following goals:
-\begin{description}
-  \item[Skip math] one should be able to skip the mathematical definitions/propositions and still get a coherent narrative.
-  \item[Well-defined physical objects] each mathematical objects need to be physically well-defined. It must shown to exist and be unique.
-\end{description}
-Non-goals:
-\begin{description}
-  \item[Can't skip physics in justifications] justifications of mathematical definitions will inherently be rooted in the preceding physics discussion.
-\end{description}
-Notation:
-\begin{description}
-  \item[state variable and space] state space (cont/discr), state variable (cont/discr), possibility (cont/discr), range of possibility (cont/discr), set of possibilities (cont/discr)
-\end{description}
+This should allow readers with different focus and background to concentrate on the parts that are more keen on.
 
@@
 
-\begin{defn}\label{statedef}
+\begin{prop}\label{statedef}
 The state space $\mathcal{S}$ of a physical system is a Hausdorff and second countable topological space.
-\end{defn}
+\end{prop}
 
@@
 
-\begin{defn}\label{topologically_distinguishable}
+\begin{prop}\label{topologically_distinguishable}
 	Any set $\mathcal{S}$ of physically distinguishable elements is a Hausdorff and second countable topological space.
-\end{defn}
+\end{prop}
 
@@
 
-\begin{defn}\label{detrevmap}
+\begin{prop}\label{detrevmap}
 A deterministic and reversible evolution map is a self-homeomorphism on the state space $\mathcal{T}_{\Delta t}:\mathcal{S} \rightarrow \mathcal{S}$.
-\end{defn}
+\end{prop}
 
@@
 
-\begin{defn}\label{continuous_map}
+\begin{prop}\label{continuous_map}
 	A map $f:\mathcal{S_1} \rightarrow \mathcal{S_2}$ between two sets of physically distinguishable elements $\mathcal{S_1}$ and $\mathcal{S_2}$ is a continuous map.
-\end{defn}
+\end{prop}
 
@@
 
-\begin{defn}\label{state_variable}
+\begin{prop}\label{state_variable}
 	A \emph{state variable} is a continuous map $q : U \rightarrow \mathbbm{L}$ where $U \subseteq \mathcal{S}$ and $\mathbbm{L}$ is the space for the possible values.
@@
 If $\mathbbm{L}\cong \mathbbm{R}$ the variable is said \emph{continuous}.
-\end{defn}
+\end{prop}
 
@@
 
-\begin{defn}\label{independent_state_variables}
+\begin{prop}\label{independent_state_variables}
 	Two state variables $q_1 : U \rightarrow \mathbbm{L}_1$ and $q_2 : U \rightarrow \mathbbm{L}_2$ are said \emph{independent} if $\exists \mathcal{s} | q_1(\mathcal{s})=l_1, q_2(\mathcal{s})=l_2 \forall l_1 \in q_1(U), l_2 \in q_2(U)$.
-\end{defn}
+\end{prop}
 
@@
 
-\begin{proof}
+\begin{justification}
 	We claim $\mathcal{S}'$ is a manifold.
@@
 $\mathcal{S}'$ is path connected.
-\end{proof}
+\end{justification}
 
@@
 
-\begin{defn}\label{evolved_transported_variable}
+\begin{prop}\label{evolved_transported_variable}
 Let $q^i$ be a set of state variables and $\mathcal{T}_{\Delta t}$ a deterministic and reversible evolution map.
@@
 The transported state variables are given by $q^i \circ \mathcal{T}_{-\Delta t}$.
-\end{defn}
+\end{prop}
 
@@
 
-\begin{proof}
+\begin{justification}
 	We claim $\#(\Delta q^i)=\#(\Delta \hat{q}^i)$.
@@
 	We claim $\#(U)=\prod\limits_{i=1}^{n}\#(\Delta q^i)$. As $q^i$ are independent variables, the states are the Cartesian product of the possibilities of each variable.
-\end{proof}
+\end{justification}
 
@@
 
-\begin{defn}\label{reducible_state_space}
+\begin{prop}\label{reducible_state_space}
 The state space $\mathcal{C}$ for decomposable systems made of a homogeneous material is an additive abelian (i.e. commutative) group.
-\end{defn}
+\end{prop}
 
@@
 
-\begin{defn}\label{classical_vector space}
+\begin{prop}\label{classical_vector space}
 The state space $\mathcal{C}$ for a homogeneous classical (i.e.
 infinitesimally reducible) material is a vector space over $\mathbbm{R}$ isomorphic to a subspace of $C(\mathcal{S}) = \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is continuous$\}$, where $\mathcal{S}$ is the state space of an infinitesimal amount of material.
-\end{defn}
+\end{prop}
 
@@
 
-\begin{defn}\label{integration}
+\begin{prop}\label{integration}
 	Let $\mathcal{S}$ be the state space for the particles of a homogeneous classical material.
@@
 That is: $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ where $L^1(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \; \int_{\mathcal{S}} |\rho| d\mu < \infty \}$.
-\end{defn}
+\end{prop}
 
@@
 
-\begin{defn}\label{classical_phase_space}
+\begin{prop}\label{classical_phase_space}
 The state space $\mathcal{S}$ for the particles of a homogeneous classical material is a symplectic manifold formed by a cotangent bundle $T^*\mathcal{Q}$ equipped with the canonical two-form $\omega$.
-\end{defn}
+\end{prop}
 
@@
 
-\begin{defn}\label{canonical_transformation}
+\begin{prop}\label{canonical_transformation}
 	A deterministic and reversible evolution map for a homogeneous classical material is self-symplectomorphism.
 That is: $\mathcal{T}_{\Delta t}: T^*\mathcal{Q} \rightarrow T^*\mathcal{Q}$ and $\mathcal{T}_{\Delta t}^*\omega = \omega$ where $\mathcal{T}_{\Delta t}^*$ is the pullback of $\mathcal{T}_{\Delta t}$.
-\end{defn}
+\end{prop}
 
@@
 
-\begin{proof}
+\begin{justification}
 We claim the vector field $S \in T\mathcal{Q}$ for the infinitesimal displacement $S = d\mathcal{s}/dt$ admits a potential $H$ such that, in component notation, $S^{a} \omega_{ab} = \partial_{b}H$.
@@
 For $a=\{n+1,...,2n\}$ we have $d_tq^i (+1) = \partial_{p_i} H$.
-\end{proof}
+\end{justification}
 
@@
 
-\section{Appendix: mathematical support}
+\section{Appendix: mathematical proofs}
 

### 2016-05-17 2015CompleteDerivation/completeDerivation.tex
Minor improvements

@@
 This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows state-variable-independent densities.
-Under arbitrary change of state variables, the density has to remain the same, which means that the Jacobian of the transformation must exist and must be unitary.
+Under an arbitrary change of state variables, the density has to remain the same, which means that the Jacobian of the transformation must exist and must be unitary.
 
@@
 Therefore a manifold that guarantees densities to be well defined is one that guarantees that state variable transformations are differentiable: a differentiable manifold.
-$\mathcal{S}$ has a differentiable structure in the sense that it is on that structure that densities are defined.
+$\mathcal{S}$ has a differentiable structure in the sense that the distributions are only defined on a set of set state variables that are linked by differential transformations.
 In the same way that discontinuous changes of state variable can be detected because they are incompatible with the topology, non differential changes of state variable can be detected because they are incompatible with state-variable-independent densities.
@@
 For example, it will be a point, a distance, two points, the ratio between two distances, the inverse of a distance: something that is expressible in the unit defined.
-As the particle state represent configurations of infinitesimal parts, it will be a local object in $\mathcal{Q}$, defined around a point.
-As we saw before, it cannot be a point itself as the density $\rho(q')_\mathcal{c}= \rho(q)_\mathcal{c} dq/dq'$ is not invariant.
-It cannot be a vector $v(q) e_q$, as the density $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$ is not invariant.
-A covector $k(q) dq$ is instead fine as $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$.
+As particle states represent configurations of infinitesimal parts, they will be local objects, defined around a point in $\mathcal{Q}$.
+As we saw before, they cannot be points themselves as the density $\rho(q')_\mathcal{c}= \rho(q)_\mathcal{c} dq/dq'$ is not invariant.
+They cannot be a vectors $v(q) e_q$, as the density $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$ is not invariant.
+They can, however, be covectors $k(q) dq$ as $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$.
 In other words: a covector is the simplest local geometrical object that allows coordinate independent densities.
@@
 This gives us insight on the physical meaning of the geometrical structure $T^*\mathcal{Q}$.
-The canonical one-form $\theta=k(q) dq$ represents the geometrical object we associate with each particle state, which is not a point in $\mathcal{Q}$.
-The canonical two-form $\omega=\Sigma \, dq^i \wedge dk_i$ is the metric that quantifies the possibilities described by two given state variables.
+The canonical one-form $\theta=k dq$ represents the geometrical object we associate with each particle state.
+The canonical two-form $\omega=\Sigma \, dq^i \wedge dp_i$ is the metric that quantifies the possibilities described by two given state variables.
 Note, instead, that the relationship $\omega = - \hbar d\theta$, while mathematically true, has no clear physical meaning as we have not defined what the exterior derivative on a state actually represents.

### 2016-05-18 2015CompleteDerivation/completeDerivation.tex
Minor improvement after Monday discussion

@@
 \begin{abstract}
-\textbf{This is manuscript is a work in progress.} Ideas are constantly reshaped to find more precise and elegant arguments. It is provided as is to stimulate discussion.
+\textbf{This manuscript is a work in progress.} Ideas are constantly reshaped to find more precise and elegant arguments. It is provided as is to stimulate discussion.
 
@@
 the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics, we now explore how to characterize a system in terms of its components.
-That is, we want to study the relationship between the state space $\mathcal{C}$ of a composite system and the state space $\mathcal{S}$ of its parts.
+That is, we want to study the relationship between the state space of a composite system and the state space of its parts.
 In general, this is quite a complicated thing to do, which requires intimate knowledge of the system at hand.
@@
 What we'll find is that under the additional assumption that the system is infinitesimally reducible (i.e.
-its state is equivalent to the state of its parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian mechanics.
+its state is equivalent to the states of its parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian mechanics.
 
@@
 \begin{prop}\label{reducible_state_space}
-The state space $\mathcal{C}$ for decomposable systems made of a homogeneous material is an additive abelian (i.e. commutative) group.
+The state space $\mathcal{C}$ for a decomposable homogeneous material is an additive abelian (i.e. commutative) group.
 \end{prop}
@@
 $\hat{\mathcal{c}} = \mathcal{c} +\delta \mathcal{c}$) we introduce an inverse $- : \mathcal{C} \rightarrow \mathcal{C}$ such that $\mathcal{c} + ( - \mathcal{c}) = 0 \forall \mathcal{c} \in \mathcal{C}$.
-Such inverse may introduce objects that do not properly represent a physical state, but a state change.
+Such inverse includes objects that do not properly represent a physical state, but a state change.
 A change of a physically distinguishable object is itself physically distinguishable therefore $\mathcal{C}$ is still a Hausdorff and second countable topological manifold as per \ref{topologically_distinguishable}.
-We'll still call $\mathcal{C}$ state space committing an abuse of terminology.
 \end{justification}
@@
 The state space $\mathcal{C}$ for a homogeneous classical (i.e.
-infinitesimally reducible) material is a vector space over $\mathbbm{R}$ isomorphic to a subspace of $C(\mathcal{S}) = \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is continuous$\}$, where $\mathcal{S}$ is the state space of an infinitesimal amount of material.
+infinitesimally reducible) material is a vector space over $\mathbbm{R}$ isomorphic to a subspace of $C(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is continuous$\}$, where $\mathcal{S}$ is the state space of an infinitesimal amount of material.
 \end{prop}
@@
 Then $\mathcal{S}$ is endowed with a Borel measure $\mu$.
-The state space $\mathcal{C}$ is isomorphic to the space of continuous functions that are integrable over any subset of $\mathcal{S}$.
+The state space $\mathcal{C}$ is isomorphic to the space of Lebesgue integrable continuous functions over $\mathcal{S}$.
 That is: $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ where $L^1(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \; \int_{\mathcal{S}} |\rho| d\mu < \infty \}$.
@@
 
+    Being able to track parts at a given scale is not even enough.
+The parts are infinitesimal, the limit of a process of infinite recursive subdivision, so we must be able to track at any scale.
+Therefore the assumption can only be validated experimentally to the smallest precision available.
+
 	The second, more conceptual, problem is that the idealizations that allow us to hold Assumption \ref{detrevass} break down as the parts get smaller.
@@
 They are the set of equation to describe the deterministic and reversible motion of the infinitesimal parts of an infinitesimally reducible homogeneous material.
-In other words: \textbf{the forces that conserve energy (i.e.
-the value of a sutiable Hamiltonian) are exactly the ones that provide deterministic and reversible motion}.
+In other words: the forces that conserve energy (i.e.
+the value of a suitable Hamiltonian) are exactly the ones that provide deterministic and reversible motion.
 The challenge in their derivation mainly lies in the necessary use of different branches of mathematics (e.g.

### 2016-05-18 2015CompleteDerivation/completeDerivation.tex
Added justification for symplectic manifold

@@
 \begin{justification}
-	TODO: we can search for omega looking from constraints. Must be linear. Must be skew-symmetric (applied to the same vector twice must be zero). Must be invariant by coord transformation.
+	We claim $\mathcal{S}$ is a differential manifold.
+A composite state $\mathcal{c}$ is a distribution $\rho_\mathcal{s} : \mathcal{S} \rightarrow \mathbbm{R}$.
+$\rho_\mathcal{c}$ is defined on any chart of $\mathcal{S}$.
+Given two charts, for any $\rho_\mathcal{c}(q)$ there is a corresponding $\rho_\mathcal{c}(q')=\rho_\mathcal{c}(q^i) | d q' / d q |$.
+The Jacobian $| d q' / d q |$ exists and is non-zero.
+The map between any two charts is differentiable.
+$\mathcal{S}$ is a differentiable manifold.
+
+	We claim the simplest state space $\mathcal{S}$ that allows state-variable-invariant densities is the symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of a one dimensional differential manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
+The description of the simplest physical object on which we can  state-variable-invariant must be given by at least one state variable $q$.
+Let $\mathcal{Q}$ be the manifold charted by $q$.
+$\mathcal{Q} \subseteq \mathcal{S}$.
+$\mathcal{Q}$ is differentiable.
+Let's assume $q$ is enough to define the unit system for the description of the physical object.
+Then the physical object is represented by a set of geometrical objects in $\mathcal{Q}$.
+The physical object is infinitesimal, therefore is represented by a local geometrical object.
+The simplest geometrical object in $\mathcal{Q}$ is a point, charted by $q$.
+A density over points is not invariant: $\rho(q')_\mathcal{c}= \rho(q)_\mathcal{c} dq/dq'$.
+Next we consider local geometrical objects charted by 2 variables.
+Densities over vectors $v(q) e_q \in T\mathcal{Q}$ are not invariant: $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$.
+Densities over covectors $k(q) dq \in T^*\mathcal{Q}$ are invariant: $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$.
+Covectors are the simplest geometrical object to allow unit independent densities.
+The state space $\mathcal{S}$ for the simplest object that allows state-variable-independent densities is $T^*\mathcal{Q}$ of a one dimensional differential manifold $\mathcal{Q}$.
+Express integration in coordinates.
+$\int_{\mathcal{S}} \rho_\mathcal{c} d\mu \propto \int_{\mathcal{S}} \rho_\mathcal{c}(q, k) dq \wedge dk$.
+$d\mu = \hbar dq \wedge dk = dq \wedge dp = \omega$ where $\hbar$ is a constant and $\omega$ is the canonical two form.
+$\omega$ is invariant under state variable change.
+$\omega$ is a symplectic metric for $\mathcal{S}=T*\mathcal{Q}$.
+$\mathcal{S} \cong(T^*\mathcal{Q}, \omega)$ as a symplectic manifold.
+
+	We claim the state space $\mathcal{S}$ for the particles of a homogeneous classical material is a symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of a one dimensional differential manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
+Let $q^i$ be a set of n continuous independent state-variable that define the units to describe a state in $U \subseteq \mathcal{S}$.
+Let $\mathcal{Q}$ be the manifold charted by $q^i$.
+Locally $\mathcal{Q} \cong \prod \mathcal{Q}^i \cong \mathbbm{R}^n$.
+Changing units of one state variable must not change the units of the other as they are independent.
+Each degree of freedom must define state-variable-distribution independently.
+For each $\mathcal{Q}^i$ we need a covector $k_i(q) dq^i \in T^*\mathcal{Q}^i$.
+Locally $\mathcal{S} \cong \prod T^*\mathcal{Q}^i \cong T^* \prod \mathcal{Q}^i \cong T^* \mathcal{Q}$.
+Integration must also be defined on an independent d.o.f.
+There must exist a non degenerate two-form $\omega$ such that $\int_{U \subset \mathcal{S}} \rho_\mathcal{c} \omega$ where $U$ is any two dimensional surface in $\mathcal{S}$.
+$\omega$ has to be form invariant under state variable change.
+The canonical two-form is the only such form.
+$\omega = \sum dq^i \wedge dp_i$.
+$\mathcal{S} \cong (T^*\mathcal{Q}, \omega)$ as a symplectic manifold.
 \end{justification}

### 2016-05-20 2015CompleteDerivation/completeDerivation.tex
Better characterization of classical particles

@@
 We will call a classical system one that is infinitesimally reducible.
-That is, it is made of arbitrarily small parts, which we call particles, and giving the state of the whole system is equivalent to giving the states of all parts.
+That is, we can keep decomposing the system indefinitely into smaller and smaller parts, each with a well defined state.
+We call particles the infinitesimal parts given by the limit of this process.
+Giving the state of the whole system is equivalent to giving the states of all particles.
 Given the state space $\mathcal{S}$ for the infinitesimal parts, which we assume to be a manifold consistently with \ref{continuous_state_space}, a composite state $\mathcal{c} \in \mathcal{C}$ will tell us the amount of material for each possible particle state.

### 2016-05-23 2015CompleteDerivation/completeDerivation.tex
Minor improvements on function spaces

@@
 This makes $\Lambda_U$ a positive linear functional.
-As the amount of material we will measure is always finite, the functional applied to any composite state $\mathcal{c}$ over any set $U$ will be finite.
+As the amount of material we measure is always finite, the functional applied to any composite state $\mathcal{c}$ over any set $U$ will be finite.
 
@@
 \end{justification}
-
-As we identified the set of Lebesque integrable continuous functions $C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as the set of distributions that are physically meaningful, it's useful to briefly show how other commonly used function spaces do not fit the bill.
-The set of continuous functions $C(\mathcal{S})$ would include functions that diverge at infinity.
-The set of continuous functions with compact support $C_C(\mathcal{S})$ would exclude Gaussians.
-The set of continuous functions that vanish at infinity $C_0(\mathcal{S})$ includes functions whose integral is infinite.
-Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a physical argument for.
-The set of Lebesgue integrable functions $L^1(\mathcal{S})$ includes discontinuous functions that for \ref{continuous_map} are unphysical.
 

### 2016-05-26 2015CompleteDerivation/completeDerivation.tex
Paragraph on functions didn't get checked in

@@
 \title{DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics}
-\author{Gabriele Carcassi}
+\author{Gabriele Carcassi, Christine A. Aidala, David J. Baker, Lydia Bieri}
 \affiliation{University of Michigan, Ann Arbor, MI 48109}
@@
 \end{justification}
+
+As we identified the set of Lebesque integrable continuous functions $C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as the set of distributions that are physically meaningful, we can better understand why other commonly used function spaces do not fit the bill.
+Some are not restrictive enough.
+The set of Lebesgue integrable functions $L^1(\mathcal{S})$ includes discontinuous functions that for \ref{continuous_map} are unphysical.
+The set of continuous functions $C(\mathcal{S})$ and the set of continuous functions that vanish at infinity $C_0(\mathcal{S})$ include functions whose integral is infinite, which would represent infinite amount of material.
+Some are too restrictive.
+The set of continuous functions with compact support $C_C(\mathcal{S})$ would exclude distributions, such as Gaussians, that span over the whole range of states.
+If we take a finite volume of an ideal gas at equilibrium, the momentum distribution spans over the whole range.
+Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a general physical justification for.
 

### 2016-05-26 2015CompleteDerivation/completeDerivation.tex
More work on relativistic mechanics. Lemma on symplectomorphism.

@@
 They cannot be a vectors $v(q) e_q$, as the density $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$ is not invariant.
-They can, however, be covectors $k(q) dq$ as $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$.
-In other words: a covector is the simplest local geometrical object that allows coordinate independent densities.
+They can, however, be covectors $k(q) dq$ as $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$.\footnote{The components $k$ are the classical analogue of the wave numbers.} In other words: a covector is the simplest local geometrical object that allows coordinate independent densities.
 We can identify $\mathcal{S}$ as the cotangent bundle $T^*\mathcal{Q}$, the space of all possible covectors.
@@
 \begin{justification}
-We claim the vector field $S \in T\mathcal{Q}$ for the infinitesimal displacement $S = d\mathcal{s}/dt$ admits a potential $H$ such that, in component notation, $S^{a} \omega_{ab} = \partial_{b}H$.
-Let $v, w \in T_P\mathcal{Q}$ be two vectors defined at a point $P \in \mathcal{Q}$.
-Let $v^a, w^b$ their components in terms of unified state variables.
-Let $v', w' \in T_{\hat{P}}\mathcal{Q}$ the vector transported by the evolution map.
-For \ref{canonical_transformation} we have $v^{a} \omega_{ab} w^{b} = v'^{a} \omega_{ab} w'^{b}$.
-The evolution is a diffeomorphism, so we have $\xi(t+dt)^a = \xi^a + S^a dt$.
-The vector components change according to $v'^a = \partial_b \xi(t+dt)^a v^b = (\delta^a_b + \partial_b S^a dt) v^b$.
-We have:
-\begin{align*}
-v^{a} \omega_{ab} w^{b} &= v'^{a} \omega_{ab} w'^{b}  \\
-&= (v^{a} + \partial_{c} S^{a} v^{c} dt) \omega_{ab} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
-&= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} \\
-&+ v^{a} \omega_{ab} \partial_{d} S^{b} w^{d}) dt + O(dt^2)
-\end{align*}
-We define $S_{b} \equiv S^{a} \omega_{ab}$.
-The metric is invariant iff $v^{c} w^{b} \partial_{c} S_{b} - v^{a} w^{d} \partial_{d} S_{a} = 0$.
-The relationship must true for and pair of vector, therefore $\partial_{a} S_{b} - \partial_{b} S_{a} = curl(S_{a}) = 0$.
-$S$ is a curl free vector field and admits a potential $H$ such that $S_{a} = \partial_{a}H$.
+We claim the vector field $S \in T\mathcal{S}$ for the infinitesimal displacement $S^a = d\xi^a/dt$ admits a potential $H$ such that $S^{a} \omega_{ab} = \partial_{b}H$.
+The state space $\mathcal{S}$ of a classical homogeneous material is a symplectic manifold by \ref{classical_phase_space}.
+The map for infinitesimal time evolution $\mathcal{T}_{dt}$ is an infinitesimal self-symplectomorphism by \ref{canonical_transformation}.
+By \ref{symplectomorphism_generator} the infinitesimal displacement $S$ admits a potential $H$ such that $S^{a} \omega_{ab} = \partial_{b}H$
 
@@
 To do this we will extend phase space to include the temporal degree of freedom.
-We'll find that the dynamics is described by the relativistic Hamiltonian mechanics.
+We'll find that the dynamics is described by relativistic Hamiltonian mechanics in the extended phase space.
 We'll also find that particle states divide into standard and anti-states depending whether time is aligned or anti-aligned with the affine parameter.
+Note that no extra assumption is needed, which makes relativistic mechanics not some additional characterization but simply the correct way to describe time dependent motion of invariant density distributions.
 
@@
 
-The point is that we can't simply stick time in the Hamiltonian and expect everything to works out.
-But we can't ignore the problem either as the three cases outlined are pretty useful situation to study.
+In all three cases we can't simply stick time in the Hamiltonian and expect everything to works out.
+But we also can't ignore the problem either as the three cases outlined are pretty common situations to study.
 So we need to go back to our definitions and amend them properly.
@@
 
+Our goal is to describe maps that take a state at a particular time and transport it to another state at another time.
+Therefore we want to extend the domain and co-domain of the evolution map to be the space of all particle states at all times.
+
+The first thing to do is to extend our $n$-dimensional $\mathcal{Q}$, the manifold that defines our units, to include the time variable $t$.
+We call $\mathcal{M}$ such extended space which is a manifold of dimension $n+1$: state variables and time variable are independent (i.e.
+any combination of state and time is valid).
+A time variable is not simply another state variables: they don't identify extra states and it's only on the former that we have defined deterministic and reversible evolution.
+When changing variables, we still need to make sure that deterministic and reversible motion can be defined on the new time variable.
+Transformations like $t'=x \; x'=t$ (exchanging time with another state variable) or $t'=\sqrt{t^2 + q^2} \; x'=\tan^{-1}(t/x)$ (polar coordinates between time and state variable) clearly do not guarantee deterministic and reversible motion over the new time variable.
+The change of time variable must preserve the ordering therefore the transformation has to be strictly monotonic.
+In this sense, the time variable is special.
+
+As before, we are interested in invariant densities which are defined on co-vectors.
+The time variable $t$ will have a conjugate quantity $E\equiv\hbar\bar{\omega}$.\footnote{$\bar{\omega}$ is the classical analogue of the wave frequency.
+We use $\bar{\omega}$ to distinguish from the phase-space metric $\omega$.} We call the combination of $(t, E)$ the temporal degree of freedom.
+As time is special, so is the temporal degree of freedom, which is treated differently when defining distribution $\rho_\mathcal{c}$ and therefore by the measure $\mu$ and the metric $\omega$.
+
+The first thing to note is that $E$ cannot be an independent variable.
+It cannot add any states, as those are defined by $T^*\mathcal{Q}$ only, and cannot add any time instants, as those are defined by $t$ alone.
+Therefore there must exist a constraint such that, locally, $E=E(t,q^i,p_i)$.
+The set of all states at all time is a sub-manifold of $T^*\mathcal{M}$ of dimension $2n+1$.
+
+A composite state $\mathcal{c}$ and its distribution $\rho_\mathcal{c}$ is no longer defined at an instant of an absolute time parameter as now we are free to mix mix $t$ and $q$'s.
+Yet, there must be at least one time variable so that the distribution is at equal time defined over a suitable $T^*\mathcal{Q} \subset T^*\mathcal{M}$.
+That is, there must be at least one set of variables for which we are able to define states and their evolution.
+Therefore our distributions $\rho_\mathcal{c}$ are defined on $2n$-dimensional sub-manifolds of $T^*\mathcal{M}$.
+In other words: densities are not densities over the time variable.
+
+Yet, as we can mix time and state variables, the volume of integration is not simply $dq^n \wedge dp_n$ as it is not invariant.
+Consider a change of variable $q'=q+vt$: the same surface may have projections of different sizes on $(q, p)$ than $(q', p')$.
+It would seem, then, that the measure for our states is different depending on the choice of variable, which cannot be.
+
+We can understand the geometry of $T^*\mathcal{M}$ in the following way.
+As we have seen, the temporal degree of freedom $(t, E)$ and the other d.o.f.
+$(q^i, p_i)$ are not independent.
+So they are not orthogonal in the extended phase space.
+States are defined on the plane where $\langle q, p \rangle$ (maximally) change: this is not the plane of constant $\langle t, E \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle q, p \rangle$ where $dt \wedge dE$ is defined.
+It is on that plane we can properly count states and define our metric.
+
+We have a right triangle-like relationship between the plane where the metric is defined and its projections on the planes defined by each d.o.f., similar to the multiple d.o.f.:
+\begin{align*}
+m.d.o.f \;\;\; &dq^1 \wedge dp_1 + dq^2 \wedge dp_2 = \omega \\
+t.d.o.f \;\;\; &dt \wedge dE + \omega = dq \wedge dp \\
+\end{align*}
+But in the previous case, the right angle was between the two independent d.o.f..
+In this case, the right angle is between the metric and the plane of constant $(q, p)$ where $dt \wedge dE$ is defined.
+We rewrite it as $dq \wedge dp - dt \wedge dE = \omega$.
+This corresponds to the Minkowski product across d.o.f.
+and the vector product within.
+In terms of unified state variables $\xi^a\equiv \{t, q^i, E, p_i\}$ we have:
+
+\begin{align*}
+\omega_{ab} =  \left[
+\begin{array}{cc}
+0 & 1 \\
+-1 & 0 \\
+\end{array}
+\right] \otimes \left[
+\begin{array}{cc}
+-1 & 0 \\
+0 & I_n \\
+\end{array}
+\right] =
+\left[
+\begin{array}{cccc}
+0 & 0 & -1 & 0 \\
+0 & 0 & 0 & I_n \\
+1 & 0 & 0 & 0 \\
+0 & -I_n & 0& 0 \\
+\end{array}
+\right] \\
+\end{align*}
+
+The metric $\omega$ is still the canonical two form expressed with the mathematically non-canonical, but physically meaningful, variable $E$.
+This allows us to better understand its physical meaning.
+The metric allows to count the possibilities of within a degree of freedom adjusting the metric on $T^*\mathcal{Q}$ to avoid the counting problems introduced by time variable changes.
+
+The metric actually fixes two other problems.
+The first is: how do we distinguish temporal from standard d.o.f.?
+Given a two-dimensional surface $U \subset T^*\mathcal{M}$, how can we tell if it should be charted by a $(t,E)$ or whether we can define a distribution on it?
+Consider $\int_U \omega$.
+The result will be positive if the integration is over a standard degree of freedom, and it will be negative if the integration is over the temporal degree of freedom.
+If the contribution in any subregion of $U$ is positive, then the whole $U$ is always oriented along a standard degree of freedom.
+
+The second problem is: how do we prevent changes in time variable that disturb determinism and reversibility?
+The invariance of $\omega$ already guarantees that.
+As we saw before the vector product is conserved within a degree of freedom is equivalent to saying that changing variables conserves the possibilities within a d.o.f.
+In fact, the flow of an infinitesimal transformation that preserve the vector product an a surface is a divergence free field.
+The conservation of the Minkowsky product across the temporal and standard d.o.f.
+means that changing variables does not introduce rotation between time and state variables.
+In face, the flow of an infinitesimal transformation that preserve the Minkowsky product on a surface is a curl free field.
+
+We can capture the above discussion by stating that the extended state space is an hypersurface of the symplectic manifold $(T^*\mathcal{M}, \omega)$, where $\omega \equiv \sum dq^i \wedge dp_i - dt \wedge dE$.
+This is the simplest manifold that allows us to define time-and-state-variable-invariant densities.
+The metric allows us to measure the cardinality of possibilities.
+
+\begin{prop}\label{classical_extended_phase_space}
+	The extended state space, the set of states at all possible times, for the particles of a homogeneous classical material is an hyper-surface of the symplectic manifold formed by a cotangent bundle $T^*\mathcal{M}$ equipped with the canonical two-form $\omega = \sum dq^i \wedge dp_i - dt \wedge dE$.
+\end{prop}
+
+\begin{justification}
 TODO
-
-% Extend Q to include time -> M. Extended phase space is a hypersurface T*M.
-
-% Call s (affine parameter) the parameter of the evolution.
-Distribution defined on a n-dimensional surface in T*M.
-Density does not change under time transformation.
-Use t'=t + \Delta t(q^i, p_i) so that t' is constant.
-Now we are at equal t': the metric is \Sigma dq^i \wedge dp_i.
-Metric is the projection on the space orthogonal to t and E.
-That is \Sigma dq^i \wedge dp_i - dt \wege dE.
-
-% definition, simplectic metric on phase space
-
 % justification.
@@
 Can't be positive, as ()t, E) is not an independent d.o.f.
+\end{justification}
 
@@
 
+As we are using time as a variable to identify state, we will no longer it as the parameter for the evolution.
+The trajectory of a particle in the extended phase space will be given by the evolved variables $\xi^a(s)$ in terms of a parameter $s$.
+As before, deterministc and reversible evolution will preserve the metric as the number of possibilities on each independent d.o.f.
+will be conserved.
+Mathematically,  deterministic and reversible evolution is a self-symplectomorphism.
+
+\begin{prop}\label{relativistic_canonical_transformation}
+	A time-dependent deterministic and reversible evolution map for a homogeneous classical material is self-symplectomorphism on the extended phase space.
+That is: $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ and $\mathcal{T}_{\Delta s}^*\omega = \omega$ where $\mathcal{T}_{\Delta s}^*$ is the pullback of $\mathcal{T}_{\Delta s}$.
+\end{prop}
+
+\begin{justification}
 	TODO
+\end{justification}
+
+Part of the trajectory describe the evolution of the time variable $t(s)$.
+Since the evolution is deterministic and reversible, for each value of $s$ we need to have one and only one value of $t$.
+$t(s)$ is invertible, strictly monotonic and $d_{s}t$ along a trajectory cannot change sign.\footnote{} This means there are two classes of states: the ones where the parametrization $s$ is aligned with time $t$, which we call standard states, and the ones where the parametrization $s$ is anti-aligned with time $t$, which we call anti-states.
+Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call standard and anti-states is also conventional.
+What is physical and not conventional, though, is that standard and anti-states cannot be connected by deterministic and reversible evolution.\footnote{This represents the classical analogue for quantum anti-particle states.}
+
+\begin{prop}\label{antistates}
+	Let $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ the time-variant deterministic and reversible evolution map for the particles of a homogeneous classical material, where $s$ is the continuous evolution parameter.
+The map partitions the extended state space into standard states, those connected by a trajectory where $d_{s}t>0$, and anti-states, those connected by a trajectory where $d_{s}t<0$.
+\end{prop}
+
+\begin{justification}
+	TODO
+\end{justification}
 
@@
 
+\begin{prop}\label{symplectomorphism_generator}
+	Let $(M, \omega)$ a symplectic manifold.
+Let $f: (M, \omega) \rightarrow (M, \omega)$ an infinitesimal self-symplectomorphism.
+Let $S \in TM \; | \; f(\xi^a(m)) = \xi^a(m) + S^a(m)dt \; \forall m \in M$ be the infinitesimal displacement.
+There exist a function $H: M \rightarrow \mathbbm{R}$ such that $S^{a} \omega_{ab} = \partial_{b}H$.
+\end{prop}
+
+\begin{justification}
+	We claim the vector field $S \in T\mathcal{Q}$ admits a potential $H$ such that $S^{a} \omega_{ab} = \partial_{b}H$.
+Let $v, w \in T_m M$ be two vectors defined at a point $m \in M$.
+Let $v^a, w^b$ their components.
+Let $v'\equiv f v, w' \in T_{f(m)}M$ be the pushforward of $v, w$ by $f$.
+Since $f$ is a symplectomorphism we have $v^{a} \omega_{ab} w^{b} = v'^{a} \omega_{ab} w'^{b}$.
+The vector components change according to $v'^a = \partial_b \xi(t+dt)^a v^b = (\delta^a_b + \partial_b S^a dt) v^b$.
+We have:
+	\begin{align*}
+	v^{a} \omega_{ab} w^{b} &= v'^{a} \omega_{ab} w'^{b}  \\
+	&= (v^{a} + \partial_{c} S^{a} v^{c} dt) \omega_{ab} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
+	&= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} \\
+	&+ v^{a} \omega_{ab} \partial_{d} S^{b} w^{d}) dt + O(dt^2)
+	\end{align*}
+	$v^{c} w^{b} \partial_{c} S_{b} - v^{a} w^{d} \partial_{d} S_{a} = 0$ where $S_{b} \equiv S^{a} \omega_{ab}$.
+The relationship must true for and pair of vector, therefore $\partial_{a} S_{b} - \partial_{b} S_{a} = curl(S_{a}) = 0$.
+$S$ is a curl free vector field and admits a potential $H$ such that $S_{a} = \partial_{a}H$.
+\end{justification}
+
 \begin{thebibliography}{0}

### 2016-06-01 2015CompleteDerivation/completeDerivation.tex
Review of classical hamiltonian

@@
 
-We will call a classical system one that is infinitesimally reducible.
+We will call a \emph{particle} the smallest amount of a material to which we can assign an independent state.
+For example, a photon is a particle of light (it is the smallest amount of light we can describe), an infinitesimal amount of water is treated classically as a point particle (the smallest amount for a continuous fluid).
+We use this definition as it allows us to retroactively talk on somewhat equal grounds about classical point-particles and quantum particles, and underline how physically (and mathematically) they are very similar yet very different objects.
+
+We will call a \emph{classical system} one that is infinitesimally reducible.
 That is, we can keep decomposing the system indefinitely into smaller and smaller parts, each with a well defined state.
-We call particles the infinitesimal parts given by the limit of this process.
+In this case, the particles are the infinitesimal parts given by the limit of this process of recursive decomposition.
 Giving the state of the whole system is equivalent to giving the states of all particles.
@@
 The amount in each tank is continuous (as we assume the water to be infinitesimally divisible) yet there are only a finite number of tanks the water can be placed into.
-As it links two physically distinguishable objects, $\rho_\mathcal{c}$ is a continuous function as discussed before.
+As it links two physically distinguishable objects, $\rho_\mathcal{c}$ is a continuous function as discussed in \ref{continuous_map}.
 
@@
 If we take a finite volume of an ideal gas at equilibrium, the momentum distribution spans over the whole range.
-Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a general physical justification for.
+Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a general physical justification for.\footnote{One can, though, make the argument that given any distribution $\rho$ one can find an infinitely smooth $\rho_{sm}$ such that the difference in description is small compared to the error already introduced by assuming the system to be homogeneous and infinitely reducible.
+}
 
@@
 The second represents the amount of material added (or taken away if negative) by the state change.
-For example, if $\rho_{\delta \mathcal{c}}$ is the change due to deterministic and reversible evolution, the amount of material does not change and therefore $\int_{\mathcal{S}} \rho_{\delta \mathcal{c}} d\mu = 0$.
+For example, if $\delta \rho$ is the change due to deterministic and reversible evolution, the amount of material does not change and therefore $\int_{\mathcal{S}} \delta \rho \, d\mu = 0$.
 The first norm would represent the total amount of material that is changing (i.e.
-being added and being removed), therefore $\int_{\mathcal{S}} |\rho_{\delta \mathcal{c}}| d\mu = 0$ means no material is moving, no change is occurring.
+being added and being removed), therefore $\int_{\mathcal{S}} | \delta \rho | \, d\mu = 0$ means no material is moving, no change is occurring.
 
@@
 This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows state-variable-independent densities.
+
 Under an arbitrary change of state variables, the density has to remain the same, which means that the Jacobian of the transformation must exist and must be unitary.
-
 For the Jacobian to exist the transformation between state variables has to be differentiable.
@@
 	We claim the simplest state space $\mathcal{S}$ that allows state-variable-invariant densities is the symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of a one dimensional differential manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
-The description of the simplest physical object on which we can  state-variable-invariant must be given by at least one state variable $q$.
+The description of the simplest physical object on which we can define a state-variable-invariant must be given by at least one continuous state variable $q$.
 Let $\mathcal{Q}$ be the manifold charted by $q$.
@@
 Let's assume $q$ is enough to define the unit system for the description of the physical object.
-Then the physical object is represented by a set of geometrical objects in $\mathcal{Q}$.
+Then the physical object is represented by a geometrical object on $\mathcal{Q}$.
 The physical object is infinitesimal, therefore is represented by a local geometrical object.
@@
 $\int_{\mathcal{S}} \rho_\mathcal{c} d\mu \propto \int_{\mathcal{S}} \rho_\mathcal{c}(q, k) dq \wedge dk$.
-$d\mu = \hbar dq \wedge dk = dq \wedge dp = \omega$ where $\hbar$ is a constant and $\omega$ is the canonical two form.
+$d\mu = \hbar dq \wedge dk = dq \wedge dp = \omega$ where $\hbar$ is a constant, $p\equiv \hbar k$  and $\omega$ is the canonical two form.
 $\omega$ is invariant under state variable change.
-$\omega$ is a symplectic metric for $\mathcal{S}=T*\mathcal{Q}$.
-$\mathcal{S} \cong(T^*\mathcal{Q}, \omega)$ as a symplectic manifold.
-
-	We claim the state space $\mathcal{S}$ for the particles of a homogeneous classical material is a symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of a one dimensional differential manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
+$\omega$ is a symplectic metric for $\mathcal{S}=T^*\mathcal{Q}$.
+$\mathcal{S} = (T^*\mathcal{Q}, \omega)$ is a symplectic manifold.
+
+	We claim the state space $\mathcal{S}$ for the particles of a homogeneous classical material is a symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of an $n$-dimensional differential manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
 Let $q^i$ be a set of n continuous independent state-variable that define the units to describe a state in $U \subseteq \mathcal{S}$.
@@
 Each degree of freedom must define state-variable-distribution independently.
-For each $\mathcal{Q}^i$ we need a covector $k_i(q) dq^i \in T^*\mathcal{Q}^i$.
+For each $\mathcal{Q}^i$ we have a covector $k_i(q) dq^i \in T^*\mathcal{Q}^i$.
 Locally $\mathcal{S} \cong \prod T^*\mathcal{Q}^i \cong T^* \prod \mathcal{Q}^i \cong T^* \mathcal{Q}$.
@@
 $\omega = \sum dq^i \wedge dp_i$.
-$\mathcal{S} \cong (T^*\mathcal{Q}, \omega)$ as a symplectic manifold.
+$\mathcal{S} = (T^*\mathcal{Q}, \omega)$ is a symplectic manifold.
 \end{justification}
@@
 	Consistently with what we said in Assumption \ref{detrevass}, if we defined a state for each particle, then a deterministic and reversible evolution on such state must exist.
-Yet, the idea that we can assign states to infinitesimal parts should be considered only a \emph{simplifying} assumption, and it is instructive to understand when and how it breaks down.
+Yet, the idea that we can assign states to infinitesimal parts should be considered only a \emph{simplifying} assumption.
+The obvious reason is that we know this does not work in practice: as we keep decomposing the material we end up with elementary particles.
+But it is more instructive to understand when and how the assumption breaks down at a more conceptual level.
 
@@
 
-    Being able to track parts at a given scale is not even enough.
-The parts are infinitesimal, the limit of a process of infinite recursive subdivision, so we must be able to track at any scale.
-Therefore the assumption can only be validated experimentally to the smallest precision available.
+    Even if we are able to track parts, the assumption requires them to be infinitesimal, the limit of a process of infinite recursive subdivision.
+The best we can do experimentally is to confirm that the assumption holds up to the smallest precision available.
+Therefore even in the best of cases it cannot be considered an experimentally validated assumption but a reasonable default assumption (e.g.
+"it worked so far").
 
@@
 
-	In the light of what we discussed, we cannot take the classical assumption to be fundamental, in the sense that we cannot take it to apply to all of physics.
-While ultimately flawed, the  assumption can be considered valid for a great number of macroscopic systems, and that is why it is very useful.
+	In light of what we discussed, we cannot take the classical assumption to be fundamental, in the sense that we cannot take it to apply to all of physics.
+While ultimately flawed, the  assumption can be considered valid for a great number of macroscopic systems, and is also useful as a default assumption of sorts.
 Understanding its shortcomings will help us later to see how the quantum assumption solves, at least partially, some of these issues.

### 2016-06-02 2015CompleteDerivation/completeDerivation.tex
Christine's comments on classical hamilotnian

@@
 Yet, the idea that we can assign states to infinitesimal parts should be considered only a \emph{simplifying} assumption.
-The obvious reason is that we know this does not work in practice: as we keep decomposing the material we end up with elementary particles.
+The obvious reason is that we know this does not work in practice: as we keep decomposing the material we end up with molecules, atoms and then subatomic particles.
 But it is more instructive to understand when and how the assumption breaks down at a more conceptual level.
@@
 
-	If we just assume the whole system is deterministic and reversible, then each part can evolve depending the states of the other parts.
+	If we just assume the whole system is deterministic and reversible, then each part can evolve depending on the states of the other parts.
 But this does not solve the problem entirely.
@@
 \begin{prop}\label{canonical_transformation}
-	A deterministic and reversible evolution map for a homogeneous classical material is self-symplectomorphism.
+	A deterministic and reversible evolution map for a homogeneous classical material is a self-symplectomorphism.
 That is: $\mathcal{T}_{\Delta t}: T^*\mathcal{Q} \rightarrow T^*\mathcal{Q}$ and $\mathcal{T}_{\Delta t}^*\omega = \omega$ where $\mathcal{T}_{\Delta t}^*$ is the pullback of $\mathcal{T}_{\Delta t}$.
@@
 	We claim $\mathcal{T}_{\Delta t}$ is a self-homeomorphism on $T^*\mathcal{Q}$.
-The state space for particles of a homogeneous classical material is $T^*\mathcal{Q}$ by \ref{classical_phase_space} $\mathcal{T}_{\Delta t}$ is a deterministic and reversible evolution map and \ref{detrevmap}.
+The state space for particles of a homogeneous classical material is $T^*\mathcal{Q}$ by \ref{classical_phase_space}.
+$\mathcal{T}_{\Delta t}$ is a deterministic and reversible evolution map and by \ref{detrevmap} is a self-homeomorphism.
 
@@
 
-In the special case of continuous time evolution we can prove the following:
+If we assume continuous time evolution we have the following:
 
 \begin{prop}\label{hamiltonian}
-	A continuous deterministic and reversible process for a homogeneous classical material admits a Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ that allows to write the laws of evolution as
+	A continuous deterministic and reversible process for a homogeneous classical material admits a Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ that allows us to write the laws of evolution as
 \begin{align*}
@@
 For $b=\{1,...,n\}$ we have $d_tp_i (-1) = \partial_{q_i} H$.
-For $a=\{n+1,...,2n\}$ we have $d_tq^i (+1) = \partial_{p_i} H$.
+For $b=\{n+1,...,2n\}$ we have $d_tq^i (+1) = \partial_{p_i} H$.
 \end{justification}
@@
 We recognize the familiar set of Hamilton's equations.
-They are the set of equation to describe the deterministic and reversible motion of the infinitesimal parts of an infinitesimally reducible homogeneous material.
+They are the set of equations to describe the deterministic and reversible motion of the infinitesimal parts of an infinitesimally reducible homogeneous material.
 In other words: the forces that conserve energy (i.e.
@@
 It's important to realize that, during the derivation, multiple mathematical features (e.g.
-invariant densities, cotangent bundle for phase space, symplectomorphism) where justified by the same physical assumption.
-As such, the math itself does not give us a conceptually unified picture, it gets lost.
-This is one of the reason we believe that starting from the physical description is objectively better if we are to come to a better understanding of our physical theories.
+invariant densities, cotangent bundle for phase space, symplectomorphism) were justified by the same physical assumption.
+The math itself gives us no indication that the different features stem from the same source; therefore, the math itself does not give us a conceptually unified picture.
+This is one of the reasons we believe that starting from the physical description is objectively better if we are to come to a better understanding of our physical theories.
 
 Note that we could have taken different approaches.
-For example, we could have appealed to information theory and asked our invariant distributions to preserve Shannon's information entropy, as no information should be gained or lost during a deterministic and reversible process.
-Or we could have appealed to statistical mechanics and required that the determinant of the covariance matrix be conserved, as deterministic and reversible process should be defined at the same level of precision.
+For example, we could have appealed to information theory and required our invariant distributions to preserve Shannon's information entropy, as no information should be gained or lost during a deterministic and reversible process.
+Or we could have appealed to statistical mechanics and required that the determinant of the covariance matrix be conserved, as a deterministic and reversible process should be defined at the same level of precision.
 With suitable treatment of independent d.o.f.
 both these approaches recover Hamiltonian mechanics as well.
-While the full treatment goes outside the scope of this work, we want to underline that the definitions used in this work go a long way to building bridges at the core of different mathematical and scientific disciplines.
+While the full treatment is outside the scope of this work, we want to underline that the definitions used in this work go a long way to building bridges at the core of different mathematical and scientific branches.
 
@@
 The discussion so far has been limited to the case where both the set of states and the evolution map are defined at equal time and never change throughout the evolution.
-This is too restrictive as there are very reasonable situation in which this does not hold.
-
-The first obvious case is that is that the map may be time dependent.
+This is too restrictive as there are very reasonable situations in which this does not hold.
+
+The first obvious case is that the map may be time dependent.
 Note that this does not affect our definition of determinism and reversibility: we still can tell final state from initial state and vice-versa.
@@
 The third case is when we perform a state variable transformation $q'=q'(q,t)$ that is time dependent (e.g.
-$x'=x+vt$).
+$q'=q+vt$).
 Our composite state distribution $\rho(q)$ becomes $\rho(q',t)$: it is no longer defined at equal time.
@@
 
-In all three cases we can't simply stick time in the Hamiltonian and expect everything to works out.
-But we also can't ignore the problem either as the three cases outlined are pretty common situations to study.
+Therefore we cannot simply stick time in the Hamiltonian and expect everything to work out.
+But we can't ignore the problem either as the three cases outlined are pretty common situations to study.
 So we need to go back to our definitions and amend them properly.
@@
 The first thing to do is to extend our $n$-dimensional $\mathcal{Q}$, the manifold that defines our units, to include the time variable $t$.
-We call $\mathcal{M}$ such extended space which is a manifold of dimension $n+1$: state variables and time variable are independent (i.e.
+We call $\mathcal{M}$ such extended space which is a manifold of dimension $n+1$: state variables and the time variable are independent (i.e.
 any combination of state and time is valid).
-A time variable is not simply another state variables: they don't identify extra states and it's only on the former that we have defined deterministic and reversible evolution.
+A time variable is not simply another state variable: it doesn't identify extra states and it's the variable on which we have defined deterministic and reversible evolution.
 When changing variables, we still need to make sure that deterministic and reversible motion can be defined on the new time variable.
-Transformations like $t'=x \; x'=t$ (exchanging time with another state variable) or $t'=\sqrt{t^2 + q^2} \; x'=\tan^{-1}(t/x)$ (polar coordinates between time and state variable) clearly do not guarantee deterministic and reversible motion over the new time variable.
-The change of time variable must preserve the ordering therefore the transformation has to be strictly monotonic.
+Transformations like $t'=q, \; q'=t$ (exchanging time with another state variable) or $t'=\sqrt{t^2 + q^2}, \; q'=\tan^{-1}(t/q)$ (polar coordinates between time and state variable) clearly do not guarantee deterministic and reversible motion over the new time variable.
+The change of time variable must preserve the time ordering therefore the transformation has to be strictly monotonic.
 In this sense, the time variable is special.
@@
 As before, we are interested in invariant densities which are defined on co-vectors.
+TODO: elaborate on covector component.
 The time variable $t$ will have a conjugate quantity $E\equiv\hbar\bar{\omega}$.\footnote{$\bar{\omega}$ is the classical analogue of the wave frequency.
@@
 Therefore there must exist a constraint such that, locally, $E=E(t,q^i,p_i)$.
-The set of all states at all time is a sub-manifold of $T^*\mathcal{M}$ of dimension $2n+1$.
+The set of all states at all times is a sub-manifold of $T^*\mathcal{M}$ of dimension $2n+1$.
 
@@
 So they are not orthogonal in the extended phase space.
-States are defined on the plane where $\langle q, p \rangle$ (maximally) change: this is not the plane of constant $\langle t, E \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle q, p \rangle$ where $dt \wedge dE$ is defined.
+States are defined on the plane where $( q, p )$ (maximally) change: this is not the plane of constant $( t, E )$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $( q, p )$ where $dt \wedge dE$ is defined.
 It is on that plane we can properly count states and define our metric.
@@
 This allows us to better understand its physical meaning.
-The metric allows to count the possibilities of within a degree of freedom adjusting the metric on $T^*\mathcal{Q}$ to avoid the counting problems introduced by time variable changes.
+The metric allows us to count the possibilities within a degree of freedom, adjusting the metric on $T^*\mathcal{Q}$ to avoid the counting problems introduced by time variable changes.
 
@@
 The result will be positive if the integration is over a standard degree of freedom, and it will be negative if the integration is over the temporal degree of freedom.
-If the contribution in any subregion of $U$ is positive, then the whole $U$ is always oriented along a standard degree of freedom.
-
-The second problem is: how do we prevent changes in time variable that disturb determinism and reversibility?
+If the contribution in every subregion of $U$ is positive, then the whole $U$ is always oriented along a standard degree of freedom.
+
+The second problem is: how do we prevent time variable changes that disturb determinism and reversibility?
 The invariance of $\omega$ already guarantees that.
-As we saw before the vector product is conserved within a degree of freedom is equivalent to saying that changing variables conserves the possibilities within a d.o.f.
-In fact, the flow of an infinitesimal transformation that preserve the vector product an a surface is a divergence free field.
+As we saw before that fact that the vector product is conserved within a degree of freedom is equivalent to saying that changing variables conserves the possibilities within a d.o.f.
+In fact, the flow of an infinitesimal transformation that preserves the vector product on a surface is a divergence free field.
 The conservation of the Minkowsky product across the temporal and standard d.o.f.
 means that changing variables does not introduce rotation between time and state variables.
-In face, the flow of an infinitesimal transformation that preserve the Minkowsky product on a surface is a curl free field.
-
-We can capture the above discussion by stating that the extended state space is an hypersurface of the symplectic manifold $(T^*\mathcal{M}, \omega)$, where $\omega \equiv \sum dq^i \wedge dp_i - dt \wedge dE$.
+In fact, the flow of an infinitesimal transformation that preserves the Minkowsky product on a surface is a curl free field.
+
+We can capture the above discussion by stating that the extended state space is a hypersurface of the symplectic manifold $(T^*\mathcal{M}, \omega)$, where $\omega \equiv \sum dq^i \wedge dp_i - dt \wedge dE$.
 This is the simplest manifold that allows us to define time-and-state-variable-invariant densities.
@@
 \begin{prop}\label{classical_extended_phase_space}
-	The extended state space, the set of states at all possible times, for the particles of a homogeneous classical material is an hyper-surface of the symplectic manifold formed by a cotangent bundle $T^*\mathcal{M}$ equipped with the canonical two-form $\omega = \sum dq^i \wedge dp_i - dt \wedge dE$.
+	The extended state space, the set of states at all possible times, for the particles of a homogeneous classical material is a hyper-surface of the symplectic manifold formed by a cotangent bundle $T^*\mathcal{M}$ equipped with the canonical two-form $\omega = \sum dq^i \wedge dp_i - dt \wedge dE$.
 \end{prop}

### 2016-06-06 2015CompleteDerivation/completeDerivation.tex
Changing numbering for equations
Concluded first pass at relativity

@@
 \usepackage{dutchcal}
+\usepackage{braket}
+
+\renewcommand{\theequation}{\thesection.\arabic{equation}}
+\numberwithin{equation}{section}
 
@@
 \renewcommand*{\theassump}{\Roman{assump}}
-\newtheorem{thrm}{Theorem}[section]
+\newtheorem{thrm}[equation]{Theorem}
 
 \theoremstyle{definition}
-\newtheorem{prop}{Proposition}[section]
+\newtheorem{prop}[equation]{Proposition}
 
@@
 
-\begin{prop}\label{hamiltonian}
+\begin{prop}\label{hamiltons_equations}
 	A continuous deterministic and reversible process for a homogeneous classical material admits a Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ that allows us to write the laws of evolution as
@@
 Physically the transformation is introducing fictitious forces that are not conservative, so they cannot be expressed by a Hamiltonian.
-But this means that we have an ill defined mathematical framework as the notion of deterministic and reversibility is not defined in a way that is independent of time transformations.
+But this means that we have an ill defined mathematical framework as the notion of determinism and reversibility is not defined in a way that is independent of time transformations.
 We need a framework that is capable of handling the fictitious forces as well.
@@
 Transformations like $t'=q, \; q'=t$ (exchanging time with another state variable) or $t'=\sqrt{t^2 + q^2}, \; q'=\tan^{-1}(t/q)$ (polar coordinates between time and state variable) clearly do not guarantee deterministic and reversible motion over the new time variable.
-The change of time variable must preserve the time ordering therefore the transformation has to be strictly monotonic.
+The change of time variable must preserve the time ordering therefore the transformation has to be strictly monotonic between the two time variables.
 In this sense, the time variable is special.
@@
 As before, we are interested in invariant densities which are defined on co-vectors.
-TODO: elaborate on covector component.
-The time variable $t$ will have a conjugate quantity $E\equiv\hbar\bar{\omega}$.\footnote{$\bar{\omega}$ is the classical analogue of the wave frequency.
-We use $\bar{\omega}$ to distinguish from the phase-space metric $\omega$.} We call the combination of $(t, E)$ the temporal degree of freedom.
+The co-vector $k_i dq^i$ is extended with a time component $\bar{\omega} dt$.\footnote{$\bar{\omega}$ is the classical analogue of the wave frequency.
+We use $\bar{\omega}$ to distinguish from the phase-space metric $\omega$.} As state variable $q^i$ have a conjugate quantity $p_i\equiv \hbar k$, the time variable $t$ will have a conjugate quantity $E\equiv\hbar\bar{\omega}$.
+We call the combination of $(t, E)$ the temporal degree of freedom.
 As time is special, so is the temporal degree of freedom, which is treated differently when defining distribution $\rho_\mathcal{c}$ and therefore by the measure $\mu$ and the metric $\omega$.
@@
 So they are not orthogonal in the extended phase space.
-States are defined on the plane where $( q, p )$ (maximally) change: this is not the plane of constant $( t, E )$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $( q, p )$ where $dt \wedge dE$ is defined.
-It is on that plane we can properly count states and define our metric.
+States are defined on the plane where $( q, p )$ (maximally) change.
+This is not the plane of constant $( t, E )$ (they are not orthogonal) where the area given by $dq \wedge dp$ is defined.
+It is the plane perpendicular to constant $( q, p )$.
+And the plane of constant $( q, p )$ is where the area given by $dt \wedge dE$ is defined.
+That is: the plane where we can properly count states and define our metric is perpendicular to $dt \wedge dE$.
 
@@
 m.d.o.f \;\;\; &dq^1 \wedge dp_1 + dq^2 \wedge dp_2 = \omega \\
-t.d.o.f \;\;\; &dt \wedge dE + \omega = dq \wedge dp \\
+t.d.o.f \;\;\; &dt \wedge dE + \omega = dq \wedge dp
 \end{align*}
@@
 
-As we are using time as a variable to identify state, we will no longer it as the parameter for the evolution.
+As we are using time as a variable to identify states, we will use a different quantity as the parameter for the evolution.
 The trajectory of a particle in the extended phase space will be given by the evolved variables $\xi^a(s)$ in terms of a parameter $s$.
@@
 
-Part of the trajectory describe the evolution of the time variable $t(s)$.
+If we assume continuous time evolution we have the following:
+
+\begin{prop}\label{relativistic_hamiltons_equations}
+	A continuous time-dependent deterministic and reversible process for a homogeneous classical material admits an extended Hamiltonian $\mathcal{H}: T^*\mathcal{M} \rightarrow \mathbbm{R}$ that allows us to write the laws of evolution as
+	\begin{align*}
+	d_{s}t &= - \partial_{E} \mathcal{H} \\
+	d_{s}E &= \partial_{t} \mathcal{H} \\
+	d_{s}q^i &= \partial_{p_i} \mathcal{H} \\
+	d_{s}p_i &= - \partial_{q^i} \mathcal{H}
+	\end{align*}
+\end{prop}
+
+\begin{justification}
+	TODO
+\end{justification}
+
+The trajectory in time $t(s)$ is of particular importance.
 Since the evolution is deterministic and reversible, for each value of $s$ we need to have one and only one value of $t$.
-$t(s)$ is invertible, strictly monotonic and $d_{s}t$ along a trajectory cannot change sign.\footnote{} This means there are two classes of states: the ones where the parametrization $s$ is aligned with time $t$, which we call standard states, and the ones where the parametrization $s$ is anti-aligned with time $t$, which we call anti-states.
+$t(s)$ is invertible, strictly monotonic and $d_{s}t$ along a trajectory cannot change sign.
+This means there are two classes of states: the ones where the parametrization $s$ is aligned with time $t$, which we call standard states, and the ones where the parametrization $s$ is anti-aligned with time $t$, which we call anti-states.
+Let us call $\lambda : T^*\mathcal{M} \rightarrow \mathbbm{R}$ the function $\lambda (\xi^a) \mapsto d_s t |_{\xi^a}$ that returns the change of $t$ along $s$.
+$\lambda > 0$ for all standard states while $\lambda < 0$ for all anti-states.
 Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call standard and anti-states is also conventional.
@@
 
-% find equation of motion
-
-% find constraint
-
-\subsection{anti-states}
-
-% relationship between s and t must be monotonic. Hamiltonian partitions extended phase space. Distribution must be always zero at the boundary.
+With the extended Hamiltonian we are able to write the equation of motions for any choice of time and state variables.
+For example: $d_t q^i = d_t s \, d_s q^i = d_s q^i / d_s t = - \partial_{p_i} \mathcal{H} / \partial_{E} \mathcal{H}$.
+These equations must be the same as the one given by a standard Hamiltonian in those coordinates.
+That is: $d_t q^i = \partial_{p_i} H = - \partial_{p_i} \mathcal{H} / \partial_{E} \mathcal{H}$.
+The partial derivatives of $H$ and $\mathcal{H}$ are related, and therefore the functions themselves are related.
+Working through the math, we find that the most general relationship is of the form: $\mathcal{H} = \lambda(t,E,q^i,p_i) \, (H(t,q^i,p_i) - E - \mathcal{h}(t))$ where $\lambda = d_s t$, $\mathcal{h}(t)$ is an arbitrary function and, most importantly, $E = H(t,q^i,p_i) + \mathcal{h}(t)$.
+
+The arbitrary function $\mathcal{h}(t)$ has no physical consequence: it is just a constant added to the Hamiltonian.
+We can either set it to zero or simply redefine the Hamiltonian to include it.
+Therefore we are left with $E = H(t,q^i,p_i)$.
+In other words: $E$ is the value of the Hamiltonian, the energy of the system.
+We have found the constraint we discussed when introducing \ref{classical_extended_phase_space}.
+
+Note that if $E = H$ then $\mathcal{H}=0$ for all states.
+For all states we also have $\lambda \neq 0$.
+Therefore $\mathcal{H}=0$ only because $E = H$.
+We can extend $\mathcal{H}$ on all of $T^* \mathcal{M}$ making sure that $\mathcal{H} \neq 0$ outside of the hyper-surface that contains states.
+This way we can directly use the constraint $\mathcal{H}=0$ to identify such hyper-surface.
+
+This way $\mathcal{H}$ gives both the constraint $\mathcal{H}=0$ to identify states and the map $d_s \xi^a=\omega^{ab} \partial_b \mathcal{H}$ to identify their evolution.
+
+\begin{prop}\label{form_of_extended_hamiltonian}
+	The form of the extended Hamiltonian is $\mathcal{H} = \lambda(t,E,q^i,p_i) \, (H(t,q^i,p_i) - E)$ where $H(t,q^i,p_i)$ is the standard Hamiltonian at each instant $t$, the conjugate time variable $E = H(t,q^i,p_i)$ represents the value of the Hamiltonian, and  $\lambda : T^*\mathcal{M} \rightarrow \mathbbm{R}$ where $\lambda(\xi^a) = d_s t$.
+\end{prop}
+
+\begin{justification}
+	TODO
+\end{justification}
+
+As an example, the Hamiltonian and extended Hamiltonian for a relativistic free particle the Hamiltonian are:
+\begin{equation}\label{free_hamiltonians}
+\begin{aligned}
+H &= c \sqrt{m^2 c^2 + p^i p_i} \\
+\mathcal{H} &= \frac{1}{2m} ( p^i p_i - (E/c)^2 + m^2c^2) \\
+ &= \frac{1}{2mc^2} (c \sqrt{m^2 c^2 + p^i p_i} + E) \\
+ &(c \sqrt{m^2 c^2 + p^i p_i} - E)
+\end{aligned}
+\end{equation}
+In the next section we'll derive and look more closely at the case of particles under potential forces.
+
+Now that we have characterized the motion of the particles, we should also characterize the motion of the composite state.
+
+If we have a distribution of material $\rho_\mathcal{c}(s)$ for a particular value of $s$, the first thing we know is that its support, the region where the density is non-zero, must be within the constraint $\mathcal{H} = 0$.
+This means $\mathcal{H} \rho_\mathcal{c}(s) = 0$ since $\mathcal{H} = 0$ where $\rho_\mathcal{c}(s) \neq 0$.
+If the distribution is differentiable, the evolution is given by the Poisson bracket $d_s \rho_\mathcal{c} (s) = \{ \rho_\mathcal{c}(s), \mathcal{H} \}$.
+
+We can also consider the distribution of material at all times $\rho_\mathcal{c}\equiv \int \rho_\mathcal{c}(s) ds$.
+The support of the distribution is also constrained so we have $\mathcal{H} \rho_\mathcal{c} = 0$.
+As the density is transported over each trajectory, its value does not change over $s$.
+Therefore, if the distribution is differentiable, we have $d_s \rho_\mathcal{c} = \{ \rho_\mathcal{c}, \mathcal{H} \} = 0$.
+
+\begin{prop}\label{relativistic_distribution_evolution}
+	Let $\rho_\mathcal{c}(s)$ the distribution of material at a particular moment in the evolution.
+$\mathcal{H} \rho_\mathcal{c}(s) = 0$ and $d_s \rho_\mathcal{c} (s) = \{ \rho_\mathcal{c}(s), \mathcal{H} \}$ where $\{ f, g \} \equiv \partial_a f \omega^{ab} \partial_b g$.
+Let $\rho_\mathcal{c}$ the distribution of material at all times.
+$\mathcal{H} \rho_\mathcal{c} = 0$ and $d_s \rho_\mathcal{c} = \{ \rho_\mathcal{c}, \mathcal{H} \} = 0$.
+\end{prop}
+
+\begin{justification}
+	TODO
+\end{justification}
+
+As we have derived relativistic motion in an unusual way, a few comments are in order.
+
+As mentioned at the beginning of the section, no new assumption was required.
+Relativistic Hamiltonian mechanics is simply the correct way of handling time dependent motion.
+As such, there was no mention of the speed of light $c$, $\mathcal{Q}$ being a Riemannian manifold or the metric $g$.
+These will be introduced in the next chapter with the kinematic equivalence assumption.
+In light of this, it is probably better to consider relativistic mechanics as a feature of the geometry of the extended phase space more than the geometry of space-time.
+
+We have also seen that anti-particle states, a feature generally associated with quantum field theory, is actually a feature of relativistic Hamiltonian mechanics.
+This is usually missed because in classical mechanics only the standard Hamiltonian is typically used.
+Note how the extended Hamiltonian in \ref{free_hamiltonians} is quadratic in $E$, which can have both positive and negative solutions on the constraint $\mathcal{H}=0$.
+
+Another important feature is that this formulation of relativistic Hamiltonian mechanics is formally very close to the Quantum case.
+Note how $\mathcal{H} \rho = 0$ and $\mathcal{H} \ket{\psi} = 0$ are formally equivalent.
+If we apply the extended Hamiltonian in \ref{free_hamiltonians} with the usual formal substitution, we have the Klein-Gordon equation using space-like convention rescaled by a factor for $2m$.
+As we'll see later, to this parallel in the mathematical formalism corresponds a parallel in the physics description.
 

### 2016-06-13 2015CompleteDerivation/completeDerivation.tex
Adding comments after Dave and Christine review

@@
 
-The notion that the system is homogeneous means that the states of the composite and of the parts are not unrelated: they are made of the same material.\footnote{Whether a system is homogeneous depends on context (e.g.
+The notion that the system is homogeneous means that the states of the composite and of each part are not unrelated: they are all made of the same material.\footnote{Whether a system is homogeneous depends on context (e.g.
 air can be thought as homogeneous if the mixture of gases does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is maintained by time evolution.} In fact, the state space $\mathcal{C}$ of all systems composed of such material will include the state of the system as well as the states of  its parts (i.e.
@@
 Note that such state difference may not describe a physical state: it may remove some material from one location to add it somewhere else.
-Yet, these state changes are still physically distinguishable objects that provide a configuration for the same homogeneous material, therefore we extend the state space $\mathcal{C}$ to include state differences.
-All combined, this gives us the structure of an abelian group.
+Yet, these state changes are still physically distinguishable objects that provide a configuration for the same homogeneous material, therefore we extend the state space $\mathcal{C}$ to include state differences.\footnote{Note that one can characterize a state difference without knowing the states themselves.
+If we take some amount of material out of a container we know how the state changed, yet we may not know the amount of material before or after.
+In certain cases one may be interesting in measuring the asymmetry of a particular state without measuring (or without being able to measure) the full state itself.} All combined, this gives us the structure of an abelian group.
 
@@
 This in general will depend on how much the state of the composite system "knows" about its parts.
-For example, the position and orientation of an ideal rigid body is enough to define where all its constituents are, while the volume/pressure/temperature of an ideal gas is not enough to determine the position and momentum of all its molecules.
+For example, the position and orientation of an ideal rigid body is enough to define where all its constituents are, while the volume/pressure/temperature of an ideal gas is not enough to determine the position and momentum of all its constituents.
 Therefore we need to characterize the system further.
@@
 
-We will call a \emph{classical system} one that is infinitesimally reducible.
+We will call a \emph{classical material} one that is infinitesimally reducible.
 That is, we can keep decomposing the system indefinitely into smaller and smaller parts, each with a well defined state.
@@
 The abelian group $\mathcal{C}$ can be extended with the operations defined by $T$, as each element $\tau \in T$ is a map $\tau : \mathcal{C} \rightarrow \mathcal{C}$.
-The map has the following properties: $(\tau_1 + \tau_2) \mathcal{c} = \tau_1 \mathcal{c} + \tau_2 \mathcal{c}$, increasing the amount of material by the sum of two constant factors is the same as combining the separate increases, and $\tau (\mathcal{c}_1 + \mathcal{c}_2) = \tau \mathcal{c}_1 + \tau \mathcal{c}_2$, increasing the amount of the total system is the same as the combination of the increased parts.
+The map has the following properties: $(\tau_1 + \tau_2) \, \mathcal{c} = \tau_1 \mathcal{c} + \tau_2 \mathcal{c} \; \forall \tau_1, \tau_2 \in T$ and $\mathcal{c} \in \mathcal{C}$, increasing the amount of material by the sum of two constant factors is the same as combining the separate increases, and $\tau \, (\mathcal{c}_1 + \mathcal{c}_2) = \tau \mathcal{c}_1 + \tau \mathcal{c}_2\; \forall \tau \in T$ and $\mathcal{c}_1, \mathcal{c}_2 \in \mathcal{C}$, increasing the amount of the total system is the same as the combination of the increased parts.
 $\mathcal{C}$ is a module over $T$, which is a field and isomorphic to $\mathbbm{R}$.

### 2016-06-13 2015CompleteDerivation/completeDerivation.tex
Adding justifications for relativistic Hamiltonian

@@
 
-%Status: conceptual work is done, need to write down
+% Review started
 
@@
 \begin{justification}
-	TODO
-\end{justification}
+	We claim $\mathcal{T}_{\Delta s}$ is a self-homeomorphism on $T^*\mathcal{M}$.
+The extended state space for particles of a homogeneous classical material is $T^*\mathcal{M}$ by \ref{classical_extended_phase_space}.
+$\mathcal{T}_{\Delta s}$ is a deterministic and reversible evolution map and by \ref{detrevmap} is a self-homeomorphism.
+
+	We claim $\mathcal{T}_{\Delta s}$ is a symplectomorphism on $(T^*\mathcal{M}, \omega)$.
+The distribution on final states must be defined.
+The Jacobian for $\mathcal{T}_{\Delta s}$ exists and is non-zero.
+$\mathcal{T}_{\Delta s}$ is a diffeomorphism.
+A deterministic and reversible process conserves the number of states and possibilities.
+$\omega$ is the metric for the cardinality of possibilities.
+$\omega$ is invariant under a deterministic and reversible evolution map.
+$\mathcal{T}_{\Delta s}$ is a symplectomorphism by definition.\end{justification}
 
@@
 \begin{justification}
-	TODO
+We claim the vector field $S \in T(T^*\mathcal{M})$ for the infinitesimal displacement $S^a = d\xi^a/dt$ admits a potential $\mathcal{H}$ such that $S^{a} \omega_{ab} = \partial_{b}\mathcal{H}$.
+The extended state space $T^*\mathcal{M}$ of a classical homogeneous material is a symplectic manifold by \ref{classical_extended_phase_space}.
+The map for infinitesimal time evolution $\mathcal{T}_{dt}$ is an infinitesimal self-symplectomorphism by \ref{relativistic_canonical_transformation}.
+By \ref{symplectomorphism_generator} the infinitesimal displacement $S$ admits a potential $\mathcal{H}$ such that $S^{a} \omega_{ab} = \partial_{b}\mathcal{H}$
+
+We claim the state variables evolve according to extended Hamilton's equations.
+$S^{a} \omega_{ab} = d_s\xi^a \omega_{ab} = \partial_{b}\mathcal{H}$.
+For $a = 0$ we have $d_s t \, (-1) = \partial_{E} \mathcal{H}$.
+For  $a=\{1,...,n\}$ we have $d_s q^i \, (+1) = \partial_{p^i} \mathcal{H}$.
+For $a=n+1$ we have $d_s E \, (+1) = \partial_{t} \mathcal{H}$.
+For $a=\{n+2,...,2n + 1\}$ we have $d_s p_i \, (-1) = \partial_{q^i} \mathcal{H}$.
 \end{justification}
@@
 \begin{justification}
-	TODO
+	We claim $t(s)$ is strictly monotonic.
+The motion is deterministic and reversible.
+At each value of $t$ there must be only one possible state: if there is more then one state the motion is non-deterministic (one state is not enough to identify them), if there is no state the motion is non-reversible (no state cannot identify a state).
+$t(s)$ is invertible.
+$t(s)$ is strictly monotonic.
+
+	We claim $d_{s}t$ partitions the space.
+Let $\mathcal{s}_1, \mathcal{s}_2 \in T^*\mathcal{M}$ two physical states connected by deterministic and reversible evolution.
+$\mathrm{sign}(d_{s}t(\mathcal{s}_1)) = \mathrm{sign}(d_{s}t(\mathcal{s}_2))$.
+Let $U \equiv \{\mathcal{s} \in T^*\mathcal{M} \; | \; d_{s}t(\mathcal{s}) > 0 \}$ and $V \equiv \{\mathcal{s} \in T^*\mathcal{M} \; | \; d_{s}t(\mathcal{s}) < 0 \}$.
+$U \cap V = \varnothing$.
+Let $\gamma : [s_0, s_1] \rightarrow T^*\mathcal{M}$ the trajectory given by deterministic and reversible evolution.
+Either $\gamma([s_0, s_1]) \subseteq U$ or $\gamma([s_0, s_1]) \subseteq V$.
 \end{justification}
@@
 	The form of the extended Hamiltonian is $\mathcal{H} = \lambda(t,E,q^i,p_i) \, (H(t,q^i,p_i) - E)$ where $H(t,q^i,p_i)$ is the standard Hamiltonian at each instant $t$, the conjugate time variable $E = H(t,q^i,p_i)$ represents the value of the Hamiltonian, and  $\lambda : T^*\mathcal{M} \rightarrow \mathbbm{R}$ where $\lambda(\xi^a) = d_s t$.
-\end{prop}
-
-\begin{justification}
-	TODO
+$\mathcal{H} = 0$ for all states at all time.
+\end{prop}
+
+\begin{justification}
+	We claim the form of the extended Hamiltonian is $\mathcal{H} = \lambda(t,E,q^i,p_i) \, (H(t,q^i,p_i) - E)$.
+Let $\mathcal{H}$ the extended Hamiltonian and $H$ be Hamiltonian for a particular choice of time and state variables.
+We have $d_s q^i = \partial_{p_i} \mathcal{H} = d_s t \, d_t q^i = - \lambda \, \partial_{p_i} H$ where $\lambda : T^*\mathcal{M} \rightarrow \mathbbm{R} \; | \; \lambda \mapsto d_s t$.
+$d_s p_i = - \partial_{q^i} \mathcal{H} = d_s t \, d_t p_i = \lambda \, \partial_{q^i} H$.
+As $\lambda \neq 0$ for all physical states, we can set $\mathcal{H} = \lambda(t,E,q^i,p_i) \, (H(t,q^i,p_i) - E + \mathcal{h}(t,E,q^i,p_i))$ without loss of generality.
+We have $\partial_{q^i} \lambda (H - E + \mathcal{h}) + \lambda (\partial_{q^i} H + \partial_{q^i} \mathcal{h}) = \lambda \, \partial_{q^i} H$.
+This holds for all choices of $t$.
+In particular for $t=s$, $\lambda (\partial_{q^i} H + \partial_{q^i} \mathcal{h}) = \lambda \, \partial_{q^i} H$.
+$\partial_{q^i} \mathcal{h} = 0$.
+$\mathcal{h}$ is not a function of $q^i$.
+Repeat the same procedure for $p_i$.
+$\mathcal{h}$ is not a function of $p_i$.
+$\lambda = d_s t = - \partial_E \mathcal{H} = - \partial_{E} \lambda (H - E + \mathcal{h}) - \lambda (-1 + \partial_{E} \mathcal{h})$.
+In particular for $t=s$, $\lambda = - \lambda (-1 + \partial_{E} \mathcal{h})$.
+$\partial_{E} \mathcal{h} = 0$.
+$\mathcal{h}$ is not a function of $E$.
+$\mathcal{h}$ is only a function of $t$.
+We can redefine $H = H + \mathcal{h}$ as the result is still a function of $q^i, p_i , t$ with the same equations of motion.
+
+	We claim $E = H(t,q^i,p_i)$ for physical states. $\lambda = d_s t = - \partial_E \mathcal{H} = - \partial_{E} \lambda (H - E + \mathcal{h}) - \lambda (-1)$ for any choice of $t$. $H - E = 0$.
+
+	We claim $\mathcal{H} = 0$ for physical states. $\mathcal{H} = \lambda (H - E) = 0$.
 \end{justification}
@@
 \begin{justification}
-	TODO
+	We claim $\mathcal{H} \rho_\mathcal{c}(s) = 0 \; \forall \mathcal{c} \in \mathcal{C} \wedge s \in \mathbbm{R}$.
+Let $\rho_\mathcal{c}(s) : T^*\mathcal{M} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c}$ at a given value of $s$.
+The distribution is defined on physical states.
+By \ref{form_of_extended_hamiltonian} $supp(\rho_\mathcal{c}(s)) \subseteq T^*\mathcal{M} \,\backslash\, supp(\mathcal{H})$.
+$supp(\mathcal{H} \rho_\mathcal{c}(s)) = supp(\mathcal{H}) \cap supp(\rho_\mathcal{c}(s)) \subseteq \varnothing$.
+
+	We claim $d_s \rho_\mathcal{c} (s) = \{ \rho_\mathcal{c}(s), \mathcal{H} \}$.
+Let $\rho_\mathcal{c} (s)$ be differentiable.
+$d_s \rho_\mathcal{c}(s) = \partial_a \rho_\mathcal{c}(s) d_s \xi^a = \partial_a \rho_\mathcal{c}(s) \omega^{ab} \partial_b \mathcal{H} = \{ \rho_\mathcal{c}(s), \mathcal{H} \}$.
+
+	We claim $\mathcal{H} \rho_\mathcal{c} = 0 \; \forall \mathcal{c}$.
+Let $\rho_\mathcal{c} : T^*\mathcal{M} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c}$ at all states at all times.
+The distribution is defined on physical states.
+By \ref{form_of_extended_hamiltonian} $supp(\rho_\mathcal{c}) \subseteq T^*\mathcal{M} \,\backslash\, supp(\mathcal{H})$.
+$supp(\mathcal{H} \rho_\mathcal{c}) = supp(\mathcal{H}) \cap supp(\rho_\mathcal{c}) \subseteq \varnothing$.
+
+	We claim $d_s \rho_\mathcal{c} = \{ \rho_\mathcal{c}(s), \mathcal{H} \} = 0$.
+Let $\rho_\mathcal{c}$ be differentiable.
+$d_s \rho_\mathcal{c}= \partial_a \rho_\mathcal{c} d_s \xi^a = \partial_a \rho_\mathcal{c}(s) \omega^{ab} \partial_b \mathcal{H} = \{ \rho_\mathcal{c}, \mathcal{H} \}$.
+Let $\gamma : \mathbbm{R} \rightarrow T^*\mathcal{M}$ a particle trajectory.
+Then $\rho_\mathcal{c}(\lambda(s)) = \rho_\mathcal{c}(\lambda(s + \Delta s))$.
+$d_s \rho_\mathcal{c} = 0$.
+
 \end{justification}

### 2016-06-15 2015CompleteDerivation/completeDerivation.tex
Another pass at invariant densitities

@@
 
-In the case where the topology of $\mathcal{S}$ is not discrete, the distribution $\rho_\mathcal{c}$ is a density, which is not something we directly measure.
+When the topology of $\mathcal{S}$ is not discrete, the distribution $\rho_\mathcal{c}$ is a density, which is not something we directly measure.
 What we do measure experimentally are finite amounts of material.
@@
 And the density will give us an indication of how big the region is: if we double the region, the density will halve.
-In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique measure $\mu$ for the size of sets of $\mathcal{S}$.
+In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique measure\footnote{Here measure is intended in the mathematical sense (a real valued function of a $\sigma$-algebra) which is distinct from other connotations in physics or probability theory.} $\mu$ for the size of sets of $\mathcal{S}$.
 From the measure we get a metric $d\mu$ and our positive linear functionals become integrals, $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
@@
 	We claim there exists a positive linear functional $\Lambda_U : \mathcal{C} \rightarrow \mathbbm{R}$ for each $U \subseteq \mathcal{S}$ such that $\Lambda_U = \Lambda_{\mathrm{int}(U)}$.
-Let $\Lambda_{U} : \mathcal{C} \rightarrow \mathbbm{R}$ be the functional that returns the amount of material in $\mathcal{c}$ compatible with the outcome associated with the open set $U \subseteq \mathcal{S}$.
+Let $\Lambda_{U} : \mathcal{C} \rightarrow \mathbbm{R}$ be the functional that returns the amount of material in $\mathcal{c}$ compatible with the outcome associated with the open set $U \subseteq \mathcal{S}$.$\Lambda_U$ is well-defined.
+$U$ is associated with a physically distinguishable outcome.
 $\Lambda_U$ is linear.
@@
 Let $U \subseteq \mathcal{S}$ not necessarily open.
-The same argument applies to $\mathrm{int}(U)$.
+$|\Lambda_{U}(\mathcal{c})| = |\Lambda_{\mathrm{int}(U)}(\mathcal{c})| < \infty$.
 
@@
 Let $U_1, U_2 \in \mathcal{S}$ not necessarily open.
-The same argument applies to their interiors.
+$\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{\mathrm{int}(U_1)} + \Lambda_{\mathrm{int}(U_2)} = \Lambda_{\mathrm{int}(U_1) \cup \mathrm{int}(U_2)} + \Lambda_{\mathrm{int}(U_1) \cap \mathrm{int}(U_2)} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2}$
 
@@
 	We claim $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$.
-As per \ref{classical_vector space}, $\mathcal{C}$ is isomorphic to a subspace $G \subseteq C(\mathcal{S})$ as a vector space over $\mathbbm{R}$.
+As per \ref{classical_vector space}, $\mathcal{C} \cong G \subseteq C(\mathcal{S})$ as a vector space over $\mathbbm{R}$.
 As $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$, then $G \subseteq \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\} = L^1(\mathcal{S}, \mu)$ by \ref{everywhere_integrable_is_lebesgue_integrable}.
@@
 The set of continuous functions with compact support $C_C(\mathcal{S})$ would exclude distributions, such as Gaussians, that span over the whole range of states.
-If we take a finite volume of an ideal gas at equilibrium, the momentum distribution spans over the whole range.
+Such distributions are physical: if we take a finite volume of an ideal gas at equilibrium, the momentum distribution spans over the whole range.
 Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a general physical justification for.\footnote{One can, though, make the argument that given any distribution $\rho$ one can find an infinitely smooth $\rho_{sm}$ such that the difference in description is small compared to the error already introduced by assuming the system to be homogeneous and infinitely reducible.
@@
 
-Another point is that while the norm associated to $L^1(\mathcal{S})$ is $\int_{\mathcal{S}} |\rho| d\mu$, the pseudo-norm $\int_{\mathcal{S}} \rho d\mu$ is perhaps more physically meaningful.
+Another consideration is that while the norm associated to $L^1(\mathcal{S})$ is $\int_{\mathcal{S}} |\rho| d\mu$, the pseudo-norm $\int_{\mathcal{S}} \rho d\mu$ is perhaps more physically meaningful.
 For proper states, the two are the same and they represent the total amount of material.
@@
 being added and being removed), therefore $\int_{\mathcal{S}} | \delta \rho | \, d\mu = 0$ means no material is moving, no change is occurring.
+
+Whatever norm one may choose, $\mathcal{C}$ is a normed vector space but not a complete one.
+It does not include the limit of all Cauchy sequences as discontinuous functions are excluded.
+Therefore $\mathcal{C}$ cannot be a Banach or a Hilbert space.
+As we mentioned before, we can extend the space for computation purposes or to study a limit case as long as we are mindful of the extension and carefully check that the mathematical results do not depend on such extension and are physically meaningful.
 
@@
 $\mathcal{S}$ has a differentiable structure in the sense that the distributions are only defined on a set of set state variables that are linked by differential transformations.
-In the same way that discontinuous changes of state variable can be detected because they are incompatible with the topology, non differential changes of state variable can be detected because they are incompatible with state-variable-independent densities.
+In the same way that discontinuous changes of state variable can be detected because they are incompatible with the topology, non differentiable changes of state variable can be detected because they are incompatible with state-variable-independent densities.
 
@@
 This means that we cannot change physical units of all the variables as we please: they are part of a unit system.
-For example, changing units of position will also change units of momentum.
+As we know, changing units of position will also change units of momentum.
 Let $\mathcal{Q}$ be the manifold charted by the minimum set of state variables that define such system of units.
@@
 They can, however, be covectors $k(q) dq$ as $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$.\footnote{The components $k$ are the classical analogue of the wave numbers.} In other words: a covector is the simplest local geometrical object that allows coordinate independent densities.
+Physically, the component $k_i$ of a covector, for example the wave number, can be expressed with a unit that is the inverse of the corresponding $q^i$.
+The product $\Delta q^i \Delta k_i$ is therefore a dimensionless quantity, invariant under unit transformation.
 We can identify $\mathcal{S}$ as the cotangent bundle $T^*\mathcal{Q}$, the space of all possible covectors.
@@
 	We claim $\mathcal{S}$ is a differential manifold.
-A composite state $\mathcal{c}$ is a distribution $\rho_\mathcal{s} : \mathcal{S} \rightarrow \mathbbm{R}$.
+Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distributions associated with a composite state $\mathcal{c} \in \mathcal{C}$.
 $\rho_\mathcal{c}$ is defined on any chart of $\mathcal{S}$.
@@
 Let $\mathcal{Q}$ be the manifold charted by $q$.
-$\mathcal{Q} \subseteq \mathcal{S}$.
+$\mathcal{Q}$ is an embedded manifold of $\mathcal{S}$.
 $\mathcal{Q}$ is differentiable.
-Let's assume $q$ is enough to define the unit system for the description of the physical object.
+Let $q$ fully define the unit system for the description of the physical object.
 Then the physical object is represented by a geometrical object on $\mathcal{Q}$.
@@
 	We claim the state space $\mathcal{S}$ for the particles of a homogeneous classical material is a symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of an $n$-dimensional differential manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
-Let $q^i$ be a set of n continuous independent state-variable that define the units to describe a state in $U \subseteq \mathcal{S}$.
+Let $q^i$ be a set of n continuous independent state-variable that define the units necessary to describe a state in $U \subseteq \mathcal{S}$.
 Let $\mathcal{Q}$ be the manifold charted by $q^i$.
@@
 $|\int_{S^+} \rho d\mu| = |\int_{S} \rho^+ d\mu| = \lVert \rho^+ \rVert < \infty$ and $|\int_{S^-} \rho d\mu| = |\int_{S} \rho^- d\mu| = \lVert \rho^- \rVert < \infty$.
-$\lVert \rho \rVert = \lVert \rho^+ \rVert_U + \lVert \rho^- \rVert_U < \infty$.
+$\lVert \rho \rVert = \lVert \rho^+ \rVert + \lVert \rho^- \rVert < \infty$.
 $\rho \in L^1(S,\mu)$.

### 2016-06-24 2015CompleteDerivation/completeDerivation.tex
Differentiable functions for classical mechanics

@@
 The state space $\mathcal{C}$ for a homogeneous classical (i.e.
-infinitesimally reducible) material is a vector space over $\mathbbm{R}$ isomorphic to a subspace of $C(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is continuous$\}$, where $\mathcal{S}$ is the state space of an infinitesimal amount of material.
+infinitesimally reducible) material is a vector space over $\mathbbm{R}$ isomorphic to a subspace of the space of continuous functions.
+That is $\mathcal{C} \cong G \subseteq C(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is continuous$\}$, where $\mathcal{S}$ is the state space of an infinitesimal amount of material.
 \end{prop}
@@
 
-The ability to associate finite amount of material with sets of states allows us to define finite regions of $\mathcal{S}$ and compare them.
+The ability to associate finite amounts of material with sets of states allows us to define finite regions of $\mathcal{S}$ and compare them.
 Consider a region of position and momentum in phase space.
@@
 Then $\mathcal{S}$ is endowed with a Borel measure $\mu$.
-The state space $\mathcal{C}$ is isomorphic to the space of Lebesgue integrable continuous functions over $\mathcal{S}$.
-That is: $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ where $L^1(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \; \int_{\mathcal{S}} |\rho| d\mu < \infty \}$.
+The state space $\mathcal{C}$ for a homogeneous classical material is isomorphic to a subspace of the space of Lebesgue integrable functions.
+That is $\mathcal{C} \cong G \subseteq L^1(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \; \int_{\mathcal{S}} |\rho| d\mu < \infty \}$.
 \end{prop}
@@
 	We claim there exists a positive linear functional $\Lambda_U : \mathcal{C} \rightarrow \mathbbm{R}$ for each $U \subseteq \mathcal{S}$ such that $\Lambda_U = \Lambda_{\mathrm{int}(U)}$.
-Let $\Lambda_{U} : \mathcal{C} \rightarrow \mathbbm{R}$ be the functional that returns the amount of material in $\mathcal{c}$ compatible with the outcome associated with the open set $U \subseteq \mathcal{S}$.$\Lambda_U$ is well-defined.
-$U$ is associated with a physically distinguishable outcome.
-$\Lambda_U$ is linear.
-$\Lambda_U(a_1 \mathcal{c}_1 + a_2 \mathcal{c}_2) = a_1 \Lambda_U(\mathcal{c}_1) + a_2 \Lambda_U(\mathcal{c}_2)$ as it has to be consistent with the operations of composing states and increasing/decreasing the amount of material by a factor.
-$\Lambda_U$ is positive.
-If the value of the distribution for each particle state is positive then the total amount of material is positive.
+Let $\Lambda_{U} : \mathcal{C} \rightarrow \mathbbm{R}$ be the functional that returns the amount of material in $\mathcal{c}$ compatible with the outcome associated with the open set $U \subseteq \mathcal{S}$.
+$\Lambda_U$ is well-defined: $U$ is associated with a physically distinguishable outcome.
+$\Lambda_U$ is linear: $\Lambda_U(a_1 \mathcal{c}_1 + a_2 \mathcal{c}_2) = a_1 \Lambda_U(\mathcal{c}_1) + a_2 \Lambda_U(\mathcal{c}_2)$ as it has to be consistent with the operations of composing states and increasing/decreasing the amount of material by a factor.
+$\Lambda_U$ is positive: if the value of the distribution for each particle state is positive then the total amount of material is positive.
 Let $U \subseteq \mathcal{S}$ not necessarily open.
@@
 
-	We claim $\mathcal{C} \cong C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$.
-As per \ref{classical_vector space}, $\mathcal{C} \cong G \subseteq C(\mathcal{S})$ as a vector space over $\mathbbm{R}$.
-As $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$, then $G \subseteq \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\} = L^1(\mathcal{S}, \mu)$ by \ref{everywhere_integrable_is_lebesgue_integrable}.
-$G \subseteq C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$.
-Let $\rho_\mathcal{c}$ be the distribution associated with the state $\mathcal{c}$.
+	We claim $\mathcal{C}$ is isomorphic to a subspace of $L^1(\mathcal{S}, \mu)$ as a vector space over $\mathbbm{R}$.
+$\forall \mathcal{c} \in \mathcal{C} \exists !
+\rho_{\mathcal{c}}:\mathcal{S} \rightarrow \mathbbm{R}$ returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$.
+As $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$, then $\rho_{\mathcal{c}} \in \mathcal{L}(S,\mu) \equiv \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\}$.
+$\mathcal{L}(S,\mu) = L^1(\mathcal{S}, \mu)$ by \ref{everywhere_integrable_is_lebesgue_integrable}.
+Let $\varrho : \mathcal{C} \rightarrow L^1(\mathcal{S}, \mu)$ such that $\varrho(\mathcal{c}) \mapsto \rho_\mathcal{c}$.
+$\varrho$ is a homomorphism as we saw in the proof for \ref{classical_vector space}.
+$\mathcal{C}$ is isomorphic to $\varrho(\mathcal{C}) \subseteq L^1(\mathcal{S}, \mu)$ as a vector space.
+\end{justification}
+
+On a state space with discrete topology and a single state variable we simply have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho_\mathcal{c}(q)$.
+For states identified by a single continuous state variable we would expect $\Lambda_U (\mathcal{c}) = \int_a^b \rho_\mathcal{c} (q) dq$.
+By a simple change of state variable, we see that this does not work in general: $\rho_\mathcal{c}(q')= \rho(q)_\mathcal{c} dq/dq'$.
+This would make $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using.
+This is inconsistent with our previous definition.
+Moreover if the transformation is not differential, the density is not even well-defined.
+This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows state-variable-independent densities.
+
+The first order of business is to guarantee that the density remains defined under an arbitrary change of state variables.
+This means the Jacobian of the transformation must exist.
+For the Jacobian to exist the transformation between state variables has to be differentiable.
+Therefore a manifold that guarantees densities to be well defined is one that guarantees that state variable transformations are differentiable: a differentiable manifold.
+$\mathcal{S}$ has a differentiable structure in the sense that the distributions and the volume element $d\mu$ are only defined on a set of state variables that are linked by differential transformations.
+In the same way that discontinuous changes of state variable can be detected because they are incompatible with the topology, non differentiable changes of state variable can be detected because they are incompatible with state-variable-independent densities.
+
+Moreover, the density itself has to be differentiable.
+The distribution $\rho_\mathcal{c}$ is a real value function of the state.
+As such, at least locally, we can use it as a state variable $q^1=\rho_\mathcal{c}$.
+As a state variable it identifies the states by the density of the material associated with them.
+This is more physically meaningful as one may first suspect, since placing physical markers is a common way to define the references for a coordinate system, which in the end is placing material at particular states.
+In a coordinate system where $q^1=\rho_\mathcal{c}$, the function is clearly differential as $\partial_1 \rho_\mathcal{c} = 1$ and $\partial_i \rho_\mathcal{c} = 0$ for $i \neq 1$.
+As we change state variables, the transformation is differentiable, therefore all the derivatives $\partial_{i'} q^1 = \partial_{i'} \rho_\mathcal{c}$ exist.
+Therefore the density itself is differentiable.
+
+\begin{prop}\label{differentiable_manifold}
+	The state space $\mathcal{S}$ for the particles of a homogeneous classical material identified by a finite number of continuous independent state variables is a differentiable manifold.
+The state space $\mathcal{C}$ for a homogeneous classical material is isomorphic to the space of Lebesgue integrable differentiable functions.
+That is $\mathcal{C} \cong C^1(\mathcal{S}) \cap L^1(\mathcal{S}, \mu)$, where $C^1(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is differentiable$\}$.
+\end{prop}
+\begin{justification}
+	TODO
+%Let $\rho_\mathcal{c}$ be the distribution associated with the state $\mathcal{c}$.
 Let $\rho \in C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ be a Lebesgue integrable continuous function.
@@
 
-As we identified the set of Lebesque integrable continuous functions $C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as the set of distributions that are physically meaningful, we can better understand why other commonly used function spaces do not fit the bill.
+As we identified the space of Lebesque integrable differentiable functions $C^1(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as the space of distributions that are physically meaningful, we can better understand why other commonly used function spaces do not fit the bill.
 Some are not restrictive enough.
 The set of Lebesgue integrable functions $L^1(\mathcal{S})$ includes discontinuous functions that for \ref{continuous_map} are unphysical.
-The set of continuous functions $C(\mathcal{S})$ and the set of continuous functions that vanish at infinity $C_0(\mathcal{S})$ include functions whose integral is infinite, which would represent infinite amount of material.
-Some are too restrictive.
-The set of continuous functions with compact support $C_C(\mathcal{S})$ would exclude distributions, such as Gaussians, that span over the whole range of states.
+The set of continuous functions $C(\mathcal{S})$ and the set of continuous functions that vanish at infinity $C_0(\mathcal{S})$ not only include non-differentiable functions, but also include functions whose integral is infinite, which would represent infinite amount of material.
+Such functions are also present in the space of differentiable functions $C^1(\mathcal{S})$.
+Some definitions are too restrictive.
+Requiring compact support (i.e.the function is different from zero only on a finite region) would exclude distributions, such as Gaussians, that span over the whole range of states.
 Such distributions are physical: if we take a finite volume of an ideal gas at equilibrium, the momentum distribution spans over the whole range.
@@
 
-We now turn our attention to the metric $d\mu$ which we need to properly identify. It turns out that before we can do so, we need to better characterize $\mathcal{S}$.
-
-On a state space with discrete topology and a single state variable we simply have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho_\mathcal{c}(q)$.
-For states identified by a single continuous state variable we would expect $\Lambda_U (\mathcal{c}) = \int_a^b \rho_\mathcal{c} (q) dq$.
-By a simple change of state variable, we see that this does not work in general: $\rho_\mathcal{c}(q')= \rho(q)_\mathcal{c} dq/dq'$.
-This would make $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using.
-This is inconsistent with our previous definition.
-Moreover if the transformation is not differential, the density is not even well-defined.
-This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows state-variable-independent densities.
-
-Under an arbitrary change of state variables, the density has to remain the same, which means that the Jacobian of the transformation must exist and must be unitary.
-For the Jacobian to exist the transformation between state variables has to be differentiable.
-Therefore a manifold that guarantees densities to be well defined is one that guarantees that state variable transformations are differentiable: a differentiable manifold.
-$\mathcal{S}$ has a differentiable structure in the sense that the distributions are only defined on a set of set state variables that are linked by differential transformations.
-In the same way that discontinuous changes of state variable can be detected because they are incompatible with the topology, non differentiable changes of state variable can be detected because they are incompatible with state-variable-independent densities.
-
-If $\rho_\mathcal{c}$ is to remain unchanged and the Jacobian to be unitary we cannot simply change one state variable as we please.
+Now that we are guaranteed that the density is properly defined for all state variables, we have to make sure that is invariant.
+The distribution should only depend on the states, and not the particular choice of state variables.
+
+If $\rho_\mathcal{c}$ is to remain invariant under state variable changes, not only the Jacobian has to exist but must be unitary.
+This means we cannot simply change one state variable as we please.
 If we change one at least another has to change in some coordinated way such that the Jacobian of the total transformation is unitary.
-This means that we cannot change physical units of all the variables as we please: they are part of a unit system.
+This means that we cannot change physical units of all the variables as we like: they are part of a unit system.
 As we know, changing units of position will also change units of momentum.
@@
 
-We should note that we are starting to be in a position similar to the one discussed in \ref{discrete_state_metric} for discrete states.
+It should be clear by now that the case of discrete topology is qualitatively different from the standard topology for $\mathbbm{R}^n$.
+The notion that the continuous case is a limit of the discrete case, that it's ``like the discrete but with more points", leads in this case to erroneous intuition.
+The question is: can we identify a single state?
+Can we have an outcome associated with only one element?
+The answer is yes in the discrete topology and is no in the continuous case.
+The consequence is that when we define the measure $\mu$ for the cardinality of continuous states, we assign a finite value to a finite region, and we assign zero to a single state.
+That is what gave us integration, densities and, ultimately, conjugate variables.
+If we were to use the discrete topology on $\mathbbm{R}^n$, if we were to assume we can identify single states, then we would assign measure one to each state and infinity to a finite region.
+We would not have integration, just a simple sum.
+Our distribution would not be a density and there would be no justification for conjugate variables.
+A finite distribution could only distribute finite amount of material in a finite number of states.
+
+In other words: this is not a case were the difference at the small scale has a small difference in the large scale.
+The two cases are radically different.
+We can tell them apart.
+We should take the use of densities and conjugate quantities as empirical evidence that the quantities like space and time are not discrete.
+Or more precisely: that the processes we use to distinguish those quantities cannot identify single instances, as that is all we can say experimentally.
+The quantum case does not change this, as the use of densities (in the form of the wave function) and conjugate quantities are even more prominent.
+
+Finally, we should note that we in a position similar to the one discussed in \ref{discrete_state_metric} for discrete states.
 We have both a topological space and a metric that allows us to count states and possibilities.

### 2016-06-24 2015CompleteDerivation/completeDerivation.tex
Adding Christine's comments on relativistic mechanics

@@
 \begin{prop}\label{relativistic_canonical_transformation}
-	A time-dependent deterministic and reversible evolution map for a homogeneous classical material is self-symplectomorphism on the extended phase space.
-That is: $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ and $\mathcal{T}_{\Delta s}^*\omega = \omega$ where $\mathcal{T}_{\Delta s}^*$ is the pullback of $\mathcal{T}_{\Delta s}$.
+	A time-dependent deterministic and reversible evolution map for a homogeneous classical material is a self-symplectomorphism on the extended phase space.
+That is: $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ such that $\mathcal{T}_{\Delta s}^*\omega = \omega$ where $\mathcal{T}_{\Delta s}^*$ is the pullback of $\mathcal{T}_{\Delta s}$.
 \end{prop}
@@
 \begin{justification}
-We claim the vector field $S \in T(T^*\mathcal{M})$ for the infinitesimal displacement $S^a = d\xi^a/dt$ admits a potential $\mathcal{H}$ such that $S^{a} \omega_{ab} = \partial_{b}\mathcal{H}$.
+We claim the vector field $S \in T(T^*\mathcal{M})$ for the infinitesimal displacement $S^a = d\xi^a/ds$ admits a potential $\mathcal{H}$ such that $S^{a} \omega_{ab} = \partial_{b}\mathcal{H}$.
 The extended state space $T^*\mathcal{M}$ of a classical homogeneous material is a symplectic manifold by \ref{classical_extended_phase_space}.
-The map for infinitesimal time evolution $\mathcal{T}_{dt}$ is an infinitesimal self-symplectomorphism by \ref{relativistic_canonical_transformation}.
+The map for infinitesimal evolution $\mathcal{T}_{ds}$ is an infinitesimal self-symplectomorphism by \ref{relativistic_canonical_transformation}.
 By \ref{symplectomorphism_generator} the infinitesimal displacement $S$ admits a potential $\mathcal{H}$ such that $S^{a} \omega_{ab} = \partial_{b}\mathcal{H}$
 
-We claim the state variables evolve according to extended Hamilton's equations.
+We claim the state variables evolve according to the extended Hamilton's equations.
 $S^{a} \omega_{ab} = d_s\xi^a \omega_{ab} = \partial_{b}\mathcal{H}$.
@@
 Since the evolution is deterministic and reversible, for each value of $s$ we need to have one and only one value of $t$.
-$t(s)$ is invertible, strictly monotonic and $d_{s}t$ along a trajectory cannot change sign.
+Therefore $t(s)$ is invertible, strictly monotonic and $d_{s}t$ along a trajectory cannot change sign.
 This means there are two classes of states: the ones where the parametrization $s$ is aligned with time $t$, which we call standard states, and the ones where the parametrization $s$ is anti-aligned with time $t$, which we call anti-states.
@@
 The motion is deterministic and reversible.
-At each value of $t$ there must be only one possible state: if there is more then one state the motion is non-deterministic (one state is not enough to identify them), if there is no state the motion is non-reversible (no state cannot identify a state).
+At each value of $t$ there must be only one possible state: if there is more than one state the motion is non-deterministic (one state is not enough to identify more than one state), if there is no state the motion is non-reversible (no state cannot identify a state).
 $t(s)$ is invertible.
@@
 $U \cap V = \varnothing$.
-Let $\gamma : [s_0, s_1] \rightarrow T^*\mathcal{M}$ the trajectory given by deterministic and reversible evolution.
+Let $\gamma : [s_0, s_1] \rightarrow T^*\mathcal{M}$ the trajectory given by deterministic and reversible evolution, where $[s_0, s_1]$ is the range of the parametrization.
 Either $\gamma([s_0, s_1]) \subseteq U$ or $\gamma([s_0, s_1]) \subseteq V$.
@@
 
-With the extended Hamiltonian we are able to write the equation of motions for any choice of time and state variables.
+With the extended Hamiltonian we are able to write the equations of motion for any choice of time and state variables.
 For example: $d_t q^i = d_t s \, d_s q^i = d_s q^i / d_s t = - \partial_{p_i} \mathcal{H} / \partial_{E} \mathcal{H}$.
-These equations must be the same as the one given by a standard Hamiltonian in those coordinates.
+These equations must be the same as the ones given by a standard Hamiltonian in those coordinates.
 That is: $d_t q^i = \partial_{p_i} H = - \partial_{p_i} \mathcal{H} / \partial_{E} \mathcal{H}$.
@@
 
-Note that if $E = H$ then $\mathcal{H}=0$ for all states.
+Note that $\mathcal{H}=0$ for all states, since $\mathcal{H}=\lambda(H-E)$ and $E = H$.
 For all states we also have $\lambda \neq 0$.

### 2016-06-27 2015CompleteDerivation/completeDerivation.tex
Adding justification for differentiability

@@
 Let $\varrho : \mathcal{C} \rightarrow L^1(\mathcal{S}, \mu)$ such that $\varrho(\mathcal{c}) \mapsto \rho_\mathcal{c}$.
-$\varrho$ is a homomorphism as we saw in the proof for \ref{classical_vector space}.
+$\varrho$ is a homomorphism, as justified in \ref{classical_vector space}.
 $\mathcal{C}$ is isomorphic to $\varrho(\mathcal{C}) \subseteq L^1(\mathcal{S}, \mu)$ as a vector space.
@@
 Moreover, the density itself has to be differentiable.
-The distribution $\rho_\mathcal{c}$ is a real value function of the state.
+The distribution $\rho_\mathcal{c}$ is a real valued function of the state.
 As such, at least locally, we can use it as a state variable $q^1=\rho_\mathcal{c}$.
@@
 As we change state variables, the transformation is differentiable, therefore all the derivatives $\partial_{i'} q^1 = \partial_{i'} \rho_\mathcal{c}$ exist.
-Therefore the density itself is differentiable.
+The density itself is differentiable.
 
@@
 \begin{justification}
-	TODO
-%Let $\rho_\mathcal{c}$ be the distribution associated with the state $\mathcal{c}$.
+	We claim $\mathcal{S}$ is a differential manifold.
+Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c} \in \mathcal{C}$.
+$\rho_\mathcal{c}$ is defined on any chart of $\mathcal{S}$.
+Given two charts, for any $\rho_\mathcal{c}(q)$ there is a corresponding $\rho_\mathcal{c}(q')=\rho_\mathcal{c}(q^i) | d q' / d q |$.
+The Jacobian $| d q' / d q |$ exists and is non-zero.
+The map between any two charts is differentiable.
+$\mathcal{S}$ is a differentiable manifold.
+
+	We claim that the distributions associated with composite states are differentiable in the regions where they are strictly monotonic.
+Let $\rho_\mathcal{c}$ be the distribution associated with state $\mathcal{c}$.
+Let $\rho_\mathcal{c}$ be strictly monotonic in a region $U\subseteq \mathcal{S}$.
+Then its fibers in $U$, the inverse images of the values, are connected hyper-surfaces.
+We can construct a chart such that $q^1=\rho_\mathcal{c}$.
+The distribution is differentiable over $U$ for that variable: $\partial_{i} \rho_\mathcal{c} = \partial_{i} q^1 = \delta_i^1$ where $\delta_i^j$ is the Kronecker delta.
+Let $q^{j}$ be another set of state variables.
+$q^1(q^{j})$ is differentiable as the manifold is differentiable.
+$\rho_\mathcal{c}(q^{j})=q^1(q^{j})$ is differentiable over $U$.
+
+	We claim that the distributions associated with composite states are differentiable.
+Let $\rho_\mathcal{c}$ be the distribution associated with state $\mathcal{c}$.
+Let $\rho_{\mathcal{c}_0}$ be the distribution associated with state $\mathcal{c}_0$.
+Let $\rho_{\mathcal{c}_0}$ be strictly monotonic in a region $U\subseteq \mathcal{S}$.
+Let $\delta \rho = \rho_{\mathcal{c}_0} - k \rho_{\mathcal{c}}$ where $k \in \mathbbm{R}$ and $k>0$.
+Let $k$ be sufficiently small so that $\delta \rho$ is strictly monotonic over $U$.
+$\rho_{\mathcal{c}_0}$ and $\delta \rho$ are differentiable over $U$ as they are strictly monotonic over $U$.
+$\rho_{\mathcal{c}}$ is a linear combination of differentiable functions over any arbitrary region $U$.
+$\rho_{\mathcal{c}}$ is differentiable.
+
+	We claim $\mathcal{C} \cong C^1(\mathcal{S}) \cap L^1(\mathcal{S}, \mu)$.
+Let $\rho_\mathcal{c}$ be the distribution associated with a state $\mathcal{c}$.
 Let $\rho \in C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ be a Lebesgue integrable continuous function.
 If $\rho_\mathcal{c} \neq \rho$ then $\exists U \in \mathcal{S} \; | \; \int_{U} \rho_\mathcal{c} d \mu \neq \int_{U} \rho d \mu$.
-There exists an outcome in $\mathcal{S}$ that can physically distinguish the two distributions.
+Both integral are finite as $\rho_\mathcal{c}, \rho \in L^1(\mathcal{S}, \mu)$.
+Both integral are expressible with the same state variable as all elements are defined on the same differential structure.
+Therefore there exists an outcome for $\mathcal{S}$ that can physically distinguish the two distributions.
 There must be a state $\mathcal{c}_1 \in \mathcal{C} \; | \; \varrho(\mathcal{c}_1)=\rho$.
 $\varrho : \mathcal{C} \rightarrow C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ is surjective.
+$\varrho$ is a homomorphism, as justified in \ref{classical_vector space}.
 $\varrho$ is an isomorphism between $\mathcal{C}$ and $C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as vector spaces.
@@
 Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a general physical justification for.\footnote{One can, though, make the argument that given any distribution $\rho$ one can find an infinitely smooth $\rho_{sm}$ such that the difference in description is small compared to the error already introduced by assuming the system to be homogeneous and infinitely reducible.
-}
+The Whitney approximation theorem guarantees that this can be done mathematically.}
 
@@
 They cannot be a vectors $v(q) e_q$, as the density $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$ is not invariant.
-They can, however, be covectors $k(q) dq$ as $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$.\footnote{The components $k$ are the classical analogue of the wave numbers.} In other words: a covector is the simplest local geometrical object that allows coordinate independent densities.
-Physically, the component $k_i$ of a covector, for example the wave number, can be expressed with a unit that is the inverse of the corresponding $q^i$.
+They can, however, be covectors $k(q) dq$ as $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$.
+Physically, the component $k_i$ of our covector, which are the classical analogue of the wave number, can be expressed with a unit that is the inverse of the corresponding $q^i$.
+If $q^i$ is in meters, $k_i$ is in inverse meters.
 The product $\Delta q^i \Delta k_i$ is therefore a dimensionless quantity, invariant under unit transformation.
-We can identify $\mathcal{S}$ as the cotangent bundle $T^*\mathcal{Q}$, the space of all possible covectors.
+A covector is the simplest local geometrical object that allows coordinate independent densities.
+The cotangent bundle $T^*\mathcal{Q}$, the space of all possible covectors, represents the simplest mathematical structure $\mathcal{S}$ can have to be able to support coordinate invariant densities.
 
@@
 In SI units the relationship $\hbar dk = dp = m du$ between kinetic and conjugate momentum, derived later, sets the relationship between units.} The cardinality for states in a finite (i.e.
-compact) region $U \subseteq \mathcal{S}$ will be given by $\int_U \hbar dq \wedge dk$.
+compact) region $U \subseteq \mathcal{S}$ will be given by $\mu(U) = \int_U \hbar dq \wedge dk$.
 
@@
 We call such a pair a degree of freedom.
-As they are independent the total number of states is the product of the possibilities of each d.o.f.: $\hbar^n dq^n \wedge dk_n = \prod \limits_{i=1}^n \hbar dq^i \wedge dk_i$.
+As they are independent the total number of states is the product of the possibilities of each d.o.f.: $d\mu = \hbar^n dq^n \wedge dk_n = \prod \limits_{i=1}^n \hbar dq^i \wedge dk_i$.
 This is equivalent to saying that independent d.o.f.
@@
 \begin{justification}
-	We claim $\mathcal{S}$ is a differential manifold.
-Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distributions associated with a composite state $\mathcal{c} \in \mathcal{C}$.
-$\rho_\mathcal{c}$ is defined on any chart of $\mathcal{S}$.
-Given two charts, for any $\rho_\mathcal{c}(q)$ there is a corresponding $\rho_\mathcal{c}(q')=\rho_\mathcal{c}(q^i) | d q' / d q |$.
-The Jacobian $| d q' / d q |$ exists and is non-zero.
-The map between any two charts is differentiable.
-$\mathcal{S}$ is a differentiable manifold.
-
 	We claim the simplest state space $\mathcal{S}$ that allows state-variable-invariant densities is the symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of a one dimensional differential manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.

### 2016-07-20 2015CompleteDerivation/completeDerivation.tex
Final cotangent bundle proof and other improvements

@@
 
-\subsection{Homogeneous decomposable systems}
+\subsection{Homogeneous decomposable systems and vector spaces}
 The notion that a system is decomposable means the states are equipped with a rule of composition that allows one to write $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$: the composite system is the sum of its parts.
@@
 
-\subsection{Classical homogeneous decomposable systems}
+\subsection{Classical material and distributions}
 We also want to be able to express the state of the composite system in terms of the state of each part.
@@
 
-\subsection{Invariant densities}
+\subsection{Integration and measure}
 
@@
 
-On a state space with discrete topology and a single state variable we simply have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho_\mathcal{c}(q)$.
-For states identified by a single continuous state variable we would expect $\Lambda_U (\mathcal{c}) = \int_a^b \rho_\mathcal{c} (q) dq$.
-By a simple change of state variable, we see that this does not work in general: $\rho_\mathcal{c}(q')= \rho(q)_\mathcal{c} dq/dq'$.
-This would make $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using.
+\subsection{Invariant densities and differentiability}
+
+On a state space with discrete topology, using a single state variable, we have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho_\mathcal{c}(q)$.
+We would expect the expression to become $\Lambda_U (\mathcal{c}) = \int_a^b \rho_\mathcal{c} (q) dq$ for states identified by a continuous state variable.
+By changing state variable, though, we see that this does not work in general: $\rho_\mathcal{c}(q')= \rho(q)_\mathcal{c} dq/dq'$.
+This makes $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using.
 This is inconsistent with our previous definition.
@@
 $\mathcal{S}$ has a differentiable structure in the sense that the distributions and the volume element $d\mu$ are only defined on a set of state variables that are linked by differential transformations.
-In the same way that discontinuous changes of state variable can be detected because they are incompatible with the topology, non differentiable changes of state variable can be detected because they are incompatible with state-variable-independent densities.
+In the same way that discontinuous changes of state variable can be detected because they tamper with physical distinguishability (i.e.
+the topology), non differentiable changes of state variable can be detected because they tamper with our ability to define state-variable-independent densities.
 
@@
 \begin{prop}\label{differentiable_manifold}
-	The state space $\mathcal{S}$ for the particles of a homogeneous classical material identified by a finite number of continuous independent state variables is a differentiable manifold.
+	The state space $\mathcal{S}$ for the particles of a  classical material identified by a finite number of continuous independent state variables is a differentiable manifold.
 The state space $\mathcal{C}$ for a homogeneous classical material is isomorphic to the space of Lebesgue integrable differentiable functions.
@@
 
-Another consideration is that while the norm associated to $L^1(\mathcal{S})$ is $\int_{\mathcal{S}} |\rho| d\mu$, the pseudo-norm $\int_{\mathcal{S}} \rho d\mu$ is perhaps more physically meaningful.
+Another consideration is that while the norm associated to $L^1(\mathcal{S})$ is $\int_{\mathcal{S}} |\rho| d\mu$, the expression $\int_{\mathcal{S}} \rho d\mu$ is perhaps more physically meaningful.
 For proper states, the two are the same and they represent the total amount of material.
@@
 
-Whatever norm one may choose, $\mathcal{C}$ is a normed vector space but not a complete one.
-It does not include the limit of all Cauchy sequences as discontinuous functions are excluded.
-Therefore $\mathcal{C}$ cannot be a Banach or a Hilbert space.
-As we mentioned before, we can extend the space for computation purposes or to study a limit case as long as we are mindful of the extension and carefully check that the mathematical results do not depend on such extension and are physically meaningful.
+Note that neither of those expressions match the norm given by the vector product defined as $\langle f, g \rangle = \int_{\mathcal{S}} fg \, d\mu$.
+Moreover, since $\mathcal{C}$ excludes discontinuous functions, it can never be a complete normed space: it will not include the limit for all Cauchy sequences.
+Such constructions, though, are still useful.
+Consider the following expressions:
+\begin{align*}
+\Lambda_{\mathcal{S}} (\mathcal{c}) &= \int_\mathcal{S} \rho_{\mathcal{c}} d \mu = \int_\mathcal{S} 1\cdot  \rho_{\mathcal{c}} d \mu \\
+\Lambda_U (\mathcal{c}) &= \int_U \rho_{\mathcal{c}} d \mu = \int_\mathcal{S} 1_U \cdot \rho_{\mathcal{c}} d \mu \\
+p_{total}(\mathcal{c}) &= \int_\mathcal{S} p \cdot \rho_{\mathcal{c}} d \mu \\
+q_{avg}(\mathcal{c}) &= 1/|\rho_{\mathcal{c}}| \int_\mathcal{S} q \cdot \rho_{\mathcal{c}} d \mu
+\end{align*}
+For each state, they represent respectively the total amount of material, the material within $U$, the total momentum and the average of a state variable.
+As they are linear functionals of $\mathcal{C}$,  they can be expressed using their dual vector over $\mathcal{S}$.
+But these vectors ($1$, $1_U$, $p$, $q/|\rho_{\mathcal{c}}|$) are \emph{not} elements of $\mathcal{C}$.
+They do not represent states, they are not continuous and integrable, and there is no single general physical meaning for them.
+Therefore, as we mentioned before, while we can extend the function space for convenience, computation purposes or to study a limit case, we need to be mindful of the extension and carefully check that the mathematical results are physically meaningful case by case.
+
+\subsection{Invariant densities and cotangent bundle}
 
@@
 This  means that we cannot change physical units of all the variables as we like: they are part of a unit system.
-As we know, changing units of position will also change units of momentum.
-Let $\mathcal{Q}$ be the manifold charted by the minimum set of state variables that define such system of units.
-An element of $\mathcal{S}$ will be a geometrical object in such space.
-For example, it will be a point, a distance, two points, the ratio between two distances, the inverse of a distance: something that is expressible in the unit defined.
-As particle states represent configurations of infinitesimal parts, they will be local objects, defined around a point in $\mathcal{Q}$.
-As we saw before, they cannot be points themselves as the density $\rho(q')_\mathcal{c}= \rho(q)_\mathcal{c} dq/dq'$ is not invariant.
-They cannot be a vectors $v(q) e_q$, as the density $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$ is not invariant.
-They can, however, be covectors $k(q) dq$ as $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$.
-Physically, the component $k_i$ of our covector, which are the classical analogue of the wave number, can be expressed with a unit that is the inverse of the corresponding $q^i$.
-If $q^i$ is in meters, $k_i$ is in inverse meters.
-The product $\Delta q^i \Delta k_i$ is therefore a dimensionless quantity, invariant under unit transformation.
-A covector is the simplest local geometrical object that allows coordinate independent densities.
-The cotangent bundle $T^*\mathcal{Q}$, the space of all possible covectors, represents the simplest mathematical structure $\mathcal{S}$ can have to be able to support coordinate invariant densities.
-
-If $\mathcal{Q}$ is one dimensional, then the state is fully identified by a pair $(q, k)$.
-These variables are independent, so the total number of possibilities in a small region of $\mathcal{S}$ is proportional to $dq \wedge dk$.
-That is if we double the range of $dq$ or $dk$ then the number of possibilities doubles.
-We set $d\mu = \hbar dq \wedge dk$, where $\hbar$ is the proportionality constant that will depend on the unit chosen to count the possibilities of the pair $(q, k)$.
+
+The simplest case is when a single variable is enough to define our units, and therefore tell how all the others must change.
+Suppose that our state space $\mathcal{S}$ is identified by $n$ independent state variables $\{q, k^i\}$ and a change of units for the first, that is $\hat{q} = \hat{q}(q)$, determines the change for all others.
+The Jacobian matrix is the block matrix:
+\begin{align*}
+J =  \left[
+\begin{array}{cc}
+d_q\hat{q} & 0 \\
+\partial_q \hat{k}^j & \partial_i \hat{k}^j \\
+\end{array}
+\right] 
+\end{align*}
+The Jacobian has to be unitary, therefore all elements $\partial_i \hat{k}^j$ must be well defined and $|\partial_i \hat{k}^j| = 1/d\hat{q}$.
+This only happens if $\partial_i \hat{k}^j$ is a $1\times1$ matrix: there is only one $k$.
+This gives means $\partial_{k} \hat{k} = d_{\hat{q}} q$ and $\hat{k} = d_{\hat{q}} q \, k + f(q)$.
+As this is a change of units, we expect the zero value of $k$ to remain constant: $\hat{k}(q, 0) = 0 = f(q)$.
+Therefore $\hat{k} = d_{\hat{q}} q \, k$ transforms as the component of a covariant vector.
+$\mathcal{S}$ is isomorphic to $T^*\mathcal{Q}$ where $\mathcal{Q}$ is the manifold that defines our unit.
+
+Physically, this means that the state variable $k$, which is the classical analogue of the wave number, is expressed with the inverse of the unit used for $q$.
+If $q$ is in meters, $k$ is in inverse meters.
+Consider now the area $dq \wedge dk$ of an infinitesimal rectangular region.
+This quantity is dimensionless and therefore invariant.
+The total number of states will be proportional to it, as  doubling the range of $dq$ or $dk$ will give us double the number of possible states.
+We have $d\mu = \hbar dq \wedge dk$, where $\hbar$ is the proportionality constant that will depend on the unit chosen to count the possibilities of the pair $(q, k)$.
 That is, a unitary range of possibilities for $q$ and $k$ will give us $\hbar$ possibilities.\footnote{The actual value and physical dimensions of $\hbar$ are determined by the system of units, and should not be taken to describe some intrinsic physical property.
@@
 compact) region $U \subseteq \mathcal{S}$ will be given by $\mu(U) = \int_U \hbar dq \wedge dk$.
-
-If $\mathcal{Q}$ is $n$-dimensional, $d\mu$ is proportional to $dq^n \wedge dk_n$.
+We are finally able to write the volume of integration in terms for state variables.
+
+We can generalize to the case where $\mathcal{Q}$ has more than one dimension.
+As we must be able to change one unit at a time, independently from the other, to each $q^i$ will correspond a $k_i$ that uses the inverse of the corresponding units.
+We have $\mathcal{S}\cong T^*\mathcal{Q}$ and $d\mu = \hbar^n dq^n \wedge dk_n$.
 But it's not just the volume that is preserved as we change state variables.
-Each pair $(q^i, k_i)$ represents an independent set of state variables, with the cardinality of possibilities given by $\hbar dq^i \wedge dk_i$.
+Each pair $(q^i, k_i)$ is expressed by an independent unit with the cardinality of possibilities given by $\hbar dq^i \wedge dk_i$.
 We call such a pair a degree of freedom.
 As they are independent the total number of states is the product of the possibilities of each d.o.f.: $d\mu = \hbar^n dq^n \wedge dk_n = \prod \limits_{i=1}^n \hbar dq^i \wedge dk_i$.
-This is equivalent to saying that independent d.o.f.
+In other words: independent d.o.f.
 are orthogonal surfaces in $\mathcal{S}$.
+The possibilities of an arbitrary degree of freedom, then, will be the sum of the projections over the $n$ orthogonal and independent d.o.f.
+defined by $(q^i, k_i)$.
 
@@
 
-We can capture the above discussion by stating that the state space $\mathcal{S}$ is the symplectic manifold $(T^*\mathcal{Q}, \omega)$.
-This is the simplest manifold that allows us to define state-variable-invariant densities.
+We can capture the above discussion by stating that the state space $\mathcal{S}$ is the symplectic manifold $(T^*\mathcal{Q}, \omega)$, where $\mathcal{Q}$ is the manifold that defines the unit system.
+This is the only manifold that allows us to define state-variable-invariant densities.
 The metric allows us to measure the cardinality of possibilities within and across independent d.o.f.
@@
 The answer is yes in the discrete topology and is no in the continuous case.
-The consequence is that when we define the measure $\mu$ for the cardinality of continuous states, we assign a finite value to a finite region, and we assign zero to a single state.
+The consequence is that when we define the measure $\mu$ for the cardinality of continuous states, we assign a finite value to a finite region, and we assign zero measure to a single state.
 That is what gave us integration, densities and, ultimately, conjugate variables.
@@
 
-In other words: this is not a case were the difference at the small scale has a small difference in the large scale.
+In other words: this is not a case were a difference at  small scale has a small difference at large scale.
 The two cases are radically different.
@@
 Finally, we should note that we in a position similar to the one discussed in \ref{discrete_state_metric} for discrete states.
-We have both a topological space and a metric that allows us to count states and possibilities.
+We have both a topological space and a measure expressible in terms of state variables that allows us to count states and possibilities.
 We recovered the metric starting from the idea of an infinitesimally reducible system.
@@
 
-Now that we have fully characterized what we mean by a classical system, we can stipulate the following:
+Now that we have fully characterized what we mean by a classical material, we can stipulate the following:
 
@@
 	In other words: the definition of a state requires some kind of asymmetry between system under study and environment.
+The environment should not affect the system, so that we can define an independent state, yet the system must affect the environment, so that we have external process that allows to define physical distinguishability.
 If we try to assume that both the system and environment are really made of infinitesimal homogeneous parts, then the claim to that asymmetry is lost.
-Therefore our justification for holding the assumption has to be that the pieces are small compared to some scale but large compared to another.
-For example, water is a homogeneous material if we consider it made up of parts sufficiently small for the scale of the problem at hand but big enough to contain a large number of molecules.
-
-	Other conceptual problems stem from the following two competing requirements: on one side we want the system under study to be isolated so that we can define independent states; on the other side we do not want the system isolated so that we can define physical processes on it.
-For example, if we were to take the state of an infinitesimal part to really be its full description, with no unstated part, then each piece would evolve independently, oblivious to the other parts.
-Each particle would essentially reside in its own separate physical universe.
-
-	If we just assume the whole system is deterministic and reversible, then each part can evolve depending on the states of the other parts.
-But this does not solve the problem entirely.
+Therefore our justification for holding the assumption has to be that the pieces are small compared to some scale (they can be regarded as infinitesimal) but large compared to another (they are little affected by the environment).
+For example, water is a homogeneous material if we consider it made up of parts sufficiently small for the scale of the problem at hand but big enough to contain a large number of molecules such that the average system is sufficiently stable.
+
+	Other conceptual problems stem from the following two competing requirements: on one side we want the system under study to be isolated so that we can define independent states; on the other side we want the environment to not be isolated so that we can define physical processes that interact with the system.
+The problem is that the same physical object could be either under study or part of the environment depending on the situation.
+In other words: we want systems to be interacting so that they can be physically aware of each other, yet we want them not interacting so that we can properly define states on them.
+We clearly can't have it both ways.
+Under our classical assumption, if we were to take the state of an infinitesimal part to really be its full description, with no unstated part, then each piece would evolve independently, oblivious to the other parts.
+Each particle would essentially reside in its own separate physical universe, as it cannot physically distinguish anything else.
+This cannot work.
+
+	If we assumed the system is deterministic and reversible only as a whole, then each part could evolve depending on the states of the other parts.
+This would seem much better, as we have pieces with no unstated part that are physically aware of each others.
+But this does not actually solve the problem of independent oblivious components.
 Let's ignore the question of how we were able to define the states of the parts.
@@
 ones that conserve energy or ones with scalar and vector potentials) are precisely the ones we derive by assuming the pieces are not interacting with each other.
-The problem is that we can always locally separate the evolution into independent degrees of freedom.
+The actual problem is that we can always locally separate the evolution into independent degrees of freedom.
 For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value, they clearly evolve independently.
@@
 In other words: we can't expect to have deterministic and reversible evolution at all levels of decomposable systems if the parts interact in any physically meaningful way.
+
+	The message is that we should \emph{always} beware of thinking that the state of a system is its full description: there is always some unstated part that allows continuous interaction between the system and the environment.
+That is the part that cannot be isolated (i.e.
+we can't take the system out of the physical universe) and therefore can't be assigned a state.
 
@@
 
-	In light of what we discussed, we cannot take the classical assumption to be fundamental, in the sense that we cannot take it to apply to all of physics.
+	In light of what we discussed, we cannot take the classical assumption to be fundamental, in the sense that we cannot take it to strictly apply to all of universe.
 While ultimately flawed, the  assumption can be considered valid for a great number of macroscopic systems, and is also useful as a default assumption of sorts.

### 2016-07-21 2015CompleteDerivation/completeDerivation.tex
Added status and reminder to download latest version
Fixed symplectic manifold justification and other edits

@@
 \begin{abstract}
-\textbf{This manuscript is a work in progress.} Ideas are constantly reshaped to find more precise and elegant arguments. It is provided as is to stimulate discussion.
+\textbf{This manuscript is a work in progress.} Ideas are constantly reshaped to find more precise and elegant arguments.
+It is provided as is to stimulate discussion.
+\textbf{Make sure you have the latest version from http://assumptionsofphysics.org}
 
@@
 
-\section{Introduction}
-
-TODO: Point out most modern physics starts by assuming some mathematical framework. Pros and cons of that.
-
-TODO: High level view of the derivation
+\section{Status}
+
+Section III - Done.
+
+Section IV - Done.
+
+Section V - Mostly done. Needs review.
+
+Section VI - Rough draft
+
+%TODO: Point out most modern physics starts by assuming some mathematical framework. Pros and cons of that.
+
+%TODO: High level view of the derivation
 
@@
 Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c} \in \mathcal{C}$.
-$\rho_\mathcal{c}$ is defined on any chart of $\mathcal{S}$.
-Given two charts, for any $\rho_\mathcal{c}(q)$ there is a corresponding $\rho_\mathcal{c}(q')=\rho_\mathcal{c}(q^i) | d q' / d q |$.
+Let $q^i$ and $\hat{q}^j$ two sets of coordinates.
+Let $\rho_\mathcal{c}(q^i)$ and $\rho_\mathcal{c}(\hat{q}^j)$ be the expression in local coordinates of the distribution.
+$\rho_\mathcal{c}(\hat{q}^j)=\rho_\mathcal{c}(q^i) | \partial_j q^i |$ as it transforms as a density.
+$\rho_\mathcal{c}(\mathcal{s}) = \rho_\mathcal{c}(q^i(\mathcal{s})) = \rho_\mathcal{c}(\hat{q}^j(\mathcal{s}))$.
 The Jacobian $| d q' / d q |$ exists and is non-zero.
@@
 Note that neither of those expressions match the norm given by the vector product defined as $\langle f, g \rangle = \int_{\mathcal{S}} fg \, d\mu$.
-Moreover, since $\mathcal{C}$ excludes discontinuous functions, it can never be a complete normed space: it will not include the limit for all Cauchy sequences.
+Moreover, since $\mathcal{C}$ excludes discontinuous functions, it can never be a complete normed space: it will not include the limit for all Cauchy sequences, it cannot be a Banach or a Hilbert space.
 Such constructions, though, are still useful.
@@
 \begin{align*}
-\Lambda_{\mathcal{S}} (\mathcal{c}) &= \int_\mathcal{S} \rho_{\mathcal{c}} d \mu = \int_\mathcal{S} 1\cdot  \rho_{\mathcal{c}} d \mu \\
-\Lambda_U (\mathcal{c}) &= \int_U \rho_{\mathcal{c}} d \mu = \int_\mathcal{S} 1_U \cdot \rho_{\mathcal{c}} d \mu \\
-p_{total}(\mathcal{c}) &= \int_\mathcal{S} p \cdot \rho_{\mathcal{c}} d \mu \\
-q_{avg}(\mathcal{c}) &= 1/|\rho_{\mathcal{c}}| \int_\mathcal{S} q \cdot \rho_{\mathcal{c}} d \mu
+\Lambda_{\mathcal{S}} (\mathcal{c}) &= \int_\mathcal{S} \rho_{\mathcal{c}} d \mu = \langle 1 , \rho_{\mathcal{c}} \rangle \\
+\Lambda_U (\mathcal{c}) &= \int_U \rho_{\mathcal{c}} d \mu = \langle 1_U , \rho_{\mathcal{c}} \rangle \\
+\rho_\mathcal{c}(q_0,p_0) &= \int_\mathcal{S} \delta_{q_0,p_0} \rho_{\mathcal{c}} d \mu = \langle \delta_{q_0,p_0} , \rho_{\mathcal{c}} \rangle \\
+p_{total}(\mathcal{c}) &= \int_\mathcal{S} p \cdot \rho_{\mathcal{c}} d \mu = \langle p , \rho_{\mathcal{c}} \rangle \\
+q_{avg}(\mathcal{c}) &= \frac{\int_\mathcal{S} q \cdot \rho_{\mathcal{c}} d \mu}{\int_\mathcal{S} \rho_{\mathcal{c}} d \mu} = \frac{\langle q , \rho_{\mathcal{c}} \rangle}{\langle 1 , \rho_{\mathcal{c}} \rangle}
 \end{align*}
-For each state, they represent respectively the total amount of material, the material within $U$, the total momentum and the average of a state variable.
+For each state, they represent respectively the total amount of material, the material within $U$, the density at $(q_0, p_0)$, the total momentum and the average of a state variable.
 As they are linear functionals of $\mathcal{C}$,  they can be expressed using their dual vector over $\mathcal{S}$.
-But these vectors ($1$, $1_U$, $p$, $q/|\rho_{\mathcal{c}}|$) are \emph{not} elements of $\mathcal{C}$.
-They do not represent states, they are not continuous and integrable, and there is no single general physical meaning for them.
+But these vectors ($1_U$, $p$, ...) are \emph{not} elements of $\mathcal{C}$.
+They do not represent states, they are not continuous and integrable, and there is no single general physical meaning for all of them.
 Therefore, as we mentioned before, while we can extend the function space for convenience, computation purposes or to study a limit case, we need to be mindful of the extension and carefully check that the mathematical results are physically meaningful case by case.
@@
 This only happens if $\partial_i \hat{k}^j$ is a $1\times1$ matrix: there is only one $k$.
-This gives means $\partial_{k} \hat{k} = d_{\hat{q}} q$ and $\hat{k} = d_{\hat{q}} q \, k + f(q)$.
+This gives $\partial_{k} \hat{k} = d_{\hat{q}} q$ and $\hat{k} = d_{\hat{q}} q \, k + f(q)$.
 As this is a change of units, we expect the zero value of $k$ to remain constant: $\hat{k}(q, 0) = 0 = f(q)$.
@@
 \begin{justification}
-	We claim the simplest state space $\mathcal{S}$ that allows state-variable-invariant densities is the symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of a one dimensional differential manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
-The description of the simplest physical object on which we can define a state-variable-invariant must be given by at least one continuous state variable $q$.
+	We claim the simplest state space $\mathcal{S}$ that allows state-variable-invariant densities is the cotangent bundle $T^*\mathcal{Q}$ of a single dimensional manifold $\mathcal{Q}$.
+Let $\mathcal{S}$ be the $n$ dimensional manifold for the particle states.
+$\mathcal{S}$ must be at least one dimensional to allow state-variable-invariant densities.
+Consider a set of state variables for $U \in \mathcal{S}$: the unit system for the simplest state space will be fully described by a single state variable.
+Let $q$ be the state variable that defines the unit system.
 Let $\mathcal{Q}$ be the manifold charted by $q$.
@@
 $\mathcal{Q}$ is differentiable.
-Let $q$ fully define the unit system for the description of the physical object.
-Then the physical object is represented by a geometrical object on $\mathcal{Q}$.
-The physical object is infinitesimal, therefore is represented by a local geometrical object.
-The simplest geometrical object in $\mathcal{Q}$ is a point, charted by $q$.
-A density over points is not invariant: $\rho(q')_\mathcal{c}= \rho(q)_\mathcal{c} dq/dq'$.
-Next we consider local geometrical objects charted by 2 variables.
-Densities over vectors $v(q) e_q \in T\mathcal{Q}$ are not invariant: $\rho_\mathcal{c}(q', v')= \rho(q, v)_\mathcal{c} (dq/dq')^2$.
-Densities over covectors $k(q) dq \in T^*\mathcal{Q}$ are invariant: $\rho_\mathcal{c}(q', k')= \rho_\mathcal{c}(q, k) (dq/dq') (dq'/dq) = \rho_\mathcal{c}(q, k)$.
-Covectors are the simplest geometrical object to allow unit independent densities.
-The state space $\mathcal{S}$ for the simplest object that allows state-variable-independent densities is $T^*\mathcal{Q}$ of a one dimensional differential manifold $\mathcal{Q}$.
+A diffeomorphism $\hat{q}=\hat{q}(q)$ fully defines a unit transformation.
+Let $(q,k^i)$ with $i=1...n-1$ be a set of state variable.
+The Jacobian determinant of the transformation is $|J| = d_q\hat{q} |\partial_i\hat{k}^j|$.
+$|J|=1$ as densities must be invariant.
+$|\partial_i \hat{k}^j| = d_{\hat{q}}q$.
+$\partial_i \hat{k}^j$ is one dimensional as its components must be fully specified by the previous relationship.
+Let $k=k^1$.
+$\hat{k}(q,0) = 0$ as a change in units does not change the $0$ value.
+$\partial_k \hat{k} = d_{\hat{q}}q$.
+$\hat{k} = d_{\hat{q}}q k$.
+$k$ changes as the component of a co-vector.
+Each state is identified by a point and a co-vector in $\mathcal{Q}$.
+The simplest state space $\mathcal{S}$ is isomorphic to $T^*\mathcal{Q}$ where $Q$ is the one dimensional manifold that defines the units.
+
+	We claim the simplest state space $\mathcal{S}$ that allows state-variable-invariant densities is the symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of a one dimensional differential manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
 Express integration in coordinates.
@@
 We recognize the familiar set of Hamilton's equations.
-They are the set of equations to describe the deterministic and reversible motion of the infinitesimal parts of an infinitesimally reducible homogeneous material.
+They are the set of equations that describe the deterministic and reversible motion of the infinitesimal parts of an infinitesimally reducible homogeneous material.
 In other words: the forces that conserve energy (i.e.

### 2016-07-22 2015CompleteDerivation/completeDerivation.tex
Changing numbering for subsection
Review classical Hamiltonian section

@@
 
+\renewcommand\thesubsection{\thesection.\Alph{subsection}}
 \renewcommand{\theequation}{\thesection.\arabic{equation}}
@@
 no amount of material).
+$\mathcal{C}$ is an additive monoid by definition.
 
 We claim $\mathcal{C}$ is an additive group.
-As we want to describe changes during the evolution (i.e.
-$\hat{\mathcal{c}} = \mathcal{c} +\delta \mathcal{c}$) we introduce an inverse $- : \mathcal{C} \rightarrow \mathcal{C}$ such that $\mathcal{c} + ( - \mathcal{c}) = 0 \forall \mathcal{c} \in \mathcal{C}$.
-Such inverse includes objects that do not properly represent a physical state, but a state change.
+We require $\mathcal{C}$ to include state changes.
 A change of a physically distinguishable object is itself physically distinguishable therefore $\mathcal{C}$ is still a Hausdorff and second countable topological manifold as per \ref{topologically_distinguishable}.
+There exists an inverse $- : \mathcal{C} \rightarrow \mathcal{C}$ such that $\mathcal{c} + ( - \mathcal{c}) = 0 \forall \mathcal{c} \in \mathcal{C}$.
+$\mathcal{C}$ is an additive group by definition.
 \end{justification}
@@
 For example, a photon is a particle of light (it is the smallest amount of light we can describe), an infinitesimal amount of water is treated classically as a point particle (the smallest amount for a continuous fluid).
-We use this definition as it allows us to retroactively talk on somewhat equal grounds about classical point-particles and quantum particles, and underline how physically (and mathematically) they are very similar yet very different objects.
-
-We will call a \emph{classical material} one that is infinitesimally reducible.
+\footnote{We use this definition as it allows us to retroactively talk on somewhat equal grounds about classical point-particles and quantum particles, and underline how physically (and mathematically) they are very similar yet very different objects.}
+
+We will call a \emph{classical material} one that is homogeneous and infinitesimally reducible.
 That is, we can keep decomposing the system indefinitely into smaller and smaller parts, each with a well defined state.
@@
 The notion that the parts are infinitesimal requires the co-domain of $\rho_\mathcal{c}$ to be a real number, as opposed to an integer.
-The state space $\mathcal{S}$, instead, can have a discrete topology.
+The state space $\mathcal{S}$ could, instead, be a discrete set.
 For example, for a system composed of different tanks connected by pipes, the state of the composite system could be the overall distribution of water among the tanks.
@@
 \begin{prop}\label{classical_vector space}
-The state space $\mathcal{C}$ for a homogeneous classical (i.e.
-infinitesimally reducible) material is a vector space over $\mathbbm{R}$ isomorphic to a subspace of the space of continuous functions.
+The state space $\mathcal{C}$ for a classical (i.e.
+homogeneous infinitesimally reducible) material is a vector space over $\mathbbm{R}$ isomorphic to a subspace of the space of continuous functions.
 That is $\mathcal{C} \cong G \subseteq C(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is continuous$\}$, where $\mathcal{S}$ is the state space of an infinitesimal amount of material.
@@
 
-Consider the set of transformations $T$ that increase or decrease the amount of material in the system by a constant factor.
-We claim $T$ is a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics or differential geometry sense (a physical quantity/tensor with a value for each point in space).} isomorphic to $\mathbbm{R}$.
+We claim $T$, the set of transformations that increase or decrease the amount of material in the system by a constant factor, is a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics or differential geometry sense (a physical quantity/tensor with a value for each point in space).} isomorphic to $\mathbbm{R}$.
 Consider $\tau: \mathbbm{R} \rightarrow T$ the mapping between a number and the transformation that increases or decreases the amount of material by that factor.
@@
 We claim $\mathcal{C}$ is isomorphic to a subspace of $C(\mathcal{S})$ as a vector space over $\mathbbm{R}$.
-$\forall \mathcal{c} \in \mathcal{C} \exists !
+$\forall \mathcal{c} \in \mathcal{C} \; \exists !
 \rho_{\mathcal{c}}:\mathcal{S} \rightarrow \mathbbm{R}$ returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$.
@@
 In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique measure\footnote{Here measure is intended in the mathematical sense (a real valued function of a $\sigma$-algebra) which is distinct from other connotations in physics or probability theory.} $\mu$ for the size of sets of $\mathcal{S}$.
-From the measure we get a metric $d\mu$ and our positive linear functionals become integrals, $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
+In terms of said measure, positive linear functionals become integrals, $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
 
@@
 \begin{prop}\label{integration}
-	Let $\mathcal{S}$ be the state space for the particles of a homogeneous classical material.
-Let each state $\mathcal{s} \in \mathcal{S}$ be fully identified by a finite number of independent state variables.
-Then $\mathcal{S}$ is endowed with a Borel measure $\mu$.
-The state space $\mathcal{C}$ for a homogeneous classical material is isomorphic to a subspace of the space of Lebesgue integrable functions.
+	The state space $\mathcal{S}$ for the particles of a classical material is endowed with a Borel measure $\mu$.
+The state space $\mathcal{C}$ for a classical material is isomorphic to a subspace of the space of Lebesgue integrable functions.
 That is $\mathcal{C} \cong G \subseteq L^1(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \; \int_{\mathcal{S}} |\rho| d\mu < \infty \}$.
@@
 	We claim that $\mathcal{S}$ is endowed with a unique Borel measure $\mu$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
-Each $\mathcal{s} \in \mathcal{S}$ is identified by a finite number of independent state variables.
-$\mathcal{S}$ is a manifold.
-$\mathcal{S}$ is locally compact.
+$\mathcal{S}$ is locally compact as it is a manifold.
 $\mathcal{C} \cong G \subseteq C(\mathcal{S})$ therefore $\Lambda_U(\mathcal{c}) \cong \Lambda_U(\rho_\mathcal{c})$.
@@
 
-On a state space with discrete topology, using a single state variable, we have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho_\mathcal{c}(q)$.
+On a state space with countable elements, using a single state variable, we have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho_\mathcal{c}(q)$.
 We would expect the expression to become $\Lambda_U (\mathcal{c}) = \int_a^b \rho_\mathcal{c} (q) dq$ for states identified by a continuous state variable.
-By changing state variable, though, we see that this does not work in general: $\rho_\mathcal{c}(q')= \rho(q)_\mathcal{c} dq/dq'$.
+By changing state variable, though, we see that this does not work in general: $\rho_\mathcal{c}(\hat{q})= \rho(q)_\mathcal{c} dq/d\hat{q}$.
 This makes $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using.
@@
 The distribution $\rho_\mathcal{c}$ is a real valued function of the state.
-As such, at least locally, we can use it as a state variable $q^1=\rho_\mathcal{c}$.
-As a state variable it identifies the states by the density of the material associated with them.
-This is more physically meaningful as one may first suspect, since placing physical markers is a common way to define the references for a coordinate system, which in the end is placing material at particular states.
+As such, at least locally, we can use it as a state variable $q^1=\rho_\mathcal{c}$: we can identify particle states by the density of the material associated with them.
+This is more physically meaningful as one may first suspect, since placing physical markers (i.e.
+placing material at particular states) is a common way to define the references for a coordinate system.
 In a coordinate system where $q^1=\rho_\mathcal{c}$, the function is clearly differential as $\partial_1 \rho_\mathcal{c} = 1$ and $\partial_i \rho_\mathcal{c} = 0$ for $i \neq 1$.
@@
 \begin{prop}\label{differentiable_manifold}
-	The state space $\mathcal{S}$ for the particles of a  classical material identified by a finite number of continuous independent state variables is a differentiable manifold.
-The state space $\mathcal{C}$ for a homogeneous classical material is isomorphic to the space of Lebesgue integrable differentiable functions.
+	The state space $\mathcal{S}$ for the particles of a  classical material is a differentiable manifold.
+The state space $\mathcal{C}$ for a classical material is isomorphic to the space of Lebesgue integrable differentiable functions.
 That is $\mathcal{C} \cong C^1(\mathcal{S}) \cap L^1(\mathcal{S}, \mu)$, where $C^1(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is differentiable$\}$.
@@
 Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c} \in \mathcal{C}$.
-Let $q^i$ and $\hat{q}^j$ two sets of coordinates.
+Let $q^i$ and $\hat{q}^j$ two sets of independent state variables in $U\subseteq \mathcal{S}$.
 Let $\rho_\mathcal{c}(q^i)$ and $\rho_\mathcal{c}(\hat{q}^j)$ be the expression in local coordinates of the distribution.
-$\rho_\mathcal{c}(\hat{q}^j)=\rho_\mathcal{c}(q^i) | \partial_j q^i |$ as it transforms as a density.
+$\rho_\mathcal{c}(\hat{q}^j)=\rho_\mathcal{c}(q^i) | \partial_j q^i |$ as $\rho_\mathcal{c}$ transforms as a density.
 $\rho_\mathcal{c}(\mathcal{s}) = \rho_\mathcal{c}(q^i(\mathcal{s})) = \rho_\mathcal{c}(\hat{q}^j(\mathcal{s}))$.
@@
 The set of Lebesgue integrable functions $L^1(\mathcal{S})$ includes discontinuous functions that for \ref{continuous_map} are unphysical.
-The set of continuous functions $C(\mathcal{S})$ and the set of continuous functions that vanish at infinity $C_0(\mathcal{S})$ not only include non-differentiable functions, but also include functions whose integral is infinite, which would represent infinite amount of material.
-Such functions are also present in the space of differentiable functions $C^1(\mathcal{S})$.
+The space of continuous functions $C(\mathcal{S})$, the space of continuous functions that vanish at infinity $C_0(\mathcal{S})$ and the space of differentiable functions $C^1(\mathcal{S})$ include functions whose integral is infinite, which would represent infinite amount of material.
 Some definitions are too restrictive.
-Requiring compact support (i.e.the function is different from zero only on a finite region) would exclude distributions, such as Gaussians, that span over the whole range of states.
+Requiring compact support (i.e.
+the function is different from zero only on a finite region) would exclude distributions, such as Gaussians, that span over the whole range of states.
 Such distributions are physical: if we take a finite volume of an ideal gas at equilibrium, the momentum distribution spans over the whole range.
 Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a general physical justification for.\footnote{One can, though, make the argument that given any distribution $\rho$ one can find an infinitely smooth $\rho_{sm}$ such that the difference in description is small compared to the error already introduced by assuming the system to be homogeneous and infinitely reducible.
-The Whitney approximation theorem guarantees that this can be done mathematically.}
+The Whitney approximation theorem makes this mathematically well defined.}
 
@@
 Note that neither of those expressions match the norm given by the vector product defined as $\langle f, g \rangle = \int_{\mathcal{S}} fg \, d\mu$.
-Moreover, since $\mathcal{C}$ excludes discontinuous functions, it can never be a complete normed space: it will not include the limit for all Cauchy sequences, it cannot be a Banach or a Hilbert space.
-Such constructions, though, are still useful.
+Moreover $\mathcal{C}$, since it excludes discontinuous functions, can never be a complete normed space: it will not include the limit for all Cauchy sequences, it cannot be a Banach or a Hilbert space.
+Such construction, though, is still useful.
 Consider the following expressions:
@@
 \Lambda_U (\mathcal{c}) &= \int_U \rho_{\mathcal{c}} d \mu = \langle 1_U , \rho_{\mathcal{c}} \rangle \\
-\rho_\mathcal{c}(q_0,p_0) &= \int_\mathcal{S} \delta_{q_0,p_0} \rho_{\mathcal{c}} d \mu = \langle \delta_{q_0,p_0} , \rho_{\mathcal{c}} \rangle \\
+\rho_\mathcal{c}(q_0,p_0) &= \int_\mathcal{S} \delta_{q_0,p_0} \cdot \rho_{\mathcal{c}} d \mu = \langle \delta_{q_0,p_0} , \rho_{\mathcal{c}} \rangle \\
 p_{total}(\mathcal{c}) &= \int_\mathcal{S} p \cdot \rho_{\mathcal{c}} d \mu = \langle p , \rho_{\mathcal{c}} \rangle \\
@@
 
-Now that we are guaranteed that the density is properly defined for all state variables, we have to make sure that is invariant.
-The distribution should only depend on the states, and not the particular choice of state variables.
+Now that we are guaranteed that the density is properly defined for all state variables, we have to make sure it is invariant.
+The distribution should only depend on the states, and not the particular choice of state variables to label them.
 
@@
 \end{align*}
-The Jacobian has to be unitary, therefore all elements $\partial_i \hat{k}^j$ must be well defined and $|\partial_i \hat{k}^j| = 1/d\hat{q}$.
+The Jacobian has to be unitary, therefore all elements $\partial_i \hat{k}^j$ must be well defined and $|\partial_i \hat{k}^j| = 1/d_q\hat{q}$.
 This only happens if $\partial_i \hat{k}^j$ is a $1\times1$ matrix: there is only one $k$.
@@
 Therefore $\hat{k} = d_{\hat{q}} q \, k$ transforms as the component of a covariant vector.
-$\mathcal{S}$ is isomorphic to $T^*\mathcal{Q}$ where $\mathcal{Q}$ is the manifold that defines our unit.
+$\mathcal{S}$ is isomorphic to $T^*\mathcal{Q}$, the space of co-vectors at a point, where $\mathcal{Q}$ is the manifold that defines our unit.
 
@@
 In SI units the relationship $\hbar dk = dp = m du$ between kinetic and conjugate momentum, derived later, sets the relationship between units.} The cardinality for states in a finite (i.e.
-compact) region $U \subseteq \mathcal{S}$ will be given by $\mu(U) = \int_U \hbar dq \wedge dk$.
-We are finally able to write the volume of integration in terms for state variables.
+compact) region $U \subseteq \mathcal{S}$ will be given by $\mu(U) = \int_U \hbar dq \wedge dk$: we are finally able to write the volume of integration in terms for state variables.
 
@@
 are orthogonal surfaces in $\mathcal{S}$.
-The possibilities of an arbitrary degree of freedom, then, will be the sum of the projections over the $n$ orthogonal and independent d.o.f.
+The possibilities of an arbitrary degree of freedom, then, will be $\sum \limits_{i=1}^n \hbar dq^i \wedge dk_i$, the sum of the projections over the $n$ orthogonal and independent d.o.f.
 defined by $(q^i, k_i)$.
@@
 This is the only manifold that allows us to define state-variable-invariant densities.
-The metric allows us to measure the cardinality of possibilities within and across independent d.o.f.
+The metric allows us to measure the cardinality of possibilities on an arbitrary d.o.f.
 
 \begin{prop}\label{classical_phase_space}
-The state space $\mathcal{S}$ for the particles of a homogeneous classical material is a symplectic manifold formed by a cotangent bundle $T^*\mathcal{Q}$ equipped with the canonical two-form $\omega$.
+The state space $\mathcal{S}$ for the particles of a classical material is a symplectic manifold formed by a cotangent bundle $T^*\mathcal{Q}$ equipped with the canonical two-form $\omega$.
 \end{prop}
@@
 	We claim the simplest state space $\mathcal{S}$ that allows state-variable-invariant densities is the cotangent bundle $T^*\mathcal{Q}$ of a single dimensional manifold $\mathcal{Q}$.
-Let $\mathcal{S}$ be the $n$ dimensional manifold for the particle states.
-$\mathcal{S}$ must be at least one dimensional to allow state-variable-invariant densities.
-Consider a set of state variables for $U \in \mathcal{S}$: the unit system for the simplest state space will be fully described by a single state variable.
+Consider a set of state variables for $U \in \mathcal{S}$: the simplest unit system is one defined by a single state variable.
 Let $q$ be the state variable that defines the unit system.
 Let $\mathcal{Q}$ be the manifold charted by $q$.
-$\mathcal{Q}$ is an embedded manifold of $\mathcal{S}$.
-$\mathcal{Q}$ is differentiable.
 A diffeomorphism $\hat{q}=\hat{q}(q)$ fully defines a unit transformation.
 Let $(q,k^i)$ with $i=1...n-1$ be a set of state variable.
-The Jacobian determinant of the transformation is $|J| = d_q\hat{q} |\partial_i\hat{k}^j|$.
-$|J|=1$ as densities must be invariant.
+The Jacobian determinant $|J| = d_q\hat{q} |\partial_i\hat{k}^j|=1$ as densities must be invariant.
 $|\partial_i \hat{k}^j| = d_{\hat{q}}q$.
@@
 Let $k=k^1$.
+$\partial_k \hat{k} = d_{\hat{q}}q$.
 $\hat{k}(q,0) = 0$ as a change in units does not change the $0$ value.
-$\partial_k \hat{k} = d_{\hat{q}}q$.
 $\hat{k} = d_{\hat{q}}q k$.
@@
 The notion that the continuous case is a limit of the discrete case, that it's ``like the discrete but with more points", leads in this case to erroneous intuition.
-The question is: can we identify a single state?
+The key question is: can we physically distinguish an isolated state?
 Can we have an outcome associated with only one element?
-The answer is yes in the discrete topology and is no in the continuous case.
+The answer is no with the standard topology on $\mathbbm{R}^n$.
 The consequence is that when we define the measure $\mu$ for the cardinality of continuous states, we assign a finite value to a finite region, and we assign zero measure to a single state.
 That is what gave us integration, densities and, ultimately, conjugate variables.
-If we were to use the discrete topology on $\mathbbm{R}^n$, if we were to assume we can identify single states, then we would assign measure one to each state and infinity to a finite region.
+If we were to use the discrete topology on $\mathbbm{R}^n$, if we were to assume we can identify single states as we do for countable states, then we would assign measure one to each state and infinity to a finite region.
 We would not have integration, just a simple sum.
@@
 We can tell them apart.
-We should take the use of densities and conjugate quantities as empirical evidence that the quantities like space and time are not discrete.
+We should take the use of densities and conjugate quantities as evidence that quantities like space and time are not discrete.
 Or more precisely: that the processes we use to distinguish those quantities cannot identify single instances, as that is all we can say experimentally.
@@
 
-Finally, we should note that we in a position similar to the one discussed in \ref{discrete_state_metric} for discrete states.
+It should also be noted that classical particles, under this light, cannot be considered point-like.
+As they are the limit of infinitesimal subdivision, their spatial extent becomes infinitesimal but not zero.\footnote{This picture is also compatible with general relativity, unlike point-like particles.
+These would not follow geodesics as their infinite mass density would significantly affect the gravitational field.} But suppose particles were truly point-like.
+Then distinguishing particles would mean distinguishing points.
+We are back to the idea of a discrete topology on $\mathbbm{R}^n$.
+In that case we would not have conjugate momentum, no $T^*\mathcal{Q}$, just the coordinates of the point in $\mathcal{Q}$.
+This would actually be more self-consistent: why wouldn't a point be enough to define the state of a point-like particle?
+As before, we should take the use of conjugate quantities as evidence that classical particles are really the limit of an infinitesimal subdivision.
+
+Finally, we should note that we are in a position similar to the one discussed in \ref{discrete_state_metric} for discrete states.
 We have both a topological space and a measure expressible in terms of state variables that allows us to count states and possibilities.
@@
 
-\begin{assump}[Classical Hamiltonian assumption]\label{classical}
+\begin{assump}[Infinitesimal reducibility]\label{classical}
 	The system under study is composed of an infinitesimally reducible homogeneous material and each part undergoes deterministic and reversible evolution.
@@
 To be able to talk about the states of a part we need a physical process that is able to distinguish between them.
-For a pool ball we can imagine marking one spot with a different color.
+For a pool ball we can imagine marking one spot with a red marker.
 This allows us to track it as the ball moves or collides with other balls.
@@
 The electromagnetic field generated can be considered determined only by our system with good approximation.
-But this also breaks down: if part of the environment is made of the same homogeneous material the charge density will be the same and the interaction between particles will be symmetrical.
+But this also breaks down: if each part of the environment is made of the same homogeneous material the charge density will be the same and the interaction between particles will be symmetrical.
 
@@
 Therefore our justification for holding the assumption has to be that the pieces are small compared to some scale (they can be regarded as infinitesimal) but large compared to another (they are little affected by the environment).
-For example, water is a homogeneous material if we consider it made up of parts sufficiently small for the scale of the problem at hand but big enough to contain a large number of molecules such that the average system is sufficiently stable.
+For example, water is a homogeneous material if we consider it made up of parts sufficiently small for the scale of the problem at hand but big enough to contain a large number of molecules such that the average motion is sufficiently stable.
 
@@
 
-	The message is that we should \emph{always} beware of thinking that the state of a system is its full description: there is always some unstated part that allows continuous interaction between the system and the environment.
+	The moral is that we should \emph{always} beware of thinking that the state of a system is its full description: there is always some unstated part that allows continuous interaction between the system and the environment.
 That is the part that cannot be isolated (i.e.
@@
 The evolution must map the distribution point-wise so that the value associated at the initial state is the same as the value at the final state.
-All of the system that starts in $\mathcal{s}$ has to end up in $\hat{\mathcal{s}}$.
+All the material that starts in $\mathcal{s}$ has to end up in $\hat{\mathcal{s}}$.
 In math terms $\rho_{\hat{\mathcal{c}}} (\hat{\mathcal{s}}) = \rho_\mathcal{c}(\mathcal{s})$.
@@
 \begin{prop}\label{canonical_transformation}
-	A deterministic and reversible evolution map for a homogeneous classical material is a self-symplectomorphism.
+	A deterministic and reversible evolution map for the particles of a classical material is a self-symplectomorphism.
 That is: $\mathcal{T}_{\Delta t}: T^*\mathcal{Q} \rightarrow T^*\mathcal{Q}$ and $\mathcal{T}_{\Delta t}^*\omega = \omega$ where $\mathcal{T}_{\Delta t}^*$ is the pullback of $\mathcal{T}_{\Delta t}$.
@@
 	We claim $\mathcal{T}_{\Delta t}$ is a self-homeomorphism on $T^*\mathcal{Q}$.
-The state space for particles of a homogeneous classical material is $T^*\mathcal{Q}$ by \ref{classical_phase_space}.
+The state space for the particles of a classical material is $T^*\mathcal{Q}$ by \ref{classical_phase_space}.
 $\mathcal{T}_{\Delta t}$ is a deterministic and reversible evolution map and by \ref{detrevmap} is a self-homeomorphism.
@@
 \begin{prop}\label{hamiltons_equations}
-	A continuous deterministic and reversible process for a homogeneous classical material admits a Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ that allows us to write the laws of evolution as
+	A continuous deterministic and reversible process for the particles of a classical material admits a Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ that allows us to write the laws of evolution as
 \begin{align*}
@@
 We claim the vector field $S \in T\mathcal{S}$ for the infinitesimal displacement $S^a = d\xi^a/dt$ admits a potential $H$ such that $S^{a} \omega_{ab} = \partial_{b}H$.
-The state space $\mathcal{S}$ of a classical homogeneous material is a symplectic manifold by \ref{classical_phase_space}.
+The state space $\mathcal{S}$ for the particles of a classical material is a symplectic manifold by \ref{classical_phase_space}.
 The map for infinitesimal time evolution $\mathcal{T}_{dt}$ is an infinitesimal self-symplectomorphism by \ref{canonical_transformation}.

### 2016-07-24 2015CompleteDerivation/completeDerivation.tex
More work on time dependent evolution

@@
 We now generalize our discussion to include time dependent dynamics.
-This is needed when the evolution map itself changes in time or when a change of state variable is time dependent (e.g.
+This is needed when the evolution map is not the same at all instants or when state variables depend on time (e.g.
 changing to a moving frame).
 
-To do this we will extend phase space to include the temporal degree of freedom.
+To do this we will redefine the particle state space to include the temporal degree of freedom.
 We'll find that the dynamics is described by relativistic Hamiltonian mechanics in the extended phase space.
-We'll also find that particle states divide into standard and anti-states depending whether time is aligned or anti-aligned with the affine parameter.
-Note that no extra assumption is needed, which makes relativistic mechanics not some additional characterization but simply the correct way to describe time dependent motion of invariant density distributions.
+We'll also find that particle states divide into standard and anti-states depending whether time is aligned or anti-aligned with the evolution parameter.
+Note that no extra assumption is needed, which makes relativistic mechanics simply the correct time dependent description of the evolution of an invariant density distribution, and not the description of a different type of system.
 
@@
 
-The second case is when we perform a transformation for the time parameter $t'=t'(t)$.
-The equations of motion transform and we have $d_{t}t' d_{t'}q^i = \partial_{p_i} H$.
-As both $q^i$ and $H$ are independent of time, they cannot be redefined to include $d_{t}t'$.
+The second case is when we perform a transformation for the time parameter $\hat{t}=\hat{t}(t)$.
+The equations of motion transform and we have $d_{t}\hat{t} d_{\hat{t}}q^i = \partial_{p_i} H$.
+As both $q^i$ and $H$ are independent of time, they cannot be redefined to include $d_{t}\hat{t}$.
 Physically the transformation is introducing fictitious forces that are not conservative, so they cannot be expressed by a Hamiltonian.
@@
 
-The third case is when we perform a state variable transformation $q'=q'(q,t)$ that is time dependent (e.g.
-$q'=q+vt$).
-Our composite state distribution $\rho(q)$ becomes $\rho(q',t)$: it is no longer defined at equal time.
-This means that both our measure $\mu$ and our metric $d\mu$ need to be modified.
+The third case is when we perform a state variable transformation $\hat{q}=\hat{q}(q,t)$ that is time dependent (e.g.
+$\hat{q}=q+vt$).
+Our composite state distribution $\rho(q)$ becomes $\rho(\hat{q},t)$: it is no longer defined at equal time.
+This means that our measure $\mu$ needs to be modified to define integrals over the time variable.
 
@@
 
-\subsection{Extended phase space}
-
-Our goal is to describe maps that take a state at a particular time and transport it to another state at another time.
-Therefore we want to extend the domain and co-domain of the evolution map to be the space of all particle states at all times.
+\subsection{Complete state space}
+
+The first thing to do is to amend our definition of state space to include all states at all times.
+It may seem like we are extending our definition of state and state space but, at a closer look, we are not.
+Since defined state as a physical configuration at a particular time, the set of all states (i.e.
+the state space) is more accurately the set of all configurations at all times.
+In the previous sections we tacitly simplified the problem by ignoring time, which made it easier to study the time independent case.
+This is common in physics and engineering but this is also the source of the above problems.
+
+We'll therefore call \emph{complete state space} the set of all configurations at all times and we'll indicate it by $\bar{\mathcal{S}}$.
+In that space, a state is an element $\mathcal{s} \in \bar{\mathcal{S}}$ while a time evolution is a line $\bar{\mathcal{s}} \subset \bar{\mathcal{S}}$ that passes once at all times.
+Specifying an evolution map means giving a set of lines such that each state is traversed once and only once.
+Note that the transported state variables are perfect to label the evolutions: as they do not change during the evolution we can take them to label not just the values at a particular time, but the evolutions themselves.
+
+With this geometric picture in mind, we can understand better the complete state space $\bar{\mathcal{C}}$ for the composite system.
+An evolution $\hat{\mathcal{c}}$ will be a distribution of material over all configuration at all times.
+Along a particle state evolution $\bar{\mathcal{s}}$, though, the distribution is constant as all the material is transported deterministically and reversibly.
+A state space $\mathcal{c}$, then, is a distribution over an hyper-surface that cuts across all the particle state evolutions.
+If we think it as a distribution over the transported state variables we are guaranteed to cover all initial conditions no matter the hyper-surface.
 
@@
 When changing variables, we still need to make sure that deterministic and reversible motion can be defined on the new time variable.
-Transformations like $t'=q, \; q'=t$ (exchanging time with another state variable) or $t'=\sqrt{t^2 + q^2}, \; q'=\tan^{-1}(t/q)$ (polar coordinates between time and state variable) clearly do not guarantee deterministic and reversible motion over the new time variable.
-The change of time variable must preserve the time ordering therefore the transformation has to be strictly monotonic between the two time variables.
+Transformations like $\hat{t}=q, \; \hat{q}=t$ (exchanging time with another state variable) or $\hat{t}=\sqrt{t^2 + q^2}, \; \hat{q}=\tan^{-1}(t/q)$ (polar coordinates between time and state variable) clearly do not guarantee deterministic and reversible motion over the new time variable.
+To be meaningful, the change of time variable must preserve the time ordering therefore the transformation has to be strictly monotonic between the two time variables.
 In this sense, the time variable is special.
@@
 \begin{prop}\label{classical_extended_phase_space}
-	The extended state space, the set of states at all possible times, for the particles of a homogeneous classical material is a hyper-surface of the symplectic manifold formed by a cotangent bundle $T^*\mathcal{M}$ equipped with the canonical two-form $\omega = \sum dq^i \wedge dp_i - dt \wedge dE$.
+	The complete state space $\hat{\mathcal{S}}$ for the particles of a classical material is a hyper-surface of the symplectic manifold formed by a cotangent bundle $T^*\mathcal{M}$ equipped with the canonical two-form $\omega = \sum dq^i \wedge dp_i - dt \wedge dE$.
 \end{prop}

### 2016-07-26 MathProblems/mathProblems.tex
Adding document with math problems to solve

\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{dutchcal}
\usepackage{braket}

\newtheorem{prop}[equation]{Proposition}

\begin{document}

\title{Math problems}

\maketitle

\section{Topology for continuous functions}

Let $X$ and $Y$ be two topological space. Let $C(X, Y) = \{f:X \rightarrow Y \; | \; f \text{ is continuous} \}$. Prove or disprove the following propositions.

\begin{prop}
	$C(\mathbbm{R}, \mathbbm{R})$ is a Hausdorff and second countable topological space. 
\end{prop}

\begin{prop}
	$C(\mathbbm{R}^n, \mathbbm{R}^m)$ is a Hausdorff and second countable topological space. 
\end{prop}

\begin{prop}
	Let $X$ be a Hausdorff and second countable manifold. $C(X, \mathbbm{R}^m)$ is a Hausdorff and second countable topological space. 
\end{prop}

\begin{prop}
	Let $X$ and $Y$ be Hausdorff and second countable manifolds. $C(X, Y)$ is a Hausdorff and second countable topological space. 
\end{prop}

\begin{prop}
	Let $X$ and $Y$ be Hausdorff and second countable topological spaces. $C(X, Y)$ is a Hausdorff and second countable topological space. 
\end{prop}

\section{Measure induced by group of self-homeomorphisms}

Prove or disprove the following proposition.

\begin{prop}
	Let X be a Hausdorff and second countable manifold. Let $G$ be a group of self-homeomorphism under composition such that, $\forall U \subseteq X$, $g(U) \subseteq U \implies g(U) = U$ and $g(U) \supseteq U \implies g(U) = U$. There exist a measure $\mu : \mathcal{B}(X) \rightarrow \mathbbm{R}$ such that, $\forall U \subseteq X, g \in G$, $\mu(U) = \mu(g(U))$. Additionally, if $\forall x_1, x_2 \in X \; \exists g \in G \;|\; x_2=g(x_1)$ then $\mu$ is unique up to a multiplicative constant.
\end{prop}

Prove or disprove if $X$ is a Hausdorff and second countable topological space.

\end{document}

### 2016-07-27 2015CompleteDerivation/completeDerivation.tex
First review for relativity

@@
 It may seem like we are extending our definition of state and state space but, at a closer look, we are not.
-Since defined state as a physical configuration at a particular time, the set of all states (i.e.
-the state space) is more accurately the set of all configurations at all times.
+Since we defined state as a physical configuration at a particular time, the set of all states (i.e.
+the state space) is more accurately defined as the set of all configurations at all times.
 In the previous sections we tacitly simplified the problem by ignoring time, which made it easier to study the time independent case.
-This is common in physics and engineering but this is also the source of the above problems.
-
-We'll therefore call \emph{complete state space} the set of all configurations at all times and we'll indicate it by $\bar{\mathcal{S}}$.
+This is common in physics and engineering, which is why we kept the standard name, but this is also the source of the above problems.
+
+We'll therefore call \emph{complete state space} of particle states the set of all configurations at all times and we'll indicate it by $\bar{\mathcal{S}}$.
 In that space, a state is an element $\mathcal{s} \in \bar{\mathcal{S}}$ while a time evolution is a line $\bar{\mathcal{s}} \subset \bar{\mathcal{S}}$ that passes once at all times.
 Specifying an evolution map means giving a set of lines such that each state is traversed once and only once.
-Note that the transported state variables are perfect to label the evolutions: as they do not change during the evolution we can take them to label not just the values at a particular time, but the evolutions themselves.
-
-With this geometric picture in mind, we can understand better the complete state space $\bar{\mathcal{C}}$ for the composite system.
-An evolution $\hat{\mathcal{c}}$ will be a distribution of material over all configuration at all times.
-Along a particle state evolution $\bar{\mathcal{s}}$, though, the distribution is constant as all the material is transported deterministically and reversibly.
-A state space $\mathcal{c}$, then, is a distribution over an hyper-surface that cuts across all the particle state evolutions.
-If we think it as a distribution over the transported state variables we are guaranteed to cover all initial conditions no matter the hyper-surface.
-
-The first thing to do is to extend our $n$-dimensional $\mathcal{Q}$, the manifold that defines our units, to include the time variable $t$.
-We call $\mathcal{M}$ such extended space which is a manifold of dimension $n+1$: state variables and the time variable are independent (i.e.
+Locally, each state will be given by $(q^i, k_i, t)$ therefore $\bar{\mathcal{S}}$ is a $2n+1$ dimensional manifold.
+A region at equal time will be a hyper-surface $\mathcal{S}_{t=t_0}$ that includes all possible configuration at a particular time.
+Such hyper-surface will need to cut across all evolutions once and only once, therefore not all functions of $\bar{\mathcal{S}}$ are suitable time variables.
+
+\begin{prop}
+	The complete state space $\bar{\mathcal{S}}$ for the particles of a classical material is a $2n+1$ dimensional differentiable manifold.
+The state space $\mathcal{S}_{t=t_0}$ at a particular time is a hyper-surface of $\bar{\mathcal{S}}$.
+\end{prop}
+
+\begin{justification}
+	TODO
+\end{justification}
+
+For the composite state we need to take a different approach.
+A distribution of material over states at equal time $\rho_\mathcal{c} : \mathcal{S}_{t=t_0} \rightarrow \mathbbm{R}$ will depend on the choice of time variable, therefore the complete state space will not be useful as it will not be variable independent.
+What we are really interested in is the \emph{evolution space} of a classical material, which we'll indicated by $\bar{\mathcal{C}}$.
+To each evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$ will correspond a distribution $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ over all particle states at all times.
+This distribution will be differentiable, it's differentiable over all equal time hyper-surfaces and time evolution is differentiable, but is not integrable over the complete state space $\bar{\mathcal{S}}$.
+Such integration will operate over time as well, giving the amount of material multiplied by time.
+If time is taken to be infinite the value is not finite.
+The distribution, instead, must be integrable over all equal time hyper-surfaces: no matter what time variable we choose, the amount of material found at a particular time has to be finite.
+
+\begin{prop}
+	The evolution state space $\bar{\mathcal{C}}$ for a classical material is isomorphic to the space of differentiable functions that are Lebesgue integrable on any equal time hyper-surface.
+That is $\bar{\mathcal{C}} \cong  C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$ where $L^1_t(\bar{\mathcal{S}}, \mu)=\{{\rho: \bar{\mathcal{S}} \rightarrow \mathbbm{R}} \; | \; {\int_{\mathcal{S}_{t=t_0}} |\rho| d\mu < \infty} \; \forall \, t, t_0 \}$.
+\end{prop}
+
+\begin{justification}
+\end{justification}
+
+Before proceeding, it's useful to get a better intuition for the geometry of these space.
+Note that the transported state variables are perfect to label the particle state evolution evolutions $\bar{\mathcal{s}} \subset \bar{\mathcal{S}}$: as they do not change during the evolution we can take them to label not just the values at a particular time, but the evolutions themselves.
+The distribution is transported deterministically and reversibly over those lines, so the value of the distribution will be constant along each particle state evolution $\bar{\mathcal{s}}$.
+To identify the full evolution, then, we just need to specify the distribution only on a region that cuts across all particle evolution, i.e.
+a hyper-surface at constant time.
+
+\subsection{Relativistic cotangent bundle}
+
+We now want to find the metric for our complete state space $\bar{\mathcal{S}}$.
+The first thing to do is to complete our $n$-dimensional $\mathcal{Q}$, the manifold that defines our units, to include the time variable $t$.
+In fact, the idea that the set of units to identify a physical configuration at a particular time did not include time should look, retrospectively, shaky at best.
+We call $\mathcal{M}$ such space which is a manifold of dimension $n+1$: state variables and the time variable are independent (i.e.
 any combination of state and time is valid).
-A time variable is not simply another state variable: it doesn't identify extra states and it's the variable on which we have defined deterministic and reversible evolution.
+As we said before, the time variable is not simply another state variable: it doesn't identify extra states and it's the variable on which we have defined deterministic and reversible evolution.
 When changing variables, we still need to make sure that deterministic and reversible motion can be defined on the new time variable.
 Transformations like $\hat{t}=q, \; \hat{q}=t$ (exchanging time with another state variable) or $\hat{t}=\sqrt{t^2 + q^2}, \; \hat{q}=\tan^{-1}(t/q)$ (polar coordinates between time and state variable) clearly do not guarantee deterministic and reversible motion over the new time variable.
-To be meaningful, the change of time variable must preserve the time ordering therefore the transformation has to be strictly monotonic between the two time variables.
-In this sense, the time variable is special.
-
-As before, we are interested in invariant densities which are defined on co-vectors.
-The co-vector $k_i dq^i$ is extended with a time component $\bar{\omega} dt$.\footnote{$\bar{\omega}$ is the classical analogue of the wave frequency.
+To be meaningful, the change of time variable must preserve time ordering therefore the transformation has to be strictly monotonic between the two time variables.
+On the other hand, the change of state variables must preserve the ability to uniquely identify states each instant in time.
+In this sense, time is not just like another state variable.\footnote{In relativity, for example, it is common to use foliations and the $3+1$ formalism to recover the special character of time when performing calculations or to gain better physical insight.}
+
+As in the time independent case, we are interested in invariant densities which are defined on co-vectors.
+The co-vector $k_i dq^i$ is completed with a time component $\bar{\omega} dt$.\footnote{$\bar{\omega}$ is the classical analogue of the wave frequency.
 We use $\bar{\omega}$ to distinguish from the phase-space metric $\omega$.} As state variable $q^i$ have a conjugate quantity $p_i\equiv \hbar k$, the time variable $t$ will have a conjugate quantity $E\equiv\hbar\bar{\omega}$.
@@
 
-The first thing to note is that $E$ cannot be an independent variable.
-It cannot add any states, as those are defined by $T^*\mathcal{Q}$ only, and cannot add any time instants, as those are defined by $t$ alone.
+Most of all, the temporal degree of freedom is not an independent degree of freedom because $E$ cannot be an independent variable.
+It cannot add any states at a particular time, as those are defined by $T^*\mathcal{Q}$ only, and cannot add any time instants, as those are defined by $t$ alone.
 Therefore there must exist a constraint such that, locally, $E=E(t,q^i,p_i)$.
-The set of all states at all times is a sub-manifold of $T^*\mathcal{M}$ of dimension $2n+1$.
-
-A composite state $\mathcal{c}$ and its distribution $\rho_\mathcal{c}$ is no longer defined at an instant of an absolute time parameter as now we are free to mix mix $t$ and $q$'s.
-Yet, there must be at least one time variable so that the distribution is at equal time defined over a suitable $T^*\mathcal{Q} \subset T^*\mathcal{M}$.
-That is, there must be at least one set of variables for which we are able to define states and their evolution.
-Therefore our distributions $\rho_\mathcal{c}$ are defined on $2n$-dimensional sub-manifolds of $T^*\mathcal{M}$.
-In other words: densities are not densities over the time variable.
-
-Yet, as we can mix time and state variables, the volume of integration is not simply $dq^n \wedge dp_n$ as it is not invariant.
-Consider a change of variable $q'=q+vt$: the same surface may have projections of different sizes on $(q, p)$ than $(q', p')$.
-It would seem, then, that the measure for our states is different depending on the choice of variable, which cannot be.
-
-We can understand the geometry of $T^*\mathcal{M}$ in the following way.
-As we have seen, the temporal degree of freedom $(t, E)$ and the other d.o.f.
-$(q^i, p_i)$ are not independent.
-So they are not orthogonal in the extended phase space.
+The complete state space set $\bar{\mathcal{S}}$ is a hyper-surface of $T^*\mathcal{M}$.
+
+As the temporal degree of freedom $(t, E)$ is not independent of the other d.o.f.
+$(q^i, p_i)$, it is not orthogonal to them in $T^*\mathcal{M}$.
 States are defined on the plane where $( q, p )$ (maximally) change.
@@
 
-We can capture the above discussion by stating that the extended state space is a hypersurface of the symplectic manifold $(T^*\mathcal{M}, \omega)$, where $\omega \equiv \sum dq^i \wedge dp_i - dt \wedge dE$.
-This is the simplest manifold that allows us to define time-and-state-variable-invariant densities.
+We can capture the above discussion by stating that the complete state space $\bar{\mathcal{S}}$ is a hypersurface of the symplectic manifold $(T^*\mathcal{M}, \omega)$, where $\omega \equiv \sum dq^i \wedge dp_i - dt \wedge dE$.
+This is the manifold that allows us to define time-and-state-variable-invariant densities.
 The metric allows us to measure the cardinality of possibilities.
@@
 \begin{prop}\label{relativistic_canonical_transformation}
-	A time-dependent deterministic and reversible evolution map for a homogeneous classical material is a self-symplectomorphism on the extended phase space.
+	A time-dependent deterministic and reversible evolution map for a homogeneous classical material is a self-symplectomorphism on $T^*\mathcal{M}$.
 That is: $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ such that $\mathcal{T}_{\Delta s}^*\omega = \omega$ where $\mathcal{T}_{\Delta s}^*$ is the pullback of $\mathcal{T}_{\Delta s}$.
@@
 
-Note that $\mathcal{H}=0$ for all states, since $\mathcal{H}=\lambda(H-E)$ and $E = H$.
+Note that $\mathcal{H}=0$ for all states $\mathcal{s} \in \bar{\mathcal{S}} \subset T^* \mathcal{M}$, since $\mathcal{H}=\lambda(H-E)$ and $E = H$.
 For all states we also have $\lambda \neq 0$.
 Therefore $\mathcal{H}=0$ only because $E = H$.
-We can extend $\mathcal{H}$ on all of $T^* \mathcal{M}$ making sure that $\mathcal{H} \neq 0$ outside of the hyper-surface that contains states.
-This way we can directly use the constraint $\mathcal{H}=0$ to identify such hyper-surface.
+We can extend $\mathcal{H}$ on all of $T^* \mathcal{M}$ making sure that $\mathcal{H} \neq 0$ outside of the complete state space $\bar{\mathcal{S}}$.
+This way we can directly use the constraint $\mathcal{H}=0$ to identify $\bar{\mathcal{S}}$ as a subspace of $T^* \mathcal{M}$.
 
@@
 	The form of the extended Hamiltonian is $\mathcal{H} = \lambda(t,E,q^i,p_i) \, (H(t,q^i,p_i) - E)$ where $H(t,q^i,p_i)$ is the standard Hamiltonian at each instant $t$, the conjugate time variable $E = H(t,q^i,p_i)$ represents the value of the Hamiltonian, and  $\lambda : T^*\mathcal{M} \rightarrow \mathbbm{R}$ where $\lambda(\xi^a) = d_s t$.
-$\mathcal{H} = 0$ for all states at all time.
+$\bar{\mathcal{S}}$ is the level set for $\mathcal{H} = 0$.
 \end{prop}
@@
 
-Now that we have characterized the motion of the particles, we should also characterize the motion of the composite state.
-
-If we have a distribution of material $\rho_\mathcal{c}(s)$ for a particular value of $s$, the first thing we know is that its support, the region where the density is non-zero, must be within the constraint $\mathcal{H} = 0$.
-This means $\mathcal{H} \rho_\mathcal{c}(s) = 0$ since $\mathcal{H} = 0$ where $\rho_\mathcal{c}(s) \neq 0$.
-If the distribution is differentiable, the evolution is given by the Poisson bracket $d_s \rho_\mathcal{c} (s) = \{ \rho_\mathcal{c}(s), \mathcal{H} \}$.
-
-We can also consider the distribution of material at all times $\rho_\mathcal{c}\equiv \int \rho_\mathcal{c}(s) ds$.
-The support of the distribution is also constrained so we have $\mathcal{H} \rho_\mathcal{c} = 0$.
+Now that we have characterized the motion of the particles, we should also characterize the evolution of composite states.
+
+Suppose we have a distribution of material $\rho_{\bar{\mathcal{c}}}(s)$ associated with an evolution $\bar{\mathcal{c}}$.
+Its support, the region where the density is non-zero, is within $\bar{\mathcal{S}}$.
+Therefore  $\mathcal{H} \rho_{\bar{\mathcal{c}}}(s) = 0$ since $\mathcal{H} = 0$ in $\bar{\mathcal{S}}$ where $\rho_{\bar{\mathcal{c}}}(s) \neq 0$.
 As the density is transported over each trajectory, its value does not change over $s$.
-Therefore, if the distribution is differentiable, we have $d_s \rho_\mathcal{c} = \{ \rho_\mathcal{c}, \mathcal{H} \} = 0$.
+Therefore we have $d_s \rho_{\bar{\mathcal{c}}} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \} = 0$.
 
 \begin{prop}\label{relativistic_distribution_evolution}
-	Let $\rho_\mathcal{c}(s)$ the distribution of material at a particular moment in the evolution.
-$\mathcal{H} \rho_\mathcal{c}(s) = 0$ and $d_s \rho_\mathcal{c} (s) = \{ \rho_\mathcal{c}(s), \mathcal{H} \}$ where $\{ f, g \} \equiv \partial_a f \omega^{ab} \partial_b g$.
-Let $\rho_\mathcal{c}$ the distribution of material at all times.
-$\mathcal{H} \rho_\mathcal{c} = 0$ and $d_s \rho_\mathcal{c} = \{ \rho_\mathcal{c}, \mathcal{H} \} = 0$.
-\end{prop}
-
-\begin{justification}
+	Let $\rho_{\bar{\mathcal{c}}}(s)$ the distribution of material associated with the evolution $\bar{\mathcal{c}}$.
+$\mathcal{H} \rho_{\bar{\mathcal{c}}} = 0$ and $d_s \rho_{\bar{\mathcal{c}}} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \} = 0$.
+\end{prop}
+
+\begin{justification}
+	TODO: language consistency
+
 	We claim $\mathcal{H} \rho_\mathcal{c}(s) = 0 \; \forall \mathcal{c} \in \mathcal{C} \wedge s \in \mathbbm{R}$.
@@
 Relativistic Hamiltonian mechanics is simply the correct way of handling time dependent motion.
-As such, there was no mention of the speed of light $c$, $\mathcal{Q}$ being a Riemannian manifold or the metric $g$.
+As such, there was no mention of the speed of light $c$, $\mathcal{M}$ being a pseudo-Riemannian manifold or its metric $g$.
 These will be introduced in the next chapter with the kinematic equivalence assumption.
-In light of this, it is probably better to consider relativistic mechanics as a feature of the geometry of the extended phase space more than the geometry of space-time.
+In light of this, it is probably better to consider relativistic mechanics as a feature of the geometry of the complete state space more than the geometry of space-time.
+Also note that this feature arises from the use of the standard topology on $\mathbbm{R}^n$ and the need of invariant densities.
+It would not have arisen with the use of a discrete topology.
 
@@
 As we'll see later, to this parallel in the mathematical formalism corresponds a parallel in the physics description.
+
+As a final thought, we note how the physics maps less elegantly to the math in the time dependent case.
+The function space for $\bar{\mathcal{C}}$ is not simply the Lebesgue integrable differentiable functions, but those that are integrable on particular hyper-surfaces.
+The complete state space $\bar{\mathcal{S}}$ is a hyper-surface of $T^*\mathcal{M}$.
+The extended Hamiltonian is needed to identify that subspace.
+This may be a hint that a better formulation is lurking.
+One where we consider the particle evolutions $\bar{\mathcal{s}}$ as the primary objects, label those (instead of the states).
+The distribution of material would simply be a distribution over the particle evolutions, a Lebesgue integrable differentiable functions.
 

### 2016-07-27 2015CompleteDerivation/completeDerivation.tex
First stab at kinematic equivalence

@@
 
-\subsection{Classical Hamiltonian assumption}
+\subsection{Infinitesimal reducibility}
 
@@
 
-\section{Kinematic equivalence}
-TODO
+\section{Elementary systems}
+
+Reductionism by itself would not help us describing a system if the state and the evolution of the pieces were more complicated to study.
+In this section we further focus our attention to those system that not only are infinitesimally reducible, but their internal dynamics can be ignored.
+That is, we will assume that the motion in space is necessary and sufficient to reconstruct the state and its dynamics.
+What we'll find is that the dynamics of each infinitesimal part follows classical Lagrangian mechanics.
+Furthermore, the Hamiltonian of the system is constrained to the one of a free particles under scalar and vector potential forces.
+
+\subsection{Kinematic equivalence}
+
+The link between kinematics, the study of motion through quantities like velocity and acceleration, and dynamics, the study of laws of evolution, is crucial in physics.
+In Newtonian physics, $\vec{F}=m\vec{a}$ provides such link.
+In Hamiltonian and Lagrangian mechanics, the link is implicitly given by the choice of the Hamiltonian and the Lagrangian.
+
+In this work we make the link conceptually explicit by assuming the following:
+
+\begin{assump}[Kinematic equivalence]\label{kinematic_equivalence}
+	Studying the kinematic of the system (i.e. trajectories in space) is equivalent to study the dynamics of the system (i.e. state trajectories in state space).
+\end{assump}
+
+\begin{rationale}
+	The idea is that the system internal dynamics has no affect on the motion and therefore can be disregarded.
+Consider the motion of our cannonball under gravitational and inertial forces: its temperature or precise chemical composition does not affect the trajectory.
+We can assume kinematic equivalence.
+Consider the motion of a helicopter under gravitational and inertial forces: the pilot and the amount of fuel in the tank do affect the trajectory.
+We cannot assume kinematic equivalence.
+
+	The assumption clearly cannot be taken in general since many systems do not posses this trait.
+Yet, as this assumption allows to recover the standard link between kinematic and dynamical quantities, there is something fundamental about it.
+That is: if we assume that the system has no discernible internal dynamics, then the only dynamics that needs to be described is the one associated with its trajectory in space: kinematic equivalence holds.
+
+	This is in fact consistent with observation, though elementary particles are quantum fields therefore some caveats are in order.
+Massless particles should be excluded as they cannot be localized, which makes it hard to talk about a proper trajectory.
+Spin is not associated to kinematic quantities: nothing is thought to be actually spinning.
+As such we can regard it as the internal dynamic of the system.
+A massive spin $0$ system is one that can be thought as localized and as having no internal structure.
+For such system we would expect kinematic equivalence to hold and it does: from kinematic momentum we can reconstruct conjugate momentum and vice-versa.
+
+	Therefore, with the appropriate caveat that we are talking about a quantum particle which we haven't yet introduced, the simplest systems are the ones that just describe a trajectory in space.
+This raises the opposite question: why all systems have \emph{at least} to describe a trajectory in space?
+This seems motivated by the previous assumption of reducibility: spatial extent is the one way that all materials are reduced.
+Therefore all particles of all materials, regardless of whether they are elementary or not, are will include spatial degrees of freedom.
+
+	It should be also noted that, once we reduced particles to be infinitesimal at a point, the rest of the description must be a local object invariant under coordinate transformations.
+This means that the remaining internal dynamics must be identified scalar, vector or tensor quantities.
+Spin is, in some sense, a quantized version of those.
+
+	While we can make an argument why kinematic equivalence is important because it is valid for elementary system, we cannot actually make an argument why nature should be reducible to elementary system.
+While it fits our physical intuition, we have no a priori argument for it.
+We can imagine, in fact, that as we reduce the behavior of one material to its constituents we find the parts to possess richer and richer states.
+For example, most macroscopic object are neutral though the parts are charged.
+We could imagine that, while dividing into parts, we find more and more types of charges and forces.
+We do not have an argument why this cannot happen.
+\end{rationale}
 

### 2016-07-28 2015CompleteDerivation/completeDerivation.tex
First pass at all justification for relativity

@@
 
-If $\rho_\mathcal{c}$ is to remain invariant under state variable changes, not only the Jacobian has to exist but must be unitary.
+If $\rho_{\mathcal{c}}$ is to remain invariant under state variable changes, not only the Jacobian has to exist but must be unitary.
 This means we cannot simply change one state variable as we please.
@@
 
-\begin{prop}\label{classical_phase_space}
+\begin{prop}\label{symplectic_manifold}
 The state space $\mathcal{S}$ for the particles of a classical material is a symplectic manifold formed by a cotangent bundle $T^*\mathcal{Q}$ equipped with the canonical two-form $\omega$.
@@
 
-\begin{assump}[Infinitesimal reducibility]\label{classical}
+\begin{assump}[Infinitesimal reducibility]\label{infinitesimal_reducibility}
 	The system under study is composed of an infinitesimally reducible homogeneous material and each part undergoes deterministic and reversible evolution.
@@
 	We claim $\mathcal{T}_{\Delta t}$ is a self-homeomorphism on $T^*\mathcal{Q}$.
-The state space for the particles of a classical material is $T^*\mathcal{Q}$ by \ref{classical_phase_space}.
+The state space for the particles of a classical material is $T^*\mathcal{Q}$ by \ref{symplectic_manifold}.
 $\mathcal{T}_{\Delta t}$ is a deterministic and reversible evolution map and by \ref{detrevmap} is a self-homeomorphism.
@@
 We claim the vector field $S \in T\mathcal{S}$ for the infinitesimal displacement $S^a = d\xi^a/dt$ admits a potential $H$ such that $S^{a} \omega_{ab} = \partial_{b}H$.
-The state space $\mathcal{S}$ for the particles of a classical material is a symplectic manifold by \ref{classical_phase_space}.
+The state space $\mathcal{S}$ for the particles of a classical material is a symplectic manifold by \ref{symplectic_manifold}.
 The map for infinitesimal time evolution $\mathcal{T}_{dt}$ is an infinitesimal self-symplectomorphism by \ref{canonical_transformation}.
@@
 
-\begin{prop}
+\begin{prop}\label{complete_particle_state_space}
 	The complete state space $\bar{\mathcal{S}}$ for the particles of a classical material is a $2n+1$ dimensional differentiable manifold.
@@
 \begin{justification}
-	TODO
+	We claim $\bar{\mathcal{S}}$ is an $2n+1$ dimensional differentiable manifold.
+Locally $\bar{\mathcal{S}} \supseteq U \cong T^*\mathcal{Q}\times \mathbbm{R}$ as a state $\mathcal{s} \in U$ exists for each configuration and each time instant.
+Locally $\bar{\mathcal{S}} \supseteq U \cong\mathbbm{R}^{2n+1}$.
+$\bar{\mathcal{S}}$ is an $2n+1$ dimensional manifold.
+$T^*\mathcal{Q}$ is a differentiable manifold.
+Time evolution is a diffeomorphism by \ref{hamiltons_equations} therefore derivatives between state variables and time variables always exist and are invertible.
+$\bar{\mathcal{S}}$ is a differentiable manifold.
+
+	We claim $\mathcal{S}_{t=t_0}$ is a hyper-surface of $\bar{\mathcal{S}}$.
+Let $t : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be a time variable.
+$\mathcal{S}_{t=t_0} = \{ \mathcal{s} \in \bar{\mathcal{S}} \; | \; t(\mathcal{s}) = t_0 \}$ is a level set.
+$t$ is monotonic as it is a time variable.
+$t$ has no critical points.
+$t_0$ is a regular value.
+$\mathcal{S}_{t=t_0}$ is an embedded submanifold of co-dimension $1$.
+$\mathcal{S}_{t=t_0}$ is an hyper-surface.
 \end{justification}
@@
 
-\begin{prop}
+\begin{prop}\label{material_evolution_state_space}
 	The evolution state space $\bar{\mathcal{C}}$ for a classical material is isomorphic to the space of differentiable functions that are Lebesgue integrable on any equal time hyper-surface.
@@
 \begin{justification}
+	We claim $\bar{\mathcal{C}} \cong  G \subseteq C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}$.
+Let $\rho_{\bar{\mathcal{c}}}$ be the distribution of material associated with an evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$.
+Let $t : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be a time variable.
+	The restriction of $\rho_{\bar{\mathcal{c}}}$ over an arbitrary level set $\mathcal{S}_{t=t_0}$ is continuous and integrable by \ref{differentiable_manifold}.
+$\rho_{\bar{\mathcal{c}}} \in L^1_t(\bar{\mathcal{S}}, \mu)$ Time evolution is a diffeomorphism by \ref{hamiltons_equations}.
+$\rho_{\bar{\mathcal{c}}}$ is differentiable along the time variable.
+$\rho_{\bar{\mathcal{c}}} \in C^1(\bar{\mathcal{S}})$
+
+	We claim $\bar{\mathcal{C}} \cong  C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$.
+As in \ref{differentiable_manifold}, if two distributions are different there exists an outcome that can tell them apart, therefore they represent two distinct physical evolutions.
 \end{justification}
@@
 
-\begin{prop}\label{classical_extended_phase_space}
+\begin{prop}\label{relativistic_symplectic_manifold}
 	The complete state space $\hat{\mathcal{S}}$ for the particles of a classical material is a hyper-surface of the symplectic manifold formed by a cotangent bundle $T^*\mathcal{M}$ equipped with the canonical two-form $\omega = \sum dq^i \wedge dp_i - dt \wedge dE$.
@@
 \begin{justification}
-TODO
-% justification.
-Need to be able to define integration for everywhere and space and on compact time.
-Metric exists.
-Must be a two-form.
-Two-form is skewsymmetric.
-All skew symmetric forms are decomposable in hyperplanes.
-T*Q is a subspace of T*M.
-Only one hyperplane left.
-Metric must be \Sigma dq^i \wedge dp_i \pm dt \wedge dE.
-Can't be positive, as ()t, E) is not an independent d.o.f.
+	We claim the space of physical objects that allows a time-state-invariant-variable densities is a symplectic manifold $(T^*\mathcal{M}, \omega)$ where $\omega = \sum dq^i \wedge dp_i - dt \wedge dE$.
+Let $\mathcal{M}$ be the manifold that defines the unit system including time.
+Locally $\mathcal{M} \cong \mathcal{Q} \times \mathbbm{R}$ as time and state variables are independent quantities.
+As in \ref{symplectic_manifold}, invariant densities are defined on the symplectic manifold $(T^*\mathcal{M}, \omega)$.
+Let $\bar{\omega}$ be the co-vector component associated with $t$.
+Let $E=\hbar \bar{\omega}$.
+Let $\omega$ be the metric.
+The terms across different d.o.f.
+must be zero as they are not invariant under unit transformations.
+$\omega=\sum dq^i \wedge dp_i \pm dt \wedge dE$.
+The $+$ case is excluded as the temporal degree is not an independent d.o.f.: $\omega = \sum dq^i \wedge dp_i - dt \wedge dE$.
+
+	We claim $\hat{\mathcal{S}}$ is a hyper-surface of $(T^*\mathcal{M}, \omega)$.
+$\hat{\mathcal{S}} \subseteq (T^*\mathcal{M}, \omega)$ as states are physical objects that allow time-state-invariant-variable densities.
+Consider the map $f : T^*\mathcal{M} \rightarrow \hat{\mathcal{S}} \; | \; (q^i, p_i, t, E) \mapsto (q^i, p_i, t)$.
+$f$ is an open map.
+The topology of $\hat{\mathcal{S}}$ is the subspace topology.
+Let $\rho : T^*\mathcal{M} \rightarrow \mathbbm{R}$ be an invariant density.
+$\rho$ is differentiable by \ref{differentiable_manifold}.
+It's restriction on $\hat{\mathcal{S}}$ is also an invariant distribution and therefore differentiable.
+The inclusion map $\hat{\mathcal{S}} \hookrightarrow T^*\mathcal{M}$ is a smooth embedding.
+$\hat{\mathcal{S}}$ is an embedded submanifold of co-dimension $1$.
+$\hat{\mathcal{S}}$ is an hypersurface.
 \end{justification}
@@
 	We claim $\mathcal{T}_{\Delta s}$ is a self-homeomorphism on $T^*\mathcal{M}$.
-The extended state space for particles of a homogeneous classical material is $T^*\mathcal{M}$ by \ref{classical_extended_phase_space}.
+The extended state space for particles of a homogeneous classical material is $T^*\mathcal{M}$ by \ref{relativistic_symplectic_manifold}.
 $\mathcal{T}_{\Delta s}$ is a deterministic and reversible evolution map and by \ref{detrevmap} is a self-homeomorphism.
@@
 $\omega$ is invariant under a deterministic and reversible evolution map.
-$\mathcal{T}_{\Delta s}$ is a symplectomorphism by definition.\end{justification}
+$\mathcal{T}_{\Delta s}$ is a symplectomorphism by definition.
+\end{justification}
 
@@
 We claim the vector field $S \in T(T^*\mathcal{M})$ for the infinitesimal displacement $S^a = d\xi^a/ds$ admits a potential $\mathcal{H}$ such that $S^{a} \omega_{ab} = \partial_{b}\mathcal{H}$.
-The extended state space $T^*\mathcal{M}$ of a classical homogeneous material is a symplectic manifold by \ref{classical_extended_phase_space}.
+The extended state space $T^*\mathcal{M}$ of a classical homogeneous material is a symplectic manifold by \ref{relativistic_symplectic_manifold}.
 The map for infinitesimal evolution $\mathcal{T}_{ds}$ is an infinitesimal self-symplectomorphism by \ref{relativistic_canonical_transformation}.
@@
 In other words: $E$ is the value of the Hamiltonian, the energy of the system.
-We have found the constraint we discussed when introducing \ref{classical_extended_phase_space}.
+We have found the constraint we discussed when introducing \ref{relativistic_symplectic_manifold}.
 
@@
 \begin{justification}
-	TODO: language consistency
-
-	We claim $\mathcal{H} \rho_\mathcal{c}(s) = 0 \; \forall \mathcal{c} \in \mathcal{C} \wedge s \in \mathbbm{R}$.
-Let $\rho_\mathcal{c}(s) : T^*\mathcal{M} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c}$ at a given value of $s$.
-The distribution is defined on physical states.
-By \ref{form_of_extended_hamiltonian} $supp(\rho_\mathcal{c}(s)) \subseteq T^*\mathcal{M} \,\backslash\, supp(\mathcal{H})$.
-$supp(\mathcal{H} \rho_\mathcal{c}(s)) = supp(\mathcal{H}) \cap supp(\rho_\mathcal{c}(s)) \subseteq \varnothing$.
-
 	We claim $d_s \rho_\mathcal{c} (s) = \{ \rho_\mathcal{c}(s), \mathcal{H} \}$.
@@
 
-	We claim $\mathcal{H} \rho_\mathcal{c} = 0 \; \forall \mathcal{c}$.
-Let $\rho_\mathcal{c} : T^*\mathcal{M} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c}$ at all states at all times.
-The distribution is defined on physical states.
-By \ref{form_of_extended_hamiltonian} $supp(\rho_\mathcal{c}) \subseteq T^*\mathcal{M} \,\backslash\, supp(\mathcal{H})$.
-$supp(\mathcal{H} \rho_\mathcal{c}) = supp(\mathcal{H}) \cap supp(\rho_\mathcal{c}) \subseteq \varnothing$.
-
-	We claim $d_s \rho_\mathcal{c} = \{ \rho_\mathcal{c}(s), \mathcal{H} \} = 0$.
-Let $\rho_\mathcal{c}$ be differentiable.
-$d_s \rho_\mathcal{c}= \partial_a \rho_\mathcal{c} d_s \xi^a = \partial_a \rho_\mathcal{c}(s) \omega^{ab} \partial_b \mathcal{H} = \{ \rho_\mathcal{c}, \mathcal{H} \}$.
-Let $\gamma : \mathbbm{R} \rightarrow T^*\mathcal{M}$ a particle trajectory.
+	We claim $\mathcal{H} \rho_{\bar{\mathcal{c}}} = 0 \; \forall \bar{\mathcal{c}} \in \bar{\mathcal{C}}$.
+Let $\rho_{\bar{\mathcal{c}}} : T^*\mathcal{M} \rightarrow \mathbbm{R}$ be the distribution associated with an evolution $\bar{\mathcal{c}}$.
+$supp(\rho_{\bar{\mathcal{c}}}) \subseteq \bar{\mathcal{S}}$ as the distribution is defined on the complete state space of the particles.
+$supp(\mathcal{H}) = T^*\mathcal{M} \,\backslash\, \bar{\mathcal{S}}$ by \ref{form_of_extended_hamiltonian}.
+$supp(\mathcal{H} \rho_\mathcal{c}) = supp(\mathcal{H}) \cap supp(\rho_\mathcal{c}) = \varnothing$.
+
+	We claim $d_s \rho_{\bar{\mathcal{c}}} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \} = 0$.
+$d_s \rho_{\bar{\mathcal{c}}}= \partial_a \rho_{\bar{\mathcal{c}}} d_s \xi^a = \partial_a \rho_{\bar{\mathcal{c}}} \omega^{ab} \partial_b \mathcal{H} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \}$.
+Let $\gamma : \mathbbm{R} \rightarrow T^*\mathcal{M}$ be a parametrization for a particle evolution $\bar{\mathcal{s}}$.
 Then $\rho_\mathcal{c}(\lambda(s)) = \rho_\mathcal{c}(\lambda(s + \Delta s))$.

### 2016-08-01 2015CompleteDerivation/completeDerivation.tex
Add math for the classical uncertainty principle
Adding a figure

@@
 \usepackage{braket}
+
+\usepackage{tikz}
+\usetikzlibrary{calc}
+\usepackage{calculator}
+\usepackage{standalone}
 
@@
 defined by $(q^i, k_i)$.
+
+\begin{figure}
+	\includestandalone[width=\columnwidth]{pictures/independentDof}
+	\caption{The metric $\omega$ and its components along two independent degrees of freedom.
+As we change state variables, the components are still orthogonal to each other.
+Note: this is a 2D conceptualization of what happens in 4D.}
+	\label{fig:independent_dof}
+\end{figure}
 
@@
 
+\section{Prelude to quantum}
+
+TODO
+
+\subsection{Classical uncertainty principle}
+
+\begin{align*}
+L = &\int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp \\
+&+ \lambda_1(\int \rho dqdp - \rho_{tot})\\
+&+ \lambda_2(- \int \rho \ln \rho \, dqdp - H_0)\\
+\delta L = &\int \delta \rho [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\ &+ \lambda_1 + \lambda_2 \ln \rho + \lambda_2 ] dqdp = 0 \\
+\rho = &\frac{\rho_{tot}}{ \sqrt{4 \pi ^2\sigma_q^2\sigma_p^2}} e^{-\frac{(q-\mu_q)^2(p-\mu_p)^2}{4\sigma_q^2\sigma_p^2}} \\
+= &\frac{\rho_{tot}}{e^{H_0 - 1}} e^{-\frac{\pi ^2(q-\mu_q)^2(p-\mu_p)^2}{\exp (2(H_0 - 1))}}
+\end{align*}
+\begin{align*}
+\sigma_q\sigma_p \geq \exp (H_0 - 1) / 2 \pi 
+\end{align*}
+
 \section{Quantum systems}

### 2016-08-01 2015CompleteDerivation/pictures/independentDof.tex
Adding a figure

\documentclass{standalone}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{calculator}
\begin{document}
	\begin{tikzpicture}[scale=.4]

	% INPUTS

	\newcommand{\xvar}{5};					% x component of omega
	\newcommand{\yvar}{6};					% y component of omega	
	\newcommand{\xprimevar}{6.4};	    % length of the bottom side of the second triangle, called x prime
	% Note: must have  x' < omega = sqrt(x^2 + y^2)

	\newcommand{\dvar}{20}; 				% space between the two drawings

	% outer rectangle and diagram name
	\draw (-8,-12) rectangle (34,14);
	\node at (13,-8) {\huge Independent d.o.f.};

	% outer rectangle and diagram name, alternate method, needs more calculations...
	%		\draw (-\yvar - 2,-12) rectangle (\dvar + \dvar, \yvar + \xvar + 2);
	%		\node at (13,-8) {\huge Independent d.o.f.};

	% calculate omega from x and y components
	\SQUARE{\xvar}{\xsquared};
	\SQUARE{\yvar}{\ysquared};
	\ADD{\xsquared}{\ysquared}{\sum};	
	\SQUAREROOT{\sum}{\omegalength};

	% calculate the angle that omega creates w.r.t. the horizontal axis
	\DIVIDE{\yvar}{\xvar}{\sol};	
	\ARCTAN{\sol}{\newsol};
	\RADtoDEG{\newsol}{\thetavar};

	% calculate y prime and theta prime from x prime 
	\SQUARE{\omegalength}{\omegasquared};
	\SQUARE{\xprimevar}{\xprimesquared};
	\SUBTRACT{\omegasquared}{\xprimesquared}{\dif};
	\SQUAREROOT{\dif}{\yprimevar};

	\DIVIDE{\yprimevar}{\xprimevar}{\sol};	
	\ARCTAN{\sol}{\newsol};
	\RADtoDEG{\newsol}{\thetaprimevar};

	% calculations
	\newcommand{\phivar}{90-\thetavar};							  % second angle 
	\newcommand{\phiprimevar}{90-\thetaprimevar};	 	   % second prime angle

	% drawing the right basic triangle 
	\filldraw[fill=red,opacity=0.5] (0,0) -- (\xvar,0) -- (\xvar,\yvar) -- cycle;

	\filldraw[fill=red,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black] (0,0) rectangle (\xvar,-\xvar) node[pos=.5] {\large $ {dq^1}\wedge {dp_1}$};

	\filldraw[fill=red,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black] (\xvar,0) rectangle (\xvar + \yvar, \yvar) node[pos=.5] {\large ${dq^2}\wedge {dp_2}$};

	\filldraw[fill=red,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar:(0,0)}]  (0,0) rectangle (\omegalength, \omegalength) node[pos=.5, rotate=\thetavar] {\large $\omega$};

	% drawing the complex right triangle  
	\filldraw[black,fill=yellow,opacity=0.5,rotate around={\thetavar:(\dvar,0)}] (\dvar + \omegalength,0) -- (\dvar,0) -- ++(-\thetaprimevar: \xprimevar) -- ++(\phiprimevar:\yprimevar);

	\filldraw[black,fill=yellow,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar - \thetaprimevar:(\dvar,0)}]  (\dvar,0) rectangle (\dvar + \xprimevar,-\xprimevar) node[pos=.5, rotate=\thetavar-\thetaprimevar] {\large $d\hat{q}^1\wedge d\hat{p}_1$};

	\filldraw[black,fill=yellow,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar - \thetaprimevar:(\dvar,0)}]  (\dvar + \xprimevar,0) rectangle (\dvar + \xprimevar + \yprimevar, \yprimevar) node[pos=.5, rotate=\thetavar-\thetaprimevar] {\large $d\hat{q}^2\wedge d\hat{p}_2$};

	\filldraw[black,fill=yellow,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar:(\dvar,0)}]  (\dvar,0) rectangle (\dvar + \omegalength, \omegalength) node[pos=.5, rotate=\thetavar] {\large $\omega$};

	\end{tikzpicture}

\end{document}

### 2016-08-01 2015CompleteDerivation/completeDerivation.tex
Reviewed relativistic parts

@@
 
-Section VI - Rough draft
+Section VI - Mostly done. Needs review.
+
+Section VII - Rough draft
 
@@
 We'll also find that particle states divide into standard and anti-states depending whether time is aligned or anti-aligned with the evolution parameter.
-Note that no extra assumption is needed, which makes relativistic mechanics simply the correct time dependent description of the evolution of an invariant density distribution, and not the description of a different type of system.
+Note that no extra assumption is needed, which makes relativistic mechanics simply the correct time dependent description of the deterministic and reversible evolution of an infinitesimally reducible system, and not the description of an altogether different physical object.
 
@@
 The first obvious case is that the map may be time dependent.
-Note that this does not affect our definition of determinism and reversibility: we still can tell final state from initial state and vice-versa.
-We just need to specify the time as well.
-At this point we have no way to do that.
-For example, the Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ is just a function of the state.
+This does not affect our definition of determinism and reversibility as we still can tell final state from initial state and vice-versa.
+But at this point we have no way to specify a time dependent process: the Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ as we derived it is just a function of the state.
 
@@
 
+As we can see, the framework we have is ill suited to handle these cases.
 Therefore we cannot simply stick time in the Hamiltonian and expect everything to work out.
@@
 In the previous sections we tacitly simplified the problem by ignoring time, which made it easier to study the time independent case.
-This is common in physics and engineering, which is why we kept the standard name, but this is also the source of the above problems.
+This approach is common in physics and engineering, which is why it fits the standard names, but this is also the source of the above problems.
 
@@
 \begin{justification}
-	We claim $\bar{\mathcal{S}}$ is an $2n+1$ dimensional differentiable manifold.
+	We claim $\bar{\mathcal{S}}$ is a $2n+1$ dimensional differentiable manifold.
 Locally $\bar{\mathcal{S}} \supseteq U \cong T^*\mathcal{Q}\times \mathbbm{R}$ as a state $\mathcal{s} \in U$ exists for each configuration and each time instant.
@@
 $T^*\mathcal{Q}$ is a differentiable manifold.
-Time evolution is a diffeomorphism by \ref{hamiltons_equations} therefore derivatives between state variables and time variables always exist and are invertible.
+Time evolution is a diffeomorphism by \ref{hamiltons_equations} therefore derivatives between state variables and time variables always exist.
 $\bar{\mathcal{S}}$ is a differentiable manifold.
@@
 For the composite state we need to take a different approach.
-A distribution of material over states at equal time $\rho_\mathcal{c} : \mathcal{S}_{t=t_0} \rightarrow \mathbbm{R}$ will depend on the choice of time variable, therefore the complete state space will not be useful as it will not be variable independent.
+A distribution of material over states at equal time $\rho_\mathcal{c} : \mathcal{S}_{t=t_0} \rightarrow \mathbbm{R}$ will depend on the choice of time variable, therefore the complete state space will not be useful as it will not be time invariant.
 What we are really interested in is the \emph{evolution space} of a classical material, which we'll indicated by $\bar{\mathcal{C}}$.
@@
 \begin{justification}
-	We claim $\bar{\mathcal{C}} \cong  G \subseteq C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}$.
-Let $\rho_{\bar{\mathcal{c}}}$ be the distribution of material associated with an evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$.
+	We claim $\bar{\mathcal{C}} \cong  G \subseteq C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$.
+Let $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be the distribution of material associated with an evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$.
 Let $t : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be a time variable.
@@
 $\rho_{\bar{\mathcal{c}}} \in L^1_t(\bar{\mathcal{S}}, \mu)$ Time evolution is a diffeomorphism by \ref{hamiltons_equations}.
-$\rho_{\bar{\mathcal{c}}}$ is differentiable along the time variable.
+$\rho_{\bar{\mathcal{c}}}$ is differentiable along the time variable as well as the state variables.
 $\rho_{\bar{\mathcal{c}}} \in C^1(\bar{\mathcal{S}})$
@@
 The first thing to do is to complete our $n$-dimensional $\mathcal{Q}$, the manifold that defines our units, to include the time variable $t$.
-In fact, the idea that the set of units to identify a physical configuration at a particular time did not include time should look, retrospectively, shaky at best.
+This, again, highlights our previous oversight: we didn't include time within the quantities needed to specify a physical configuration at a particular time.
 We call $\mathcal{M}$ such space which is a manifold of dimension $n+1$: state variables and the time variable are independent (i.e.
-any combination of state and time is valid).
-As we said before, the time variable is not simply another state variable: it doesn't identify extra states and it's the variable on which we have defined deterministic and reversible evolution.
+any combination of state and time variables is valid).
+
+As we said before, the time variable is not simply another state variable: it doesn't identify extra configurations and it's the variable on which we have defined deterministic and reversible evolution.
 When changing variables, we still need to make sure that deterministic and reversible motion can be defined on the new time variable.
@@
 To be meaningful, the change of time variable must preserve time ordering therefore the transformation has to be strictly monotonic between the two time variables.
-On the other hand, the change of state variables must preserve the ability to uniquely identify states each instant in time.
+On the other hand, the change of state variables must preserve the ability to uniquely identify states at each instant in time.
 In this sense, time is not just like another state variable.\footnote{In relativity, for example, it is common to use foliations and the $3+1$ formalism to recover the special character of time when performing calculations or to gain better physical insight.}
@@
 Most of all, the temporal degree of freedom is not an independent degree of freedom because $E$ cannot be an independent variable.
-It cannot add any states at a particular time, as those are defined by $T^*\mathcal{Q}$ only, and cannot add any time instants, as those are defined by $t$ alone.
+It cannot add any configurations, as those are defined by $T^*\mathcal{Q}$ only, and cannot add any time instants, as those are defined by $t$ alone.
 Therefore there must exist a constraint such that, locally, $E=E(t,q^i,p_i)$.
@@
 This allows us to better understand its physical meaning.
-The metric allows us to count the possibilities within a degree of freedom, adjusting the metric on $T^*\mathcal{Q}$ to avoid the counting problems introduced by time variable changes.
+The metric allows us to count the possibilities within a degree of freedom, adjusting the metric on $T^*\mathcal{M}$ to avoid the counting problems introduced by time variable changes.
 
@@
 The invariance of $\omega$ already guarantees that.
-As we saw before that fact that the vector product is conserved within a degree of freedom is equivalent to saying that changing variables conserves the possibilities within a d.o.f.
+As we saw before that fact that the vector product is conserved within a degree of freedom is equivalent to saying that changing variables does not change the possibilities within a d.o.f.
 In fact, the flow of an infinitesimal transformation that preserves the vector product on a surface is a divergence free field.
@@
 In fact, the flow of an infinitesimal transformation that preserves the Minkowsky product on a surface is a curl free field.
+
+TODO: add proofs in the appendix
 
@@
 Let $\omega$ be the metric.
-The terms across different d.o.f.
+The components of $\omega$ across different d.o.f.
 must be zero as they are not invariant under unit transformations.
@@
 
-	We claim $\hat{\mathcal{S}}$ is a hyper-surface of $(T^*\mathcal{M}, \omega)$.
-$\hat{\mathcal{S}} \subseteq (T^*\mathcal{M}, \omega)$ as states are physical objects that allow time-state-invariant-variable densities.
-Consider the map $f : T^*\mathcal{M} \rightarrow \hat{\mathcal{S}} \; | \; (q^i, p_i, t, E) \mapsto (q^i, p_i, t)$.
+	We claim $\bar{\mathcal{S}}$ is a hyper-surface of $(T^*\mathcal{M}, \omega)$.
+$\bar{\mathcal{S}} \subseteq (T^*\mathcal{M}, \omega)$ as states are physical objects that allow time-state-invariant-variable densities.
+Consider the map $f : T^*\mathcal{M} \rightarrow \bar{\mathcal{S}} \; | \; (q^i, p_i, t, E) \mapsto (q^i, p_i, t)$.
 $f$ is an open map.
-The topology of $\hat{\mathcal{S}}$ is the subspace topology.
+The topology of $\bar{\mathcal{S}}$ is the subspace topology.
 Let $\rho : T^*\mathcal{M} \rightarrow \mathbbm{R}$ be an invariant density.
 $\rho$ is differentiable by \ref{differentiable_manifold}.
-It's restriction on $\hat{\mathcal{S}}$ is also an invariant distribution and therefore differentiable.
-The inclusion map $\hat{\mathcal{S}} \hookrightarrow T^*\mathcal{M}$ is a smooth embedding.
-$\hat{\mathcal{S}}$ is an embedded submanifold of co-dimension $1$.
-$\hat{\mathcal{S}}$ is an hypersurface.
+It's restriction on $\bar{\mathcal{S}}$ is also an invariant distribution and therefore differentiable.
+The inclusion map $\bar{\mathcal{S}} \hookrightarrow T^*\mathcal{M}$ is a smooth embedding.
+$\bar{\mathcal{S}}$ is an embedded submanifold of co-dimension $1$.
+$\bar{\mathcal{S}}$ is an hypersurface.
 \end{justification}
@@
 	We claim $\mathcal{T}_{\Delta s}$ is a self-homeomorphism on $T^*\mathcal{M}$.
-The extended state space for particles of a homogeneous classical material is $T^*\mathcal{M}$ by \ref{relativistic_symplectic_manifold}.
+The complete state space for particles of a classical material is $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ by \ref{relativistic_symplectic_manifold}.
 $\mathcal{T}_{\Delta s}$ is a deterministic and reversible evolution map and by \ref{detrevmap} is a self-homeomorphism.
+We can extend the map over $T^*\mathcal{M}$.
 
@@
 \begin{prop}\label{relativistic_hamiltons_equations}
-	A continuous time-dependent deterministic and reversible process for a homogeneous classical material admits an extended Hamiltonian $\mathcal{H}: T^*\mathcal{M} \rightarrow \mathbbm{R}$ that allows us to write the laws of evolution as
+	A continuous time-dependent deterministic and reversible process for a homogeneous classical material admits an invariant Hamiltonian $\mathcal{H}: T^*\mathcal{M} \rightarrow \mathbbm{R}$ that allows us to write the laws of evolution as
 	\begin{align*}
@@
 We claim the vector field $S \in T(T^*\mathcal{M})$ for the infinitesimal displacement $S^a = d\xi^a/ds$ admits a potential $\mathcal{H}$ such that $S^{a} \omega_{ab} = \partial_{b}\mathcal{H}$.
-The extended state space $T^*\mathcal{M}$ of a classical homogeneous material is a symplectic manifold by \ref{relativistic_symplectic_manifold}.
 The map for infinitesimal evolution $\mathcal{T}_{ds}$ is an infinitesimal self-symplectomorphism by \ref{relativistic_canonical_transformation}.
@@
 \begin{prop}\label{antistates}
-	Let $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ the time-variant deterministic and reversible evolution map for the particles of a homogeneous classical material, where $s$ is the continuous evolution parameter.
+	Let $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ the time dependent deterministic and reversible evolution map for the particles of a classical material.
 The map partitions the extended state space into standard states, those connected by a trajectory where $d_{s}t>0$, and anti-states, those connected by a trajectory where $d_{s}t<0$.
@@
 
-With the extended Hamiltonian we are able to write the equations of motion for any choice of time and state variables.
+With the invariant Hamiltonian we are able to write the equations of motion for any choice of time and state variables.
 For example: $d_t q^i = d_t s \, d_s q^i = d_s q^i / d_s t = - \partial_{p_i} \mathcal{H} / \partial_{E} \mathcal{H}$.
@@
 \begin{prop}\label{form_of_extended_hamiltonian}
-	The form of the extended Hamiltonian is $\mathcal{H} = \lambda(t,E,q^i,p_i) \, (H(t,q^i,p_i) - E)$ where $H(t,q^i,p_i)$ is the standard Hamiltonian at each instant $t$, the conjugate time variable $E = H(t,q^i,p_i)$ represents the value of the Hamiltonian, and  $\lambda : T^*\mathcal{M} \rightarrow \mathbbm{R}$ where $\lambda(\xi^a) = d_s t$.
+	The form of the invariant Hamiltonian is $\mathcal{H} = \lambda(t,E,q^i,p_i) \, (H(t,q^i,p_i) - E)$ where $H(t,q^i,p_i)$ is the standard Hamiltonian at each instant $t$, the conjugate time variable $E = H(t,q^i,p_i)$ represents the value of the Hamiltonian, and  $\lambda : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ where $\lambda(\xi^a) = d_s t$.
 $\bar{\mathcal{S}}$ is the level set for $\mathcal{H} = 0$.
@@
 	We claim the form of the extended Hamiltonian is $\mathcal{H} = \lambda(t,E,q^i,p_i) \, (H(t,q^i,p_i) - E)$.
-Let $\mathcal{H}$ the extended Hamiltonian and $H$ be Hamiltonian for a particular choice of time and state variables.
+Let $\mathcal{H}$ be the invariant Hamiltonian and $H$ be the Hamiltonian for a particular choice of time and state variables.
 We have $d_s q^i = \partial_{p_i} \mathcal{H} = d_s t \, d_t q^i = - \lambda \, \partial_{p_i} H$ where $\lambda : T^*\mathcal{M} \rightarrow \mathbbm{R} \; | \; \lambda \mapsto d_s t$.
@@
 We can redefine $H = H + \mathcal{h}$ as the result is still a function of $q^i, p_i , t$ with the same equations of motion.
-
-	We claim $E = H(t,q^i,p_i)$ for physical states. $\lambda = d_s t = - \partial_E \mathcal{H} = - \partial_{E} \lambda (H - E + \mathcal{h}) - \lambda (-1)$ for any choice of $t$. $H - E = 0$.
-
-	We claim $\mathcal{H} = 0$ for physical states. $\mathcal{H} = \lambda (H - E) = 0$.
-\end{justification}
-
-As an example, the Hamiltonian and extended Hamiltonian for a relativistic free particle the Hamiltonian are:
+$\mathcal{H}=\lambda (H - E)$
+
+	We claim $E = H(t,q^i,p_i)$ for physical states.
+$\lambda = d_s t = - \partial_E \mathcal{H} = - \partial_{E} \lambda (H - E + \mathcal{h}) - \lambda (-1)$ regardless of the choice of t and therefore $\lambda$.
+$H - E = 0$.
+
+	We claim $\bar{\mathcal{S}}$ is the level set for $\mathcal{H} = 0$.
+$\mathcal{H} = \lambda (H - E) = 0$ over $\bar{\mathcal{S}}$.
+Extend $\mathcal{H}$ over $T^* \mathcal{M}$ such that $\mathcal{H} \neq 0$ over $T^*\mathcal{M} \,\backslash\, \bar{\mathcal{S}}$.
+$\bar{\mathcal{S}}$ is the level set for $\mathcal{H} = 0$
+\end{justification}
+
+As an example, the Hamiltonian and invariant Hamiltonian for a relativistic free particle are:
 \begin{equation}\label{free_hamiltonians}
@@
 \end{equation}
-In the next section we'll derive and look more closely at the case of particles under potential forces.
+In the next section we'll derive the above expression and look more closely at the case of particles under potential forces.
 
@@
 \begin{prop}\label{relativistic_distribution_evolution}
-	Let $\rho_{\bar{\mathcal{c}}}(s)$ the distribution of material associated with the evolution $\bar{\mathcal{c}}$.
+	Let $\rho_{\bar{\mathcal{c}}}$ be the distribution of material associated with the evolution $\bar{\mathcal{c}}$.
 $\mathcal{H} \rho_{\bar{\mathcal{c}}} = 0$ and $d_s \rho_{\bar{\mathcal{c}}} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \} = 0$.
@@
 \begin{justification}
-	We claim $d_s \rho_\mathcal{c} (s) = \{ \rho_\mathcal{c}(s), \mathcal{H} \}$.
-Let $\rho_\mathcal{c} (s)$ be differentiable.
-$d_s \rho_\mathcal{c}(s) = \partial_a \rho_\mathcal{c}(s) d_s \xi^a = \partial_a \rho_\mathcal{c}(s) \omega^{ab} \partial_b \mathcal{H} = \{ \rho_\mathcal{c}(s), \mathcal{H} \}$.
-
 	We claim $\mathcal{H} \rho_{\bar{\mathcal{c}}} = 0 \; \forall \bar{\mathcal{c}} \in \bar{\mathcal{C}}$.
@@
 Let $\gamma : \mathbbm{R} \rightarrow T^*\mathcal{M}$ be a parametrization for a particle evolution $\bar{\mathcal{s}}$.
-Then $\rho_\mathcal{c}(\lambda(s)) = \rho_\mathcal{c}(\lambda(s + \Delta s))$.
+Then $\rho_\mathcal{c}(\lambda(s)) = \rho_\mathcal{c}(\lambda(s + \Delta s))$ as the material does not change over the particle evolution.
 $d_s \rho_\mathcal{c} = 0$.
@@
 This is usually missed because in classical mechanics only the standard Hamiltonian is typically used.
-Note how the extended Hamiltonian in \ref{free_hamiltonians} is quadratic in $E$, which can have both positive and negative solutions on the constraint $\mathcal{H}=0$.
+Note how the invariant Hamiltonian in \ref{free_hamiltonians} is quadratic in $E$, which can have both positive and negative solutions on the constraint $\mathcal{H}=0$.
 
@@
 Note how $\mathcal{H} \rho = 0$ and $\mathcal{H} \ket{\psi} = 0$ are formally equivalent.
-If we apply the extended Hamiltonian in \ref{free_hamiltonians} with the usual formal substitution, we have the Klein-Gordon equation using space-like convention rescaled by a factor for $2m$.
-As we'll see later, to this parallel in the mathematical formalism corresponds a parallel in the physics description.
+If we apply to the invariant Hamiltonian in \ref{free_hamiltonians} the usual formal classical/quantum substitution, we have the Klein-Gordon equation using space-like convention rescaled by a factor for $2m$.
+As we'll see later, to this parallel in the mathematical formalism corresponds a parallel in the physical description.
 
@@
 The complete state space $\bar{\mathcal{S}}$ is a hyper-surface of $T^*\mathcal{M}$.
-The extended Hamiltonian is needed to identify that subspace.
+The invariant Hamiltonian is needed to identify that subspace.
 This may be a hint that a better formulation is lurking.
@@
 The distribution of material would simply be a distribution over the particle evolutions, a Lebesgue integrable differentiable functions.
+Yet, the need to be able to describe the evolution in terms of physically meaningful state variables will most likely re-impose the extra structure.
 

### 2016-08-02 2015CompleteDerivation/completeDerivation.tex
Reviewed language for classical assumption

@@
 
-	The second, more conceptual, problem is that the idealizations that allow us to hold Assumption \ref{detrevass} break down as the parts get smaller.
-One such idealization is when the system under study is "big enough".
+	The second, more conceptual, problem is that the idealizations that allowed us to define states in Assumption \ref{detrevass} may not hold as the parts get smaller.
 Recall the cannonball whose motion is sufficiently unaffected by the photons that scatter off its surface.
-Clearly this breaks down for infinitesimal parts: as we keep dividing the material, we will find an amount that will be affected by the interaction with the environment, and therefore not sufficiently isolated.
-The other idealization is when the interaction between system and environment is not symmetrical.
+As we consider smaller and smaller parts, at some point we will find an amount of material that is affected by the interaction with the air or photons scattering off of it.
+At that point the parts are no longer sufficiently isolated to define an independent state, their evolution depends on the environment.
+Therefore the deterministic evolution of the whole cannot be reduced to the deterministic evolution of its parts.
+
+	%The other idealization is when the interaction between system and environment is not symmetrical.
 For example, suppose the electric charge of the system under study is much larger than the charges in the environment.
@@
 
-	In other words: the definition of a state requires some kind of asymmetry between system under study and environment.
-The environment should not affect the system, so that we can define an independent state, yet the system must affect the environment, so that we have external process that allows to define physical distinguishability.
-If we try to assume that both the system and environment are really made of infinitesimal homogeneous parts, then the claim to that asymmetry is lost.
-Therefore our justification for holding the assumption has to be that the pieces are small compared to some scale (they can be regarded as infinitesimal) but large compared to another (they are little affected by the environment).
+	Another issues is that defining a state requires some kind of asymmetry between system under study and environment.
+The system evolution is not affected by the state of the environment, so that we can define an independent state, yet the environment evolution is affected by the state of the system, so that we have external processes that allow us to define physical distinguishability.
+If we try to assume that both the system and environment are really made of the same classical material, then the claim to that asymmetry is lost: there is no difference in size, everything is made of infinitesimal pieces, and there is no difference in the laws of evolution, as the pieces are homogeneous.
+
+	Another issue arises if we include the probe in our description.
+Say $a \in A$ is the state of our system and $b \in B$ the state of our probe, the future state $\hat{a}=f(a)$ because the system is deterministic, but $\hat{b}=g(a,b)$ because the probe needs to be affected by the system.
+This means $(a, b) \mapsto (\hat{a}, \hat{b})$ is not a reversible map.
+If we want determinisms and reversibility for the combined system, we either have two isolated systems $\hat{a}=f(a)$ and $\hat{b}=g(b)$, which means the probe can't learn anything about the system, or two interacting systems $\hat{a}=f(a, b)$ and $\hat{b}=g(a, b)$, which means our system is not deterministic as its future state is determined by the state of the probe.
+In other words: we can't expect to have deterministic and reversible evolution at all levels of aggregations and have the parts interact in any physically meaningful way.
+
+	%Therefore our justification for holding the assumption has to be that the pieces are small compared to some scale (they can be regarded as infinitesimal) but large compared to another (they are little affected by the environment).
 For example, water is a homogeneous material if we consider it made up of parts sufficiently small for the scale of the problem at hand but big enough to contain a large number of molecules such that the average motion is sufficiently stable.
 
-	Other conceptual problems stem from the following two competing requirements: on one side we want the system under study to be isolated so that we can define independent states; on the other side we want the environment to not be isolated so that we can define physical processes that interact with the system.
+	%Other conceptual problems stem from the following two competing requirements: on one side we want the system under study to be isolated so that we can define independent states; on the other side we want the environment to not be isolated so that we can define physical processes that interact with the system.
 The problem is that the same physical object could be either under study or part of the environment depending on the situation.
@@
 We clearly can't have it both ways.
+
+	The same problem of not meaningful interaction surfaces when we consider the states of infinitesimal parts.
 Under our classical assumption, if we were to take the state of an infinitesimal part to really be its full description, with no unstated part, then each piece would evolve independently, oblivious to the other parts.
 Each particle would essentially reside in its own separate physical universe, as it cannot physically distinguish anything else.
-This cannot work.
+This is not physically meaningful.
 
 	If we assumed the system is deterministic and reversible only as a whole, then each part could evolve depending on the states of the other parts.
-This would seem much better, as we have pieces with no unstated part that are physically aware of each others.
+This would seem much better, as the state of the pieces is still exhaustive yet they are physically connected to each others.
 But this does not actually solve the problem of independent oblivious components.
-Let's ignore the question of how we were able to define the states of the parts.
-Let's also ignore that the type of forces one would typically use (i.e.
-ones that conserve energy or ones with scalar and vector potentials) are precisely the ones we derive by assuming the pieces are not interacting with each other.
-The actual problem is that we can always locally separate the evolution into independent degrees of freedom.
+First, for this to work, one would have to specify how the states of the parts were defined since their evolution is no longer deterministic (the future of each part depends on the state of other parts).
+One would also need to explain what type of interaction exists between the parts: simply using conservative forces, for example, would smell of a circular argument since those are precisely the force that we derive in the case where the pieces do evolve independently.
+But ignoring these two issue, the bigger problem is that we can always locally separate the evolution into independent degrees of freedom.
 For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value, they clearly evolve independently.
@@
 While such description may be cumbersome to achieve in practice, conceptually it is still possible.
-So, even if the particles are in principle interacting, the system is described by degrees of freedom that evolve independently, oblivious to each other.
-
-	The same type of problem arises on the whole system when we consider its interaction with our probe or other parts of the environment.
-Say $a \in A$ is the state of our system and $b \in B$ the state of our probe, the future state $\hat{a}=f(a)$ because the system is deterministic, but $\hat{b}=g(a,b)$ because the probe needs to be affected by the system.
-This means $(a, b) \mapsto (\hat{a}, \hat{b})$ is not a reversible map.
-In other words: we can't expect to have deterministic and reversible evolution at all levels of decomposable systems if the parts interact in any physically meaningful way.
-
-	The moral is that we should \emph{always} beware of thinking that the state of a system is its full description: there is always some unstated part that allows continuous interaction between the system and the environment.
-That is the part that cannot be isolated (i.e.
-we can't take the system out of the physical universe) and therefore can't be assigned a state.
-
-	One final conceptual problem is with time itself: a deterministic and reversible system cannot tell time.
+So, even if each particle evolution depends on the state of the others, the system is described by degrees of freedom that evolve independently, oblivious to each other.
+
+	The moral is that the classical idea of being able to assign to all systems a state which represents its full description does not work.
+The state is only what we can describe and it can't be the full description.
+As the division between system and environment is subjective, each system must be able to function as both.
+Therefore it will have a part whose evolution depends only on the system itself, the state, and a part whole evolution depends on other systems, the unstated part.
+While they may not be the same in all circumstances, they must exist.
+The state is what gives the identity to the system, its the subject of our discussion.
+The unstated part is what allows continuous interaction between the system and the environment, its what allows us to have the discussion.
+As we'll see later, it is precisely this problem that quantum system addresses better conceptually than classical mechanics.
+
+	One final problem is with time itself: an infinitesimally reducible deterministic and reversible system cannot tell time.
 If it did, we would be able to find a quantity that changes through time but is invariant under state variable changes.
@@
 
-	In light of what we discussed, we cannot take the classical assumption to be fundamental, in the sense that we cannot take it to strictly apply to all of universe.
+	In light of what we discussed, we cannot take the classical assumption to be fundamental, in the sense that we cannot take it to strictly apply to all of the universe.
 While ultimately flawed, the  assumption can be considered valid for a great number of macroscopic systems, and is also useful as a default assumption of sorts.

### 2016-08-03 2015CompleteDerivation/completeDerivation.tex
Reviewed classical assumption with Christine

@@
 To be able to talk about the states of a part we need a physical process that is able to distinguish between them.
-For a pool ball we can imagine marking one spot with a red marker.
+For a billiard ball we can imagine marking one spot with a red marker.
 This allows us to track it as the ball moves or collides with other balls.
@@
 
-	%The other idealization is when the interaction between system and environment is not symmetrical.
-For example, suppose the electric charge of the system under study is much larger than the charges in the environment.
-The electromagnetic field generated can be considered determined only by our system with good approximation.
-But this also breaks down: if each part of the environment is made of the same homogeneous material the charge density will be the same and the interaction between particles will be symmetrical.
-
-	Another issues is that defining a state requires some kind of asymmetry between system under study and environment.
-The system evolution is not affected by the state of the environment, so that we can define an independent state, yet the environment evolution is affected by the state of the system, so that we have external processes that allow us to define physical distinguishability.
+	Another issue is that defining a state requires some kind of asymmetry between system under study and environment.
+The future state of the system does not depend on the state of the environment, so that we can define an independent state, yet the future state of the environment is affected by the state of the system, so that we can have external processes that allow us to define physical distinguishability.
 If we try to assume that both the system and environment are really made of the same classical material, then the claim to that asymmetry is lost: there is no difference in size, everything is made of infinitesimal pieces, and there is no difference in the laws of evolution, as the pieces are homogeneous.
@@
 If we want determinisms and reversibility for the combined system, we either have two isolated systems $\hat{a}=f(a)$ and $\hat{b}=g(b)$, which means the probe can't learn anything about the system, or two interacting systems $\hat{a}=f(a, b)$ and $\hat{b}=g(a, b)$, which means our system is not deterministic as its future state is determined by the state of the probe.
-In other words: we can't expect to have deterministic and reversible evolution at all levels of aggregations and have the parts interact in any physically meaningful way.
-
-	%Therefore our justification for holding the assumption has to be that the pieces are small compared to some scale (they can be regarded as infinitesimal) but large compared to another (they are little affected by the environment).
-For example, water is a homogeneous material if we consider it made up of parts sufficiently small for the scale of the problem at hand but big enough to contain a large number of molecules such that the average motion is sufficiently stable.
-
-	%Other conceptual problems stem from the following two competing requirements: on one side we want the system under study to be isolated so that we can define independent states; on the other side we want the environment to not be isolated so that we can define physical processes that interact with the system.
-The problem is that the same physical object could be either under study or part of the environment depending on the situation.
-In other words: we want systems to be interacting so that they can be physically aware of each other, yet we want them not interacting so that we can properly define states on them.
-We clearly can't have it both ways.
-
-	The same problem of not meaningful interaction surfaces when we consider the states of infinitesimal parts.
+In other words: we can't expect to have deterministic and reversible evolution at all levels of aggregation and have the parts interact in any physically meaningful way.
+
+	The same problem of physically meaningless interaction surfaces when we consider the states of infinitesimal parts.
 Under our classical assumption, if we were to take the state of an infinitesimal part to really be its full description, with no unstated part, then each piece would evolve independently, oblivious to the other parts.
@@
 	If we assumed the system is deterministic and reversible only as a whole, then each part could evolve depending on the states of the other parts.
-This would seem much better, as the state of the pieces is still exhaustive yet they are physically connected to each others.
+This would seem much better, as the state of the pieces is still exhaustive yet they are physically connected to each other.
 But this does not actually solve the problem of independent oblivious components.
 First, for this to work, one would have to specify how the states of the parts were defined since their evolution is no longer deterministic (the future of each part depends on the state of other parts).
-One would also need to explain what type of interaction exists between the parts: simply using conservative forces, for example, would smell of a circular argument since those are precisely the force that we derive in the case where the pieces do evolve independently.
-But ignoring these two issue, the bigger problem is that we can always locally separate the evolution into independent degrees of freedom.
+One would also need to explain what type of interaction exists between the parts: simply using conservative forces, for example, would hint at a circular argument since those are precisely the forces that we derive in the case where the pieces do evolve independently.
+But ignoring these two issues, the bigger problem is that we can always locally separate the evolution into independent degrees of freedom.
 For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value, they clearly evolve independently.
@@
 
-	The moral is that the classical idea of being able to assign to all systems a state which represents its full description does not work.
+	The moral is that the classical idea of being able to assign to all systems a state which represents their full description does not work.
 The state is only what we can describe and it can't be the full description.
 As the division between system and environment is subjective, each system must be able to function as both.
-Therefore it will have a part whose evolution depends only on the system itself, the state, and a part whole evolution depends on other systems, the unstated part.
+Therefore it will have a part whose evolution depends only on the system itself, the state, and a part whose evolution depends on other systems, the unstated part.
 While they may not be the same in all circumstances, they must exist.
-The state is what gives the identity to the system, its the subject of our discussion.
-The unstated part is what allows continuous interaction between the system and the environment, its what allows us to have the discussion.
-As we'll see later, it is precisely this problem that quantum system addresses better conceptually than classical mechanics.
-
-	One final problem is with time itself: an infinitesimally reducible deterministic and reversible system cannot tell time.
+The state is what gives the identity to the system, the part we can study and describe.
+The unstated part is what allows continuous interaction between the system and the environment, it's what allows us to study and describe the state.
+As we'll see later, it is precisely this problem that quantum mechanics addresses better conceptually than classical mechanics.
+
+	One final problem is with time itself: an infinitesimally reducible system undergoing deterministic and reversible evolution cannot tell time.
 If it did, we would be able to find a quantity that changes through time but is invariant under state variable changes.
@@
 So, all quantities that are invariant under state variable changes are also invariant under time evolution.
+As the very definition of deterministic and reversible processes relies on time, this has to come from elsewhere (i.e.
+the environment).
 
@@
 
-\section{Elementary systems}
+\section{Kinematics and elementary systems}
 
@@
 In Newtonian physics, $\vec{F}=m\vec{a}$ provides such link.
-In Hamiltonian and Lagrangian mechanics, the link is implicitly given by the choice of the Hamiltonian and the Lagrangian.
-
-In this work we make the link conceptually explicit by assuming the following:
+In Hamiltonian (and Lagrangian) mechanics, the link is implicitly given by the choice of the Hamiltonian (or Lagrangian).
+
+Here we make the link conceptually explicit by assuming the following:
 
 \begin{assump}[Kinematic equivalence]\label{kinematic_equivalence}
-	Studying the kinematic of the system (i.e. trajectories in space) is equivalent to study the dynamics of the system (i.e. state trajectories in state space).
+	For the given system, studying its kinematics  (i.e. trajectories in space) is equivalent to study its dynamics  (i.e. state trajectories in state space).
 \end{assump}
@@
 \end{rationale}
+
+\subsection{Space-time manifold and Lagrangian mechanics}
+
+As we need to describe trajectory, we introduce $\mathcal{M}$ as the manifold that defines space and time.
+We call $x^i$ the spatial coordinates and $t$ the time coordinate.
+A trajectory will be given by a set of functions $x^i(s)$ and $t(s)$.
+What we need to understand is the link between these trajectories and the complete state space of the particles of a classical material.
+
+We first note that the system of units for the trajectories is fully defined by $\mathcal{M}$.
+Any trajectory can be specified once a set of coordinates $(t, x^i)$ is chosen.
+Under the kinematic assumption, these units must also be sufficient to define the dynamics.
+Therefore $\mathcal{M}$ has conceptually two functions: it is the space where the motion unfolds and it is the space that defines the system of units for our states.
+This already tells us that the our state space is a hyper-surface of $T^*\mathcal{M}$.
+
+Under the kinematic equivalence, given a trajectory we must identify a state and vice-versa.
+There must be a homeomorphism between the complete state space and the space of possible trajectories.
+We already know what the complete state space is.
+We just need to find a way to label the trajectories so that we can express the correspondence.
+
+Because of the double role of $\mathcal{M}$, we can set half of our state variables to be equal to the space-time variables: $(t, q^i) = (t, x^i)$.
+This already tells us that the trajectory is differentiable since $q^i(t)$ is differentiable.
+The four velocity $(u^t, u^i) = (d_s t, d_s x^i)$ is therefore well defined and it can be used to identify trajectories.
+The magnitude of the four velocity can be rescaled by changing the evolution parameter $s$, and is therefore not physical.
+Only the direction is physical so the four velocity gives us another $n$ independent variables to identify trajectories.
+The combined position and velocity, which we call \emph{initial conditions}, identify $2n + 1$ trajectories using a hyper-surface of the tangent bundle $T\mathcal{M}$.
+This is exactly the dimensionality of the complete state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$.
+We cannot add more curves, as a homeomorphism preserve dimensionality, therefore we found the suitable space for our trajectories.
+
+As $\omega(dt,dq^i,dE,dp_i)$ allows us to count possibilities in terms of state variable, we will have a metric $g(dt,dx^i,du^t,du^i)$ that allows us to count the possibilities in terms of initial conditions.
+Since the both $(dt,dx^i)$ and $(du^t,du^i)$ are four vectors and the metric must be linear, $g$ is a rank two tensor on $\mathcal{M}$.
+$g$ is also symmetric.
+Suppose $(t, x^i)$ are orthogonal directions, then a change of unit of one coordinate will only change the component of position and velocity for that coordinate.
+That is: the component of position and velocity form an independent d.o.f..
+For that coordinate system it must be $g(dx^i, du^j)=0=g(dx^j, du^i)$ for $i\neq j$.
+And the same must hold across temporal and standard d.o.f..
+As $g$ is a tensor, the symmetry applies to all choices of coordinates.
+
+Mathematically, under the kinematic assumption the state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is isomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$ of a pseudo-Riemannian manifold $(\mathcal{M}, g)$.
+
+\begin{prop}\label{riemannian_manifold}
+	The state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is isomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$ of a pseudo-Riemannian manifold $(\mathcal{M}, g)$.
+\end{prop}
+
+\begin{justification}
+	TODO
+\end{justification}
+
+\subsection{Scalar and vector potential forces}
 

### 2016-08-04 2015CompleteDerivation/completeDerivation.tex
First pass at kinematic assumption

@@
 
-As we need to describe trajectory, we introduce $\mathcal{M}$ as the manifold that defines space and time.
-We call $x^i$ the spatial coordinates and $t$ the time coordinate.
-A trajectory will be given by a set of functions $x^i(s)$ and $t(s)$.
+As we need to describe trajectories, we introduce $\mathcal{M}$ as the manifold that defines space and time.
+We call $x^i$ the spatial coordinates, $t$ the time coordinate and $x^\alpha = (t, x^i)$ their combination.
+A trajectory will be given by a set of functions $x^\alpha(s)$.
 What we need to understand is the link between these trajectories and the complete state space of the particles of a classical material.
@@
 We first note that the system of units for the trajectories is fully defined by $\mathcal{M}$.
-Any trajectory can be specified once a set of coordinates $(t, x^i)$ is chosen.
+Any trajectory can be specified once a set of coordinates $x^\alpha$ is chosen.
 Under the kinematic assumption, these units must also be sufficient to define the dynamics.
@@
 
-Because of the double role of $\mathcal{M}$, we can set half of our state variables to be equal to the space-time variables: $(t, q^i) = (t, x^i)$.
+Because of the double role of $\mathcal{M}$, we can set half of our state variables to be equal to the space-time variables: $(t, q^i) = x^\alpha$.
 This already tells us that the trajectory is differentiable since $q^i(t)$ is differentiable.
-The four velocity $(u^t, u^i) = (d_s t, d_s x^i)$ is therefore well defined and it can be used to identify trajectories.
-The magnitude of the four velocity can be rescaled by changing the evolution parameter $s$, and is therefore not physical.
+The four velocity $u^\alpha = d_s x^\alpha$ is therefore well defined and it can be used to identify trajectories.
+The magnitude of the four velocity can be rescaled by changing the evolution parameter, $u^\alpha = d_s x^\alpha = d_s \hat{s} \, d_{\hat{s}} x^\alpha = d_s \hat{s} \, \hat{u}^\alpha$, and is therefore not physical.
 Only the direction is physical so the four velocity gives us another $n$ independent variables to identify trajectories.
@@
 
-As $\omega(dt,dq^i,dE,dp_i)$ allows us to count possibilities in terms of state variable, we will have a metric $g(dt,dx^i,du^t,du^i)$ that allows us to count the possibilities in terms of initial conditions.
-Since the both $(dt,dx^i)$ and $(du^t,du^i)$ are four vectors and the metric must be linear, $g$ is a rank two tensor on $\mathcal{M}$.
-$g$ is also symmetric.
-Suppose $(t, x^i)$ are orthogonal directions, then a change of unit of one coordinate will only change the component of position and velocity for that coordinate.
-That is: the component of position and velocity form an independent d.o.f..
-For that coordinate system it must be $g(dx^i, du^j)=0=g(dx^j, du^i)$ for $i\neq j$.
-And the same must hold across temporal and standard d.o.f..
-As $g$ is a tensor, the symmetry applies to all choices of coordinates.
-
 Mathematically, under the kinematic assumption the state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is isomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$ of a pseudo-Riemannian manifold $(\mathcal{M}, g)$.
 
+\begin{prop}\label{initial_conditions}
+	Let $\mathcal{M}$ be the $n+1$ dimensional space-time manifolds over which trajectories are defined.
+The state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is isomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$.
+\end{prop}
+
+\begin{justification}
+	TODO
+\end{justification}
+
+This relationship allows us to write the initial conditions in terms of the state variable.
+In particular we have $u^\alpha=u^\alpha(t, q^i, E, p_i)$.
+As the relationship is invertible, these functions must be monotonic in $E$ and $p_i$.
+Since $\partial_{p_i} \mathcal{H}=d_s q^i= u^i$ and $- \partial_{E} \mathcal{H}= d_s t= u^0$, $\mathcal{H}$ is convex in $E$ and $p_i$.
+Under such conditions we can define a Legendre transformation $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$.
+Such transformation is the Lagrangian of the system and laws of motions are given by the Euler-Lagrange equation.
+
+\begin{prop}\label{lagrangian_mechanics}
+	Let $\mathcal{H}$ be an extended Hamiltonian defined on $\mathbf{T}^*\mathcal{M}$.
+Under the kinematic assumption, we can define the Legendre transform $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$ which we call the \emph{extended Lagrangian}.
+For the time invariant case, we define the \emph{Lagrangian} $L=v^i p_i - H$.
+\end{prop}
+
+\begin{justification}
+	TODO
+\end{justification}
+
+Being able to convert the state variable for the particles into initial conditions: we also need to be able to express the distributions of the composite states over position and velocity.
+To do that, we need to be able to count trajectories in the same way we are able to count states.
+There must exist a metric $g$, consistent with $\omega$, that allows to calculate the cardinality of possibilities for initial conditions.
+That is: $\omega(dt,dq^i,dE,dp_i) \propto g(dx^\alpha,du^\alpha)$.
+
+Since the both $dx^\alpha$ and $du^\alpha$ are four vectors and the metric must be linear, $g$ is a rank two tensor on $\mathcal{M}$.
+This means that a each direction in space-time $dx^\alpha$ is associated with a direction in the space of velocities $du^\alpha$ to form an independent d.o.f.
+Across independent d.o.f.
+we have $\omega(dq^i, dp_j) = 0 =\omega(dq^j, dp_i)$.
+Therefore if $dx^i$ and $dx^j$ do not form a d.o.f., $g(dx^i, du^j) = 0 = g(dx^j, du^i)$: $g$ is reflexive.
+We also know that $g$ cannot be alternating, that is $g(dx^i, du^i) \neq 0$: a choice of units for $x^i$ is a choice of units for $u^i$ and therefore they cannot belong to different d.o.f.
+This means $g$ is symmetric.
+$\mathcal{M}$ is a pseudo-Riemannian manifold.
+
+Since $g$ is symmetric, it can be diagonalized: $\omega= dq^i \wedge dp_i - dt \wedge dE \propto g = dx^i g_{ii} du^i + dt g_{00} du^0$.
+This tells us $g_{ii} > 0$ and $g_{00} < 0$.
+We can further change the units along the spatial directions so that $g_{ii}=1$: if we use the same units for the spatial coordinates, each direction will contribute the same number of possibilities.
+Since time must use a different unit, we set $g_{00}=-c^2$.
+Therefore, locally $g(dx^i, dx^i) = (dx^i)^2 - c^2 dt^2$ is Minkowskian.
+
 \begin{prop}\label{riemannian_manifold}
-	The state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is isomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$ of a pseudo-Riemannian manifold $(\mathcal{M}, g)$.
+	The space-time manifold $\mathcal{M}$ is a pseudo-Riemannian manifold $(\mathcal{M}, g)$ and locally Mikowskian. That is, locally $g = (dx^i)^2 - c^2 dt^2$.
 \end{prop}
@@
 
-\subsection{Scalar and vector potential forces}
+The metric tensor $g$ and the constant $c$ found here are indeed the ones used in the context of relativity.
+Yet, they have acquired a different meaning.
+$g$ does not measure space-time distances: it measure cardinality of possibilities, like $\omega$.
+Directions in space are orthogonal not because the angle between them is $90^\circ$, but because they identify independent d.o.f..
+Space and time, instead, are not independent and the geometry between them is different.
+In other words: $g$ inherits it's properties from $\omega$.
+
+This also allows to give $c$ a more insightful physical meaning.
+It is not a velocity: the metric is defined at each point without motion.
+It is the constant of proportionality that maps a range of initial condition from the time variable $t$ to the space variable $x^i$.
+In the same way that $1$km would correspond to a thousand times more initial conditions than $1$m, $1$sec corresponds to $299,792,458$ more initial conditions than $1$m.
+Therefore when we write $x^\alpha = (ct, x^i)$ or $p_\alpha = (- E/c, p_i)$ we are really expressing everything in units that are proportional to the cardinality of state and possibilities.
+Naturally, these units make the book keeping of number of states much easier which, as we saw, is the only thing that Hamiltonian and Lagrangian mechanics are doing.
+
+This also allows to understand conceptually why the speed of a deterministic and reversible signal cannot be greater than $c$.
+Consider the trajectory of a point particle.
+Over an interval $dt$ it will travel $dx$ in space.
+Since the motion is deterministic and reversible, it will travel through all the states defined in the interval $dx$.
+It will also travel through every moment in time, which corresponds to $cdt$ possibilities.
+The number of states that will travel in space cannot exceed the number of states it travels in time as it can only be at one position for each instant.
+Therefore we have:
+\begin{align*}
+\frac{|n_x(dx)|}{|n_t(dt)|} &=  \frac{|dx|}{|c dt|} \leq 1 \\
+\frac{|dx|}{|dt|} & \leq c \\
+\end{align*}
+Note, though, that this limit applies to deterministic and reversible motion.
+It was under that premise that we were able to show that the trajectories are continuous, that we could define $\omega$ and therefore $g$.
+Non-local quantum effects, such as the ones features in EPR experiments, are not examples of deterministic and reversible motion.
+Therefore they do not have anything to violate: the maximum speed, simply put, does not apply.
+
+\subsection{Massive particles under scalar and vector potential forces}
+
+The link between $\omega$ and $g$ allows us to constrain the kind of homemorphism between state variables and initial conditions.
+If we express the possibilities in terms of both state variables and initial conditions we have: $dq^\alpha \wedge dp_\alpha \propto dx^\alpha du_\alpha$.
+Since  $p_\alpha=p_\alpha(x, u)$, we have $dp_\alpha = m du_\alpha$, where $m$ is the constant of proportionality.
+We recognize $m$ as the inertial mass.
+Integrating the expression, we have $p_\alpha = m u_\alpha + A_\alpha(x)$ where $A_\alpha(x)$ is an arbitrary function.
+This is the most general link between velocity and conjugate momentum, which is the sum of kinetic momentum and a scalar potential.
+
+This constrains the form of the extended Hamiltonian.
+$\frac{1}{m}g^{\alpha\beta}(p_\beta-A_\beta) = u^\alpha = d_s x^\alpha = d_s q^\alpha = \partial_{p_\alpha} \mathcal{H}$.
+Integrating we have $\mathcal{H}=\frac{1}{2m}(p_\alpha-A_\alpha(x))g^{\alpha\beta}(p_\beta-A_\beta(x))+V(x)$ where $V(x)$ is an arbitrary function.
+We can further constrain the Hamiltonian by choosing the natural parameterization for $s$ (also called unit speed or arc length parametrization) for which the modulus of the velocity is constant.
+We have $u^\alpha u_\alpha = - c^2 = \frac{1}{m^2}(p^\alpha-A^\alpha)(p_\alpha-A_\alpha)$.
+$\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha(x)g^{\alpha\beta}(p_\beta-A_\beta(x) + m^2 c^2)$.
+Applying the same procedure in the time independent case yields the Hamiltonian $H=\frac{1}{2m}(p^i-A^i(x))(p_i-A_i(x))+V(x)$.
+
+We recognize the Hamiltonian for massive particles under scalar and vector potential forces.
+
+\begin{prop}\label{kineticHamiltonian}
+The extended Hamiltonian for the particles of a classical material under the kinematic assumption is $\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta) + m^2 c^2)$.
+\end{prop}
+
+\begin{justification}
+	TODO
+\end{justification}
+
+The above discussion gives the inertial mass a new physical meaning: it is the conversion constant that tells how to convert a range of velocity to a range of conjugate momentum with the same number of possibilities.
+In other words: the more massive the particles, the more states per unit velocity.
+Intuitively, a more massive body is harder to accelerate because it has to go through more states for the same change in velocity.
+Note that the Hamiltonian excludes massless particles and this makes absolute sense.
+A massless particle always has the same velocity, $c$, therefore the trajectory is not enough to tell us the dynamics.
+The motion of the photon is not enough to tell us its energy or its momentum.
+Therefore it makes sense that, by assuming kinematic equivalence we implicitly required massive particles.
+
+We have seen that the inertial mass $m$, the invariant speed $c$ and the reduced Plank constant $\hbar$ all serve the same purpose: convert one quantity to another preserving the number of possibilities.
+$dp=\hbar dk$ converts between ranges of wave number and conjugate momentum.
+$dx = c dt$ converts between ranges of space and time.
+$dp = m du$ converts between ranges of velocity and conjugate momentum.
+They allow to keep proper book keeping when mapping sets of states.
+There is a difference, though: while $c$ and $\hbar$ are simply determined by the unit system, $m$ also depends on the type of material.
+That is, different homogeneous materials can have different inertial mass.
+
+Another important thing to notice is that the relationship between conjugate momentum and velocity, $p_\alpha = m u_\alpha + A_\alpha(x)$, includes an arbitrary function.
+As it is arbitrary, we cannot give $p$ a well defined physical meaning in terms of the space-time trajectory.
+Suppose, in fact, we changed $A$ and $p$ while leaving $u$ unchanged, the physical description in terms of trajectories remains the same.
+Therefore we should ask whether any choice of $A$ actually represent a different case leading to different trajectories.
+
+The simplest and more insightful way to study this problem is simply to use $x^\alpha$ and $u_\alpha$ as state variable.
+They are not going to be conjugates of each other but they do capture the entire physical description.
+Any change in $A$ that does not affect $x^\alpha$ and $mu_\alpha$ does not change the physics.
+Expressing the Hamiltonian and the Poisson brackets in terms of the initial conditions we have:
+\begin{align*}
+\mathcal{H}&=\frac{1}{2m}(mu_\alpha mu^\alpha + m^2 c^2) \\
+\{x^\alpha, x^\beta\} &= 0 \\
+\{x^\alpha, mu_\beta\} &= \delta^\alpha_\beta \\
+\{mu_\alpha, mu_\beta\} &= \partial_\alpha A_\beta - \partial_\beta A_\alpha = F_{\alpha\beta}
+\end{align*}
+Note how the Poisson bracket between kinetic momentum is the force tensor.
+The weaker the force, the more $(x^i, mu^i)$ become independent d.o.f.
+The force tensor, then, tells how much $(x^i, mu^i)$ fail to be independent d.o.f..
+With these expressions, the commutator between any function of position and kinetic momentum can now be calculated.
+For example:
+\begin{align*}
+d_s x^\alpha &= \{x^\alpha, \mathcal{H}\} = u^\alpha \\
+d_s (mu^\alpha) &= \{mu^\alpha, \mathcal{H}\} = F^{\alpha\beta} u_\beta
+\end{align*}
+where we recognize the covariant form of the Lorentz force.
+
+Any change of $A$ that leaves $F_{\alpha\beta}$ unchanged does not affect the motion of any function of position and velocity. Such transformation is what we call a gauge transformation.
+
+As we mentioned before, as the extended Hamiltonian is quadratic in the energy, negative energy states are possible.
+As we fixed the natural parametrization, $u^\alpha u_\alpha = - c^2$, there are two cases: the parametrization is aligned with time, therefore $s=\tau$ where $\tau$ is proper time, or the parametrization is anti-aligned with time, therefore $s=-\tau$.
+The trajectory velocity $u^\alpha= \pm d_\tau x^\alpha$ is plus or minus the physical velocity depending on whether it's a standard or anti-particle state.
+Expanding the equation for the force, we have:
+\begin{align*}
+\pm d_\tau (\pm m d_\tau x^\alpha) &= m d_\tau^2 x^\alpha = \pm  F^\alpha_\beta d_\tau x^\beta
+\end{align*}
+Therefore the force act on anti-particle states in the opposite way: they act as they have opposite charge. Note that nothing physical is actually going backwards in time, just the parametrization.
+
+In this setting, charge conjugation $C$ is simply changing the parametrization from $s$ to $-s$, following the trajectories in the opposite way.
+This will permute each particle anti-state with the corresponding standard state and vice-versa.
+But if we apply parity $P$ and time reversal $T$, changing $x^\alpha$ to $-x^\alpha$, we would have the same net effect: the parametrization stays the same, but it now moves backwards in space and time.
+As they have the same net effect, if we apply all three we go back where we started.
+This gives us a very intuitive classical particle mechanics analogue for the $CPT$ theorem.
+
+To sum up, the kinematic assumption has more inherent consequences one would expect at first.
+It means the particles are identified by position and velocity, the system of equation is second order.
+It links the space of initial conditions on $T\mathcal{M}$ with the space of states on $T^*\mathcal{M}$.
+Such link allows state space metric $\omega$ to induce the space-time metric $g$.
+It allows us to define Lagrangians.
+It constrains the motion to massive particles under scalar and vector potential forces.
+Moreover, it allowed to re-understand basic concepts, like invariant speed $c$, the inertial mass $m$ and the force tensor $F_{\alpha\beta}$ in terms of state counting and d.o.f.
+independence.
+
+We now have all the building block of classical mechanics, the consequences of three seemingly simple assumption.
+But that's not all: we have placed ourselves in a very good position to move forward to quantum systems as well by using the new physical insights we have gained.
 

### 2016-08-05 2015CompleteDerivation/completeDerivation.tex
Second pass at kinematcs

@@
 
-Reductionism by itself would not help us describing a system if the state and the evolution of the pieces were more complicated to study.
-In this section we further focus our attention to those system that not only are infinitesimally reducible, but their internal dynamics can be ignored.
-That is, we will assume that the motion in space is necessary and sufficient to reconstruct the state and its dynamics.
+We will now focus our attention to the simplest possible classical material: one for which the motion in space is necessary and sufficient to reconstruct the state and its evolution.
+That is, the internal dynamics can be ignored.
 What we'll find is that the dynamics of each infinitesimal part follows classical Lagrangian mechanics.
@@
 
-The link between kinematics, the study of motion through quantities like velocity and acceleration, and dynamics, the study of laws of evolution, is crucial in physics.
+The link between kinematics, the study of motion through quantities like velocity and acceleration, and dynamics, the study of state evolution and its causes, is crucial in physics.
 In Newtonian physics, $\vec{F}=m\vec{a}$ provides such link.
@@
 
-Here we make the link conceptually explicit by assuming the following:
+Here we make the link conceptually explicit with the following assumption:
 
 \begin{assump}[Kinematic equivalence]\label{kinematic_equivalence}
-	For the given system, studying its kinematics  (i.e. trajectories in space) is equivalent to study its dynamics  (i.e. state trajectories in state space).
+	For the given classical material, studying its kinematics  (i.e. trajectories in space) is equivalent to study its dynamics  (i.e. states and their evolution).
 \end{assump}
@@
 
-	The assumption clearly cannot be taken in general since many systems do not posses this trait.
-Yet, as this assumption allows to recover the standard link between kinematic and dynamical quantities, there is something fundamental about it.
-That is: if we assume that the system has no discernible internal dynamics, then the only dynamics that needs to be described is the one associated with its trajectory in space: kinematic equivalence holds.
-
-	This is in fact consistent with observation, though elementary particles are quantum fields therefore some caveats are in order.
-Massless particles should be excluded as they cannot be localized, which makes it hard to talk about a proper trajectory.
-Spin is not associated to kinematic quantities: nothing is thought to be actually spinning.
-As such we can regard it as the internal dynamic of the system.
-A massive spin $0$ system is one that can be thought as localized and as having no internal structure.
-For such system we would expect kinematic equivalence to hold and it does: from kinematic momentum we can reconstruct conjugate momentum and vice-versa.
-
-	Therefore, with the appropriate caveat that we are talking about a quantum particle which we haven't yet introduced, the simplest systems are the ones that just describe a trajectory in space.
-This raises the opposite question: why all systems have \emph{at least} to describe a trajectory in space?
+	The assumption clearly cannot be taken in general.
+Not all material will be elementary in the sense that they may have some internal dynamics.
+That is: position in space is not their only state variable.
+Yet, the assumption is fundamental because what we learn about the link between kinematic and dynamical quantities applies to all materials, even the non elementary ones.
+
+	The first natural question to ask is: why do all systems have \emph{at least} to describe a trajectory in space?
 This seems motivated by the previous assumption of reducibility: spatial extent is the one way that all materials are reduced.
-Therefore all particles of all materials, regardless of whether they are elementary or not, are will include spatial degrees of freedom.
+Therefore the description of all particles of all materials, regardless of whether they are elementary or not, will include spatial degrees of freedom.
+Kinematic variables will form a set of independent degrees of freedom that are universal.
+
+	When the system is not elementary, the particles will be described by the kinematic variables plus some additional independent state variables.
+Yet, the set of all possible evolutions will include those for which kinematic variables evolve independently from the other.
+For example, the evolutions for which the internal d.o.f.
+do not change.
+In other words: the overall dynamics will be the spatial degrees of freedom coupled with the internal degrees of freedom.
+Therefore by studying an elementary material, we are also studying a component of the dynamics that is present in all materials.
 
 	It should be also noted that, once we reduced particles to be infinitesimal at a point, the rest of the description must be a local object invariant under coordinate transformations.
-This means that the remaining internal dynamics must be identified scalar, vector or tensor quantities.
-Spin is, in some sense, a quantized version of those.
-
-	While we can make an argument why kinematic equivalence is important because it is valid for elementary system, we cannot actually make an argument why nature should be reducible to elementary system.
-While it fits our physical intuition, we have no a priori argument for it.
+This means that the remaining internal dynamics must be identified by scalar, vector or tensor quantities.
+This may be used in the future to justify the extension of this work to field theories.
+
+	To be clear: we are not saying that kinematic equivalence is fundamental because everything must reduce to a material with no internal dynamics.
+While this statement seems reasonable and fits our physical intuition, we have no a priori argument for it.
 We can imagine, in fact, that as we reduce the behavior of one material to its constituents we find the parts to possess richer and richer states.
 For example, most macroscopic object are neutral though the parts are charged.
-We could imagine that, while dividing into parts, we find more and more types of charges and forces.
+We could imagine that, while dividing into parts, we find more and more types of charges, forces and internal dynamics.
 We do not have an argument why this cannot happen.
+
+	What we are saying is that any classical material is \emph{at least} an elementary material and will have posses the dynamic of one coupled with the internal dynamics.
 \end{rationale}
@@
 
-As we need to describe trajectories, we introduce $\mathcal{M}$ as the manifold that defines space and time.
-We call $x^i$ the spatial coordinates, $t$ the time coordinate and $x^\alpha = (t, x^i)$ their combination.
-A trajectory will be given by a set of functions $x^\alpha(s)$.
-What we need to understand is the link between these trajectories and the complete state space of the particles of a classical material.
+As we need to describe trajectories, we introduce $\mathcal{M}$ as the $n+1$ dimensional manifold that defines space and time.
+A point in $\mathcal{M}$ will be identified by $x^\alpha = (t, x^i)$, its time and space coordinates.
+A trajectory will be given by a set of functions $x^\alpha(s)$ where $s$ is an arbitrary parametrization of the curve.
+What we need to understand is the link between the state space of all the possible trajectories and the complete state space of the particles of a classical material.
 
@@
 Any trajectory can be specified once a set of coordinates $x^\alpha$ is chosen.
-Under the kinematic assumption, these units must also be sufficient to define the dynamics.
-Therefore $\mathcal{M}$ has conceptually two functions: it is the space where the motion unfolds and it is the space that defines the system of units for our states.
+Under the kinematic assumption, these units must also be sufficient to describe the dynamics.
+Therefore $\mathcal{M}$ has conceptually two roles: it is the space where the motion unfolds and it is the space that defines the system of units for our states.
 This already tells us that the our state space is a hyper-surface of $T^*\mathcal{M}$.
@@
 
-Because of the double role of $\mathcal{M}$, we can set half of our state variables to be equal to the space-time variables: $(t, q^i) = x^\alpha$.
+Because of the double role of $\mathcal{M}$, we can set half of our state variables to be equal to the space-time variables: $q^\alpha = (t, q^i) = x^\alpha$.
 This already tells us that the trajectory is differentiable since $q^i(t)$ is differentiable.
@@
 This is exactly the dimensionality of the complete state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$.
-We cannot add more curves, as a homeomorphism preserve dimensionality, therefore we found the suitable space for our trajectories.
-
-Mathematically, under the kinematic assumption the state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is isomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$ of a pseudo-Riemannian manifold $(\mathcal{M}, g)$.
+We cannot add more curves, as a homeomorphism preserves dimensionality, therefore we found the suitable space for the trajectories of particles of a classical material under the kinematic assumption.
+
+Being able to identify particle through both state variables and initial conditions is not enough: we also need to be able to express the distributions of the composite states over position and velocity.
+As such, the differentiable distribution over state variables must become a differentiable distribution over initial condition.
+This tells us the transformation is differentiable.
+
+Mathematically, the state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is diffeomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$ of a pseudo-Riemannian manifold $(\mathcal{M}, g)$.
 
 \begin{prop}\label{initial_conditions}
-	Let $\mathcal{M}$ be the $n+1$ dimensional space-time manifolds over which trajectories are defined.
-The state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is isomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$.
+	Let $\mathcal{M}$ be the $n+1$ dimensional space-time manifold over which trajectories are defined.
+The state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is diffeomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$.
 \end{prop}
@@
 
-Being able to convert the state variable for the particles into initial conditions: we also need to be able to express the distributions of the composite states over position and velocity.
+Not only we must be able to express distributions in terms of both state variable and initial conditions, but integration has to carry over as well.
 To do that, we need to be able to count trajectories in the same way we are able to count states.
@@
 Since the both $dx^\alpha$ and $du^\alpha$ are four vectors and the metric must be linear, $g$ is a rank two tensor on $\mathcal{M}$.
-This means that a each direction in space-time $dx^\alpha$ is associated with a direction in the space of velocities $du^\alpha$ to form an independent d.o.f.
+To each direction in space-time $dx^\alpha$ is associated a direction in the tangent space of velocities $du^\alpha$ with which it forms an independent d.o.f..
 Across independent d.o.f.
 we have $\omega(dq^i, dp_j) = 0 =\omega(dq^j, dp_i)$.
-Therefore if $dx^i$ and $dx^j$ do not form a d.o.f., $g(dx^i, du^j) = 0 = g(dx^j, du^i)$: $g$ is reflexive.
+Therefore if $dx^i$ and $du^j$ do not form a d.o.f., $g(dx^i, du^j) = 0 = g(dx^j, du^i)$: $g$ is reflexive.
 We also know that $g$ cannot be alternating, that is $g(dx^i, du^i) \neq 0$: a choice of units for $x^i$ is a choice of units for $u^i$ and therefore they cannot belong to different d.o.f.
@@
 We can further change the units along the spatial directions so that $g_{ii}=1$: if we use the same units for the spatial coordinates, each direction will contribute the same number of possibilities.
-Since time must use a different unit, we set $g_{00}=-c^2$.
+Since time must use a different unit, we set $g_{00}=-c^2$ where $c$ is the constant that allows to change an range in time to a range in space of equal cardinality of possibilities (i.e.
+that counts the same number of states).
 Therefore, locally $g(dx^i, dx^i) = (dx^i)^2 - c^2 dt^2$ is Minkowskian.
@@
 Space and time, instead, are not independent and the geometry between them is different.
-In other words: $g$ inherits it's properties from $\omega$.
+In other words: $g$ inherits it's signature from $\omega$.
 
@@
 \end{align*}
-Note, though, that this limit applies to deterministic and reversible motion.
+Note, though, that this constraint applies only to deterministic and reversible motion.
 It was under that premise that we were able to show that the trajectories are continuous, that we could define $\omega$ and therefore $g$.
 Non-local quantum effects, such as the ones features in EPR experiments, are not examples of deterministic and reversible motion.
-Therefore they do not have anything to violate: the maximum speed, simply put, does not apply.
+Therefore they do not have anything to violate: the maximum speed does not apply to them.
 
@@
 
-The link between $\omega$ and $g$ allows us to constrain the kind of homemorphism between state variables and initial conditions.
+The link between $\omega$ and $g$ allows us to constrain the kind of diffeomorphism between state variables and initial conditions.
 If we express the possibilities in terms of both state variables and initial conditions we have: $dq^\alpha \wedge dp_\alpha \propto dx^\alpha du_\alpha$.
@@
 $\frac{1}{m}g^{\alpha\beta}(p_\beta-A_\beta) = u^\alpha = d_s x^\alpha = d_s q^\alpha = \partial_{p_\alpha} \mathcal{H}$.
-Integrating we have $\mathcal{H}=\frac{1}{2m}(p_\alpha-A_\alpha(x))g^{\alpha\beta}(p_\beta-A_\beta(x))+V(x)$ where $V(x)$ is an arbitrary function.
+Integrating we have $\mathcal{H}=\frac{1}{2m}(p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta)+V(x)$ where $V(x)$ is an arbitrary function.
 We can further constrain the Hamiltonian by choosing the natural parameterization for $s$ (also called unit speed or arc length parametrization) for which the modulus of the velocity is constant.
 We have $u^\alpha u_\alpha = - c^2 = \frac{1}{m^2}(p^\alpha-A^\alpha)(p_\alpha-A_\alpha)$.
-$\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha(x)g^{\alpha\beta}(p_\beta-A_\beta(x) + m^2 c^2)$.
-Applying the same procedure in the time independent case yields the Hamiltonian $H=\frac{1}{2m}(p^i-A^i(x))(p_i-A_i(x))+V(x)$.
+$\mathcal{H}=\frac{1}{2m}((p^\alpha-A^\alpha)(p_\alpha-A_\alpha) + m^2 c^2)$.
+Applying the same procedure in the time independent case yields the Hamiltonian $H=\frac{1}{2m}(p^i-A^i)(p_i-A_i)+V$.
 
@@
 
-The above discussion gives the inertial mass a new physical meaning: it is the conversion constant that tells how to convert a range of velocity to a range of conjugate momentum with the same number of possibilities.
+The above discussion gives the inertial mass a new physical meaning: it is the conversion constant that tells us how to convert a range of velocity to a range of conjugate momentum with the same number of possibilities.
 In other words: the more massive the particles, the more states per unit velocity.
 Intuitively, a more massive body is harder to accelerate because it has to go through more states for the same change in velocity.
-Note that the Hamiltonian excludes massless particles and this makes absolute sense.
+Note that the Hamiltonian excludes massless particles because the kinematic assumption does that implicitly.
 A massless particle always has the same velocity, $c$, therefore the trajectory is not enough to tell us the dynamics.
@@
 $dp=\hbar dk$ converts between ranges of wave number and conjugate momentum.
-$dx = c dt$ converts between ranges of space and time.
+$dx = c dt$ converts between ranges of time and space.
 $dp = m du$ converts between ranges of velocity and conjugate momentum.
 They allow to keep proper book keeping when mapping sets of states.
+Using natural units means measuring all these quantities through the cardinality of possibilities.
 There is a difference, though: while $c$ and $\hbar$ are simply determined by the unit system, $m$ also depends on the type of material.
@@
 Suppose, in fact, we changed $A$ and $p$ while leaving $u$ unchanged, the physical description in terms of trajectories remains the same.
-Therefore we should ask whether any choice of $A$ actually represent a different case leading to different trajectories.
-
-The simplest and more insightful way to study this problem is simply to use $x^\alpha$ and $u_\alpha$ as state variable.
+
+We should also ask whether any choice of $A$ actually represent a different case leading to different trajectories.
+The simplest and more insightful way to study this problem is to use $x^\alpha$ and $mu_\alpha$ as state variable.
 They are not going to be conjugates of each other but they do capture the entire physical description.
-Any change in $A$ that does not affect $x^\alpha$ and $mu_\alpha$ does not change the physics.
+Any change in $A$ that does not affect the relationship between $x^\alpha$ and $mu_\alpha$ does not change the physics.
 Expressing the Hamiltonian and the Poisson brackets in terms of the initial conditions we have:
@@
 Note how the Poisson bracket between kinetic momentum is the force tensor.
-The weaker the force, the more $(x^i, mu^i)$ become independent d.o.f.
-The force tensor, then, tells how much $(x^i, mu^i)$ fail to be independent d.o.f..
+The weaker the force, the more $(x^i, mu_i)$ become independent d.o.f.
+The force tensor, then, tells how much $(x^i, mu_i)$ fail to be independent d.o.f..
 With these expressions, the commutator between any function of position and kinetic momentum can now be calculated.
@@
 
-Any change of $A$ that leaves $F_{\alpha\beta}$ unchanged does not affect the motion of any function of position and velocity. Such transformation is what we call a gauge transformation.
-
-As we mentioned before, as the extended Hamiltonian is quadratic in the energy, negative energy states are possible.
+Any change of $A$ that leaves $F_{\alpha\beta}$ unchanged does not affect the motion of any function of position and velocity.
+Such transformation is what we call a gauge transformation.
+Note the same group of transformations is found by fixing $A$ and changing $p$ such that the laws of motion don't change.
+In other words: $A$ and $p$ are defined relatively to the each other.
+$A$ defines the $0$ for $p$ and vice versa.
+The gauge group expresses the arbitrariness of such definition.
+
+The extended Hamiltonian we found is quadratic in energy, therefore negative energy states are possible.
+These correspond to the particle anti-states mentioned in \ref{antistates}.
 As we fixed the natural parametrization, $u^\alpha u_\alpha = - c^2$, there are two cases: the parametrization is aligned with time, therefore $s=\tau$ where $\tau$ is proper time, or the parametrization is anti-aligned with time, therefore $s=-\tau$.
-The trajectory velocity $u^\alpha= \pm d_\tau x^\alpha$ is plus or minus the physical velocity depending on whether it's a standard or anti-particle state.
+The trajectory velocity $u^\alpha= \pm d_\tau x^\alpha$ is plus or minus the physical velocity depending on whether it's a standard particle state or an anti-state.
 Expanding the equation for the force, we have:
@@
 \end{align*}
-Therefore the force act on anti-particle states in the opposite way: they act as they have opposite charge. Note that nothing physical is actually going backwards in time, just the parametrization.
+Therefore the force act on anti-particle states in the opposite way: they act as if they have opposite charge. Note that nothing physical is actually going backwards in time, just the parametrization.
 
@@
 To sum up, the kinematic assumption has more inherent consequences one would expect at first.
-It means the particles are identified by position and velocity, the system of equation is second order.
+It means the particles are identified by position and velocity and therefore the system of equation is second order.
 It links the space of initial conditions on $T\mathcal{M}$ with the space of states on $T^*\mathcal{M}$.
@@
 It constrains the motion to massive particles under scalar and vector potential forces.
-Moreover, it allowed to re-understand basic concepts, like invariant speed $c$, the inertial mass $m$ and the force tensor $F_{\alpha\beta}$ in terms of state counting and d.o.f.
+Moreover, it allowed to re-understand basic concepts, like the invariant speed $c$, the inertial mass $m$ and the force tensor $F_{\alpha\beta}$ in terms of state counting and d.o.f.
 independence.
@@
 We now have all the building block of classical mechanics, the consequences of three seemingly simple assumption.
-But that's not all: we have placed ourselves in a very good position to move forward to quantum systems as well by using the new physical insights we have gained.
+But that's not all: we have also placed ourselves in a very good position to move forward to quantum systems by using the new physical insights we have gained.
 

### 2016-08-09 2015CompleteDerivation/completeDerivation.tex
Fix measurement device argument

@@
 
-	Another issue arises if we include the probe in our description.
-Say $a \in A$ is the state of our system and $b \in B$ the state of our probe, the future state $\hat{a}=f(a)$ because the system is deterministic, but $\hat{b}=g(a,b)$ because the probe needs to be affected by the system.
+	Another issue arises if we include the measuring device in our description.
+Say $a \in A$ is the state of our system and $b \in B$ the state of our measuring device, the future state $\hat{a}=\hat{a}(a)$ because the system is deterministic, $\hat{b}=\hat{b}(a,b)$ because the measuring device needs to be affected by the system and $a=f(\hat{b})$ because we should be able to know the original state by looking at the measuring device.
 This means $(a, b) \mapsto (\hat{a}, \hat{b})$ is not a reversible map.
-If we want determinisms and reversibility for the combined system, we either have two isolated systems $\hat{a}=f(a)$ and $\hat{b}=g(b)$, which means the probe can't learn anything about the system, or two interacting systems $\hat{a}=f(a, b)$ and $\hat{b}=g(a, b)$, which means our system is not deterministic as its future state is determined by the state of the probe.
+We would have $b = b(\hat{a}(f(\hat{b})), \hat{b}) = b(\hat{b})$ and $a=f(\hat{b})$: the whole past state cannot be reconstructed just by looking at the future state of the measuring device.
+If we want determinisms and reversibility for the combined system, either both systems $\hat{a}=\hat{a}(a, b)$ and $\hat{b}=\hat{b}(a, b)$ are not independently deterministic or the second is not a measuring device $a \neq f(\hat{b})$.
 In other words: we can't expect to have deterministic and reversible evolution at all levels of aggregation and have the parts interact in any physically meaningful way.

### 2016-08-09 2015CompleteDerivation/completeDerivation.tex
Removing mention that conservative force and circular arguments
Too distracting

@@
 But this does not actually solve the problem of independent oblivious components.
-First, for this to work, one would have to specify how the states of the parts were defined since their evolution is no longer deterministic (the future of each part depends on the state of other parts).
-One would also need to explain what type of interaction exists between the parts: simply using conservative forces, for example, would hint at a circular argument since those are precisely the forces that we derive in the case where the pieces do evolve independently.
-But ignoring these two issues, the bigger problem is that we can always locally separate the evolution into independent degrees of freedom.
+First, for this to work, one would have to specify how the states of the parts were defined since their evolution is individually no longer deterministic (the future of each part depends on the state of other parts).
+But ignoring that issues, the bigger problem is that we can always locally separate the evolution into independent degrees of freedom.
 For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value, they clearly evolve independently.

### 2016-08-09 2015CompleteDerivation/completeDerivation.tex
Another pass at kinematic

@@
 
-\section{Kinematics and elementary systems}
-
-We will now focus our attention to the simplest possible classical material: one for which the motion in space is necessary and sufficient to reconstruct the state and its evolution.
-That is, the internal dynamics can be ignored.
+\section{Kinematics and minimal materials}
+
+We will now focus our attention to the simplest possible classical material: one that is fully described by its motion in space, the internal dynamics can be ignored.
 What we'll find is that the dynamics of each infinitesimal part follows classical Lagrangian mechanics.
@@
 
-Here we make the link conceptually explicit with the following assumption:
+Our aim is to make that link more conceptually crisp and explicit with the following assumption:
 
@@
 
-	The assumption clearly cannot be taken in general.
-Not all material will be elementary in the sense that they may have some internal dynamics.
-That is: position in space is not their only state variable.
-Yet, the assumption is fundamental because what we learn about the link between kinematic and dynamical quantities applies to all materials, even the non elementary ones.
-
-	The first natural question to ask is: why do all systems have \emph{at least} to describe a trajectory in space?
-This seems motivated by the previous assumption of reducibility: spatial extent is the one way that all materials are reduced.
-Therefore the description of all particles of all materials, regardless of whether they are elementary or not, will include spatial degrees of freedom.
-Kinematic variables will form a set of independent degrees of freedom that are universal.
-
-	When the system is not elementary, the particles will be described by the kinematic variables plus some additional independent state variables.
-Yet, the set of all possible evolutions will include those for which kinematic variables evolve independently from the other.
-For example, the evolutions for which the internal d.o.f.
-do not change.
-In other words: the overall dynamics will be the spatial degrees of freedom coupled with the internal degrees of freedom.
-Therefore by studying an elementary material, we are also studying a component of the dynamics that is present in all materials.
-
-	It should be also noted that, once we reduced particles to be infinitesimal at a point, the rest of the description must be a local object invariant under coordinate transformations.
+	The assumption clearly cannot be taken in general: kinematic variables may not fully identify the particle state of a material.
+Yet, the assumption is fundamental because particles of any material will \emph{at least} describe a trajectory in space and what we learn about the link between kinematic and dynamical quantities applies to all materials, even the ones where the kinematic assumption does not apply.
+
+	The first natural question to ask is: why does state evolution always describe a trajectory in space?
+Because spatial extent is the one way that all materials can be decomposed.
+As we divide materials in smaller pieces, physical location is one of the state variable for the particles of all materials.
+If we assume kinematic equivalence, the kinematic variables will be enough to identify a particle state.
+In such a case, we will call the material \emph{minimal} to highlight the idea that the state is described by the smallest possible number of d.o.f..
+
+	When the system is not minimal, the particles will be described by some additional independent state variables.
+The overall dynamics will be the spatial d.o.f.
+coupled with the internal d.o.f..
+Therefore by studying a minimal material, we are also studying a component of the dynamics that is present in all materials.
+
+	To be clear: we are not saying that kinematic equivalence is fundamental because everything must reduce to a material with no internal dynamics.
+That is: we are not saying elementary materials are minimal.
+While this statement seems reasonable and may fit our physical intuition, we have no a priori argument for it.
+We can imagine, in fact, that as we reduce the behavior of one material to its constituents we find the parts to possess richer and richer states.
+In a similar way that the elementary particles have electromagnetic and color charges, while their aggregation is typically neutral, we could imagine finding more and more richness in the internal dynamics.
+While this does not seem to happen in practice, we do not have an argument why this cannot happen.
+
+	What we are saying is that any material is \emph{at least} a minimal material and will possess the dynamic of one coupled with the internal dynamics.
+
+	It should also be noted that, once we reduced particles to be infinitesimal at a point, the rest of the description must be a local object invariant under coordinate transformations.
 This means that the remaining internal dynamics must be identified by scalar, vector or tensor quantities.
 This may be used in the future to justify the extension of this work to field theories.
-
-	To be clear: we are not saying that kinematic equivalence is fundamental because everything must reduce to a material with no internal dynamics.
-While this statement seems reasonable and fits our physical intuition, we have no a priori argument for it.
-We can imagine, in fact, that as we reduce the behavior of one material to its constituents we find the parts to possess richer and richer states.
-For example, most macroscopic object are neutral though the parts are charged.
-We could imagine that, while dividing into parts, we find more and more types of charges, forces and internal dynamics.
-We do not have an argument why this cannot happen.
-
-	What we are saying is that any classical material is \emph{at least} an elementary material and will have posses the dynamic of one coupled with the internal dynamics.
 \end{rationale}
@@
 A trajectory will be given by a set of functions $x^\alpha(s)$ where $s$ is an arbitrary parametrization of the curve.
-What we need to understand is the link between the state space of all the possible trajectories and the complete state space of the particles of a classical material.
+What we need to understand is the link between the complete state space of the particles of a classical material and the space of all the possible trajectories under the kinematic assumption.
 
@@
 
-Under the kinematic equivalence, given a trajectory we must identify a state and vice-versa.
+Under kinematic equivalence, given a trajectory we must identify a state and vice-versa.
 There must be a homeomorphism between the complete state space and the space of possible trajectories.
@@
 Only the direction is physical so the four velocity gives us another $n$ independent variables to identify trajectories.
-The combined position and velocity, which we call \emph{initial conditions}, identify $2n + 1$ trajectories using a hyper-surface of the tangent bundle $T\mathcal{M}$.
+The combined position and velocity, the kinematic variables, identify $2n + 1$ trajectories using a hyper-surface of the tangent bundle $T\mathcal{M}$.
 This is exactly the dimensionality of the complete state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$.
-We cannot add more curves, as a homeomorphism preserves dimensionality, therefore we found the suitable space for the trajectories of particles of a classical material under the kinematic assumption.
-
-Being able to identify particle through both state variables and initial conditions is not enough: we also need to be able to express the distributions of the composite states over position and velocity.
-As such, the differentiable distribution over state variables must become a differentiable distribution over initial condition.
+We cannot add more curves, as a homeomorphism preserves dimensionality, therefore we found the suitable space for the trajectories of particles of a classical material under kinematic equivalence.
+
+Being able to identify particles through either state or kinematic variables is not enough: we also need to be able to express the distributions of the composite states over position and velocity.
+As such, the differentiable distribution over state variables must become a differentiable distribution over kinematic variables.
 This tells us the transformation is differentiable.
@@
 
-This relationship allows us to write the initial conditions in terms of the state variable.
+This relationship allows us to write the kinematic variables in terms of the state variable.
 In particular we have $u^\alpha=u^\alpha(t, q^i, E, p_i)$.
@@
 \begin{justification}
-	TODO
-\end{justification}
-
-Not only we must be able to express distributions in terms of both state variable and initial conditions, but integration has to carry over as well.
+	We claim $\mathcal{H}$ is convex in $E$ and $p_i$.
+By \ref{initial_conditions} we can write the velocity $u^\alpha=u^\alpha(q^\alpha, E, p_i)$ as a differentiable function of the state variables.
+At fixed $q^\alpha$, $u^\alpha$ must be invertible.
+$u^\alpha$ are monotonic in $E$ and $p_i$.
+$\partial_{p_i} \mathcal{H}=d_s q^i= u^i$ and $- \partial_{E} \mathcal{H}= d_s t= u^0$.
+The derivatives of $\mathcal{H}$ in $E$ and $p_i$ are monotonic in $E$ and $p_i$.
+$\mathcal{H}$ is convex in $E$ and $p_i$.
+
+	We claim there exist a Lagrangian $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$.
+$\mathcal{H}$ is convex in $E$ and $p_i$.
+It admits a Legendre transform.
+The Lagrangian $\mathcal{L}$ is such transform.
+\end{justification}
+
+Not only we must be able to express distributions in terms of both state and kinematic variables, but integration has to carry over as well.
 To do that, we need to be able to count trajectories in the same way we are able to count states.
@@
 It is not a velocity: the metric is defined at each point without motion.
-It is the constant of proportionality that maps a range of initial condition from the time variable $t$ to the space variable $x^i$.
-In the same way that $1$km would correspond to a thousand times more initial conditions than $1$m, $1$sec corresponds to $299,792,458$ more initial conditions than $1$m.
+It is the constant of proportionality that maps a range of initial conditions from the time variable $t$ to the space variable $x^i$.
+In the same way that $1$km would correspond to a thousand times more possible states than $1$m, $1$sec corresponds to $299,792,458$ more possible states than $1$m.
 Therefore when we write $x^\alpha = (ct, x^i)$ or $p_\alpha = (- E/c, p_i)$ we are really expressing everything in units that are proportional to the cardinality of state and possibilities.
@@
 
-The link between $\omega$ and $g$ allows us to constrain the kind of diffeomorphism between state variables and initial conditions.
+The link between $\omega$ and $g$ allows us to constrain the kind of diffeomorphism between state variables and kinematic variables.
 If we express the possibilities in terms of both state variables and initial conditions we have: $dq^\alpha \wedge dp_\alpha \propto dx^\alpha du_\alpha$.
@@
 The motion of the photon is not enough to tell us its energy or its momentum.
-Therefore it makes sense that, by assuming kinematic equivalence we implicitly required massive particles.
-
-We have seen that the inertial mass $m$, the invariant speed $c$ and the reduced Plank constant $\hbar$ all serve the same purpose: convert one quantity to another preserving the number of possibilities.
+Therefore it makes sense that, by assuming kinematic equivalence we derived equation valid only for massive particles.
+
+We have seen that the inertial mass $m$, the invariant speed $c$ and the reduced Plank constant $\hbar$ all serve the similar purposes: convert one quantity to another preserving the number of possibilities.
 $dp=\hbar dk$ converts between ranges of wave number and conjugate momentum.
@@
 We should also ask whether any choice of $A$ actually represent a different case leading to different trajectories.
-The simplest and more insightful way to study this problem is to use $x^\alpha$ and $mu_\alpha$ as state variable.
+The simplest and more insightful way to study this problem is to use the kinematic quantities $x^\alpha$ and $mu_\alpha$ as state variable.
 They are not going to be conjugates of each other but they do capture the entire physical description.
@@
 
-Any change of $A$ that leaves $F_{\alpha\beta}$ unchanged does not affect the motion of any function of position and velocity.
+Any change of $A$ that leaves $F_{\alpha\beta}$ unchanged does not affect the evolution of any function of position and velocity.
 Such transformation is what we call a gauge transformation.
@@
 \end{align*}
-Therefore the force act on anti-particle states in the opposite way: they act as if they have opposite charge. Note that nothing physical is actually going backwards in time, just the parametrization.
-
-In this setting, charge conjugation $C$ is simply changing the parametrization from $s$ to $-s$, following the trajectories in the opposite way.
-This will permute each particle anti-state with the corresponding standard state and vice-versa.
+Therefore the force acts on anti-particle states in the opposite way: they act as if they have opposite charge.
+Note that nothing physical is actually going backwards in time, just the parametrization.
+
+In this setting, charge conjugation $C$ is simply changing the parametrization from $s$ to $-s$.
+This will permute particle standard and anti-state as they follow the trajectories in the opposite direction.
 But if we apply parity $P$ and time reversal $T$, changing $x^\alpha$ to $-x^\alpha$, we would have the same net effect: the parametrization stays the same, but it now moves backwards in space and time.
@@
 To sum up, the kinematic assumption has more inherent consequences one would expect at first.
-It means the particles are identified by position and velocity and therefore the system of equation is second order.
-It links the space of initial conditions on $T\mathcal{M}$ with the space of states on $T^*\mathcal{M}$.
+It means the particles are identified by position and velocity and therefore the system of equations is second order.
+It links the space of trajectories on $T\mathcal{M}$ with the space of states on $T^*\mathcal{M}$.
 Such link allows state space metric $\omega$ to induce the space-time metric $g$.
@@
 It constrains the motion to massive particles under scalar and vector potential forces.
-Moreover, it allowed to re-understand basic concepts, like the invariant speed $c$, the inertial mass $m$ and the force tensor $F_{\alpha\beta}$ in terms of state counting and d.o.f.
+Moreover, it allowed to re-understand basic concepts, like the invariant speed $c$, the inertial mass $m$ and the force tensor $F_{\alpha\beta}$ in terms of state counting possibilities and d.o.f.
 independence.
@@
 We now have all the building block of classical mechanics, the consequences of three seemingly simple assumption.
-But that's not all: we have also placed ourselves in a very good position to move forward to quantum systems by using the new physical insights we have gained.
+We have more insight of why these blocks fit together the way they do.
+But we have also placed ourselves in a very good position to move forward to quantum systems by using the new physical insights we have gained.
 
 \section{Prelude to quantum}
+
+In principle, we should be able to simply introduce assumption \ref{???} and work our way to quantum mechanics.
+Unfortunately, given that all sorts of things have been said on quantum system, as well as their opposites, it is best to spend some time clarifying our take on some of these matters.
+This will decrease the chance of being misconstrued later.
+
+We will spend this section talking about the similarities between classical and quantum system.
+We will see that classical distributions already possess qualitative properties that are typically associated with quantum systems.
+Therefore it is more natural to think of quantum systems, to a first approximation, as composite states then particle states.
+This will allow us in the next chapter to truly pinpoint how and why classical and quantum systems are different.
+
+\subsection{Particles and distributions (waves)}
+
+The core problem of much confusion for quantum system is identifying them as something akin to classical particles. This alone leads
 

### 2016-08-09 2015CompleteDerivation/completeDerivation.tex
Another pass at the kinematics text

@@
 	The idea is that the system internal dynamics has no affect on the motion and therefore can be disregarded.
-Consider the motion of our cannonball under gravitational and inertial forces: its temperature or precise chemical composition does not affect the trajectory.
+Consider the motion of our cannonball under gravitational and inertial forces: its temperature or precise chemical composition does not affect the overall trajectory.
 We can assume kinematic equivalence.
@@
 
-	The assumption clearly cannot be taken in general: kinematic variables may not fully identify the particle state of a material.
-Yet, the assumption is fundamental because particles of any material will \emph{at least} describe a trajectory in space and what we learn about the link between kinematic and dynamical quantities applies to all materials, even the ones where the kinematic assumption does not apply.
+	Kinematic equivalence can't therefore be assumed for all systems.
+Yet, the assumption is fundamental because the particles of any material will \emph{at least} describe a trajectory in space and what we learn about the link between kinematic and dynamical quantities applies to all materials, even the ones where the kinematic assumption does not apply.
 
 	The first natural question to ask is: why does state evolution always describe a trajectory in space?
-Because spatial extent is the one way that all materials can be decomposed.
-As we divide materials in smaller pieces, physical location is one of the state variable for the particles of all materials.
+Because spatial extent is the one way in which all materials can be decomposed.
+As we divide materials in smaller pieces, physical location is one state variable we can assign to the particles of all materials.
 If we assume kinematic equivalence, the kinematic variables will be enough to identify a particle state.
@@
 Therefore by studying a minimal material, we are also studying a component of the dynamics that is present in all materials.
-
-	To be clear: we are not saying that kinematic equivalence is fundamental because everything must reduce to a material with no internal dynamics.
-That is: we are not saying elementary materials are minimal.
-While this statement seems reasonable and may fit our physical intuition, we have no a priori argument for it.
-We can imagine, in fact, that as we reduce the behavior of one material to its constituents we find the parts to possess richer and richer states.
-In a similar way that the elementary particles have electromagnetic and color charges, while their aggregation is typically neutral, we could imagine finding more and more richness in the internal dynamics.
-While this does not seem to happen in practice, we do not have an argument why this cannot happen.
-
-	What we are saying is that any material is \emph{at least} a minimal material and will possess the dynamic of one coupled with the internal dynamics.
+In other words: any material is \emph{at least} a minimal material and will possess the dynamic needed to describe its motion in space.
 
@@
 
-As we need to describe trajectories, we introduce $\mathcal{M}$ as the $n+1$ dimensional manifold that defines space and time.
+As we need to describe trajectories, we introduce $\mathcal{M}$ as the $n+1$ dimensional manifold that defines physical space and time.
 A point in $\mathcal{M}$ will be identified by $x^\alpha = (t, x^i)$, its time and space coordinates.
@@
 This is exactly the dimensionality of the complete state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$.
-We cannot add more curves, as a homeomorphism preserves dimensionality, therefore we found the suitable space for the trajectories of particles of a classical material under kinematic equivalence.
-
-Being able to identify particles through either state or kinematic variables is not enough: we also need to be able to express the distributions of the composite states over position and velocity.
+We cannot add more curves, as a homeomorphism preserves dimensionality, therefore we identified the suitable space for the trajectories of particles of a classical material under kinematic equivalence.
+
+Being able to identify particles through either state or kinematic variables is not enough: we also need to be able to express a composite state as the distribution of particle states over position and velocity.
 As such, the differentiable distribution over state variables must become a differentiable distribution over kinematic variables.
@@
 In other words: $g$ inherits it's signature from $\omega$.
+The geometry of space-time is the geometry of the ranges for possible states.
 
@@
 The link between $\omega$ and $g$ allows us to constrain the kind of diffeomorphism between state variables and kinematic variables.
-If we express the possibilities in terms of both state variables and initial conditions we have: $dq^\alpha \wedge dp_\alpha \propto dx^\alpha du_\alpha$.
+If we express the possibilities in terms of both state variables and kinematic variables we have: $dq^\alpha \wedge dp_\alpha \propto dx^\alpha du_\alpha$.
 Since  $p_\alpha=p_\alpha(x, u)$, we have $dp_\alpha = m du_\alpha$, where $m$ is the constant of proportionality.
@@
 Intuitively, a more massive body is harder to accelerate because it has to go through more states for the same change in velocity.
-Note that the Hamiltonian excludes massless particles because the kinematic assumption does that implicitly.
+Also note that the Hamiltonian excludes massless particles because the kinematic assumption does that implicitly.
 A massless particle always has the same velocity, $c$, therefore the trajectory is not enough to tell us the dynamics.
@@
 
-We have seen that the inertial mass $m$, the invariant speed $c$ and the reduced Plank constant $\hbar$ all serve the similar purposes: convert one quantity to another preserving the number of possibilities.
+We have seen that the inertial mass $m$, the invariant speed $c$ and the reduced Plank constant $\hbar$ all serve a similar purposes: convert one quantity to another preserving the number of possibilities.
 $dp=\hbar dk$ converts between ranges of wave number and conjugate momentum.
@@
 Using natural units means measuring all these quantities through the cardinality of possibilities.
-There is a difference, though: while $c$ and $\hbar$ are simply determined by the unit system, $m$ also depends on the type of material.
+There is a difference among them, though: while $c$ and $\hbar$ are simply determined by the unit system, $m$ also depends on the type of material.
 That is, different homogeneous materials can have different inertial mass.
@@
 \{x^\alpha, mu_\beta\} &= \delta^\alpha_\beta \\
-\{mu_\alpha, mu_\beta\} &= \partial_\alpha A_\beta - \partial_\beta A_\alpha = F_{\alpha\beta}
+\{mu_\alpha, mu_\beta\} &= \{p_\alpha, - A_\beta\} + \{-A_\alpha, p_\beta\} \\
+&= \partial_\alpha A_\beta - \partial_\beta A_\alpha = F_{\alpha\beta}
 \end{align*}
@@
 The force tensor, then, tells how much $(x^i, mu_i)$ fail to be independent d.o.f..
-With these expressions, the commutator between any function of position and kinetic momentum can now be calculated.
+With these expressions, the Poisson bracket between any function of position and kinetic momentum can now be calculated.
 For example:
@@
 It constrains the motion to massive particles under scalar and vector potential forces.
-Moreover, it allowed to re-understand basic concepts, like the invariant speed $c$, the inertial mass $m$ and the force tensor $F_{\alpha\beta}$ in terms of state counting possibilities and d.o.f.
+Moreover, it allowed to re-understand basic concepts, like the invariant speed $c$, the inertial mass $m$ and the force tensor $F_{\alpha\beta}$ in terms of state counting and d.o.f.
 independence.
+Finally, it allowed to construct classical equivalent for anti-particles and the CPT theorem.
 

### 2016-08-09 2015CompleteDerivation/completeDerivation.tex
Included Christine's comments

@@
 If we take some amount of material out of a container we know how the state changed, yet we may not know the amount of material before or after.
-In certain cases one may be interesting in measuring the asymmetry of a particular state without measuring (or without being able to measure) the full state itself.} All combined, this gives us the structure of an abelian group.
+In certain cases one may be interested in measuring the asymmetry of a particular state without measuring (or without being able to measure) the full state itself.} All combined, this gives us the structure of an abelian group.
 
@@
 We will call a \emph{particle} the smallest amount of a material to which we can assign an independent state.
-For example, a photon is a particle of light (it is the smallest amount of light we can describe), an infinitesimal amount of water is treated classically as a point particle (the smallest amount for a continuous fluid).
-\footnote{We use this definition as it allows us to retroactively talk on somewhat equal grounds about classical point-particles and quantum particles, and underline how physically (and mathematically) they are very similar yet very different objects.}
+For example, a photon is a particle of light (it is the smallest amount of light we can describe), an infinitesimal amount of water is treated classically as a point particle (the smallest amount for a continuous fluid).\footnote{We use this definition as it allows us to retroactively talk on somewhat equal grounds about classical point-particles and quantum particles, and underline how physically (and mathematically) they are very similar yet very different objects.}
 
@@
 As such, at least locally, we can use it as a state variable $q^1=\rho_\mathcal{c}$: we can identify particle states by the density of the material associated with them.
-This is more physically meaningful as one may first suspect, since placing physical markers (i.e.
+This is more physically meaningful than one may first suspect, since placing physical markers (i.e.
 placing material at particular states) is a common way to define the references for a coordinate system.
@@
 The set of Lebesgue integrable functions $L^1(\mathcal{S})$ includes discontinuous functions that for \ref{continuous_map} are unphysical.
-The space of continuous functions $C(\mathcal{S})$, the space of continuous functions that vanish at infinity $C_0(\mathcal{S})$ and the space of differentiable functions $C^1(\mathcal{S})$ include functions whose integral is infinite, which would represent infinite amount of material.
+The space of continuous functions $C(\mathcal{S})$, the space of continuous functions that vanish at infinity $C_0(\mathcal{S})$ and the space of differentiable functions $C^1(\mathcal{S})$ include functions whose integral is infinite, which would represent an infinite amount of material.
 Some definitions are too restrictive.
@@
 
-Note that neither of those expressions match the norm given by the vector product defined as $\langle f, g \rangle = \int_{\mathcal{S}} fg \, d\mu$.
+Note that neither of those expressions matches the norm given by the inner product defined as $\langle f, g \rangle = \int_{\mathcal{S}} fg \, d\mu$.
 Moreover $\mathcal{C}$, since it excludes discontinuous functions, can never be a complete normed space: it will not include the limit for all Cauchy sequences, it cannot be a Banach or a Hilbert space.
@@
 
-If $\rho_{\mathcal{c}}$ is to remain invariant under state variable changes, not only the Jacobian has to exist but must be unitary.
+If $\rho_{\mathcal{c}}$ is to remain invariant under state variable changes, not only does the Jacobian have to exist but it must be unitary.
 This means we cannot simply change one state variable as we please.
@@
 (i.e.
-the wedge product within a d.o.f) and the orthogonality of different d.o.f.
+the wedge product within a d.o.f.) and the orthogonality of different d.o.f.
 (i.e.
@@
 A diffeomorphism $\hat{q}=\hat{q}(q)$ fully defines a unit transformation.
-Let $(q,k^i)$ with $i=1...n-1$ be a set of state variable.
+Let $(q,k^i)$ with $i=1...n-1$ be a set of state variables.
 The Jacobian determinant $|J| = d_q\hat{q} |\partial_i\hat{k}^j|=1$ as densities must be invariant.
@@
 	We claim the state space $\mathcal{S}$ for the particles of a homogeneous classical material is a symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of an $n$-dimensional differential manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
-Let $q^i$ be a set of n continuous independent state-variable that define the units necessary to describe a state in $U \subseteq \mathcal{S}$.
+Let $q^i$ be a set of n continuous independent state variables that define the units necessary to describe a state in $U \subseteq \mathcal{S}$.
 Let $\mathcal{Q}$ be the manifold charted by $q^i$.
@@
 Changing units of one state variable must not change the units of the other as they are independent.
-Each degree of freedom must define state-variable-distribution independently.
+State-variable-distribution can  be defined separately on each degree of freedom.
 For each $\mathcal{Q}^i$ we have a covector $k_i(q) dq^i \in T^*\mathcal{Q}^i$.
@@
 Integration must also be defined on an independent d.o.f.
-There must exist a non degenerate two-form $\omega$ such that $\int_{U \subset \mathcal{S}} \rho_\mathcal{c} \omega$ where $U$ is any two dimensional surface in $\mathcal{S}$.
+There must exist a non-degenerate two-form $\omega$ such that $\int_{U \subset \mathcal{S}} \rho_\mathcal{c} \omega$ where $U$ is any two dimensional surface in $\mathcal{S}$.
 $\omega$ has to be form invariant under state variable change.
@@
 Our distribution would not be a density and there would be no justification for conjugate variables.
-A finite distribution could only distribute finite amount of material in a finite number of states.
-
-In other words: this is not a case were a difference at  small scale has a small difference at large scale.
+A finite distribution could only distribute a finite amount of material in a finite number of states.
+
+In other words: this is not a case where a difference at  small scales has a small difference at large scales.
 The two cases are radically different.

### 2016-08-12 2015CompleteDerivation/completeDerivation.tex
Another pass at kinematic

@@
 We can capture the above discussion by stating that the complete state space $\bar{\mathcal{S}}$ is a hypersurface of the symplectic manifold $(T^*\mathcal{M}, \omega)$, where $\omega \equiv \sum dq^i \wedge dp_i - dt \wedge dE$.
+And if we set $q^\alpha = (t, q^i)$ and $p_\alpha = (-E, p_i)$, we can write $\omega \equiv \sum dq^\alpha \wedge dp_\alpha$.
 This is the manifold that allows us to define time-and-state-variable-invariant densities.
@@
 \begin{prop}\label{relativistic_symplectic_manifold}
-	The complete state space $\hat{\mathcal{S}}$ for the particles of a classical material is a hyper-surface of the symplectic manifold formed by a cotangent bundle $T^*\mathcal{M}$ equipped with the canonical two-form $\omega = \sum dq^i \wedge dp_i - dt \wedge dE$.
+	The complete state space $\hat{\mathcal{S}}$ for the particles of a classical material is a hyper-surface of the symplectic manifold formed by a cotangent bundle $T^*\mathcal{M}$ equipped with the canonical two-form $\omega = \sum dq^i \wedge dp_i - dt \wedge dE = \sum dq^\alpha \wedge dp_\alpha$.
 \end{prop}
@@
 We will now focus our attention to the simplest possible classical material: one that is fully described by its motion in space, the internal dynamics can be ignored.
-What we'll find is that the dynamics of each infinitesimal part follows classical Lagrangian mechanics.
+What we'll find is that the dynamics of its infinitesimal parts follows classical Lagrangian mechanics.
 Furthermore, the Hamiltonian of the system is constrained to the one of a free particles under scalar and vector potential forces.
@@
 \begin{rationale}
-	The idea is that the system internal dynamics has no affect on the motion and therefore can be disregarded.
+	The idea is that the system internal dynamics has no effect on the motion and therefore can be disregarded.
 Consider the motion of our cannonball under gravitational and inertial forces: its temperature or precise chemical composition does not affect the overall trajectory.
@@
 	Kinematic equivalence can't therefore be assumed for all systems.
-Yet, the assumption is fundamental because the particles of any material will \emph{at least} describe a trajectory in space and what we learn about the link between kinematic and dynamical quantities applies to all materials, even the ones where the kinematic assumption does not apply.
+Yet, the assumption is fundamental because the particles of any material will \emph{at least} describe a trajectory in space and what we learn about the link between kinematic and dynamical quantities applies to all materials, even the ones where the kinematic assumption does not hold.
 
@@
 
-	When the system is not minimal, the particles will be described by some additional independent state variables.
+	When the system is not minimal, the state of the particles will be described by some additional independent state variables.
 The overall dynamics will be the spatial d.o.f.
@@
 	It should also be noted that, once we reduced particles to be infinitesimal at a point, the rest of the description must be a local object invariant under coordinate transformations.
-This means that the remaining internal dynamics must be identified by scalar, vector or tensor quantities.
+This means that the remaining internal state must be identified by scalar, vector or tensor quantities.
 This may be used in the future to justify the extension of this work to field theories.
@@
 
-Mathematically, the state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is diffeomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$ of a pseudo-Riemannian manifold $(\mathcal{M}, g)$.
+Mathematically, the state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is diffeomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$ of the space-time manifold.
 
@@
 	Let $\mathcal{M}$ be the $n+1$ dimensional space-time manifold over which trajectories are defined.
-The state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is diffeomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$.
-\end{prop}
-
-\begin{justification}
-	TODO
+The complete state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ for the particles of a minimal classical material is diffeomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$.
+\end{prop}
+
+\begin{justification}
+	We claim the complete state space is a hyper-surface of the cotangent bundle of the space-time manifold, $\bar{\mathcal{S}} \subset T^*\mathcal{M}$.
+Let $\bar{\mathcal{s}} \in \bar{\mathcal{S}}$ be the evolution of a particle for a minimal classical material.
+Let $f(\bar{\mathcal{s}}) \mapsto x^\alpha(s)$ be the function that returns the space-time trajectory associated with the evolution.
+$f$ is injective as the material is minimal.
+$\bar{\mathcal{s}}$ can be identified by a trajectory $x^\alpha(s)$ which can be fully described by coordinates defined on $\mathcal{M}$.
+Let $\mathcal{s} \in \bar{\mathcal{S}}$ be the particle state for a minimal classical material.
+Let $f(\bar{\mathcal{s}}, t) \mapsto \mathcal{s}$ the function that returns the state at a particular time $t$ along an evolution $\bar{\mathcal{s}}$.
+Such function is a bijection as the evolution is deterministic and reversible.
+$\mathcal{s}$ can be identified by an evolution $\bar{\mathcal{s}}$ and an instant $t$, both fully described by coordinates defined on $\mathcal{M}$.
+$\mathcal{M}$ defines the unit system.
+$\bar{\mathcal{S}} \subset T^*\mathcal{M}$ by \ref{relativistic_symplectic_manifold}.
+
+	We claim there exists a homeomorphism between $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ and a hyper-surface of $T\mathcal{M}$.
+Let $\mathcal{s} \in \bar{\mathcal{S}}$ be the particle state for a minimal classical material.
+Let $x^\alpha(s)$ the evolution associated with that state.
+Let $t(\mathcal{s})$ be the instant of time at which $\mathcal{s}$ is defined.
+$x^\alpha(s(t(\mathcal{s}))) = f(\mathcal{s}) = q^\alpha$ as the initial position cannot be derived by the law of evolution.
+$q^\alpha$ is differentiable therefore $d_s x^\alpha$ exists.
+Suppose $d_s x^\alpha$ is not part of the state.
+$d_s x^\alpha = d_s q^\alpha$ is determined by the law of evolution.
+$x^\alpha(s)$ is determined by the law of evolution.
+$q^\alpha$ determines the trajectory $x^\alpha(s)$ and the state $\mathcal{s}$, as the material is minimal.
+This is a contradiction: $\mathcal{s}$ is not identified by just $q^\alpha$.
+$u^\alpha=d_s x^\alpha$ is part of the state.
+$p_\alpha = p_\alpha(x^\alpha, u^\alpha)$.
+The map $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is continuous by \ref{continuous_map} and is bijective as the material is minimal.
+It is a homemorphism between $T^*\mathcal{M}$ and $T\mathcal{M}$.
+$\bar{\mathcal{S}}$ is a hyper-surface of $T^*\mathcal{M}$ therefore its image is a hyper-surface of $T\mathcal{M}$.
+
+	We claim $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is a diffeomorphism.
+Let $\bar{\mathcal{c}}$ be the evolution of a minimal classical material.
+Let $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be the distribution associated with $\bar{\mathcal{c}}$.
+$\rho_{\bar{\mathcal{c}}}(\mathcal{s}(x^\alpha, u^\alpha))$ as the material is minimal.
+$\rho_{\bar{\mathcal{c}}}(x^\alpha, u^\alpha)$ is differential for the same arguments expressed in \ref{differentiable_manifold}.
+$(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is a diffeomorphism as it as a homeomorphism that maps a differentiable function to a differentiable function.
 \end{justification}
@@
 This relationship allows us to write the kinematic variables in terms of the state variable.
-In particular we have $u^\alpha=u^\alpha(t, q^i, E, p_i)$.
-As the relationship is invertible, these functions must be monotonic in $E$ and $p_i$.
-Since $\partial_{p_i} \mathcal{H}=d_s q^i= u^i$ and $- \partial_{E} \mathcal{H}= d_s t= u^0$, $\mathcal{H}$ is convex in $E$ and $p_i$.
+In particular we have $u^\alpha=u^\alpha(q^\alpha, p_\alpha)$.
+As the relationship is invertible, these functions must be monotonic in $p_\alpha$.
+Since $\partial_{p_\alpha} \mathcal{H}=d_s q^\alpha= u^\alpha$, $\mathcal{H}$ is convex in $p_\alpha$.
 Under such conditions we can define a Legendre transformation $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$.
-Such transformation is the Lagrangian of the system and laws of motions are given by the Euler-Lagrange equation.
+Such transformation is the Lagrangian of the system and the laws of motions are given by the Euler-Lagrange equation.
+For the time independent case, we define the \emph{Lagrangian} $L=v^i p_i - H$ where $v^i=d_t x^i$.
 
@@
 Under the kinematic assumption, we can define the Legendre transform $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$ which we call the \emph{extended Lagrangian}.
-For the time invariant case, we define the \emph{Lagrangian} $L=v^i p_i - H$.
-\end{prop}
-
-\begin{justification}
-	We claim $\mathcal{H}$ is convex in $E$ and $p_i$.
+\end{prop}
+
+\begin{justification}
+	We claim $\mathcal{H}$ is convex in $p_\alpha$.
 By \ref{initial_conditions} we can write the velocity $u^\alpha=u^\alpha(q^\alpha, E, p_i)$ as a differentiable function of the state variables.
 At fixed $q^\alpha$, $u^\alpha$ must be invertible.
-$u^\alpha$ are monotonic in $E$ and $p_i$.
-$\partial_{p_i} \mathcal{H}=d_s q^i= u^i$ and $- \partial_{E} \mathcal{H}= d_s t= u^0$.
-The derivatives of $\mathcal{H}$ in $E$ and $p_i$ are monotonic in $E$ and $p_i$.
-$\mathcal{H}$ is convex in $E$ and $p_i$.
-
-	We claim there exist a Lagrangian $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$.
-$\mathcal{H}$ is convex in $E$ and $p_i$.
-It admits a Legendre transform.
-The Lagrangian $\mathcal{L}$ is such transform.
+The components $u^\alpha$ are monotonic in $p_\alpha$.
+$\partial_{p_\alpha} \mathcal{H}=d_s q^\alpha= u^\alpha$.
+The derivatives of $\mathcal{H}$ in $p_\alpha$ are monotonic in $p_\alpha$.
+$\mathcal{H}$ is convex in $p_\alpha$.
+
+	We claim there exist a Lagrangian $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$. $\mathcal{H}$ is convex in $p_\alpha$. It admits a Legendre transform. The Lagrangian $\mathcal{L}$ is such transform.
 \end{justification}
@@
 There must exist a metric $g$, consistent with $\omega$, that allows to calculate the cardinality of possibilities for initial conditions.
-That is: $\omega(dt,dq^i,dE,dp_i) \propto g(dx^\alpha,du^\alpha)$.
+That is: $\omega(dq^\alpha,dp_\alpha) \propto g(dx^\alpha,du^\alpha)$.
 
 Since the both $dx^\alpha$ and $du^\alpha$ are four vectors and the metric must be linear, $g$ is a rank two tensor on $\mathcal{M}$.
-To each direction in space-time $dx^\alpha$ is associated a direction in the tangent space of velocities $du^\alpha$ with which it forms an independent d.o.f..
-Across independent d.o.f.
-we have $\omega(dq^i, dp_j) = 0 =\omega(dq^j, dp_i)$.
-Therefore if $dx^i$ and $du^j$ do not form a d.o.f., $g(dx^i, du^j) = 0 = g(dx^j, du^i)$: $g$ is reflexive.
-We also know that $g$ cannot be alternating, that is $g(dx^i, du^i) \neq 0$: a choice of units for $x^i$ is a choice of units for $u^i$ and therefore they cannot belong to different d.o.f.
+To each direction in space-time $dx^\alpha$ is associated a direction in the tangent space of velocities $du^\alpha$ with which it forms a d.o.f..
+Across d.o.f.
+we have $\omega(dq^\alpha, dp_\beta) = 0 =\omega(dq^\beta, dp_\alpha)$.
+Therefore if $dx^\alpha$ and $du^\beta$ do not form a d.o.f., $g(dx^\alpha, du^\beta) = 0 = g(dx^\beta, du^\alpha)$: $g$ is reflexive.
+We also know that $g$ cannot be alternating, that is $g(dx^\alpha, du^\alpha) \neq 0$: a choice of units for $x^\alpha$ is a choice of units for $u^\alpha$ and therefore they cannot belong to different d.o.f.
 This means $g$ is symmetric.
@@
 We can further change the units along the spatial directions so that $g_{ii}=1$: if we use the same units for the spatial coordinates, each direction will contribute the same number of possibilities.
-Since time must use a different unit, we set $g_{00}=-c^2$ where $c$ is the constant that allows to change an range in time to a range in space of equal cardinality of possibilities (i.e.
-that counts the same number of states).
+Since time must use a different unit, we set $g_{00}=-c^2$ where $c^2$ is the constant that allows to convert the possibilities over the temporal d.o.f.
+into the possibilities of a spatial d.o.f..
 Therefore, locally $g(dx^i, dx^i) = (dx^i)^2 - c^2 dt^2$ is Minkowskian.
@@
 \begin{prop}\label{riemannian_manifold}
-	The space-time manifold $\mathcal{M}$ is a pseudo-Riemannian manifold $(\mathcal{M}, g)$ and locally Mikowskian. That is, locally $g = (dx^i)^2 - c^2 dt^2$.
-\end{prop}
-
-\begin{justification}
-	TODO
+	The space-time manifold $\mathcal{M}$ is a pseudo-Riemannian manifold $(\mathcal{M}, g)$. $g$ has signature $(3,1)$ and locally is $g = (dx^i)^2 - c^2 dt^2$.
+\end{prop}
+
+\begin{justification}
+	N.B.
+Though $du^\alpha \in TT\mathcal{M}$, its components change as the components of vectors in $T\mathcal{M}$ under coordinate transformation.
+There is a canonical way to associate between elements of $TT\mathcal{M}$ and $T\mathcal{M}$.
+Therefore we consider $du^\alpha \in T\mathcal{M}$ a vector as it is standard in physics.
+
+	We claim $(\mathcal{M}, g)$ is a pseudo-Riemannian manifold where $g$ is a metric induced by $\omega$ under kinematic equivalence.
+Let $(dx^\alpha, du^\alpha)$ an infinitesimal range of possibilities for the kinematic variables of the particles of a minimal classical material.
+There exists a map $g : T\mathcal{M} \times T\mathcal{M} \rightarrow \mathbbm{R}$ such that $g(dx^\alpha, du^\alpha) \propto \omega(dq^\alpha, dp_\alpha)$.
+The map $g$ is linear, since a linear combination of infinitesimal ranges leads to the linear combination of possibilities.
+$g$ is a rank two tensor.
+Suppose $g(dx^\alpha, du^\beta) = 0$.
+$g(dx^\alpha, du^\beta) \propto \omega(dq^\alpha, dp_\beta) = 0 = \omega(dq^\beta, dp_\alpha) \propto g(dx^\beta, du^\alpha) = 0$.
+$g$ is reflexive.
+Position and velocity along the same direction use related units.
+They cannot form independent d.o.f..
+$g(dx^\alpha, du^\alpha) \neq 0$.
+$g$ is not alternating.
+$g$ is symmetric.
+$g$ is a symmetric rank two tensor.
+$(\mathcal{M}, g)$ is a pseudo-Riemannian manifold.
+
+	We claim $g$ is locally Minkowskian.
+Let $x^\alpha$ be a choice of coordinate that diagonalizes $g$.
+$x^\alpha$ exists as $g$ is symmetric.
+$\omega(dq^\alpha, dp_\beta) = dq^i \wedge dp_i - dt \wedge dE \propto dq^i g_{ii} du^i + dt g_{00} du^0$.
+$dq^\alpha = dx^\alpha$.
+$g_{ii} > 0$ and $g_{00}$.
+$g$ has signature $(3, 1)$.
+Choosing the same unit for all $dx^i$ and a temporal unit for $dt$ we can choose coordinates such that locally $g = (dx^i)^2 - c^2 dt^2$.
+$g$ is locally Minkowskian.
 \end{justification}
@@
 Directions in space are orthogonal not because the angle between them is $90^\circ$, but because they identify independent d.o.f..
-Space and time, instead, are not independent and the geometry between them is different.
+That is, the choice of units along one direction is independent from the choice of unit along orthogonal directions.
+Space and time, instead, are not independent d.o.f.
+and the geometry between them is different.
 In other words: $g$ inherits it's signature from $\omega$.
@@
 In the same way that $1$km would correspond to a thousand times more possible states than $1$m, $1$sec corresponds to $299,792,458$ more possible states than $1$m.
-Therefore when we write $x^\alpha = (ct, x^i)$ or $p_\alpha = (- E/c, p_i)$ we are really expressing everything in units that are proportional to the cardinality of state and possibilities.
-Naturally, these units make the book keeping of number of states much easier which, as we saw, is the only thing that Hamiltonian and Lagrangian mechanics are doing.
-
-This also allows to understand conceptually why the speed of a deterministic and reversible signal cannot be greater than $c$.
+Therefore when we write $x^\alpha = (ct, x^i)$ or $p_\alpha = (- E/c, p_i)$ we are really expressing everything in quantities that are proportional to the cardinality of state and possibilities.
+These coordinates make the book keeping of number of states much easier which, as we saw, is the only thing that Hamiltonian and Lagrangian mechanics are doing.
+
+This also allows to understand conceptually why the speed of deterministic and reversible motion cannot be greater than $c$.
 Consider the trajectory of a point particle.
 Over an interval $dt$ it will travel $dx$ in space.
-Since the motion is deterministic and reversible, it will travel through all the states defined in the interval $dx$.
+Since the motion is deterministic and reversible, and therefore continuous, it will travel through all the states defined in the interval $dx$.
 It will also travel through every moment in time, which corresponds to $cdt$ possibilities.
-The number of states that will travel in space cannot exceed the number of states it travels in time as it can only be at one position for each instant.
+As it moves, time will always change but not necessarily space.
+The number of states traveled in space cannot exceed the number of states traveled in time.
 Therefore we have:
@@
 It was under that premise that we were able to show that the trajectories are continuous, that we could define $\omega$ and therefore $g$.
-Non-local quantum effects, such as the ones features in EPR experiments, are not examples of deterministic and reversible motion.
+Non-local quantum effects, such as the ones featured in EPR experiments, are not examples of deterministic and reversible motion.
 Therefore they do not have anything to violate: the maximum speed does not apply to them.
@@
 
-The link between $\omega$ and $g$ allows us to constrain the kind of diffeomorphism between state variables and kinematic variables.
+The link between $\omega$ and $g$ allows us to constrain the possible diffeomorphisms between state variables and kinematic variables.
 If we express the possibilities in terms of both state variables and kinematic variables we have: $dq^\alpha \wedge dp_\alpha \propto dx^\alpha du_\alpha$.
@@
 We recognize $m$ as the inertial mass.
-Integrating the expression, we have $p_\alpha = m u_\alpha + A_\alpha(x)$ where $A_\alpha(x)$ is an arbitrary function.
+Integrating the expression, we have $p_\alpha = m u_\alpha + A_\alpha(x)$ where $A_\alpha(x)$ is an arbitrary vector field.
 This is the most general link between velocity and conjugate momentum, which is the sum of kinetic momentum and a scalar potential.
@@
 \begin{justification}
-	TODO
-\end{justification}
-
-The above discussion gives the inertial mass a new physical meaning: it is the conversion constant that tells us how to convert a range of velocity to a range of conjugate momentum with the same number of possibilities.
+	We claim the conjugate momentum expressed as a function of the kinematic variables is $p_\alpha = m u_\alpha + A_\alpha(x)$ where $A_\alpha \in \mathfrak{X}(\mathcal{M})$ is a differentiable vector field.
+$\omega(dq^\alpha, dp_\beta) = dq^\alpha \wedge dp_\alpha \propto g(dx^\alpha, du^\beta) = dx^\alpha du_\alpha$.
+Let $m$ be the proportionality constant.
+$dq^\alpha \wedge dp_\alpha = dx^\alpha \wedge \partial_{u_\beta} p_\alpha du_\beta = m dx^\alpha du_\alpha$.
+$\partial_{u_\beta} p_\alpha = m \delta_\alpha^\beta$.
+After integration $p_\alpha = m u_\alpha + A_\alpha(x)$ where $A$ is an arbitrary vector field.
+
+	We claim the Hamiltonian is $\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta) + m^2 c^2)$.
+$\frac{1}{m}g^{\alpha\beta}(p_\beta-A_\beta) = u^\alpha = d_s x^\alpha = d_s q^\alpha = \partial_{p_\alpha} \mathcal{H}$.
+After integration $\mathcal{H}=\frac{1}{2m}(p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta)+V(x)$ where $V$ is an arbitrary function.
+Let $s$ be the natural parametrization such that $u^\alpha u_\alpha = - c^2$.
+$V= \frac{1}{2}m c^2$ as $\mathcal{H}=0$ by \ref{form_of_extended_hamiltonian}.
+$\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta) + m^2 c^2)$.
+
+\end{justification}
+
+The above discussion gives the inertial mass $m$ a new physical meaning: it is the conversion constant that tells us how to convert a range of velocity to a range of conjugate momentum with the same number of possibilities.
 In other words: the more massive the particles, the more states per unit velocity.
@@
 That is, different homogeneous materials can have different inertial mass.
+The reason is that the choice of units for $x^\alpha$ already determines the others ($x^\alpha \rightarrow u^\alpha$ and $x^\alpha \rightarrow q^\alpha \rightarrow k_\alpha \rightarrow p_\alpha$) therefore the link between units of $u^\alpha$ and $p_\alpha$ is not a new unit conversion.
 
@@
 
-We should also ask whether any choice of $A$ actually represent a different case leading to different trajectories.
+We should also ask whether any choice of $A$ actually represents a different case leading to different trajectories.
 The simplest and more insightful way to study this problem is to use the kinematic quantities $x^\alpha$ and $mu_\alpha$ as state variable.
@@
 Any change in $A$ that does not affect the relationship between $x^\alpha$ and $mu_\alpha$ does not change the physics.
-Expressing the Hamiltonian and the Poisson brackets in terms of the initial conditions we have:
+Expressing the Hamiltonian and the Poisson brackets in terms of the kinematic variables we have:
 \begin{align*}
-\mathcal{H}&=\frac{1}{2m}(mu_\alpha mu^\alpha + m^2 c^2) \\
+\mathcal{H}&=\frac{1}{2}mu_\alpha u^\alpha + \frac{1}{2} m c^2 \\
 \{x^\alpha, x^\beta\} &= 0 \\
@@
 independence.
-Finally, it allowed to construct classical equivalent for anti-particles and the CPT theorem.
-
-We now have all the building block of classical mechanics, the consequences of three seemingly simple assumption.
+Finally, it allowed to construct classical equivalents for anti-particles and the CPT theorem.
+
+We now have all the building block of classical mechanics, the consequences of three seemingly simple assumptions.
 We have more insight of why these blocks fit together the way they do.
@@
 
-\subsection{Particles and distributions (waves)}
-
-The core problem of much confusion for quantum system is identifying them as something akin to classical particles. This alone leads
-
-TODO
+\subsection{Particles and distributions}
+
+As we saw in the previous sections, classical particles should not be understood as point-like finite mass objects as this would open a series of problems and inconsistencies.
+The issues created by such misidentification are only amplified within the context of quantum mechanics.
+In fact, many issues preventing a satisfying intuitive picture stem from this confusion.
+
+The problem goes along these line.
+One assumes, implicitly or explicitly, that quantum particles are point-like.
+But they don't behave like one, they are described by a distribution.
+In some cases, that distribution experimentally becomes a probability distribution.
+So the state of a quantum particles is a probability distribution.
+Except that it self-interacts.
+So its a weird quantum probability distribution.
+The quantum particle behaves like its in more than one place at once.
+We learn this, we get used to it, we can get our calculations done...
+but it's still preposterous.
+
+The first thing to note here is that most objects behave like they are in more than one place at once.
+The earth is both under Milano and Ann Arbor.
+The readers right index finger is probably at a different place from his left eye.
+In fact, based on what we saw before, any object of finite mass is a distribution over a region, as discontinuous distributions are ruled out.
+An electron is an object of finite mass.
+Why should it be any different?
+In fact, this we are claiming that is different from any other object, shouldn't there be a high burden of proof?
+
+Part of the confusion seem to be the idea is that the electron is an elementary particle, and therefore should have no spatial extent.
+But that does not follow.
+First of all, because one can construct a counter example.
+Suppose you have objects constrained on the surface of a sphere.
+All their dynamics can be expressed by their polar and azimuthal angles.
+If we scale the sphere, the dynamics stays the same.
+We can shrink the whole sphere to a point: it is now a point-like object that is not elementary.
+Second of all, if elementary objects are point-like, they cannot know how many spatial dimensions are there.
+Now can the whole universe be reduced to elementary elements that have no idea how many spatial dimensions are there?
+
+Entertaining the thought that an electron is not a point-like object described by weird quantum probability, but an amount of material spread in space like any other, may seem, at this point, weird.
+Mainly because its not how we are taught to think about it, and trying out a new thought is hard work.
+But hopefully it should be evident that this new thought is not at all more preposterous than the old one.
+
+In fact, if we accept it, we start to see that another critical feature usually associated with quantum systems is already there for classical distributions.
 
 \subsection{Classical uncertainty principle}
+
+Suppose we have a finite amount of classical material distributed over a single d.o.f.
+$a_m = \int \rho_\mathcal{c} dq^n \wedge dp_n$.
+The Shannon entropy associated with the distribution, which can be though as the amount of information required to identify a particle of the distribution, is finite as well.
+$I(\rho_\mathcal{c}) = \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq^n \wedge dp_n = I_0$.
+TODO: prove in appendix.
+
+As the state evolves into its final state, the amount of material remains the same: $a_m = \int \rho_{\hat{\mathcal{c}}} dq^n \wedge dp_n$.
+The final entropy will also remain the same since, like the distribution itself, is just another integrable function: $I(\rho_{\hat{\mathcal{c}}}) = I_0$.
+
+In the same way that each particle is constrained to move along states at equal energy, the composite state is constrained to move along distributions at equal entropy.
+This makes sense: if the motion is deterministic and reversible, the same amount of information is required to identify a particle of material.
+
+We now ask the following question: once we fix the entropy, what is the relationship between the spread of the marginal distributions along $q$ and $p$?
+What is the distribution $\rho$ that minimizes the product of the variance $\sigma_q^2 \sigma_p^2 = \int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp$?
+
+Suppose that we let interact it with an entropy source for enough time. The marginal entropy
 

### 2016-08-24 2015CompleteDerivation/completeDerivation.tex
Update status
More work on prelude to quantum

@@
 
-Section VII - Rough draft
+Section VII - Mostly done. Needs review.
+
+Section VIII - Scoping ideas
 
@@
 
-In principle, we should be able to simply introduce assumption \ref{???} and work our way to quantum mechanics.
-Unfortunately, given that all sorts of things have been said on quantum system, as well as their opposites, it is best to spend some time clarifying our take on some of these matters.
-This will decrease the chance of being misconstrued later.
-
-We will spend this section talking about the similarities between classical and quantum system.
-We will see that classical distributions already possess qualitative properties that are typically associated with quantum systems.
-Therefore it is more natural to think of quantum systems, to a first approximation, as composite states then particle states.
-This will allow us in the next chapter to truly pinpoint how and why classical and quantum systems are different.
+In principle, we could simply introduce assumption \ref{???} and work our way to quantum mechanics.
+Unfortunately, given the history and baggage associated with quantum systems, there is a high chance of being misconstrued.
+The intuitive pictures people have constructed or inherited are varied, may start from different premises and therefore it is easy to misinterpret what one is saying.
+Therefore we will first ground in classical mechanics the intuition that will be needed later, making the step into full quantum mechanics a gentler, and hopefully less confusing, one.
+
+We will see that, as a first approximation, quantum particles are more keenly described by classical composite states (i.e.
+distributions) as, unlike classical particles, they are made of a finite amount of material.
+We will also see that classical distributions already possess qualitative properties that are typically associated with quantum systems and construct a classical analogue of the uncertainty principle.
 
@@
 One assumes, implicitly or explicitly, that quantum particles are point-like.
-But they don't behave like one, they are described by a distribution.
-In some cases, that distribution experimentally becomes a probability distribution.
+But they don't always behave like one, they also behave like a wave, a distribution.
+During measurements, it coincides experimentally with a probability distribution.
 So the state of a quantum particles is a probability distribution.
@@
 The quantum particle behaves like its in more than one place at once.
-We learn this, we get used to it, we can get our calculations done...
-but it's still preposterous.
-
-The first thing to note here is that most objects behave like they are in more than one place at once.
-The earth is both under Milano and Ann Arbor.
-The readers right index finger is probably at a different place from his left eye.
+We learn this picture and, sometimes begrudgingly, we get used to it.
+
+Here we contend that this picture is not only confusing and unnecessary, it also prevents to understand fully the deep links between classical and quantum mechanics, in the same way that thinking of classical particles as point-like objects prevented to see the holistic picture in classical mechanics.
+
+The first thing to note here is that all objects behave like they are in more than one place at once.
+The earth is both under Milano and New York.
+Your right index finger is not, hopefully, at the same position of your his left eye.
 In fact, based on what we saw before, any object of finite mass is a distribution over a region, as discontinuous distributions are ruled out.
@@
 Why should it be any different?
-In fact, this we are claiming that is different from any other object, shouldn't there be a high burden of proof?
+In fact, shouldn't the burden of proof be on the claim that the electron is point-like, unlike any other finite mass object?
 
@@
 But that does not follow.
-First of all, because one can construct a counter example.
-Suppose you have objects constrained on the surface of a sphere.
+Lack of spatial extend and lack of substructure are independent concepts.
+In fact: we can construct an infinitesimal object that is not elementary.
+Suppose you have dots constrained on the surface of a sphere.
 All their dynamics can be expressed by their polar and azimuthal angles.
 If we scale the sphere, the dynamics stays the same.
-We can shrink the whole sphere to a point: it is now a point-like object that is not elementary.
-Second of all, if elementary objects are point-like, they cannot know how many spatial dimensions are there.
-Now can the whole universe be reduced to elementary elements that have no idea how many spatial dimensions are there?
-
-Entertaining the thought that an electron is not a point-like object described by weird quantum probability, but an amount of material spread in space like any other, may seem, at this point, weird.
-Mainly because its not how we are taught to think about it, and trying out a new thought is hard work.
-But hopefully it should be evident that this new thought is not at all more preposterous than the old one.
-
-In fact, if we accept it, we start to see that another critical feature usually associated with quantum systems is already there for classical distributions.
+We can shrink the whole sphere to a point: it is now a point-like object that is not elementary, as there is motion on its surface.
+
+Not only point-like object don't have to be elementary, but it's not clear how they can be elementary.
+How can the whole universe be reduced to elementary elements that have no idea how many spatial dimensions are there?
+Mathematically: the set of points by itself does not define the topology.
+How is it supposed to work?
+
+Entertaining the thought that an electron is actually physically distributed in space may seem preposterous, probably because it's not how we are taught to think about it.
+But it's no more preposterous than to think it point-like, described by a probability distribution that self-interacts.
+
+The first payoff of accepting that an electron is a finite amount of material distributed in space is that this is more consistent with classical mechanics as we have defined it\footnote{It is also more consistent with general relativity.}.
+The second payoff is that some features of classical mechanics already explain part of the behavior that we thought was inherently quantum.
 
@@
 As the state evolves into its final state, the amount of material remains the same: $a_m = \int \rho_{\hat{\mathcal{c}}} dq^n \wedge dp_n$.
-The final entropy will also remain the same since, like the distribution itself, is just another integrable function: $I(\rho_{\hat{\mathcal{c}}}) = I_0$.
+The final entropy will also remain the same since the entropy, like the distribution itself, is just another integrable function: $I(\rho_{\hat{\mathcal{c}}}) = I_0$.
 
@@
 We now ask the following question: once we fix the entropy, what is the relationship between the spread of the marginal distributions along $q$ and $p$?
-What is the distribution $\rho$ that minimizes the product of the variance $\sigma_q^2 \sigma_p^2 = \int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp$?
-
-Suppose that we let interact it with an entropy source for enough time. The marginal entropy
-
+Using Lagrange multipliers, we fix the total amount of material $a_m$ and the total entropy $I_0$ and find the distribution that minimizes the the product of the variance $\sigma_q^2 \sigma_p^2 = \int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp$.
 \begin{align*}
@@
 \end{align*}
+We find that the distribution that minimizes the spread is the product of two independent Gaussians.
+In quantum mechanics, the Gaussian wave packet is the state that minimizes uncertainty: this is the classical analogue.
+
+As the distribution evolves, the entropy is conserved and therefore the product $\sigma_q^2 \sigma_p^2$ can never be less the one given by the Gaussian distribution of the same entropy. We have:
 \begin{align*}
@@
 \end{align*}
+This constitute the classical equivalent of the Heisenberg uncertainty principle.
+
+The generalization to multiple d.o.f.
+is outside of the scope of this work.
+Note that this result is similar in spirit to the Gromov non-squeezing theorem and the symplectic camel, which are already generalized to multiple d.o.f., but the link here with information theory gives us further insight.
+In fact, it allows to understand the following picture.
+
+Suppose the overall motion is deterministc and reversible but that each particle is undergoing non deterministic motion.
+Suppose $\rho$ is a distribution at equilibrium of such motion.
+While each particle moves randomly, the overall distribution $\rho$ stays the same.
+The non deterministic motion is effectively just reshuffling the particles within the same distribution.
+
+Suppose now $\rho$ evolves adiabatically.
+At each instant we can assume it to be at equilibrium.
+At each instant the particles are reshuffled but the amount of entropy of the distribution remains the same.
+The overall evolution is still Hamiltonian, but the particle trajectories are now averages.
+The value of the entropy of the distribution is set by the non deterministic motion of the particles.
+
+If we assumed that all materials were interacting with the same background that caused all particles to exhibit a random walk of the same magnitude, as defined by the information entropy, then the classical uncertainty principle would apply to all materials with the same constant.
+It would also be in line to address some of the problems mentioned in \ref{infinitesimal_reducibility}, as determinism and reversibility at all scales conceptually does not work.
+
+It should now be clear that in classical mechanics we can already find some of the ingredients that are typically assumed to be characteristics of just quantum mechanics.
+Starting from well defined physical concepts is what allowed us to make such connection in simple and precise terms.
+Classical particles are not point-like objects, they are the limit of infinitesimal reductionism, and any finite amount of material is a distribution over phase space.
+Quantum particles are made of a finite amount of material, therefore will inherit some of their behavior from classical distributions.
+We are now ready to see where they get the remaining behavior.
 

### 2016-08-26 2015CompleteDerivation/completeDerivation.tex
Another pass for quantum prelude

@@
 
-As we saw in the previous sections, classical particles should not be understood as point-like finite mass objects as this would open a series of problems and inconsistencies.
+As we saw in the previous sections, classical particles should not be understood as point-like objects of finite mass as this would open a series of problems and inconsistencies.
 The issues created by such misidentification are only amplified within the context of quantum mechanics.
@@
 But they don't always behave like one, they also behave like a wave, a distribution.
-During measurements, it coincides experimentally with a probability distribution.
-So the state of a quantum particles is a probability distribution.
+During measurements, this distribution coincides experimentally with a probability distribution.
+So the state of a quantum particle is a probability distribution.
 Except that it self-interacts.
-So its a weird quantum probability distribution.
-The quantum particle behaves like its in more than one place at once.
+So its a special quantum probability distribution as the quantum particle behaves like its in more than one place at once.
 We learn this picture and, sometimes begrudgingly, we get used to it.
 
-Here we contend that this picture is not only confusing and unnecessary, it also prevents to understand fully the deep links between classical and quantum mechanics, in the same way that thinking of classical particles as point-like objects prevented to see the holistic picture in classical mechanics.
-
-The first thing to note here is that all objects behave like they are in more than one place at once.
+Here we contend that this picture is not only confusing and unnecessary, it also prevents to understand fully the deep links between classical and quantum mechanics, in the same way that thinking of classical particles as point-like objects prevented to see a more holistic picture in classical mechanics.
+
+The first thing to note is that all physical objects behave like they are in more than one place at once.
 The earth is both under Milano and New York.
-Your right index finger is not, hopefully, at the same position of your his left eye.
+Your right index finger is not, hopefully, at the same position of your left eye.
 In fact, based on what we saw before, any object of finite mass is a distribution over a region, as discontinuous distributions are ruled out.
@@
 Why should it be any different?
-In fact, shouldn't the burden of proof be on the claim that the electron is point-like, unlike any other finite mass object?
-
-Part of the confusion seem to be the idea is that the electron is an elementary particle, and therefore should have no spatial extent.
+In fact, if one claimed an electron to be point-like, he would be claiming an electron behaves in a different way than all other physical objects.
+Shouldn't \emph{he} have the burden of proof?
+
+Part of the confusion seem to be the idea that, since the electron is an elementary particle, it should have no spatial extent.
 But that does not follow.
@@
 In fact: we can construct an infinitesimal object that is not elementary.
-Suppose you have dots constrained on the surface of a sphere.
+Suppose we have a sphere with dots moving on its surface.
 All their dynamics can be expressed by their polar and azimuthal angles.
@@
 Not only point-like object don't have to be elementary, but it's not clear how they can be elementary.
-How can the whole universe be reduced to elementary elements that have no idea how many spatial dimensions are there?
-Mathematically: the set of points by itself does not define the topology.
-How is it supposed to work?
+How can the whole universe be reduced to  elements that have no idea how many spatial dimensions are there?
+Mathematically a set of points by itself does not define the dimension of a space: $\mathbbm{R}^n$ is the same order of infinity for any $n$.
+One needs a topology, which is an additional structure.
+If everything is just points, where does this come from?
 
@@
 
-The first payoff of accepting that an electron is a finite amount of material distributed in space is that this is more consistent with classical mechanics as we have defined it\footnote{It is also more consistent with general relativity.}.
-The second payoff is that some features of classical mechanics already explain part of the behavior that we thought was inherently quantum.
+The first payoff of accepting that a quantum particle is a finite amount of material distributed in space is that this is more consistent with classical mechanics as we have defined it\footnote{It is also more consistent with general relativity.}.
+The second payoff is that some features of classical mechanics already explain part of the behavior that is usually thought as inherent to quantum mechanics.
 
@@
 Suppose we have a finite amount of classical material distributed over a single d.o.f.
-$a_m = \int \rho_\mathcal{c} dq^n \wedge dp_n$.
+$a_m = \int \rho_\mathcal{c} dq \wedge dp$.
 The Shannon entropy associated with the distribution, which can be though as the amount of information required to identify a particle of the distribution, is finite as well.
-$I(\rho_\mathcal{c}) = \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq^n \wedge dp_n = I_0$.
+$I(\rho_\mathcal{c}) = \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq \wedge dp = I_0$.
 TODO: prove in appendix.
 
-As the state evolves into its final state, the amount of material remains the same: $a_m = \int \rho_{\hat{\mathcal{c}}} dq^n \wedge dp_n$.
+As the state evolves into its final state, the amount of material remains the same: $a_m = \int \rho_{\hat{\mathcal{c}}} dq \wedge dp$.
 The final entropy will also remain the same since the entropy, like the distribution itself, is just another integrable function: $I(\rho_{\hat{\mathcal{c}}}) = I_0$.
@@
 \rho = &\frac{\rho_{tot}}{ \sqrt{4 \pi ^2\sigma_q^2\sigma_p^2}} e^{-\frac{(q-\mu_q)^2(p-\mu_p)^2}{4\sigma_q^2\sigma_p^2}} \\
-= &\frac{\rho_{tot}}{e^{H_0 - 1}} e^{-\frac{\pi ^2(q-\mu_q)^2(p-\mu_p)^2}{\exp (2(H_0 - 1))}}
+= &\frac{\rho_{tot}}{e^{H_0 - 1}} e^{-\frac{\pi ^2(q-\mu_q)^2(p-\mu_p)^2}{\exp (2(I_0 - 1))}}
 \end{align*}
 We find that the distribution that minimizes the spread is the product of two independent Gaussians.
-In quantum mechanics, the Gaussian wave packet is the state that minimizes uncertainty: this is the classical analogue.
-
-As the distribution evolves, the entropy is conserved and therefore the product $\sigma_q^2 \sigma_p^2$ can never be less the one given by the Gaussian distribution of the same entropy. We have:
+Recall that in quantum mechanics, the Gaussian wave packet is the state that minimizes uncertainty: this is the classical analogue.
+
+As the distribution evolves, the entropy is conserved and therefore the product $\sigma_q^2 \sigma_p^2$ can never be less than the one given by the Gaussian distribution of the same entropy. We have:
 \begin{align*}
-\sigma_q\sigma_p \geq \exp (H_0 - 1) / 2 \pi 
+\sigma_q\sigma_p \geq \exp (I_0 - 1) / 2 \pi 
 \end{align*}
-This constitute the classical equivalent of the Heisenberg uncertainty principle.
+This is strikingly similar to the Heisenberg uncertainty principle, except that the value depends on the initial state. This can be addressed with some extra assumptions.
+
+Suppose the evolution of the distribution is deterministc and reversible but that each particle is undergoing non deterministic motion.
+Suppose $\rho$ is a distribution at equilibrium of such motion.
+While each particle moves randomly, the overall distribution $\rho$ stays the same: the non deterministic motion is effectively just reshuffling the particles within the same distribution.
+The value of the entropy of the distribution is set by the non deterministic motion of the particles: the stronger it is, the greater the entropy.
+
+Suppose now $\rho$ evolves adiabatically.
+At each instant we can assume $\rho$ to be at equilibrium.
+At each instant the particles are reshuffled but the amount of entropy of the distribution remains the same.
+The overall evolution is still Hamiltonian, but the trajectories predicted by the framework are now the average motion of the particles.
+
+If we assumed that all materials were interacting with the same background that caused all particles to exhibit a random walk of the same magnitude, as defined by the information entropy, then the spread of all distributions for all material will satisfy the same inequality.
+This is gives us the classical uncertainty principle: $\sigma_q\sigma_p \geq \exp (I_\varnothing - 1) / 2 \pi $ where $I_\varnothing$ is the entropy associated with the background.
+
+Assuming such a background is not far fetched.
+In fact, it serves as a remainder of the issues mentioned in assumption \ref{infinitesimal_reducibility}: as determinism and reversibility at all scales conceptually does not work, there exists a minimum constant interaction to all particles: the system is not perfectly isolated.
+If the entropy of our distribution is comparable to the entropy introduced by the background, we are likely in the range where classical mechanics is not valid as the random motion of each particle is not negligible.
 
@@
 is outside of the scope of this work.
-Note that this result is similar in spirit to the Gromov non-squeezing theorem and the symplectic camel, which are already generalized to multiple d.o.f., but the link here with information theory gives us further insight.
-In fact, it allows to understand the following picture.
-
-Suppose the overall motion is deterministc and reversible but that each particle is undergoing non deterministic motion.
-Suppose $\rho$ is a distribution at equilibrium of such motion.
-While each particle moves randomly, the overall distribution $\rho$ stays the same.
-The non deterministic motion is effectively just reshuffling the particles within the same distribution.
-
-Suppose now $\rho$ evolves adiabatically.
-At each instant we can assume it to be at equilibrium.
-At each instant the particles are reshuffled but the amount of entropy of the distribution remains the same.
-The overall evolution is still Hamiltonian, but the particle trajectories are now averages.
-The value of the entropy of the distribution is set by the non deterministic motion of the particles.
-
-If we assumed that all materials were interacting with the same background that caused all particles to exhibit a random walk of the same magnitude, as defined by the information entropy, then the classical uncertainty principle would apply to all materials with the same constant.
-It would also be in line to address some of the problems mentioned in \ref{infinitesimal_reducibility}, as determinism and reversibility at all scales conceptually does not work.
-
-It should now be clear that in classical mechanics we can already find some of the ingredients that are typically assumed to be characteristics of just quantum mechanics.
-Starting from well defined physical concepts is what allowed us to make such connection in simple and precise terms.
-Classical particles are not point-like objects, they are the limit of infinitesimal reductionism, and any finite amount of material is a distribution over phase space.
-Quantum particles are made of a finite amount of material, therefore will inherit some of their behavior from classical distributions.
-We are now ready to see where they get the remaining behavior.
+Note, though, that this result is similar in spirit to the Gromov non-squeezing theorem and the symplectic camel, which are already generalized to multiple d.o.f..
+It is the link with information theory that gives us further physical insight.
+
+It should now be clear how in classical mechanics we can already find some of the ingredients that are typically assumed to be characteristics of quantum mechanics.
+It has its own flavor of anti-particles, CPT theorem and uncertainty principle.
+These are general features of an evolution that is deterministic, reversible and adiabatic (i.e.
+Hamiltonian dynamics).
+Quantum mechanics, as we'll see, just inherits these traits.
 
 \section{Quantum systems}
+
+We now focus on a different type of homogeneous material: one that is irreducible.
+That is, the state and the evolution of the whole does not tell us anything about the state and the evolution of the parts.
+As the internal dynamics is undetermined, we'll find that each irreducible element is suitably described by the standard framework of quantum mechanics.
 

### 2016-08-30 2015CompleteDerivation/completeDerivation.tex
Using namespace for labels?

@@
 
-\begin{assump}[Determinism and reversibility]\label{detrevass}
+\begin{assump}[Determinism and reversibility]\label{ass:determinism}
 The state of the physical system under study undergoes deterministic and reversible evolution.
@@
 
-\begin{prop}\label{differentiable_manifold}
+\begin{prop}\label{prop:differentiable_manifold}
 	The state space $\mathcal{S}$ for the particles of a  classical material is a differentiable manifold.
@@
 
-\begin{assump}[Infinitesimal reducibility]\label{infinitesimal_reducibility}
+\begin{assump}[Infinitesimal reducibility]\label{ass:infinitesimal_reducibility}
 	The system under study is composed of an infinitesimally reducible homogeneous material and each part undergoes deterministic and reversible evolution.
@@
 
-	Consistently with what we said in Assumption \ref{detrevass}, if we defined a state for each particle, then a deterministic and reversible evolution on such state must exist.
+	Consistently with what we said in assumption \ref{ass:determinism}, if we defined a state for each particle, then a deterministic and reversible evolution on such state must exist.
 Yet, the idea that we can assign states to infinitesimal parts should be considered only a \emph{simplifying} assumption.
@@
 
-	The second, more conceptual, problem is that the idealizations that allowed us to define states in Assumption \ref{detrevass} may not hold as the parts get smaller.
+	The second, more conceptual, problem is that the idealizations that allowed us to define states in assumption \ref{ass:determinism} may not hold as the parts get smaller.
 Recall the cannonball whose motion is sufficiently unaffected by the photons that scatter off its surface.
@@
 Let $t : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be a time variable.
-	The restriction of $\rho_{\bar{\mathcal{c}}}$ over an arbitrary level set $\mathcal{S}_{t=t_0}$ is continuous and integrable by \ref{differentiable_manifold}.
+	The restriction of $\rho_{\bar{\mathcal{c}}}$ over an arbitrary level set $\mathcal{S}_{t=t_0}$ is continuous and integrable by \ref{prop:differentiable_manifold}.
 $\rho_{\bar{\mathcal{c}}} \in L^1_t(\bar{\mathcal{S}}, \mu)$ Time evolution is a diffeomorphism by \ref{hamiltons_equations}.
@@
 	We claim $\bar{\mathcal{C}} \cong  C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$.
-As in \ref{differentiable_manifold}, if two distributions are different there exists an outcome that can tell them apart, therefore they represent two distinct physical evolutions.
+As in \ref{prop:differentiable_manifold}, if two distributions are different there exists an outcome that can tell them apart, therefore they represent two distinct physical evolutions.
 \end{justification}
@@
 Let $\rho : T^*\mathcal{M} \rightarrow \mathbbm{R}$ be an invariant density.
-$\rho$ is differentiable by \ref{differentiable_manifold}.
+$\rho$ is differentiable by \ref{prop:differentiable_manifold}.
 It's restriction on $\bar{\mathcal{S}}$ is also an invariant distribution and therefore differentiable.
@@
 
-\begin{prop}\label{antistates}
+\begin{prop}\label{prop:antistates}
 	Let $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ the time dependent deterministic and reversible evolution map for the particles of a classical material.
@@
 
-\begin{assump}[Kinematic equivalence]\label{kinematic_equivalence}
+\begin{assump}[Kinematic equivalence]\label{ass:kinematic_equivalence}
 	For the given classical material, studying its kinematics  (i.e. trajectories in space) is equivalent to study its dynamics  (i.e. states and their evolution).
@@
 $\rho_{\bar{\mathcal{c}}}(\mathcal{s}(x^\alpha, u^\alpha))$ as the material is minimal.
-$\rho_{\bar{\mathcal{c}}}(x^\alpha, u^\alpha)$ is differential for the same arguments expressed in \ref{differentiable_manifold}.
+$\rho_{\bar{\mathcal{c}}}(x^\alpha, u^\alpha)$ is differential for the same arguments expressed in \ref{prop:differentiable_manifold}.
 $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is a diffeomorphism as it as a homeomorphism that maps a differentiable function to a differentiable function.
@@
 The extended Hamiltonian we found is quadratic in energy, therefore negative energy states are possible.
-These correspond to the particle anti-states mentioned in \ref{antistates}.
+These correspond to the particle anti-states mentioned in \ref{prop:antistates}.
 As we fixed the natural parametrization, $u^\alpha u_\alpha = - c^2$, there are two cases: the parametrization is aligned with time, therefore $s=\tau$ where $\tau$ is proper time, or the parametrization is anti-aligned with time, therefore $s=-\tau$.
@@
 Assuming such a background is not far fetched.
-In fact, it serves as a remainder of the issues mentioned in assumption \ref{infinitesimal_reducibility}: as determinism and reversibility at all scales conceptually does not work, there exists a minimum constant interaction to all particles: the system is not perfectly isolated.
+In fact, it serves as a remainder of the issues mentioned in assumption \ref{ass:infinitesimal_reducibility}: as determinism and reversibility at all scales conceptually does not work, there exists a minimum constant interaction to all particles: the system is not perfectly isolated.
 If the entropy of our distribution is comparable to the entropy introduced by the background, we are likely in the range where classical mechanics is not valid as the random motion of each particle is not negligible.

### 2016-09-09 2015CompleteDerivation/completeDerivation.tex
Relativity review with Christine

@@
 We'll also find that particle states divide into standard and anti-states depending whether time is aligned or anti-aligned with the evolution parameter.
-Note that no extra assumption is needed, which makes relativistic mechanics simply the correct time dependent description of the deterministic and reversible evolution of an infinitesimally reducible system, and not the description of an altogether different physical object.
+Note that no extra assumption is needed, which makes relativistic mechanics simply the correct time dependent description of the deterministic and reversible evolution of an infinitesimally reducible system.
 
@@
 
-The discussion so far has been limited to the case where both the set of states and the evolution map are defined at equal time and never change throughout the evolution.
+The discussion so far has been limited to the case where both the set of states and the evolution map never change throughout the evolution.
 This is too restrictive as there are very reasonable situations in which this does not hold.
@@
 The second case is when we perform a transformation for the time parameter $\hat{t}=\hat{t}(t)$.
-The equations of motion transform and we have $d_{t}\hat{t} d_{\hat{t}}q^i = \partial_{p_i} H$.
-As both $q^i$ and $H$ are independent of time, they cannot be redefined to include $d_{t}\hat{t}$.
+The equations of motion transform and we have $d_{t}\hat{t} \, d_{\hat{t}}q^i = \partial_{p_i} H$.
+$q^i: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ and $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ are real functions of the state space and cannot be redefined to include $d_{t}\hat{t}$.
 Physically the transformation is introducing fictitious forces that are not conservative, so they cannot be expressed by a Hamiltonian.
@@
 Locally, each state will be given by $(q^i, k_i, t)$ therefore $\bar{\mathcal{S}}$ is a $2n+1$ dimensional manifold.
-A region at equal time will be a hyper-surface $\mathcal{S}_{t=t_0}$ that includes all possible configuration at a particular time.
-Such hyper-surface will need to cut across all evolutions once and only once, therefore not all functions of $\bar{\mathcal{S}}$ are suitable time variables.
+A region at equal time will be a hyper-surface $\mathcal{S}_{t=t_0} \subset \bar{\mathcal{S}}$ that includes all possible configurations at a particular time.
+Such hyper-surface will need to cut across all evolutions once and only once, therefore not all functions $t : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ are suitable time variables.
 
@@
 Locally $\bar{\mathcal{S}} \supseteq U \cong\mathbbm{R}^{2n+1}$.
-$\bar{\mathcal{S}}$ is an $2n+1$ dimensional manifold.
+$\bar{\mathcal{S}}$ is a $2n+1$ dimensional manifold.
 $T^*\mathcal{Q}$ is a differentiable manifold.
@@
 $\mathcal{S}_{t=t_0}$ is an embedded submanifold of co-dimension $1$.
-$\mathcal{S}_{t=t_0}$ is an hyper-surface.
+$\mathcal{S}_{t=t_0}$ is a hyper-surface.
 \end{justification}
@@
 A distribution of material over states at equal time $\rho_\mathcal{c} : \mathcal{S}_{t=t_0} \rightarrow \mathbbm{R}$ will depend on the choice of time variable, therefore the complete state space will not be useful as it will not be time invariant.
-What we are really interested in is the \emph{evolution space} of a classical material, which we'll indicated by $\bar{\mathcal{C}}$.
+What we are really interested in is the \emph{evolution space} of a classical material, which we'll indicate by $\bar{\mathcal{C}}$.
 To each evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$ will correspond a distribution $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ over all particle states at all times.
-This distribution will be differentiable, it's differentiable over all equal time hyper-surfaces and time evolution is differentiable, but is not integrable over the complete state space $\bar{\mathcal{S}}$.
+This distribution will be differentiable: it's differentiable over all equal time hyper-surfaces and time evolution is differentiable.
+The distribution, though, is not integrable over the complete state space $\bar{\mathcal{S}}$.
 Such integration will operate over time as well, giving the amount of material multiplied by time.
@@
 
-Before proceeding, it's useful to get a better intuition for the geometry of these space.
-Note that the transported state variables are perfect to label the particle state evolution evolutions $\bar{\mathcal{s}} \subset \bar{\mathcal{S}}$: as they do not change during the evolution we can take them to label not just the values at a particular time, but the evolutions themselves.
+Before proceeding, it's useful to get a better intuition for the geometry of these spaces.
+Note that the transported state variables are perfect to label the particle state evolution $\bar{\mathcal{s}} \subset \bar{\mathcal{S}}$: as they do not change during the evolution we can take them to label not just the values at a particular time, but the evolutions themselves.
 The distribution is transported deterministically and reversibly over those lines, so the value of the distribution will be constant along each particle state evolution $\bar{\mathcal{s}}$.
-To identify the full evolution, then, we just need to specify the distribution only on a region that cuts across all particle evolution, i.e.
+To identify the full evolution, then, we just need to specify the distribution only on a region that cuts across all particle evolutions, i.e.
 a hyper-surface at constant time.
@@
 The co-vector $k_i dq^i$ is completed with a time component $\bar{\omega} dt$.\footnote{$\bar{\omega}$ is the classical analogue of the wave frequency.
-We use $\bar{\omega}$ to distinguish from the phase-space metric $\omega$.} As state variable $q^i$ have a conjugate quantity $p_i\equiv \hbar k$, the time variable $t$ will have a conjugate quantity $E\equiv\hbar\bar{\omega}$.
+We use $\bar{\omega}$ to distinguish from the phase-space metric $\omega$.} As each state variable $q^i$ has a conjugate quantity $p_i\equiv \hbar k$, the time variable $t$ will have a conjugate quantity $E\equiv\hbar\bar{\omega}$.
 We call the combination of $(t, E)$ the temporal degree of freedom.
-As time is special, so is the temporal degree of freedom, which is treated differently when defining distribution $\rho_\mathcal{c}$ and therefore by the measure $\mu$ and the metric $\omega$.
+As time is special, so is the temporal degree of freedom, which is treated differently when defining a distribution $\rho_\mathcal{c}$ and therefore by the measure $\mu$ and the metric $\omega$.
 
@@
 Therefore there must exist a constraint such that, locally, $E=E(t,q^i,p_i)$.
-The complete state space set $\bar{\mathcal{S}}$ is a hyper-surface of $T^*\mathcal{M}$.
+The complete state space $\bar{\mathcal{S}}$ is a hyper-surface of $T^*\mathcal{M}$.
 
@@
 That is: the plane where we can properly count states and define our metric is perpendicular to $dt \wedge dE$.
+Intuitively, states are defined independently of time, therefore they are defined on a surface perpendicular to the temporal d.o.f.
 
@@
 The invariance of $\omega$ already guarantees that.
-As we saw before that fact that the vector product is conserved within a degree of freedom is equivalent to saying that changing variables does not change the possibilities within a d.o.f.
-In fact, the flow of an infinitesimal transformation that preserves the vector product on a surface is a divergence free field.
-The conservation of the Minkowsky product across the temporal and standard d.o.f.
+As we saw before, the fact that the vector product is conserved within a degree of freedom is equivalent to saying that changing variables does not change the possibilities within a d.o.f.
+In fact, the infinitesimal displacement $d_s x^i$ for a general infinitesimal transformation that preserves the vector product on a surface is a divergence free field.
+The conservation of the Minkowski product across the temporal and standard d.o.f.
 means that changing variables does not introduce rotation between time and state variables.
-In fact, the flow of an infinitesimal transformation that preserves the Minkowsky product on a surface is a curl free field.
+In fact, the infinitesimal displacement $d_s x^i$ for a general infinitesimal transformation that preserves the Minkowski product on a surface is a curl free field.
 
@@
 \begin{prop}\label{relativistic_symplectic_manifold}
-	The complete state space $\hat{\mathcal{S}}$ for the particles of a classical material is a hyper-surface of the symplectic manifold formed by a cotangent bundle $T^*\mathcal{M}$ equipped with the canonical two-form $\omega = \sum dq^i \wedge dp_i - dt \wedge dE = \sum dq^\alpha \wedge dp_\alpha$.
-\end{prop}
-
-\begin{justification}
-	We claim the space of physical objects that allows a time-state-invariant-variable densities is a symplectic manifold $(T^*\mathcal{M}, \omega)$ where $\omega = \sum dq^i \wedge dp_i - dt \wedge dE$.
+	The complete state space $\bar{\mathcal{S}}$ for the particles of a classical material is a hyper-surface of the symplectic manifold formed by a cotangent bundle $T^*\mathcal{M}$ equipped with the canonical two-form $\omega = \sum dq^i \wedge dp_i - dt \wedge dE = \sum dq^\alpha \wedge dp_\alpha$.
+\end{prop}
+
+\begin{justification}
+	We claim the space of physical objects that allows time-and-state-variable-invariant densities is a symplectic manifold $(T^*\mathcal{M}, \omega)$ where $\omega = \sum dq^i \wedge dp_i - dt \wedge dE$.
 Let $\mathcal{M}$ be the manifold that defines the unit system including time.
@@
 Let $\bar{\omega}$ be the co-vector component associated with $t$.
-Let $E=\hbar \bar{\omega}$.
+Let $E\equiv\hbar \bar{\omega}$.
 Let $\omega$ be the metric.
@@
 	We claim $\bar{\mathcal{S}}$ is a hyper-surface of $(T^*\mathcal{M}, \omega)$.
-$\bar{\mathcal{S}} \subseteq (T^*\mathcal{M}, \omega)$ as states are physical objects that allow time-state-invariant-variable densities.
+$\bar{\mathcal{S}} \subseteq (T^*\mathcal{M}, \omega)$ as states are physical objects that allow time-and-state-variable-invariant densities.
 Consider the map $f : T^*\mathcal{M} \rightarrow \bar{\mathcal{S}} \; | \; (q^i, p_i, t, E) \mapsto (q^i, p_i, t)$.
@@
 $\rho$ is differentiable by \ref{prop:differentiable_manifold}.
-It's restriction on $\bar{\mathcal{S}}$ is also an invariant distribution and therefore differentiable.
+Its restriction on $\bar{\mathcal{S}}$ is also an invariant distribution and therefore differentiable.
 The inclusion map $\bar{\mathcal{S}} \hookrightarrow T^*\mathcal{M}$ is a smooth embedding.
 $\bar{\mathcal{S}}$ is an embedded submanifold of co-dimension $1$.
-$\bar{\mathcal{S}}$ is an hypersurface.
+$\bar{\mathcal{S}}$ is a hypersurface.
 \end{justification}
@@
 The trajectory of a particle in the extended phase space will be given by the evolved variables $\xi^a(s)$ in terms of a parameter $s$.
-As before, deterministc and reversible evolution will preserve the metric as the number of possibilities on each independent d.o.f.
+As before, deterministic and reversible evolution will preserve the metric as the number of possibilities on each independent d.o.f.
 will be conserved.
@@
 \begin{justification}
-We claim the vector field $S \in T(T^*\mathcal{M})$ for the infinitesimal displacement $S^a = d\xi^a/ds$ admits a potential $\mathcal{H}$ such that $S^{a} \omega_{ab} = \partial_{b}\mathcal{H}$.
+We claim the vector field $S \in T(T^*\mathcal{M})$ for the infinitesimal displacement $S^a = d_s\xi^a$ admits a potential $\mathcal{H}$ such that $S^{a} \omega_{ab} = \partial_{b}\mathcal{H}$.
 The map for infinitesimal evolution $\mathcal{T}_{ds}$ is an infinitesimal self-symplectomorphism by \ref{relativistic_canonical_transformation}.
@@
 
-We claim the state variables evolve according to the extended Hamilton's equations.
+We claim the state variables evolve according to the extended Hamilton equations.
 $S^{a} \omega_{ab} = d_s\xi^a \omega_{ab} = \partial_{b}\mathcal{H}$.
@@
 \begin{prop}\label{prop:antistates}
-	Let $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ the time dependent deterministic and reversible evolution map for the particles of a classical material.
+	Let $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ be the time dependent deterministic and reversible evolution map for the particles of a classical material.
 The map partitions the extended state space into standard states, those connected by a trajectory where $d_{s}t>0$, and anti-states, those connected by a trajectory where $d_{s}t<0$.
@@
 	We claim $d_{s}t$ partitions the space.
-Let $\mathcal{s}_1, \mathcal{s}_2 \in T^*\mathcal{M}$ two physical states connected by deterministic and reversible evolution.
+Let $\mathcal{s}_1, \mathcal{s}_2 \in T^*\mathcal{M}$ be two physical states connected by deterministic and reversible evolution.
 $\mathrm{sign}(d_{s}t(\mathcal{s}_1)) = \mathrm{sign}(d_{s}t(\mathcal{s}_2))$.
@@
 The partial derivatives of $H$ and $\mathcal{H}$ are related, and therefore the functions themselves are related.
-Working through the math, we find that the most general relationship is of the form: $\mathcal{H} = \lambda(t,E,q^i,p_i) \, (H(t,q^i,p_i) - E - \mathcal{h}(t))$ where $\lambda = d_s t$, $\mathcal{h}(t)$ is an arbitrary function and, most importantly, $E = H(t,q^i,p_i) + \mathcal{h}(t)$.
+Working through the math, we find that the most general relationship is of the form: $\mathcal{H} = \lambda(t,q^i,E,p_i) \, (H(t,q^i,p_i) - E - \mathcal{h}(t))$ where $\lambda = d_s t$, $\mathcal{h}(t)$ is an arbitrary function and, most importantly, $E = H(t,q^i,p_i) + \mathcal{h}(t)$.
 
@@
 Therefore $\mathcal{H}=0$ only because $E = H$.
-We can extend $\mathcal{H}$ on all of $T^* \mathcal{M}$ making sure that $\mathcal{H} \neq 0$ outside of the complete state space $\bar{\mathcal{S}}$.
+We can extend $\mathcal{H}$ on all of $T^* \mathcal{M}$ such that $\mathcal{H} \neq 0$ outside of the complete state space $\bar{\mathcal{S}}$.
 This way we can directly use the constraint $\mathcal{H}=0$ to identify $\bar{\mathcal{S}}$ as a subspace of $T^* \mathcal{M}$.
@@
 \begin{prop}\label{form_of_extended_hamiltonian}
-	The form of the invariant Hamiltonian is $\mathcal{H} = \lambda(t,E,q^i,p_i) \, (H(t,q^i,p_i) - E)$ where $H(t,q^i,p_i)$ is the standard Hamiltonian at each instant $t$, the conjugate time variable $E = H(t,q^i,p_i)$ represents the value of the Hamiltonian, and  $\lambda : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ where $\lambda(\xi^a) = d_s t$.
+	The form of the invariant Hamiltonian is $\mathcal{H} = \lambda(t,q^i,E,p_i) \, (H(t,q^i,p_i) - E)$ where $H(t,q^i,p_i)$ is the standard Hamiltonian at each instant $t$, the conjugate time variable $E = H(t,q^i,p_i)$ represents the value of the Hamiltonian, and  $\lambda : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ where $\lambda(\xi^a) = d_s t$.
 $\bar{\mathcal{S}}$ is the level set for $\mathcal{H} = 0$.
@@
 $d_s p_i = - \partial_{q^i} \mathcal{H} = d_s t \, d_t p_i = \lambda \, \partial_{q^i} H$.
-As $\lambda \neq 0$ for all physical states, we can set $\mathcal{H} = \lambda(t,E,q^i,p_i) \, (H(t,q^i,p_i) - E + \mathcal{h}(t,E,q^i,p_i))$ without loss of generality.
-We have $\partial_{q^i} \lambda (H - E + \mathcal{h}) + \lambda (\partial_{q^i} H + \partial_{q^i} \mathcal{h}) = \lambda \, \partial_{q^i} H$.
+As $\lambda \neq 0$ for all physical states, we can set $\mathcal{H} = \lambda(t,q^i,E,p_i) \, (H(t,q^i,p_i) - E + \mathcal{h}(t,q^i,E,p_i))$ without loss of generality.
+We have $\partial_{q^i} \lambda \, (H - E + \mathcal{h}) + \lambda \, (\partial_{q^i} H + \partial_{q^i} \mathcal{h}) = \lambda \, \partial_{q^i} H$.
 This holds for all choices of $t$.
-In particular for $t=s$, $\lambda (\partial_{q^i} H + \partial_{q^i} \mathcal{h}) = \lambda \, \partial_{q^i} H$.
+In particular for $t=s$, $\lambda \, (\partial_{q^i} H + \partial_{q^i} \mathcal{h}) = \lambda \, \partial_{q^i} H$.
 $\partial_{q^i} \mathcal{h} = 0$.
@@
 $\mathcal{h}$ is not a function of $p_i$.
-$\lambda = d_s t = - \partial_E \mathcal{H} = - \partial_{E} \lambda (H - E + \mathcal{h}) - \lambda (-1 + \partial_{E} \mathcal{h})$.
+$\lambda = d_s t = - \partial_E \mathcal{H} = - \partial_{E} \lambda (H - E + \mathcal{h}) - \lambda \, (-1 + \partial_{E} \mathcal{h})$.
 In particular for $t=s$, $\lambda = - \lambda (-1 + \partial_{E} \mathcal{h})$.

### 2016-09-11 2015CompleteDerivation/completeDerivation.tex
First part for quantum

@@
 \section{Composite systems and reducibility}
+\label{sec:reducibility}
 
@@
 
-\subsection{Homogeneous decomposable systems and vector spaces}
+\subsection{Homogeneous decomposable systems and real vector spaces}
 The notion that a system is decomposable means the states are equipped with a rule of composition that allows one to write $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$: the composite system is the sum of its parts.
@@
 
-\begin{prop}\label{reducible_state_space}
+\begin{prop}\label{prop:abelian_group}
 The state space $\mathcal{C}$ for a decomposable homogeneous material is an additive abelian (i.e. commutative) group.
@@
 \begin{justification}
-We claim $\mathcal{C}$ is an abelian group. $\mathcal{C}$ is the state space for a decomposable homogeneous system. $\mathcal{C}$ is an abelian group  by \ref{reducible_state_space}.
+We claim $\mathcal{C}$ is an abelian group. $\mathcal{C}$ is the state space for a decomposable homogeneous system. $\mathcal{C}$ is an abelian group  by \ref{prop:abelian_group}.
 
@@
 
-\section{Quantum systems}
-
-We now focus on a different type of homogeneous material: one that is irreducible.
-That is, the state and the evolution of the whole does not tell us anything about the state and the evolution of the parts.
+\section{Irreducible systems}
+
+In this section we turn to a different case in which the relationship between the whole material and its parts is different from the one described in section \ref{sec:reducibility}.
+The system is still infinitesimally decomposable, made of infinitesimal parts as before, but this time it is irreducible.
+That is: the state and the evolution of the whole does not tell us anything about the state and the evolution of the parts.
 As the internal dynamics is undetermined, we'll find that each irreducible element is suitably described by the standard framework of quantum mechanics.
 
+\subsection{Quantized material}
+
+In section \ref{sec:reducibility} we introduced the idea of a classical material, one that is homogeneous and infinitesimally reducible, which represented a way to relate the state of the whole system to the state of its parts.
+In this section we explore an alternative case.
+
+We will call a \emph{quantized material} a homogeneous material that is finitely reducible.
+That is, as we decompose the system we can only assign a state to finite amounts of material, which we call \emph{quantum particle}.
+As the material is homogeneous, all its particles carry the same amount of material and are described by the same state space.
+Each quantum particle is irreducible: its state does not tell us anything about the state of its parts.
+While this idea should be familiar enough, it is best to iron out some details to avoid possible misconceptions.
+
+The first possible source of confusion is between divisibility, the idea that an object can be divided into parts, and reducibility, the idea that the state of the whole is equivalent to the state of the parts.
+Most of the time these two concepts overlap, but they are separate as we can see in the following examples.
+
+Consider the a planarian flatworm.
+If we cut it into multiple pieces, each part will grow into a new worm.
+Yet, the state of each worm is not the state of multiple worms put together.
+The planarian is divisible into multiple worms but it is not reducible to multiple worms.
+Consider a magnet.
+Its state is described, to a first approximation, by the position of the north and south poles.
+Yet, we cannot divide the magnet into a north and south pole.
+The magnet is reducible to the poles but is not divisible into the poles.
+
+Consider a muon.
+It will decay into an electron and two neutrinos.
+But the state of a muon is not the state of an electron and two neutrinos.
+The muon is divisible into the three particles but it is not reducible into them.
+Consider a proton.
+Its state is given by the state of its three valence quarks.
+Yet, we cannot divide the proton into the three quarks.
+The proton is reducible to quarks, but divisible into quarks.
+
+The fundamental property of a quantum particle, as introduced here, is not indivisibility: it's irreducibility. It is the state that cannot be further partitioned, not the material itself.
+
+Even more subtle is the difference between divisibility and decomposability, the idea that the object has parts.
+While we are agnostic on whether a quantum particle can be divided, we are implicitly assuming that it can be decomposed.
+Irreducibility describes the relationship between the state of the whole and the state of the parts, therefore parts are implicit in the definition.
+While these parts must exist, we cannot isolate them and act on them independently: if we were able to do that we could assign them a state.
+
+While we must be able to talk about parts of a quantum particles, not all questions will make sense.
+Can we ask what fraction of a quantum particle is in a particular region of space?
+Yes, since the particle is decomposed some part may be inside the volume.
+Can we ask where that same fraction will be at a later time?
+No, since the particle is irreducible we cannot describe the evolution of a part.
+
+The proper way to think of a quantized material is therefore the following: a continuum of infinitesimal parts clumped together to form quantum particles, the smallest amount of material to which we can assign a state.
+With this picture in mind, we can see, even before developing the math, how some features typical of quantum behaviors emerge.
+
+The first feature is the importance of statistics in the quantum particle state.
+Consider a part of a quantum particle in an initial state.
+As we cannot track its motion, it can be any part of the particle in its final state.
+That is, the motion of a part of a particle is a random walk within the evolution of a particle.
+
+The second feature is non locality.
+Suppose we have a quantum particle state distributed in space.
+Suppose we exert a force only on a subregion of its distribution.
+As the particle is irreducible, we cannot interact with only part of it: we affect also the part where no force is exerted.
+We cannot simply interact with the part at a specific region, as this would allow to assign a state to that part.
+
+Wave particle duality should be clear at this point.
+The quantum particle behaves like a wave in the sense that its parts form a distribution in space.
+Yet, it behaves as a single unit, as the state is one unit and cannot be separated.
+
+The most striking feature, though, is that we can predict the functional form for particle decay.
+Suppose a quantum particle spontaneously decays into two or more particles.
+Such process cannot be deterministic as the d.o.f.
+increase, therefore is not determined by the state of the quantum particle.
+As the decay is spontaneous, it is not triggered by the environment (i.e.
+external state).
+It is therefore triggered by the internal unstated dynamic of the parts.
+As we assumed the internal dynamic is unknown, we cannot learn anything about it by looking at the product of the decay.
+In particular, all states have to have equal chance of decay or we could tell that the internal dynamic is more stable for some states.
+The probability of decay is constant in time.
+The probability of survival follows the exponential distribution.
+
+As we see qualitatively how the features typical of quantum mechanics naturally emerge from this picture, we are ready to make this more quantitatively and mathematically precise.
+
+\subsection{Quantum particle and complex vector spaces}
+
+Turning these concept into a precise mathematical formulation is tricky as we need to formally characterize the unstated part which is not physically distinguishable.
+While many previous results do not apply, the classical derivation still can serve as a template.
+
+As in the classical case, the quantum particle is thought to be decomposable: different parts will be in different places, the evolution of the whole is the composition of the evolution of parts.
+Therefore \ref{prop:abelian_group} applies and the state space $\Psi$ of an irreducible material is an abelian group.
+
+Note, though, that we can't think of the parts as independent components.
+As the unstated part is statistical in nature, it will combine as statistical variables do.
+Suppose $X$ and $Y$ are two random variables, $\sigma^2_{X+Y} = \sigma^2_{X} + \sigma^2_{Y} + 2 \, \sigma_{X} \sigma_{Y} \rho_{X,Y}$, where $\sigma^2$ is the variance, $\sigma$ is the standard deviation, and $\rho_{X,Y}=COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient.
+As $-1\leq\rho_{X,Y}\leq+1$, the total variance can be greater or smaller then the sum of the independent parts, and the strength of the uncertainty within the unstated part is something physically distinguishable.
+
+Therefore, while the motion of the unstated part is not itself physically relevant, the correlation between the motion of the unstated part of two combined parts is very relevant.
+We'll see how interference can be understood in this context: it is constructive if the randomness is correlated while it is destructive if it is anti-correlated.
+
+So, how can we characterized these relationship between different configurations of the unstated parts?
+We can look at the classical derivation for guidance.
+In that case, we looked at the set of transformation that increase/decrease the amount of material.
+Those are the only type of transformation that we can apply on a classical system that do not change the overall state given by the distribution.
+
+As the classical material is homogeneous, permutations between classical particles in the same state $\mathcal{s}$ do not matter: the ordering is unstated.
+In a way, increasing/decreasing the amount of material increases/decreases the unstated part of the system.
+
+For the irreducible material we follow the same strategy: we find the state transformations that affect the unstated part while keeping the overall state, the distribution, unchanged.
+
+Suppose we have $\psi_1$ a state for an irreducible material characterized by one d.o.f.
+The material will be spread over the configurations of the infinitesimal fragments.
+As the material is homogeneous, the material is spread evenly among the such possible configuration.
+Counting configurations is therefore equivalent to counting amount of material.
+Each fragment would be identified by a pair of independent random variables $(A, B)$, as a d.o.f.
+is two-dimensional.
+The shape of the distribution and the actual values are not physically distinguishable: these are just dummy variables we assign to count the fragment configurations.
+We choose A and B to be uniform distributions from 0 to 1, so that the area is a square of unit size, which represent the unit of unstated  configurations the material is spread over.
+
+Suppose we have another state $\psi_2$, where only unstated part is changed.
+The unstated part can also be characterized by a square formed by two independent uniform distributions (C, D).
+Counting states on (C,D) is equivalent to counting them on (A,B), as they are simply different arrangements.
+As all distributions are uniform, they can only be connected by a linear operation.
+That is: $C=aA+bB$ and $D= -bA + aB$.
+We can write $\psi_2 = \tau(a,b) \psi_1$ as $a$ and $b$ are the only parameters of the transformation.
+
+The transformation $\tau(a,b)$ increases/decreases the area by the factor $a^2+b^2$ and rotates the square by $arctan(x/y)$.
+If we set $c=a+\imath b$ those same quantities can be expressed as $\rho^2=|c|^2$ and $\theta = arg(z)$, the modulus and the phase of a complex number.
+In other words: complex numbers provide a representation for the group of rotations and magnitude rescaling.
+
+For multiple independent d.o.f., we have one square each.
+As the material is irreducible, all areas need to be always the same since relative changes would allow to learn something about the motion of the infinitesimal fragments.
+This means that the transformation $\tau(c)$ has to be the same for all d.o.f.
+In other words, the group of transformations is the same regardless of the number of independent d.o.f.
+
+Mathematically, the state space $\Psi$ for an irreducible material is a complex vector space.
+
+\begin{prop}\label{prop:complex_vector_space}
+	The state space $\Psi$ for an irreducible material is a vector space over $\mathbbm{C}$.
+\end{prop}
+
+\begin{justification}
 	TODO
+\end{justification}
+
+If $\mathcal{Q}$ is the space that defines our units, a state $\psi$ must tell us for each point $\mathcal{q} \in \mathcal{Q}$ how much material is there and the the correlation of the unstated part.
+That is: the state will be identified by $\psi : \mathcal{Q} \rightarrow \mathbbm{C}$, a complex function over the $\mathcal{Q}$.
+
+The square modulus at each point will give us the density of the material.
+As in the classical case, such density mush be integrable to give us the total amount of material.
+Therefore the function $\psi \in L^2(\mathcal{Q}, \mu)$ is square integrable.
+The norm $|\psi|^2 = \int_{\mathcal{Q}} |\psi(q)|^2 dq$ will give us the amount of material associated with the state.
+
+As the material is homogeneous, all particles of the material are made of the same amount.
+We can set, by convention, the amount of material for a particle to be unitary, therefore quantum particle states are the subset of $\Psi$ such that $|\psi|^2 = 1$.
+Note that changing the phase by the same amount everywhere is only changes the unstated part.
+Such change does not change the state, therefore distribution whose differ by a total phase are not physically distinguishable and represent the same physical state.
+
+\begin{prop}\label{prop:wavefuntion}
+	The state space $\Psi$ for an irreducible material is isomorphic to the space of Lebesgue square integrable differentiable functions over $\mathcal{Q}$.
+That is $\Psi \cong C^1(\mathcal{Q}) \cap L^1(\mathcal{Q}, \mu)$.
+The state space for a quantum particle is the projective space $\mathsf{P}(\Psi)$.
+\end{prop}
+\begin{justification}
+	TODO
+\end{justification}
+
+The observation made after \ref{prop:differentiable_manifold} in classical case are still valid: $\Psi$ is not a complete metric space.
+
+\subsection{Irreducibility}
+
+\subsection{Quantum mechanics}
 

### 2016-09-21 2015CompleteDerivation/completeDerivation.tex
Status update
Adding Christine's feedback

@@
 
-Section V - Mostly done. Needs review.
+Section V - Done.
 
@@
 
-Section VIII - Scoping ideas
+Section VIII - Preliminary work
+
+Section IX - Preliminary work
 
@@
 We can redefine $H = H + \mathcal{h}$ as the result is still a function of $q^i, p_i , t$ with the same equations of motion.
-$\mathcal{H}=\lambda (H - E)$
+$\mathcal{H}=\lambda (H - E)$.
 
 	We claim $E = H(t,q^i,p_i)$ for physical states.
-$\lambda = d_s t = - \partial_E \mathcal{H} = - \partial_{E} \lambda (H - E + \mathcal{h}) - \lambda (-1)$ regardless of the choice of t and therefore $\lambda$.
+$\lambda = d_s t = - \partial_E \mathcal{H} = - \partial_{E} \lambda (H - E) - \lambda (-1)$ regardless of the choice of t and therefore for all $\lambda$.
 $H - E = 0$.
@@
 $\mathcal{H} = \lambda (H - E) = 0$ over $\bar{\mathcal{S}}$.
-Extend $\mathcal{H}$ over $T^* \mathcal{M}$ such that $\mathcal{H} \neq 0$ over $T^*\mathcal{M} \,\backslash\, \bar{\mathcal{S}}$.
+We can extend $\mathcal{H}$ over $T^* \mathcal{M}$ such that $\mathcal{H} \neq 0$ over $T^*\mathcal{M} \,\backslash\, \bar{\mathcal{S}}$.
 $\bar{\mathcal{S}}$ is the level set for $\mathcal{H} = 0$
@@
  &= \frac{1}{2mc^2} (c \sqrt{m^2 c^2 + p^i p_i} + E) \\
- &(c \sqrt{m^2 c^2 + p^i p_i} - E)
+ &(c \sqrt{m^2 c^2 + p^i p_i} - E) \\
+\lambda &= \frac{1}{2mc^2} (c \sqrt{m^2 c^2 + p^i p_i} + E)
 \end{aligned}
 \end{equation}
+Therefore we can see that the invariant Hamiltonian has the form $\mathcal{H}=\lambda (H - E)$.
 In the next section we'll derive the above expression and look more closely at the case of particles under potential forces.
@@
 As the density is transported over each trajectory, its value does not change over $s$.
-Therefore we have $d_s \rho_{\bar{\mathcal{c}}} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \} = 0$.
+Therefore we have $d_s \rho_{\bar{\mathcal{c}}}= d_s \xi^a \partial_a \rho_{\bar{\mathcal{c}}} = \omega^{ab} \partial_b \mathcal{H} \partial_a \rho_{\bar{\mathcal{c}}} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \} = 0$ where $\{f , g\} = \partial_a f \omega^{ab} \partial_b g$ is the Poisson bracket.
 
@@
 Let $\gamma : \mathbbm{R} \rightarrow T^*\mathcal{M}$ be a parametrization for a particle evolution $\bar{\mathcal{s}}$.
-Then $\rho_\mathcal{c}(\lambda(s)) = \rho_\mathcal{c}(\lambda(s + \Delta s))$ as the material does not change over the particle evolution.
-$d_s \rho_\mathcal{c} = 0$.
+Then $\rho_{\bar{\mathcal{c}}}(\lambda(s)) = \rho_{\bar{\mathcal{c}}}(\lambda(s + \Delta s))$ as the material does not change over the particle evolution.
+$d_s \rho_{\bar{\mathcal{c}}} = 0$.
 
@@
 As such, there was no mention of the speed of light $c$, $\mathcal{M}$ being a pseudo-Riemannian manifold or its metric $g$.
-These will be introduced in the next chapter with the kinematic equivalence assumption.
+These will be introduced in the next section with the kinematic equivalence assumption.
 In light of this, it is probably better to consider relativistic mechanics as a feature of the geometry of the complete state space more than the geometry of space-time.
-Also note that this feature arises from the use of the standard topology on $\mathbbm{R}^n$ and the need of invariant densities.
+Also note that this feature arises from the use of the standard topology on $\mathbbm{R}^n$ and the need for invariant densities.
 It would not have arisen with the use of a discrete topology.
 
-We have also seen that anti-particle states, a feature generally associated with quantum field theory, is actually a feature of relativistic Hamiltonian mechanics.
+We have also seen that anti-particle states, generally associated with quantum field theory, are actually a feature of relativistic Hamiltonian mechanics.
 This is usually missed because in classical mechanics only the standard Hamiltonian is typically used.
-Note how the invariant Hamiltonian in \ref{free_hamiltonians} is quadratic in $E$, which can have both positive and negative solutions on the constraint $\mathcal{H}=0$.
-
-Another important feature is that this formulation of relativistic Hamiltonian mechanics is formally very close to the Quantum case.
+Note how the invariant Hamiltonian in \ref{free_hamiltonians} is quadratic in $E$, which can have both positive and negative solutions given the constraint $\mathcal{H}=0$.
+
+Another important feature is that this formulation of relativistic Hamiltonian mechanics is formally very close to the quantum case.
 Note how $\mathcal{H} \rho = 0$ and $\mathcal{H} \ket{\psi} = 0$ are formally equivalent.
-If we apply to the invariant Hamiltonian in \ref{free_hamiltonians} the usual formal classical/quantum substitution, we have the Klein-Gordon equation using space-like convention rescaled by a factor for $2m$.
+If we apply to the invariant Hamiltonian in \ref{free_hamiltonians} the usual formal classical/quantum substitution, we have the Klein-Gordon equation using space-like convention rescaled by a factor of $2m$.
 As we'll see later, to this parallel in the mathematical formalism corresponds a parallel in the physical description.
@@
 The invariant Hamiltonian is needed to identify that subspace.
-This may be a hint that a better formulation is lurking.
-One where we consider the particle evolutions $\bar{\mathcal{s}}$ as the primary objects, label those (instead of the states).
-The distribution of material would simply be a distribution over the particle evolutions, a Lebesgue integrable differentiable functions.
+This may be a hint that a better formulation is lurking, one where we consider the particle evolutions $\bar{\mathcal{s}}$ as the primary objects and label those (instead of the states).
+The distribution of material would simply be a distribution over the particle evolutions, a Lebesgue integrable differentiable function.
 Yet, the need to be able to describe the evolution in terms of physically meaningful state variables will most likely re-impose the extra structure.
@@
 
-We will now focus our attention to the simplest possible classical material: one that is fully described by its motion in space, the internal dynamics can be ignored.
+We will now focus our attention on the simplest possible classical material: one that is fully described by its motion in space, i.e.~the internal dynamics can be ignored.
 What we'll find is that the dynamics of its infinitesimal parts follows classical Lagrangian mechanics.
-Furthermore, the Hamiltonian of the system is constrained to the one of a free particles under scalar and vector potential forces.
+Furthermore, the Hamiltonian of the system is constrained to the one of a free particle under scalar and vector potential forces.
 
@@
 \begin{assump}[Kinematic equivalence]\label{ass:kinematic_equivalence}
-	For the given classical material, studying its kinematics  (i.e. trajectories in space) is equivalent to study its dynamics  (i.e. states and their evolution).
+	For the system under investigation, studying its kinematics  (i.e. trajectories in space) is equivalent to studying its dynamics  (i.e. states and their evolution).
 \end{assump}
@@
 	The first natural question to ask is: why does state evolution always describe a trajectory in space?
-Because spatial extent is the one way in which all materials can be decomposed.
+It's because spatial extent is the one way in which all materials can be decomposed.
 As we divide materials in smaller pieces, physical location is one state variable we can assign to the particles of all materials.

### 2016-09-21 2015CompleteDerivation/completeDerivation.tex
Starting work on inner product and operators

@@
 	The state space $\Psi$ for an irreducible material is isomorphic to the space of Lebesgue square integrable differentiable functions over $\mathcal{Q}$.
-That is $\Psi \cong C^1(\mathcal{Q}) \cap L^1(\mathcal{Q}, \mu)$.
+That is $\Psi \cong C^1(\mathcal{Q}) \cap L^2(\mathcal{Q}, \mu)$.
 The state space for a quantum particle is the projective space $\mathsf{P}(\Psi)$.
@@
 
-The observation made after \ref{prop:differentiable_manifold} in classical case are still valid: $\Psi$ is not a complete metric space.
+\subsection{Inner product and operators}
+
+Both in the classical and quantum state the full state is identified by distributions on $\mathcal{Q}$.
+The $q^i$ do not label the states, but the infinitesimal parts.
+For a classical material, the infinitesimal parts can be given a state, have their own topological space and a set of local state variables $(q^i, p_i)$ to identify them.
+We can use the richness of theory of differentiable manifolds and metric spaces.
+For an irreducible material, the infinitesimal parts cannot be given a state, cannot be identified by a set of local variables.
+We have to develop other mathematical tools.
+
+To be on par with classical mechanics we need at least address three problems.
+We need to be able to compare two quantum states and ask what fraction of them share the same configuration: this will give us an inner product.
+We need to be able to express state variable distributions and expectations: this will give us linear operators.
+We need to identify the conjugate quantities $k_i$ and $p_i$.
+And we need to be mindful that the observations made after \ref{prop:differentiable_manifold} for the  classical case are still valid: $\Psi$ cannot be a complete metric space.
+Let's start with the first.
+
+The idea is that, given two states $\psi_1$ and $\psi_2$, we should be able to tell how much of $\psi_1$ is prepared in the same way of $\psi_2$.
+That is $\psi_1 = \alpha \psi_2 + \psi_3$ is the combination of the part that is prepared like $\psi_2$ (e.g.
+same distributions in position and momentum) and a part that is prepared completely differently.
+Let $\mathsf{P}_{\psi_2} : \Psi \rightarrow \Psi$ be the operator that gives us the equally prepared part $\mathsf{P}_{\psi_2} (\psi_1) \equiv \alpha \psi_2$.
+This operator must be linear in $\psi_1$: combining parts that are equally prepared still gives us parts that are equally prepared.
+It also gives us the same result if applied once or twice: once we isolated the part that is equally prepared, it remains the part that is equally prepared.
+$\mathsf{P}_{\psi_2}$ is therefore a projection: a linear operator such that $\mathsf{P}^2_{\psi_2}= \mathsf{P}_{\psi_2}$.
+
+From the norm and the projection we can define a product $< \cdot |\cdot> : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{<\psi_2|\psi_1>}{|\psi_2|^2} \psi_2$.
+If the arguments are the same, the projection returns the same state, therefore $<\psi_1|\psi_1> = |\psi_1|^2$: the product is positive definite.
+Since the projection is linear, the product is linear as well.
+
+The coefficient $\alpha$ is a function of both $\psi_1$ and $\psi_2$, so we write it as $\alpha = <\psi_2|\psi_1>$.
+It is linear in $\psi_1$ as $\mathsf{P}_{\psi_2}$ is linear.
+The square module $|<\psi_2|\psi_1>|^2$ gives us the amount of material in $\psi_1$ that has the same distribution of $\psi_2$.
+$|<\psi_1|\psi_1>|^2$ As all states
+
+Suppose we are .
 

### 2016-09-26 2015CompleteDerivation/completeDerivation.tex
Feedback from Jiahua

@@
 
-Suppose we study the motion of a cannonball; its state under gravitational and inertial forces will be properly described by the position and momentum of the center of mass.
+Suppose we study the motion of a cannonball; its state under gravitational and (inertial) inertial forces will be properly described by the position and momentum of the center of mass.
 While light and air molecules may scatter off its surface unpredictably, its trajectory is not greatly affected as it is a massive rigid body.
@@
 Setting up an experiment means choosing a particular process with desired outcomes and forcing the system under study to interact with it one or more times.
-After that, there is no special role played by the "observer" in making outcomes come about.
+After that, there is no special role played by the ``observer" in making outcomes come about.
 
@@
 First, to make sure the process is indeed correlated to the states.
-Second, because "a single take does not a measurement make".
+Second, because ``a single take does not a measurement make".
 One has to gather enough statistics.
@@
 Let $\mathcal{s_1}, \mathcal{s_2} \in \mathcal{S}$.
-As states are physically distinguishable, there must exist a physical process with two outcomes $V_1, V_2 \in \mathsf{T}$ for which $s_1 \in V_1, s_2 \in V_2, V_1 \cap V_2 = \varnothing$.
+As states are physically distinguishable, there must exist a physical process with two possible outcomes $V_1, V_2 \in \mathsf{T}$ for which $s_1 \in V_1, s_2 \in V_2, V_1 \cap V_2 = \varnothing$.
 $\mathcal{S}$ is Hausdorff by definition.

### 2016-09-26 2015CompleteDerivation/completeDerivation.tex
Feedback from Christine

@@
 \begin{rationale}
-	The idea is that the system internal dynamics has no effect on the motion and therefore can be disregarded.
+	The idea is that the system internal dynamics has no effect on the overall motion and therefore can be disregarded.
 Consider the motion of our cannonball under gravitational and inertial forces: its temperature or precise chemical composition does not affect the overall trajectory.
 We can assume kinematic equivalence.
-Consider the motion of a helicopter under gravitational and inertial forces: the pilot and the amount of fuel in the tank do affect the trajectory.
+Suppose we now add magnetic forces to our consideration.
+The magnetization of the cannonball may now affect the motion and therefore we cannot always reconstruct the full state by simply looking at the trajectory.
 We cannot assume kinematic equivalence.
@@
 	Kinematic equivalence can't therefore be assumed for all systems.
-Yet, the assumption is fundamental because the particles of any material will \emph{at least} describe a trajectory in space and what we learn about the link between kinematic and dynamical quantities applies to all materials, even the ones where the kinematic assumption does not hold.
-
-	The first natural question to ask is: why does state evolution always describe a trajectory in space?
+Yet, the assumption is fundamental because the particles of any material, during their evolution, will \emph{at least} describe a trajectory in space and what we learn about the link between kinematic and dynamical quantities applies to all materials, even the ones where the kinematic assumption does not hold.
+
+	The first natural question to ask is: why does state evolution of a particle of a classical material always describe a trajectory in space?
 It's because spatial extent is the one way in which all materials can be decomposed.
-As we divide materials in smaller pieces, physical location is one state variable we can assign to the particles of all materials.
+As we divide materials into smaller pieces, physical location is one state variable we can assign to the particles of all materials.
 If we assume kinematic equivalence, the kinematic variables will be enough to identify a particle state.
@@
 Therefore by studying a minimal material, we are also studying a component of the dynamics that is present in all materials.
-In other words: any material is \emph{at least} a minimal material and will possess the dynamic needed to describe its motion in space.
+In other words: any material is an extension to a minimal material.
 
@@
 Therefore $\mathcal{M}$ has conceptually two roles: it is the space where the motion unfolds and it is the space that defines the system of units for our states.
-This already tells us that the our state space is a hyper-surface of $T^*\mathcal{M}$.
+This already tells us that the state space for the particles of the material is a hyper-surface of $T^*\mathcal{M}$.
 
@@
 Because of the double role of $\mathcal{M}$, we can set half of our state variables to be equal to the space-time variables: $q^\alpha = (t, q^i) = x^\alpha$.
-This already tells us that the trajectory is differentiable since $q^i(t)$ is differentiable.
+This already tells us that the trajectory is differentiable since $q^\alpha(s)$ is differentiable.
 The four velocity $u^\alpha = d_s x^\alpha$ is therefore well defined and it can be used to identify trajectories.
@@
 Let $\mathcal{s} \in \bar{\mathcal{S}}$ be the particle state for a minimal classical material.
-Let $x^\alpha(s)$ the evolution associated with that state.
+Let $x^\alpha(s)$ be the evolution associated with that state.
 Let $t(\mathcal{s})$ be the instant of time at which $\mathcal{s}$ is defined.
-$x^\alpha(s(t(\mathcal{s}))) = f(\mathcal{s}) = q^\alpha$ as the initial position cannot be derived by the law of evolution.
+$q^\alpha(\mathcal{s}) \equiv x^\alpha(s(t(\mathcal{s})))$ are state variables as they are real function of the state.
 $q^\alpha$ is differentiable therefore $d_s x^\alpha$ exists.

### 2016-09-26 2015CompleteDerivation/completeDerivation.tex
Adding quantum inner product and projections

@@
 We need to identify the conjugate quantities $k_i$ and $p_i$.
-And we need to be mindful that the observations made after \ref{prop:differentiable_manifold} for the  classical case are still valid: $\Psi$ cannot be a complete metric space.
 Let's start with the first.
@@
 
-From the norm and the projection we can define a product $< \cdot |\cdot> : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{<\psi_2|\psi_1>}{|\psi_2|^2} \psi_2$.
-If the arguments are the same, the projection returns the same state, therefore $<\psi_1|\psi_1> = |\psi_1|^2$: the product is positive definite.
+This projection has other useful properties.
+If $\psi_1$ is already prepared like $\psi_2$, then its projection on $\psi_2$ will be equal to $\psi_1$.
+Conversely, if the projection of $\psi_1$ is equal to itself, then it means that the state is already prepared like $\psi_2$.
+That is: $\mathsf{P}_{\psi_2} (\psi_1) = \psi_1 \Leftrightarrow \psi_1 = \alpha \psi_2$.
+Moreover, because the material is irreducible, if $23\%$ of $\psi_1$ is prepared like $\psi_2$, then $23\%$ of $\psi_2$ is prepared like $\psi_1$.
+If not, we could learn more (or less) about $\psi_1$ by studying the part of $\psi_2$ that is prepared like $\psi_1$.
+Therefore we have $\frac{|\mathsf{P}_{\psi_2} (\psi_1)|^2}{|\psi_1|^2}=\frac{|\mathsf{P}_{\psi_1} (\psi_2)|^2}{|\psi_2|^2}$.
+Lastly, if we take the projection of $\psi_1$ along $\psi_2$ and we project it again on $\psi_1$, we end up with a part of $\psi_1$.
+As this is a proper part of $\psi_1$, its state is just $\psi_1$ rescaled, no change in correlation.
+That is: $\mathsf{P}_{\psi_1} (\mathsf{P}_{\psi_2} (\psi_1)) = a \psi_1$ with $a \in \mathbbm{R}$.
+
+From the norm and the projection we can define a complex product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{\langle \psi_2|\psi_1 \rangle}{|\psi_2|^2} \psi_2$.
+Because of the properties of the norm and the projections, the product is positive definite, linear and conjugate symmetric.
+This means the state space $\Psi$ for an irreducible material is an inner product space.
+
+\begin{prop}\label{prop:inner_product}
+	The state space $\Psi$ for an irreducible material is an inner product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$.
+\end{prop}
+\begin{justification}
+	TODO
+\end{justification}
+
+Note that the inner product only has an indirect physical definition, based on the norm and the projections. We'll therefore want to base our physical intuition on the latter concepts.
+
+Also note that the observations made after \ref{prop:differentiable_manifold} for the  classical case are still valid.
+$\Psi$ cannot be a complete metric space as it does not include discontinuous functions.
+Therefore it is only a pre-Hilbert space.
+For mathematical convenience, one can take its completion with respect to the norm induced by the inner product and obtain a Hilbert space.
+Yet, such object does no longer represent a physically meaningful entity.
+
+As such, we need to make sure that the rest of our definitions and justifications do not require a complete metric space to be valid or they would not be physically meaningful.
+In particular, we cannot write the usual link $\psi(q) = \langle q | \psi \rangle$ between the wave function and the inner product as the material cannot be prepared at a specific value if $q$ is a continuous variable.
+
+%If the arguments are the same, the projection returns the same state, therefore $<\psi_1|\psi_1> = |\psi_1|^2$: the product is positive definite.
 Since the projection is linear, the product is linear as well.
 
-The coefficient $\alpha$ is a function of both $\psi_1$ and $\psi_2$, so we write it as $\alpha = <\psi_2|\psi_1>$.
+%The coefficient $\alpha$ is a function of both $\psi_1$ and $\psi_2$, so we write it as $\alpha = <\psi_2|\psi_1>$.
 It is linear in $\psi_1$ as $\mathsf{P}_{\psi_2}$ is linear.
@@
 $|<\psi_1|\psi_1>|^2$ As all states
-
-Suppose we are .
 
@@
 In this section we include mathematical demonstration in support of the physical arguments of the paper.
+
+\begin{thrm}\label{thrm:function_topology}
+	Let $X$ and $Y$ be Hausdorff and second countable topological spaces.
+Let $C(X, Y) \equiv \{f: X \rightarrow Y \; | \; f$ is continuous$\}$ be the space of continuous functions endowed with the Compact-Open Topology $\mathsf{T}$.
+Then $C(X, Y)$ is a Hausdorff and second countable topological space.
+\end{thrm}
+
+\begin{proof}
+	We claim $C(X, Y)$ is Haudorff. Let $f, g \in C(X, Y)$ such that $f \neq g$. $\exists x \in X$ such that $f(x)\neq g(x)$. Let $K = \{x\}$. $K$ is compact.
+\end{proof}
 
@@
 
-\begin{justification}
+\begin{proof}
 	We claim the vector field $S \in T\mathcal{Q}$ admits a potential $H$ such that $S^{a} \omega_{ab} = \partial_{b}H$.
@@
 $S$ is a curl free vector field and admits a potential $H$ such that $S_{a} = \partial_{a}H$.
-\end{justification}
+\end{proof}
+
+\begin{thrm}
+	Scalar product
+\end{thrm}
+
+\begin{proof}
+	We claim $\langle x , y \rangle = \langle x, \mathsf{P}_x(y)\rangle$.
+Let $x, y \in X$.
+$\langle x , y \rangle x = |x|^2 \mathsf{P}_x(y) = |x|^2 \mathsf{P}_x(\mathsf{P}_x(y)) = \langle x, \mathsf{P}_x(y)\rangle$.
+
+	We claim $\mathsf{P}_x(y) = a x$ for some $a \in \mathbbm{C}$. Let $x, y \in X$. $\mathsf{P}_x(\mathsf{P}_x(y)) = \mathsf{P}_x(y)$. $\mathsf{P}_x(y) = a x$ for some $a \in \mathbbm{C}$
+
+	We claim $\langle a x , y \rangle = a^\dagger \langle x , y \rangle$.
+Let $x, y \in X$ and $a \in \mathbbm{C}$.
+$\langle a x , y \rangle = \langle a x, \mathsf{P}_x(y)\rangle = \langle a x, b x\rangle$ for some $b \in \mathbbm{C}$.
+$\langle a x, b x\rangle = \frac{b}{a} \langle a x, a x\rangle = \frac{b}{a} |a x|^2 = \frac{b}{a} |a|^2 |x|^2 = \frac{a^\dagger ab}{a} \langle x, x\rangle = a^\dagger \langle x, b x\rangle = a^\dagger \langle x, \mathsf{P}_x(y)\rangle = a^\dagger \langle x , y \rangle$
+\end{proof}
 

### 2016-09-30 2015CompleteDerivation/completeDerivation.tex
Christine comments on kinematics

@@
 $\rho_{\bar{\mathcal{c}}}(\mathcal{s}(x^\alpha, u^\alpha))$ as the material is minimal.
-$\rho_{\bar{\mathcal{c}}}(x^\alpha, u^\alpha)$ is differential for the same arguments expressed in \ref{prop:differentiable_manifold}.
+$\rho_{\bar{\mathcal{c}}}(x^\alpha, u^\alpha)$ is differential by the same logic expressed in \ref{prop:differentiable_manifold}.
 $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is a diffeomorphism as it as a homeomorphism that maps a differentiable function to a differentiable function.
@@
 
-This relationship allows us to write the kinematic variables in terms of the state variable.
+This relationship allows us to write the kinematic variables in terms of the state variables.
 In particular we have $u^\alpha=u^\alpha(q^\alpha, p_\alpha)$.
@@
 Under such conditions we can define a Legendre transformation $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$.
-Such transformation is the Lagrangian of the system and the laws of motions are given by the Euler-Lagrange equation.
-For the time independent case, we define the \emph{Lagrangian} $L=v^i p_i - H$ where $v^i=d_t x^i$.
+Such transformation is the Lagrangian of the system and the motion is given by the Euler-Lagrange equation.
+For the time independent case, we define $L=v^i p_i - H$ where $v^i=d_t x^i$.
 
@@
 	We claim $\mathcal{H}$ is convex in $p_\alpha$.
-By \ref{initial_conditions} we can write the velocity $u^\alpha=u^\alpha(q^\alpha, E, p_i)$ as a differentiable function of the state variables.
+By \ref{initial_conditions} we can write the velocity $u^\alpha=u^\alpha(q^\alpha, p_\alpha)$ as a differentiable function of the state variables.
 At fixed $q^\alpha$, $u^\alpha$ must be invertible.
@@
 
-	We claim there exist a Lagrangian $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$. $\mathcal{H}$ is convex in $p_\alpha$. It admits a Legendre transform. The Lagrangian $\mathcal{L}$ is such transform.
-\end{justification}
-
-Not only we must be able to express distributions in terms of both state and kinematic variables, but integration has to carry over as well.
+	We claim there exists an extended Lagrangian $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$.
+$\mathcal{H}$ is convex in $p_\alpha$.
+It admits a Legendre transform.
+The extended Lagrangian $\mathcal{L}$ is such transform.
+\end{justification}
+
+Not only must we be able to express distributions in terms of both state and kinematic variables, but integration has to carry over as well.
 To do that, we need to be able to count trajectories in the same way we are able to count states.
-There must exist a metric $g$, consistent with $\omega$, that allows to calculate the cardinality of possibilities for initial conditions.
+There must exist a metric $g$, consistent with $\omega$, that allows one to calculate the cardinality of possibilities for initial conditions.
 That is: $\omega(dq^\alpha,dp_\alpha) \propto g(dx^\alpha,du^\alpha)$.
 
-Since the both $dx^\alpha$ and $du^\alpha$ are four vectors and the metric must be linear, $g$ is a rank two tensor on $\mathcal{M}$.
+Since both $dx^\alpha$ and $du^\alpha$ are four vectors and the metric must be linear, $g$ is a rank two tensor on $\mathcal{M}$.
 To each direction in space-time $dx^\alpha$ is associated a direction in the tangent space of velocities $du^\alpha$ with which it forms a d.o.f..
-Across d.o.f.
-we have $\omega(dq^\alpha, dp_\beta) = 0 =\omega(dq^\beta, dp_\alpha)$.
+Across d.o.f., that is $\alpha \neq \beta$, we have $\omega(dq^\alpha, dp_\beta) = 0 =\omega(dq^\beta, dp_\alpha)$.
 Therefore if $dx^\alpha$ and $du^\beta$ do not form a d.o.f., $g(dx^\alpha, du^\beta) = 0 = g(dx^\beta, du^\alpha)$: $g$ is reflexive.
@@
 We can further change the units along the spatial directions so that $g_{ii}=1$: if we use the same units for the spatial coordinates, each direction will contribute the same number of possibilities.
-Since time must use a different unit, we set $g_{00}=-c^2$ where $c^2$ is the constant that allows to convert the possibilities over the temporal d.o.f.
+Since time must use a different unit, we set $g_{00}=-c^2$ where $c^2$ is the constant that allows one to convert the possibilities over the temporal d.o.f.
 into the possibilities of a spatial d.o.f..
@@
 	We claim $(\mathcal{M}, g)$ is a pseudo-Riemannian manifold where $g$ is a metric induced by $\omega$ under kinematic equivalence.
-Let $(dx^\alpha, du^\alpha)$ an infinitesimal range of possibilities for the kinematic variables of the particles of a minimal classical material.
+Let $(dx^\alpha, du^\alpha)$ be an infinitesimal range of possibilities for the kinematic variables of the particles of a minimal classical material.
 There exists a map $g : T\mathcal{M} \times T\mathcal{M} \rightarrow \mathbbm{R}$ such that $g(dx^\alpha, du^\alpha) \propto \omega(dq^\alpha, dp_\alpha)$.
@@
 $dq^\alpha = dx^\alpha$.
-$g_{ii} > 0$ and $g_{00}$.
+$g_{ii} > 0$ and $g_{00} < 0$.
 $g$ has signature $(3, 1)$.
@@
 Directions in space are orthogonal not because the angle between them is $90^\circ$, but because they identify independent d.o.f..
-That is, the choice of units along one direction is independent from the choice of unit along orthogonal directions.
+That is, the choice of units along one direction is independent from the choice of units along orthogonal directions.
 Space and time, instead, are not independent d.o.f.
 and the geometry between them is different.
-In other words: $g$ inherits it's signature from $\omega$.
+In other words: $g$ inherits its signature from $\omega$.
 The geometry of space-time is the geometry of the ranges for possible states.
 
-This also allows to give $c$ a more insightful physical meaning.
+This also allows us to give $c$ a more insightful physical meaning.
 It is not a velocity: the metric is defined at each point without motion.
 It is the constant of proportionality that maps a range of initial conditions from the time variable $t$ to the space variable $x^i$.
-In the same way that $1$km would correspond to a thousand times more possible states than $1$m, $1$sec corresponds to $299,792,458$ more possible states than $1$m.
-Therefore when we write $x^\alpha = (ct, x^i)$ or $p_\alpha = (- E/c, p_i)$ we are really expressing everything in quantities that are proportional to the cardinality of state and possibilities.
-These coordinates make the book keeping of number of states much easier which, as we saw, is the only thing that Hamiltonian and Lagrangian mechanics are doing.
-
-This also allows to understand conceptually why the speed of deterministic and reversible motion cannot be greater than $c$.
+For example, once the SI is adopted, $1$km would correspond to a thousand times more possible states than $1$m, and, in the same way, $1$sec would correspond to $299,792,458$ more possible states than $1$m.
+Therefore when we write $x^\alpha = (ct, x^i)$ or $p_\alpha = (- E/c, p_i)$ we are really expressing everything in quantities that are proportional to the cardinality of states and possibilities.
+These coordinates make the bookkeeping of number of states much easier which, as we saw, is the only thing that Hamiltonian and Lagrangian mechanics are doing.
+
+This also allows us to understand conceptually why the speed of deterministic and reversible motion cannot be greater than $c$.
 Consider the trajectory of a point particle.

### 2016-09-30 2015CompleteDerivation/completeDerivation.tex
Proof for covariant operator is derivation

@@
 
+\begin{thrm}
+	Let $C^1(\mathcal{Q}, \mathbbm{C})$ the space of differentiable complex functions defined over $\mathcal{Q}$.
+Let $q : \mathcal{Q} \rightarrow \mathbbm{R}$ a differentiable function.
+Let $K_q$ a covariant linear operator: given an invertible differentiable function $\hat{q} : \mathbbm{R} \rightarrow \mathbbm{R}$ $K_{\hat{q}} = \partial_{\hat{q}} q K_{q}$.
+Then $K_q = a \partial_q$ for some $a \in \mathbbm{C}$.
+\end{thrm}
+
+\begin{proof}
+	We claim $K_q(f(q))= \partial_q f K_q(q)$.
+Let $I : \mathbbm{R} \rightarrow \mathbbm{R}$ be the identify function, that is $I \circ f = f \; \forall f : \mathbbm{R} \rightarrow \mathbbm{R}$.
+$K_q(q) = K_q(I(q)) = \partial_q I(q) K_{I}(I) = K_{I}(I)$ by covariance.
+	Let $f : \mathbbm{R} \rightarrow \mathbbm{R}$ be a differentiable invertible function.
+$K_q(f(q)) =K_q(I(f(q))) = \partial_q f K_{f}(I(f)) = \partial_q f \partial_f I(f) K_{I}(I) = \partial_q f K_q(q)$.
+Now let $f : \mathbbm{R} \rightarrow \mathbbm{R}$ be a differentiable function non necessarily invertible.
+It can be written as $f = f_1 + f_2$ with $f_1, f_2$ differentiable and invertible.
+By linearity $K_q(f)=K_q(f_1) + K_q(f_2) = \partial_q f_1 K_q(q) + \partial_q f_2 K_q(q)  = \partial_q f K_q(q)$.
+
+	We claim $K_q$ is a derivation.
+Let $f,g,h$ be differentiable functions such that $h(q)=f(q)g(q)$.
+$K_q(ln(h(q)))=\partial_q ln(h(q)) K_q(q) = \frac{1}{h} \partial_q h K_q(q) = \frac{1}{h} K_q(h(q)) = K_q(ln(f(q)g(q))) = K_q(ln(f(q))) + K_q(ln(g(q))) = \frac{1}{f} K_q(f(q)) + \frac{1}{g} K_q(g(q))$.
+Multiplying by $h=fg$ we have $K_q(h(q)) = g K_q(f(q)) + f K_q(g(q))$.
+$K_q$ is a linear operator that satisfies the product rule.
+$K_q$ is a derivation by definition.
+
+	We claim $K_q = a \partial_q$ for some $a \in \mathbbm{C}$.
+$K_q$ is a derivation.
+As $K_q(f(q))= \partial_q f K_q(q)$ it is a derivation along $q$, which can be expressed as $K_q(q)=a \partial_q$ for some $a \in \mathbbm{C}$.
+\end{proof}
+
 \begin{thebibliography}{0}

### 2016-09-30 2015CompleteDerivation/completeDerivation.tex
More work on quantum

@@
 Note the same group of transformations is found by fixing $A$ and changing $p$ such that the laws of motion don't change.
-In other words: $A$ and $p$ are defined relatively to the each other.
+In other words: $A$ and $p$ are defined relatively to each other.
 $A$ defines the $0$ for $p$ and vice versa.
@@
 \begin{prop}\label{prop:inner_product}
-	The state space $\Psi$ for an irreducible material is an inner product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$.
+	The state space $\Psi$ for an irreducible material is an inner product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q) \psi_2(q) dq$.
 \end{prop}
@@
 In particular, we cannot write the usual link $\psi(q) = \langle q | \psi \rangle$ between the wave function and the inner product as the material cannot be prepared at a specific value if $q$ is a continuous variable.
+For the same reason, we cannot show that to each state variable is associated an hermitian operator by constructing it from its eigenstates.
+
+The proceed in the following way.
+Suppose we have a state variable $f : \mathcal{Q} \rightarrow \mathbbm{R}$.
+Consider the product $f (q) \psi (q)$: it returns the value of $f$ weighted by the wave function.
+We can think of $f$ as an operator that takes a complex function of $\mathcal{Q}$ and returns another complex function of $\mathcal{Q}$.
+The operator is linear since $f (q) (a\psi_1 + b\psi_2) = a f \psi_1 + b f  \psi_2$.
+The operator is hermitian since $f^\dagger = f$.
+Consider $\langle \psi | F \psi \rangle = \langle F \psi | \psi \rangle = \int_\mathcal{Q} \psi^\dagger (q) f(q) \psi(q) dq = f(q) | \psi(q)|^2$.
+It integrates the value of $f$ weighted by the amount of the material: it is the expectation of the state variable $f$ over the distribution.
+
+While $\mathcal{Q}$ is enough to define the system of units, it cannot be enough to define all the variables over which we can define the distributions.
+The conjugate variables $p_i=\hbar k_i$ cannot be expressed as a function of $q^i$, which reiterates the idea that we cannot specify all variables for each fragment.
+Suppose there is another set of state variables that provide another way to subdivide the fragments.
+Then those variables will form a manifold $\hat{\mathcal{Q}}$ such that each $\psi$ can be expressed as a wave function $\psi(\hat{q})$ over $\hat{\mathcal{Q}}$.
+
+As each state $\psi$ can be identified by a wavefunction on either space, there must be an operator $\Upsilon^q_{\hat{q}}(\psi(q)) = \psi(\hat{q})$ that converts the wave function on one space to the other.
+As to a linear combination of states corresponds a linear combination of wave function, $\Upsilon^q_{\hat{q}}$ is a linear operator.
+As converting the space again must give us the original wave function, $\Upsilon^{\hat{q}}_{q} \Upsilon^q_{\hat{q}}= 1$ is unitary.
+We can now define a have linear operator $\hat{F}$ corresponding to state variable $\hat{f} : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
+Using $\Upsilon^{\hat{q}}_{q}$ we can express $\hat{F}$ as an operator over complex functions of $\mathcal{Q}$ It will be still be linear and hermitian.
+
+That is: for each state variable we have a corresponding linear hermitian operator which can be expressed as function multiplication in a suitable choice of $\hat{\mathcal{Q}}$.
+
+\begin{prop}\label{prop:self_adjoint_operators}
+	For each fragment parameter $f$ there exist a corresponding self-adjoint linear operator $F : \Psi \rightarrow \Psi$.
+There exists a manifold $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
+There exist an injective map $\Psi_{\hat{\mathcal{Q}}} : \Psi \rightarrow C^1(\hat{\mathcal{Q}}, \mathbbm{C})$.
+\end{prop}
+\begin{justification}
+	TODO
+\end{justification}
+
+The last thing to do is identify the operators corresponding to $k_i$ and $p_i$, expressed over the space of functions of $\mathcal{Q}$.
+We know that the operators have to be covariant: under a change of variables $K_{\hat{q}} = \partial_{\hat{q}} q K_{q}$.
+This is essentially the chain rule: the operator we are looking for is derivation by $q$.
+As it has to be self-adjoint, it will be of the form $K_{q} = a \imath \partial_q$ where $a$ is a real number.
+We chose $a$ to be negative by convention and to be unitary so that $K_{q}$ is expressed in units of inverse $q$ as in the classical case.
+Therefore we have $K_{i} = -\imath \partial_i$ and $P_{i} = -\imath \hbar \partial_i$.
 

### 2016-09-30 2015CompleteDerivation/completeDerivation.tex
Comments from Jiahua

@@
 $U$ is an open set in the (final) state space topology by definition.
-Consider the process $P_0$ that evolves the initial states and then distinguishes the final state with $P$.
+Consider the process $P_0$ that first evolves the initial states and then distinguishes the final state with $P$.
 $P_0$ is a process that distinguishes initial states.
@@
 \begin{justification}
-	Same justification for continuity as in \ref{detrevmap} with "initial states" and "final states" replaced by "elements" of $\mathcal{S_1}$ and $\mathcal{S_2}$ respectively.
+	Same justification for continuity as in \ref{detrevmap} with ``initial states" and ``final states" replaced by "elements" of $\mathcal{S_1}$ and $\mathcal{S_2}$ respectively.
 \end{justification}
@@
 The choice of spin is independent from the rest.
-The choice of $l$, though, depends on the choice of $n$: for $n=1$ only $l=s$ is available; for $n=2$ we can choose $l=s$ or $l=p$.
+The choice of $l$, though, depends on the choice of $n$: for $n=1$ only $l=0$ is available; for $n=2$ we can choose $l=0$ or $l=1$.
 The choices are not independent.
@@
 \begin{prop}\label{independent_state_variables}
-	Two state variables $q_1 : U \rightarrow \mathbbm{L}_1$ and $q_2 : U \rightarrow \mathbbm{L}_2$ are said \emph{independent} if $\exists \mathcal{s} | q_1(\mathcal{s})=l_1, q_2(\mathcal{s})=l_2 \forall l_1 \in q_1(U), l_2 \in q_2(U)$.
+	Two state variables $q_1 : U \rightarrow \mathbbm{L}_1$ and $q_2 : U \rightarrow \mathbbm{L}_2$ are said \emph{independent} if $\exists \mathcal{s}$ such that $q_1(\mathcal{s})=l_1, q_2(\mathcal{s})=l_2 \forall l_1 \in q_1(U), l_2 \in q_2(U)$.
 \end{prop}
@@
 Giving the state of the whole system is equivalent to giving the states of all particles.
-Given the state space $\mathcal{S}$ for the infinitesimal parts, which we assume to be a manifold consistently with \ref{continuous_state_space}, a composite state $\mathcal{c} \in \mathcal{C}$ will tell us the amount of material for each possible particle state.
+Given the state space $\mathcal{S}$ for the infinitesimal parts, which we assume consistently with \ref{continuous_state_space} to be a manifold, a composite state $\mathcal{c} \in \mathcal{C}$ will tell us the amount of material for each possible particle state.
 That is, each state is fully identified by a function $\rho_\mathcal{c}: \mathcal{S} \to \mathbbm{R}$ that returns the amount of material in the composite state $\mathcal{c}$ that is prepared in the particle state $\mathcal{s}$.
@@
 This is inconsistent with our previous definition.
-Moreover if the transformation is not differential, the density is not even well-defined.
+Moreover if the transformation is not differentiable, the density is not even well-defined.
 This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows state-variable-independent densities.
@@
 Therefore a manifold that guarantees densities to be well defined is one that guarantees that state variable transformations are differentiable: a differentiable manifold.
-$\mathcal{S}$ has a differentiable structure in the sense that the distributions and the volume element $d\mu$ are only defined on a set of state variables that are linked by differential transformations.
+$\mathcal{S}$ has a differential structure in the sense that the distributions and the volume element $d\mu$ are only defined on a set of state variables that are linked by differentiable transformations.
 In the same way that discontinuous changes of state variable can be detected because they tamper with physical distinguishability (i.e.
@@
 placing material at particular states) is a common way to define the references for a coordinate system.
-In a coordinate system where $q^1=\rho_\mathcal{c}$, the function is clearly differential as $\partial_1 \rho_\mathcal{c} = 1$ and $\partial_i \rho_\mathcal{c} = 0$ for $i \neq 1$.
+In a coordinate system where $q^1=\rho_\mathcal{c}$, the function is clearly differentiable as $\partial_1 \rho_\mathcal{c} = 1$ and $\partial_i \rho_\mathcal{c} = 0$ for $i \neq 1$.
 As we change state variables, the transformation is differentiable, therefore all the derivatives $\partial_{i'} q^1 = \partial_{i'} \rho_\mathcal{c}$ exist.
@@
 \begin{justification}
-	We claim $\mathcal{S}$ is a differential manifold.
+	We claim $\mathcal{S}$ is a differentiable manifold.
 Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c} \in \mathcal{C}$.
@@
 We first define conjugate momentum as $p_i=\hbar k_i$ and unified state variables as $\xi^a\equiv \{q^i, p_i\}$.
-Then we consider the canonical two-form $\omega$ given by the following components:
+Then we consider the canonical two-form $\omega =\Sigma \, dq^i \wedge dp_i$ given by the following components:
 \begin{align*}
@@
 The canonical one-form $\theta=k dq$ represents the geometrical object we associate with each particle state.
-The canonical two-form $\omega=\Sigma \, dq^i \wedge dp_i$ is the metric that quantifies the possibilities described by two given state variables.
+The canonical two-form $\omega$ is the metric that quantifies the possibilities described by two given state variables.
 Note, instead, that the relationship $\omega = - \hbar d\theta$, while mathematically true, has no clear physical meaning as we have not defined what the exterior derivative on a state actually represents.
@@
 Let $\mathcal{Q}$ be the manifold charted by $q$.
-A diffeomorphism $\hat{q}=\hat{q}(q)$ fully defines a unit transformation.
+A diffeomorphism $\hat{q}=\hat{q}(q)$ fully defines a change of units.
 Let $(q,k^i)$ with $i=1...n-1$ be a set of state variables.
@@
 
-	We claim the simplest state space $\mathcal{S}$ that allows state-variable-invariant densities is the symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of a one dimensional differential manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
+	We claim the simplest state space $\mathcal{S}$ that allows state-variable-invariant densities is the symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of a one dimensional differentiable manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
 Express integration in coordinates.
@@
 
-	We claim the state space $\mathcal{S}$ for the particles of a homogeneous classical material is a symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of an $n$-dimensional differential manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
+	We claim the state space $\mathcal{S}$ for the particles of a homogeneous classical material is a symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of an $n$-dimensional differentiable manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
 Let $q^i$ be a set of n continuous independent state variables that define the units necessary to describe a state in $U \subseteq \mathcal{S}$.
@@
 The components of $\omega$ across different d.o.f.
-must be zero as they are not invariant under unit transformations.
+must be zero as they are not invariant under change of units.
 $\omega=\sum dq^i \wedge dp_i \pm dt \wedge dE$.
@@
 $\rho_{\bar{\mathcal{c}}}(\mathcal{s}(x^\alpha, u^\alpha))$ as the material is minimal.
-$\rho_{\bar{\mathcal{c}}}(x^\alpha, u^\alpha)$ is differential by the same logic expressed in \ref{prop:differentiable_manifold}.
+$\rho_{\bar{\mathcal{c}}}(x^\alpha, u^\alpha)$ is differentiable by the same logic expressed in \ref{prop:differentiable_manifold}.
 $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is a diffeomorphism as it as a homeomorphism that maps a differentiable function to a differentiable function.

### 2016-09-30 2015CompleteDerivation/completeDerivation.tex
Including figures for d.o.f.

@@
 defined by $(q^i, k_i)$.
-
-\begin{figure}
-	\includestandalone[width=\columnwidth]{pictures/independentDof}
-	\caption{The metric $\omega$ and its components along two independent degrees of freedom.
-As we change state variables, the components are still orthogonal to each other.
-Note: this is a 2D conceptualization of what happens in 4D.}
-	\label{fig:independent_dof}
-\end{figure}
 
@@
 
+\begin{figure}
+	\includestandalone[width=\columnwidth]{pictures/independentDof}
+	\caption{The metric $\omega$ and its components along two independent degrees of freedom.
+As we change state variables, the new d.o.f.
+are also independent and therefore orthogonal to each other.
+Note: this is a 2D conceptualization of 4D space.}
+	\label{fig:independent_dof}
+\end{figure}
+
 This gives us insight on the physical meaning of the geometrical structure $T^*\mathcal{Q}$.
@@
 \end{align*}
+
+\begin{figure}
+	\includestandalone[width=\columnwidth]{pictures/temporalDof}
+	\caption{The metric $\omega$ and its components along temporal and standard degrees of freedom.
+The two d.o.f.
+are not independent and therefore not orthogonal.
+The temporal d.o.f.
+is orthogonal to the metric as states are defined (and counted) independently of time.
+Compare with FIG.~\ref{fig:independent_dof}.
+Note: this is a 2D conceptualization of 4D space.}
+	\label{fig:temporal_dof}
+\end{figure}
 

### 2016-09-30 2015CompleteDerivation/pictures/temporalDof.tex
Including figures for d.o.f.

\documentclass{standalone}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{calculator}
\begin{document}
	\begin{tikzpicture}[scale=0.4]

		% INPUTS:
		\newcommand{\avar}{5};		% omega horizontal length
		\newcommand{\bvar}{6};		% height of first dt dE 
		\newcommand{\dvar}{4};			 % height of second dt dE
		\newcommand{\xvar}{20};			% space between the two

		% outer rectangle and name of diagram
		\draw (-8,-7) rectangle (34,12.5);

		% the different points that will be connected to draw the shapes
		\coordinate (p2) at (\avar,0);
		\coordinate (p3) at (\avar,\bvar);
		\coordinate (p5) at (\avar + \bvar,\bvar);	
	    \coordinate (p7) at (\avar,- \avar);
	    \coordinate (p8) at (\avar - \bvar,\avar + \bvar);

		% angle made by the coordinates to rotate the first dq dp square
		\newcommand{\thetavar}{atan2(\bvar , \avar)};

		% drawing the first shape
		\filldraw[black,fill=blue,opacity=0.5] (0,0) -- ($(p2)$) -- ($(p3)$) -- cycle;

		\filldraw[black,fill=blue,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={0:(0,0)}] (0,0) rectangle ($(p7)$) node[pos=.5] {\large $\omega$};

		\draw[black,fill=blue,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar:(0,0)}] (0,0) rectangle ($(p8)$) node[pos=.5, rotate=\thetavar] {\large $dq\wedge dp$};

		\draw[black,fill=blue,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={0:(0,0)}] ($(p2)$) rectangle ($(p5)$) node[pos=.5] {\large $dt\wedge dE$};

		% coordinates to the second shape (to the right)
		\coordinate (h0) at (\xvar,0);
		\coordinate (h2) at (\xvar + \avar,0); 
		\coordinate (h3) at (\xvar + \avar, \dvar);
		\coordinate (h5) at (\xvar + \avar + \dvar, \dvar);
		\coordinate (h7) at (\xvar + \avar,- \avar);
		\coordinate (h8) at (\xvar + \avar - \dvar,\avar + \dvar);

		% angle made by the coordinates to rotate the first dq dp square
		\newcommand{\phivar}{atan2(\dvar , \avar)};

		%drawing the shape on the right
		\filldraw[black,fill=green,opacity=0.5] ($(h0)$) -- ($(h2)$) -- ($(h3)$) -- cycle;

		\filldraw[black,fill=green,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={0:($(h0)$)}] ($(h0)$) rectangle ($(h7)$) node[pos=.5] {\large $\omega$};

		\draw[black,fill=green,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\phivar:($(h0)$)}] ($(h0)$) rectangle ($(h8)$) node[pos=.5, rotate=\phivar] {\large $dq\wedge dp$};

		\draw[black,fill=green,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={0:($(h0)$)}] ($(h2)$) rectangle ($(h5)$) node[pos=.5] {\large $dt\wedge dE$};

	\end{tikzpicture}
\end{document}

### 2016-10-03 2015CompleteDerivation/pictures/independentDof.tex
Including figures for d.o.f.
No need to use calculator package

@@
 \usetikzlibrary{calc}
-\usepackage{calculator}
 \begin{document}
@@
 	% outer rectangle and diagram name
-	\draw (-8,-12) rectangle (34,14);
-	\node at (13,-8) {\huge Independent d.o.f.};
+	\draw (-8,-8) rectangle (34,12.5);
 
@@
 	% calculate omega from x and y components
-	\SQUARE{\xvar}{\xsquared};
-	\SQUARE{\yvar}{\ysquared};
-	\ADD{\xsquared}{\ysquared}{\sum};	
-	\SQUAREROOT{\sum}{\omegalength};
+	\newcommand{\omegalength}{sqrt(\xvar*\xvar+\yvar*\yvar)}
 
 	% calculate the angle that omega creates w.r.t. the horizontal axis
-	\DIVIDE{\yvar}{\xvar}{\sol};	
-	\ARCTAN{\sol}{\newsol};
-	\RADtoDEG{\newsol}{\thetavar};
+	\newcommand{\thetavar}{atan2(\yvar,\xvar)}
 
-	% calculate y prime and theta prime from x prime 
-	\SQUARE{\omegalength}{\omegasquared};
-	\SQUARE{\xprimevar}{\xprimesquared};
-	\SUBTRACT{\omegasquared}{\xprimesquared}{\dif};
-	\SQUAREROOT{\dif}{\yprimevar};
+	\newcommand{\yprimevar}{sqrt(\omegalength*\omegalength - \xprimevar*\xprimevar)}
 
-	\DIVIDE{\yprimevar}{\xprimevar}{\sol};	
-	\ARCTAN{\sol}{\newsol};
-	\RADtoDEG{\newsol}{\thetaprimevar};
+	\newcommand{\thetaprimevar}{atan2(\yprimevar,\xprimevar)}
 
@@
 
-	\filldraw[fill=red,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar:(0,0)}]  (0,0) rectangle (\omegalength, \omegalength) node[pos=.5, rotate=\thetavar] {\large $\omega$};
+	\filldraw[fill=red,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar:(0,0)}]  (0,0) rectangle ({\omegalength}, {\omegalength}) node[pos=.5, rotate={\thetavar}] {\large $\omega$};
 
 	% drawing the complex right triangle  
-	\filldraw[black,fill=yellow,opacity=0.5,rotate around={\thetavar:(\dvar,0)}] (\dvar + \omegalength,0) -- (\dvar,0) -- ++(-\thetaprimevar: \xprimevar) -- ++(\phiprimevar:\yprimevar);
+	\filldraw[black,fill=yellow,opacity=0.5,rotate around={\thetavar:(\dvar,0)}] ({\dvar + \omegalength},0) -- ({\dvar},0) -- ++({-\thetaprimevar}: {\xprimevar}) -- ++({\phiprimevar}:{\yprimevar});
 
@@
 
-	\filldraw[black,fill=yellow,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar - \thetaprimevar:(\dvar,0)}]  (\dvar + \xprimevar,0) rectangle (\dvar + \xprimevar + \yprimevar, \yprimevar) node[pos=.5, rotate=\thetavar-\thetaprimevar] {\large $d\hat{q}^2\wedge d\hat{p}_2$};
+	\filldraw[black,fill=yellow,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar - \thetaprimevar:(\dvar,0)}]  ({\dvar + \xprimevar},0) rectangle ({\dvar + \xprimevar + \yprimevar}, {\yprimevar}) node[pos=.5, rotate={\thetavar-\thetaprimevar}] {\large $d\hat{q}^2\wedge d\hat{p}_2$};
 
-	\filldraw[black,fill=yellow,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar:(\dvar,0)}]  (\dvar,0) rectangle (\dvar + \omegalength, \omegalength) node[pos=.5, rotate=\thetavar] {\large $\omega$};
+	\filldraw[black,fill=yellow,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar:(\dvar,0)}]  (\dvar,0) rectangle ({\dvar + \omegalength}, {\omegalength}) node[pos=.5, rotate=\thetavar] {\large $\omega$};
 

### 2016-10-03 2015CompleteDerivation/completeDerivation.tex
Cleaned the math for quantum

@@
 \begin{thrm}
-	Scalar product
+	Let $V$ be a complex vector space.
+Let $| \cdot | ^2 : V \rightarrow \mathbbm{R}$ such that $|x|^2 \ge 0 \; \forall x \in V$ and $|x|^2 = 0 \Leftrightarrow x = 0$.
+Let $x \in V$ and $\mathsf{P}_x : V \rightarrow V$ be a family of projections (i.e.
+$\mathsf{P}_x$ linear and $\mathsf{P}_x \circ \mathsf{P}_x = \mathsf{P}_x$) such that: $\mathsf{P}_x (y) = y \Leftrightarrow y = a x$ for some $a \in \mathbbm{C}$; $|y|^2|\mathsf{P}_x (y)|^2=|x|^2|\mathsf{P}_y (x)|^2$; $\mathsf{P}_x (\mathsf{P}_y (x)) = r x$ for some $r \in \mathbbm{R}$.
+	Then $V$ is an inner product space with the product $\langle \cdot , \cdot \rangle : V \times V \rightarrow \mathbbm{C}$ defined by $\langle x , y \rangle x = |x|^2 \mathsf{P}_x (y)$.
 \end{thrm}
@@
 \begin{proof}
-	We claim $\langle x , y \rangle = \langle x, \mathsf{P}_x(y)\rangle$.
-Let $x, y \in X$.
-$\langle x , y \rangle x = |x|^2 \mathsf{P}_x(y) = |x|^2 \mathsf{P}_x(\mathsf{P}_x(y)) = \langle x, \mathsf{P}_x(y)\rangle$.
-
-	We claim $\mathsf{P}_x(y) = a x$ for some $a \in \mathbbm{C}$. Let $x, y \in X$. $\mathsf{P}_x(\mathsf{P}_x(y)) = \mathsf{P}_x(y)$. $\mathsf{P}_x(y) = a x$ for some $a \in \mathbbm{C}$
-
-	We claim $\langle a x , y \rangle = a^\dagger \langle x , y \rangle$.
-Let $x, y \in X$ and $a \in \mathbbm{C}$.
-$\langle a x , y \rangle = \langle a x, \mathsf{P}_x(y)\rangle = \langle a x, b x\rangle$ for some $b \in \mathbbm{C}$.
-$\langle a x, b x\rangle = \frac{b}{a} \langle a x, a x\rangle = \frac{b}{a} |a x|^2 = \frac{b}{a} |a|^2 |x|^2 = \frac{a^\dagger ab}{a} \langle x, x\rangle = a^\dagger \langle x, b x\rangle = a^\dagger \langle x, \mathsf{P}_x(y)\rangle = a^\dagger \langle x , y \rangle$
+	We claim $\langle \cdot , \cdot \rangle$ is well defined.
+Let $x, y \in V$.
+$\mathsf{P}_x (\mathsf{P}_x (y)) =\mathsf{P}_x(y)$ as $\mathsf{P}_x$ is a projection.
+$\mathsf{P}_x(y) = a x $ for some $a \in \mathbbm{C}$ by the first property of $\mathsf{P}_x$.
+$\langle x , y \rangle = |x|^2 a$ exists and is unique.
+
+	We claim $\langle \cdot , \cdot \rangle$ is positive definite.
+Let $x \in V$.
+$\langle x , x \rangle x = |x|^2 \mathsf{P}_x(x) = |x|^2 x $.
+Therefore $\langle x , x \rangle = |x|^2 \ge 0$ and $\langle x , x \rangle = |x|^2 = 0 \Leftrightarrow x = 0$.
+
+	We claim $\langle \cdot , \cdot \rangle$ is linear in the second argument.
+Let $x,y,z \in V$ and $a,b \in \mathbbm{C}$.
+$\langle x , a y + b z \rangle x = |x|^2 \mathsf{P}_x(a y + b z)=a  |x|^2 \mathsf{P}_x(y) + b  |x|^2 \mathsf{P}_x(z) = a \langle x , y \rangle x + b \langle x, z \rangle x$ by linearity of $\mathsf{P}_x$.
+Therefore $\langle x , a y + b z \rangle = a \langle x , y \rangle + b \langle x, z \rangle$.
+
+	We claim $\langle \cdot , \cdot \rangle$ is conjugate symmetric.
+Let $x,y \in V$.
+$ | |x|^2\mathsf{P}_x(y)|^2 = |x|^4|\mathsf{P}_x(y)|^2 = |\langle x , y \rangle|^2 |x|^2$.
+Using the second property of the projection we have $|\langle x , y \rangle|^2 = |x|^2|\mathsf{P}_x(y)|^2 =  |y|^2|\mathsf{P}_y(x)|^2 = |\langle y , x \rangle|^2$.
+The modulus of the product is symmetric as a function of the arguments.
+Now consider $|x|^2|y|^2\mathsf{P}_x(\mathsf{P}_y(x)) = |x|^2 \langle y , x \rangle \mathsf{P}_x(y) = \langle y , x \rangle \langle x , y \rangle x$.
+The product $\langle x , y \rangle \langle y , x \rangle \in \mathbbm{R}$ by the third property of the projection.
+$\arg(\langle x , y \rangle) + \arg(\langle y , x \rangle)$ is $0$ or $\pi$.
+$\pi$ is excluded since $\langle x , x \rangle = |x|^2 \ge 0$.
+$\langle y , x \rangle$ and $\langle x , y \rangle$ have equal modulus and opposite phase.
+$\langle x , y \rangle = \langle y , x \rangle^\dagger$
 \end{proof}
@@
 \begin{thrm}
-	Let $C^1(\mathcal{Q}, \mathbbm{C})$ the space of differentiable complex functions defined over $\mathcal{Q}$.
+	Let $C^1(\mathcal{Q}, \mathbbm{C})$ be the space of differentiable complex functions defined over $\mathcal{Q}$.
 Let $q : \mathcal{Q} \rightarrow \mathbbm{R}$ a differentiable function.
-Let $K_q$ a covariant linear operator: given an invertible differentiable function $\hat{q} : \mathbbm{R} \rightarrow \mathbbm{R}$ $K_{\hat{q}} = \partial_{\hat{q}} q K_{q}$.
+Let $K_q$ be a covariant linear operator, that is given an invertible differentiable function $\hat{q} : \mathbbm{R} \rightarrow \mathbbm{R}$ $K_{\hat{q}} = \partial_{\hat{q}} q K_{q}$.
 Then $K_q = a \partial_q$ for some $a \in \mathbbm{C}$.
@@
 
+\begin{thrm}
+	Let $G \equiv C^1(\mathcal{Q}, \mathbbm{C}) \cap L^1(\mathcal{Q}, \mathbbm{C})$ be the space of differentiable and integrable complex functions defined over $\mathcal{Q}$.
+Let $K$ be a linear adjoint operator of the form $a \partial_q$ with $a \in \mathbbm{C}$.
+Let $K$ be self-adjoint with respect to the inner product $\langle f, g \rangle =
+	\int_{\mathcal{Q}} f^\dagger(q) g(q) dq$. Then $K= \imath r \partial_q$ with $r \in \mathbbm{R}$.
+\end{thrm}
+
+\begin{proof}
+	We claim $K= \imath r \partial_q$ with $r \in \mathbbm{R}$.
+Let $f,g \in G$.
+We have $\langle f, K g \rangle = \int_{\mathcal{Q}}f^\dagger a \partial_q g dq = f^\dagger g |_{\partial \mathcal{Q}}  - \int_{\mathcal{Q}} a \partial_qf^\dagger  g dq= \int_{\mathcal{Q}} (- a\partial_q f^\dagger) g dq = \int_{\mathcal{Q}} (- a^\dagger\partial_q f)^\dagger g dq = \langle - a^\dagger\partial_q f, g \rangle$ using integration by part and integrability of $f$ and $g$.
+$\langle - a^\dagger\partial_q f, g \rangle =\langle K f, g \rangle$ as $K$ is self-adjoint.
+$- a^\dagger\partial_q = a \partial_q$.
+$a = - a^\dagger$.
+$a = \imath r$ with $r \in \mathbbm{R}$.
+\end{proof}
+
 \begin{thebibliography}{0}

### 2016-10-03 2015CompleteDerivation/completeDerivation.tex
Adding figure for discontinuous maps

@@
 It is physical consistency that limits us to continuous functions.
+
+\begin{figure}
+	\includestandalone[width=\columnwidth]{pictures/continuousMap}
+	\caption{A continuous function always maps finite precision knowledge in the vertical axis (i.e.
+an open set) to finite precision knowledge in the horizontal axis.
+A discontinuous function may map to finite precision (i.e.
+a closed set).
+If finite precision identification is ruled out (i.e.
+use of standard topology) then discontinuous functions must also be ruled out.}
+	\label{fig:continuous_map}
+\end{figure}
 

### 2016-10-03 2015CompleteDerivation/pictures/continuousMap.tex
Adding figure for discontinuous maps

\documentclass{standalone}
\usepackage{tikz}
\usetikzlibrary{calc}
\begin{document}
	\begin{tikzpicture}[scale=.4]

		% INPUTS

		\newcommand{\displace}{23};			 % distance between the two plots, origin to origin

		\newcommand{\axislength}{10};	   	 % axis lengeth
		\newcommand{\axisheight}{8};		 % axis height

		\newcommand{\horlinelength}{7};		% length of all horizontal lines
		\newcommand{\vertlinelength}{7};	 % length of all vertical lines

		\newcommand{\lineheighta}{4};		  % y-coordinate of first horizontal line in both plots
		\newcommand{\lineheightb}{5.5};		 % y-coordinate of second horizontal line in both plots
		\newcommand{\functionheight}{2};  % y-coordinate of function in right plot

		\newcommand{\linea}{3};					  % x-coordinate of first vertical line in left plot
		\newcommand{\lineb}{5};				   	  % x-coordinate of second vertical line in left plot
		\newcommand{\linec}{4};					  % x-coordinate of vertical line in right plot

		\newcommand{\hmargin}{2};
		\newcommand{\vmargin}{1.5};

		% styles		
	    \tikzstyle{axis}=[thick,color=black];				
	    \tikzstyle{line}=[color=black,dashed];  % dashed better than dotted?
	    \tikzstyle{function}=[ultra thick,color=red];

		% outer rectangle and name of diagram
		\draw ({-\hmargin - \axislength},{- \axisheight - \vmargin}) rectangle ({\displace + \axislength + \hmargin},{\axisheight + \vmargin});

		% axes for left plot
		\draw [style=axis] (0,-\axisheight) -- (0,\axisheight);		
		\draw [style=axis] (-\axislength,0) -- (\axislength,0);

		% lines in left plot
		\draw [style=line] (\lineb,-\vertlinelength) -- (\lineb,\vertlinelength);
		\draw [style=line] (\linea,-\vertlinelength) -- (\linea,\vertlinelength);

		\draw [style=line] (-\horlinelength,\lineheighta) -- (\horlinelength,\lineheighta);		
		\draw [style=line] (-\horlinelength,\lineheightb) -- (\horlinelength,\lineheightb);

		% function in left plot		
		\draw [style=function] plot [smooth] coordinates {(-8,-3) (0,0) (\linea,\lineheighta) (\lineb,\lineheightb) (8,6)};

		% axes for right plot
		\draw [style=axis] (\displace + 0,-\axisheight) -- (\displace + 0,\axisheight);		
		\draw [style=axis] (\displace + -\axislength,0) -- (\displace + \axislength,0);

		% lines in right plot
		\draw [style=line] (\displace + \linec,-\vertlinelength) -- (\displace + \linec,\vertlinelength);

		\draw [style=line] (\displace + -\horlinelength,\lineheighta) -- (\displace + \horlinelength,\lineheighta);		
		\draw [style=line] (\displace + -\horlinelength,\lineheightb) -- (\displace + \horlinelength,\lineheightb);

		% function including dot in right plot		
		\draw [style=function]	(\displace + -8,\functionheight) -- (\displace + 8,\functionheight);
		\fill [style=function] (\displace + \linec,4.75) circle (.25);
		\filldraw [draw=red,fill=white,thick] (\displace + \linec,\functionheight) circle (.25);

	\end{tikzpicture}

\end{document}

### 2016-10-04 2015CompleteDerivation/pictures/temporalDof.tex
No need to use calculator package
Missing hats in the left part

@@
 
-		\draw[black,fill=green,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\phivar:($(h0)$)}] ($(h0)$) rectangle ($(h8)$) node[pos=.5, rotate=\phivar] {\large $dq\wedge dp$};
+		\draw[black,fill=green,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\phivar:($(h0)$)}] ($(h0)$) rectangle ($(h8)$) node[pos=.5, rotate=\phivar] {\large $d\hat{q}\wedge d\hat{p}$};
 
-		\draw[black,fill=green,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={0:($(h0)$)}] ($(h2)$) rectangle ($(h5)$) node[pos=.5] {\large $dt\wedge dE$};
+		\draw[black,fill=green,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={0:($(h0)$)}] ($(h2)$) rectangle ($(h5)$) node[pos=.5] {\large $d\hat{t}\wedge d\hat{E}$};
 

### 2016-10-07 2015CompleteDerivation/completeDerivation.tex
Closing the loop for Hamiltonian/determinism as Dave suggested

@@
 \begin{prop}\label{hamiltons_equations}
-	A continuous deterministic and reversible process for the particles of a classical material admits a Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ that allows us to write the laws of evolution as
+	A continuous deterministic and reversible process for the particles of a classical material admits a Hamiltonian $H \in C^2(T^*\mathcal{Q}, \mathbbm{R})$ that allows us to write the laws of evolution as
 \begin{align*}
@@
 They are the set of equations that describe the deterministic and reversible motion of the infinitesimal parts of an infinitesimally reducible homogeneous material.
+Note that the argument goes the other way as well.
+A system governed by Hamilton's equations is deterministic and reversible: the equation of motion given by $H$ are differentiable since $H$ is twice differentiable.
+Therefore the equations are Lipschitz continuous and, by the Picard-Lindel\"{o}f theorem, a unique solution exists.
+
 In other words: the forces that conserve energy (i.e.
@@
 topology, measure theory, differential geometry, symplectic geometry).
-The conceptual meaning, on the other hand, is hopefully straight forward.
+The conceptual meaning, on the other hand, is hopefully straightforward.
 
@@
 Let $S \in TM \; | \; f(\xi^a(m)) = \xi^a(m) + S^a(m)dt \; \forall m \in M$ be the infinitesimal displacement.
-There exist a function $H: M \rightarrow \mathbbm{R}$ such that $S^{a} \omega_{ab} = \partial_{b}H$.
+There exists a function $H: M \rightarrow \mathbbm{R}$ such that $S^{a} \omega_{ab} = \partial_{b}H$ and $H \in C^2(M, \mathbbm{R})$.
 \end{prop}
@@
 $S$ is a curl free vector field and admits a potential $H$ such that $S_{a} = \partial_{a}H$.
+
+	We claim $H \in C^2(M, \mathbbm{R})$.
+$S$ is differentiable as it is the displacement of a symplectomorphism.
+The derivatives of $H$ are differentiable as $S_{a} = \partial_{a}H$.
+$H$ is twice differentiable.
 \end{proof}

### 2016-10-07 2015CompleteDerivation/completeDerivation.tex
Feedback from Josh

@@
 That is: $\delta\mathcal{c}=\hat{\mathcal{c}}-\mathcal{c}$.
-Note that such state difference may not describe a physical state: it may remove some material from one location to add it somewhere else.
+Note that such a state difference may not describe a physical state: it may remove some material from one location to add it somewhere else.
 Yet, these state changes are still physically distinguishable objects that provide a configuration for the same homogeneous material, therefore we extend the state space $\mathcal{C}$ to include state differences.\footnote{Note that one can characterize a state difference without knowing the states themselves.
@@
 For example, given the state of a fluid we'll want to know how the material is distributed in space.
-This in general will depend on how much the state of the composite system "knows" about its parts.
+This in general will depend on how much the state of the composite system ``knows" about its parts.
 For example, the position and orientation of an ideal rigid body is enough to define where all its constituents are, while the volume/pressure/temperature of an ideal gas is not enough to determine the position and momentum of all its constituents.
@@
 We will call a \emph{particle} the smallest amount of a material to which we can assign an independent state.
-For example, a photon is a particle of light (it is the smallest amount of light we can describe), an infinitesimal amount of water is treated classically as a point particle (the smallest amount for a continuous fluid).\footnote{We use this definition as it allows us to retroactively talk on somewhat equal grounds about classical point-particles and quantum particles, and underline how physically (and mathematically) they are very similar yet very different objects.}
+For example, a photon is a particle of light (it is the smallest amount of light we can describe), an infinitesimal amount of water is treated classically as a point particle (the smallest amount for a continuous fluid).\footnote{We use this definition as it allows us to retroactively talk on somewhat equal grounds about classical point-particles and quantum particles, and underscore how physically (and mathematically) they are very similar yet very different objects.}
 
@@
 \begin{justification}
-We claim $\mathcal{C}$ is an abelian group. $\mathcal{C}$ is the state space for a decomposable homogeneous system. $\mathcal{C}$ is an abelian group  by \ref{prop:abelian_group}.
+We claim $\mathcal{C}$ is an abelian group. $\mathcal{C}$ is the state space for a decomposable homogeneous system and is therefore an abelian group  by \ref{prop:abelian_group}.
 
@@
 Let $\mathcal{c}_1=\tau(a)\mathcal{c}_2$, then $\varrho(\mathcal{c}_1)=a \varrho(\mathcal{c}_2)$ as $\tau(a)$ increases the amount of material by the factor $a$.
-$\varrho$ is a homomorphism.
-$\mathcal{C}$ is isomorphic to $\varrho(\mathcal{C}) \subseteq C(\mathcal{S})$ as a vector space.
+$\varrho$ is a homomorphism as a vector space.
+$\mathcal{C}$ is isomorphic to $\varrho(\mathcal{C}) \subseteq C(\mathcal{S})$.
 \end{justification}
@@
 \begin{justification}
-	We claim there exists a positive linear functional $\Lambda_U : \mathcal{C} \rightarrow \mathbbm{R}$ for each $U \subseteq \mathcal{S}$ such that $\Lambda_U = \Lambda_{\mathrm{int}(U)}$.
+	We claim there exists a positive linear functional $\Lambda_U : \mathcal{C} \rightarrow \mathbbm{R}$ for each $U \subseteq \mathcal{S}$ such that $\Lambda_U = \Lambda_{\mathrm{int}(U)}$ where $\mathrm{int}(U)$ is the interior of $U$.
 Let $\Lambda_{U} : \mathcal{C} \rightarrow \mathbbm{R}$ be the functional that returns the amount of material in $\mathcal{c}$ compatible with the outcome associated with the open set $U \subseteq \mathcal{S}$.
@@
 Let $U \subseteq \mathcal{S}$ be an open set of particle states associated with an outcome.
-Let $\mathcal{c} \in \mathcal{C}$ a composite state.
+Let $\mathcal{c} \in \mathcal{C}$ be a composite state.
 The amount of material of $\mathcal{c}$ associated with the outcome $U$ must be finite, as physically we always work with finite quantities.
@@
 We would expect the expression to become $\Lambda_U (\mathcal{c}) = \int_a^b \rho_\mathcal{c} (q) dq$ for states identified by a continuous state variable.
-By changing state variable, though, we see that this does not work in general: $\rho_\mathcal{c}(\hat{q})= \rho(q)_\mathcal{c} dq/d\hat{q}$.
+By changing state variable, though, we see that this does not work in general: $\rho_\mathcal{c}(\hat{q})= \rho_\mathcal{c}(q) dq/d\hat{q}$.
 This makes $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using.
@@
 Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c} \in \mathcal{C}$.
-Let $q^i$ and $\hat{q}^j$ two sets of independent state variables in $U\subseteq \mathcal{S}$.
+Let $q^i$ and $\hat{q}^j$ be two sets of independent state variables in $U\subseteq \mathcal{S}$.
 Let $\rho_\mathcal{c}(q^i)$ and $\rho_\mathcal{c}(\hat{q}^j)$ be the expression in local coordinates of the distribution.
@@
 
-As we identified the space of Lebesque integrable differentiable functions $C^1(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as the space of distributions that are physically meaningful, we can better understand why other commonly used function spaces do not fit the bill.
+As we identified the space of Lebesgue integrable differentiable functions $C^1(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as the space of distributions that are physically meaningful, we can better understand why other commonly used function spaces do not fit the bill.
 Some are not restrictive enough.
@@
 Note that neither of those expressions matches the norm given by the inner product defined as $\langle f, g \rangle = \int_{\mathcal{S}} fg \, d\mu$.
-Moreover $\mathcal{C}$, since it excludes discontinuous functions, can never be a complete normed space: it will not include the limit for all Cauchy sequences, it cannot be a Banach or a Hilbert space.
-Such construction, though, is still useful.
+Moreover $\mathcal{C}$, since it excludes discontinuous functions, can never be a complete normed space: it will not include the limit for all Cauchy sequences; it cannot be a Banach or a Hilbert space.
+Such a construction, though, is still useful.
 Consider the following expressions:
@@
 The Jacobian has to be unitary, therefore all elements $\partial_i \hat{k}^j$ must be well defined and $|\partial_i \hat{k}^j| = 1/d_q\hat{q}$.
-This only happens if $\partial_i \hat{k}^j$ is a $1\times1$ matrix: there is only one $k$.
+As this can only happen if $\partial_i \hat{k}^j$ is a $1\times1$ matrix, there can only be one $k$.
 This gives $\partial_{k} \hat{k} = d_{\hat{q}} q$ and $\hat{k} = d_{\hat{q}} q \, k + f(q)$.
@@
 Only dimensionless relationships to other physical constants, such as the fine structure constant $\alpha = e^2/\hbar c$, can possess that trait.
-That is why one can choose "natural units" for which $\hbar=1$.
+That is why one can choose ``natural units" for which $\hbar=1$.
 In SI units the relationship $\hbar dk = dp = m du$ between kinetic and conjugate momentum, derived later, sets the relationship between units.} The cardinality for states in a finite (i.e.
-compact) region $U \subseteq \mathcal{S}$ will be given by $\mu(U) = \int_U \hbar dq \wedge dk$: we are finally able to write the volume of integration in terms for state variables.
+compact) region $U \subseteq \mathcal{S}$ will be given by $\mu(U) = \int_U \hbar dq \wedge dk$: we are finally able to write the volume of integration in terms of state variables.
 
@@
 We can tell them apart.
-We should take the use of densities and conjugate quantities as evidence that quantities like space and time are not discrete.
-Or more precisely: that the processes we use to distinguish those quantities cannot identify single instances, as that is all we can say experimentally.
-The quantum case does not change this, as the use of densities (in the form of the wave function) and conjugate quantities are even more prominent.
+We should take the use of densities and conjugate quantities as evidence that quantities like space and time are not discrete in the sense that the processes we use to distinguish those quantities cannot identify single instances.
+The quantum case does not change this, as the use of densities (in the form of the wave function) and conjugate quantities is even more prominent.\footnote{One may still posit that space is fundamentally discrete at scales where both classical and quantum mechanics cease to be valid.
+For that to work, the measure defined on such a discrete space has to become the measure on the classical phase space.
+That is: it must be a discrete version of $T^*\mathcal{Q}$ and not $\mathcal{Q}$ by itself.}
 
@@
 
-	Consistently with what we said in assumption \ref{ass:determinism}, if we defined a state for each particle, then a deterministic and reversible evolution on such state must exist.
+	Consistent with what we said in assumption \ref{ass:determinism}, if we defined a state for each particle, then a deterministic and reversible evolution on that state must exist.
 Yet, the idea that we can assign states to infinitesimal parts should be considered only a \emph{simplifying} assumption.
 The obvious reason is that we know this does not work in practice: as we keep decomposing the material we end up with molecules, atoms and then subatomic particles.
-But it is more instructive to understand when and how the assumption breaks down at a more conceptual level.
+But it is instructive to understand when and how the assumption breaks down at a more conceptual level.
 
@@
 The future state of the system does not depend on the state of the environment, so that we can define an independent state, yet the future state of the environment is affected by the state of the system, so that we can have external processes that allow us to define physical distinguishability.
-If we try to assume that both the system and environment are really made of the same classical material, then the claim to that asymmetry is lost: there is no difference in size, everything is made of infinitesimal pieces, and there is no difference in the laws of evolution, as the pieces are homogeneous.
+If we try to assume that both the system and environment are really made of the same classical material, then the claim to that asymmetry is lost: the infinitesimal parts of both system and environment must affect one another in the same way.
 
@@
 First, for this to work, one would have to specify how the states of the parts were defined since their evolution is individually no longer deterministic (the future of each part depends on the state of other parts).
-But ignoring that issues, the bigger problem is that we can always locally separate the evolution into independent degrees of freedom.
-For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value, they clearly evolve independently.
+But ignoring that issue, the bigger problem is that we can always locally separate the evolution into independent degrees of freedom.
+For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value; they clearly evolve independently.
 And since the pieces are infinitesimal, a local decomposition is all that is needed.
@@
 The state is what gives the identity to the system, the part we can study and describe.
-The unstated part is what allows continuous interaction between the system and the environment, it's what allows us to study and describe the state.
-As we'll see later, it is precisely this problem that quantum mechanics addresses better conceptually than classical mechanics.
+The unstated part is what allows continuous interaction between the system and the environment; it's what allows us to study and describe the state.
+As we'll see later, it is precisely this problem that quantum mechanics conceptually addresses better than classical mechanics.
 
@@
 So, all quantities that are invariant under state variable changes are also invariant under time evolution.
-As the very definition of deterministic and reversible processes relies on time, this has to come from elsewhere (i.e.
+As the very definition of deterministic and reversible processes relies on time, time has to come from elsewhere (i.e.
 the environment).
@@
 	We claim $\mathcal{T}_{\Delta t}$ is a symplectomorphism on $(T^*\mathcal{Q}, \omega)$.
-The distribution on final states must be defined.
-The Jacobian for $\mathcal{T}_{\Delta t}$ exists and is non-zero.
-$\mathcal{T}_{\Delta t}$ is a diffeomorphism.
+The distribution on final states must be defined  therefore the Jacobian for $\mathcal{T}_{\Delta t}$ exists and is non-zero: $\mathcal{T}_{\Delta t}$ is a diffeomorphism.
 A deterministic and reversible process conserves the number of states and possibilities.

### 2016-10-14 2015CompleteDerivation/completeDerivation.tex
Feedback from Josn
Incorporating comments from Jiahua

@@
 If $\rho_\mathcal{c} \neq \rho$ then $\exists U \in \mathcal{S} \; | \; \int_{U} \rho_\mathcal{c} d \mu \neq \int_{U} \rho d \mu$.
-Both integral are finite as $\rho_\mathcal{c}, \rho \in L^1(\mathcal{S}, \mu)$.
-Both integral are expressible with the same state variable as all elements are defined on the same differential structure.
+Both integrals are finite as $\rho_\mathcal{c}, \rho \in L^1(\mathcal{S}, \mu)$.
+Both integrals are expressible with the same state variable as all elements are defined on the same differential structure.
 Therefore there exists an outcome for $\mathcal{S}$ that can physically distinguish the two distributions.
@@
 \begin{justification}
-We claim the vector field $S \in T\mathcal{S}$ for the infinitesimal displacement $S^a = d\xi^a/dt$ admits a potential $H$ such that $S^{a} \omega_{ab} = \partial_{b}H$.
+We claim the vector field $S \in T\mathcal{S}$ for the infinitesimal displacement $S^a = d\xi^a/dt$ admits a potential $H \in C^2(T^*\mathcal{Q}, \mathbbm{R})$ such that $S^{a} \omega_{ab} = \partial_{b}H$.
 The state space $\mathcal{S}$ for the particles of a classical material is a symplectic manifold by \ref{symplectic_manifold}.
 The map for infinitesimal time evolution $\mathcal{T}_{dt}$ is an infinitesimal self-symplectomorphism by \ref{canonical_transformation}.
-By \ref{symplectomorphism_generator} the infinitesimal displacement $S$ admits a potential $H$ such that $S^{a} \omega_{ab} = \partial_{b}H$
+By \ref{symplectomorphism_generator} the infinitesimal displacement $S$ admits a potential $H \in C^2(T^*\mathcal{Q}, \mathbbm{R})$ such that $S^{a} \omega_{ab} = \partial_{b}H$
 
@@
 $T^*\mathcal{Q}$ is a differentiable manifold.
-Time evolution is a diffeomorphism by \ref{hamiltons_equations} therefore derivatives between state variables and time variables always exist.
+Time evolution is a diffeomorphism by \ref{canonical_transformation} therefore derivatives between state variables and time variables always exist.
 $\bar{\mathcal{S}}$ is a differentiable manifold.
@@
 This is the manifold that allows us to define time-and-state-variable-invariant densities.
-The metric allows us to measure the cardinality of possibilities.
+The metric $\omega$ allows us to measure the cardinality of possibilities.
 
@@
 Let $\mathcal{H}$ be the invariant Hamiltonian and $H$ be the Hamiltonian for a particular choice of time and state variables.
-We have $d_s q^i = \partial_{p_i} \mathcal{H} = d_s t \, d_t q^i = - \lambda \, \partial_{p_i} H$ where $\lambda : T^*\mathcal{M} \rightarrow \mathbbm{R} \; | \; \lambda \mapsto d_s t$.
+We have $d_s q^i = \partial_{p_i} \mathcal{H} = d_s t \, d_t q^i = - \lambda \, \partial_{p_i} H$ where $\lambda : T^*\mathcal{M} \rightarrow \mathbbm{R}$ such that $\lambda \equiv d_s t$.
 $d_s p_i = - \partial_{q^i} \mathcal{H} = d_s t \, d_t p_i = \lambda \, \partial_{q^i} H$.
@@
 There must exist a metric $g$, consistent with $\omega$, that allows one to calculate the cardinality of possibilities for initial conditions.
-That is: $\omega(dq^\alpha,dp_\alpha) \propto g(dx^\alpha,du^\alpha)$.
-
-Since both $dx^\alpha$ and $du^\alpha$ are four vectors and the metric must be linear, $g$ is a rank two tensor on $\mathcal{M}$.
+That is: $\omega(dq^\alpha,dp_\alpha) \propto g(dx^\alpha,du^\alpha)$ where the arguments are the value ranges for the respective quantities.
+
+Since both $dx^\alpha$ and $du^\alpha$ are four vector components and the metric must be linear, $g$ is a rank two tensor on $\mathcal{M}$.
 To each direction in space-time $dx^\alpha$ is associated a direction in the tangent space of velocities $du^\alpha$ with which it forms a d.o.f..
@@
 
-Since $g$ is symmetric, it can be diagonalized: $\omega= dq^i \wedge dp_i - dt \wedge dE \propto g = dx^i g_{ii} du^i + dt g_{00} du^0$.
-This tells us $g_{ii} > 0$ and $g_{00} < 0$.
+Since $g$ is symmetric, it can be diagonalized: $\omega(dq^\alpha, dp_\beta)= dq^i \omega_{q^ip_i} dp_i + dt \omega_{tE} dE \propto g(dx^\alpha, du^\beta) = dx^i g_{ii} du^i + dt g_{00} du^0$.
+As $\omega_{q^ip_i} > 0$ we must have $g_{ii} > 0$.
+As $\omega_{tE} < 0$ we must have $g_{00} < 0$.
 We can further change the units along the spatial directions so that $g_{ii}=1$: if we use the same units for the spatial coordinates, each direction will contribute the same number of possibilities.
@@
 $x^\alpha$ exists as $g$ is symmetric.
-$\omega(dq^\alpha, dp_\beta) = dq^i \wedge dp_i - dt \wedge dE \propto dq^i g_{ii} du^i + dt g_{00} du^0$.
+$\omega(dq^\alpha, dp_\beta) = dq^i \omega_{q^ip_i} dp_i + dt \omega_{tE} dE \propto g(dx^\alpha, du^\beta) \propto dq^i g_{ii} du^i + dt g_{00} du^0$.
 $dq^\alpha = dx^\alpha$.
-$g_{ii} > 0$ and $g_{00} < 0$.
+We require $g(dx^\alpha, du^\beta)>0$ when the arguments are positive.
+Therefore $g_{ii} > 0$ and $g_{00} < 0$.
 $g$ has signature $(3, 1)$.
@@
 Let $v, w \in T_m M$ be two vectors defined at a point $m \in M$.
-Let $v^a, w^b$ their components.
+Let $v^a, w^b$ be their components.
 Let $v'\equiv f v, w' \in T_{f(m)}M$ be the pushforward of $v, w$ by $f$.

### 2016-10-14 2015CompleteDerivation/completeDerivation.tex
More work on quantum

@@
 \begin{prop}\label{prop:wavefuntion}
-	The state space $\Psi$ for an irreducible material is isomorphic to the space of Lebesgue square integrable differentiable functions over $\mathcal{Q}$.
-That is $\Psi \cong C^1(\mathcal{Q}) \cap L^2(\mathcal{Q}, \mu)$.
+	The state space $\Psi$ for an irreducible material is isomorphic to a subspace of the space of Lebesgue square integrable differentiable.
+That is $\Psi \cong \Psi(\mathcal{Q}, dp^i) \subseteq C^1(\mathcal{Q}) \cap L^2(\mathcal{Q}, \mu)$ under the isomorphism $\Upsilon_q : \Psi \leftrightarrow \Psi(\mathcal{Q}, dp^i)$ where $\Psi(\mathcal{Q}, dp^i)$ is the space of wave functions over $q^i$.
 The state space for a quantum particle is the projective space $\mathsf{P}(\Psi)$.
@@
 \begin{prop}\label{prop:inner_product}
-	The state space $\Psi$ for an irreducible material is an inner product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q) \psi_2(q) dq$.
+	The state space $\Psi$ for an irreducible material is an inner product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$.
 \end{prop}
@@
 
+TODO: ranges of $q^i$ and $\hat{q}^i$ must be the same: there must be a bijective function between them. Bijection is continuous only if $\hat{q}^i(q^i)$. Otherwise can be discontinuous.
+
 That is: for each state variable we have a corresponding linear hermitian operator which can be expressed as function multiplication in a suitable choice of $\hat{\mathcal{Q}}$.
@@
 \begin{prop}\label{prop:self_adjoint_operators}
-	For each fragment parameter $f$ there exist a corresponding self-adjoint linear operator $F : \Psi \rightarrow \Psi$.
-There exists a manifold $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
-There exist an injective map $\Psi_{\hat{\mathcal{Q}}} : \Psi \rightarrow C^1(\hat{\mathcal{Q}}, \mathbbm{C})$.
+	For each fragment variable $f$ there exists a corresponding self-adjoint linear operator $F : \Psi \rightarrow \Psi$.
+There exist a manifold $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ and $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ is an isomorphism.
 \end{prop}
@@
 As it has to be self-adjoint, it will be of the form $K_{q} = a \imath \partial_q$ where $a$ is a real number.
-We chose $a$ to be negative by convention and to be unitary so that $K_{q}$ is expressed in units of inverse $q$ as in the classical case.
+We choose $a$ to be negative by convention and to be unitary so that $K_{q}$ is expressed in units of inverse $q$ as in the classical case.
 Therefore we have $K_{i} = -\imath \partial_i$ and $P_{i} = -\imath \hbar \partial_i$.
+
+TODO: Schar
+
+\begin{prop}\label{prop:momentum_operator}
+	The operator $P_i$ associated with the conjugate quantity of $q^i$ is $P_i = - \imath \hbar \partial_{i}$.
+The operator that changes the variable of the wave function from $q^i$ to $p^i$ is $\Upsilon^{q_i}_{p^i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^3} \int_{\mathcal{Q}} e^{\frac{q_i p^i }{\imath \hbar}} \psi(q^i) d^n q $.
+\end{prop}
+\begin{justification}
+	TODO
+\end{justification}
+
+We now have all the basic features of the state space of a quantum particle.
+And we have them by re-deriving them from first principles instead of using some kind of formal analogy.
+The similarity with the classical framework are simply due to some of the similar starting points.
+
+We reached the goal despite $\Psi$ is not being a complete metric space.
 
@@
 
+Now that we have fully characterized what we mean by an irreducible material, we can stipulate the following:
+
+\begin{assump}[Infinitesimal reducibility]\label{ass:infinitesimal_reducibility}
+	The system under study is composed of an irreducible homogeneous material and as a whole undergoes deterministic and reversible evolution.
+\end{assump}
+
+\begin{rationale}
+	The idea is that each amount of material has a state, the physically distinguishable configuration of the distribution, and an unstated part, the physically indistinguishable configuration of the fragments of the material.
+The system is deterministic and reversible in the sense outlined in section ???: given the state of the system at one time (i.e.
+the whole distribution) we can predict/reconstruct the future/past states.
+But we cannot tells anything about the evolution of the unstated part (i.e.
+the motion of the fragments).
+
+	As we saw, the state space recovered under this assumption is the one used in quantum mechanics.
+In the following sections we will also derive two types of state evolution: the deterministic and reversible one, corresponding to unitary evolution (e.g.
+the Schroedinger equation), and a non deterministic one useful to describe measurement interactions, corresponding to the projection postulate.
+Therefore, as the infinitesimal reducibility assumptions leads to classical mechanics, the irreducibility assumption leads to quantum mechanics.
+
+	As we previously saw in section ???, the infinitesimal reducibility assumptions had its problem. How about the irreducibility assumption? How do they compare?
+
+	The methodological assumes a different connotation.
+In order for the assumption to hold, we need to show that there are no processes at our disposal that can probe the constituents of a quantum particle.
+For example: an electron as a whole interacts with a photon as a whole.
+Note that this does not require showing that no such process exists: conceptually specific parts of the photon may be interacting with specific parts of the electron.
+But if those processes are not available to us, say because we can only manipulate whole photons and they always interacts with a whole electron, they cannot be used to create a more detailed picture.
+So we simply have to enumerate the processes available to us, which is, at least in principle, feasible.
+
+	The issues rising from a lack of perfect isolation all but disappear.
+Each degree of freedom comes with its own unstated part (i.e.
+the motion of the fragments within that d.o.f.) therefore these is no pretense of perfect isolation.
+It is built into the description.
+
+	The problem of physically meaningless interaction is also considerably lessened.
+Even if we have multiple particles each with their state evolving deterministically, we can still imagine their unstated parts as interacting with each other.
+Therefore each quantum particle would not essentially reside in its own separate universe.
+
+	In these way, the irreducibility assumption fares much better than the infinitesimal reducibility assumption. Yet, it is still plagued with some of the same problems.
+
+	The problem of incompatibility of measurements and deterministic evolution at all levels still remains.
+This will lead us to require a different non deterministic evolution, that will allow us to recover the statistical interpretation of the wavefunction.
+
+	The inability to define time by following a deterministic and reversible problem is still present.
+
+	There is one additional problem, though, not present in the classical case.
+The quantum states, as we have defined them, are states at equilibrium: they all provide the same level of description for the fragments; they are all reachable from one another through unitary evolution (which we'll see corresponds to deterministic and reversible evolution).
+Therefore they cannot and should not be expected to provide a description under non deterministic evolution.
+Consider a muon decay: the initial state of the muon and the final states of the electron and neutrinos can certainly be well described by a quantum states.
+We have no reason to expect, though, that the while the muon is decaying into the resulting particles the state is always in equilibrium and therefore well described mathematically by a vector in a Hilbert space.
+After all, the state of a single spin one half particle cannot determine the future state of three spin on half particles.
+
+	Using unitary evolution in those cases to make predictions, though, may still make physical sense.
+Quantities that are not affected by the particular evolution of the unstated part will have the same value under non-deterministic evolution and an equivalent quasi-static evolution.
+The distribution of final states is one such quantities by the irreducibility assumption: they don't care about the unstated part evolution.
+But it stands to reason that, at some point, the unstated part will have actual physical consequences.
+For example: it would be extremely reasonable to expect that is indeed the evolution of the unstated part that decides the moment in which the muon would decay.
+
+	In light of what we discussed, the irreducibility assumption is less flawed but it still cannot be fundamental, in the sense that we cannot take it to strictly apply to all of the universe.
+We should not expect to solve open problems, such as the arrow of time and the measurement problem, within the assumption.
+\end{rationale}
+
 \subsection{Quantum mechanics}
+
+We are now ready to write the equation of motion for an irreducible material.
+
+\begin{prop}\label{unitary_evolution}
+	A deterministic and reversible evolution map $\mathcal{T}_{\Delta t}: \Psi \rightarrow \Psi$ for an irreducible material is a unitary operator.
+That is: $\langle \mathcal{T}_{\Delta t} \psi_1, \mathcal{T}_{\Delta t} \psi_2 \rangle = \langle \psi_1 , \psi_2 \rangle$ where $\psi_1, \psi_2 \in \Psi$.
+The time evolution for the wave function $\Upsilon_q \mathcal{T}_{\Delta t} \Upsilon^q : S(\mathcal{Q}, dq) \rightarrow S(\mathcal{Q}, dq)$ is a $C^{\infty}$-diffeomorphism.
+\end{prop}
+
+\begin{justification}
+\end{justification}
+
+\begin{prop}\label{prop:schroedinger_equation}
+	A continuous deterministic and reversible evolution for an irreducible material admits a Hamiltonian operator $H : \Psi \rightarrow \Psi$ that allows us to write the laws of evolution as
+	\begin{align*}
+	\imath \hbar \partial_t \psi = H \psi
+	\end{align*}
+\end{prop}
+
+\begin{justification}
+\end{justification}
+
+Note: we get time dependent Schroedinger. No basis for time independent.
 
@@
 \begin{thrm}[Extended Riesz theorem]\label{extended_riesz_theorem}
-	Let $(S, \tau)$ a locally compact Hausdorff space.
+	Let $(S, \tau)$ be a locally compact Hausdorff space.
 Let $\Lambda = \{\Lambda_U : C(S) \rightarrow \mathbbm{R}\}_{U \subseteq S}$ a family of positive linear functionals such that $\forall U \subseteq S \; \Lambda_U = \Lambda_{\mathrm{int}(U)}$ and $\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2} \; \forall U_1, U_2 \subseteq S$.
@@
 
-\begin{thrm}
+\begin{thrm}\label{thrm:inner_product}
 	Let $V$ be a complex vector space.
@@
 
-\begin{thrm}
+\begin{thrm}\label{thrm:covariant_operator}
 	Let $C^1(\mathcal{Q}, \mathbbm{C})$ be the space of differentiable complex functions defined over $\mathcal{Q}$.

### 2016-10-14 2015CompleteDerivation/completeDerivation.tex
More feedback from Josh

@@
 \usepackage{braket}
+\usepackage{enumitem}
 
@@
 
-	Another issue arises if we include the measuring device in our description.
-Say $a \in A$ is the state of our system and $b \in B$ the state of our measuring device, the future state $\hat{a}=\hat{a}(a)$ because the system is deterministic, $\hat{b}=\hat{b}(a,b)$ because the measuring device needs to be affected by the system and $a=f(\hat{b})$ because we should be able to know the original state by looking at the measuring device.
-This means $(a, b) \mapsto (\hat{a}, \hat{b})$ is not a reversible map.
-We would have $b = b(\hat{a}(f(\hat{b})), \hat{b}) = b(\hat{b})$ and $a=f(\hat{b})$: the whole past state cannot be reconstructed just by looking at the future state of the measuring device.
-If we want determinisms and reversibility for the combined system, either both systems $\hat{a}=\hat{a}(a, b)$ and $\hat{b}=\hat{b}(a, b)$ are not independently deterministic or the second is not a measuring device $a \neq f(\hat{b})$.
+	Another issue arises if we include the measuring device in our description. Ideally, we'd like to require the following:
+	\begin{itemize}[noitemsep]
+		\item The system under study is deterministic and reversible
+		\item The measuring device ascertains the state of the system
+		\item The system as a whole is deterministic and reversible
+	\end{itemize}
+	Unfortunately, these three requirements together are inconsistent.
+Say $a \in A$ is the initial state of our system and $b \in B$ the initial state of our measuring device: the final state $\hat{a}=\hat{a}(a)$ because the system is deterministic; $a=f(\hat{b})$ because we are able to know the initial state of the system by looking at the final state of the measuring device; $b=b(\hat{a},\hat{b})$ because the system as a whole is reversible.
+Since $\hat{a}$ is determined by $a$, and $a$ can be known from $\hat{b}$, $b = b(\hat{a}(f(\hat{b})), \hat{b}) = b(\hat{b})$ can be determined from just $\hat{b}$.
+But since $a=f(\hat{b})$, the whole past state could be reconstructed just by looking at the future state of the measuring device.
+But $(\hat{b}) \rightarrow (a, b)$ cannot be an injection, therefore we arrive at a contradiction.
+
+	Of those three requirements we can only pick two.
+If we want determinisms and reversibility for the combined system, either both systems are not independently deterministic or the second is not a measuring device.
 In other words: we can't expect to have deterministic and reversible evolution at all levels of aggregation and have the parts interact in any physically meaningful way.
@@
 So, all quantities that are invariant under state variable changes are also invariant under time evolution.
-As the very definition of deterministic and reversible processes relies on time, time has to come from elsewhere (i.e.
-the environment).
+The ironic result is that while time is essential for the very definition of deterministic and reversible evolution, time itself has to be defined by some other type of process.\footnote{For example, if the phase-space volume occupied by a distribution increased, it would give us a coordinate-invariant quantity that changes in time.
+This, however, is not possible under Hamiltonian evolution.}
 

### 2016-10-14 2015CompleteDerivation/completeDerivation.tex
Forgot to save

@@
 So, all quantities that are invariant under state variable changes are also invariant under time evolution.
-The ironic result is that while time is essential for the very definition of deterministic and reversible evolution, time itself has to be defined by some other type of process.\footnote{For example, if the phase-space volume occupied by a distribution increased, it would give us a coordinate-invariant quantity that changes in time.
+The somewhat ironic result is that while time is essential for the very definition of deterministic and reversible evolution, time itself has to be defined by some other type of process.\footnote{For example, if the phase-space volume occupied by a distribution increased, it would give us a coordinate-invariant quantity that changes in time.
 This, however, is not possible under Hamiltonian evolution.}

### 2016-10-27 2015CompleteDerivation/completeDerivation.tex
Comments from Eric

@@
 For example, if the cannonball deformed the ground by this amount then its kinetic energy at impact was within this range; if the electron follows a certain path, then its state is among the ones that have spin $+1/2$.
+
+Note that the outcome of the process (e.g.
+the electron followed a particular path) is conceptually distinct from the set of states compatible with it (the states associated with the path) and the way we label them (spin=$+1/2$).
+Yet,  we'll conflate the concepts as this makes the language closer to experimental physics (``the outcome of the measurement is spin=$+1/2$").
+Context will allow us to recover the distinction.
 
@@
 
-We claim that $\mathcal{S}$ is Hausdorff.
+We claim that $\mathsf{T}$ is Hausdorff.
 Let $\mathcal{s_1}, \mathcal{s_2} \in \mathcal{S}$.
@@
 
-We claim that $\mathcal{S}$ is second countable.
+We claim that $\mathsf{T}$ is second countable.
 Consider the set of outcomes of a process.

### 2016-10-27 2015CompleteDerivation/completeDerivation.tex
Comments from Josh

@@
 \begin{justification}
-We claim $\mathcal{C}$ is an abelian group. $\mathcal{C}$ is the state space for a decomposable homogeneous system and is therefore an abelian group  by \ref{prop:abelian_group}.
+We claim $\mathcal{C}$ is an abelian group. $\mathcal{C}$ is the state space for a decomposable homogeneous material and is therefore an abelian group  by \ref{prop:abelian_group}.
 
@@
 		\item The measuring device ascertains the state of the system
-		\item The system as a whole is deterministic and reversible
+		\item System and measuring device, as a whole, is deterministic and reversible
 	\end{itemize}

### 2016-10-27 2015CompleteDerivation/completeDerivation.tex
Review with Dave and Christine

@@
 We'll therefore call \emph{complete state space} of particle states the set of all configurations at all times and we'll indicate it by $\bar{\mathcal{S}}$.
-In that space, a state is an element $\mathcal{s} \in \bar{\mathcal{S}}$ while a time evolution is a line $\bar{\mathcal{s}} \subset \bar{\mathcal{S}}$ that passes once at all times.
-Specifying an evolution map means giving a set of lines such that each state is traversed once and only once.
-Locally, each state will be given by $(q^i, k_i, t)$ therefore $\bar{\mathcal{S}}$ is a $2n+1$ dimensional manifold.
+In that space, a state $\mathcal{s} \in \bar{\mathcal{S}}$ will be given by $(q^i, p_i, t)$, making $\bar{\mathcal{S}}$ a $2n+1$ dimensional manifold.
 A region at equal time will be a hyper-surface $\mathcal{S}_{t=t_0} \subset \bar{\mathcal{S}}$ that includes all possible configurations at a particular time.
 Such hyper-surface will need to cut across all evolutions once and only once, therefore not all functions $t : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ are suitable time variables.
+A time evolution is a line  $\bar{\mathcal{s}} \subset \bar{\mathcal{S}}$ that passes once through each equal time surfaces.
+Specifying an evolution map means giving a set of lines such that each state is traversed once and only once.
 
@@
 	The restriction of $\rho_{\bar{\mathcal{c}}}$ over an arbitrary level set $\mathcal{S}_{t=t_0}$ is continuous and integrable by \ref{prop:differentiable_manifold}.
-$\rho_{\bar{\mathcal{c}}} \in L^1_t(\bar{\mathcal{S}}, \mu)$ Time evolution is a diffeomorphism by \ref{hamiltons_equations}.
+$\rho_{\bar{\mathcal{c}}} \in L^1_t(\bar{\mathcal{S}}, \mu)$.
+Time evolution is a diffeomorphism by \ref{hamiltons_equations}.
 $\rho_{\bar{\mathcal{c}}}$ is differentiable along the time variable as well as the state variables.
-$\rho_{\bar{\mathcal{c}}} \in C^1(\bar{\mathcal{S}})$
+$\rho_{\bar{\mathcal{c}}} \in C^1(\bar{\mathcal{S}})$.
 
@@
 
-As the temporal degree of freedom $(t, E)$ is not independent of the other d.o.f.
-$(q^i, p_i)$, it is not orthogonal to them in $T^*\mathcal{M}$.
+As the temporal degree of freedom $(t, E)$ is not independent of the other d.o.f.~$(q^i, p_i)$; it is not orthogonal to them in $T^*\mathcal{M}$.
 States are defined on the plane where $( q, p )$ (maximally) change.
@@
 \begin{align*}
-m.d.o.f \;\;\; &dq^1 \wedge dp_1 + dq^2 \wedge dp_2 = \omega \\
-t.d.o.f \;\;\; &dt \wedge dE + \omega = dq \wedge dp
+\textrm{multiple d.o.f.} \;\;\; &dq^1 \wedge dp_1 + dq^2 \wedge dp_2 = \omega \\
+\textrm{temporal d.o.f.} \;\;\; &dt \wedge dE + \omega = dq \wedge dp
 \end{align*}

### 2016-10-27 2015CompleteDerivation/completeDerivation.tex
First pass at all quantum justifications

@@
 \begin{justification}
-	TODO
+	We claim $\Psi$ is an abelian group. $\Psi$ is the state space for a decomposable homogeneous material and is therefore an abelian group  by \ref{prop:abelian_group}.
+
+	We claim the set of physically distinguishable transformations $T$ that change the unstated part of an irreducible material is isomorphic to $\mathbbm{C}$.
+Consider an amount $a_m \in \mathbbm{R}$ of irreducible material characterized by one d.o.f.
+The configuration of a fragment is characterized by two random variables $A, B: \Omega \rightarrow \mathbbm{R}$ where $\Omega$ is the sample space for the configuration of the fragments.
+As the configuration of each fragment is physically indistinguishable, $A$ and $B$ can be chosen such that the distribution $\rho$ of material over such variables is uniform and unitary.
+We can also rescale $A$ and $B$ such that they are defined on the range $[0, \sqrt{a_m}]$.
+We have $a_m=\int \rho dA \wedge dB = \int dA \wedge dB$.
+Consider a second amount $\hat{a}_m$ of irreducible material in the same configuration.
+It also can be characterized by a uniform distribution over two variables $C, D: \Omega \rightarrow \mathbbm{R}$ over a range of $\sqrt{\hat{a}_m}$.
+As the two amounts of material are in the same configuration we have $(C,D)$ is a function of $(A,B)$.
+We have $\hat{a}_m = \int dC \wedge dD = \int |J| dA \wedge dB$ where $|J|$ is the Jacobian determinant.
+As both distributions are uniform, $|J|$ has to be constant.
+The transformation has to be linear: $C=aA + bB$ and $D=-bA + aB$.
+As the amount of material is the only physically distinguishable quantity and is fully specified by $a$ and $b$, no other transformation is physically relevant.
+Let $\tau: \mathbbm{C} \rightarrow T$ represent such a transformation parametrized by $c=a+\imath b$.
+$|c|^2 = (a^2 + b^2) = \hat{a}_m / a_m$ is physically distinguishable.
+Define on $T$ an addition $+: T \times T \rightarrow T$ and a multiplication $*: T \times T \rightarrow T$ such that $\tau(a) + \tau(b) = \tau(a+b)$ and $\tau(a) * \tau(b) = \tau(a*b)$, $a,b \in \mathbbm{C}$, so that the sum and product of the transformation is equal to the sum and product of their respective factors.
+The phase $arg(c)$ under addition is physically distinguishable as it can affect the modulus of the result.
+$\tau$ is an isomorphism between $T$ and $\mathbbm{C}$ as fields.
+Generalizing to multiple independent d.o.f.
+the set of transformation remains the same.
+The transformation on each d.o.f.
+is similarly defined by $c_i \in \mathbbm{C}$.
+As $\hat{a}_m / a_m$ is the same under all d.o.f.
+$|c_i|=|c_j|$.
+That has to always hold under addition as well, therefore $arg(c_i) = arg(c_j)$.
+$T\cong \mathbbm{C}$ even in multiple d.o.f.
+
+	We claim $\Psi$ is a vector space over $\mathbbm{C}$.
+The abelian group $\Psi$ can be extended with the operations defined by $T$ as done in \ref{classical_vector space} for the infinitesimally reducible case.
+
 \end{justification}
@@
 That is $\Psi \cong \Psi(\mathcal{Q}, dp^i) \subseteq C^1(\mathcal{Q}) \cap L^2(\mathcal{Q}, \mu)$ under the isomorphism $\Upsilon_q : \Psi \leftrightarrow \Psi(\mathcal{Q}, dp^i)$ where $\Psi(\mathcal{Q}, dp^i)$ is the space of wave functions over $q^i$.
-The state space for a quantum particle is the projective space $\mathsf{P}(\Psi)$.
-\end{prop}
-\begin{justification}
-	TODO
+The state space for a quantum particle is isomorphic to the projective space $\mathsf{P}(\Psi)$.
+\end{prop}
+
+\begin{justification}
+	We claim the manifold $\mathcal{Q}$ that defines the units to describe a fragment of an irreducible material also provides a maximal description for the fragment configuration, that is: no other state variables can be added.
+Let $\mathcal{Q}$ be the manifold that defines the units.
+Then for each state $\psi \in \Psi$ there exists a distribution $\rho_\psi(q^i)$ that represents the quantity of material at each point.
+Suppose this distribution didn't represent a maximal description for the fragments.
+Then another independent variable $k$ would exist such that for each state $\psi \in \Psi$ we would have a distribution $\rho_\psi(q^i, k)$.
+$k$ cannot be part of a new d.o.f.
+as it would use a unit defined independently from $\mathcal{Q}$.
+This is a contradiction as $\mathcal{Q}$ defines all units.
+$k$ cannot lie in the same d.o.f.
+of one of the $q^i$ or it would provide a complete description for the fragment.
+This is a contradiction as the material is irreducible.
+$\mathcal{Q}$ provides a maximal description.
+
+	We claim $\Psi \cong \Psi(\mathcal{Q}, dp^i) \subseteq C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu, \mathbbm{C})$ as a vector space.
+The maximal physical description of a set of fragments is given by a point in $\mathcal{Q}$ parametrized by $q^i$ and by a transformation in $T$ parametrized by $c \in \mathbbm{C}$.
+For each $\psi \in \Psi$ there must exist a wave function $\psi(q^i) : \mathcal{Q} \rightarrow \mathbbm{C}$ that provides the maximal physical description for the fragments.
+$\rho_\psi(q^i) = |\psi(q^i)|^2$ as the modulus of $\psi(q^i)$ is the increase/decrease from the unit amount of material.
+$\rho_\psi(q^i) \in C(\mathcal{Q}, \mathbbm{R})$ by \ref{continuous_map}.
+Therefore $\psi(q^i) \in C(\mathcal{Q}, \mathbbm{R})$.
+$\rho_\psi(q^i) \in L^1(\mathcal{Q}, \mu, \mathbbm{R})$ by the same arguments in the justification for \ref{integration}.
+Therefore $\psi(q^i) \in L^2(\mathcal{Q}, \mu, \mathbbm{C})$.
+Let $\Upsilon_{q^i} : \Psi \rightarrow C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu, \mathbbm{C})$ be the function that given a state returns its wave function.
+As the wave function capture the maximal description of the system, two states must represent different distributions.
+$\forall \psi_1, \psi_2 \in \Psi, \psi_1 \neq \psi_2 \implies \Upsilon_{q^i}(\psi_1) \neq \Upsilon_{q^i}(\psi_2)$.
+$\Upsilon_{q^i}$ is injective.
+$\Upsilon_{q^i}$ is a bijection between $\Psi$ and $\Psi(\mathcal{Q}, dp^i)\equiv\Upsilon_{q^i}(\Psi)$.
+Let $\psi=\psi_1+\psi_2$, then $\Upsilon_{q^i}(\psi)=\Upsilon_{q^i}(\psi_1)+\Upsilon_{q^i}(\psi_2)$ as combining the distributions means combining the statistical description of the fragments.
+Let $\psi_1=\tau(c)\psi_2$, then $\Upsilon_{q^i}(\psi_1)=c \Upsilon_{q^i}(\psi_2))$ as $\tau(c)$ transforms the fragment configuration by the factor $C$.
+$\Upsilon_{q^i}$ is a homomorphism as a vector space.
+$\Psi$ is isomorphic to $\Psi(\mathcal{Q}, dp^i) \subseteq C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu, \mathbbm{C})$.
+
+	We claim the state space for a quantum particle is isomorphic to the projective space $\mathsf{P}(\Psi)$.
+Let $\psi \in \Psi$ be the state of a quantum particle.
+$e^{\imath \theta} \psi$ with $\theta \in \mathbbm{R}$ is not physically distinguishable from $\psi$ as phases are only physically distinguishable through addition.
+$\sqrt{\rho} \psi$ with $\sqrt{\rho} \in \mathbbm{R}^+$ cannot represent a quantum particle as it as a different amount of material.
+There exists only one distinguishable state for a quantum particle in the equivalence class $\psi \sim c \psi$ with $c \in \mathbbm{C}$.
+The state with no amount of material does not correspond to any particle state.
+The quotient space $\mathsf{P}(\Psi) \equiv (\Psi - \{0\}) / \sim$ is isomorphic to the state of quantum particles.
 \end{justification}
@@
 \begin{prop}\label{prop:inner_product}
-	The state space $\Psi$ for an irreducible material is an inner product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$.
-\end{prop}
-\begin{justification}
-	TODO
+	The state space $\Psi$ for an irreducible material is an inner product with $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$.
+\end{prop}
+\begin{justification}
+	We claim there is exists a map $| \cdot | ^2 : \Psi \rightarrow \mathbbm{R}$ such that $|\psi|^2 \ge 0 \; \forall \psi \in \Psi$ and $|\psi|^2 = 0 \Leftrightarrow \psi = 0$.
+Let $| \cdot | ^2 : \Psi \rightarrow \mathbbm{R}$ be the map that returns the amount of material for a given state.
+Let $\psi \in \Psi$ and $c \in \mathbbm{C}$.
+$|\tau(c) \psi|^2 = |c|^2 |\psi|^2$ as $\tau(c)$ increases the amount of material by $|c|^2$, which is always positive.
+$|\psi|^2 \ge 0 \; \forall \psi \in \Psi$.
+$|\psi|^2 = 0 \Leftrightarrow \psi = 0$ as the only state with no material is the empty state.
+
+	We claim $\forall \psi \in \Psi$ there exist a projection $\mathsf{P}_\psi : \Psi \rightarrow \Psi$ with $\psi \in \Psi$.
+Let $\psi \in \Psi$ be a state.
+Let $\mathsf{P}_\psi : \Psi \rightarrow \Psi$ be an operator that returns the part of a given state that is prepared as $\psi$.
+$\mathsf{P}_{\psi}(c_1 \phi_1 + c_2 \phi_2) = c_1 \mathsf{P}_{\psi}(\phi_1) + c_2 \mathsf{P}_{\psi}(\phi_2)$ is linear as the part of the composition prepared like $\psi$ is the composition of the parts so prepared.
+$\mathsf{P}_\psi \circ \mathsf{P}_\psi = \mathsf{P}_\psi$ is a projection because the whole $\mathsf{P}_\psi$  is prepared as $\psi$.
+
+	We claim $\mathsf{P}_{\psi} (\phi) = \phi \Leftrightarrow \phi = c \psi$ for some $c \in \mathbbm{C}$.
+Let $\psi, \phi \in \Psi$ such that $\mathsf{P}_{\psi} (\phi) = \phi$.
+The whole state $\phi$ is prepared as $\psi$.
+We can obtain $\phi$ from $\psi$ by a transformation in the unstated part.
+$\phi = c \psi$ for some $c \in \mathbbm{C}$.
+Let $\psi, \phi \in \Psi$ such that $\phi = c \psi$ for some $c \in \mathbbm{C}$.
+Then all of $\phi$ is prepared as $\psi$.
+$\mathsf{P}_{\psi} (\phi) = \phi$.
+Therefore $\mathsf{P}_{\psi} (\phi) = \phi \Leftrightarrow \phi = c \psi$.
+
+	We claim $|\psi|^2|\mathsf{P}_{\psi} (\phi)|^2=|\phi|^2|\mathsf{P}_{\phi} (\psi)|^2$.
+Let $\psi, \phi \in \Psi$.
+As the material is irreducible, $\psi$ and $\phi$ cannot be used to describe smaller parts.
+The fraction of $\phi$ that is prepared as $\psi$ must be equal to the fraction of $\psi$ that is prepared as $\phi$.
+$\frac{|\mathsf{P}_{\psi} (\phi)|^2}{|\phi|^2}=\frac{|\mathsf{P}_{\phi} (\psi)|^2}{|\psi|^2}$.
+
+	We claim $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi)) = r \psi$ for some $r \in \mathbbm{R}$.
+Let $\psi, \phi \in \Psi$.
+Consider $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi))$: we first take the part of $\psi$ prepared like $\phi$ and then take the part of that prepared like $\psi$.
+The overall operation is simply returning a part of $\psi$.
+These must correspond to the same exact configuration for the fragments.
+The only thing that can change is the amount of material.
+Therefore $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi)) = r \psi$ for some $r \in \mathbbm{R}^+$.
+
+	We claim that the state space $\Psi$ for an irreducible material has an inner product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$.
+$| \cdot | ^2$ and $\mathsf{P}_{\phi}$ satisfy the requirements in \ref{thrm:inner_product}.
+$\Psi$ is an inner product space with $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\langle \psi |\phi \rangle \psi = | \psi | ^2\mathsf{P}_{\psi}(\phi)$.
+
+	We claim that the inner product can be expressed in terms of the wave functions as $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$.
+The amount of material associated with a state $\psi \in \Psi$ is $\langle \psi | \psi \rangle = | \psi |^2 = \int_\mathcal{Q} \psi^\dagger (q^i) \psi(q^i) dq^i$.
+The inner product associated with the $L^2$ norm is $\int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$ with $\psi_1, \psi_2 \in \Psi$.
+Therefore $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$.
+
 \end{justification}
@@
 \begin{prop}\label{prop:self_adjoint_operators}
-	For each fragment variable $f$ there exists a corresponding self-adjoint linear operator $F : \Psi \rightarrow \Psi$.
-There exist a manifold $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ and $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ is an isomorphism.
-\end{prop}
-\begin{justification}
-	TODO
+	For each fragment variable $f$ there exists a corresponding self-adjoint linear operator $F : \Psi \rightarrow \Psi$ such that $\langle \psi | F \psi \rangle$ is the expectation of $f$ multiplied by the amount of material.
+There exist a topological space $\hat{\mathcal{Q}}$ such that $q^1 \equiv f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ and a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ where $\Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ is the space of wave functions over $\hat{\mathcal{Q}}$.
+\end{prop}
+\begin{justification}
+	We claim for each fragment variable $f$ there exist a topological space $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
+Let $f$ be a fragment variable.
+Let $\hat{q}^i$ be a set of fragment variables such that $\hat{q}^1\equiv f$.
+If $\hat{q}^i$ does not provide a maximal description of $f$ then there exists another set of fragment variables.
+Extend $\hat{q}^i$ with such variables.
+As the number of possibilities charted by $\hat{q}^i$ cannot exceed the number of possibilities charted by $q^i$, at some point we'll reach a maximal description.
+We call $\hat{\mathcal{Q}}$ the topological space charted by the fragment variables $\hat{q}^i$.
+
+	We claim there exists a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$.
+Let $\psi \in \Psi$ be a state.
+$q^i$ provides a maximal description for the fragments.
+There must exist a $\psi(q^i) : \mathbbm{R}^n \rightarrow \mathbbm{C}$ associated with $\psi$ such that it gives the fragment configuration at a particular value for $q^i$.
+Such wave function provides a full description for the state.
+Let $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \rightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ the map that associates a state with its wave function over $\hat{\mathcal{Q}}$.
+$\Upsilon_{\hat{\mathcal{q}}^i}$ is a vector space isomorphism for the same reason discussed in \ref{prop:wavefuntion}
+
+	We claim there exist a self-adjoint linear operator $F : \Psi \rightarrow \Psi$ such that $\langle \psi | F \psi \rangle$ is the expectation of $f$ multiplied by the amount of material.
+Let $f$ be a fragment variable.
+The expectation value multiplied by the amount of material is given by $\int f \rho(f) df$ where $\rho(f)$ is the density of material in $f$.
+Let $q^i$ a set of fragment variables that provides a maximal description for the fragments.
+Let $q^1 = f$ and $k=2,...,n$.
+We have $\rho(f) = \int \psi^\dagger(f, q^k) \psi(f, q^k) dq^{n-1}$.
+$\int f df \int \psi^\dagger(f, q^k) \psi(f, q^k) dq^{n-1} = \int \psi^\dagger(q^i) f(q^i) \psi(q^i) dq^{n}$.
+$f$ is a linear operator on the space of the wave function over $\hat{\mathcal{Q}}$ that increases the module at each point.
+TODO.
+Therefore $f(q^i) \psi(q^i) \in \Psi(\hat{\mathcal{Q}})$.
+As $\Upsilon_{\hat{\mathcal{q}}^i}$ is an isomorphism, there exists a linear operator $F : \Psi \rightarrow \Psi$ such that $f(q^i) \psi(q^i) = f(q^i) \Upsilon_{\hat{\mathcal{q}}^i}(\psi) = \Upsilon_{\hat{\mathcal{q}}^i}(F \psi)$.
+$\langle F \psi | \psi \rangle = \int (f(q^i) \psi(q^i))^\dagger \psi(q^i) = \int \psi^\dagger(q^i) f(q^i) \psi(q^i) = \langle \psi | F \psi \rangle$: $F$ is self-adjoint.
 \end{justification}
@@
 	The operator $P_i$ associated with the conjugate quantity of $q^i$ is $P_i = - \imath \hbar \partial_{i}$.
-The operator that changes the variable of the wave function from $q^i$ to $p^i$ is $\Upsilon^{q_i}_{p^i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^3} \int_{\mathcal{Q}} e^{\frac{q_i p^i }{\imath \hbar}} \psi(q^i) d^n q $.
-\end{prop}
-\begin{justification}
-	TODO
+The operator that changes the variable of the wave function from $q^i$ to $p_i$ is $\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^3} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
+The space of wave functions $\Psi(\mathcal{Q}, dq^i) = S(\mathcal{Q}, q^i)$ where $S(\mathcal{Q}, q^i)$ is the Schwartz space of rapidly decreasing smooth functions.
+\end{prop}
+\begin{justification}
+	We claim the operator associated with $k_i$ is $K_i = - \imath \partial_{i}$.
+Let $K_i : \Psi(\mathcal{Q}, dq^i) \rightarrow \Psi(\mathcal{Q}, dq^i)$ the linear operator associated with $k_i$ as express on the space of wave functions over $q^i$.
+Such operator must be covariant under the transformation $\hat{q}^i = \hat{q}^i(q^i)$.
+That is: $K_{\hat{q}^i} = \partial_{\hat{q}^i}q^iK_{q^i}$.
+By \ref{thrm:covariant_operator} $K_i=c\partial_i$ for some $c \in \mathbbm{C}$.
+$K_i$ is hermitian: $K_i=\imath r\partial_i$ for some $r \in \mathbbm{R}$ by \ref{thrm:antihermitian_derivative}.
+$k_i$ is expressed in units of inverse $q^i$ therefore $|r|=1$.
+By convention, we choose $r$ to be negative.
+$K_i = - \imath \partial_{i}$.
+
+	We claim the operator associated with $p_i$ is $P_i = - \imath \hbar \partial_{i}$. As $p_i=\hbar k_i$, their expectation is proportional by $\hbar$. $P_i=\hbar K_i= -\imath \hbar \partial_i$.
+
+	We claim the operator that changes the variable of the wave function from $q^i$ to $p_i$ is $\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^3} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
+Let $\Upsilon^{q_i}_{p^i} : \Psi(\mathcal{Q}, dq^i) \rightarrow \Psi(\mathcal{P}, dp^i)$ the operator that maps a wave function over $\mathcal{Q}$ into a wave function over $\mathcal{P}$, the space charted by $p_i$.
+$\Upsilon^{q_i}_{p^i}$ must exist and be unique (up to a total phase) as $\Upsilon^{q_i}_{p^i} = \Upsilon_{p^i}\Upsilon^{q_i} = \Upsilon_{p^i}(\Upsilon_{q_i})^{-1}$ is an isomorphism.
+$\Upsilon^{q_i}_{p^i}$ must be a linear transform that takes the operator $P_i= - \imath \hbar \partial_i$ as expressed over $q^i$ to $P_i = p_i$ as expressed over $p_i$.
+Such transformation is the Fourier transform.
+$\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^3} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
+
+	We claim $\Psi(\mathcal{Q}, dq^i) \subseteq S(\mathcal{Q}, q^i)$.
+Let $\psi(q^i) \in \Psi(\mathcal{Q}, dq^i)$.
+$P_i^n \psi(q^i) = (-\imath \hbar \partial_i)^n \psi(q^i)$ represents the n-th power of the conjugate momentum operator applied to the given state.
+As the quantity is physical, such operation must be defined for any $n$ therefore $\psi(q^i) \in C^\infty(\mathcal{Q})$.
+Let $\psi(p_i) = \Upsilon^{q^i}_{p_i} (\psi(q^i))$ be the wave function over momentum space.
+$(Q^i)^n \psi(p_i) = (+\imath \hbar \partial_{p_i})^n \psi(p_i)$ represents the n-th power of the $q^i$ operator applied to the given state.
+As the quantity is physical, such operation must be defined for any $n$ therefore $\psi(p_i) \in C^\infty(\mathcal{P})$.
+As $\psi(q^i)$ is the Fourier transform of an infinitely smooth function, it decreases faster then any inverse power of $q^i$.
+That is: $\forall m \in \mathbbm{Z}^+ \exists c \in \mathbbm{C}$ such that  $|\psi(q^i)|<\frac{c}{|q^i|^m} \; \forall |q^i|> 1$.
+$\Psi(\mathcal{Q}, dq^i) \subseteq S(\mathcal{Q}, q^i)$.
+
+	We claim $\Psi(\mathcal{Q}, dq^i) = S(\mathcal{Q}, q^i)$.
+Let $\psi(q^i) \in \Psi(\mathcal{Q}, dq^i)$ and $\hat{\psi}(q^i) \in S(\mathcal{Q}, q^i)$ such that $\psi(q^i) \neq \hat{\psi}(q^i)$.
+There exists $a, b \in \mathbbm{Z}^+$ such that, for a given $i$, $\lVert \psi(q^i) - \hat{\psi}(q^i) \rVert_{i,a,b} \neq 0$  where $\lVert \psi(q^i) \rVert = sup|(q^i)^a (\partial_i) ^b \psi(q^i)|$.
+The operator $(Q^i)^a(P_i)^b$ applied to the two functions will provide different distributions for the associated fragment variables.
+$\hat{\psi}(q^i)$ is physically distinguishable and is associated with a physical state.
+$\hat{\psi}(q^i) \in \Psi(\mathcal{Q}, dq^i)$
 \end{justification}
@@
 Let $x \in V$ and $\mathsf{P}_x : V \rightarrow V$ be a family of projections (i.e.
-$\mathsf{P}_x$ linear and $\mathsf{P}_x \circ \mathsf{P}_x = \mathsf{P}_x$) such that: $\mathsf{P}_x (y) = y \Leftrightarrow y = a x$ for some $a \in \mathbbm{C}$; $|y|^2|\mathsf{P}_x (y)|^2=|x|^2|\mathsf{P}_y (x)|^2$; $\mathsf{P}_x (\mathsf{P}_y (x)) = r x$ for some $r \in \mathbbm{R}$.
+$\mathsf{P}_x$ linear and $\mathsf{P}_x \circ \mathsf{P}_x = \mathsf{P}_x$) such that: $\mathsf{P}_x (y) = y \Leftrightarrow y = a x$ for some $a \in \mathbbm{C}$; $|x|^2|\mathsf{P}_x (y)|^2=|y|^2|\mathsf{P}_y (x)|^2$; $\mathsf{P}_x (\mathsf{P}_y (x)) = r x$ for some $r \in \mathbbm{R}$.
 	Then $V$ is an inner product space with the product $\langle \cdot , \cdot \rangle : V \times V \rightarrow \mathbbm{C}$ defined by $\langle x , y \rangle x = |x|^2 \mathsf{P}_x (y)$.
@@
 
-\begin{thrm}
+\begin{thrm}\label{thrm:antihermitian_derivative}
 	Let $G \equiv C^1(\mathcal{Q}, \mathbbm{C}) \cap L^1(\mathcal{Q}, \mathbbm{C})$ be the space of differentiable and integrable complex functions defined over $\mathcal{Q}$.

### 2016-11-01 2015CompleteDerivation/completeDerivation.tex
First pass at quantum

@@
 
-\begin{assump}[Infinitesimal reducibility]\label{ass:infinitesimal_reducibility}
+\begin{assump}[Infinitesimal reducibility]\label{ass:irreducibility}
 	The system under study is composed of an irreducible homogeneous material and as a whole undergoes deterministic and reversible evolution.
@@
 
-\begin{prop}\label{unitary_evolution}
+\begin{prop}\label{prop:unitary_evolution}
 	A deterministic and reversible evolution map $\mathcal{T}_{\Delta t}: \Psi \rightarrow \Psi$ for an irreducible material is a unitary operator.
 That is: $\langle \mathcal{T}_{\Delta t} \psi_1, \mathcal{T}_{\Delta t} \psi_2 \rangle = \langle \psi_1 , \psi_2 \rangle$ where $\psi_1, \psi_2 \in \Psi$.
-The time evolution for the wave function $\Upsilon_q \mathcal{T}_{\Delta t} \Upsilon^q : S(\mathcal{Q}, dq) \rightarrow S(\mathcal{Q}, dq)$ is a $C^{\infty}$-diffeomorphism.
-\end{prop}
-
-\begin{justification}
+\end{prop}
+
+\begin{justification}
+	We claim $\mathcal{T}_{\Delta t}: \Psi \rightarrow \Psi$ is a unitary operator.
+Let $\psi \in \Psi$ be a state.
+$|\psi|^2 = |\mathcal{T}_{\Delta t} \psi|^2$ because deterministic and reversible evolution does not change the amount of material.
+Let $\psi_1, \psi_2 \in \Psi$ be two states.
+$\mathcal{T}_{\Delta t}(a\psi_1+b\psi_2) = a \mathcal{T}_{\Delta t} \psi_1 + b \mathcal{T}_{\Delta t} \psi_2$ with $a,b \in \mathcal{C}$ as the composition of the evolution is the evolution of the composition: $\mathcal{T}_{\Delta t}$ is a linear operator.
+$\mathcal{T}_{\Delta t} \mathsf{P}_{\psi_1} (\psi_2) = \mathsf{P}_{\mathcal{T}_{\Delta t}\psi_1} (\mathcal{T}_{\Delta t}\psi_2)$ as the evolved part of $\psi_2$ that was prepared like $\psi_1$ is the part of the evolved $\psi_2$ that ends prepared like the evolved $\psi_1$.
+We have $\mathcal{T}_{\Delta t} \mathsf{P}_{\psi_1} (\psi_2) = \mathcal{T}_{\Delta t} \langle \psi_1 , \psi_2 \rangle \psi_1 = \langle \psi_1 , \psi_2 \rangle \mathcal{T}_{\Delta t} \psi_1 = \mathsf{P}_{\mathcal{T}_{\Delta t}\psi_1} (\mathcal{T}_{\Delta t}\psi_2) = \langle \mathcal{T}_{\Delta t}\psi_1 , \mathcal{T}_{\Delta t}\psi_2 \rangle \mathcal{T}_{\Delta t} \psi_1$.
+$\langle \psi_1 , \psi_2 \rangle = \langle \mathcal{T}_{\Delta t}\psi_1 , \mathcal{T}_{\Delta t}\psi_2 \rangle$.
+$\mathcal{T}_{\Delta t}$ is unitary.
 \end{justification}
@@
 \begin{justification}
-\end{justification}
-
-Note: we get time dependent Schroedinger. No basis for time independent.
+	We claim the state evolves according to Schroedinger's equation.
+Let  $\mathcal{T}_{dt}: \Psi \rightarrow \Psi$ be the evolution for an infinitesimal time interval $dt$.
+As it is unitary, it can be expressed as $\mathcal{T}_{dt} = I + \frac{H dt}{\imath \hbar}$ where $H : \psi \rightarrow \psi$ is a self-adjoint operator.
+Let $\psi_t \in \Psi$ be a state and $\psi_{t+dt} \in \Psi$ its evolution after an infinitesimal time interval.
+We have $\mathcal{T}_{dt} \psi_t = \psi_{t+dt} = \psi_t + \frac{H dt}{\imath \hbar} \psi_t$.
+$\imath \hbar \frac{\psi_{t+dt} - \psi_t}{dt} = \imath \hbar \partial_t \psi_t = H\psi_t$.
+\end{justification}
+
+We recognize the Schroedinger equation, the time evolution equation for a quantum particle state.
+
+Note, though, that nothing tells us that the system must be in an eigenfunction of $H$ or that it the ground state is somehow preferred.
+In fact, the eigenfunctions of $H$ may not be part of $S(\mathcal{Q}, q^i)$ and therefore may not be physical.
+In other words: we have no basis for the time independent Schroedinger equation.
+This is actually consistent.
+For example, for a system to reach its ground state it has to be able to radiate energy: it cannot isolated.
+Therefore this dynamic cannot be described under the deterministic and reversible assumption.
+This already hints that deterministic and reversible evolution is not the only dynamic possible.
+Other considerations outside of the scope of this work must be used to recover this behavior (e.g.
+thermodynamics or decoherence).
+
+\subsection{Time dependence and kinematic assumption}
+
+The next step would be to integrate the quantum description with time dependent evolution and the kinematic assumption.
+We will only provide a sketch, though, without going through the details.
+These would take a significant amount of space while not providing any significantly new insight.
+Also, given that we are limited to single particles (i.e.
+no fields) without spin (i.e.
+minimal material only described by position), our results are limited to finding the Klein-Gordon equation modified with electromagnetic interactions.
+While this is still of note, the equation is of limited physical use and therefore may not warrant the space necessary for a more rigorous derivation.
+
+As for the classical case, the first thing to do is to extend $\mathcal{Q}$ to include time.
+This gives us the same $\mathcal{M}$ as before.
+A quantum state is a complex valued distribution $\psi : \mathcal{M}_{t=t_0} \rightarrow \mathbbm{C}$ over an hypersurface at constant time.
+An evolution $\bar{\psi} : \mathcal{M} \rightarrow \mathbbm{C}$ is a complex valued function defined on the whole $\mathcal{M}$.
+The operator for the quantity conjugate of time is $E=\imath\hbar\partial_t$.
+
+Next we need to extend integration.
+In the quantum case, this means to make sure the inner product $\langle \psi | \phi \rangle$ remains unchanged under all variable changes, including time.
+This allows to describe the amounts of material and their distribution over fragments in a way that is independent of our description.
+
+As in the classical case, we use $s$ as the evolution parameter.
+Deterministic and reversible evolution $\mathcal{T}_{ds}$ will be a unitary transformation defined over this extended inner product, as the amounts of material get conserved.
+Continuous evolution will admit an invariant Hamiltonian $\mathcal{H}$ such that $\imath \hbar \partial_s \psi = \mathcal{H} \psi$.
+As for the composite state evolution in the classical case, $\bar{\psi}$ remains unchanged by the evolution as it transports the distribution along $s$.
+Therefore $\mathcal{H} \hat{\psi} = \mathcal{h} \hat{\psi}$ is an eigenfunction of $\mathcal{H}$.
+As in the classical case, we use $\mathcal{H}$ to be zero to identify the state space: the temporal d.o.f.
+is not independent and therefore there is a constraint between the variables.
+Therefore we have $\mathcal{H} \psi = 0$.
+This is the time dependent evolution equation.
+
+Note that states are indeed eigenfunctions of the invariant Hamiltonian $\mathcal{H}$: here we do have a justification, as opposed to the standard Hamiltonian $H$ case.
+For example, a massive particle may have fragments with different energy, but they all must have the same mass.
+Also note that $\mathcal{H} \psi = 0$ is basically the form of all equations (i.e.
+Klein-Gordon, Dirac, Maxwell, ...) in quantum field theory.
+The respective invariant Hamiltonian $\mathcal{H}$ is the operator associated with each equation.
+
+Having handled the time dependent case, we can turn our attention to the kinematic equivalence assumption.
+While the trajectory of a fragment is not well defined in either phase space or physical space, what we can describe remains the same under the assumption.
+Therefore the distributions still need to be transported from state space to kinematic variables up to a constant factor.
+Which means we still have a linear relationship between velocity and conjugate momentum.
+In terms of the quantum operators: $m\, d_s Q^\alpha = g^{\alpha\beta} P_{\beta} - A^\alpha = m\frac{[Q^\alpha, \mathcal{H}]}{\imath \hbar}$.
+
+This leads to the same general for the invariant Hamiltonian $\mathcal{H}=\frac{1}{2m}((P_\alpha-A_\alpha)g^{\alpha\beta}(P_\beta-A_\beta) + m^2 c^2) = \frac{1}{2m}((-\imath\hbar\partial_\alpha-A_\alpha)g^{\alpha\beta}(-\imath\hbar\partial_\beta-A_\beta) + m^2 c^2)$.
+In the free particle case, we have $\frac{1}{2m}(-\hbar^2 \partial^\alpha \partial_\alpha + m^2 c^2) \bar{\psi} = 0$ which is the Klein-Gordon equation.
+We can also write the operator for kinetic momentum as $mU^\alpha = P^\alpha - A^\alpha = i\hbar D^\alpha$ where $D^\alpha = \partial^\alpha - \frac{A^\alpha}{\imath \hbar}$ is the covariant gauge derivative.
+We have $(\frac{1}{2}mU^\alpha U_\alpha + \frac{1}{2} m c^2) \bar{\psi} = \frac{1}{2m}(-\hbar^2 D^\alpha D_\alpha + m^2 c^2) \bar{\psi} = 0$ which is the Klein-Gordon equation with electromagnetic interaction.
+
+Note how the covariant derivative is, up to a constant, kinetic momentum.
+This motivates why it is such an important operator.
+As for the Klein-Gordon equation, the only thing it does is to impose that the norm of the four velocity is the speed of light, a condition much more trivial one may have anticipated.
+
+It should be clear, even without the details, that all the assumptions come together even in the quantum case.
+They give more insight to all the pieces in a comprehensive way and help tear down the artificial walls between the different physical theories.
+This allows to form a picture of fundamental physics that is more unified, something that has been sorely missing.
+
+\subsection{Non deterministic evolution and the projection postulate}
+
+As we made it clear from the start, deterministic and reversible evolution is an assumption.
+And as we seen in the justifications for assumptions \ref{ass:infinitesimal_reducibility} and \ref{ass:irreducibility} we cannot have determinism at all levels of decomposition while transferring information from one system to another.
+
+In this section we will see how non deterministic processing become necessary when describing measurements for irreducible systems.
+But we'll see that, under the same assumption of irreducibility, we are able to make a connection between the output of deterministic and non deterministic processes.
+
+Suppose we have a quantum particle state $\psi$ we want to identify.
+Suppose that we have a detector that is able to capture the particle.
+If it's captured, we will know that the particle was in the region of space enclosed by the detector.
+
+Suppose that we create an array of such detectors. This will allow to detect the position of the particle with a resolution given by the aperture of each detector.
+
+Suppose now, though, that the quantum particle state is spread over a region significantly greater than a single detector but smaller than the array. What will it happen?
+
+It is clear that it has to be absorbed as it has no way to escape.
+It is also clear that it must be absorbed whole: as it is an irreducible system, the detector cannot partially interact with it, it cannot half absorb it.
+So a single detector will capture it.
+It is also clear that it cannot always be the same detector: that corresponds to states where the particle is spread within the aperture of that detector.
+In other words: as the particle hits the array, one detector triggers randomly based on the distribution of the particle.
+
+The process is non deterministic in the sense that the state of the particle is not enough to predict which detector will trigger.
+Maybe it depends on the state of the detector, maybe it depends on the unstated part of the particle.
+The point is that a single take does not tell us much about the incoming particle states.
+
+We then repeat the process, making sure the incoming particle is in the same state.
+Given a single "pure" incoming particle state, we get a "mixed" statistical distribution describing the output of the array.\footnote{Note that this is different from the model used often to describe the quantum measurement problem, where the incoming system becomes entangled with the measuring device.
+We do not make any claim on what the final state of the particle should be: in many cases, like in a photomultiplier tube, the particle is gone and there is no state.
+We simply have one incoming state for the particle, prepared over and over exactly the same, and the outgoing state for the measurement device, which is a classical statistical distribution over the different takes.} The question is: how is the statistical distribution coming out of the detector related to the material distribution coming in with the particle?
+
+If the detector does not introduce biases, then the final distribution is only a function of the incoming particle: all unbiased detectors will give the same result.
+The unstated part also cannot affect the final distribution of outcomes: it is not physically distinguishable.
+The final statistical distribution of the detector must match the incoming material distribution of the particle.
+That is if $N$ is the total number of takes and $N_A$ is the number time the detector with aperture in the $A$ region captured the particle, the ration $\frac{N_A}{N}$ will be approximately the amount of material within the aperture $\int_A \psi^\dagger(q^i) \psi(q^i) dq^i$.
+If we assume an infinite amount of takes, we expect them to coincide: $\int_A \psi^\dagger(q^i) \psi(q^i) dq^i = \lim\limits_{N\rightarrow\infty} \frac{N_A}{N}$.
+But this is also the frequentist definition of probability.
+Therefore we have $\int_A \psi^\dagger(q^i) \psi(q^i) dq^i = P(\textrm{measuring particle in } A)$ thus recovering the probabilistic interpretation of the wave function.
+
+We can conceptually extend this picture to the generic case.
+Suppose we know the outcomes of our measurement are parametrized by a fragment variable $\hat{q}$.
+We will have a linear functional $\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0} : \Psi \rightarrow \mathbbm{C}$ such that $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2$ returns the amount of material prepared with the particular value $\hat{q}_0$.
+We can re-express this quantity in terms of the dual and the inner product extended to all Lebesgue integrable functions: $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2 = |\langle \hat{q}_0 | \psi \rangle|^2 $.
+This is also the probability that the outcome associated with $\hat{q}_0$ is selected by the measuring device, thus recovering the statistical nature of the quantum predictions.
+
+Note that while this process is of a non-deterministic nature, it is of a specific non-deterministic nature.
+It requires that no other state can influence the final distribution or the distribution of measurement outcomes would not match the distribution of incoming material.
+That is: it cannot be understood as an averaging over some unknown external state.
+
+Also note that while we have a way to predict the distribution over final outcomes from the initial state, we have no way of describing what happens in between.
+When we assume that only the initial state matters, we are saying that we can disregard the particulars of the non-deterministic evolution.
+So much so that we can substitute it with a deterministic one to compute the probability distribution.
+In particular, we have no way of predicting which set of final outcomes we will be projecting to, nor when does that actually happen.
+
+While this may seem like an incomplete description, it is as complete as the irreducibility assumption allows.
+The state is known and determines the final distribution.
+The unstated part cannot be known, and influences the non-deterministic evolution.
+If we were able to say something more about the non-deterministic evolution, we would be able to describe an aspect of the unstated part, which we cannot do without violating the assumption.
+If we cannot learn anything about the unstated part from the final distribution, then we are consistent with our assumption and we are able to use deterministic evolution and projection to determine the final mixed state.
+In other words: the inability to describe the unstated part must be matched with a corresponding inability to describe its evolution.
+
+We therefore have arrived to all the major underpinning of quantum mechanics and see how they stem from assumption \ref{ass:irreducibility} of irreducibility.
+We have somewhat a closure, probably not satisfactory to many, for some of the open problems in quantum mechanics.
+Not because we solved them, we haven't at all, but because we have good reason to have them.
+The desire to know ``what is really going on" is hindered by assuming that no description can be given below the level of a particle.
+
+On the upside, irreducibility is what it is: an assumption.
+As with infinitesimal reducibility, it is a reasonable initial assumption assumption.
+There is hope that, if we abandon it, we can have a more satisfying and complete description.
+What should be clear, though, that abandoning irreducibility may entail much much more than just finding another equation of motion.
 

### 2016-11-01 2015CompleteDerivation/completeDerivation.tex
Second pass for prelude to quantum

@@
 
-In principle, we could simply introduce assumption \ref{???} and work our way to quantum mechanics.
-Unfortunately, given the history and baggage associated with quantum systems, there is a high chance of being misconstrued.
-The intuitive pictures people have constructed or inherited are varied, may start from different premises and therefore it is easy to misinterpret what one is saying.
-Therefore we will first ground in classical mechanics the intuition that will be needed later, making the step into full quantum mechanics a gentler, and hopefully less confusing, one.
+While in principle, we could simply introduce assumption \ref{ass:irreducibility} and work our way to quantum mechanics, there is a non insignificant chance of being misconstrued.
+Different people come with different background, to the point that the same words trigger different intuitions and expectations.
+Therefore we will first ground in classical mechanics the intuition that will be needed later, making the step into quantum mechanics a gentler, and hopefully less confusing, one.
+It will also help us in showing that classical and quantum mechanics are not, after all, so different.
 
 We will see that, as a first approximation, quantum particles are more keenly described by classical composite states (i.e.
-distributions) as, unlike classical particles, they are made of a finite amount of material.
-We will also see that classical distributions already possess qualitative properties that are typically associated with quantum systems and construct a classical analogue of the uncertainty principle.
-
-\subsection{Particles and distributions}
-
-As we saw in the previous sections, classical particles should not be understood as point-like objects of finite mass as this would open a series of problems and inconsistencies.
-The issues created by such misidentification are only amplified within the context of quantum mechanics.
-In fact, many issues preventing a satisfying intuitive picture stem from this confusion.
-
-The problem goes along these line.
-One assumes, implicitly or explicitly, that quantum particles are point-like.
-But they don't always behave like one, they also behave like a wave, a distribution.
-During measurements, this distribution coincides experimentally with a probability distribution.
-So the state of a quantum particle is a probability distribution.
-Except that it self-interacts.
-So its a special quantum probability distribution as the quantum particle behaves like its in more than one place at once.
-We learn this picture and, sometimes begrudgingly, we get used to it.
-
-Here we contend that this picture is not only confusing and unnecessary, it also prevents to understand fully the deep links between classical and quantum mechanics, in the same way that thinking of classical particles as point-like objects prevented to see a more holistic picture in classical mechanics.
-
-The first thing to note is that all physical objects behave like they are in more than one place at once.
+a distribution in phase space) as they are made of a finite amount of material.
+In fact, classical distributions already possess qualitative properties that are typically associated with quantum systems such as a classical analogue of the uncertainty principle.
+
+\subsection{The case against point particles}
+
+As we defined them in previous sections, classical particles should not be understood as point-like objects of finite mass.
+It was only by assuming they were infinitesimal parts of an invariant density distribution that gave us conjugate pairs of state variables.
+Hamiltonian mechanics would not have been justified otherwise.
+
+Any finite amount of classical material is described by a distribution.
+Therefore, to a first approximation, an electron, which is a finite amount of charge/mass, must be described by a distribution as well.
+This may seem preposterous at first, but we contend that it is no more preposterous than to think it point-like, described by a probability distribution that self-interacts.
+In fact, the idea of an electron being a distribution of material is already conceptually more compatible with two issues.
+
+The first one is that an electron does not seem to have a well defined position or momentum.
+If we misidentify the electron with a classical particle, this would indeed be worth of note and puzzlement.
+But if the electron is a classical distribution, there is not issue.
 The earth is both under Milano and New York.
 Your right index finger is not, hopefully, at the same position of your left eye.
-In fact, based on what we saw before, any object of finite mass is a distribution over a region, as discontinuous distributions are ruled out.
-An electron is an object of finite mass.
-Why should it be any different?
-In fact, if one claimed an electron to be point-like, he would be claiming an electron behaves in a different way than all other physical objects.
+In fact, based on what we saw before, any object of finite mass is a distribution over a region, and discontinuous distributions are ruled out by \ref{continuous_map}.
+An electron is an object of finite mass: why should it be any different?
+In fact, if one claimed an electron to be point-like, he would be claiming an electron behaves in a different way than all other physical objects of finite mass (which are the only ones we experimentally work with).
 Shouldn't \emph{he} have the burden of proof?
 
-Part of the confusion seem to be the idea that, since the electron is an elementary particle, it should have no spatial extent.
-But that does not follow.
-Lack of spatial extend and lack of substructure are independent concepts.
-In fact: we can construct an infinitesimal object that is not elementary.
+The other issue is particle-wave duality.
+If the spread of the distribution is narrow enough for the problem at hand, the motion can be described by the position and center of mass, making it behave like a classical particle.
+If the spread is not negligible (e.g.
+it encounters obstacles of a smaller scale) then we have to use the full distribution.
+Therefore, even in this classical approximation, the electron would have this dual nature.
+
+While this picture is quite reasonable, some people may reject it because they believe that a fundamental object must have no size, it must be a point.
+But that does not follow: lack of spatial extend and lack of substructure are independent concepts.
+
+In fact, we can construct an infinitesimal object that is not elementary.
 Suppose we have a sphere with dots moving on its surface.
@@
 If we scale the sphere, the dynamics stays the same.
-We can shrink the whole sphere to a point: it is now a point-like object that is not elementary, as there is motion on its surface.
-
-Not only point-like object don't have to be elementary, but it's not clear how they can be elementary.
+We can shrink the whole sphere to a point: it is now a point-like object that is not elementary, as there is motion on its surface detectable along different angle of incidence.
+
+Not only point-like objects don't have to be elementary, but it's unclear how can they be elementary.
 How can the whole universe be reduced to  elements that have no idea how many spatial dimensions are there?
 Mathematically a set of points by itself does not define the dimension of a space: $\mathbbm{R}^n$ is the same order of infinity for any $n$.
-One needs a topology, which is an additional structure.
-If everything is just points, where does this come from?
-
-Entertaining the thought that an electron is actually physically distributed in space may seem preposterous, probably because it's not how we are taught to think about it.
-But it's no more preposterous than to think it point-like, described by a probability distribution that self-interacts.
-
-The first payoff of accepting that a quantum particle is a finite amount of material distributed in space is that this is more consistent with classical mechanics as we have defined it\footnote{It is also more consistent with general relativity.}.
-The second payoff is that some features of classical mechanics already explain part of the behavior that is usually thought as inherent to quantum mechanics.
+One needs a topology to differentiate, which is an additional structure.
+It's the neighborhoods that define the dimensionality.
+If everything is just points, where does the topology come from?
+
+If we abandon the idea of point-like objects of finite mass, we avoid these problems and get a picture more consistent across the different realms of physics.
+And we can also see that classical mechanics already has some features that are thought to be inherent to quantum mechanics.
 
@@
 
-Suppose we have a finite amount of classical material distributed over a single d.o.f.
-$a_m = \int \rho_\mathcal{c} dq \wedge dp$.
-The Shannon entropy associated with the distribution, which can be though as the amount of information required to identify a particle of the distribution, is finite as well.
-$I(\rho_\mathcal{c}) = \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq \wedge dp = I_0$.
+Suppose we have our classical electron described by a finite amount of classical material distributed over a single d.o.f.
+$\int \rho_\mathcal{c} dq \wedge dp = 1$ where $\rho_\mathcal{c}$ is the normalized distribution.
+In this context, we will call \emph{fragment} an infinitesimal part of our classical electron.
+The Shannon entropy associated with the distribution, which can be though as the amount of information required to identify a fragment given the distribution, is finite as well.\footnote{To be precise, Shannon's differential entropy is actually the number of bits needed relative to a uniform distribution of unitary range on all axis.
+The intuitive meaning, though, is the same.} $I(\rho_\mathcal{c}) = \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq \wedge dp = I_0$.
 TODO: prove in appendix.
 
-As the state evolves into its final state, the amount of material remains the same: $a_m = \int \rho_{\hat{\mathcal{c}}} dq \wedge dp$.
-The final entropy will also remain the same since the entropy, like the distribution itself, is just another integrable function: $I(\rho_{\hat{\mathcal{c}}}) = I_0$.
-
-In the same way that each particle is constrained to move along states at equal energy, the composite state is constrained to move along distributions at equal entropy.
-This makes sense: if the motion is deterministic and reversible, the same amount of information is required to identify a particle of material.
+As the state evolves into its final state, the amount of material remains the same: $\int \rho_{\hat{\mathcal{c}}} dq \wedge dp = \int \rho_{\mathcal{c}} dq \wedge dp$.
+The final entropy will also remain the same: $I(\rho_{\hat{\mathcal{c}}}) = I_0$.
+Mathematically because $\rho_\mathcal{c} \log (\rho_\mathcal{c})$, like the distribution itself, is just another integrable function.
+Physically, because the amount of information required to identify a fragment given the distribution has to remain the same under deterministic and reversible evolution.
+Therefore, in the same way that each fragment is constrained to move along states at equal energy, the composite state is constrained to move along distributions at equal entropy.
 
 We now ask the following question: once we fix the entropy, what is the relationship between the spread of the marginal distributions along $q$ and $p$?
-Using Lagrange multipliers, we fix the total amount of material $a_m$ and the total entropy $I_0$ and find the distribution that minimizes the the product of the variance $\sigma_q^2 \sigma_p^2 = \int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp$.
+Using Lagrange multipliers, we fix the total amount of material to $1$ and the total entropy $I_0$ and find the distribution that minimizes the the product of the variance $\sigma_q^2 \sigma_p^2 = \int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp$.
 \begin{align*}
 L = &\int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp \\
-&+ \lambda_1(\int \rho dqdp - \rho_{tot})\\
-&+ \lambda_2(- \int \rho \ln \rho \, dqdp - H_0)\\
+&+ \lambda_1(\int \rho dqdp - 1)\\
+&+ \lambda_2(- \int \rho \ln \rho \, dqdp - I_0)\\
 \delta L = &\int \delta \rho [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\ &+ \lambda_1 + \lambda_2 \ln \rho + \lambda_2 ] dqdp = 0 \\
-\rho = &\frac{\rho_{tot}}{ \sqrt{4 \pi ^2\sigma_q^2\sigma_p^2}} e^{-\frac{(q-\mu_q)^2(p-\mu_p)^2}{4\sigma_q^2\sigma_p^2}} \\
-= &\frac{\rho_{tot}}{e^{H_0 - 1}} e^{-\frac{\pi ^2(q-\mu_q)^2(p-\mu_p)^2}{\exp (2(I_0 - 1))}}
+\rho = &\frac{1}{ \sqrt{4 \pi ^2\sigma_q^2\sigma_p^2}} e^{-\frac{(q-\mu_q)^2(p-\mu_p)^2}{4\sigma_q^2\sigma_p^2}} \\
+= &\frac{1}{e^{I_0 - 1}} e^{-\frac{\pi ^2(q-\mu_q)^2(p-\mu_p)^2}{\exp (2(I_0 - 1))}}
 \end{align*}
@@
 \end{align*}
-This is strikingly similar to the Heisenberg uncertainty principle, except that the value depends on the initial state. This can be addressed with some extra assumptions.
-
-Suppose the evolution of the distribution is deterministc and reversible but that each particle is undergoing non deterministic motion.
+This is strikingly similar to the Heisenberg uncertainty principle, except that the value depends on the initial state $I_0$.
+
+Suppose, though, the evolution of the distribution is deterministc and reversible while each fragment is undergoing non deterministic motion.
 Suppose $\rho$ is a distribution at equilibrium of such motion.
-While each particle moves randomly, the overall distribution $\rho$ stays the same: the non deterministic motion is effectively just reshuffling the particles within the same distribution.
-The value of the entropy of the distribution is set by the non deterministic motion of the particles: the stronger it is, the greater the entropy.
-
-Suppose now $\rho$ evolves adiabatically.
+While each fragment moves randomly, the overall distribution $\rho$ stays the same: the non deterministic motion is effectively just reshuffling the fragments within the same distribution.
+The value of the entropy of the distribution, and therefore its spread, is set by the non deterministic motion of the particles: the stronger it is, the greater the entropy.
+
+Suppose now the classical electron undergoes quasi-static evolution.
 At each instant we can assume $\rho$ to be at equilibrium.
-At each instant the particles are reshuffled but the amount of entropy of the distribution remains the same.
-The overall evolution is still Hamiltonian, but the trajectories predicted by the framework are now the average motion of the particles.
-
-If we assumed that all materials were interacting with the same background that caused all particles to exhibit a random walk of the same magnitude, as defined by the information entropy, then the spread of all distributions for all material will satisfy the same inequality.
-This is gives us the classical uncertainty principle: $\sigma_q\sigma_p \geq \exp (I_\varnothing - 1) / 2 \pi $ where $I_\varnothing$ is the entropy associated with the background.
+At each instant the fragments are reshuffled but the amount of entropy of the distribution remains the same.
+The overall evolution is still Hamiltonian, but the trajectories predicted by the framework are now just the average motion of the fragments.
+
+If we assumed that all our classical electrons were interacting with the same background that caused all fragments to exhibit a random walk of the same magnitude, as defined by the information entropy, then the spread of all distributions for all material would satisfy the same inequality.
+This gives us the classical uncertainty principle: $\sigma_q\sigma_p \geq \exp (I_\varnothing - 1) / 2 \pi $ where $I_\varnothing$ is the entropy associated with the background.
 
 Assuming such a background is not far fetched.
-In fact, it serves as a remainder of the issues mentioned in assumption \ref{ass:infinitesimal_reducibility}: as determinism and reversibility at all scales conceptually does not work, there exists a minimum constant interaction to all particles: the system is not perfectly isolated.
-If the entropy of our distribution is comparable to the entropy introduced by the background, we are likely in the range where classical mechanics is not valid as the random motion of each particle is not negligible.
+In fact, it serves as a remainder of the issues mentioned in assumption \ref{ass:infinitesimal_reducibility}: as determinism and reversibility at all scales conceptually does not work, there exists a minimum constant interaction: the system cannot be perfectly isolated.
+If the entropy of our distribution is comparable to the entropy introduced by the background, we are likely in the range where classical mechanics is not valid as the random motion of each fragment is not negligible.
 
@@
 Note, though, that this result is similar in spirit to the Gromov non-squeezing theorem and the symplectic camel, which are already generalized to multiple d.o.f..
-It is the link with information theory that gives us further physical insight.
-
-It should now be clear how in classical mechanics we can already find some of the ingredients that are typically assumed to be characteristics of quantum mechanics.
+It is the link with information theory that gives us further physical insight and is of particular interest given the strong connection between thermodynamics and Shannon's entropy.
+This is something to be developed in a future work.
+
+At any rate, this discussion shows, once again, how in classical mechanics we can already find some of the ingredients that are typically assumed to be characteristics of quantum mechanics.
 It has its own flavor of anti-particles, CPT theorem and uncertainty principle.
-These are general features of an evolution that is deterministic, reversible and adiabatic (i.e.
-Hamiltonian dynamics).
-Quantum mechanics, as we'll see, just inherits these traits.
+These should really be considered general features of deterministic and reversible evolution of distributions (i.e.
+Hamiltonian mechanics).
+Quantum mechanics just inherits these traits.
+
+The discussion also gives us an intuitive physical picture that we will develop further in the next section: a quantum particle is a well determined distribution of not well determined fragments.
+The distribution as a whole can be studied and evolves deterministically while the fragments do not.
 
@@
 
-In this section we turn to a different case in which the relationship between the whole material and its parts is different from the one described in section \ref{sec:reducibility}.
+In this section we explore a different relationship between the whole material and its parts from the one described in section \ref{sec:reducibility}.
 The system is still infinitesimally decomposable, made of infinitesimal parts as before, but this time it is irreducible.

### 2016-11-02 2015CompleteDerivation/completeDerivation.tex
Another pass at quantum prelude

@@
 
-While in principle, we could simply introduce assumption \ref{ass:irreducibility} and work our way to quantum mechanics, there is a non insignificant chance of being misconstrued.
-Different people come with different background, to the point that the same words trigger different intuitions and expectations.
-Therefore we will first ground in classical mechanics the intuition that will be needed later, making the step into quantum mechanics a gentler, and hopefully less confusing, one.
-It will also help us in showing that classical and quantum mechanics are not, after all, so different.
+Before introducing assumption \ref{ass:irreducibility} and deriving quantum mechanics as its consequence, we first give a semi-classical account of the conceptual model we will be using.
+This allows to build the intuition more gradually and to show that classical and quantum mechanics are not as different as one may first imagine.
 
 We will see that, as a first approximation, quantum particles are more keenly described by classical composite states (i.e.
-a distribution in phase space) as they are made of a finite amount of material.
-In fact, classical distributions already possess qualitative properties that are typically associated with quantum systems such as a classical analogue of the uncertainty principle.
+a distribution in phase space) like any other object made of a finite amount of material.
+We'll also see that classical distributions already possess qualitative properties that are typically associated with quantum systems, including a classical analogue of the uncertainty principle.
 
@@
 As we defined them in previous sections, classical particles should not be understood as point-like objects of finite mass.
-It was only by assuming they were infinitesimal parts of an invariant density distribution that gave us conjugate pairs of state variables.
-Hamiltonian mechanics would not have been justified otherwise.
-
-Any finite amount of classical material is described by a distribution.
-Therefore, to a first approximation, an electron, which is a finite amount of charge/mass, must be described by a distribution as well.
-This may seem preposterous at first, but we contend that it is no more preposterous than to think it point-like, described by a probability distribution that self-interacts.
+It was their being infinitesimal and spread around a small neighborhood that gave us conjugate pairs of state variables and Hamiltonian evolution.
+Any finite amount of classical material, instead, is described by a distribution.
+
+If we take a quantum particle, say an electron, it has a finite amount of charge/mass/energy.
+Therefore, to a first approximation, it must be described by a distribution as well.
+This may, to some, seem preposterous at first: an electron is not a point particle.
+But we contend that it is no more preposterous than to think it point-like, described by a probability distribution that self-interacts.
 In fact, the idea of an electron being a distribution of material is already conceptually more compatible with two issues.
@@
 If we misidentify the electron with a classical particle, this would indeed be worth of note and puzzlement.
-But if the electron is a classical distribution, there is not issue.
+But if the electron is a classical distribution, there is no issue.
 The earth is both under Milano and New York.
@@
 An electron is an object of finite mass: why should it be any different?
-In fact, if one claimed an electron to be point-like, he would be claiming an electron behaves in a different way than all other physical objects of finite mass (which are the only ones we experimentally work with).
+In fact, if one claimed an electron to be point-like, he would be claiming an electron behaves in a different way than all other physical objects of finite mass/energy (which are the only ones we experimentally work with).
 Shouldn't \emph{he} have the burden of proof?
@@
 The other issue is particle-wave duality.
-If the spread of the distribution is narrow enough for the problem at hand, the motion can be described by the position and center of mass, making it behave like a classical particle.
+If the spread of the distribution is narrow enough for the problem at hand, the motion can be described by just the position and momentum of the center of mass, making it behave like a classical particle.
 If the spread is not negligible (e.g.
 it encounters obstacles of a smaller scale) then we have to use the full distribution.
-Therefore, even in this classical approximation, the electron would have this dual nature.
+Therefore, even in this classical approximation, the electron would have this dual nature depending on the problem.
 
@@
 
-In fact, we can construct an infinitesimal object that is not elementary.
-Suppose we have a sphere with dots moving on its surface.
-All their dynamics can be expressed by their polar and azimuthal angles.
-If we scale the sphere, the dynamics stays the same.
-We can shrink the whole sphere to a point: it is now a point-like object that is not elementary, as there is motion on its surface detectable along different angle of incidence.
+In fact, we can conceptually construct an infinitesimal object that is not elementary.
+Suppose we have dots moving on the surface of a sphere.
+Their motion can be expressed in terms of their polar and azimuthal angles.
+If we shrink the sphere, the dynamics and its description stays the same.
+We can collapse the whole sphere to a point: it is now a point-like object that is not elementary, as there is motion detectable along different angle of incidence.
 
@@
 In this context, we will call \emph{fragment} an infinitesimal part of our classical electron.
-The Shannon entropy associated with the distribution, which can be though as the amount of information required to identify a fragment given the distribution, is finite as well.\footnote{To be precise, Shannon's differential entropy is actually the number of bits needed relative to a uniform distribution of unitary range on all axis.
-The intuitive meaning, though, is the same.} $I(\rho_\mathcal{c}) = \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq \wedge dp = I_0$.
-TODO: prove in appendix.
+The Shannon entropy associated with the distribution is finite as well.
+$I(\rho_\mathcal{c}) = - \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq \wedge dp = I_0$.
+Intuitively, this value can be though as the amount of information required to identify a fragment within the distribution.\footnote{To be precise, Shannon's differential entropy is actually the number of bits required by the given distribution relative to a uniform distribution of unitary range on all axis.
+The intuitive meaning, though, is the same.} TODO: prove in appendix.
 
@@
 The final entropy will also remain the same: $I(\rho_{\hat{\mathcal{c}}}) = I_0$.
-Mathematically because $\rho_\mathcal{c} \log (\rho_\mathcal{c})$, like the distribution itself, is just another integrable function.
+Mathematically because $-\rho_\mathcal{c} \log (\rho_\mathcal{c})$, like the distribution itself, is just another integrable function whose integral is conserved by Hamiltonian evolution.
 Physically, because the amount of information required to identify a fragment given the distribution has to remain the same under deterministic and reversible evolution.
-Therefore, in the same way that each fragment is constrained to move along states at equal energy, the composite state is constrained to move along distributions at equal entropy.
+Therefore, in the same way that each fragment is constrained to move through states at equal energy, the composite state is constrained to move through distributions at equal entropy.
 
@@
 &+ \lambda_2(- \int \rho \ln \rho \, dqdp - I_0)\\
-\delta L = &\int \delta \rho [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\ &+ \lambda_1 + \lambda_2 \ln \rho + \lambda_2 ] dqdp = 0 \\
-\rho = &\frac{1}{ \sqrt{4 \pi ^2\sigma_q^2\sigma_p^2}} e^{-\frac{(q-\mu_q)^2(p-\mu_p)^2}{4\sigma_q^2\sigma_p^2}} \\
-= &\frac{1}{e^{I_0 - 1}} e^{-\frac{\pi ^2(q-\mu_q)^2(p-\mu_p)^2}{\exp (2(I_0 - 1))}}
+\delta L = &\int \delta \rho [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\ &+ \lambda_1 - \lambda_2 \ln \rho - \lambda_2 ] dqdp = 0 \\
+\lambda_2 \ln \rho = &\lambda_1 - \lambda_2 + (q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\
+\rho = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(q-\mu_q)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_q^2 (p-\mu_p)^2}{\lambda_2}}\\
+\rho = &\frac{1}{ 2 \pi \sigma_q \sigma_p} e^{-\frac{(q-\mu_q)^2}{2\sigma_q^2}} e^{-\frac{(p-\mu_p)^2}{2\sigma_p^2}} \\
+I_0 = &\ln (2\sigma_q\sigma_p\pi) + 1
 \end{align*}
-We find that the distribution that minimizes the spread is the product of two independent Gaussians.
+(TODO: move the details of the math) We find that the distribution that minimizes the spread is the product of two independent Gaussians.
 Recall that in quantum mechanics, the Gaussian wave packet is the state that minimizes uncertainty: this is the classical analogue.
@@
 
-Suppose, though, the evolution of the distribution is deterministc and reversible while each fragment is undergoing non deterministic motion.
+Suppose, though, that while the evolution of the distribution of our electron electron as a whole is deterministc and reversible, each fragment is undergoing non deterministic motion.
 Suppose $\rho$ is a distribution at equilibrium of such motion.
@@
 
-Suppose now the classical electron undergoes quasi-static evolution.
+Suppose now our classical electron undergoes quasi-static evolution.
 At each instant we can assume $\rho$ to be at equilibrium.
@@
 
-If we assumed that all our classical electrons were interacting with the same background that caused all fragments to exhibit a random walk of the same magnitude, as defined by the information entropy, then the spread of all distributions for all material would satisfy the same inequality.
+If we assumed that all our classical electrons were interacting with the same background, all fragments would exhibit a random walk of the same magnitude as defined by the information entropy.
+The spread of all distributions for all material would satisfy the same inequality.
 This gives us the classical uncertainty principle: $\sigma_q\sigma_p \geq \exp (I_\varnothing - 1) / 2 \pi $ where $I_\varnothing$ is the entropy associated with the background.
@@
 Assuming such a background is not far fetched.
-In fact, it serves as a remainder of the issues mentioned in assumption \ref{ass:infinitesimal_reducibility}: as determinism and reversibility at all scales conceptually does not work, there exists a minimum constant interaction: the system cannot be perfectly isolated.
+We could justify it to address the issues mentioned in assumption \ref{ass:infinitesimal_reducibility}: as determinism and reversibility at all scales conceptually does not work, we posit a minimum constant interaction: the system cannot be perfectly isolated.
 If the entropy of our distribution is comparable to the entropy introduced by the background, we are likely in the range where classical mechanics is not valid as the random motion of each fragment is not negligible.
@@
 It is the link with information theory that gives us further physical insight and is of particular interest given the strong connection between thermodynamics and Shannon's entropy.
-This is something to be developed in a future work.
 

### 2016-11-11 2015CompleteDerivation/completeDerivation.tex
Fixing imprecisions about transformation and tangent spaces

@@
 
-The metric actually fixes two other problems.
-The first is: how do we distinguish temporal from standard d.o.f.?
+The metric $\omega$ also allows us to distinguish temporal from standard d.o.f.
 Given a two-dimensional surface $U \subset T^*\mathcal{M}$, how can we tell if it should be charted by a $(t,E)$ or whether we can define a distribution on it?
@@
 
-The second problem is: how do we prevent time variable changes that disturb determinism and reversibility?
-The invariance of $\omega$ already guarantees that.
-As we saw before, the fact that the vector product is conserved within a degree of freedom is equivalent to saying that changing variables does not change the possibilities within a d.o.f.
-In fact, the infinitesimal displacement $d_s x^i$ for a general infinitesimal transformation that preserves the vector product on a surface is a divergence free field.
-The conservation of the Minkowski product across the temporal and standard d.o.f.
-means that changing variables does not introduce rotation between time and state variables.
-In fact, the infinitesimal displacement $d_s x^i$ for a general infinitesimal transformation that preserves the Minkowski product on a surface is a curl free field.
-
-TODO: add proofs in the appendix
+The invariance of $\omega$ also prevents mixing between standard and temporal d.o.f.
+The Euclidean signature across standard d.o.f.
+allows to move ranges of possibilities from one d.o.f.
+to the other while conserving the total: the generators of rotation are in fact divergence free.
+The Minkowski signature, instead, only allows hyperbolic rotation (i.e.
+Lorentz boosts) across standard and temporal d.o.f., whose generators are curl free.
+Therefore, as required, we cannot transform a distribution over a standard d.o.f.
+into a distribution over a temporal one.
 
@@
 \begin{justification}
-	N.B.
-Though $du^\alpha \in TT\mathcal{M}$, its components change as the components of vectors in $T\mathcal{M}$ under coordinate transformation.
-There is a canonical way to associate between elements of $TT\mathcal{M}$ and $T\mathcal{M}$.
-Therefore we consider $du^\alpha \in T\mathcal{M}$ a vector as it is standard in physics.
-
 	We claim $(\mathcal{M}, g)$ is a pseudo-Riemannian manifold where $g$ is a metric induced by $\omega$ under kinematic equivalence.

### 2016-11-11 2015CompleteDerivation/completeDerivation.tex
Minor improvements on prelude for quantum

@@
 Before introducing assumption \ref{ass:irreducibility} and deriving quantum mechanics as its consequence, we first give a semi-classical account of the conceptual model we will be using.
-This allows to build the intuition more gradually and to show that classical and quantum mechanics are not as different as one may first imagine.
+This allows us to build the intuition more gradually and to show that classical and quantum mechanics are not as different as one may first imagine.
 
 We will see that, as a first approximation, quantum particles are more keenly described by classical composite states (i.e.
-a distribution in phase space) like any other object made of a finite amount of material.
+distributions in phase space) like any other object made of a finite amount of material.
 We'll also see that classical distributions already possess qualitative properties that are typically associated with quantum systems, including a classical analogue of the uncertainty principle.
@@
 \rho = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(q-\mu_q)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_q^2 (p-\mu_p)^2}{\lambda_2}}\\
+\end{align*}
+We solve the multipliers and have:
+\begin{align*}
 \rho = &\frac{1}{ 2 \pi \sigma_q \sigma_p} e^{-\frac{(q-\mu_q)^2}{2\sigma_q^2}} e^{-\frac{(p-\mu_p)^2}{2\sigma_p^2}} \\
@@
 \end{align*}
-(TODO: move the details of the math) We find that the distribution that minimizes the spread is the product of two independent Gaussians.
+We find that the distribution that minimizes the spread is the product of two independent Gaussians.
 Recall that in quantum mechanics, the Gaussian wave packet is the state that minimizes uncertainty: this is the classical analogue.

### 2016-11-14 2015CompleteDerivation/completeDerivation.tex
Prelude to quantum ready for review

@@
 
-Section VIII - Preliminary work
+Section VIII - Mostly done. Needs review.
 
@@
 
-Before introducing assumption \ref{ass:irreducibility} and deriving quantum mechanics as its consequence, we first give a semi-classical account of the conceptual model we will be using.
+Before introducing assumption \ref{ass:irreducibility} and deriving quantum mechanics as its consequence, we first give a semi-classical account of the conceptual picture that emerges.
 This allows us to build the intuition more gradually and to show that classical and quantum mechanics are not as different as one may first imagine.
@@
 It was their being infinitesimal and spread around a small neighborhood that gave us conjugate pairs of state variables and Hamiltonian evolution.
-Any finite amount of classical material, instead, is described by a distribution.
+Any finite amount of classical material, therefore, is described by a distribution.
 
@@
 Therefore, to a first approximation, it must be described by a distribution as well.
-This may, to some, seem preposterous at first: an electron is not a point particle.
+The idea that an electron is not a point particle may, to some, seem preposterous at first.
 But we contend that it is no more preposterous than to think it point-like, described by a probability distribution that self-interacts.
-In fact, the idea of an electron being a distribution of material is already conceptually more compatible with two issues.
-
-The first one is that an electron does not seem to have a well defined position or momentum.
+In fact, the idea of an electron being a distribution of material is already conceptually more compatible with two features of quantum mechanics.
+
+The first one is that quantum particles are not associated with a well defined value for position and momentum.
 If we misidentify the electron with a classical particle, this would indeed be worth of note and puzzlement.
@@
 Your right index finger is not, hopefully, at the same position of your left eye.
-In fact, based on what we saw before, any object of finite mass is a distribution over a region, and discontinuous distributions are ruled out by \ref{continuous_map}.
+In fact, based on what we saw before, any object of finite mass is a distribution over a region of phase space, and discontinuous distributions are ruled out by \ref{continuous_map}.
 An electron is an object of finite mass: why should it be any different?
@@
 
-The other issue is particle-wave duality.
+The other feature is particle-wave duality.
 If the spread of the distribution is narrow enough for the problem at hand, the motion can be described by just the position and momentum of the center of mass, making it behave like a classical particle.
@@
 $I(\rho_\mathcal{c}) = - \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq \wedge dp = I_0$.
-Intuitively, this value can be though as the amount of information required to identify a fragment within the distribution.\footnote{To be precise, Shannon's differential entropy is actually the number of bits required by the given distribution relative to a uniform distribution of unitary range on all axis.
-The intuitive meaning, though, is the same.} TODO: prove in appendix.
+Intuitively, this value can be though as the amount of information required to identify a fragment within the distribution.\footnote{To be precise, Shannon's differential entropy is actually the number of bits required to identify an element of the given distribution relative to one of a uniform distribution of unitary range on all axis.}
 
@@
 The final entropy will also remain the same: $I(\rho_{\hat{\mathcal{c}}}) = I_0$.
-Mathematically because $-\rho_\mathcal{c} \log (\rho_\mathcal{c})$, like the distribution itself, is just another integrable function whose integral is conserved by Hamiltonian evolution.
+Mathematically because $-\rho_\mathcal{c} \log (\rho_\mathcal{c})$, like the distribution itself, is just another integrable function and therefore its integral is conserved by Hamiltonian evolution.
 Physically, because the amount of information required to identify a fragment given the distribution has to remain the same under deterministic and reversible evolution.
@@
 
-If we assumed that all our classical electrons were interacting with the same background, all fragments would exhibit a random walk of the same magnitude as defined by the information entropy.
-The spread of all distributions for all material would satisfy the same inequality.
-This gives us the classical uncertainty principle: $\sigma_q\sigma_p \geq \exp (I_\varnothing - 1) / 2 \pi $ where $I_\varnothing$ is the entropy associated with the background.
-
-Assuming such a background is not far fetched.
-We could justify it to address the issues mentioned in assumption \ref{ass:infinitesimal_reducibility}: as determinism and reversibility at all scales conceptually does not work, we posit a minimum constant interaction: the system cannot be perfectly isolated.
-If the entropy of our distribution is comparable to the entropy introduced by the background, we are likely in the range where classical mechanics is not valid as the random motion of each fragment is not negligible.
+If we assumed that all our classical electrons were affected by the same source of randomness, all fragments would exhibit a random walk of the same magnitude as defined by the information entropy.
+The spread of all distributions for the material would satisfy the same inequality.
+This gives us the classical uncertainty principle: $\sigma_q\sigma_p \geq \exp (I_\varnothing - 1) / 2 \pi $ where $I_\varnothing$ is the entropy associated with the source.
+
+Such an assumption is not far fetched.
+The source of entropy could be identified in the environment (i.e.
+a minimum interaction that cannot be avoided as perfect isolation cannot be achieved) or the internal self interaction of the fragments (i.e.
+the evolution of the unstated part).
+Whatever it may be, if it characterized by an entropy comparable to the one of the distribution, that is we are trying to localize the electron better than the source can allow, the classical assumption breaks down as the random motion of each fragment is not negligible.
 
@@
 is outside of the scope of this work.
-Note, though, that this result is similar in spirit to the Gromov non-squeezing theorem and the symplectic camel, which are already generalized to multiple d.o.f..
+Note, though, that this result is similar to the Gromov non-squeezing theorem and the symplectic camel, which are already generalized to multiple d.o.f..
 It is the link with information theory that gives us further physical insight and is of particular interest given the strong connection between thermodynamics and Shannon's entropy.
 
-At any rate, this discussion shows, once again, how in classical mechanics we can already find some of the ingredients that are typically assumed to be characteristics of quantum mechanics.
+This discussion shows, once again, how we can already find in classical mechanics some of the ingredients that are typically assumed to be characteristics of quantum mechanics.
 It has its own flavor of anti-particles, CPT theorem and uncertainty principle.

### 2016-11-17 2015CompleteDerivation/completeDerivation.tex
Reviewed first to subsections for quantum systems

@@
 In general, this is quite a complicated thing to do, which requires intimate knowledge of the system at hand.
-So we simplify our problem and study a system made of infinitesimal homogeneous parts.
-What we'll find is that under the additional assumption that the system is infinitesimally reducible (i.e.
+So we simplify our problem and study a material made of infinitesimal homogeneous parts.
+What we'll find is that under the additional assumption that the material is infinitesimally reducible (i.e.
 its state is equivalent to the states of its parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian mechanics.
@@
 
-In this section we explore a different relationship between the whole material and its parts from the one described in section \ref{sec:reducibility}.
-The system is still infinitesimally decomposable, made of infinitesimal parts as before, but this time it is irreducible.
-That is: the state and the evolution of the whole does not tell us anything about the state and the evolution of the parts.
-As the internal dynamics is undetermined, we'll find that each irreducible element is suitably described by the standard framework of quantum mechanics.
+In this section we explore a different relationship between the whole and its parts.
+We will consider an irreducible material, one where the state and the evolution of the whole does not tell us anything about the state and the evolution of the parts.
+The material is still conceptually infinitesimally decomposable, made of infinitesimal parts which we call fragments.
+But each fragment cannot be studied independently and therefore cannot be assigned a state: the internal dynamics is undetermined.
+We'll find that the evolution of an irreducible material is suitably described by the standard framework of quantum mechanics.
 
@@
 The first possible source of confusion is between divisibility, the idea that an object can be divided into parts, and reducibility, the idea that the state of the whole is equivalent to the state of the parts.
+More precisely, a system is divisible if there is a time evolution $\mathcal{T}_{\Delta t}: \mathcal{S} \rightarrow \mathcal{S}_1 \times \mathcal{S}_2$ that takes the state of a system and returns the state of two separate isolated systems.
+It is reducible if $\mathcal{S} = \mathcal{S}_1 \times \mathcal{S}_2$, the state of the system is the state of the two subsystems.
 Most of the time these two concepts overlap, but they are separate as we can see in the following examples.
@@
 But the state of a muon is not the state of an electron and two neutrinos.
-The muon is divisible into the three particles but it is not reducible into them.
+The muon is divisible into the three particles but it is not reducible to them.
 Consider a proton.
-Its state is given by the state of its three valence quarks.
-Yet, we cannot divide the proton into the three quarks.
-The proton is reducible to quarks, but divisible into quarks.
+Its state is given by the state of its three valence quarks and the gluon field.
+Yet, we cannot divide the proton into the three isolated quarks.
+The proton is reducible to quarks, but not divisible into quarks.
 
@@
 Even more subtle is the difference between divisibility and decomposability, the idea that the object has parts.
-While we are agnostic on whether a quantum particle can be divided, we are implicitly assuming that it can be decomposed.
-Irreducibility describes the relationship between the state of the whole and the state of the parts, therefore parts are implicit in the definition.
-While these parts must exist, we cannot isolate them and act on them independently: if we were able to do that we could assign them a state.
-
-While we must be able to talk about parts of a quantum particles, not all questions will make sense.
+That is, a system is decomposable if there is a law $+ : \mathcal{S} \times \mathcal{S} \rightarrow \mathcal{S}$ to combine two states to get a third.
+While irreducibility says nothing on whether a quantum particle can be divided, it implicitly assumes that it can be decomposed.
+Irreducibility describes the relationship between the state of the whole and the state of the parts, therefore the existence of parts are implicit in the definition.
+But it also tells us that we are limited in what description we can give to these parts: we cannot isolate them as we want and act on them independently because if we were able to do that we could assign them an independent state.
+
+Therefore, while we must be able to talk about parts of a quantum particles, not all questions will make sense.
 Can we ask what fraction of a quantum particle is in a particular region of space?
-Yes, since the particle is decomposed some part may be inside the volume.
+Yes, since the particle can be spatially decomposed some amount of material may be inside the volume.
 Can we ask where that same fraction will be at a later time?
@@
 
-The proper way to think of a quantized material is therefore the following: a continuum of infinitesimal parts clumped together to form quantum particles, the smallest amount of material to which we can assign a state.
+Note that if the material is infinitesimally reducible, reducibility/decomposability/divisibility are all possible.
+This is why one may not dwell on their difference and use the word ``part" interchangeably in the three cases.
+Now that we assume irreducibility, it is better to use more precise language to avoid confusion.
+Therefore we will call \emph{fragments} the infinitesimal parts of the material that cannot be described independently.
+We will call \emph{components} the parts that the material can be decomposed into.
+A component will have a state associated with it, and therefore will always be a collection of fragments.
+As in the classical case, a component may be the difference between two states, therefore one should not imagine it always ``contained" in the composed system.
+
+The proper way to think of a quantized material is therefore the following: a continuum of infinitesimal fragments clumped together to form quantum particles, the smallest amount of material to which we can assign a state.
 With this picture in mind, we can see, even before developing the math, how some features typical of quantum behaviors emerge.
 
-The first feature is the importance of statistics in the quantum particle state.
-Consider a part of a quantum particle in an initial state.
-As we cannot track its motion, it can be any part of the particle in its final state.
-That is, the motion of a part of a particle is a random walk within the evolution of a particle.
+The first feature is the importance of statistics.
+Consider a fragment of a quantum particle in an initial state.
+As we cannot track its motion, it can correspond to any fragment of the particle in its final state.
+That is, the motion of a fragment of a particle is a random walk within the evolution of a particle.
 
 The second feature is non locality.
-Suppose we have a quantum particle state distributed in space.
+Suppose we have a quantum particle distributed in space.
 Suppose we exert a force only on a subregion of its distribution.
-As the particle is irreducible, we cannot interact with only part of it: we affect also the part where no force is exerted.
+As the particle is irreducible, we cannot interact with only part of it: we must also affect the part where no force is exerted.
 We cannot simply interact with the part at a specific region, as this would allow to assign a state to that part.
 
-Wave particle duality should be clear at this point.
+Wave particle duality is also accounted for.
 The quantum particle behaves like a wave in the sense that its parts form a distribution in space.
-Yet, it behaves as a single unit, as the state is one unit and cannot be separated.
-
-The most striking feature, though, is that we can predict the functional form for particle decay.
+Yet, it behaves as a single unit, as the state is irreducible and cannot be split.
+
+The most striking feature, though, is that we can predict the functional form for particle decay over time.
 Suppose a quantum particle spontaneously decays into two or more particles.
-Such process cannot be deterministic as the d.o.f.
-increase, therefore is not determined by the state of the quantum particle.
+Such process cannot be deterministic as the number of d.o.f.
+increases.
 As the decay is spontaneous, it is not triggered by the environment (i.e.
 external state).
-It is therefore triggered by the internal unstated dynamic of the parts.
+It must be triggered by the internal dynamic of the fragments.
 As we assumed the internal dynamic is unknown, we cannot learn anything about it by looking at the product of the decay.
 In particular, all states have to have equal chance of decay or we could tell that the internal dynamic is more stable for some states.
-The probability of decay is constant in time.
+The probability of decay is constant for all states at all times.
 The probability of survival follows the exponential distribution.
@@
 
-\subsection{Quantum particle and complex vector spaces}
-
-Turning these concept into a precise mathematical formulation is tricky as we need to formally characterize the unstated part which is not physically distinguishable.
-While many previous results do not apply, the classical derivation still can serve as a template.
-
-As in the classical case, the quantum particle is thought to be decomposable: different parts will be in different places, the evolution of the whole is the composition of the evolution of parts.
+\subsection{Irreducible material and complex vector spaces}
+
+It is time to turn these concepts into a precise mathematical formulation following the template used for classical mechanics.
+We want to formalize the state space for an \emph{irreducible material} as a quantum particle will be a finite amount of such material.
+
+As for the classical material, the irreducible material is thought to be decomposable: each state can seen as the sum of different components (e.g.
+spatially distributed part) such that the evolution of the whole is the composition of their evolution.
 Therefore \ref{prop:abelian_group} applies and the state space $\Psi$ of an irreducible material is an abelian group.
 
-Note, though, that we can't think of the parts as independent components.
-As the unstated part is statistical in nature, it will combine as statistical variables do.
-Suppose $X$ and $Y$ are two random variables, $\sigma^2_{X+Y} = \sigma^2_{X} + \sigma^2_{Y} + 2 \, \sigma_{X} \sigma_{Y} \rho_{X,Y}$, where $\sigma^2$ is the variance, $\sigma$ is the standard deviation, and $\rho_{X,Y}=COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient.
-As $-1\leq\rho_{X,Y}\leq+1$, the total variance can be greater or smaller then the sum of the independent parts, and the strength of the uncertainty within the unstated part is something physically distinguishable.
-
-Therefore, while the motion of the unstated part is not itself physically relevant, the correlation between the motion of the unstated part of two combined parts is very relevant.
-We'll see how interference can be understood in this context: it is constructive if the randomness is correlated while it is destructive if it is anti-correlated.
-
-So, how can we characterized these relationship between different configurations of the unstated parts?
+Note, though, that the components are in general not independent.
+As the motion of each fragment is statistical in nature, it will combine as statistical variables do.
+Suppose $X$ and $Y$ are two random variables,
+\begin{equation}\label{eqn:correlation}
+\sigma^2_{X+Y} = \sigma^2_{X} + \sigma^2_{Y} + 2 \, \sigma_{X} \sigma_{Y} \rho_{X,Y}
+\end{equation}
+where $\sigma^2$ is the variance, $\sigma$ is the standard deviation, and $\rho_{X,Y}=COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient.
+As $-1\leq\rho_{X,Y}\leq+1$, the total variance can be greater or smaller then the sum of of the variance of the components.
+The strength of the chaotic motion of the fragments is something physically distinguishable as it contributes to the spread of the distribution.
+Therefore while the motion of each fragment is not itself physically distinguishable, it falls within the unstated part, the strength and correlation of the chaotic motion must be, they fall within the state.
+
+How can we characterize the aspects of the unstated part that are captured by the state?
 We can look at the classical derivation for guidance.
 In that case, we looked at the set of transformation that increase/decrease the amount of material.
-Those are the only type of transformation that we can apply on a classical system that do not change the overall state given by the distribution.
-
-As the classical material is homogeneous, permutations between classical particles in the same state $\mathcal{s}$ do not matter: the ordering is unstated.
-In a way, increasing/decreasing the amount of material increases/decreases the unstated part of the system.
-
-For the irreducible material we follow the same strategy: we find the state transformations that affect the unstated part while keeping the overall state, the distribution, unchanged.
-
-Suppose we have $\psi_1$ a state for an irreducible material characterized by one d.o.f.
+We can think them as the transformations that change the unstated part of the system.
+As the classical material is homogeneous, permutations between classical particles in the same state $\mathcal{s}$ do not matter (i.e.
+the ordering is unstated).
+Increasing/decreasing the amount of material increases/decreases the number of permutations and therefore changes the unstated part of the system.
+For the irreducible material we follow the same strategy: we look for the transformations that affect the unstated part while keeping the overall state, the distribution, unchanged.
+
+Suppose we have $\psi_1$ a state for an amount $a_m$ of irreducible material characterized by one d.o.f.
 The material will be spread over the configurations of the infinitesimal fragments.
-As the material is homogeneous, the material is spread evenly among the such possible configuration.
-Counting configurations is therefore equivalent to counting amount of material.
+As the material is homogeneous, the material is spread evenly among such possible configurations.
+Counting fragment configurations is therefore equivalent to counting amount of material.
 Each fragment would be identified by a pair of independent random variables $(A, B)$, as a d.o.f.
 is two-dimensional.
-The shape of the distribution and the actual values are not physically distinguishable: these are just dummy variables we assign to count the fragment configurations.
-We choose A and B to be uniform distributions from 0 to 1, so that the area is a square of unit size, which represent the unit of unstated  configurations the material is spread over.
-
-Suppose we have another state $\psi_2$, where only unstated part is changed.
-The unstated part can also be characterized by a square formed by two independent uniform distributions (C, D).
-Counting states on (C,D) is equivalent to counting them on (A,B), as they are simply different arrangements.
-As all distributions are uniform, they can only be connected by a linear operation.
+The shape of the distribution and the actual values are not physically distinguishable: these are just dummy variables we assign to count the fragments.
+We can choose A and B to be uniform distributions of interval $\sqrt{a_m}$, so that they identify a square of area $a_m$, the amount of material.
+
+Suppose we have another configuration $\psi_2$, where only the unstated part is changed.
+The unstated part of $\psi_2$ can also be characterized by a square formed by two independent uniform distributions (C, D).
+Counting states on (C,D) is equivalent to counting them on (A,B), as they are simply different arrangements of fragments within the same state.
+The relationship between the two pairs of variables must be linear since all distributions are uniform.
 That is: $C=aA+bB$ and $D= -bA + aB$.
 We can write $\psi_2 = \tau(a,b) \psi_1$ as $a$ and $b$ are the only parameters of the transformation.
-
-The transformation $\tau(a,b)$ increases/decreases the area by the factor $a^2+b^2$ and rotates the square by $arctan(x/y)$.
-If we set $c=a+\imath b$ those same quantities can be expressed as $\rho^2=|c|^2$ and $\theta = arg(z)$, the modulus and the phase of a complex number.
-In other words: complex numbers provide a representation for the group of rotations and magnitude rescaling.
-
-For multiple independent d.o.f., we have one square each.
-As the material is irreducible, all areas need to be always the same since relative changes would allow to learn something about the motion of the infinitesimal fragments.
-This means that the transformation $\tau(c)$ has to be the same for all d.o.f.
-In other words, the group of transformations is the same regardless of the number of independent d.o.f.
+Or more compactly, $\psi_2 = \tau(c) \psi_1$ where $c=a+\imath b$.
+
+The transformation $\tau(c)$ increases/decreases the area by the modulus $|c|^2=a^2+b^2$, which physically corresponds to the increase/decrease of amount of material.
+It also rotates the square by the phase $\theta_c \equiv arg(c) = arctan(a/b)$, which corresponds in a change in correlation: if $\theta_c = 0$, $C = |c| A$ are clearly correlated while if $\theta_c = \pi$, $C = - |c| A$ are clearly anti-correlated.
+
+To better understand the link between phase and correlation, suppose $\psi$ is the state for an irreducible material.
+We can write $\tau(c) \psi = \tau(c_1 + c_2) \psi = \tau(c_1) \psi + \tau(c_2) \psi$ as the composition of two separate transformations.
+We have
+\begin{equation}\label{eqn:correlation_state}
+|c|^2=|c_1|^2 + |c_2|^2 + 2 |c_1||c_2|\cos(\theta_{c_2} - \theta_{c_1})
+\end{equation}
+Compare to \ref{eqn:correlation}.
+The material is spread evenly among fragment configurations.
+Also, since it is homogeneous, the ratio of fragment per configuration is the same.
+Therefore the amount of material is proportional to the strength of the chaotic motion of the fragment, which is measured by the variance.
+The increase/decrease in amount of material therefore combines like the variance of a statistical variable.
+The cosine of the phase difference takes the place of the Pearson correlation coefficient.
+A more meaningful name for the phase, then, would be \emph{correlation angle}.
+Also note how correlation for a single variable is a meaningless quantity: it is the correlation between two random variables that is significant.
+That is why in quantum mechanics phase differences have physical meaning, as it gives the correlation between two states, while absolute phases do not.
+
+If we increase the number of independent d.o.f.
+the group of transformations remains the same.
+As the material is homogeneous, the same amount of material needs to be spread over the same number of fragment configurations along each independent d.o.f.
+This means that, as we combine components, the strength of the chaotic motion has to be the same in all directions.
+This means that the increase/decrease of material and the change in correlation has to be the same across all d.o.f.
 
 Mathematically, the state space $\Psi$ for an irreducible material is a complex vector space.
+It is so because it is the state space for a decomposable material (i.e.
+an abelian group) where we can increase/decrease the amount and change the correlation of the internal stochastic motion (i.e.
+a set of transformation conveniently parametrized by a complex number).
 
@@
 
+\subsection{Complex valued distributions}
+
 If $\mathcal{Q}$ is the space that defines our units, a state $\psi$ must tell us for each point $\mathcal{q} \in \mathcal{Q}$ how much material is there and the the correlation of the unstated part.

### 2016-11-18 2015CompleteDerivation/completeDerivation.tex
Adding Christine's comments

@@
 
-We will see that, as a first approximation, quantum particles are more keenly described by classical composite states (i.e.
+We will see that, as a first approximation, quantum particles are more suitably described by classical composite states (i.e.
 distributions in phase space) like any other object made of a finite amount of material.
@@
 The first one is that quantum particles are not associated with a well defined value for position and momentum.
-If we misidentify the electron with a classical particle, this would indeed be worth of note and puzzlement.
-But if the electron is a classical distribution, there is no issue.
-The earth is both under Milano and New York.
-Your right index finger is not, hopefully, at the same position of your left eye.
+If we misidentify the electron with a classical particle (i.e.
+the infinitesimal part of a classical material), this would indeed be worthy of note and puzzlement.
+But if the electron is a distribution, there is no issue as a phase space distribution does not have a single value for position or momentum.
+The earth is, so to speak, in both North America and Europe at the same time.
 In fact, based on what we saw before, any object of finite mass is a distribution over a region of phase space, and discontinuous distributions are ruled out by \ref{continuous_map}.
 An electron is an object of finite mass: why should it be any different?
-In fact, if one claimed an electron to be point-like, he would be claiming an electron behaves in a different way than all other physical objects of finite mass/energy (which are the only ones we experimentally work with).
+In fact, if one claimed an electron to be point-like, he would be claiming an electron is fundamentally different than all other physical objects of finite mass/energy (which are the only ones we experimentally work with).
 Shouldn't \emph{he} have the burden of proof?
@@
 While this picture is quite reasonable, some people may reject it because they believe that a fundamental object must have no size, it must be a point.
-But that does not follow: lack of spatial extend and lack of substructure are independent concepts.
+But that does not follow: lack of spatial extent and lack of substructure are independent concepts.
 
@@
 If we shrink the sphere, the dynamics and its description stays the same.
-We can collapse the whole sphere to a point: it is now a point-like object that is not elementary, as there is motion detectable along different angle of incidence.
-
-Not only point-like objects don't have to be elementary, but it's unclear how can they be elementary.
+We can collapse the whole sphere to a point: it is now a point-like object that is not elementary, as there is motion detectable along different angles of incidence.
+
+Not only do point-like objects not have to be elementary, but it's unclear how they can be elementary.
 How can the whole universe be reduced to  elements that have no idea how many spatial dimensions are there?
@@
 If we abandon the idea of point-like objects of finite mass, we avoid these problems and get a picture more consistent across the different realms of physics.
-And we can also see that classical mechanics already has some features that are thought to be inherent to quantum mechanics.
+And we can also see that classical mechanics already has some features that are commonly ascribed to quantum mechanics.
 
@@
 $I(\rho_\mathcal{c}) = - \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq \wedge dp = I_0$.
-Intuitively, this value can be though as the amount of information required to identify a fragment within the distribution.\footnote{To be precise, Shannon's differential entropy is actually the number of bits required to identify an element of the given distribution relative to one of a uniform distribution of unitary range on all axis.}
+Intuitively, this value can be thought of as the amount of information required to identify a fragment within the distribution.\footnote{To be precise, Shannon's differential entropy is actually the number of bits required to identify an element of the given distribution relative to one of a uniform distribution of unitary range on all axes.}
 
@@
 The final entropy will also remain the same: $I(\rho_{\hat{\mathcal{c}}}) = I_0$.
-Mathematically because $-\rho_\mathcal{c} \log (\rho_\mathcal{c})$, like the distribution itself, is just another integrable function and therefore its integral is conserved by Hamiltonian evolution.
+Mathematically, because $-\rho_\mathcal{c} \log (\rho_\mathcal{c})$, like the distribution itself, is just another integrable function and therefore its integral is conserved by Hamiltonian evolution.
 Physically, because the amount of information required to identify a fragment given the distribution has to remain the same under deterministic and reversible evolution.
@@
 We now ask the following question: once we fix the entropy, what is the relationship between the spread of the marginal distributions along $q$ and $p$?
-Using Lagrange multipliers, we fix the total amount of material to $1$ and the total entropy $I_0$ and find the distribution that minimizes the the product of the variance $\sigma_q^2 \sigma_p^2 = \int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp$.
+Using Lagrange multipliers, we fix the total amount of material to $1$ and the total entropy $I_0$ and find the distribution that minimizes the product of the variance $\sigma_q^2 \sigma_p^2 = \int (q-\mu_q)^2 \rho_{\mathcal{c}} \, dqdp \int (p-\mu_p)^2 \rho_{\mathcal{c}} \, dqdp$.
 \begin{align*}
-L = &\int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp \\
-&+ \lambda_1(\int \rho dqdp - 1)\\
-&+ \lambda_2(- \int \rho \ln \rho \, dqdp - I_0)\\
-\delta L = &\int \delta \rho [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\ &+ \lambda_1 - \lambda_2 \ln \rho - \lambda_2 ] dqdp = 0 \\
-\lambda_2 \ln \rho = &\lambda_1 - \lambda_2 + (q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\
-\rho = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(q-\mu_q)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_q^2 (p-\mu_p)^2}{\lambda_2}}\\
+L = &\int (q-\mu_q)^2 \rho_{\mathcal{c}} \, dqdp \int (p-\mu_p)^2 \rho_{\mathcal{c}} \, dqdp \\
+&+ \lambda_1(\int \rho_{\mathcal{c}} dqdp - 1)\\
+&+ \lambda_2(- \int \rho_{\mathcal{c}} \ln \rho_{\mathcal{c}} \, dqdp - I_0)\\
+\delta L = &\int \delta \rho_{\mathcal{c}} [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\ &+ \lambda_1 - \lambda_2 \ln \rho_{\mathcal{c}} - \lambda_2 ] dqdp = 0 \\
+\lambda_2 \ln \rho_{\mathcal{c}} = &\lambda_1 - \lambda_2 + (q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\
+\rho_{\mathcal{c}} = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(q-\mu_q)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_q^2 (p-\mu_p)^2}{\lambda_2}}\\
 \end{align*}
@@
 \begin{align*}
-\rho = &\frac{1}{ 2 \pi \sigma_q \sigma_p} e^{-\frac{(q-\mu_q)^2}{2\sigma_q^2}} e^{-\frac{(p-\mu_p)^2}{2\sigma_p^2}} \\
-I_0 = &\ln (2\sigma_q\sigma_p\pi) + 1
+\rho_{\mathcal{c}} = &\frac{1}{ 2 \pi \sigma_q \sigma_p} e^{-\frac{(q-\mu_q)^2}{2\sigma_q^2}} e^{-\frac{(p-\mu_p)^2}{2\sigma_p^2}} \\
+I_0 = &\ln (2\pi\sigma_q\sigma_p) + 1
 \end{align*}
@@
 
-Suppose, though, that while the evolution of the distribution of our electron electron as a whole is deterministc and reversible, each fragment is undergoing non deterministic motion.
+Suppose, though, that while the evolution of the distribution of our electron as a whole is deterministc and reversible, each fragment is undergoing non deterministic motion.
 Suppose $\rho$ is a distribution at equilibrium of such motion.
@@
 
-If we assumed that all our classical electrons were affected by the same source of randomness, all fragments would exhibit a random walk of the same magnitude as defined by the information entropy.
+If we assumed that any classical electron were affected by the same source of randomness, all fragments would exhibit a random walk of the same magnitude as defined by the information entropy.
 The spread of all distributions for the material would satisfy the same inequality.
@@
 the evolution of the unstated part).
-Whatever it may be, if it characterized by an entropy comparable to the one of the distribution, that is we are trying to localize the electron better than the source can allow, the classical assumption breaks down as the random motion of each fragment is not negligible.
+Whatever it may be, if it is characterized by an entropy comparable to the one of the distribution, that is we are trying to localize the electron better than the source can allow, the classical assumption breaks down as the random motion of each fragment is not negligible.
 
@@
 is outside of the scope of this work.
-Note, though, that this result is similar to the Gromov non-squeezing theorem and the symplectic camel, which are already generalized to multiple d.o.f..
+Note, though, that this result is related to the Gromov non-squeezing theorem and the symplectic camel, which are already generalized to multiple d.o.f..
 It is the link with information theory that gives us further physical insight and is of particular interest given the strong connection between thermodynamics and Shannon's entropy.
@@
 
-Consider the a planarian flatworm.
+Consider a planarian flatworm.
 If we cut it into multiple pieces, each part will grow into a new worm.
@@
 While irreducibility says nothing on whether a quantum particle can be divided, it implicitly assumes that it can be decomposed.
-Irreducibility describes the relationship between the state of the whole and the state of the parts, therefore the existence of parts are implicit in the definition.
+Irreducibility describes the relationship between the state of the whole and the state of the parts, therefore the existence of parts is implicit in the definition.
 But it also tells us that we are limited in what description we can give to these parts: we cannot isolate them as we want and act on them independently because if we were able to do that we could assign them an independent state.
 
-Therefore, while we must be able to talk about parts of a quantum particles, not all questions will make sense.
+Therefore, while we must be able to talk about parts of a quantum particle, not all questions will make sense.
 Can we ask what fraction of a quantum particle is in a particular region of space?
@@
 
-Note that if the material is infinitesimally reducible, reducibility/decomposability/divisibility are all possible.
+Note that if the material is infinitesimally reducible (i.e.
+the classical case), reducibility/decomposability/divisibility are all possible.
 This is why one may not dwell on their difference and use the word ``part" interchangeably in the three cases.
@@
 As the particle is irreducible, we cannot interact with only part of it: we must also affect the part where no force is exerted.
-We cannot simply interact with the part at a specific region, as this would allow to assign a state to that part.
-
-Wave particle duality is also accounted for.
+We cannot simply interact with the part at a specific region, as this would allow us to assign a state to that part.
+
+Wave-particle duality is also accounted for.
 The quantum particle behaves like a wave in the sense that its parts form a distribution in space.
@@
 external state).
-It must be triggered by the internal dynamic of the fragments.
-As we assumed the internal dynamic is unknown, we cannot learn anything about it by looking at the product of the decay.
-In particular, all states have to have equal chance of decay or we could tell that the internal dynamic is more stable for some states.
+It must be triggered by the internal dynamics of the fragments.
+As we assumed the internal dynamics is unknown, we cannot learn anything about it by looking at the product of the decay.
+In particular, all states have to have equal chance of decay or we could tell that the internal dynamics is more stable for some states.
 The probability of decay is constant for all states at all times.
@@
 
-As for the classical material, the irreducible material is thought to be decomposable: each state can seen as the sum of different components (e.g.
-spatially distributed part) such that the evolution of the whole is the composition of their evolution.
+As for the classical material, the irreducible material is thought to be decomposable: each state can be seen as the sum of different components (e.g.
+spatially distributed parts) such that the evolution of the whole is the composition of their evolution.
 Therefore \ref{prop:abelian_group} applies and the state space $\Psi$ of an irreducible material is an abelian group.
@@
 where $\sigma^2$ is the variance, $\sigma$ is the standard deviation, and $\rho_{X,Y}=COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient.
-As $-1\leq\rho_{X,Y}\leq+1$, the total variance can be greater or smaller then the sum of of the variance of the components.
+As $-1\leq\rho_{X,Y}\leq+1$, the total variance can be greater or smaller than the sum of the variance of the components.
 The strength of the chaotic motion of the fragments is something physically distinguishable as it contributes to the spread of the distribution.
@@
 We can look at the classical derivation for guidance.
-In that case, we looked at the set of transformation that increase/decrease the amount of material.
-We can think them as the transformations that change the unstated part of the system.
+In that case, we looked at the set of transformations that increase/decrease the amount of material.
+We can think of them as the transformations that change the unstated part of the system.
 As the classical material is homogeneous, permutations between classical particles in the same state $\mathcal{s}$ do not matter (i.e.
@@
 The material will be spread over the configurations of the infinitesimal fragments.
-As the material is homogeneous, the material is spread evenly among such possible configurations.
-Counting fragment configurations is therefore equivalent to counting amount of material.
+As the material is homogeneous, it is spread evenly among such possible configurations.
+Counting fragment configurations is therefore equivalent to counting the amount of material.
 Each fragment would be identified by a pair of independent random variables $(A, B)$, as a d.o.f.
@@
 The transformation $\tau(c)$ increases/decreases the area by the modulus $|c|^2=a^2+b^2$, which physically corresponds to the increase/decrease of amount of material.
-It also rotates the square by the phase $\theta_c \equiv arg(c) = arctan(a/b)$, which corresponds in a change in correlation: if $\theta_c = 0$, $C = |c| A$ are clearly correlated while if $\theta_c = \pi$, $C = - |c| A$ are clearly anti-correlated.
+It also rotates the square by the phase $\theta_c \equiv arg(c) = arctan(a/b)$, which corresponds to a change in correlation: if $\theta_c = 0$, $C = |c| A$ are clearly correlated while if $\theta_c = \pi$, $C = - |c| A$ are clearly anti-correlated.
 
@@
 The material is spread evenly among fragment configurations.
-Also, since it is homogeneous, the ratio of fragment per configuration is the same.
+Also, since it is homogeneous, the ratio of fragments per configuration is the same.
 Therefore the amount of material is proportional to the strength of the chaotic motion of the fragment, which is measured by the variance.
@@
 an abelian group) where we can increase/decrease the amount and change the correlation of the internal stochastic motion (i.e.
-a set of transformation conveniently parametrized by a complex number).
+a set of transformations conveniently parametrized by a complex number).
 

### 2016-11-22 2015CompleteDerivation/completeDerivation.tex
More work on the quantum part

@@
 We can extend $\mathcal{H}$ over $T^* \mathcal{M}$ such that $\mathcal{H} \neq 0$ over $T^*\mathcal{M} \,\backslash\, \bar{\mathcal{S}}$.
-$\bar{\mathcal{S}}$ is the level set for $\mathcal{H} = 0$
+$\bar{\mathcal{S}}$ is the level set for $\mathcal{H} = 0$.
 \end{justification}
@@
 Consider an amount $a_m \in \mathbbm{R}$ of irreducible material characterized by one d.o.f.
-The configuration of a fragment is characterized by two random variables $A, B: \Omega \rightarrow \mathbbm{R}$ where $\Omega$ is the sample space for the configuration of the fragments.
+The configuration of a fragment of the material is characterized by two random variables $A, B: \Omega \rightarrow \mathbbm{R}$ where $\Omega$ is the sample space for the configuration of the fragments.
 As the configuration of each fragment is physically indistinguishable, $A$ and $B$ can be chosen such that the distribution $\rho$ of material over such variables is uniform and unitary.
 We can also rescale $A$ and $B$ such that they are defined on the range $[0, \sqrt{a_m}]$.
-We have $a_m=\int \rho dA \wedge dB = \int dA \wedge dB$.
+We have $a_m=\int_{\Omega} \rho dA \wedge dB = \int_{\Omega} dA \wedge dB$.
 Consider a second amount $\hat{a}_m$ of irreducible material in the same configuration.
 It also can be characterized by a uniform distribution over two variables $C, D: \Omega \rightarrow \mathbbm{R}$ over a range of $\sqrt{\hat{a}_m}$.
-As the two amounts of material are in the same configuration we have $(C,D)$ is a function of $(A,B)$.
+As the two amounts of material are in the same configuration, $(C,D)$ is a function of $(A,B)$.
 We have $\hat{a}_m = \int dC \wedge dD = \int |J| dA \wedge dB$ where $|J|$ is the Jacobian determinant.
@@
 Let $\tau: \mathbbm{C} \rightarrow T$ represent such a transformation parametrized by $c=a+\imath b$.
-$|c|^2 = (a^2 + b^2) = \hat{a}_m / a_m$ is physically distinguishable.
+$|c|^2 = (a^2 + b^2) = \hat{a}_m / a_m$ is the ratio of the two amounts of materials and is physically distinguishable.
 Define on $T$ an addition $+: T \times T \rightarrow T$ and a multiplication $*: T \times T \rightarrow T$ such that $\tau(a) + \tau(b) = \tau(a+b)$ and $\tau(a) * \tau(b) = \tau(a*b)$, $a,b \in \mathbbm{C}$, so that the sum and product of the transformation is equal to the sum and product of their respective factors.
@@
 $\tau$ is an isomorphism between $T$ and $\mathbbm{C}$ as fields.
-Generalizing to multiple independent d.o.f.
-the set of transformation remains the same.
+Assume now that the material is characterized by multiple independent d.o.f.
 The transformation on each d.o.f.
@@
 That has to always hold under addition as well, therefore $arg(c_i) = arg(c_j)$.
-$T\cong \mathbbm{C}$ even in multiple d.o.f.
+$T\cong \mathbbm{C}$ even in the case of multiple d.o.f.
 
 	We claim $\Psi$ is a vector space over $\mathbbm{C}$.
-The abelian group $\Psi$ can be extended with the operations defined by $T$ as done in \ref{classical_vector space} for the infinitesimally reducible case.
+The abelian group $\Psi$ can be extended with the operations defined by $T$ as was done in \ref{classical_vector space} for the infinitesimally reducible case.
 
@@
 
-If $\mathcal{Q}$ is the space that defines our units, a state $\psi$ must tell us for each point $\mathcal{q} \in \mathcal{Q}$ how much material is there and the the correlation of the unstated part.
+In the case of a reducible material, we saw how a finite amount of it was described by a distribution over the state variables of infinitesimal parts.
+For an irreducible material, we cannot assign states to the fragments, therefore we cannot properly talk about state variables.
+Therefore we call \emph{fragment variables} the quantities that we use to label the fragment configurations, and it will be over these variables that we will define the distribution of irreducible material.
+
+Naturally, we cannot give a complete description of the configuration of the fragments, therefore we cannot give a joint distribution for all fragment variables at the same time.
+That is: irreducibility contains the notion of complementarity.
+We can, though, find a maximal set of fragment variables: one that provides the best description possible.
+While it should be clear that there cannot be a single unique set, we can show that one such set is provided by the fragment variables $q^i : \mathcal{Q} \rightarrow \mathbbm{R}$ that define the units required to describe the system.
+
+In fact, we cannot add a new d.o.f.
+(as this would mean changing $\mathcal{Q}$) and we cannot add a new fragment variable in an existing d.o.f.
+(as this would fully specify said d.o.f.).
+Therefore the set of fragment variable $q^i$ already provide a maximal description for the fragments.
+This means that a state $\psi$ must tell us for each point in $\mathcal{Q}$ how much material is there and the correlation of the unstated part.
 That is: the state will be identified by $\psi : \mathcal{Q} \rightarrow \mathbbm{C}$, a complex function over the $\mathcal{Q}$.
@@
 As in the classical case, such density mush be integrable to give us the total amount of material.
-Therefore the function $\psi \in L^2(\mathcal{Q}, \mu)$ is square integrable.
+Therefore the function $\psi(q) \in L^2(\mathcal{Q}, \mu)$ is square integrable.
 The norm $|\psi|^2 = \int_{\mathcal{Q}} |\psi(q)|^2 dq$ will give us the amount of material associated with the state.
@@
 We can set, by convention, the amount of material for a particle to be unitary, therefore quantum particle states are the subset of $\Psi$ such that $|\psi|^2 = 1$.
-Note that changing the phase by the same amount everywhere is only changes the unstated part.
-Such change does not change the state, therefore distribution whose differ by a total phase are not physically distinguishable and represent the same physical state.
+As we have seen, changing the phase by a constant only affects the unstated part.
+Therefore unitary distributions whose differ by a total phase are not physically distinguishable and represent the same quantum particle state.
 
@@
 Let $\Upsilon_{q^i} : \Psi \rightarrow C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu, \mathbbm{C})$ be the function that given a state returns its wave function.
-As the wave function capture the maximal description of the system, two states must represent different distributions.
+As the wave function captures the maximal description of the system, two states must represent different distributions.
 $\forall \psi_1, \psi_2 \in \Psi, \psi_1 \neq \psi_2 \implies \Upsilon_{q^i}(\psi_1) \neq \Upsilon_{q^i}(\psi_2)$.
@@
 Let $\psi=\psi_1+\psi_2$, then $\Upsilon_{q^i}(\psi)=\Upsilon_{q^i}(\psi_1)+\Upsilon_{q^i}(\psi_2)$ as combining the distributions means combining the statistical description of the fragments.
-Let $\psi_1=\tau(c)\psi_2$, then $\Upsilon_{q^i}(\psi_1)=c \Upsilon_{q^i}(\psi_2))$ as $\tau(c)$ transforms the fragment configuration by the factor $C$.
+Let $\psi_1=\tau(c)\psi_2$, then $\Upsilon_{q^i}(\psi_1)=c \Upsilon_{q^i}(\psi_2)$ as $\tau(c)$ transforms the fragment configuration by the factor $C$.
 $\Upsilon_{q^i}$ is a homomorphism as a vector space.
@@
 Let $\psi \in \Psi$ be the state of a quantum particle.
-$e^{\imath \theta} \psi$ with $\theta \in \mathbbm{R}$ is not physically distinguishable from $\psi$ as phases are only physically distinguishable through addition.
-$\sqrt{\rho} \psi$ with $\sqrt{\rho} \in \mathbbm{R}^+$ cannot represent a quantum particle as it as a different amount of material.
-There exists only one distinguishable state for a quantum particle in the equivalence class $\psi \sim c \psi$ with $c \in \mathbbm{C}$.
+$\tau(e^{\imath \theta}) \psi$ with $\theta \in \mathbbm{R}$ is not physically distinguishable from $\psi$ as phases are only physically distinguishable through addition.
+$\tau(\sqrt{\rho}) \psi$ with $\sqrt{\rho} \in \mathbbm{R}^+$ cannot represent a quantum particle as it as a different amount of material.
+There exists only one distinguishable quantum particle state in the equivalence class $\psi \sim \tau(c) \psi$ with $c \in \mathbbm{C}$.
 The state with no amount of material does not correspond to any particle state.
@@
 
-Both in the classical and quantum state the full state is identified by distributions on $\mathcal{Q}$.
-The $q^i$ do not label the states, but the infinitesimal parts.
+In both the reducible and irreducible case, the material is distributed over $\mathcal{Q}$.
+The $q^i$ do not label the states of the material itself, but the configurations for the infinitesimal parts.
 For a classical material, the infinitesimal parts can be given a state, have their own topological space and a set of local state variables $(q^i, p_i)$ to identify them.
@@
 For an irreducible material, the infinitesimal parts cannot be given a state, cannot be identified by a set of local variables.
-We have to develop other mathematical tools.
-
-To be on par with classical mechanics we need at least address three problems.
-We need to be able to compare two quantum states and ask what fraction of them share the same configuration: this will give us an inner product.
-We need to be able to express state variable distributions and expectations: this will give us linear operators.
+We have to use other mathematical tools, such as vector spaces.
+As we will often use linear combination of states and we will write, for brevity, $c\psi$ instead of $\tau(c)\psi$.
+That is: each complex numbers still represents a state transformation.
+
+For the tool set to be on par with classical mechanics we need to address three problems.
+We need to be able to compare two states and ask what fraction of them share the same configuration: this will give us an inner product.
+We need to be able to express distributions and expectations for different quantities: this will give us linear operators.
 We need to identify the conjugate quantities $k_i$ and $p_i$.
@@
 The idea is that, given two states $\psi_1$ and $\psi_2$, we should be able to tell how much of $\psi_1$ is prepared in the same way of $\psi_2$.
-That is $\psi_1 = \alpha \psi_2 + \psi_3$ is the combination of the part that is prepared like $\psi_2$ (e.g.
-same distributions in position and momentum) and a part that is prepared completely differently.
-Let $\mathsf{P}_{\psi_2} : \Psi \rightarrow \Psi$ be the operator that gives us the equally prepared part $\mathsf{P}_{\psi_2} (\psi_1) \equiv \alpha \psi_2$.
-This operator must be linear in $\psi_1$: combining parts that are equally prepared still gives us parts that are equally prepared.
-It also gives us the same result if applied once or twice: once we isolated the part that is equally prepared, it remains the part that is equally prepared.
+That is $\psi_1 = \alpha \psi_2 + \psi_3$ is the combination of the component that is prepared like $\psi_2$ (e.g.
+same distributions in position and momentum) and a component that is prepared completely differently.
+Let $\mathsf{P}_{\psi_2} : \Psi \rightarrow \Psi$ be the operator that gives us the equally prepared component $\mathsf{P}_{\psi_2} (\psi_1) \equiv \alpha \psi_2$.
+This operator must be linear in $\psi_1$: combining components that are equally prepared still gives us components that are equally prepared.
+It also gives us the same result if applied once or twice: the component that is equally prepared is already fully equally prepared.
 $\mathsf{P}_{\psi_2}$ is therefore a projection: a linear operator such that $\mathsf{P}^2_{\psi_2}= \mathsf{P}_{\psi_2}$.
@@
 That is: $\mathsf{P}_{\psi_2} (\psi_1) = \psi_1 \Leftrightarrow \psi_1 = \alpha \psi_2$.
-Moreover, because the material is irreducible, if $23\%$ of $\psi_1$ is prepared like $\psi_2$, then $23\%$ of $\psi_2$ is prepared like $\psi_1$.
-If not, we could learn more (or less) about $\psi_1$ by studying the part of $\psi_2$ that is prepared like $\psi_1$.
+Moreover, because the material is irreducible, if some fraction of $\psi_1$ is prepared like $\psi_2$, then the same fraction of $\psi_2$ is prepared like $\psi_1$.
+If not, we could learn more (or less) about $\psi_1$ by studying the component of $\psi_2$ that is prepared like $\psi_1$.
 Therefore we have $\frac{|\mathsf{P}_{\psi_2} (\psi_1)|^2}{|\psi_1|^2}=\frac{|\mathsf{P}_{\psi_1} (\psi_2)|^2}{|\psi_2|^2}$.
-Lastly, if we take the projection of $\psi_1$ along $\psi_2$ and we project it again on $\psi_1$, we end up with a part of $\psi_1$.
-As this is a proper part of $\psi_1$, its state is just $\psi_1$ rescaled, no change in correlation.
+Lastly, if we take the projection of $\psi_1$ along $\psi_2$ and we project it again on $\psi_1$, we end up with a component of $\psi_1$.
+As this is a part of $\psi_1$ prepared as $\psi_1$, its state is just $\psi_1$ rescaled, with no change in correlation.
 That is: $\mathsf{P}_{\psi_1} (\mathsf{P}_{\psi_2} (\psi_1)) = a \psi_1$ with $a \in \mathbbm{R}$.
@@
 Let $\psi \in \Psi$ be a state.
-Let $\mathsf{P}_\psi : \Psi \rightarrow \Psi$ be an operator that returns the part of a given state that is prepared as $\psi$.
-$\mathsf{P}_{\psi}(c_1 \phi_1 + c_2 \phi_2) = c_1 \mathsf{P}_{\psi}(\phi_1) + c_2 \mathsf{P}_{\psi}(\phi_2)$ is linear as the part of the composition prepared like $\psi$ is the composition of the parts so prepared.
-$\mathsf{P}_\psi \circ \mathsf{P}_\psi = \mathsf{P}_\psi$ is a projection because the whole $\mathsf{P}_\psi$  is prepared as $\psi$.
+Let $\mathsf{P}_\psi : \Psi \rightarrow \Psi$ be an operator that returns the component of a given state that is prepared as $\psi$.
+$\mathsf{P}_{\psi}(c_1 \phi_1 + c_2 \phi_2) = c_1 \mathsf{P}_{\psi}(\phi_1) + c_2 \mathsf{P}_{\psi}(\phi_2)$ is linear as the component of the composition prepared like $\psi$ is the composition of the components so prepared.
+$\mathsf{P}_\psi \circ \mathsf{P}_\psi = \mathsf{P}_\psi$ is a projection because the result of $\mathsf{P}_\psi$  is prepared as $\psi$.
 
@@
 Let $\psi, \phi \in \Psi$.
-Consider $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi))$: we first take the part of $\psi$ prepared like $\phi$ and then take the part of that prepared like $\psi$.
+Consider $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi))$: we first take the component of $\psi$ prepared like $\phi$ and then take the component of that prepared like $\psi$.
 The overall operation is simply returning a part of $\psi$.
@@
 Also note that the observations made after \ref{prop:differentiable_manifold} for the  classical case are still valid.
-$\Psi$ cannot be a complete metric space as it does not include discontinuous functions.
+$\Psi$ cannot be a complete metric space (for the norm induced by the inner product) as it does not include discontinuous functions.
 Therefore it is only a pre-Hilbert space.
 For mathematical convenience, one can take its completion with respect to the norm induced by the inner product and obtain a Hilbert space.
-Yet, such object does no longer represent a physically meaningful entity.
+Yet, such object does no longer represent a physically meaningful set of states.
 
@@
 
-The proceed in the following way.
-Suppose we have a state variable $f : \mathcal{Q} \rightarrow \mathbbm{R}$.
+We proceed in the following way.
+Suppose we have a fragment variable $f : \mathcal{Q} \rightarrow \mathbbm{R}$.
 Consider the product $f (q) \psi (q)$: it returns the value of $f$ weighted by the wave function.
@@
 The conjugate variables $p_i=\hbar k_i$ cannot be expressed as a function of $q^i$, which reiterates the idea that we cannot specify all variables for each fragment.
-Suppose there is another set of state variables that provide another way to subdivide the fragments.
-Then those variables will form a manifold $\hat{\mathcal{Q}}$ such that each $\psi$ can be expressed as a wave function $\psi(\hat{q})$ over $\hat{\mathcal{Q}}$.
-
-As each state $\psi$ can be identified by a wavefunction on either space, there must be an operator $\Upsilon^q_{\hat{q}}(\psi(q)) = \psi(\hat{q})$ that converts the wave function on one space to the other.
-As to a linear combination of states corresponds a linear combination of wave function, $\Upsilon^q_{\hat{q}}$ is a linear operator.
+Suppose there is another set of state variables that provide another maximal description for the fragments.
+Then those variables will form a topological space $\hat{\mathcal{Q}}$ such that each $\psi$ can be expressed as a wave function $\psi(\hat{q}) :\hat{\mathcal{Q}} \rightarrow \mathbbm{C}$ over $\hat{\mathcal{Q}}$.
+
+As each state $\psi$ can be identified by a wave function on either space, there must be an operator $\Upsilon^q_{\hat{q}}(\psi(q)) = \psi(\hat{q})$ that converts the wave function on one space to the other.
+As to a linear combination of states corresponds a linear combination of wave functions, $\Upsilon^q_{\hat{q}}$ is a linear operator.
 As converting the space again must give us the original wave function, $\Upsilon^{\hat{q}}_{q} \Upsilon^q_{\hat{q}}= 1$ is unitary.
 We can now define a have linear operator $\hat{F}$ corresponding to state variable $\hat{f} : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
-Using $\Upsilon^{\hat{q}}_{q}$ we can express $\hat{F}$ as an operator over complex functions of $\mathcal{Q}$ It will be still be linear and hermitian.
+Using $\Upsilon^{\hat{q}}_{q}$ we can express $\hat{F}$ as an operator over complex functions of $\mathcal{Q}$.
+It will be still be linear and hermitian.
 
@@
 
-That is: for each state variable we have a corresponding linear hermitian operator which can be expressed as function multiplication in a suitable choice of $\hat{\mathcal{Q}}$.
+That is: for each state variable $f$ we have a corresponding linear hermitian operator $F$ which can be expressed as function multiplication in a suitable choice of $\hat{\mathcal{Q}}$ and whose expectations is given by $\langle \psi | F \psi \rangle$.
 

### 2016-12-01 2015CompleteDerivation/completeDerivation.tex
Continuing the first pass on quantum

@@
 The configuration of a fragment of the material is characterized by two random variables $A, B: \Omega \rightarrow \mathbbm{R}$ where $\Omega$ is the sample space for the configuration of the fragments.
-As the configuration of each fragment is physically indistinguishable, $A$ and $B$ can be chosen such that the distribution $\rho$ of material over such variables is uniform and unitary.
-We can also rescale $A$ and $B$ such that they are defined on the range $[0, \sqrt{a_m}]$.
-We have $a_m=\int_{\Omega} \rho dA \wedge dB = \int_{\Omega} dA \wedge dB$.
-Consider a second amount $\hat{a}_m$ of irreducible material in the same configuration.
+As the configuration of each fragment is physically indistinguishable, $A$ and $B$ can be chosen such that the distribution $\rho$ of material over such variables is uniform over a range of $\sqrt{a_m}$ for both variables.
+We have $a_m=\int \rho dA \wedge dB = \int dA \wedge dB$.
+Consider a second amount $\hat{a}_m$ of irreducible material in the same state.
 It also can be characterized by a uniform distribution over two variables $C, D: \Omega \rightarrow \mathbbm{R}$ over a range of $\sqrt{\hat{a}_m}$.
-As the two amounts of material are in the same configuration, $(C,D)$ is a function of $(A,B)$.
+As the two amounts of material are in the same state, $(C,D)$ is a function of $(A,B)$.
 We have $\hat{a}_m = \int dC \wedge dD = \int |J| dA \wedge dB$ where $|J|$ is the Jacobian determinant.
@@
 The transformation has to be linear: $C=aA + bB$ and $D=-bA + aB$.
-As the amount of material is the only physically distinguishable quantity and is fully specified by $a$ and $b$, no other transformation is physically relevant.
+As the change in amount of material, the only physically distinguishable quantity associated with the fragments, is fully specified by $a$ and $b$, no other transformation is physically relevant.
 Let $\tau: \mathbbm{C} \rightarrow T$ represent such a transformation parametrized by $c=a+\imath b$.
-$|c|^2 = (a^2 + b^2) = \hat{a}_m / a_m$ is the ratio of the two amounts of materials and is physically distinguishable.
-Define on $T$ an addition $+: T \times T \rightarrow T$ and a multiplication $*: T \times T \rightarrow T$ such that $\tau(a) + \tau(b) = \tau(a+b)$ and $\tau(a) * \tau(b) = \tau(a*b)$, $a,b \in \mathbbm{C}$, so that the sum and product of the transformation is equal to the sum and product of their respective factors.
+$|c|^2 = |J| = (a^2 + b^2) = \hat{a}_m / a_m$ is the ratio of the two amounts of materials and is physically distinguishable.
+Define on $T$ an addition $+: T \times T \rightarrow T$ and a multiplication $*: T \times T \rightarrow T$ such that $\tau(a) + \tau(b) = \tau(a+b)$ and $\tau(a) * \tau(b) = \tau(a*b)$, $a,b \in \mathbbm{C}$, the sum and product of the transformation is equal to the sum and product of their respective factors.
 The phase $arg(c)$ under addition is physically distinguishable as it can affect the modulus of the result.
@@
 The transformation on each d.o.f.
-is similarly defined by $c_i \in \mathbbm{C}$.
+is defined independently by $c_i \in \mathbbm{C}$.
 As $\hat{a}_m / a_m$ is the same under all d.o.f.
 $|c_i|=|c_j|$.
-That has to always hold under addition as well, therefore $arg(c_i) = arg(c_j)$.
-$T\cong \mathbbm{C}$ even in the case of multiple d.o.f.
+The relationship has to always hold under addition as well, therefore $arg(c_i) = arg(c_j)$.
+$T\cong \mathbbm{C}$ also for multiple d.o.f.
 
@@
 
-Naturally, we cannot give a complete description of the configuration of the fragments, therefore we cannot give a joint distribution for all fragment variables at the same time.
-That is: irreducibility contains the notion of complementarity.
-We can, though, find a maximal set of fragment variables: one that provides the best description possible.
+Naturally, we cannot give a complete description for the configuration of a fragment, therefore we cannot give a joint distribution for all fragment variables at the same time.\footnote{This recovers the notion of complementarity introduced by Bohr.} We can, though, find a maximal set of fragment variables: one that provides the best description possible.
 While it should be clear that there cannot be a single unique set, we can show that one such set is provided by the fragment variables $q^i : \mathcal{Q} \rightarrow \mathbbm{R}$ that define the units required to describe the system.
@@
 (as this would fully specify said d.o.f.).
-Therefore the set of fragment variable $q^i$ already provide a maximal description for the fragments.
+Therefore the set of fragment variables $q^i$ already provide a maximal description for the fragments.
 This means that a state $\psi$ must tell us for each point in $\mathcal{Q}$ how much material is there and the correlation of the unstated part.
-That is: the state will be identified by $\psi : \mathcal{Q} \rightarrow \mathbbm{C}$, a complex function over the $\mathcal{Q}$.
+That is: the state will be identified by $\psi : \mathcal{Q} \rightarrow \mathbbm{C}$, the wave function over the $\mathcal{Q}$.
 
 The square modulus at each point will give us the density of the material.
-As in the classical case, such density mush be integrable to give us the total amount of material.
+As in the classical case, such density must be integrable to give us the total amount of material.
 Therefore the function $\psi(q) \in L^2(\mathcal{Q}, \mu)$ is square integrable.
@@
 
-As the material is homogeneous, all particles of the material are made of the same amount.
+As the material is homogeneous, all its quantum particles are made of the same amount.
 We can set, by convention, the amount of material for a particle to be unitary, therefore quantum particle states are the subset of $\Psi$ such that $|\psi|^2 = 1$.
@@
 \begin{prop}\label{prop:wavefuntion}
-	The state space $\Psi$ for an irreducible material is isomorphic to a subspace of the space of Lebesgue square integrable differentiable.
+	The state space $\Psi$ for an irreducible material is isomorphic to a subspace of the space of Lebesgue square integrable differentiable functions.
 That is $\Psi \cong \Psi(\mathcal{Q}, dp^i) \subseteq C^1(\mathcal{Q}) \cap L^2(\mathcal{Q}, \mu)$ under the isomorphism $\Upsilon_q : \Psi \leftrightarrow \Psi(\mathcal{Q}, dp^i)$ where $\Psi(\mathcal{Q}, dp^i)$ is the space of wave functions over $q^i$.
@@
 \begin{justification}
-	We claim the manifold $\mathcal{Q}$ that defines the units to describe a fragment of an irreducible material also provides a maximal description for the fragment configuration, that is: no other state variables can be added.
+	We claim the manifold $\mathcal{Q}$ that defines the units to describe a fragment of an irreducible material also provides a maximal description for the fragment configuration, that is: no other fragment variable can be added.
 Let $\mathcal{Q}$ be the manifold that defines the units.
@@
 $\rho_\psi(q^i) \in C(\mathcal{Q}, \mathbbm{R})$ by \ref{continuous_map}.
-Therefore $\psi(q^i) \in C(\mathcal{Q}, \mathbbm{R})$.
+Therefore $\psi(q^i) \in C(\mathcal{Q}, \mathbbm{C})$.
 $\rho_\psi(q^i) \in L^1(\mathcal{Q}, \mu, \mathbbm{R})$ by the same arguments in the justification for \ref{integration}.
@@
 Let $\psi=\psi_1+\psi_2$, then $\Upsilon_{q^i}(\psi)=\Upsilon_{q^i}(\psi_1)+\Upsilon_{q^i}(\psi_2)$ as combining the distributions means combining the statistical description of the fragments.
-Let $\psi_1=\tau(c)\psi_2$, then $\Upsilon_{q^i}(\psi_1)=c \Upsilon_{q^i}(\psi_2)$ as $\tau(c)$ transforms the fragment configuration by the factor $C$.
+Let $\psi_1=\tau(c)\psi_2$, then $\Upsilon_{q^i}(\psi_1)=c \Upsilon_{q^i}(\psi_2)$ as $\tau(c)$ transforms the fragment configuration by the factor $c$.
 $\Upsilon_{q^i}$ is a homomorphism as a vector space.
@@
 $\tau(e^{\imath \theta}) \psi$ with $\theta \in \mathbbm{R}$ is not physically distinguishable from $\psi$ as phases are only physically distinguishable through addition.
-$\tau(\sqrt{\rho}) \psi$ with $\sqrt{\rho} \in \mathbbm{R}^+$ cannot represent a quantum particle as it as a different amount of material.
+$\tau(\sqrt{\rho}) \psi$ with $\sqrt{\rho} \in \mathbbm{R}^+ \backslash \{1\}$ cannot represent a quantum particle as it as a different amount of material.
 There exists only one distinguishable quantum particle state in the equivalence class $\psi \sim \tau(c) \psi$ with $c \in \mathbbm{C}$.
 The state with no amount of material does not correspond to any particle state.
-The quotient space $\mathsf{P}(\Psi) \equiv (\Psi - \{0\}) / \sim$ is isomorphic to the state of quantum particles.
+The quotient space $\mathsf{P}(\Psi) \equiv (\Psi \backslash \{0\}) / \sim$ is isomorphic to the state of quantum particles.
 \end{justification}
@@
 For a classical material, the infinitesimal parts can be given a state, have their own topological space and a set of local state variables $(q^i, p_i)$ to identify them.
-We can use the richness of theory of differentiable manifolds and metric spaces.
-For an irreducible material, the infinitesimal parts cannot be given a state, cannot be identified by a set of local variables.
+We can use the richness of differential geometry and metric spaces.
+For an irreducible material, the infinitesimal parts cannot be given a state, cannot be fully identified by a set of variables.
 We have to use other mathematical tools, such as vector spaces.
-As we will often use linear combination of states and we will write, for brevity, $c\psi$ instead of $\tau(c)\psi$.
+As we will often use linear combination of states, we will write, for brevity, $c\psi$ instead of $\tau(c)\psi$.
 That is: each complex numbers still represents a state transformation.
 
-For the tool set to be on par with classical mechanics we need to address three problems.
+For the tool set to be on par with classical mechanics we need to address three requirements.
 We need to be able to compare two states and ask what fraction of them share the same configuration: this will give us an inner product.
-We need to be able to express distributions and expectations for different quantities: this will give us linear operators.
+We need to be able to express distributions and expectations for different quantities: this will give us self-adjoint linear operators.
 We need to identify the conjugate quantities $k_i$ and $p_i$.
@@
 The idea is that, given two states $\psi_1$ and $\psi_2$, we should be able to tell how much of $\psi_1$ is prepared in the same way of $\psi_2$.
-That is $\psi_1 = \alpha \psi_2 + \psi_3$ is the combination of the component that is prepared like $\psi_2$ (e.g.
+That is $\psi_1 = \alpha \psi_2 + \psi_3$ is the combination of the component that is prepared like $\psi_2$ (e.q.
 same distributions in position and momentum) and a component that is prepared completely differently.
-Let $\mathsf{P}_{\psi_2} : \Psi \rightarrow \Psi$ be the operator that gives us the equally prepared component $\mathsf{P}_{\psi_2} (\psi_1) \equiv \alpha \psi_2$.
-This operator must be linear in $\psi_1$: combining components that are equally prepared still gives us components that are equally prepared.
-It also gives us the same result if applied once or twice: the component that is equally prepared is already fully equally prepared.
+Let $\mathsf{P}_{\psi_2} : \Psi \rightarrow \Psi$ be the operator that gives us the component $\mathsf{P}_{\psi_2} (\psi_1) \equiv \alpha \psi_2$ prepared like $\psi_2$.
+This operator is linear in $\psi_1$: the component of the combined system that is prepared like $\psi_2$ is the sum of the components of the parts that are prepared like $\psi_2$.
+It also gives us the same result if applied once or twice: once we found the component that is prepared like $\psi_2$, it remains the same.
 $\mathsf{P}_{\psi_2}$ is therefore a projection: a linear operator such that $\mathsf{P}^2_{\psi_2}= \mathsf{P}_{\psi_2}$.
@@
 \begin{justification}
-	We claim there is exists a map $| \cdot | ^2 : \Psi \rightarrow \mathbbm{R}$ such that $|\psi|^2 \ge 0 \; \forall \psi \in \Psi$ and $|\psi|^2 = 0 \Leftrightarrow \psi = 0$.
+	We claim there exists a map $| \cdot | ^2 : \Psi \rightarrow \mathbbm{R}$ such that $|\psi|^2 \ge 0 \; \forall \psi \in \Psi$ and $|\psi|^2 = 0 \Leftrightarrow \psi = 0$.
 Let $| \cdot | ^2 : \Psi \rightarrow \mathbbm{R}$ be the map that returns the amount of material for a given state.
-Let $\psi \in \Psi$ and $c \in \mathbbm{C}$.
-$|\tau(c) \psi|^2 = |c|^2 |\psi|^2$ as $\tau(c)$ increases the amount of material by $|c|^2$, which is always positive.
-$|\psi|^2 \ge 0 \; \forall \psi \in \Psi$.
+Let $\psi \in \Psi$.
+$|\psi|^2 = \int \rho(q^i) d^nq = \int \psi^\dagger(q^i) \psi(q^i) d^nq \ge 0$.
 $|\psi|^2 = 0 \Leftrightarrow \psi = 0$ as the only state with no material is the empty state.
@@
 The overall operation is simply returning a part of $\psi$.
-These must correspond to the same exact configuration for the fragments.
+This must correspond to the same exact configuration for the fragments.
 The only thing that can change is the amount of material.
@@
 The inner product associated with the $L^2$ norm is $\int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$ with $\psi_1, \psi_2 \in \Psi$.
-Therefore $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$.
+$\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$.
 
@@
 
-Note that the inner product only has an indirect physical definition, based on the norm and the projections. We'll therefore want to base our physical intuition on the latter concepts.
+Note that the inner product only has an indirect physical definition, based on the norm and the projections.
+The former concepts are therefore more useful to gain physical intuition as the latter does not have a straightforward physical meaning.
 
@@
 As such, we need to make sure that the rest of our definitions and justifications do not require a complete metric space to be valid or they would not be physically meaningful.
-In particular, we cannot write the usual link $\psi(q) = \langle q | \psi \rangle$ between the wave function and the inner product as the material cannot be prepared at a specific value if $q$ is a continuous variable.
-For the same reason, we cannot show that to each state variable is associated an hermitian operator by constructing it from its eigenstates.
+In particular, we don't want to use the usual link $\psi(q) = \langle q | \psi \rangle$ between the wave function and the inner product as the material cannot be prepared at a specific value when $q$ is a continuous variable.
+Similarly, we cannot show that to each state variable is associated an hermitian operator simply by constructing it from its eigenstates.
 
@@
 Suppose we have a fragment variable $f : \mathcal{Q} \rightarrow \mathbbm{R}$.
-Consider the product $f (q) \psi (q)$: it returns the value of $f$ weighted by the wave function.
-We can think of $f$ as an operator that takes a complex function of $\mathcal{Q}$ and returns another complex function of $\mathcal{Q}$.
+Consider the product $f (q) \psi (q)$: it returns the wave function weighted by the value of $f$.
+We can think of $f$ as an operator that takes a complex function of $\mathcal{Q}$ and returns another complex function of $\mathcal{Q}$ that represents the distribution of the expectation over the fragments.
 The operator is linear since $f (q) (a\psi_1 + b\psi_2) = a f \psi_1 + b f  \psi_2$.
 The operator is hermitian since $f^\dagger = f$.
-Consider $\langle \psi | F \psi \rangle = \langle F \psi | \psi \rangle = \int_\mathcal{Q} \psi^\dagger (q) f(q) \psi(q) dq = f(q) | \psi(q)|^2$.
+Consider $\int_\mathcal{Q} \psi^\dagger (q) f(q) \psi(q) dq = \int f(q) | \psi(q)|^2 dq$.
 It integrates the value of $f$ weighted by the amount of the material: it is the expectation of the state variable $f$ over the distribution.
 
-While $\mathcal{Q}$ is enough to define the system of units, it cannot be enough to define all the variables over which we can define the distributions.
+While $\mathcal{Q}$ is enough to define the system of units, it cannot be enough to define all the variables over which we can define the distributions of irreducible material.
 The conjugate variables $p_i=\hbar k_i$ cannot be expressed as a function of $q^i$, which reiterates the idea that we cannot specify all variables for each fragment.
@@
 
-As each state $\psi$ can be identified by a wave function on either space, there must be an operator $\Upsilon^q_{\hat{q}}(\psi(q)) = \psi(\hat{q})$ that converts the wave function on one space to the other.
-As to a linear combination of states corresponds a linear combination of wave functions, $\Upsilon^q_{\hat{q}}$ is a linear operator.
+As each state $\psi$ can be identified by a wave function on either space, there must be an operator $\Upsilon^q_{\hat{q}}(\psi(q)) = \psi(\hat{q})$ that converts the wave function on one space to the other.\footnote{Note that a neighbourhood of $q^i$ is not necesserily a neighbourood of $\hat{q}^i$.
+They are only if $q^i$ and $\hat{q}^i$ can be expressed as a bijection of each other $\hat{q}^i(q^i)$: they are different coordinates of the same manifold.} As to a linear combination of states corresponds a linear combination of wave functions, $\Upsilon^q_{\hat{q}}$ is a linear operator.
 As converting the space again must give us the original wave function, $\Upsilon^{\hat{q}}_{q} \Upsilon^q_{\hat{q}}= 1$ is unitary.
-We can now define a have linear operator $\hat{F}$ corresponding to state variable $\hat{f} : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
-Using $\Upsilon^{\hat{q}}_{q}$ we can express $\hat{F}$ as an operator over complex functions of $\mathcal{Q}$.
+We can now define a have linear operator $F_{[\hat{q}^i]}$ corresponding to state variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
+Using $\Upsilon^{q}_{\hat{q}}$ we can express $F_{[q^i]} = \Upsilon^{\hat{q}}_{q} F_{[\hat{q}^i]} \Upsilon^{q}_{\hat{q}}$ as an operator over complex functions of $\mathcal{Q}$.
 It will be still be linear and hermitian.
 
-TODO: ranges of $q^i$ and $\hat{q}^i$ must be the same: there must be a bijective function between them. Bijection is continuous only if $\hat{q}^i(q^i)$. Otherwise can be discontinuous.
-
-That is: for each state variable $f$ we have a corresponding linear hermitian operator $F$ which can be expressed as function multiplication in a suitable choice of $\hat{\mathcal{Q}}$ and whose expectations is given by $\langle \psi | F \psi \rangle$.
+That is: for each state variable $f$ we have a corresponding linear hermitian operator $F_{[q^i]}$ that acts on wave functions over $q^i$, which can be expressed as function multiplication in a suitable choice of $\hat{\mathcal{Q}}$.
+The expectation multiplied by the amount of material is given by $\int \psi^\dagger(q^i) F_{[q^i]} (\psi(q^i)) d^nq$.
 
 \begin{prop}\label{prop:self_adjoint_operators}
-	For each fragment variable $f$ there exists a corresponding self-adjoint linear operator $F : \Psi \rightarrow \Psi$ such that $\langle \psi | F \psi \rangle$ is the expectation of $f$ multiplied by the amount of material.
-There exist a topological space $\hat{\mathcal{Q}}$ such that $q^1 \equiv f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ and a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ where $\Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ is the space of wave functions over $\hat{\mathcal{Q}}$.
+	For each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ and a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ where $\Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ is the space of wave functions over $\hat{\mathcal{Q}}$.
+There also exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbbm{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i))d^nq $ is the expectation of $f$ multiplied by the amount of material.
 \end{prop}
 \begin{justification}
-	We claim for each fragment variable $f$ there exist a topological space $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
+	We claim for each fragment variable $f$ there exist a topological space $\hat{\mathcal{Q}}$ such that  $q^i : \mathcal{Q} \rightarrow \mathbbm{R}^n$ provides a maximal description and $f : \mathcal{Q} \rightarrow \mathbbm{R}$.
 Let $f$ be a fragment variable.
-Let $\hat{q}^i$ be a set of fragment variables such that $\hat{q}^1\equiv f$.
-If $\hat{q}^i$ does not provide a maximal description of $f$ then there exists another set of fragment variables.
+Let $\hat{q}^i$ be a set of fragment variables such that $f=f(\hat{q}^i)$.
+If $\hat{q}^i$ does not provide a maximal description for the fragments then there exists another set of fragment variables that refine the description.
 Extend $\hat{q}^i$ with such variables.
+If still not maximal, continue extending.
 As the number of possibilities charted by $\hat{q}^i$ cannot exceed the number of possibilities charted by $q^i$, at some point we'll reach a maximal description.
@@
 Let $\psi \in \Psi$ be a state.
-$q^i$ provides a maximal description for the fragments.
-There must exist a $\psi(q^i) : \mathbbm{R}^n \rightarrow \mathbbm{C}$ associated with $\psi$ such that it gives the fragment configuration at a particular value for $q^i$.
+$\hat{q}^i$ provides a maximal description for the fragments.
+There must exist a wave function $\psi(\hat{q}^i) : \mathbbm{R}^n \rightarrow \mathbbm{C}$ associated with $\psi$ such that it gives the fragment configuration at a particular value for $\hat{q}^i$.
 Such wave function provides a full description for the state.
@@
 
-	We claim there exist a self-adjoint linear operator $F : \Psi \rightarrow \Psi$ such that $\langle \psi | F \psi \rangle$ is the expectation of $f$ multiplied by the amount of material.
+	We claim that for each fragment variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ there exists a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ such that $\int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]}( \psi(\hat{q}^i)) d^n \hat{q}$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
+The expectation value of $f$ multiplied by the amount of material is given by $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q}$ where $\hat{q}^i$ provide a maximal description such that $f=f(\hat{q})$.
+We have $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q} = \int \psi^\dagger(\hat{q}^i) f(\hat{q}^i) \psi(\hat{q}^i) d^{n}\hat{q}$.
+Let $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ such that $F_{[\hat{q}^i]} (\psi(\hat{q}^i)) = f(\hat{q}^i) \psi(\hat{q}^i)$.
+$F_{[\hat{q}^i]}$ is a linear operator.
+$F_{[\hat{q}^i]}$ is self-adjoint as $\int (F_{[\hat{q}^i]}(\psi(\hat{q}^i)))^\dagger \psi(\hat{q}^i) d^{n}\hat{q} = \int (f(\hat{q}^i) \psi(\hat{q}^i))^\dagger \psi(\hat{q}^i) d^{n}\hat{q} = \int \psi^\dagger(\hat{q}^i) f(\hat{q}^i) \psi(\hat{q}^i) d^{n}\hat{q} = \int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]} (\psi(\hat{q}^i)) d^{n}\hat{q}$.
+
+	We claim that for any fragment variable $f$ there exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbbm{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i)) d^nq$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
 Let $f$ be a fragment variable.
-The expectation value multiplied by the amount of material is given by $\int f \rho(f) df$ where $\rho(f)$ is the density of material in $f$.
-Let $q^i$ a set of fragment variables that provides a maximal description for the fragments.
-Let $q^1 = f$ and $k=2,...,n$.
-We have $\rho(f) = \int \psi^\dagger(f, q^k) \psi(f, q^k) dq^{n-1}$.
-$\int f df \int \psi^\dagger(f, q^k) \psi(f, q^k) dq^{n-1} = \int \psi^\dagger(q^i) f(q^i) \psi(q^i) dq^{n}$.
-$f$ is a linear operator on the space of the wave function over $\hat{\mathcal{Q}}$ that increases the module at each point.
-TODO.
-Therefore $f(q^i) \psi(q^i) \in \Psi(\hat{\mathcal{Q}})$.
-As $\Upsilon_{\hat{\mathcal{q}}^i}$ is an isomorphism, there exists a linear operator $F : \Psi \rightarrow \Psi$ such that $f(q^i) \psi(q^i) = f(q^i) \Upsilon_{\hat{\mathcal{q}}^i}(\psi) = \Upsilon_{\hat{\mathcal{q}}^i}(F \psi)$.
-$\langle F \psi | \psi \rangle = \int (f(q^i) \psi(q^i))^\dagger \psi(q^i) = \int \psi^\dagger(q^i) f(q^i) \psi(q^i) = \langle \psi | F \psi \rangle$: $F$ is self-adjoint.
+There exist a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ where $\hat{\mathcal{Q}}$ provides a maximal description.
+As both $\hat{\mathcal{Q}}$ and $\mathcal{Q}$ provide a maximal description, we can define $\Upsilon^{q^i}_{\hat{q}^i} \equiv \Upsilon_{\hat{q}^i} \circ \Upsilon_{q^i}^{-1}$ which transforms a wave function over $\mathcal{Q}$ into a wave function over $\hat{\mathcal{Q}}$.
+$\Upsilon^{q^i}_{\hat{q}^i}$ is an isomorphism as $\Upsilon_{\hat{q}^i}$ and $\Upsilon_{q^i}$ are isomorphisms.
+$\Upsilon^{q^i}_{\hat{q}^i}$ is unitary as $\Upsilon^{\hat{q}^i}_{q^i} \circ \Upsilon^{q^i}_{\hat{q}^i} = (\Upsilon^{q^i}_{\hat{q}^i})^{-1} \circ \Upsilon^{q^i}_{\hat{q}^i} = 1$.
+We can extend $\Upsilon^{q^i}_{\hat{q}^i}$ on all continuous functions and define $F_{[q^i]} \equiv \Upsilon^{\hat{q}^i}_{q^i} \circ F_{[\hat{q}^i]} \circ \Upsilon^{q^i}_{\hat{q}^i}$.
+$F_{[q^i]}$ is a self-adjoint linear operator.
+$\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i)) d^nq = \int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]}( \psi(\hat{q}^i)) d^n \hat{q}$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
+
 \end{justification}
 
-The last thing to do is identify the operators corresponding to $k_i$ and $p_i$, expressed over the space of functions of $\mathcal{Q}$.
-We know that the operators have to be covariant: under a change of variables $K_{\hat{q}} = \partial_{\hat{q}} q K_{q}$.
-This is essentially the chain rule: the operator we are looking for is derivation by $q$.
-As it has to be self-adjoint, it will be of the form $K_{q} = a \imath \partial_q$ where $a$ is a real number.
-We choose $a$ to be negative by convention and to be unitary so that $K_{q}$ is expressed in units of inverse $q$ as in the classical case.
-Therefore we have $K_{i} = -\imath \partial_i$ and $P_{i} = -\imath \hbar \partial_i$.
-
-TODO: Schar
+The last thing to do is characterize the conjugate variables $k_i$ and $p_i$.
+To be precise, we want to find $K_{i [q^i]}$, the operators that act on the wave functions over $q^i$ corresponding to fragment variables $k_i$.
+We know that the operators have to be covariant: under a change of units $\hat{q}^j = \hat{q}^j(q^i)$ we must have $K_{\hat{q}^j[q^i]} = \partial_{\hat{q}^j} q^i K_{q^i [q^i]}$.
+We are looking for a linear operator that obeys the chain rule, which turns out to be the derivation by the corresponding $q^i$.
+As it has to be self-adjoint, it will be of the form $K_{i [q^i]} = a \imath \partial_{i}$ where $a$ is a real number.
+We choose $a$ to be negative by convention and to be unitary so that $k_i$ is expressed in units of inverse $q$ as in the classical case.
+Therefore we have $K_{i [q^i]} = -\imath \partial_i$ and $P_{i [q^i]} = -\imath \hbar \partial_i$.
+
+If we assume $\mathcal{Q}$ to be $\mathbbm{R}^n$, the functional $\Upsilon^{q^i}_{p_i} : \Psi(\mathcal{Q}, dq^i) \leftrightarrow \Psi(\mathcal{P}, dq^i)$ that converts a wave function in position to a wave function in momentum is the Fourier transform.
+As the fragment variables for any power of $q^i$ and $p^i$ must exist, the wave functions in both expressions are infinitely differentiable: the space of the wave functions $\Psi(\mathcal{Q}, dq^i) = S(\mathcal{Q}, q^i)$ is the Schwartz space.
 
 \begin{prop}\label{prop:momentum_operator}
-	The operator $P_i$ associated with the conjugate quantity of $q^i$ is $P_i = - \imath \hbar \partial_{i}$.
-The operator that changes the variable of the wave function from $q^i$ to $p_i$ is $\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^3} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
+	Let $\mathcal{Q}=\mathbbm{R}^n$.
+The operator $P_{i [q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbbm{C})$ associated with the conjugate quantity of $q^i$ is $P_{i [q^i]} = - \imath \hbar \partial_{i}$.
+The isomorphism $\Upsilon^{q^i}_{p_i} : \Psi(\mathcal{Q}, dq^i) \leftrightarrow \Psi(\mathcal{P}, dq^i)$ that changes the variable of the wave function from $q^i$ to $p_i$ is $\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^3} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
 The space of wave functions $\Psi(\mathcal{Q}, dq^i) = S(\mathcal{Q}, q^i)$ where $S(\mathcal{Q}, q^i)$ is the Schwartz space of rapidly decreasing smooth functions.
@@
 \begin{justification}
-	We claim the operator associated with $k_i$ is $K_i = - \imath \partial_{i}$.
-Let $K_i : \Psi(\mathcal{Q}, dq^i) \rightarrow \Psi(\mathcal{Q}, dq^i)$ the linear operator associated with $k_i$ as express on the space of wave functions over $q^i$.
-Such operator must be covariant under the transformation $\hat{q}^i = \hat{q}^i(q^i)$.
-That is: $K_{\hat{q}^i} = \partial_{\hat{q}^i}q^iK_{q^i}$.
-By \ref{thrm:covariant_operator} $K_i=c\partial_i$ for some $c \in \mathbbm{C}$.
-$K_i$ is hermitian: $K_i=\imath r\partial_i$ for some $r \in \mathbbm{R}$ by \ref{thrm:antihermitian_derivative}.
+	We claim the operator associated with $k_i$ is $K_{i [q^i]} = - \imath \partial_{i}$.
+Let $K_{i [q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, dq^i)$ be the linear operator associated with $k_i$ as express on the space of wave functions over $q^i$.
+Such operator must be covariant under the transformation $\hat{q}^j = \hat{q}^j(q^i)$.
+That is: $K_{\hat{q}^j [q^i]} = \partial_{\hat{q}^j}q^iK_{q^i [q^i]}$.
+In particular, it needs to be covariant when changing only one variable: $\hat{q}^1 = \hat{q}^1(q^1)$, $\hat{q}^j = q^j$ where $j=2...n$.
+By \ref{thrm:covariant_operator} $K_{i [q^i]}=c\partial_i$ for some $c \in \mathbbm{C}$.
+$K_{i [q^i]}$ is hermitian: $K_{i  [q^i]}=\imath r\partial_i$ for some $r \in \mathbbm{R}$ by \ref{thrm:antihermitian_derivative}.
 $k_i$ is expressed in units of inverse $q^i$ therefore $|r|=1$.
 By convention, we choose $r$ to be negative.
-$K_i = - \imath \partial_{i}$.
-
-	We claim the operator associated with $p_i$ is $P_i = - \imath \hbar \partial_{i}$. As $p_i=\hbar k_i$, their expectation is proportional by $\hbar$. $P_i=\hbar K_i= -\imath \hbar \partial_i$.
+$K_{i [q^i]} = - \imath \partial_{i}$.
+
+	We claim the operator associated with $p_i$ is $P_{i [q^i]} = - \imath \hbar \partial_{i}$.
+As $p_i=\hbar k_i$, their expectation is proportional by $\hbar$.
+$P_{i [q^i]}=\hbar K_{i [q^i]}= -\imath \hbar \partial_i$.
 
 	We claim the operator that changes the variable of the wave function from $q^i$ to $p_i$ is $\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^3} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
-Let $\Upsilon^{q_i}_{p^i} : \Psi(\mathcal{Q}, dq^i) \rightarrow \Psi(\mathcal{P}, dp^i)$ the operator that maps a wave function over $\mathcal{Q}$ into a wave function over $\mathcal{P}$, the space charted by $p_i$.
+Let $\Upsilon^{q_i}_{p^i} : \Psi(\mathcal{Q}, dq^i) \rightarrow \Psi(\mathcal{P}, dp^i)$ be the operator that maps a wave function over $\mathcal{Q}$ into a wave function over $\mathcal{P}$, the space charted by $p_i$.
 $\Upsilon^{q_i}_{p^i}$ must exist and be unique (up to a total phase) as $\Upsilon^{q_i}_{p^i} = \Upsilon_{p^i}\Upsilon^{q_i} = \Upsilon_{p^i}(\Upsilon_{q_i})^{-1}$ is an isomorphism.
-$\Upsilon^{q_i}_{p^i}$ must be a linear transform that takes the operator $P_i= - \imath \hbar \partial_i$ as expressed over $q^i$ to $P_i = p_i$ as expressed over $p_i$.
+$\Upsilon^{q_i}_{p^i}$ must be a linear transform that takes the operator $P_{i [q^i]}= - \imath \hbar \partial_i$ as expressed over $q^i$ to $P_{i [p_i]} = p_i$ as expressed over $p_i$.
 Such transformation is the Fourier transform.
@@
 Let $\psi(q^i) \in \Psi(\mathcal{Q}, dq^i)$.
-$P_i^n \psi(q^i) = (-\imath \hbar \partial_i)^n \psi(q^i)$ represents the n-th power of the conjugate momentum operator applied to the given state.
-As the quantity is physical, such operation must be defined for any $n$ therefore $\psi(q^i) \in C^\infty(\mathcal{Q})$.
+$P_{i [q^i]}^n \psi(q^i) = (-\imath \hbar \partial_i)^n \psi(q^i)$ represents the n-th power of the conjugate momentum operator applied to the given state.
+As the quantity is physically meaningful, such operation must be defined for any $n$ therefore $\psi(q^i) \in C^\infty(\mathcal{Q})$.
 Let $\psi(p_i) = \Upsilon^{q^i}_{p_i} (\psi(q^i))$ be the wave function over momentum space.
-$(Q^i)^n \psi(p_i) = (+\imath \hbar \partial_{p_i})^n \psi(p_i)$ represents the n-th power of the $q^i$ operator applied to the given state.
-As the quantity is physical, such operation must be defined for any $n$ therefore $\psi(p_i) \in C^\infty(\mathcal{P})$.
+$(Q^i_{[p_i]})^n \psi(p_i) = (+\imath \hbar \partial_{p_i})^n \psi(p_i)$ represents the n-th power of the $q^i$ operator applied to the given state.
+As the quantity is physically meaningful, such operation must be defined for any $n$ therefore $\psi(p_i) \in C^\infty(\mathcal{P})$.
 As $\psi(q^i)$ is the Fourier transform of an infinitely smooth function, it decreases faster then any inverse power of $q^i$.
@@
 	We claim $\Psi(\mathcal{Q}, dq^i) = S(\mathcal{Q}, q^i)$.
-Let $\psi(q^i) \in \Psi(\mathcal{Q}, dq^i)$ and $\hat{\psi}(q^i) \in S(\mathcal{Q}, q^i)$ such that $\psi(q^i) \neq \hat{\psi}(q^i)$.
-There exists $a, b \in \mathbbm{Z}^+$ such that, for a given $i$, $\lVert \psi(q^i) - \hat{\psi}(q^i) \rVert_{i,a,b} \neq 0$  where $\lVert \psi(q^i) \rVert = sup|(q^i)^a (\partial_i) ^b \psi(q^i)|$.
-The operator $(Q^i)^a(P_i)^b$ applied to the two functions will provide different distributions for the associated fragment variables.
-$\hat{\psi}(q^i)$ is physically distinguishable and is associated with a physical state.
-$\hat{\psi}(q^i) \in \Psi(\mathcal{Q}, dq^i)$
+Let $\psi(q^i) \in \Psi(\mathcal{Q}, dq^i)$ and $\phi(q^i) \in S(\mathcal{Q}, q^i)$ such that $\psi(q^i) \neq \phi(q^i)$.
+There exists $a, b, j, k \in \mathbbm{Z}^+$, $j,k < n$ such that $\lVert \psi(q^i) - \phi(q^i) \rVert_{j,a,k,b} \neq 0$  where $\lVert \psi(q^i) \rVert_{j,a,k,b} = sup|(q^j)^a (\partial_k) ^b \psi(q^i)|$ as $S(\mathcal{Q}, q^i)$ is a Frechet space induced by that family of seminorms.
+The operator $(Q^j_{[q^i]})^a(P_{k [q^i]})^b$ applied to the two functions will provide different distributions for the associated fragment variables.
+$\phi(q^i)$ is physically distinguishable and is associated with a physical state.
+$\phi(q^i) \in \Psi(\mathcal{Q}, dq^i)$
 \end{justification}
@@
 We now have all the basic features of the state space of a quantum particle.
-And we have them by re-deriving them from first principles instead of using some kind of formal analogy.
-The similarity with the classical framework are simply due to some of the similar starting points.
-
-We reached the goal despite $\Psi$ is not being a complete metric space.
+And we obtained them by re-deriving them from first principles instead of using some formal analogy.
+The similarity with the classical framework are simply due to the similarity of the starting points.
+
+It should be clear, though, how the correspondence between mathematical and physical objects is not as satisfying as in the classical case.
+For instance, while any differentiable function of $T^*\mathcal{Q}$ can be a state variable for a classical particle, not all self-adjoint linear operators can be associated with a fragment variable.
+Consider $f: \mathbbm{R} \rightarrow \mathbbm{R}$ for which $f(r) = -r$ if $r$ is an integer and $f(r) = r$ otherwise.
+Clearly $f(q)$ does not preserve the topology and cannot be used as a fragment variable.
+Yet, we can still create a self-adjoint linear operator, with the same spectra as the one associated with $q$, just with the eigenvalues switched.
+It is not clear what plays the role of the differentiable structure of $T^*\mathcal{Q}$ in the quantum case.
+
+The situation is worsened by the fact that the topology of $\mathcal{P}$, the space charted by conjugate momentum, depends globally on the topology of $\mathcal{Q}$.
+For example, if $\mathcal{Q}$ is an n-dimensional torus, there is no infinity and $\mathcal{P}$ has a discrete topology.
+That is why, in the end, we restricted ourselves to $\mathbbm{R}^n$.
+Moreover, we obtained the space of functions not because we required them to fall off at infinity, but because we required the ability to express polynomials of $q^i$ and $p_i$.
+It is not immediately clear how this generalizes over a manifold.
+
+While the link between math and physics may not be as elegant and as general as one would like, we still managed the goal: we identified the states of an irreducible material and saw how these are related to the states in quantum mechanics.
 
@@
 	The idea is that each amount of material has a state, the physically distinguishable configuration of the distribution, and an unstated part, the physically indistinguishable configuration of the fragments of the material.
-The system is deterministic and reversible in the sense outlined in section ???: given the state of the system at one time (i.e.
+The system is deterministic and reversible in the sense outlined in assumption \ref{ass:determinism}: given the state of the system at one time (i.e.
 the whole distribution) we can predict/reconstruct the future/past states.
-But we cannot tells anything about the evolution of the unstated part (i.e.
+But we cannot tell much about the evolution of the unstated part (i.e.
 the motion of the fragments).
@@
 	As we saw, the state space recovered under this assumption is the one used in quantum mechanics.
-In the following sections we will also derive two types of state evolution: the deterministic and reversible one, corresponding to unitary evolution (e.g.
-the Schroedinger equation), and a non deterministic one useful to describe measurement interactions, corresponding to the projection postulate.
+In the following sections we will also derive two types of state evolution: the deterministic and reversible one, corresponding to unitary evolution (i.e.
+Schroedinger's equation), and a non deterministic one useful to describe measurement interactions, corresponding to the projection postulate.
 Therefore, as the infinitesimal reducibility assumptions leads to classical mechanics, the irreducibility assumption leads to quantum mechanics.
 
-	As we previously saw in section ???, the infinitesimal reducibility assumptions had its problem. How about the irreducibility assumption? How do they compare?
-
-	The methodological assumes a different connotation.
+	As we saw in assumption \ref{ass:infinitesimal_reducibility}, infinitesimal reducibility had its problems. How about the irreducibility assumption? How do they compare?
+
+	The methodological problem assumes a different connotation.
 In order for the assumption to hold, we need to show that there are no processes at our disposal that can probe the constituents of a quantum particle.
@@
 So we simply have to enumerate the processes available to us, which is, at least in principle, feasible.
-
-	The issues rising from a lack of perfect isolation all but disappear.
+Yet, we can never rule out that someone in the future may discover a new process.
+
+	The issues rising from a lack of perfect isolation are addressed.
 Each degree of freedom comes with its own unstated part (i.e.
@@
 
-	In these way, the irreducibility assumption fares much better than the infinitesimal reducibility assumption. Yet, it is still plagued with some of the same problems.
+	In these ways, the irreducibility assumption fares much better than the infinitesimal reducibility assumption. Yet, it is still plagued with some of the same problems.
 
 	The problem of incompatibility of measurements and deterministic evolution at all levels still remains.
-This will lead us to require a different non deterministic evolution, that will allow us to recover the statistical interpretation of the wavefunction.
+The unstated part of an irreducible system does not change the picture: it is completely undetermined and cannot be used to pass information from one system to another.
 
 	The inability to define time by following a deterministic and reversible problem is still present.
+The unstated part cannot give us a measurable quantity that changes in time as that would be a description of the unstated part.
 
 	There is one additional problem, though, not present in the classical case.
-The quantum states, as we have defined them, are states at equilibrium: they all provide the same level of description for the fragments; they are all reachable from one another through unitary evolution (which we'll see corresponds to deterministic and reversible evolution).
-Therefore they cannot and should not be expected to provide a description under non deterministic evolution.
-Consider a muon decay: the initial state of the muon and the final states of the electron and neutrinos can certainly be well described by a quantum states.
-We have no reason to expect, though, that the while the muon is decaying into the resulting particles the state is always in equilibrium and therefore well described mathematically by a vector in a Hilbert space.
-After all, the state of a single spin one half particle cannot determine the future state of three spin on half particles.
+Quantum particle states, as we have defined them, are states at equilibrium: they all provide the same level of description for the fragments.
+In fact if the internal dynamics were out of equilibrium, the material would not be homogeneous which would contradict the assumption.
+Deterministic and reversible evolution (which we'll see corresponds to unitary evolution) is necessarily quasi static: at each moment a state must be well defined, at each moment the system is in equilibrium.
+Therefore quantum states cannot and should not be expected to provide a description under non deterministic evolution.
+
+	In light of this, consider a muon decay: the initial state of the muon and the final states of the electron and neutrinos can certainly be well described by quantum particle states.
+We have no reason to expect, though, that the while muon is decaying into the resulting particles the state is always in equilibrium and therefore well described mathematically by a vector in a Hilbert space.
+After all, the state of a single spin one half particle cannot determine the future state of three spin one half particles.
 
@@
 Quantities that are not affected by the particular evolution of the unstated part will have the same value under non-deterministic evolution and an equivalent quasi-static evolution.
-The distribution of final states is one such quantities by the irreducibility assumption: they don't care about the unstated part evolution.
-But it stands to reason that, at some point, the unstated part will have actual physical consequences.
-For example: it would be extremely reasonable to expect that is indeed the evolution of the unstated part that decides the moment in which the muon would decay.
-
-	In light of what we discussed, the irreducibility assumption is less flawed but it still cannot be fundamental, in the sense that we cannot take it to strictly apply to all of the universe.
+For example, the overall probability distribution resulting from the decay of a muon is something that cannot depend on the detail motion of the fragment (otherwise we would be able to characterize it).
+In these cases, the use of deterministic and reversible evolution as a ``stand-in" to calculate final statistical distributions is justified.
+But it should not be taken literally as a physical model of what actually happens.
+Because what happens is that, at a particular moment in time for causes the state cannot predict, the muon ceases to be a homogeneous material, and morphs into 3 inhomogeneous pieces, whose final state is not determined by the initial state.
+There is no justification for assuming this process to be quasi-static deterministic and reversible evolution of quantum particle states.
+
+	In light of what we discussed, the irreducibility assumption is less flawed but it still cannot be fundamental, in the sense that we cannot take it to strictly apply to all systems.
 We should not expect to solve open problems, such as the arrow of time and the measurement problem, within the assumption.
@@
 We are now ready to write the equation of motion for an irreducible material.
+Deterministic and reversible evolution for the whole distribution necessarily means quasi static evolution: there is a well defined state at each moment in time.
+Moreover the components must evolve independently as we cannot learn more about one component by observing its evolution in different combinations: the evolution of the composition is the composition of the evolution.
+In other words: the evolution must preserve the inner product.
+
+Note that formally there is a strong analogy with the reducible material case.
+In classical mechanics the state space structure was captured by a symplectic manifold.
+Deterministic and reversible evolution was a transformation that preserved that structure, a symplectomorphism.
+In the quantum case the state space structure is captured by an inner product space.
+Deterministic and reversible evolution is a transformation that preserves that structure, a unitary transformation.
+That is: once we have defined what mathematical structure captures the definition of a state, deterministic and reversible evolution must necessarily preserve such structure.
 
@@
 Let $q : \mathcal{Q} \rightarrow \mathbbm{R}$ a differentiable function.
-Let $K_q$ be a covariant linear operator, that is given an invertible differentiable function $\hat{q} : \mathbbm{R} \rightarrow \mathbbm{R}$ $K_{\hat{q}} = \partial_{\hat{q}} q K_{q}$.
+Let $K_q : C^1(\mathcal{Q}, \mathbbm{C}) \rightarrow C^1(\mathcal{Q}, \mathbbm{C})$ be a covariant linear operator, that is given an invertible differentiable function $\hat{q} : \mathbbm{R} \rightarrow \mathbbm{R}$ $K_{\hat{q}} = \partial_{\hat{q}} q K_{q}$.
 Then $K_q = a \partial_q$ for some $a \in \mathbbm{C}$.

### 2016-12-05 2015CompleteDerivation/completeDerivation.tex
Minor improvemens on quantum

@@
 
-\begin{assump}[Infinitesimal reducibility]\label{ass:irreducibility}
+\begin{assump}[Irreducibility]\label{ass:irreducibility}
 	The system under study is composed of an irreducible homogeneous material and as a whole undergoes deterministic and reversible evolution.
@@
 the whole distribution) we can predict/reconstruct the future/past states.
-But we cannot tell much about the evolution of the unstated part (i.e.
+But we cannot describe the evolution of the unstated part (i.e.
 the motion of the fragments).
@@
 	As we saw, the state space recovered under this assumption is the one used in quantum mechanics.
-In the following sections we will also derive two types of state evolution: the deterministic and reversible one, corresponding to unitary evolution (i.e.
+In the following sections we will derive two types of state evolution: the deterministic and reversible one, corresponding to unitary evolution (i.e.
 Schroedinger's equation), and a non deterministic one useful to describe measurement interactions, corresponding to the projection postulate.
@@
 Each degree of freedom comes with its own unstated part (i.e.
-the motion of the fragments within that d.o.f.) therefore these is no pretense of perfect isolation.
-It is built into the description.
+the motion of the fragments within that d.o.f.) therefore these is no pretense of a perfect description under perfect isolation.
 
@@
 
-	The inability to define time by following a deterministic and reversible problem is still present.
+	The inability to define time using a deterministic and reversible evolution is still a problem.
 The unstated part cannot give us a measurable quantity that changes in time as that would be a description of the unstated part.
@@
 
-	In light of this, consider a muon decay: the initial state of the muon and the final states of the electron and neutrinos can certainly be well described by quantum particle states.
-We have no reason to expect, though, that the while muon is decaying into the resulting particles the state is always in equilibrium and therefore well described mathematically by a vector in a Hilbert space.
+	In light of this, consider a muon decay: the initial state of the muon and the final states of the electron and neutrinos can certainly be well described by quantum particle states.\footnote{In this context, a state is in equilibrium regardless of it being in a local minimum, easily overcome by fluctuations, or a global minimum, that persists indefinitely.} We have no reason to expect, though, that the while muon is decaying into the resulting particles the state is always in equilibrium and therefore well described mathematically by a vector in a Hilbert space.
 After all, the state of a single spin one half particle cannot determine the future state of three spin one half particles.
@@
 	Using unitary evolution in those cases to make predictions, though, may still make physical sense.
-Quantities that are not affected by the particular evolution of the unstated part will have the same value under non-deterministic evolution and an equivalent quasi-static evolution.
+Quantities that are not affected by the particular evolution of the unstated part will have, by their very nature, the same value under non-deterministic evolution and an equivalent quasi-static evolution.
 For example, the overall probability distribution resulting from the decay of a muon is something that cannot depend on the detail motion of the fragment (otherwise we would be able to characterize it).
@@
 But it should not be taken literally as a physical model of what actually happens.
-Because what happens is that, at a particular moment in time for causes the state cannot predict, the muon ceases to be a homogeneous material, and morphs into 3 inhomogeneous pieces, whose final state is not determined by the initial state.
 There is no justification for assuming this process to be quasi-static deterministic and reversible evolution of quantum particle states.
 
-	In light of what we discussed, the irreducibility assumption is less flawed but it still cannot be fundamental, in the sense that we cannot take it to strictly apply to all systems.
+	In other words, the assumption of irreducibility already contains the seeds of its own demise.
+It incorporates the unstated part, yet it does not allow to characterize its full, out of equilibrium, dynamics.
+
+	In light of what we discussed, the irreducibility assumption is less flawed but it still cannot be taken as fundamental, in the sense that we cannot take it to strictly apply to all systems.
 We should not expect to solve open problems, such as the arrow of time and the measurement problem, within the assumption.
@@
 
-We are now ready to write the equation of motion for an irreducible material.
+We are now ready to write the equations of motion for an irreducible material.
 Deterministic and reversible evolution for the whole distribution necessarily means quasi static evolution: there is a well defined state at each moment in time.
@@
 	We claim $\mathcal{T}_{\Delta t}: \Psi \rightarrow \Psi$ is a unitary operator.
-Let $\psi \in \Psi$ be a state.
-$|\psi|^2 = |\mathcal{T}_{\Delta t} \psi|^2$ because deterministic and reversible evolution does not change the amount of material.
 Let $\psi_1, \psi_2 \in \Psi$ be two states.
@@
 
-Note, though, that nothing tells us that the system must be in an eigenfunction of $H$ or that it the ground state is somehow preferred.
+Note, though, that nothing tells us that the system must be in an eigenfunction of $H$ or that the ground state is somehow preferred.
 In fact, the eigenfunctions of $H$ may not be part of $S(\mathcal{Q}, q^i)$ and therefore may not be physical.
 In other words: we have no basis for the time independent Schroedinger equation.
-This is actually consistent.
+This is actually consistent with our assumptions.
 For example, for a system to reach its ground state it has to be able to radiate energy: it cannot isolated.
@@
 This already hints that deterministic and reversible evolution is not the only dynamic possible.
-Other considerations outside of the scope of this work must be used to recover this behavior (e.g.
+Other considerations, outside of the scope of this work, can be used to recover this behavior (e.g.
 thermodynamics or decoherence).
@@
 As in the classical case, we use $s$ as the evolution parameter.
-Deterministic and reversible evolution $\mathcal{T}_{ds}$ will be a unitary transformation defined over this extended inner product, as the amounts of material get conserved.
+Deterministic and reversible evolution $\mathcal{T}_{ds}$ will be a unitary transformation defined over this extended inner product, as the structure of the state space is conserved.
 Continuous evolution will admit an invariant Hamiltonian $\mathcal{H}$ such that $\imath \hbar \partial_s \psi = \mathcal{H} \psi$.
@@
 This is the time dependent evolution equation.
+It means that the wave function does not change along $s$, the parameter of the infinitesimal transformation generated by $\mathcal{H}$.
 

### 2016-12-05 2015CompleteDerivation/completeDerivation.tex
Complete rewrite of complex vector space justification

@@
 
-\subsection{Homogeneous decomposable systems and real vector spaces}
+\subsection{Homogeneous decomposable systems and vector spaces}
 The notion that a system is decomposable means the states are equipped with a rule of composition that allows one to write $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$: the composite system is the sum of its parts.
@@
 
-It is time to turn these concepts into a precise mathematical formulation following the template used for classical mechanics.
+It is time to turn these concepts into a precise mathematical formulation as we did for classical mechanics.
 We want to formalize the state space for an \emph{irreducible material} as a quantum particle will be a finite amount of such material.
 
-As for the classical material, the irreducible material is thought to be decomposable: each state can be seen as the sum of different components (e.g.
-spatially distributed parts) such that the evolution of the whole is the composition of their evolution.
+As for the classical material, the irreducible material is thought to be decomposable: each amount can be seen as the sum of different components (e.g.
+spatially distributed).
 Therefore \ref{prop:abelian_group} applies and the state space $\Psi$ of an irreducible material is an abelian group.
 
-Note, though, that the components are in general not independent.
-As the motion of each fragment is statistical in nature, it will combine as statistical variables do.
+Yet, the description of each component cannot be more accurate than the description of the whole.
+To capture this idea precisely we need to characterize the aspects of the unstated part that influence the state definition.
+How do we do that?
+We can look at the classical derivation for guidance.
+In that case, we looked at the set of transformations that increased/decreased the amount of material.
+The idea is that we can think of these transformations as characterizing what the state can and cannot say about the infinitesimal parts.
+
+Suppose we have a state $\mathcal{c}_1$ for an amount $a_m$ of classical material.
+Suppose we assigned a numeric identifier to each particle of the material.
+This identifier would be just a dummy label we assign, not a physically quantity.
+As the classical material is homogeneous, permutations between classical particles within the same state have no effect.
+Therefore the ordering given by the identifier does not matter.
+Identifying a classical particle within a composite state is equivalent to picking a number from a random variable $A$ uniformly distributed over an interval of size $a_m$.
+
+Suppose we have another state $\mathcal{c}_2 = \tau(a) \mathcal{c}_1$ where the only difference is an increase/decrease of the amount of material by a factor $a$.
+The random variable $B$ that identifies a particle of $\mathcal{c}_2$ is now a uniform distribution $B=aA$ where the range is the new amount of material: there are now proportionally more/less labels to assign.
+This is also the only possible transformation that can link the two uniform random variables $A$ and $B$.
+Therefore increase/decrease of material is the only operation that is going to affect the labeling of the fragments that is captured by the state.
+In other words: the classical state does not care about the permutation of the fragments, but it cares how many fragments are there.
+
+For the irreducible material we follow the same strategy and study the transformations that affect the labeling of what the irreducible state cannot tell us: the configuration of each fragment.
+Similarly, the irreducible state will not care about permutations of fragment configurations, but it will care how many fragment configurations are there.
+
+Suppose we have a state $\psi_1$ for an amount $a_m$ of irreducible material characterized by one d.o.f.
+The configurations of the infinitesimal fragments will be identified by two values, as a d.o.f.
+is two-dimensional.
+Each fragment configuration is identified by a pair of independent random variables $(A, B)$, which are, as before, just dummy variables we use to name the fragment configurations.
+As the material is homogeneous, the distribution is uniform or we could tell something about the fragment configurations.
+We choose A and B to be uniform distributions of interval $\sqrt{a_m}$: identifying a fragment configuration is like picking a point at random over a square of area $a_m$, the amount of material.
+
+Suppose we have another state $\psi_2$, where the labeling for the fragment configurations has changed.
+We can label the fragment configurations of $\psi_2$ in the same way, with two independent uniform distributions $(C, D)$.
+But since the labeling is the only thing that has changed, we must be able to relate the new distribution over $(C,D)$ to the old distribution over $(A,B)$.
+The relationship between the two pairs of variables must be linear since all distributions are uniform.
+$C=aA+bB$.
+To preserve the square distribution we must then have $D= -bA + aB$.
+We can write $\psi_2 = \tau(a,b) \psi_1$ as $a$ and $b$ are the only parameters of the transformation.
+Or more compactly, $\psi_2 = \tau(c) \psi_1$ where $c=a+\imath b$.
+
+The transformation $\tau(c)$ increases/decreases the area by the modulus $|c|^2=a^2+b^2$, which physically corresponds to the increase/decrease of amount of material.
+It also rotates the square by the phase $\theta_c \equiv arg(c)$, which corresponds to a change in correlation: if $\theta_c = 0$, $C = |c| A$ are clearly correlated while if $\theta_c = \pi$, $C = - |c| A$ are clearly anti-correlated.
+
+This line of reasoning is fairly abstract and it may seem to have no bearing with quantum states.
+Therefore, jumping a bit ahead, let's see if we can already find quantitative consequences of this model and see whether they are consistent with our knowledge of quantum mechanics.
+
+If the amount of material is connected to the area of two uniform distributions, it should combine like the variance of a random variable.
+And in fact it does.
 Suppose $X$ and $Y$ are two random variables,
@@
 As $-1\leq\rho_{X,Y}\leq+1$, the total variance can be greater or smaller than the sum of the variance of the components.
-The strength of the chaotic motion of the fragments is something physically distinguishable as it contributes to the spread of the distribution.
-Therefore while the motion of each fragment is not itself physically distinguishable, it falls within the unstated part, the strength and correlation of the chaotic motion must be, they fall within the state.
-
-How can we characterize the aspects of the unstated part that are captured by the state?
-We can look at the classical derivation for guidance.
-In that case, we looked at the set of transformations that increase/decrease the amount of material.
-We can think of them as the transformations that change the unstated part of the system.
-As the classical material is homogeneous, permutations between classical particles in the same state $\mathcal{s}$ do not matter (i.e.
-the ordering is unstated).
-Increasing/decreasing the amount of material increases/decreases the number of permutations and therefore changes the unstated part of the system.
-For the irreducible material we follow the same strategy: we look for the transformations that affect the unstated part while keeping the overall state, the distribution, unchanged.
-
-Suppose we have $\psi_1$ a state for an amount $a_m$ of irreducible material characterized by one d.o.f.
-The material will be spread over the configurations of the infinitesimal fragments.
-As the material is homogeneous, it is spread evenly among such possible configurations.
-Counting fragment configurations is therefore equivalent to counting the amount of material.
-Each fragment would be identified by a pair of independent random variables $(A, B)$, as a d.o.f.
-is two-dimensional.
-The shape of the distribution and the actual values are not physically distinguishable: these are just dummy variables we assign to count the fragments.
-We can choose A and B to be uniform distributions of interval $\sqrt{a_m}$, so that they identify a square of area $a_m$, the amount of material.
-
-Suppose we have another configuration $\psi_2$, where only the unstated part is changed.
-The unstated part of $\psi_2$ can also be characterized by a square formed by two independent uniform distributions (C, D).
-Counting states on (C,D) is equivalent to counting them on (A,B), as they are simply different arrangements of fragments within the same state.
-The relationship between the two pairs of variables must be linear since all distributions are uniform.
-That is: $C=aA+bB$ and $D= -bA + aB$.
-We can write $\psi_2 = \tau(a,b) \psi_1$ as $a$ and $b$ are the only parameters of the transformation.
-Or more compactly, $\psi_2 = \tau(c) \psi_1$ where $c=a+\imath b$.
-
-The transformation $\tau(c)$ increases/decreases the area by the modulus $|c|^2=a^2+b^2$, which physically corresponds to the increase/decrease of amount of material.
-It also rotates the square by the phase $\theta_c \equiv arg(c) = arctan(a/b)$, which corresponds to a change in correlation: if $\theta_c = 0$, $C = |c| A$ are clearly correlated while if $\theta_c = \pi$, $C = - |c| A$ are clearly anti-correlated.
-
-To better understand the link between phase and correlation, suppose $\psi$ is the state for an irreducible material.
-We can write $\tau(c) \psi = \tau(c_1 + c_2) \psi = \tau(c_1) \psi + \tau(c_2) \psi$ as the composition of two separate transformations.
-We have
+Now suppose $\psi$ is the state for a unit amount of irreducible material (i.e.
+a quantum particle).
+We can write $c \psi = (c_1 + c_2) \psi = c_1 \psi + c_2 \psi$.
+We have:
 \begin{equation}\label{eqn:correlation_state}
-|c|^2=|c_1|^2 + |c_2|^2 + 2 |c_1||c_2|\cos(\theta_{c_2} - \theta_{c_1})
+|c_1+c_2|^2=|c_1|^2 + |c_2|^2 + 2 |c_1||c_2|\cos(\theta_{c_2} - \theta_{c_1})
 \end{equation}
-Compare to \ref{eqn:correlation}.
-The material is spread evenly among fragment configurations.
-Also, since it is homogeneous, the ratio of fragments per configuration is the same.
-Therefore the amount of material is proportional to the strength of the chaotic motion of the fragment, which is measured by the variance.
-The increase/decrease in amount of material therefore combines like the variance of a statistical variable.
-The cosine of the phase difference takes the place of the Pearson correlation coefficient.
+The two expressions are formally identical and it shouldn't come as a surprise.
+The more material we have, the more labels for the fragment configurations to choose from.
+$-1\leq\cos(\theta_{c_2} - \theta_{c_1})\leq+1$ takes the place of the Pearson correlation coefficient.
+Quantum interference can then be understood in terms of the combination of correlated/anti-correlated distributions.
+
+If the phase determines the correlation between two states, then it should only be meaningful when comparing two states.
+And in fact only differences in phases are physically distinguishable in quantum mechanics.
 A more meaningful name for the phase, then, would be \emph{correlation angle}.
-Also note how correlation for a single variable is a meaningless quantity: it is the correlation between two random variables that is significant.
-That is why in quantum mechanics phase differences have physical meaning, as it gives the correlation between two states, while absolute phases do not.
+
+If quantum states are uniform distributions over fragment configurations, we should see a hint of that when computing distributions over observables.
+And in fact we do.
+Suppose we have $Q\psi = q \psi$, an eigenstate of some observable.
+Then it is also an eigenstate of the infinitesimal transformation of the conjugate variable $1+\frac{Qdp}{\imath\hbar}$.
+Since it has a symmetry over $p$, the distribution is uniform on that parameter.
+
+If quantum particle states are really just unit amount of material uniformly distributed, we should be able to calculate the Shannon entropy associated with them.
+Using our dummy variables $I(\psi_i) = - \int \rho ln (\rho) dA dB = - \int 1 ln (1) dA dB = 0$.
+For a quantum mixed state $\tilde{\rho} = \sum \tilde{\rho}_i |\psi_i\rangle \langle \psi_i |$, we have $I(\tilde{\rho}) = - \sum \tilde{\rho}_i ln \tilde{\rho}_i - \tilde{\rho}_i I(\psi_i) = - \sum \tilde{\rho}_i ln \tilde{\rho}_i = -tr(\tilde{\rho} ln \tilde{\rho})$.
+And in fact this is the Von Neumann entropy used in quantum information, that implicitly assumes all states provide the same amount of entropy (i.e.
+zero).
+
+If two quantum particle states $\psi_1$ and $\psi_2$ are really just two states at equilibrium with the same entropy, we should always be able to find a deterministic and reversible evolution to connect the two.
+And in fact we can.
+Take the plane identified by $\psi_1$ and $\psi_2$.
+Find the infinitesimal rotation $1 + \frac{Od\alpha}{\imath\hbar}$ along that plane.
+$\psi_2 = e^{\frac{O\alpha}{\imath\hbar}} \psi_1$ is the unitary evolution of the other.
+Note that this is \emph{not} always possible in classical mechanics, as two states $\mathcal{c}_1$ and $\mathcal{c}_2$ may have different entropy.
+
+In other words, not only our simple model is consistent, but it has predictive power: it provides intuition that is well supported by these findings. It is not ``just a mathematical trick".
 
@@
 the group of transformations remains the same.
-As the material is homogeneous, the same amount of material needs to be spread over the same number of fragment configurations along each independent d.o.f.
-This means that, as we combine components, the strength of the chaotic motion has to be the same in all directions.
-This means that the increase/decrease of material and the change in correlation has to be the same across all d.o.f.
+While we have more fragment configurations to label, each d.o.f.
+is labeled independently.
+Therefore we have $n$ uniform distributions, each over a square of the same amount of material.
+This means that, as we combine components, the labels have to combine in the same way along the different d.o.f.: the increase/decrease of material and the change in correlation has to be the same across all d.o.f.
 
@@
 It is so because it is the state space for a decomposable material (i.e.
-an abelian group) where we can increase/decrease the amount and change the correlation of the internal stochastic motion (i.e.
+an abelian group) where we can increase/decrease the amount and change the correlation of the internal description (i.e.
 a set of transformations conveniently parametrized by a complex number).
@@
 The configuration of a fragment of the material is characterized by two random variables $A, B: \Omega \rightarrow \mathbbm{R}$ where $\Omega$ is the sample space for the configuration of the fragments.
-As the configuration of each fragment is physically indistinguishable, $A$ and $B$ can be chosen such that the distribution $\rho$ of material over such variables is uniform over a range of $\sqrt{a_m}$ for both variables.
+As the material is homogeneous, $A$ and $B$ can be chosen such that the distribution $\rho$ of material over such variables is uniform over a range of $\sqrt{a_m}$ for both variables.
 We have $a_m=\int \rho dA \wedge dB = \int dA \wedge dB$.
 Consider a second amount $\hat{a}_m$ of irreducible material in the same state.
-It also can be characterized by a uniform distribution over two variables $C, D: \Omega \rightarrow \mathbbm{R}$ over a range of $\sqrt{\hat{a}_m}$.
+Its fragment configurations can also be characterized by a uniform distribution over two variables $C, D: \Omega \rightarrow \mathbbm{R}$ over a range of $\sqrt{\hat{a}_m}$.
 As the two amounts of material are in the same state, $(C,D)$ is a function of $(A,B)$.
@@
 The transformation has to be linear: $C=aA + bB$ and $D=-bA + aB$.
-As the change in amount of material, the only physically distinguishable quantity associated with the fragments, is fully specified by $a$ and $b$, no other transformation is physically relevant.
+As the change in amount of material, the only physically distinguishable quantity associated with the fragment configurations, is fully specified by $a$ and $b$, no other transformation is physically relevant for the state.
 Let $\tau: \mathbbm{C} \rightarrow T$ represent such a transformation parametrized by $c=a+\imath b$.
@@
 They are only if $q^i$ and $\hat{q}^i$ can be expressed as a bijection of each other $\hat{q}^i(q^i)$: they are different coordinates of the same manifold.} As to a linear combination of states corresponds a linear combination of wave functions, $\Upsilon^q_{\hat{q}}$ is a linear operator.
-As converting the space again must give us the original wave function, $\Upsilon^{\hat{q}}_{q} \Upsilon^q_{\hat{q}}= 1$ is unitary.
+As converting the space again must give us the original wave function, $\Upsilon^{\hat{q}}_{q} \Upsilon^q_{\hat{q}}= I$ is unitary.
 We can now define a have linear operator $F_{[\hat{q}^i]}$ corresponding to state variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
@@
 $\Upsilon^{q^i}_{\hat{q}^i}$ is an isomorphism as $\Upsilon_{\hat{q}^i}$ and $\Upsilon_{q^i}$ are isomorphisms.
-$\Upsilon^{q^i}_{\hat{q}^i}$ is unitary as $\Upsilon^{\hat{q}^i}_{q^i} \circ \Upsilon^{q^i}_{\hat{q}^i} = (\Upsilon^{q^i}_{\hat{q}^i})^{-1} \circ \Upsilon^{q^i}_{\hat{q}^i} = 1$.
+$\Upsilon^{q^i}_{\hat{q}^i}$ is unitary as $\Upsilon^{\hat{q}^i}_{q^i} \circ \Upsilon^{q^i}_{\hat{q}^i} = (\Upsilon^{q^i}_{\hat{q}^i})^{-1} \circ \Upsilon^{q^i}_{\hat{q}^i} = I$.
 We can extend $\Upsilon^{q^i}_{\hat{q}^i}$ on all continuous functions and define $F_{[q^i]} \equiv \Upsilon^{\hat{q}^i}_{q^i} \circ F_{[\hat{q}^i]} \circ \Upsilon^{q^i}_{\hat{q}^i}$.

### 2016-12-05 2015CompleteDerivation/completeDerivation.tex
Making the rest of quantum consistent with the new language

@@
 
-In the case of a reducible material, we saw how a finite amount of it was described by a distribution over the state variables of infinitesimal parts.
+In the previous sections we saw how a finite amount of classical material was described by a distribution over the state variables of infinitesimal parts.
 For an irreducible material, we cannot assign states to the fragments, therefore we cannot properly talk about state variables.
@@
 
-Naturally, we cannot give a complete description for the configuration of a fragment, therefore we cannot give a joint distribution for all fragment variables at the same time.\footnote{This recovers the notion of complementarity introduced by Bohr.} We can, though, find a maximal set of fragment variables: one that provides the best description possible.
-While it should be clear that there cannot be a single unique set, we can show that one such set is provided by the fragment variables $q^i : \mathcal{Q} \rightarrow \mathbbm{R}$ that define the units required to describe the system.
+Naturally, we cannot give a complete description for the configuration of a fragment, therefore we cannot give a joint distribution for all fragment variables at the same time.\footnote{This recovers the notion of complementarity introduced by Bohr.} We can, though, find a maximal set of fragment variables: one that can be used to provide the best description possible.
+While there isn't a unique maximal set, we can show that one such set is provided by the fragment variables $q^i : \mathcal{Q} \rightarrow \mathbbm{R}$ that define the units required to describe the system.
 
@@
 (as this would fully specify said d.o.f.).
-Therefore the set of fragment variables $q^i$ already provide a maximal description for the fragments.
-This means that a state $\psi$ must tell us for each point in $\mathcal{Q}$ how much material is there and the correlation of the unstated part.
+Therefore the set of fragment variables $q^i$ already provides a maximal set of fragment variables.
+This means that a state $\psi$ must tell us for each point in $\mathcal{Q}$ the amount and correlation angle of the material.
 That is: the state will be identified by $\psi : \mathcal{Q} \rightarrow \mathbbm{C}$, the wave function over the $\mathcal{Q}$.
@@
 The square modulus at each point will give us the density of the material.
-As in the classical case, such density must be integrable to give us the total amount of material.
+As in the classical case, such density must be integrable to give us the total amount.
 Therefore the function $\psi(q) \in L^2(\mathcal{Q}, \mu)$ is square integrable.
@@
 As we have seen, changing the phase by a constant only affects the unstated part.
-Therefore unitary distributions whose differ by a total phase are not physically distinguishable and represent the same quantum particle state.
+Therefore normalized distributions whose differ by a total phase are not physically distinguishable and represent the same quantum particle state.
 
 \begin{prop}\label{prop:wavefuntion}
-	The state space $\Psi$ for an irreducible material is isomorphic to a subspace of the space of Lebesgue square integrable differentiable functions.
-That is $\Psi \cong \Psi(\mathcal{Q}, dp^i) \subseteq C^1(\mathcal{Q}) \cap L^2(\mathcal{Q}, \mu)$ under the isomorphism $\Upsilon_q : \Psi \leftrightarrow \Psi(\mathcal{Q}, dp^i)$ where $\Psi(\mathcal{Q}, dp^i)$ is the space of wave functions over $q^i$.
+	The state space $\Psi$ for an irreducible material is isomorphic to a subspace of the space of Lebesgue square integrable continuous functions.
+That is $\Psi \cong \Psi(\mathcal{Q}, dp^i) \subseteq C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu)$ under the isomorphism $\Upsilon_q : \Psi \leftrightarrow \Psi(\mathcal{Q}, dp^i)$ where $\Psi(\mathcal{Q}, dp^i)$ is the space of wave functions over $q^i$.
 The state space for a quantum particle is isomorphic to the projective space $\mathsf{P}(\Psi)$.
@@
 \begin{justification}
-	We claim the manifold $\mathcal{Q}$ that defines the units to describe a fragment of an irreducible material also provides a maximal description for the fragment configuration, that is: no other fragment variable can be added.
+	We claim the manifold $\mathcal{Q}$ that defines the units to describe a fragment of an irreducible material also provides a maximal set of fragment variable: no other fragment variable can be added.
 Let $\mathcal{Q}$ be the manifold that defines the units.
 Then for each state $\psi \in \Psi$ there exists a distribution $\rho_\psi(q^i)$ that represents the quantity of material at each point.
-Suppose this distribution didn't represent a maximal description for the fragments.
+Suppose this distribution could be further refined.
 Then another independent variable $k$ would exist such that for each state $\psi \in \Psi$ we would have a distribution $\rho_\psi(q^i, k)$.
@@
 	We claim $\Psi \cong \Psi(\mathcal{Q}, dp^i) \subseteq C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu, \mathbbm{C})$ as a vector space.
-The maximal physical description of a set of fragments is given by a point in $\mathcal{Q}$ parametrized by $q^i$ and by a transformation in $T$ parametrized by $c \in \mathbbm{C}$.
-For each $\psi \in \Psi$ there must exist a wave function $\psi(q^i) : \mathcal{Q} \rightarrow \mathbbm{C}$ that provides the maximal physical description for the fragments.
+The maximal physical description of a fragment configuration is given by a point in $\mathcal{Q}$ parametrized by $q^i$ and by a transformation $\tau \in T$ parametrized by $c \in \mathbbm{C}$.
+For each $\psi \in \Psi$ there must exist a wave function $\psi(q^i) : \mathcal{Q} \rightarrow \mathbbm{C}$ that provides the maximal physical description for the state.
 $\rho_\psi(q^i) = |\psi(q^i)|^2$ as the modulus of $\psi(q^i)$ is the increase/decrease from the unit amount of material.
 $\rho_\psi(q^i) \in C(\mathcal{Q}, \mathbbm{R})$ by \ref{continuous_map}.
+$\rho_\psi(q^i) \in C^1(\mathcal{Q}, \mathbbm{R})$ by the same arguments in the justification for \ref{prop:differentiable_manifold}.
 Therefore $\psi(q^i) \in C(\mathcal{Q}, \mathbbm{C})$.
@@
 $\Upsilon_{q^i}$ is a bijection between $\Psi$ and $\Psi(\mathcal{Q}, dp^i)\equiv\Upsilon_{q^i}(\Psi)$.
-Let $\psi=\psi_1+\psi_2$, then $\Upsilon_{q^i}(\psi)=\Upsilon_{q^i}(\psi_1)+\Upsilon_{q^i}(\psi_2)$ as combining the distributions means combining the statistical description of the fragments.
-Let $\psi_1=\tau(c)\psi_2$, then $\Upsilon_{q^i}(\psi_1)=c \Upsilon_{q^i}(\psi_2)$ as $\tau(c)$ transforms the fragment configuration by the factor $c$.
+Let $\psi=\psi_1+\psi_2$, then $\Upsilon_{q^i}(\psi)=\Upsilon_{q^i}(\psi_1)+\Upsilon_{q^i}(\psi_2)$ as combining the distributions means combining the statistical description of the fragment configurations.
+Let $\psi_1=\tau(c)\psi_2$, then $\Upsilon_{q^i}(\psi_1)=c \Upsilon_{q^i}(\psi_2)$ as $\tau(c)$ transforms the fragment configuration description by the factor $c$.
 $\Upsilon_{q^i}$ is a homomorphism as a vector space.
@@
 $\tau(e^{\imath \theta}) \psi$ with $\theta \in \mathbbm{R}$ is not physically distinguishable from $\psi$ as phases are only physically distinguishable through addition.
-$\tau(\sqrt{\rho}) \psi$ with $\sqrt{\rho} \in \mathbbm{R}^+ \backslash \{1\}$ cannot represent a quantum particle as it as a different amount of material.
+$\tau(\sqrt{\rho}) \psi$ with $\sqrt{\rho} \in \mathbbm{R}^+ \backslash \{1\}$ cannot represent a quantum particle as it has a different amount of material.
 There exists only one distinguishable quantum particle state in the equivalence class $\psi \sim \tau(c) \psi$ with $c \in \mathbbm{C}$.
@@
 In both the reducible and irreducible case, the material is distributed over $\mathcal{Q}$.
-The $q^i$ do not label the states of the material itself, but the configurations for the infinitesimal parts.
+The $q^i$ do not label the states of a finite amount of material, but the configurations for the infinitesimal parts.
 For a classical material, the infinitesimal parts can be given a state, have their own topological space and a set of local state variables $(q^i, p_i)$ to identify them.
@@
 For an irreducible material, the infinitesimal parts cannot be given a state, cannot be fully identified by a set of variables.
-We have to use other mathematical tools, such as vector spaces.
+We have to use other mathematical tools, typically vector spaces.
 As we will often use linear combination of states, we will write, for brevity, $c\psi$ instead of $\tau(c)\psi$.
-That is: each complex numbers still represents a state transformation.
+That is: each complex number still represents a state transformation.
 
@@
 The idea is that, given two states $\psi_1$ and $\psi_2$, we should be able to tell how much of $\psi_1$ is prepared in the same way of $\psi_2$.
-That is $\psi_1 = \alpha \psi_2 + \psi_3$ is the combination of the component that is prepared like $\psi_2$ (e.q.
+That is $\psi_1 = \alpha \psi_2 + \psi_3$ is the combination of the component that is prepared like $\psi_2$ (e.g.
 same distributions in position and momentum) and a component that is prepared completely differently.
 Let $\mathsf{P}_{\psi_2} : \Psi \rightarrow \Psi$ be the operator that gives us the component $\mathsf{P}_{\psi_2} (\psi_1) \equiv \alpha \psi_2$ prepared like $\psi_2$.
-This operator is linear in $\psi_1$: the component of the combined system that is prepared like $\psi_2$ is the sum of the components of the parts that are prepared like $\psi_2$.
+This operator is linear in $\psi_1$: the component of the combined system that is prepared like $\psi_2$ is the sum of the subcomponents that are prepared like $\psi_2$.
 It also gives us the same result if applied once or twice: once we found the component that is prepared like $\psi_2$, it remains the same.
-$\mathsf{P}_{\psi_2}$ is therefore a projection: a linear operator such that $\mathsf{P}^2_{\psi_2}= \mathsf{P}_{\psi_2}$.
+$\mathsf{P}_{\psi_2}$ is therefore a projection: a linear operator such that $\mathsf{P}_{\psi_2}(\mathsf{P}_{\psi_2} (\psi_1))= \mathsf{P}_{\psi_2} (\psi_1)$.
 
@@
 The whole state $\phi$ is prepared as $\psi$.
-We can obtain $\phi$ from $\psi$ by a transformation in the unstated part.
+We can obtain $\phi$ from $\psi$ by a relabeling of the fragment configurations.
 $\phi = c \psi$ for some $c \in \mathbbm{C}$.
@@
 Also note that the observations made after \ref{prop:differentiable_manifold} for the  classical case are still valid.
-$\Psi$ cannot be a complete metric space (for the norm induced by the inner product) as it does not include discontinuous functions.
+$\Psi$ cannot be a complete metric space for the norm induced by the inner product as it does not include discontinuous functions.
 Therefore it is only a pre-Hilbert space.
@@
 
-As such, we need to make sure that the rest of our definitions and justifications do not require a complete metric space to be valid or they would not be physically meaningful.
+Therefore we need to make sure that the rest of our definitions and justifications do not require a complete metric space to be valid or they would not be physically meaningful.
 In particular, we don't want to use the usual link $\psi(q) = \langle q | \psi \rangle$ between the wave function and the inner product as the material cannot be prepared at a specific value when $q$ is a continuous variable.
-Similarly, we cannot show that to each state variable is associated an hermitian operator simply by constructing it from its eigenstates.
+Similarly, we cannot show that to each state variable is associated an hermitian operator simply by constructing it from its eigenstates as they may not represent physical states.
 
@@
 
-While $\mathcal{Q}$ is enough to define the system of units, it cannot be enough to define all the variables over which we can define the distributions of irreducible material.
+While $\mathcal{Q}$ is enough to define the system of units, it is not enough to define all the fragment variables over which we can define the distributions of irreducible material.
 The conjugate variables $p_i=\hbar k_i$ cannot be expressed as a function of $q^i$, which reiterates the idea that we cannot specify all variables for each fragment.
-Suppose there is another set of state variables that provide another maximal description for the fragments.
+Suppose there is another maximal set of fragment variables.
 Then those variables will form a topological space $\hat{\mathcal{Q}}$ such that each $\psi$ can be expressed as a wave function $\psi(\hat{q}) :\hat{\mathcal{Q}} \rightarrow \mathbbm{C}$ over $\hat{\mathcal{Q}}$.
 
-As each state $\psi$ can be identified by a wave function on either space, there must be an operator $\Upsilon^q_{\hat{q}}(\psi(q)) = \psi(\hat{q})$ that converts the wave function on one space to the other.\footnote{Note that a neighbourhood of $q^i$ is not necesserily a neighbourood of $\hat{q}^i$.
+As each state $\psi$ can be identified by a wave function on either space, there must be an operator $\Upsilon^q_{\hat{q}}(\psi(q)) = \psi(\hat{q})$ that converts the wave function on one space to the other.\footnote{Note that a neighbourhood of $q^i$ is not necessarily a neighbourood of $\hat{q}^i$.
 They are only if $q^i$ and $\hat{q}^i$ can be expressed as a bijection of each other $\hat{q}^i(q^i)$: they are different coordinates of the same manifold.} As to a linear combination of states corresponds a linear combination of wave functions, $\Upsilon^q_{\hat{q}}$ is a linear operator.
-As converting the space again must give us the original wave function, $\Upsilon^{\hat{q}}_{q} \Upsilon^q_{\hat{q}}= I$ is unitary.
-We can now define a have linear operator $F_{[\hat{q}^i]}$ corresponding to state variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
-Using $\Upsilon^{q}_{\hat{q}}$ we can express $F_{[q^i]} = \Upsilon^{\hat{q}}_{q} F_{[\hat{q}^i]} \Upsilon^{q}_{\hat{q}}$ as an operator over complex functions of $\mathcal{Q}$.
+As converting the space again must give us the original wave function, $\Upsilon^{\hat{q}}_{q} \circ \Upsilon^q_{\hat{q}}= I$ is unitary.
+We can now define a linear operator $F_{[\hat{q}^i]}$ corresponding to state variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
+Using $\Upsilon^{q}_{\hat{q}}$ we can express $F_{[q^i]} = \Upsilon^{\hat{q}}_{q} \circ f(\hat{q}^i) \Upsilon^{q}_{\hat{q}}$ as an operator over complex functions of $\mathcal{Q}$.
 It will be still be linear and hermitian.
@@
 Let $\hat{q}^i$ be a set of fragment variables such that $f=f(\hat{q}^i)$.
-If $\hat{q}^i$ does not provide a maximal description for the fragments then there exists another set of fragment variables that refine the description.
+If $\hat{q}^i$ does not provide a maximal set of fragment variables then there exists other fragment variables that refine the description.
 Extend $\hat{q}^i$ with such variables.
 If still not maximal, continue extending.
-As the number of possibilities charted by $\hat{q}^i$ cannot exceed the number of possibilities charted by $q^i$, at some point we'll reach a maximal description.
+As the number of possibilities charted by $\hat{q}^i$ cannot exceed the number of possibilities charted by $q^i$, at some point we'll reach a maximal set.
 We call $\hat{\mathcal{Q}}$ the topological space charted by the fragment variables $\hat{q}^i$.
@@
 Let $\psi \in \Psi$ be a state.
-$\hat{q}^i$ provides a maximal description for the fragments.
-There must exist a wave function $\psi(\hat{q}^i) : \mathbbm{R}^n \rightarrow \mathbbm{C}$ associated with $\psi$ such that it gives the fragment configuration at a particular value for $\hat{q}^i$.
-Such wave function provides a full description for the state.
+$\hat{q}^i$ provides a maximal set of fragment variables.
+There must exist a wave function $\psi(\hat{q}^i) : \mathbbm{R}^n \rightarrow \mathbbm{C}$ associated with $\psi$ such that it gives the maximal description for the state.
 Let $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \rightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ the map that associates a state with its wave function over $\hat{\mathcal{Q}}$.
@@
 	We claim that for each fragment variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ there exists a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ such that $\int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]}( \psi(\hat{q}^i)) d^n \hat{q}$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
-The expectation value of $f$ multiplied by the amount of material is given by $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q}$ where $\hat{q}^i$ provide a maximal description such that $f=f(\hat{q})$.
+The expectation value of $f$ multiplied by the amount of material is given by $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q}$ where $\hat{q}^i$ provide a maximal set of fragment variables such that $f=f(\hat{q})$.
 We have $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q} = \int \psi^\dagger(\hat{q}^i) f(\hat{q}^i) \psi(\hat{q}^i) d^{n}\hat{q}$.
@@
 Let $f$ be a fragment variable.
-There exist a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ where $\hat{\mathcal{Q}}$ provides a maximal description.
-As both $\hat{\mathcal{Q}}$ and $\mathcal{Q}$ provide a maximal description, we can define $\Upsilon^{q^i}_{\hat{q}^i} \equiv \Upsilon_{\hat{q}^i} \circ \Upsilon_{q^i}^{-1}$ which transforms a wave function over $\mathcal{Q}$ into a wave function over $\hat{\mathcal{Q}}$.
+There exist a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ where $\hat{\mathcal{Q}}$ provides a maximal set of fragment variables.
+As both $\hat{\mathcal{Q}}$ and $\mathcal{Q}$ provide a maximal set, we can define $\Upsilon^{q^i}_{\hat{q}^i} \equiv \Upsilon_{\hat{q}^i} \circ \Upsilon_{q^i}^{-1}$ which transforms a wave function over $\mathcal{Q}$ into a wave function over $\hat{\mathcal{Q}}$.
 $\Upsilon^{q^i}_{\hat{q}^i}$ is an isomorphism as $\Upsilon_{\hat{q}^i}$ and $\Upsilon_{q^i}$ are isomorphisms.
@@
 We know that the operators have to be covariant: under a change of units $\hat{q}^j = \hat{q}^j(q^i)$ we must have $K_{\hat{q}^j[q^i]} = \partial_{\hat{q}^j} q^i K_{q^i [q^i]}$.
-We are looking for a linear operator that obeys the chain rule, which turns out to be the derivation by the corresponding $q^i$.
-As it has to be self-adjoint, it will be of the form $K_{i [q^i]} = a \imath \partial_{i}$ where $a$ is a real number.
-We choose $a$ to be negative by convention and to be unitary so that $k_i$ is expressed in units of inverse $q$ as in the classical case.
+We are looking for linear operators that obey the chain rule, which turn out to be the derivation by the corresponding $q^i$.
+They have to be self-adjoint, so they will be of the form $K_{i [q^i]} = a \imath \partial_{i}$ where $a$ is a real number.
+We choose $a$ to be negative by convention and to be unitary so that $k_i$ is expressed in units of inverse $q^i$ as in the classical case.
 Therefore we have $K_{i [q^i]} = -\imath \partial_i$ and $P_{i [q^i]} = -\imath \hbar \partial_i$.
@@
 If we assume $\mathcal{Q}$ to be $\mathbbm{R}^n$, the functional $\Upsilon^{q^i}_{p_i} : \Psi(\mathcal{Q}, dq^i) \leftrightarrow \Psi(\mathcal{P}, dq^i)$ that converts a wave function in position to a wave function in momentum is the Fourier transform.
-As the fragment variables for any power of $q^i$ and $p^i$ must exist, the wave functions in both expressions are infinitely differentiable: the space of the wave functions $\Psi(\mathcal{Q}, dq^i) = S(\mathcal{Q}, q^i)$ is the Schwartz space.
+As the fragment variables for any power of $q^i$ and $p_i$ must exist, the wave functions in both expressions are infinitely differentiable: the space of the wave functions $\Psi(\mathcal{Q}, dq^i) = S(\mathcal{Q}, q^i)$ is the Schwartz space.
 
@@
 	The problem of incompatibility of measurements and deterministic evolution at all levels still remains.
-The unstated part of an irreducible system does not change the picture: it is completely undetermined and cannot be used to pass information from one system to another.
+The unstated part of multiple irreducible particles may be interacting, but it cannot be used to pass information from one subsystem to another without making itself distinguishable.
 
 	The inability to define time using a deterministic and reversible evolution is still a problem.
-The unstated part cannot give us a measurable quantity that changes in time as that would be a description of the unstated part.
+As the system is homogeneous and irreducible, the unstated part cannot give us a measurable quantity that changes in time as that would provide a description of the unstated part.
 
 	There is one additional problem, though, not present in the classical case.
-Quantum particle states, as we have defined them, are states at equilibrium: they all provide the same level of description for the fragments.
+Quantum particle states, as we have defined them, are states at equilibrium: they all provide the same level of description for the fragments, they carry the same information entropy.
 In fact if the internal dynamics were out of equilibrium, the material would not be homogeneous which would contradict the assumption.
-Deterministic and reversible evolution (which we'll see corresponds to unitary evolution) is necessarily quasi static: at each moment a state must be well defined, at each moment the system is in equilibrium.
+Deterministic and reversible evolution (which we'll see corresponds to unitary evolution) is necessarily quasi static: at each moment a state must be well defined, at each moment the system is in equilibrium, the system changes slowly.
 Therefore quantum states cannot and should not be expected to provide a description under non deterministic evolution.

### 2016-12-06 2015CompleteDerivation/completeDerivation.tex
Comments from Christine on complex vector space

@@
 Suppose we assigned a numeric identifier to each particle of the material.
-This identifier would be just a dummy label we assign, not a physically quantity.
+This identifier would be just a dummy label we assign, not a physical quantity.
 As the classical material is homogeneous, permutations between classical particles within the same state have no effect.
@@
 This is also the only possible transformation that can link the two uniform random variables $A$ and $B$.
-Therefore increase/decrease of material is the only operation that is going to affect the labeling of the fragments that is captured by the state.
+Therefore increase/decrease of material is the only operation that is captured by the state among those that affect the labeling of the fragments.
 In other words: the classical state does not care about the permutation of the fragments, but it cares how many fragments are there.
@@
 The two expressions are formally identical and it shouldn't come as a surprise.
-The more material we have, the more labels for the fragment configurations to choose from.
+The more material we have, the more labels for the fragment configurations to choose from, the bigger the spread of the random variables that assign the labels.
 $-1\leq\cos(\theta_{c_2} - \theta_{c_1})\leq+1$ takes the place of the Pearson correlation coefficient.
@@
 Then it is also an eigenstate of the infinitesimal transformation of the conjugate variable $1+\frac{Qdp}{\imath\hbar}$.
-Since it has a symmetry over $p$, the distribution is uniform on that parameter.
-
-If quantum particle states are really just unit amount of material uniformly distributed, we should be able to calculate the Shannon entropy associated with them.
-Using our dummy variables $I(\psi_i) = - \int \rho ln (\rho) dA dB = - \int 1 ln (1) dA dB = 0$.
-For a quantum mixed state $\tilde{\rho} = \sum \tilde{\rho}_i |\psi_i\rangle \langle \psi_i |$, we have $I(\tilde{\rho}) = - \sum \tilde{\rho}_i ln \tilde{\rho}_i - \tilde{\rho}_i I(\psi_i) = - \sum \tilde{\rho}_i ln \tilde{\rho}_i = -tr(\tilde{\rho} ln \tilde{\rho})$.
+Since it has a symmetry over $p$, the distribution has to be equal at each value of $p$: the distribution is uniform.
+
+If quantum particle states are really just unit amounts of material uniformly distributed, we should be able to calculate the Shannon entropy associated with them.
+Using our dummy variables, $I(\psi) = - \int \rho \ln (\rho) dA dB = - \int 1 ln (1) dA dB = 0$.
+For a quantum mixed state $\tilde{\rho} = \sum \tilde{\rho}_i |\psi_i\rangle \langle \psi_i |$, we have $I(\tilde{\rho}) = - \sum \tilde{\rho}_i \ln \tilde{\rho}_i - \tilde{\rho}_i I(\psi_i) = - \sum \tilde{\rho}_i \ln \tilde{\rho}_i = -tr(\tilde{\rho} \ln \tilde{\rho})$.
 And in fact this is the Von Neumann entropy used in quantum information, that implicitly assumes all states provide the same amount of entropy (i.e.
@@
 
-In other words, not only our simple model is consistent, but it has predictive power: it provides intuition that is well supported by these findings. It is not ``just a mathematical trick".
+In other words, our simple model is not only consistent with quantum mechanics but it has predictive power: it provides intuition that is well supported by these findings.
+It is not ``just a mathematical trick".
 

### 2016-12-07 2015CompleteDerivation/completeDerivation.tex
Quantum should be ready for review

@@
 
-Section IX - Preliminary work
+Section IX - Mostly done. Needs review.
 
@@
 Yet, we can still create a self-adjoint linear operator, with the same spectra as the one associated with $q$, just with the eigenvalues switched.
-It is not clear what plays the role of the differentiable structure of $T^*\mathcal{Q}$ in the quantum case.
-
-The situation is worsened by the fact that the topology of $\mathcal{P}$, the space charted by conjugate momentum, depends globally on the topology of $\mathcal{Q}$.
+The space of physically meaningful operator, then, should be further constrained.
+A reasonable requirement would be that they preserve the differentiable/integrable structure of the Schwartz space $S(\mathcal{Q}, q^i)$ as classical state variables preserved the differentiable structure of $T^*\mathcal{Q}$.
+Unfortunately, this is not as easy to justify.
+
+The topology of $\mathcal{P}$, the space charted by conjugate momentum, depends globally on the topology of $\mathcal{Q}$.
 For example, if $\mathcal{Q}$ is an n-dimensional torus, there is no infinity and $\mathcal{P}$ has a discrete topology.
 That is why, in the end, we restricted ourselves to $\mathbbm{R}^n$.
-Moreover, we obtained the space of functions not because we required them to fall off at infinity, but because we required the ability to express polynomials of $q^i$ and $p_i$.
-It is not immediately clear how this generalizes over a manifold.
+Moreover, we characterized the space of functions not because we required them to fall off at infinity, but because we required the ability to express polynomials of $q^i$ and $p_i$.
+It is not immediately clear how this generalizes over a manifold: further work is needed.
 
@@
 Note that formally there is a strong analogy with the reducible material case.
-In classical mechanics the state space structure was captured by a symplectic manifold.
+In classical mechanics the state space structure was captured by a distribution over a symplectic manifold.
 Deterministic and reversible evolution was a transformation that preserved that structure, a symplectomorphism.
@@
 This is actually consistent with our assumptions.
-For example, for a system to reach its ground state it has to be able to radiate energy: it cannot isolated.
+For example, for a system to reach its ground state it has to be able to radiate energy: it cannot be isolated.
 Therefore this dynamic cannot be described under the deterministic and reversible assumption.
@@
 Other considerations, outside of the scope of this work, can be used to recover this behavior (e.g.
-thermodynamics or decoherence).
+thermodynamics, decoherence, ...).
+
+Also note that $H$ operates on $\Psi$ which, in the case of $\mathcal{Q}=\mathbbm{R}^n$, corresponds to the space $S(\mathcal{Q}, q^i)$ of Schwartz functions.
+$H$ therefore preserve smoothness and integrability.
+This, again, reminds us that the space of all self-adjoint operators is far too vast, and does not map well to physical meaningful quantities.
 
@@
 As for the composite state evolution in the classical case, $\bar{\psi}$ remains unchanged by the evolution as it transports the distribution along $s$.
-Therefore $\mathcal{H} \hat{\psi} = \mathcal{h} \hat{\psi}$ is an eigenfunction of $\mathcal{H}$.
-As in the classical case, we use $\mathcal{H}$ to be zero to identify the state space: the temporal d.o.f.
+Therefore $\mathcal{H} \bar{\psi} = \mathcal{h} \bar{\psi}$ is an eigenfunction of $\mathcal{H}$.
+As in the classical case, we use $\mathcal{H}=0$ to identify the state space: the temporal d.o.f.
 is not independent and therefore there is a constraint between the variables.
@@
 
-Note how the covariant derivative is, up to a constant, kinetic momentum.
+Note how the gauge covariant derivative is, up to a constant, kinetic momentum.
 This motivates why it is such an important operator.
@@
 
-It is clear that it has to be absorbed as it has no way to escape.
+It is clear that it has to be absorbed as it is within the aperture of the whole array.
 It is also clear that it must be absorbed whole: as it is an irreducible system, the detector cannot partially interact with it, it cannot half absorb it.
@@
 Maybe it depends on the state of the detector, maybe it depends on the unstated part of the particle.
-The point is that a single take does not tell us much about the incoming particle states.
+The point is that a single measurement take does not tell us much about the incoming particle states.
 
@@
 We do not make any claim on what the final state of the particle should be: in many cases, like in a photomultiplier tube, the particle is gone and there is no state.
-We simply have one incoming state for the particle, prepared over and over exactly the same, and the outgoing state for the measurement device, which is a classical statistical distribution over the different takes.} The question is: how is the statistical distribution coming out of the detector related to the material distribution coming in with the particle?
+We simply have one incoming state for the particle, prepared over and over exactly the same, and the outgoing state for the measurement device, which is a classical statistical distribution over the different measurement takes.} The question is: how is the statistical distribution coming out of the detector related to the material distribution coming in with the particle?
 
@@
 The final statistical distribution of the detector must match the incoming material distribution of the particle.
-That is if $N$ is the total number of takes and $N_A$ is the number time the detector with aperture in the $A$ region captured the particle, the ration $\frac{N_A}{N}$ will be approximately the amount of material within the aperture $\int_A \psi^\dagger(q^i) \psi(q^i) dq^i$.
+That is if $N$ is the total number of takes and $N_A$ is the number of times the detector with aperture in the $A$ region captured the particle, the ratio $\frac{N_A}{N}$ will be approximately the amount of material within the aperture $\int_A \psi^\dagger(q^i) \psi(q^i) dq^i$.
 If we assume an infinite amount of takes, we expect them to coincide: $\int_A \psi^\dagger(q^i) \psi(q^i) dq^i = \lim\limits_{N\rightarrow\infty} \frac{N_A}{N}$.
@@
 We therefore have arrived to all the major underpinning of quantum mechanics and see how they stem from assumption \ref{ass:irreducibility} of irreducibility.
-We have somewhat a closure, probably not satisfactory to many, for some of the open problems in quantum mechanics.
-Not because we solved them, we haven't at all, but because we have good reason to have them.
-The desire to know ``what is really going on" is hindered by assuming that no description can be given below the level of a particle.
-
-On the upside, irreducibility is what it is: an assumption.
-As with infinitesimal reducibility, it is a reasonable initial assumption assumption.
-There is hope that, if we abandon it, we can have a more satisfying and complete description.
-What should be clear, though, that abandoning irreducibility may entail much much more than just finding another equation of motion.
+This gives us a more satisfactory account on what quantum mechanics describes, though it does not claim to solve many of the open problems.
+It does, though, give good reason why we have those problems and how the desire to know ``what is really going on" is hindered by the very assumption that the quantum description is based on.
 
@@
 In the classical case $\mathcal{e}(i)$ form a basis of a real vector space, where $i$ is the state vector for an infinitesimal part, and $a_i$ a real valued coefficient that represents the amount of particles in each state, which describes how the parts are combined.
+
+\section{Discussion and further work}
+
+%Time arrow
+
+%Particle excitations
+
+%Mixed state as fundamental (law of evolution defined on the mixed state)
+
+%Black hole as irreducible object
+
+\section{Conclusion}
+
+%We also have a more unified picture of fundamental physics which, after all ideas are digested, helps the intuition to gain insight that span across the different field.
+It also illustrate a new approach in fundamental physics: one where we better clarify our starting physical assumption and definition, where physics comes first, and they are later captured with mathematical definitions.
 

### 2016-12-09 2015CompleteDerivation/completeDerivation.tex
Comments from Christine

@@
 
-	We claim the set of physically distinguishable transformations $T$ that change the unstated part of an irreducible material is isomorphic to $\mathbbm{C}$.
+	We claim the set of physically distinguishable transformations $T$ that relabel the fragment configurations is isomorphic to $\mathbbm{C}$.
 Consider an amount $a_m \in \mathbbm{R}$ of irreducible material characterized by one d.o.f.
@@
 We have $a_m=\int \rho dA \wedge dB = \int dA \wedge dB$.
-Consider a second amount $\hat{a}_m$ of irreducible material in the same state.
+Consider a second amount $\hat{a}_m$ of irreducible material where we changed the labeling of the fragment configurations.
 Its fragment configurations can also be characterized by a uniform distribution over two variables $C, D: \Omega \rightarrow \mathbbm{R}$ over a range of $\sqrt{\hat{a}_m}$.
-As the two amounts of material are in the same state, $(C,D)$ is a function of $(A,B)$.
+As the two amounts of material differ just by the labeling of fragment configuration, $(C,D)$ is a function of $(A,B)$.
 We have $\hat{a}_m = \int dC \wedge dD = \int |J| dA \wedge dB$ where $|J|$ is the Jacobian determinant.
@@
 $|c|^2 = |J| = (a^2 + b^2) = \hat{a}_m / a_m$ is the ratio of the two amounts of materials and is physically distinguishable.
-Define on $T$ an addition $+: T \times T \rightarrow T$ and a multiplication $*: T \times T \rightarrow T$ such that $\tau(a) + \tau(b) = \tau(a+b)$ and $\tau(a) * \tau(b) = \tau(a*b)$, $a,b \in \mathbbm{C}$, the sum and product of the transformation is equal to the sum and product of their respective factors.
+Define on $T$ an addition $+: T \times T \rightarrow T$ and a multiplication $*: T \times T \rightarrow T$ such that $\tau(z_1) + \tau(z_2) = \tau(z_1+z_2)$ and $\tau(z_1) * \tau(z_2) = \tau(z_1*z_2)$, $z_1,z_2 \in \mathbbm{C}$, the sum and product of the transformation is equal to the sum and product of their respective factors.
 The phase $arg(c)$ under addition is physically distinguishable as it can affect the modulus of the result.
@@
 This means that a state $\psi$ must tell us for each point in $\mathcal{Q}$ the amount and correlation angle of the material.
-That is: the state will be identified by $\psi : \mathcal{Q} \rightarrow \mathbbm{C}$, the wave function over the $\mathcal{Q}$.
+That is: the state will be identified by $\psi : \mathcal{Q} \rightarrow \mathbbm{C}$, the wave function over $\mathcal{Q}$.
 
@@
 As we have seen, changing the phase by a constant only affects the unstated part.
-Therefore normalized distributions whose differ by a total phase are not physically distinguishable and represent the same quantum particle state.
+Therefore normalized distributions that differ by a total phase are not physically distinguishable and represent the same quantum particle state.
 
@@
 	The state space $\Psi$ for an irreducible material is isomorphic to a subspace of the space of Lebesgue square integrable continuous functions.
-That is $\Psi \cong \Psi(\mathcal{Q}, dp^i) \subseteq C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu)$ under the isomorphism $\Upsilon_q : \Psi \leftrightarrow \Psi(\mathcal{Q}, dp^i)$ where $\Psi(\mathcal{Q}, dp^i)$ is the space of wave functions over $q^i$.
+That is $\Psi \cong \Psi(\mathcal{Q}, dq^i) \subseteq C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu, \mathbbm{C})$ under the isomorphism $\Upsilon_q : \Psi \leftrightarrow \Psi(\mathcal{Q}, dq^i)$ where $\Psi(\mathcal{Q}, dq^i)$ is the space of wave functions over $q^i$.
 The state space for a quantum particle is isomorphic to the projective space $\mathsf{P}(\Psi)$.
@@
 \begin{justification}
-	We claim the manifold $\mathcal{Q}$ that defines the units to describe a fragment of an irreducible material also provides a maximal set of fragment variable: no other fragment variable can be added.
+	We claim the manifold $\mathcal{Q}$ that defines the units to describe a fragment of an irreducible material also provides a maximal set of fragment variables: no other fragment variable can be added.
 Let $\mathcal{Q}$ be the manifold that defines the units.
@@
 $k$ cannot lie in the same d.o.f.
-of one of the $q^i$ or it would provide a complete description for the fragment.
+as of one of the $q^i$ or it would provide a complete description for the fragment.
 This is a contradiction as the material is irreducible.
-$\mathcal{Q}$ provides a maximal description.
+$\mathcal{Q}$ provides a maximal set of fragment variables.
 
@@
 For each $\psi \in \Psi$ there must exist a wave function $\psi(q^i) : \mathcal{Q} \rightarrow \mathbbm{C}$ that provides the maximal physical description for the state.
-$\rho_\psi(q^i) = |\psi(q^i)|^2$ as the modulus of $\psi(q^i)$ is the increase/decrease from the unit amount of material.
+$\rho_\psi(q^i) = |\psi(q^i)|^2$ as the square modulus of $\psi(q^i)$ is the increase/decrease from the unit amount of material.
 $\rho_\psi(q^i) \in C(\mathcal{Q}, \mathbbm{R})$ by \ref{continuous_map}.
@@
 Let $\Upsilon_{q^i} : \Psi \rightarrow C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu, \mathbbm{C})$ be the function that given a state returns its wave function.
-As the wave function captures the maximal description of the system, two states must represent different distributions.
+As the wave function captures the maximal description of the system, two different states must represent different distributions.
 $\forall \psi_1, \psi_2 \in \Psi, \psi_1 \neq \psi_2 \implies \Upsilon_{q^i}(\psi_1) \neq \Upsilon_{q^i}(\psi_2)$.
 $\Upsilon_{q^i}$ is injective.
-$\Upsilon_{q^i}$ is a bijection between $\Psi$ and $\Psi(\mathcal{Q}, dp^i)\equiv\Upsilon_{q^i}(\Psi)$.
+$\Upsilon_{q^i}$ is a bijection between $\Psi$ and $\Psi(\mathcal{Q}, dq^i)\equiv\Upsilon_{q^i}(\Psi)$.
 Let $\psi=\psi_1+\psi_2$, then $\Upsilon_{q^i}(\psi)=\Upsilon_{q^i}(\psi_1)+\Upsilon_{q^i}(\psi_2)$ as combining the distributions means combining the statistical description of the fragment configurations.
 Let $\psi_1=\tau(c)\psi_2$, then $\Upsilon_{q^i}(\psi_1)=c \Upsilon_{q^i}(\psi_2)$ as $\tau(c)$ transforms the fragment configuration description by the factor $c$.
-$\Upsilon_{q^i}$ is a homomorphism as a vector space.
-$\Psi$ is isomorphic to $\Psi(\mathcal{Q}, dp^i) \subseteq C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu, \mathbbm{C})$.
+$\Upsilon_{q^i}$ is a homomorphism between vector spaces.
+$\Psi$ is isomorphic to $\Psi(\mathcal{Q}, dq^i) \subseteq C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu, \mathbbm{C})$.
 
@@
 The state with no amount of material does not correspond to any particle state.
-The quotient space $\mathsf{P}(\Psi) \equiv (\Psi \backslash \{0\}) / \sim$ is isomorphic to the state of quantum particles.
+The quotient space $\mathsf{P}(\Psi) \equiv (\Psi \backslash \{0\}) / \sim$ is isomorphic to the state space for quantum particles.
 \end{justification}
@@
 We have to use other mathematical tools, typically vector spaces.
-As we will often use linear combination of states, we will write, for brevity, $c\psi$ instead of $\tau(c)\psi$.
+As we will often use linear combinations of states, we will write, for brevity, $c\psi$ instead of $\tau(c)\psi$.
 That is: each complex number still represents a state transformation.
@@
 We need to be able to express distributions and expectations for different quantities: this will give us self-adjoint linear operators.
-We need to identify the conjugate quantities $k_i$ and $p_i$.
+We then need to identify the self-adjoint linear operators for the quantities $k_i$ and $p_i$ and to be able express the distribution of material in terms of them.
 Let's start with the first.
 
-The idea is that, given two states $\psi_1$ and $\psi_2$, we should be able to tell how much of $\psi_1$ is prepared in the same way of $\psi_2$.
+The idea is that, given two states $\psi_1$ and $\psi_2$, we should be able to tell how much of $\psi_1$ is prepared in the same way as of $\psi_2$.
 That is $\psi_1 = \alpha \psi_2 + \psi_3$ is the combination of the component that is prepared like $\psi_2$ (e.g.
@@
 
-This projection has other useful properties.
 If $\psi_1$ is already prepared like $\psi_2$, then its projection on $\psi_2$ will be equal to $\psi_1$.
@@
 That is: $\mathsf{P}_{\psi_2} (\psi_1) = \psi_1 \Leftrightarrow \psi_1 = \alpha \psi_2$.
-Moreover, because the material is irreducible, if some fraction of $\psi_1$ is prepared like $\psi_2$, then the same fraction of $\psi_2$ is prepared like $\psi_1$.
-If not, we could learn more (or less) about $\psi_1$ by studying the component of $\psi_2$ that is prepared like $\psi_1$.
+Because of this, can define a non degenerate complex product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{\langle \psi_2|\psi_1 \rangle}{|\psi_2|^2} \psi_2$.
+
+The projections have other useful properties.
+Because the material is irreducible, if some fraction of $\psi_1$ is prepared like $\psi_2$, then the same fraction of $\psi_2$ is prepared like $\psi_1$.
+If not, studying the fragments of one state would give us a better (or worse) description for the fragments of the other state.
 Therefore we have $\frac{|\mathsf{P}_{\psi_2} (\psi_1)|^2}{|\psi_1|^2}=\frac{|\mathsf{P}_{\psi_1} (\psi_2)|^2}{|\psi_2|^2}$.
+This tells us that $|\langle \psi_2|\psi_1 \rangle|^2 = |\langle \psi_1|\psi_2 \rangle|^2$.
+
 Lastly, if we take the projection of $\psi_1$ along $\psi_2$ and we project it again on $\psi_1$, we end up with a component of $\psi_1$.
@@
 That is: $\mathsf{P}_{\psi_1} (\mathsf{P}_{\psi_2} (\psi_1)) = a \psi_1$ with $a \in \mathbbm{R}$.
-
-From the norm and the projection we can define a complex product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{\langle \psi_2|\psi_1 \rangle}{|\psi_2|^2} \psi_2$.
+This tells us that $\langle \psi_2|\psi_1 \rangle = \langle \psi_1|\psi_2 \rangle^\dagger$.
+
 Because of the properties of the norm and the projections, the product is positive definite, linear and conjugate symmetric.
-This means the state space $\Psi$ for an irreducible material is an inner product space.
+In other words, the state space $\Psi$ for an irreducible material is an inner product space derived by the norm and projections.
 

### 2016-12-12 2015CompleteDerivation/completeDerivation.tex
Incorporate discussion with Dave and Lydia on state counting

@@
 We'll see that very similar relationships are what define Hamiltonian and Lagrangian mechanics.
-But recovering them in the continuous case is trickier, and in fact will be part of the challenge in the following sections.
+But recovering them in the continuous case is not as straight forward.
 Consider the map $q'=aq$ with $0<a<1$.
@@
 In the limit where we apply the map an infinite amount of times, any value of $q$ will be brought infinitely close to $0$, which also does not sound like a reversible process.
-The issue is that while deterministic evolution will map points to points, not all point to point maps can be considered deterministic and reversible.
-Discrete sets already come with a way to measure and compare the number of elements which is preserved by any bijective map.
-With continuous sets, we need to do extra work to properly define a measure (or metric) that allows one to count states and possibilities.
+
+The issue is that while deterministic evolution will map states to states, not all bijective state to state maps can be considered deterministic and reversible.
+This is a crucial point where this work is different from more usual treatments of dynamical systems.
+If one focuses only on the states themselves and counts them, it reaches the conclusion that a range of $1$m contains the same number of initial conditions of $1$Km: they have the same infinite number of elements.
+This conclusion has a number of problems: it does not match our physical intuition; it's not consistent with all types of topology; it does not explain why the laws of evolution are very often, not only continuous, but differentiable.
+In short: a bijection does not capture the entirety of the physical description.
+
+If we work with a discrete topology then, by all means, the way to measure the size of a set is to count the elements in a set.
+But if we use the standard topology on $\mathbbm{R}$, isolated points have measure zero: what defines the number of initial conditions is the range of the state variables.
+As we'll see, the geometrical structures in Hamiltonian/Lagrangian classical/quantum mechanics are there to allow us to properly count states and, more specifically, possibilities along independent state variables.
+If we think that counting states is equivalent to counting elements in the set, we miss the whole point of that geometrical structure.
+
+What makes the matter even more confusing is that the single physical idea of counting states and possibilities along independent state variable does not map to a single mathematical concept.
+Sometimes we need a measure, sometimes we need a symplectic form and sometimes we need a metric tensor.
+It will depend, case by case, on the state space of the system under study.
+But, however defined mathematically, deterministic and reversible motion must preserve that count.
+This is the crucial insight that gives physical meaning to those structures, that justifies the use of differential equations for state evolution, that tells us why symplectic and Riemannian geometry are so important in physics.
+Properly characterizing this count will be one of the key goals of the following sections.
 
@@
 As we can see, the framework we have is ill suited to handle these cases.
-Therefore we cannot simply stick time in the Hamiltonian and expect everything to work out.
-But we can't ignore the problem either as the three cases outlined are pretty common situations to study.
+Therefore we cannot simply stick time in the Hamiltonian and expect everything to work out.\footnote{Looking ahead, another way to frame the problem is that Galilean transformations do not preserve phase space volume.
+Lorentz transformations, on the other hand, do and it is this requirements plus the kinematic assumption that will recover them later.} But we can't ignore the problem either as the three cases outlined are pretty common situations to study.
 So we need to go back to our definitions and amend them properly.

### 2016-12-12 2015CompleteDerivation/completeDerivation.tex
First pass at discussion and conclusion

@@
 
-%Time arrow
-
-%Particle excitations
-
-%Mixed state as fundamental (law of evolution defined on the mixed state)
-
-%Black hole as irreducible object
+As we went through many details in the derivation, it is time to step back and conclude with a few general remark.
+
+The first remarkable aspect of this work is that it can exists.
+Regardless of whether all the minor details are addressed in the best possible way, the idea that so much can be derived from so little is unexpected.
+
+It often happened that the inability to justify one detail forced a partial shift in thinking that enabled to connect more ideas more simply.
+The ability to constrain the Hamiltonian to vector potential came as a surprise after we realized we needed distributions as a core concept to justify Hamilton's equations.
+The fact that relationships must be continuous functions came as a surprise after we connected topology with the notion of physical distinguishability.
+It was this constant simplification that shows us that we were on the right track and that, in the end, decided the boundary of the work as the initial goals were very modest in comparison.
+
+The second striking aspect is a shift in perspective that this work forces upon us.
+The usual reductionist view is that you start with the rules of the small parts and then build up the composite whole.
+This works shift that perspective upside down: it's not what the parts can tell you about the whole, it's what the whole can tell you about the parts.
+
+The state space for classical particles is the one that allows to express invariant distributions: that is a requirement that comes from the whole.
+The fragments of quantum particles do not even have a well defined description: the whole tells us nothing about them.
+In retrospect, this is not be surprising.
+Nobody is going to hand us down the laws of infinitesimal parts: we induce them by manipulating finite systems.
+
+The third notable aspect is that the assumptions have more to do with our ability to create a reliable description rather than essential features of the system itself.
+A system is not deterministic per se: it is a particular choice for its description (i.e.
+state) that may be deterministic under some evolution.
+In the same vein, a material is not homogeneous per se.
+Or infinitesimally reducible.
+In this context, the state, its properties and the law of evolution are the description of the system under our idealization, not some intrinsic feature of the system.
+
+The fourth intriguing aspect is that, when framing classical and quantum mechanics in this way, there is a progression.
+First we assume a homogeneous infinitesimally reducible material.
+Then we assume an homogeneous irreducible material.
+Connecting the dots, the next assumption would be a inhomogeneous irreducible material: where the material is not in constant equilibrium and its particles are not composed of the same amount.
+This different starting point may lead to the resolution of some of the problems we mentioned.
+
+As different particle can transform into one another, they could be seen as different stable states of the same inhomogeneous material.
+In particular, since $\sigma_x \sigma_v \geq \frac{\hbar}{2m}$, particles in different generations (i.e.
+electron vs muon) could be similar stable configurations where the fragments of the material are more "packed" in terms of states labeled by the kinematic variables.
+
+As we are not always at equilibrium, the state of the system will be in general a statistical mixture of distributions of material.
+But the evolution may also take pure states (i.e.
+a mixed state with a single state with $100\%$ probability) and evolve them into mixed states, thus giving an account of what happens during a measurement or a decay.
+
+The law of evolution may still be deterministic and reversible, in the sense that given a particular mixed state we can predict/reconstruct future/past mixed state, but it will not be time symmetric.
+We can then tell that time flows in the direction where information is exchanged, where entropy increases.
+This would give us a time arrow.
+
+These ideas are appealing because they seem to fit well in the narrative, but at this point they are just speculation.
+The only thing we can say at this point is that the state space would be a vector space on a field that is neither $\mathbbm{R}$ or $\mathbbm{C}$.
+
+In the near future it would seem appropriate to extend this work to the remaining part of well-established physics.
+The ideas toward that goal are more concrete.
+The state space for a quantized material, already introduced here, should be captured by the Fock space.
+The idea of a non-minimal material will force us to consider vector and tensor quantities at each point in the distribution, whose state should be captured by spin systems.
+Assuming that the vector potential forces themselves are dependent on the state of a material should hopefully give us the Hamiltonian for the fields.
 
 \section{Conclusion}
+
+We have seen that much of fundamental physics can be derived from a few simple physical assumption (and lots of hard work).
+In doing so we obtained a more unified picture of the different branches of fundamental physics centered around the idea of deterministic and reversible evolution of distributions of material.
+The main physical concepts can be summarized by the following points:
+
+\begin{itemize}
+
+\item The state space is not just a collection of states but also captures its physical properties such as physical distinguishability (represented by a topology), the ability to count states (represented by a measure, symplectic form or metric tensor) or decomposability (represented by a vector space structure).
+
+\item Deterministic and reversible evolution is a one-to-one map that preserves the nature of the system (e.g.
+physical distinguishability, state count and decomposability).
+Therefore it is an isomorphism on the mathematical structure defined by the state space (e.g.
+homeomorphism, canonical transformation, unitary transformation), not just a state-to-state bijection.
+
+\item Classical Hamiltonian mechanics describes the evolution of a homogeneous material for which the state is equivalent to the states of all its infinitesimal parts, the classical particles.
+The state space for classical particles is the symplectic manifold $(T^*\mathcal{Q}, \omega)$; deterministic and reversible evolution is a symplectomorphism (i.e.
+a canonical transformation).
+
+\item If the classical particle states are fully characterized by their motion in space, then the states also evolve according to Lagrangian mechanics and the motion is that of a massive particle under scalar/vector potential forces.
+
+\item Relativistic motion arises by properly handling time dependent deterministic and reversible evolution for the material (i.e. distributions over phase space).
+
+\item Quantum (Hamiltonian) mechanics describes the evolution of a homogeneous material for which the state can only be reduced up to finite parts, the quantum particles, whose internal motion is undetermined.
+The state space for quantum particles is an inner vector product space $(\Psi, \langle \cdot , \cdot \rangle)$; deterministic and reversible evolution is a unitary transformation.
+\end{itemize}
+
+While we are sure that some mathematical details can be improved, the feel that the overall ``conceptual castle" is solid.
+The starting concepts are simple, the conceptual mapping of those concepts into math is often straight forward and it often gave intuitive ideas that connects to different branches of math and physics: it cannot be a simple coincidence.
+
+While the scope of this work is limited to re-founding and re-understanding the already known physical theories, it is our hope that this insight may also lead to new ideas to address some of the open problems in the foundations of physics.
 

### 2016-12-15 2015CompleteDerivation/completeDerivation.tex
First pass at the introduction

@@
 
-We work strives as much as possible to be philosophically consistent, physically meaningful and mathematically precise.
+This work strives as much as possible to be philosophically consistent, physically meaningful and mathematically precise.
 \end{abstract}
@@
 
-\section{Status}
+\textbf{\emph{Status}}
 
@@
 Section IX - Mostly done. Needs review.
+
+\section{Introduction}
+
+While some branches of physics (e.g.
+Newtonian mechanics, thermodynamics, special relativity) are derived from physical assumptions (e.g.
+Newton's laws, laws of thermodynamics, invariance of the speed of light), others start by simply setting a mathematical framework.
+It is the case for classical Lagrangian mechanics, where we start by assuming the existence of a Lagrangian function whose path integral is minimized along the motion.
+It is the case for classical Hamiltonian mechanics, where we start by assuming the state is described by conjugate pairs $(q^i, p_i)$ and evolves according to Hamilton's equations.
+It is the case for quantum mechanics, where the state space is a complex inner product vector space and the state evolves according to Schroedinger's equations.
+That is, at the very core of many of our physical theories we have mathematical definitions instead of physical concepts.
+
+While we have grown accustomed to this state of affair, it is nonetheless troublesome as it has problematic consequences.
+Suppose you have a system in front of you, how can you tell whether it's Hamiltonian?
+Or classical?
+We know that these theories have different range of applicability.
+To describe a cannonball we should be using classical mechanics while for an electron we should be using quantum mechanics.
+Not all Hamiltonians admit a corresponding Lagrangian.
+How do we pick?
+
+If all we have are mathematical definitions, the only answer is that it is a Hamiltonian system if it obeys Hamiltonian mechanics.
+While many are indeed satisfied by such answer, one has to admit that it begs the question.
+We do not have a clear idea of what an Hamiltonian system is and how it different from a Lagrangian one.
+Or what makes a system classical and how it is different form a quantum system.
+
+We started this work with the unpretentious goal of exploring these questions.
+The idea was to create a math-to-physics dictionary so that each mathematical concept was clearly mapped to a crisp physical concept.
+But to make sure that such dictionary was complete, we had to make sure that all the relevant mathematical framework could be recovered.
+This lead to the mission of re-deriving as much as possible starting from physical assumptions.
+
+This approach pushed us further and further back to the basic definitions.
+For example, if we truly wanted to justify why the state space of Hamiltonian mechanics is charted by conjugate variables $q^i$ and $p_i$, we had to justify why it was a manifold as well, which in turn meant we needed to justify why it was a topological space.
+But when we did find physical concepts that mapped to these more basic mathematical definitions, we realized that this link was valid for a far wider range of cases: the concepts were more general than we expected.
+
+To our surprise, this pushed us to bring in other areas that were not within our horizon.
+The same link that allowed to justify Hamiltonian dynamics also gave us relativistic motion on the extended phase space.
+The same link that allowed to justify Lagrangian mechanics also gave us constrained the motion to free particles under scalar/vector potential forces.
+Since this happened repeatedly, it lead to think that, maybe, the picture we were developing was not simply ``a nicer way to think about things", but a fundamental framework to understand and link different parts of basic science and mathematics.
+
+%It felt like, when trying to assemble a jigsaw puzzle, you realize that the reason you can't find a piece is because you assembled the previous parts incorrectly and you end with a larger part that fits together a lot better.
+
+These unexpected connections lead to the realization that sometimes different mathematical concepts are expression of the same physical concept.
+That is: by putting physics first we can achieve a sort of unification of ideas that is not possible by math alone.
+
+The goal of this particular work is to report on this progress.
+The claim is that the whole architectural framework is solid enough to be convinced that it is possible to derive classical/quantum Hamiltonian/Lagrangian mechanics starting from a few physical principle.
+This is not just an exercise in style: this approach allows to make intuitive connections between different areas of physics and mathematics.
+
+The claim is that, starting from a few simple physical assumptions, we can derive classical/quantum Hamiltonian/Lagrangian mechanics.
+The claim is not that we have achieved perfection and that every proposition and justification cannot be improved.
+The claim is, simply, that it \emph{can} be done, and that doing so provides better insight and intuition that clarifies the physics being described and that allows to strengthen connections between different areas of math and physics.
+We will see that sometimes different mathematical concepts are the expression of one physical concept.
+Therefore, by putting physics, first we can achieve a sort of unification of ideas that is not possible by math alone.
+
+While trying to re-derive classical, quantum and (to some extent) relativistic mechanics may seem too broad a scope, it is necessary to ensure that the concepts we develop are truly general, that they can be applied to many different cases.
+If we could derive classical Hamiltonian mechanics on assumptions that are inapplicable to the quantum case, what good would it do?
+Quantum system are Hamiltonian systems, so clearly we haven't understood what makes a system Hamiltonian.
+The ability to apply the same physical concepts far and wide is what gives us confidence that these are truly general ideas and they capture the physical aspects we are describing.
+
+As brief overview, there are four assumptions we will consider.
+Assuming deterministic and reversible evolution will give us a dynamical system: a state and a law of evolution.
+Additionally assuming infinitesimal reducibility, studying the whole system is equivalent to studying its infinitesimal parts, gives us Hamiltonian mechanics.
+Additionally assuming kinematic equivalence, trajectories in space tell us everything about the system, gives us Lagrangian mechanics and constrains the motion to massive particles under scalar/vector potential forces.
+Relativistic motion arises when studying time dependent laws of evolution with no further assumption.
+Alternatively, as the second assumption we can take irreducibility, studying the whole systems tells us nothing about its parts, which leads to quantum (Hamiltonian) mechanics.
+
+In section \ref{sec:fundamental_model} we will describe in general what it means to study a physical system and how the assumption of determinism and reversibility plays into it.
+We will develop an abstract conceptual model that we will use when discussing the later assumptions.
+
+In section \ref{sec:dynamical_systems} we will show how the physical concept of physical distinguishability maps to the mathematical concept of topological spaces.
+Deterministic and reversible is an invertible continuous map (i.e.
+a homeomorphism) as it needs to preserve what is physically distinguishable.
+
+In section \ref{sec:reducibility} we will introduce the notion of a classical material, one for which we assume infinitesimal reducibility.
+The state of an arbitrary amount of material is a distribution over the states of the infinitesimal parts, the classical particles.
+As such distribution must be invariant under the arbitrary choice of units, for each variable $q^i$ that defines a unit there is a corresponding variable $k^i$ defined on the inverse of such unit.
+Classical particle states can be identified by conjugate pairs of state variable $(q^i, p_i)$ (i.e.
+a point in the cotangent bundle $T^*\mathcal{Q}$) and the number of possible initial conditions for each pair is given by $dq^i \wedge dp_i$ (i.e.
+the canonical symplectic form $\omega$).
+During evolution, the distribution must be mapped point by point which leads to Hamiltonian mechanics (i.e.
+a symplectomorphism on $T^*\mathcal{Q}$).
+
+In section \ref{sec:relativity} we will show that a relativistic version of Hamilton's equations arises naturally when we consider time dependent laws of evolution for distributions of classical material.
+This also gives the notion of deterministic and reversible evolution a characterization that is independent of any transformation that mixes time and state variables $(q^i, p_i)$.
+
+In section \ref{sec:Lagrangian} we will introduce the assumption of kinematic equivalence, studying the motion in space is equivalent to studying the evolution of the state.
+Under that assumption there is a link between the state variables $(q^i, p_i)$ and the kinematic variable $(x^i, u^i)$ (i.e.
+position and velocity).
+This link will allow us to express the number of possible initial conditions for d.o.f.
+in terms of $dx^i$ and $du^i$ (i.e.
+a metric tensor $g$).
+Such expression allows to constrain the Hamiltonian to the one for massive particles under scalar potential forces.
+Such assumption also allows for a Lagrangian formulation.
+
+In section \ref{sec:quantum_prelude} we will introduce some ideas in classical mechanics to more gently transition over quantum mechanics.
+We will see how classical materials (i.e.
+distributions over phase space) already have features that are typically associated with quantum systems and we will show how to construct a classical uncertainty principle that helps to build intuition used in the later section.
+
+In section \ref{sec:irreducibility} we will introduce the notion of an irreducible material, one for which studying the whole distribution does not tell us anything about the motion of its infinitesimal parts, the fragment.
+A unit amount of such material will correspond to a quantum particle.
+As this means that the states of such material are invariant under permutations of the configurations for the fragments, we will show that the state space of an irreducible material is a complex vector space.
+As we will still be able to tell how similar or different two states are, this will gives an inner product to the state space.
+Such comparisons will be preserved by deterministic and reversible evolution which leads to Schroedinger equation (i.e.
+a unitary transformation).
+
+In section \ref{sec:discussion} we will discuss of how this work could be extended in different directions.
 
@@
 While the style of a justification is formal, it is not a pure mathematical proof as it necessarily has to mix physical and mathematical arguments.
-
 Before or after each proposition there is a \textbf{discussion} that conceptually explains the motivation and the result in a less formal, more approachable way.
+
+  Justifications and discussions are designed to be independent to allow two separate types of reading style: one focused on higher level concepts and one focused on making sure the details follow.
+This inevitably leads to some repetition.
 
@@
 \section{On studying a physical system}
+\label{sec:fundamental_model}
 
@@
 \section{States and state space}
+\label{sec:dynamical_systems}
 
@@
 \section{Time dependent evolution}
+\label{sec:relativity}
 
@@
 \section{Kinematics and minimal materials}
+\label{sec:Lagrangian}
 
@@
 \section{Prelude to quantum}
+\label{sec:quantum_prelude}
 
@@
 \section{Irreducible systems}
+\label{sec:irreducibility}
 
@@
 \section{Discussion and further work}
+\label{sec:discussion}
 
@@
 We have seen that much of fundamental physics can be derived from a few simple physical assumption (and lots of hard work).
-In doing so we obtained a more unified picture of the different branches of fundamental physics centered around the idea of deterministic and reversible evolution of distributions of material.
+In doing so we obtained a more unified picture of the different branches of fundamental physics centered around the idea of deterministic and reversible evolution of distributions of a homogeneous material.
 The main physical concepts can be summarized by the following points:
@@
 
-\item The state space is not just a collection of states but also captures its physical properties such as physical distinguishability (represented by a topology), the ability to count states (represented by a measure, symplectic form or metric tensor) or decomposability (represented by a vector space structure).
+\item A state space is not just a collection of states for the system under study but it must also capture physical properties such as physical distinguishability (represented by a topology), the ability to count states (represented by a measure, symplectic form or metric tensor) or decomposability (represented by a vector space structure).
 
@@
 
-\item Classical Hamiltonian mechanics describes the evolution of a homogeneous material for which the state is equivalent to the states of all its infinitesimal parts, the classical particles.
-The state space for classical particles is the symplectic manifold $(T^*\mathcal{Q}, \omega)$; deterministic and reversible evolution is a symplectomorphism (i.e.
+\item Classical Hamiltonian mechanics describes the evolution of a homogeneous material whose state is equivalent to the states of all its infinitesimal parts, the classical particles.
+The state space for classical particles is a symplectic manifold $(T^*\mathcal{Q}, \omega)$ as this allows coordinate invariant density distributions; deterministic and reversible evolution for such state space is a symplectomorphism (i.e.
 a canonical transformation).
 
-\item If the classical particle states are fully characterized by their motion in space, then the states also evolve according to Lagrangian mechanics and the motion is that of a massive particle under scalar/vector potential forces.
+\item Assuming classical particle states are fully characterized by their trajectory in space, the states also evolve according to Lagrangian mechanics and the motion is that of a massive particle under scalar/vector potential forces.
 
@@
 \item Quantum (Hamiltonian) mechanics describes the evolution of a homogeneous material for which the state can only be reduced up to finite parts, the quantum particles, whose internal motion is undetermined.
-The state space for quantum particles is an inner vector product space $(\Psi, \langle \cdot , \cdot \rangle)$; deterministic and reversible evolution is a unitary transformation.
+The state space for quantum particles is an inner vector product space $(\Psi, \langle \cdot , \cdot \rangle)$; deterministic and reversible evolution for such state space is a unitary transformation.
 \end{itemize}
 
-While we are sure that some mathematical details can be improved, the feel that the overall ``conceptual castle" is solid.
-The starting concepts are simple, the conceptual mapping of those concepts into math is often straight forward and it often gave intuitive ideas that connects to different branches of math and physics: it cannot be a simple coincidence.
-
-While the scope of this work is limited to re-founding and re-understanding the already known physical theories, it is our hope that this insight may also lead to new ideas to address some of the open problems in the foundations of physics.
+While we are sure that some mathematical details can be improved, we feel that the overall ``conceptual castle" is solid.
+The starting concepts are simple, the conceptual mapping of those concepts into math is often straight forward and the intuitive ideas that emerge help connecting different branches of math and physics.
+This is unlikely a matter of simple coincidence.
+
+While the scope of this work is limited to re-founding and re-understanding physical theories already experimentally established, it may also has the potential to provide new insights and ideas on how to address some of the open problems in the foundations of physics.
 

### 2016-12-19 2015CompleteDerivation/completeDerivation.tex
Reviewed the first two sections

@@
 While some branches of physics (e.g.
-Newtonian mechanics, thermodynamics, special relativity) are derived from physical assumptions (e.g.
+Newtonian mechanics, thermodynamics, special relativity) are derived from physical principles (e.g.
 Newton's laws, laws of thermodynamics, invariance of the speed of light), others start by simply setting a mathematical framework.
@@
 It is the case for quantum mechanics, where the state space is a complex inner product vector space and the state evolves according to Schroedinger's equations.
-That is, at the very core of many of our physical theories we have mathematical definitions instead of physical concepts.
-
-While we have grown accustomed to this state of affair, it is nonetheless troublesome as it has problematic consequences.
-Suppose you have a system in front of you, how can you tell whether it's Hamiltonian?
-Or classical?
-We know that these theories have different range of applicability.
-To describe a cannonball we should be using classical mechanics while for an electron we should be using quantum mechanics.
-Not all Hamiltonians admit a corresponding Lagrangian.
-How do we pick?
-
-If all we have are mathematical definitions, the only answer is that it is a Hamiltonian system if it obeys Hamiltonian mechanics.
-While many are indeed satisfied by such answer, one has to admit that it begs the question.
-We do not have a clear idea of what an Hamiltonian system is and how it different from a Lagrangian one.
-Or what makes a system classical and how it is different form a quantum system.
-
-We started this work with the unpretentious goal of exploring these questions.
+That is, at the very core of some of our physical theories we have mathematical definitions instead of physical ideas.
+
+If we ask ``when is a particular physical system suitably described by Hamiltonian (or by classical) mechanics?" the only answer we have is ``when it fits the mathematical framework".
+While many seem to be content with this answer (e.g.
+``you have to start from somewhere, it may as well be the math which at least is precise"), it begs the question: it does not tell us how and what physical properties of the system under study are being captured.
+
+The fundamental problem is that mathematical structures are not enough to characterize the physical system being described.
+For example, the same math of linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems.
+The mathematical treatment is the same because it captures aspects of the description that are common to all cases.
+But for the same very reason, we cannot infer what system is being described if all we are given is the mathematical model.
+In the jargon of linguistic/computer science/phylosophy: math captures the syntax (the relationships between the objects) but does not capture the semantic (the meaning of each object).
+
+We started this work with the unpretentious goal of exploring simple questions such as the following.
+What do these mathematical symbols used in classical/quantum Hamiltonian represent?
+What is the physical meaning attached to it?
+Why is the state of a quantum system described by a wave function (i.e.
+a vector in a complex inner product space), while the state of a classical hamiltonian system is described by the three pairs $(q^i, p_i)$ (i.e.
+a point in a cotangent bundle)?
 The idea was to create a math-to-physics dictionary so that each mathematical concept was clearly mapped to a crisp physical concept.
-But to make sure that such dictionary was complete, we had to make sure that all the relevant mathematical framework could be recovered.
-This lead to the mission of re-deriving as much as possible starting from physical assumptions.
-
-This approach pushed us further and further back to the basic definitions.
+But to make sure that such dictionary was complete, we realized we had to make sure that the whole relevant mathematical framework could be recovered.
+
+This redefined our mission to re-deriving as much as possible starting from physical assumptions.
+It also forced us to push further and further back toward the basic definitions.
 For example, if we truly wanted to justify why the state space of Hamiltonian mechanics is charted by conjugate variables $q^i$ and $p_i$, we had to justify why it was a manifold as well, which in turn meant we needed to justify why it was a topological space.
-But when we did find physical concepts that mapped to these more basic mathematical definitions, we realized that this link was valid for a far wider range of cases: the concepts were more general than we expected.
-
-To our surprise, this pushed us to bring in other areas that were not within our horizon.
+But when we did find physical concepts that mapped to these more basic mathematical definitions, we realized that this link was valid for a far wider range of cases: the concepts were more general than we anticipated.
+For example, we realized that the mathematical notion of a topology maps incredibly well to the idea of physical distinguishability, but states of Hamiltonian system are not the only thing that are physically distinguishable.
+
+To our surprise, this naturally expanded the boundary of the work.
 The same link that allowed to justify Hamiltonian dynamics also gave us relativistic motion on the extended phase space.
-The same link that allowed to justify Lagrangian mechanics also gave us constrained the motion to free particles under scalar/vector potential forces.
-Since this happened repeatedly, it lead to think that, maybe, the picture we were developing was not simply ``a nicer way to think about things", but a fundamental framework to understand and link different parts of basic science and mathematics.
-
-%It felt like, when trying to assemble a jigsaw puzzle, you realize that the reason you can't find a piece is because you assembled the previous parts incorrectly and you end with a larger part that fits together a lot better.
-
-These unexpected connections lead to the realization that sometimes different mathematical concepts are expression of the same physical concept.
-That is: by putting physics first we can achieve a sort of unification of ideas that is not possible by math alone.
-
-The goal of this particular work is to report on this progress.
-The claim is that the whole architectural framework is solid enough to be convinced that it is possible to derive classical/quantum Hamiltonian/Lagrangian mechanics starting from a few physical principle.
-This is not just an exercise in style: this approach allows to make intuitive connections between different areas of physics and mathematics.
-
-The claim is that, starting from a few simple physical assumptions, we can derive classical/quantum Hamiltonian/Lagrangian mechanics.
-The claim is not that we have achieved perfection and that every proposition and justification cannot be improved.
-The claim is, simply, that it \emph{can} be done, and that doing so provides better insight and intuition that clarifies the physics being described and that allows to strengthen connections between different areas of math and physics.
-We will see that sometimes different mathematical concepts are the expression of one physical concept.
-Therefore, by putting physics, first we can achieve a sort of unification of ideas that is not possible by math alone.
+The same link that allowed to justify Lagrangian mechanics also onstrained the motion to that of free particles under scalar/vector potential forces.
+We also realized that sometimes different mathematical concepts could be justified from the same physical starting point.
+That is: by putting physics first we could achieve a sort of unification of ideas that was not practicable by math alone.
+
+The success of the approach lead us to think that, maybe, the picture we were developing was not simply ``a nicer way to think about things", but instead a better framework to understand and link different parts of basic science and mathematics.
+After many years of work and different incremental iterations we think it is time to checkpoint and report the progress.
+
+The goal of this paper is to show that it is possible to derive classical/quantum Hamiltonian/Lagrangian mechanics starting from a few physical principle.
+We do \emph{not} claim that we have perfectly achieved this result, with no room from improvement.
+We do claim that the whole architectural framework is solid enough to be convinced that this is indeed possible, and that it is a worthwhile endeavor as it allows to make intuitive connections between different areas of physics and mathematics.
+That is: it \emph{can} and \emph{should} be done.
 
@@
 Quantum system are Hamiltonian systems, so clearly we haven't understood what makes a system Hamiltonian.
-The ability to apply the same physical concepts far and wide is what gives us confidence that these are truly general ideas and they capture the physical aspects we are describing.
-
-As brief overview, there are four assumptions we will consider.
+The ability to apply the same physical ideas far and wide is what gives us confidence that these are truly general concept that capture the physical aspects we are describing.
+The breath itself is what makes the work interesting and more than a curiosity.
+
+As a brief overview, there are four assumptions we will consider.
 Assuming deterministic and reversible evolution will give us a dynamical system: a state and a law of evolution.
 Additionally assuming infinitesimal reducibility, studying the whole system is equivalent to studying its infinitesimal parts, gives us Hamiltonian mechanics.
-Additionally assuming kinematic equivalence, trajectories in space tell us everything about the system, gives us Lagrangian mechanics and constrains the motion to massive particles under scalar/vector potential forces.
+Additionally assuming kinematic equivalence, trajectories in space tell us everything about the state of the system, gives us Lagrangian mechanics and constrains the motion to massive particles under scalar/vector potential forces.
 Relativistic motion arises when studying time dependent laws of evolution with no further assumption.
@@
 
-In section \ref{sec:fundamental_model} we will describe in general what it means to study a physical system and how the assumption of determinism and reversibility plays into it.
+As a more detailed overview: in section \ref{sec:fundamental_model} we will describe in general what it means to study a physical system and how the assumption of determinism and reversibility plays into it.
 We will develop an abstract conceptual model that we will use when discussing the later assumptions.
@@
 In section \ref{sec:dynamical_systems} we will show how the physical concept of physical distinguishability maps to the mathematical concept of topological spaces.
-Deterministic and reversible is an invertible continuous map (i.e.
+Deterministic and reversible evolution is an invertible continuous map (i.e.
 a homeomorphism) as it needs to preserve what is physically distinguishable.
@@
 The state of an arbitrary amount of material is a distribution over the states of the infinitesimal parts, the classical particles.
-As such distribution must be invariant under the arbitrary choice of units, for each variable $q^i$ that defines a unit there is a corresponding variable $k^i$ defined on the inverse of such unit.
+As such distribution must be invariant under the arbitrary choice of units, we'll see that for each variable $q^i$ that defines a unit there is a corresponding variable $k^i$ defined on the inverse of such unit.
 Classical particle states can be identified by conjugate pairs of state variable $(q^i, p_i)$ (i.e.
@@
 the canonical symplectic form $\omega$).
-During evolution, the distribution must be mapped point by point which leads to Hamiltonian mechanics (i.e.
+During evolution, the distribution (and marginal distributions) must be mapped point by point which leads to Hamiltonian mechanics (i.e.
 a symplectomorphism on $T^*\mathcal{Q}$).
@@
 This also gives the notion of deterministic and reversible evolution a characterization that is independent of any transformation that mixes time and state variables $(q^i, p_i)$.
+It will also give us a classical equivalent for anti-particles states.
 
@@
 This link will allow us to express the number of possible initial conditions for d.o.f.
-in terms of $dx^i$ and $du^i$ (i.e.
-a metric tensor $g$).
-Such expression allows to constrain the Hamiltonian to the one for massive particles under scalar potential forces.
-Such assumption also allows for a Lagrangian formulation.
+in terms of $dx^i$ and $du^i$ (leading to the metric tensor $g$).
+Such expression also allows to constrain the Hamiltonian to the one for massive particles under scalar potential forces and an equivalent Lagrangian formulation.
 
@@
 
-In section \ref{sec:irreducibility} we will introduce the notion of an irreducible material, one for which studying the whole distribution does not tell us anything about the motion of its infinitesimal parts, the fragment.
-A unit amount of such material will correspond to a quantum particle.
+In section \ref{sec:irreducibility} we will introduce the notion of an irreducible material, one for which studying the whole distribution does not tell us anything about the motion of its infinitesimal parts, the fragments.
+A unit amount will correspond to a quantum particle.
 As this means that the states of such material are invariant under permutations of the configurations for the fragments, we will show that the state space of an irreducible material is a complex vector space.
-As we will still be able to tell how similar or different two states are, this will gives an inner product to the state space.
+These states can still be compared to each other in a way that gives the state space an inner product.
 Such comparisons will be preserved by deterministic and reversible evolution which leads to Schroedinger equation (i.e.
@@
 
-The work is organized into:
+The overall layout of the work is a balance between the need to show that the derivation follows (which requires starting from basic principles and reaching conclusion using formal arguments) and the need to give some intuitive understanding (which often requires starting from specific cases and then generalizing).
+
+The deductive aspect is formally organized into the following:
 \begin{description}
@@
 While the style of a justification is formal, it is not a pure mathematical proof as it necessarily has to mix physical and mathematical arguments.
-Before or after each proposition there is a \textbf{discussion} that conceptually explains the motivation and the result in a less formal, more approachable way.
-
-  Justifications and discussions are designed to be independent to allow two separate types of reading style: one focused on higher level concepts and one focused on making sure the details follow.
-This inevitably leads to some repetition.
 
@@
 \end{description}
-This should allow readers with different focus and background to concentrate on the parts that are more keen on.
+
+As these sections are clearly demarcated, this allows us to make sure that each justification only references past propositions.
+It also makes it easier to read the work at the desired level of detail.
+
+Interspersed between each step of the more formal treatment, there are numerous discussions that conceptually explain motivations and results in a more approachable way.
+These can use more concrete and intuitive examples without the fear of circular arguments.
+As both the formal and more intuitive part need to cover the same material, this will inevitably lead to some amount of repetition.
+
+The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology, measure theory, differential geometry (symplectic and Rimannian), group theory and vector spaces, statistics, information theory, Hamiltonian mechanics, Lagrangian mechanics, quantum mechanics, thermodynamics, special relativity and so on.
+Connecting these areas, in fact, is part of the insight given by this work.
+While this is a necessity, it does create a number of problems.
+
+The first one is terminology and prerequisite knowledge.
+Since we touch a good number of areas, we cannot expect all readers to be equally comfortable with all the technical terms.
+Yet we cannot introduce all of them either.
+The compromise we adopted is to keep the informal discussion more accessible while keeping the formal derivation more precise.
+Even in the formal derivation, though, we will seldom need more than the basic definitions within a specific area and few major results.
+
+The second problem is notation.
+As each field has developed its own set of conventions, the common usage is now an empty set.
+We ended up with a compromise that we believe strikes a good balance.
+Especially in the informal discussion, we crafted something that in general feels familiar to a physicist.
+Within formal parts, though, we may use notation and terminology more specific to the area of math being used.
+But then again: one would need to be familiar to that area anyway to be able to follow the details.
+
+We have tested the prose on a number of students and colleagues to be confirm its accessibility at different levels.
+We hope the reader enjoys this journey and comes out with a few new insights and ideas.
 

### 2016-12-20 2015CompleteDerivation/completeDerivation.tex
New pass at the introduction

@@
 Newton's laws, laws of thermodynamics, invariance of the speed of light), others start by simply setting a mathematical framework.
-It is the case for classical Lagrangian mechanics, where we start by assuming the existence of a Lagrangian function whose path integral is minimized along the motion.
+It is the case for classical Lagrangian mechanics, where we start by assuming the existence of a Lagrangian function whose path integral is minimized during the motion.
 It is the case for classical Hamiltonian mechanics, where we start by assuming the state is described by conjugate pairs $(q^i, p_i)$ and evolves according to Hamilton's equations.
@@
 
-If we ask ``when is a particular physical system suitably described by Hamiltonian (or by classical) mechanics?" the only answer we have is ``when it fits the mathematical framework".
+If we ask ``when can a particular physical system be suitably described by Hamiltonian (or by classical) mechanics?" the only answer we have is ``when it fits the mathematical framework".
 While many seem to be content with this answer (e.g.
@@
 
-The fundamental problem is that mathematical structures are not enough to characterize the physical system being described.
-For example, the same math of linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems.
+This is a problem since mathematical structures are not enough to characterize the physical system being described.
+For example, the same math for linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems.
 The mathematical treatment is the same because it captures aspects of the description that are common to all cases.
-But for the same very reason, we cannot infer what system is being described if all we are given is the mathematical model.
+But for the very same reason, we cannot infer what system is being described if all we are given is the mathematical model.
 In the jargon of linguistic/computer science/phylosophy: math captures the syntax (the relationships between the objects) but does not capture the semantic (the meaning of each object).
@@
 We started this work with the unpretentious goal of exploring simple questions such as the following.
-What do these mathematical symbols used in classical/quantum Hamiltonian represent?
+What do the mathematical symbols used in classical/quantum Hamiltonian represent?
 What is the physical meaning attached to it?
-Why is the state of a quantum system described by a wave function (i.e.
-a vector in a complex inner product space), while the state of a classical hamiltonian system is described by the three pairs $(q^i, p_i)$ (i.e.
+Why is the state of a quantum particle described by a wave function (i.e.
+a vector in a complex inner product space), while the state of a classical particle is described by the pairs $(q^i, p_i)$ (i.e.
 a point in a cotangent bundle)?
@@
 
-To our surprise, this naturally expanded the boundary of the work.
+Each new link often shrank the complexity of the arguments while expanding the scope of the work in a way that was not at all anticipated.
 The same link that allowed to justify Hamiltonian dynamics also gave us relativistic motion on the extended phase space.
-The same link that allowed to justify Lagrangian mechanics also onstrained the motion to that of free particles under scalar/vector potential forces.
+The same link that allowed to justify Lagrangian mechanics also constrained the motion to that of free particles under scalar/vector potential forces.
 We also realized that sometimes different mathematical concepts could be justified from the same physical starting point.
@@
 
-The success of the approach lead us to think that, maybe, the picture we were developing was not simply ``a nicer way to think about things", but instead a better framework to understand and link different parts of basic science and mathematics.
-After many years of work and different incremental iterations we think it is time to checkpoint and report the progress.
-
-The goal of this paper is to show that it is possible to derive classical/quantum Hamiltonian/Lagrangian mechanics starting from a few physical principle.
+The unexpected success of the approach lead us to think that, maybe, the picture we were developing was not simply ``a nicer way to think about things", but instead a better framework to understand and link different parts of basic science and mathematics.
+After many years of work and different incremental iterations, we think the arguments are well developed enough to report the progress.
+
+The goal of this paper is to show that it is possible to derive classical/quantum Hamiltonian/Lagrangian mechanics starting from few physical assumptions.
 We do \emph{not} claim that we have perfectly achieved this result, with no room from improvement.
@@
 
-While trying to re-derive classical, quantum and (to some extent) relativistic mechanics may seem too broad a scope, it is necessary to ensure that the concepts we develop are truly general, that they can be applied to many different cases.
+While trying to re-derive classical, quantum and (to some extent) relativistic mechanics may seem too broad a scope, it is necessary to show that the concepts we develop are truly general, that they can be applied to many different cases.
 If we could derive classical Hamiltonian mechanics on assumptions that are inapplicable to the quantum case, what good would it do?
-Quantum system are Hamiltonian systems, so clearly we haven't understood what makes a system Hamiltonian.
+Quantum system are Hamiltonian systems, so clearly we wouldn't have understood what makes a system Hamiltonian.
 The ability to apply the same physical ideas far and wide is what gives us confidence that these are truly general concept that capture the physical aspects we are describing.
@@
 As a brief overview, there are four assumptions we will consider.
-Assuming deterministic and reversible evolution will give us a dynamical system: a state and a law of evolution.
+Assuming deterministic and reversible evolution will give us a dynamical system: states and a law of evolution.
 Additionally assuming infinitesimal reducibility, studying the whole system is equivalent to studying its infinitesimal parts, gives us Hamiltonian mechanics.
@@
 
-In section \ref{sec:dynamical_systems} we will show how the physical concept of physical distinguishability maps to the mathematical concept of topological spaces.
+In section \ref{sec:dynamical_systems} we will show how the physical concept of physical distinguishability leads to the mathematical concept of topological spaces.
 Deterministic and reversible evolution is an invertible continuous map (i.e.
@@
 position and velocity).
-This link will allow us to express the number of possible initial conditions for d.o.f.
-in terms of $dx^i$ and $du^i$ (leading to the metric tensor $g$).
+This link will allow us to express the number of possible initial conditions for a d.o.f.
+in terms of $dx^i$ and $du^i$, leading to the metric tensor $g$.
 Such expression also allows to constrain the Hamiltonian to the one for massive particles under scalar potential forces and an equivalent Lagrangian formulation.
@@
 
-The overall layout of the work is a balance between the need to show that the derivation follows (which requires starting from basic principles and reaching conclusion using formal arguments) and the need to give some intuitive understanding (which often requires starting from specific cases and then generalizing).
+The overall layout of this work is a balance between the need to show that the derivation follows (which requires starting from basic principles and proceeding using formal arguments) and the need to give an intuitive understanding (which often requires starting from specific cases and then generalizing).
 
@@
 A \textbf{rationale} follows each assumption, which uses physical and sometimes philosophical arguments to motivate why (or why not) such an assumption makes sense (in a particular case).
+
   \item[Propositions] these characterize the mathematical properties of the physical concepts we are studying.
-A \textbf{justification} follows each proposition to show why such characterization follows from the physical properties of the system.
+A \textbf{justification} follows each proposition to show why such characterization captures the particular physical properties of the system.
 While the style of a justification is formal, it is not a pure mathematical proof as it necessarily has to mix physical and mathematical arguments.
@@
 
-As these sections are clearly demarcated, this allows us to make sure that each justification only references past propositions.
-It also makes it easier to read the work at the desired level of detail.
+As these sections are clearly demarcated, they allow us to make sure that each justification only references past propositions. It also makes it easier to read the work at the desired level of detail.
 
@@
 The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology, measure theory, differential geometry (symplectic and Rimannian), group theory and vector spaces, statistics, information theory, Hamiltonian mechanics, Lagrangian mechanics, quantum mechanics, thermodynamics, special relativity and so on.
-Connecting these areas, in fact, is part of the insight given by this work.
-While this is a necessity, it does create a number of problems.
+Using idea from all these areas is necessary, in fact unraveling their connections is part of the value of this work, but it does create a number of problems.
 
@@
 Yet we cannot introduce all of them either.
-The compromise we adopted is to keep the informal discussion more accessible while keeping the formal derivation more precise.
-Even in the formal derivation, though, we will seldom need more than the basic definitions within a specific area and few major results.
+The compromise we adopted is to keep the informal discussion more accessible while keeping the formal derivation more precise, without needing more than the basic definitions and few major results within each specific area.
+This may leave an expert in a particular field not fully satisfied but, as we mentioned in the introduction, our goal is to lay out the general picture.
 
@@
 As each field has developed its own set of conventions, the common usage is now an empty set.
-We ended up with a compromise that we believe strikes a good balance.
-Especially in the informal discussion, we crafted something that in general feels familiar to a physicist.
-Within formal parts, though, we may use notation and terminology more specific to the area of math being used.
-But then again: one would need to be familiar to that area anyway to be able to follow the details.
-
-We have tested the prose on a number of students and colleagues to be confirm its accessibility at different levels.
+We ended up with a compromise that hopefully feels natural enough to a broad audience of physicist, mathematical physicists and philosophers of physics.
+Within formal parts we may use notation and terminology more specific to the area of math being used, but one would need to be familiar with them anyway to be able to follow the details.
+
+We have tested the prose on a number of colleagues and students to confirm its accessibility at different levels and believe we have struck a decent compromise.
 We hope the reader enjoys this journey and comes out with a few new insights and ideas.

### 2016-12-20 2015CompleteDerivation/completeDerivation.tex
Another pass at discussion and conclusions

@@
 
-The first remarkable aspect of this work is that it can exists.
-Regardless of whether all the minor details are addressed in the best possible way, the idea that so much can be derived from so little is unexpected.
-
-It often happened that the inability to justify one detail forced a partial shift in thinking that enabled to connect more ideas more simply.
-The ability to constrain the Hamiltonian to vector potential came as a surprise after we realized we needed distributions as a core concept to justify Hamilton's equations.
-The fact that relationships must be continuous functions came as a surprise after we connected topology with the notion of physical distinguishability.
-It was this constant simplification that shows us that we were on the right track and that, in the end, decided the boundary of the work as the initial goals were very modest in comparison.
-
-The second striking aspect is a shift in perspective that this work forces upon us.
+The first remarkable aspect of this work is that something along these lines can be done.
+Regardless of whether all the minor details are addressed in the best possible way, the idea that so much can be derived from so little was not something we expected.
+In hindsight, the idea that physics should be founded on physical principles is nothing new: it just hasn't been thoroughly applied to basic mechanics.
+As these mathematical frameworks have developed across two centuries, it maybe should not be so surprising that they are ripe for some reorganization.
+
+The second striking aspect is a shift in perspective that this work seems to suggest.
 The usual reductionist view is that you start with the rules of the small parts and then build up the composite whole.
@@
 A system is not deterministic per se: it is a particular choice for its description (i.e.
-state) that may be deterministic under some evolution.
+state) that may undergo deterministic evolution in some cases.
 In the same vein, a material is not homogeneous per se.
@@
 Connecting the dots, the next assumption would be a inhomogeneous irreducible material: where the material is not in constant equilibrium and its particles are not composed of the same amount.
-This different starting point may lead to the resolution of some of the problems we mentioned.
-
-As different particle can transform into one another, they could be seen as different stable states of the same inhomogeneous material.
+This different starting point may have potential connections to some of the problems we mentioned in the rationales for the assumptions and to other interesting physics questions in general.
+
+As different particles can transform into one another, they could be seen as different stable states of the same inhomogeneous material.
 In particular, since $\sigma_x \sigma_v \geq \frac{\hbar}{2m}$, particles in different generations (i.e.
@@
 
-In the near future it would seem appropriate to extend this work to the remaining part of well-established physics.
-The ideas toward that goal are more concrete.
+We limit the discussion to these few remark as the work is already long.
+We hope to extended it to field theories in the near future.
 The state space for a quantized material, already introduced here, should be captured by the Fock space.
@@
 Assuming that the vector potential forces themselves are dependent on the state of a material should hopefully give us the Hamiltonian for the fields.
+Whether this can be done will be a matter of working through the details, which, we are sure, are likely to bring other unexpected surprises.
 
@@
 
-We have seen that much of fundamental physics can be derived from a few simple physical assumption (and lots of hard work).
-In doing so we obtained a more unified picture of the different branches of fundamental physics centered around the idea of deterministic and reversible evolution of distributions of a homogeneous material.
+We have seen that much of fundamental physics can be derived from few simple physical assumption (and lots of hard work).
+In doing so we obtained a more unified picture of the different branches of fundamental physics centered around the idea of deterministic and reversible evolution of distributions of homogeneous materials.
 The main physical concepts can be summarized by the following points:
@@
 Therefore it is an isomorphism on the mathematical structure defined by the state space (e.g.
-homeomorphism, canonical transformation, unitary transformation), not just a state-to-state bijection.
+homeomorphism, canonical transformation, unitary transformation), not just a state-to-state bijection (i.e.
+isomorphism of sets).
 
 \item Classical Hamiltonian mechanics describes the evolution of a homogeneous material whose state is equivalent to the states of all its infinitesimal parts, the classical particles.
-The state space for classical particles is a symplectic manifold $(T^*\mathcal{Q}, \omega)$ as this allows coordinate invariant density distributions; deterministic and reversible evolution for such state space is a symplectomorphism (i.e.
+The state space for classical particles is a symplectic manifold $(T^*\mathcal{Q}, \omega)$ as this allows to define coordinate invariant density distributions; deterministic and reversible evolution for such state space is a symplectomorphism (i.e.
 a canonical transformation).
@@
 \item Quantum (Hamiltonian) mechanics describes the evolution of a homogeneous material for which the state can only be reduced up to finite parts, the quantum particles, whose internal motion is undetermined.
-The state space for quantum particles is an inner vector product space $(\Psi, \langle \cdot , \cdot \rangle)$; deterministic and reversible evolution for such state space is a unitary transformation.
+The state space for quantum particles is a complex inner product space $(\Psi, \langle \cdot , \cdot \rangle)$; deterministic and reversible evolution for such state space is a unitary transformation.
 \end{itemize}
@@
 While we are sure that some mathematical details can be improved, we feel that the overall ``conceptual castle" is solid.
-The starting concepts are simple, the conceptual mapping of those concepts into math is often straight forward and the intuitive ideas that emerge help connecting different branches of math and physics.
+The starting concepts are simple, the conceptual mapping of those concepts into math is often straight forward and the intuitive ideas that emerge help to connect different branches of math and physics.
 This is unlikely a matter of simple coincidence.

### 2016-12-22 2015CompleteDerivation/completeDerivation.tex
A global pass to make the language more homogeneous throughout

@@
 
-We can also conceptually combine two different processes into a single one.
-That is, having a way to measure quantity $x$ and a way to measure quantity $y$ gives us a way to measure the combination $(x,y)$.
+We can also conceptually combine two different processes into a single one.\footnote{That is, take the logical and between outcomes of different processes.} That is, having a way to measure quantity $x$ and a way to measure quantity $y$ gives us a way to measure the combination $(x,y)$.
 For ensembles, if we can measure the marginal distribution $\rho_x(x)$ and the marginal distribution $\rho_y(y)$, we know the joint distribution $\rho(x,y)$ has to be compatible with both.
@@
 
-We can also coarsen a single process.
-That is, having a way to measure $(x,y)$ gives us a way to measure $x$ alone (or any $f(x,y)$).
+We can also coarsen a single process.\footnote{That is, take the logical or between outcomes of the same process.} That is, having a way to measure $(x,y)$ gives us a way to measure $x$ alone (or any $f(x,y)$).
 Formally, the outcomes of the coarsened process are given by performing the union of some outcomes of the original process.
@@
 With our state space defined, deterministic and reversible evolution corresponds to a bijective map between initial and final states.
-Moreover, a physical process that can distinguish final states is also a physical process that can distinguish initial states.
-That is: the topology is mapped and preserved by the deterministic and reversible evolution (i.e.
-initial and final states are equally distinguishable).
+But that's not enough.
+A physical process that can distinguish final states can be used, together with the evolution, to distinguish initial states.
+We prepare the initial state, let it evolve deterministically, measure the final state and use reversibility to convert the measure to one on the initial state.
+Therefore, not only the evolution is a bijection, but it also maps outcomes to outcomes.
+That is: the topology is mapped and preserved by the deterministic and reversible evolution because physical distinguishability must remain unchanged.
 The evolution is then a self-homeomorphism on the state space.
@@
 
-The generality of this result explains why in physics one always assumes functions to be well behaved.
+The generality of this result explains why in physics one always assumes functions to be ``well behaved".
 As the result was derived from our notion of physical distinguishability, this is not a matter of practical convenience.
@@
 But it also allows us to create an outcome compatible with only a single point.
-A finite precision measurement of the force provides us infinite precision of space, which we ruled out.\footnote{This assumes that we are able to position the probe perfectly, which we could do if we were able to manipulate forces at that precision.} This is what the math is telling us, that if we said before that outcomes can't distinguish isolated points (i.e.
-the topology is second countable) then neither can maps.
+A finite precision measurement of the force provides us infinite precision of space, which we ruled out.\footnote{This assumes that we are able to position the probe perfectly, which we could do if we were able to manipulate forces at that precision.} This is what the math is telling us, that if we claim that outcomes can't distinguish isolated points (i.e.
+standard topology on $\mathbbm{R}^n$) then neither can maps.
 It is physical consistency that limits us to continuous functions.
@@
 We call a state variable \emph{discrete} or \emph{continuous} if the possibilities are integer or real numbers respectively.
-From now on, we are going to consider state spaces whose states can be identified, at least within a region, by a finite set of discrete and continuous state variables (i.e.
-the state space is locally isomorphic to $\bigcup\limits_{1 \leq i \leq n} \mathbbm{R}^{m_i}$).\footnote{TODO: Is there a more physically meaningful requirement?
-Could we have a set of states that are not locally described by a set of quantities?}
+The topology used reflects the notion that integers can be measured perfectly (i.e.
+discrete topology) while real numbers can be measured up to a finite precision (i.e.
+standard topology on $\mathbbm{R}$) From now on, we are going to consider state spaces whose states can be identified, at least within a region, by a finite set of discrete and continuous state variables (i.e.
+the state space is locally isomorphic to $\bigcup\limits_{1 \leq i \leq n} \mathbbm{R}^{m_i}$).
 
@@
 The only requirements for state variables is that they identify states.
-This means the possible values for state variables are physically distinguishable and are defined at equal time (since states are defined and mapped at a particular time).
-\footnote{This is the main reason that a quantity like velocity is not a suitable state variable, as it is  defined over an interval of time.
+This means the possible values for state variables are physically distinguishable and are defined at equal time (since states are defined and mapped at a particular time).\footnote{This is the main reason that a quantity like velocity is not a suitable state variable, as it is  defined over an interval of time.
 Therefore velocity is always physically well defined but is not a state variable in general, while conjugate momentum is always a state variable but is not physically well defined by itself (it requires the vector potential to be specified as well).
-As we'll see later, if there exists a one to one map between velocity and conjugate momentum they can both be physically well defined state variables.}
+As we'll see later, when there exists a one to one map between velocity and conjugate momentum they can both be physically well defined state variables.}
 
@@
 The choices are not independent.
-That is: two or more variables are independent if there always exists a state for any possible combination, if the total number of states is the cartesian product of the possibilities of each variable.
+That is: two or more variables are independent if there always exists a state for any possible combination, if the total number of states is the product of the possibilities of each variable.
 We call \emph{state vector} a collection of independent state variables that fully identify a state.
@@
 So we introduce a new set of variables for which $\hat{q}^i(\hat{\mathcal{s}})=q^i(\mathcal{s})$, which we can always do as the evolution is deterministic and reversible.
-The level sets of $\hat{q}^i$ are the evolved level sets of $q^i$ (e.g.
+The hyper-surfaces at constant $\hat{q}^i$ are the evolved hyper-surfaces at constant of $q^i$ (e.g.
 all the states that started with a particular value for pressure) which allows us to study how groups of states evolve in time.
@@
 In our case, the state is always changing.
-The choice is between tracking the change using different state variables or different values.}
+The choice is between tracking the change with different values of the same state variable or different state variables that give the same value.}
 
@@
 
-To show the usefulness of transported state variables, we can prove the following:
+One reason that transported variables will be useful is that, during deterministic and reversible evolution, we need to make sure that the number of initial and final possibilities across independent state variables remains the same.
+This is more easily done using transported variable as the actual values do not change.
+In the discrete case, this is easy to show.
 
@@
 
-The issue is that while deterministic evolution will map states to states, not all bijective state to state maps can be considered deterministic and reversible.
-This is a crucial point where this work is different from more usual treatments of dynamical systems.
+The issue is that while deterministic evolution will map states to states, not all bijective state to state maps can be considered deterministic and reversible.\footnote{In fact, we have already seen that the map must at least be continuous.} This is a crucial point where this work is different from more usual treatments of dynamical systems.
 If one focuses only on the states themselves and counts them, it reaches the conclusion that a range of $1$m contains the same number of initial conditions of $1$Km: they have the same infinite number of elements.
@@
 If we work with a discrete topology then, by all means, the way to measure the size of a set is to count the elements in a set.
-But if we use the standard topology on $\mathbbm{R}$, isolated points have measure zero: what defines the number of initial conditions is the range of the state variables.
+But if we use the standard topology on $\mathbbm{R}$, isolated points have measure zero (i.e.
+we can't physically distinguish them): what defines the number of initial conditions is the range of the state variables.
 As we'll see, the geometrical structures in Hamiltonian/Lagrangian classical/quantum mechanics are there to allow us to properly count states and, more specifically, possibilities along independent state variables.
@@
 What we'll find is that under the additional assumption that the material is infinitesimally reducible (i.e.
-its state is equivalent to the states of its parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian mechanics.
+its state is equivalent to the states of its infinitesimal parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian mechanics.
 
 \subsection{Homogeneous decomposable systems and vector spaces}
-The notion that a system is decomposable means the states are equipped with a rule of composition that allows one to write $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$: the composite system is the sum of its parts.
+The notion that a system is decomposable means the states are equipped with a rule of composition that allows one to write $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$: the composite system is the sum of its components.
 For example, the state of a ball is equal to the state of its top and bottom parts.
@@
 The domain and codomain match because the material is homogeneous.
-The law is commutative $\mathcal{c}_1 +\mathcal{c}_2 = \mathcal{c}_2+\mathcal{c}_1$ and associative $(\mathcal{c}_1 + \mathcal{c}_2) + \mathcal{c}_3 = \mathcal{c}_1 + (\mathcal{c}_2 + \mathcal{c}_3)$, as it does not matter in what order we physically compose the parts.
+The law is commutative $\mathcal{c}_1 +\mathcal{c}_2 = \mathcal{c}_2+\mathcal{c}_1$ and associative $(\mathcal{c}_1 + \mathcal{c}_2) + \mathcal{c}_3 = \mathcal{c}_1 + (\mathcal{c}_2 + \mathcal{c}_3)$, as it does not matter in what order we physically compose the components.
 There exists a unique zero element $\mathcal{c} + 0 = \mathcal{c}$ and it represents the physically empty state (i.e.
@@
 We will call a \emph{particle} the smallest amount of a material to which we can assign an independent state.
-For example, a photon is a particle of light (it is the smallest amount of light we can describe), an infinitesimal amount of water is treated classically as a point particle (the smallest amount for a continuous fluid).\footnote{We use this definition as it allows us to retroactively talk on somewhat equal grounds about classical point-particles and quantum particles, and underscore how physically (and mathematically) they are very similar yet very different objects.}
+For example, a photon is a particle of light (it is the smallest amount of light we can describe), an infinitesimal amount of water is treated classically as a point particle (the smallest amount for a continuous fluid).\footnote{We use this definition as it allows us to retroactively talk on somewhat equal grounds about classical particles and quantum particles, and underscore how physically (and mathematically) they are very similar yet very different objects.}
 
@@
 That is, we can keep decomposing the system indefinitely into smaller and smaller parts, each with a well defined state.
-In this case, the particles are the infinitesimal parts given by the limit of this process of recursive decomposition.
+In this case, the particles are the infinitesimal parts given by the limit of this process of recursive reduction.
 Giving the state of the whole system is equivalent to giving the states of all particles.
@@
 And the density will give us an indication of how big the region is: if we double the region, the density will halve.
-In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique measure\footnote{Here measure is intended in the mathematical sense (a real valued function of a $\sigma$-algebra) which is distinct from other connotations in physics or probability theory.} $\mu$ for the size of sets of $\mathcal{S}$.
+In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique measure\footnote{Here measure is intended in the mathematical sense (a real valued function of a $\sigma$-algebra) which is distinct from other connotations in physics.} $\mu$ that allows to count the number of states in each set of $\mathcal{S}$.
 In terms of said measure, positive linear functionals become integrals, $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
@@
 In the same way that discontinuous changes of state variable can be detected because they tamper with physical distinguishability (i.e.
-the topology), non differentiable changes of state variable can be detected because they tamper with our ability to define state-variable-independent densities.
+the topology), non differentiable changes of state variable can be detected because they tamper with our ability to define state-variable-independent densities and count states.
 
@@
 
-Another consideration is that while the norm associated to $L^1(\mathcal{S})$ is $\int_{\mathcal{S}} |\rho| d\mu$, the expression $\int_{\mathcal{S}} \rho d\mu$ is perhaps more physically meaningful.
+Another consideration is that while the norm associated to $L^1(\mathcal{S})$ is $\int_{\mathcal{S}} |\rho| d\mu$, the expression $\int_{\mathcal{S}} \rho d\mu$ may be more physically meaningful in some cases.
 For proper states, the two are the same and they represent the total amount of material.
@@
 For each state, they represent respectively the total amount of material, the material within $U$, the density at $(q_0, p_0)$, the total momentum and the average of a state variable.
-As they are linear functionals of $\mathcal{C}$,  they can be expressed using their dual vector over $\mathcal{S}$.
+As they are linear functionals of $\mathcal{C}$,  they can be expressed using their dual vector.
 But these vectors ($1_U$, $p$, ...) are \emph{not} elements of $\mathcal{C}$.
@@
 That is why one can choose ``natural units" for which $\hbar=1$.
-In SI units the relationship $\hbar dk = dp = m du$ between kinetic and conjugate momentum, derived later, sets the relationship between units.} The cardinality for states in a finite (i.e.
+In SI units the relationship $\hbar dk = dp = m du$ between kinetic and conjugate momentum, derived later, sets the relationship between units.} The state count in a finite (i.e.
 compact) region $U \subseteq \mathcal{S}$ will be given by $\mu(U) = \int_U \hbar dq \wedge dk$: we are finally able to write the volume of integration in terms of state variables.
@@
 We have $\mathcal{S}\cong T^*\mathcal{Q}$ and $d\mu = \hbar^n dq^n \wedge dk_n$.
-But it's not just the volume that is preserved as we change state variables.
-Each pair $(q^i, k_i)$ is expressed by an independent unit with the cardinality of possibilities given by $\hbar dq^i \wedge dk_i$.
+But it's not just the volume, the state count, that is preserved as we change state variables.
+Each pair $(q^i, k_i)$ is expressed by an independent unit with the count of possibilities given by $\hbar dq^i \wedge dk_i$.
 We call such a pair a degree of freedom.
-As they are independent the total number of states is the product of the possibilities of each d.o.f.: $d\mu = \hbar^n dq^n \wedge dk_n = \prod \limits_{i=1}^n \hbar dq^i \wedge dk_i$.
+As they are independent the total state count is the product of the possibilities of each d.o.f.: $d\mu = \hbar^n dq^n \wedge dk_n = \prod \limits_{i=1}^n \hbar dq^i \wedge dk_i$.
 In other words: independent d.o.f.
 are orthogonal surfaces in $\mathcal{S}$.
-The possibilities of an arbitrary degree of freedom, then, will be $\sum \limits_{i=1}^n \hbar dq^i \wedge dk_i$, the sum of the projections over the $n$ orthogonal and independent d.o.f.
+The possibility count of an arbitrary degree of freedom, then, will be $\sum \limits_{i=1}^n \hbar dq^i \wedge dk_i$, the sum of the projections over the $n$ orthogonal and independent d.o.f.
 defined by $(q^i, k_i)$.
 
-The cardinality of possibilities for each d.o.f.
+The possibility count for each d.o.f.
 (i.e.
@@
 We first define conjugate momentum as $p_i=\hbar k_i$ and unified state variables as $\xi^a\equiv \{q^i, p_i\}$.
-Then we consider the canonical two-form $\omega =\Sigma \, dq^i \wedge dp_i$ given by the following components:
+Then we consider the canonical symplectic two-form $\omega =\Sigma \, dq^i \wedge dp_i$ given by the following components:
 \begin{align*}
@@
 and the scalar product across.
-Requiring the invariance of this metric under state variable changes assures us the cardinality of states and possibilities is well defined.
+Requiring the invariance of this metric under state variable changes assures us the count of states and possibilities is well defined.
 
@@
 The canonical one-form $\theta=k dq$ represents the geometrical object we associate with each particle state.
-The canonical two-form $\omega$ is the metric that quantifies the possibilities described by two given state variables.
+The symplectic form $\omega$ is the bilinear that quantifies the possibilities described by two given state variables.
 Note, instead, that the relationship $\omega = - \hbar d\theta$, while mathematically true, has no clear physical meaning as we have not defined what the exterior derivative on a state actually represents.
@@
 This is the only manifold that allows us to define state-variable-invariant densities.
-The metric allows us to measure the cardinality of possibilities on an arbitrary d.o.f.
+The symplectic form allows us to count possibilities on an arbitrary d.o.f.
 
@@
 We should take the use of densities and conjugate quantities as evidence that quantities like space and time are not discrete in the sense that the processes we use to distinguish those quantities cannot identify single instances.
-The quantum case does not change this, as the use of densities (in the form of the wave function) and conjugate quantities is even more prominent.\footnote{One may still posit that space is fundamentally discrete at scales where both classical and quantum mechanics cease to be valid.
-For that to work, the measure defined on such a discrete space has to become the measure on the classical phase space.
-That is: it must be a discrete version of $T^*\mathcal{Q}$ and not $\mathcal{Q}$ by itself.}
+The quantum case does not change this, as the use of densities (in the form of the wave function) and conjugate quantities is even more prominent.\footnote{If one wants to posit that space is fundamentally discrete at scales where both classical and quantum mechanics cease to be valid, the better strategy would be to work with a discrete version of $T^*\mathcal{Q}$ and not $\mathcal{Q}$ by itself.
+Though it is not clear how a discrete topology would become non discrete in the limit, at least the measure should be workable.}
 
@@
 Finally, we should note that we are in a position similar to the one discussed in \ref{discrete_state_metric} for discrete states.
-We have both a topological space and a measure expressible in terms of state variables that allows us to count states and possibilities.
-We recovered the metric starting from the idea of an infinitesimally reducible system.
-That led to a state space $\mathcal{S}$ for the infinitesimal parts, on which we must be able to define state-variable-independent distributions, which in turn gave us degrees of freedom made by pairs of conjugate variables and the symplectic metric typical of classical phase space.
-In short: being able to measure the amount of material is what allows us to measure the cardinality of states.
+We have a topological space plus a measure that allows to count states and a symplectic form that allows to count possibilities, both expressible in terms of state variables.
+We recovered this structure starting from the idea of an infinitesimally reducible system.
+That led to a state space $\mathcal{S}$ for the infinitesimal parts, on which we must be able to define state-variable-independent distributions, which in turn gave us degrees of freedom made by pairs of conjugate variables and the symplectic form typical of classical phase space.
+In short: being able to measure the amount of material is what allows us to count states.
 
@@
 
-But probably the best way to look at it is that initial and final sets of states have to possess the same cardinality of states and possibilities, which were defined by our metric $\omega$.
+But probably the easiest way to look at it is that initial and final sets of states have to possess the same count of states and possibilities as defined by the symplectic form $\omega$.
 Therefore the area within a degree of freedom, which represents the number of possibilities in said d.o.f., needs to be mapped to an equal area within the transported d.o.f.
@@
 Mathematically, under the classical Hamiltonian assumption,  deterministic and reversible evolution is a self-symplectomorphism (or self-isometry or canonical transformation depending on your math training).
-That is, it does not just preserve the topology but also the symplectic metric $\omega$.
+That is, it does not just preserve the topology but also the symplectic form $\omega$.
 
@@
 
-The discussion so far has been limited to the case where both the set of states and the evolution map never change throughout the evolution.
+The discussion so far has been limited to the case where both the state variables and the evolution map never change throughout the evolution.
 This is too restrictive as there are very reasonable situations in which this does not hold.
@@
 This does not affect our definition of determinism and reversibility as we still can tell final state from initial state and vice-versa.
-But at this point we have no way to specify a time dependent process: the Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ as we derived it is just a function of the state.
+But at this point we have no way to specify a time dependent evolution: the Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ as we derived it is just a function of the state.
 
@@
 Our composite state distribution $\rho(q)$ becomes $\rho(\hat{q},t)$: it is no longer defined at equal time.
-This means that our measure $\mu$ needs to be modified to define integrals over the time variable.
+This means that our measure $\mu$ needs to be modified to define integrals over the time variable.\footnote{Looking ahead, another way to frame the problem is that Galilean transformations do not preserve phase space volume.
+Lorentz transformations, on the other hand, do and it is this requirements plus the kinematic assumption that will recover them later.}
 
 As we can see, the framework we have is ill suited to handle these cases.
-Therefore we cannot simply stick time in the Hamiltonian and expect everything to work out.\footnote{Looking ahead, another way to frame the problem is that Galilean transformations do not preserve phase space volume.
-Lorentz transformations, on the other hand, do and it is this requirements plus the kinematic assumption that will recover them later.} But we can't ignore the problem either as the three cases outlined are pretty common situations to study.
+Therefore we cannot simply stick time in the Hamiltonian and expect everything to work out.
+But we can't ignore the problem either as the three cases outlined are pretty common situations to study.
 So we need to go back to our definitions and amend them properly.
@@
 
-We now want to find the metric for our complete state space $\bar{\mathcal{S}}$.
+We now want to extend the symplectic form so we can count state and possibilities over the complete state space $\bar{\mathcal{S}}$.
 The first thing to do is to complete our $n$-dimensional $\mathcal{Q}$, the manifold that defines our units, to include the time variable $t$.
@@
 The co-vector $k_i dq^i$ is completed with a time component $\bar{\omega} dt$.\footnote{$\bar{\omega}$ is the classical analogue of the wave frequency.
-We use $\bar{\omega}$ to distinguish from the phase-space metric $\omega$.} As each state variable $q^i$ has a conjugate quantity $p_i\equiv \hbar k$, the time variable $t$ will have a conjugate quantity $E\equiv\hbar\bar{\omega}$.
+We use $\bar{\omega}$ to distinguish from the symplectic form $\omega$.} As each state variable $q^i$ has a conjugate quantity $p_i\equiv \hbar k$, the time variable $t$ will have a conjugate quantity $E\equiv\hbar\bar{\omega}$.
 We call the combination of $(t, E)$ the temporal degree of freedom.
-As time is special, so is the temporal degree of freedom, which is treated differently when defining a distribution $\rho_\mathcal{c}$ and therefore by the measure $\mu$ and the metric $\omega$.
+As time is special, so is the temporal degree of freedom, which is treated differently when defining a distribution $\rho_\mathcal{c}$ and therefore by the measure $\mu$ and the symplectic form $\omega$.
 
@@
 And the plane of constant $( q, p )$ is where the area given by $dt \wedge dE$ is defined.
-That is: the plane where we can properly count states and define our metric is perpendicular to $dt \wedge dE$.
+That is: the plane where we can properly count states and define our symplectic form $\omega$ is perpendicular to $dt \wedge dE$.
 Intuitively, states are defined independently of time, therefore they are defined on a surface perpendicular to the temporal d.o.f.
 
-We have a right triangle-like relationship between the plane where the metric is defined and its projections on the planes defined by each d.o.f., similar to the multiple d.o.f.:
+We have a right triangle-like relationship between the plane where $\omega$ is defined and its projections on the planes defined by each d.o.f., similar to the multiple d.o.f.:
 \begin{align*}
@@
 But in the previous case, the right angle was between the two independent d.o.f..
-In this case, the right angle is between the metric and the plane of constant $(q, p)$ where $dt \wedge dE$ is defined.
+In this case, the right angle is between $\omega$ and the plane of constant $(q, p)$ where $dt \wedge dE$ is defined.
 We rewrite it as $dq \wedge dp - dt \wedge dE = \omega$.
@@
 	\includestandalone[width=\columnwidth]{pictures/temporalDof}
-	\caption{The metric $\omega$ and its components along temporal and standard degrees of freedom.
+	\caption{The symplectic form $\omega$ and its components along temporal and standard degrees of freedom.
 The two d.o.f.
@@
 The temporal d.o.f.
-is orthogonal to the metric as states are defined (and counted) independently of time.
+is orthogonal to the symplectic form as states are defined (and counted) independently of time.
 Compare with FIG.~\ref{fig:independent_dof}.
@@
 
-The metric $\omega$ is still the canonical two form expressed with the mathematically non-canonical, but physically meaningful, variable $E$.
+The symplectic form $\omega$ is still the canonical two form expressed with the mathematically non-canonical, but physically meaningful, variable $E$.
 This allows us to better understand its physical meaning.
-The metric allows us to count the possibilities within a degree of freedom, adjusting the metric on $T^*\mathcal{M}$ to avoid the counting problems introduced by time variable changes.
-
-The metric $\omega$ also allows us to distinguish temporal from standard d.o.f.
+$\omega$ allows us to count the possibilities within a degree of freedom, adjusting it on $T^*\mathcal{M}$ to avoid the counting problems introduced by time variable changes.
+
+The symplectic form $\omega$ also allows us to distinguish temporal from standard d.o.f.
 Given a two-dimensional surface $U \subset T^*\mathcal{M}$, how can we tell if it should be charted by a $(t,E)$ or whether we can define a distribution on it?
@@
 This is the manifold that allows us to define time-and-state-variable-invariant densities.
-The metric $\omega$ allows us to measure the cardinality of possibilities.
+The symplectic form $\omega$ allows us to measure the cardinality of possibilities as in the time independent case.
 
@@
 Let $E\equiv\hbar \bar{\omega}$.
-Let $\omega$ be the metric.
+Let $\omega$ be the symplect form.
 The components of $\omega$ across different d.o.f.
@@
 The trajectory of a particle in the extended phase space will be given by the evolved variables $\xi^a(s)$ in terms of a parameter $s$.
-As before, deterministic and reversible evolution will preserve the metric as the number of possibilities on each independent d.o.f.
+As before, deterministic and reversible evolution will preserve $\omega$ as the possibility count on each independent d.o.f.
 will be conserved.
@@
 \begin{prop}\label{lagrangian_mechanics}
-	Let $\mathcal{H}$ be an extended Hamiltonian defined on $\mathbf{T}^*\mathcal{M}$.
+	Let $\mathcal{H}$ be an extended Hamiltonian defined on $T^*\mathcal{M}$.
 Under the kinematic assumption, we can define the Legendre transform $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$ which we call the \emph{extended Lagrangian}.
@@
 To do that, we need to be able to count trajectories in the same way we are able to count states.
-There must exist a metric $g$, consistent with $\omega$, that allows one to calculate the cardinality of possibilities for initial conditions.
-That is: $\omega(dq^\alpha,dp_\alpha) \propto g(dx^\alpha,du^\alpha)$ where the arguments are the value ranges for the respective quantities.
-
-Since both $dx^\alpha$ and $du^\alpha$ are four vector components and the metric must be linear, $g$ is a rank two tensor on $\mathcal{M}$.
+Just as the symplectic form $\omega$ gives us the count of possibilities for a d.o.f.
+in terms of $dq^\alpha$ and $dp_\alpha$, the ranges of the state variables, there should be a bilinear form $g$ that gives the the count of possibilities in terms of $dx^\alpha$ and $du^\alpha$, the ranges of the kinematic variables.
+Under the kinematic assumption, these counts must be consistent with each other.
+That is: $\omega(dq^\alpha,dp_\alpha) \propto g(dx^\alpha,du^\alpha)$.
+
+Since both $dx^\alpha$ and $du^\alpha$ are four vector components, $g$ is a rank two tensor on $\mathcal{M}$.
 To each direction in space-time $dx^\alpha$ is associated a direction in the tangent space of velocities $du^\alpha$ with which it forms a d.o.f..
@@
 We also know that $g$ cannot be alternating, that is $g(dx^\alpha, du^\alpha) \neq 0$: a choice of units for $x^\alpha$ is a choice of units for $u^\alpha$ and therefore they cannot belong to different d.o.f.
-This means $g$ is symmetric.
-$\mathcal{M}$ is a pseudo-Riemannian manifold.
+This means $g$ is a symmetric non-degenerate tensor.
+We recognize it as the metric tensor and therefore $\mathcal{M}$ is a pseudo-Riemannian manifold.
 
@@
 We can further change the units along the spatial directions so that $g_{ii}=1$: if we use the same units for the spatial coordinates, each direction will contribute the same number of possibilities.
-Since time must use a different unit, we set $g_{00}=-c^2$ where $c^2$ is the constant that allows one to convert the possibilities over the temporal d.o.f.
-into the possibilities of a spatial d.o.f..
+Since time must use a different unit, we set $g_{00}=-c^2$ where $c^2$ is the constant that allows one to convert the possibility count over the temporal d.o.f.
+into the possibility count of a spatial d.o.f..
 Therefore, locally $g(dx^i, dx^i) = (dx^i)^2 - c^2 dt^2$ is Minkowskian.
@@
 
-The metric tensor $g$ and the constant $c$ found here are indeed the ones used in the context of relativity.
+The metric tensor $g$ and the constant $c$ found here are indeed the ones used in the context of special relativity.
 Yet, they have acquired a different meaning.
-$g$ does not measure space-time distances: it measure cardinality of possibilities, like $\omega$.
+$g$ does not measure space-time distances: it counts possibilities of a d.o.f., like $\omega$.
 Directions in space are orthogonal not because the angle between them is $90^\circ$, but because they identify independent d.o.f..
@@
 Therefore they do not have anything to violate: the maximum speed does not apply to them.
+On the other hand, only deterministic and reversible motion can carry information and can establish before/after relationships, therefore everything remains consistent.
 
@@
 They allow to keep proper book keeping when mapping sets of states.
-Using natural units means measuring all these quantities through the cardinality of possibilities.
+Using natural units means measuring all these quantities through the possibility count.
 There is a difference among them, though: while $c$ and $\hbar$ are simply determined by the unit system, $m$ also depends on the type of material.
@@
 
-Another important thing to notice is that the relationship between conjugate momentum and velocity, $p_\alpha = m u_\alpha + A_\alpha(x)$, includes an arbitrary function.
+Another important thing to note is that the relationship between conjugate momentum and velocity, $p_\alpha = m u_\alpha + A_\alpha(x)$, includes an arbitrary function.
 As it is arbitrary, we cannot give $p$ a well defined physical meaning in terms of the space-time trajectory.
@@
 \end{align*}
+Note how the constraint $\mathcal{H}=0$ simply sets the rest kinetic energy.
 Note how the Poisson bracket between kinetic momentum is the force tensor.
@@
 
+Again we want to stress just how the math takes a very simple and physically meaningful form when studying the motion in phase space using position and kinematic momentum.
+It is most expressive because, by using the kinematic variables of a deterministic and reversible motion of an infinitesimal amount of material, it makes use of all the assumptions about the system we are studying.
+
 To sum up, the kinematic assumption has more inherent consequences one would expect at first.
 It means the particles are identified by position and velocity and therefore the system of equations is second order.
-It links the space of trajectories on $T\mathcal{M}$ with the space of states on $T^*\mathcal{M}$.
-Such link allows state space metric $\omega$ to induce the space-time metric $g$.
+It links the space of initial conditions on $T\mathcal{M}$ with the space of states on $T^*\mathcal{M}$.
+Such link allows phase space symplectic form $\omega$ to induce the space-time metric tensor $g$.
 It allows us to define Lagrangians.
@@
 Any finite amount of classical material, therefore, is described by a distribution.
-
 If we take a quantum particle, say an electron, it has a finite amount of charge/mass/energy.
 Therefore, to a first approximation, it must be described by a distribution as well.
+
 The idea that an electron is not a point particle may, to some, seem preposterous at first.
@@
 
-As the state evolves into its final state, the amount of material remains the same: $\int \rho_{\hat{\mathcal{c}}} dq \wedge dp = \int \rho_{\mathcal{c}} dq \wedge dp$.
+As the state evolves into its final state, the amount of material remains the same: $\int \rho_{\hat{\mathcal{c}}} dq \wedge dp = 1$.
 The final entropy will also remain the same: $I(\rho_{\hat{\mathcal{c}}}) = I_0$.
@@
 At each instant we can assume $\rho$ to be at equilibrium.
-At each instant the fragments are reshuffled but the amount of entropy of the distribution remains the same.
+At each instant the fragments are reshuffled.
+At each instant the entropy of the distribution is the same one associated with the random motion.
 The overall evolution is still Hamiltonian, but the trajectories predicted by the framework are now just the average motion of the fragments.
@@
 It is time to turn these concepts into a precise mathematical formulation as we did for classical mechanics.
-We want to formalize the state space for an \emph{irreducible material} as a quantum particle will be a finite amount of such material.
+We want to formalize the state space for an \emph{irreducible material} as a quantum particle will be a unit amount of such material.
 
@@
 For a classical material, the infinitesimal parts can be given a state, have their own topological space and a set of local state variables $(q^i, p_i)$ to identify them.
-We can use the richness of differential geometry and metric spaces.
+We can use the richness of differential geometry.
 For an irreducible material, the infinitesimal parts cannot be given a state, cannot be fully identified by a set of variables.
@@
 We need to be able to express distributions and expectations for different quantities: this will give us self-adjoint linear operators.
-We then need to identify the self-adjoint linear operators for the quantities $k_i$ and $p_i$ and to be able express the distribution of material in terms of them.
+We then need to identify the self-adjoint linear operators for the quantities $k_i$ and $p_i$ and to be able express the distribution of material in terms of those.
 Let's start with the first.
@@
 We know that the operators have to be covariant: under a change of units $\hat{q}^j = \hat{q}^j(q^i)$ we must have $K_{\hat{q}^j[q^i]} = \partial_{\hat{q}^j} q^i K_{q^i [q^i]}$.
-We are looking for linear operators that obey the chain rule, which turn out to be the derivation by the corresponding $q^i$.
+We are looking for linear operators that obey the chain rule, which turns out to be the derivation by the corresponding $q^i$.
 They have to be self-adjoint, so they will be of the form $K_{i [q^i]} = a \imath \partial_{i}$ where $a$ is a real number.
@@
 
-It should be clear, though, how the correspondence between mathematical and physical objects is not as satisfying as in the classical case.
+It should be evident, though, how the correspondence between mathematical and physical objects is not as satisfying as in the classical case.
 For instance, while any differentiable function of $T^*\mathcal{Q}$ can be a state variable for a classical particle, not all self-adjoint linear operators can be associated with a fragment variable.
@@
 
-	Using unitary evolution in those cases to make predictions, though, may still make physical sense.
+	As we'll discuss in a later subsection, using unitary evolution in those cases to make predictions, though, may still make physical sense.
 Quantities that are not affected by the particular evolution of the unstated part will have, by their very nature, the same value under non-deterministic evolution and an equivalent quasi-static evolution.
@@
 Also note that $H$ operates on $\Psi$ which, in the case of $\mathcal{Q}=\mathbbm{R}^n$, corresponds to the space $S(\mathcal{Q}, q^i)$ of Schwartz functions.
-$H$ therefore preserve smoothness and integrability.
+$H$ therefore preserve infinite smoothness and integrability.
 This, again, reminds us that the space of all self-adjoint operators is far too vast, and does not map well to physical meaningful quantities.
@@
 While this may seem like an incomplete description, it is as complete as the irreducibility assumption allows.
-The state is known and determines the final distribution.
+The state is known and determines the final statistical distribution.
 The unstated part cannot be known, and influences the non-deterministic evolution.
@@
 In hindsight, the idea that physics should be founded on physical principles is nothing new: it just hasn't been thoroughly applied to basic mechanics.
-As these mathematical frameworks have developed across two centuries, it maybe should not be so surprising that they are ripe for some reorganization.
+If this work would have been done incrementally from the start, these concepts would have been developed gradually and the results would feel more mundane.
 
@@
 Nobody is going to hand us down the laws of infinitesimal parts: we induce them by manipulating finite systems.
+Therefore what the whole allows to infer about its part is physically fundamental.
 
@@
 Then we assume an homogeneous irreducible material.
-Connecting the dots, the next assumption would be a inhomogeneous irreducible material: where the material is not in constant equilibrium and its particles are not composed of the same amount.
+Connecting the dots, the next assumption would be a inhomogeneous irreducible material: where the material is not in constant equilibrium and its particles do not all look the same.
 This different starting point may have potential connections to some of the problems we mentioned in the rationales for the assumptions and to other interesting physics questions in general.
 
-As different particles can transform into one another, they could be seen as different stable states of the same inhomogeneous material.
+As there would be different particles, they could be seen as different stable states of the same inhomogeneous material.
 In particular, since $\sigma_x \sigma_v \geq \frac{\hbar}{2m}$, particles in different generations (i.e.
@@
 As we are not always at equilibrium, the state of the system will be in general a statistical mixture of distributions of material.
-But the evolution may also take pure states (i.e.
+The evolution may also take pure states (i.e.
 a mixed state with a single state with $100\%$ probability) and evolve them into mixed states, thus giving an account of what happens during a measurement or a decay.
@@
 These ideas are appealing because they seem to fit well in the narrative, but at this point they are just speculation.
-The only thing we can say at this point is that the state space would be a vector space on a field that is neither $\mathbbm{R}$ or $\mathbbm{C}$.
-
-We limit the discussion to these few remark as the work is already long.
-We hope to extended it to field theories in the near future.
+The only thing we can say for now is that the state space under this new assumption would be a vector space on a field that is neither $\mathbbm{R}$ or $\mathbbm{C}$.
+But it is not unreasonable that these will inspire new work in the future.
+
+In the foreseeable future, however, we will work on the more mundane task of extending this work to field theories, for which a more concrete path is available.
 The state space for a quantized material, already introduced here, should be captured by the Fock space.
@@
 In doing so we obtained a more unified picture of the different branches of fundamental physics centered around the idea of deterministic and reversible evolution of distributions of homogeneous materials.
-The main physical concepts can be summarized by the following points:
+Though the work is extensive we can summarize a few important conclusions with the following points:
 
@@
 
-\item Relativistic motion arises by properly handling time dependent deterministic and reversible evolution for the material (i.e. distributions over phase space).
+\item Relativistic motion arises by properly handling time dependent deterministic and reversible evolution for the material (i.e. distributions over phase space) without further assumptions.
 
@@
 
-While the scope of this work is limited to re-founding and re-understanding physical theories already experimentally established, it may also has the potential to provide new insights and ideas on how to address some of the open problems in the foundations of physics.
+While the scope of this work is limited to re-founding and re-understanding physical theories already experimentally established, it would not be unexpected that a better grasp of old theories may lead to insights and ideas for new ones.
 
@@
 In this section we include mathematical demonstration in support of the physical arguments of the paper.
-
-\begin{thrm}\label{thrm:function_topology}
-	Let $X$ and $Y$ be Hausdorff and second countable topological spaces.
-Let $C(X, Y) \equiv \{f: X \rightarrow Y \; | \; f$ is continuous$\}$ be the space of continuous functions endowed with the Compact-Open Topology $\mathsf{T}$.
-Then $C(X, Y)$ is a Hausdorff and second countable topological space.
-\end{thrm}
-
-\begin{proof}
-	We claim $C(X, Y)$ is Haudorff. Let $f, g \in C(X, Y)$ such that $f \neq g$. $\exists x \in X$ such that $f(x)\neq g(x)$. Let $K = \{x\}$. $K$ is compact.
-\end{proof}
 

### 2017-01-03 2015CompleteDerivation/completeDerivation.tex
Reviewing with Christine

@@
 We need to be able to express distributions and expectations for different quantities: this will give us self-adjoint linear operators.
-We then need to identify the self-adjoint linear operators for the quantities $k_i$ and $p_i$ and to be able express the distribution of material in terms of those.
+We then need to identify the self-adjoint linear operators for the quantities $k_i$ and $p_i$ and to be able to express the distribution of material in terms of those.
 Let's start with the first.
 
-The idea is that, given two states $\psi_1$ and $\psi_2$, we should be able to tell how much of $\psi_1$ is prepared in the same way as of $\psi_2$.
+The idea is that, given two states $\psi_1$ and $\psi_2$, we should be able to tell how much of $\psi_1$ is prepared in the same way as $\psi_2$.
 That is $\psi_1 = \alpha \psi_2 + \psi_3$ is the combination of the component that is prepared like $\psi_2$ (e.g.
@@
 That is: $\mathsf{P}_{\psi_2} (\psi_1) = \psi_1 \Leftrightarrow \psi_1 = \alpha \psi_2$.
-Because of this, can define a non degenerate complex product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{\langle \psi_2|\psi_1 \rangle}{|\psi_2|^2} \psi_2$.
+Because of this, we can define a non degenerate complex product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{\langle \psi_2|\psi_1 \rangle}{|\psi_2|^2} \psi_2$.
 
@@
 As this is a part of $\psi_1$ prepared as $\psi_1$, its state is just $\psi_1$ rescaled, with no change in correlation.
-That is: $\mathsf{P}_{\psi_1} (\mathsf{P}_{\psi_2} (\psi_1)) = a \psi_1$ with $a \in \mathbbm{R}$.
+That is: $\mathsf{P}_{\psi_1} (\mathsf{P}_{\psi_2} (\psi_1)) = a \psi_1$ with $a \in \mathbbm{R}^+$.
 This tells us that $\langle \psi_2|\psi_1 \rangle = \langle \psi_1|\psi_2 \rangle^\dagger$.
@@
 \begin{prop}\label{prop:inner_product}
-	The state space $\Psi$ for an irreducible material is an inner product with $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$.
+	The state space $\Psi$ for an irreducible material is an inner product with $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$.
 \end{prop}
@@
 Let $| \cdot | ^2 : \Psi \rightarrow \mathbbm{R}$ be the map that returns the amount of material for a given state.
-Let $\psi \in \Psi$.
+Let $\psi \in \Psi$ be a state.
 $|\psi|^2 = \int \rho(q^i) d^nq = \int \psi^\dagger(q^i) \psi(q^i) d^nq \ge 0$.
@@
 
-	We claim $\forall \psi \in \Psi$ there exist a projection $\mathsf{P}_\psi : \Psi \rightarrow \Psi$ with $\psi \in \Psi$.
+	We claim $\forall \psi \in \Psi$ there exists a projection $\mathsf{P}_\psi : \Psi \rightarrow \Psi$ with $\psi \in \Psi$.
 Let $\psi \in \Psi$ be a state.
@@
 $\frac{|\mathsf{P}_{\psi} (\phi)|^2}{|\phi|^2}=\frac{|\mathsf{P}_{\phi} (\psi)|^2}{|\psi|^2}$.
-
-	We claim $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi)) = r \psi$ for some $r \in \mathbbm{R}$.
+Multiplying by $|\psi|^2|\phi|^2$ we have $|\psi|^2|\mathsf{P}_{\psi} (\phi)|^2=|\phi|^2|\mathsf{P}_{\phi} (\psi)|^2$.
+This holds for the empty state as well.
+Let $\phi$ be the empty state.
+$|\phi|^2=0$ because it has no material and $|\mathsf{P}_{\psi} (\phi)|^2$ as no component of the empty state is prepared like $\psi$.
+
+	We claim $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi)) = r \psi$ for some $r \in \mathbbm{R}^+$.
 Let $\psi, \phi \in \Psi$.
@@
 
-	We claim that the inner product can be expressed in terms of the wave functions as $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$.
-The amount of material associated with a state $\psi \in \Psi$ is $\langle \psi | \psi \rangle = | \psi |^2 = \int_\mathcal{Q} \psi^\dagger (q^i) \psi(q^i) dq^i$.
-The inner product associated with the $L^2$ norm is $\int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$ with $\psi_1, \psi_2 \in \Psi$.
-$\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) dq^i$.
+	We claim that the inner product can be expressed in terms of the wave functions as $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$.
+The amount of material associated with a state $\psi \in \Psi$ is $\langle \psi | \psi \rangle = | \psi |^2 = \int_\mathcal{Q} \psi^\dagger (q^i) \psi(q^i) d^nq$.
+The inner product associated with the $L^2$ norm is $\int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$ with $\psi_1, \psi_2 \in \Psi$.
+$\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$.
 
@@
 Note that the inner product only has an indirect physical definition, based on the norm and the projections.
-The former concepts are therefore more useful to gain physical intuition as the latter does not have a straightforward physical meaning.
+These two concepts are therefore more useful to gain physical intuition as the first does not have a straightforward physical meaning.
 
@@
 For mathematical convenience, one can take its completion with respect to the norm induced by the inner product and obtain a Hilbert space.
-Yet, such object does no longer represent a physically meaningful set of states.
+Yet, such object no longer represents a physically meaningful set of states.
 
@@
 In particular, we don't want to use the usual link $\psi(q) = \langle q | \psi \rangle$ between the wave function and the inner product as the material cannot be prepared at a specific value when $q$ is a continuous variable.
-Similarly, we cannot show that to each state variable is associated an hermitian operator simply by constructing it from its eigenstates as they may not represent physical states.
+Similarly, we cannot show that to each state variable is associated a self-adjoint operator $Q = \sum q |q\rangle \langle q |$ simply by expressing it in terms of eigenstates as they may not represent physical states.
 
@@
 Consider the product $f (q) \psi (q)$: it returns the wave function weighted by the value of $f$.
-We can think of $f$ as an operator that takes a complex function of $\mathcal{Q}$ and returns another complex function of $\mathcal{Q}$ that represents the distribution of the expectation over the fragments.
-The operator is linear since $f (q) (a\psi_1 + b\psi_2) = a f \psi_1 + b f  \psi_2$.
-The operator is hermitian since $f^\dagger = f$.
+We can also think of $f$ as an operator that takes a complex function of $\mathcal{Q}$ and returns another complex function of $\mathcal{Q}$ that represents the distribution of the expectation over the fragments.
+The operator is linear since $f (q) (a\psi_1(q) + b\psi_2(q)) = a f(q) \psi_1(q) + b f(q)  \psi_2(q)$.
+The operator is self-adjoint since $f^\dagger = f$.
 Consider $\int_\mathcal{Q} \psi^\dagger (q) f(q) \psi(q) dq = \int f(q) | \psi(q)|^2 dq$.
-It integrates the value of $f$ weighted by the amount of the material: it is the expectation of the state variable $f$ over the distribution.
+It integrates the value of $f$ weighted by the amount of the material: it is the expectation of the fragment variable $f$ over the distribution.
 
@@
 Suppose there is another maximal set of fragment variables.
-Then those variables will form a topological space $\hat{\mathcal{Q}}$ such that each $\psi$ can be expressed as a wave function $\psi(\hat{q}) :\hat{\mathcal{Q}} \rightarrow \mathbbm{C}$ over $\hat{\mathcal{Q}}$.
-
-As each state $\psi$ can be identified by a wave function on either space, there must be an operator $\Upsilon^q_{\hat{q}}(\psi(q)) = \psi(\hat{q})$ that converts the wave function on one space to the other.\footnote{Note that a neighbourhood of $q^i$ is not necessarily a neighbourood of $\hat{q}^i$.
-They are only if $q^i$ and $\hat{q}^i$ can be expressed as a bijection of each other $\hat{q}^i(q^i)$: they are different coordinates of the same manifold.} As to a linear combination of states corresponds a linear combination of wave functions, $\Upsilon^q_{\hat{q}}$ is a linear operator.
+Then those variables will form a topological space $\hat{\mathcal{Q}}$ such that each $\psi$ can be expressed as a wave function $\psi(\hat{q}) :\hat{\mathcal{Q}} \rightarrow \mathbbm{C}$.
+
+As each state $\psi$ can be identified by a wave function on either space, there must be an operator $\Upsilon^q_{\hat{q}}(\psi(q)) = \psi(\hat{q})$ that converts the wave function on one space to the other.\footnote{Note that a neighborhood of $q$ is not necessarily a neighborhood of $\hat{q}$.
+They are only if $q$ and $\hat{q}$ can be expressed as a bijection of each other $\hat{q}^i(q^i)$: they are different coordinates of the same manifold.} As a linear combination of states corresponds to a linear combination of wave functions, $\Upsilon^q_{\hat{q}}$ is a linear operator.
 As converting the space again must give us the original wave function, $\Upsilon^{\hat{q}}_{q} \circ \Upsilon^q_{\hat{q}}= I$ is unitary.
-We can now define a linear operator $F_{[\hat{q}^i]}$ corresponding to state variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
-Using $\Upsilon^{q}_{\hat{q}}$ we can express $F_{[q^i]} = \Upsilon^{\hat{q}}_{q} \circ f(\hat{q}^i) \Upsilon^{q}_{\hat{q}}$ as an operator over complex functions of $\mathcal{Q}$.
-It will be still be linear and hermitian.
-
-That is: for each state variable $f$ we have a corresponding linear hermitian operator $F_{[q^i]}$ that acts on wave functions over $q^i$, which can be expressed as function multiplication in a suitable choice of $\hat{\mathcal{Q}}$.
-The expectation multiplied by the amount of material is given by $\int \psi^\dagger(q^i) F_{[q^i]} (\psi(q^i)) d^nq$.
+We can now define a linear operator $F_{[\hat{q}]} \equiv f(\hat{q})$ corresponding to fragment variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
+Using $\Upsilon^{q}_{\hat{q}}$ we can express $F_{[q]} = \Upsilon^{\hat{q}}_{q} \circ F_{[\hat{q}]} \circ  \Upsilon^{q}_{\hat{q}}$ as an operator over complex functions of $\mathcal{Q}$.
+It will still be linear and self-adjoint.
+
+That is: for each fragment variable $f$ we have a corresponding linear self-adjoint operator $F_{[q^i]}$ that acts on wave functions over $q^i$, which can be expressed as function multiplication in a suitable choice of $\hat{\mathcal{Q}}$.
+The (normalized) expectation of $f$ is given by $\frac{1}{|\psi|^2}\int \psi^\dagger(q^i) F_{[q^i]} (\psi(q^i)) d^nq$.
 
 \begin{prop}\label{prop:self_adjoint_operators}
-	For each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ and a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ where $\Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ is the space of wave functions over $\hat{\mathcal{Q}}$.
+	For each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ and a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ that takes a state $\psi \in \Psi$ and returns a wave function $\psi(\hat{q}^i)$ from $\Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ the space of wave functions over $\hat{\mathcal{Q}}$.
 There also exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbbm{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i))d^nq $ is the expectation of $f$ multiplied by the amount of material.
@@
 \begin{justification}
-	We claim for each fragment variable $f$ there exist a topological space $\hat{\mathcal{Q}}$ such that  $q^i : \mathcal{Q} \rightarrow \mathbbm{R}^n$ provides a maximal description and $f : \mathcal{Q} \rightarrow \mathbbm{R}$.
+	We claim for each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that  $\hat{q}^i : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}^n$ provides a maximal description and $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
 Let $f$ be a fragment variable.
 Let $\hat{q}^i$ be a set of fragment variables such that $f=f(\hat{q}^i)$.
-If $\hat{q}^i$ does not provide a maximal set of fragment variables then there exists other fragment variables that refine the description.
+If $\hat{q}^i$ does not provide a maximal set of fragment variables then there exist other fragment variables that refine the description.
 Extend $\hat{q}^i$ with such variables.
@@
 There must exist a wave function $\psi(\hat{q}^i) : \mathbbm{R}^n \rightarrow \mathbbm{C}$ associated with $\psi$ such that it gives the maximal description for the state.
-Let $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \rightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ the map that associates a state with its wave function over $\hat{\mathcal{Q}}$.
+Let $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \rightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ be the map that associates a state with its wave function over $\hat{\mathcal{Q}}$.
 $\Upsilon_{\hat{\mathcal{q}}^i}$ is a vector space isomorphism for the same reason discussed in \ref{prop:wavefuntion}
@@
 	We claim that for each fragment variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ there exists a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ such that $\int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]}( \psi(\hat{q}^i)) d^n \hat{q}$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
-The expectation value of $f$ multiplied by the amount of material is given by $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q}$ where $\hat{q}^i$ provide a maximal set of fragment variables such that $f=f(\hat{q})$.
+The expectation value of $f$ multiplied by the amount of material is given by $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q}$ where $\hat{q}^i$ provide a maximal set of fragment variables such that $f=f(\hat{q}^i)$.
 We have $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q} = \int \psi^\dagger(\hat{q}^i) f(\hat{q}^i) \psi(\hat{q}^i) d^{n}\hat{q}$.
@@
 Let $f$ be a fragment variable.
-There exist a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ where $\hat{\mathcal{Q}}$ provides a maximal set of fragment variables.
+There exists a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ where $\hat{\mathcal{Q}}$ provides a maximal set of fragment variables.
 As both $\hat{\mathcal{Q}}$ and $\mathcal{Q}$ provide a maximal set, we can define $\Upsilon^{q^i}_{\hat{q}^i} \equiv \Upsilon_{\hat{q}^i} \circ \Upsilon_{q^i}^{-1}$ which transforms a wave function over $\mathcal{Q}$ into a wave function over $\hat{\mathcal{Q}}$.
@@
 To be precise, we want to find $K_{i [q^i]}$, the operators that act on the wave functions over $q^i$ corresponding to fragment variables $k_i$.
-We know that the operators have to be covariant: under a change of units $\hat{q}^j = \hat{q}^j(q^i)$ we must have $K_{\hat{q}^j[q^i]} = \partial_{\hat{q}^j} q^i K_{q^i [q^i]}$.
-We are looking for linear operators that obey the chain rule, which turns out to be the derivation by the corresponding $q^i$.
+We know that the operators have to be covariant: under a change of units $\hat{q}^j = \hat{q}^j(q^i)$ we must have $\hat{K}_{j[q^i]} = \partial_{\hat{q}^j} q^i K_{i [q^i]}$.
+In other words: we are looking for the set of linear operators that obey the chain rule.
+This turns out to be the space of derivations, therefore $K_{i [q^i]}$ will be the derivatives in the direction of the corresponding $q^i$.
 They have to be self-adjoint, so they will be of the form $K_{i [q^i]} = a \imath \partial_{i}$ where $a$ is a real number.
@@
 
-If we assume $\mathcal{Q}$ to be $\mathbbm{R}^n$, the functional $\Upsilon^{q^i}_{p_i} : \Psi(\mathcal{Q}, dq^i) \leftrightarrow \Psi(\mathcal{P}, dq^i)$ that converts a wave function in position to a wave function in momentum is the Fourier transform.
-As the fragment variables for any power of $q^i$ and $p_i$ must exist, the wave functions in both expressions are infinitely differentiable: the space of the wave functions $\Psi(\mathcal{Q}, dq^i) = S(\mathcal{Q}, q^i)$ is the Schwartz space.
+If we assume $\mathcal{Q}$ to be $\mathbbm{R}^n$, the functional $\Upsilon^{q^i}_{p_i} : \Psi(\mathcal{Q}, dq^i) \leftrightarrow \Psi(\mathcal{P}, dp_i)$ that converts a wave function in position to a wave function in conjugate momentum is the Fourier transform.
+As the fragment variables for any power of $q^i$ and $p_i$ must exist, the wave functions in both expressions are infinitely differentiable: the space of the wave functions $\Psi(\mathcal{Q}, dq^i) = S(\mathcal{Q}, q^i)$ is the Schwartz space of rapidly decreasing smooth functions.
 
@@
 The operator $P_{i [q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbbm{C})$ associated with the conjugate quantity of $q^i$ is $P_{i [q^i]} = - \imath \hbar \partial_{i}$.
-The isomorphism $\Upsilon^{q^i}_{p_i} : \Psi(\mathcal{Q}, dq^i) \leftrightarrow \Psi(\mathcal{P}, dq^i)$ that changes the variable of the wave function from $q^i$ to $p_i$ is $\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^3} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
+The isomorphism $\Upsilon^{q^i}_{p_i} : \Psi(\mathcal{Q}, dq^i) \leftrightarrow \Psi(\mathcal{P}, dp_i)$ that changes the variable of the wave function from $q^i$ to $p_i$ is $\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^n} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
 The space of wave functions $\Psi(\mathcal{Q}, dq^i) = S(\mathcal{Q}, q^i)$ where $S(\mathcal{Q}, q^i)$ is the Schwartz space of rapidly decreasing smooth functions.
@@
 	We claim the operator associated with $k_i$ is $K_{i [q^i]} = - \imath \partial_{i}$.
-Let $K_{i [q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, dq^i)$ be the linear operator associated with $k_i$ as express on the space of wave functions over $q^i$.
-Such operator must be covariant under the transformation $\hat{q}^j = \hat{q}^j(q^i)$.
-That is: $K_{\hat{q}^j [q^i]} = \partial_{\hat{q}^j}q^iK_{q^i [q^i]}$.
+Let $K_{i [q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, dq^i)$ be the linear operator associated with $k_i$ as expressed on the space of wave functions over $q^i$.
+Such an operator must be covariant under the transformation $\hat{q}^j = \hat{q}^j(q^i)$.
+That is: $\hat{K}_{j [q^i]} = \partial_{\hat{q}^j}q^iK_{i [q^i]}$.
 In particular, it needs to be covariant when changing only one variable: $\hat{q}^1 = \hat{q}^1(q^1)$, $\hat{q}^j = q^j$ where $j=2...n$.
 By \ref{thrm:covariant_operator} $K_{i [q^i]}=c\partial_i$ for some $c \in \mathbbm{C}$.
-$K_{i [q^i]}$ is hermitian: $K_{i  [q^i]}=\imath r\partial_i$ for some $r \in \mathbbm{R}$ by \ref{thrm:antihermitian_derivative}.
+$K_{i [q^i]}$ is self-adjoint: $K_{i  [q^i]}=\imath r\partial_i$ for some $r \in \mathbbm{R}$ by \ref{thrm:antihermitian_derivative}.
 $k_i$ is expressed in units of inverse $q^i$ therefore $|r|=1$.
@@
 	We claim the operator associated with $p_i$ is $P_{i [q^i]} = - \imath \hbar \partial_{i}$.
-As $p_i=\hbar k_i$, their expectation is proportional by $\hbar$.
+As $p_i=\hbar k_i$, their expectation is proportional by a factor of $\hbar$.
 $P_{i [q^i]}=\hbar K_{i [q^i]}= -\imath \hbar \partial_i$.
 
-	We claim the operator that changes the variable of the wave function from $q^i$ to $p_i$ is $\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^3} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
-Let $\Upsilon^{q_i}_{p^i} : \Psi(\mathcal{Q}, dq^i) \rightarrow \Psi(\mathcal{P}, dp^i)$ be the operator that maps a wave function over $\mathcal{Q}$ into a wave function over $\mathcal{P}$, the space charted by $p_i$.
-$\Upsilon^{q_i}_{p^i}$ must exist and be unique (up to a total phase) as $\Upsilon^{q_i}_{p^i} = \Upsilon_{p^i}\Upsilon^{q_i} = \Upsilon_{p^i}(\Upsilon_{q_i})^{-1}$ is an isomorphism.
-$\Upsilon^{q_i}_{p^i}$ must be a linear transform that takes the operator $P_{i [q^i]}= - \imath \hbar \partial_i$ as expressed over $q^i$ to $P_{i [p_i]} = p_i$ as expressed over $p_i$.
+	We claim the operator that changes the variable of the wave function from $q^i$ to $p_i$ is $\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^n} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
+Let $\Upsilon^{q^i}_{p_i} : \Psi(\mathcal{Q}, dq^i) \rightarrow \Psi(\mathcal{P}, dp_i)$ be the operator that maps a wave function over $\mathcal{Q}$ into a wave function over $\mathcal{P}$, the space charted by $p_i$.
+$\Upsilon^{q^i}_{p_i}$ must exist and be unique (up to a total phase) as $\Upsilon^{q^i}_{p_i} = \Upsilon_{p_i}\circ\Upsilon^{q^i} = \Upsilon_{p_i}(\Upsilon_{q^i})^{-1}$ is an isomorphism.
+$\Upsilon^{q^i}_{p_i}$ must be a linear transform that takes the operator $P_{i [q^i]}= - \imath \hbar \partial_i$ as expressed over $q^i$ to $P_{i [p_i]} = p_i$ as expressed over $p_i$.
 Such transformation is the Fourier transform.
-$\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^3} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
+$\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^n} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
 
@@
 Let $\psi(q^i) \in \Psi(\mathcal{Q}, dq^i)$.
-$P_{i [q^i]}^n \psi(q^i) = (-\imath \hbar \partial_i)^n \psi(q^i)$ represents the n-th power of the conjugate momentum operator applied to the given state.
-As the quantity is physically meaningful, such operation must be defined for any $n$ therefore $\psi(q^i) \in C^\infty(\mathcal{Q})$.
+$(P_{i [q^i]})^m \psi(q^i) = (-\imath \hbar \partial_i)^m \psi(q^i)$ represents the $m^{th}$ power of the conjugate momentum operator applied to the given state.
+As the quantity is physically meaningful, such an operation must be defined for any $m$ therefore $\psi(q^i) \in C^\infty(\mathcal{Q})$.
 Let $\psi(p_i) = \Upsilon^{q^i}_{p_i} (\psi(q^i))$ be the wave function over momentum space.
-$(Q^i_{[p_i]})^n \psi(p_i) = (+\imath \hbar \partial_{p_i})^n \psi(p_i)$ represents the n-th power of the $q^i$ operator applied to the given state.
-As the quantity is physically meaningful, such operation must be defined for any $n$ therefore $\psi(p_i) \in C^\infty(\mathcal{P})$.
-As $\psi(q^i)$ is the Fourier transform of an infinitely smooth function, it decreases faster then any inverse power of $q^i$.
-That is: $\forall m \in \mathbbm{Z}^+ \exists c \in \mathbbm{C}$ such that  $|\psi(q^i)|<\frac{c}{|q^i|^m} \; \forall |q^i|> 1$.
+$(Q^i_{[p_i]})^m \psi(p_i) = (+\imath \hbar \partial_{p_i})^m \psi(p_i)$ represents the $m^{th}$ power of the $q^i$ operator applied to the given state.
+As the quantity is physically meaningful, such an operation must be defined for any $m$ therefore $\psi(p_i) \in C^\infty(\mathcal{P})$.
+As $\psi(q^i)$ is the Fourier transform of an infinitely smooth function, it decreases faster than any inverse power of $q^i$.
+That is: $\forall m \in \mathbbm{Z}^+ \exists c \in \mathbbm{C}$ such that  $|\psi(q^i)|<\frac{|c|}{|q^i|^m} \; \forall |q^i|> 1$.
 $\Psi(\mathcal{Q}, dq^i) \subseteq S(\mathcal{Q}, q^i)$.
@@
 Let $\psi(q^i) \in \Psi(\mathcal{Q}, dq^i)$ and $\phi(q^i) \in S(\mathcal{Q}, q^i)$ such that $\psi(q^i) \neq \phi(q^i)$.
-There exists $a, b, j, k \in \mathbbm{Z}^+$, $j,k < n$ such that $\lVert \psi(q^i) - \phi(q^i) \rVert_{j,a,k,b} \neq 0$  where $\lVert \psi(q^i) \rVert_{j,a,k,b} = sup|(q^j)^a (\partial_k) ^b \psi(q^i)|$ as $S(\mathcal{Q}, q^i)$ is a Frechet space induced by that family of seminorms.
-The operator $(Q^j_{[q^i]})^a(P_{k [q^i]})^b$ applied to the two functions will provide different distributions for the associated fragment variables.
+There exists a choice of $a, b, j, k \in \mathbbm{Z}^+$, $j,k \leq n$ such that $\lVert \psi(q^i) - \phi(q^i) \rVert_{j,a,k,b} \neq 0$  where $\lVert \psi(q^i) \rVert_{j,a,k,b} = sup|(q^j)^a (\partial_k) ^b \psi(q^i)|$ as $S(\mathcal{Q}, q^i)$ is a Frechet space induced by that family of seminorms.
+The operator $(Q^j_{[q^i]})^a(P_{k [q^i]})^b$ applied to $\psi(q^i)$ and $\phi(q^i)$ will provide different distributions for the associated fragment variables.
 $\phi(q^i)$ is physically distinguishable and is associated with a physical state.
-$\phi(q^i) \in \Psi(\mathcal{Q}, dq^i)$
+$\phi(q^i) \in \Psi(\mathcal{Q}, dq^i)$.
 \end{justification}
@@
 And we obtained them by re-deriving them from first principles instead of using some formal analogy.
-The similarity with the classical framework are simply due to the similarity of the starting points.
+The similarities with the classical framework are simply due to the similarities of the starting points.
 
@@
 Yet, we can still create a self-adjoint linear operator, with the same spectra as the one associated with $q$, just with the eigenvalues switched.
-The space of physically meaningful operator, then, should be further constrained.
+The space of physically meaningful operators, then, should be further constrained.
 A reasonable requirement would be that they preserve the differentiable/integrable structure of the Schwartz space $S(\mathcal{Q}, q^i)$ as classical state variables preserved the differentiable structure of $T^*\mathcal{Q}$.
@@
 Schroedinger's equation), and a non deterministic one useful to describe measurement interactions, corresponding to the projection postulate.
-Therefore, as the infinitesimal reducibility assumptions leads to classical mechanics, the irreducibility assumption leads to quantum mechanics.
+Therefore, as the infinitesimal reducibility assumption leads to classical mechanics, the irreducibility assumption leads to quantum mechanics.
 
@@
 	The methodological problem assumes a different connotation.
-In order for the assumption to hold, we need to show that there are no processes at our disposal that can probe the constituents of a quantum particle.
+In order for the assumption to hold, we need to show that there are no processes at our disposal that can probe the fragments of a quantum particle.
 For example: an electron as a whole interacts with a photon as a whole.
 Note that this does not require showing that no such process exists: conceptually specific parts of the photon may be interacting with specific parts of the electron.
-But if those processes are not available to us, say because we can only manipulate whole photons and they always interacts with a whole electron, they cannot be used to create a more detailed picture.
+But if those processes are not available to us, say because we can only manipulate whole photons and they always interact with a whole electron, they cannot be used to create a more detailed picture.
 So we simply have to enumerate the processes available to us, which is, at least in principle, feasible.
@@
 
-	The issues rising from a lack of perfect isolation are addressed.
+	The issues arising from a lack of perfect isolation are addressed.
 Each degree of freedom comes with its own unstated part (i.e.
-the motion of the fragments within that d.o.f.) therefore these is no pretense of a perfect description under perfect isolation.
-
-	The problem of physically meaningless interaction is also considerably lessened.
+the motion of the fragments within that d.o.f.), therefore there is no pretense of a perfect description under perfect isolation.
+Since we cannot describe the nature of the chaotic motion, we cannot say whether it is internally or externally driven.
+
+	The problem of physically meaningless interactions is also considerably lessened.
 Even if we have multiple particles each with their state evolving deterministically, we can still imagine their unstated parts as interacting with each other.
@@
 
-	The problem of incompatibility of measurements and deterministic evolution at all levels still remains.
+	The problem of incompatibility of measurements with deterministic evolution at all levels still remains.
 The unstated part of multiple irreducible particles may be interacting, but it cannot be used to pass information from one subsystem to another without making itself distinguishable.
@@
 	There is one additional problem, though, not present in the classical case.
-Quantum particle states, as we have defined them, are states at equilibrium: they all provide the same level of description for the fragments, they carry the same information entropy.
+Under the irreducibility assumption states are necessarily at equilibrium: they all provide the same level of description for the fragments, they carry the same information entropy.
 In fact if the internal dynamics were out of equilibrium, the material would not be homogeneous which would contradict the assumption.
 Deterministic and reversible evolution (which we'll see corresponds to unitary evolution) is necessarily quasi static: at each moment a state must be well defined, at each moment the system is in equilibrium, the system changes slowly.
-Therefore quantum states cannot and should not be expected to provide a description under non deterministic evolution.
-
-	In light of this, consider a muon decay: the initial state of the muon and the final states of the electron and neutrinos can certainly be well described by quantum particle states.\footnote{In this context, a state is in equilibrium regardless of it being in a local minimum, easily overcome by fluctuations, or a global minimum, that persists indefinitely.} We have no reason to expect, though, that the while muon is decaying into the resulting particles the state is always in equilibrium and therefore well described mathematically by a vector in a Hilbert space.
-After all, the state of a single spin one half particle cannot determine the future state of three spin one half particles.
-
-	As we'll discuss in a later subsection, using unitary evolution in those cases to make predictions, though, may still make physical sense.
-Quantities that are not affected by the particular evolution of the unstated part will have, by their very nature, the same value under non-deterministic evolution and an equivalent quasi-static evolution.
-For example, the overall probability distribution resulting from the decay of a muon is something that cannot depend on the detail motion of the fragment (otherwise we would be able to characterize it).
+Therefore quantum states, as formulated here, cannot and should not be expected to provide a valid description during non-deterministic evolution.
+
+	In light of this, consider a muon decay: the initial state of the muon and the final states of the electron and neutrinos can certainly be well described by quantum particle states.\footnote{In this context, a state is in equilibrium regardless of its being in a local minimum, easily overcome by fluctuations, or a global minimum, that persists indefinitely.} We have no reason to expect, though, that while the muon is decaying into the resulting particles the state is always in equilibrium and therefore well described mathematically by a vector in a complex inner product space.
+After all, the state of a single spin-one-half particle cannot determine the future state of three spin-one-half particles.
+
+	This notion may seem in contradiction to the current practice.
+For example, the S-matrix, one of the main devices for calculating scattering amplitudes and cross sections, is derived by assuming unitary evolution.
+Doesn't it imply we are assuming that deterministic and reversible evolution is happening at every instant?
+
+	If we look more closely, we realize that the process that the S-matrix captures is a limit where initial states are at $t \rightarrow - \infty$ and final states are at $t \rightarrow + \infty$, one that takes an infinite amount of time.
+That is, to calculate predictions for a fast out of equilibrium process we are approximating it using slow quasi-static evolution.
+This is actually not so uncommon: it is a standard first approximation in thermodynamics.
+The question is why would it work here?
+As we'll see in more detail when talking about the projection postulate, it's because of the irreducibility assumption itself.
+
+	Suppose we are trying to calculate a quantity that does not depend on the particular motion of the unstated part.
+Then, by its very nature, its value will be the same under an equivalent quasi-static evolution, because this too does not depend on the particular motion of the fragments.
+The probability distribution of the out-states during scattering is one such quantity.
+As the final statistical distribution is determined only by the initial state, it is not dependent on the particulars of the unstated part.
 In these cases, the use of deterministic and reversible evolution as a ``stand-in" to calculate final statistical distributions is justified.
-But it should not be taken literally as a physical model of what actually happens.
+However, it should not be taken literally as a physical model of what actually happens.
 There is no justification for assuming this process to be quasi-static deterministic and reversible evolution of quantum particle states.
@@
 	In other words, the assumption of irreducibility already contains the seeds of its own demise.
-It incorporates the unstated part, yet it does not allow to characterize its full, out of equilibrium, dynamics.
-
-	In light of what we discussed, the irreducibility assumption is less flawed but it still cannot be taken as fundamental, in the sense that we cannot take it to strictly apply to all systems.
-We should not expect to solve open problems, such as the arrow of time and the measurement problem, within the assumption.
+On one hand, it tells us that there is a component of chaotic motion within the system.
+On the other hand, it can only characterize the equilibrium dynamics, which clearly can't be the full range.
+Yet it also offers a way out: since the motion of the fragments is not accessible, we can predict statistical distributions that are independent of it.
+
+	In light of what we discussed, the irreducibility assumption is less flawed than infinitesimal reducibility but it still cannot be taken as fundamental, in the sense that we cannot take it to strictly apply to all systems.
+We should not expect to solve open problems, such as the arrow of time and the measurement problem, under the assumption as formulated in this work.
 \end{rationale}
@@
 We are now ready to write the equations of motion for an irreducible material.
-Deterministic and reversible evolution for the whole distribution necessarily means quasi static evolution: there is a well defined state at each moment in time.
+Deterministic and reversible evolution for the whole distribution necessarily means quasi-static evolution: there is a well defined state at each moment in time.
 Moreover the components must evolve independently as we cannot learn more about one component by observing its evolution in different combinations: the evolution of the composition is the composition of the evolution.
@@
 Let $\psi_1, \psi_2 \in \Psi$ be two states.
-$\mathcal{T}_{\Delta t}(a\psi_1+b\psi_2) = a \mathcal{T}_{\Delta t} \psi_1 + b \mathcal{T}_{\Delta t} \psi_2$ with $a,b \in \mathcal{C}$ as the composition of the evolution is the evolution of the composition: $\mathcal{T}_{\Delta t}$ is a linear operator.
+$\mathcal{T}_{\Delta t}(a\psi_1+b\psi_2) = a \mathcal{T}_{\Delta t} \psi_1 + b \mathcal{T}_{\Delta t} \psi_2$ with $a,b \in \mathbbm{C}$ as the composition of the evolution is the evolution of the composition: $\mathcal{T}_{\Delta t}$ is a linear operator.
 $\mathcal{T}_{\Delta t} \mathsf{P}_{\psi_1} (\psi_2) = \mathsf{P}_{\mathcal{T}_{\Delta t}\psi_1} (\mathcal{T}_{\Delta t}\psi_2)$ as the evolved part of $\psi_2$ that was prepared like $\psi_1$ is the part of the evolved $\psi_2$ that ends prepared like the evolved $\psi_1$.
@@
 For example, for a system to reach its ground state it has to be able to radiate energy: it cannot be isolated.
-Therefore this dynamic cannot be described under the deterministic and reversible assumption.
-This already hints that deterministic and reversible evolution is not the only dynamic possible.
+Therefore this dynamics cannot be described under the deterministic and reversible assumption.
+This already hints that deterministic and reversible evolution is not the only dynamics possible.
 Other considerations, outside of the scope of this work, can be used to recover this behavior (e.g.
@@
 Also note that $H$ operates on $\Psi$ which, in the case of $\mathcal{Q}=\mathbbm{R}^n$, corresponds to the space $S(\mathcal{Q}, q^i)$ of Schwartz functions.
-$H$ therefore preserve infinite smoothness and integrability.
-This, again, reminds us that the space of all self-adjoint operators is far too vast, and does not map well to physical meaningful quantities.
+$H$ therefore preserves infinite smoothness and integrability.
+This, again, reminds us that the space of all self-adjoint operators is far too vast, and does not map well to physically meaningful quantities.
 
@@
 The final statistical distribution of the detector must match the incoming material distribution of the particle.
-That is if $N$ is the total number of takes and $N_A$ is the number of times the detector with aperture in the $A$ region captured the particle, the ratio $\frac{N_A}{N}$ will be approximately the amount of material within the aperture $\int_A \psi^\dagger(q^i) \psi(q^i) dq^i$.
-If we assume an infinite amount of takes, we expect them to coincide: $\int_A \psi^\dagger(q^i) \psi(q^i) dq^i = \lim\limits_{N\rightarrow\infty} \frac{N_A}{N}$.
+That is if $N$ is the total number of takes and $N_A$ is the number of times the detector with aperture in the $A$ region captured the particle, the ratio $\frac{N_A}{N}$ will be approximately the amount of material within the aperture $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq$.
+If we assume an infinite amount of takes, we expect them to coincide: $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq = \lim\limits_{N\rightarrow\infty} \frac{N_A}{N}$.
 But this is also the frequentist definition of probability.
-Therefore we have $\int_A \psi^\dagger(q^i) \psi(q^i) dq^i = P(\textrm{measuring particle in } A)$ thus recovering the probabilistic interpretation of the wave function.
+Therefore we have $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq = P(\textrm{measuring particle in } A)$ thus recovering the probabilistic interpretation of the wave function.
 

### 2017-01-03 2015CompleteDerivation/completeDerivation.tex
Reviewed topology justifications

@@
 \renewenvironment{proof}{\emph{Proof}.}{\qed}
+
+%TODO: decide on identity function typesetting
+\newcommand{\id}{\textrm{id}}
 
@@
 
-\begin{prop}\label{statedef}
+\begin{prop}\label{prop:state_topology}
 The state space $\mathcal{S}$ of a physical system is a Hausdorff and second countable topological space.
@@
 Let $\mathsf{T}$ be the set of all sets associated with all physical outcomes.
-$\varnothing \in \mathsf{T}$ and is associated with impossible outcomes.
-$\mathcal{S} \in \mathsf{T}$ and is associated with unavoidable outcomes.
+$\mathcal{S} \in \mathsf{T}$ and is associated with the outcome ``the system exists".
+$\varnothing \in \mathsf{T}$ and is associated with the outcome ``the system doesn't exists".
 Let $V_1, V_2 \in \mathsf{T}$.
 Then, by definition, there exists a process $P_1$ that admits $V_1$ as an outcome and a process $P_2$ that admits $V_2$ as an outcome.
-Consider the process $P$ that combines the outcomes of $P_1$ and $P_2$.
+Consider the process $P$ that combines the outcomes of $P_1$ and $P_2$ with a logical and.
 This always exists physically as we can prepare the same state multiple times and let it interact with each process separately.
@@
 If they are physically distinguishable, then there exists a physical process $P$ that admits both as outcomes.
-Given $P$, we can always construct the process $P_0$ that combines $V_1, V_2$ into a single outcome $V$ by "forgetting" which of the two was given.
+Given $P$, we can always construct the process $P_0$ that combines $V_1, V_2$ with a logical or into a single outcome $V$, by ``forgetting" which of the two was given.
 The states compatible with $V$ must be in either $V_1$ or $V_2$, that is $V = V_1 \cup V_2$.
@@
 Consider the set of outcomes of a process.
-As each is identified by a finite number of digits, the set of outcomes is countable.
+As each can provide only a finite amount of information (e.g.
+is identified by a finite number of digits), the set of outcomes is countable.
 This remains true when combining processes: the number of processes we can combine is finite as is the number of times we can prepare the same state and the set of intersections obtained from the countable outcomes of a finite number of processes is still countable.
@@
 
-\begin{prop}\label{topologically_distinguishable}
+\begin{prop}\label{prop:topology}
 	Any set $\mathcal{S}$ of physically distinguishable elements is a Hausdorff and second countable topological space.
@@
 \begin{justification}
-	Same justification as in \ref{statedef} with "state" replaced by "element" of the set $\mathcal{S}$.
+	Same justification as in \ref{prop:state_topology} with ``state" replaced by ``element" of the set $\mathcal{S}$.
 \end{justification}
@@
 
-\begin{prop}\label{detrevmap}
+\begin{prop}\label{prop:homeomorphism}
 A deterministic and reversible evolution map is a self-homeomorphism on the state space $\mathcal{T}_{\Delta t}:\mathcal{S} \rightarrow \mathcal{S}$.
@@
 The system is reversible: there exists a map $\mathcal{T}_{-\Delta t}:\mathcal{S} \rightarrow \mathcal{S}$ that returns the initial state given the final state.
-$\mathcal{T}_{-\Delta t} \circ \mathcal{T}_{\Delta t} = \mathcal{T}_{\Delta t} \circ \mathcal{T}_{-\Delta t} = id$ as mapping forward and then backward or backwards and then forward must return the original element.
+$\mathcal{T}_{-\Delta t} \circ \mathcal{T}_{\Delta t} = \mathcal{T}_{\Delta t} \circ \mathcal{T}_{-\Delta t} = \id_{\mathcal{S}}$ as mapping forward and then backward or backwards and then forward must return the original element.
 $\mathcal{T}_{\Delta t}$ admits $\mathcal{T}_{-\Delta t}$ as an inverse.
@@
 
-\begin{prop}\label{continuous_map}
+\begin{prop}\label{prop:continuity}
 	A map $f:\mathcal{S_1} \rightarrow \mathcal{S_2}$ between two sets of physically distinguishable elements $\mathcal{S_1}$ and $\mathcal{S_2}$ is a continuous map.
@@
 \begin{justification}
-	Same justification for continuity as in \ref{detrevmap} with ``initial states" and ``final states" replaced by "elements" of $\mathcal{S_1}$ and $\mathcal{S_2}$ respectively.
+	Same justification for continuity as in \ref{prop:homeomorphism} with ``initial states" and ``final states" replaced by ``elements" of $\mathcal{S_1}$ and $\mathcal{S_2}$ respectively.
 \end{justification}
@@
 use of standard topology) then discontinuous functions must also be ruled out.}
-	\label{fig:continuous_map}
+	\label{fig:continuity}
 \end{figure}
@@
 \begin{justification}
-	We claim $\mathbbm{L}$ is a topological space. $\mathbbm{L}$ is a set of physically distinguishable possibilities. $\mathbbm{L}$ is a topological space because of \ref{topologically_distinguishable}.
-
-	We claim $q$ is a continuous map. $q$ is a map between two sets of physically distinguishable elements. $q$ is continuous because of \ref{continuous_map}.
+	We claim $\mathbbm{L}$ is a topological space. $\mathbbm{L}$ is a set of physically distinguishable possibilities. $\mathbbm{L}$ is a topological space because of \ref{prop:topology}.
+
+	We claim $q$ is a continuous map. $q$ is a map between two sets of physically distinguishable elements. $q$ is continuous because of \ref{prop:continuity}.
 \end{justification}
@@
 $\lambda$ is a map between two physically distinguishable quantities, time and states.
-$\lambda$ is continuous by \ref{continuous_map}.
+$\lambda$ is continuous by \ref{prop:continuity}.
 All $\hat{\mathcal{s}} \in \mathcal{S}'$ are path connected to $\mathcal{s}$.
@@
 We require $\mathcal{C}$ to include state changes.
-A change of a physically distinguishable object is itself physically distinguishable therefore $\mathcal{C}$ is still a Hausdorff and second countable topological manifold as per \ref{topologically_distinguishable}.
+A change of a physically distinguishable object is itself physically distinguishable therefore $\mathcal{C}$ is still a Hausdorff and second countable topological manifold as per \ref{prop:topology}.
 There exists an inverse $- : \mathcal{C} \rightarrow \mathcal{C}$ such that $\mathcal{c} + ( - \mathcal{c}) = 0 \forall \mathcal{c} \in \mathcal{C}$.
@@
 The amount in each tank is continuous (as we assume the water to be infinitesimally divisible) yet there are only a finite number of tanks the water can be placed into.
-As it links two physically distinguishable objects, $\rho_\mathcal{c}$ is a continuous function as discussed in \ref{continuous_map}.
+As it links two physically distinguishable objects, $\rho_\mathcal{c}$ is a continuous function as discussed in \ref{prop:continuity}.
 
@@
 \rho_{\mathcal{c}}:\mathcal{S} \rightarrow \mathbbm{R}$ returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$.
-$\rho_{\mathcal{c}} \in C(\mathcal{S})$ by \ref{continuous_map}.
+$\rho_{\mathcal{c}} \in C(\mathcal{S})$ by \ref{prop:continuity}.
 Let $\varrho : \mathcal{C} \rightarrow C(\mathcal{S})$ such that $\varrho(\mathcal{c}) \mapsto \rho_\mathcal{c}$.
@@
 Some are not restrictive enough.
-The set of Lebesgue integrable functions $L^1(\mathcal{S})$ includes discontinuous functions that for \ref{continuous_map} are unphysical.
+The set of Lebesgue integrable functions $L^1(\mathcal{S})$ includes discontinuous functions that for \ref{prop:continuity} are unphysical.
 The space of continuous functions $C(\mathcal{S})$, the space of continuous functions that vanish at infinity $C_0(\mathcal{S})$ and the space of differentiable functions $C^1(\mathcal{S})$ include functions whose integral is infinite, which would represent an infinite amount of material.
@@
 
-We are finally ready to write the equations of motion. As per \ref{detrevmap} our evolution is at least a self-homeomorphism $f:\mathcal{S} \leftrightarrow \mathcal{S}$. But this is not enough.
+We are finally ready to write the equations of motion.
+As per \ref{prop:homeomorphism} our evolution is at least a self-homeomorphism $f:\mathcal{S} \leftrightarrow \mathcal{S}$.
+But this is not enough.
 
@@
 The state space for the particles of a classical material is $T^*\mathcal{Q}$ by \ref{symplectic_manifold}.
-$\mathcal{T}_{\Delta t}$ is a deterministic and reversible evolution map and by \ref{detrevmap} is a self-homeomorphism.
+$\mathcal{T}_{\Delta t}$ is a deterministic and reversible evolution map and by \ref{prop:homeomorphism} is a self-homeomorphism.
 
@@
 The complete state space for particles of a classical material is $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ by \ref{relativistic_symplectic_manifold}.
-$\mathcal{T}_{\Delta s}$ is a deterministic and reversible evolution map and by \ref{detrevmap} is a self-homeomorphism.
+$\mathcal{T}_{\Delta s}$ is a deterministic and reversible evolution map and by \ref{prop:homeomorphism} is a self-homeomorphism.
 We can extend the map over $T^*\mathcal{M}$.
@@
 $p_\alpha = p_\alpha(x^\alpha, u^\alpha)$.
-The map $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is continuous by \ref{continuous_map} and is bijective as the material is minimal.
+The map $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is continuous by \ref{prop:continuity} and is bijective as the material is minimal.
 It is a homemorphism between $T^*\mathcal{M}$ and $T\mathcal{M}$.
@@
 The earth is, so to speak, in both North America and Europe at the same time.
-In fact, based on what we saw before, any object of finite mass is a distribution over a region of phase space, and discontinuous distributions are ruled out by \ref{continuous_map}.
+In fact, based on what we saw before, any object of finite mass is a distribution over a region of phase space, and discontinuous distributions are ruled out by \ref{prop:continuity}.
 An electron is an object of finite mass: why should it be any different?
@@
 $\rho_\psi(q^i) = |\psi(q^i)|^2$ as the square modulus of $\psi(q^i)$ is the increase/decrease from the unit amount of material.
-$\rho_\psi(q^i) \in C(\mathcal{Q}, \mathbbm{R})$ by \ref{continuous_map}.
+$\rho_\psi(q^i) \in C(\mathcal{Q}, \mathbbm{R})$ by \ref{prop:continuity}.
 $\rho_\psi(q^i) \in C^1(\mathcal{Q}, \mathbbm{R})$ by the same arguments in the justification for \ref{prop:differentiable_manifold}.

### 2017-01-03 2015CompleteDerivation/completeDerivation.tex
Reviewed propositions for all classical

@@
 
-\begin{prop}\label{state_variable}
+\begin{prop}\label{prop:state_variable}
 	A \emph{state variable} is a continuous map $q : U \rightarrow \mathbbm{L}$ where $U \subseteq \mathcal{S}$ and $\mathbbm{L}$ is the space for the possible values.
@@
 
-\begin{prop}\label{independent_state_variables}
+\begin{prop}\label{prop:independent_state_variables}
 	Two state variables $q_1 : U \rightarrow \mathbbm{L}_1$ and $q_2 : U \rightarrow \mathbbm{L}_2$ are said \emph{independent} if $\exists \mathcal{s}$ such that $q_1(\mathcal{s})=l_1, q_2(\mathcal{s})=l_2 \forall l_1 \in q_1(U), l_2 \in q_2(U)$.
@@
 
-\begin{prop}\label{continuous_state_space}
+\begin{prop}\label{prop:manifold}
 	Let $\mathcal{s} \in \mathcal{S}$ a state within a state space.
@@
 
-\begin{prop}\label{evolved_transported_variable}
+\begin{prop}\label{prop:evolved_transported_variable}
 Let $q^i$ be a set of state variables and $\mathcal{T}_{\Delta t}$ a deterministic and reversible evolution map.
@@
 
-\begin{prop}\label{discrete_state_metric}
+\begin{prop}\label{prop:discrete_measure}
 	Let $U \subseteq \mathcal{S}$ a set of states fully identified by a set of $n$ independent discrete state variables $q^i$.
 Let $\Delta q^i \equiv q^i(U)$ the range of possibilities of each variable.
-Then $\#(\Delta q^i)=\#(\Delta \hat{q}^i) \; \forall i$ and $\#(U)=\prod\limits_{i=1}^{n}\#(\Delta q^i)=\prod\limits_{i=1}^{n}\#(\Delta \hat{q}^i)=\#(\hat{U})$ where $\hat{U}=\mathcal{T}_{\Delta t}(U)$ and $\#$ denotes the cardinality of the given set.
+Then $\#(\Delta q^i)=\#(\Delta \hat{q}^i) \; \forall i$ and $\#(U)=\prod\limits_{i=1}^{n}\#(\Delta q^i)=\prod\limits_{i=1}^{n}\#(\Delta \hat{q}^i)=\#(\hat{U})$ where $\hat{U}=\mathcal{T}_{\Delta t}(U)$ and $\#$ denotes the number of elements in the given set.
 \end{prop}
@@
 Giving the state of the whole system is equivalent to giving the states of all particles.
-Given the state space $\mathcal{S}$ for the infinitesimal parts, which we assume consistently with \ref{continuous_state_space} to be a manifold, a composite state $\mathcal{c} \in \mathcal{C}$ will tell us the amount of material for each possible particle state.
+Given the state space $\mathcal{S}$ for the infinitesimal parts, which we assume consistently with \ref{prop:manifold} to be a manifold, a composite state $\mathcal{c} \in \mathcal{C}$ will tell us the amount of material for each possible particle state.
 That is, each state is fully identified by a function $\rho_\mathcal{c}: \mathcal{S} \to \mathbbm{R}$ that returns the amount of material in the composite state $\mathcal{c}$ that is prepared in the particle state $\mathcal{s}$.
@@
 
-\begin{prop}\label{classical_vector space}
+\begin{prop}\label{prop:real_vector_space}
 The state space $\mathcal{C}$ for a classical (i.e.
@@
 
-\begin{prop}\label{integration}
+\begin{prop}\label{prop:integration}
 	The state space $\mathcal{S}$ for the particles of a classical material is endowed with a Borel measure $\mu$.
@@
 Let $\varrho : \mathcal{C} \rightarrow L^1(\mathcal{S}, \mu)$ such that $\varrho(\mathcal{c}) \mapsto \rho_\mathcal{c}$.
-$\varrho$ is a homomorphism, as justified in \ref{classical_vector space}.
+$\varrho$ is a homomorphism, as justified in \ref{prop:real_vector_space}.
 $\mathcal{C}$ is isomorphic to $\varrho(\mathcal{C}) \subseteq L^1(\mathcal{S}, \mu)$ as a vector space.
@@
 Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c} \in \mathcal{C}$.
-Let $q^i$ and $\hat{q}^j$ be two sets of independent state variables in $U\subseteq \mathcal{S}$.
+Let $q^i$ and $\hat{q}^j$ be two sets of independent state variables that fully identify states within $U\subseteq \mathcal{S}$.
 Let $\rho_\mathcal{c}(q^i)$ and $\rho_\mathcal{c}(\hat{q}^j)$ be the expression in local coordinates of the distribution.
@@
 Let $\rho \in C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ be a Lebesgue integrable continuous function.
-If $\rho_\mathcal{c} \neq \rho$ then $\exists U \in \mathcal{S} \; | \; \int_{U} \rho_\mathcal{c} d \mu \neq \int_{U} \rho d \mu$.
+If $\rho_\mathcal{c} \neq \rho$ then $\exists U \in \mathcal{S}$ such that $\int_{U} \rho_\mathcal{c} d \mu \neq \int_{U} \rho d \mu$.
 Both integrals are finite as $\rho_\mathcal{c}, \rho \in L^1(\mathcal{S}, \mu)$.
-Both integrals are expressible with the same state variable as all elements are defined on the same differential structure.
+Both integrals are expressible with the same state variables as all elements are defined on the same differential structure.
 Therefore there exists an outcome for $\mathcal{S}$ that can physically distinguish the two distributions.
@@
 $\varrho : \mathcal{C} \rightarrow C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ is surjective.
-$\varrho$ is a homomorphism, as justified in \ref{classical_vector space}.
+$\varrho$ is a homomorphism, as justified in \ref{prop:real_vector_space}.
 $\varrho$ is an isomorphism between $\mathcal{C}$ and $C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as vector spaces.
@@
 
-\begin{prop}\label{symplectic_manifold}
+\begin{prop}\label{prop:symplectic_manifold}
 The state space $\mathcal{S}$ for the particles of a classical material is a symplectic manifold formed by a cotangent bundle $T^*\mathcal{Q}$ equipped with the canonical two-form $\omega$.
@@
 \begin{justification}
-	We claim the simplest state space $\mathcal{S}$ that allows state-variable-invariant densities is the cotangent bundle $T^*\mathcal{Q}$ of a single dimensional manifold $\mathcal{Q}$.
+	We claim the simplest (i.e.
+lowest dimension) state space $\mathcal{S}$ that allows state-variable-invariant densities is the cotangent bundle $T^*\mathcal{Q}$ of a single dimensional manifold $\mathcal{Q}$.
 Consider a set of state variables for $U \in \mathcal{S}$: the simplest unit system is one defined by a single state variable.
@@
 $\omega$ is invariant under state variable change.
-$\omega$ is a symplectic metric for $\mathcal{S}=T^*\mathcal{Q}$.
+$\omega$ is the symplectic form for $\mathcal{S}=T^*\mathcal{Q}$.
 $\mathcal{S} = (T^*\mathcal{Q}, \omega)$ is a symplectic manifold.
@@
 
-Finally, we should note that we are in a position similar to the one discussed in \ref{discrete_state_metric} for discrete states.
+Finally, we should note that we are in a position similar to the one discussed in \ref{prop:discrete_measure} for discrete states.
 We have a topological space plus a measure that allows to count states and a symplectic form that allows to count possibilities, both expressible in terms of state variables.
@@
 
-\begin{prop}\label{canonical_transformation}
+\begin{prop}\label{prop:symplectomorphism}
 	A deterministic and reversible evolution map for the particles of a classical material is a self-symplectomorphism.
@@
 	We claim $\mathcal{T}_{\Delta t}$ is a self-homeomorphism on $T^*\mathcal{Q}$.
-The state space for the particles of a classical material is $T^*\mathcal{Q}$ by \ref{symplectic_manifold}.
+The state space for the particles of a classical material is $T^*\mathcal{Q}$ by \ref{prop:symplectic_manifold}.
 $\mathcal{T}_{\Delta t}$ is a deterministic and reversible evolution map and by \ref{prop:homeomorphism} is a self-homeomorphism.
@@
 A deterministic and reversible process conserves the number of states and possibilities.
-$\omega$ is the metric for the cardinality of possibilities.
-$\omega$ is invariant under a deterministic and reversible evolution map.
+$\omega$ is the two-form that returns the count of possibilities.
+$\omega$ is invariant under deterministic and reversible evolution.
 $\mathcal{T}_{\Delta t}$ is a symplectomorphism by definition.
@@
 
-\begin{prop}\label{hamiltons_equations}
-	A continuous deterministic and reversible process for the particles of a classical material admits a Hamiltonian $H \in C^2(T^*\mathcal{Q}, \mathbbm{R})$ that allows us to write the laws of evolution as
+\begin{prop}\label{prop:hamiltons_equations}
+	A continuous deterministic and reversible evolution for the particles of a classical material admits a Hamiltonian $H \in C^2(T^*\mathcal{Q}, \mathbbm{R})$ that allows us to write the laws of evolution as
 \begin{align*}
@@
 We claim the vector field $S \in T\mathcal{S}$ for the infinitesimal displacement $S^a = d\xi^a/dt$ admits a potential $H \in C^2(T^*\mathcal{Q}, \mathbbm{R})$ such that $S^{a} \omega_{ab} = \partial_{b}H$.
-The state space $\mathcal{S}$ for the particles of a classical material is a symplectic manifold by \ref{symplectic_manifold}.
-The map for infinitesimal time evolution $\mathcal{T}_{dt}$ is an infinitesimal self-symplectomorphism by \ref{canonical_transformation}.
+The state space $\mathcal{S}$ for the particles of a classical material is a symplectic manifold by \ref{prop:symplectic_manifold}.
+The map for infinitesimal time evolution $\mathcal{T}_{dt}$ is an infinitesimal self-symplectomorphism by \ref{prop:symplectomorphism}.
 By \ref{symplectomorphism_generator} the infinitesimal displacement $S$ admits a potential $H \in C^2(T^*\mathcal{Q}, \mathbbm{R})$ such that $S^{a} \omega_{ab} = \partial_{b}H$
@@
 
-\begin{prop}\label{complete_particle_state_space}
+\begin{prop}\label{prop:complete_particle_state_space}
 	The complete state space $\bar{\mathcal{S}}$ for the particles of a classical material is a $2n+1$ dimensional differentiable manifold.
@@
 $T^*\mathcal{Q}$ is a differentiable manifold.
-Time evolution is a diffeomorphism by \ref{canonical_transformation} therefore derivatives between state variables and time variables always exist.
+Time evolution is a diffeomorphism by \ref{prop:symplectomorphism} therefore derivatives between state variables and time variables always exist.
 $\bar{\mathcal{S}}$ is a differentiable manifold.
@@
 A distribution of material over states at equal time $\rho_\mathcal{c} : \mathcal{S}_{t=t_0} \rightarrow \mathbbm{R}$ will depend on the choice of time variable, therefore the complete state space will not be useful as it will not be time invariant.
-What we are really interested in is the \emph{evolution space} of a classical material, which we'll indicate by $\bar{\mathcal{C}}$.
-To each evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$ will correspond a distribution $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ over all particle states at all times.
+What we are really interested in is the \emph{state evolution space} of a classical material, which we'll indicate by $\bar{\mathcal{C}}$.
+To each state evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$ will correspond a distribution $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ over all particle states at all times.
 This distribution will be differentiable: it's differentiable over all equal time hyper-surfaces and time evolution is differentiable.
@@
 
-\begin{prop}\label{material_evolution_state_space}
-	The evolution state space $\bar{\mathcal{C}}$ for a classical material is isomorphic to the space of differentiable functions that are Lebesgue integrable on any equal time hyper-surface.
+\begin{prop}\label{prop:state_evolution_space}
+	The state evolution space $\bar{\mathcal{C}}$ for a classical material is isomorphic to the space of differentiable functions that are Lebesgue integrable on any equal time hyper-surface.
 That is $\bar{\mathcal{C}} \cong  C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$ where $L^1_t(\bar{\mathcal{S}}, \mu)=\{{\rho: \bar{\mathcal{S}} \rightarrow \mathbbm{R}} \; | \; {\int_{\mathcal{S}_{t=t_0}} |\rho| d\mu < \infty} \; \forall \, t, t_0 \}$.
@@
 $\rho_{\bar{\mathcal{c}}} \in L^1_t(\bar{\mathcal{S}}, \mu)$.
-Time evolution is a diffeomorphism by \ref{hamiltons_equations}.
+Time evolution is a diffeomorphism by \ref{prop:hamiltons_equations}.
 $\rho_{\bar{\mathcal{c}}}$ is differentiable along the time variable as well as the state variables.
@@
 
-\begin{prop}\label{relativistic_symplectic_manifold}
+\begin{prop}\label{prop:relativistic_symplectic_manifold}
 	The complete state space $\bar{\mathcal{S}}$ for the particles of a classical material is a hyper-surface of the symplectic manifold formed by a cotangent bundle $T^*\mathcal{M}$ equipped with the canonical two-form $\omega = \sum dq^i \wedge dp_i - dt \wedge dE = \sum dq^\alpha \wedge dp_\alpha$.
@@
 Locally $\mathcal{M} \cong \mathcal{Q} \times \mathbbm{R}$ as time and state variables are independent quantities.
-As in \ref{symplectic_manifold}, invariant densities are defined on the symplectic manifold $(T^*\mathcal{M}, \omega)$.
+As in \ref{prop:symplectic_manifold}, invariant densities are defined on the symplectic manifold $(T^*\mathcal{M}, \omega)$.
 Let $\bar{\omega}$ be the co-vector component associated with $t$.
@@
 
-\begin{prop}\label{relativistic_canonical_transformation}
-	A time-dependent deterministic and reversible evolution map for a homogeneous classical material is a self-symplectomorphism on $T^*\mathcal{M}$.
+\begin{prop}\label{prop:relativistic_symplectomorphism}
+	A time-dependent deterministic and reversible evolution map for the particles of a classical material is a self-symplectomorphism on $T^*\mathcal{M}$.
 That is: $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ such that $\mathcal{T}_{\Delta s}^*\omega = \omega$ where $\mathcal{T}_{\Delta s}^*$ is the pullback of $\mathcal{T}_{\Delta s}$.
@@
 	We claim $\mathcal{T}_{\Delta s}$ is a self-homeomorphism on $T^*\mathcal{M}$.
-The complete state space for particles of a classical material is $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ by \ref{relativistic_symplectic_manifold}.
+The complete state space for particles of a classical material is $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ by \ref{prop:relativistic_symplectic_manifold}.
 $\mathcal{T}_{\Delta s}$ is a deterministic and reversible evolution map and by \ref{prop:homeomorphism} is a self-homeomorphism.
@@
 A deterministic and reversible process conserves the number of states and possibilities.
-$\omega$ is the metric for the cardinality of possibilities.
-$\omega$ is invariant under a deterministic and reversible evolution map.
+$\omega$ is the two-form that returns the count of possibilities.
+$\omega$ is invariant under deterministic and reversible evolution.
 $\mathcal{T}_{\Delta s}$ is a symplectomorphism by definition.
@@
 
-\begin{prop}\label{relativistic_hamiltons_equations}
-	A continuous time-dependent deterministic and reversible process for a homogeneous classical material admits an invariant Hamiltonian $\mathcal{H}: T^*\mathcal{M} \rightarrow \mathbbm{R}$ that allows us to write the laws of evolution as
+\begin{prop}\label{prop:relativistic_hamiltons_equations}
+	A continuous time-dependent deterministic and reversible process for the particles of a classical material admits an invariant Hamiltonian $\mathcal{H}: T^*\mathcal{M} \rightarrow \mathbbm{R}$ that allows us to write the laws of evolution as
 	\begin{align*}
@@
 We claim the vector field $S \in T(T^*\mathcal{M})$ for the infinitesimal displacement $S^a = d_s\xi^a$ admits a potential $\mathcal{H}$ such that $S^{a} \omega_{ab} = \partial_{b}\mathcal{H}$.
-The map for infinitesimal evolution $\mathcal{T}_{ds}$ is an infinitesimal self-symplectomorphism by \ref{relativistic_canonical_transformation}.
+The map for infinitesimal evolution $\mathcal{T}_{ds}$ is an infinitesimal self-symplectomorphism by \ref{prop:relativistic_symplectomorphism}.
 By \ref{symplectomorphism_generator} the infinitesimal displacement $S$ admits a potential $\mathcal{H}$ such that $S^{a} \omega_{ab} = \partial_{b}\mathcal{H}$
@@
 In other words: $E$ is the value of the Hamiltonian, the energy of the system.
-We have found the constraint we discussed when introducing \ref{relativistic_symplectic_manifold}.
+We have found the constraint we discussed when introducing \ref{prop:relativistic_symplectic_manifold}.
 
@@
 
-\begin{prop}\label{form_of_extended_hamiltonian}
+\begin{prop}\label{prop:form_of_invariant_hamiltonian}
 	The form of the invariant Hamiltonian is $\mathcal{H} = \lambda(t,q^i,E,p_i) \, (H(t,q^i,p_i) - E)$ where $H(t,q^i,p_i)$ is the standard Hamiltonian at each instant $t$, the conjugate time variable $E = H(t,q^i,p_i)$ represents the value of the Hamiltonian, and  $\lambda : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ where $\lambda(\xi^a) = d_s t$.
@@
 
-\begin{prop}\label{relativistic_distribution_evolution}
+\begin{prop}\label{prop:relativistic_state_evolution}
 	Let $\rho_{\bar{\mathcal{c}}}$ be the distribution of material associated with the evolution $\bar{\mathcal{c}}$.
@@
 $supp(\rho_{\bar{\mathcal{c}}}) \subseteq \bar{\mathcal{S}}$ as the distribution is defined on the complete state space of the particles.
-$supp(\mathcal{H}) = T^*\mathcal{M} \,\backslash\, \bar{\mathcal{S}}$ by \ref{form_of_extended_hamiltonian}.
+$supp(\mathcal{H}) = T^*\mathcal{M} \,\backslash\, \bar{\mathcal{S}}$ by \ref{prop:form_of_invariant_hamiltonian}.
 $supp(\mathcal{H} \rho_\mathcal{c}) = supp(\mathcal{H}) \cap supp(\rho_\mathcal{c}) = \varnothing$.
@@
 
-\begin{prop}\label{initial_conditions}
+\begin{prop}\label{prop:tangent_bundle}
 	Let $\mathcal{M}$ be the $n+1$ dimensional space-time manifold over which trajectories are defined.
@@
 $\mathcal{M}$ defines the unit system.
-$\bar{\mathcal{S}} \subset T^*\mathcal{M}$ by \ref{relativistic_symplectic_manifold}.
+$\bar{\mathcal{S}} \subset T^*\mathcal{M}$ by \ref{prop:relativistic_symplectic_manifold}.
 
@@
 Suppose $d_s x^\alpha$ is not part of the state.
-$d_s x^\alpha = d_s q^\alpha$ is determined by the law of evolution.
-$x^\alpha(s)$ is determined by the law of evolution.
-$q^\alpha$ determines the trajectory $x^\alpha(s)$ and the state $\mathcal{s}$, as the material is minimal.
+Then $d_s x^\alpha = d_s q^\alpha$ must be determined by the law of evolution.
+$x^\alpha(s)$ is determined by the law of evolution therefore $q^\alpha$ are the only state variables needed from the trajectory.
+As the material is minimal, $q^\alpha$ determines the full state $\mathcal{s}$.
 This is a contradiction: $\mathcal{s}$ is not identified by just $q^\alpha$.
@@
 $\rho_{\bar{\mathcal{c}}}(x^\alpha, u^\alpha)$ is differentiable by the same logic expressed in \ref{prop:differentiable_manifold}.
-$(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is a diffeomorphism as it as a homeomorphism that maps a differentiable function to a differentiable function.
+$(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is a diffeomorphism as it is a homeomorphism that maps a differentiable function to a differentiable function.
 \end{justification}
@@
 
-\begin{prop}\label{lagrangian_mechanics}
-	Let $\mathcal{H}$ be an extended Hamiltonian defined on $T^*\mathcal{M}$.
+\begin{prop}\label{prop:lagrangian}
+	Let $\mathcal{H}$ be an invariant Hamiltonian defined on $T^*\mathcal{M}$.
 Under the kinematic assumption, we can define the Legendre transform $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$ which we call the \emph{extended Lagrangian}.
@@
 	We claim $\mathcal{H}$ is convex in $p_\alpha$.
-By \ref{initial_conditions} we can write the velocity $u^\alpha=u^\alpha(q^\alpha, p_\alpha)$ as a differentiable function of the state variables.
+By \ref{prop:tangent_bundle} we can write the velocity $u^\alpha=u^\alpha(q^\alpha, p_\alpha)$ as a differentiable function of the state variables.
 At fixed $q^\alpha$, $u^\alpha$ must be invertible.
@@
 
-\begin{prop}\label{riemannian_manifold}
-	The space-time manifold $\mathcal{M}$ is a pseudo-Riemannian manifold $(\mathcal{M}, g)$. $g$ has signature $(3,1)$ and locally is $g = (dx^i)^2 - c^2 dt^2$.
+\begin{prop}\label{prop:riemannian_manifold}
+	The space-time manifold $\mathcal{M}$ is a pseudo-Riemannian manifold $(\mathcal{M}, g)$. $g$ has signature $(n,1)$ and locally is $g = (dx^i)^2 - c^2 dt^2$.
 \end{prop}
@@
 \begin{justification}
-	We claim $(\mathcal{M}, g)$ is a pseudo-Riemannian manifold where $g$ is a metric induced by $\omega$ under kinematic equivalence.
+	We claim $(\mathcal{M}, g)$ is a pseudo-Riemannian manifold where $g$ is a metric tensor induced by $\omega$ under kinematic equivalence.
 Let $(dx^\alpha, du^\alpha)$ be an infinitesimal range of possibilities for the kinematic variables of the particles of a minimal classical material.
@@
 Therefore $g_{ii} > 0$ and $g_{00} < 0$.
-$g$ has signature $(3, 1)$.
+$g$ has signature $(n, 1)$.
 Choosing the same unit for all $dx^i$ and a temporal unit for $dt$ we can choose coordinates such that locally $g = (dx^i)^2 - c^2 dt^2$.
@@
 
-\begin{prop}\label{kineticHamiltonian}
+\begin{prop}\label{prop:kinetic_hamiltonian}
 The extended Hamiltonian for the particles of a classical material under the kinematic assumption is $\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta) + m^2 c^2)$.
@@
 $\partial_{u_\beta} p_\alpha = m \delta_\alpha^\beta$.
-After integration $p_\alpha = m u_\alpha + A_\alpha(x)$ where $A$ is an arbitrary vector field.
+After integration $p_\alpha = m u_\alpha + A_\alpha(x)$ where $A_\alpha$ is an arbitrary vector field.
 
@@
 Let $s$ be the natural parametrization such that $u^\alpha u_\alpha = - c^2$.
-$V= \frac{1}{2}m c^2$ as $\mathcal{H}=0$ by \ref{form_of_extended_hamiltonian}.
+$V= \frac{1}{2}m c^2$ as $\mathcal{H}=0$ by \ref{prop:form_of_invariant_hamiltonian}.
 $\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta) + m^2 c^2)$.
@@
 	We claim $\Psi$ is a vector space over $\mathbbm{C}$.
-The abelian group $\Psi$ can be extended with the operations defined by $T$ as was done in \ref{classical_vector space} for the infinitesimally reducible case.
+The abelian group $\Psi$ can be extended with the operations defined by $T$ as was done in \ref{prop:real_vector_space} for the infinitesimally reducible case.
 
@@
 Therefore $\psi(q^i) \in C(\mathcal{Q}, \mathbbm{C})$.
-$\rho_\psi(q^i) \in L^1(\mathcal{Q}, \mu, \mathbbm{R})$ by the same arguments in the justification for \ref{integration}.
+$\rho_\psi(q^i) \in L^1(\mathcal{Q}, \mu, \mathbbm{R})$ by the same arguments in the justification for \ref{prop:integration}.
 Therefore $\psi(q^i) \in L^2(\mathcal{Q}, \mu, \mathbbm{C})$.

### 2017-01-04 2015CompleteDerivation/completeDerivation.tex
Cleanup and added biblio items

@@
 While some branches of physics (e.g.
-Newtonian mechanics, thermodynamics, special relativity) are derived from physical principles (e.g.
+Newtonian mechanics(ref?), thermodynamics(ref?), special relativity(ref?)) are derived from physical principles (e.g.
 Newton's laws, laws of thermodynamics, invariance of the speed of light), others start by simply setting a mathematical framework.
-It is the case for classical Lagrangian mechanics, where we start by assuming the existence of a Lagrangian function whose path integral is minimized during the motion.
-It is the case for classical Hamiltonian mechanics, where we start by assuming the state is described by conjugate pairs $(q^i, p_i)$ and evolves according to Hamilton's equations.
-It is the case for quantum mechanics, where the state space is a complex inner product vector space and the state evolves according to Schroedinger's equations.
+It is the case for classical Lagrangian mechanics(ref?), where we start by assuming the existence of a Lagrangian function whose path integral is minimized during the motion.
+It is the case for classical Hamiltonian mechanics(ref?), where we start by assuming the state is described by conjugate pairs $(q^i, p_i)$ and evolves according to Hamilton's equations.
+It is the case for quantum mechanics(ref?), where the state space is a complex inner product vector space and the state evolves according to Schroedinger's equations.
 That is, at the very core of some of our physical theories we have mathematical definitions instead of physical ideas.
 
-If we ask ``when can a particular physical system be suitably described by Hamiltonian (or by classical) mechanics?" the only answer we have is ``when it fits the mathematical framework".
+If we ask ``when can a particular physical system be suitably described by classical Hamiltonian mechanics?" the only answer we have is ``when it fits the mathematical framework".
 While many seem to be content with this answer (e.g.
@@
 This is a problem since mathematical structures are not enough to characterize the physical system being described.
-For example, the same math for linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems.
-The mathematical treatment is the same because it captures aspects of the description that are common to all cases.
+For example, the same mathematical framework for linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems.\cite{Roland,Barron,Borutzky} The mathematical treatment is the same because it captures aspects of the description that are common to all cases.
 But for the very same reason, we cannot infer what system is being described if all we are given is the mathematical model.
-In the jargon of linguistic/computer science/phylosophy: math captures the syntax (the relationships between the objects) but does not capture the semantic (the meaning of each object).
+In the jargon of linguistic/computer science/phylosophy\cite{Freiding,Friedman}: math captures the syntax (the relationships between the objects) but does not capture the semantic (the meaning of each object).
 
@@
 The idea was to create a math-to-physics dictionary so that each mathematical concept was clearly mapped to a crisp physical concept.
-But to make sure that such dictionary was complete, we realized we had to make sure that the whole relevant mathematical framework could be recovered.
-
-This redefined our mission to re-deriving as much as possible starting from physical assumptions.
+But to make sure that such dictionary was complete, we realized we had to make sure that the whole relevant mathematical framework could be recovered: the translation needs to work both ways.
+
+This redefined our mission to re-deriving as much as possible of the mathematical structure by starting from the smallest set of physical assumptions.
 It also forced us to push further and further back toward the basic definitions.
@@
 We also realized that sometimes different mathematical concepts could be justified from the same physical starting point.
-That is: by putting physics first we could achieve a sort of unification of ideas that was not practicable by math alone.
+That is: by putting physics first we could achieve a sort of unification of ideas that was not evident in the math alone.
 
@@
 The ability to apply the same physical ideas far and wide is what gives us confidence that these are truly general concept that capture the physical aspects we are describing.
-The breath itself is what makes the work interesting and more than a curiosity.
+The breath itself is what makes the work interesting and more than a mere curiosity.
 
@@
 
-%TODO: Point out most modern physics starts by assuming some mathematical framework. Pros and cons of that.
-
-%TODO: High level view of the derivation
-
-%While some branches of fundamental physics (e.g.
-Newtonian mechanics and special relativity) are founded on physical laws or principles others (e.g.
-Lagrangian/Hamiltonian/Quantum mechanics) start by postulating mathematical frameworks or relationships.
-The latter approach presents a significant drawback: the mathematical structure is not enough to define the physical system it describes.
-We are left asking: under what physical conditions a particular system is described by Lagrangian (or Quantum) mechanics?
-Naturally one \emph{can} define a Lagrangian system (or a Quantum system) as one that obeys those rules, but that begs the question.
-Another issue is that each element of the mathematical structure may or may not correspond to some physical concept.
-Suppose the earning (or losses) of two companies are governed by Hamiltonian's equations, which variable is on the cotangent space of the other?
-In other words: if we want to fully appreciate what the fundamental theories of physics are describing, what each mathematical concept corresponds to in the physical world, we need to start by characterizing the physical system we are describing in a precise enough way that it can be encoded in mathematical language.
-Naturally, the physical insight we gain will be relevant for only that type of system (i.e.
-other different systems may \emph{happen} to use the same equations), but what we'll see is that the assumptions we make are minimal and can apply to a variety of systems, especially the ones that are considered fundamental.
-It is indeed surprising how so much can be derived by imposing so little.
-
-%This work aims to re-organize the known elements and equations in a more consistent and comprehensive way, leading to better insight on why the fundamental concepts and laws are what they are.
-We will start with physical assumptions, that clarify the models we impose on the physical world, and give arguments on when those assumptions can be considered valid.
-We will justify our mathematical definitions, based on the formal properties of the objects of our discussion.
-And will, as a consequence, re-derive known results and theories.
-We will strive to do this in a way that is mathematically meaningful, philosophically consistent and mathematically precise.
-
-%We'll use concepts from different disciplines, such as set theory, differential geometry, relativity, Hamiltonian and Lagrangian mechanics, and we'll find interesting connections among them.
-We'll keep names and notation as consistent as possible to current use across the different disciplines.
-This may sometimes lead to some non sequitur as it will not be immediately clear why the new definitions are equivalent to the standard ones.
-These are typically resolved by subsequent derivation of the expected properties.
-
-%No mathematical breakthrough should be expected: the goal, after all, is to derive the \emph{known} framework from a set of \emph{simple} definitions in the most \emph{obvious} way possible.
-No proof is longer than a couple of paragraphs, so the word \emph{theorem} is avoided in favor of \emph{proposition} and \emph{corollary}.
-The novel, and surprising, result is how so much can be derived from so little.
-
 \section{Organization and style}
@@
 
-The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology, measure theory, differential geometry (symplectic and Rimannian), group theory and vector spaces, statistics, information theory, Hamiltonian mechanics, Lagrangian mechanics, quantum mechanics, thermodynamics, special relativity and so on.
+The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology\cite{LeeTM}, measure theory (ref?), differential geometry (symplectic and Rimannian)\cite{Lee}, group theory and vector spaces\cite{Lang,Young}, statistics (ref?), information theory (ref?), Hamiltonian mechanics, Lagrangian mechanics\cite{classical_dynamics}, quantum mechanics (ref?), thermodynamics (ref?), special relativity (ref?) and so on.
 Using idea from all these areas is necessary, in fact unraveling their connections is part of the value of this work, but it does create a number of problems.
@@
 
-This link between state definition and deterministic processes should not be too surprising as the state, in the context of thermodynamics and systems theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system}.
-As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can indeed be calculated, measured and prepared.
+This link between state definition and deterministic processes should not be too surprising as the state, in the context of control theory and dynamical systems theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system}.(ref?) As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can indeed be calculated, measured and prepared.
 We can also describe the evolution of each element of the distribution provided that: we have a way to isolate it and study it under deterministic and reversible motion (so that we can define microstates); the non-deterministic motion does not alter the system (the set of microstates is preserved by the evolution).
@@
 
-However precise our measurements are, we can only gather a finite amount of statistics and each outcome is expressed by a finite set of digits; therefore, the set of outcomes is countable.\footnote{The information provided by the process as measured by Shannon's entropy is finite.} Note that different outcomes for the same process can overlap.
+However precise our measurements are, we can only gather a finite amount of statistics and each outcome is expressed by a finite set of digits; therefore, the set of outcomes is countable.\footnote{The information provided by the process as measured by Shannon's entropy\cite{Shannon} is finite.} Note that different outcomes for the same process can overlap.
 For example, $4.12 \pm 0.05$ cm and $4.13 \pm 0.05$ cm are both legitimate possible outputs of the same measurement device.
@@
 Given that scientific reductionism (i.e.
-the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics, we now explore how to characterize a system in terms of its components.
+the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics,(ref?) we now explore how to characterize a system in terms of its components.
 That is, we want to study the relationship between the state space of a composite system and the state space of its parts.
@@
 Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a general physical justification for.\footnote{One can, though, make the argument that given any distribution $\rho$ one can find an infinitely smooth $\rho_{sm}$ such that the difference in description is small compared to the error already introduced by assuming the system to be homogeneous and infinitely reducible.
-The Whitney approximation theorem makes this mathematically well defined.}
+The Whitney approximation theorem(ref?) makes this mathematically well defined.}
 
@@
 is also conserved.
-These statements are the physical justification of Gromov's non-squeezing theorem and Liouville's theorem.
+These statements are the physical justification of Gromov's non-squeezing theorem \cite{Gromov,deGosson,Stewart} and Liouville's theorem.
 And they give intuitive insight on the geometry of Hamiltonian systems.
@@
 This means that our measure $\mu$ needs to be modified to define integrals over the time variable.\footnote{Looking ahead, another way to frame the problem is that Galilean transformations do not preserve phase space volume.
-Lorentz transformations, on the other hand, do and it is this requirements plus the kinematic assumption that will recover them later.}
+Lorentz transformations, on the other hand, do\cite{Treumann} and it is this requirements plus the kinematic assumption that will recover them later.}
 
@@
 On the other hand, the change of state variables must preserve the ability to uniquely identify states at each instant in time.
-In this sense, time is not just like another state variable.\footnote{In relativity, for example, it is common to use foliations and the $3+1$ formalism to recover the special character of time when performing calculations or to gain better physical insight.}
+In this sense, time is not just like another state variable.\footnote{In relativity, for example, it is common to use foliations and the $3+1$ formalism to recover the special character of time when performing calculations or to gain better physical insight.\cite{Gourgoulhon}}
 
@@
 
+We recognize the equations as those for Hamiltonian mechanics on the extended phase space.\cite{Synge, Struckmeier}
+
 The trajectory in time $t(s)$ is of particular importance.
@@
 In Newtonian physics, $\vec{F}=m\vec{a}$ provides such link.
-In Hamiltonian (and Lagrangian) mechanics, the link is implicitly given by the choice of the Hamiltonian (or Lagrangian).
+In Hamiltonian (and Lagrangian) mechanics, the link is implicitly given by the choice of the Hamiltonian (or Lagrangian).(ref?)
 
@@
 It was under that premise that we were able to show that the trajectories are continuous, that we could define $\omega$ and therefore $g$.
-Non-local quantum effects, such as the ones featured in EPR experiments, are not examples of deterministic and reversible motion.
+Non-local quantum effects, such as the ones featured in EPR\cite{EPR} experiments, are not examples of deterministic and reversible motion.
 Therefore they do not have anything to violate: the maximum speed does not apply to them.
@@
 In this context, we will call \emph{fragment} an infinitesimal part of our classical electron.
-The Shannon entropy associated with the distribution is finite as well.
+The Shannon entropy\cite{Shannon} associated with the distribution is finite as well.
 $I(\rho_\mathcal{c}) = - \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq \wedge dp = I_0$.
@@
 	This notion may seem in contradiction to the current practice.
-For example, the S-matrix, one of the main devices for calculating scattering amplitudes and cross sections, is derived by assuming unitary evolution.
-Doesn't it imply we are assuming that deterministic and reversible evolution is happening at every instant?
+For example, the S-matrix, one of the main devices for calculating scattering amplitudes and cross sections, is derived by assuming unitary evolution.(ref?) Doesn't it imply we are assuming that deterministic and reversible evolution is happening at every instant?
 
@@
 It does, though, give good reason why we have those problems and how the desire to know ``what is really going on" is hindered by the very assumption that the quantum description is based on.
-
-%Note on composite systems and distinguishability.
-When putting together two classical system, just sum distributions.
-When putting together quantum system we have to clarify: combine to create one quantum system (sum in vector space) or two separate quantum system (symmetrized tensor product).
-In classical mechanics, 2a + 2b = (a+b) + (a+b).
-In quantum note |a> x |b> = |a>|b> + |b>|a> but (|a> + |b>) x (|a> + |b>) = |a>|a> + ...
-That is two quantum states in two different states is not the same as two quantum states spread equally over those two states.
-In classical they are.)
-
-% Muon example.
-Consider a muon and its decay into an electron and two neutrinos: it is clear that the three outgoing particles have a state and trajectory of their own, it is clear that the resulting total mass and energy came from the muon.
-Yet, before the decay, we cannot ascribe an internal independent state and evolution to each part.
-The state of a muon is not some combination of the state of an electron and two neutrinos.
-That is: the unstated part is not just microstates.
-
-%TODO: moved from classical
-%This intuitive picture is unfortunately not suitable to be generalized to different contexts, so we develop another.
-
-%Consider the following expression:
-%\begin{align*}
-%\mathcal{c} = \sum\limits_{i \in \mathbbm{I}} a_i %\mathcal{e}(i)
-%\end{align*}
-%$\mathbbm{I}$ represents all the possible values of a state variable representing the best description of an infinitesimal part that a composite system can give.
-$\mathcal{e}(i)$ represents a composite system made of a unit amount of particles, all prepared in the state determined by $i$.
-$a_i$ represents a transformation that changes the whole state, without affecting the value $i$ for any infinitesimal part.
-For a classical system, $i$ represents the full information about the infinitesimal part, and therefore includes the value of all state variables (i.e.
-the state vector); the only transformation left is increasing/decreasing the amount of particles by a scalar multiple.
-For a different system, where $i$ does not represent the full state of the parts, $\mathcal{e}(i)$ will be an ensemble and $a_i$ also represents an internal transformation of such ensemble.
-This is intuitively how we decouple the composite state into the part $i$ that describes the infinitesimal parts, and the part $a_i$ that describes how the parts are composed.
-Giving a full description of the state space of a decomposable system means fully characterizing these components: the state variable $\mathbbm{I}$ and the set of transformations $A$.
-In the classical case $\mathcal{e}(i)$ form a basis of a real vector space, where $i$ is the state vector for an infinitesimal part, and $a_i$ a real valued coefficient that represents the amount of particles in each state, which describes how the parts are combined.
 
@@
 \bibitem{Struckmeier} Struckmeier, J., ``Hamiltonian dynamics on the symplectic extended phase space for autonomous and non-autonomous systems'', J. Phys. A: Math. Gen. 38, pp. 1257--1278, (2005).
+\bibitem{Lee} Lee, J. M.: Introduction to Smooth Manifolds. Springer (2003)
+\bibitem{LeeTM} Lee, J. M.: Introduction to Topological Manifolds. Springer (2000)
+\bibitem{Lang} Lang, S. L.: Algebra. Springer (2002)
+\bibitem{Young} Young, N.: An Introduction to Hilbert Space. Cambridge Univ. Press, Cambridge, UK (1988)
+
+\bibitem{Barron} Barron, R. F.: 8.2 Lumped-parameter analysis. In: Barron, R. F.: Industrial Noise Control and Acoustics. CRC Press (2002)
+%online at http://ssu.ac.ir/cms/fileadmin/user_upload/Daneshkadaha/dbehdasht/behdasht_herfei/e_book/_Randall_F._Barons_Industrial_Noise_Control_and_.org__1_.pdf
+
+\bibitem{Roland}  
+ Thomas, R. E., Rosa, A. J.: The analysis and design of linear circuits. Wiley (2001)
+
+\bibitem{Borutzky} Borutzky, W.: Bond graph methodology: development and analysis of multidisciplinary dynamic system models.
+Springer (2010)
+
+\bibitem{Freiding} Freidin, R., Lasnik, H.: Syntax critical concepts in linguistics. Routledge (2006)
+
+\bibitem{Friedman} Friedman, D. P., Wand, M., Haynes, C.T.: Essentials of Programming Languages. MIT Press (2001)
+
+\bibitem{Treumann} Treumann, R. A., Nakamura, R., Baumjohann,  W.: Relativistic transformation of phase-space distributions. Annales Geophysicae (2011)
+
+\bibitem{Gourgoulhon} Gourgoulhon, E.: 3+1 Formalism and Bases of Numerical Relativity. arXiv. http://arxiv.org/abs/gr-qc/0703035 (2007)
+
+\bibitem{EPR} Einstein, A., Podolsky, B.,  Rosen, N.: Can quantum-mechanical description of physical reality be considered complete? Phys. Rev. 47 777 (1935)
 

### 2017-01-05 2015CompleteDerivation/completeDerivation.tex
Adding Christine's comments

@@
 
-The next step would be to integrate the quantum description with time dependent evolution and the kinematic assumption.
+The next step would be to integrate the quantum description with time dependent evolution (i.e.
+relativistic mechanics) and the kinematic assumption (i.e.
+Lagrangian mechanics).
 We will only provide a sketch, though, without going through the details.
-These would take a significant amount of space while not providing any significantly new insight.
-Also, given that we are limited to single particles (i.e.
+Given that we are limited to single particles (i.e.
 no fields) without spin (i.e.
@@
 
-As for the classical case, the first thing to do is to extend $\mathcal{Q}$ to include time.
+As in the classical case, the first thing to do is to extend $\mathcal{Q}$ to include time.
 This gives us the same $\mathcal{M}$ as before.
-A quantum state is a complex valued distribution $\psi : \mathcal{M}_{t=t_0} \rightarrow \mathbbm{C}$ over an hypersurface at constant time.
-An evolution $\bar{\psi} : \mathcal{M} \rightarrow \mathbbm{C}$ is a complex valued function defined on the whole $\mathcal{M}$.
-The operator for the quantity conjugate of time is $E=\imath\hbar\partial_t$.
+A quantum state is a complex valued distribution $\psi : \mathcal{M}_{t=t_0} \rightarrow \mathbbm{C}$ over a hypersurface at constant time.
+An evolution $\widetilde{\psi} : \mathcal{M} \rightarrow \mathbbm{C}$ is a complex valued function defined on the whole $\mathcal{M}$.
+The operator for the quantity conjugate to time is $E=\imath\hbar\partial_t$.
 
@@
 In the quantum case, this means to make sure the inner product $\langle \psi | \phi \rangle$ remains unchanged under all variable changes, including time.
-This allows to describe the amounts of material and their distribution over fragments in a way that is independent of our description.
+This allows us to describe the amounts of material and their distribution over fragments in a way that is independent of our description.
 
@@
 Continuous evolution will admit an invariant Hamiltonian $\mathcal{H}$ such that $\imath \hbar \partial_s \psi = \mathcal{H} \psi$.
-As for the composite state evolution in the classical case, $\bar{\psi}$ remains unchanged by the evolution as it transports the distribution along $s$.
-Therefore $\mathcal{H} \bar{\psi} = \mathcal{h} \bar{\psi}$ is an eigenfunction of $\mathcal{H}$.
+As for the composite state evolution in the classical case, $\widetilde{\psi}$ remains unchanged by the evolution as it transports the distribution along $s$.
+Therefore $\mathcal{H} \widetilde{\psi} = \mathcal{h} \widetilde{\psi}$ is an eigenfunction of $\mathcal{H}$.
 As in the classical case, we use $\mathcal{H}=0$ to identify the state space: the temporal d.o.f.
-is not independent and therefore there is a constraint between the variables.
+is not independent and therefore there is a constraint among the variables.
 Therefore we have $\mathcal{H} \psi = 0$.
@@
 Note that states are indeed eigenfunctions of the invariant Hamiltonian $\mathcal{H}$: here we do have a justification, as opposed to the standard Hamiltonian $H$ case.
-For example, a massive particle may have fragments with different energy, but they all must have the same mass.
+For example, a massive particle may have fragments with different energy, but they all must carry the same fraction of mass.
 Also note that $\mathcal{H} \psi = 0$ is basically the form of all equations (i.e.
@@
 Therefore the distributions still need to be transported from state space to kinematic variables up to a constant factor.
-Which means we still have a linear relationship between velocity and conjugate momentum.
+This means we still have a linear relationship between velocity and conjugate momentum.
 In terms of the quantum operators: $m\, d_s Q^\alpha = g^{\alpha\beta} P_{\beta} - A^\alpha = m\frac{[Q^\alpha, \mathcal{H}]}{\imath \hbar}$.
 
-This leads to the same general for the invariant Hamiltonian $\mathcal{H}=\frac{1}{2m}((P_\alpha-A_\alpha)g^{\alpha\beta}(P_\beta-A_\beta) + m^2 c^2) = \frac{1}{2m}((-\imath\hbar\partial_\alpha-A_\alpha)g^{\alpha\beta}(-\imath\hbar\partial_\beta-A_\beta) + m^2 c^2)$.
-In the free particle case, we have $\frac{1}{2m}(-\hbar^2 \partial^\alpha \partial_\alpha + m^2 c^2) \bar{\psi} = 0$ which is the Klein-Gordon equation.
-We can also write the operator for kinetic momentum as $mU^\alpha = P^\alpha - A^\alpha = i\hbar D^\alpha$ where $D^\alpha = \partial^\alpha - \frac{A^\alpha}{\imath \hbar}$ is the covariant gauge derivative.
-We have $(\frac{1}{2}mU^\alpha U_\alpha + \frac{1}{2} m c^2) \bar{\psi} = \frac{1}{2m}(-\hbar^2 D^\alpha D_\alpha + m^2 c^2) \bar{\psi} = 0$ which is the Klein-Gordon equation with electromagnetic interaction.
+This leads to the same form for the invariant Hamiltonian $\mathcal{H}=\frac{1}{2m}((P_\alpha-A_\alpha)g^{\alpha\beta}(P_\beta-A_\beta) + m^2 c^2) = \frac{1}{2m}((-\imath\hbar\partial_\alpha-A_\alpha)g^{\alpha\beta}(-\imath\hbar\partial_\beta-A_\beta) + m^2 c^2)$.
+In the free particle case, we have $\frac{1}{2m}(-\hbar^2 \partial^\alpha \partial_\alpha + m^2 c^2) \widetilde{\psi} = 0$ which is the Klein-Gordon equation.
+We can also write the operator for kinetic momentum as $mU^\alpha = P^\alpha - A^\alpha = -i\hbar D^\alpha$ where $D^\alpha = \partial^\alpha + \frac{A^\alpha}{\imath \hbar}$ is the gauge covariant derivative.
+We have $(\frac{1}{2}mU^\alpha U_\alpha + \frac{1}{2} m c^2) \widetilde{\psi} = \frac{1}{2m}(-\hbar^2 D^\alpha D_\alpha + m^2 c^2) \widetilde{\psi} = 0$ which is the Klein-Gordon equation with electromagnetic interaction.
 
@@
 This motivates why it is such an important operator.
-As for the Klein-Gordon equation, the only thing it does is to impose that the norm of the four velocity is the speed of light, a condition much more trivial one may have anticipated.
+As for the Klein-Gordon equation, the only thing it does is to impose that the norm of the four velocity is the speed of light, a condition much more trivial than one may have anticipated.
 
@@
 They give more insight to all the pieces in a comprehensive way and help tear down the artificial walls between the different physical theories.
-This allows to form a picture of fundamental physics that is more unified, something that has been sorely missing.
-
-\subsection{Non deterministic evolution and the projection postulate}
+This allows us to form a picture of fundamental physics that is more unified, something that has been sorely missing.
+
+\subsection{Non-deterministic evolution and the projection postulate}
 
 As we made it clear from the start, deterministic and reversible evolution is an assumption.
-And as we seen in the justifications for assumptions \ref{ass:infinitesimal_reducibility} and \ref{ass:irreducibility} we cannot have determinism at all levels of decomposition while transferring information from one system to another.
-
-In this section we will see how non deterministic processing become necessary when describing measurements for irreducible systems.
-But we'll see that, under the same assumption of irreducibility, we are able to make a connection between the output of deterministic and non deterministic processes.
+And as we have seen in the justifications for assumptions \ref{ass:infinitesimal_reducibility} and \ref{ass:irreducibility} we cannot have determinism at all levels of decomposition while transferring information from one system to another.
+
+In this section we will see how non-deterministic processes become necessary when describing measurements for irreducible systems.
+But we'll see that, under the same assumption of irreducibility, we are able to make a connection between the output of deterministic and non-deterministic processes.
 
@@
 
-Suppose that we create an array of such detectors. This will allow to detect the position of the particle with a resolution given by the aperture of each detector.
-
-Suppose now, though, that the quantum particle state is spread over a region significantly greater than a single detector but smaller than the array. What will it happen?
+Suppose that we create an array of such detectors. This will allow us to detect the position of the particle with a resolution given by the aperture of each detector.
+
+Suppose now, though, that the quantum particle state is spread over a region significantly greater than a single detector but smaller than the array. What will happen?
 
@@
 
-The process is non deterministic in the sense that the state of the particle is not enough to predict which detector will trigger.
+The process is non-deterministic in the sense that the state of the particle is not enough to predict which detector will trigger.
 Maybe it depends on the state of the detector, maybe it depends on the unstated part of the particle.
-The point is that a single measurement take does not tell us much about the incoming particle states.
+The point is that a single measurement does not tell us much about the state of the incoming particle.
 

### 2017-01-05 2015CompleteDerivation/completeDerivation.tex
Using mathbb instead of mathbbm

@@
 
-\usepackage{bbm}
 \usepackage{amsmath}
@@
 A finite precision measurement of the force provides us infinite precision of space, which we ruled out.\footnote{This assumes that we are able to position the probe perfectly, which we could do if we were able to manipulate forces at that precision.} This is what the math is telling us, that if we claim that outcomes can't distinguish isolated points (i.e.
-standard topology on $\mathbbm{R}^n$) then neither can maps.
+standard topology on $\mathbb{R}^n$) then neither can maps.
 It is physical consistency that limits us to continuous functions.
@@
 discrete topology) while real numbers can be measured up to a finite precision (i.e.
-standard topology on $\mathbbm{R}$) From now on, we are going to consider state spaces whose states can be identified, at least within a region, by a finite set of discrete and continuous state variables (i.e.
-the state space is locally isomorphic to $\bigcup\limits_{1 \leq i \leq n} \mathbbm{R}^{m_i}$).
+standard topology on $\mathbb{R}$) From now on, we are going to consider state spaces whose states can be identified, at least within a region, by a finite set of discrete and continuous state variables (i.e.
+the state space is locally isomorphic to $\bigcup\limits_{1 \leq i \leq n} \mathbb{R}^{m_i}$).
 
@@
 \begin{prop}\label{prop:state_variable}
-	A \emph{state variable} is a continuous map $q : U \rightarrow \mathbbm{L}$ where $U \subseteq \mathcal{S}$ and $\mathbbm{L}$ is the space for the possible values.
-If $\mathbbm{L}\cong \mathbbm{Z}$ the variable is said \emph{discrete}.
-If $\mathbbm{L}\cong \mathbbm{R}$ the variable is said \emph{continuous}.
+	A \emph{state variable} is a continuous map $q : U \rightarrow \mathbb{L}$ where $U \subseteq \mathcal{S}$ and $\mathbb{L}$ is the space for the possible values.
+If $\mathbb{L}\cong \mathbb{Z}$ the variable is said \emph{discrete}.
+If $\mathbb{L}\cong \mathbb{R}$ the variable is said \emph{continuous}.
 \end{prop}
@@
 \begin{justification}
-	We claim $\mathbbm{L}$ is a topological space. $\mathbbm{L}$ is a set of physically distinguishable possibilities. $\mathbbm{L}$ is a topological space because of \ref{prop:topology}.
+	We claim $\mathbb{L}$ is a topological space. $\mathbb{L}$ is a set of physically distinguishable possibilities. $\mathbb{L}$ is a topological space because of \ref{prop:topology}.
 
@@
 \begin{prop}\label{prop:independent_state_variables}
-	Two state variables $q_1 : U \rightarrow \mathbbm{L}_1$ and $q_2 : U \rightarrow \mathbbm{L}_2$ are said \emph{independent} if $\exists \mathcal{s}$ such that $q_1(\mathcal{s})=l_1, q_2(\mathcal{s})=l_2 \forall l_1 \in q_1(U), l_2 \in q_2(U)$.
+	Two state variables $q_1 : U \rightarrow \mathbb{L}_1$ and $q_2 : U \rightarrow \mathbb{L}_2$ are said \emph{independent} if $\exists \mathcal{s}$ such that $q_1(\mathcal{s})=l_1, q_2(\mathcal{s})=l_2 \forall l_1 \in q_1(U), l_2 \in q_2(U)$.
 \end{prop}
@@
 But deterministic and reversible evolution cannot take us from a different number of continuous independent state variables (i.e.
-there is no homeomorphism between $\mathbbm{R}^n$ and $\mathbbm{R}^m$).
+there is no homeomorphism between $\mathbb{R}^n$ and $\mathbb{R}^m$).
 That is the number of balls cannot change under deterministic and reversible evolution.
 Therefore we can restrict ourselves to the case where the number of continuous independent state variables is constant without loss of generality.
-This means that, at least locally, the state space is always homeomorphic to $\mathbbm{R}^n$, and is therefore a manifold.
+This means that, at least locally, the state space is always homeomorphic to $\mathbb{R}^n$, and is therefore a manifold.
 
 Discrete variables do not present such problems, as any number of them can be flattened out in a single one (i.e.
-$\mathbbm{Z}^n$ is homeomorphic to $\mathbbm{Z}$).
+$\mathbb{Z}^n$ is homeomorphic to $\mathbb{Z}$).
 They will determine the number of connected components of the state space.
@@
 Let $n$ be the number of independent continuous variables needed to identify $\mathcal{s}$.
-There exists a neighborhood $U$ around $\mathcal{s}$ where we have $(q_1,...,q_n):U\rightarrow \mathbbm{R}^n$.
+There exists a neighborhood $U$ around $\mathcal{s}$ where we have $(q_1,...,q_n):U\rightarrow \mathbb{R}^n$.
 This map is a bijection as $\mathcal{s}$ is identified by those variables.
-$\mathcal{S}$ is homeomorphic to $\mathbbm{R}^n$ around  $\mathcal{s}$.
+$\mathcal{S}$ is homeomorphic to $\mathbb{R}^n$ around  $\mathcal{s}$.
 Let $\mathcal{S}'$ be the set of all states potentially reachable by deterministic and reversible evolution from $\mathcal{s}$.
@@
 $\forall \hat{\mathcal{s}} \in \mathcal{S}'$ there exist a map $\mathcal{T}_{\Delta t}:\mathcal{S} \rightarrow \mathcal{S}$ such that $\hat{\mathcal{s}} =\mathcal{T}_{\Delta t}(\mathcal{s})$.
-$\hat{\mathcal{s}} \in \mathcal{T}_{\Delta t}(U)$ and $\mathcal{T}_{\Delta t}(U)$ is isomorphic to $\mathbbm{R}^n$.
-$\mathcal{S}'$ equipped with the subspace topology is a Hausdorff, second countable topological space everywhere homeomorphic to $\mathbbm{R}^n$.
+$\hat{\mathcal{s}} \in \mathcal{T}_{\Delta t}(U)$ and $\mathcal{T}_{\Delta t}(U)$ is isomorphic to $\mathbb{R}^n$.
+$\mathcal{S}'$ equipped with the subspace topology is a Hausdorff, second countable topological space everywhere homeomorphic to $\mathbb{R}^n$.
 $\mathcal{S}'$ is a manifold of dimension $n$.
@@
 If we work with a discrete topology then, by all means, the way to measure the size of a set is to count the elements in a set.
-But if we use the standard topology on $\mathbbm{R}$, isolated points have measure zero (i.e.
+But if we use the standard topology on $\mathbb{R}$, isolated points have measure zero (i.e.
 we can't physically distinguish them): what defines the number of initial conditions is the range of the state variables.
@@
 Given the state space $\mathcal{S}$ for the infinitesimal parts, which we assume consistently with \ref{prop:manifold} to be a manifold, a composite state $\mathcal{c} \in \mathcal{C}$ will tell us the amount of material for each possible particle state.
-That is, each state is fully identified by a function $\rho_\mathcal{c}: \mathcal{S} \to \mathbbm{R}$ that returns the amount of material in the composite state $\mathcal{c}$ that is prepared in the particle state $\mathcal{s}$.
+That is, each state is fully identified by a function $\rho_\mathcal{c}: \mathcal{S} \to \mathbb{R}$ that returns the amount of material in the composite state $\mathcal{c}$ that is prepared in the particle state $\mathcal{s}$.
 For example, given a certain configuration of gas, we can tell the density of material that is at a specific point in space with a specific value of momentum.
@@
 The state space $\mathcal{C}$ for a classical (i.e.
-homogeneous infinitesimally reducible) material is a vector space over $\mathbbm{R}$ isomorphic to a subspace of the space of continuous functions.
-That is $\mathcal{C} \cong G \subseteq C(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is continuous$\}$, where $\mathcal{S}$ is the state space of an infinitesimal amount of material.
+homogeneous infinitesimally reducible) material is a vector space over $\mathbb{R}$ isomorphic to a subspace of the space of continuous functions.
+That is $\mathcal{C} \cong G \subseteq C(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbb{R} \; | \; \rho$ is continuous$\}$, where $\mathcal{S}$ is the state space of an infinitesimal amount of material.
 \end{prop}
@@
 
-We claim $T$, the set of transformations that increase or decrease the amount of material in the system by a constant factor, is a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics or differential geometry sense (a physical quantity/tensor with a value for each point in space).} isomorphic to $\mathbbm{R}$.
-Consider $\tau: \mathbbm{R} \rightarrow T$ the mapping between a number and the transformation that increases or decreases the amount of material by that factor.
+We claim $T$, the set of transformations that increase or decrease the amount of material in the system by a constant factor, is a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics or differential geometry sense (a physical quantity/tensor with a value for each point in space).} isomorphic to $\mathbb{R}$.
+Consider $\tau: \mathbb{R} \rightarrow T$ the mapping between a number and the transformation that increases or decreases the amount of material by that factor.
 This transformation exists: the system is infinitesimally decomposable and the amount can be changed continuously.
-Define on $T$ an addition $+: T \times T \rightarrow T$ and a multiplication $*: T \times T \rightarrow T$ such that $\tau(a) + \tau(b) = \tau(a+b)$ and $\tau(a) * \tau(b) = \tau(a*b)$, $a,b \in \mathbbm{R}$, so that the sum and product of the transformation is equal to the sum and product of their respective factors.
-$\tau$ is an isomorphism between $T$ and $\mathbbm{R}$ as fields.
-
-We claim $\mathcal{C}$ is a vector space over $\mathbbm{R}$.
+Define on $T$ an addition $+: T \times T \rightarrow T$ and a multiplication $*: T \times T \rightarrow T$ such that $\tau(a) + \tau(b) = \tau(a+b)$ and $\tau(a) * \tau(b) = \tau(a*b)$, $a,b \in \mathbb{R}$, so that the sum and product of the transformation is equal to the sum and product of their respective factors.
+$\tau$ is an isomorphism between $T$ and $\mathbb{R}$ as fields.
+
+We claim $\mathcal{C}$ is a vector space over $\mathbb{R}$.
 The abelian group $\mathcal{C}$ can be extended with the operations defined by $T$, as each element $\tau \in T$ is a map $\tau : \mathcal{C} \rightarrow \mathcal{C}$.
 The map has the following properties: $(\tau_1 + \tau_2) \, \mathcal{c} = \tau_1 \mathcal{c} + \tau_2 \mathcal{c} \; \forall \tau_1, \tau_2 \in T$ and $\mathcal{c} \in \mathcal{C}$, increasing the amount of material by the sum of two constant factors is the same as combining the separate increases, and $\tau \, (\mathcal{c}_1 + \mathcal{c}_2) = \tau \mathcal{c}_1 + \tau \mathcal{c}_2\; \forall \tau \in T$ and $\mathcal{c}_1, \mathcal{c}_2 \in \mathcal{C}$, increasing the amount of the total system is the same as the combination of the increased parts.
-$\mathcal{C}$ is a module over $T$, which is a field and isomorphic to $\mathbbm{R}$.
+$\mathcal{C}$ is a module over $T$, which is a field and isomorphic to $\mathbb{R}$.
 $\mathcal{C}$ is (isomorphic to) a real vector space.
 
-We claim $\mathcal{C}$ is isomorphic to a subspace of $C(\mathcal{S})$ as a vector space over $\mathbbm{R}$.
+We claim $\mathcal{C}$ is isomorphic to a subspace of $C(\mathcal{S})$ as a vector space over $\mathbb{R}$.
 $\forall \mathcal{c} \in \mathcal{C} \; \exists !
-\rho_{\mathcal{c}}:\mathcal{S} \rightarrow \mathbbm{R}$ returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$.
+\rho_{\mathcal{c}}:\mathcal{S} \rightarrow \mathbb{R}$ returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$.
 $\rho_{\mathcal{c}} \in C(\mathcal{S})$ by \ref{prop:continuity}.
@@
 This means that given a composite state $\mathcal{c}$ and a set $U \subseteq \mathcal{S}$ of particle states compatible with an outcome of a process, we must be able to tell the amount of material that we will find associated with that outcome.
-That is for each open set $U \subseteq \mathcal{S}$ there will be a functional $\Lambda_U : \mathcal{C} \rightarrow \mathbbm{R}$.
+That is for each open set $U \subseteq \mathcal{S}$ there will be a functional $\Lambda_U : \mathcal{C} \rightarrow \mathbb{R}$.
 
@@
 The state space $\mathcal{C}$ for a classical material is isomorphic to a subspace of the space of Lebesgue integrable functions.
-That is $\mathcal{C} \cong G \subseteq L^1(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbbm{R} \; | \; \int_{\mathcal{S}} |\rho| d\mu < \infty \}$.
+That is $\mathcal{C} \cong G \subseteq L^1(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbb{R} \; | \; \int_{\mathcal{S}} |\rho| d\mu < \infty \}$.
 \end{prop}
@@
 \begin{justification}
-	We claim there exists a positive linear functional $\Lambda_U : \mathcal{C} \rightarrow \mathbbm{R}$ for each $U \subseteq \mathcal{S}$ such that $\Lambda_U = \Lambda_{\mathrm{int}(U)}$ where $\mathrm{int}(U)$ is the interior of $U$.
-Let $\Lambda_{U} : \mathcal{C} \rightarrow \mathbbm{R}$ be the functional that returns the amount of material in $\mathcal{c}$ compatible with the outcome associated with the open set $U \subseteq \mathcal{S}$.
+	We claim there exists a positive linear functional $\Lambda_U : \mathcal{C} \rightarrow \mathbb{R}$ for each $U \subseteq \mathcal{S}$ such that $\Lambda_U = \Lambda_{\mathrm{int}(U)}$ where $\mathrm{int}(U)$ is the interior of $U$.
+Let $\Lambda_{U} : \mathcal{C} \rightarrow \mathbb{R}$ be the functional that returns the amount of material in $\mathcal{c}$ compatible with the outcome associated with the open set $U \subseteq \mathcal{S}$.
 $\Lambda_U$ is well-defined: $U$ is associated with a physically distinguishable outcome.
@@
 $\mathcal{C} \cong G \subseteq C(\mathcal{S})$ therefore $\Lambda_U(\mathcal{c}) \cong \Lambda_U(\rho_\mathcal{c})$.
-$\Lambda = \{\Lambda_U : C(\mathcal{S}) \rightarrow \mathbbm{R}\}_{U \subseteq \mathcal{S}}$ is a family of positive linear functionals such that $\forall U \subseteq \mathcal{S} \; \Lambda_U = \Lambda_{\mathrm{int}(U)}$ and $\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2} \; \forall U_1, U_2 \subseteq \mathcal{S}$.
+$\Lambda = \{\Lambda_U : C(\mathcal{S}) \rightarrow \mathbb{R}\}_{U \subseteq \mathcal{S}}$ is a family of positive linear functionals such that $\forall U \subseteq \mathcal{S} \; \Lambda_U = \Lambda_{\mathrm{int}(U)}$ and $\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2} \; \forall U_1, U_2 \subseteq \mathcal{S}$.
 By the extension \ref{extended_riesz_theorem} of the Riesz representation theorem for linear functionals
@@
 
-	We claim $\mathcal{C}$ is isomorphic to a subspace of $L^1(\mathcal{S}, \mu)$ as a vector space over $\mathbbm{R}$.
+	We claim $\mathcal{C}$ is isomorphic to a subspace of $L^1(\mathcal{S}, \mu)$ as a vector space over $\mathbb{R}$.
 $\forall \mathcal{c} \in \mathcal{C} \exists !
-\rho_{\mathcal{c}}:\mathcal{S} \rightarrow \mathbbm{R}$ returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$.
-As $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$, then $\rho_{\mathcal{c}} \in \mathcal{L}(S,\mu) \equiv \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\}$.
+\rho_{\mathcal{c}}:\mathcal{S} \rightarrow \mathbb{R}$ returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$.
+As $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$, then $\rho_{\mathcal{c}} \in \mathcal{L}(S,\mu) \equiv \{ \rho : S \rightarrow \mathbb{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\}$.
 $\mathcal{L}(S,\mu) = L^1(\mathcal{S}, \mu)$ by \ref{everywhere_integrable_is_lebesgue_integrable}.
@@
 The state space $\mathcal{C}$ for a classical material is isomorphic to the space of Lebesgue integrable differentiable functions.
-That is $\mathcal{C} \cong C^1(\mathcal{S}) \cap L^1(\mathcal{S}, \mu)$, where $C^1(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is differentiable$\}$.
+That is $\mathcal{C} \cong C^1(\mathcal{S}) \cap L^1(\mathcal{S}, \mu)$, where $C^1(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbb{R} \; | \; \rho$ is differentiable$\}$.
 \end{prop}
@@
 	We claim $\mathcal{S}$ is a differentiable manifold.
-Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c} \in \mathcal{C}$.
+Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbb{R}$ be the distribution associated with a composite state $\mathcal{c} \in \mathcal{C}$.
 Let $q^i$ and $\hat{q}^j$ be two sets of independent state variables that fully identify states within $U\subseteq \mathcal{S}$.
@@
 Let $\rho_{\mathcal{c}_0}$ be strictly monotonic in a region $U\subseteq \mathcal{S}$.
-Let $\delta \rho = \rho_{\mathcal{c}_0} - k \rho_{\mathcal{c}}$ where $k \in \mathbbm{R}$ and $k>0$.
+Let $\delta \rho = \rho_{\mathcal{c}_0} - k \rho_{\mathcal{c}}$ where $k \in \mathbb{R}$ and $k>0$.
 Let $k$ be sufficiently small so that $\delta \rho$ is strictly monotonic over $U$.
@@
 Let $\mathcal{Q}$ be the manifold charted by $q^i$.
-Locally $\mathcal{Q} \cong \prod \mathcal{Q}^i \cong \mathbbm{R}^n$.
+Locally $\mathcal{Q} \cong \prod \mathcal{Q}^i \cong \mathbb{R}^n$.
 Changing units of one state variable must not change the units of the other as they are independent.
@@
 
-It should be clear by now that the case of discrete topology is qualitatively different from the standard topology for $\mathbbm{R}^n$.
+It should be clear by now that the case of discrete topology is qualitatively different from the standard topology for $\mathbb{R}^n$.
 The notion that the continuous case is a limit of the discrete case, that it's ``like the discrete but with more points", leads in this case to erroneous intuition.
@@
 Can we have an outcome associated with only one element?
-The answer is no with the standard topology on $\mathbbm{R}^n$.
+The answer is no with the standard topology on $\mathbb{R}^n$.
 The consequence is that when we define the measure $\mu$ for the cardinality of continuous states, we assign a finite value to a finite region, and we assign zero measure to a single state.
 That is what gave us integration, densities and, ultimately, conjugate variables.
-If we were to use the discrete topology on $\mathbbm{R}^n$, if we were to assume we can identify single states as we do for countable states, then we would assign measure one to each state and infinity to a finite region.
+If we were to use the discrete topology on $\mathbb{R}^n$, if we were to assume we can identify single states as we do for countable states, then we would assign measure one to each state and infinity to a finite region.
 We would not have integration, just a simple sum.
@@
 Then distinguishing particles would mean distinguishing points.
-We are back to the idea of a discrete topology on $\mathbbm{R}^n$.
+We are back to the idea of a discrete topology on $\mathbb{R}^n$.
 In that case we would not have conjugate momentum, no $T^*\mathcal{Q}$, just the coordinates of the point in $\mathcal{Q}$.
@@
 \begin{prop}\label{prop:hamiltons_equations}
-	A continuous deterministic and reversible evolution for the particles of a classical material admits a Hamiltonian $H \in C^2(T^*\mathcal{Q}, \mathbbm{R})$ that allows us to write the laws of evolution as
+	A continuous deterministic and reversible evolution for the particles of a classical material admits a Hamiltonian $H \in C^2(T^*\mathcal{Q}, \mathbb{R})$ that allows us to write the laws of evolution as
 \begin{align*}
@@
 \begin{justification}
-We claim the vector field $S \in T\mathcal{S}$ for the infinitesimal displacement $S^a = d\xi^a/dt$ admits a potential $H \in C^2(T^*\mathcal{Q}, \mathbbm{R})$ such that $S^{a} \omega_{ab} = \partial_{b}H$.
+We claim the vector field $S \in T\mathcal{S}$ for the infinitesimal displacement $S^a = d\xi^a/dt$ admits a potential $H \in C^2(T^*\mathcal{Q}, \mathbb{R})$ such that $S^{a} \omega_{ab} = \partial_{b}H$.
 The state space $\mathcal{S}$ for the particles of a classical material is a symplectic manifold by \ref{prop:symplectic_manifold}.
 The map for infinitesimal time evolution $\mathcal{T}_{dt}$ is an infinitesimal self-symplectomorphism by \ref{prop:symplectomorphism}.
-By \ref{symplectomorphism_generator} the infinitesimal displacement $S$ admits a potential $H \in C^2(T^*\mathcal{Q}, \mathbbm{R})$ such that $S^{a} \omega_{ab} = \partial_{b}H$
+By \ref{symplectomorphism_generator} the infinitesimal displacement $S$ admits a potential $H \in C^2(T^*\mathcal{Q}, \mathbb{R})$ such that $S^{a} \omega_{ab} = \partial_{b}H$
 
@@
 This does not affect our definition of determinism and reversibility as we still can tell final state from initial state and vice-versa.
-But at this point we have no way to specify a time dependent evolution: the Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ as we derived it is just a function of the state.
+But at this point we have no way to specify a time dependent evolution: the Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbb{R}$ as we derived it is just a function of the state.
 
@@
 The equations of motion transform and we have $d_{t}\hat{t} \, d_{\hat{t}}q^i = \partial_{p_i} H$.
-$q^i: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ and $H: T^*\mathcal{Q} \rightarrow \mathbbm{R}$ are real functions of the state space and cannot be redefined to include $d_{t}\hat{t}$.
+$q^i: T^*\mathcal{Q} \rightarrow \mathbb{R}$ and $H: T^*\mathcal{Q} \rightarrow \mathbb{R}$ are real functions of the state space and cannot be redefined to include $d_{t}\hat{t}$.
 Physically the transformation is introducing fictitious forces that are not conservative, so they cannot be expressed by a Hamiltonian.
@@
 A region at equal time will be a hyper-surface $\mathcal{S}_{t=t_0} \subset \bar{\mathcal{S}}$ that includes all possible configurations at a particular time.
-Such hyper-surface will need to cut across all evolutions once and only once, therefore not all functions $t : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ are suitable time variables.
+Such hyper-surface will need to cut across all evolutions once and only once, therefore not all functions $t : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ are suitable time variables.
 A time evolution is a line  $\bar{\mathcal{s}} \subset \bar{\mathcal{S}}$ that passes once through each equal time surfaces.
@@
 	We claim $\bar{\mathcal{S}}$ is a $2n+1$ dimensional differentiable manifold.
-Locally $\bar{\mathcal{S}} \supseteq U \cong T^*\mathcal{Q}\times \mathbbm{R}$ as a state $\mathcal{s} \in U$ exists for each configuration and each time instant.
-Locally $\bar{\mathcal{S}} \supseteq U \cong\mathbbm{R}^{2n+1}$.
+Locally $\bar{\mathcal{S}} \supseteq U \cong T^*\mathcal{Q}\times \mathbb{R}$ as a state $\mathcal{s} \in U$ exists for each configuration and each time instant.
+Locally $\bar{\mathcal{S}} \supseteq U \cong\mathbb{R}^{2n+1}$.
 $\bar{\mathcal{S}}$ is a $2n+1$ dimensional manifold.
@@
 	We claim $\mathcal{S}_{t=t_0}$ is a hyper-surface of $\bar{\mathcal{S}}$.
-Let $t : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be a time variable.
+Let $t : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ be a time variable.
 $\mathcal{S}_{t=t_0} = \{ \mathcal{s} \in \bar{\mathcal{S}} \; | \; t(\mathcal{s}) = t_0 \}$ is a level set.
@@
 For the composite state we need to take a different approach.
-A distribution of material over states at equal time $\rho_\mathcal{c} : \mathcal{S}_{t=t_0} \rightarrow \mathbbm{R}$ will depend on the choice of time variable, therefore the complete state space will not be useful as it will not be time invariant.
+A distribution of material over states at equal time $\rho_\mathcal{c} : \mathcal{S}_{t=t_0} \rightarrow \mathbb{R}$ will depend on the choice of time variable, therefore the complete state space will not be useful as it will not be time invariant.
 What we are really interested in is the \emph{state evolution space} of a classical material, which we'll indicate by $\bar{\mathcal{C}}$.
-To each state evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$ will correspond a distribution $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ over all particle states at all times.
+To each state evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$ will correspond a distribution $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ over all particle states at all times.
 This distribution will be differentiable: it's differentiable over all equal time hyper-surfaces and time evolution is differentiable.
@@
 	The state evolution space $\bar{\mathcal{C}}$ for a classical material is isomorphic to the space of differentiable functions that are Lebesgue integrable on any equal time hyper-surface.
-That is $\bar{\mathcal{C}} \cong  C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$ where $L^1_t(\bar{\mathcal{S}}, \mu)=\{{\rho: \bar{\mathcal{S}} \rightarrow \mathbbm{R}} \; | \; {\int_{\mathcal{S}_{t=t_0}} |\rho| d\mu < \infty} \; \forall \, t, t_0 \}$.
+That is $\bar{\mathcal{C}} \cong  C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$ where $L^1_t(\bar{\mathcal{S}}, \mu)=\{{\rho: \bar{\mathcal{S}} \rightarrow \mathbb{R}} \; | \; {\int_{\mathcal{S}_{t=t_0}} |\rho| d\mu < \infty} \; \forall \, t, t_0 \}$.
 \end{prop}
@@
 	We claim $\bar{\mathcal{C}} \cong  G \subseteq C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$.
-Let $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be the distribution of material associated with an evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$.
-Let $t : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be a time variable.
+Let $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ be the distribution of material associated with an evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$.
+Let $t : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ be a time variable.
 	The restriction of $\rho_{\bar{\mathcal{c}}}$ over an arbitrary level set $\mathcal{S}_{t=t_0}$ is continuous and integrable by \ref{prop:differentiable_manifold}.
@@
 Let $\mathcal{M}$ be the manifold that defines the unit system including time.
-Locally $\mathcal{M} \cong \mathcal{Q} \times \mathbbm{R}$ as time and state variables are independent quantities.
+Locally $\mathcal{M} \cong \mathcal{Q} \times \mathbb{R}$ as time and state variables are independent quantities.
 As in \ref{prop:symplectic_manifold}, invariant densities are defined on the symplectic manifold $(T^*\mathcal{M}, \omega)$.
@@
 The topology of $\bar{\mathcal{S}}$ is the subspace topology.
-Let $\rho : T^*\mathcal{M} \rightarrow \mathbbm{R}$ be an invariant density.
+Let $\rho : T^*\mathcal{M} \rightarrow \mathbb{R}$ be an invariant density.
 $\rho$ is differentiable by \ref{prop:differentiable_manifold}.
@@
 \begin{prop}\label{prop:relativistic_hamiltons_equations}
-	A continuous time-dependent deterministic and reversible process for the particles of a classical material admits an invariant Hamiltonian $\mathcal{H}: T^*\mathcal{M} \rightarrow \mathbbm{R}$ that allows us to write the laws of evolution as
+	A continuous time-dependent deterministic and reversible process for the particles of a classical material admits an invariant Hamiltonian $\mathcal{H}: T^*\mathcal{M} \rightarrow \mathbb{R}$ that allows us to write the laws of evolution as
 	\begin{align*}
@@
 This means there are two classes of states: the ones where the parametrization $s$ is aligned with time $t$, which we call standard states, and the ones where the parametrization $s$ is anti-aligned with time $t$, which we call anti-states.
-Let us call $\lambda : T^*\mathcal{M} \rightarrow \mathbbm{R}$ the function $\lambda (\xi^a) \mapsto d_s t |_{\xi^a}$ that returns the change of $t$ along $s$.
+Let us call $\lambda : T^*\mathcal{M} \rightarrow \mathbb{R}$ the function $\lambda (\xi^a) \mapsto d_s t |_{\xi^a}$ that returns the change of $t$ along $s$.
 $\lambda > 0$ for all standard states while $\lambda < 0$ for all anti-states.
@@
 \begin{prop}\label{prop:form_of_invariant_hamiltonian}
-	The form of the invariant Hamiltonian is $\mathcal{H} = \lambda(t,q^i,E,p_i) \, (H(t,q^i,p_i) - E)$ where $H(t,q^i,p_i)$ is the standard Hamiltonian at each instant $t$, the conjugate time variable $E = H(t,q^i,p_i)$ represents the value of the Hamiltonian, and  $\lambda : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ where $\lambda(\xi^a) = d_s t$.
+	The form of the invariant Hamiltonian is $\mathcal{H} = \lambda(t,q^i,E,p_i) \, (H(t,q^i,p_i) - E)$ where $H(t,q^i,p_i)$ is the standard Hamiltonian at each instant $t$, the conjugate time variable $E = H(t,q^i,p_i)$ represents the value of the Hamiltonian, and  $\lambda : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ where $\lambda(\xi^a) = d_s t$.
 $\bar{\mathcal{S}}$ is the level set for $\mathcal{H} = 0$.
@@
 Let $\mathcal{H}$ be the invariant Hamiltonian and $H$ be the Hamiltonian for a particular choice of time and state variables.
-We have $d_s q^i = \partial_{p_i} \mathcal{H} = d_s t \, d_t q^i = - \lambda \, \partial_{p_i} H$ where $\lambda : T^*\mathcal{M} \rightarrow \mathbbm{R}$ such that $\lambda \equiv d_s t$.
+We have $d_s q^i = \partial_{p_i} \mathcal{H} = d_s t \, d_t q^i = - \lambda \, \partial_{p_i} H$ where $\lambda : T^*\mathcal{M} \rightarrow \mathbb{R}$ such that $\lambda \equiv d_s t$.
 $d_s p_i = - \partial_{q^i} \mathcal{H} = d_s t \, d_t p_i = \lambda \, \partial_{q^i} H$.
@@
 	We claim $\mathcal{H} \rho_{\bar{\mathcal{c}}} = 0 \; \forall \bar{\mathcal{c}} \in \bar{\mathcal{C}}$.
-Let $\rho_{\bar{\mathcal{c}}} : T^*\mathcal{M} \rightarrow \mathbbm{R}$ be the distribution associated with an evolution $\bar{\mathcal{c}}$.
+Let $\rho_{\bar{\mathcal{c}}} : T^*\mathcal{M} \rightarrow \mathbb{R}$ be the distribution associated with an evolution $\bar{\mathcal{c}}$.
 $supp(\rho_{\bar{\mathcal{c}}}) \subseteq \bar{\mathcal{S}}$ as the distribution is defined on the complete state space of the particles.
@@
 $d_s \rho_{\bar{\mathcal{c}}}= \partial_a \rho_{\bar{\mathcal{c}}} d_s \xi^a = \partial_a \rho_{\bar{\mathcal{c}}} \omega^{ab} \partial_b \mathcal{H} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \}$.
-Let $\gamma : \mathbbm{R} \rightarrow T^*\mathcal{M}$ be a parametrization for a particle evolution $\bar{\mathcal{s}}$.
+Let $\gamma : \mathbb{R} \rightarrow T^*\mathcal{M}$ be a parametrization for a particle evolution $\bar{\mathcal{s}}$.
 Then $\rho_{\bar{\mathcal{c}}}(\lambda(s)) = \rho_{\bar{\mathcal{c}}}(\lambda(s + \Delta s))$ as the material does not change over the particle evolution.
@@
 In light of this, it is probably better to consider relativistic mechanics as a feature of the geometry of the complete state space more than the geometry of space-time.
-Also note that this feature arises from the use of the standard topology on $\mathbbm{R}^n$ and the need for invariant densities.
+Also note that this feature arises from the use of the standard topology on $\mathbb{R}^n$ and the need for invariant densities.
 It would not have arisen with the use of a discrete topology.
@@
 Let $\bar{\mathcal{c}}$ be the evolution of a minimal classical material.
-Let $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be the distribution associated with $\bar{\mathcal{c}}$.
+Let $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ be the distribution associated with $\bar{\mathcal{c}}$.
 $\rho_{\bar{\mathcal{c}}}(\mathcal{s}(x^\alpha, u^\alpha))$ as the material is minimal.
@@
 Let $(dx^\alpha, du^\alpha)$ be an infinitesimal range of possibilities for the kinematic variables of the particles of a minimal classical material.
-There exists a map $g : T\mathcal{M} \times T\mathcal{M} \rightarrow \mathbbm{R}$ such that $g(dx^\alpha, du^\alpha) \propto \omega(dq^\alpha, dp_\alpha)$.
+There exists a map $g : T\mathcal{M} \times T\mathcal{M} \rightarrow \mathbb{R}$ such that $g(dx^\alpha, du^\alpha) \propto \omega(dq^\alpha, dp_\alpha)$.
 The map $g$ is linear, since a linear combination of infinitesimal ranges leads to the linear combination of possibilities.
@@
 How can the whole universe be reduced to  elements that have no idea how many spatial dimensions are there?
-Mathematically a set of points by itself does not define the dimension of a space: $\mathbbm{R}^n$ is the same order of infinity for any $n$.
+Mathematically a set of points by itself does not define the dimension of a space: $\mathbb{R}^n$ is the same order of infinity for any $n$.
 One needs a topology to differentiate, which is an additional structure.
@@
 \begin{prop}\label{prop:complex_vector_space}
-	The state space $\Psi$ for an irreducible material is a vector space over $\mathbbm{C}$.
+	The state space $\Psi$ for an irreducible material is a vector space over $\mathbb{C}$.
 \end{prop}
@@
 
-	We claim the set of physically distinguishable transformations $T$ that relabel the fragment configurations is isomorphic to $\mathbbm{C}$.
-Consider an amount $a_m \in \mathbbm{R}$ of irreducible material characterized by one d.o.f.
-The configuration of a fragment of the material is characterized by two random variables $A, B: \Omega \rightarrow \mathbbm{R}$ where $\Omega$ is the sample space for the configuration of the fragments.
+	We claim the set of physically distinguishable transformations $T$ that relabel the fragment configurations is isomorphic to $\mathbb{C}$.
+Consider an amount $a_m \in \mathbb{R}$ of irreducible material characterized by one d.o.f.
+The configuration of a fragment of the material is characterized by two random variables $A, B: \Omega \rightarrow \mathbb{R}$ where $\Omega$ is the sample space for the configuration of the fragments.
 As the material is homogeneous, $A$ and $B$ can be chosen such that the distribution $\rho$ of material over such variables is uniform over a range of $\sqrt{a_m}$ for both variables.
@@
 Consider a second amount $\hat{a}_m$ of irreducible material where we changed the labeling of the fragment configurations.
-Its fragment configurations can also be characterized by a uniform distribution over two variables $C, D: \Omega \rightarrow \mathbbm{R}$ over a range of $\sqrt{\hat{a}_m}$.
+Its fragment configurations can also be characterized by a uniform distribution over two variables $C, D: \Omega \rightarrow \mathbb{R}$ over a range of $\sqrt{\hat{a}_m}$.
 As the two amounts of material differ just by the labeling of fragment configuration, $(C,D)$ is a function of $(A,B)$.
@@
 As the change in amount of material, the only physically distinguishable quantity associated with the fragment configurations, is fully specified by $a$ and $b$, no other transformation is physically relevant for the state.
-Let $\tau: \mathbbm{C} \rightarrow T$ represent such a transformation parametrized by $c=a+\imath b$.
+Let $\tau: \mathbb{C} \rightarrow T$ represent such a transformation parametrized by $c=a+\imath b$.
 $|c|^2 = |J| = (a^2 + b^2) = \hat{a}_m / a_m$ is the ratio of the two amounts of materials and is physically distinguishable.
-Define on $T$ an addition $+: T \times T \rightarrow T$ and a multiplication $*: T \times T \rightarrow T$ such that $\tau(z_1) + \tau(z_2) = \tau(z_1+z_2)$ and $\tau(z_1) * \tau(z_2) = \tau(z_1*z_2)$, $z_1,z_2 \in \mathbbm{C}$, the sum and product of the transformation is equal to the sum and product of their respective factors.
+Define on $T$ an addition $+: T \times T \rightarrow T$ and a multiplication $*: T \times T \rightarrow T$ such that $\tau(z_1) + \tau(z_2) = \tau(z_1+z_2)$ and $\tau(z_1) * \tau(z_2) = \tau(z_1*z_2)$, $z_1,z_2 \in \mathbb{C}$, the sum and product of the transformation is equal to the sum and product of their respective factors.
 The phase $arg(c)$ under addition is physically distinguishable as it can affect the modulus of the result.
-$\tau$ is an isomorphism between $T$ and $\mathbbm{C}$ as fields.
+$\tau$ is an isomorphism between $T$ and $\mathbb{C}$ as fields.
 Assume now that the material is characterized by multiple independent d.o.f.
 The transformation on each d.o.f.
-is defined independently by $c_i \in \mathbbm{C}$.
+is defined independently by $c_i \in \mathbb{C}$.
 As $\hat{a}_m / a_m$ is the same under all d.o.f.
@@
 The relationship has to always hold under addition as well, therefore $arg(c_i) = arg(c_j)$.
-$T\cong \mathbbm{C}$ also for multiple d.o.f.
-
-	We claim $\Psi$ is a vector space over $\mathbbm{C}$.
+$T\cong \mathbb{C}$ also for multiple d.o.f.
+
+	We claim $\Psi$ is a vector space over $\mathbb{C}$.
 The abelian group $\Psi$ can be extended with the operations defined by $T$ as was done in \ref{prop:real_vector_space} for the infinitesimally reducible case.
@@
 Naturally, we cannot give a complete description for the configuration of a fragment, therefore we cannot give a joint distribution for all fragment variables at the same time.\footnote{This recovers the notion of complementarity introduced by Bohr.} We can, though, find a maximal set of fragment variables: one that can be used to provide the best description possible.
-While there isn't a unique maximal set, we can show that one such set is provided by the fragment variables $q^i : \mathcal{Q} \rightarrow \mathbbm{R}$ that define the units required to describe the system.
+While there isn't a unique maximal set, we can show that one such set is provided by the fragment variables $q^i : \mathcal{Q} \rightarrow \mathbb{R}$ that define the units required to describe the system.
 
@@
 This means that a state $\psi$ must tell us for each point in $\mathcal{Q}$ the amount and correlation angle of the material.
-That is: the state will be identified by $\psi : \mathcal{Q} \rightarrow \mathbbm{C}$, the wave function over $\mathcal{Q}$.
+That is: the state will be identified by $\psi : \mathcal{Q} \rightarrow \mathbb{C}$, the wave function over $\mathcal{Q}$.
 
@@
 	The state space $\Psi$ for an irreducible material is isomorphic to a subspace of the space of Lebesgue square integrable continuous functions.
-That is $\Psi \cong \Psi(\mathcal{Q}, dq^i) \subseteq C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu, \mathbbm{C})$ under the isomorphism $\Upsilon_q : \Psi \leftrightarrow \Psi(\mathcal{Q}, dq^i)$ where $\Psi(\mathcal{Q}, dq^i)$ is the space of wave functions over $q^i$.
+That is $\Psi \cong \Psi(\mathcal{Q}, dq^i) \subseteq C(\mathcal{Q}, \mathbb{C}) \cap L^2(\mathcal{Q}, \mu, \mathbb{C})$ under the isomorphism $\Upsilon_q : \Psi \leftrightarrow \Psi(\mathcal{Q}, dq^i)$ where $\Psi(\mathcal{Q}, dq^i)$ is the space of wave functions over $q^i$.
 The state space for a quantum particle is isomorphic to the projective space $\mathsf{P}(\Psi)$.
@@
 
-	We claim $\Psi \cong \Psi(\mathcal{Q}, dp^i) \subseteq C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu, \mathbbm{C})$ as a vector space.
-The maximal physical description of a fragment configuration is given by a point in $\mathcal{Q}$ parametrized by $q^i$ and by a transformation $\tau \in T$ parametrized by $c \in \mathbbm{C}$.
-For each $\psi \in \Psi$ there must exist a wave function $\psi(q^i) : \mathcal{Q} \rightarrow \mathbbm{C}$ that provides the maximal physical description for the state.
+	We claim $\Psi \cong \Psi(\mathcal{Q}, dp^i) \subseteq C(\mathcal{Q}, \mathbb{C}) \cap L^2(\mathcal{Q}, \mu, \mathbb{C})$ as a vector space.
+The maximal physical description of a fragment configuration is given by a point in $\mathcal{Q}$ parametrized by $q^i$ and by a transformation $\tau \in T$ parametrized by $c \in \mathbb{C}$.
+For each $\psi \in \Psi$ there must exist a wave function $\psi(q^i) : \mathcal{Q} \rightarrow \mathbb{C}$ that provides the maximal physical description for the state.
 $\rho_\psi(q^i) = |\psi(q^i)|^2$ as the square modulus of $\psi(q^i)$ is the increase/decrease from the unit amount of material.
-$\rho_\psi(q^i) \in C(\mathcal{Q}, \mathbbm{R})$ by \ref{prop:continuity}.
-$\rho_\psi(q^i) \in C^1(\mathcal{Q}, \mathbbm{R})$ by the same arguments in the justification for \ref{prop:differentiable_manifold}.
-Therefore $\psi(q^i) \in C(\mathcal{Q}, \mathbbm{C})$.
-$\rho_\psi(q^i) \in L^1(\mathcal{Q}, \mu, \mathbbm{R})$ by the same arguments in the justification for \ref{prop:integration}.
-Therefore $\psi(q^i) \in L^2(\mathcal{Q}, \mu, \mathbbm{C})$.
-Let $\Upsilon_{q^i} : \Psi \rightarrow C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu, \mathbbm{C})$ be the function that given a state returns its wave function.
+$\rho_\psi(q^i) \in C(\mathcal{Q}, \mathbb{R})$ by \ref{prop:continuity}.
+$\rho_\psi(q^i) \in C^1(\mathcal{Q}, \mathbb{R})$ by the same arguments in the justification for \ref{prop:differentiable_manifold}.
+Therefore $\psi(q^i) \in C(\mathcal{Q}, \mathbb{C})$.
+$\rho_\psi(q^i) \in L^1(\mathcal{Q}, \mu, \mathbb{R})$ by the same arguments in the justification for \ref{prop:integration}.
+Therefore $\psi(q^i) \in L^2(\mathcal{Q}, \mu, \mathbb{C})$.
+Let $\Upsilon_{q^i} : \Psi \rightarrow C(\mathcal{Q}, \mathbb{C}) \cap L^2(\mathcal{Q}, \mu, \mathbb{C})$ be the function that given a state returns its wave function.
 As the wave function captures the maximal description of the system, two different states must represent different distributions.
@@
 $\Upsilon_{q^i}$ is a homomorphism between vector spaces.
-$\Psi$ is isomorphic to $\Psi(\mathcal{Q}, dq^i) \subseteq C(\mathcal{Q}, \mathbbm{C}) \cap L^2(\mathcal{Q}, \mu, \mathbbm{C})$.
+$\Psi$ is isomorphic to $\Psi(\mathcal{Q}, dq^i) \subseteq C(\mathcal{Q}, \mathbb{C}) \cap L^2(\mathcal{Q}, \mu, \mathbb{C})$.
 
@@
 Let $\psi \in \Psi$ be the state of a quantum particle.
-$\tau(e^{\imath \theta}) \psi$ with $\theta \in \mathbbm{R}$ is not physically distinguishable from $\psi$ as phases are only physically distinguishable through addition.
-$\tau(\sqrt{\rho}) \psi$ with $\sqrt{\rho} \in \mathbbm{R}^+ \backslash \{1\}$ cannot represent a quantum particle as it has a different amount of material.
-There exists only one distinguishable quantum particle state in the equivalence class $\psi \sim \tau(c) \psi$ with $c \in \mathbbm{C}$.
+$\tau(e^{\imath \theta}) \psi$ with $\theta \in \mathbb{R}$ is not physically distinguishable from $\psi$ as phases are only physically distinguishable through addition.
+$\tau(\sqrt{\rho}) \psi$ with $\sqrt{\rho} \in \mathbb{R}^+ \backslash \{1\}$ cannot represent a quantum particle as it has a different amount of material.
+There exists only one distinguishable quantum particle state in the equivalence class $\psi \sim \tau(c) \psi$ with $c \in \mathbb{C}$.
 The state with no amount of material does not correspond to any particle state.
@@
 That is: $\mathsf{P}_{\psi_2} (\psi_1) = \psi_1 \Leftrightarrow \psi_1 = \alpha \psi_2$.
-Because of this, we can define a non degenerate complex product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{\langle \psi_2|\psi_1 \rangle}{|\psi_2|^2} \psi_2$.
+Because of this, we can define a non degenerate complex product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbb{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{\langle \psi_2|\psi_1 \rangle}{|\psi_2|^2} \psi_2$.
 
@@
 As this is a part of $\psi_1$ prepared as $\psi_1$, its state is just $\psi_1$ rescaled, with no change in correlation.
-That is: $\mathsf{P}_{\psi_1} (\mathsf{P}_{\psi_2} (\psi_1)) = a \psi_1$ with $a \in \mathbbm{R}^+$.
+That is: $\mathsf{P}_{\psi_1} (\mathsf{P}_{\psi_2} (\psi_1)) = a \psi_1$ with $a \in \mathbb{R}^+$.
 This tells us that $\langle \psi_2|\psi_1 \rangle = \langle \psi_1|\psi_2 \rangle^\dagger$.
@@
 \begin{prop}\label{prop:inner_product}
-	The state space $\Psi$ for an irreducible material is an inner product with $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$.
+	The state space $\Psi$ for an irreducible material is an inner product with $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbb{C}$ such that $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$.
 \end{prop}
 \begin{justification}
-	We claim there exists a map $| \cdot | ^2 : \Psi \rightarrow \mathbbm{R}$ such that $|\psi|^2 \ge 0 \; \forall \psi \in \Psi$ and $|\psi|^2 = 0 \Leftrightarrow \psi = 0$.
-Let $| \cdot | ^2 : \Psi \rightarrow \mathbbm{R}$ be the map that returns the amount of material for a given state.
+	We claim there exists a map $| \cdot | ^2 : \Psi \rightarrow \mathbb{R}$ such that $|\psi|^2 \ge 0 \; \forall \psi \in \Psi$ and $|\psi|^2 = 0 \Leftrightarrow \psi = 0$.
+Let $| \cdot | ^2 : \Psi \rightarrow \mathbb{R}$ be the map that returns the amount of material for a given state.
 Let $\psi \in \Psi$ be a state.
@@
 
-	We claim $\mathsf{P}_{\psi} (\phi) = \phi \Leftrightarrow \phi = c \psi$ for some $c \in \mathbbm{C}$.
+	We claim $\mathsf{P}_{\psi} (\phi) = \phi \Leftrightarrow \phi = c \psi$ for some $c \in \mathbb{C}$.
 Let $\psi, \phi \in \Psi$ such that $\mathsf{P}_{\psi} (\phi) = \phi$.
@@
 We can obtain $\phi$ from $\psi$ by a relabeling of the fragment configurations.
-$\phi = c \psi$ for some $c \in \mathbbm{C}$.
-Let $\psi, \phi \in \Psi$ such that $\phi = c \psi$ for some $c \in \mathbbm{C}$.
+$\phi = c \psi$ for some $c \in \mathbb{C}$.
+Let $\psi, \phi \in \Psi$ such that $\phi = c \psi$ for some $c \in \mathbb{C}$.
 Then all of $\phi$ is prepared as $\psi$.
@@
 
-	We claim $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi)) = r \psi$ for some $r \in \mathbbm{R}^+$.
+	We claim $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi)) = r \psi$ for some $r \in \mathbb{R}^+$.
 Let $\psi, \phi \in \Psi$.
@@
 The only thing that can change is the amount of material.
-Therefore $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi)) = r \psi$ for some $r \in \mathbbm{R}^+$.
-
-	We claim that the state space $\Psi$ for an irreducible material has an inner product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$.
+Therefore $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi)) = r \psi$ for some $r \in \mathbb{R}^+$.
+
+	We claim that the state space $\Psi$ for an irreducible material has an inner product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbb{C}$.
 $| \cdot | ^2$ and $\mathsf{P}_{\phi}$ satisfy the requirements in \ref{thrm:inner_product}.
-$\Psi$ is an inner product space with $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbbm{C}$ such that $\langle \psi |\phi \rangle \psi = | \psi | ^2\mathsf{P}_{\psi}(\phi)$.
+$\Psi$ is an inner product space with $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbb{C}$ such that $\langle \psi |\phi \rangle \psi = | \psi | ^2\mathsf{P}_{\psi}(\phi)$.
 
@@
 We proceed in the following way.
-Suppose we have a fragment variable $f : \mathcal{Q} \rightarrow \mathbbm{R}$.
+Suppose we have a fragment variable $f : \mathcal{Q} \rightarrow \mathbb{R}$.
 Consider the product $f (q) \psi (q)$: it returns the wave function weighted by the value of $f$.
@@
 Suppose there is another maximal set of fragment variables.
-Then those variables will form a topological space $\hat{\mathcal{Q}}$ such that each $\psi$ can be expressed as a wave function $\psi(\hat{q}) :\hat{\mathcal{Q}} \rightarrow \mathbbm{C}$.
+Then those variables will form a topological space $\hat{\mathcal{Q}}$ such that each $\psi$ can be expressed as a wave function $\psi(\hat{q}) :\hat{\mathcal{Q}} \rightarrow \mathbb{C}$.
 
@@
 As converting the space again must give us the original wave function, $\Upsilon^{\hat{q}}_{q} \circ \Upsilon^q_{\hat{q}}= I$ is unitary.
-We can now define a linear operator $F_{[\hat{q}]} \equiv f(\hat{q})$ corresponding to fragment variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
+We can now define a linear operator $F_{[\hat{q}]} \equiv f(\hat{q})$ corresponding to fragment variable $f : \hat{\mathcal{Q}} \rightarrow \mathbb{R}$.
 Using $\Upsilon^{q}_{\hat{q}}$ we can express $F_{[q]} = \Upsilon^{\hat{q}}_{q} \circ F_{[\hat{q}]} \circ  \Upsilon^{q}_{\hat{q}}$ as an operator over complex functions of $\mathcal{Q}$.
@@
 \begin{prop}\label{prop:self_adjoint_operators}
-	For each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ and a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ that takes a state $\psi \in \Psi$ and returns a wave function $\psi(\hat{q}^i)$ from $\Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ the space of wave functions over $\hat{\mathcal{Q}}$.
-There also exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbbm{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i))d^nq $ is the expectation of $f$ multiplied by the amount of material.
+	For each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbb{R}$ and a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ that takes a state $\psi \in \Psi$ and returns a wave function $\psi(\hat{q}^i)$ from $\Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ the space of wave functions over $\hat{\mathcal{Q}}$.
+There also exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbb{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i))d^nq $ is the expectation of $f$ multiplied by the amount of material.
 \end{prop}
 \begin{justification}
-	We claim for each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that  $\hat{q}^i : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}^n$ provides a maximal description and $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$.
+	We claim for each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that  $\hat{q}^i : \hat{\mathcal{Q}} \rightarrow \mathbb{R}^n$ provides a maximal description and $f : \hat{\mathcal{Q}} \rightarrow \mathbb{R}$.
 Let $f$ be a fragment variable.
@@
 $\hat{q}^i$ provides a maximal set of fragment variables.
-There must exist a wave function $\psi(\hat{q}^i) : \mathbbm{R}^n \rightarrow \mathbbm{C}$ associated with $\psi$ such that it gives the maximal description for the state.
+There must exist a wave function $\psi(\hat{q}^i) : \mathbb{R}^n \rightarrow \mathbb{C}$ associated with $\psi$ such that it gives the maximal description for the state.
 Let $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \rightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ be the map that associates a state with its wave function over $\hat{\mathcal{Q}}$.
@@
 
-	We claim that for each fragment variable $f : \hat{\mathcal{Q}} \rightarrow \mathbbm{R}$ there exists a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ such that $\int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]}( \psi(\hat{q}^i)) d^n \hat{q}$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
+	We claim that for each fragment variable $f : \hat{\mathcal{Q}} \rightarrow \mathbb{R}$ there exists a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbb{C})$ such that $\int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]}( \psi(\hat{q}^i)) d^n \hat{q}$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
 The expectation value of $f$ multiplied by the amount of material is given by $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q}$ where $\hat{q}^i$ provide a maximal set of fragment variables such that $f=f(\hat{q}^i)$.
 We have $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q} = \int \psi^\dagger(\hat{q}^i) f(\hat{q}^i) \psi(\hat{q}^i) d^{n}\hat{q}$.
-Let $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ such that $F_{[\hat{q}^i]} (\psi(\hat{q}^i)) = f(\hat{q}^i) \psi(\hat{q}^i)$.
+Let $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbb{C})$ such that $F_{[\hat{q}^i]} (\psi(\hat{q}^i)) = f(\hat{q}^i) \psi(\hat{q}^i)$.
 $F_{[\hat{q}^i]}$ is a linear operator.
@@
 
-	We claim that for any fragment variable $f$ there exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbbm{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i)) d^nq$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
+	We claim that for any fragment variable $f$ there exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbb{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i)) d^nq$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
 Let $f$ be a fragment variable.
-There exists a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbbm{C})$ where $\hat{\mathcal{Q}}$ provides a maximal set of fragment variables.
+There exists a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbb{C})$ where $\hat{\mathcal{Q}}$ provides a maximal set of fragment variables.
 As both $\hat{\mathcal{Q}}$ and $\mathcal{Q}$ provide a maximal set, we can define $\Upsilon^{q^i}_{\hat{q}^i} \equiv \Upsilon_{\hat{q}^i} \circ \Upsilon_{q^i}^{-1}$ which transforms a wave function over $\mathcal{Q}$ into a wave function over $\hat{\mathcal{Q}}$.
@@
 
-If we assume $\mathcal{Q}$ to be $\mathbbm{R}^n$, the functional $\Upsilon^{q^i}_{p_i} : \Psi(\mathcal{Q}, dq^i) \leftrightarrow \Psi(\mathcal{P}, dp_i)$ that converts a wave function in position to a wave function in conjugate momentum is the Fourier transform.
+If we assume $\mathcal{Q}$ to be $\mathbb{R}^n$, the functional $\Upsilon^{q^i}_{p_i} : \Psi(\mathcal{Q}, dq^i) \leftrightarrow \Psi(\mathcal{P}, dp_i)$ that converts a wave function in position to a wave function in conjugate momentum is the Fourier transform.
 As the fragment variables for any power of $q^i$ and $p_i$ must exist, the wave functions in both expressions are infinitely differentiable: the space of the wave functions $\Psi(\mathcal{Q}, dq^i) = S(\mathcal{Q}, q^i)$ is the Schwartz space of rapidly decreasing smooth functions.
@@
 \begin{prop}\label{prop:momentum_operator}
-	Let $\mathcal{Q}=\mathbbm{R}^n$.
-The operator $P_{i [q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbbm{C})$ associated with the conjugate quantity of $q^i$ is $P_{i [q^i]} = - \imath \hbar \partial_{i}$.
+	Let $\mathcal{Q}=\mathbb{R}^n$.
+The operator $P_{i [q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbb{C})$ associated with the conjugate quantity of $q^i$ is $P_{i [q^i]} = - \imath \hbar \partial_{i}$.
 The isomorphism $\Upsilon^{q^i}_{p_i} : \Psi(\mathcal{Q}, dq^i) \leftrightarrow \Psi(\mathcal{P}, dp_i)$ that changes the variable of the wave function from $q^i$ to $p_i$ is $\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^n} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
@@
 In particular, it needs to be covariant when changing only one variable: $\hat{q}^1 = \hat{q}^1(q^1)$, $\hat{q}^j = q^j$ where $j=2...n$.
-By \ref{thrm:covariant_operator} $K_{i [q^i]}=c\partial_i$ for some $c \in \mathbbm{C}$.
-$K_{i [q^i]}$ is self-adjoint: $K_{i  [q^i]}=\imath r\partial_i$ for some $r \in \mathbbm{R}$ by \ref{thrm:antihermitian_derivative}.
+By \ref{thrm:covariant_operator} $K_{i [q^i]}=c\partial_i$ for some $c \in \mathbb{C}$.
+$K_{i [q^i]}$ is self-adjoint: $K_{i  [q^i]}=\imath r\partial_i$ for some $r \in \mathbb{R}$ by \ref{thrm:antihermitian_derivative}.
 $k_i$ is expressed in units of inverse $q^i$ therefore $|r|=1$.
@@
 As $\psi(q^i)$ is the Fourier transform of an infinitely smooth function, it decreases faster than any inverse power of $q^i$.
-That is: $\forall m \in \mathbbm{Z}^+ \exists c \in \mathbbm{C}$ such that  $|\psi(q^i)|<\frac{|c|}{|q^i|^m} \; \forall |q^i|> 1$.
+That is: $\forall m \in \mathbb{Z}^+ \exists c \in \mathbb{C}$ such that  $|\psi(q^i)|<\frac{|c|}{|q^i|^m} \; \forall |q^i|> 1$.
 $\Psi(\mathcal{Q}, dq^i) \subseteq S(\mathcal{Q}, q^i)$.
@@
 Let $\psi(q^i) \in \Psi(\mathcal{Q}, dq^i)$ and $\phi(q^i) \in S(\mathcal{Q}, q^i)$ such that $\psi(q^i) \neq \phi(q^i)$.
-There exists a choice of $a, b, j, k \in \mathbbm{Z}^+$, $j,k \leq n$ such that $\lVert \psi(q^i) - \phi(q^i) \rVert_{j,a,k,b} \neq 0$  where $\lVert \psi(q^i) \rVert_{j,a,k,b} = sup|(q^j)^a (\partial_k) ^b \psi(q^i)|$ as $S(\mathcal{Q}, q^i)$ is a Frechet space induced by that family of seminorms.
+There exists a choice of $a, b, j, k \in \mathbb{Z}^+$, $j,k \leq n$ such that $\lVert \psi(q^i) - \phi(q^i) \rVert_{j,a,k,b} \neq 0$  where $\lVert \psi(q^i) \rVert_{j,a,k,b} = sup|(q^j)^a (\partial_k) ^b \psi(q^i)|$ as $S(\mathcal{Q}, q^i)$ is a Frechet space induced by that family of seminorms.
 The operator $(Q^j_{[q^i]})^a(P_{k [q^i]})^b$ applied to $\psi(q^i)$ and $\phi(q^i)$ will provide different distributions for the associated fragment variables.
@@
 For instance, while any differentiable function of $T^*\mathcal{Q}$ can be a state variable for a classical particle, not all self-adjoint linear operators can be associated with a fragment variable.
-Consider $f: \mathbbm{R} \rightarrow \mathbbm{R}$ for which $f(r) = -r$ if $r$ is an integer and $f(r) = r$ otherwise.
+Consider $f: \mathbb{R} \rightarrow \mathbb{R}$ for which $f(r) = -r$ if $r$ is an integer and $f(r) = r$ otherwise.
 Clearly $f(q)$ does not preserve the topology and cannot be used as a fragment variable.
@@
 For example, if $\mathcal{Q}$ is an n-dimensional torus, there is no infinity and $\mathcal{P}$ has a discrete topology.
-That is why, in the end, we restricted ourselves to $\mathbbm{R}^n$.
+That is why, in the end, we restricted ourselves to $\mathbb{R}^n$.
 Moreover, we characterized the space of functions not because we required them to fall off at infinity, but because we required the ability to express polynomials of $q^i$ and $p_i$.
@@
 Let $\psi_1, \psi_2 \in \Psi$ be two states.
-$\mathcal{T}_{\Delta t}(a\psi_1+b\psi_2) = a \mathcal{T}_{\Delta t} \psi_1 + b \mathcal{T}_{\Delta t} \psi_2$ with $a,b \in \mathbbm{C}$ as the composition of the evolution is the evolution of the composition: $\mathcal{T}_{\Delta t}$ is a linear operator.
+$\mathcal{T}_{\Delta t}(a\psi_1+b\psi_2) = a \mathcal{T}_{\Delta t} \psi_1 + b \mathcal{T}_{\Delta t} \psi_2$ with $a,b \in \mathbb{C}$ as the composition of the evolution is the evolution of the composition: $\mathcal{T}_{\Delta t}$ is a linear operator.
 $\mathcal{T}_{\Delta t} \mathsf{P}_{\psi_1} (\psi_2) = \mathsf{P}_{\mathcal{T}_{\Delta t}\psi_1} (\mathcal{T}_{\Delta t}\psi_2)$ as the evolved part of $\psi_2$ that was prepared like $\psi_1$ is the part of the evolved $\psi_2$ that ends prepared like the evolved $\psi_1$.
@@
 
-Also note that $H$ operates on $\Psi$ which, in the case of $\mathcal{Q}=\mathbbm{R}^n$, corresponds to the space $S(\mathcal{Q}, q^i)$ of Schwartz functions.
+Also note that $H$ operates on $\Psi$ which, in the case of $\mathcal{Q}=\mathbb{R}^n$, corresponds to the space $S(\mathcal{Q}, q^i)$ of Schwartz functions.
 $H$ therefore preserves infinite smoothness and integrability.
@@
 This gives us the same $\mathcal{M}$ as before.
-A quantum state is a complex valued distribution $\psi : \mathcal{M}_{t=t_0} \rightarrow \mathbbm{C}$ over a hypersurface at constant time.
-An evolution $\widetilde{\psi} : \mathcal{M} \rightarrow \mathbbm{C}$ is a complex valued function defined on the whole $\mathcal{M}$.
+A quantum state is a complex valued distribution $\psi : \mathcal{M}_{t=t_0} \rightarrow \mathbb{C}$ over a hypersurface at constant time.
+An evolution $\widetilde{\psi} : \mathcal{M} \rightarrow \mathbb{C}$ is a complex valued function defined on the whole $\mathcal{M}$.
 The operator for the quantity conjugate to time is $E=\imath\hbar\partial_t$.
@@
 Suppose we know the outcomes of our measurement are parametrized by a fragment variable $\hat{q}$.
-We will have a linear functional $\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0} : \Psi \rightarrow \mathbbm{C}$ such that $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2$ returns the amount of material prepared with the particular value $\hat{q}_0$.
+We will have a linear functional $\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0} : \Psi \rightarrow \mathbb{C}$ such that $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2$ returns the amount of material prepared with the particular value $\hat{q}_0$.
 We can re-express this quantity in terms of the dual and the inner product extended to all Lebesgue integrable functions: $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2 = |\langle \hat{q}_0 | \psi \rangle|^2 $.
@@
 These ideas are appealing because they seem to fit well in the narrative, but at this point they are just speculation.
-The only thing we can say for now is that the state space under this new assumption would be a vector space on a field that is neither $\mathbbm{R}$ or $\mathbbm{C}$.
+The only thing we can say for now is that the state space under this new assumption would be a vector space on a field that is neither $\mathbb{R}$ or $\mathbb{C}$.
 But it is not unreasonable that these will inspire new work in the future.
@@
 	Let $(S, \tau)$ be a locally compact Hausdorff space.
-Let $\Lambda = \{\Lambda_U : C(S) \rightarrow \mathbbm{R}\}_{U \subseteq S}$ a family of positive linear functionals such that $\forall U \subseteq S \; \Lambda_U = \Lambda_{\mathrm{int}(U)}$ and $\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2} \; \forall U_1, U_2 \subseteq S$.
+Let $\Lambda = \{\Lambda_U : C(S) \rightarrow \mathbb{R}\}_{U \subseteq S}$ a family of positive linear functionals such that $\forall U \subseteq S \; \Lambda_U = \Lambda_{\mathrm{int}(U)}$ and $\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2} \; \forall U_1, U_2 \subseteq S$.
 Then there exists a unique Borel measure $\mu$ such that $\Lambda_U (\rho) = \int_{U} \rho d\mu \; \forall \rho \in C(S)$.
@@
 $\forall s \in S \; \exists U \subseteq S$ such that $U$ is compact.
-$U$ with the subspace topology is a compact Hausdorff topological space on which is defined a positive linear functional $\Lambda_U : C(S) \rightarrow \mathbbm{R}$.
+$U$ with the subspace topology is a compact Hausdorff topological space on which is defined a positive linear functional $\Lambda_U : C(S) \rightarrow \mathbb{R}$.
 For the Riesz representation theorem for linear functionals, there exists a unique Borel measure $\mu_U$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho d \mu_U \; \forall \rho \in C(S)$.
@@
 \begin{thrm}\label{everywhere_integrable_is_lebesgue_integrable}
-	Let $\mathcal{L}(S,\mu) \equiv \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\}$.
-$\mathcal{L}(S,\mu)=L^1(S,\mu)$ where $L^1(S,\mu) = \{ \rho : S \rightarrow \mathbbm{R} \; | \;\; \int_{S} |\rho| d\mu < \infty \}$
+	Let $\mathcal{L}(S,\mu) \equiv \{ \rho : S \rightarrow \mathbb{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\}$.
+$\mathcal{L}(S,\mu)=L^1(S,\mu)$ where $L^1(S,\mu) = \{ \rho : S \rightarrow \mathbb{R} \; | \;\; \int_{S} |\rho| d\mu < \infty \}$
 \end{thrm}
@@
 Let $S \in TM \; | \; f(\xi^a(m)) = \xi^a(m) + S^a(m)dt \; \forall m \in M$ be the infinitesimal displacement.
-There exists a function $H: M \rightarrow \mathbbm{R}$ such that $S^{a} \omega_{ab} = \partial_{b}H$ and $H \in C^2(M, \mathbbm{R})$.
+There exists a function $H: M \rightarrow \mathbb{R}$ such that $S^{a} \omega_{ab} = \partial_{b}H$ and $H \in C^2(M, \mathbb{R})$.
 \end{prop}
@@
 
-	We claim $H \in C^2(M, \mathbbm{R})$.
+	We claim $H \in C^2(M, \mathbb{R})$.
 $S$ is differentiable as it is the displacement of a symplectomorphism.
@@
 	Let $V$ be a complex vector space.
-Let $| \cdot | ^2 : V \rightarrow \mathbbm{R}$ such that $|x|^2 \ge 0 \; \forall x \in V$ and $|x|^2 = 0 \Leftrightarrow x = 0$.
+Let $| \cdot | ^2 : V \rightarrow \mathbb{R}$ such that $|x|^2 \ge 0 \; \forall x \in V$ and $|x|^2 = 0 \Leftrightarrow x = 0$.
 Let $x \in V$ and $\mathsf{P}_x : V \rightarrow V$ be a family of projections (i.e.
-$\mathsf{P}_x$ linear and $\mathsf{P}_x \circ \mathsf{P}_x = \mathsf{P}_x$) such that: $\mathsf{P}_x (y) = y \Leftrightarrow y = a x$ for some $a \in \mathbbm{C}$; $|x|^2|\mathsf{P}_x (y)|^2=|y|^2|\mathsf{P}_y (x)|^2$; $\mathsf{P}_x (\mathsf{P}_y (x)) = r x$ for some $r \in \mathbbm{R}$.
-	Then $V$ is an inner product space with the product $\langle \cdot , \cdot \rangle : V \times V \rightarrow \mathbbm{C}$ defined by $\langle x , y \rangle x = |x|^2 \mathsf{P}_x (y)$.
+$\mathsf{P}_x$ linear and $\mathsf{P}_x \circ \mathsf{P}_x = \mathsf{P}_x$) such that: $\mathsf{P}_x (y) = y \Leftrightarrow y = a x$ for some $a \in \mathbb{C}$; $|x|^2|\mathsf{P}_x (y)|^2=|y|^2|\mathsf{P}_y (x)|^2$; $\mathsf{P}_x (\mathsf{P}_y (x)) = r x$ for some $r \in \mathbb{R}$.
+	Then $V$ is an inner product space with the product $\langle \cdot , \cdot \rangle : V \times V \rightarrow \mathbb{C}$ defined by $\langle x , y \rangle x = |x|^2 \mathsf{P}_x (y)$.
 \end{thrm}
@@
 $\mathsf{P}_x (\mathsf{P}_x (y)) =\mathsf{P}_x(y)$ as $\mathsf{P}_x$ is a projection.
-$\mathsf{P}_x(y) = a x $ for some $a \in \mathbbm{C}$ by the first property of $\mathsf{P}_x$.
+$\mathsf{P}_x(y) = a x $ for some $a \in \mathbb{C}$ by the first property of $\mathsf{P}_x$.
 $\langle x , y \rangle = |x|^2 a$ exists and is unique.
@@
 	We claim $\langle \cdot , \cdot \rangle$ is linear in the second argument.
-Let $x,y,z \in V$ and $a,b \in \mathbbm{C}$.
+Let $x,y,z \in V$ and $a,b \in \mathbb{C}$.
 $\langle x , a y + b z \rangle x = |x|^2 \mathsf{P}_x(a y + b z)=a  |x|^2 \mathsf{P}_x(y) + b  |x|^2 \mathsf{P}_x(z) = a \langle x , y \rangle x + b \langle x, z \rangle x$ by linearity of $\mathsf{P}_x$.
@@
 Now consider $|x|^2|y|^2\mathsf{P}_x(\mathsf{P}_y(x)) = |x|^2 \langle y , x \rangle \mathsf{P}_x(y) = \langle y , x \rangle \langle x , y \rangle x$.
-The product $\langle x , y \rangle \langle y , x \rangle \in \mathbbm{R}$ by the third property of the projection.
+The product $\langle x , y \rangle \langle y , x \rangle \in \mathbb{R}$ by the third property of the projection.
 $\arg(\langle x , y \rangle) + \arg(\langle y , x \rangle)$ is $0$ or $\pi$.
@@
 \begin{thrm}\label{thrm:covariant_operator}
-	Let $C^1(\mathcal{Q}, \mathbbm{C})$ be the space of differentiable complex functions defined over $\mathcal{Q}$.
-Let $q : \mathcal{Q} \rightarrow \mathbbm{R}$ a differentiable function.
-Let $K_q : C^1(\mathcal{Q}, \mathbbm{C}) \rightarrow C^1(\mathcal{Q}, \mathbbm{C})$ be a covariant linear operator, that is given an invertible differentiable function $\hat{q} : \mathbbm{R} \rightarrow \mathbbm{R}$ $K_{\hat{q}} = \partial_{\hat{q}} q K_{q}$.
-Then $K_q = a \partial_q$ for some $a \in \mathbbm{C}$.
+	Let $C^1(\mathcal{Q}, \mathbb{C})$ be the space of differentiable complex functions defined over $\mathcal{Q}$.
+Let $q : \mathcal{Q} \rightarrow \mathbb{R}$ a differentiable function.
+Let $K_q : C^1(\mathcal{Q}, \mathbb{C}) \rightarrow C^1(\mathcal{Q}, \mathbb{C})$ be a covariant linear operator, that is given an invertible differentiable function $\hat{q} : \mathbb{R} \rightarrow \mathbb{R}$ $K_{\hat{q}} = \partial_{\hat{q}} q K_{q}$.
+Then $K_q = a \partial_q$ for some $a \in \mathbb{C}$.
 \end{thrm}
@@
 	We claim $K_q(f(q))= \partial_q f K_q(q)$.
-Let $I : \mathbbm{R} \rightarrow \mathbbm{R}$ be the identify function, that is $I \circ f = f \; \forall f : \mathbbm{R} \rightarrow \mathbbm{R}$.
+Let $I : \mathbb{R} \rightarrow \mathbb{R}$ be the identify function, that is $I \circ f = f \; \forall f : \mathbb{R} \rightarrow \mathbb{R}$.
 $K_q(q) = K_q(I(q)) = \partial_q I(q) K_{I}(I) = K_{I}(I)$ by covariance.
-	Let $f : \mathbbm{R} \rightarrow \mathbbm{R}$ be a differentiable invertible function.
+	Let $f : \mathbb{R} \rightarrow \mathbb{R}$ be a differentiable invertible function.
 $K_q(f(q)) =K_q(I(f(q))) = \partial_q f K_{f}(I(f)) = \partial_q f \partial_f I(f) K_{I}(I) = \partial_q f K_q(q)$.
-Now let $f : \mathbbm{R} \rightarrow \mathbbm{R}$ be a differentiable function non necessarily invertible.
+Now let $f : \mathbb{R} \rightarrow \mathbb{R}$ be a differentiable function non necessarily invertible.
 It can be written as $f = f_1 + f_2$ with $f_1, f_2$ differentiable and invertible.
@@
 
-	We claim $K_q = a \partial_q$ for some $a \in \mathbbm{C}$.
+	We claim $K_q = a \partial_q$ for some $a \in \mathbb{C}$.
 $K_q$ is a derivation.
-As $K_q(f(q))= \partial_q f K_q(q)$ it is a derivation along $q$, which can be expressed as $K_q(q)=a \partial_q$ for some $a \in \mathbbm{C}$.
+As $K_q(f(q))= \partial_q f K_q(q)$ it is a derivation along $q$, which can be expressed as $K_q(q)=a \partial_q$ for some $a \in \mathbb{C}$.
 \end{proof}
@@
 \begin{thrm}\label{thrm:antihermitian_derivative}
-	Let $G \equiv C^1(\mathcal{Q}, \mathbbm{C}) \cap L^1(\mathcal{Q}, \mathbbm{C})$ be the space of differentiable and integrable complex functions defined over $\mathcal{Q}$.
-Let $K$ be a linear adjoint operator of the form $a \partial_q$ with $a \in \mathbbm{C}$.
+	Let $G \equiv C^1(\mathcal{Q}, \mathbb{C}) \cap L^1(\mathcal{Q}, \mathbb{C})$ be the space of differentiable and integrable complex functions defined over $\mathcal{Q}$.
+Let $K$ be a linear adjoint operator of the form $a \partial_q$ with $a \in \mathbb{C}$.
 Let $K$ be self-adjoint with respect to the inner product $\langle f, g \rangle =
-	\int_{\mathcal{Q}} f^\dagger(q) g(q) dq$. Then $K= \imath r \partial_q$ with $r \in \mathbbm{R}$.
+	\int_{\mathcal{Q}} f^\dagger(q) g(q) dq$. Then $K= \imath r \partial_q$ with $r \in \mathbb{R}$.
 \end{thrm}
@@
 \begin{proof}
-	We claim $K= \imath r \partial_q$ with $r \in \mathbbm{R}$.
+	We claim $K= \imath r \partial_q$ with $r \in \mathbb{R}$.
 Let $f,g \in G$.
@@
 $a = - a^\dagger$.
-$a = \imath r$ with $r \in \mathbbm{R}$.
+$a = \imath r$ with $r \in \mathbb{R}$.
 \end{proof}

### 2017-01-05 2015CompleteDerivation/completeDerivation.tex
Changing style to match Foundations of Physics

@@
-\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
+%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
+\documentclass[draft,smallextended,nospthms]{svjour3}
 
@@
 
-\renewcommand\thesubsection{\thesection.\Alph{subsection}}
-\renewcommand{\theequation}{\thesection.\arabic{equation}}
+%\renewcommand\thesubsection{\thesection.\Alph{subsection}}
+%\renewcommand{\theequation}{\thesection.\arabic{equation}}
 \numberwithin{equation}{section}
@@
 \author{Gabriele Carcassi, Christine A. Aidala, David J. Baker, Lydia Bieri}
-\affiliation{University of Michigan, Ann Arbor, MI 48109}
-\email{carcassi@umich.edu}
+%\affiliation{University of Michigan, Ann Arbor, MI 48109}
+
+\institute{G. Carcassi \at
+	University of Michigan, Ann Arbor, MI 48109 \\
+%	Tel.: +123-45-678910\\
+%	Fax: +123-45-678910\\
+	\email{carcassi@umich.edu}           %  \\
+	%             \emph{Present address:} of F. Author  %  if needed
+}
+
+\titlerunning{From physical principles to classical and quantum particle mechanics}
+
 \date{\today}
+
+\maketitle
 
@@
 \end{abstract}
-\maketitle
-
-\textbf{\emph{Status}}
-
-Section III - Done.
-
-Section IV - Done.
-
-Section V - Done.
-
-Section VI - Mostly done. Needs review.
-
-Section VII - Mostly done. Needs review.
-
-Section VIII - Mostly done. Needs review.
-
-Section IX - Mostly done. Needs review.
 
 \section{Introduction}
+
+Status:
+\begin{itemize}
+\item Section \ref{sec:fundamental_model}: Done.
+
+\item Section \ref{sec:dynamical_systems}: Done.
+
+\item Section \ref{sec:reducibility}: Done.
+
+\item Section \ref{sec:relativity}: Done.
+
+\item Section \ref{sec:Lagrangian}: Mostly done. Needs review.
+
+\item Section \ref{sec:quantum_prelude}: Mostly done. Needs review.
+
+\item Section \ref{sec:irreducibility}: Mostly done. Needs review.
+\end{itemize}
 
@@
 
-We recognize the equations as those for Hamiltonian mechanics on the extended phase space.\cite{Synge, Struckmeier}
+We recognize the equations as those for Hamiltonian mechanics on the extended phase space.\cite{Synge,Struckmeier}
 
@@
 \begin{prop}\label{prop:form_of_invariant_hamiltonian}
-	The form of the invariant Hamiltonian is $\mathcal{H} = \lambda(t,q^i,E,p_i) \, (H(t,q^i,p_i) - E)$ where $H(t,q^i,p_i)$ is the standard Hamiltonian at each instant $t$, the conjugate time variable $E = H(t,q^i,p_i)$ represents the value of the Hamiltonian, and  $\lambda : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ where $\lambda(\xi^a) = d_s t$.
+	The form of the invariant Hamiltonian is given by $\mathcal{H} = \lambda(t,q^i,E,p_i) \,  (H(t,q^i,p_i) - E)$ where $H(t,q^i,p_i)$ is the standard Hamiltonian at each instant $t$, the conjugate time variable $E = H(t,q^i,p_i)$ represents the value of the Hamiltonian, and  $\lambda : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ where $\lambda(\xi^a) = d_s t$.
 $\bar{\mathcal{S}}$ is the level set for $\mathcal{H} = 0$.
@@
 \end{equation}
-where $\sigma^2$ is the variance, $\sigma$ is the standard deviation, and $\rho_{X,Y}=COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient.
+where $\sigma^2$ is the variance, $\sigma$ the standard deviation, $\rho_{X,Y}= COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient.
 As $-1\leq\rho_{X,Y}\leq+1$, the total variance can be greater or smaller than the sum of the variance of the components.
@@
 \bibitem{Lanczos} Lanczos, C., ``The variational principles of mechanics'', University of Toronto Press, (1949).
-\bibitem{Synge} Synge, J. L., ``Classical dynamics'', Encyclopedia of Physics Vol 3/1, Springer (1960).
+\bibitem{Synge} Synge, J. L.: Classical dynamics. Encyclopedia of Physics Vol 3/1, Springer (1960)
+
 \bibitem{Struckmeier} Struckmeier, J., ``Hamiltonian dynamics on the symplectic extended phase space for autonomous and non-autonomous systems'', J. Phys. A: Math. Gen. 38, pp. 1257--1278, (2005).

### 2017-01-05 2015CompleteDerivation/completeDerivation.tex
Formatting and reordering references

@@
 %\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
-\documentclass[draft,smallextended,nospthms]{svjour3}
+\documentclass[smallextended,nospthms]{svjour3}
 
@@
 
-However precise our measurements are, we can only gather a finite amount of statistics and each outcome is expressed by a finite set of digits; therefore, the set of outcomes is countable.\footnote{The information provided by the process as measured by Shannon's entropy\cite{Shannon} is finite.} Note that different outcomes for the same process can overlap.
+However precise our measurements are, we can only gather a finite amount of statistics and each outcome is expressed by a finite set of digits; therefore, the set of outcomes is countable.\footnote{The information provided by the process as measured by Shannon's entropy\cite{Shannon,Jaynes} is finite.} Note that different outcomes for the same process can overlap.
 For example, $4.12 \pm 0.05$ cm and $4.13 \pm 0.05$ cm are both legitimate possible outputs of the same measurement device.
@@
 
-We recognize the equations as those for Hamiltonian mechanics on the extended phase space.\cite{Synge,Struckmeier}
+We recognize the equations as those for Hamiltonian mechanics on the extended phase space.\cite{Synge,Lanczos,Struckmeier}
 
@@
 \begin{thebibliography}{0}
-
-\bibitem{Shannon} Shannon, C. E., ``A mathematical theory of communication'', The Bell System Technical Journal, Vol. 27, pp. 379--423, 623--656, (1948).
-\bibitem{Jaynes} Jaynes, E. T., ``Information theory and statistical mechanics'', Statistical Physics 3, pp. 181--218, (1963).
-\bibitem{classical_dynamics} J. V. Jos\'{e}, E. J. Saletan, ``Classical Dynamics'', Cambridge University Press, (1998).
-\bibitem{Gromov} Gromov, M. L., ``Pseudo holomorphic curves in symplectic manifolds'', Inventiones Mathematicae 82, pp. 307--347, (1985).
-\bibitem{deGosson} de Gosson, M. A., ``The symplectic camel and the uncertainty principle: the tip of an iceberg?'', Foundations of Physics 39, pp. 194--214, (2009).
-\bibitem{Stewart} Stewart, I., ``The symplectic camel'', Nature 329, pp. 17--18, (1987).
-\bibitem{Lanczos} Lanczos, C., ``The variational principles of mechanics'', University of Toronto Press, (1949).
-\bibitem{Synge} Synge, J. L.: Classical dynamics. Encyclopedia of Physics Vol 3/1, Springer (1960)
-
-\bibitem{Struckmeier} Struckmeier, J., ``Hamiltonian dynamics on the symplectic extended phase space for autonomous and non-autonomous systems'', J. Phys. A: Math. Gen. 38, pp. 1257--1278, (2005).
+\bibitem{Roland}  
+Thomas, R. E., Rosa, A. J.: The analysis and design of linear circuits. Wiley (2001)
+
+\bibitem{Barron} Barron, R. F.: 8.2 Lumped-parameter analysis. In: Barron, R. F.: Industrial Noise Control and Acoustics. CRC Press (2002)
+%online at http://ssu.ac.ir/cms/fileadmin/user_upload/Daneshkadaha/dbehdasht/behdasht_herfei/e_book/_Randall_F._Barons_Industrial_Noise_Control_and_.org__1_.pdf
+
+\bibitem{Borutzky} Borutzky, W.: Bond graph methodology: development and analysis of multidisciplinary dynamic system models.
+Springer (2010)
+
+\bibitem{Freiding} Freidin, R., Lasnik, H.: Syntax critical concepts in linguistics. Routledge (2006)
+
+\bibitem{Friedman} Friedman, D. P., Wand, M., Haynes, C.T.: Essentials of Programming Languages. MIT Press (2001)
+
+\bibitem{LeeTM} Lee, J. M.: Introduction to Topological Manifolds. Springer (2000)
+
 \bibitem{Lee} Lee, J. M.: Introduction to Smooth Manifolds. Springer (2003)
-\bibitem{LeeTM} Lee, J. M.: Introduction to Topological Manifolds. Springer (2000)
+
 \bibitem{Lang} Lang, S. L.: Algebra. Springer (2002)
@@
 
-\bibitem{Barron} Barron, R. F.: 8.2 Lumped-parameter analysis. In: Barron, R. F.: Industrial Noise Control and Acoustics. CRC Press (2002)
-%online at http://ssu.ac.ir/cms/fileadmin/user_upload/Daneshkadaha/dbehdasht/behdasht_herfei/e_book/_Randall_F._Barons_Industrial_Noise_Control_and_.org__1_.pdf
-
-\bibitem{Roland}  
- Thomas, R. E., Rosa, A. J.: The analysis and design of linear circuits. Wiley (2001)
-
-\bibitem{Borutzky} Borutzky, W.: Bond graph methodology: development and analysis of multidisciplinary dynamic system models.
-Springer (2010)
-
-\bibitem{Freiding} Freidin, R., Lasnik, H.: Syntax critical concepts in linguistics. Routledge (2006)
-
-\bibitem{Friedman} Friedman, D. P., Wand, M., Haynes, C.T.: Essentials of Programming Languages. MIT Press (2001)
+\bibitem{classical_dynamics} Jos\'{e}, J. V., Saletan, E. J.: Classical Dynamics. Cambridge University Press, (1998)
+
+\bibitem{Shannon} Shannon, C. E.: A mathematical theory of communication. The Bell System Technical Journal, Vol. 27, pp. 379--423, 623--656, (1948)
+
+\bibitem{Jaynes} Jaynes, E. T.: Information theory and statistical mechanics. Statistical Physics 3, pp. 181--218, (1963)
+
+\bibitem{Gromov} Gromov, M. L.: Pseudo holomorphic curves in symplectic manifolds. Inventiones Mathematicae 82, pp. 307--347, (1985)
+
+\bibitem{deGosson} de Gosson, M. A.: The symplectic camel and the uncertainty principle: the tip of an iceberg?. Foundations of Physics 39, pp. 194--214, (2009)
+
+\bibitem{Stewart} Stewart, I.: The symplectic camel, Nature 329, pp. 17--18, (1987)
 
@@
 
+\bibitem{Synge} Synge, J. L.: Classical dynamics. Encyclopedia of Physics Vol 3/1, Springer (1960)
+
+\bibitem{Lanczos} Lanczos, C.: The variational principles of mechanics, University of Toronto Press, (1949)
+
+\bibitem{Struckmeier} Struckmeier, J.: Hamiltonian dynamics on the symplectic extended phase space for autonomous and non-autonomous systems. J. Phys. A: Math. Gen. 38, pp. 1257--1278, (2005)
+
 \bibitem{EPR} Einstein, A., Podolsky, B.,  Rosen, N.: Can quantum-mechanical description of physical reality be considered complete? Phys. Rev. 47 777 (1935)

### 2017-01-05 2015CompleteDerivation/completeDerivation.tex
Some equations can now fit in one line

@@
 \mathcal{H} &= \frac{1}{2m} ( p^i p_i - (E/c)^2 + m^2c^2) \\
- &= \frac{1}{2mc^2} (c \sqrt{m^2 c^2 + p^i p_i} + E) \\
- &(c \sqrt{m^2 c^2 + p^i p_i} - E) \\
+ &= \frac{1}{2mc^2} (c \sqrt{m^2 c^2 + p^i p_i} + E) (c \sqrt{m^2 c^2 + p^i p_i} - E) \\
 \lambda &= \frac{1}{2mc^2} (c \sqrt{m^2 c^2 + p^i p_i} + E)
@@
 L = &\int (q-\mu_q)^2 \rho_{\mathcal{c}} \, dqdp \int (p-\mu_p)^2 \rho_{\mathcal{c}} \, dqdp \\
-&+ \lambda_1(\int \rho_{\mathcal{c}} dqdp - 1)\\
-&+ \lambda_2(- \int \rho_{\mathcal{c}} \ln \rho_{\mathcal{c}} \, dqdp - I_0)\\
-\delta L = &\int \delta \rho_{\mathcal{c}} [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\ &+ \lambda_1 - \lambda_2 \ln \rho_{\mathcal{c}} - \lambda_2 ] dqdp = 0 \\
+&+ \lambda_1(\int \rho_{\mathcal{c}} dqdp - 1) + \lambda_2(- \int \rho_{\mathcal{c}} \ln \rho_{\mathcal{c}} \, dqdp - I_0)\\
+\delta L = &\int \delta \rho_{\mathcal{c}} [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 + \lambda_1 - \lambda_2 \ln \rho_{\mathcal{c}} - \lambda_2 ] dqdp = 0 \\
 \lambda_2 \ln \rho_{\mathcal{c}} = &\lambda_1 - \lambda_2 + (q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\
@@
 	&= (v^{a} + \partial_{c} S^{a} v^{c} dt) \omega_{ab} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
-	&= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} \\
-	&+ v^{a} \omega_{ab} \partial_{d} S^{b} w^{d}) dt + O(dt^2)
+	&= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} + v^{a} \omega_{ab} \partial_{d} S^{b} w^{d}) dt + O(dt^2)
 	\end{align*}
 	$v^{c} w^{b} \partial_{c} S_{b} - v^{a} w^{d} \partial_{d} S_{a} = 0$ where $S_{b} \equiv S^{a} \omega_{ab}$.
-The relationship must true for and pair of vector, therefore $\partial_{a} S_{b} - \partial_{b} S_{a} = curl(S_{a}) = 0$.
+The relationship must be true for and pair of vector, therefore $\partial_{a} S_{b} - \partial_{b} S_{a} = curl(S_{a}) = 0$.
 $S$ is a curl free vector field and admits a potential $H$ such that $S_{a} = \partial_{a}H$.

### 2017-01-06 2015CompleteDerivation/completeDerivation.tex
Finished reviewing quantum with Christine

@@
 As we made it clear from the start, deterministic and reversible evolution is an assumption.
-And as we have seen in the justifications for assumptions \ref{ass:infinitesimal_reducibility} and \ref{ass:irreducibility} we cannot have determinism at all levels of decomposition while transferring information from one system to another.
+And as we have seen in the rationale for assumptions \ref{ass:infinitesimal_reducibility} and \ref{ass:irreducibility} we cannot have determinism at all levels of decomposition while transferring information from one system to another.
 
@@
 It is also clear that it cannot always be the same detector: that corresponds to states where the particle is spread within the aperture of that detector.
-In other words: as the particle hits the array, one detector triggers randomly based on the distribution of the particle.
+In other words: as the particle hits the array, one detector triggers non-deterministically based on the distribution of the particle.
 
 The process is non-deterministic in the sense that the state of the particle is not enough to predict which detector will trigger.
-Maybe it depends on the state of the detector, maybe it depends on the unstated part of the particle.
+Maybe it depends on the state or unstated part of the detector, maybe it depends on the unstated part of the particle or maybe on the environment.
 The point is that a single measurement does not tell us much about the state of the incoming particle.
 
-We then repeat the process, making sure the incoming particle is in the same state.
+We then repeat the process, making sure the incoming particle is prepared in the same way.
 Given a single "pure" incoming particle state, we get a "mixed" statistical distribution describing the output of the array.\footnote{Note that this is different from the model used often to describe the quantum measurement problem, where the incoming system becomes entangled with the measuring device.
@@
 That is if $N$ is the total number of takes and $N_A$ is the number of times the detector with aperture in the $A$ region captured the particle, the ratio $\frac{N_A}{N}$ will be approximately the amount of material within the aperture $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq$.
-If we assume an infinite amount of takes, we expect them to coincide: $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq = \lim\limits_{N\rightarrow\infty} \frac{N_A}{N}$.
+If we assume an infinite number of takes, we expect them to coincide: $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq = \lim\limits_{N\rightarrow\infty} \frac{N_A}{N}$.
 But this is also the frequentist definition of probability.
@@
 We will have a linear functional $\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0} : \Psi \rightarrow \mathbb{C}$ such that $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2$ returns the amount of material prepared with the particular value $\hat{q}_0$.
-We can re-express this quantity in terms of the dual and the inner product extended to all Lebesgue integrable functions: $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2 = |\langle \hat{q}_0 | \psi \rangle|^2 $.
+We can re-express this quantity in terms of the dual vector $\hat{q}_0$ corresponding to the linear functional $\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}$ and the inner product extended to all Lebesgue integrable functions: $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2 = |\langle \hat{q}_0 | \psi \rangle|^2 $.
 This is also the probability that the outcome associated with $\hat{q}_0$ is selected by the measuring device, thus recovering the statistical nature of the quantum predictions.
@@
 Note that while this process is of a non-deterministic nature, it is of a specific non-deterministic nature.
-It requires that no other state can influence the final distribution or the distribution of measurement outcomes would not match the distribution of incoming material.
-That is: it cannot be understood as an averaging over some unknown external state.
+It requires that no other state can influence the final distribution as the match between the physical material distribution and the statistical measurement distribution was recovered only under the assumption that the detector does not introduce any bias.
+That is: the non-deterministic nature cannot be understood as an averaging over some unknown external state.
 
@@
 So much so that we can substitute it with a deterministic one to compute the probability distribution.
-In particular, we have no way of predicting which set of final outcomes we will be projecting to, nor when does that actually happen.
+In particular, given only the initial state, we have no way of predicting which set of final outcomes the system will project to as that depends on the actual choice of detector setup, nor when that projection actually happens.
 
 While this may seem like an incomplete description, it is as complete as the irreducibility assumption allows.
-The state is known and determines the final statistical distribution.
-The unstated part cannot be known, and influences the non-deterministic evolution.
-If we were able to say something more about the non-deterministic evolution, we would be able to describe an aspect of the unstated part, which we cannot do without violating the assumption.
-If we cannot learn anything about the unstated part from the final distribution, then we are consistent with our assumption and we are able to use deterministic evolution and projection to determine the final mixed state.
-In other words: the inability to describe the unstated part must be matched with a corresponding inability to describe its evolution.
-
-We therefore have arrived to all the major underpinning of quantum mechanics and see how they stem from assumption \ref{ass:irreducibility} of irreducibility.
-This gives us a more satisfactory account on what quantum mechanics describes, though it does not claim to solve many of the open problems.
-It does, though, give good reason why we have those problems and how the desire to know ``what is really going on" is hindered by the very assumption that the quantum description is based on.
+The state can be known and determines any final statistical distribution.
+The unstated part cannot be known, and is what gives the statistical distribution in the first place.
+If we were able to say something more about the evolution of the unstated part, we would be able to further describe it, which we cannot do without violating the assumption.
+If we cannot learn anything about the unstated part from the final distribution, then we are consistent with our assumption and we are able to use deterministic evolution and projection to determine the final statistical distribution.
+In other words: the inability to describe the unstated part not only constrains the state space and deterministic evolution, but also the set of non-deterministic processes by excluding those for which the final statistical distribution would depend on the unstated part itself.
+
+We therefore have arrived at all the major underpinnings of quantum mechanics and see how they stem from assumption \ref{ass:irreducibility} of irreducibility.
+This gives us a more satisfactory account of what quantum mechanics describes, although it does not claim to solve many of the open problems.
+It does, however, give good reasons as to why we have those problems and how the desire to know ``what is really going on" is hindered by the very assumption that the quantum description is based on.
 
@@
 
-As we went through many details in the derivation, it is time to step back and conclude with a few general remark.
+As we went through many details in the derivation, it is time to step back and conclude with a few general remarks.
 
@@
 In hindsight, the idea that physics should be founded on physical principles is nothing new: it just hasn't been thoroughly applied to basic mechanics.
-If this work would have been done incrementally from the start, these concepts would have been developed gradually and the results would feel more mundane.
+If this work had been done incrementally from the start, these concepts would have been developed gradually and the results would feel more mundane.
 
@@
 The usual reductionist view is that you start with the rules of the small parts and then build up the composite whole.
-This works shift that perspective upside down: it's not what the parts can tell you about the whole, it's what the whole can tell you about the parts.
-
-The state space for classical particles is the one that allows to express invariant distributions: that is a requirement that comes from the whole.
-The fragments of quantum particles do not even have a well defined description: the whole tells us nothing about them.
+This work flips that perspective: it's not what the parts can tell you about the whole, it's what the whole can tell you about the parts.
+
+The state space for classical particles is the one that allows us to express invariant distributions: that is a requirement that comes from the whole.
+The fragments of quantum particles do not even have a well defined description: the whole tells us nothing about their configuration.
 In retrospect, this is not be surprising.
 Nobody is going to hand us down the laws of infinitesimal parts: we induce them by manipulating finite systems.
-Therefore what the whole allows to infer about its part is physically fundamental.
+Therefore what the whole allows us to infer about its part is physically fundamental.
 
@@
 
-In this section we include mathematical demonstration in support of the physical arguments of the paper.
+In this section we include mathematical demonstration in support of the physical justifications of the paper.
 
@@
 
-\begin{prop}\label{symplectomorphism_generator}
+\begin{thrm}\label{symplectomorphism_generator}
 	Let $(M, \omega)$ a symplectic manifold.
@@
 There exists a function $H: M \rightarrow \mathbb{R}$ such that $S^{a} \omega_{ab} = \partial_{b}H$ and $H \in C^2(M, \mathbb{R})$.
-\end{prop}
+\end{thrm}
 

### 2017-01-08 2015CompleteDerivation/completeDerivation.tex
Using theorem env provided by journal
Reviewed intro/conclusions with Christine

@@
 %\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
-\documentclass[smallextended,nospthms]{svjour3}
+\documentclass[smallextended]{svjour3}
 
@@
 \usepackage{amsfonts}
-\usepackage{amsthm}
 \usepackage{dutchcal}
@@
 
+%----
+
+% Theorem definitions using amsthm
+
+%\usepackage{amsthm}
+
 %\renewcommand\thesubsection{\thesection.\Alph{subsection}}
@@
 
-\newtheorem{assump}{Assumption}
-\renewcommand*{\theassump}{\Roman{assump}}
-\newtheorem{thrm}[equation]{Theorem}
-
-\theoremstyle{definition}
-\newtheorem{prop}[equation]{Proposition}
-
-\newenvironment{rationale}{\emph{Rationale}.}{\qed}
-\newenvironment{justification}{\emph{Justification}.}{\qed}
-\renewenvironment{proof}{\emph{Proof}.}{\qed}
+%\newtheorem{assump}{Assumption}
+%\renewcommand*{\theassump}{\Roman{assump}}
+%\newtheorem{thrm}[equation]{Theorem}
+
+%\theoremstyle{definition}
+%\newtheorem{prop}[equation]{Proposition}
+
+%\newenvironment{rationale}{\emph{Rationale}.}{\qed}
+%\newenvironment{justification}{\emph{Justification}.}{\qed}
+%\renewenvironment{proof}{\emph{Proof}.}{\qed}
+
+%----
+
+\spnewtheorem{assump}{Assumption}{\bf}{\it}
+\renewcommand{\theassump}{\Roman{assump}}
+\spnewtheorem{prop}[equation]{Proposition}{\bf}{\rm}
+\spnewtheorem{thrm}[equation]{Theorem}{\bf}{\it}
+
+\newenvironment{rationale}{\emph{Rationale}.}{\hfill\(\qed\)}
+\newenvironment{justification}{\emph{Justification}.}{\hfill\(\qed\)}
+\renewenvironment{proof}{\emph{Proof}.}{\hfill\(\qed\)}
 
@@
 	University of Michigan, Ann Arbor, MI 48109 \\
-%	Tel.: +123-45-678910\\
-%	Fax: +123-45-678910\\
 	\email{carcassi@umich.edu}           %  \\
-	%             \emph{Present address:} of F. Author  %  if needed
 }
 
+% As the page margins are so large, the full title does not fit
 \titlerunning{From physical principles to classical and quantum particle mechanics}
@@
 While some branches of physics (e.g.
-Newtonian mechanics(ref?), thermodynamics(ref?), special relativity(ref?)) are derived from physical principles (e.g.
+Newtonian mechanics, thermodynamics, special relativity) are derived from physical principles (e.g.
 Newton's laws, laws of thermodynamics, invariance of the speed of light), others start by simply setting a mathematical framework.
-It is the case for classical Lagrangian mechanics(ref?), where we start by assuming the existence of a Lagrangian function whose path integral is minimized during the motion.
-It is the case for classical Hamiltonian mechanics(ref?), where we start by assuming the state is described by conjugate pairs $(q^i, p_i)$ and evolves according to Hamilton's equations.
-It is the case for quantum mechanics(ref?), where the state space is a complex inner product vector space and the state evolves according to Schroedinger's equations.
+This is the case for classical Lagrangian mechanics, where we start by assuming the existence of a Lagrangian function whose path integral is minimized during the motion, for classical Hamiltonian mechanics, where we start by assuming the state is described by conjugate pairs $(q^i, p_i)$ and evolves according to Hamilton's equations, and for quantum mechanics, where the state space is a complex inner product space and the state evolves according to Schroedinger's equation.
 That is, at the very core of some of our physical theories we have mathematical definitions instead of physical ideas.
 
-If we ask ``when can a particular physical system be suitably described by classical Hamiltonian mechanics?" the only answer we have is ``when it fits the mathematical framework".
+If we ask ``when can a particular physical system be suitably described by classical Hamiltonian mechanics?" the only rigorous answer we have is ``when it fits the mathematical framework".
 While many seem to be content with this answer (e.g.
@@
 This is a problem since mathematical structures are not enough to characterize the physical system being described.
-For example, the same mathematical framework for linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems.\cite{Roland,Barron,Borutzky} The mathematical treatment is the same because it captures aspects of the description that are common to all cases.
+For example, the same mathematical framework for linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems~\cite{Roland,Barron,Borutzky}.
+The mathematical treatment is the same because it captures aspects of the description that are common to all cases.
 But for the very same reason, we cannot infer what system is being described if all we are given is the mathematical model.
-In the jargon of linguistic/computer science/phylosophy\cite{Freiding,Friedman}: math captures the syntax (the relationships between the objects) but does not capture the semantic (the meaning of each object).
+In the jargon of linguistics/computer science/philosophy~\cite{Freidin,Friedman}: math captures the syntax (the relationships between the objects) but does not capture the semantics (the meaning of each object).
 
 We started this work with the unpretentious goal of exploring simple questions such as the following.
-What do the mathematical symbols used in classical/quantum Hamiltonian represent?
-What is the physical meaning attached to it?
+What do the mathematical symbols used in classical/quantum Hamiltonian mechanics represent?
+What is the physical meaning attached to them?
 Why is the state of a quantum particle described by a wave function (i.e.
-a vector in a complex inner product space), while the state of a classical particle is described by the pairs $(q^i, p_i)$ (i.e.
+a vector in a complex inner product space), while the state of a classical particle is described by position and conjugate momentum (i.e.
 a point in a cotangent bundle)?
 The idea was to create a math-to-physics dictionary so that each mathematical concept was clearly mapped to a crisp physical concept.
-But to make sure that such dictionary was complete, we realized we had to make sure that the whole relevant mathematical framework could be recovered: the translation needs to work both ways.
-
-This redefined our mission to re-deriving as much as possible of the mathematical structure by starting from the smallest set of physical assumptions.
+But to make sure that such a dictionary was complete, we realized we had to make sure that the whole relevant mathematical framework could be recovered: the translation needs to work both ways.
+
+This redefined our mission to rederiving as much as possible of the mathematical structure by starting from the smallest set of physical assumptions.
 It also forced us to push further and further back toward the basic definitions.
@@
 But when we did find physical concepts that mapped to these more basic mathematical definitions, we realized that this link was valid for a far wider range of cases: the concepts were more general than we anticipated.
-For example, we realized that the mathematical notion of a topology maps incredibly well to the idea of physical distinguishability, but states of Hamiltonian system are not the only thing that are physically distinguishable.
+For example, we realized that the mathematical notion of a topology maps incredibly well to the idea of physical distinguishability, but states of Hamiltonian system are not the only things that are physically distinguishable.
 
 Each new link often shrank the complexity of the arguments while expanding the scope of the work in a way that was not at all anticipated.
-The same link that allowed to justify Hamiltonian dynamics also gave us relativistic motion on the extended phase space.
-The same link that allowed to justify Lagrangian mechanics also constrained the motion to that of free particles under scalar/vector potential forces.
+The same link that allowed us to justify Hamiltonian dynamics also gave us relativistic motion on the extended phase space.
+The same link that allowed us to justify Lagrangian mechanics also constrained the motion to that of massive particles under scalar/vector potential forces.
 We also realized that sometimes different mathematical concepts could be justified from the same physical starting point.
@@
 
-The unexpected success of the approach lead us to think that, maybe, the picture we were developing was not simply ``a nicer way to think about things", but instead a better framework to understand and link different parts of basic science and mathematics.
+The unexpected success of the approach led us to think that, maybe, the picture we were developing was not simply ``a nicer way to think about things", but instead a better framework to understand and link different parts of basic science and mathematics.
 After many years of work and different incremental iterations, we think the arguments are well developed enough to report the progress.
 
-The goal of this paper is to show that it is possible to derive classical/quantum Hamiltonian/Lagrangian mechanics starting from few physical assumptions.
-We do \emph{not} claim that we have perfectly achieved this result, with no room from improvement.
-We do claim that the whole architectural framework is solid enough to be convinced that this is indeed possible, and that it is a worthwhile endeavor as it allows to make intuitive connections between different areas of physics and mathematics.
+The goal of this paper is to show that it is possible to derive classical/quantum Hamiltonian/Lagrangian particle mechanics starting from few physical assumptions.
+We do \emph{not} claim that we have perfectly achieved this result, with no room for improvement.
+We do claim that the whole architectural framework is solid enough to be convinced that this is indeed possible, and that it is a worthwhile endeavor as it allows us to make intuitive connections between different areas of physics and mathematics.
 That is: it \emph{can} and \emph{should} be done.
 
-While trying to re-derive classical, quantum and (to some extent) relativistic mechanics may seem too broad a scope, it is necessary to show that the concepts we develop are truly general, that they can be applied to many different cases.
+While trying to rederive classical, quantum and (to some extent) relativistic mechanics may seem too broad a scope, it is necessary to show that the concepts we develop are truly general, that they can be applied to many different cases.
 If we could derive classical Hamiltonian mechanics on assumptions that are inapplicable to the quantum case, what good would it do?
-Quantum system are Hamiltonian systems, so clearly we wouldn't have understood what makes a system Hamiltonian.
-The ability to apply the same physical ideas far and wide is what gives us confidence that these are truly general concept that capture the physical aspects we are describing.
-The breath itself is what makes the work interesting and more than a mere curiosity.
+Quantum systems are Hamiltonian systems, so clearly we wouldn't have understood what makes a system Hamiltonian.
+The ability to apply the same physical ideas far and wide is what gives us confidence that these are truly general concepts that capture the physical aspects we are describing.
+The breadth itself is what makes the work interesting and more than a mere curiosity.
 
 As a brief overview, there are four assumptions we will consider.
-Assuming deterministic and reversible evolution will give us a dynamical system: states and a law of evolution.
+Assuming deterministic and reversible evolution gives us a dynamical system: states and a law of evolution.
 Additionally assuming infinitesimal reducibility, studying the whole system is equivalent to studying its infinitesimal parts, gives us Hamiltonian mechanics.
-Additionally assuming kinematic equivalence, trajectories in space tell us everything about the state of the system, gives us Lagrangian mechanics and constrains the motion to massive particles under scalar/vector potential forces.
-Relativistic motion arises when studying time dependent laws of evolution with no further assumption.
-Alternatively, as the second assumption we can take irreducibility, studying the whole systems tells us nothing about its parts, which leads to quantum (Hamiltonian) mechanics.
-
-As a more detailed overview: in section \ref{sec:fundamental_model} we will describe in general what it means to study a physical system and how the assumption of determinism and reversibility plays into it.
-We will develop an abstract conceptual model that we will use when discussing the later assumptions.
-
-In section \ref{sec:dynamical_systems} we will show how the physical concept of physical distinguishability leads to the mathematical concept of topological spaces.
+Further assuming kinematic equivalence, trajectories in space tell us everything about the state of the system, gives us Lagrangian mechanics and constrains the motion to massive particles under scalar/vector potential forces.
+Relativistic motion arises when studying time dependent laws of evolution with no further assumptions.
+Alternatively, as the second assumption we can instead take irreducibility, studying the whole system tells us nothing about its parts, which leads to quantum (Hamiltonian) mechanics.
+
+As a more detailed overview: in section \ref{sec:fundamental_model} we describe in general what it means to study a physical system and how the assumption of determinism and reversibility plays into it.
+We develop an abstract conceptual model that we use when discussing the later assumptions.
+
+In section \ref{sec:dynamical_systems} we show how the physical concept of physical distinguishability leads to the mathematical concept of topological spaces.
 Deterministic and reversible evolution is an invertible continuous map (i.e.
@@
 
-In section \ref{sec:reducibility} we will introduce the notion of a classical material, one for which we assume infinitesimal reducibility.
+In section \ref{sec:reducibility} we introduce the notion of a classical material, one for which we assume infinitesimal reducibility.
 The state of an arbitrary amount of material is a distribution over the states of the infinitesimal parts, the classical particles.
-As such distribution must be invariant under the arbitrary choice of units, we'll see that for each variable $q^i$ that defines a unit there is a corresponding variable $k^i$ defined on the inverse of such unit.
-Classical particle states can be identified by conjugate pairs of state variable $(q^i, p_i)$ (i.e.
+As such a distribution must be invariant under the arbitrary choice of units, we see that for each variable $q^i$ that defines a unit there is a corresponding variable $k^i$ defined on the inverse of that unit.
+Classical particle states can be identified by conjugate pairs of state variables $(q^i, p_i)$ (i.e.
 a point in the cotangent bundle $T^*\mathcal{Q}$) and the number of possible initial conditions for each pair is given by $dq^i \wedge dp_i$ (i.e.
@@
 
-In section \ref{sec:relativity} we will show that a relativistic version of Hamilton's equations arises naturally when we consider time dependent laws of evolution for distributions of classical material.
+In section \ref{sec:relativity} we show that a relativistic version of Hamilton's equations arises naturally when we consider time dependent laws of evolution for distributions of classical material.
 This also gives the notion of deterministic and reversible evolution a characterization that is independent of any transformation that mixes time and state variables $(q^i, p_i)$.
-It will also give us a classical equivalent for anti-particles states.
-
-In section \ref{sec:Lagrangian} we will introduce the assumption of kinematic equivalence, studying the motion in space is equivalent to studying the evolution of the state.
-Under that assumption there is a link between the state variables $(q^i, p_i)$ and the kinematic variable $(x^i, u^i)$ (i.e.
+Moreover it gives us a classical equivalent for anti-particle states.
+
+In section \ref{sec:Lagrangian} we introduce the assumption of kinematic equivalence, studying the motion in space is equivalent to studying the evolution of the state.
+Under that assumption there is a link between the state variables $(q^i, p_i)$ and the kinematic variables $(x^i, u^i)$ (i.e.
 position and velocity).
-This link will allow us to express the number of possible initial conditions for a d.o.f.
-in terms of $dx^i$ and $du^i$, leading to the metric tensor $g$.
-Such expression also allows to constrain the Hamiltonian to the one for massive particles under scalar potential forces and an equivalent Lagrangian formulation.
-
-In section \ref{sec:quantum_prelude} we will introduce some ideas in classical mechanics to more gently transition over quantum mechanics.
-We will see how classical materials (i.e.
-distributions over phase space) already have features that are typically associated with quantum systems and we will show how to construct a classical uncertainty principle that helps to build intuition used in the later section.
-
-In section \ref{sec:irreducibility} we will introduce the notion of an irreducible material, one for which studying the whole distribution does not tell us anything about the motion of its infinitesimal parts, the fragments.
-A unit amount will correspond to a quantum particle.
-As this means that the states of such material are invariant under permutations of the configurations for the fragments, we will show that the state space of an irreducible material is a complex vector space.
+This link allows us to express the number of possible initial conditions for a degree of freedom (d.o.f.) in terms of $dx^i$ and $du^i$, leading to the metric tensor $g$.
+Such an expression also allows us to constrain the Hamiltonian to the one for massive particles under scalar/vector potential forces and an equivalent Lagrangian formulation.
+
+In section \ref{sec:quantum_prelude} we introduce some ideas in classical mechanics to more gently transition to quantum mechanics.
+We see how distributions over phase space already have features that are typically associated with quantum systems and we will show how to construct a classical uncertainty principle that helps to build intuition used in the later section.
+
+In section \ref{sec:irreducibility} we introduce the notion of an irreducible material, one for which studying the whole distribution does not tell us anything about the motion of its infinitesimal parts, the fragments.
+A unit amount of this material will correspond to a quantum particle.
+As this means that the states of such a material are invariant under permutations of the configurations of the fragments, we show that the state space of an irreducible material is a complex vector space.
 These states can still be compared to each other in a way that gives the state space an inner product.
-Such comparisons will be preserved by deterministic and reversible evolution which leads to Schroedinger equation (i.e.
+Such comparisons will be preserved by deterministic and reversible evolution which leads to Schroedinger's equation (i.e.
 a unitary transformation).
 
-In section \ref{sec:discussion} we will discuss of how this work could be extended in different directions.
+In section \ref{sec:discussion} we present a few higher level remarks about the work.
 
@@
 By contradiction, $\mathcal{S}$ must be second countable.
-
 \end{justification}
@@
 $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i)) d^nq = \int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]}( \psi(\hat{q}^i)) d^n \hat{q}$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
-
 \end{justification}
@@
 The fragments of quantum particles do not even have a well defined description: the whole tells us nothing about their configuration.
-In retrospect, this is not be surprising.
-Nobody is going to hand us down the laws of infinitesimal parts: we induce them by manipulating finite systems.
-Therefore what the whole allows us to infer about its part is physically fundamental.
+In retrospect, this should not be surprising.
+Nobody is going to hand us down the laws of infinitesimal parts: we induce  what we can by manipulating finite systems.
+Therefore what the whole allows us to infer about its parts is physically fundamental.
 
@@
 Or infinitesimally reducible.
-In this context, the state, its properties and the law of evolution are the description of the system under our idealization, not some intrinsic feature of the system.
+In this context, the state, its properties and the laws of evolution are the description of the system under our idealization, not some intrinsic feature of the system.
 
@@
 First we assume a homogeneous infinitesimally reducible material.
-Then we assume an homogeneous irreducible material.
-Connecting the dots, the next assumption would be a inhomogeneous irreducible material: where the material is not in constant equilibrium and its particles do not all look the same.
+Then we assume a homogeneous irreducible material.
+Extrapolating from this progression, the next assumption would be an inhomogeneous irreducible material: where the material is not in constant equilibrium and its particles do not all look the same.
+
 This different starting point may have potential connections to some of the problems we mentioned in the rationales for the assumptions and to other interesting physics questions in general.
-
-As there would be different particles, they could be seen as different stable states of the same inhomogeneous material.
-In particular, since $\sigma_x \sigma_v \geq \frac{\hbar}{2m}$, particles in different generations (i.e.
-electron vs muon) could be similar stable configurations where the fragments of the material are more "packed" in terms of states labeled by the kinematic variables.
-
-As we are not always at equilibrium, the state of the system will be in general a statistical mixture of distributions of material.
-The evolution may also take pure states (i.e.
-a mixed state with a single state with $100\%$ probability) and evolve them into mixed states, thus giving an account of what happens during a measurement or a decay.
-
-The law of evolution may still be deterministic and reversible, in the sense that given a particular mixed state we can predict/reconstruct future/past mixed state, but it will not be time symmetric.
-We can then tell that time flows in the direction where information is exchanged, where entropy increases.
-This would give us a time arrow.
-
-These ideas are appealing because they seem to fit well in the narrative, but at this point they are just speculation.
-The only thing we can say for now is that the state space under this new assumption would be a vector space on a field that is neither $\mathbb{R}$ or $\mathbb{C}$.
-But it is not unreasonable that these will inspire new work in the future.
-
 In the foreseeable future, however, we will work on the more mundane task of extending this work to field theories, for which a more concrete path is available.
-The state space for a quantized material, already introduced here, should be captured by the Fock space.
-The idea of a non-minimal material will force us to consider vector and tensor quantities at each point in the distribution, whose state should be captured by spin systems.
-Assuming that the vector potential forces themselves are dependent on the state of a material should hopefully give us the Hamiltonian for the fields.
-Whether this can be done will be a matter of working through the details, which, we are sure, are likely to bring other unexpected surprises.
 
@@
 
-We have seen that much of fundamental physics can be derived from few simple physical assumption (and lots of hard work).
+We have seen that much of fundamental physics can be derived from few simple physical assumptions.
 In doing so we obtained a more unified picture of the different branches of fundamental physics centered around the idea of deterministic and reversible evolution of distributions of homogeneous materials.
@@
 Therefore it is an isomorphism on the mathematical structure defined by the state space (e.g.
-homeomorphism, canonical transformation, unitary transformation), not just a state-to-state bijection (i.e.
+homeomorphism, symplectomorphism, unitary transformation), not just a state-to-state bijection (i.e.
 isomorphism of sets).
@@
 \item Classical Hamiltonian mechanics describes the evolution of a homogeneous material whose state is equivalent to the states of all its infinitesimal parts, the classical particles.
-The state space for classical particles is a symplectic manifold $(T^*\mathcal{Q}, \omega)$ as this allows to define coordinate invariant density distributions; deterministic and reversible evolution for such state space is a symplectomorphism (i.e.
+The state space for classical particles is a symplectic manifold $(T^*\mathcal{Q}, \omega)$ as this allows us to define state-variable-invariant density distributions; deterministic and reversible evolution for such a state space is a symplectomorphism (i.e.
 a canonical transformation).
@@
 \item Quantum (Hamiltonian) mechanics describes the evolution of a homogeneous material for which the state can only be reduced up to finite parts, the quantum particles, whose internal motion is undetermined.
-The state space for quantum particles is a complex inner product space $(\Psi, \langle \cdot , \cdot \rangle)$; deterministic and reversible evolution for such state space is a unitary transformation.
+The state space for quantum particles is a complex inner product space $(\Psi, \langle \cdot , \cdot \rangle)$; deterministic and reversible evolution for such a state space is a unitary transformation.
 \end{itemize}
@@
 While we are sure that some mathematical details can be improved, we feel that the overall ``conceptual castle" is solid.
-The starting concepts are simple, the conceptual mapping of those concepts into math is often straight forward and the intuitive ideas that emerge help to connect different branches of math and physics.
-This is unlikely a matter of simple coincidence.
-
-While the scope of this work is limited to re-founding and re-understanding physical theories already experimentally established, it would not be unexpected that a better grasp of old theories may lead to insights and ideas for new ones.
+The starting physical assumptions are simple, the translation into math is often straightforward and the intuitive ideas that emerge help unveil connections within and across different branches of math and physics.
+It is the tight fitting of these different pieces into a broader picture that validates the merit of the approach.
+
+While the scope of this work is limited to refounding and reunderstanding physical theories already experimentally established, it would not be unexpected for a better grasp of old theories to lead to insights and ideas for new ones.
 
@@
 
-\bibitem{Freiding} Freidin, R., Lasnik, H.: Syntax critical concepts in linguistics. Routledge (2006)
+\bibitem{Freidin} Freidin, R., Lasnik, H.: Syntax: critical concepts in linguistics. Routledge (2006)
 

### 2017-01-08 2015CompleteDerivation/completeDerivation.tex
Reviewed intro with Christine

@@
 
-While some branches of physics (e.g.
+%While some branches of physics (e.g.
 Newtonian mechanics, thermodynamics, special relativity) are derived from physical principles (e.g.
@@
 
-If we ask ``when can a particular physical system be suitably described by classical Hamiltonian mechanics?" the only rigorous answer we have is ``when it fits the mathematical framework".
+%If we ask ``when can a particular physical system be suitably described by classical Hamiltonian mechanics?" the only rigorous answer we have is ``when it fits the mathematical framework".
 While many seem to be content with this answer (e.g.
@@
 
-This is a problem since mathematical structures are not enough to characterize the physical system being described.
+The past century has seen a resounding success in the use of mathematics in fundamental physics.
+In many cases, it was mathematical ideas that allowed progress in theoretical physics that ultimately led to experimental discoveries.
+This success should, and indeed is, celebrated but, however, we should keep in mind that it comes at a price.
+
+The issue is that mathematical structures are not enough to characterize what physical system is being studied.
 For example, the same mathematical framework for linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems~\cite{Roland,Barron,Borutzky}.
-The mathematical treatment is the same because it captures aspects of the description that are common to all cases.
+The mathematical treatment is the same because it captures aspects of the description that are common to all cases, that indeed is its power.
 But for the very same reason, we cannot infer what system is being described if all we are given is the mathematical model.
@@
 
-We started this work with the unpretentious goal of exploring simple questions such as the following.
+The real trouble comes when the mathematical description essentially becomes the foundation of a physical theory.
+While some branches of physics, such as Newtonian mechanics, thermodynamics and special relativity, are founded on physical laws or assumptions, others, such as Hamiltonian, Lagrangian and quantum mechanics, start by setting their mathematical structure.
+Nothing tells us why there should be  a Lagrangian function whose path integral is minimized during the motion, or when a state is described by conjugate pairs $(q^i, p_i)$ instead of a vector in a complex inner product space.
+Those are taken as given.
+
+This means that if we have a system in front of us, we do not have a rigorous way to conceptually establish whether the system is Hamiltonian or not, whether it is classical or quantum.
+We have some heuristics we can apply, but, at the end the day, a system is Hamiltonian because it follows the Hamiltonian framework.\footnote{The lack of a properly defined, and uniquely accepted, semantic is of course very evident in quantum mechanics with its numerous interpretations.}
+
+Unsatisfied with this situation, we started creating a sort of math-to-physics dictionary so that each mathematical concept could be clearly associated to a crisp physical concept.
+As we pushed furthered back toward the basic mathematical definitions, we ended up with physical concepts of a more general nature and found that, sometimes, a single physical idea could have more than one mathematical consequence.
+That is: by putting physics at the center, we could achieve a sort of unification of ideas that was not evident in the math alone.
+The unexpected success of the approach led us to think that, maybe, the picture we were developing was a better framework to understand and link different parts of basic science and mathematics.
+This work is the systematized result of this endeavor.
+
+%We started this work with the unpretentious goal of exploring simple questions such as the following.
 What do the mathematical symbols used in classical/quantum Hamiltonian mechanics represent?
@@
 
-This redefined our mission to rederiving as much as possible of the mathematical structure by starting from the smallest set of physical assumptions.
+%This redefined our mission to rederiving as much as possible of the mathematical structure by starting from the smallest set of physical assumptions.
 It also forced us to push further and further back toward the basic definitions.
@@
 
-Each new link often shrank the complexity of the arguments while expanding the scope of the work in a way that was not at all anticipated.
+%Each new link often shrank the complexity of the arguments while expanding the scope of the work in a way that was not at all anticipated.
 The same link that allowed us to justify Hamiltonian dynamics also gave us relativistic motion on the extended phase space.
@@
 
-The unexpected success of the approach led us to think that, maybe, the picture we were developing was not simply ``a nicer way to think about things", but instead a better framework to understand and link different parts of basic science and mathematics.
+%The unexpected success of the approach led us to think that, maybe, the picture we were developing was not simply ``a nicer way to think about things", but instead a better framework to understand and link different parts of basic science and mathematics.
 After many years of work and different incremental iterations, we think the arguments are well developed enough to report the progress.
 
-The goal of this paper is to show that it is possible to derive classical/quantum Hamiltonian/Lagrangian particle mechanics starting from few physical assumptions.
+The main goal is to show that it is possible to derive classical/quantum Hamiltonian/Lagrangian particle mechanics starting from few physical assumptions.
 We do \emph{not} claim that we have perfectly achieved this result, with no room for improvement.
-We do claim that the whole architectural framework is solid enough to be convinced that this is indeed possible, and that it is a worthwhile endeavor as it allows us to make intuitive connections between different areas of physics and mathematics.
-That is: it \emph{can} and \emph{should} be done.
-
-While trying to rederive classical, quantum and (to some extent) relativistic mechanics may seem too broad a scope, it is necessary to show that the concepts we develop are truly general, that they can be applied to many different cases.
+We \emph{do} claim that the whole architectural framework is solid enough to be convinced that not only this is indeed possible, but it allows a more unified picture that facilitates intuitive connections between different areas of physics and mathematics.
+It enables a consistent narrative across several domains of knowledge.
+
+Rederiving classical, quantum and (to some extent) relativistic mechanics within the same work is therefore crucial to show that the concepts we develop are truly general, that they can be applied to many different cases.
 If we could derive classical Hamiltonian mechanics on assumptions that are inapplicable to the quantum case, what good would it do?
 Quantum systems are Hamiltonian systems, so clearly we wouldn't have understood what makes a system Hamiltonian.
-The ability to apply the same physical ideas far and wide is what gives us confidence that these are truly general concepts that capture the physical aspects we are describing.
-The breadth itself is what makes the work interesting and more than a mere curiosity.
+The breadth of applicability is what makes the work valuable.
 
@@
 
-The overall layout of this work is a balance between the need to show that the derivation follows (which requires starting from basic principles and proceeding using formal arguments) and the need to give an intuitive understanding (which often requires starting from specific cases and then generalizing).
+The overall layout of this work is a balance between the need to show that the derivation follows, which requires starting from basic principles and proceeding using formal arguments, and the need to give an intuitive understanding, which often requires starting from specific cases and then generalizing.
 
@@
 \begin{description}
-  \item[Assumptions] these characterize the physical system we are studying and constitute the premise of our discussion.
-A \textbf{rationale} follows each assumption, which uses physical and sometimes philosophical arguments to motivate why (or why not) such an assumption makes sense (in a particular case).
-
-  \item[Propositions] these characterize the mathematical properties of the physical concepts we are studying.
-A \textbf{justification} follows each proposition to show why such characterization captures the particular physical properties of the system.
+  \item[Assumptions:] these characterize the physical system we are studying and constitute the premise of our discussion.
+A \textbf{rationale} follows each assumption, which uses physical and sometimes philosophical arguments to motivate why (or not) such an assumption makes sense in a particular case.
+
+  \item[Propositions:] these capture the properties of our physical objects into mathematical language.
+A \textbf{justification} follows each proposition to show the necessity of such a characterization.
 While the style of a justification is formal, it is not a pure mathematical proof as it necessarily has to mix physical and mathematical arguments.
 
-  \item[Theorems] these are pure mathematical statements that are used to simplify propositions.
-A mathematical \textbf{proof} follows each proposition.
+  \item[Theorems:] these are pure mathematical statements that are used to simplify propositions.
+A mathematical \textbf{proof} follows each theorem.
 No mathematical breakthrough should be expected as we mostly use well known results from different fields.
@@
 
-As these sections are clearly demarcated, they allow us to make sure that each justification only references past propositions. It also makes it easier to read the work at the desired level of detail.
-
+As these sections are clearly demarcated, they allow us to make sure that each justification only references past propositions and avoid circular arguments.
+It also makes it easier to read the work at the desired level of detail.
 Interspersed between each step of the more formal treatment, there are numerous discussions that conceptually explain motivations and results in a more approachable way.
-These can use more concrete and intuitive examples without the fear of circular arguments.
-As both the formal and more intuitive part need to cover the same material, this will inevitably lead to some amount of repetition.
-
-The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology\cite{LeeTM}, measure theory (ref?), differential geometry (symplectic and Rimannian)\cite{Lee}, group theory and vector spaces\cite{Lang,Young}, statistics (ref?), information theory (ref?), Hamiltonian mechanics, Lagrangian mechanics\cite{classical_dynamics}, quantum mechanics (ref?), thermodynamics (ref?), special relativity (ref?) and so on.
-Using idea from all these areas is necessary, in fact unraveling their connections is part of the value of this work, but it does create a number of problems.
+These can use more concrete and intuitive examples which may, in some cases, make forward references.
+As both the formal and more intuitive parts need to cover the same material, this will inevitably lead to some repetition.
+
+The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology \cite{LeeTM}, measure theory (ref?), differential geometry (symplectic and Riemannian) \cite{Lee}, group theory and vector spaces \cite{Lang,Young}, statistics (ref?), information theory (ref?), Hamiltonian mechanics, Lagrangian mechanics \cite{classical_dynamics}, quantum mechanics (ref?), thermodynamics (ref?), special relativity (ref?) and so on.\footnote{The references given provide the reader a list of resources consistent with the development of this work.} Using ideas from all these areas is necessary and valuable, but it does create a number of problems.
 
 The first one is terminology and prerequisite knowledge.
-Since we touch a good number of areas, we cannot expect all readers to be equally comfortable with all the technical terms.
+Since we touch numerous areas, we cannot expect all readers to be equally comfortable with all the technical terms.
 Yet we cannot introduce all of them either.
 The compromise we adopted is to keep the informal discussion more accessible while keeping the formal derivation more precise, without needing more than the basic definitions and few major results within each specific area.
-This may leave an expert in a particular field not fully satisfied but, as we mentioned in the introduction, our goal is to lay out the general picture.
+This may leave an expert in a particular field not fully satisfied but, as stated in the introduction, our goal is to lay out the general picture.
 
@@
 As each field has developed its own set of conventions, the common usage is now an empty set.
-We ended up with a compromise that hopefully feels natural enough to a broad audience of physicist, mathematical physicists and philosophers of physics.
-Within formal parts we may use notation and terminology more specific to the area of math being used, but one would need to be familiar with them anyway to be able to follow the details.
-
-We have tested the prose on a number of colleagues and students to confirm its accessibility at different levels and believe we have struck a decent compromise.
-We hope the reader enjoys this journey and comes out with a few new insights and ideas.
+We ended up with a compromise that hopefully feels natural enough to a broad audience of physicists, mathematical physicists and philosophers of physics.
+Within formal parts we may use notation and terminology more specific to the appropriate area of math.
+
+We have tested the prose on a number of colleagues and students in physics, philosophy and mathematics to confirm its accessibility at different levels and believe we have struck a reasonable compromise.
 

### 2017-01-12 2015CompleteDerivation/completeDerivation.tex
Another pass at the intro with Christine

@@
 \begin{itemize}
-\item Section \ref{sec:fundamental_model}: Done.
-
-\item Section \ref{sec:dynamical_systems}: Done.
-
-\item Section \ref{sec:reducibility}: Done.
-
-\item Section \ref{sec:relativity}: Done.
-
-\item Section \ref{sec:Lagrangian}: Mostly done. Needs review.
-
-\item Section \ref{sec:quantum_prelude}: Mostly done. Needs review.
-
-\item Section \ref{sec:irreducibility}: Mostly done. Needs review.
+%\item Section \ref{sec:fundamental_model}: Done.
+
+%\item Section \ref{sec:dynamical_systems}: Done.
+
+%\item Section \ref{sec:reducibility}: Done.
+
+%\item Section \ref{sec:relativity}: Done.
+
+\item Section \ref{sec:Lagrangian}: Needs review.
+
+\item Section \ref{sec:quantum_prelude}: Needs review.
+
+\item Section \ref{sec:irreducibility}: Needs review.
 \end{itemize}
 
-%While some branches of physics (e.g.
-Newtonian mechanics, thermodynamics, special relativity) are derived from physical principles (e.g.
-Newton's laws, laws of thermodynamics, invariance of the speed of light), others start by simply setting a mathematical framework.
-This is the case for classical Lagrangian mechanics, where we start by assuming the existence of a Lagrangian function whose path integral is minimized during the motion, for classical Hamiltonian mechanics, where we start by assuming the state is described by conjugate pairs $(q^i, p_i)$ and evolves according to Hamilton's equations, and for quantum mechanics, where the state space is a complex inner product space and the state evolves according to Schroedinger's equation.
-That is, at the very core of some of our physical theories we have mathematical definitions instead of physical ideas.
-
-%If we ask ``when can a particular physical system be suitably described by classical Hamiltonian mechanics?" the only rigorous answer we have is ``when it fits the mathematical framework".
-While many seem to be content with this answer (e.g.
-``you have to start from somewhere, it may as well be the math which at least is precise"), it begs the question: it does not tell us how and what physical properties of the system under study are being captured.
-
 The past century has seen a resounding success in the use of mathematics in fundamental physics.
-In many cases, it was mathematical ideas that allowed progress in theoretical physics that ultimately led to experimental discoveries.
-This success should, and indeed is, celebrated but, however, we should keep in mind that it comes at a price.
+In many cases, mathematical ideas were what allowed progress in theoretical physics and ultimately led to experimental discoveries.
+While these ideas are indeed useful---anything that can advance our understanding is welcome---they cannot, however, be the whole story.
 
@@
 For example, the same mathematical framework for linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems~\cite{Roland,Barron,Borutzky}.
-The mathematical treatment is the same because it captures aspects of the description that are common to all cases, that indeed is its power.
+The mathematical treatment is the same because it captures aspects of the description that are common to all cases; that indeed is its power.
 But for the very same reason, we cannot infer what system is being described if all we are given is the mathematical model.
@@
 This means that if we have a system in front of us, we do not have a rigorous way to conceptually establish whether the system is Hamiltonian or not, whether it is classical or quantum.
-We have some heuristics we can apply, but, at the end the day, a system is Hamiltonian because it follows the Hamiltonian framework.\footnote{The lack of a properly defined, and uniquely accepted, semantic is of course very evident in quantum mechanics with its numerous interpretations.}
+We have some heuristics we can apply, but, at the end the day, a system is Hamiltonian because it follows the Hamiltonian framework.\footnote{The lack of a properly defined, and uniquely accepted, semantics is of course very evident in quantum mechanics with its numerous interpretations.}
 
 Unsatisfied with this situation, we started creating a sort of math-to-physics dictionary so that each mathematical concept could be clearly associated to a crisp physical concept.
-As we pushed furthered back toward the basic mathematical definitions, we ended up with physical concepts of a more general nature and found that, sometimes, a single physical idea could have more than one mathematical consequence.
-That is: by putting physics at the center, we could achieve a sort of unification of ideas that was not evident in the math alone.
-The unexpected success of the approach led us to think that, maybe, the picture we were developing was a better framework to understand and link different parts of basic science and mathematics.
+As we pushed further back toward the basic mathematical definitions, we ended up with physical concepts of a more general nature and found that, in some cases, a single physical idea could have more than one mathematical consequence.
+That is: by putting physics at the center, we could achieve a unification of ideas that was not evident in the math alone.
+The unexpected success of the approach led us to think that, perhaps, the picture we were developing was a better framework to understand and link different areas of basic science and mathematics.
 This work is the systematized result of this endeavor.
-
-%We started this work with the unpretentious goal of exploring simple questions such as the following.
-What do the mathematical symbols used in classical/quantum Hamiltonian mechanics represent?
-What is the physical meaning attached to them?
-Why is the state of a quantum particle described by a wave function (i.e.
-a vector in a complex inner product space), while the state of a classical particle is described by position and conjugate momentum (i.e.
-a point in a cotangent bundle)?
-The idea was to create a math-to-physics dictionary so that each mathematical concept was clearly mapped to a crisp physical concept.
-But to make sure that such a dictionary was complete, we realized we had to make sure that the whole relevant mathematical framework could be recovered: the translation needs to work both ways.
-
-%This redefined our mission to rederiving as much as possible of the mathematical structure by starting from the smallest set of physical assumptions.
-It also forced us to push further and further back toward the basic definitions.
-For example, if we truly wanted to justify why the state space of Hamiltonian mechanics is charted by conjugate variables $q^i$ and $p_i$, we had to justify why it was a manifold as well, which in turn meant we needed to justify why it was a topological space.
-But when we did find physical concepts that mapped to these more basic mathematical definitions, we realized that this link was valid for a far wider range of cases: the concepts were more general than we anticipated.
-For example, we realized that the mathematical notion of a topology maps incredibly well to the idea of physical distinguishability, but states of Hamiltonian system are not the only things that are physically distinguishable.
-
-%Each new link often shrank the complexity of the arguments while expanding the scope of the work in a way that was not at all anticipated.
-The same link that allowed us to justify Hamiltonian dynamics also gave us relativistic motion on the extended phase space.
-The same link that allowed us to justify Lagrangian mechanics also constrained the motion to that of massive particles under scalar/vector potential forces.
-We also realized that sometimes different mathematical concepts could be justified from the same physical starting point.
-That is: by putting physics first we could achieve a sort of unification of ideas that was not evident in the math alone.
-
-%The unexpected success of the approach led us to think that, maybe, the picture we were developing was not simply ``a nicer way to think about things", but instead a better framework to understand and link different parts of basic science and mathematics.
-After many years of work and different incremental iterations, we think the arguments are well developed enough to report the progress.
 
@@
 We do \emph{not} claim that we have perfectly achieved this result, with no room for improvement.
-We \emph{do} claim that the whole architectural framework is solid enough to be convinced that not only this is indeed possible, but it allows a more unified picture that facilitates intuitive connections between different areas of physics and mathematics.
+We \emph{do} claim that the whole architectural framework is solid enough to be convinced that not only is this indeed possible, but it allows a more unified picture that facilitates intuitive connections within and across different areas of physics and mathematics.
 It enables a consistent narrative across several domains of knowledge.
@@
 Quantum systems are Hamiltonian systems, so clearly we wouldn't have understood what makes a system Hamiltonian.
-The breadth of applicability is what makes the work valuable.
+This breadth of applicability is what makes the work valuable.
 

### 2017-01-12 2015CompleteDerivation/completeDerivation.tex
Making single switch to change format
Removing second countability

@@
-%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
+\newif\ifjournal
+%\journaltrue
+\journalfalse
+
+\ifjournal
 	\documentclass[smallextended]{svjour3} 
+\else 
+	\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
+\fi
 
@@
 
-%----
-
-% Theorem definitions using amsthm
-
-%\usepackage{amsthm}
-
-%\renewcommand\thesubsection{\thesection.\Alph{subsection}}
-%\renewcommand{\theequation}{\thesection.\arabic{equation}}
 \numberwithin{equation}{section}
 
-%\newtheorem{assump}{Assumption}
-%\renewcommand*{\theassump}{\Roman{assump}}
-%\newtheorem{thrm}[equation]{Theorem}
-
-%\theoremstyle{definition}
-%\newtheorem{prop}[equation]{Proposition}
-
-%\newenvironment{rationale}{\emph{Rationale}.}{\qed}
-%\newenvironment{justification}{\emph{Justification}.}{\qed}
-%\renewenvironment{proof}{\emph{Proof}.}{\qed}
-
-%----
-
+\ifjournal
 	\spnewtheorem{assump}{Assumption}{\bf}{\it}
@@
 	\renewenvironment{proof}{\emph{Proof}.}{\hfill\(\qed\)}
+\else
+	% Theorem definitions using amsthm
+
+	\usepackage{amsthm}
+
+	\renewcommand\thesubsection{\thesection.\Alph{subsection}}
+	\renewcommand{\theequation}{\thesection.\arabic{equation}}
+
+	\newtheorem{assump}{Assumption}
+	\renewcommand*{\theassump}{\Roman{assump}}
+	\newtheorem{thrm}[equation]{Theorem}
+
+	\theoremstyle{definition}
+	\newtheorem{prop}[equation]{Proposition}
+
+	\newenvironment{rationale}{\emph{Rationale}.}{\qed}
+	\newenvironment{justification}{\emph{Justification}.}{\qed}
+	\renewenvironment{proof}{\emph{Proof}.}{\qed}
+
+\fi
 
@@
 \author{Gabriele Carcassi, Christine A. Aidala, David J. Baker, Lydia Bieri}
-%\affiliation{University of Michigan, Ann Arbor, MI 48109}
-
+
+\ifjournal
 	\institute{G. Carcassi \at
@@
 	}
-
+\else
+	\affiliation{University of Michigan, Ann Arbor, MI 48109}
+\fi
+
+\ifjournal
 	% As the page margins are so large, the full title does not fit
 	\titlerunning{From physical principles to classical and quantum particle mechanics}
+\fi
 
@@
 
+\ifjournal
+    % Journal format has this before the abstract
 	\maketitle
+\fi
 
@@
 \end{abstract}
+
+\ifjournal
+\else
+	\maketitle
+\fi
 
@@
 We now proceed to characterize states and physical distinguishability in more precise terms so that we can capture their description mathematically.
-What we'll see is that the outcomes of all physical processes that can be used to gain information about the system, that can be used to perform a measurement, induce a topology on the set of states that is Hausdorff and second countable.
+What we'll see is that the outcomes of all physical processes that can be used to gain information about the system, that can be used to perform a measurement, induce a topology on the set of states that is at least Hausdorff.
 That is, physical distinguishability is mathematically captured by topological distinguishability.
@@
 The fact that two elements of the set can be distinguished requires the space to be Hausdorff.
-The fact that the potential outcomes of a process are countable requires the space to be second countable.
 
 \begin{prop}\label{prop:state_topology}
-The state space $\mathcal{S}$ of a physical system is a Hausdorff and second countable topological space.
+The state space $\mathcal{S}$ of a physical system is a Hausdorff topological space.
 \end{prop}
@@
 $\mathcal{S}$ is Hausdorff by definition.
-
-We claim that $\mathsf{T}$ is second countable.
-Consider the set of outcomes of a process.
-As each can provide only a finite amount of information (e.g.
-is identified by a finite number of digits), the set of outcomes is countable.
-This remains true when combining processes: the number of processes we can combine is finite as is the number of times we can prepare the same state and the set of intersections obtained from the countable outcomes of a finite number of processes is still countable.
-Consider a basis for the topology.
-Suppose it is of cardinality greater than countable.
-Then the basis would be able to distinguish among states in a way that no physical process ever could.
-The topology could distinguish among states that are not physically distinguishable.
-The basis for the topology cannot be of cardinality greater than countable.
-By contradiction, $\mathcal{S}$ must be second countable.
 \end{justification}
@@
 \begin{prop}\label{prop:topology}
-	Any set $\mathcal{S}$ of physically distinguishable elements is a Hausdorff and second countable topological space.
+	Any set $\mathcal{S}$ of physically distinguishable elements is a Hausdorff topological space.
 \end{prop}
@@
 discrete topology) while real numbers can be measured up to a finite precision (i.e.
-standard topology on $\mathbb{R}$) From now on, we are going to consider state spaces whose states can be identified, at least within a region, by a finite set of discrete and continuous state variables (i.e.
+standard topology on $\mathbb{R}$).
+From now on, we are going to consider state spaces whose states can be identified, at least within a region, by a finite set of discrete and continuous state variables (i.e.
 the state space is locally isomorphic to $\bigcup\limits_{1 \leq i \leq n} \mathbb{R}^{m_i}$).
@@
 $\hat{\mathcal{s}} \in \mathcal{T}_{\Delta t}(U)$ and $\mathcal{T}_{\Delta t}(U)$ is isomorphic to $\mathbb{R}^n$.
-$\mathcal{S}'$ equipped with the subspace topology is a Hausdorff, second countable topological space everywhere homeomorphic to $\mathbb{R}^n$.
+$\mathcal{S}'$ equipped with the subspace topology is a topological space everywhere homeomorphic to $\mathbb{R}^n$.
 $\mathcal{S}'$ is a manifold of dimension $n$.
@@
 We require $\mathcal{C}$ to include state changes.
-A change of a physically distinguishable object is itself physically distinguishable therefore $\mathcal{C}$ is still a Hausdorff and second countable topological manifold as per \ref{prop:topology}.
+A change of a physically distinguishable object is itself physically distinguishable therefore $\mathcal{C}$ is still a Hausdorff topological manifold as per \ref{prop:topology}.
 There exists an inverse $- : \mathcal{C} \rightarrow \mathcal{C}$ such that $\mathcal{c} + ( - \mathcal{c}) = 0 \forall \mathcal{c} \in \mathcal{C}$.

### 2017-01-12 2015CompleteDerivation/completeDerivation.tex
Added macro for conditional text based on style

@@
 \newcommand{\id}{\textrm{id}}
+
+\newcommand{\journal}[1]{\ifjournal#1\fi}
+\newcommand{\arxiv}[1]{\ifjournal\else#1\fi}
+
+\renewcommand{\descriptionlabel}[1]{%
+	\hspace\labelsep \upshape\bfseries #1:%
+}
 
@@
 
-\ifjournal
 % Journal format has this before the abstract
-	\maketitle
-\fi
+\journal{\maketitle}
 
@@
 
-\ifjournal
-\else
-	\maketitle
-\fi
+\arxiv{\maketitle}
 
@@
 \begin{description}
-  \item[Assumptions:] these characterize the physical system we are studying and constitute the premise of our discussion.
+  \item[Assumptions] these characterize the physical system we are studying and constitute the premise of our discussion.
 A \textbf{rationale} follows each assumption, which uses physical and sometimes philosophical arguments to motivate why (or not) such an assumption makes sense in a particular case.
 
-  \item[Propositions:] these capture the properties of our physical objects into mathematical language.
+  \item[Propositions] these capture the properties of our physical objects into mathematical language.
 A \textbf{justification} follows each proposition to show the necessity of such a characterization.
@@
 
-  \item[Theorems:] these are pure mathematical statements that are used to simplify propositions.
+  \item[Theorems] these are pure mathematical statements that are used to simplify propositions.
 A mathematical \textbf{proof} follows each theorem.
@@
 	\includestandalone[width=\columnwidth]{pictures/independentDof}
-	\caption{The metric $\omega$ and its components along two independent degrees of freedom.
+	\caption{The symplectic form $\omega$ and its components along two independent degrees of freedom.
 As we change state variables, the new d.o.f.
@@
 L = &\int (q-\mu_q)^2 \rho_{\mathcal{c}} \, dqdp \int (p-\mu_p)^2 \rho_{\mathcal{c}} \, dqdp \\
-&+ \lambda_1(\int \rho_{\mathcal{c}} dqdp - 1) + \lambda_2(- \int \rho_{\mathcal{c}} \ln \rho_{\mathcal{c}} \, dqdp - I_0)\\
-\delta L = &\int \delta \rho_{\mathcal{c}} [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 + \lambda_1 - \lambda_2 \ln \rho_{\mathcal{c}} - \lambda_2 ] dqdp = 0 \\
+\journal{ 
+	&+ \lambda_1(\int \rho_{\mathcal{c}} dqdp - 1) + \lambda_2(- \int \rho_{\mathcal{c}} \ln \rho_{\mathcal{c}} \, dqdp - I_0)\\ }
+\arxiv{ 
+	&+ \lambda_1(\int \rho_{\mathcal{c}} dqdp - 1) \\ &+ \lambda_2(- \int \rho_{\mathcal{c}} \ln \rho_{\mathcal{c}} \, dqdp - I_0)\\ }
+\delta L = &\int \delta \rho_{\mathcal{c}} [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 + \\ &\lambda_1 - \lambda_2 \ln \rho_{\mathcal{c}} - \lambda_2 ] dqdp = 0 \\
 \lambda_2 \ln \rho_{\mathcal{c}} = &\lambda_1 - \lambda_2 + (q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\
@@
 	&= (v^{a} + \partial_{c} S^{a} v^{c} dt) \omega_{ab} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
-	&= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} + v^{a} \omega_{ab} \partial_{d} S^{b} w^{d}) dt + O(dt^2)
+	&= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} + v^{a} \omega_{ab} \partial_{d} S^{b} w^{d}) dt \\ &+ O(dt^2)
 	\end{align*}

### 2017-01-13 2015CompleteDerivation/completeDerivation.tex
Adding comments from Lydia

@@
 The answer is no with the standard topology on $\mathbb{R}^n$.
-The consequence is that when we define the measure $\mu$ for the cardinality of continuous states, we assign a finite value to a finite region, and we assign zero measure to a single state.
+The consequence is that when we define the measure $\mu$ for the count of continuous states, we assign a finite value to a finite region, and we assign zero measure to a single state.
 That is what gave us integration, densities and, ultimately, conjugate variables.
@@
 This is the manifold that allows us to define time-and-state-variable-invariant densities.
-The symplectic form $\omega$ allows us to measure the cardinality of possibilities as in the time independent case.
+The symplectic form $\omega$ allows us to measure the count of possibilities as in the time independent case.
 
@@
 This is a contradiction: $\mathcal{s}$ is not identified by just $q^\alpha$.
-$u^\alpha=d_s x^\alpha$ is part of the state.
+Therefore $u^\alpha=d_s x^\alpha$ is part of the state.
 $p_\alpha = p_\alpha(x^\alpha, u^\alpha)$.
@@
 For example, once the SI is adopted, $1$km would correspond to a thousand times more possible states than $1$m, and, in the same way, $1$sec would correspond to $299,792,458$ more possible states than $1$m.
-Therefore when we write $x^\alpha = (ct, x^i)$ or $p_\alpha = (- E/c, p_i)$ we are really expressing everything in quantities that are proportional to the cardinality of states and possibilities.
+Therefore when we write $x^\alpha = (ct, x^i)$ or $p_\alpha = (- E/c, p_i)$ we are really expressing everything in quantities that are proportional to the count of states and possibilities.
 These coordinates make the bookkeeping of number of states much easier which, as we saw, is the only thing that Hamiltonian and Lagrangian mechanics are doing.
@@
 It was under that premise that we were able to show that the trajectories are continuous, that we could define $\omega$ and therefore $g$.
-Non-local quantum effects, such as the ones featured in EPR\cite{EPR} experiments, are not examples of deterministic and reversible motion.
-Therefore they do not have anything to violate: the maximum speed does not apply to them.
-On the other hand, only deterministic and reversible motion can carry information and can establish before/after relationships, therefore everything remains consistent.
+Non-local quantum effects, such as the ones featured in EPR\cite{EPR} experiments, are not examples of deterministic and reversible evolution.
+Therefore they do not have anything to violate: the maximum speed does not apply to their ``motion".\footnote{The very notion of velocity becomes ill defined in such cases precisely because we do not have a unique well-defined evolution.} On the other hand, only deterministic and reversible motion can carry information and can establish before/after relationships, therefore everything remains consistent.
 

### 2017-01-13 2015CompleteDerivation/completeDerivation.tex
Reviewing the abstract

@@
 The aim of this work is to show that particle mechanics, both classical and quantum, Hamiltonian and Lagrangian, can be derived from few simple physical assumptions.
-
-Assuming deterministic and reversible time evolution will give us a dynamical system, whose set of states form a topological space and whose law of evolution is a self-homeomorphism.
-Assuming the system is infinitesimally reducible, giving the state and the dynamics of the whole system is equivalent to giving the state and the dynamics of its infinitesimal parts, will give us a classical Hamiltonian system.
-Assuming the system is irreducible, giving the state and the dynamics of the whole system tells us nothing about the state and the dynamics of its substructure, will give us a quantum Hamiltonian system.
-Assuming kinematic equivalence, studying trajectories is equivalent to studying state evolution, will give us Lagrangian mechanics together and limit the form of the Hamiltonian/Lagrangian to the one with scalar and vector potential forces.
-
-This work strives as much as possible to be philosophically consistent, physically meaningful and mathematically precise.
+Assuming deterministic and reversible time evolution will give us a dynamical system whose set of states form a topological space and whose law of evolution is a self-homeomorphism.
+Assuming the system is infinitesimally reducible, that giving the state and the dynamics of the whole system is equivalent to giving the state and the dynamics of its infinitesimal parts, will give us a classical Hamiltonian system.
+Assuming the system is irreducible, that giving the state and the dynamics of the whole system tells us nothing about the state and the dynamics of its substructure, will give us a quantum Hamiltonian system.
+Assuming kinematic equivalence, that studying trajectories is equivalent to studying state evolution, will give us Lagrangian mechanics and limit the form of the Hamiltonian/Lagrangian to the one with scalar and vector potential forces.
 \end{abstract}

### 2017-01-19 2015CompleteDerivation/completeDerivation.tex
Added acknowledgement section fixed appendix format

@@
 
-\section{Appendix: mathematical proofs}
+\begin{acknowledgements}
+We'd like to thank Erik Curiel, Jiahua Gu, Josh Hunt, Isaac Mooney and Pavel Okun for providing feedback on drafts of this manuscript, Mark Greenfield for further insights in the topological aspects and Alan Phillips for contributing figures.
+We additionally thank Paul Stankus for interest and support in the earlier stages of this project.
+Funding for this work was provided in part by the MCubed program of the University of Michigan.
+\end{acknowledgements}
+
+\appendix
+\section*{Appendix A. Mathematical proofs}
+\stepcounter{section}
 

### 2017-01-24 2015CompleteDerivation/completeDerivation.tex
Adding ref places and other minor fixes from Christine

@@
 \journal{\maketitle}
+
+%TODO: Check hyper-surface or hypersurface
 
@@
 
-We can also conceptually combine two different processes into a single one.\footnote{That is, take the logical and between outcomes of different processes.} That is, having a way to measure quantity $x$ and a way to measure quantity $y$ gives us a way to measure the combination $(x,y)$.
+We can also conceptually combine two different processes into a single one.\footnote{That is, take the logical \textsc{and} between outcomes of different processes.} That is, having a way to measure quantity $x$ and a way to measure quantity $y$ gives us a way to measure the combination $(x,y)$.
 For ensembles, if we can measure the marginal distribution $\rho_x(x)$ and the marginal distribution $\rho_y(y)$, we know the joint distribution $\rho(x,y)$ has to be compatible with both.
@@
 
-We can also coarsen a single process.\footnote{That is, take the logical or between outcomes of the same process.} That is, having a way to measure $(x,y)$ gives us a way to measure $x$ alone (or any $f(x,y)$).
+We can also coarsen a single process.\footnote{That is, take the logical \textsc{or} between outcomes of the same process.} That is, having a way to measure $(x,y)$ gives us a way to measure $x$ alone (or any $f(x,y)$).
 Formally, the outcomes of the coarsened process are given by performing the union of some outcomes of the original process.
@@
 $\mathcal{S} \in \mathsf{T}$ and is associated with the outcome ``the system exists".
-$\varnothing \in \mathsf{T}$ and is associated with the outcome ``the system doesn't exists".
+$\varnothing \in \mathsf{T}$ and is associated with the outcome ``the system doesn't exist".
 Let $V_1, V_2 \in \mathsf{T}$.
 Then, by definition, there exists a process $P_1$ that admits $V_1$ as an outcome and a process $P_2$ that admits $V_2$ as an outcome.
-Consider the process $P$ that combines the outcomes of $P_1$ and $P_2$ with a logical and.
+Consider the process $P$ that combines the outcomes of $P_1$ and $P_2$ with a logical \textsc{and}.
 This always exists physically as we can prepare the same state multiple times and let it interact with each process separately.
@@
 If they are physically distinguishable, then there exists a physical process $P$ that admits both as outcomes.
-Given $P$, we can always construct the process $P_0$ that combines $V_1, V_2$ with a logical or into a single outcome $V$, by ``forgetting" which of the two was given.
+Given $P$, we can always construct the process $P_0$ that combines $V_1, V_2$ with a logical \textsc{or} into a single outcome $V$, by ``forgetting" which of the two was given.
 The states compatible with $V$ must be in either $V_1$ or $V_2$, that is $V = V_1 \cup V_2$.
@@
 States, though, are not the only objects with that property.
-In fact: any element of a set of physical objects (e.g.
+In fact, any element of a set of physical objects (e.g.
 forces, physical properties such as mass or charge, time) needs to be physically distinguishable to be well defined.
@@
 an open set) to finite precision knowledge in the horizontal axis.
-A discontinuous function may map to finite precision (i.e.
+A discontinuous function may map to infinite precision (i.e.
 a closed set).
-If finite precision identification is ruled out (i.e.
+If infinite precision identification is ruled out (i.e.
 use of standard topology) then discontinuous functions must also be ruled out.}
@@
 
-While functions with few discontinuities are useful and used in physics and engineering, they are employed for idealized cases  (e.g.~a signal change is fast enough, a charge distribution is small enough) that are often treated as a special case (e.g.~propagation in material discontinuities).
+While functions with few discontinuities are useful and used in physics and engineering, they are employed for idealized cases  (e.g.~a signal change is fast enough, a charge distribution is small enough) that are often treated as a special case (e.g.~propagation across material discontinuities).
 While this may be obvious and intuitive to the physicist, it may be troubling to the mathematician as the proper use of many mathematical techniques requires the inclusion of discontinuous functions.
@@
 conjugate momentum in a gauge theory).
-The only requirements for state variables is that they identify states.
+The only requirement for state variables is that they identify states.
 This means the possible values for state variables are physically distinguishable and are defined at equal time (since states are defined and mapped at a particular time).\footnote{This is the main reason that a quantity like velocity is not a suitable state variable, as it is  defined over an interval of time.
@@
 One reason that transported variables will be useful is that, during deterministic and reversible evolution, we need to make sure that the number of initial and final possibilities across independent state variables remains the same.
-This is more easily done using transported variable as the actual values do not change.
+This is more easily done using transported variables as the actual values do not change.
 In the discrete case, this is easy to show.
@@
 We'll see that very similar relationships are what define Hamiltonian and Lagrangian mechanics.
-But recovering them in the continuous case is not as straight forward.
+But recovering them in the continuous case is not as straightforward.
 Consider the map $q'=aq$ with $0<a<1$.
@@
 
-What makes the matter even more confusing is that the single physical idea of counting states and possibilities along independent state variable does not map to a single mathematical concept.
+What makes the matter even more confusing is that the single physical idea of counting states and possibilities along independent state variables does not map to a single mathematical concept.
 Sometimes we need a measure, sometimes we need a symplectic form and sometimes we need a metric tensor.
@@
 The notion that the system is homogeneous means that the states of the composite and of each part are not unrelated: they are all made of the same material.\footnote{Whether a system is homogeneous depends on context (e.g.
-air can be thought as homogeneous if the mixture of gases does not change in space or in time due to phase transitions or chemical processes) and one must check that such property is maintained by time evolution.} In fact, the state space $\mathcal{C}$ of all systems composed of such material will include the state of the system as well as the states of  its parts (i.e.
+air can be thought as homogeneous if the mixture of gases does not change in space or in time due to phase transitions or chemical processes) and one must check that such a property is maintained by time evolution.} In fact, the state space $\mathcal{C}$ of all systems composed of such material will include the state of the system as well as the states of  its parts (i.e.
 $\mathcal{c}, \mathcal{c}_1, \mathcal{c}_2 \in \mathcal{C}$).
@@
 And the density will give us an indication of how big the region is: if we double the region, the density will halve.
-In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique measure\footnote{Here measure is intended in the mathematical sense (a real valued function of a $\sigma$-algebra) which is distinct from other connotations in physics.} $\mu$ that allows to count the number of states in each set of $\mathcal{S}$.
+In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique measure\footnote{Here measure is intended in the mathematical sense (a real valued function of a $\sigma$-algebra) which is distinct from other connotations in physics.} $\mu$ that allows us to count the number of states in each set of $\mathcal{S}$.
 In terms of said measure, positive linear functionals become integrals, $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
 
-This intuitive picture is formalized mathematically by the Riesz representation theorem for linear functionals, which gives $\mathcal{S}$ the structure of a measure space, with a suitable Borel $\sigma$-algebra and measure $\mu$.
+This intuitive picture is formalized mathematically by the Riesz representation theorem for linear functionals(ref?), which gives $\mathcal{S}$ the structure of a measure space, with a suitable Borel $\sigma$-algebra and measure $\mu$.
 
@@
 
-In other words: this is not a case where a difference at  small scales has a small difference at large scales.
+In other words: this is not a case where a difference at small scales leads to a small difference at large scales.
 The two cases are radically different.
@@
 The quantum case does not change this, as the use of densities (in the form of the wave function) and conjugate quantities is even more prominent.\footnote{If one wants to posit that space is fundamentally discrete at scales where both classical and quantum mechanics cease to be valid, the better strategy would be to work with a discrete version of $T^*\mathcal{Q}$ and not $\mathcal{Q}$ by itself.
-Though it is not clear how a discrete topology would become non discrete in the limit, at least the measure should be workable.}
+Though it is not clear how a discrete topology would become non-discrete in the limit, at least the measure may be workable.}
 
@@
 	Of those three requirements we can only pick two.
-If we want determinisms and reversibility for the combined system, either both systems are not independently deterministic or the second is not a measuring device.
+If we want determinism and reversibility for the combined system, either both systems are not independently deterministic or the second is not a measuring device.
 In other words: we can't expect to have deterministic and reversible evolution at all levels of aggregation and have the parts interact in any physically meaningful way.
@@
 But ignoring that issue, the bigger problem is that we can always locally separate the evolution into independent degrees of freedom.
-For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value; they clearly evolve independently.
+For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (ref?)} We can always find such a local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value; they clearly evolve independently.
 And since the pieces are infinitesimal, a local decomposition is all that is needed.
@@
 	As a final note: we caution against automatically thinking that the classical assumption is valid for all macroscopic systems.
-It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, in which case the assumption would not hold.
+It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, in which case the assumption would not hold.(ref?)
 \end{rationale}
@@
 A system governed by Hamilton's equations is deterministic and reversible: the equation of motion given by $H$ are differentiable since $H$ is twice differentiable.
-Therefore the equations are Lipschitz continuous and, by the Picard-Lindel\"{o}f theorem, a unique solution exists.
+Therefore the equations are Lipschitz continuous and, by the Picard-Lindel\"{o}f theorem(ref?), a unique solution exists.
 
@@
 Note that we could have taken different approaches.
-For example, we could have appealed to information theory and required our invariant distributions to preserve Shannon's information entropy, as no information should be gained or lost during a deterministic and reversible process.
+For example, we could have appealed to information theory and required our invariant distributions to preserve Shannon's information entropy\cite{Shannon}, as no information should be gained or lost during a deterministic and reversible process.
 Or we could have appealed to statistical mechanics and required that the determinant of the covariance matrix be conserved, as a deterministic and reversible process should be defined at the same level of precision.
@@
 This means that our measure $\mu$ needs to be modified to define integrals over the time variable.\footnote{Looking ahead, another way to frame the problem is that Galilean transformations do not preserve phase space volume.
-Lorentz transformations, on the other hand, do\cite{Treumann} and it is this requirements plus the kinematic assumption that will recover them later.}
+Lorentz transformations, on the other hand, do\cite{Treumann} and it is this requirement plus the kinematic assumption that will recover them later.}
 
@@
 Therefore we cannot simply stick time in the Hamiltonian and expect everything to work out.
-But we can't ignore the problem either as the three cases outlined are pretty common situations to study.
+But we can't ignore the problem either as the three cases outlined are common situations to study.
 So we need to go back to our definitions and amend them properly.
@@
 This, again, highlights our previous oversight: we didn't include time within the quantities needed to specify a physical configuration at a particular time.
-We call $\mathcal{M}$ such space which is a manifold of dimension $n+1$: state variables and the time variable are independent (i.e.
+We call $\mathcal{M}$ such a space which is a manifold of dimension $n+1$: state variables and the time variable are independent (i.e.
 any combination of state and time variables is valid).
@@
 On the other hand, the change of state variables must preserve the ability to uniquely identify states at each instant in time.
-In this sense, time is not just like another state variable.\footnote{In relativity, for example, it is common to use foliations and the $3+1$ formalism to recover the special character of time when performing calculations or to gain better physical insight.\cite{Gourgoulhon}}
+In this sense, time is not just like another state variable.\footnote{In relativity, for example, it is common to use foliations and the $3+1$ formalism to recover the special character of time when performing calculations or to gain better physical insight\cite{Gourgoulhon}.}
 
@@
 
-The symplectic form $\omega$ is still the canonical two form expressed with the mathematically non-canonical, but physically meaningful, variable $E$.
+The symplectic form $\omega$ is still the canonical two-form expressed with the mathematically non-canonical, but physically meaningful, variable $E$.
 This allows us to better understand its physical meaning.
@@
 The Euclidean signature across standard d.o.f.
-allows to move ranges of possibilities from one d.o.f.
+allows us to move ranges of possibilities from one d.o.f.
 to the other while conserving the total: the generators of rotation are in fact divergence free.
@@
 The trajectory of a particle in the extended phase space will be given by the evolved variables $\xi^a(s)$ in terms of a parameter $s$.
-As before, deterministic and reversible evolution will preserve $\omega$ as the possibility count on each independent d.o.f.
+As before, deterministic and reversible evolution will preserve $\omega$, as the possibility count on each independent d.o.f.
 will be conserved.
@@
 
-We recognize the equations as those for Hamiltonian mechanics on the extended phase space.\cite{Synge,Lanczos,Struckmeier}
+We recognize the equations as those for Hamiltonian mechanics on the extended phase space\cite{Synge,Lanczos,Struckmeier}.
 
@@
 The distribution of material would simply be a distribution over the particle evolutions, a Lebesgue integrable differentiable function.
-Yet, the need to be able to describe the evolution in terms of physically meaningful state variables will most likely re-impose the extra structure.
+Yet, the need to be able to describe the evolution in terms of physically meaningful state variables will most likely reimpose the extra structure.
 
@@
 The link between kinematics, the study of motion through quantities like velocity and acceleration, and dynamics, the study of state evolution and its causes, is crucial in physics.
-In Newtonian physics, $\vec{F}=m\vec{a}$ provides such link.
-In Hamiltonian (and Lagrangian) mechanics, the link is implicitly given by the choice of the Hamiltonian (or Lagrangian).(ref?)
+In Newtonian physics, $\vec{F}=m\vec{a}$ provides such a link.
+In Hamiltonian (and Lagrangian) mechanics, the link is implicitly given by the choice of the Hamiltonian (or Lagrangian).
 
@@
 This already tells us that the trajectory is differentiable since $q^\alpha(s)$ is differentiable.
-The four velocity $u^\alpha = d_s x^\alpha$ is therefore well defined and it can be used to identify trajectories.
-The magnitude of the four velocity can be rescaled by changing the evolution parameter, $u^\alpha = d_s x^\alpha = d_s \hat{s} \, d_{\hat{s}} x^\alpha = d_s \hat{s} \, \hat{u}^\alpha$, and is therefore not physical.
-Only the direction is physical so the four velocity gives us another $n$ independent variables to identify trajectories.
+The four-velocity $u^\alpha = d_s x^\alpha$ is therefore well defined and it can be used to identify trajectories.
+The magnitude of the four-velocity can be rescaled by changing the evolution parameter, $u^\alpha = d_s x^\alpha = d_s \hat{s} \, d_{\hat{s}} x^\alpha = d_s \hat{s} \, \hat{u}^\alpha$, and is therefore not physical.
+Only the direction is physical so the four-velocity gives us another $n$ independent variables to identify trajectories.
 The combined position and velocity, the kinematic variables, identify $2n + 1$ trajectories using a hyper-surface of the tangent bundle $T\mathcal{M}$.
@@
 Integrating we have $\mathcal{H}=\frac{1}{2m}(p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta)+V(x)$ where $V(x)$ is an arbitrary function.
-We can further constrain the Hamiltonian by choosing the natural parameterization for $s$ (also called unit speed or arc length parametrization) for which the modulus of the velocity is constant.
+We can further constrain the Hamiltonian by choosing the natural parameterization for $s$ (also called unit speed or arc length parametrization (ref?)) for which the modulus of the velocity is constant.
 We have $u^\alpha u_\alpha = - c^2 = \frac{1}{m^2}(p^\alpha-A^\alpha)(p_\alpha-A_\alpha)$.
@@
 The motion of the photon is not enough to tell us its energy or its momentum.
-Therefore it makes sense that, by assuming kinematic equivalence we derived equation valid only for massive particles.
-
-We have seen that the inertial mass $m$, the invariant speed $c$ and the reduced Plank constant $\hbar$ all serve a similar purposes: convert one quantity to another preserving the number of possibilities.
+Therefore it makes sense that, by assuming kinematic equivalence we derived an equation valid only for massive particles.
+
+We have seen that the inertial mass $m$, the invariant speed $c$ and the reduced Plank constant $\hbar$ all serve similar purposes: convert one quantity to another preserving the number of possibilities.
 $dp=\hbar dk$ converts between ranges of wave number and conjugate momentum.
@@
 $dp = m du$ converts between ranges of velocity and conjugate momentum.
-They allow to keep proper book keeping when mapping sets of states.
+They allow us to keep proper bookkeeping when mapping sets of states.
 Using natural units means measuring all these quantities through the possibility count.
@@
 We should also ask whether any choice of $A$ actually represents a different case leading to different trajectories.
-The simplest and more insightful way to study this problem is to use the kinematic quantities $x^\alpha$ and $mu_\alpha$ as state variable.
+The simplest and more insightful way to study this problem is to use the kinematic quantities $x^\alpha$ and $mu_\alpha$ as state variables.
 They are not going to be conjugates of each other but they do capture the entire physical description.
@@
 In other words: $A$ and $p$ are defined relatively to each other.
-$A$ defines the $0$ for $p$ and vice versa.
-The gauge group expresses the arbitrariness of such definition.
+$A$ defines the $0$ for $p$ and vice-versa.
+The gauge group expresses the arbitrariness of such a definition.
 
@@
 In this setting, charge conjugation $C$ is simply changing the parametrization from $s$ to $-s$.
-This will permute particle standard and anti-state as they follow the trajectories in the opposite direction.
+This will permute particle standard and anti-states as they follow the trajectories in the opposite direction.
 But if we apply parity $P$ and time reversal $T$, changing $x^\alpha$ to $-x^\alpha$, we would have the same net effect: the parametrization stays the same, but it now moves backwards in space and time.
 As they have the same net effect, if we apply all three we go back where we started.
-This gives us a very intuitive classical particle mechanics analogue for the $CPT$ theorem.
+This gives us a very intuitive classical particle mechanics analogue for the $CPT$ theorem(ref?).
 
@@
 
-To sum up, the kinematic assumption has more inherent consequences one would expect at first.
+To sum up, the kinematic assumption has more inherent consequences than one would expect at first.
 It means the particles are identified by position and velocity and therefore the system of equations is second order.
 It links the space of initial conditions on $T\mathcal{M}$ with the space of states on $T^*\mathcal{M}$.
-Such link allows phase space symplectic form $\omega$ to induce the space-time metric tensor $g$.
+Such a link allows phase space symplectic form $\omega$ to induce the space-time metric tensor $g$.
 It allows us to define Lagrangians.
 It constrains the motion to massive particles under scalar and vector potential forces.
-Moreover, it allowed to re-understand basic concepts, like the invariant speed $c$, the inertial mass $m$ and the force tensor $F_{\alpha\beta}$ in terms of state counting and d.o.f.
+Moreover, it allowed us to reunderstand basic concepts, like the invariant speed $c$, the inertial mass $m$ and the force tensor $F_{\alpha\beta}$ in terms of state counting and d.o.f.
 independence.
-Finally, it allowed to construct classical equivalents for anti-particles and the CPT theorem.
+Finally, it allowed us to construct classical equivalents for anti-particles and the $CPT$ theorem.
 
@@
 \end{align*}
-This is strikingly similar to the Heisenberg uncertainty principle, except that the value depends on the initial state $I_0$.
-
-Suppose, though, that while the evolution of the distribution of our electron as a whole is deterministc and reversible, each fragment is undergoing non deterministic motion.
+This is strikingly similar to the Heisenberg uncertainty principle, except that the value depends on the initial entropy $I_0$.
+
+Suppose, though, that while the evolution of the distribution of our electron as a whole is deterministc and reversible, each fragment is undergoing non-deterministic motion.
 Suppose $\rho$ is a distribution at equilibrium of such motion.
-While each fragment moves randomly, the overall distribution $\rho$ stays the same: the non deterministic motion is effectively just reshuffling the fragments within the same distribution.
-The value of the entropy of the distribution, and therefore its spread, is set by the non deterministic motion of the particles: the stronger it is, the greater the entropy.
+While each fragment moves randomly, the overall distribution $\rho$ stays the same: the non-deterministic motion is effectively just reshuffling the fragments within the same distribution.
+The value of the entropy of the distribution, and therefore its spread, is set by the non-deterministic motion of the particles: the stronger it is, the greater the entropy.
 
@@
 This discussion shows, once again, how we can already find in classical mechanics some of the ingredients that are typically assumed to be characteristics of quantum mechanics.
-It has its own flavor of anti-particles, CPT theorem and uncertainty principle.
+It has its own flavor of anti-particles, $CPT$ theorem and uncertainty principle.
 These should really be considered general features of deterministic and reversible evolution of distributions (i.e.
@@
 We will call a \emph{quantized material} a homogeneous material that is finitely reducible.
-That is, as we decompose the system we can only assign a state to finite amounts of material, which we call \emph{quantum particle}.
+That is, as we decompose the system we can only assign a state to finite amounts of material, which we call \emph{quantum particles}.
 As the material is homogeneous, all its particles carry the same amount of material and are described by the same state space.
@@
 
-The second feature is non locality.
+The second feature is non-locality.
 Suppose we have a quantum particle distributed in space.
@@
 Suppose a quantum particle spontaneously decays into two or more particles.
-Such process cannot be deterministic as the number of d.o.f.
+Such a process cannot be deterministic as the number of d.o.f.
 increases.
@@
 \end{equation}
-where $\sigma^2$ is the variance, $\sigma$ the standard deviation, $\rho_{X,Y}= COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient.
+where $\sigma^2$ is the variance, $\sigma$ the standard deviation, $\rho_{X,Y}= COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient(ref?).
 As $-1\leq\rho_{X,Y}\leq+1$, the total variance can be greater or smaller than the sum of the variance of the components.
@@
 For a quantum mixed state $\tilde{\rho} = \sum \tilde{\rho}_i |\psi_i\rangle \langle \psi_i |$, we have $I(\tilde{\rho}) = - \sum \tilde{\rho}_i \ln \tilde{\rho}_i - \tilde{\rho}_i I(\psi_i) = - \sum \tilde{\rho}_i \ln \tilde{\rho}_i = -tr(\tilde{\rho} \ln \tilde{\rho})$.
-And in fact this is the Von Neumann entropy used in quantum information, that implicitly assumes all states provide the same amount of entropy (i.e.
+And in fact this is the Von Neumann entropy (ref?) used in quantum information, that implicitly assumes all states provide the same amount of entropy (i.e.
 zero).
@@
 
-Naturally, we cannot give a complete description for the configuration of a fragment, therefore we cannot give a joint distribution for all fragment variables at the same time.\footnote{This recovers the notion of complementarity introduced by Bohr.} We can, though, find a maximal set of fragment variables: one that can be used to provide the best description possible.
+Naturally, we cannot give a complete description for the configuration of a fragment, therefore we cannot give a joint distribution for all fragment variables at the same time.\footnote{This recovers the notion of complementarity introduced by Bohr(ref?).} We can, though, find a maximal set of fragment variables: one that can be used to provide the best description possible.
 While there isn't a unique maximal set, we can show that one such set is provided by the fragment variables $q^i : \mathcal{Q} \rightarrow \mathbb{R}$ that define the units required to describe the system.
@@
 In the following sections we will derive two types of state evolution: the deterministic and reversible one, corresponding to unitary evolution (i.e.
-Schroedinger's equation), and a non deterministic one useful to describe measurement interactions, corresponding to the projection postulate.
+Schroedinger's equation), and a non-deterministic one useful to describe measurement interactions, corresponding to the projection postulate.
 Therefore, as the infinitesimal reducibility assumption leads to classical mechanics, the irreducibility assumption leads to quantum mechanics.
@@
 This motivates why it is such an important operator.
-As for the Klein-Gordon equation, the only thing it does is to impose that the norm of the four velocity is the speed of light, a condition much more trivial than one may have anticipated.
+As for the Klein-Gordon equation, the only thing it does is to impose that the norm of the four-velocity is the speed of light, a condition much more trivial than one may have anticipated.
 
@@
 
-In this section we include mathematical demonstration in support of the physical justifications of the paper.
+In this section we include mathematical demonstrations in support of the physical justifications of the paper.
 

### 2017-01-25 2015CompleteDerivation/completeDerivation.tex
Adding Christine comments

@@
 \journal{\maketitle}
-
-%TODO: Check hyper-surface or hypersurface
 
@@
 Finding new sets of states with new laws of evolution is, in fact, what leads to new physics.
-We will therefore call \emph{fundamental model of physics} the triad of state, unstated part and environment, with the assumption that the state may undergo deterministic and reversible evolution, and the unstated part undergoes non-deterministic non-reversible evolution that does not alter the set of states.
+We therefore call \emph{fundamental model of physics} the triad of state, unstated part and environment, with the assumption that the state may undergo deterministic and reversible evolution, and the unstated part undergoes non-deterministic non-reversible evolution that does not alter the set of states.
 \end{rationale}
@@
 So we introduce a new set of variables for which $\hat{q}^i(\hat{\mathcal{s}})=q^i(\mathcal{s})$, which we can always do as the evolution is deterministic and reversible.
-The hyper-surfaces at constant $\hat{q}^i$ are the evolved hyper-surfaces at constant of $q^i$ (e.g.
+The hypersurfaces at constant $\hat{q}^i$ are the evolved hypersurfaces at constant of $q^i$ (e.g.
 all the states that started with a particular value for pressure) which allows us to study how groups of states evolve in time.
@@
 Let $\rho_\mathcal{c}$ be strictly monotonic in a region $U\subseteq \mathcal{S}$.
-Then its fibers in $U$, the inverse images of the values, are connected hyper-surfaces.
+Then its fibers in $U$, the inverse images of the values, are connected hypersurfaces.
 We can construct a chart such that $q^1=\rho_\mathcal{c}$.
@@
 In that space, a state $\mathcal{s} \in \bar{\mathcal{S}}$ will be given by $(q^i, p_i, t)$, making $\bar{\mathcal{S}}$ a $2n+1$ dimensional manifold.
-A region at equal time will be a hyper-surface $\mathcal{S}_{t=t_0} \subset \bar{\mathcal{S}}$ that includes all possible configurations at a particular time.
-Such hyper-surface will need to cut across all evolutions once and only once, therefore not all functions $t : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ are suitable time variables.
+A region at equal time will be a hypersurface $\mathcal{S}_{t=t_0} \subset \bar{\mathcal{S}}$ that includes all possible configurations at a particular time.
+Such hypersurface will need to cut across all evolutions once and only once, therefore not all functions $t : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ are suitable time variables.
 A time evolution is a line  $\bar{\mathcal{s}} \subset \bar{\mathcal{S}}$ that passes once through each equal time surfaces.
@@
 	The complete state space $\bar{\mathcal{S}}$ for the particles of a classical material is a $2n+1$ dimensional differentiable manifold.
-The state space $\mathcal{S}_{t=t_0}$ at a particular time is a hyper-surface of $\bar{\mathcal{S}}$.
+The state space $\mathcal{S}_{t=t_0}$ at a particular time is a hypersurface of $\bar{\mathcal{S}}$.
 \end{prop}
@@
 
-	We claim $\mathcal{S}_{t=t_0}$ is a hyper-surface of $\bar{\mathcal{S}}$.
+	We claim $\mathcal{S}_{t=t_0}$ is a hypersurface of $\bar{\mathcal{S}}$.
 Let $t : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ be a time variable.
@@
 $\mathcal{S}_{t=t_0}$ is an embedded submanifold of co-dimension $1$.
-$\mathcal{S}_{t=t_0}$ is a hyper-surface.
+$\mathcal{S}_{t=t_0}$ is a hypersurface.
 \end{justification}
@@
 To each state evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$ will correspond a distribution $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ over all particle states at all times.
-This distribution will be differentiable: it's differentiable over all equal time hyper-surfaces and time evolution is differentiable.
+This distribution will be differentiable: it's differentiable over all equal time hypersurfaces and time evolution is differentiable.
 The distribution, though, is not integrable over the complete state space $\bar{\mathcal{S}}$.
@@
 If time is taken to be infinite the value is not finite.
-The distribution, instead, must be integrable over all equal time hyper-surfaces: no matter what time variable we choose, the amount of material found at a particular time has to be finite.
+The distribution, instead, must be integrable over all equal time hypersurfaces: no matter what time variable we choose, the amount of material found at a particular time has to be finite.
 
 \begin{prop}\label{prop:state_evolution_space}
-	The state evolution space $\bar{\mathcal{C}}$ for a classical material is isomorphic to the space of differentiable functions that are Lebesgue integrable on any equal time hyper-surface.
+	The state evolution space $\bar{\mathcal{C}}$ for a classical material is isomorphic to the space of differentiable functions that are Lebesgue integrable on any equal time hypersurface.
 That is $\bar{\mathcal{C}} \cong  C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$ where $L^1_t(\bar{\mathcal{S}}, \mu)=\{{\rho: \bar{\mathcal{S}} \rightarrow \mathbb{R}} \; | \; {\int_{\mathcal{S}_{t=t_0}} |\rho| d\mu < \infty} \; \forall \, t, t_0 \}$.
@@
 To identify the full evolution, then, we just need to specify the distribution only on a region that cuts across all particle evolutions, i.e.
-a hyper-surface at constant time.
+a hypersurface at constant time.
 
@@
 Therefore there must exist a constraint such that, locally, $E=E(t,q^i,p_i)$.
-The complete state space $\bar{\mathcal{S}}$ is a hyper-surface of $T^*\mathcal{M}$.
+The complete state space $\bar{\mathcal{S}}$ is a hypersurface of $T^*\mathcal{M}$.
 
@@
 \begin{prop}\label{prop:relativistic_symplectic_manifold}
-	The complete state space $\bar{\mathcal{S}}$ for the particles of a classical material is a hyper-surface of the symplectic manifold formed by a cotangent bundle $T^*\mathcal{M}$ equipped with the canonical two-form $\omega = \sum dq^i \wedge dp_i - dt \wedge dE = \sum dq^\alpha \wedge dp_\alpha$.
+	The complete state space $\bar{\mathcal{S}}$ for the particles of a classical material is a hypersurface of the symplectic manifold formed by a cotangent bundle $T^*\mathcal{M}$ equipped with the canonical two-form $\omega = \sum dq^i \wedge dp_i - dt \wedge dE = \sum dq^\alpha \wedge dp_\alpha$.
 \end{prop}
@@
 
-	We claim $\bar{\mathcal{S}}$ is a hyper-surface of $(T^*\mathcal{M}, \omega)$.
+	We claim $\bar{\mathcal{S}}$ is a hypersurface of $(T^*\mathcal{M}, \omega)$.
 $\bar{\mathcal{S}} \subseteq (T^*\mathcal{M}, \omega)$ as states are physical objects that allow time-and-state-variable-invariant densities.
@@
 As a final thought, we note how the physics maps less elegantly to the math in the time dependent case.
-The function space for $\bar{\mathcal{C}}$ is not simply the Lebesgue integrable differentiable functions, but those that are integrable on particular hyper-surfaces.
-The complete state space $\bar{\mathcal{S}}$ is a hyper-surface of $T^*\mathcal{M}$.
+The function space for $\bar{\mathcal{C}}$ is not simply the Lebesgue integrable differentiable functions, but those that are integrable on particular hypersurfaces.
+The complete state space $\bar{\mathcal{S}}$ is a hypersurface of $T^*\mathcal{M}$.
 The invariant Hamiltonian is needed to identify that subspace.
@@
 Therefore $\mathcal{M}$ has conceptually two roles: it is the space where the motion unfolds and it is the space that defines the system of units for our states.
-This already tells us that the state space for the particles of the material is a hyper-surface of $T^*\mathcal{M}$.
+This already tells us that the state space for the particles of the material is a hypersurface of $T^*\mathcal{M}$.
 
@@
 Only the direction is physical so the four-velocity gives us another $n$ independent variables to identify trajectories.
-The combined position and velocity, the kinematic variables, identify $2n + 1$ trajectories using a hyper-surface of the tangent bundle $T\mathcal{M}$.
+The combined position and velocity, the kinematic variables, identify $2n + 1$ trajectories using a hypersurface of the tangent bundle $T\mathcal{M}$.
 This is exactly the dimensionality of the complete state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$.
@@
 
-Mathematically, the state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is diffeomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$ of the space-time manifold.
+Mathematically, the state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is diffeomorphic to a hypersurface of the tangent bundle $T\mathcal{M}$ of the space-time manifold.
 
@@
 	Let $\mathcal{M}$ be the $n+1$ dimensional space-time manifold over which trajectories are defined.
-The complete state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ for the particles of a minimal classical material is diffeomorphic to a hyper-surface of the tangent bundle $T\mathcal{M}$.
+The complete state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ for the particles of a minimal classical material is diffeomorphic to a hypersurface of the tangent bundle $T\mathcal{M}$.
 \end{prop}
@@
 \begin{justification}
-	We claim the complete state space is a hyper-surface of the cotangent bundle of the space-time manifold, $\bar{\mathcal{S}} \subset T^*\mathcal{M}$.
+	We claim the complete state space is a hypersurface of the cotangent bundle of the space-time manifold, $\bar{\mathcal{S}} \subset T^*\mathcal{M}$.
 Let $\bar{\mathcal{s}} \in \bar{\mathcal{S}}$ be the evolution of a particle for a minimal classical material.
@@
 
-	We claim there exists a homeomorphism between $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ and a hyper-surface of $T\mathcal{M}$.
+	We claim there exists a homeomorphism between $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ and a hypersurface of $T\mathcal{M}$.
 Let $\mathcal{s} \in \bar{\mathcal{S}}$ be the particle state for a minimal classical material.
@@
 It is a homemorphism between $T^*\mathcal{M}$ and $T\mathcal{M}$.
-$\bar{\mathcal{S}}$ is a hyper-surface of $T^*\mathcal{M}$ therefore its image is a hyper-surface of $T\mathcal{M}$.
+$\bar{\mathcal{S}}$ is a hypersurface of $T^*\mathcal{M}$ therefore its image is a hypersurface of $T\mathcal{M}$.
 
@@
 
-We have seen that the inertial mass $m$, the invariant speed $c$ and the reduced Plank constant $\hbar$ all serve similar purposes: convert one quantity to another preserving the number of possibilities.
+We have seen that the inertial mass $m$, the invariant speed $c$ and the reduced Planck constant $\hbar$ all serve similar purposes: convert one quantity to another preserving the number of possibilities.
 $dp=\hbar dk$ converts between ranges of wave number and conjugate momentum.
@@
 We now ask the following question: once we fix the entropy, what is the relationship between the spread of the marginal distributions along $q$ and $p$?
-Using Lagrange multipliers, we fix the total amount of material to $1$ and the total entropy $I_0$ and find the distribution that minimizes the product of the variance $\sigma_q^2 \sigma_p^2 = \int (q-\mu_q)^2 \rho_{\mathcal{c}} \, dqdp \int (p-\mu_p)^2 \rho_{\mathcal{c}} \, dqdp$.
+Using Lagrange multipliers, we fix the total amount of material to $1$ and the total entropy $I_0$ and find the distribution that minimizes the product of the variance $\sigma_q^2 \sigma_p^2 \equiv \int (q-\mu_q)^2 \rho_{\mathcal{c}} \, dqdp \int (p-\mu_p)^2 \rho_{\mathcal{c}} \, dqdp$.
 \begin{align*}

### 2017-01-25 2015CompleteDerivation/completeDerivation.tex
Including Lydia's comments

@@
 A \textbf{justification} follows each proposition to show the necessity of such a characterization.
-While the style of a justification is formal, it is not a pure mathematical proof as it necessarily has to mix physical and mathematical arguments.
+While it is a formal argument, it is not a strictly mathematical proof as it contains physical arguments.
 
@@
 
-This also allows us to understand conceptually why the speed of deterministic and reversible motion cannot be greater than $c$.
+This also allows us to understand heuristically why the speed of deterministic and reversible motion cannot be greater than $c$.
 Consider the trajectory of a point particle.

### 2017-01-25 2015CompleteDerivation/completeDerivation.tex
Filled in all the biblio suggestions

@@
 
-The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology \cite{LeeTM}, measure theory (ref?), differential geometry (symplectic and Riemannian) \cite{Lee}, group theory and vector spaces \cite{Lang,Young}, statistics (ref?), information theory (ref?), Hamiltonian mechanics, Lagrangian mechanics \cite{classical_dynamics}, quantum mechanics (ref?), thermodynamics (ref?), special relativity (ref?) and so on.\footnote{The references given provide the reader a list of resources consistent with the development of this work.} Using ideas from all these areas is necessary and valuable, but it does create a number of problems.
+The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology \cite{LeeTM}, measure theory \cite{Halmos}, differential geometry (symplectic and Riemannian) \cite{Lee}, group theory and vector spaces \cite{Lang,Young}, statistics \cite{Grimmet}, information theory \cite{Pierce}, Hamiltonian mechanics, Lagrangian mechanics \cite{classical_dynamics}, quantum mechanics \cite{Landau, Weinberg}, thermodynamics \cite{Van Ness}, special relativity \cite{French} and so on.\footnote{The references given provide the reader a list of resources consistent with the development of this work.} Using ideas from all these areas is necessary and valuable, but it does create a number of problems.
 
@@
 
-This link between state definition and deterministic processes should not be too surprising as the state, in the context of control theory and dynamical systems theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system}.(ref?) As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can indeed be calculated, measured and prepared.
+This link between state definition and deterministic processes should not be too surprising as the state, in the context of control theory and dynamical systems theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system} \cite{Katok, Sontang}.
+As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can indeed be calculated, measured and prepared.
 We can also describe the evolution of each element of the distribution provided that: we have a way to isolate it and study it under deterministic and reversible motion (so that we can define microstates); the non-deterministic motion does not alter the system (the set of microstates is preserved by the evolution).
@@
 Given that scientific reductionism (i.e.
-the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics,(ref?) we now explore how to characterize a system in terms of its components.
+the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics \cite{Ruse}, we now explore how to characterize a system in terms of its components.
 That is, we want to study the relationship between the state space of a composite system and the state space of its parts.
@@
 
-This intuitive picture is formalized mathematically by the Riesz representation theorem for linear functionals(ref?), which gives $\mathcal{S}$ the structure of a measure space, with a suitable Borel $\sigma$-algebra and measure $\mu$.
+This intuitive picture is formalized mathematically by the Riesz representation theorem for linear functionals \cite{Halmos}, which gives $\mathcal{S}$ the structure of a measure space, with a suitable Borel $\sigma$-algebra and measure $\mu$.
 
@@
 Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a general physical justification for.\footnote{One can, though, make the argument that given any distribution $\rho$ one can find an infinitely smooth $\rho_{sm}$ such that the difference in description is small compared to the error already introduced by assuming the system to be homogeneous and infinitely reducible.
-The Whitney approximation theorem(ref?) makes this mathematically well defined.}
+The Whitney approximation theorem \cite{Lee} makes this mathematically well defined.}
 
@@
 But ignoring that issue, the bigger problem is that we can always locally separate the evolution into independent degrees of freedom.
-For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (ref?)} We can always find such a local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value; they clearly evolve independently.
+For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle \cite{Linares}} We can always find such a local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value; they clearly evolve independently.
 And since the pieces are infinitesimal, a local decomposition is all that is needed.
@@
 	As a final note: we caution against automatically thinking that the classical assumption is valid for all macroscopic systems.
-It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, in which case the assumption would not hold.(ref?)
+It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, in which case the assumption would not hold.\footnote{This is the case in Bose-Einstein condensates \cite{BEC}, where each part becomes indistinguishable from the rest.}
 \end{rationale}
@@
 A system governed by Hamilton's equations is deterministic and reversible: the equation of motion given by $H$ are differentiable since $H$ is twice differentiable.
-Therefore the equations are Lipschitz continuous and, by the Picard-Lindel\"{o}f theorem(ref?), a unique solution exists.
+Therefore the equations are Lipschitz continuous and, by the Picard-Lindel\"{o}f theorem, a unique solution exists \cite{Grant}.
 
@@
 Integrating we have $\mathcal{H}=\frac{1}{2m}(p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta)+V(x)$ where $V(x)$ is an arbitrary function.
-We can further constrain the Hamiltonian by choosing the natural parameterization for $s$ (also called unit speed or arc length parametrization (ref?)) for which the modulus of the velocity is constant.
+We can further constrain the Hamiltonian by choosing the natural parameterization for $s$ (also called unit speed \cite{Lee} or arc length parametrization) for which the modulus of the velocity is constant.
 We have $u^\alpha u_\alpha = - c^2 = \frac{1}{m^2}(p^\alpha-A^\alpha)(p_\alpha-A_\alpha)$.
@@
 As they have the same net effect, if we apply all three we go back where we started.
-This gives us a very intuitive classical particle mechanics analogue for the $CPT$ theorem(ref?).
+This gives us a very intuitive classical particle mechanics analogue for the $CPT$ theorem \cite{Weinberg}.
 
@@
 \end{equation}
-where $\sigma^2$ is the variance, $\sigma$ the standard deviation, $\rho_{X,Y}= COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient(ref?).
+where $\sigma^2$ is the variance, $\sigma$ the standard deviation, $\rho_{X,Y}= COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient \cite{Grimmet}.
 As $-1\leq\rho_{X,Y}\leq+1$, the total variance can be greater or smaller than the sum of the variance of the components.
@@
 For a quantum mixed state $\tilde{\rho} = \sum \tilde{\rho}_i |\psi_i\rangle \langle \psi_i |$, we have $I(\tilde{\rho}) = - \sum \tilde{\rho}_i \ln \tilde{\rho}_i - \tilde{\rho}_i I(\psi_i) = - \sum \tilde{\rho}_i \ln \tilde{\rho}_i = -tr(\tilde{\rho} \ln \tilde{\rho})$.
-And in fact this is the Von Neumann entropy (ref?) used in quantum information, that implicitly assumes all states provide the same amount of entropy (i.e.
+And in fact this is the Von Neumann entropy \cite{Von Neumann} used in quantum information, that implicitly assumes all states provide the same amount of entropy (i.e.
 zero).
@@
 
-Naturally, we cannot give a complete description for the configuration of a fragment, therefore we cannot give a joint distribution for all fragment variables at the same time.\footnote{This recovers the notion of complementarity introduced by Bohr(ref?).} We can, though, find a maximal set of fragment variables: one that can be used to provide the best description possible.
+Naturally, we cannot give a complete description for the configuration of a fragment, therefore we cannot give a joint distribution for all fragment variables at the same time.\footnote{This recovers the notion of complementarity introduced by Bohr \cite{Bohr}.} We can, though, find a maximal set of fragment variables: one that can be used to provide the best description possible.
 While there isn't a unique maximal set, we can show that one such set is provided by the fragment variables $q^i : \mathcal{Q} \rightarrow \mathbb{R}$ that define the units required to describe the system.
@@
 	This notion may seem in contradiction to the current practice.
-For example, the S-matrix, one of the main devices for calculating scattering amplitudes and cross sections, is derived by assuming unitary evolution.(ref?) Doesn't it imply we are assuming that deterministic and reversible evolution is happening at every instant?
+For example, the S-matrix, one of the main devices for calculating scattering amplitudes and cross sections, is derived by assuming unitary evolution \cite{Weinberg}.
+Doesn't it imply we are assuming that deterministic and reversible evolution is happening at every instant?
 
@@
 
+\bibitem{Halmos} Halmos, P: Measure Theory. D. van Nostrand and Co. (1950)
+
 \bibitem{Lee} Lee, J. M.: Introduction to Smooth Manifolds. Springer (2003)
@@
 
+\bibitem{Grimmet} Grimmett, G. R., Stirzaker, D. R.: Probability and random processes. Oxford Univ. Press 3. ed. (2001)
+
+\bibitem{Pierce} Pierce, J. R.: An Introduction to Information Theory: Symbols, Signals and Noise. Dover Publications (1980)
+
 \bibitem{classical_dynamics} Jos\'{e}, J. V., Saletan, E. J.: Classical Dynamics. Cambridge University Press, (1998)
 
+\bibitem{Landau} Landau, L.D., Lifshitz, E.M.: Quantum Mechanics: Non-Relativistic Theory. Vol. 3 (3rd ed.). Pergamon Press (1977)
+
+\bibitem{Weinberg} Weinberg, S.: The Quantum Theory of Fields, vol I. Cambridge University Press (2002)
+
+\bibitem{Van Ness} Van Ness, H. C.: Understanding Thermodynamics. Dover Publications (1983)
+
+\bibitem{French} French, A. P.: Special Relativity. W. W. Norton \& Company (1968)
+
+\bibitem{Katok} Katok, A., Hasselblatt, B.: Introduction to the modern theory of dynamical systems. Cambridge (1996)
+
+\bibitem{Sontang} Sontag, E. D.: Mathematical Control Theory: Deterministic Finite Dimensional Systems (2nd ed.). Springer (1999)
+
 \bibitem{Shannon} Shannon, C. E.: A mathematical theory of communication. The Bell System Technical Journal, Vol. 27, pp. 379--423, 623--656, (1948)
@@
 
+\bibitem{Ruse} Ruse, M.: Entry for ``Reductionism". The Oxford Companion to Philosophy (2nd ed.). Oxford University Press. p. 793 (2005)
+
+\bibitem{Linares} Galindo-Linares, E., Navarro-Morale, E., Silva-Ortigoza, G., Suárez-Xique, R., Marciano-Melchor, M., Silva-Ortigoza, R., Román-Hernández, E.: Any Hamiltonian System Is Locally Equivalent to a Free Particle.
+World Journal of Mechanics, 2012, 2, pp.
+246--252 (2012)
+
+\bibitem{BEC} Pitaevskii, L. P., Stringari, S.: Bose–Einstein Condensation. Clarendon Press, Oxford (2003)
+
 \bibitem{Gromov} Gromov, M. L.: Pseudo holomorphic curves in symplectic manifolds. Inventiones Mathematicae 82, pp. 307--347, (1985)
@@
 
+\bibitem{Grant} Grant, C. P.: Theory of Ordinary Differential Equations. CreateSpace Independent Publishing Platform (2014)
+
 \bibitem{Treumann} Treumann, R. A., Nakamura, R., Baumjohann,  W.: Relativistic transformation of phase-space distributions. Annales Geophysicae (2011)
@@
 
+\bibitem{Von Neumann} Von Neumann, J.: Mathematical Foundations of Quantum Mechanics. Princeton University Press (1955)
+
+\bibitem{Bohr} Bohr, N.: The Quantum Postulate and the Recent Development of Atomic Theory. Nature 121, pp. 580-–590 (1928)
 \end{thebibliography}

### 2017-01-26 2015CompleteDerivation/completeDerivation.tex
Minor fixes

@@
 	As a final note: we caution against automatically thinking that the classical assumption is valid for all macroscopic systems.
-It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, in which case the assumption would not hold.\footnote{This is the case in Bose-Einstein condensates \cite{BEC}, where each part becomes indistinguishable from the rest.}
+It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, in which case the assumption would not hold.\footnote{For example, this is the case in Bose-Einstein condensates \cite{BEC}.}
 \end{rationale}
@@
 
-\bibitem{BEC} Pitaevskii, L. P., Stringari, S.: Bose–Einstein Condensation. Clarendon Press, Oxford (2003)
+\bibitem{BEC} Pitaevskii, L. P., Stringari, S.: Bose–-Einstein Condensation. Clarendon Press, Oxford (2003)
 

### 2017-02-03 2015CompleteDerivation/completeDerivation.tex
Comments from Dave and Josh

@@
 This means that our measure $\mu$ needs to be modified to define integrals over the time variable.\footnote{Looking ahead, another way to frame the problem is that Galilean transformations do not preserve phase space volume.
-Lorentz transformations, on the other hand, do\cite{Treumann} and it is this requirement plus the kinematic assumption that will recover them later.}
+Lorentz transformations, on the other hand, do \cite{Treumann}.}
 
@@
 Such hypersurface will need to cut across all evolutions once and only once, therefore not all functions $t : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ are suitable time variables.
-A time evolution is a line  $\bar{\mathcal{s}} \subset \bar{\mathcal{S}}$ that passes once through each equal time surfaces.
+A time evolution is a line  $\bar{\mathcal{s}} \subset \bar{\mathcal{S}}$ that passes once through each equal time surface.
 Specifying an evolution map means giving a set of lines such that each state is traversed once and only once.
@@
 	We claim $\bar{\mathcal{S}}$ is a $2n+1$ dimensional differentiable manifold.
-Locally $\bar{\mathcal{S}} \supseteq U \cong T^*\mathcal{Q}\times \mathbb{R}$ as a state $\mathcal{s} \in U$ exists for each configuration and each time instant.
+Locally there exists $U\subseteq \bar{\mathcal{S}}$ such that $U \cong T^*\mathcal{Q}\times \mathbb{R}$ as a state $\mathcal{s} \in U$ exists for each configuration and each time instant.
 Locally $\bar{\mathcal{S}} \supseteq U \cong\mathbb{R}^{2n+1}$.
@@
 
-We now want to extend the symplectic form so we can count state and possibilities over the complete state space $\bar{\mathcal{S}}$.
+We now want to extend the symplectic form so we can count states and possibilities over the complete state space $\bar{\mathcal{S}}$.
 The first thing to do is to complete our $n$-dimensional $\mathcal{Q}$, the manifold that defines our units, to include the time variable $t$.
@@
 Let $E\equiv\hbar \bar{\omega}$.
-Let $\omega$ be the symplect form.
+Let $\omega$ be the symplectic form.
 The components of $\omega$ across different d.o.f.
@@
 
-As we are using time as a variable to identify states, we will use a different quantity as the parameter for the evolution.
+As we are using time as a variable to label states, we will use a different quantity as the parameter for the evolution.
 The trajectory of a particle in the extended phase space will be given by the evolved variables $\xi^a(s)$ in terms of a parameter $s$.
@@
 In other words: $E$ is the value of the Hamiltonian, the energy of the system.
-We have found the constraint we discussed when introducing \ref{prop:relativistic_symplectic_manifold}.
+We have found the constraint that identifies the hypersuface of $T^*\mathcal{M}$ introduced in \ref{prop:relativistic_symplectic_manifold}.
 
@@
 Therefore by studying a minimal material, we are also studying a component of the dynamics that is present in all materials.
-In other words: any material is an extension to a minimal material.
-
-	It should also be noted that, once we reduced particles to be infinitesimal at a point, the rest of the description must be a local object invariant under coordinate transformations.
+In other words: any material is an extension of a minimal material.
+
+	It should also be noted that, once we reduce particles to be infinitesimal at a point, the rest of the description must be a local object invariant under coordinate transformations.
 This means that the remaining internal state must be identified by scalar, vector or tensor quantities.
@@
 Under the kinematic assumption, these units must also be sufficient to describe the dynamics.
-Therefore $\mathcal{M}$ has conceptually two roles: it is the space where the motion unfolds and it is the space that defines the system of units for our states.
+Therefore $\mathcal{M}$ has conceptually two roles: it is the space where the motion unfolds and it is the space on which we define the system of units for our states.
 This already tells us that the state space for the particles of the material is a hypersurface of $T^*\mathcal{M}$.
@@
 We recognize $m$ as the inertial mass.
-Integrating the expression, we have $p_\alpha = m u_\alpha + A_\alpha(x)$ where $A_\alpha(x)$ is an arbitrary vector field.
-This is the most general link between velocity and conjugate momentum, which is the sum of kinetic momentum and a scalar potential.
+Integrating the expression, we have $p_\alpha = m u_\alpha + A_\alpha(x)$ where $A_\alpha(x)$ is an arbitrary vector field.\footnote{Note that $A_\alpha(x)$ corresponds to the product the vector potential and the charge.
+This simplifies the notation and avoids having to find a symbol for the charge itself.} This is the most general link between velocity and conjugate momentum, which is the sum of kinetic momentum and a scalar potential.
 
@@
 The reason is that the choice of units for $x^\alpha$ already determines the others ($x^\alpha \rightarrow u^\alpha$ and $x^\alpha \rightarrow q^\alpha \rightarrow k_\alpha \rightarrow p_\alpha$) therefore the link between units of $u^\alpha$ and $p_\alpha$ is not a new unit conversion.
+In other words: $m$ is not just another arbitrary constant set by the choice of units.
 
@@
 
-We now have all the building block of classical mechanics, the consequences of three seemingly simple assumptions.
-We have more insight of why these blocks fit together the way they do.
+We now have all the building blocks of classical mechanics, the consequences of three seemingly simple assumptions.
+We have more insight into why these blocks fit together the way they do.
 But we have also placed ourselves in a very good position to move forward to quantum systems by using the new physical insights we have gained.
@@
 
-Before introducing assumption \ref{ass:irreducibility} and deriving quantum mechanics as its consequence, we first give a semi-classical account of the conceptual picture that emerges.
+In section \ref{sec:reducibility} we made the key assumption of infinitesimal reducibility which led to classical Hamiltonian mechanics.
+We are now going back to that point, that is we set aside the kinematic assumption and time dependent evolution, and take a different path.
+But before introducing assumption \ref{ass:irreducibility} and deriving quantum mechanics as its consequence, we first give a semi-classical account of the conceptual picture that emerges.
 This allows us to build the intuition more gradually and to show that classical and quantum mechanics are not as different as one may first imagine.
@@
 
-As we defined them in previous sections, classical particles should not be understood as point-like objects of finite mass.
+As we defined them in previous sections, classical particles are the infinitesimal limit of subsequent subdivisions and therefore should not be understood as point-like objects of finite mass.
 It was their being infinitesimal and spread around a small neighborhood that gave us conjugate pairs of state variables and Hamiltonian evolution.
@@
 the infinitesimal part of a classical material), this would indeed be worthy of note and puzzlement.
-But if the electron is a distribution, there is no issue as a phase space distribution does not have a single value for position or momentum.
+But if the electron is a distribution, there is no issue because a phase space distribution does not have a single value for position or momentum.
 The earth is, so to speak, in both North America and Europe at the same time.
@@
 If the spread is not negligible (e.g.
-it encounters obstacles of a smaller scale) then we have to use the full distribution.
+the object encounters obstacles of a smaller scale) then we have to use the full distribution.
 Therefore, even in this classical approximation, the electron would have this dual nature depending on the problem.
@@
 
-Not only do point-like objects not have to be elementary, but it's unclear how they can be elementary.
-How can the whole universe be reduced to  elements that have no idea how many spatial dimensions are there?
-Mathematically a set of points by itself does not define the dimension of a space: $\mathbb{R}^n$ is the same order of infinity for any $n$.
-One needs a topology to differentiate, which is an additional structure.
-It's the neighborhoods that define the dimensionality.
-If everything is just points, where does the topology come from?
+Whether a physical system has a distinguishable substructure depends on whether there exists a physical process that can interact with one of its parts independently of the others.
+If we are always forced to interact with the whole system, then we can't find any discernible part.
+The fact that the system is extended in space is a separate property, which does not necessarily imply the existence of such a process.
 
@@
 $\int \rho_\mathcal{c} dq \wedge dp = 1$ where $\rho_\mathcal{c}$ is the normalized distribution.
-In this context, we will call \emph{fragment} an infinitesimal part of our classical electron.
-The Shannon entropy\cite{Shannon} associated with the distribution is finite as well.
+In this context, we will call a \emph{fragment} an infinitesimal part of our classical electron.
+The Shannon entropy \cite{Shannon} associated with the distribution is finite as well.
 $I(\rho_\mathcal{c}) = - \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq \wedge dp = I_0$.
@@
 The final entropy will also remain the same: $I(\rho_{\hat{\mathcal{c}}}) = I_0$.
-Mathematically, because $-\rho_\mathcal{c} \log (\rho_\mathcal{c})$, like the distribution itself, is just another integrable function and therefore its integral is conserved by Hamiltonian evolution.
-Physically, because the amount of information required to identify a fragment given the distribution has to remain the same under deterministic and reversible evolution.
+Mathematically, because $-\rho_\mathcal{c} \log (\rho_\mathcal{c})$, like the distribution itself, is just another integrable function, its integral is conserved by Hamiltonian evolution.
+Physically, the entropy remains unchanged because the amount of information required to identify a fragment given the distribution has to remain the same under deterministic and reversible evolution.
 Therefore, in the same way that each fragment is constrained to move through states at equal energy, the composite state is constrained to move through distributions at equal entropy.
@@
 the evolution of the unstated part).
-Whatever it may be, if it is characterized by an entropy comparable to the one of the distribution, that is we are trying to localize the electron better than the source can allow, the classical assumption breaks down as the random motion of each fragment is not negligible.
+Whatever it may be, if it is characterized by an entropy comparable to the one of the distribution (i.e.
+if we are trying to localize the electron better than the source can allow) the classical assumption breaks down as the random motion of each fragment is not negligible.
 
@@
 
-\item Assuming classical particle states are fully characterized by their trajectory in space, the states also evolve according to Lagrangian mechanics and the motion is that of a massive particle under scalar/vector potential forces.
-
-\item Relativistic motion arises by properly handling time dependent deterministic and reversible evolution for the material (i.e. distributions over phase space) without further assumptions.
+\item Lagrangian mechanics describes the case where the trajectory in space fully characterizes the dynamics of the system.
+If we want distributions in phase space to be expressed in terms of kinematic variables (i.e.
+position and velocity), we find that $(\mathcal{Q}, g)$ is a Riemannian manifold and the motion is that of a massive particle under scalar/vector potential forces.
+
+\item Relativistic motion arises by properly handling time dependent deterministic and reversible evolution for the material (i.e.
+distributions over phase space) without further assumptions.
+That is, we derive the geometrical structure of space-time from the geometrical structure of phase space, inverting the typical approach.
 

### 2017-02-13 2015CompleteDerivation/completeDerivation.tex
Removing mention of Galilean/Lorenz transformations

@@
 Our composite state distribution $\rho(q)$ becomes $\rho(\hat{q},t)$: it is no longer defined at equal time.
-This means that our measure $\mu$ needs to be modified to define integrals over the time variable.\footnote{Looking ahead, another way to frame the problem is that Galilean transformations do not preserve phase space volume.
-Lorentz transformations, on the other hand, do \cite{Treumann}.}
+This means that our measure $\mu$ needs to be modified to define integrals over the time variable.
 
@@
 
-\bibitem{Treumann} Treumann, R. A., Nakamura, R., Baumjohann,  W.: Relativistic transformation of phase-space distributions. Annales Geophysicae (2011)
+%\bibitem{Treumann} Treumann, R. A., Nakamura, R., Baumjohann,  W.: Relativistic transformation of phase-space distributions. Annales Geophysicae (2011)
 

### 2017-02-16 2015CompleteDerivation/completeDerivation.tex
Comments from Stephen

@@
 
-The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology \cite{LeeTM}, measure theory \cite{Halmos}, differential geometry (symplectic and Riemannian) \cite{Lee}, group theory and vector spaces \cite{Lang,Young}, statistics \cite{Grimmet}, information theory \cite{Pierce}, Hamiltonian mechanics, Lagrangian mechanics \cite{classical_dynamics}, quantum mechanics \cite{Landau, Weinberg}, thermodynamics \cite{Van Ness}, special relativity \cite{French} and so on.\footnote{The references given provide the reader a list of resources consistent with the development of this work.} Using ideas from all these areas is necessary and valuable, but it does create a number of problems.
+The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology~\cite{LeeTM}, measure theory~\cite{Halmos}, differential geometry (symplectic and Riemannian)~\cite{Lee}, group theory and vector spaces~\cite{Lang,Young}, statistics~\cite{Grimmet}, information theory~\cite{Pierce}, Hamiltonian mechanics, Lagrangian mechanics~\cite{classical_dynamics}, quantum mechanics~\cite{Landau, Weinberg}, thermodynamics~\cite{Van Ness}, special relativity \cite{French} and so on.\footnote{The references given provide the reader a list of resources consistent with the development of this work.} Using ideas from all these areas is necessary and valuable, but it does create a number of problems.
 
@@
 Yes, but that would be a different evolution defined on a different state.
-Does it mean the unstated part is always evolving chaotically?
+Does it mean the unstated part is always evolving randomly?
 Not at all.
@@
 
-As the term "measurement" has become particularly loaded, let's first characterize what we mean by physical distinguishability in our context.
+As the term ``measurement" has become particularly loaded, let's first characterize what we mean by physical distinguishability in our context.
 
 Consider the motion of a cannonball under inertial and gravitational forces.
-Light will scatter off of it; as it lands the ground will be deformed and the impact will make the temperature slightly rise.
+Light will scatter off of it; as it lands, the ground will be deformed and the impact will make the temperature rise slightly.
 Those external physical processes, which happen no matter what we do, can be used to distinguish the motion of the cannonball as their outcomes are correlated.
@@
 First, to make sure the process is indeed correlated to the states.
-Second, because ``a single take does not a measurement make".
-One has to gather enough statistics.
+Second, because ``a single take does not a measurement make." One has to gather enough statistics.
 Note that the number of takes influences the outcomes: with greater statistics the precision and number of distinguishable cases increases.
@@
 Let $\mathsf{T}$ be the set of all sets associated with all physical outcomes.
-$\mathcal{S} \in \mathsf{T}$ and is associated with the outcome ``the system exists".
-$\varnothing \in \mathsf{T}$ and is associated with the outcome ``the system doesn't exist".
-Let $V_1, V_2 \in \mathsf{T}$.
+$\mathcal{S} \in \mathsf{T}$ and is associated with the outcome ``the system exists." $\varnothing \in \mathsf{T}$ and is associated with the outcome ``the system doesn't exist." Let $V_1, V_2 \in \mathsf{T}$.
 Then, by definition, there exists a process $P_1$ that admits $V_1$ as an outcome and a process $P_2$ that admits $V_2$ as an outcome.
@@
 
-The generality of this result explains why in physics one always assumes functions to be ``well behaved".
-As the result was derived from our notion of physical distinguishability, this is not a matter of practical convenience.
+The generality of this result explains why in physics one always assumes functions to be ``well behaved." As the result was derived from our notion of physical distinguishability, this is not a matter of practical convenience.
 Suppose we were able to prepare a force field that was zero everywhere in space except at a single point.
@@
 
-Evolved state variables are useful to write equations of motions, study how physical quantities change, form a physical picture of what happens.
+Evolved state variables are useful to write equations of motion, study how physical quantities change, form a physical picture of what happens.
 Transported state variables are useful to write invariants, study state space trajectories, form a geometric picture for the state space.\footnote{One should not confuse evolved/transported state variable with active/passive transformations or with Schroedinger/Heisenberg pictures.
@@
 By changing state variable, though, we see that this does not work in general: $\rho_\mathcal{c}(\hat{q})= \rho_\mathcal{c}(q) dq/d\hat{q}$.
+For example, if we changed units from $g/m$ to $g/Km$, the value of the density would increase by a factor of $1000$.
 This makes $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using.
@@
 Therefore even in the best of cases it cannot be considered an experimentally validated assumption but a reasonable default assumption (e.g.
-"it worked so far").
+``it worked so far").
 
@@
 \end{align*}
-But in the previous case, the right angle was between the two independent d.o.f..
+But in the previous case, the right angle was between the two independent d.o.f.
 In this case, the right angle is between $\omega$ and the plane of constant $(q, p)$ where $dt \wedge dE$ is defined.
@@
 $d_s \rho_{\bar{\mathcal{c}}} = 0$.
-
 \end{justification}
@@
 If we assume kinematic equivalence, the kinematic variables will be enough to identify a particle state.
-In such a case, we will call the material \emph{minimal} to highlight the idea that the state is described by the smallest possible number of d.o.f..
+In such a case, we will call the material \emph{minimal} to highlight the idea that the state is described by the smallest possible number of d.o.f.
 
@@
 The overall dynamics will be the spatial d.o.f.
-coupled with the internal d.o.f..
+coupled with the internal d.o.f.
 Therefore by studying a minimal material, we are also studying a component of the dynamics that is present in all materials.
@@
 Since both $dx^\alpha$ and $du^\alpha$ are four vector components, $g$ is a rank two tensor on $\mathcal{M}$.
-To each direction in space-time $dx^\alpha$ is associated a direction in the tangent space of velocities $du^\alpha$ with which it forms a d.o.f..
+To each direction in space-time $dx^\alpha$ is associated a direction in the tangent space of velocities $du^\alpha$ with which it forms a d.o.f.
 Across d.o.f., that is $\alpha \neq \beta$, we have $\omega(dq^\alpha, dp_\beta) = 0 =\omega(dq^\beta, dp_\alpha)$.
@@
 Since time must use a different unit, we set $g_{00}=-c^2$ where $c^2$ is the constant that allows one to convert the possibility count over the temporal d.o.f.
-into the possibility count of a spatial d.o.f..
+into the possibility count of a spatial d.o.f.
 Therefore, locally $g(dx^i, dx^i) = (dx^i)^2 - c^2 dt^2$ is Minkowskian.
@@
 Position and velocity along the same direction use related units.
-They cannot form independent d.o.f..
+They cannot form independent d.o.f.
 $g(dx^\alpha, du^\alpha) \neq 0$.
@@
 $g$ does not measure space-time distances: it counts possibilities of a d.o.f., like $\omega$.
-Directions in space are orthogonal not because the angle between them is $90^\circ$, but because they identify independent d.o.f..
+Directions in space are orthogonal not because the angle between them is $90^\circ$, but because they identify independent d.o.f.
 That is, the choice of units along one direction is independent from the choice of units along orthogonal directions.
@@
 Non-local quantum effects, such as the ones featured in EPR\cite{EPR} experiments, are not examples of deterministic and reversible evolution.
-Therefore they do not have anything to violate: the maximum speed does not apply to their ``motion".\footnote{The very notion of velocity becomes ill defined in such cases precisely because we do not have a unique well-defined evolution.} On the other hand, only deterministic and reversible motion can carry information and can establish before/after relationships, therefore everything remains consistent.
+Therefore they do not have anything to violate: the maximum speed does not apply to their ``motion."\footnote{The very notion of velocity becomes ill defined in such cases precisely because we do not have a unique well-defined evolution.} On the other hand, only deterministic and reversible motion can carry information and can establish before/after relationships, therefore everything remains consistent.
 
@@
 $\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta) + m^2 c^2)$.
-
 \end{justification}
@@
 The weaker the force, the more $(x^i, mu_i)$ become independent d.o.f.
-The force tensor, then, tells how much $(x^i, mu_i)$ fail to be independent d.o.f..
+The force tensor, then, tells how much $(x^i, mu_i)$ fail to be independent d.o.f.
 With these expressions, the Poisson bracket between any function of position and kinetic momentum can now be calculated.
@@
 If the spread is not negligible (e.g.
-the object encounters obstacles of a smaller scale) then we have to use the full distribution.
+the object encounters obstacles or slits of a smaller scale) then we have to use the full distribution.
 Therefore, even in this classical approximation, the electron would have this dual nature depending on the problem.
@@
 While each fragment moves randomly, the overall distribution $\rho$ stays the same: the non-deterministic motion is effectively just reshuffling the fragments within the same distribution.
-The value of the entropy of the distribution, and therefore its spread, is set by the non-deterministic motion of the particles: the stronger it is, the greater the entropy.
+The value of the entropy of the distribution, and therefore its spread, is set by the non-deterministic motion of the fragments: the stronger it is, the greater the entropy.
 
@@
 is outside of the scope of this work.
-Note, though, that this result is related to the Gromov non-squeezing theorem and the symplectic camel, which are already generalized to multiple d.o.f..
+Note, though, that this result is related to the Gromov non-squeezing theorem and the symplectic camel, which are already generalized to multiple d.o.f.
 It is the link with information theory that gives us further physical insight and is of particular interest given the strong connection between thermodynamics and Shannon's entropy.
@@
 Yes, since the particle can be spatially decomposed some amount of material may be inside the volume.
-Can we ask where that same fraction will be at a later time?
+Can we ask where the fragments in that same fraction will be at a later time?
 No, since the particle is irreducible we cannot describe the evolution of a part.
@@
 A component will have a state associated with it, and therefore will always be a collection of fragments.
-As in the classical case, a component may be the difference between two states, therefore one should not imagine it always ``contained" in the composed system.
 
@@
 In other words, our simple model is not only consistent with quantum mechanics but it has predictive power: it provides intuition that is well supported by these findings.
-It is not ``just a mathematical trick".
+It is not ``just a mathematical trick."
 
@@
 The abelian group $\Psi$ can be extended with the operations defined by $T$ as was done in \ref{prop:real_vector_space} for the infinitesimally reducible case.
-
 \end{justification}
@@
 
-For the tool set to be on par with classical mechanics we need to address three requirements.
+For the tool set to be on par with classical mechanics, such that we can write the equations of motion, we need to address three requirements.
 We need to be able to compare two states and ask what fraction of them share the same configuration: this will give us an inner product.
@@
 That is $\psi_1 = \alpha \psi_2 + \psi_3$ is the combination of the component that is prepared like $\psi_2$ (e.g.
-same distributions in position and momentum) and a component that is prepared completely differently.
+same distributions in position and momentum) and the component $\psi_3$ that is prepared completely differently.
 Let $\mathsf{P}_{\psi_2} : \Psi \rightarrow \Psi$ be the operator that gives us the component $\mathsf{P}_{\psi_2} (\psi_1) \equiv \alpha \psi_2$ prepared like $\psi_2$.
@@
 $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$.
-
 \end{justification}
@@
 
-While $\mathcal{Q}$ is enough to define the system of units, it is not enough to define all the fragment variables over which we can define the distributions of irreducible material.
+While $\mathcal{Q}$ is enough to define the system of units, it is not enough to define all the fragment variables over which we can express the distributions of irreducible material.
 The conjugate variables $p_i=\hbar k_i$ cannot be expressed as a function of $q^i$, which reiterates the idea that we cannot specify all variables for each fragment.
@@
 On one hand, it tells us that there is a component of chaotic motion within the system.
-On the other hand, it can only characterize the equilibrium dynamics, which clearly can't be the full range.
+On the other hand, it can only characterize the equilibrium dynamics, which clearly can't be the full range of dynamics.
 Yet it also offers a way out: since the motion of the fragments is not accessible, we can predict statistical distributions that are independent of it.
@@
 We then repeat the process, making sure the incoming particle is prepared in the same way.
-Given a single "pure" incoming particle state, we get a "mixed" statistical distribution describing the output of the array.\footnote{Note that this is different from the model used often to describe the quantum measurement problem, where the incoming system becomes entangled with the measuring device.
+Given a single ``pure" incoming particle state, we get a ``mixed" statistical distribution describing the output of the array.\footnote{Note that this is different from the model used often to describe the quantum measurement problem, where the incoming system becomes entangled with the measuring device.
 We do not make any claim on what the final state of the particle should be: in many cases, like in a photomultiplier tube, the particle is gone and there is no state.
-We simply have one incoming state for the particle, prepared over and over exactly the same, and the outgoing state for the measurement device, which is a classical statistical distribution over the different measurement takes.} The question is: how is the statistical distribution coming out of the detector related to the material distribution coming in with the particle?
+We simply have one incoming state for the particle, prepared over and over exactly the same, and the outgoing state for the measurement device, which is a classical statistical distribution over the different measurement takes.} The question is, how is the statistical distribution coming out of the detector related to the material distribution coming in with the particle?
 
@@
 \begin{acknowledgements}
-We'd like to thank Erik Curiel, Jiahua Gu, Josh Hunt, Isaac Mooney and Pavel Okun for providing feedback on drafts of this manuscript, Mark Greenfield for further insights in the topological aspects and Alan Phillips for contributing figures.
+We'd like to thank Erik Curiel, Stephen DiIorio, Jiahua Gu, Josh Hunt, Isaac Mooney and Pavel Okun for providing feedback on drafts of this manuscript, Mark Greenfield for further insights in the topological aspects and Alan Phillips for contributing figures.
 We additionally thank Paul Stankus for interest and support in the earlier stages of this project.

### 2017-02-16 2015CompleteDerivation/completeDerivation.tex
Preparing title page for submission

@@
 \newif\ifjournal
-%\journaltrue
-\journalfalse
+\journaltrue
+%\journalfalse
 
@@
 
-\title{DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics}
+\title{From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics}
 \author{Gabriele Carcassi, Christine A. Aidala, David J. Baker, Lydia Bieri}
@@
 \ifjournal
-	\institute{G. Carcassi \at
-		University of Michigan, Ann Arbor, MI 48109 \\
-		\email{carcassi@umich.edu}           %  \\
+	\institute{G. Carcassi, C. A. Aidala \at
+		Physics Department, University of Michigan, Ann Arbor, MI 48109 \\
+		\email{carcassi@umich.edu}
+		\and
+		D. J. Baker \at
+		Philosophy Department, University of Michigan, Ann Arbor, MI 48109
+		\and
+		L. Bieri \at
+		Mathematics Department, University of Michigan, Ann Arbor, MI 48109 \\
 	}
@@
 \begin{abstract}
-\textbf{This manuscript is a work in progress.} Ideas are constantly reshaped to find more precise and elegant arguments.
-It is provided as is to stimulate discussion.
-\textbf{Make sure you have the latest version from http://assumptionsofphysics.org}
-
 The aim of this work is to show that particle mechanics, both classical and quantum, Hamiltonian and Lagrangian, can be derived from few simple physical assumptions.
@@
 Assuming kinematic equivalence, that studying trajectories is equivalent to studying state evolution, will give us Lagrangian mechanics and limit the form of the Hamiltonian/Lagrangian to the one with scalar and vector potential forces.
+\keywords{Hamiltonian mechanics \and Lagrangian mechanics \and Quantum mechanics \and Fundamental assumptions}
 \end{abstract}
@@
 \section{Introduction}
-
-Status:
-\begin{itemize}
-%\item Section \ref{sec:fundamental_model}: Done.
-
-%\item Section \ref{sec:dynamical_systems}: Done.
-
-%\item Section \ref{sec:reducibility}: Done.
-
-%\item Section \ref{sec:relativity}: Done.
-
-\item Section \ref{sec:Lagrangian}: Needs review.
-
-\item Section \ref{sec:quantum_prelude}: Needs review.
-
-\item Section \ref{sec:irreducibility}: Needs review.
-\end{itemize}
 

### 2017-02-16 2015CompleteDerivation/completeDerivation.tex
Fixed space before citation

@@
 
-The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology~\cite{LeeTM}, measure theory~\cite{Halmos}, differential geometry (symplectic and Riemannian)~\cite{Lee}, group theory and vector spaces~\cite{Lang,Young}, statistics~\cite{Grimmet}, information theory~\cite{Pierce}, Hamiltonian mechanics, Lagrangian mechanics~\cite{classical_dynamics}, quantum mechanics~\cite{Landau, Weinberg}, thermodynamics~\cite{Van Ness}, special relativity \cite{French} and so on.\footnote{The references given provide the reader a list of resources consistent with the development of this work.} Using ideas from all these areas is necessary and valuable, but it does create a number of problems.
+The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology~\cite{LeeTM}, measure theory~\cite{Halmos}, differential geometry (symplectic and Riemannian)~\cite{Lee}, group theory and vector spaces~\cite{Lang,Young}, statistics~\cite{Grimmet}, information theory~\cite{Pierce}, Hamiltonian mechanics, Lagrangian mechanics~\cite{classical_dynamics}, quantum mechanics~\cite{Landau,Weinberg}, thermodynamics~\cite{Van Ness}, special relativity~\cite{French} and so on.\footnote{The references given provide the reader a list of resources consistent with the development of this work.} Using ideas from all these areas is necessary and valuable, but it does create a number of problems.
 
@@
 
-This link between state definition and deterministic processes should not be too surprising as the state, in the context of control theory and dynamical systems theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system} \cite{Katok, Sontang}.
+This link between state definition and deterministic processes should not be too surprising as the state, in the context of control theory and dynamical systems theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system}~\cite{Katok,Sontang}.
 As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can indeed be calculated, measured and prepared.
@@
 
-However precise our measurements are, we can only gather a finite amount of statistics and each outcome is expressed by a finite set of digits; therefore, the set of outcomes is countable.\footnote{The information provided by the process as measured by Shannon's entropy\cite{Shannon,Jaynes} is finite.} Note that different outcomes for the same process can overlap.
+However precise our measurements are, we can only gather a finite amount of statistics and each outcome is expressed by a finite set of digits; therefore, the set of outcomes is countable.\footnote{The information provided by the process as measured by Shannon's entropy~\cite{Shannon,Jaynes} is finite.} Note that different outcomes for the same process can overlap.
 For example, $4.12 \pm 0.05$ cm and $4.13 \pm 0.05$ cm are both legitimate possible outputs of the same measurement device.
@@
 Given that scientific reductionism (i.e.
-the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics \cite{Ruse}, we now explore how to characterize a system in terms of its components.
+the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics~\cite{Ruse}, we now explore how to characterize a system in terms of its components.
 That is, we want to study the relationship between the state space of a composite system and the state space of its parts.
@@
 
-This intuitive picture is formalized mathematically by the Riesz representation theorem for linear functionals \cite{Halmos}, which gives $\mathcal{S}$ the structure of a measure space, with a suitable Borel $\sigma$-algebra and measure $\mu$.
+This intuitive picture is formalized mathematically by the Riesz representation theorem for linear functionals~\cite{Halmos}, which gives $\mathcal{S}$ the structure of a measure space, with a suitable Borel $\sigma$-algebra and measure $\mu$.
 
@@
 Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a general physical justification for.\footnote{One can, though, make the argument that given any distribution $\rho$ one can find an infinitely smooth $\rho_{sm}$ such that the difference in description is small compared to the error already introduced by assuming the system to be homogeneous and infinitely reducible.
-The Whitney approximation theorem \cite{Lee} makes this mathematically well defined.}
+The Whitney approximation theorem~\cite{Lee} makes this mathematically well defined.}
 
@@
 But ignoring that issue, the bigger problem is that we can always locally separate the evolution into independent degrees of freedom.
-For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle \cite{Linares}} We can always find such a local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value; they clearly evolve independently.
+For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle~\cite{Linares}} We can always find such a local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value; they clearly evolve independently.
 And since the pieces are infinitesimal, a local decomposition is all that is needed.
@@
 	As a final note: we caution against automatically thinking that the classical assumption is valid for all macroscopic systems.
-It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, in which case the assumption would not hold.\footnote{For example, this is the case in Bose-Einstein condensates \cite{BEC}.}
+It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, in which case the assumption would not hold.\footnote{For example, this is the case in Bose-Einstein condensates~\cite{BEC}.}
 \end{rationale}
@@
 is also conserved.
-These statements are the physical justification of Gromov's non-squeezing theorem \cite{Gromov,deGosson,Stewart} and Liouville's theorem.
+These statements are the physical justification of Gromov's non-squeezing theorem~\cite{Gromov,deGosson,Stewart} and Liouville's theorem.
 And they give intuitive insight on the geometry of Hamiltonian systems.
@@
 A system governed by Hamilton's equations is deterministic and reversible: the equation of motion given by $H$ are differentiable since $H$ is twice differentiable.
-Therefore the equations are Lipschitz continuous and, by the Picard-Lindel\"{o}f theorem, a unique solution exists \cite{Grant}.
+Therefore the equations are Lipschitz continuous and, by the Picard-Lindel\"{o}f theorem, a unique solution exists~\cite{Grant}.
 
@@
 Note that we could have taken different approaches.
-For example, we could have appealed to information theory and required our invariant distributions to preserve Shannon's information entropy\cite{Shannon}, as no information should be gained or lost during a deterministic and reversible process.
+For example, we could have appealed to information theory and required our invariant distributions to preserve Shannon's information entropy~\cite{Shannon}, as no information should be gained or lost during a deterministic and reversible process.
 Or we could have appealed to statistical mechanics and required that the determinant of the covariance matrix be conserved, as a deterministic and reversible process should be defined at the same level of precision.
@@
 On the other hand, the change of state variables must preserve the ability to uniquely identify states at each instant in time.
-In this sense, time is not just like another state variable.\footnote{In relativity, for example, it is common to use foliations and the $3+1$ formalism to recover the special character of time when performing calculations or to gain better physical insight\cite{Gourgoulhon}.}
+In this sense, time is not just like another state variable.\footnote{In relativity, for example, it is common to use foliations and the $3+1$ formalism to recover the special character of time when performing calculations or to gain better physical insight~\cite{Gourgoulhon}.}
 
@@
 
-We recognize the equations as those for Hamiltonian mechanics on the extended phase space\cite{Synge,Lanczos,Struckmeier}.
+We recognize the equations as those for Hamiltonian mechanics on the extended phase space~\cite{Synge,Lanczos,Struckmeier}.
 
@@
 It was under that premise that we were able to show that the trajectories are continuous, that we could define $\omega$ and therefore $g$.
-Non-local quantum effects, such as the ones featured in EPR\cite{EPR} experiments, are not examples of deterministic and reversible evolution.
+Non-local quantum effects, such as the ones featured in EPR~\cite{EPR} experiments, are not examples of deterministic and reversible evolution.
 Therefore they do not have anything to violate: the maximum speed does not apply to their ``motion."\footnote{The very notion of velocity becomes ill defined in such cases precisely because we do not have a unique well-defined evolution.} On the other hand, only deterministic and reversible motion can carry information and can establish before/after relationships, therefore everything remains consistent.
@@
 Integrating we have $\mathcal{H}=\frac{1}{2m}(p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta)+V(x)$ where $V(x)$ is an arbitrary function.
-We can further constrain the Hamiltonian by choosing the natural parameterization for $s$ (also called unit speed \cite{Lee} or arc length parametrization) for which the modulus of the velocity is constant.
+We can further constrain the Hamiltonian by choosing the natural parameterization for $s$ (also called unit speed~\cite{Lee} or arc length parametrization) for which the modulus of the velocity is constant.
 We have $u^\alpha u_\alpha = - c^2 = \frac{1}{m^2}(p^\alpha-A^\alpha)(p_\alpha-A_\alpha)$.
@@
 As they have the same net effect, if we apply all three we go back where we started.
-This gives us a very intuitive classical particle mechanics analogue for the $CPT$ theorem \cite{Weinberg}.
+This gives us a very intuitive classical particle mechanics analogue for the $CPT$ theorem~\cite{Weinberg}.
 
@@
 In this context, we will call a \emph{fragment} an infinitesimal part of our classical electron.
-The Shannon entropy \cite{Shannon} associated with the distribution is finite as well.
+The Shannon entropy~\cite{Shannon} associated with the distribution is finite as well.
 $I(\rho_\mathcal{c}) = - \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq \wedge dp = I_0$.
@@
 \end{equation}
-where $\sigma^2$ is the variance, $\sigma$ the standard deviation, $\rho_{X,Y}= COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient \cite{Grimmet}.
+where $\sigma^2$ is the variance, $\sigma$ the standard deviation, $\rho_{X,Y}= COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient~\cite{Grimmet}.
 As $-1\leq\rho_{X,Y}\leq+1$, the total variance can be greater or smaller than the sum of the variance of the components.
@@
 For a quantum mixed state $\tilde{\rho} = \sum \tilde{\rho}_i |\psi_i\rangle \langle \psi_i |$, we have $I(\tilde{\rho}) = - \sum \tilde{\rho}_i \ln \tilde{\rho}_i - \tilde{\rho}_i I(\psi_i) = - \sum \tilde{\rho}_i \ln \tilde{\rho}_i = -tr(\tilde{\rho} \ln \tilde{\rho})$.
-And in fact this is the Von Neumann entropy \cite{Von Neumann} used in quantum information, that implicitly assumes all states provide the same amount of entropy (i.e.
+And in fact this is the Von Neumann entropy~\cite{Von Neumann} used in quantum information, that implicitly assumes all states provide the same amount of entropy (i.e.
 zero).
@@
 
-Naturally, we cannot give a complete description for the configuration of a fragment, therefore we cannot give a joint distribution for all fragment variables at the same time.\footnote{This recovers the notion of complementarity introduced by Bohr \cite{Bohr}.} We can, though, find a maximal set of fragment variables: one that can be used to provide the best description possible.
+Naturally, we cannot give a complete description for the configuration of a fragment, therefore we cannot give a joint distribution for all fragment variables at the same time.\footnote{This recovers the notion of complementarity introduced by Bohr~\cite{Bohr}.} We can, though, find a maximal set of fragment variables: one that can be used to provide the best description possible.
 While there isn't a unique maximal set, we can show that one such set is provided by the fragment variables $q^i : \mathcal{Q} \rightarrow \mathbb{R}$ that define the units required to describe the system.
@@
 	This notion may seem in contradiction to the current practice.
-For example, the S-matrix, one of the main devices for calculating scattering amplitudes and cross sections, is derived by assuming unitary evolution \cite{Weinberg}.
+For example, the S-matrix, one of the main devices for calculating scattering amplitudes and cross sections, is derived by assuming unitary evolution~\cite{Weinberg}.
 Doesn't it imply we are assuming that deterministic and reversible evolution is happening at every instant?

### 2017-02-16 2015CompleteDerivation/completeDerivation.tex
Fixing overfull boxes in journal style

@@
 \begin{prop}\label{prop:homeomorphism}
-A deterministic and reversible evolution map is a self-homeomorphism on the state space $\mathcal{T}_{\Delta t}:\mathcal{S} \rightarrow \mathcal{S}$.
+A deterministic and reversible evolution map is a self-\journal{\break}homeomorphism on the state space $\mathcal{T}_{\Delta t}:\mathcal{S} \rightarrow \mathcal{S}$.
 \end{prop}
@@
 
-Another consideration is that while the norm associated to $L^1(\mathcal{S})$ is $\int_{\mathcal{S}} |\rho| d\mu$, the expression $\int_{\mathcal{S}} \rho d\mu$ may be more physically meaningful in some cases.
+Another consideration is that while the norm associated to $L^1(\mathcal{S})$ is\journal{\break} $\int_{\mathcal{S}} |\rho| d\mu$, the expression $\int_{\mathcal{S}} \rho d\mu$ may be more physically meaningful in some cases.
 For proper states, the two are the same and they represent the total amount of material.
@@
 	We claim $E = H(t,q^i,p_i)$ for physical states.
-$\lambda = d_s t = - \partial_E \mathcal{H} = - \partial_{E} \lambda (H - E) - \lambda (-1)$ regardless of the choice of t and therefore for all $\lambda$.
+$\lambda = d_s t = - \partial_E \mathcal{H} =\journal{\break} - \partial_{E} \lambda (H - E) - \lambda (-1)$ regardless of the choice of t and therefore for all $\lambda$.
 $H - E = 0$.
@@
 \mathcal{H} &= \frac{1}{2m} ( p^i p_i - (E/c)^2 + m^2c^2) \\
+\ifjournal
  &= \frac{1}{2mc^2} (c \sqrt{m^2 c^2 + p^i p_i} + E) (c \sqrt{m^2 c^2 + p^i p_i} - E) \\
+\else
+ &= \frac{1}{2mc^2} (c \sqrt{m^2 c^2 + p^i p_i} + E)\\ &(c \sqrt{m^2 c^2 + p^i p_i} - E) \\
+\fi
 \lambda &= \frac{1}{2mc^2} (c \sqrt{m^2 c^2 + p^i p_i} + E)
@@
 $F_{[\hat{q}^i]}$ is a linear operator.
-$F_{[\hat{q}^i]}$ is self-adjoint as $\int (F_{[\hat{q}^i]}(\psi(\hat{q}^i)))^\dagger \psi(\hat{q}^i) d^{n}\hat{q} = \int (f(\hat{q}^i) \psi(\hat{q}^i))^\dagger \psi(\hat{q}^i) d^{n}\hat{q} = \int \psi^\dagger(\hat{q}^i) f(\hat{q}^i) \psi(\hat{q}^i) d^{n}\hat{q} = \int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]} (\psi(\hat{q}^i)) d^{n}\hat{q}$.
-
-	We claim that for any fragment variable $f$ there exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbb{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i)) d^nq$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
+$F_{[\hat{q}^i]}$ is also self-adjoint since\journal{\break} $\int (F_{[\hat{q}^i]}(\psi(\hat{q}^i)))^\dagger \psi(\hat{q}^i) d^{n}\hat{q} = \int (f(\hat{q}^i) \psi(\hat{q}^i))^\dagger \psi(\hat{q}^i) d^{n}\hat{q} = \int \psi^\dagger(\hat{q}^i) f(\hat{q}^i) \psi(\hat{q}^i) d^{n}\hat{q} = \int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]} (\psi(\hat{q}^i)) d^{n}\hat{q}$.
+
+	We claim that for any fragment variable $f$ there exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbb{C})$ such that the integral $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i)) d^nq$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
 Let $f$ be a fragment variable.
@@
 Let $\psi(p_i) = \Upsilon^{q^i}_{p_i} (\psi(q^i))$ be the wave function over momentum space.
-$(Q^i_{[p_i]})^m \psi(p_i) = (+\imath \hbar \partial_{p_i})^m \psi(p_i)$ represents the $m^{th}$ power of the $q^i$ operator applied to the given state.
+$(Q^i_{[p_i]})^m \psi(p_i) =\journal{\break} (+\imath \hbar \partial_{p_i})^m \psi(p_i)$ represents the $m^{th}$ power of the $q^i$ operator applied to the given state.
 As the quantity is physically meaningful, such an operation must be defined for any $m$ therefore $\psi(p_i) \in C^\infty(\mathcal{P})$.
@@
 Also $\Lambda_V (\rho) = \Lambda_U (\rho) + \Lambda_{V\backslash U} (\rho) = \int_U \rho d \mu_U + \int_{V \backslash U} \rho d \mu_{\overline{V \backslash U}} $.
-Therefore $\int_{U} \rho_{\mathcal{c}} d \mu_U = \int_{U} \rho_{\mathcal{c}} d \mu_V$.
+Therefore\journal{\break} $\int_{U} \rho_{\mathcal{c}} d \mu_U = \int_{U} \rho_{\mathcal{c}} d \mu_V$.
 $d\mu$ is unique and does not depend on the choice of neighborhood.
@@
 Let $f,g,h$ be differentiable functions such that $h(q)=f(q)g(q)$.
-$K_q(ln(h(q)))=\partial_q ln(h(q)) K_q(q) = \frac{1}{h} \partial_q h K_q(q) = \frac{1}{h} K_q(h(q)) = K_q(ln(f(q)g(q))) = K_q(ln(f(q))) + K_q(ln(g(q))) = \frac{1}{f} K_q(f(q)) + \frac{1}{g} K_q(g(q))$.
+$K_q(ln(h(q)))=\partial_q ln(h(q)) K_q(q) = \frac{1}{h} \partial_q h K_q(q) = \frac{1}{h} K_q(h(q)) = K_q(ln(f(q)g(q)))\journal{\break} = K_q(ln(f(q))) + K_q(ln(g(q))) = \frac{1}{f} K_q(f(q)) + \frac{1}{g} K_q(g(q))$.
 Multiplying by $h=fg$ we have $K_q(h(q)) = g K_q(f(q)) + f K_q(g(q))$.

### 2017-02-16 2015CompleteDerivation/pictures/temporalDof.tex
Updating colors for figures

--- 2016-09-30 2015CompleteDerivation/pictures/temporalDof.tex
@@
 		\newcommand{\xvar}{20};			% space between the two
+
+		% styles
+		\tikzstyle{triangle}=[ultra thick,fill=white,opacity=0.5];
+		\tikzstyle{projection} = [fill=yellow,fill opacity=0.1,text opacity =1,text=black];
 
@@
 		% drawing the first shape
-		\filldraw[black,fill=blue,opacity=0.5] (0,0) -- ($(p2)$) -- ($(p3)$) -- cycle;
+		\filldraw[style=triangle] (0,0) -- ($(p2)$) -- ($(p3)$) -- cycle;
 
-		\filldraw[black,fill=blue,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={0:(0,0)}] (0,0) rectangle ($(p7)$) node[pos=.5] {\large $\omega$};
+		\filldraw[style=projection,rotate around={0:(0,0)}] (0,0) rectangle ($(p7)$) node[pos=.5] {\large $\omega$};
 
-		\draw[black,fill=blue,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar:(0,0)}] (0,0) rectangle ($(p8)$) node[pos=.5, rotate=\thetavar] {\large $dq\wedge dp$};
+		\draw[style=projection,rotate around={\thetavar:(0,0)}] (0,0) rectangle ($(p8)$) node[pos=.5, rotate=\thetavar] {\large $dq\wedge dp$};
 
-		\draw[black,fill=blue,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={0:(0,0)}] ($(p2)$) rectangle ($(p5)$) node[pos=.5] {\large $dt\wedge dE$};
+		\draw[style=projection,rotate around={0:(0,0)}] ($(p2)$) rectangle ($(p5)$) node[pos=.5] {\large $dt\wedge dE$};
 
@@
 		%drawing the shape on the right
-		\filldraw[black,fill=green,opacity=0.5] ($(h0)$) -- ($(h2)$) -- ($(h3)$) -- cycle;
+		\filldraw[style=triangle] ($(h0)$) -- ($(h2)$) -- ($(h3)$) -- cycle;
 
-		\filldraw[black,fill=green,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={0:($(h0)$)}] ($(h0)$) rectangle ($(h7)$) node[pos=.5] {\large $\omega$};
+		\filldraw[style=projection,rotate around={0:($(h0)$)}] ($(h0)$) rectangle ($(h7)$) node[pos=.5] {\large $\omega$};
 
-		\draw[black,fill=green,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\phivar:($(h0)$)}] ($(h0)$) rectangle ($(h8)$) node[pos=.5, rotate=\phivar] {\large $dq\wedge dp$};
+		\draw[style=projection,rotate around={\phivar:($(h0)$)}] ($(h0)$) rectangle ($(h8)$) node[pos=.5, rotate=\phivar] {\large $d\hat{q}\wedge d\hat{p}$};
 
-		\draw[black,fill=green,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={0:($(h0)$)}] ($(h2)$) rectangle ($(h5)$) node[pos=.5] {\large $dt\wedge dE$};
+		\draw[style=projection,rotate around={0:($(h0)$)}] ($(h2)$) rectangle ($(h5)$) node[pos=.5] {\large $d\hat{t}\wedge d\hat{E}$};
 

### 2017-02-16 2015CompleteDerivation/pictures/independentDof.tex
Updating colors for figures

@@
 
+	% styles
+	\tikzstyle{triangle}=[ultra thick,fill=white,opacity=0.5];
+	\tikzstyle{projection} = [fill=yellow,fill opacity=0.1,text opacity =1,text=black];
+
 	% drawing the right basic triangle 
-	\filldraw[fill=red,opacity=0.5] (0,0) -- (\xvar,0) -- (\xvar,\yvar) -- cycle;
+	\filldraw[style=triangle] (0,0) -- (\xvar,0) -- (\xvar,\yvar) -- cycle;
 
-	\filldraw[fill=red,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black] (0,0) rectangle (\xvar,-\xvar) node[pos=.5] {\large $ {dq^1}\wedge {dp_1}$};
+	\filldraw[style=projection] (0,0) rectangle (\xvar,-\xvar) node[pos=.5] {\large $ {dq^1}\wedge {dp_1}$};
 
-	\filldraw[fill=red,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black] (\xvar,0) rectangle (\xvar + \yvar, \yvar) node[pos=.5] {\large ${dq^2}\wedge {dp_2}$};
+	\filldraw[style=projection] (\xvar,0) rectangle (\xvar + \yvar, \yvar) node[pos=.5] {\large ${dq^2}\wedge {dp_2}$};
 
-	\filldraw[fill=red,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar:(0,0)}]  (0,0) rectangle ({\omegalength}, {\omegalength}) node[pos=.5, rotate={\thetavar}] {\large $\omega$};
+	\filldraw[style=projection,rotate around={\thetavar:(0,0)}]  (0,0) rectangle ({\omegalength}, {\omegalength}) node[pos=.5, rotate={\thetavar}] {\large $\omega$};
 
 	% drawing the complex right triangle  
-	\filldraw[black,fill=yellow,opacity=0.5,rotate around={\thetavar:(\dvar,0)}] ({\dvar + \omegalength},0) -- ({\dvar},0) -- ++({-\thetaprimevar}: {\xprimevar}) -- ++({\phiprimevar}:{\yprimevar});
+	\filldraw[style=triangle,rotate around={\thetavar:(\dvar,0)}] ({\dvar + \omegalength},0) -- ({\dvar},0) -- ++({-\thetaprimevar}: {\xprimevar}) -- ++({\phiprimevar}:{\yprimevar});
 
-	\filldraw[black,fill=yellow,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar - \thetaprimevar:(\dvar,0)}]  (\dvar,0) rectangle (\dvar + \xprimevar,-\xprimevar) node[pos=.5, rotate=\thetavar-\thetaprimevar] {\large $d\hat{q}^1\wedge d\hat{p}_1$};
+	\filldraw[style=projection,rotate around={\thetavar - \thetaprimevar:(\dvar,0)}]  (\dvar,0) rectangle (\dvar + \xprimevar,-\xprimevar) node[pos=.5, rotate=\thetavar-\thetaprimevar] {\large $d\hat{q}^1\wedge d\hat{p}_1$};
 
-	\filldraw[black,fill=yellow,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar - \thetaprimevar:(\dvar,0)}]  ({\dvar + \xprimevar},0) rectangle ({\dvar + \xprimevar + \yprimevar}, {\yprimevar}) node[pos=.5, rotate={\thetavar-\thetaprimevar}] {\large $d\hat{q}^2\wedge d\hat{p}_2$};
+	\filldraw[style=projection,rotate around={\thetavar - \thetaprimevar:(\dvar,0)}]  ({\dvar + \xprimevar},0) rectangle ({\dvar + \xprimevar + \yprimevar}, {\yprimevar}) node[pos=.5, rotate={\thetavar-\thetaprimevar}] {\large $d\hat{q}^2\wedge d\hat{p}_2$};
 
-	\filldraw[black,fill=yellow,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar:(\dvar,0)}]  (\dvar,0) rectangle ({\dvar + \omegalength}, {\omegalength}) node[pos=.5, rotate=\thetavar] {\large $\omega$};
+	\filldraw[style=projection,rotate around={\thetavar:(\dvar,0)}]  (\dvar,0) rectangle ({\dvar + \omegalength}, {\omegalength}) node[pos=.5, rotate=\thetavar] {\large $\omega$};
 

### 2017-02-16 2017HamiltonianInformation/hamiltonianInformation.tex
Updating colors for figures

--- 2017-02-16 2015CompleteDerivation/completeDerivation.tex
@@
 \newif\ifjournal
-\journaltrue
-%\journalfalse
+%\journaltrue
+\journalfalse
 
@@
 
-\title{From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics}
-\author{Gabriele Carcassi, Christine A. Aidala, David J. Baker, Lydia Bieri}
+\title{DRAFT \\ Hamiltonian mechanics is conservation of information}
+\author{Gabriele Carcassi, Christine A. Aidala}
 
 \ifjournal
-	\institute{G. Carcassi, C. A. Aidala \at
-		Physics Department, University of Michigan, Ann Arbor, MI 48109 \\
-		\email{carcassi@umich.edu}
-		\and
-		D. J. Baker \at
-		Philosophy Department, University of Michigan, Ann Arbor, MI 48109
-		\and
-		L. Bieri \at
-		Mathematics Department, University of Michigan, Ann Arbor, MI 48109 \\
+	\institute{G. Carcassi \at
+		University of Michigan, Ann Arbor, MI 48109 \\
+		\email{carcassi@umich.edu}           %  \\
 	}
@@
 
-\ifjournal
-	% As the page margins are so large, the full title does not fit
-	\titlerunning{From physical principles to classical and quantum particle mechanics}
-\fi
+%\ifjournal
+%	% As the page margins are so large, the full title does not fit
+%	\titlerunning{From physical principles to classical and quantum particle mechanics}
+%\fi
 
@@
 \begin{abstract}
-The aim of this work is to show that particle mechanics, both classical and quantum, Hamiltonian and Lagrangian, can be derived from few simple physical assumptions.
-Assuming deterministic and reversible time evolution will give us a dynamical system whose set of states form a topological space and whose law of evolution is a self-homeomorphism.
-Assuming the system is infinitesimally reducible, that giving the state and the dynamics of the whole system is equivalent to giving the state and the dynamics of its infinitesimal parts, will give us a classical Hamiltonian system.
-Assuming the system is irreducible, that giving the state and the dynamics of the whole system tells us nothing about the state and the dynamics of its substructure, will give us a quantum Hamiltonian system.
-Assuming kinematic equivalence, that studying trajectories is equivalent to studying state evolution, will give us Lagrangian mechanics and limit the form of the Hamiltonian/Lagrangian to the one with scalar and vector potential forces.
-\keywords{Hamiltonian mechanics \and Lagrangian mechanics \and Quantum mechanics \and Fundamental assumptions}
+\textbf{This manuscript is a work in progress.} Ideas are constantly reshaped to find more precise and elegant arguments.
+It is provided as is to stimulate discussion.
+\textbf{Make sure you have the latest version from http://assumptionsofphysics.org}
+
+In this work we show that the canonical transformations of classical Hamiltonian mechanics are exactly the transformations that preserve information.
 \end{abstract}
@@
 
-The past century has seen a resounding success in the use of mathematics in fundamental physics.
-In many cases, mathematical ideas were what allowed progress in theoretical physics and ultimately led to experimental discoveries.
-While these ideas are indeed useful---anything that can advance our understanding is welcome---they cannot, however, be the whole story.
+\section{Something}
 
-The issue is that mathematical structures are not enough to characterize what physical system is being studied.
-For example, the same mathematical framework for linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems~\cite{Roland,Barron,Borutzky}.
-The mathematical treatment is the same because it captures aspects of the description that are common to all cases; that indeed is its power.
-But for the very same reason, we cannot infer what system is being described if all we are given is the mathematical model.
-In the jargon of linguistics/computer science/philosophy~\cite{Freidin,Friedman}: math captures the syntax (the relationships between the objects) but does not capture the semantics (the meaning of each object).
-
-The real trouble comes when the mathematical description essentially becomes the foundation of a physical theory.
-While some branches of physics, such as Newtonian mechanics, thermodynamics and special relativity, are founded on physical laws or assumptions, others, such as Hamiltonian, Lagrangian and quantum mechanics, start by setting their mathematical structure.
-Nothing tells us why there should be  a Lagrangian function whose path integral is minimized during the motion, or when a state is described by conjugate pairs $(q^i, p_i)$ instead of a vector in a complex inner product space.
-Those are taken as given.
-
-This means that if we have a system in front of us, we do not have a rigorous way to conceptually establish whether the system is Hamiltonian or not, whether it is classical or quantum.
-We have some heuristics we can apply, but, at the end the day, a system is Hamiltonian because it follows the Hamiltonian framework.\footnote{The lack of a properly defined, and uniquely accepted, semantics is of course very evident in quantum mechanics with its numerous interpretations.}
-
-Unsatisfied with this situation, we started creating a sort of math-to-physics dictionary so that each mathematical concept could be clearly associated to a crisp physical concept.
-As we pushed further back toward the basic mathematical definitions, we ended up with physical concepts of a more general nature and found that, in some cases, a single physical idea could have more than one mathematical consequence.
-That is: by putting physics at the center, we could achieve a unification of ideas that was not evident in the math alone.
-The unexpected success of the approach led us to think that, perhaps, the picture we were developing was a better framework to understand and link different areas of basic science and mathematics.
-This work is the systematized result of this endeavor.
-
-The main goal is to show that it is possible to derive classical/quantum Hamiltonian/Lagrangian particle mechanics starting from few physical assumptions.
-We do \emph{not} claim that we have perfectly achieved this result, with no room for improvement.
-We \emph{do} claim that the whole architectural framework is solid enough to be convinced that not only is this indeed possible, but it allows a more unified picture that facilitates intuitive connections within and across different areas of physics and mathematics.
-It enables a consistent narrative across several domains of knowledge.
-
-Rederiving classical, quantum and (to some extent) relativistic mechanics within the same work is therefore crucial to show that the concepts we develop are truly general, that they can be applied to many different cases.
-If we could derive classical Hamiltonian mechanics on assumptions that are inapplicable to the quantum case, what good would it do?
-Quantum systems are Hamiltonian systems, so clearly we wouldn't have understood what makes a system Hamiltonian.
-This breadth of applicability is what makes the work valuable.
-
-As a brief overview, there are four assumptions we will consider.
-Assuming deterministic and reversible evolution gives us a dynamical system: states and a law of evolution.
-Additionally assuming infinitesimal reducibility, studying the whole system is equivalent to studying its infinitesimal parts, gives us Hamiltonian mechanics.
-Further assuming kinematic equivalence, trajectories in space tell us everything about the state of the system, gives us Lagrangian mechanics and constrains the motion to massive particles under scalar/vector potential forces.
-Relativistic motion arises when studying time dependent laws of evolution with no further assumptions.
-Alternatively, as the second assumption we can instead take irreducibility, studying the whole system tells us nothing about its parts, which leads to quantum (Hamiltonian) mechanics.
-
-As a more detailed overview: in section \ref{sec:fundamental_model} we describe in general what it means to study a physical system and how the assumption of determinism and reversibility plays into it.
-We develop an abstract conceptual model that we use when discussing the later assumptions.
-
-In section \ref{sec:dynamical_systems} we show how the physical concept of physical distinguishability leads to the mathematical concept of topological spaces.
-Deterministic and reversible evolution is an invertible continuous map (i.e.
-a homeomorphism) as it needs to preserve what is physically distinguishable.
-
-In section \ref{sec:reducibility} we introduce the notion of a classical material, one for which we assume infinitesimal reducibility.
-The state of an arbitrary amount of material is a distribution over the states of the infinitesimal parts, the classical particles.
-As such a distribution must be invariant under the arbitrary choice of units, we see that for each variable $q^i$ that defines a unit there is a corresponding variable $k^i$ defined on the inverse of that unit.
-Classical particle states can be identified by conjugate pairs of state variables $(q^i, p_i)$ (i.e.
-a point in the cotangent bundle $T^*\mathcal{Q}$) and the number of possible initial conditions for each pair is given by $dq^i \wedge dp_i$ (i.e.
-the canonical symplectic form $\omega$).
-During evolution, the distribution (and marginal distributions) must be mapped point by point which leads to Hamiltonian mechanics (i.e.
-a symplectomorphism on $T^*\mathcal{Q}$).
-
-In section \ref{sec:relativity} we show that a relativistic version of Hamilton's equations arises naturally when we consider time dependent laws of evolution for distributions of classical material.
-This also gives the notion of deterministic and reversible evolution a characterization that is independent of any transformation that mixes time and state variables $(q^i, p_i)$.
-Moreover it gives us a classical equivalent for anti-particle states.
-
-In section \ref{sec:Lagrangian} we introduce the assumption of kinematic equivalence, studying the motion in space is equivalent to studying the evolution of the state.
-Under that assumption there is a link between the state variables $(q^i, p_i)$ and the kinematic variables $(x^i, u^i)$ (i.e.
-position and velocity).
-This link allows us to express the number of possible initial conditions for a degree of freedom (d.o.f.) in terms of $dx^i$ and $du^i$, leading to the metric tensor $g$.
-Such an expression also allows us to constrain the Hamiltonian to the one for massive particles under scalar/vector potential forces and an equivalent Lagrangian formulation.
-
-In section \ref{sec:quantum_prelude} we introduce some ideas in classical mechanics to more gently transition to quantum mechanics.
-We see how distributions over phase space already have features that are typically associated with quantum systems and we will show how to construct a classical uncertainty principle that helps to build intuition used in the later section.
-
-In section \ref{sec:irreducibility} we introduce the notion of an irreducible material, one for which studying the whole distribution does not tell us anything about the motion of its infinitesimal parts, the fragments.
-A unit amount of this material will correspond to a quantum particle.
-As this means that the states of such a material are invariant under permutations of the configurations of the fragments, we show that the state space of an irreducible material is a complex vector space.
-These states can still be compared to each other in a way that gives the state space an inner product.
-Such comparisons will be preserved by deterministic and reversible evolution which leads to Schroedinger's equation (i.e.
-a unitary transformation).
-
-In section \ref{sec:discussion} we present a few higher level remarks about the work.
-
-\section{Organization and style}
-
-The overall layout of this work is a balance between the need to show that the derivation follows, which requires starting from basic principles and proceeding using formal arguments, and the need to give an intuitive understanding, which often requires starting from specific cases and then generalizing.
-
-The deductive aspect is formally organized into the following:
-\begin{description}
-  \item[Assumptions] these characterize the physical system we are studying and constitute the premise of our discussion.
-A \textbf{rationale} follows each assumption, which uses physical and sometimes philosophical arguments to motivate why (or not) such an assumption makes sense in a particular case.
-
-  \item[Propositions] these capture the properties of our physical objects into mathematical language.
-A \textbf{justification} follows each proposition to show the necessity of such a characterization.
-While it is a formal argument, it is not a strictly mathematical proof as it contains physical arguments.
-
-  \item[Theorems] these are pure mathematical statements that are used to simplify propositions.
-A mathematical \textbf{proof} follows each theorem.
-No mathematical breakthrough should be expected as we mostly use well known results from different fields.
-Theorems are grouped in the appendix.
-\end{description}
-
-As these sections are clearly demarcated, they allow us to make sure that each justification only references past propositions and avoid circular arguments.
-It also makes it easier to read the work at the desired level of detail.
-Interspersed between each step of the more formal treatment, there are numerous discussions that conceptually explain motivations and results in a more approachable way.
-These can use more concrete and intuitive examples which may, in some cases, make forward references.
-As both the formal and more intuitive parts need to cover the same material, this will inevitably lead to some repetition.
-
-The work uses ideas and tools from a broad range of mathematical and scientific subjects: topology~\cite{LeeTM}, measure theory~\cite{Halmos}, differential geometry (symplectic and Riemannian)~\cite{Lee}, group theory and vector spaces~\cite{Lang,Young}, statistics~\cite{Grimmet}, information theory~\cite{Pierce}, Hamiltonian mechanics, Lagrangian mechanics~\cite{classical_dynamics}, quantum mechanics~\cite{Landau, Weinberg}, thermodynamics~\cite{Van Ness}, special relativity \cite{French} and so on.\footnote{The references given provide the reader a list of resources consistent with the development of this work.} Using ideas from all these areas is necessary and valuable, but it does create a number of problems.
-
-The first one is terminology and prerequisite knowledge.
-Since we touch numerous areas, we cannot expect all readers to be equally comfortable with all the technical terms.
-Yet we cannot introduce all of them either.
-The compromise we adopted is to keep the informal discussion more accessible while keeping the formal derivation more precise, without needing more than the basic definitions and few major results within each specific area.
-This may leave an expert in a particular field not fully satisfied but, as stated in the introduction, our goal is to lay out the general picture.
-
-The second problem is notation.
-As each field has developed its own set of conventions, the common usage is now an empty set.
-We ended up with a compromise that hopefully feels natural enough to a broad audience of physicists, mathematical physicists and philosophers of physics.
-Within formal parts we may use notation and terminology more specific to the appropriate area of math.
-
-We have tested the prose on a number of colleagues and students in physics, philosophy and mathematics to confirm its accessibility at different levels and believe we have struck a reasonable compromise.
-
-\section{On studying a physical system}
-\label{sec:fundamental_model}
-
-% Status: Conceptual work done. Text ready. Incorporated Dave, Isaac and Christine feedback
-
-Our first task is to develop a conceptual model that applies to all realms of physics we'll be considering: classical, statistical and, later, quantum mechanics.
-We will take the standard picture of system plus environment and extend it to differentiate between the state of the system (i.e.
-the aspects under study) from the unstated part of the system (i.e.
-the aspects missing from our description).
-
-We will assume that the state evolves according to a deterministic and reversible law (i.e.
-for each present state there is one and only one future state).
-While the unstated part does not influence the state evolution, we'll see how it constrains what states are available and what type of description can or cannot be given to the system.
-
-\subsection{States and their evolution}
-
-We start by fixing a \emph{physical system}, meaning something we can interact with and perform measurements on (e.g.
-a planet, a fluid, ...).
-We call \emph{environment} everything else.
-We set what particular aspect we want to study (e.g.
-the motion around a star, the flow in a pipe, ...).
-We call \emph{state} a physically distinguishable configuration of the aspect under study at a particular time (e.g.
-position/momentum of center of mass, velocity field, ...).
-Since the state does not, in general, exhaust the description of the system, a part remains \emph{unstated}, and as such we'll call it, for lack of a better word (e.g.
-the chemical composition, the motion of each of its molecules, ...).
-Note that the environment plays an essential role here as it's what allows us to define two states as physically distinguishable: we can find an external process (i.e.
-part of the environment) whose outcome changes depending on the different state.
-As such processes are what we can use to perform measurements, we consider the experimental apparatus (and us performing measurements) independent of the system, part of the environment.\footnote{This is true even in the case of general relativity, where we can imagine multiple researchers on small spaceships collecting data without greatly influencing the motion of stars and planets.
-The case where the physical system is the whole universe and there is no environment presents practical problems  and conceptual challenges that the current physical theories do not seem to be equipped to address, and therefore will be absent from our discussion.
-For example, what physical device can we use to store and process the state of the whole universe to make predictions and compare?
-How do we define physically distinguishable?
-Do the physical laws determine which measurements we are going to make and does that limit what is actually distinguishable?}
-
-In this context, we call the evolution of the system \emph{deterministic} if the state at a given time uniquely identifies states at future times, and \emph{reversible} if it uniquely identifies states at past times.
-While this is a common enough definition, we need to be clear how this applies to the unstated part.
-Note that the concept of determinism outlined here is context dependent because the state itself represents only the part of the system that we choose to (or can) describe.
-In this sense, the unstated part is \emph{always} non-deterministic (and non-reversible) in the sense that the state of the system does not determine its evolution.
-For example, suppose we define the state as position and momentum of the center of mass of a cannonball.
-Suppose that the evolution is deterministic on that state.
-What does it tell us about the cannonball temperature, or about the motion of each of its atoms?
-Nothing.
-In this sense, the unstated part is non-deterministic and non-reversible.
-Could we extend the state and the laws of evolution to account for temperature?
-Yes, but that would be a different evolution defined on a different state.
-Does it mean the unstated part is always evolving randomly?
-Not at all.
-The temperature may remain constant throughout the motion of the cannonball.
-Yet we wouldn't know, since we are not studying it: the evolution is deterministic and/or reversible only as far as the state is concerned.
-With this in mind, we will restrict ourselves to the cases where the following is valid:
-
-\begin{assump}[Determinism and reversibility]\label{ass:determinism}
-The state of the physical system under study undergoes deterministic and reversible evolution.
-\end{assump}
-
-\begin{rationale}
-As it is an assumption, we first need to discuss when it is valid.
-More specifically, we need to understand that the non-deterministic/non-reversible evolution of the unstated part plays as much of a fundamental role as the deterministic/reversible evolution of the state.
-In fact, the non-deterministic part contributes in determining what states are available to the system.
-
-Suppose we study the motion of a cannonball; its state under gravitational and (inertial) inertial forces will be properly described by the position and momentum of the center of mass.
-While light and air molecules may scatter off its surface unpredictably, its trajectory is not greatly affected as it is a massive rigid body.
-Suppose we study the motion of a small particle, small enough that the random scattering does influence the trajectory and it undergoes Brownian motion: its state will be a probability distribution for position and momentum of the center of mass.
-Gravitational and inertial forces have not changed, yet the states have changed from ``pure" to statistical ensembles.
-In other words, the set of states must be closed under both the deterministic evolution of the state and the non-deterministic evolution of the unstated part.
-If the Brownian motion is not negligible, we do not end in a well defined position/momentum pair, even if we start from one.
-
-A similar more drastic effect: consider a book and its motion under gravitation and inertial forces, its state being the position and momentum of the center of mass.
-As we increase the temperature of the air around the book, its motion remains unaffected until, at some point, the book burns.
-Clearly, the non-deterministic evolution has pushed one of the states outside the set of states, to the point that the system is no longer recognizable.
-
-As we have hinted, sometimes the state is identified by a distribution (either statistical or actual).
-Even in this case, the state can be deterministic and reversible.
-That is, given the distribution at one time we can determine the distribution at future times.
-The shape and the parameters of the distribution can be deterministic, even if the evolution of the parts are not as they fall within the unstated part.
-Note that we \emph{cannot} assume trajectories and states are always defined for the unstated part, as this includes also the unknown unknowns.
-We will return to this aspect when discussing quantum systems.
-
-It should also be clear that what constitutes state and unstated part does not depend only on the system under study, but also on the processes we are considering.
-In some circumstances, the chemical composition of a fluid may be relevant, in others it may not.
-The choices of environment, state and unstated part are not independent from each other.
-By choosing a particular set of states, we are not only saying that the state evolution is well approximated by a deterministic/reversible map from initial to final state, but we are also saying that the non-deterministic/non-reversible evolution of the unstated part does not change the nature of the system, and processes that do not satisfy these conditions are not under consideration.
-
-As with all assumptions, we should also ask whether it is necessary.
-That is, could we define a set of physically distinguishable states and yet have no deterministic and reversible processes defined on them?
-The claim is that this assumption is indeed needed, as without it we cannot properly define states or write useful physics laws.
-We can provide different arguments that point in the same direction.
-
-First, to be able to identify the system, we must be able to tell it apart from anything else.
-Intuitively, we can distinguish between two chairs because we can move the first to another room and sit on it without having touched the second.
-We can manipulate the state of the first system without affecting the second, and vice-versa.
-So, to identify a system it has to be sufficiently isolated from everything else.
-This means that the system future and past states are with good approximation determined only by its own state: the state undergoes deterministic and reversible evolution.
-
-Second, the aim of physics is to write laws that can be used to make predictions that can be validated experimentally.
-If I drop an anvil from a tower, it will accelerate at $9.81$~m/s$^2$; if I want the anvil to reach the ground at $x$~m/s I have to drop it from $y$~m.
-To the extent that we want to make predictions in time, we need to have a correspondence between initial and final states.
-
-Third, operationally we must reliably prepare and measure states.
-That is, we need a process for which the input settings of our preparing device determine the outgoing state of the system; and a process for which the incoming state of the system can be reconstructed by the output of the measuring device.
-That is, our system must, at least in some cases, be able to participate in a deterministic and reversible process with the preparing and measuring device.
-Without it we wouldn't be able to calibrate our experimental apparatus.
-
-Fourth, to be able to ascribe a property to a system we need to claim that, at least for a finite interval of time, the system either held or did not hold such property.
-That is, there is a deterministic and reversible process for that finite period of time for such property.
-
-This link between state definition and deterministic processes should not be too surprising as the state, in the context of control theory and dynamical systems theory, is often defined as \emph{the set of variables needed to determine the future evolution of the system} \cite{Katok, Sontang}.
-As we saw before, this applies also to statistical processes: the distribution (the ensemble) as a whole can indeed be calculated, measured and prepared.
-We can also describe the evolution of each element of the distribution provided that: we have a way to isolate it and study it under deterministic and reversible motion (so that we can define microstates); the non-deterministic motion does not alter the system (the set of microstates is preserved by the evolution).
-
-As with many assumptions, we should stress that it's an idealization: it can never be completely achieved in practice.
-A system can be prepared or measured up to a certain level of precision.
-Perfect determinism and isolation of a system is impossible both practically (e.g.
-black-body radiation, gravity, ...) and conceptually (e.g.
-if the system is perfectly isolated, we cannot interact with it: how can it be physically distinguished?).
-It's a simplifying assumption that can only be taken if the environment and the internal dynamics of the system interact in such a way that they little affect and are little affected by the aspect we are studying.
-As we saw before, for example, assuming that the state consists of the position and momentum of the center of mass requires assuming that the Brownian motion of the body is negligible.
-
-Yet, this is a fundamental assumption in the sense that it is needed.
-If a particular set of states does not satisfy deterministic and reversible evolution under certain conditions, what we do is to keep at it until we find a set that does.
-That is, we work to restore the assumption.
-Finding new sets of states with new laws of evolution is, in fact, what leads to new physics.
-We therefore call \emph{fundamental model of physics} the triad of state, unstated part and environment, with the assumption that the state may undergo deterministic and reversible evolution, and the unstated part undergoes non-deterministic non-reversible evolution that does not alter the set of states.
-\end{rationale}
-
-\section{States and state space}
-\label{sec:dynamical_systems}
-
-% Status: First subsection. Conceptual work done. Text ready. Incorporated Christine feedback
-% Second subsection. Conceptual work almost done. Need to finish the text
-
-%Open questions: smoothness (obviously required to define densities later, is it required in general? What does it means that physical processes/measurements don't give a smooth topology?)
-
-We now proceed to characterize states and physical distinguishability in more precise terms so that we can capture their description mathematically.
-What we'll see is that the outcomes of all physical processes that can be used to gain information about the system, that can be used to perform a measurement, induce a topology on the set of states that is at least Hausdorff.
-That is, physical distinguishability is mathematically captured by topological distinguishability.
-Deterministic and reversible evolution will then preserve the topology, and they will be mathematically captured by self-homeomorphisms.
-
-We'll focus on state spaces that can be described by a set of independent state variables, either discrete or continuous, and see what can be said in general on the evolution of state variables.
-
-\subsection{States and topology}
-
-As the term ``measurement" has become particularly loaded, let's first characterize what we mean by physical distinguishability in our context.
-
-Consider the motion of a cannonball under inertial and gravitational forces.
-Light will scatter off of it; as it lands, the ground will be deformed and the impact will make the temperature rise slightly.
-Those external physical processes, which happen no matter what we do, can be used to distinguish the motion of the cannonball as their outcomes are correlated.
-Therefore we can learn the position by looking at the reflected light, and learn the final kinetic energy by looking at the deformation of the ground.
-In principle, any external process that has a correlation with the states under study can be used to perform a measurement, and any measurement is based on such a process.
-That is, for us a measurement is simply a physical process that we can use to distinguish states.
-Setting up an experiment means choosing a particular process with desired outcomes and forcing the system under study to interact with it one or more times.
-After that, there is no special role played by the ``observer" in making outcomes come about.
-
-This external process may be quite complicated: when a particle enters a calorimeter, a shower of particles is produced, photons are captured and are directed to photomultipliers, a current is read out, the current is then digitized, and so on.
-Some processes interfere with the system, they affect its dynamics, and others are destructive, the system no longer can be described by the original set of states.
-A tracking chamber is an example of the first (the magnetic field curves the motion of a charged particle); burning a substance to determine its caloric content is an example of the latter.
-Therefore intimate knowledge of the process is always needed to ensure that one makes the proper link between outcomes and the \emph{original} states, and properly accounts for systematic uncertainties that would skew that link.
-
-Repeatability is also fundamental.
-First, to make sure the process is indeed correlated to the states.
-Second, because ``a single take does not a measurement make." One has to gather enough statistics.
-Note that the number of takes influences the outcomes: with greater statistics the precision and number of distinguishable cases increases.
-Therefore the processes, as we defined them, may require repeated interactions with similarly prepared states.
-They may even be a combination of different kinds of interactions that, taken all together, create a set of distinguishable outcomes.
-But in the end, however complicated it is, the conceptual model remains the same: each process has a set of possible outcomes, and each possible outcome will be associated with a set of states consistent with that outcome.
-For example, if the cannonball deformed the ground by this amount then its kinetic energy at impact was within this range; if the electron follows a certain path, then its state is among the ones that have spin $+1/2$.
-
-Note that the outcome of the process (e.g.
-the electron followed a particular path) is conceptually distinct from the set of states compatible with it (the states associated with the path) and the way we label them (spin=$+1/2$).
-Yet,  we'll conflate the concepts as this makes the language closer to experimental physics (``the outcome of the measurement is spin=$+1/2$").
-Context will allow us to recover the distinction.
-
-However precise our measurements are, we can only gather a finite amount of statistics and each outcome is expressed by a finite set of digits; therefore, the set of outcomes is countable.\footnote{The information provided by the process as measured by Shannon's entropy\cite{Shannon,Jaynes} is finite.} Note that different outcomes for the same process can overlap.
-For example, $4.12 \pm 0.05$ cm and $4.13 \pm 0.05$ cm are both legitimate possible outputs of the same measurement device.
-But since all states must be distinguishable, given two arbitrary states there must be a process precise enough to tell them apart.
-That is, the potential outcomes associated with the two states do not overlap.
-For example, $4.12 \pm 0.0005$ cm and $4.13 \pm 0.0005$ cm do not overlap anymore.
-
-We can also conceptually combine two different processes into a single one.\footnote{That is, take the logical \textsc{and} between outcomes of different processes.} That is, having a way to measure quantity $x$ and a way to measure quantity $y$ gives us a way to measure the combination $(x,y)$.
-For ensembles, if we can measure the marginal distribution $\rho_x(x)$ and the marginal distribution $\rho_y(y)$, we know the joint distribution $\rho(x,y)$ has to be compatible with both.
-Note, though, that this does not provide a way to fully measure $\rho(x,y)$ as we know nothing about the correlation between the two variables.\footnote{The quantum case is similar: we measure marginal distributions and rule out states that are incompatible with them.} Formally, the states compatible with the outcomes of the combined process will be the intersections of the states compatible with each pair of outcomes of the original processes.
-
-We can also coarsen a single process.\footnote{That is, take the logical \textsc{or} between outcomes of the same process.} That is, having a way to measure $(x,y)$ gives us a way to measure $x$ alone (or any $f(x,y)$).
-Formally, the outcomes of the coarsened process are given by performing the union of some outcomes of the original process.
-
-This model maps very naturally to a topological space.
-The states are elements of a set and the physical outcomes provide a topology on that set.
-The fact that two elements of the set can be distinguished requires the space to be Hausdorff.
-
-\begin{prop}\label{prop:state_topology}
-The state space $\mathcal{S}$ of a physical system is a Hausdorff topological space.
+\begin{prop}
+	Transported coordinates are canonical. Distributions are function of the point and are transported.
 \end{prop}
 
-\begin{justification}
-We claim $\mathcal{S}$ is a set. Each state is well defined as it is physically distinguishable. The collection of all possible states forms a set.
+$(T^*\mathcal{Q}, \omega)$, $(q^i, p_j)$ set of canonical coordinates, $F: T^*\mathcal{Q} \rightarrow T^*\mathcal{Q}$ such that $\omega = F^* \omega$ where $F^*$ is the pullback.
+$\hat{q}^i = q^i \circ F^{-1}$ and $\hat{p}_i = p_i \circ F^{-1}$.
 
-We claim $\mathcal{S}$ has a topology $\mathsf{T}$.
-Consider the set of all possible physical outcomes associated with all physical processes.
-Each possible outcome is associated with a set of states that are compatible with that outcome.
-Let $\mathsf{T}$ be the set of all sets associated with all physical outcomes.
-$\mathcal{S} \in \mathsf{T}$ and is associated with the outcome ``the system exists." $\varnothing \in \mathsf{T}$ and is associated with the outcome ``the system doesn't exist." Let $V_1, V_2 \in \mathsf{T}$.
-Then, by definition, there exists a process $P_1$ that admits $V_1$ as an outcome and a process $P_2$ that admits $V_2$ as an outcome.
-Consider the process $P$ that combines the outcomes of $P_1$ and $P_2$ with a logical \textsc{and}.
-This always exists physically as we can prepare the same state multiple times and let it interact with each process separately.
-$P$ will have a possible outcome $V$ corresponding to the case where $P_1$ gave outcome $V_1$ and $P_2$ gave outcome $V_2$.
-The states compatible with $V$ must be in both $V_1$ and $V_2$, that is $V = V_1 \cap V_2$.
-Therefore $\mathsf{T}$ is closed under intersection.
-Let $V_1, V_2 \in \mathsf{T}$.
-If they are physically distinguishable, then there exists a physical process $P$ that admits both as outcomes.
-Given $P$, we can always construct the process $P_0$ that combines $V_1, V_2$ with a logical \textsc{or} into a single outcome $V$, by ``forgetting" which of the two was given.
-The states compatible with $V$ must be in either $V_1$ or $V_2$, that is $V = V_1 \cup V_2$.
-Therefore $\mathsf{T}$ is closed under union.
-$\mathsf{T}$ is a topology by definition.
+$F^* (\sum d\hat{q}^i \wedge d\hat{p}_i) =\sum F^* (d\hat{q}^i \wedge d\hat{p}_i) =  \sum (F^* d\hat{q}^i)\wedge (F^* d\hat{p}_i) = \sum d(\hat{q}^i \circ F)\wedge d(\hat{p}_i \circ F) = \sum d(q^i \circ F^{-1} \circ F)\wedge d(p_i \circ F^{-1} \circ F) = \sum dq^i\wedge dp_i = \omega = F^* \omega$.
+$\omega =  \sum d\hat{q}^i \wedge d\hat{p}_i$.
+$(\hat{q}^i, \hat{p}_j)$ is a set of canonical coordinates.
 
-We claim that $\mathsf{T}$ is Hausdorff.
-Let $\mathcal{s_1}, \mathcal{s_2} \in \mathcal{S}$.
-As states are physically distinguishable, there must exist a physical process with two possible outcomes $V_1, V_2 \in \mathsf{T}$ for which $s_1 \in V_1, s_2 \in V_2, V_1 \cap V_2 = \varnothing$.
-$\mathcal{S}$ is Hausdorff by definition.
-\end{justification}
+$\rho(q^i, p_j) : \mathbb{R}^{2n} \rightarrow \mathbb{R}$ integrable and differentiable.
+$\rho (q^i, p_j) \, dq^n \wedge dp_n = \rho (q^i, p_j) \, \omega^n = \rho (q^i(\hat{q}^i, \hat{p}_j), p_j(\hat{q}^i, \hat{p}_j)) \, d\hat{q}^n \wedge d\hat{p}_n$.
+$\rho$, defined on the symplectic structure, transforms like a function.
+We can write $\rho : (T^*\mathcal{Q}, \omega) \rightarrow \mathbb{R}$.
 
-Note that the arguments that led to the topological space had nothing to do with states per se, just that they are physically distinguishable.
-States, though, are not the only objects with that property.
-In fact, any element of a set of physical objects (e.g.
-forces, physical properties such as mass or charge, time) needs to be physically distinguishable to be well defined.
-We can generalize the above justification: any set of physically distinguishable elements is a topological space.
+$\hat{\rho} : (T^*\mathcal{Q}, \omega) \rightarrow \mathbb{R}$ such that $\hat{\rho} = \rho \circ F^{-1 }$.
+$\hat{\rho}(\hat{q}^i, \hat{p}_j) = \hat{\rho} \circ (\hat{q}^i, \hat{p}_j)^{-1} = \rho \circ F^{-1} \circ (q^i \circ F^{-1}, p_i \circ F^{-1}) ^ {-1}= \rho \circ F^{-1} \circ F \circ (q^i, p_j) ^ {-1} = \rho \circ (q^i, p_j) ^ {-1} = \rho (q^i, p_j)$.
 
-\begin{prop}\label{prop:topology}
-	Any set $\mathcal{S}$ of physically distinguishable elements is a Hausdorff topological space.
+\begin{prop}
+	Canonical transformations transport marginal distributions and preserve independence.
 \end{prop}
 
-\begin{justification}
-	Same justification as in \ref{prop:state_topology} with ``state" replaced by ``element" of the set $\mathcal{S}$.
-\end{justification}
-
-With our state space defined, deterministic and reversible evolution corresponds to a bijective map between initial and final states.
-But that's not enough.
-A physical process that can distinguish final states can be used, together with the evolution, to distinguish initial states.
-We prepare the initial state, let it evolve deterministically, measure the final state and use reversibility to convert the measure to one on the initial state.
-Therefore, not only the evolution is a bijection, but it also maps outcomes to outcomes.
-That is: the topology is mapped and preserved by the deterministic and reversible evolution because physical distinguishability must remain unchanged.
-The evolution is then a self-homeomorphism on the state space.
-
-\begin{prop}\label{prop:homeomorphism}
-A deterministic and reversible evolution map is a self-homeomorphism on the state space $\mathcal{T}_{\Delta t}:\mathcal{S} \rightarrow \mathcal{S}$.
+\begin{prop}
+	Canonical transformations transport mutual information.
 \end{prop}
 
-\begin{justification}
-We claim that $\mathcal{T}_{\Delta t}$ exists.
-The system is deterministic: given an initial state $\mathcal{s} \in \mathcal{S}$ there exists a well defined final state $\mathcal{T}_{\Delta t}(\mathcal{s})=\hat{\mathcal{s}} \in \mathcal{S}$.
+\begin{thebibliography}{0}
 
-We claim $\mathcal{T}_{\Delta t}$ is continuous.
-Let $U \subseteq \mathcal{S}$ represent an outcome of a process $P$ on the final states.
-$U$ is an open set in the (final) state space topology by definition.
-Consider the process $P_0$ that first evolves the initial states and then distinguishes the final state with $P$.
-$P_0$ is a process that distinguishes initial states.
-The set of initial states compatible with $U$ are $\mathcal{T}_{\Delta t}^{-1}(U)$.
-$\mathcal{T}_{\Delta t}^{-1}(U)$ is an open set in the (initial) state space topology by definition.
-$\mathcal{T}_{\Delta t}$ is a continuous map.
-
-We claim that $\mathcal{T}_{\Delta t}$ is a bijection.
-The system is reversible: there exists a map $\mathcal{T}_{-\Delta t}:\mathcal{S} \rightarrow \mathcal{S}$ that returns the initial state given the final state.
-$\mathcal{T}_{-\Delta t} \circ \mathcal{T}_{\Delta t} = \mathcal{T}_{\Delta t} \circ \mathcal{T}_{-\Delta t} = \id_{\mathcal{S}}$ as mapping forward and then backward or backwards and then forward must return the original element.
-$\mathcal{T}_{\Delta t}$ admits $\mathcal{T}_{-\Delta t}$ as an inverse.
-$\mathcal{T}_{\Delta t}$ is a bijection.
-
-We claim that $\mathcal{T}_{\Delta t}$ is a self-homeomorphism as it is a continuous bijection.
-\end{justification}
-
-Again, we note that the arguments that lead to continuity had nothing to do with states per se, just that the relationship is between physically distinguishable objects.
-We can then generalize the above justification.
-
-\begin{prop}\label{prop:continuity}
-	A map $f:\mathcal{S_1} \rightarrow \mathcal{S_2}$ between two sets of physically distinguishable elements $\mathcal{S_1}$ and $\mathcal{S_2}$ is a continuous map.
-\end{prop}
-
-\begin{justification}
-	Same justification for continuity as in \ref{prop:homeomorphism} with ``initial states" and ``final states" replaced by ``elements" of $\mathcal{S_1}$ and $\mathcal{S_2}$ respectively.
-\end{justification}
-
-The generality of this result explains why in physics one always assumes functions to be ``well behaved." As the result was derived from our notion of physical distinguishability, this is not a matter of practical convenience.
-Suppose we were able to prepare a force field that was zero everywhere in space except at a single point.
-This gives us a way to tag a specific point.
-But it also allows us to create an outcome compatible with only a single point.
-A finite precision measurement of the force provides us infinite precision of space, which we ruled out.\footnote{This assumes that we are able to position the probe perfectly, which we could do if we were able to manipulate forces at that precision.} This is what the math is telling us, that if we claim that outcomes can't distinguish isolated points (i.e.
-standard topology on $\mathbb{R}^n$) then neither can maps.
-It is physical consistency that limits us to continuous functions.
-
-\begin{figure}
-	\includestandalone[width=\columnwidth]{pictures/continuousMap}
-	\caption{A continuous function always maps finite precision knowledge in the vertical axis (i.e.
-an open set) to finite precision knowledge in the horizontal axis.
-A discontinuous function may map to infinite precision (i.e.
-a closed set).
-If infinite precision identification is ruled out (i.e.
-use of standard topology) then discontinuous functions must also be ruled out.}
-	\label{fig:continuity}
-\end{figure}
-
-While functions with few discontinuities are useful and used in physics and engineering, they are employed for idealized cases  (e.g.~a signal change is fast enough, a charge distribution is small enough) that are often treated as a special case (e.g.~propagation across material discontinuities).
-While this may be obvious and intuitive to the physicist, it may be troubling to the mathematician as the proper use of many mathematical techniques requires the inclusion of discontinuous functions.
-This is less of a problem than it would seem at first.
-Once we made sure that the objects and their relationships are physically meaningful, we can extend our mathematical spaces for the purpose of math computations.
-Our physically meaningful continuous function can be expanded into a sum of discontinuous functions.
-One just has to be mindful of the extension and be wary that mathematical results that depend on such extension may or may not be physically meaningful.
-
-\subsection{Manifolds and labeling states}
-
-To identify and name states one uses a set of quantities, typically numbers.
-For example, the orbital of an electron in a hydrogen atom is identified by the quantum numbers $n$, $l$, $m$ and $s$.
-We call each of these quantities \emph{state variables}.
-We call a \emph{possibility} a possible value that can be taken by a state variable.
-We call a state variable \emph{discrete} or \emph{continuous} if the possibilities are integer or real numbers respectively.
-The topology used reflects the notion that integers can be measured perfectly (i.e.
-discrete topology) while real numbers can be measured up to a finite precision (i.e.
-standard topology on $\mathbb{R}$).
-From now on, we are going to consider state spaces whose states can be identified, at least within a region, by a finite set of discrete and continuous state variables (i.e.
-the state space is locally isomorphic to $\bigcup\limits_{1 \leq i \leq n} \mathbb{R}^{m_i}$).
-
-We purposely use the term state variables instead of coordinates (even though that's what they are mathematically) as it would create confusion with space-time coordinates.
-We also avoid the term observable or measurable, as not all state variables may be directly physically tangible (e.g.
-conjugate momentum in a gauge theory).
-The only requirement for state variables is that they identify states.
-This means the possible values for state variables are physically distinguishable and are defined at equal time (since states are defined and mapped at a particular time).\footnote{This is the main reason that a quantity like velocity is not a suitable state variable, as it is  defined over an interval of time.
-Therefore velocity is always physically well defined but is not a state variable in general, while conjugate momentum is always a state variable but is not physically well defined by itself (it requires the vector potential to be specified as well).
-As we'll see later, when there exists a one to one map between velocity and conjugate momentum they can both be physically well defined state variables.}
-
-\begin{prop}\label{prop:state_variable}
-	A \emph{state variable} is a continuous map $q : U \rightarrow \mathbb{L}$ where $U \subseteq \mathcal{S}$ and $\mathbb{L}$ is the space for the possible values.
-If $\mathbb{L}\cong \mathbb{Z}$ the variable is said \emph{discrete}.
-If $\mathbb{L}\cong \mathbb{R}$ the variable is said \emph{continuous}.
-\end{prop}
-
-\begin{justification}
-	We claim $\mathbb{L}$ is a topological space. $\mathbb{L}$ is a set of physically distinguishable possibilities. $\mathbb{L}$ is a topological space because of \ref{prop:topology}.
-
-	We claim $q$ is a continuous map. $q$ is a map between two sets of physically distinguishable elements. $q$ is continuous because of \ref{prop:continuity}.
-\end{justification}
-
-When combining multiple state variables, it is important to understand how they relate to each other.
-Consider the orbital of an electron in a hydrogen atom, which is identified by the quantum numbers $n$, $l$, $m$ and $s$.
-For each combination of $n$, $l$ and $m$, the spin $s$ can have two values.
-The choice of spin is independent from the rest.
-The choice of $l$, though, depends on the choice of $n$: for $n=1$ only $l=0$ is available; for $n=2$ we can choose $l=0$ or $l=1$.
-The choices are not independent.
-That is: two or more variables are independent if there always exists a state for any possible combination, if the total number of states is the product of the possibilities of each variable.
-We call \emph{state vector} a collection of independent state variables that fully identify a state.
-
-\begin{prop}\label{prop:independent_state_variables}
-	Two state variables $q_1 : U \rightarrow \mathbb{L}_1$ and $q_2 : U \rightarrow \mathbb{L}_2$ are said \emph{independent} if $\exists \mathcal{s}$ such that $q_1(\mathcal{s})=l_1, q_2(\mathcal{s})=l_2 \forall l_1 \in q_1(U), l_2 \in q_2(U)$.
-\end{prop}
-
-In general, the entire state space may not be identified by a predetermined set of independent state variables.
-Consider the state of a pool table, determined by the number of balls together with position and momentum of the center of mass: the number of state variables is not fixed as it depends on the number of balls.
-But deterministic and reversible evolution cannot take us from a different number of continuous independent state variables (i.e.
-there is no homeomorphism between $\mathbb{R}^n$ and $\mathbb{R}^m$).
-That is the number of balls cannot change under deterministic and reversible evolution.
-Therefore we can restrict ourselves to the case where the number of continuous independent state variables is constant without loss of generality.
-This means that, at least locally, the state space is always homeomorphic to $\mathbb{R}^n$, and is therefore a manifold.
-
-Discrete variables do not present such problems, as any number of them can be flattened out in a single one (i.e.
-$\mathbb{Z}^n$ is homeomorphic to $\mathbb{Z}$).
-They will determine the number of connected components of the state space.
-Once we introduce a continuous parameter for time evolution, though, these become irrelevant as continuous time evolution requires continuous trajectories that cannot move states across disconnected components.
-This justifies the special interest in path connected manifolds, as this is where trajectories for deterministic and reversible continuous evolution live.
-
-\begin{prop}\label{prop:manifold}
-	Let $\mathcal{s} \in \mathcal{S}$ a state within a state space.
-The set of states $\mathcal{S}'\subseteq\mathcal{S}$ potentially reachable from $\mathcal{s}$ by deterministic and reversible continuous evolution is a path connected manifold of dimension equal to the number of independent continuous state variables necessary to identify it.
-\end{prop}
-
-\begin{justification}
-	We claim $\mathcal{S}'$ is a manifold.
-Let $\mathcal{s} \in \mathcal{S}$.
-Let $n$ be the number of independent continuous variables needed to identify $\mathcal{s}$.
-There exists a neighborhood $U$ around $\mathcal{s}$ where we have $(q_1,...,q_n):U\rightarrow \mathbb{R}^n$.
-This map is a bijection as $\mathcal{s}$ is identified by those variables.
-$\mathcal{S}$ is homeomorphic to $\mathbb{R}^n$ around  $\mathcal{s}$.
-Let $\mathcal{S}'$ be the set of all states potentially reachable by deterministic and reversible evolution from $\mathcal{s}$.
-Deterministic and reversible evolution is a homeomorphism between initial and final states.
-$\forall \hat{\mathcal{s}} \in \mathcal{S}'$ there exist a map $\mathcal{T}_{\Delta t}:\mathcal{S} \rightarrow \mathcal{S}$ such that $\hat{\mathcal{s}} =\mathcal{T}_{\Delta t}(\mathcal{s})$.
-$\hat{\mathcal{s}} \in \mathcal{T}_{\Delta t}(U)$ and $\mathcal{T}_{\Delta t}(U)$ is isomorphic to $\mathbb{R}^n$.
-$\mathcal{S}'$ equipped with the subspace topology is a topological space everywhere homeomorphic to $\mathbb{R}^n$.
-$\mathcal{S}'$ is a manifold of dimension $n$.
-
-	We claim $\mathcal{S}'$ is path connected.
-Let $\hat{\mathcal{s}} \in \mathcal{S}'$.
-There exist map $\lambda : [t_0,t_1] \rightarrow \mathcal{S}'$ where $\lambda(t_0)=\mathcal{s}$, $\lambda(t_1)=\hat{\mathcal{s}}$ and $t_0$ and $t_1$ are the initial and final time respectively.
-$\lambda$ is a map between two physically distinguishable quantities, time and states.
-$\lambda$ is continuous by \ref{prop:continuity}.
-All $\hat{\mathcal{s}} \in \mathcal{S}'$ are path connected to $\mathcal{s}$.
-$\mathcal{S}'$ is path connected.
-\end{justification}
-
-\subsection{Evolution of state variables}
-
-To study time evolution, we need to describe how state variables change under deterministic and reversible evolution.
-There are two ways to do it, and we'll need both.
-The first approach is to \emph{evolve} the state variables from the initial value $q^i(\mathcal{s})$ to the final value $q^i(\hat{\mathcal{s}})$.
-The result is a trajectory $q^i(t)=q^i(\lambda(t))$ which is especially useful when state variables correspond to physically meaningful quantities.
-For example, we track how the temperature or the pressure of an ideal gas changes.
-Evolved state variables, however, make it hard to find relationships that are invariant and common to all deterministic and reversible processes.
-
-The second approach is to transport the state variables.
-The idea is to keep the connection to the initial state by labeling the future state by the original value of $q^i$, instead of by the future values.
-For example, the evolved state variable will not tell us the current value of pressure, but the one for the initial conditions.
-So we introduce a new set of variables for which $\hat{q}^i(\hat{\mathcal{s}})=q^i(\mathcal{s})$, which we can always do as the evolution is deterministic and reversible.
-The hypersurfaces at constant $\hat{q}^i$ are the evolved hypersurfaces at constant of $q^i$ (e.g.
-all the states that started with a particular value for pressure) which allows us to study how groups of states evolve in time.
-Given that the value of the transported state variables does not change during evolution, and that it is unique for each initial state, transported state variables also provide a way to label the trajectories themselves.
-
-Evolved state variables are useful to write equations of motion, study how physical quantities change, form a physical picture of what happens.
-Transported state variables are useful to write invariants, study state space trajectories, form a geometric picture for the state space.\footnote{One should not confuse evolved/transported state variable with active/passive transformations or with Schroedinger/Heisenberg pictures.
-In those cases, the choice is between changing the state or the coordinates/observables.
-In our case, the state is always changing.
-The choice is between tracking the change with different values of the same state variable or different state variables that give the same value.}
-
-\begin{prop}\label{prop:evolved_transported_variable}
-Let $q^i$ be a set of state variables and $\mathcal{T}_{\Delta t}$ a deterministic and reversible evolution map.
-The evolved state variables are given by $q^i \circ \mathcal{T}_{\Delta t}$.
-The transported state variables are given by $q^i \circ \mathcal{T}_{-\Delta t}$.
-\end{prop}
-
-One reason that transported variables will be useful is that, during deterministic and reversible evolution, we need to make sure that the number of initial and final possibilities across independent state variables remains the same.
-This is more easily done using transported variables as the actual values do not change.
-In the discrete case, this is easy to show.
-
-\begin{prop}\label{prop:discrete_measure}
-	Let $U \subseteq \mathcal{S}$ a set of states fully identified by a set of $n$ independent discrete state variables $q^i$.
-Let $\Delta q^i \equiv q^i(U)$ the range of possibilities of each variable.
-Then $\#(\Delta q^i)=\#(\Delta \hat{q}^i) \; \forall i$ and $\#(U)=\prod\limits_{i=1}^{n}\#(\Delta q^i)=\prod\limits_{i=1}^{n}\#(\Delta \hat{q}^i)=\#(\hat{U})$ where $\hat{U}=\mathcal{T}_{\Delta t}(U)$ and $\#$ denotes the number of elements in the given set.
-\end{prop}
-
-\begin{justification}
-	We claim $\#(\Delta q^i)=\#(\Delta \hat{q}^i)$.
-For each $\mathcal{s} \in U$, let $\hat{\mathcal{s}}=\mathcal{T}_{\Delta t}(\mathcal{s})$.
-We have $\hat{q}^i(\hat{\mathcal{s}}) = q^i(\mathcal{s})$.
-Therefore $\Delta q^i = q^i(U) = \hat{q}^i(\hat{U})=\Delta \hat{q}^i$.
-$\#(\Delta q^i)=\#(\Delta \hat{q}^i)$.
-
-	We claim $\#(U)=\prod\limits_{i=1}^{n}\#(\Delta q^i)$. As $q^i$ are independent variables, the states are the Cartesian product of the possibilities of each variable.
-\end{justification}
-
-What happens is that, because the variables are independent, the total number of states is the product of the number of possibilities for each variable.
-If $\Delta q^1$ and $\Delta q^2$ are ranges of possibilities for two independent variables, the total number of possibilities is $\#(\Delta q^1) \#(\Delta q^2)$.
-By construction, the set $\Delta q^i$ of possibilities for each independent variable is the same as the set $\Delta \hat{q}^i$ of possibilities for the transported variable.
-Also, the transported variables remain independent.
-Therefore the relationship $\#(\Delta \hat{q}^1) \#(\Delta \hat{q}^2) = \#(\Delta q^1) \#(\Delta q^2)$ is valid throughout the evolution.
-
-We'll see that very similar relationships are what define Hamiltonian and Lagrangian mechanics.
-But recovering them in the continuous case is not as straightforward.
-Consider the map $q'=aq$ with $0<a<1$.
-At first glance, it's a bijective continuous map so we may think it can represent a deterministic and reversible evolution.
-Yet, $\Delta q \equiv [-b, b] \supset \Delta q'$: a set of states is mapped to a proper subset (i.e.
-to fewer states) which does not make sense for a reversible process.
-In the limit where we apply the map an infinite amount of times, any value of $q$ will be brought infinitely close to $0$, which also does not sound like a reversible process.
-
-The issue is that while deterministic evolution will map states to states, not all bijective state to state maps can be considered deterministic and reversible.\footnote{In fact, we have already seen that the map must at least be continuous.} This is a crucial point where this work is different from more usual treatments of dynamical systems.
-If one focuses only on the states themselves and counts them, it reaches the conclusion that a range of $1$m contains the same number of initial conditions of $1$Km: they have the same infinite number of elements.
-This conclusion has a number of problems: it does not match our physical intuition; it's not consistent with all types of topology; it does not explain why the laws of evolution are very often, not only continuous, but differentiable.
-In short: a bijection does not capture the entirety of the physical description.
-
-If we work with a discrete topology then, by all means, the way to measure the size of a set is to count the elements in a set.
-But if we use the standard topology on $\mathbb{R}$, isolated points have measure zero (i.e.
-we can't physically distinguish them): what defines the number of initial conditions is the range of the state variables.
-As we'll see, the geometrical structures in Hamiltonian/Lagrangian classical/quantum mechanics are there to allow us to properly count states and, more specifically, possibilities along independent state variables.
-If we think that counting states is equivalent to counting elements in the set, we miss the whole point of that geometrical structure.
-
-What makes the matter even more confusing is that the single physical idea of counting states and possibilities along independent state variables does not map to a single mathematical concept.
-Sometimes we need a measure, sometimes we need a symplectic form and sometimes we need a metric tensor.
-It will depend, case by case, on the state space of the system under study.
-But, however defined mathematically, deterministic and reversible motion must preserve that count.
-This is the crucial insight that gives physical meaning to those structures, that justifies the use of differential equations for state evolution, that tells us why symplectic and Riemannian geometry are so important in physics.
-Properly characterizing this count will be one of the key goals of the following sections.
-
-\section{Composite systems and reducibility}
-\label{sec:reducibility}
-
-% Review started
-
-Given that scientific reductionism (i.e.
-the idea of reducing physical systems and interactions to the sum of their constituent parts in order to make them easier to study) is at the heart of fundamental physics \cite{Ruse}, we now explore how to characterize a system in terms of its components.
-That is, we want to study the relationship between the state space of a composite system and the state space of its parts.
-In general, this is quite a complicated thing to do, which requires intimate knowledge of the system at hand.
-So we simplify our problem and study a material made of infinitesimal homogeneous parts.
-What we'll find is that under the additional assumption that the material is infinitesimally reducible (i.e.
-its state is equivalent to the states of its infinitesimal parts) and that each part undergoes deterministic and reversible evolution, the motion is suitably described by the standard framework of classical Hamiltonian mechanics.
-
-\subsection{Homogeneous decomposable systems and vector spaces}
-The notion that a system is decomposable means the states are equipped with a rule of composition that allows one to write $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$: the composite system is the sum of its components.
-For example, the state of a ball is equal to the state of its top and bottom parts.
-
-The notion that the system is homogeneous means that the states of the composite and of each part are not unrelated: they are all made of the same material.\footnote{Whether a system is homogeneous depends on context (e.g.
-air can be thought as homogeneous if the mixture of gases does not change in space or in time due to phase transitions or chemical processes) and one must check that such a property is maintained by time evolution.} In fact, the state space $\mathcal{C}$ of all systems composed of such material will include the state of the system as well as the states of  its parts (i.e.
-$\mathcal{c}, \mathcal{c}_1, \mathcal{c}_2 \in \mathcal{C}$).
-Since combining any two systems made of a homogeneous material will always give us a system made of the same homogeneous material (e.g.
-combining elements made of water gives us another element made of water), the state space $\mathcal{C}$ is closed under composition.
-
-As we study the system under evolution, we'll also want to study state changes.
-For example, if a stable mixture of gas expands, we'll want to know the difference between the initial and final distributions.
-That is: $\delta\mathcal{c}=\hat{\mathcal{c}}-\mathcal{c}$.
-Note that such a state difference may not describe a physical state: it may remove some material from one location to add it somewhere else.
-Yet, these state changes are still physically distinguishable objects that provide a configuration for the same homogeneous material, therefore we extend the state space $\mathcal{C}$ to include state differences.\footnote{Note that one can characterize a state difference without knowing the states themselves.
-If we take some amount of material out of a container we know how the state changed, yet we may not know the amount of material before or after.
-In certain cases one may be interested in measuring the asymmetry of a particular state without measuring (or without being able to measure) the full state itself.} All combined, this gives us the structure of an abelian group.
-
-\begin{prop}\label{prop:abelian_group}
-The state space $\mathcal{C}$ for a decomposable homogeneous material is an additive abelian (i.e. commutative) group.
-\end{prop}
-
-\begin{justification}
-We claim $\mathcal{C}$ is an additive monoid.
-There exists a law of composition $+ : \mathcal{C} \times \mathcal{C} \rightarrow \mathcal{C}$ that takes two states and returns one that is the physical composition of the two.
-The domain and codomain match because the material is homogeneous.
-The law is commutative $\mathcal{c}_1 +\mathcal{c}_2 = \mathcal{c}_2+\mathcal{c}_1$ and associative $(\mathcal{c}_1 + \mathcal{c}_2) + \mathcal{c}_3 = \mathcal{c}_1 + (\mathcal{c}_2 + \mathcal{c}_3)$, as it does not matter in what order we physically compose the components.
-There exists a unique zero element $\mathcal{c} + 0 = \mathcal{c}$ and it represents the physically empty state (i.e.
-no amount of material).
-$\mathcal{C}$ is an additive monoid by definition.
-
-We claim $\mathcal{C}$ is an additive group.
-We require $\mathcal{C}$ to include state changes.
-A change of a physically distinguishable object is itself physically distinguishable therefore $\mathcal{C}$ is still a Hausdorff topological manifold as per \ref{prop:topology}.
-There exists an inverse $- : \mathcal{C} \rightarrow \mathcal{C}$ such that $\mathcal{c} + ( - \mathcal{c}) = 0 \forall \mathcal{c} \in \mathcal{C}$.
-$\mathcal{C}$ is an additive group by definition.
-\end{justification}
-
-\subsection{Classical material and distributions}
-We also want to be able to express the state of the composite system in terms of the state of each part.
-For example, given the state of a fluid we'll want to know how the material is distributed in space.
-This in general will depend on how much the state of the composite system ``knows" about its parts.
-For example, the position and orientation of an ideal rigid body is enough to define where all its constituents are, while the volume/pressure/temperature of an ideal gas is not enough to determine the position and momentum of all its constituents.
-Therefore we need to characterize the system further.
-
-We will call a \emph{particle} the smallest amount of a material to which we can assign an independent state.
-For example, a photon is a particle of light (it is the smallest amount of light we can describe), an infinitesimal amount of water is treated classically as a point particle (the smallest amount for a continuous fluid).\footnote{We use this definition as it allows us to retroactively talk on somewhat equal grounds about classical particles and quantum particles, and underscore how physically (and mathematically) they are very similar yet very different objects.}
-
-We will call a \emph{classical material} one that is homogeneous and infinitesimally reducible.
-That is, we can keep decomposing the system indefinitely into smaller and smaller parts, each with a well defined state.
-In this case, the particles are the infinitesimal parts given by the limit of this process of recursive reduction.
-Giving the state of the whole system is equivalent to giving the states of all particles.
-Given the state space $\mathcal{S}$ for the infinitesimal parts, which we assume consistently with \ref{prop:manifold} to be a manifold, a composite state $\mathcal{c} \in \mathcal{C}$ will tell us the amount of material for each possible particle state.
-That is, each state is fully identified by a function $\rho_\mathcal{c}: \mathcal{S} \to \mathbb{R}$ that returns the amount of material in the composite state $\mathcal{c}$ that is prepared in the particle state $\mathcal{s}$.
-For example, given a certain configuration of gas, we can tell the density of material that is at a specific point in space with a specific value of momentum.
-
-The notion that the parts are infinitesimal requires the co-domain of $\rho_\mathcal{c}$ to be a real number, as opposed to an integer.
-The state space $\mathcal{S}$ could, instead, be a discrete set.
-For example, for a system composed of different tanks connected by pipes, the state of the composite system could be the overall distribution of water among the tanks.
-The amount in each tank is continuous (as we assume the water to be infinitesimally divisible) yet there are only a finite number of tanks the water can be placed into.
-As it links two physically distinguishable objects, $\rho_\mathcal{c}$ is a continuous function as discussed in \ref{prop:continuity}.
-
-As we combine the states of different parts, we sum the distributions over particle states $\rho_{\mathcal{c}}(\mathcal{s})=\rho_{\mathcal{c}_1}(\mathcal{s})+\rho_{\mathcal{c}_2}(\mathcal{s})$: the amount of material in the composite state is the sum of the amount of material of the parts.
-We can also increase or decrease the amount of material by a constant factor $\rho_{\mathcal{c}_1}(\mathcal{s})=a\rho_{\mathcal{c}_2}(\mathcal{s})$.
-This gives us the structure of a real vector space that is isomorphic to a subspace of continuous functions.
-
-\begin{prop}\label{prop:real_vector_space}
-The state space $\mathcal{C}$ for a classical (i.e.
-homogeneous infinitesimally reducible) material is a vector space over $\mathbb{R}$ isomorphic to a subspace of the space of continuous functions.
-That is $\mathcal{C} \cong G \subseteq C(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbb{R} \; | \; \rho$ is continuous$\}$, where $\mathcal{S}$ is the state space of an infinitesimal amount of material.
-\end{prop}
-
-\begin{justification}
-We claim $\mathcal{C}$ is an abelian group. $\mathcal{C}$ is the state space for a decomposable homogeneous material and is therefore an abelian group  by \ref{prop:abelian_group}.
-
-We claim $T$, the set of transformations that increase or decrease the amount of material in the system by a constant factor, is a field\footnote{Here field is intended in the abstract algebraic sense (a nonzero commutative division ring) which has no relationship to the field in the physics or differential geometry sense (a physical quantity/tensor with a value for each point in space).} isomorphic to $\mathbb{R}$.
-Consider $\tau: \mathbb{R} \rightarrow T$ the mapping between a number and the transformation that increases or decreases the amount of material by that factor.
-This transformation exists: the system is infinitesimally decomposable and the amount can be changed continuously.
-Define on $T$ an addition $+: T \times T \rightarrow T$ and a multiplication $*: T \times T \rightarrow T$ such that $\tau(a) + \tau(b) = \tau(a+b)$ and $\tau(a) * \tau(b) = \tau(a*b)$, $a,b \in \mathbb{R}$, so that the sum and product of the transformation is equal to the sum and product of their respective factors.
-$\tau$ is an isomorphism between $T$ and $\mathbb{R}$ as fields.
-
-We claim $\mathcal{C}$ is a vector space over $\mathbb{R}$.
-The abelian group $\mathcal{C}$ can be extended with the operations defined by $T$, as each element $\tau \in T$ is a map $\tau : \mathcal{C} \rightarrow \mathcal{C}$.
-The map has the following properties: $(\tau_1 + \tau_2) \, \mathcal{c} = \tau_1 \mathcal{c} + \tau_2 \mathcal{c} \; \forall \tau_1, \tau_2 \in T$ and $\mathcal{c} \in \mathcal{C}$, increasing the amount of material by the sum of two constant factors is the same as combining the separate increases, and $\tau \, (\mathcal{c}_1 + \mathcal{c}_2) = \tau \mathcal{c}_1 + \tau \mathcal{c}_2\; \forall \tau \in T$ and $\mathcal{c}_1, \mathcal{c}_2 \in \mathcal{C}$, increasing the amount of the total system is the same as the combination of the increased parts.
-$\mathcal{C}$ is a module over $T$, which is a field and isomorphic to $\mathbb{R}$.
-$\mathcal{C}$ is (isomorphic to) a real vector space.
-
-We claim $\mathcal{C}$ is isomorphic to a subspace of $C(\mathcal{S})$ as a vector space over $\mathbb{R}$.
-$\forall \mathcal{c} \in \mathcal{C} \; \exists !
-\rho_{\mathcal{c}}:\mathcal{S} \rightarrow \mathbb{R}$ returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$.
-$\rho_{\mathcal{c}} \in C(\mathcal{S})$ by \ref{prop:continuity}.
-Let $\varrho : \mathcal{C} \rightarrow C(\mathcal{S})$ such that $\varrho(\mathcal{c}) \mapsto \rho_\mathcal{c}$.
-As the system is reducible, two distinct composite states must represent different distributions.
-$\forall \mathcal{c_1}, \mathcal{c_2} \in \mathcal{C}, \mathcal{c_1} \neq \mathcal{c_2} \implies \varrho(\mathcal{c_1}) \neq \varrho(\mathcal{c_2})$.
-$\varrho$ is injective.
-$\varrho$ is a bijection between $\mathcal{C}$ and $\varrho(\mathcal{C})$.
-Let $\mathcal{c}=\mathcal{c}_1+\mathcal{c}_2$, then $\varrho(\mathcal{c})=\varrho(\mathcal{c}_1)+\varrho(\mathcal{c}_2)$ as the amount of material of the composition of two states is the sum of the individual amounts.
-Let $\mathcal{c}_1=\tau(a)\mathcal{c}_2$, then $\varrho(\mathcal{c}_1)=a \varrho(\mathcal{c}_2)$ as $\tau(a)$ increases the amount of material by the factor $a$.
-$\varrho$ is a homomorphism as a vector space.
-$\mathcal{C}$ is isomorphic to $\varrho(\mathcal{C}) \subseteq C(\mathcal{S})$.
-\end{justification}
-
-\subsection{Integration and measure}
-
-When the topology of $\mathcal{S}$ is not discrete, the distribution $\rho_\mathcal{c}$ is a density, which is not something we directly measure.
-What we do measure experimentally are finite amounts of material.
-For example, the amount of fluid in a particular volume within a particular range of momentum specified in some units (e.g.
-moles, kg, ...).
-This means that given a composite state $\mathcal{c}$ and a set $U \subseteq \mathcal{S}$ of particle states compatible with an outcome of a process, we must be able to tell the amount of material that we will find associated with that outcome.
-That is for each open set $U \subseteq \mathcal{S}$ there will be a functional $\Lambda_U : \mathcal{C} \rightarrow \mathbb{R}$.
-
-As we combine parts, or increase the amount of material in each part, the total amount of material found will have to be consistent with those operations.
-That is: $\Lambda_U(a_1 \mathcal{c}_1 + a_2 \mathcal{c}_2) = a_1 \Lambda_U(\mathcal{c}_1) + a_2 \Lambda_U(\mathcal{c}_2)$.
-For a proper state, one that is not the difference of two states, we will expect a positive amount of material.
-So, if a density $\rho_\mathcal{c}$ is positive everywhere, then the total amount is also positive.
-This makes $\Lambda_U$ a positive linear functional.
-As the amount of material we measure is always finite, the functional applied to any composite state $\mathcal{c}$ over any set $U$ will be finite.
-
-The ability to associate finite amounts of material with sets of states allows us to define finite regions of $\mathcal{S}$ and compare them.
-Consider a region of position and momentum in phase space.
-If we are able to spread a finite amount of material into a non-infinitesimal uniform distribution, then we know we have a finite region.
-And the density will give us an indication of how big the region is: if we double the region, the density will halve.
-In other words: because we are describing densities over $\mathcal{S}$, we are able to give a unique measure\footnote{Here measure is intended in the mathematical sense (a real valued function of a $\sigma$-algebra) which is distinct from other connotations in physics.} $\mu$ that allows us to count the number of states in each set of $\mathcal{S}$.
-In terms of said measure, positive linear functionals become integrals, $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
-
-This intuitive picture is formalized mathematically by the Riesz representation theorem for linear functionals \cite{Halmos}, which gives $\mathcal{S}$ the structure of a measure space, with a suitable Borel $\sigma$-algebra and measure $\mu$.
-
-\begin{prop}\label{prop:integration}
-	The state space $\mathcal{S}$ for the particles of a classical material is endowed with a Borel measure $\mu$.
-The state space $\mathcal{C}$ for a classical material is isomorphic to a subspace of the space of Lebesgue integrable functions.
-That is $\mathcal{C} \cong G \subseteq L^1(\mathcal{S}, \mu) = \{ \rho : \mathcal{S} \rightarrow \mathbb{R} \; | \; \int_{\mathcal{S}} |\rho| d\mu < \infty \}$.
-\end{prop}
-
-\begin{justification}
-	We claim there exists a positive linear functional $\Lambda_U : \mathcal{C} \rightarrow \mathbb{R}$ for each $U \subseteq \mathcal{S}$ such that $\Lambda_U = \Lambda_{\mathrm{int}(U)}$ where $\mathrm{int}(U)$ is the interior of $U$.
-Let $\Lambda_{U} : \mathcal{C} \rightarrow \mathbb{R}$ be the functional that returns the amount of material in $\mathcal{c}$ compatible with the outcome associated with the open set $U \subseteq \mathcal{S}$.
-$\Lambda_U$ is well-defined: $U$ is associated with a physically distinguishable outcome.
-$\Lambda_U$ is linear: $\Lambda_U(a_1 \mathcal{c}_1 + a_2 \mathcal{c}_2) = a_1 \Lambda_U(\mathcal{c}_1) + a_2 \Lambda_U(\mathcal{c}_2)$ as it has to be consistent with the operations of composing states and increasing/decreasing the amount of material by a factor.
-$\Lambda_U$ is positive: if the value of the distribution for each particle state is positive then the total amount of material is positive.
-Let $U \subseteq \mathcal{S}$ not necessarily open.
-Define $\Lambda_U$ as $\Lambda_{\mathrm{int}(U)}$.
-
-	We claim that $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$.
-Let $U \subseteq \mathcal{S}$ be an open set of particle states associated with an outcome.
-Let $\mathcal{c} \in \mathcal{C}$ be a composite state.
-The amount of material of $\mathcal{c}$ associated with the outcome $U$ must be finite, as physically we always work with finite quantities.
-$|\Lambda_{U}(\mathcal{c})| < \infty$.
-Let $U \subseteq \mathcal{S}$ not necessarily open.
-$|\Lambda_{U}(\mathcal{c})| = |\Lambda_{\mathrm{int}(U)}(\mathcal{c})| < \infty$.
-
-	We claim that $\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2}$ $\forall $ $U_1, U_2 \in \mathcal{S}$.
-Let $U_1, U_2 \in \mathcal{S}$ be open sets of particle states associated with two outcomes.
-Suppose $U_1 \cap U_2 = 0$, $\Lambda_{U_1 \cup U_2} = \Lambda_{U_1} + \Lambda_{U_2}$ as the material found in $U_1 \cup U_2$ must be either in $U_1$ or $U_2$.
-Suppose $U_1 \cap U_2 \neq 0$, $\Lambda_{U_1 \cup U_2} = \Lambda_{U_1} + \Lambda_{U_2} - \Lambda_{U_1 \cap U_2}$ as the sum of the material associated with each outcome will double count the intersection.
-Let $U_1, U_2 \in \mathcal{S}$ not necessarily open.
-$\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{\mathrm{int}(U_1)} + \Lambda_{\mathrm{int}(U_2)} = \Lambda_{\mathrm{int}(U_1) \cup \mathrm{int}(U_2)} + \Lambda_{\mathrm{int}(U_1) \cap \mathrm{int}(U_2)} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2}$
-
-	We claim that $\mathcal{S}$ is endowed with a unique Borel measure $\mu$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho_{\mathcal{c}} d \mu$.
-$\mathcal{S}$ is locally compact as it is a manifold.
-$\mathcal{C} \cong G \subseteq C(\mathcal{S})$ therefore $\Lambda_U(\mathcal{c}) \cong \Lambda_U(\rho_\mathcal{c})$.
-$\Lambda = \{\Lambda_U : C(\mathcal{S}) \rightarrow \mathbb{R}\}_{U \subseteq \mathcal{S}}$ is a family of positive linear functionals such that $\forall U \subseteq \mathcal{S} \; \Lambda_U = \Lambda_{\mathrm{int}(U)}$ and $\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2} \; \forall U_1, U_2 \subseteq \mathcal{S}$.
-By the extension \ref{extended_riesz_theorem} of the Riesz representation theorem for linear functionals
-	there exists a unique Borel measure $\mu$ such that $\Lambda_U (\mathcal{c}) = \int_{U} \rho_\mathcal{c} d\mu$.
-
-	We claim $\mathcal{C}$ is isomorphic to a subspace of $L^1(\mathcal{S}, \mu)$ as a vector space over $\mathbb{R}$.
-$\forall \mathcal{c} \in \mathcal{C} \exists !
-\rho_{\mathcal{c}}:\mathcal{S} \rightarrow \mathbb{R}$ returning the amount of material for each state $\mathcal{s} \in \mathcal{S}$.
-As $|\Lambda_{U}(\mathcal{c})| < \infty \; \forall \mathcal{c} \in \mathcal{C} \; \forall U \subseteq \mathcal{S}$, then $\rho_{\mathcal{c}} \in \mathcal{L}(S,\mu) \equiv \{ \rho : S \rightarrow \mathbb{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\}$.
-$\mathcal{L}(S,\mu) = L^1(\mathcal{S}, \mu)$ by \ref{everywhere_integrable_is_lebesgue_integrable}.
-Let $\varrho : \mathcal{C} \rightarrow L^1(\mathcal{S}, \mu)$ such that $\varrho(\mathcal{c}) \mapsto \rho_\mathcal{c}$.
-$\varrho$ is a homomorphism, as justified in \ref{prop:real_vector_space}.
-$\mathcal{C}$ is isomorphic to $\varrho(\mathcal{C}) \subseteq L^1(\mathcal{S}, \mu)$ as a vector space.
-\end{justification}
-
-\subsection{Invariant densities and differentiability}
-
-On a state space with countable elements, using a single state variable, we have $\Lambda_U (\mathcal{c}) = \sum \limits_{q=a}^b \rho_\mathcal{c}(q)$.
-We would expect the expression to become $\Lambda_U (\mathcal{c}) = \int_a^b \rho_\mathcal{c} (q) dq$ for states identified by a continuous state variable.
-By changing state variable, though, we see that this does not work in general: $\rho_\mathcal{c}(\hat{q})= \rho_\mathcal{c}(q) dq/d\hat{q}$.
-For example, if we changed units from $g/m$ to $g/Km$, the value of the density would increase by a factor of $1000$.
-This makes $\rho_\mathcal{c}$ a function not just of the state $\mathcal{s}$, but also of the state variables we are using.
-This is inconsistent with our previous definition.
-Moreover if the transformation is not differentiable, the density is not even well-defined.
-This tells us that $\mathcal{S}$ cannot be any manifold: it has to be one that allows state-variable-independent densities.
-
-The first order of business is to guarantee that the density remains defined under an arbitrary change of state variables.
-This means the Jacobian of the transformation must exist.
-For the Jacobian to exist the transformation between state variables has to be differentiable.
-Therefore a manifold that guarantees densities to be well defined is one that guarantees that state variable transformations are differentiable: a differentiable manifold.
-$\mathcal{S}$ has a differential structure in the sense that the distributions and the volume element $d\mu$ are only defined on a set of state variables that are linked by differentiable transformations.
-In the same way that discontinuous changes of state variable can be detected because they tamper with physical distinguishability (i.e.
-the topology), non differentiable changes of state variable can be detected because they tamper with our ability to define state-variable-independent densities and count states.
-
-Moreover, the density itself has to be differentiable.
-The distribution $\rho_\mathcal{c}$ is a real valued function of the state.
-As such, at least locally, we can use it as a state variable $q^1=\rho_\mathcal{c}$: we can identify particle states by the density of the material associated with them.
-This is more physically meaningful than one may first suspect, since placing physical markers (i.e.
-placing material at particular states) is a common way to define the references for a coordinate system.
-In a coordinate system where $q^1=\rho_\mathcal{c}$, the function is clearly differentiable as $\partial_1 \rho_\mathcal{c} = 1$ and $\partial_i \rho_\mathcal{c} = 0$ for $i \neq 1$.
-As we change state variables, the transformation is differentiable, therefore all the derivatives $\partial_{i'} q^1 = \partial_{i'} \rho_\mathcal{c}$ exist.
-The density itself is differentiable.
-
-\begin{prop}\label{prop:differentiable_manifold}
-	The state space $\mathcal{S}$ for the particles of a  classical material is a differentiable manifold.
-The state space $\mathcal{C}$ for a classical material is isomorphic to the space of Lebesgue integrable differentiable functions.
-That is $\mathcal{C} \cong C^1(\mathcal{S}) \cap L^1(\mathcal{S}, \mu)$, where $C^1(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbb{R} \; | \; \rho$ is differentiable$\}$.
-\end{prop}
-\begin{justification}
-	We claim $\mathcal{S}$ is a differentiable manifold.
-Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbb{R}$ be the distribution associated with a composite state $\mathcal{c} \in \mathcal{C}$.
-Let $q^i$ and $\hat{q}^j$ be two sets of independent state variables that fully identify states within $U\subseteq \mathcal{S}$.
-Let $\rho_\mathcal{c}(q^i)$ and $\rho_\mathcal{c}(\hat{q}^j)$ be the expression in local coordinates of the distribution.
-$\rho_\mathcal{c}(\hat{q}^j)=\rho_\mathcal{c}(q^i) | \partial_j q^i |$ as $\rho_\mathcal{c}$ transforms as a density.
-$\rho_\mathcal{c}(\mathcal{s}) = \rho_\mathcal{c}(q^i(\mathcal{s})) = \rho_\mathcal{c}(\hat{q}^j(\mathcal{s}))$.
-The Jacobian $| d q' / d q |$ exists and is non-zero.
-The map between any two charts is differentiable.
-$\mathcal{S}$ is a differentiable manifold.
-
-	We claim that the distributions associated with composite states are differentiable in the regions where they are strictly monotonic.
-Let $\rho_\mathcal{c}$ be the distribution associated with state $\mathcal{c}$.
-Let $\rho_\mathcal{c}$ be strictly monotonic in a region $U\subseteq \mathcal{S}$.
-Then its fibers in $U$, the inverse images of the values, are connected hypersurfaces.
-We can construct a chart such that $q^1=\rho_\mathcal{c}$.
-The distribution is differentiable over $U$ for that variable: $\partial_{i} \rho_\mathcal{c} = \partial_{i} q^1 = \delta_i^1$ where $\delta_i^j$ is the Kronecker delta.
-Let $q^{j}$ be another set of state variables.
-$q^1(q^{j})$ is differentiable as the manifold is differentiable.
-$\rho_\mathcal{c}(q^{j})=q^1(q^{j})$ is differentiable over $U$.
-
-	We claim that the distributions associated with composite states are differentiable.
-Let $\rho_\mathcal{c}$ be the distribution associated with state $\mathcal{c}$.
-Let $\rho_{\mathcal{c}_0}$ be the distribution associated with state $\mathcal{c}_0$.
-Let $\rho_{\mathcal{c}_0}$ be strictly monotonic in a region $U\subseteq \mathcal{S}$.
-Let $\delta \rho = \rho_{\mathcal{c}_0} - k \rho_{\mathcal{c}}$ where $k \in \mathbb{R}$ and $k>0$.
-Let $k$ be sufficiently small so that $\delta \rho$ is strictly monotonic over $U$.
-$\rho_{\mathcal{c}_0}$ and $\delta \rho$ are differentiable over $U$ as they are strictly monotonic over $U$.
-$\rho_{\mathcal{c}}$ is a linear combination of differentiable functions over any arbitrary region $U$.
-$\rho_{\mathcal{c}}$ is differentiable.
-
-	We claim $\mathcal{C} \cong C^1(\mathcal{S}) \cap L^1(\mathcal{S}, \mu)$.
-Let $\rho_\mathcal{c}$ be the distribution associated with a state $\mathcal{c}$.
-Let $\rho \in C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ be a Lebesgue integrable continuous function.
-If $\rho_\mathcal{c} \neq \rho$ then $\exists U \in \mathcal{S}$ such that $\int_{U} \rho_\mathcal{c} d \mu \neq \int_{U} \rho d \mu$.
-Both integrals are finite as $\rho_\mathcal{c}, \rho \in L^1(\mathcal{S}, \mu)$.
-Both integrals are expressible with the same state variables as all elements are defined on the same differential structure.
-Therefore there exists an outcome for $\mathcal{S}$ that can physically distinguish the two distributions.
-There must be a state $\mathcal{c}_1 \in \mathcal{C} \; | \; \varrho(\mathcal{c}_1)=\rho$.
-$\varrho : \mathcal{C} \rightarrow C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ is surjective.
-$\varrho$ is a homomorphism, as justified in \ref{prop:real_vector_space}.
-$\varrho$ is an isomorphism between $\mathcal{C}$ and $C(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as vector spaces.
-\end{justification}
-
-As we identified the space of Lebesgue integrable differentiable functions $C^1(\mathcal{S}) \cap  L^1(\mathcal{S}, \mu)$ as the space of distributions that are physically meaningful, we can better understand why other commonly used function spaces do not fit the bill.
-Some are not restrictive enough.
-The set of Lebesgue integrable functions $L^1(\mathcal{S})$ includes discontinuous functions that for \ref{prop:continuity} are unphysical.
-The space of continuous functions $C(\mathcal{S})$, the space of continuous functions that vanish at infinity $C_0(\mathcal{S})$ and the space of differentiable functions $C^1(\mathcal{S})$ include functions whose integral is infinite, which would represent an infinite amount of material.
-Some definitions are too restrictive.
-Requiring compact support (i.e.
-the function is different from zero only on a finite region) would exclude distributions, such as Gaussians, that span over the whole range of states.
-Such distributions are physical: if we take a finite volume of an ideal gas at equilibrium, the momentum distribution spans over the whole range.
-Schwartz space $S(\mathcal{S})$ excludes functions that are not infinitely smooth which we don't have a general physical justification for.\footnote{One can, though, make the argument that given any distribution $\rho$ one can find an infinitely smooth $\rho_{sm}$ such that the difference in description is small compared to the error already introduced by assuming the system to be homogeneous and infinitely reducible.
-The Whitney approximation theorem \cite{Lee} makes this mathematically well defined.}
-
-Another consideration is that while the norm associated to $L^1(\mathcal{S})$ is $\int_{\mathcal{S}} |\rho| d\mu$, the expression $\int_{\mathcal{S}} \rho d\mu$ may be more physically meaningful in some cases.
-For proper states, the two are the same and they represent the total amount of material.
-For state changes they differ.
-The second represents the amount of material added (or taken away if negative) by the state change.
-For example, if $\delta \rho$ is the change due to deterministic and reversible evolution, the amount of material does not change and therefore $\int_{\mathcal{S}} \delta \rho \, d\mu = 0$.
-The first norm would represent the total amount of material that is changing (i.e.
-being added and being removed), therefore $\int_{\mathcal{S}} | \delta \rho | \, d\mu = 0$ means no material is moving, no change is occurring.
-
-Note that neither of those expressions matches the norm given by the inner product defined as $\langle f, g \rangle = \int_{\mathcal{S}} fg \, d\mu$.
-Moreover $\mathcal{C}$, since it excludes discontinuous functions, can never be a complete normed space: it will not include the limit for all Cauchy sequences; it cannot be a Banach or a Hilbert space.
-Such a construction, though, is still useful.
-Consider the following expressions:
-\begin{align*}
-\Lambda_{\mathcal{S}} (\mathcal{c}) &= \int_\mathcal{S} \rho_{\mathcal{c}} d \mu = \langle 1 , \rho_{\mathcal{c}} \rangle \\
-\Lambda_U (\mathcal{c}) &= \int_U \rho_{\mathcal{c}} d \mu = \langle 1_U , \rho_{\mathcal{c}} \rangle \\
-\rho_\mathcal{c}(q_0,p_0) &= \int_\mathcal{S} \delta_{q_0,p_0} \cdot \rho_{\mathcal{c}} d \mu = \langle \delta_{q_0,p_0} , \rho_{\mathcal{c}} \rangle \\
-p_{total}(\mathcal{c}) &= \int_\mathcal{S} p \cdot \rho_{\mathcal{c}} d \mu = \langle p , \rho_{\mathcal{c}} \rangle \\
-q_{avg}(\mathcal{c}) &= \frac{\int_\mathcal{S} q \cdot \rho_{\mathcal{c}} d \mu}{\int_\mathcal{S} \rho_{\mathcal{c}} d \mu} = \frac{\langle q , \rho_{\mathcal{c}} \rangle}{\langle 1 , \rho_{\mathcal{c}} \rangle}
-\end{align*}
-For each state, they represent respectively the total amount of material, the material within $U$, the density at $(q_0, p_0)$, the total momentum and the average of a state variable.
-As they are linear functionals of $\mathcal{C}$,  they can be expressed using their dual vector.
-But these vectors ($1_U$, $p$, ...) are \emph{not} elements of $\mathcal{C}$.
-They do not represent states, they are not continuous and integrable, and there is no single general physical meaning for all of them.
-Therefore, as we mentioned before, while we can extend the function space for convenience, computation purposes or to study a limit case, we need to be mindful of the extension and carefully check that the mathematical results are physically meaningful case by case.
-
-\subsection{Invariant densities and cotangent bundle}
-
-Now that we are guaranteed that the density is properly defined for all state variables, we have to make sure it is invariant.
-The distribution should only depend on the states, and not the particular choice of state variables to label them.
-
-If $\rho_{\mathcal{c}}$ is to remain invariant under state variable changes, not only does the Jacobian have to exist but it must be unitary.
-This means we cannot simply change one state variable as we please.
-If we change one at least another has to change in some coordinated way such that the Jacobian of the total transformation is unitary.
-This  means that we cannot change physical units of all the variables as we like: they are part of a unit system.
-
-The simplest case is when a single variable is enough to define our units, and therefore tell how all the others must change.
-Suppose that our state space $\mathcal{S}$ is identified by $n$ independent state variables $\{q, k^i\}$ and a change of units for the first, that is $\hat{q} = \hat{q}(q)$, determines the change for all others.
-The Jacobian matrix is the block matrix:
-\begin{align*}
-J =  \left[
-\begin{array}{cc}
-d_q\hat{q} & 0 \\
-\partial_q \hat{k}^j & \partial_i \hat{k}^j \\
-\end{array}
-\right] 
-\end{align*}
-The Jacobian has to be unitary, therefore all elements $\partial_i \hat{k}^j$ must be well defined and $|\partial_i \hat{k}^j| = 1/d_q\hat{q}$.
-As this can only happen if $\partial_i \hat{k}^j$ is a $1\times1$ matrix, there can only be one $k$.
-This gives $\partial_{k} \hat{k} = d_{\hat{q}} q$ and $\hat{k} = d_{\hat{q}} q \, k + f(q)$.
-As this is a change of units, we expect the zero value of $k$ to remain constant: $\hat{k}(q, 0) = 0 = f(q)$.
-Therefore $\hat{k} = d_{\hat{q}} q \, k$ transforms as the component of a covariant vector.
-$\mathcal{S}$ is isomorphic to $T^*\mathcal{Q}$, the space of co-vectors at a point, where $\mathcal{Q}$ is the manifold that defines our unit.
-
-Physically, this means that the state variable $k$, which is the classical analogue of the wave number, is expressed with the inverse of the unit used for $q$.
-If $q$ is in meters, $k$ is in inverse meters.
-Consider now the area $dq \wedge dk$ of an infinitesimal rectangular region.
-This quantity is dimensionless and therefore invariant.
-The total number of states will be proportional to it, as  doubling the range of $dq$ or $dk$ will give us double the number of possible states.
-We have $d\mu = \hbar dq \wedge dk$, where $\hbar$ is the proportionality constant that will depend on the unit chosen to count the possibilities of the pair $(q, k)$.
-That is, a unitary range of possibilities for $q$ and $k$ will give us $\hbar$ possibilities.\footnote{The actual value and physical dimensions of $\hbar$ are determined by the system of units, and should not be taken to describe some intrinsic physical property.
-Only dimensionless relationships to other physical constants, such as the fine structure constant $\alpha = e^2/\hbar c$, can possess that trait.
-That is why one can choose ``natural units" for which $\hbar=1$.
-In SI units the relationship $\hbar dk = dp = m du$ between kinetic and conjugate momentum, derived later, sets the relationship between units.} The state count in a finite (i.e.
-compact) region $U \subseteq \mathcal{S}$ will be given by $\mu(U) = \int_U \hbar dq \wedge dk$: we are finally able to write the volume of integration in terms of state variables.
-
-We can generalize to the case where $\mathcal{Q}$ has more than one dimension.
-As we must be able to change one unit at a time, independently from the other, to each $q^i$ will correspond a $k_i$ that uses the inverse of the corresponding units.
-We have $\mathcal{S}\cong T^*\mathcal{Q}$ and $d\mu = \hbar^n dq^n \wedge dk_n$.
-But it's not just the volume, the state count, that is preserved as we change state variables.
-Each pair $(q^i, k_i)$ is expressed by an independent unit with the count of possibilities given by $\hbar dq^i \wedge dk_i$.
-We call such a pair a degree of freedom.
-As they are independent the total state count is the product of the possibilities of each d.o.f.: $d\mu = \hbar^n dq^n \wedge dk_n = \prod \limits_{i=1}^n \hbar dq^i \wedge dk_i$.
-In other words: independent d.o.f.
-are orthogonal surfaces in $\mathcal{S}$.
-The possibility count of an arbitrary degree of freedom, then, will be $\sum \limits_{i=1}^n \hbar dq^i \wedge dk_i$, the sum of the projections over the $n$ orthogonal and independent d.o.f.
-defined by $(q^i, k_i)$.
-
-The possibility count for each d.o.f.
-(i.e.
-the wedge product within a d.o.f.) and the orthogonality of different d.o.f.
-(i.e.
-the scalar product across d.o.f.) must be the same regardless of the choice of state variables.
-We can express both requirements mathematically in a compact way.
-We first define conjugate momentum as $p_i=\hbar k_i$ and unified state variables as $\xi^a\equiv \{q^i, p_i\}$.
-Then we consider the canonical symplectic two-form $\omega =\Sigma \, dq^i \wedge dp_i$ given by the following components:
-\begin{align*}
-\omega_{ab} =  \left[
-\begin{array}{cc}
-0 & 1 \\
--1 & 0 \\
-\end{array}
-\right] \otimes I_n =
-\left[
-\begin{array}{cc}
-0 & I_n \\
--I_n & 0 \\
-\end{array}
-\right] \\
-\end{align*}
-It returns the wedge product within a d.o.f.
-and the scalar product across.
-Requiring the invariance of this metric under state variable changes assures us the count of states and possibilities is well defined.
-
-\begin{figure}
-	\includestandalone[width=\columnwidth]{pictures/independentDof}
-	\caption{The symplectic form $\omega$ and its components along two independent degrees of freedom.
-As we change state variables, the new d.o.f.
-are also independent and therefore orthogonal to each other.
-Note: this is a 2D conceptualization of 4D space.}
-	\label{fig:independent_dof}
-\end{figure}
-
-This gives us insight on the physical meaning of the geometrical structure $T^*\mathcal{Q}$.
-The canonical one-form $\theta=k dq$ represents the geometrical object we associate with each particle state.
-The symplectic form $\omega$ is the bilinear that quantifies the possibilities described by two given state variables.
-Note, instead, that the relationship $\omega = - \hbar d\theta$, while mathematically true, has no clear physical meaning as we have not defined what the exterior derivative on a state actually represents.
-
-We can capture the above discussion by stating that the state space $\mathcal{S}$ is the symplectic manifold $(T^*\mathcal{Q}, \omega)$, where $\mathcal{Q}$ is the manifold that defines the unit system.
-This is the only manifold that allows us to define state-variable-invariant densities.
-The symplectic form allows us to count possibilities on an arbitrary d.o.f.
-
-\begin{prop}\label{prop:symplectic_manifold}
-The state space $\mathcal{S}$ for the particles of a classical material is a symplectic manifold formed by a cotangent bundle $T^*\mathcal{Q}$ equipped with the canonical two-form $\omega$.
-\end{prop}
-
-\begin{justification}
-	We claim the simplest (i.e.
-lowest dimension) state space $\mathcal{S}$ that allows state-variable-invariant densities is the cotangent bundle $T^*\mathcal{Q}$ of a single dimensional manifold $\mathcal{Q}$.
-Consider a set of state variables for $U \in \mathcal{S}$: the simplest unit system is one defined by a single state variable.
-Let $q$ be the state variable that defines the unit system.
-Let $\mathcal{Q}$ be the manifold charted by $q$.
-A diffeomorphism $\hat{q}=\hat{q}(q)$ fully defines a change of units.
-Let $(q,k^i)$ with $i=1...n-1$ be a set of state variables.
-The Jacobian determinant $|J| = d_q\hat{q} |\partial_i\hat{k}^j|=1$ as densities must be invariant.
-$|\partial_i \hat{k}^j| = d_{\hat{q}}q$.
-$\partial_i \hat{k}^j$ is one dimensional as its components must be fully specified by the previous relationship.
-Let $k=k^1$.
-$\partial_k \hat{k} = d_{\hat{q}}q$.
-$\hat{k}(q,0) = 0$ as a change in units does not change the $0$ value.
-$\hat{k} = d_{\hat{q}}q k$.
-$k$ changes as the component of a co-vector.
-Each state is identified by a point and a co-vector in $\mathcal{Q}$.
-The simplest state space $\mathcal{S}$ is isomorphic to $T^*\mathcal{Q}$ where $Q$ is the one dimensional manifold that defines the units.
-
-	We claim the simplest state space $\mathcal{S}$ that allows state-variable-invariant densities is the symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of a one dimensional differentiable manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
-Express integration in coordinates.
-$\int_{\mathcal{S}} \rho_\mathcal{c} d\mu \propto \int_{\mathcal{S}} \rho_\mathcal{c}(q, k) dq \wedge dk$.
-$d\mu = \hbar dq \wedge dk = dq \wedge dp = \omega$ where $\hbar$ is a constant, $p\equiv \hbar k$  and $\omega$ is the canonical two form.
-$\omega$ is invariant under state variable change.
-$\omega$ is the symplectic form for $\mathcal{S}=T^*\mathcal{Q}$.
-$\mathcal{S} = (T^*\mathcal{Q}, \omega)$ is a symplectic manifold.
-
-	We claim the state space $\mathcal{S}$ for the particles of a homogeneous classical material is a symplectic manifold $(T^*\mathcal{Q}, \omega)$ where $T^*\mathcal{Q}$ is the cotangent bundle of an $n$-dimensional differentiable manifold $\mathcal{Q}$ and $\omega$ is the canonical two-form.
-Let $q^i$ be a set of n continuous independent state variables that define the units necessary to describe a state in $U \subseteq \mathcal{S}$.
-Let $\mathcal{Q}$ be the manifold charted by $q^i$.
-Locally $\mathcal{Q} \cong \prod \mathcal{Q}^i \cong \mathbb{R}^n$.
-Changing units of one state variable must not change the units of the other as they are independent.
-State-variable-distribution can  be defined separately on each degree of freedom.
-For each $\mathcal{Q}^i$ we have a covector $k_i(q) dq^i \in T^*\mathcal{Q}^i$.
-Locally $\mathcal{S} \cong \prod T^*\mathcal{Q}^i \cong T^* \prod \mathcal{Q}^i \cong T^* \mathcal{Q}$.
-Integration must also be defined on an independent d.o.f.
-There must exist a non-degenerate two-form $\omega$ such that $\int_{U \subset \mathcal{S}} \rho_\mathcal{c} \omega$ where $U$ is any two dimensional surface in $\mathcal{S}$.
-$\omega$ has to be form invariant under state variable change.
-The canonical two-form is the only such form.
-$\omega = \sum dq^i \wedge dp_i$.
-$\mathcal{S} = (T^*\mathcal{Q}, \omega)$ is a symplectic manifold.
-\end{justification}
-
-It should be clear by now that the case of discrete topology is qualitatively different from the standard topology for $\mathbb{R}^n$.
-The notion that the continuous case is a limit of the discrete case, that it's ``like the discrete but with more points", leads in this case to erroneous intuition.
-The key question is: can we physically distinguish an isolated state?
-Can we have an outcome associated with only one element?
-The answer is no with the standard topology on $\mathbb{R}^n$.
-The consequence is that when we define the measure $\mu$ for the count of continuous states, we assign a finite value to a finite region, and we assign zero measure to a single state.
-That is what gave us integration, densities and, ultimately, conjugate variables.
-If we were to use the discrete topology on $\mathbb{R}^n$, if we were to assume we can identify single states as we do for countable states, then we would assign measure one to each state and infinity to a finite region.
-We would not have integration, just a simple sum.
-Our distribution would not be a density and there would be no justification for conjugate variables.
-A finite distribution could only distribute a finite amount of material in a finite number of states.
-
-In other words: this is not a case where a difference at small scales leads to a small difference at large scales.
-The two cases are radically different.
-We can tell them apart.
-We should take the use of densities and conjugate quantities as evidence that quantities like space and time are not discrete in the sense that the processes we use to distinguish those quantities cannot identify single instances.
-The quantum case does not change this, as the use of densities (in the form of the wave function) and conjugate quantities is even more prominent.\footnote{If one wants to posit that space is fundamentally discrete at scales where both classical and quantum mechanics cease to be valid, the better strategy would be to work with a discrete version of $T^*\mathcal{Q}$ and not $\mathcal{Q}$ by itself.
-Though it is not clear how a discrete topology would become non-discrete in the limit, at least the measure may be workable.}
-
-It should also be noted that classical particles, under this light, cannot be considered point-like.
-As they are the limit of infinitesimal subdivision, their spatial extent becomes infinitesimal but not zero.\footnote{This picture is also compatible with general relativity, unlike point-like particles.
-These would not follow geodesics as their infinite mass density would significantly affect the gravitational field.} But suppose particles were truly point-like.
-Then distinguishing particles would mean distinguishing points.
-We are back to the idea of a discrete topology on $\mathbb{R}^n$.
-In that case we would not have conjugate momentum, no $T^*\mathcal{Q}$, just the coordinates of the point in $\mathcal{Q}$.
-This would actually be more self-consistent: why wouldn't a point be enough to define the state of a point-like particle?
-As before, we should take the use of conjugate quantities as evidence that classical particles are really the limit of an infinitesimal subdivision.
-
-Finally, we should note that we are in a position similar to the one discussed in \ref{prop:discrete_measure} for discrete states.
-We have a topological space plus a measure that allows to count states and a symplectic form that allows to count possibilities, both expressible in terms of state variables.
-We recovered this structure starting from the idea of an infinitesimally reducible system.
-That led to a state space $\mathcal{S}$ for the infinitesimal parts, on which we must be able to define state-variable-independent distributions, which in turn gave us degrees of freedom made by pairs of conjugate variables and the symplectic form typical of classical phase space.
-In short: being able to measure the amount of material is what allows us to count states.
-
-\subsection{Infinitesimal reducibility}
-
-Now that we have fully characterized what we mean by a classical material, we can stipulate the following:
-
-\begin{assump}[Infinitesimal reducibility]\label{ass:infinitesimal_reducibility}
-	The system under study is composed of an infinitesimally reducible homogeneous material and each part undergoes deterministic and reversible evolution.
-\end{assump}
-
-\begin{rationale}
-	The idea is that time evolution specifies a map for the state space $\mathcal{S}$ of each infinitesimal part.
-Knowing how the parts evolve tells us how a composite state $\mathcal{c}$ evolves as well.
-
-	Consistent with what we said in assumption \ref{ass:determinism}, if we defined a state for each particle, then a deterministic and reversible evolution on that state must exist.
-Yet, the idea that we can assign states to infinitesimal parts should be considered only a \emph{simplifying} assumption.
-The obvious reason is that we know this does not work in practice: as we keep decomposing the material we end up with molecules, atoms and then subatomic particles.
-But it is instructive to understand when and how the assumption breaks down at a more conceptual level.
-
-	The first problem is methodological.
-To be able to talk about the states of a part we need a physical process that is able to distinguish between them.
-For a billiard ball we can imagine marking one spot with a red marker.
-This allows us to track it as the ball moves or collides with other balls.
-For an electron, instead, we do not have a way to mark a piece.
-In fact when two electrons scatter we can't even tell which is which, let alone what portion went where.
-The classical assumption may not hold because we do not have suitable physical processes at our disposal.
-
-    Even if we are able to track parts, the assumption requires them to be infinitesimal, the limit of a process of infinite recursive subdivision.
-The best we can do experimentally is to confirm that the assumption holds up to the smallest precision available.
-Therefore even in the best of cases it cannot be considered an experimentally validated assumption but a reasonable default assumption (e.g.
-``it worked so far").
-
-	The second, more conceptual, problem is that the idealizations that allowed us to define states in assumption \ref{ass:determinism} may not hold as the parts get smaller.
-Recall the cannonball whose motion is sufficiently unaffected by the photons that scatter off its surface.
-As we consider smaller and smaller parts, at some point we will find an amount of material that is affected by the interaction with the air or photons scattering off of it.
-At that point the parts are no longer sufficiently isolated to define an independent state, their evolution depends on the environment.
-Therefore the deterministic evolution of the whole cannot be reduced to the deterministic evolution of its parts.
-
-	Another issue is that defining a state requires some kind of asymmetry between system under study and environment.
-The future state of the system does not depend on the state of the environment, so that we can define an independent state, yet the future state of the environment is affected by the state of the system, so that we can have external processes that allow us to define physical distinguishability.
-If we try to assume that both the system and environment are really made of the same classical material, then the claim to that asymmetry is lost: the infinitesimal parts of both system and environment must affect one another in the same way.
-
-	Another issue arises if we include the measuring device in our description. Ideally, we'd like to require the following:
-	\begin{itemize}[noitemsep]
-		\item The system under study is deterministic and reversible
-		\item The measuring device ascertains the state of the system
-		\item System and measuring device, as a whole, is deterministic and reversible
-	\end{itemize}
-	Unfortunately, these three requirements together are inconsistent.
-Say $a \in A$ is the initial state of our system and $b \in B$ the initial state of our measuring device: the final state $\hat{a}=\hat{a}(a)$ because the system is deterministic; $a=f(\hat{b})$ because we are able to know the initial state of the system by looking at the final state of the measuring device; $b=b(\hat{a},\hat{b})$ because the system as a whole is reversible.
-Since $\hat{a}$ is determined by $a$, and $a$ can be known from $\hat{b}$, $b = b(\hat{a}(f(\hat{b})), \hat{b}) = b(\hat{b})$ can be determined from just $\hat{b}$.
-But since $a=f(\hat{b})$, the whole past state could be reconstructed just by looking at the future state of the measuring device.
-But $(\hat{b}) \rightarrow (a, b)$ cannot be an injection, therefore we arrive at a contradiction.
-
-	Of those three requirements we can only pick two.
-If we want determinism and reversibility for the combined system, either both systems are not independently deterministic or the second is not a measuring device.
-In other words: we can't expect to have deterministic and reversible evolution at all levels of aggregation and have the parts interact in any physically meaningful way.
-
-	The same problem of physically meaningless interaction surfaces when we consider the states of infinitesimal parts.
-Under our classical assumption, if we were to take the state of an infinitesimal part to really be its full description, with no unstated part, then each piece would evolve independently, oblivious to the other parts.
-Each particle would essentially reside in its own separate physical universe, as it cannot physically distinguish anything else.
-This is not physically meaningful.
-
-	If we assumed the system is deterministic and reversible only as a whole, then each part could evolve depending on the states of the other parts.
-This would seem much better, as the state of the pieces is still exhaustive yet they are physically connected to each other.
-But this does not actually solve the problem of independent oblivious components.
-First, for this to work, one would have to specify how the states of the parts were defined since their evolution is individually no longer deterministic (the future of each part depends on the state of other parts).
-But ignoring that issue, the bigger problem is that we can always locally separate the evolution into independent degrees of freedom.
-For example, the position and momentum of two particles may affect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle \cite{Linares}} We can always find such a local decomposition, and the easiest way to see that is in terms of the transported variables: they retain the original value; they clearly evolve independently.
-And since the pieces are infinitesimal, a local decomposition is all that is needed.
-While such description may be cumbersome to achieve in practice, conceptually it is still possible.
-So, even if each particle evolution depends on the state of the others, the system is described by degrees of freedom that evolve independently, oblivious to each other.
-
-	The moral is that the classical idea of being able to assign to all systems a state which represents their full description does not work.
-The state is only what we can describe and it can't be the full description.
-As the division between system and environment is subjective, each system must be able to function as both.
-Therefore it will have a part whose evolution depends only on the system itself, the state, and a part whose evolution depends on other systems, the unstated part.
-While they may not be the same in all circumstances, they must exist.
-The state is what gives the identity to the system, the part we can study and describe.
-The unstated part is what allows continuous interaction between the system and the environment; it's what allows us to study and describe the state.
-As we'll see later, it is precisely this problem that quantum mechanics conceptually addresses better than classical mechanics.
-
-	One final problem is with time itself: an infinitesimally reducible system undergoing deterministic and reversible evolution cannot tell time.
-If it did, we would be able to find a quantity that changes through time but is invariant under state variable changes.
-The problem is that deterministic and reversible time evolution is equivalent to a state variable change.
-In fact, at a fixed time, we can choose to describe the system with the transported state variables of any past or future times.
-So, all quantities that are invariant under state variable changes are also invariant under time evolution.
-The somewhat ironic result is that while time is essential for the very definition of deterministic and reversible evolution, time itself has to be defined by some other type of process.\footnote{For example, if the phase-space volume occupied by a distribution increased, it would give us a coordinate-invariant quantity that changes in time.
-This, however, is not possible under Hamiltonian evolution.}
-
-	In light of what we discussed, we cannot take the classical assumption to be fundamental, in the sense that we cannot take it to strictly apply to all of the universe.
-While ultimately flawed, the  assumption can be considered valid for a great number of macroscopic systems, and is also useful as a default assumption of sorts.
-Understanding its shortcomings will help us later to see how the quantum assumption solves, at least partially, some of these issues.
-
-	As a final note: we caution against automatically thinking that the classical assumption is valid for all macroscopic systems.
-It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, in which case the assumption would not hold.\footnote{For example, this is the case in Bose-Einstein condensates \cite{BEC}.}
-\end{rationale}
-
-\subsection{Hamiltonian mechanics}
-
-We are finally ready to write the equations of motion.
-As per \ref{prop:homeomorphism} our evolution is at least a self-homeomorphism $f:\mathcal{S} \leftrightarrow \mathcal{S}$.
-But this is not enough.
-
-The evolution must map the distribution point-wise so that the value associated at the initial state is the same as the value at the final state.
-All the material that starts in $\mathcal{s}$ has to end up in $\hat{\mathcal{s}}$.
-In math terms $\rho_{\hat{\mathcal{c}}} (\hat{\mathcal{s}}) = \rho_\mathcal{c}(\mathcal{s})$.
-
-In the same way we expect the total amount of material to be conserved.
-If $U \in \mathcal{S}$ is a set of initial particle states and $\Lambda_U(\mathcal{c})$ is the amount of material associated with that set, then we expect it to be equal to the amount of material $\Lambda_{\hat{U}}(\hat{\mathcal{c}})$ associated with the set of final states $\hat{U} \in \mathcal{S}$.
-
-But probably the easiest way to look at it is that initial and final sets of states have to possess the same count of states and possibilities as defined by the symplectic form $\omega$.
-Therefore the area within a degree of freedom, which represents the number of possibilities in said d.o.f., needs to be mapped to an equal area within the transported d.o.f.
-(i.e.
-the d.o.f.
-defined by the transported state variables).
-Independent d.o.f.
-must remain independent, and therefore transported orthogonal d.o.f.
-remain orthogonal.
-This means that the product of possibilities of independent d.o.f.
-is also conserved.
-These statements are the physical justification of Gromov's non-squeezing theorem \cite{Gromov,deGosson,Stewart} and Liouville's theorem.
-And they give intuitive insight on the geometry of Hamiltonian systems.
-
-Mathematically, under the classical Hamiltonian assumption,  deterministic and reversible evolution is a self-symplectomorphism (or self-isometry or canonical transformation depending on your math training).
-That is, it does not just preserve the topology but also the symplectic form $\omega$.
-
-\begin{prop}\label{prop:symplectomorphism}
-	A deterministic and reversible evolution map for the particles of a classical material is a self-symplectomorphism.
-That is: $\mathcal{T}_{\Delta t}: T^*\mathcal{Q} \rightarrow T^*\mathcal{Q}$ and $\mathcal{T}_{\Delta t}^*\omega = \omega$ where $\mathcal{T}_{\Delta t}^*$ is the pullback of $\mathcal{T}_{\Delta t}$.
-\end{prop}
-
-\begin{justification}
-	We claim $\mathcal{T}_{\Delta t}$ is a self-homeomorphism on $T^*\mathcal{Q}$.
-The state space for the particles of a classical material is $T^*\mathcal{Q}$ by \ref{prop:symplectic_manifold}.
-$\mathcal{T}_{\Delta t}$ is a deterministic and reversible evolution map and by \ref{prop:homeomorphism} is a self-homeomorphism.
-
-	We claim $\mathcal{T}_{\Delta t}$ is a symplectomorphism on $(T^*\mathcal{Q}, \omega)$.
-The distribution on final states must be defined  therefore the Jacobian for $\mathcal{T}_{\Delta t}$ exists and is non-zero: $\mathcal{T}_{\Delta t}$ is a diffeomorphism.
-A deterministic and reversible process conserves the number of states and possibilities.
-$\omega$ is the two-form that returns the count of possibilities.
-$\omega$ is invariant under deterministic and reversible evolution.
-$\mathcal{T}_{\Delta t}$ is a symplectomorphism by definition.
-\end{justification}
-
-If we assume continuous time evolution we have the following:
-
-\begin{prop}\label{prop:hamiltons_equations}
-	A continuous deterministic and reversible evolution for the particles of a classical material admits a Hamiltonian $H \in C^2(T^*\mathcal{Q}, \mathbb{R})$ that allows us to write the laws of evolution as
-\begin{align*}
-d_{t}q^i &= \partial_{p_i} H \\
-d_{t}p_i &= - \partial_{q^i} H
-\end{align*}
-\end{prop}
-
-\begin{justification}
-We claim the vector field $S \in T\mathcal{S}$ for the infinitesimal displacement $S^a = d\xi^a/dt$ admits a potential $H \in C^2(T^*\mathcal{Q}, \mathbb{R})$ such that $S^{a} \omega_{ab} = \partial_{b}H$.
-The state space $\mathcal{S}$ for the particles of a classical material is a symplectic manifold by \ref{prop:symplectic_manifold}.
-The map for infinitesimal time evolution $\mathcal{T}_{dt}$ is an infinitesimal self-symplectomorphism by \ref{prop:symplectomorphism}.
-By \ref{symplectomorphism_generator} the infinitesimal displacement $S$ admits a potential $H \in C^2(T^*\mathcal{Q}, \mathbb{R})$ such that $S^{a} \omega_{ab} = \partial_{b}H$
-
-We claim the state variables evolve according to Hamilton's equations.
-$S^{a} \omega_{ab} = d_t\xi^a \omega_{ab} = \partial_{b}H$.
-For $b=\{1,...,n\}$ we have $d_tp_i (-1) = \partial_{q_i} H$.
-For $b=\{n+1,...,2n\}$ we have $d_tq^i (+1) = \partial_{p_i} H$.
-\end{justification}
-
-We recognize the familiar set of Hamilton's equations.
-They are the set of equations that describe the deterministic and reversible motion of the infinitesimal parts of an infinitesimally reducible homogeneous material.
-Note that the argument goes the other way as well.
-A system governed by Hamilton's equations is deterministic and reversible: the equation of motion given by $H$ are differentiable since $H$ is twice differentiable.
-Therefore the equations are Lipschitz continuous and, by the Picard-Lindel\"{o}f theorem, a unique solution exists \cite{Grant}.
-
-In other words: the forces that conserve energy (i.e.
-the value of a suitable Hamiltonian) are exactly the ones that provide deterministic and reversible motion.
-The challenge in their derivation mainly lies in the necessary use of different branches of mathematics (e.g.
-topology, measure theory, differential geometry, symplectic geometry).
-The conceptual meaning, on the other hand, is hopefully straightforward.
-
-It's important to realize that, during the derivation, multiple mathematical features (e.g.
-invariant densities, cotangent bundle for phase space, symplectomorphism) were justified by the same physical assumption.
-The math itself gives us no indication that the different features stem from the same source; therefore, the math itself does not give us a conceptually unified picture.
-This is one of the reasons we believe that starting from the physical description is objectively better if we are to come to a better understanding of our physical theories.
-
-Note that we could have taken different approaches.
-For example, we could have appealed to information theory and required our invariant distributions to preserve Shannon's information entropy\cite{Shannon}, as no information should be gained or lost during a deterministic and reversible process.
-Or we could have appealed to statistical mechanics and required that the determinant of the covariance matrix be conserved, as a deterministic and reversible process should be defined at the same level of precision.
-With suitable treatment of independent d.o.f.
-both these approaches recover Hamiltonian mechanics as well.
-While the full treatment is outside the scope of this work, we want to underline that the definitions used in this work go a long way to building bridges at the core of different mathematical and scientific branches.
-
-\section{Time dependent evolution}
-\label{sec:relativity}
-
-We now generalize our discussion to include time dependent dynamics.
-This is needed when the evolution map is not the same at all instants or when state variables depend on time (e.g.
-changing to a moving frame).
-
-To do this we will redefine the particle state space to include the temporal degree of freedom.
-We'll find that the dynamics is described by relativistic Hamiltonian mechanics in the extended phase space.
-We'll also find that particle states divide into standard and anti-states depending whether time is aligned or anti-aligned with the evolution parameter.
-Note that no extra assumption is needed, which makes relativistic mechanics simply the correct time dependent description of the deterministic and reversible evolution of an infinitesimally reducible system.
-
-\subsection{Time changes}
-
-The discussion so far has been limited to the case where both the state variables and the evolution map never change throughout the evolution.
-This is too restrictive as there are very reasonable situations in which this does not hold.
-
-The first obvious case is that the map may be time dependent.
-This does not affect our definition of determinism and reversibility as we still can tell final state from initial state and vice-versa.
-But at this point we have no way to specify a time dependent evolution: the Hamiltonian $H: T^*\mathcal{Q} \rightarrow \mathbb{R}$ as we derived it is just a function of the state.
-
-The second case is when we perform a transformation for the time parameter $\hat{t}=\hat{t}(t)$.
-The equations of motion transform and we have $d_{t}\hat{t} \, d_{\hat{t}}q^i = \partial_{p_i} H$.
-$q^i: T^*\mathcal{Q} \rightarrow \mathbb{R}$ and $H: T^*\mathcal{Q} \rightarrow \mathbb{R}$ are real functions of the state space and cannot be redefined to include $d_{t}\hat{t}$.
-Physically the transformation is introducing fictitious forces that are not conservative, so they cannot be expressed by a Hamiltonian.
-But this means that we have an ill defined mathematical framework as the notion of determinism and reversibility is not defined in a way that is independent of time transformations.
-We need a framework that is capable of handling the fictitious forces as well.
-
-The third case is when we perform a state variable transformation $\hat{q}=\hat{q}(q,t)$ that is time dependent (e.g.
-$\hat{q}=q+vt$).
-Our composite state distribution $\rho(q)$ becomes $\rho(\hat{q},t)$: it is no longer defined at equal time.
-This means that our measure $\mu$ needs to be modified to define integrals over the time variable.
-
-As we can see, the framework we have is ill suited to handle these cases.
-Therefore we cannot simply stick time in the Hamiltonian and expect everything to work out.
-But we can't ignore the problem either as the three cases outlined are common situations to study.
-So we need to go back to our definitions and amend them properly.
-
-\subsection{Complete state space}
-
-The first thing to do is to amend our definition of state space to include all states at all times.
-It may seem like we are extending our definition of state and state space but, at a closer look, we are not.
-Since we defined state as a physical configuration at a particular time, the set of all states (i.e.
-the state space) is more accurately defined as the set of all configurations at all times.
-In the previous sections we tacitly simplified the problem by ignoring time, which made it easier to study the time independent case.
-This approach is common in physics and engineering, which is why it fits the standard names, but this is also the source of the above problems.
-
-We'll therefore call \emph{complete state space} of particle states the set of all configurations at all times and we'll indicate it by $\bar{\mathcal{S}}$.
-In that space, a state $\mathcal{s} \in \bar{\mathcal{S}}$ will be given by $(q^i, p_i, t)$, making $\bar{\mathcal{S}}$ a $2n+1$ dimensional manifold.
-A region at equal time will be a hypersurface $\mathcal{S}_{t=t_0} \subset \bar{\mathcal{S}}$ that includes all possible configurations at a particular time.
-Such hypersurface will need to cut across all evolutions once and only once, therefore not all functions $t : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ are suitable time variables.
-A time evolution is a line  $\bar{\mathcal{s}} \subset \bar{\mathcal{S}}$ that passes once through each equal time surface.
-Specifying an evolution map means giving a set of lines such that each state is traversed once and only once.
-
-\begin{prop}\label{prop:complete_particle_state_space}
-	The complete state space $\bar{\mathcal{S}}$ for the particles of a classical material is a $2n+1$ dimensional differentiable manifold.
-The state space $\mathcal{S}_{t=t_0}$ at a particular time is a hypersurface of $\bar{\mathcal{S}}$.
-\end{prop}
-
-\begin{justification}
-	We claim $\bar{\mathcal{S}}$ is a $2n+1$ dimensional differentiable manifold.
-Locally there exists $U\subseteq \bar{\mathcal{S}}$ such that $U \cong T^*\mathcal{Q}\times \mathbb{R}$ as a state $\mathcal{s} \in U$ exists for each configuration and each time instant.
-Locally $\bar{\mathcal{S}} \supseteq U \cong\mathbb{R}^{2n+1}$.
-$\bar{\mathcal{S}}$ is a $2n+1$ dimensional manifold.
-$T^*\mathcal{Q}$ is a differentiable manifold.
-Time evolution is a diffeomorphism by \ref{prop:symplectomorphism} therefore derivatives between state variables and time variables always exist.
-$\bar{\mathcal{S}}$ is a differentiable manifold.
-
-	We claim $\mathcal{S}_{t=t_0}$ is a hypersurface of $\bar{\mathcal{S}}$.
-Let $t : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ be a time variable.
-$\mathcal{S}_{t=t_0} = \{ \mathcal{s} \in \bar{\mathcal{S}} \; | \; t(\mathcal{s}) = t_0 \}$ is a level set.
-$t$ is monotonic as it is a time variable.
-$t$ has no critical points.
-$t_0$ is a regular value.
-$\mathcal{S}_{t=t_0}$ is an embedded submanifold of co-dimension $1$.
-$\mathcal{S}_{t=t_0}$ is a hypersurface.
-\end{justification}
-
-For the composite state we need to take a different approach.
-A distribution of material over states at equal time $\rho_\mathcal{c} : \mathcal{S}_{t=t_0} \rightarrow \mathbb{R}$ will depend on the choice of time variable, therefore the complete state space will not be useful as it will not be time invariant.
-What we are really interested in is the \emph{state evolution space} of a classical material, which we'll indicate by $\bar{\mathcal{C}}$.
-To each state evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$ will correspond a distribution $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ over all particle states at all times.
-This distribution will be differentiable: it's differentiable over all equal time hypersurfaces and time evolution is differentiable.
-The distribution, though, is not integrable over the complete state space $\bar{\mathcal{S}}$.
-Such integration will operate over time as well, giving the amount of material multiplied by time.
-If time is taken to be infinite the value is not finite.
-The distribution, instead, must be integrable over all equal time hypersurfaces: no matter what time variable we choose, the amount of material found at a particular time has to be finite.
-
-\begin{prop}\label{prop:state_evolution_space}
-	The state evolution space $\bar{\mathcal{C}}$ for a classical material is isomorphic to the space of differentiable functions that are Lebesgue integrable on any equal time hypersurface.
-That is $\bar{\mathcal{C}} \cong  C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$ where $L^1_t(\bar{\mathcal{S}}, \mu)=\{{\rho: \bar{\mathcal{S}} \rightarrow \mathbb{R}} \; | \; {\int_{\mathcal{S}_{t=t_0}} |\rho| d\mu < \infty} \; \forall \, t, t_0 \}$.
-\end{prop}
-
-\begin{justification}
-	We claim $\bar{\mathcal{C}} \cong  G \subseteq C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$.
-Let $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ be the distribution of material associated with an evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$.
-Let $t : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ be a time variable.
-	The restriction of $\rho_{\bar{\mathcal{c}}}$ over an arbitrary level set $\mathcal{S}_{t=t_0}$ is continuous and integrable by \ref{prop:differentiable_manifold}.
-$\rho_{\bar{\mathcal{c}}} \in L^1_t(\bar{\mathcal{S}}, \mu)$.
-Time evolution is a diffeomorphism by \ref{prop:hamiltons_equations}.
-$\rho_{\bar{\mathcal{c}}}$ is differentiable along the time variable as well as the state variables.
-$\rho_{\bar{\mathcal{c}}} \in C^1(\bar{\mathcal{S}})$.
-
-	We claim $\bar{\mathcal{C}} \cong  C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$.
-As in \ref{prop:differentiable_manifold}, if two distributions are different there exists an outcome that can tell them apart, therefore they represent two distinct physical evolutions.
-\end{justification}
-
-Before proceeding, it's useful to get a better intuition for the geometry of these spaces.
-Note that the transported state variables are perfect to label the particle state evolution $\bar{\mathcal{s}} \subset \bar{\mathcal{S}}$: as they do not change during the evolution we can take them to label not just the values at a particular time, but the evolutions themselves.
-The distribution is transported deterministically and reversibly over those lines, so the value of the distribution will be constant along each particle state evolution $\bar{\mathcal{s}}$.
-To identify the full evolution, then, we just need to specify the distribution only on a region that cuts across all particle evolutions, i.e.
-a hypersurface at constant time.
-
-\subsection{Relativistic cotangent bundle}
-
-We now want to extend the symplectic form so we can count states and possibilities over the complete state space $\bar{\mathcal{S}}$.
-The first thing to do is to complete our $n$-dimensional $\mathcal{Q}$, the manifold that defines our units, to include the time variable $t$.
-This, again, highlights our previous oversight: we didn't include time within the quantities needed to specify a physical configuration at a particular time.
-We call $\mathcal{M}$ such a space which is a manifold of dimension $n+1$: state variables and the time variable are independent (i.e.
-any combination of state and time variables is valid).
-
-As we said before, the time variable is not simply another state variable: it doesn't identify extra configurations and it's the variable on which we have defined deterministic and reversible evolution.
-When changing variables, we still need to make sure that deterministic and reversible motion can be defined on the new time variable.
-Transformations like $\hat{t}=q, \; \hat{q}=t$ (exchanging time with another state variable) or $\hat{t}=\sqrt{t^2 + q^2}, \; \hat{q}=\tan^{-1}(t/q)$ (polar coordinates between time and state variable) clearly do not guarantee deterministic and reversible motion over the new time variable.
-To be meaningful, the change of time variable must preserve time ordering therefore the transformation has to be strictly monotonic between the two time variables.
-On the other hand, the change of state variables must preserve the ability to uniquely identify states at each instant in time.
-In this sense, time is not just like another state variable.\footnote{In relativity, for example, it is common to use foliations and the $3+1$ formalism to recover the special character of time when performing calculations or to gain better physical insight\cite{Gourgoulhon}.}
-
-As in the time independent case, we are interested in invariant densities which are defined on co-vectors.
-The co-vector $k_i dq^i$ is completed with a time component $\bar{\omega} dt$.\footnote{$\bar{\omega}$ is the classical analogue of the wave frequency.
-We use $\bar{\omega}$ to distinguish from the symplectic form $\omega$.} As each state variable $q^i$ has a conjugate quantity $p_i\equiv \hbar k$, the time variable $t$ will have a conjugate quantity $E\equiv\hbar\bar{\omega}$.
-We call the combination of $(t, E)$ the temporal degree of freedom.
-As time is special, so is the temporal degree of freedom, which is treated differently when defining a distribution $\rho_\mathcal{c}$ and therefore by the measure $\mu$ and the symplectic form $\omega$.
-
-Most of all, the temporal degree of freedom is not an independent degree of freedom because $E$ cannot be an independent variable.
-It cannot add any configurations, as those are defined by $T^*\mathcal{Q}$ only, and cannot add any time instants, as those are defined by $t$ alone.
-Therefore there must exist a constraint such that, locally, $E=E(t,q^i,p_i)$.
-The complete state space $\bar{\mathcal{S}}$ is a hypersurface of $T^*\mathcal{M}$.
-
-As the temporal degree of freedom $(t, E)$ is not independent of the other d.o.f.~$(q^i, p_i)$; it is not orthogonal to them in $T^*\mathcal{M}$.
-States are defined on the plane where $( q, p )$ (maximally) change.
-This is not the plane of constant $( t, E )$ (they are not orthogonal) where the area given by $dq \wedge dp$ is defined.
-It is the plane perpendicular to constant $( q, p )$.
-And the plane of constant $( q, p )$ is where the area given by $dt \wedge dE$ is defined.
-That is: the plane where we can properly count states and define our symplectic form $\omega$ is perpendicular to $dt \wedge dE$.
-Intuitively, states are defined independently of time, therefore they are defined on a surface perpendicular to the temporal d.o.f.
-
-We have a right triangle-like relationship between the plane where $\omega$ is defined and its projections on the planes defined by each d.o.f., similar to the multiple d.o.f.:
-\begin{align*}
-\textrm{multiple d.o.f.} \;\;\; &dq^1 \wedge dp_1 + dq^2 \wedge dp_2 = \omega \\
-\textrm{temporal d.o.f.} \;\;\; &dt \wedge dE + \omega = dq \wedge dp
-\end{align*}
-But in the previous case, the right angle was between the two independent d.o.f.
-In this case, the right angle is between $\omega$ and the plane of constant $(q, p)$ where $dt \wedge dE$ is defined.
-We rewrite it as $dq \wedge dp - dt \wedge dE = \omega$.
-This corresponds to the Minkowski product across d.o.f.
-and the vector product within.
-In terms of unified state variables $\xi^a\equiv \{t, q^i, E, p_i\}$ we have:
-
-\begin{align*}
-\omega_{ab} =  \left[
-\begin{array}{cc}
-0 & 1 \\
--1 & 0 \\
-\end{array}
-\right] \otimes \left[
-\begin{array}{cc}
--1 & 0 \\
-0 & I_n \\
-\end{array}
-\right] =
-\left[
-\begin{array}{cccc}
-0 & 0 & -1 & 0 \\
-0 & 0 & 0 & I_n \\
-1 & 0 & 0 & 0 \\
-0 & -I_n & 0& 0 \\
-\end{array}
-\right] \\
-\end{align*}
-
-\begin{figure}
-	\includestandalone[width=\columnwidth]{pictures/temporalDof}
-	\caption{The symplectic form $\omega$ and its components along temporal and standard degrees of freedom.
-The two d.o.f.
-are not independent and therefore not orthogonal.
-The temporal d.o.f.
-is orthogonal to the symplectic form as states are defined (and counted) independently of time.
-Compare with FIG.~\ref{fig:independent_dof}.
-Note: this is a 2D conceptualization of 4D space.}
-	\label{fig:temporal_dof}
-\end{figure}
-
-The symplectic form $\omega$ is still the canonical two-form expressed with the mathematically non-canonical, but physically meaningful, variable $E$.
-This allows us to better understand its physical meaning.
-$\omega$ allows us to count the possibilities within a degree of freedom, adjusting it on $T^*\mathcal{M}$ to avoid the counting problems introduced by time variable changes.
-
-The symplectic form $\omega$ also allows us to distinguish temporal from standard d.o.f.
-Given a two-dimensional surface $U \subset T^*\mathcal{M}$, how can we tell if it should be charted by a $(t,E)$ or whether we can define a distribution on it?
-Consider $\int_U \omega$.
-The result will be positive if the integration is over a standard degree of freedom, and it will be negative if the integration is over the temporal degree of freedom.
-If the contribution in every subregion of $U$ is positive, then the whole $U$ is always oriented along a standard degree of freedom.
-
-The invariance of $\omega$ also prevents mixing between standard and temporal d.o.f.
-The Euclidean signature across standard d.o.f.
-allows us to move ranges of possibilities from one d.o.f.
-to the other while conserving the total: the generators of rotation are in fact divergence free.
-The Minkowski signature, instead, only allows hyperbolic rotation (i.e.
-Lorentz boosts) across standard and temporal d.o.f., whose generators are curl free.
-Therefore, as required, we cannot transform a distribution over a standard d.o.f.
-into a distribution over a temporal one.
-
-We can capture the above discussion by stating that the complete state space $\bar{\mathcal{S}}$ is a hypersurface of the symplectic manifold $(T^*\mathcal{M}, \omega)$, where $\omega \equiv \sum dq^i \wedge dp_i - dt \wedge dE$.
-And if we set $q^\alpha = (t, q^i)$ and $p_\alpha = (-E, p_i)$, we can write $\omega \equiv \sum dq^\alpha \wedge dp_\alpha$.
-This is the manifold that allows us to define time-and-state-variable-invariant densities.
-The symplectic form $\omega$ allows us to measure the count of possibilities as in the time independent case.
-
-\begin{prop}\label{prop:relativistic_symplectic_manifold}
-	The complete state space $\bar{\mathcal{S}}$ for the particles of a classical material is a hypersurface of the symplectic manifold formed by a cotangent bundle $T^*\mathcal{M}$ equipped with the canonical two-form $\omega = \sum dq^i \wedge dp_i - dt \wedge dE = \sum dq^\alpha \wedge dp_\alpha$.
-\end{prop}
-
-\begin{justification}
-	We claim the space of physical objects that allows time-and-state-variable-invariant densities is a symplectic manifold $(T^*\mathcal{M}, \omega)$ where $\omega = \sum dq^i \wedge dp_i - dt \wedge dE$.
-Let $\mathcal{M}$ be the manifold that defines the unit system including time.
-Locally $\mathcal{M} \cong \mathcal{Q} \times \mathbb{R}$ as time and state variables are independent quantities.
-As in \ref{prop:symplectic_manifold}, invariant densities are defined on the symplectic manifold $(T^*\mathcal{M}, \omega)$.
-Let $\bar{\omega}$ be the co-vector component associated with $t$.
-Let $E\equiv\hbar \bar{\omega}$.
-Let $\omega$ be the symplectic form.
-The components of $\omega$ across different d.o.f.
-must be zero as they are not invariant under change of units.
-$\omega=\sum dq^i \wedge dp_i \pm dt \wedge dE$.
-The $+$ case is excluded as the temporal degree is not an independent d.o.f.: $\omega = \sum dq^i \wedge dp_i - dt \wedge dE$.
-
-	We claim $\bar{\mathcal{S}}$ is a hypersurface of $(T^*\mathcal{M}, \omega)$.
-$\bar{\mathcal{S}} \subseteq (T^*\mathcal{M}, \omega)$ as states are physical objects that allow time-and-state-variable-invariant densities.
-Consider the map $f : T^*\mathcal{M} \rightarrow \bar{\mathcal{S}} \; | \; (q^i, p_i, t, E) \mapsto (q^i, p_i, t)$.
-$f$ is an open map.
-The topology of $\bar{\mathcal{S}}$ is the subspace topology.
-Let $\rho : T^*\mathcal{M} \rightarrow \mathbb{R}$ be an invariant density.
-$\rho$ is differentiable by \ref{prop:differentiable_manifold}.
-Its restriction on $\bar{\mathcal{S}}$ is also an invariant distribution and therefore differentiable.
-The inclusion map $\bar{\mathcal{S}} \hookrightarrow T^*\mathcal{M}$ is a smooth embedding.
-$\bar{\mathcal{S}}$ is an embedded submanifold of co-dimension $1$.
-$\bar{\mathcal{S}}$ is a hypersurface.
-\end{justification}
-
-\subsection{Relativistic Hamiltonian mechanics}
-
-As we are using time as a variable to label states, we will use a different quantity as the parameter for the evolution.
-The trajectory of a particle in the extended phase space will be given by the evolved variables $\xi^a(s)$ in terms of a parameter $s$.
-As before, deterministic and reversible evolution will preserve $\omega$, as the possibility count on each independent d.o.f.
-will be conserved.
-Mathematically, deterministic and reversible evolution is a self-symplectomorphism.
-
-\begin{prop}\label{prop:relativistic_symplectomorphism}
-	A time-dependent deterministic and reversible evolution map for the particles of a classical material is a self-symplectomorphism on $T^*\mathcal{M}$.
-That is: $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ such that $\mathcal{T}_{\Delta s}^*\omega = \omega$ where $\mathcal{T}_{\Delta s}^*$ is the pullback of $\mathcal{T}_{\Delta s}$.
-\end{prop}
-
-\begin{justification}
-	We claim $\mathcal{T}_{\Delta s}$ is a self-homeomorphism on $T^*\mathcal{M}$.
-The complete state space for particles of a classical material is $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ by \ref{prop:relativistic_symplectic_manifold}.
-$\mathcal{T}_{\Delta s}$ is a deterministic and reversible evolution map and by \ref{prop:homeomorphism} is a self-homeomorphism.
-We can extend the map over $T^*\mathcal{M}$.
-
-	We claim $\mathcal{T}_{\Delta s}$ is a symplectomorphism on $(T^*\mathcal{M}, \omega)$.
-The distribution on final states must be defined.
-The Jacobian for $\mathcal{T}_{\Delta s}$ exists and is non-zero.
-$\mathcal{T}_{\Delta s}$ is a diffeomorphism.
-A deterministic and reversible process conserves the number of states and possibilities.
-$\omega$ is the two-form that returns the count of possibilities.
-$\omega$ is invariant under deterministic and reversible evolution.
-$\mathcal{T}_{\Delta s}$ is a symplectomorphism by definition.
-\end{justification}
-
-If we assume continuous time evolution we have the following:
-
-\begin{prop}\label{prop:relativistic_hamiltons_equations}
-	A continuous time-dependent deterministic and reversible process for the particles of a classical material admits an invariant Hamiltonian $\mathcal{H}: T^*\mathcal{M} \rightarrow \mathbb{R}$ that allows us to write the laws of evolution as
-	\begin{align*}
-	d_{s}t &= - \partial_{E} \mathcal{H} \\
-	d_{s}E &= \partial_{t} \mathcal{H} \\
-	d_{s}q^i &= \partial_{p_i} \mathcal{H} \\
-	d_{s}p_i &= - \partial_{q^i} \mathcal{H}
-	\end{align*}
-\end{prop}
-
-\begin{justification}
-We claim the vector field $S \in T(T^*\mathcal{M})$ for the infinitesimal displacement $S^a = d_s\xi^a$ admits a potential $\mathcal{H}$ such that $S^{a} \omega_{ab} = \partial_{b}\mathcal{H}$.
-The map for infinitesimal evolution $\mathcal{T}_{ds}$ is an infinitesimal self-symplectomorphism by \ref{prop:relativistic_symplectomorphism}.
-By \ref{symplectomorphism_generator} the infinitesimal displacement $S$ admits a potential $\mathcal{H}$ such that $S^{a} \omega_{ab} = \partial_{b}\mathcal{H}$
-
-We claim the state variables evolve according to the extended Hamilton equations.
-$S^{a} \omega_{ab} = d_s\xi^a \omega_{ab} = \partial_{b}\mathcal{H}$.
-For $a = 0$ we have $d_s t \, (-1) = \partial_{E} \mathcal{H}$.
-For  $a=\{1,...,n\}$ we have $d_s q^i \, (+1) = \partial_{p^i} \mathcal{H}$.
-For $a=n+1$ we have $d_s E \, (+1) = \partial_{t} \mathcal{H}$.
-For $a=\{n+2,...,2n + 1\}$ we have $d_s p_i \, (-1) = \partial_{q^i} \mathcal{H}$.
-\end{justification}
-
-We recognize the equations as those for Hamiltonian mechanics on the extended phase space\cite{Synge,Lanczos,Struckmeier}.
-
-The trajectory in time $t(s)$ is of particular importance.
-Since the evolution is deterministic and reversible, for each value of $s$ we need to have one and only one value of $t$.
-Therefore $t(s)$ is invertible, strictly monotonic and $d_{s}t$ along a trajectory cannot change sign.
-This means there are two classes of states: the ones where the parametrization $s$ is aligned with time $t$, which we call standard states, and the ones where the parametrization $s$ is anti-aligned with time $t$, which we call anti-states.
-Let us call $\lambda : T^*\mathcal{M} \rightarrow \mathbb{R}$ the function $\lambda (\xi^a) \mapsto d_s t |_{\xi^a}$ that returns the change of $t$ along $s$.
-$\lambda > 0$ for all standard states while $\lambda < 0$ for all anti-states.
-Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call standard and anti-states is also conventional.
-What is physical and not conventional, though, is that standard and anti-states cannot be connected by deterministic and reversible evolution.\footnote{This represents the classical analogue for quantum anti-particle states.}
-
-\begin{prop}\label{prop:antistates}
-	Let $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ be the time dependent deterministic and reversible evolution map for the particles of a classical material.
-The map partitions the extended state space into standard states, those connected by a trajectory where $d_{s}t>0$, and anti-states, those connected by a trajectory where $d_{s}t<0$.
-\end{prop}
-
-\begin{justification}
-	We claim $t(s)$ is strictly monotonic.
-The motion is deterministic and reversible.
-At each value of $t$ there must be only one possible state: if there is more than one state the motion is non-deterministic (one state is not enough to identify more than one state), if there is no state the motion is non-reversible (no state cannot identify a state).
-$t(s)$ is invertible.
-$t(s)$ is strictly monotonic.
-
-	We claim $d_{s}t$ partitions the space.
-Let $\mathcal{s}_1, \mathcal{s}_2 \in T^*\mathcal{M}$ be two physical states connected by deterministic and reversible evolution.
-$\mathrm{sign}(d_{s}t(\mathcal{s}_1)) = \mathrm{sign}(d_{s}t(\mathcal{s}_2))$.
-Let $U \equiv \{\mathcal{s} \in T^*\mathcal{M} \; | \; d_{s}t(\mathcal{s}) > 0 \}$ and $V \equiv \{\mathcal{s} \in T^*\mathcal{M} \; | \; d_{s}t(\mathcal{s}) < 0 \}$.
-$U \cap V = \varnothing$.
-Let $\gamma : [s_0, s_1] \rightarrow T^*\mathcal{M}$ the trajectory given by deterministic and reversible evolution, where $[s_0, s_1]$ is the range of the parametrization.
-Either $\gamma([s_0, s_1]) \subseteq U$ or $\gamma([s_0, s_1]) \subseteq V$.
-\end{justification}
-
-With the invariant Hamiltonian we are able to write the equations of motion for any choice of time and state variables.
-For example: $d_t q^i = d_t s \, d_s q^i = d_s q^i / d_s t = - \partial_{p_i} \mathcal{H} / \partial_{E} \mathcal{H}$.
-These equations must be the same as the ones given by a standard Hamiltonian in those coordinates.
-That is: $d_t q^i = \partial_{p_i} H = - \partial_{p_i} \mathcal{H} / \partial_{E} \mathcal{H}$.
-The partial derivatives of $H$ and $\mathcal{H}$ are related, and therefore the functions themselves are related.
-Working through the math, we find that the most general relationship is of the form: $\mathcal{H} = \lambda(t,q^i,E,p_i) \, (H(t,q^i,p_i) - E - \mathcal{h}(t))$ where $\lambda = d_s t$, $\mathcal{h}(t)$ is an arbitrary function and, most importantly, $E = H(t,q^i,p_i) + \mathcal{h}(t)$.
-
-The arbitrary function $\mathcal{h}(t)$ has no physical consequence: it is just a constant added to the Hamiltonian.
-We can either set it to zero or simply redefine the Hamiltonian to include it.
-Therefore we are left with $E = H(t,q^i,p_i)$.
-In other words: $E$ is the value of the Hamiltonian, the energy of the system.
-We have found the constraint that identifies the hypersuface of $T^*\mathcal{M}$ introduced in \ref{prop:relativistic_symplectic_manifold}.
-
-Note that $\mathcal{H}=0$ for all states $\mathcal{s} \in \bar{\mathcal{S}} \subset T^* \mathcal{M}$, since $\mathcal{H}=\lambda(H-E)$ and $E = H$.
-For all states we also have $\lambda \neq 0$.
-Therefore $\mathcal{H}=0$ only because $E = H$.
-We can extend $\mathcal{H}$ on all of $T^* \mathcal{M}$ such that $\mathcal{H} \neq 0$ outside of the complete state space $\bar{\mathcal{S}}$.
-This way we can directly use the constraint $\mathcal{H}=0$ to identify $\bar{\mathcal{S}}$ as a subspace of $T^* \mathcal{M}$.
-
-This way $\mathcal{H}$ gives both the constraint $\mathcal{H}=0$ to identify states and the map $d_s \xi^a=\omega^{ab} \partial_b \mathcal{H}$ to identify their evolution.
-
-\begin{prop}\label{prop:form_of_invariant_hamiltonian}
-	The form of the invariant Hamiltonian is given by $\mathcal{H} = \lambda(t,q^i,E,p_i) \,  (H(t,q^i,p_i) - E)$ where $H(t,q^i,p_i)$ is the standard Hamiltonian at each instant $t$, the conjugate time variable $E = H(t,q^i,p_i)$ represents the value of the Hamiltonian, and  $\lambda : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ where $\lambda(\xi^a) = d_s t$.
-$\bar{\mathcal{S}}$ is the level set for $\mathcal{H} = 0$.
-\end{prop}
-
-\begin{justification}
-	We claim the form of the extended Hamiltonian is $\mathcal{H} = \lambda(t,E,q^i,p_i) \, (H(t,q^i,p_i) - E)$.
-Let $\mathcal{H}$ be the invariant Hamiltonian and $H$ be the Hamiltonian for a particular choice of time and state variables.
-We have $d_s q^i = \partial_{p_i} \mathcal{H} = d_s t \, d_t q^i = - \lambda \, \partial_{p_i} H$ where $\lambda : T^*\mathcal{M} \rightarrow \mathbb{R}$ such that $\lambda \equiv d_s t$.
-$d_s p_i = - \partial_{q^i} \mathcal{H} = d_s t \, d_t p_i = \lambda \, \partial_{q^i} H$.
-As $\lambda \neq 0$ for all physical states, we can set $\mathcal{H} = \lambda(t,q^i,E,p_i) \, (H(t,q^i,p_i) - E + \mathcal{h}(t,q^i,E,p_i))$ without loss of generality.
-We have $\partial_{q^i} \lambda \, (H - E + \mathcal{h}) + \lambda \, (\partial_{q^i} H + \partial_{q^i} \mathcal{h}) = \lambda \, \partial_{q^i} H$.
-This holds for all choices of $t$.
-In particular for $t=s$, $\lambda \, (\partial_{q^i} H + \partial_{q^i} \mathcal{h}) = \lambda \, \partial_{q^i} H$.
-$\partial_{q^i} \mathcal{h} = 0$.
-$\mathcal{h}$ is not a function of $q^i$.
-Repeat the same procedure for $p_i$.
-$\mathcal{h}$ is not a function of $p_i$.
-$\lambda = d_s t = - \partial_E \mathcal{H} = - \partial_{E} \lambda (H - E + \mathcal{h}) - \lambda \, (-1 + \partial_{E} \mathcal{h})$.
-In particular for $t=s$, $\lambda = - \lambda (-1 + \partial_{E} \mathcal{h})$.
-$\partial_{E} \mathcal{h} = 0$.
-$\mathcal{h}$ is not a function of $E$.
-$\mathcal{h}$ is only a function of $t$.
-We can redefine $H = H + \mathcal{h}$ as the result is still a function of $q^i, p_i , t$ with the same equations of motion.
-$\mathcal{H}=\lambda (H - E)$.
-
-	We claim $E = H(t,q^i,p_i)$ for physical states.
-$\lambda = d_s t = - \partial_E \mathcal{H} = - \partial_{E} \lambda (H - E) - \lambda (-1)$ regardless of the choice of t and therefore for all $\lambda$.
-$H - E = 0$.
-
-	We claim $\bar{\mathcal{S}}$ is the level set for $\mathcal{H} = 0$.
-$\mathcal{H} = \lambda (H - E) = 0$ over $\bar{\mathcal{S}}$.
-We can extend $\mathcal{H}$ over $T^* \mathcal{M}$ such that $\mathcal{H} \neq 0$ over $T^*\mathcal{M} \,\backslash\, \bar{\mathcal{S}}$.
-$\bar{\mathcal{S}}$ is the level set for $\mathcal{H} = 0$.
-\end{justification}
-
-As an example, the Hamiltonian and invariant Hamiltonian for a relativistic free particle are:
-\begin{equation}\label{free_hamiltonians}
-\begin{aligned}
-H &= c \sqrt{m^2 c^2 + p^i p_i} \\
-\mathcal{H} &= \frac{1}{2m} ( p^i p_i - (E/c)^2 + m^2c^2) \\
- &= \frac{1}{2mc^2} (c \sqrt{m^2 c^2 + p^i p_i} + E) (c \sqrt{m^2 c^2 + p^i p_i} - E) \\
-\lambda &= \frac{1}{2mc^2} (c \sqrt{m^2 c^2 + p^i p_i} + E)
-\end{aligned}
-\end{equation}
-Therefore we can see that the invariant Hamiltonian has the form $\mathcal{H}=\lambda (H - E)$.
-In the next section we'll derive the above expression and look more closely at the case of particles under potential forces.
-
-Now that we have characterized the motion of the particles, we should also characterize the evolution of composite states.
-
-Suppose we have a distribution of material $\rho_{\bar{\mathcal{c}}}(s)$ associated with an evolution $\bar{\mathcal{c}}$.
-Its support, the region where the density is non-zero, is within $\bar{\mathcal{S}}$.
-Therefore  $\mathcal{H} \rho_{\bar{\mathcal{c}}}(s) = 0$ since $\mathcal{H} = 0$ in $\bar{\mathcal{S}}$ where $\rho_{\bar{\mathcal{c}}}(s) \neq 0$.
-As the density is transported over each trajectory, its value does not change over $s$.
-Therefore we have $d_s \rho_{\bar{\mathcal{c}}}= d_s \xi^a \partial_a \rho_{\bar{\mathcal{c}}} = \omega^{ab} \partial_b \mathcal{H} \partial_a \rho_{\bar{\mathcal{c}}} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \} = 0$ where $\{f , g\} = \partial_a f \omega^{ab} \partial_b g$ is the Poisson bracket.
-
-\begin{prop}\label{prop:relativistic_state_evolution}
-	Let $\rho_{\bar{\mathcal{c}}}$ be the distribution of material associated with the evolution $\bar{\mathcal{c}}$.
-$\mathcal{H} \rho_{\bar{\mathcal{c}}} = 0$ and $d_s \rho_{\bar{\mathcal{c}}} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \} = 0$.
-\end{prop}
-
-\begin{justification}
-	We claim $\mathcal{H} \rho_{\bar{\mathcal{c}}} = 0 \; \forall \bar{\mathcal{c}} \in \bar{\mathcal{C}}$.
-Let $\rho_{\bar{\mathcal{c}}} : T^*\mathcal{M} \rightarrow \mathbb{R}$ be the distribution associated with an evolution $\bar{\mathcal{c}}$.
-$supp(\rho_{\bar{\mathcal{c}}}) \subseteq \bar{\mathcal{S}}$ as the distribution is defined on the complete state space of the particles.
-$supp(\mathcal{H}) = T^*\mathcal{M} \,\backslash\, \bar{\mathcal{S}}$ by \ref{prop:form_of_invariant_hamiltonian}.
-$supp(\mathcal{H} \rho_\mathcal{c}) = supp(\mathcal{H}) \cap supp(\rho_\mathcal{c}) = \varnothing$.
-
-	We claim $d_s \rho_{\bar{\mathcal{c}}} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \} = 0$.
-$d_s \rho_{\bar{\mathcal{c}}}= \partial_a \rho_{\bar{\mathcal{c}}} d_s \xi^a = \partial_a \rho_{\bar{\mathcal{c}}} \omega^{ab} \partial_b \mathcal{H} = \{ \rho_{\bar{\mathcal{c}}}, \mathcal{H} \}$.
-Let $\gamma : \mathbb{R} \rightarrow T^*\mathcal{M}$ be a parametrization for a particle evolution $\bar{\mathcal{s}}$.
-Then $\rho_{\bar{\mathcal{c}}}(\lambda(s)) = \rho_{\bar{\mathcal{c}}}(\lambda(s + \Delta s))$ as the material does not change over the particle evolution.
-$d_s \rho_{\bar{\mathcal{c}}} = 0$.
-\end{justification}
-
-As we have derived relativistic motion in an unusual way, a few comments are in order.
-
-As mentioned at the beginning of the section, no new assumption was required.
-Relativistic Hamiltonian mechanics is simply the correct way of handling time dependent motion.
-As such, there was no mention of the speed of light $c$, $\mathcal{M}$ being a pseudo-Riemannian manifold or its metric $g$.
-These will be introduced in the next section with the kinematic equivalence assumption.
-In light of this, it is probably better to consider relativistic mechanics as a feature of the geometry of the complete state space more than the geometry of space-time.
-Also note that this feature arises from the use of the standard topology on $\mathbb{R}^n$ and the need for invariant densities.
-It would not have arisen with the use of a discrete topology.
-
-We have also seen that anti-particle states, generally associated with quantum field theory, are actually a feature of relativistic Hamiltonian mechanics.
-This is usually missed because in classical mechanics only the standard Hamiltonian is typically used.
-Note how the invariant Hamiltonian in \ref{free_hamiltonians} is quadratic in $E$, which can have both positive and negative solutions given the constraint $\mathcal{H}=0$.
-
-Another important feature is that this formulation of relativistic Hamiltonian mechanics is formally very close to the quantum case.
-Note how $\mathcal{H} \rho = 0$ and $\mathcal{H} \ket{\psi} = 0$ are formally equivalent.
-If we apply to the invariant Hamiltonian in \ref{free_hamiltonians} the usual formal classical/quantum substitution, we have the Klein-Gordon equation using space-like convention rescaled by a factor of $2m$.
-As we'll see later, to this parallel in the mathematical formalism corresponds a parallel in the physical description.
-
-As a final thought, we note how the physics maps less elegantly to the math in the time dependent case.
-The function space for $\bar{\mathcal{C}}$ is not simply the Lebesgue integrable differentiable functions, but those that are integrable on particular hypersurfaces.
-The complete state space $\bar{\mathcal{S}}$ is a hypersurface of $T^*\mathcal{M}$.
-The invariant Hamiltonian is needed to identify that subspace.
-This may be a hint that a better formulation is lurking, one where we consider the particle evolutions $\bar{\mathcal{s}}$ as the primary objects and label those (instead of the states).
-The distribution of material would simply be a distribution over the particle evolutions, a Lebesgue integrable differentiable function.
-Yet, the need to be able to describe the evolution in terms of physically meaningful state variables will most likely reimpose the extra structure.
-
-\section{Kinematics and minimal materials}
-\label{sec:Lagrangian}
-
-We will now focus our attention on the simplest possible classical material: one that is fully described by its motion in space, i.e.~the internal dynamics can be ignored.
-What we'll find is that the dynamics of its infinitesimal parts follows classical Lagrangian mechanics.
-Furthermore, the Hamiltonian of the system is constrained to the one of a free particle under scalar and vector potential forces.
-
-\subsection{Kinematic equivalence}
-
-The link between kinematics, the study of motion through quantities like velocity and acceleration, and dynamics, the study of state evolution and its causes, is crucial in physics.
-In Newtonian physics, $\vec{F}=m\vec{a}$ provides such a link.
-In Hamiltonian (and Lagrangian) mechanics, the link is implicitly given by the choice of the Hamiltonian (or Lagrangian).
-
-Our aim is to make that link more conceptually crisp and explicit with the following assumption:
-
-\begin{assump}[Kinematic equivalence]\label{ass:kinematic_equivalence}
-	For the system under investigation, studying its kinematics  (i.e. trajectories in space) is equivalent to studying its dynamics  (i.e. states and their evolution).
-\end{assump}
-
-\begin{rationale}
-	The idea is that the system internal dynamics has no effect on the overall motion and therefore can be disregarded.
-Consider the motion of our cannonball under gravitational and inertial forces: its temperature or precise chemical composition does not affect the overall trajectory.
-We can assume kinematic equivalence.
-Suppose we now add magnetic forces to our consideration.
-The magnetization of the cannonball may now affect the motion and therefore we cannot always reconstruct the full state by simply looking at the trajectory.
-We cannot assume kinematic equivalence.
-
-	Kinematic equivalence can't therefore be assumed for all systems.
-Yet, the assumption is fundamental because the particles of any material, during their evolution, will \emph{at least} describe a trajectory in space and what we learn about the link between kinematic and dynamical quantities applies to all materials, even the ones where the kinematic assumption does not hold.
-
-	The first natural question to ask is: why does state evolution of a particle of a classical material always describe a trajectory in space?
-It's because spatial extent is the one way in which all materials can be decomposed.
-As we divide materials into smaller pieces, physical location is one state variable we can assign to the particles of all materials.
-If we assume kinematic equivalence, the kinematic variables will be enough to identify a particle state.
-In such a case, we will call the material \emph{minimal} to highlight the idea that the state is described by the smallest possible number of d.o.f.
-
-	When the system is not minimal, the state of the particles will be described by some additional independent state variables.
-The overall dynamics will be the spatial d.o.f.
-coupled with the internal d.o.f.
-Therefore by studying a minimal material, we are also studying a component of the dynamics that is present in all materials.
-In other words: any material is an extension of a minimal material.
-
-	It should also be noted that, once we reduce particles to be infinitesimal at a point, the rest of the description must be a local object invariant under coordinate transformations.
-This means that the remaining internal state must be identified by scalar, vector or tensor quantities.
-This may be used in the future to justify the extension of this work to field theories.
-\end{rationale}
-
-\subsection{Space-time manifold and Lagrangian mechanics}
-
-As we need to describe trajectories, we introduce $\mathcal{M}$ as the $n+1$ dimensional manifold that defines physical space and time.
-A point in $\mathcal{M}$ will be identified by $x^\alpha = (t, x^i)$, its time and space coordinates.
-A trajectory will be given by a set of functions $x^\alpha(s)$ where $s$ is an arbitrary parametrization of the curve.
-What we need to understand is the link between the complete state space of the particles of a classical material and the space of all the possible trajectories under the kinematic assumption.
-
-We first note that the system of units for the trajectories is fully defined by $\mathcal{M}$.
-Any trajectory can be specified once a set of coordinates $x^\alpha$ is chosen.
-Under the kinematic assumption, these units must also be sufficient to describe the dynamics.
-Therefore $\mathcal{M}$ has conceptually two roles: it is the space where the motion unfolds and it is the space on which we define the system of units for our states.
-This already tells us that the state space for the particles of the material is a hypersurface of $T^*\mathcal{M}$.
-
-Under kinematic equivalence, given a trajectory we must identify a state and vice-versa.
-There must be a homeomorphism between the complete state space and the space of possible trajectories.
-We already know what the complete state space is.
-We just need to find a way to label the trajectories so that we can express the correspondence.
-
-Because of the double role of $\mathcal{M}$, we can set half of our state variables to be equal to the space-time variables: $q^\alpha = (t, q^i) = x^\alpha$.
-This already tells us that the trajectory is differentiable since $q^\alpha(s)$ is differentiable.
-The four-velocity $u^\alpha = d_s x^\alpha$ is therefore well defined and it can be used to identify trajectories.
-The magnitude of the four-velocity can be rescaled by changing the evolution parameter, $u^\alpha = d_s x^\alpha = d_s \hat{s} \, d_{\hat{s}} x^\alpha = d_s \hat{s} \, \hat{u}^\alpha$, and is therefore not physical.
-Only the direction is physical so the four-velocity gives us another $n$ independent variables to identify trajectories.
-The combined position and velocity, the kinematic variables, identify $2n + 1$ trajectories using a hypersurface of the tangent bundle $T\mathcal{M}$.
-This is exactly the dimensionality of the complete state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$.
-We cannot add more curves, as a homeomorphism preserves dimensionality, therefore we identified the suitable space for the trajectories of particles of a classical material under kinematic equivalence.
-
-Being able to identify particles through either state or kinematic variables is not enough: we also need to be able to express a composite state as the distribution of particle states over position and velocity.
-As such, the differentiable distribution over state variables must become a differentiable distribution over kinematic variables.
-This tells us the transformation is differentiable.
-
-Mathematically, the state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ is diffeomorphic to a hypersurface of the tangent bundle $T\mathcal{M}$ of the space-time manifold.
-
-\begin{prop}\label{prop:tangent_bundle}
-	Let $\mathcal{M}$ be the $n+1$ dimensional space-time manifold over which trajectories are defined.
-The complete state space $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ for the particles of a minimal classical material is diffeomorphic to a hypersurface of the tangent bundle $T\mathcal{M}$.
-\end{prop}
-
-\begin{justification}
-	We claim the complete state space is a hypersurface of the cotangent bundle of the space-time manifold, $\bar{\mathcal{S}} \subset T^*\mathcal{M}$.
-Let $\bar{\mathcal{s}} \in \bar{\mathcal{S}}$ be the evolution of a particle for a minimal classical material.
-Let $f(\bar{\mathcal{s}}) \mapsto x^\alpha(s)$ be the function that returns the space-time trajectory associated with the evolution.
-$f$ is injective as the material is minimal.
-$\bar{\mathcal{s}}$ can be identified by a trajectory $x^\alpha(s)$ which can be fully described by coordinates defined on $\mathcal{M}$.
-Let $\mathcal{s} \in \bar{\mathcal{S}}$ be the particle state for a minimal classical material.
-Let $f(\bar{\mathcal{s}}, t) \mapsto \mathcal{s}$ the function that returns the state at a particular time $t$ along an evolution $\bar{\mathcal{s}}$.
-Such function is a bijection as the evolution is deterministic and reversible.
-$\mathcal{s}$ can be identified by an evolution $\bar{\mathcal{s}}$ and an instant $t$, both fully described by coordinates defined on $\mathcal{M}$.
-$\mathcal{M}$ defines the unit system.
-$\bar{\mathcal{S}} \subset T^*\mathcal{M}$ by \ref{prop:relativistic_symplectic_manifold}.
-
-	We claim there exists a homeomorphism between $\bar{\mathcal{S}} \subset T^*\mathcal{M}$ and a hypersurface of $T\mathcal{M}$.
-Let $\mathcal{s} \in \bar{\mathcal{S}}$ be the particle state for a minimal classical material.
-Let $x^\alpha(s)$ be the evolution associated with that state.
-Let $t(\mathcal{s})$ be the instant of time at which $\mathcal{s}$ is defined.
-$q^\alpha(\mathcal{s}) \equiv x^\alpha(s(t(\mathcal{s})))$ are state variables as they are real function of the state.
-$q^\alpha$ is differentiable therefore $d_s x^\alpha$ exists.
-Suppose $d_s x^\alpha$ is not part of the state.
-Then $d_s x^\alpha = d_s q^\alpha$ must be determined by the law of evolution.
-$x^\alpha(s)$ is determined by the law of evolution therefore $q^\alpha$ are the only state variables needed from the trajectory.
-As the material is minimal, $q^\alpha$ determines the full state $\mathcal{s}$.
-This is a contradiction: $\mathcal{s}$ is not identified by just $q^\alpha$.
-Therefore $u^\alpha=d_s x^\alpha$ is part of the state.
-$p_\alpha = p_\alpha(x^\alpha, u^\alpha)$.
-The map $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is continuous by \ref{prop:continuity} and is bijective as the material is minimal.
-It is a homemorphism between $T^*\mathcal{M}$ and $T\mathcal{M}$.
-$\bar{\mathcal{S}}$ is a hypersurface of $T^*\mathcal{M}$ therefore its image is a hypersurface of $T\mathcal{M}$.
-
-	We claim $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is a diffeomorphism.
-Let $\bar{\mathcal{c}}$ be the evolution of a minimal classical material.
-Let $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbb{R}$ be the distribution associated with $\bar{\mathcal{c}}$.
-$\rho_{\bar{\mathcal{c}}}(\mathcal{s}(x^\alpha, u^\alpha))$ as the material is minimal.
-$\rho_{\bar{\mathcal{c}}}(x^\alpha, u^\alpha)$ is differentiable by the same logic expressed in \ref{prop:differentiable_manifold}.
-$(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is a diffeomorphism as it is a homeomorphism that maps a differentiable function to a differentiable function.
-\end{justification}
-
-This relationship allows us to write the kinematic variables in terms of the state variables.
-In particular we have $u^\alpha=u^\alpha(q^\alpha, p_\alpha)$.
-As the relationship is invertible, these functions must be monotonic in $p_\alpha$.
-Since $\partial_{p_\alpha} \mathcal{H}=d_s q^\alpha= u^\alpha$, $\mathcal{H}$ is convex in $p_\alpha$.
-Under such conditions we can define a Legendre transformation $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$.
-Such transformation is the Lagrangian of the system and the motion is given by the Euler-Lagrange equation.
-For the time independent case, we define $L=v^i p_i - H$ where $v^i=d_t x^i$.
-
-\begin{prop}\label{prop:lagrangian}
-	Let $\mathcal{H}$ be an invariant Hamiltonian defined on $T^*\mathcal{M}$.
-Under the kinematic assumption, we can define the Legendre transform $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$ which we call the \emph{extended Lagrangian}.
-\end{prop}
-
-\begin{justification}
-	We claim $\mathcal{H}$ is convex in $p_\alpha$.
-By \ref{prop:tangent_bundle} we can write the velocity $u^\alpha=u^\alpha(q^\alpha, p_\alpha)$ as a differentiable function of the state variables.
-At fixed $q^\alpha$, $u^\alpha$ must be invertible.
-The components $u^\alpha$ are monotonic in $p_\alpha$.
-$\partial_{p_\alpha} \mathcal{H}=d_s q^\alpha= u^\alpha$.
-The derivatives of $\mathcal{H}$ in $p_\alpha$ are monotonic in $p_\alpha$.
-$\mathcal{H}$ is convex in $p_\alpha$.
-
-	We claim there exists an extended Lagrangian $\mathcal{L}=u^\alpha p_\alpha - \mathcal{H}$.
-$\mathcal{H}$ is convex in $p_\alpha$.
-It admits a Legendre transform.
-The extended Lagrangian $\mathcal{L}$ is such transform.
-\end{justification}
-
-Not only must we be able to express distributions in terms of both state and kinematic variables, but integration has to carry over as well.
-To do that, we need to be able to count trajectories in the same way we are able to count states.
-Just as the symplectic form $\omega$ gives us the count of possibilities for a d.o.f.
-in terms of $dq^\alpha$ and $dp_\alpha$, the ranges of the state variables, there should be a bilinear form $g$ that gives the the count of possibilities in terms of $dx^\alpha$ and $du^\alpha$, the ranges of the kinematic variables.
-Under the kinematic assumption, these counts must be consistent with each other.
-That is: $\omega(dq^\alpha,dp_\alpha) \propto g(dx^\alpha,du^\alpha)$.
-
-Since both $dx^\alpha$ and $du^\alpha$ are four vector components, $g$ is a rank two tensor on $\mathcal{M}$.
-To each direction in space-time $dx^\alpha$ is associated a direction in the tangent space of velocities $du^\alpha$ with which it forms a d.o.f.
-Across d.o.f., that is $\alpha \neq \beta$, we have $\omega(dq^\alpha, dp_\beta) = 0 =\omega(dq^\beta, dp_\alpha)$.
-Therefore if $dx^\alpha$ and $du^\beta$ do not form a d.o.f., $g(dx^\alpha, du^\beta) = 0 = g(dx^\beta, du^\alpha)$: $g$ is reflexive.
-We also know that $g$ cannot be alternating, that is $g(dx^\alpha, du^\alpha) \neq 0$: a choice of units for $x^\alpha$ is a choice of units for $u^\alpha$ and therefore they cannot belong to different d.o.f.
-This means $g$ is a symmetric non-degenerate tensor.
-We recognize it as the metric tensor and therefore $\mathcal{M}$ is a pseudo-Riemannian manifold.
-
-Since $g$ is symmetric, it can be diagonalized: $\omega(dq^\alpha, dp_\beta)= dq^i \omega_{q^ip_i} dp_i + dt \omega_{tE} dE \propto g(dx^\alpha, du^\beta) = dx^i g_{ii} du^i + dt g_{00} du^0$.
-As $\omega_{q^ip_i} > 0$ we must have $g_{ii} > 0$.
-As $\omega_{tE} < 0$ we must have $g_{00} < 0$.
-We can further change the units along the spatial directions so that $g_{ii}=1$: if we use the same units for the spatial coordinates, each direction will contribute the same number of possibilities.
-Since time must use a different unit, we set $g_{00}=-c^2$ where $c^2$ is the constant that allows one to convert the possibility count over the temporal d.o.f.
-into the possibility count of a spatial d.o.f.
-Therefore, locally $g(dx^i, dx^i) = (dx^i)^2 - c^2 dt^2$ is Minkowskian.
-
-\begin{prop}\label{prop:riemannian_manifold}
-	The space-time manifold $\mathcal{M}$ is a pseudo-Riemannian manifold $(\mathcal{M}, g)$. $g$ has signature $(n,1)$ and locally is $g = (dx^i)^2 - c^2 dt^2$.
-\end{prop}
-
-\begin{justification}
-	We claim $(\mathcal{M}, g)$ is a pseudo-Riemannian manifold where $g$ is a metric tensor induced by $\omega$ under kinematic equivalence.
-Let $(dx^\alpha, du^\alpha)$ be an infinitesimal range of possibilities for the kinematic variables of the particles of a minimal classical material.
-There exists a map $g : T\mathcal{M} \times T\mathcal{M} \rightarrow \mathbb{R}$ such that $g(dx^\alpha, du^\alpha) \propto \omega(dq^\alpha, dp_\alpha)$.
-The map $g$ is linear, since a linear combination of infinitesimal ranges leads to the linear combination of possibilities.
-$g$ is a rank two tensor.
-Suppose $g(dx^\alpha, du^\beta) = 0$.
-$g(dx^\alpha, du^\beta) \propto \omega(dq^\alpha, dp_\beta) = 0 = \omega(dq^\beta, dp_\alpha) \propto g(dx^\beta, du^\alpha) = 0$.
-$g$ is reflexive.
-Position and velocity along the same direction use related units.
-They cannot form independent d.o.f.
-$g(dx^\alpha, du^\alpha) \neq 0$.
-$g$ is not alternating.
-$g$ is symmetric.
-$g$ is a symmetric rank two tensor.
-$(\mathcal{M}, g)$ is a pseudo-Riemannian manifold.
-
-	We claim $g$ is locally Minkowskian.
-Let $x^\alpha$ be a choice of coordinate that diagonalizes $g$.
-$x^\alpha$ exists as $g$ is symmetric.
-$\omega(dq^\alpha, dp_\beta) = dq^i \omega_{q^ip_i} dp_i + dt \omega_{tE} dE \propto g(dx^\alpha, du^\beta) \propto dq^i g_{ii} du^i + dt g_{00} du^0$.
-$dq^\alpha = dx^\alpha$.
-We require $g(dx^\alpha, du^\beta)>0$ when the arguments are positive.
-Therefore $g_{ii} > 0$ and $g_{00} < 0$.
-$g$ has signature $(n, 1)$.
-Choosing the same unit for all $dx^i$ and a temporal unit for $dt$ we can choose coordinates such that locally $g = (dx^i)^2 - c^2 dt^2$.
-$g$ is locally Minkowskian.
-\end{justification}
-
-The metric tensor $g$ and the constant $c$ found here are indeed the ones used in the context of special relativity.
-Yet, they have acquired a different meaning.
-$g$ does not measure space-time distances: it counts possibilities of a d.o.f., like $\omega$.
-Directions in space are orthogonal not because the angle between them is $90^\circ$, but because they identify independent d.o.f.
-That is, the choice of units along one direction is independent from the choice of units along orthogonal directions.
-Space and time, instead, are not independent d.o.f.
-and the geometry between them is different.
-In other words: $g$ inherits its signature from $\omega$.
-The geometry of space-time is the geometry of the ranges for possible states.
-
-This also allows us to give $c$ a more insightful physical meaning.
-It is not a velocity: the metric is defined at each point without motion.
-It is the constant of proportionality that maps a range of initial conditions from the time variable $t$ to the space variable $x^i$.
-For example, once the SI is adopted, $1$km would correspond to a thousand times more possible states than $1$m, and, in the same way, $1$sec would correspond to $299,792,458$ more possible states than $1$m.
-Therefore when we write $x^\alpha = (ct, x^i)$ or $p_\alpha = (- E/c, p_i)$ we are really expressing everything in quantities that are proportional to the count of states and possibilities.
-These coordinates make the bookkeeping of number of states much easier which, as we saw, is the only thing that Hamiltonian and Lagrangian mechanics are doing.
-
-This also allows us to understand heuristically why the speed of deterministic and reversible motion cannot be greater than $c$.
-Consider the trajectory of a point particle.
-Over an interval $dt$ it will travel $dx$ in space.
-Since the motion is deterministic and reversible, and therefore continuous, it will travel through all the states defined in the interval $dx$.
-It will also travel through every moment in time, which corresponds to $cdt$ possibilities.
-As it moves, time will always change but not necessarily space.
-The number of states traveled in space cannot exceed the number of states traveled in time.
-Therefore we have:
-\begin{align*}
-\frac{|n_x(dx)|}{|n_t(dt)|} &=  \frac{|dx|}{|c dt|} \leq 1 \\
-\frac{|dx|}{|dt|} & \leq c \\
-\end{align*}
-Note, though, that this constraint applies only to deterministic and reversible motion.
-It was under that premise that we were able to show that the trajectories are continuous, that we could define $\omega$ and therefore $g$.
-Non-local quantum effects, such as the ones featured in EPR\cite{EPR} experiments, are not examples of deterministic and reversible evolution.
-Therefore they do not have anything to violate: the maximum speed does not apply to their ``motion."\footnote{The very notion of velocity becomes ill defined in such cases precisely because we do not have a unique well-defined evolution.} On the other hand, only deterministic and reversible motion can carry information and can establish before/after relationships, therefore everything remains consistent.
-
-\subsection{Massive particles under scalar and vector potential forces}
-
-The link between $\omega$ and $g$ allows us to constrain the possible diffeomorphisms between state variables and kinematic variables.
-If we express the possibilities in terms of both state variables and kinematic variables we have: $dq^\alpha \wedge dp_\alpha \propto dx^\alpha du_\alpha$.
-Since  $p_\alpha=p_\alpha(x, u)$, we have $dp_\alpha = m du_\alpha$, where $m$ is the constant of proportionality.
-We recognize $m$ as the inertial mass.
-Integrating the expression, we have $p_\alpha = m u_\alpha + A_\alpha(x)$ where $A_\alpha(x)$ is an arbitrary vector field.\footnote{Note that $A_\alpha(x)$ corresponds to the product the vector potential and the charge.
-This simplifies the notation and avoids having to find a symbol for the charge itself.} This is the most general link between velocity and conjugate momentum, which is the sum of kinetic momentum and a scalar potential.
-
-This constrains the form of the extended Hamiltonian.
-$\frac{1}{m}g^{\alpha\beta}(p_\beta-A_\beta) = u^\alpha = d_s x^\alpha = d_s q^\alpha = \partial_{p_\alpha} \mathcal{H}$.
-Integrating we have $\mathcal{H}=\frac{1}{2m}(p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta)+V(x)$ where $V(x)$ is an arbitrary function.
-We can further constrain the Hamiltonian by choosing the natural parameterization for $s$ (also called unit speed \cite{Lee} or arc length parametrization) for which the modulus of the velocity is constant.
-We have $u^\alpha u_\alpha = - c^2 = \frac{1}{m^2}(p^\alpha-A^\alpha)(p_\alpha-A_\alpha)$.
-$\mathcal{H}=\frac{1}{2m}((p^\alpha-A^\alpha)(p_\alpha-A_\alpha) + m^2 c^2)$.
-Applying the same procedure in the time independent case yields the Hamiltonian $H=\frac{1}{2m}(p^i-A^i)(p_i-A_i)+V$.
-
-We recognize the Hamiltonian for massive particles under scalar and vector potential forces.
-
-\begin{prop}\label{prop:kinetic_hamiltonian}
-The extended Hamiltonian for the particles of a classical material under the kinematic assumption is $\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta) + m^2 c^2)$.
-\end{prop}
-
-\begin{justification}
-	We claim the conjugate momentum expressed as a function of the kinematic variables is $p_\alpha = m u_\alpha + A_\alpha(x)$ where $A_\alpha \in \mathfrak{X}(\mathcal{M})$ is a differentiable vector field.
-$\omega(dq^\alpha, dp_\beta) = dq^\alpha \wedge dp_\alpha \propto g(dx^\alpha, du^\beta) = dx^\alpha du_\alpha$.
-Let $m$ be the proportionality constant.
-$dq^\alpha \wedge dp_\alpha = dx^\alpha \wedge \partial_{u_\beta} p_\alpha du_\beta = m dx^\alpha du_\alpha$.
-$\partial_{u_\beta} p_\alpha = m \delta_\alpha^\beta$.
-After integration $p_\alpha = m u_\alpha + A_\alpha(x)$ where $A_\alpha$ is an arbitrary vector field.
-
-	We claim the Hamiltonian is $\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta) + m^2 c^2)$.
-$\frac{1}{m}g^{\alpha\beta}(p_\beta-A_\beta) = u^\alpha = d_s x^\alpha = d_s q^\alpha = \partial_{p_\alpha} \mathcal{H}$.
-After integration $\mathcal{H}=\frac{1}{2m}(p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta)+V(x)$ where $V$ is an arbitrary function.
-Let $s$ be the natural parametrization such that $u^\alpha u_\alpha = - c^2$.
-$V= \frac{1}{2}m c^2$ as $\mathcal{H}=0$ by \ref{prop:form_of_invariant_hamiltonian}.
-$\mathcal{H}=\frac{1}{2m}((p_\alpha-A_\alpha)g^{\alpha\beta}(p_\beta-A_\beta) + m^2 c^2)$.
-\end{justification}
-
-The above discussion gives the inertial mass $m$ a new physical meaning: it is the conversion constant that tells us how to convert a range of velocity to a range of conjugate momentum with the same number of possibilities.
-In other words: the more massive the particles, the more states per unit velocity.
-Intuitively, a more massive body is harder to accelerate because it has to go through more states for the same change in velocity.
-Also note that the Hamiltonian excludes massless particles because the kinematic assumption does that implicitly.
-A massless particle always has the same velocity, $c$, therefore the trajectory is not enough to tell us the dynamics.
-The motion of the photon is not enough to tell us its energy or its momentum.
-Therefore it makes sense that, by assuming kinematic equivalence we derived an equation valid only for massive particles.
-
-We have seen that the inertial mass $m$, the invariant speed $c$ and the reduced Planck constant $\hbar$ all serve similar purposes: convert one quantity to another preserving the number of possibilities.
-$dp=\hbar dk$ converts between ranges of wave number and conjugate momentum.
-$dx = c dt$ converts between ranges of time and space.
-$dp = m du$ converts between ranges of velocity and conjugate momentum.
-They allow us to keep proper bookkeeping when mapping sets of states.
-Using natural units means measuring all these quantities through the possibility count.
-There is a difference among them, though: while $c$ and $\hbar$ are simply determined by the unit system, $m$ also depends on the type of material.
-That is, different homogeneous materials can have different inertial mass.
-The reason is that the choice of units for $x^\alpha$ already determines the others ($x^\alpha \rightarrow u^\alpha$ and $x^\alpha \rightarrow q^\alpha \rightarrow k_\alpha \rightarrow p_\alpha$) therefore the link between units of $u^\alpha$ and $p_\alpha$ is not a new unit conversion.
-In other words: $m$ is not just another arbitrary constant set by the choice of units.
-
-Another important thing to note is that the relationship between conjugate momentum and velocity, $p_\alpha = m u_\alpha + A_\alpha(x)$, includes an arbitrary function.
-As it is arbitrary, we cannot give $p$ a well defined physical meaning in terms of the space-time trajectory.
-Suppose, in fact, we changed $A$ and $p$ while leaving $u$ unchanged, the physical description in terms of trajectories remains the same.
-
-We should also ask whether any choice of $A$ actually represents a different case leading to different trajectories.
-The simplest and more insightful way to study this problem is to use the kinematic quantities $x^\alpha$ and $mu_\alpha$ as state variables.
-They are not going to be conjugates of each other but they do capture the entire physical description.
-Any change in $A$ that does not affect the relationship between $x^\alpha$ and $mu_\alpha$ does not change the physics.
-Expressing the Hamiltonian and the Poisson brackets in terms of the kinematic variables we have:
-\begin{align*}
-\mathcal{H}&=\frac{1}{2}mu_\alpha u^\alpha + \frac{1}{2} m c^2 \\
-\{x^\alpha, x^\beta\} &= 0 \\
-\{x^\alpha, mu_\beta\} &= \delta^\alpha_\beta \\
-\{mu_\alpha, mu_\beta\} &= \{p_\alpha, - A_\beta\} + \{-A_\alpha, p_\beta\} \\
-&= \partial_\alpha A_\beta - \partial_\beta A_\alpha = F_{\alpha\beta}
-\end{align*}
-Note how the constraint $\mathcal{H}=0$ simply sets the rest kinetic energy.
-Note how the Poisson bracket between kinetic momentum is the force tensor.
-The weaker the force, the more $(x^i, mu_i)$ become independent d.o.f.
-The force tensor, then, tells how much $(x^i, mu_i)$ fail to be independent d.o.f.
-With these expressions, the Poisson bracket between any function of position and kinetic momentum can now be calculated.
-For example:
-\begin{align*}
-d_s x^\alpha &= \{x^\alpha, \mathcal{H}\} = u^\alpha \\
-d_s (mu^\alpha) &= \{mu^\alpha, \mathcal{H}\} = F^{\alpha\beta} u_\beta
-\end{align*}
-where we recognize the covariant form of the Lorentz force.
-
-Any change of $A$ that leaves $F_{\alpha\beta}$ unchanged does not affect the evolution of any function of position and velocity.
-Such transformation is what we call a gauge transformation.
-Note the same group of transformations is found by fixing $A$ and changing $p$ such that the laws of motion don't change.
-In other words: $A$ and $p$ are defined relatively to each other.
-$A$ defines the $0$ for $p$ and vice-versa.
-The gauge group expresses the arbitrariness of such a definition.
-
-The extended Hamiltonian we found is quadratic in energy, therefore negative energy states are possible.
-These correspond to the particle anti-states mentioned in \ref{prop:antistates}.
-As we fixed the natural parametrization, $u^\alpha u_\alpha = - c^2$, there are two cases: the parametrization is aligned with time, therefore $s=\tau$ where $\tau$ is proper time, or the parametrization is anti-aligned with time, therefore $s=-\tau$.
-The trajectory velocity $u^\alpha= \pm d_\tau x^\alpha$ is plus or minus the physical velocity depending on whether it's a standard particle state or an anti-state.
-Expanding the equation for the force, we have:
-\begin{align*}
-\pm d_\tau (\pm m d_\tau x^\alpha) &= m d_\tau^2 x^\alpha = \pm  F^\alpha_\beta d_\tau x^\beta
-\end{align*}
-Therefore the force acts on anti-particle states in the opposite way: they act as if they have opposite charge.
-Note that nothing physical is actually going backwards in time, just the parametrization.
-
-In this setting, charge conjugation $C$ is simply changing the parametrization from $s$ to $-s$.
-This will permute particle standard and anti-states as they follow the trajectories in the opposite direction.
-But if we apply parity $P$ and time reversal $T$, changing $x^\alpha$ to $-x^\alpha$, we would have the same net effect: the parametrization stays the same, but it now moves backwards in space and time.
-As they have the same net effect, if we apply all three we go back where we started.
-This gives us a very intuitive classical particle mechanics analogue for the $CPT$ theorem \cite{Weinberg}.
-
-Again we want to stress just how the math takes a very simple and physically meaningful form when studying the motion in phase space using position and kinematic momentum.
-It is most expressive because, by using the kinematic variables of a deterministic and reversible motion of an infinitesimal amount of material, it makes use of all the assumptions about the system we are studying.
-
-To sum up, the kinematic assumption has more inherent consequences than one would expect at first.
-It means the particles are identified by position and velocity and therefore the system of equations is second order.
-It links the space of initial conditions on $T\mathcal{M}$ with the space of states on $T^*\mathcal{M}$.
-Such a link allows phase space symplectic form $\omega$ to induce the space-time metric tensor $g$.
-It allows us to define Lagrangians.
-It constrains the motion to massive particles under scalar and vector potential forces.
-Moreover, it allowed us to reunderstand basic concepts, like the invariant speed $c$, the inertial mass $m$ and the force tensor $F_{\alpha\beta}$ in terms of state counting and d.o.f.
-independence.
-Finally, it allowed us to construct classical equivalents for anti-particles and the $CPT$ theorem.
-
-We now have all the building blocks of classical mechanics, the consequences of three seemingly simple assumptions.
-We have more insight into why these blocks fit together the way they do.
-But we have also placed ourselves in a very good position to move forward to quantum systems by using the new physical insights we have gained.
-
-\section{Prelude to quantum}
-\label{sec:quantum_prelude}
-
-In section \ref{sec:reducibility} we made the key assumption of infinitesimal reducibility which led to classical Hamiltonian mechanics.
-We are now going back to that point, that is we set aside the kinematic assumption and time dependent evolution, and take a different path.
-But before introducing assumption \ref{ass:irreducibility} and deriving quantum mechanics as its consequence, we first give a semi-classical account of the conceptual picture that emerges.
-This allows us to build the intuition more gradually and to show that classical and quantum mechanics are not as different as one may first imagine.
-
-We will see that, as a first approximation, quantum particles are more suitably described by classical composite states (i.e.
-distributions in phase space) like any other object made of a finite amount of material.
-We'll also see that classical distributions already possess qualitative properties that are typically associated with quantum systems, including a classical analogue of the uncertainty principle.
-
-\subsection{The case against point particles}
-
-As we defined them in previous sections, classical particles are the infinitesimal limit of subsequent subdivisions and therefore should not be understood as point-like objects of finite mass.
-It was their being infinitesimal and spread around a small neighborhood that gave us conjugate pairs of state variables and Hamiltonian evolution.
-Any finite amount of classical material, therefore, is described by a distribution.
-If we take a quantum particle, say an electron, it has a finite amount of charge/mass/energy.
-Therefore, to a first approximation, it must be described by a distribution as well.
-
-The idea that an electron is not a point particle may, to some, seem preposterous at first.
-But we contend that it is no more preposterous than to think it point-like, described by a probability distribution that self-interacts.
-In fact, the idea of an electron being a distribution of material is already conceptually more compatible with two features of quantum mechanics.
-
-The first one is that quantum particles are not associated with a well defined value for position and momentum.
-If we misidentify the electron with a classical particle (i.e.
-the infinitesimal part of a classical material), this would indeed be worthy of note and puzzlement.
-But if the electron is a distribution, there is no issue because a phase space distribution does not have a single value for position or momentum.
-The earth is, so to speak, in both North America and Europe at the same time.
-In fact, based on what we saw before, any object of finite mass is a distribution over a region of phase space, and discontinuous distributions are ruled out by \ref{prop:continuity}.
-An electron is an object of finite mass: why should it be any different?
-In fact, if one claimed an electron to be point-like, he would be claiming an electron is fundamentally different than all other physical objects of finite mass/energy (which are the only ones we experimentally work with).
-Shouldn't \emph{he} have the burden of proof?
-
-The other feature is particle-wave duality.
-If the spread of the distribution is narrow enough for the problem at hand, the motion can be described by just the position and momentum of the center of mass, making it behave like a classical particle.
-If the spread is not negligible (e.g.
-the object encounters obstacles or slits of a smaller scale) then we have to use the full distribution.
-Therefore, even in this classical approximation, the electron would have this dual nature depending on the problem.
-
-While this picture is quite reasonable, some people may reject it because they believe that a fundamental object must have no size, it must be a point.
-But that does not follow: lack of spatial extent and lack of substructure are independent concepts.
-
-In fact, we can conceptually construct an infinitesimal object that is not elementary.
-Suppose we have dots moving on the surface of a sphere.
-Their motion can be expressed in terms of their polar and azimuthal angles.
-If we shrink the sphere, the dynamics and its description stays the same.
-We can collapse the whole sphere to a point: it is now a point-like object that is not elementary, as there is motion detectable along different angles of incidence.
-
-Whether a physical system has a distinguishable substructure depends on whether there exists a physical process that can interact with one of its parts independently of the others.
-If we are always forced to interact with the whole system, then we can't find any discernible part.
-The fact that the system is extended in space is a separate property, which does not necessarily imply the existence of such a process.
-
-If we abandon the idea of point-like objects of finite mass, we avoid these problems and get a picture more consistent across the different realms of physics.
-And we can also see that classical mechanics already has some features that are commonly ascribed to quantum mechanics.
-
-\subsection{Classical uncertainty principle}
-
-Suppose we have our classical electron described by a finite amount of classical material distributed over a single d.o.f.
-$\int \rho_\mathcal{c} dq \wedge dp = 1$ where $\rho_\mathcal{c}$ is the normalized distribution.
-In this context, we will call a \emph{fragment} an infinitesimal part of our classical electron.
-The Shannon entropy \cite{Shannon} associated with the distribution is finite as well.
-$I(\rho_\mathcal{c}) = - \int \rho_\mathcal{c} \log (\rho_\mathcal{c})  dq \wedge dp = I_0$.
-Intuitively, this value can be thought of as the amount of information required to identify a fragment within the distribution.\footnote{To be precise, Shannon's differential entropy is actually the number of bits required to identify an element of the given distribution relative to one of a uniform distribution of unitary range on all axes.}
-
-As the state evolves into its final state, the amount of material remains the same: $\int \rho_{\hat{\mathcal{c}}} dq \wedge dp = 1$.
-The final entropy will also remain the same: $I(\rho_{\hat{\mathcal{c}}}) = I_0$.
-Mathematically, because $-\rho_\mathcal{c} \log (\rho_\mathcal{c})$, like the distribution itself, is just another integrable function, its integral is conserved by Hamiltonian evolution.
-Physically, the entropy remains unchanged because the amount of information required to identify a fragment given the distribution has to remain the same under deterministic and reversible evolution.
-Therefore, in the same way that each fragment is constrained to move through states at equal energy, the composite state is constrained to move through distributions at equal entropy.
-
-We now ask the following question: once we fix the entropy, what is the relationship between the spread of the marginal distributions along $q$ and $p$?
-Using Lagrange multipliers, we fix the total amount of material to $1$ and the total entropy $I_0$ and find the distribution that minimizes the product of the variance $\sigma_q^2 \sigma_p^2 \equiv \int (q-\mu_q)^2 \rho_{\mathcal{c}} \, dqdp \int (p-\mu_p)^2 \rho_{\mathcal{c}} \, dqdp$.
-\begin{align*}
-L = &\int (q-\mu_q)^2 \rho_{\mathcal{c}} \, dqdp \int (p-\mu_p)^2 \rho_{\mathcal{c}} \, dqdp \\
-\journal{ 
-	&+ \lambda_1(\int \rho_{\mathcal{c}} dqdp - 1) + \lambda_2(- \int \rho_{\mathcal{c}} \ln \rho_{\mathcal{c}} \, dqdp - I_0)\\ }
-\arxiv{ 
-	&+ \lambda_1(\int \rho_{\mathcal{c}} dqdp - 1) \\ &+ \lambda_2(- \int \rho_{\mathcal{c}} \ln \rho_{\mathcal{c}} \, dqdp - I_0)\\ }
-\delta L = &\int \delta \rho_{\mathcal{c}} [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 + \\ &\lambda_1 - \lambda_2 \ln \rho_{\mathcal{c}} - \lambda_2 ] dqdp = 0 \\
-\lambda_2 \ln \rho_{\mathcal{c}} = &\lambda_1 - \lambda_2 + (q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\
-\rho_{\mathcal{c}} = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(q-\mu_q)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_q^2 (p-\mu_p)^2}{\lambda_2}}\\
-\end{align*}
-We solve the multipliers and have:
-\begin{align*}
-\rho_{\mathcal{c}} = &\frac{1}{ 2 \pi \sigma_q \sigma_p} e^{-\frac{(q-\mu_q)^2}{2\sigma_q^2}} e^{-\frac{(p-\mu_p)^2}{2\sigma_p^2}} \\
-I_0 = &\ln (2\pi\sigma_q\sigma_p) + 1
-\end{align*}
-We find that the distribution that minimizes the spread is the product of two independent Gaussians.
-Recall that in quantum mechanics, the Gaussian wave packet is the state that minimizes uncertainty: this is the classical analogue.
-
-As the distribution evolves, the entropy is conserved and therefore the product $\sigma_q^2 \sigma_p^2$ can never be less than the one given by the Gaussian distribution of the same entropy. We have:
-\begin{align*}
-\sigma_q\sigma_p \geq \exp (I_0 - 1) / 2 \pi 
-\end{align*}
-This is strikingly similar to the Heisenberg uncertainty principle, except that the value depends on the initial entropy $I_0$.
-
-Suppose, though, that while the evolution of the distribution of our electron as a whole is deterministc and reversible, each fragment is undergoing non-deterministic motion.
-Suppose $\rho$ is a distribution at equilibrium of such motion.
-While each fragment moves randomly, the overall distribution $\rho$ stays the same: the non-deterministic motion is effectively just reshuffling the fragments within the same distribution.
-The value of the entropy of the distribution, and therefore its spread, is set by the non-deterministic motion of the fragments: the stronger it is, the greater the entropy.
-
-Suppose now our classical electron undergoes quasi-static evolution.
-At each instant we can assume $\rho$ to be at equilibrium.
-At each instant the fragments are reshuffled.
-At each instant the entropy of the distribution is the same one associated with the random motion.
-The overall evolution is still Hamiltonian, but the trajectories predicted by the framework are now just the average motion of the fragments.
-
-If we assumed that any classical electron were affected by the same source of randomness, all fragments would exhibit a random walk of the same magnitude as defined by the information entropy.
-The spread of all distributions for the material would satisfy the same inequality.
-This gives us the classical uncertainty principle: $\sigma_q\sigma_p \geq \exp (I_\varnothing - 1) / 2 \pi $ where $I_\varnothing$ is the entropy associated with the source.
-
-Such an assumption is not far fetched.
-The source of entropy could be identified in the environment (i.e.
-a minimum interaction that cannot be avoided as perfect isolation cannot be achieved) or the internal self interaction of the fragments (i.e.
-the evolution of the unstated part).
-Whatever it may be, if it is characterized by an entropy comparable to the one of the distribution (i.e.
-if we are trying to localize the electron better than the source can allow) the classical assumption breaks down as the random motion of each fragment is not negligible.
-
-The generalization to multiple d.o.f.
-is outside of the scope of this work.
-Note, though, that this result is related to the Gromov non-squeezing theorem and the symplectic camel, which are already generalized to multiple d.o.f.
-It is the link with information theory that gives us further physical insight and is of particular interest given the strong connection between thermodynamics and Shannon's entropy.
-
-This discussion shows, once again, how we can already find in classical mechanics some of the ingredients that are typically assumed to be characteristics of quantum mechanics.
-It has its own flavor of anti-particles, $CPT$ theorem and uncertainty principle.
-These should really be considered general features of deterministic and reversible evolution of distributions (i.e.
-Hamiltonian mechanics).
-Quantum mechanics just inherits these traits.
-
-The discussion also gives us an intuitive physical picture that we will develop further in the next section: a quantum particle is a well determined distribution of not well determined fragments.
-The distribution as a whole can be studied and evolves deterministically while the fragments do not.
-
-\section{Irreducible systems}
-\label{sec:irreducibility}
-
-In this section we explore a different relationship between the whole and its parts.
-We will consider an irreducible material, one where the state and the evolution of the whole does not tell us anything about the state and the evolution of the parts.
-The material is still conceptually infinitesimally decomposable, made of infinitesimal parts which we call fragments.
-But each fragment cannot be studied independently and therefore cannot be assigned a state: the internal dynamics is undetermined.
-We'll find that the evolution of an irreducible material is suitably described by the standard framework of quantum mechanics.
-
-\subsection{Quantized material}
-
-In section \ref{sec:reducibility} we introduced the idea of a classical material, one that is homogeneous and infinitesimally reducible, which represented a way to relate the state of the whole system to the state of its parts.
-In this section we explore an alternative case.
-
-We will call a \emph{quantized material} a homogeneous material that is finitely reducible.
-That is, as we decompose the system we can only assign a state to finite amounts of material, which we call \emph{quantum particles}.
-As the material is homogeneous, all its particles carry the same amount of material and are described by the same state space.
-Each quantum particle is irreducible: its state does not tell us anything about the state of its parts.
-While this idea should be familiar enough, it is best to iron out some details to avoid possible misconceptions.
-
-The first possible source of confusion is between divisibility, the idea that an object can be divided into parts, and reducibility, the idea that the state of the whole is equivalent to the state of the parts.
-More precisely, a system is divisible if there is a time evolution $\mathcal{T}_{\Delta t}: \mathcal{S} \rightarrow \mathcal{S}_1 \times \mathcal{S}_2$ that takes the state of a system and returns the state of two separate isolated systems.
-It is reducible if $\mathcal{S} = \mathcal{S}_1 \times \mathcal{S}_2$, the state of the system is the state of the two subsystems.
-Most of the time these two concepts overlap, but they are separate as we can see in the following examples.
-
-Consider a planarian flatworm.
-If we cut it into multiple pieces, each part will grow into a new worm.
-Yet, the state of each worm is not the state of multiple worms put together.
-The planarian is divisible into multiple worms but it is not reducible to multiple worms.
-Consider a magnet.
-Its state is described, to a first approximation, by the position of the north and south poles.
-Yet, we cannot divide the magnet into a north and south pole.
-The magnet is reducible to the poles but is not divisible into the poles.
-
-Consider a muon.
-It will decay into an electron and two neutrinos.
-But the state of a muon is not the state of an electron and two neutrinos.
-The muon is divisible into the three particles but it is not reducible to them.
-Consider a proton.
-Its state is given by the state of its three valence quarks and the gluon field.
-Yet, we cannot divide the proton into the three isolated quarks.
-The proton is reducible to quarks, but not divisible into quarks.
-
-The fundamental property of a quantum particle, as introduced here, is not indivisibility: it's irreducibility. It is the state that cannot be further partitioned, not the material itself.
-
-Even more subtle is the difference between divisibility and decomposability, the idea that the object has parts.
-That is, a system is decomposable if there is a law $+ : \mathcal{S} \times \mathcal{S} \rightarrow \mathcal{S}$ to combine two states to get a third.
-While irreducibility says nothing on whether a quantum particle can be divided, it implicitly assumes that it can be decomposed.
-Irreducibility describes the relationship between the state of the whole and the state of the parts, therefore the existence of parts is implicit in the definition.
-But it also tells us that we are limited in what description we can give to these parts: we cannot isolate them as we want and act on them independently because if we were able to do that we could assign them an independent state.
-
-Therefore, while we must be able to talk about parts of a quantum particle, not all questions will make sense.
-Can we ask what fraction of a quantum particle is in a particular region of space?
-Yes, since the particle can be spatially decomposed some amount of material may be inside the volume.
-Can we ask where the fragments in that same fraction will be at a later time?
-No, since the particle is irreducible we cannot describe the evolution of a part.
-
-Note that if the material is infinitesimally reducible (i.e.
-the classical case), reducibility/decomposability/divisibility are all possible.
-This is why one may not dwell on their difference and use the word ``part" interchangeably in the three cases.
-Now that we assume irreducibility, it is better to use more precise language to avoid confusion.
-Therefore we will call \emph{fragments} the infinitesimal parts of the material that cannot be described independently.
-We will call \emph{components} the parts that the material can be decomposed into.
-A component will have a state associated with it, and therefore will always be a collection of fragments.
-
-The proper way to think of a quantized material is therefore the following: a continuum of infinitesimal fragments clumped together to form quantum particles, the smallest amount of material to which we can assign a state.
-With this picture in mind, we can see, even before developing the math, how some features typical of quantum behaviors emerge.
-
-The first feature is the importance of statistics.
-Consider a fragment of a quantum particle in an initial state.
-As we cannot track its motion, it can correspond to any fragment of the particle in its final state.
-That is, the motion of a fragment of a particle is a random walk within the evolution of a particle.
-
-The second feature is non-locality.
-Suppose we have a quantum particle distributed in space.
-Suppose we exert a force only on a subregion of its distribution.
-As the particle is irreducible, we cannot interact with only part of it: we must also affect the part where no force is exerted.
-We cannot simply interact with the part at a specific region, as this would allow us to assign a state to that part.
-
-Wave-particle duality is also accounted for.
-The quantum particle behaves like a wave in the sense that its parts form a distribution in space.
-Yet, it behaves as a single unit, as the state is irreducible and cannot be split.
-
-The most striking feature, though, is that we can predict the functional form for particle decay over time.
-Suppose a quantum particle spontaneously decays into two or more particles.
-Such a process cannot be deterministic as the number of d.o.f.
-increases.
-As the decay is spontaneous, it is not triggered by the environment (i.e.
-external state).
-It must be triggered by the internal dynamics of the fragments.
-As we assumed the internal dynamics is unknown, we cannot learn anything about it by looking at the product of the decay.
-In particular, all states have to have equal chance of decay or we could tell that the internal dynamics is more stable for some states.
-The probability of decay is constant for all states at all times.
-The probability of survival follows the exponential distribution.
-
-As we see qualitatively how the features typical of quantum mechanics naturally emerge from this picture, we are ready to make this more quantitatively and mathematically precise.
-
-\subsection{Irreducible material and complex vector spaces}
-
-It is time to turn these concepts into a precise mathematical formulation as we did for classical mechanics.
-We want to formalize the state space for an \emph{irreducible material} as a quantum particle will be a unit amount of such material.
-
-As for the classical material, the irreducible material is thought to be decomposable: each amount can be seen as the sum of different components (e.g.
-spatially distributed).
-Therefore \ref{prop:abelian_group} applies and the state space $\Psi$ of an irreducible material is an abelian group.
-
-Yet, the description of each component cannot be more accurate than the description of the whole.
-To capture this idea precisely we need to characterize the aspects of the unstated part that influence the state definition.
-How do we do that?
-We can look at the classical derivation for guidance.
-In that case, we looked at the set of transformations that increased/decreased the amount of material.
-The idea is that we can think of these transformations as characterizing what the state can and cannot say about the infinitesimal parts.
-
-Suppose we have a state $\mathcal{c}_1$ for an amount $a_m$ of classical material.
-Suppose we assigned a numeric identifier to each particle of the material.
-This identifier would be just a dummy label we assign, not a physical quantity.
-As the classical material is homogeneous, permutations between classical particles within the same state have no effect.
-Therefore the ordering given by the identifier does not matter.
-Identifying a classical particle within a composite state is equivalent to picking a number from a random variable $A$ uniformly distributed over an interval of size $a_m$.
-
-Suppose we have another state $\mathcal{c}_2 = \tau(a) \mathcal{c}_1$ where the only difference is an increase/decrease of the amount of material by a factor $a$.
-The random variable $B$ that identifies a particle of $\mathcal{c}_2$ is now a uniform distribution $B=aA$ where the range is the new amount of material: there are now proportionally more/less labels to assign.
-This is also the only possible transformation that can link the two uniform random variables $A$ and $B$.
-Therefore increase/decrease of material is the only operation that is captured by the state among those that affect the labeling of the fragments.
-In other words: the classical state does not care about the permutation of the fragments, but it cares how many fragments are there.
-
-For the irreducible material we follow the same strategy and study the transformations that affect the labeling of what the irreducible state cannot tell us: the configuration of each fragment.
-Similarly, the irreducible state will not care about permutations of fragment configurations, but it will care how many fragment configurations are there.
-
-Suppose we have a state $\psi_1$ for an amount $a_m$ of irreducible material characterized by one d.o.f.
-The configurations of the infinitesimal fragments will be identified by two values, as a d.o.f.
-is two-dimensional.
-Each fragment configuration is identified by a pair of independent random variables $(A, B)$, which are, as before, just dummy variables we use to name the fragment configurations.
-As the material is homogeneous, the distribution is uniform or we could tell something about the fragment configurations.
-We choose A and B to be uniform distributions of interval $\sqrt{a_m}$: identifying a fragment configuration is like picking a point at random over a square of area $a_m$, the amount of material.
-
-Suppose we have another state $\psi_2$, where the labeling for the fragment configurations has changed.
-We can label the fragment configurations of $\psi_2$ in the same way, with two independent uniform distributions $(C, D)$.
-But since the labeling is the only thing that has changed, we must be able to relate the new distribution over $(C,D)$ to the old distribution over $(A,B)$.
-The relationship between the two pairs of variables must be linear since all distributions are uniform.
-$C=aA+bB$.
-To preserve the square distribution we must then have $D= -bA + aB$.
-We can write $\psi_2 = \tau(a,b) \psi_1$ as $a$ and $b$ are the only parameters of the transformation.
-Or more compactly, $\psi_2 = \tau(c) \psi_1$ where $c=a+\imath b$.
-
-The transformation $\tau(c)$ increases/decreases the area by the modulus $|c|^2=a^2+b^2$, which physically corresponds to the increase/decrease of amount of material.
-It also rotates the square by the phase $\theta_c \equiv arg(c)$, which corresponds to a change in correlation: if $\theta_c = 0$, $C = |c| A$ are clearly correlated while if $\theta_c = \pi$, $C = - |c| A$ are clearly anti-correlated.
-
-This line of reasoning is fairly abstract and it may seem to have no bearing with quantum states.
-Therefore, jumping a bit ahead, let's see if we can already find quantitative consequences of this model and see whether they are consistent with our knowledge of quantum mechanics.
-
-If the amount of material is connected to the area of two uniform distributions, it should combine like the variance of a random variable.
-And in fact it does.
-Suppose $X$ and $Y$ are two random variables,
-\begin{equation}\label{eqn:correlation}
-\sigma^2_{X+Y} = \sigma^2_{X} + \sigma^2_{Y} + 2 \, \sigma_{X} \sigma_{Y} \rho_{X,Y}
-\end{equation}
-where $\sigma^2$ is the variance, $\sigma$ the standard deviation, $\rho_{X,Y}= COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient \cite{Grimmet}.
-As $-1\leq\rho_{X,Y}\leq+1$, the total variance can be greater or smaller than the sum of the variance of the components.
-Now suppose $\psi$ is the state for a unit amount of irreducible material (i.e.
-a quantum particle).
-We can write $c \psi = (c_1 + c_2) \psi = c_1 \psi + c_2 \psi$.
-We have:
-\begin{equation}\label{eqn:correlation_state}
-|c_1+c_2|^2=|c_1|^2 + |c_2|^2 + 2 |c_1||c_2|\cos(\theta_{c_2} - \theta_{c_1})
-\end{equation}
-The two expressions are formally identical and it shouldn't come as a surprise.
-The more material we have, the more labels for the fragment configurations to choose from, the bigger the spread of the random variables that assign the labels.
-$-1\leq\cos(\theta_{c_2} - \theta_{c_1})\leq+1$ takes the place of the Pearson correlation coefficient.
-Quantum interference can then be understood in terms of the combination of correlated/anti-correlated distributions.
-
-If the phase determines the correlation between two states, then it should only be meaningful when comparing two states.
-And in fact only differences in phases are physically distinguishable in quantum mechanics.
-A more meaningful name for the phase, then, would be \emph{correlation angle}.
-
-If quantum states are uniform distributions over fragment configurations, we should see a hint of that when computing distributions over observables.
-And in fact we do.
-Suppose we have $Q\psi = q \psi$, an eigenstate of some observable.
-Then it is also an eigenstate of the infinitesimal transformation of the conjugate variable $1+\frac{Qdp}{\imath\hbar}$.
-Since it has a symmetry over $p$, the distribution has to be equal at each value of $p$: the distribution is uniform.
-
-If quantum particle states are really just unit amounts of material uniformly distributed, we should be able to calculate the Shannon entropy associated with them.
-Using our dummy variables, $I(\psi) = - \int \rho \ln (\rho) dA dB = - \int 1 ln (1) dA dB = 0$.
-For a quantum mixed state $\tilde{\rho} = \sum \tilde{\rho}_i |\psi_i\rangle \langle \psi_i |$, we have $I(\tilde{\rho}) = - \sum \tilde{\rho}_i \ln \tilde{\rho}_i - \tilde{\rho}_i I(\psi_i) = - \sum \tilde{\rho}_i \ln \tilde{\rho}_i = -tr(\tilde{\rho} \ln \tilde{\rho})$.
-And in fact this is the Von Neumann entropy \cite{Von Neumann} used in quantum information, that implicitly assumes all states provide the same amount of entropy (i.e.
-zero).
-
-If two quantum particle states $\psi_1$ and $\psi_2$ are really just two states at equilibrium with the same entropy, we should always be able to find a deterministic and reversible evolution to connect the two.
-And in fact we can.
-Take the plane identified by $\psi_1$ and $\psi_2$.
-Find the infinitesimal rotation $1 + \frac{Od\alpha}{\imath\hbar}$ along that plane.
-$\psi_2 = e^{\frac{O\alpha}{\imath\hbar}} \psi_1$ is the unitary evolution of the other.
-Note that this is \emph{not} always possible in classical mechanics, as two states $\mathcal{c}_1$ and $\mathcal{c}_2$ may have different entropy.
-
-In other words, our simple model is not only consistent with quantum mechanics but it has predictive power: it provides intuition that is well supported by these findings.
-It is not ``just a mathematical trick."
-
-If we increase the number of independent d.o.f.
-the group of transformations remains the same.
-While we have more fragment configurations to label, each d.o.f.
-is labeled independently.
-Therefore we have $n$ uniform distributions, each over a square of the same amount of material.
-This means that, as we combine components, the labels have to combine in the same way along the different d.o.f.: the increase/decrease of material and the change in correlation has to be the same across all d.o.f.
-
-Mathematically, the state space $\Psi$ for an irreducible material is a complex vector space.
-It is so because it is the state space for a decomposable material (i.e.
-an abelian group) where we can increase/decrease the amount and change the correlation of the internal description (i.e.
-a set of transformations conveniently parametrized by a complex number).
-
-\begin{prop}\label{prop:complex_vector_space}
-	The state space $\Psi$ for an irreducible material is a vector space over $\mathbb{C}$.
-\end{prop}
-
-\begin{justification}
-	We claim $\Psi$ is an abelian group. $\Psi$ is the state space for a decomposable homogeneous material and is therefore an abelian group  by \ref{prop:abelian_group}.
-
-	We claim the set of physically distinguishable transformations $T$ that relabel the fragment configurations is isomorphic to $\mathbb{C}$.
-Consider an amount $a_m \in \mathbb{R}$ of irreducible material characterized by one d.o.f.
-The configuration of a fragment of the material is characterized by two random variables $A, B: \Omega \rightarrow \mathbb{R}$ where $\Omega$ is the sample space for the configuration of the fragments.
-As the material is homogeneous, $A$ and $B$ can be chosen such that the distribution $\rho$ of material over such variables is uniform over a range of $\sqrt{a_m}$ for both variables.
-We have $a_m=\int \rho dA \wedge dB = \int dA \wedge dB$.
-Consider a second amount $\hat{a}_m$ of irreducible material where we changed the labeling of the fragment configurations.
-Its fragment configurations can also be characterized by a uniform distribution over two variables $C, D: \Omega \rightarrow \mathbb{R}$ over a range of $\sqrt{\hat{a}_m}$.
-As the two amounts of material differ just by the labeling of fragment configuration, $(C,D)$ is a function of $(A,B)$.
-We have $\hat{a}_m = \int dC \wedge dD = \int |J| dA \wedge dB$ where $|J|$ is the Jacobian determinant.
-As both distributions are uniform, $|J|$ has to be constant.
-The transformation has to be linear: $C=aA + bB$ and $D=-bA + aB$.
-As the change in amount of material, the only physically distinguishable quantity associated with the fragment configurations, is fully specified by $a$ and $b$, no other transformation is physically relevant for the state.
-Let $\tau: \mathbb{C} \rightarrow T$ represent such a transformation parametrized by $c=a+\imath b$.
-$|c|^2 = |J| = (a^2 + b^2) = \hat{a}_m / a_m$ is the ratio of the two amounts of materials and is physically distinguishable.
-Define on $T$ an addition $+: T \times T \rightarrow T$ and a multiplication $*: T \times T \rightarrow T$ such that $\tau(z_1) + \tau(z_2) = \tau(z_1+z_2)$ and $\tau(z_1) * \tau(z_2) = \tau(z_1*z_2)$, $z_1,z_2 \in \mathbb{C}$, the sum and product of the transformation is equal to the sum and product of their respective factors.
-The phase $arg(c)$ under addition is physically distinguishable as it can affect the modulus of the result.
-$\tau$ is an isomorphism between $T$ and $\mathbb{C}$ as fields.
-Assume now that the material is characterized by multiple independent d.o.f.
-The transformation on each d.o.f.
-is defined independently by $c_i \in \mathbb{C}$.
-As $\hat{a}_m / a_m$ is the same under all d.o.f.
-$|c_i|=|c_j|$.
-The relationship has to always hold under addition as well, therefore $arg(c_i) = arg(c_j)$.
-$T\cong \mathbb{C}$ also for multiple d.o.f.
-
-	We claim $\Psi$ is a vector space over $\mathbb{C}$.
-The abelian group $\Psi$ can be extended with the operations defined by $T$ as was done in \ref{prop:real_vector_space} for the infinitesimally reducible case.
-\end{justification}
-
-\subsection{Complex valued distributions}
-
-In the previous sections we saw how a finite amount of classical material was described by a distribution over the state variables of infinitesimal parts.
-For an irreducible material, we cannot assign states to the fragments, therefore we cannot properly talk about state variables.
-Therefore we call \emph{fragment variables} the quantities that we use to label the fragment configurations, and it will be over these variables that we will define the distribution of irreducible material.
-
-Naturally, we cannot give a complete description for the configuration of a fragment, therefore we cannot give a joint distribution for all fragment variables at the same time.\footnote{This recovers the notion of complementarity introduced by Bohr \cite{Bohr}.} We can, though, find a maximal set of fragment variables: one that can be used to provide the best description possible.
-While there isn't a unique maximal set, we can show that one such set is provided by the fragment variables $q^i : \mathcal{Q} \rightarrow \mathbb{R}$ that define the units required to describe the system.
-
-In fact, we cannot add a new d.o.f.
-(as this would mean changing $\mathcal{Q}$) and we cannot add a new fragment variable in an existing d.o.f.
-(as this would fully specify said d.o.f.).
-Therefore the set of fragment variables $q^i$ already provides a maximal set of fragment variables.
-This means that a state $\psi$ must tell us for each point in $\mathcal{Q}$ the amount and correlation angle of the material.
-That is: the state will be identified by $\psi : \mathcal{Q} \rightarrow \mathbb{C}$, the wave function over $\mathcal{Q}$.
-
-The square modulus at each point will give us the density of the material.
-As in the classical case, such density must be integrable to give us the total amount.
-Therefore the function $\psi(q) \in L^2(\mathcal{Q}, \mu)$ is square integrable.
-The norm $|\psi|^2 = \int_{\mathcal{Q}} |\psi(q)|^2 dq$ will give us the amount of material associated with the state.
-
-As the material is homogeneous, all its quantum particles are made of the same amount.
-We can set, by convention, the amount of material for a particle to be unitary, therefore quantum particle states are the subset of $\Psi$ such that $|\psi|^2 = 1$.
-As we have seen, changing the phase by a constant only affects the unstated part.
-Therefore normalized distributions that differ by a total phase are not physically distinguishable and represent the same quantum particle state.
-
-\begin{prop}\label{prop:wavefuntion}
-	The state space $\Psi$ for an irreducible material is isomorphic to a subspace of the space of Lebesgue square integrable continuous functions.
-That is $\Psi \cong \Psi(\mathcal{Q}, dq^i) \subseteq C(\mathcal{Q}, \mathbb{C}) \cap L^2(\mathcal{Q}, \mu, \mathbb{C})$ under the isomorphism $\Upsilon_q : \Psi \leftrightarrow \Psi(\mathcal{Q}, dq^i)$ where $\Psi(\mathcal{Q}, dq^i)$ is the space of wave functions over $q^i$.
-The state space for a quantum particle is isomorphic to the projective space $\mathsf{P}(\Psi)$.
-\end{prop}
-
-\begin{justification}
-	We claim the manifold $\mathcal{Q}$ that defines the units to describe a fragment of an irreducible material also provides a maximal set of fragment variables: no other fragment variable can be added.
-Let $\mathcal{Q}$ be the manifold that defines the units.
-Then for each state $\psi \in \Psi$ there exists a distribution $\rho_\psi(q^i)$ that represents the quantity of material at each point.
-Suppose this distribution could be further refined.
-Then another independent variable $k$ would exist such that for each state $\psi \in \Psi$ we would have a distribution $\rho_\psi(q^i, k)$.
-$k$ cannot be part of a new d.o.f.
-as it would use a unit defined independently from $\mathcal{Q}$.
-This is a contradiction as $\mathcal{Q}$ defines all units.
-$k$ cannot lie in the same d.o.f.
-as of one of the $q^i$ or it would provide a complete description for the fragment.
-This is a contradiction as the material is irreducible.
-$\mathcal{Q}$ provides a maximal set of fragment variables.
-
-	We claim $\Psi \cong \Psi(\mathcal{Q}, dp^i) \subseteq C(\mathcal{Q}, \mathbb{C}) \cap L^2(\mathcal{Q}, \mu, \mathbb{C})$ as a vector space.
-The maximal physical description of a fragment configuration is given by a point in $\mathcal{Q}$ parametrized by $q^i$ and by a transformation $\tau \in T$ parametrized by $c \in \mathbb{C}$.
-For each $\psi \in \Psi$ there must exist a wave function $\psi(q^i) : \mathcal{Q} \rightarrow \mathbb{C}$ that provides the maximal physical description for the state.
-$\rho_\psi(q^i) = |\psi(q^i)|^2$ as the square modulus of $\psi(q^i)$ is the increase/decrease from the unit amount of material.
-$\rho_\psi(q^i) \in C(\mathcal{Q}, \mathbb{R})$ by \ref{prop:continuity}.
-$\rho_\psi(q^i) \in C^1(\mathcal{Q}, \mathbb{R})$ by the same arguments in the justification for \ref{prop:differentiable_manifold}.
-Therefore $\psi(q^i) \in C(\mathcal{Q}, \mathbb{C})$.
-$\rho_\psi(q^i) \in L^1(\mathcal{Q}, \mu, \mathbb{R})$ by the same arguments in the justification for \ref{prop:integration}.
-Therefore $\psi(q^i) \in L^2(\mathcal{Q}, \mu, \mathbb{C})$.
-Let $\Upsilon_{q^i} : \Psi \rightarrow C(\mathcal{Q}, \mathbb{C}) \cap L^2(\mathcal{Q}, \mu, \mathbb{C})$ be the function that given a state returns its wave function.
-As the wave function captures the maximal description of the system, two different states must represent different distributions.
-$\forall \psi_1, \psi_2 \in \Psi, \psi_1 \neq \psi_2 \implies \Upsilon_{q^i}(\psi_1) \neq \Upsilon_{q^i}(\psi_2)$.
-$\Upsilon_{q^i}$ is injective.
-$\Upsilon_{q^i}$ is a bijection between $\Psi$ and $\Psi(\mathcal{Q}, dq^i)\equiv\Upsilon_{q^i}(\Psi)$.
-Let $\psi=\psi_1+\psi_2$, then $\Upsilon_{q^i}(\psi)=\Upsilon_{q^i}(\psi_1)+\Upsilon_{q^i}(\psi_2)$ as combining the distributions means combining the statistical description of the fragment configurations.
-Let $\psi_1=\tau(c)\psi_2$, then $\Upsilon_{q^i}(\psi_1)=c \Upsilon_{q^i}(\psi_2)$ as $\tau(c)$ transforms the fragment configuration description by the factor $c$.
-$\Upsilon_{q^i}$ is a homomorphism between vector spaces.
-$\Psi$ is isomorphic to $\Psi(\mathcal{Q}, dq^i) \subseteq C(\mathcal{Q}, \mathbb{C}) \cap L^2(\mathcal{Q}, \mu, \mathbb{C})$.
-
-	We claim the state space for a quantum particle is isomorphic to the projective space $\mathsf{P}(\Psi)$.
-Let $\psi \in \Psi$ be the state of a quantum particle.
-$\tau(e^{\imath \theta}) \psi$ with $\theta \in \mathbb{R}$ is not physically distinguishable from $\psi$ as phases are only physically distinguishable through addition.
-$\tau(\sqrt{\rho}) \psi$ with $\sqrt{\rho} \in \mathbb{R}^+ \backslash \{1\}$ cannot represent a quantum particle as it has a different amount of material.
-There exists only one distinguishable quantum particle state in the equivalence class $\psi \sim \tau(c) \psi$ with $c \in \mathbb{C}$.
-The state with no amount of material does not correspond to any particle state.
-The quotient space $\mathsf{P}(\Psi) \equiv (\Psi \backslash \{0\}) / \sim$ is isomorphic to the state space for quantum particles.
-\end{justification}
-
-\subsection{Inner product and operators}
-
-In both the reducible and irreducible case, the material is distributed over $\mathcal{Q}$.
-The $q^i$ do not label the states of a finite amount of material, but the configurations for the infinitesimal parts.
-For a classical material, the infinitesimal parts can be given a state, have their own topological space and a set of local state variables $(q^i, p_i)$ to identify them.
-We can use the richness of differential geometry.
-For an irreducible material, the infinitesimal parts cannot be given a state, cannot be fully identified by a set of variables.
-We have to use other mathematical tools, typically vector spaces.
-As we will often use linear combinations of states, we will write, for brevity, $c\psi$ instead of $\tau(c)\psi$.
-That is: each complex number still represents a state transformation.
-
-For the tool set to be on par with classical mechanics, such that we can write the equations of motion, we need to address three requirements.
-We need to be able to compare two states and ask what fraction of them share the same configuration: this will give us an inner product.
-We need to be able to express distributions and expectations for different quantities: this will give us self-adjoint linear operators.
-We then need to identify the self-adjoint linear operators for the quantities $k_i$ and $p_i$ and to be able to express the distribution of material in terms of those.
-Let's start with the first.
-
-The idea is that, given two states $\psi_1$ and $\psi_2$, we should be able to tell how much of $\psi_1$ is prepared in the same way as $\psi_2$.
-That is $\psi_1 = \alpha \psi_2 + \psi_3$ is the combination of the component that is prepared like $\psi_2$ (e.g.
-same distributions in position and momentum) and the component $\psi_3$ that is prepared completely differently.
-Let $\mathsf{P}_{\psi_2} : \Psi \rightarrow \Psi$ be the operator that gives us the component $\mathsf{P}_{\psi_2} (\psi_1) \equiv \alpha \psi_2$ prepared like $\psi_2$.
-This operator is linear in $\psi_1$: the component of the combined system that is prepared like $\psi_2$ is the sum of the subcomponents that are prepared like $\psi_2$.
-It also gives us the same result if applied once or twice: once we found the component that is prepared like $\psi_2$, it remains the same.
-$\mathsf{P}_{\psi_2}$ is therefore a projection: a linear operator such that $\mathsf{P}_{\psi_2}(\mathsf{P}_{\psi_2} (\psi_1))= \mathsf{P}_{\psi_2} (\psi_1)$.
-
-If $\psi_1$ is already prepared like $\psi_2$, then its projection on $\psi_2$ will be equal to $\psi_1$.
-Conversely, if the projection of $\psi_1$ is equal to itself, then it means that the state is already prepared like $\psi_2$.
-That is: $\mathsf{P}_{\psi_2} (\psi_1) = \psi_1 \Leftrightarrow \psi_1 = \alpha \psi_2$.
-Because of this, we can define a non degenerate complex product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbb{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{\langle \psi_2|\psi_1 \rangle}{|\psi_2|^2} \psi_2$.
-
-The projections have other useful properties.
-Because the material is irreducible, if some fraction of $\psi_1$ is prepared like $\psi_2$, then the same fraction of $\psi_2$ is prepared like $\psi_1$.
-If not, studying the fragments of one state would give us a better (or worse) description for the fragments of the other state.
-Therefore we have $\frac{|\mathsf{P}_{\psi_2} (\psi_1)|^2}{|\psi_1|^2}=\frac{|\mathsf{P}_{\psi_1} (\psi_2)|^2}{|\psi_2|^2}$.
-This tells us that $|\langle \psi_2|\psi_1 \rangle|^2 = |\langle \psi_1|\psi_2 \rangle|^2$.
-
-Lastly, if we take the projection of $\psi_1$ along $\psi_2$ and we project it again on $\psi_1$, we end up with a component of $\psi_1$.
-As this is a part of $\psi_1$ prepared as $\psi_1$, its state is just $\psi_1$ rescaled, with no change in correlation.
-That is: $\mathsf{P}_{\psi_1} (\mathsf{P}_{\psi_2} (\psi_1)) = a \psi_1$ with $a \in \mathbb{R}^+$.
-This tells us that $\langle \psi_2|\psi_1 \rangle = \langle \psi_1|\psi_2 \rangle^\dagger$.
-
-Because of the properties of the norm and the projections, the product is positive definite, linear and conjugate symmetric.
-In other words, the state space $\Psi$ for an irreducible material is an inner product space derived by the norm and projections.
-
-\begin{prop}\label{prop:inner_product}
-	The state space $\Psi$ for an irreducible material is an inner product with $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbb{C}$ such that $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$.
-\end{prop}
-\begin{justification}
-	We claim there exists a map $| \cdot | ^2 : \Psi \rightarrow \mathbb{R}$ such that $|\psi|^2 \ge 0 \; \forall \psi \in \Psi$ and $|\psi|^2 = 0 \Leftrightarrow \psi = 0$.
-Let $| \cdot | ^2 : \Psi \rightarrow \mathbb{R}$ be the map that returns the amount of material for a given state.
-Let $\psi \in \Psi$ be a state.
-$|\psi|^2 = \int \rho(q^i) d^nq = \int \psi^\dagger(q^i) \psi(q^i) d^nq \ge 0$.
-$|\psi|^2 = 0 \Leftrightarrow \psi = 0$ as the only state with no material is the empty state.
-
-	We claim $\forall \psi \in \Psi$ there exists a projection $\mathsf{P}_\psi : \Psi \rightarrow \Psi$ with $\psi \in \Psi$.
-Let $\psi \in \Psi$ be a state.
-Let $\mathsf{P}_\psi : \Psi \rightarrow \Psi$ be an operator that returns the component of a given state that is prepared as $\psi$.
-$\mathsf{P}_{\psi}(c_1 \phi_1 + c_2 \phi_2) = c_1 \mathsf{P}_{\psi}(\phi_1) + c_2 \mathsf{P}_{\psi}(\phi_2)$ is linear as the component of the composition prepared like $\psi$ is the composition of the components so prepared.
-$\mathsf{P}_\psi \circ \mathsf{P}_\psi = \mathsf{P}_\psi$ is a projection because the result of $\mathsf{P}_\psi$  is prepared as $\psi$.
-
-	We claim $\mathsf{P}_{\psi} (\phi) = \phi \Leftrightarrow \phi = c \psi$ for some $c \in \mathbb{C}$.
-Let $\psi, \phi \in \Psi$ such that $\mathsf{P}_{\psi} (\phi) = \phi$.
-The whole state $\phi$ is prepared as $\psi$.
-We can obtain $\phi$ from $\psi$ by a relabeling of the fragment configurations.
-$\phi = c \psi$ for some $c \in \mathbb{C}$.
-Let $\psi, \phi \in \Psi$ such that $\phi = c \psi$ for some $c \in \mathbb{C}$.
-Then all of $\phi$ is prepared as $\psi$.
-$\mathsf{P}_{\psi} (\phi) = \phi$.
-Therefore $\mathsf{P}_{\psi} (\phi) = \phi \Leftrightarrow \phi = c \psi$.
-
-	We claim $|\psi|^2|\mathsf{P}_{\psi} (\phi)|^2=|\phi|^2|\mathsf{P}_{\phi} (\psi)|^2$.
-Let $\psi, \phi \in \Psi$.
-As the material is irreducible, $\psi$ and $\phi$ cannot be used to describe smaller parts.
-The fraction of $\phi$ that is prepared as $\psi$ must be equal to the fraction of $\psi$ that is prepared as $\phi$.
-$\frac{|\mathsf{P}_{\psi} (\phi)|^2}{|\phi|^2}=\frac{|\mathsf{P}_{\phi} (\psi)|^2}{|\psi|^2}$.
-Multiplying by $|\psi|^2|\phi|^2$ we have $|\psi|^2|\mathsf{P}_{\psi} (\phi)|^2=|\phi|^2|\mathsf{P}_{\phi} (\psi)|^2$.
-This holds for the empty state as well.
-Let $\phi$ be the empty state.
-$|\phi|^2=0$ because it has no material and $|\mathsf{P}_{\psi} (\phi)|^2$ as no component of the empty state is prepared like $\psi$.
-
-	We claim $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi)) = r \psi$ for some $r \in \mathbb{R}^+$.
-Let $\psi, \phi \in \Psi$.
-Consider $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi))$: we first take the component of $\psi$ prepared like $\phi$ and then take the component of that prepared like $\psi$.
-The overall operation is simply returning a part of $\psi$.
-This must correspond to the same exact configuration for the fragments.
-The only thing that can change is the amount of material.
-Therefore $\mathsf{P}_{\psi} (\mathsf{P}_{\phi} (\psi)) = r \psi$ for some $r \in \mathbb{R}^+$.
-
-	We claim that the state space $\Psi$ for an irreducible material has an inner product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbb{C}$.
-$| \cdot | ^2$ and $\mathsf{P}_{\phi}$ satisfy the requirements in \ref{thrm:inner_product}.
-$\Psi$ is an inner product space with $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbb{C}$ such that $\langle \psi |\phi \rangle \psi = | \psi | ^2\mathsf{P}_{\psi}(\phi)$.
-
-	We claim that the inner product can be expressed in terms of the wave functions as $\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$.
-The amount of material associated with a state $\psi \in \Psi$ is $\langle \psi | \psi \rangle = | \psi |^2 = \int_\mathcal{Q} \psi^\dagger (q^i) \psi(q^i) d^nq$.
-The inner product associated with the $L^2$ norm is $\int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$ with $\psi_1, \psi_2 \in \Psi$.
-$\langle \psi_1 | \psi_2 \rangle = \int_\mathcal{Q} \psi_1^\dagger (q^i) \psi_2(q^i) d^nq$.
-\end{justification}
-
-Note that the inner product only has an indirect physical definition, based on the norm and the projections.
-These two concepts are therefore more useful to gain physical intuition as the first does not have a straightforward physical meaning.
-
-Also note that the observations made after \ref{prop:differentiable_manifold} for the  classical case are still valid.
-$\Psi$ cannot be a complete metric space for the norm induced by the inner product as it does not include discontinuous functions.
-Therefore it is only a pre-Hilbert space.
-For mathematical convenience, one can take its completion with respect to the norm induced by the inner product and obtain a Hilbert space.
-Yet, such object no longer represents a physically meaningful set of states.
-
-Therefore we need to make sure that the rest of our definitions and justifications do not require a complete metric space to be valid or they would not be physically meaningful.
-In particular, we don't want to use the usual link $\psi(q) = \langle q | \psi \rangle$ between the wave function and the inner product as the material cannot be prepared at a specific value when $q$ is a continuous variable.
-Similarly, we cannot show that to each state variable is associated a self-adjoint operator $Q = \sum q |q\rangle \langle q |$ simply by expressing it in terms of eigenstates as they may not represent physical states.
-
-We proceed in the following way.
-Suppose we have a fragment variable $f : \mathcal{Q} \rightarrow \mathbb{R}$.
-Consider the product $f (q) \psi (q)$: it returns the wave function weighted by the value of $f$.
-We can also think of $f$ as an operator that takes a complex function of $\mathcal{Q}$ and returns another complex function of $\mathcal{Q}$ that represents the distribution of the expectation over the fragments.
-The operator is linear since $f (q) (a\psi_1(q) + b\psi_2(q)) = a f(q) \psi_1(q) + b f(q)  \psi_2(q)$.
-The operator is self-adjoint since $f^\dagger = f$.
-Consider $\int_\mathcal{Q} \psi^\dagger (q) f(q) \psi(q) dq = \int f(q) | \psi(q)|^2 dq$.
-It integrates the value of $f$ weighted by the amount of the material: it is the expectation of the fragment variable $f$ over the distribution.
-
-While $\mathcal{Q}$ is enough to define the system of units, it is not enough to define all the fragment variables over which we can express the distributions of irreducible material.
-The conjugate variables $p_i=\hbar k_i$ cannot be expressed as a function of $q^i$, which reiterates the idea that we cannot specify all variables for each fragment.
-Suppose there is another maximal set of fragment variables.
-Then those variables will form a topological space $\hat{\mathcal{Q}}$ such that each $\psi$ can be expressed as a wave function $\psi(\hat{q}) :\hat{\mathcal{Q}} \rightarrow \mathbb{C}$.
-
-As each state $\psi$ can be identified by a wave function on either space, there must be an operator $\Upsilon^q_{\hat{q}}(\psi(q)) = \psi(\hat{q})$ that converts the wave function on one space to the other.\footnote{Note that a neighborhood of $q$ is not necessarily a neighborhood of $\hat{q}$.
-They are only if $q$ and $\hat{q}$ can be expressed as a bijection of each other $\hat{q}^i(q^i)$: they are different coordinates of the same manifold.} As a linear combination of states corresponds to a linear combination of wave functions, $\Upsilon^q_{\hat{q}}$ is a linear operator.
-As converting the space again must give us the original wave function, $\Upsilon^{\hat{q}}_{q} \circ \Upsilon^q_{\hat{q}}= I$ is unitary.
-We can now define a linear operator $F_{[\hat{q}]} \equiv f(\hat{q})$ corresponding to fragment variable $f : \hat{\mathcal{Q}} \rightarrow \mathbb{R}$.
-Using $\Upsilon^{q}_{\hat{q}}$ we can express $F_{[q]} = \Upsilon^{\hat{q}}_{q} \circ F_{[\hat{q}]} \circ  \Upsilon^{q}_{\hat{q}}$ as an operator over complex functions of $\mathcal{Q}$.
-It will still be linear and self-adjoint.
-
-That is: for each fragment variable $f$ we have a corresponding linear self-adjoint operator $F_{[q^i]}$ that acts on wave functions over $q^i$, which can be expressed as function multiplication in a suitable choice of $\hat{\mathcal{Q}}$.
-The (normalized) expectation of $f$ is given by $\frac{1}{|\psi|^2}\int \psi^\dagger(q^i) F_{[q^i]} (\psi(q^i)) d^nq$.
-
-\begin{prop}\label{prop:self_adjoint_operators}
-	For each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbb{R}$ and a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ that takes a state $\psi \in \Psi$ and returns a wave function $\psi(\hat{q}^i)$ from $\Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ the space of wave functions over $\hat{\mathcal{Q}}$.
-There also exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbb{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i))d^nq $ is the expectation of $f$ multiplied by the amount of material.
-\end{prop}
-\begin{justification}
-	We claim for each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that  $\hat{q}^i : \hat{\mathcal{Q}} \rightarrow \mathbb{R}^n$ provides a maximal description and $f : \hat{\mathcal{Q}} \rightarrow \mathbb{R}$.
-Let $f$ be a fragment variable.
-Let $\hat{q}^i$ be a set of fragment variables such that $f=f(\hat{q}^i)$.
-If $\hat{q}^i$ does not provide a maximal set of fragment variables then there exist other fragment variables that refine the description.
-Extend $\hat{q}^i$ with such variables.
-If still not maximal, continue extending.
-As the number of possibilities charted by $\hat{q}^i$ cannot exceed the number of possibilities charted by $q^i$, at some point we'll reach a maximal set.
-We call $\hat{\mathcal{Q}}$ the topological space charted by the fragment variables $\hat{q}^i$.
-
-	We claim there exists a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$.
-Let $\psi \in \Psi$ be a state.
-$\hat{q}^i$ provides a maximal set of fragment variables.
-There must exist a wave function $\psi(\hat{q}^i) : \mathbb{R}^n \rightarrow \mathbb{C}$ associated with $\psi$ such that it gives the maximal description for the state.
-Let $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \rightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ be the map that associates a state with its wave function over $\hat{\mathcal{Q}}$.
-$\Upsilon_{\hat{\mathcal{q}}^i}$ is a vector space isomorphism for the same reason discussed in \ref{prop:wavefuntion}
-
-	We claim that for each fragment variable $f : \hat{\mathcal{Q}} \rightarrow \mathbb{R}$ there exists a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbb{C})$ such that $\int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]}( \psi(\hat{q}^i)) d^n \hat{q}$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
-The expectation value of $f$ multiplied by the amount of material is given by $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q}$ where $\hat{q}^i$ provide a maximal set of fragment variables such that $f=f(\hat{q}^i)$.
-We have $\int f(\hat{q}^i) \rho(\hat{q}^i) d^n\hat{q} = \int \psi^\dagger(\hat{q}^i) f(\hat{q}^i) \psi(\hat{q}^i) d^{n}\hat{q}$.
-Let $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbb{C})$ such that $F_{[\hat{q}^i]} (\psi(\hat{q}^i)) = f(\hat{q}^i) \psi(\hat{q}^i)$.
-$F_{[\hat{q}^i]}$ is a linear operator.
-$F_{[\hat{q}^i]}$ is self-adjoint as $\int (F_{[\hat{q}^i]}(\psi(\hat{q}^i)))^\dagger \psi(\hat{q}^i) d^{n}\hat{q} = \int (f(\hat{q}^i) \psi(\hat{q}^i))^\dagger \psi(\hat{q}^i) d^{n}\hat{q} = \int \psi^\dagger(\hat{q}^i) f(\hat{q}^i) \psi(\hat{q}^i) d^{n}\hat{q} = \int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]} (\psi(\hat{q}^i)) d^{n}\hat{q}$.
-
-	We claim that for any fragment variable $f$ there exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbb{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i)) d^nq$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
-Let $f$ be a fragment variable.
-There exists a corresponding self-adjoint linear operator $F_{[\hat{q}^i]} : \Psi(\hat{\mathcal{Q}}, d\hat{q}^i) \rightarrow C(\hat{\mathcal{Q}}, \mathbb{C})$ where $\hat{\mathcal{Q}}$ provides a maximal set of fragment variables.
-As both $\hat{\mathcal{Q}}$ and $\mathcal{Q}$ provide a maximal set, we can define $\Upsilon^{q^i}_{\hat{q}^i} \equiv \Upsilon_{\hat{q}^i} \circ \Upsilon_{q^i}^{-1}$ which transforms a wave function over $\mathcal{Q}$ into a wave function over $\hat{\mathcal{Q}}$.
-$\Upsilon^{q^i}_{\hat{q}^i}$ is an isomorphism as $\Upsilon_{\hat{q}^i}$ and $\Upsilon_{q^i}$ are isomorphisms.
-$\Upsilon^{q^i}_{\hat{q}^i}$ is unitary as $\Upsilon^{\hat{q}^i}_{q^i} \circ \Upsilon^{q^i}_{\hat{q}^i} = (\Upsilon^{q^i}_{\hat{q}^i})^{-1} \circ \Upsilon^{q^i}_{\hat{q}^i} = I$.
-We can extend $\Upsilon^{q^i}_{\hat{q}^i}$ on all continuous functions and define $F_{[q^i]} \equiv \Upsilon^{\hat{q}^i}_{q^i} \circ F_{[\hat{q}^i]} \circ \Upsilon^{q^i}_{\hat{q}^i}$.
-$F_{[q^i]}$ is a self-adjoint linear operator.
-$\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i)) d^nq = \int \psi^\dagger(\hat{q}^i) F_{[\hat{q}^i]}( \psi(\hat{q}^i)) d^n \hat{q}$ is, if it exists, the expectation of $f$ multiplied by the amount of material.
-\end{justification}
-
-The last thing to do is characterize the conjugate variables $k_i$ and $p_i$.
-To be precise, we want to find $K_{i [q^i]}$, the operators that act on the wave functions over $q^i$ corresponding to fragment variables $k_i$.
-We know that the operators have to be covariant: under a change of units $\hat{q}^j = \hat{q}^j(q^i)$ we must have $\hat{K}_{j[q^i]} = \partial_{\hat{q}^j} q^i K_{i [q^i]}$.
-In other words: we are looking for the set of linear operators that obey the chain rule.
-This turns out to be the space of derivations, therefore $K_{i [q^i]}$ will be the derivatives in the direction of the corresponding $q^i$.
-They have to be self-adjoint, so they will be of the form $K_{i [q^i]} = a \imath \partial_{i}$ where $a$ is a real number.
-We choose $a$ to be negative by convention and to be unitary so that $k_i$ is expressed in units of inverse $q^i$ as in the classical case.
-Therefore we have $K_{i [q^i]} = -\imath \partial_i$ and $P_{i [q^i]} = -\imath \hbar \partial_i$.
-
-If we assume $\mathcal{Q}$ to be $\mathbb{R}^n$, the functional $\Upsilon^{q^i}_{p_i} : \Psi(\mathcal{Q}, dq^i) \leftrightarrow \Psi(\mathcal{P}, dp_i)$ that converts a wave function in position to a wave function in conjugate momentum is the Fourier transform.
-As the fragment variables for any power of $q^i$ and $p_i$ must exist, the wave functions in both expressions are infinitely differentiable: the space of the wave functions $\Psi(\mathcal{Q}, dq^i) = S(\mathcal{Q}, q^i)$ is the Schwartz space of rapidly decreasing smooth functions.
-
-\begin{prop}\label{prop:momentum_operator}
-	Let $\mathcal{Q}=\mathbb{R}^n$.
-The operator $P_{i [q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbb{C})$ associated with the conjugate quantity of $q^i$ is $P_{i [q^i]} = - \imath \hbar \partial_{i}$.
-The isomorphism $\Upsilon^{q^i}_{p_i} : \Psi(\mathcal{Q}, dq^i) \leftrightarrow \Psi(\mathcal{P}, dp_i)$ that changes the variable of the wave function from $q^i$ to $p_i$ is $\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^n} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
-The space of wave functions $\Psi(\mathcal{Q}, dq^i) = S(\mathcal{Q}, q^i)$ where $S(\mathcal{Q}, q^i)$ is the Schwartz space of rapidly decreasing smooth functions.
-\end{prop}
-\begin{justification}
-	We claim the operator associated with $k_i$ is $K_{i [q^i]} = - \imath \partial_{i}$.
-Let $K_{i [q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, dq^i)$ be the linear operator associated with $k_i$ as expressed on the space of wave functions over $q^i$.
-Such an operator must be covariant under the transformation $\hat{q}^j = \hat{q}^j(q^i)$.
-That is: $\hat{K}_{j [q^i]} = \partial_{\hat{q}^j}q^iK_{i [q^i]}$.
-In particular, it needs to be covariant when changing only one variable: $\hat{q}^1 = \hat{q}^1(q^1)$, $\hat{q}^j = q^j$ where $j=2...n$.
-By \ref{thrm:covariant_operator} $K_{i [q^i]}=c\partial_i$ for some $c \in \mathbb{C}$.
-$K_{i [q^i]}$ is self-adjoint: $K_{i  [q^i]}=\imath r\partial_i$ for some $r \in \mathbb{R}$ by \ref{thrm:antihermitian_derivative}.
-$k_i$ is expressed in units of inverse $q^i$ therefore $|r|=1$.
-By convention, we choose $r$ to be negative.
-$K_{i [q^i]} = - \imath \partial_{i}$.
-
-	We claim the operator associated with $p_i$ is $P_{i [q^i]} = - \imath \hbar \partial_{i}$.
-As $p_i=\hbar k_i$, their expectation is proportional by a factor of $\hbar$.
-$P_{i [q^i]}=\hbar K_{i [q^i]}= -\imath \hbar \partial_i$.
-
-	We claim the operator that changes the variable of the wave function from $q^i$ to $p_i$ is $\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^n} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
-Let $\Upsilon^{q^i}_{p_i} : \Psi(\mathcal{Q}, dq^i) \rightarrow \Psi(\mathcal{P}, dp_i)$ be the operator that maps a wave function over $\mathcal{Q}$ into a wave function over $\mathcal{P}$, the space charted by $p_i$.
-$\Upsilon^{q^i}_{p_i}$ must exist and be unique (up to a total phase) as $\Upsilon^{q^i}_{p_i} = \Upsilon_{p_i}\circ\Upsilon^{q^i} = \Upsilon_{p_i}(\Upsilon_{q^i})^{-1}$ is an isomorphism.
-$\Upsilon^{q^i}_{p_i}$ must be a linear transform that takes the operator $P_{i [q^i]}= - \imath \hbar \partial_i$ as expressed over $q^i$ to $P_{i [p_i]} = p_i$ as expressed over $p_i$.
-Such transformation is the Fourier transform.
-$\Upsilon^{q^i}_{p_i} (\psi(q^i)) = \frac{1}{(\sqrt{2\pi})^n} \int_{\mathcal{Q}} e^{\frac{q^i p_i }{\imath \hbar}} \psi(q^i) d^n q $.
-
-	We claim $\Psi(\mathcal{Q}, dq^i) \subseteq S(\mathcal{Q}, q^i)$.
-Let $\psi(q^i) \in \Psi(\mathcal{Q}, dq^i)$.
-$(P_{i [q^i]})^m \psi(q^i) = (-\imath \hbar \partial_i)^m \psi(q^i)$ represents the $m^{th}$ power of the conjugate momentum operator applied to the given state.
-As the quantity is physically meaningful, such an operation must be defined for any $m$ therefore $\psi(q^i) \in C^\infty(\mathcal{Q})$.
-Let $\psi(p_i) = \Upsilon^{q^i}_{p_i} (\psi(q^i))$ be the wave function over momentum space.
-$(Q^i_{[p_i]})^m \psi(p_i) = (+\imath \hbar \partial_{p_i})^m \psi(p_i)$ represents the $m^{th}$ power of the $q^i$ operator applied to the given state.
-As the quantity is physically meaningful, such an operation must be defined for any $m$ therefore $\psi(p_i) \in C^\infty(\mathcal{P})$.
-As $\psi(q^i)$ is the Fourier transform of an infinitely smooth function, it decreases faster than any inverse power of $q^i$.
-That is: $\forall m \in \mathbb{Z}^+ \exists c \in \mathbb{C}$ such that  $|\psi(q^i)|<\frac{|c|}{|q^i|^m} \; \forall |q^i|> 1$.
-$\Psi(\mathcal{Q}, dq^i) \subseteq S(\mathcal{Q}, q^i)$.
-
-	We claim $\Psi(\mathcal{Q}, dq^i) = S(\mathcal{Q}, q^i)$.
-Let $\psi(q^i) \in \Psi(\mathcal{Q}, dq^i)$ and $\phi(q^i) \in S(\mathcal{Q}, q^i)$ such that $\psi(q^i) \neq \phi(q^i)$.
-There exists a choice of $a, b, j, k \in \mathbb{Z}^+$, $j,k \leq n$ such that $\lVert \psi(q^i) - \phi(q^i) \rVert_{j,a,k,b} \neq 0$  where $\lVert \psi(q^i) \rVert_{j,a,k,b} = sup|(q^j)^a (\partial_k) ^b \psi(q^i)|$ as $S(\mathcal{Q}, q^i)$ is a Frechet space induced by that family of seminorms.
-The operator $(Q^j_{[q^i]})^a(P_{k [q^i]})^b$ applied to $\psi(q^i)$ and $\phi(q^i)$ will provide different distributions for the associated fragment variables.
-$\phi(q^i)$ is physically distinguishable and is associated with a physical state.
-$\phi(q^i) \in \Psi(\mathcal{Q}, dq^i)$.
-\end{justification}
-
-We now have all the basic features of the state space of a quantum particle.
-And we obtained them by re-deriving them from first principles instead of using some formal analogy.
-The similarities with the classical framework are simply due to the similarities of the starting points.
-
-It should be evident, though, how the correspondence between mathematical and physical objects is not as satisfying as in the classical case.
-For instance, while any differentiable function of $T^*\mathcal{Q}$ can be a state variable for a classical particle, not all self-adjoint linear operators can be associated with a fragment variable.
-Consider $f: \mathbb{R} \rightarrow \mathbb{R}$ for which $f(r) = -r$ if $r$ is an integer and $f(r) = r$ otherwise.
-Clearly $f(q)$ does not preserve the topology and cannot be used as a fragment variable.
-Yet, we can still create a self-adjoint linear operator, with the same spectra as the one associated with $q$, just with the eigenvalues switched.
-The space of physically meaningful operators, then, should be further constrained.
-A reasonable requirement would be that they preserve the differentiable/integrable structure of the Schwartz space $S(\mathcal{Q}, q^i)$ as classical state variables preserved the differentiable structure of $T^*\mathcal{Q}$.
-Unfortunately, this is not as easy to justify.
-
-The topology of $\mathcal{P}$, the space charted by conjugate momentum, depends globally on the topology of $\mathcal{Q}$.
-For example, if $\mathcal{Q}$ is an n-dimensional torus, there is no infinity and $\mathcal{P}$ has a discrete topology.
-That is why, in the end, we restricted ourselves to $\mathbb{R}^n$.
-Moreover, we characterized the space of functions not because we required them to fall off at infinity, but because we required the ability to express polynomials of $q^i$ and $p_i$.
-It is not immediately clear how this generalizes over a manifold: further work is needed.
-
-While the link between math and physics may not be as elegant and as general as one would like, we still managed the goal: we identified the states of an irreducible material and saw how these are related to the states in quantum mechanics.
-
-%If the arguments are the same, the projection returns the same state, therefore $<\psi_1|\psi_1> = |\psi_1|^2$: the product is positive definite.
-Since the projection is linear, the product is linear as well.
-
-%The coefficient $\alpha$ is a function of both $\psi_1$ and $\psi_2$, so we write it as $\alpha = <\psi_2|\psi_1>$.
-It is linear in $\psi_1$ as $\mathsf{P}_{\psi_2}$ is linear.
-The square module $|<\psi_2|\psi_1>|^2$ gives us the amount of material in $\psi_1$ that has the same distribution of $\psi_2$.
-$|<\psi_1|\psi_1>|^2$ As all states
-
-\subsection{Irreducibility}
-
-Now that we have fully characterized what we mean by an irreducible material, we can stipulate the following:
-
-\begin{assump}[Irreducibility]\label{ass:irreducibility}
-	The system under study is composed of an irreducible homogeneous material and as a whole undergoes deterministic and reversible evolution.
-\end{assump}
-
-\begin{rationale}
-	The idea is that each amount of material has a state, the physically distinguishable configuration of the distribution, and an unstated part, the physically indistinguishable configuration of the fragments of the material.
-The system is deterministic and reversible in the sense outlined in assumption \ref{ass:determinism}: given the state of the system at one time (i.e.
-the whole distribution) we can predict/reconstruct the future/past states.
-But we cannot describe the evolution of the unstated part (i.e.
-the motion of the fragments).
-
-	As we saw, the state space recovered under this assumption is the one used in quantum mechanics.
-In the following sections we will derive two types of state evolution: the deterministic and reversible one, corresponding to unitary evolution (i.e.
-Schroedinger's equation), and a non-deterministic one useful to describe measurement interactions, corresponding to the projection postulate.
-Therefore, as the infinitesimal reducibility assumption leads to classical mechanics, the irreducibility assumption leads to quantum mechanics.
-
-	As we saw in assumption \ref{ass:infinitesimal_reducibility}, infinitesimal reducibility had its problems. How about the irreducibility assumption? How do they compare?
-
-	The methodological problem assumes a different connotation.
-In order for the assumption to hold, we need to show that there are no processes at our disposal that can probe the fragments of a quantum particle.
-For example: an electron as a whole interacts with a photon as a whole.
-Note that this does not require showing that no such process exists: conceptually specific parts of the photon may be interacting with specific parts of the electron.
-But if those processes are not available to us, say because we can only manipulate whole photons and they always interact with a whole electron, they cannot be used to create a more detailed picture.
-So we simply have to enumerate the processes available to us, which is, at least in principle, feasible.
-Yet, we can never rule out that someone in the future may discover a new process.
-
-	The issues arising from a lack of perfect isolation are addressed.
-Each degree of freedom comes with its own unstated part (i.e.
-the motion of the fragments within that d.o.f.), therefore there is no pretense of a perfect description under perfect isolation.
-Since we cannot describe the nature of the chaotic motion, we cannot say whether it is internally or externally driven.
-
-	The problem of physically meaningless interactions is also considerably lessened.
-Even if we have multiple particles each with their state evolving deterministically, we can still imagine their unstated parts as interacting with each other.
-Therefore each quantum particle would not essentially reside in its own separate universe.
-
-	In these ways, the irreducibility assumption fares much better than the infinitesimal reducibility assumption. Yet, it is still plagued with some of the same problems.
-
-	The problem of incompatibility of measurements with deterministic evolution at all levels still remains.
-The unstated part of multiple irreducible particles may be interacting, but it cannot be used to pass information from one subsystem to another without making itself distinguishable.
-
-	The inability to define time using a deterministic and reversible evolution is still a problem.
-As the system is homogeneous and irreducible, the unstated part cannot give us a measurable quantity that changes in time as that would provide a description of the unstated part.
-
-	There is one additional problem, though, not present in the classical case.
-Under the irreducibility assumption states are necessarily at equilibrium: they all provide the same level of description for the fragments, they carry the same information entropy.
-In fact if the internal dynamics were out of equilibrium, the material would not be homogeneous which would contradict the assumption.
-Deterministic and reversible evolution (which we'll see corresponds to unitary evolution) is necessarily quasi static: at each moment a state must be well defined, at each moment the system is in equilibrium, the system changes slowly.
-Therefore quantum states, as formulated here, cannot and should not be expected to provide a valid description during non-deterministic evolution.
-
-	In light of this, consider a muon decay: the initial state of the muon and the final states of the electron and neutrinos can certainly be well described by quantum particle states.\footnote{In this context, a state is in equilibrium regardless of its being in a local minimum, easily overcome by fluctuations, or a global minimum, that persists indefinitely.} We have no reason to expect, though, that while the muon is decaying into the resulting particles the state is always in equilibrium and therefore well described mathematically by a vector in a complex inner product space.
-After all, the state of a single spin-one-half particle cannot determine the future state of three spin-one-half particles.
-
-	This notion may seem in contradiction to the current practice.
-For example, the S-matrix, one of the main devices for calculating scattering amplitudes and cross sections, is derived by assuming unitary evolution \cite{Weinberg}.
-Doesn't it imply we are assuming that deterministic and reversible evolution is happening at every instant?
-
-	If we look more closely, we realize that the process that the S-matrix captures is a limit where initial states are at $t \rightarrow - \infty$ and final states are at $t \rightarrow + \infty$, one that takes an infinite amount of time.
-That is, to calculate predictions for a fast out of equilibrium process we are approximating it using slow quasi-static evolution.
-This is actually not so uncommon: it is a standard first approximation in thermodynamics.
-The question is why would it work here?
-As we'll see in more detail when talking about the projection postulate, it's because of the irreducibility assumption itself.
-
-	Suppose we are trying to calculate a quantity that does not depend on the particular motion of the unstated part.
-Then, by its very nature, its value will be the same under an equivalent quasi-static evolution, because this too does not depend on the particular motion of the fragments.
-The probability distribution of the out-states during scattering is one such quantity.
-As the final statistical distribution is determined only by the initial state, it is not dependent on the particulars of the unstated part.
-In these cases, the use of deterministic and reversible evolution as a ``stand-in" to calculate final statistical distributions is justified.
-However, it should not be taken literally as a physical model of what actually happens.
-There is no justification for assuming this process to be quasi-static deterministic and reversible evolution of quantum particle states.
-
-	In other words, the assumption of irreducibility already contains the seeds of its own demise.
-On one hand, it tells us that there is a component of chaotic motion within the system.
-On the other hand, it can only characterize the equilibrium dynamics, which clearly can't be the full range of dynamics.
-Yet it also offers a way out: since the motion of the fragments is not accessible, we can predict statistical distributions that are independent of it.
-
-	In light of what we discussed, the irreducibility assumption is less flawed than infinitesimal reducibility but it still cannot be taken as fundamental, in the sense that we cannot take it to strictly apply to all systems.
-We should not expect to solve open problems, such as the arrow of time and the measurement problem, under the assumption as formulated in this work.
-\end{rationale}
-
-\subsection{Quantum mechanics}
-
-We are now ready to write the equations of motion for an irreducible material.
-Deterministic and reversible evolution for the whole distribution necessarily means quasi-static evolution: there is a well defined state at each moment in time.
-Moreover the components must evolve independently as we cannot learn more about one component by observing its evolution in different combinations: the evolution of the composition is the composition of the evolution.
-In other words: the evolution must preserve the inner product.
-
-Note that formally there is a strong analogy with the reducible material case.
-In classical mechanics the state space structure was captured by a distribution over a symplectic manifold.
-Deterministic and reversible evolution was a transformation that preserved that structure, a symplectomorphism.
-In the quantum case the state space structure is captured by an inner product space.
-Deterministic and reversible evolution is a transformation that preserves that structure, a unitary transformation.
-That is: once we have defined what mathematical structure captures the definition of a state, deterministic and reversible evolution must necessarily preserve such structure.
-
-\begin{prop}\label{prop:unitary_evolution}
-	A deterministic and reversible evolution map $\mathcal{T}_{\Delta t}: \Psi \rightarrow \Psi$ for an irreducible material is a unitary operator.
-That is: $\langle \mathcal{T}_{\Delta t} \psi_1, \mathcal{T}_{\Delta t} \psi_2 \rangle = \langle \psi_1 , \psi_2 \rangle$ where $\psi_1, \psi_2 \in \Psi$.
-\end{prop}
-
-\begin{justification}
-	We claim $\mathcal{T}_{\Delta t}: \Psi \rightarrow \Psi$ is a unitary operator.
-Let $\psi_1, \psi_2 \in \Psi$ be two states.
-$\mathcal{T}_{\Delta t}(a\psi_1+b\psi_2) = a \mathcal{T}_{\Delta t} \psi_1 + b \mathcal{T}_{\Delta t} \psi_2$ with $a,b \in \mathbb{C}$ as the composition of the evolution is the evolution of the composition: $\mathcal{T}_{\Delta t}$ is a linear operator.
-$\mathcal{T}_{\Delta t} \mathsf{P}_{\psi_1} (\psi_2) = \mathsf{P}_{\mathcal{T}_{\Delta t}\psi_1} (\mathcal{T}_{\Delta t}\psi_2)$ as the evolved part of $\psi_2$ that was prepared like $\psi_1$ is the part of the evolved $\psi_2$ that ends prepared like the evolved $\psi_1$.
-We have $\mathcal{T}_{\Delta t} \mathsf{P}_{\psi_1} (\psi_2) = \mathcal{T}_{\Delta t} \langle \psi_1 , \psi_2 \rangle \psi_1 = \langle \psi_1 , \psi_2 \rangle \mathcal{T}_{\Delta t} \psi_1 = \mathsf{P}_{\mathcal{T}_{\Delta t}\psi_1} (\mathcal{T}_{\Delta t}\psi_2) = \langle \mathcal{T}_{\Delta t}\psi_1 , \mathcal{T}_{\Delta t}\psi_2 \rangle \mathcal{T}_{\Delta t} \psi_1$.
-$\langle \psi_1 , \psi_2 \rangle = \langle \mathcal{T}_{\Delta t}\psi_1 , \mathcal{T}_{\Delta t}\psi_2 \rangle$.
-$\mathcal{T}_{\Delta t}$ is unitary.
-\end{justification}
-
-\begin{prop}\label{prop:schroedinger_equation}
-	A continuous deterministic and reversible evolution for an irreducible material admits a Hamiltonian operator $H : \Psi \rightarrow \Psi$ that allows us to write the laws of evolution as
-	\begin{align*}
-	\imath \hbar \partial_t \psi = H \psi
-	\end{align*}
-\end{prop}
-
-\begin{justification}
-	We claim the state evolves according to Schroedinger's equation.
-Let  $\mathcal{T}_{dt}: \Psi \rightarrow \Psi$ be the evolution for an infinitesimal time interval $dt$.
-As it is unitary, it can be expressed as $\mathcal{T}_{dt} = I + \frac{H dt}{\imath \hbar}$ where $H : \psi \rightarrow \psi$ is a self-adjoint operator.
-Let $\psi_t \in \Psi$ be a state and $\psi_{t+dt} \in \Psi$ its evolution after an infinitesimal time interval.
-We have $\mathcal{T}_{dt} \psi_t = \psi_{t+dt} = \psi_t + \frac{H dt}{\imath \hbar} \psi_t$.
-$\imath \hbar \frac{\psi_{t+dt} - \psi_t}{dt} = \imath \hbar \partial_t \psi_t = H\psi_t$.
-\end{justification}
-
-We recognize the Schroedinger equation, the time evolution equation for a quantum particle state.
-
-Note, though, that nothing tells us that the system must be in an eigenfunction of $H$ or that the ground state is somehow preferred.
-In fact, the eigenfunctions of $H$ may not be part of $S(\mathcal{Q}, q^i)$ and therefore may not be physical.
-In other words: we have no basis for the time independent Schroedinger equation.
-This is actually consistent with our assumptions.
-For example, for a system to reach its ground state it has to be able to radiate energy: it cannot be isolated.
-Therefore this dynamics cannot be described under the deterministic and reversible assumption.
-This already hints that deterministic and reversible evolution is not the only dynamics possible.
-Other considerations, outside of the scope of this work, can be used to recover this behavior (e.g.
-thermodynamics, decoherence, ...).
-
-Also note that $H$ operates on $\Psi$ which, in the case of $\mathcal{Q}=\mathbb{R}^n$, corresponds to the space $S(\mathcal{Q}, q^i)$ of Schwartz functions.
-$H$ therefore preserves infinite smoothness and integrability.
-This, again, reminds us that the space of all self-adjoint operators is far too vast, and does not map well to physically meaningful quantities.
-
-\subsection{Time dependence and kinematic assumption}
-
-The next step would be to integrate the quantum description with time dependent evolution (i.e.
-relativistic mechanics) and the kinematic assumption (i.e.
-Lagrangian mechanics).
-We will only provide a sketch, though, without going through the details.
-Given that we are limited to single particles (i.e.
-no fields) without spin (i.e.
-minimal material only described by position), our results are limited to finding the Klein-Gordon equation modified with electromagnetic interactions.
-While this is still of note, the equation is of limited physical use and therefore may not warrant the space necessary for a more rigorous derivation.
-
-As in the classical case, the first thing to do is to extend $\mathcal{Q}$ to include time.
-This gives us the same $\mathcal{M}$ as before.
-A quantum state is a complex valued distribution $\psi : \mathcal{M}_{t=t_0} \rightarrow \mathbb{C}$ over a hypersurface at constant time.
-An evolution $\widetilde{\psi} : \mathcal{M} \rightarrow \mathbb{C}$ is a complex valued function defined on the whole $\mathcal{M}$.
-The operator for the quantity conjugate to time is $E=\imath\hbar\partial_t$.
-
-Next we need to extend integration.
-In the quantum case, this means to make sure the inner product $\langle \psi | \phi \rangle$ remains unchanged under all variable changes, including time.
-This allows us to describe the amounts of material and their distribution over fragments in a way that is independent of our description.
-
-As in the classical case, we use $s$ as the evolution parameter.
-Deterministic and reversible evolution $\mathcal{T}_{ds}$ will be a unitary transformation defined over this extended inner product, as the structure of the state space is conserved.
-Continuous evolution will admit an invariant Hamiltonian $\mathcal{H}$ such that $\imath \hbar \partial_s \psi = \mathcal{H} \psi$.
-As for the composite state evolution in the classical case, $\widetilde{\psi}$ remains unchanged by the evolution as it transports the distribution along $s$.
-Therefore $\mathcal{H} \widetilde{\psi} = \mathcal{h} \widetilde{\psi}$ is an eigenfunction of $\mathcal{H}$.
-As in the classical case, we use $\mathcal{H}=0$ to identify the state space: the temporal d.o.f.
-is not independent and therefore there is a constraint among the variables.
-Therefore we have $\mathcal{H} \psi = 0$.
-This is the time dependent evolution equation.
-It means that the wave function does not change along $s$, the parameter of the infinitesimal transformation generated by $\mathcal{H}$.
-
-Note that states are indeed eigenfunctions of the invariant Hamiltonian $\mathcal{H}$: here we do have a justification, as opposed to the standard Hamiltonian $H$ case.
-For example, a massive particle may have fragments with different energy, but they all must carry the same fraction of mass.
-Also note that $\mathcal{H} \psi = 0$ is basically the form of all equations (i.e.
-Klein-Gordon, Dirac, Maxwell, ...) in quantum field theory.
-The respective invariant Hamiltonian $\mathcal{H}$ is the operator associated with each equation.
-
-Having handled the time dependent case, we can turn our attention to the kinematic equivalence assumption.
-While the trajectory of a fragment is not well defined in either phase space or physical space, what we can describe remains the same under the assumption.
-Therefore the distributions still need to be transported from state space to kinematic variables up to a constant factor.
-This means we still have a linear relationship between velocity and conjugate momentum.
-In terms of the quantum operators: $m\, d_s Q^\alpha = g^{\alpha\beta} P_{\beta} - A^\alpha = m\frac{[Q^\alpha, \mathcal{H}]}{\imath \hbar}$.
-
-This leads to the same form for the invariant Hamiltonian $\mathcal{H}=\frac{1}{2m}((P_\alpha-A_\alpha)g^{\alpha\beta}(P_\beta-A_\beta) + m^2 c^2) = \frac{1}{2m}((-\imath\hbar\partial_\alpha-A_\alpha)g^{\alpha\beta}(-\imath\hbar\partial_\beta-A_\beta) + m^2 c^2)$.
-In the free particle case, we have $\frac{1}{2m}(-\hbar^2 \partial^\alpha \partial_\alpha + m^2 c^2) \widetilde{\psi} = 0$ which is the Klein-Gordon equation.
-We can also write the operator for kinetic momentum as $mU^\alpha = P^\alpha - A^\alpha = -i\hbar D^\alpha$ where $D^\alpha = \partial^\alpha + \frac{A^\alpha}{\imath \hbar}$ is the gauge covariant derivative.
-We have $(\frac{1}{2}mU^\alpha U_\alpha + \frac{1}{2} m c^2) \widetilde{\psi} = \frac{1}{2m}(-\hbar^2 D^\alpha D_\alpha + m^2 c^2) \widetilde{\psi} = 0$ which is the Klein-Gordon equation with electromagnetic interaction.
-
-Note how the gauge covariant derivative is, up to a constant, kinetic momentum.
-This motivates why it is such an important operator.
-As for the Klein-Gordon equation, the only thing it does is to impose that the norm of the four-velocity is the speed of light, a condition much more trivial than one may have anticipated.
-
-It should be clear, even without the details, that all the assumptions come together even in the quantum case.
-They give more insight to all the pieces in a comprehensive way and help tear down the artificial walls between the different physical theories.
-This allows us to form a picture of fundamental physics that is more unified, something that has been sorely missing.
-
-\subsection{Non-deterministic evolution and the projection postulate}
-
-As we made it clear from the start, deterministic and reversible evolution is an assumption.
-And as we have seen in the rationale for assumptions \ref{ass:infinitesimal_reducibility} and \ref{ass:irreducibility} we cannot have determinism at all levels of decomposition while transferring information from one system to another.
-
-In this section we will see how non-deterministic processes become necessary when describing measurements for irreducible systems.
-But we'll see that, under the same assumption of irreducibility, we are able to make a connection between the output of deterministic and non-deterministic processes.
-
-Suppose we have a quantum particle state $\psi$ we want to identify.
-Suppose that we have a detector that is able to capture the particle.
-If it's captured, we will know that the particle was in the region of space enclosed by the detector.
-
-Suppose that we create an array of such detectors. This will allow us to detect the position of the particle with a resolution given by the aperture of each detector.
-
-Suppose now, though, that the quantum particle state is spread over a region significantly greater than a single detector but smaller than the array. What will happen?
-
-It is clear that it has to be absorbed as it is within the aperture of the whole array.
-It is also clear that it must be absorbed whole: as it is an irreducible system, the detector cannot partially interact with it, it cannot half absorb it.
-So a single detector will capture it.
-It is also clear that it cannot always be the same detector: that corresponds to states where the particle is spread within the aperture of that detector.
-In other words: as the particle hits the array, one detector triggers non-deterministically based on the distribution of the particle.
-
-The process is non-deterministic in the sense that the state of the particle is not enough to predict which detector will trigger.
-Maybe it depends on the state or unstated part of the detector, maybe it depends on the unstated part of the particle or maybe on the environment.
-The point is that a single measurement does not tell us much about the state of the incoming particle.
-
-We then repeat the process, making sure the incoming particle is prepared in the same way.
-Given a single ``pure" incoming particle state, we get a ``mixed" statistical distribution describing the output of the array.\footnote{Note that this is different from the model used often to describe the quantum measurement problem, where the incoming system becomes entangled with the measuring device.
-We do not make any claim on what the final state of the particle should be: in many cases, like in a photomultiplier tube, the particle is gone and there is no state.
-We simply have one incoming state for the particle, prepared over and over exactly the same, and the outgoing state for the measurement device, which is a classical statistical distribution over the different measurement takes.} The question is, how is the statistical distribution coming out of the detector related to the material distribution coming in with the particle?
-
-If the detector does not introduce biases, then the final distribution is only a function of the incoming particle: all unbiased detectors will give the same result.
-The unstated part also cannot affect the final distribution of outcomes: it is not physically distinguishable.
-The final statistical distribution of the detector must match the incoming material distribution of the particle.
-That is if $N$ is the total number of takes and $N_A$ is the number of times the detector with aperture in the $A$ region captured the particle, the ratio $\frac{N_A}{N}$ will be approximately the amount of material within the aperture $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq$.
-If we assume an infinite number of takes, we expect them to coincide: $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq = \lim\limits_{N\rightarrow\infty} \frac{N_A}{N}$.
-But this is also the frequentist definition of probability.
-Therefore we have $\int_A \psi^\dagger(q^i) \psi(q^i) d^nq = P(\textrm{measuring particle in } A)$ thus recovering the probabilistic interpretation of the wave function.
-
-We can conceptually extend this picture to the generic case.
-Suppose we know the outcomes of our measurement are parametrized by a fragment variable $\hat{q}$.
-We will have a linear functional $\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0} : \Psi \rightarrow \mathbb{C}$ such that $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2$ returns the amount of material prepared with the particular value $\hat{q}_0$.
-We can re-express this quantity in terms of the dual vector $\hat{q}_0$ corresponding to the linear functional $\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}$ and the inner product extended to all Lebesgue integrable functions: $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2 = |\langle \hat{q}_0 | \psi \rangle|^2 $.
-This is also the probability that the outcome associated with $\hat{q}_0$ is selected by the measuring device, thus recovering the statistical nature of the quantum predictions.
-
-Note that while this process is of a non-deterministic nature, it is of a specific non-deterministic nature.
-It requires that no other state can influence the final distribution as the match between the physical material distribution and the statistical measurement distribution was recovered only under the assumption that the detector does not introduce any bias.
-That is: the non-deterministic nature cannot be understood as an averaging over some unknown external state.
-
-Also note that while we have a way to predict the distribution over final outcomes from the initial state, we have no way of describing what happens in between.
-When we assume that only the initial state matters, we are saying that we can disregard the particulars of the non-deterministic evolution.
-So much so that we can substitute it with a deterministic one to compute the probability distribution.
-In particular, given only the initial state, we have no way of predicting which set of final outcomes the system will project to as that depends on the actual choice of detector setup, nor when that projection actually happens.
-
-While this may seem like an incomplete description, it is as complete as the irreducibility assumption allows.
-The state can be known and determines any final statistical distribution.
-The unstated part cannot be known, and is what gives the statistical distribution in the first place.
-If we were able to say something more about the evolution of the unstated part, we would be able to further describe it, which we cannot do without violating the assumption.
-If we cannot learn anything about the unstated part from the final distribution, then we are consistent with our assumption and we are able to use deterministic evolution and projection to determine the final statistical distribution.
-In other words: the inability to describe the unstated part not only constrains the state space and deterministic evolution, but also the set of non-deterministic processes by excluding those for which the final statistical distribution would depend on the unstated part itself.
-
-We therefore have arrived at all the major underpinnings of quantum mechanics and see how they stem from assumption \ref{ass:irreducibility} of irreducibility.
-This gives us a more satisfactory account of what quantum mechanics describes, although it does not claim to solve many of the open problems.
-It does, however, give good reasons as to why we have those problems and how the desire to know ``what is really going on" is hindered by the very assumption that the quantum description is based on.
-
-\section{Discussion and further work}
-\label{sec:discussion}
-
-As we went through many details in the derivation, it is time to step back and conclude with a few general remarks.
-
-The first remarkable aspect of this work is that something along these lines can be done.
-Regardless of whether all the minor details are addressed in the best possible way, the idea that so much can be derived from so little was not something we expected.
-In hindsight, the idea that physics should be founded on physical principles is nothing new: it just hasn't been thoroughly applied to basic mechanics.
-If this work had been done incrementally from the start, these concepts would have been developed gradually and the results would feel more mundane.
-
-The second striking aspect is a shift in perspective that this work seems to suggest.
-The usual reductionist view is that you start with the rules of the small parts and then build up the composite whole.
-This work flips that perspective: it's not what the parts can tell you about the whole, it's what the whole can tell you about the parts.
-
-The state space for classical particles is the one that allows us to express invariant distributions: that is a requirement that comes from the whole.
-The fragments of quantum particles do not even have a well defined description: the whole tells us nothing about their configuration.
-In retrospect, this should not be surprising.
-Nobody is going to hand us down the laws of infinitesimal parts: we induce  what we can by manipulating finite systems.
-Therefore what the whole allows us to infer about its parts is physically fundamental.
-
-The third notable aspect is that the assumptions have more to do with our ability to create a reliable description rather than essential features of the system itself.
-A system is not deterministic per se: it is a particular choice for its description (i.e.
-state) that may undergo deterministic evolution in some cases.
-In the same vein, a material is not homogeneous per se.
-Or infinitesimally reducible.
-In this context, the state, its properties and the laws of evolution are the description of the system under our idealization, not some intrinsic feature of the system.
-
-The fourth intriguing aspect is that, when framing classical and quantum mechanics in this way, there is a progression.
-First we assume a homogeneous infinitesimally reducible material.
-Then we assume a homogeneous irreducible material.
-Extrapolating from this progression, the next assumption would be an inhomogeneous irreducible material: where the material is not in constant equilibrium and its particles do not all look the same.
-
-This different starting point may have potential connections to some of the problems we mentioned in the rationales for the assumptions and to other interesting physics questions in general.
-In the foreseeable future, however, we will work on the more mundane task of extending this work to field theories, for which a more concrete path is available.
-
-\section{Conclusion}
-
-We have seen that much of fundamental physics can be derived from few simple physical assumptions.
-In doing so we obtained a more unified picture of the different branches of fundamental physics centered around the idea of deterministic and reversible evolution of distributions of homogeneous materials.
-Though the work is extensive we can summarize a few important conclusions with the following points:
-
-\begin{itemize}
-
-\item A state space is not just a collection of states for the system under study but it must also capture physical properties such as physical distinguishability (represented by a topology), the ability to count states (represented by a measure, symplectic form or metric tensor) or decomposability (represented by a vector space structure).
-
-\item Deterministic and reversible evolution is a one-to-one map that preserves the nature of the system (e.g.
-physical distinguishability, state count and decomposability).
-Therefore it is an isomorphism on the mathematical structure defined by the state space (e.g.
-homeomorphism, symplectomorphism, unitary transformation), not just a state-to-state bijection (i.e.
-isomorphism of sets).
-
-\item Classical Hamiltonian mechanics describes the evolution of a homogeneous material whose state is equivalent to the states of all its infinitesimal parts, the classical particles.
-The state space for classical particles is a symplectic manifold $(T^*\mathcal{Q}, \omega)$ as this allows us to define state-variable-invariant density distributions; deterministic and reversible evolution for such a state space is a symplectomorphism (i.e.
-a canonical transformation).
-
-\item Lagrangian mechanics describes the case where the trajectory in space fully characterizes the dynamics of the system.
-If we want distributions in phase space to be expressed in terms of kinematic variables (i.e.
-position and velocity), we find that $(\mathcal{Q}, g)$ is a Riemannian manifold and the motion is that of a massive particle under scalar/vector potential forces.
-
-\item Relativistic motion arises by properly handling time dependent deterministic and reversible evolution for the material (i.e.
-distributions over phase space) without further assumptions.
-That is, we derive the geometrical structure of space-time from the geometrical structure of phase space, inverting the typical approach.
-
-\item Quantum (Hamiltonian) mechanics describes the evolution of a homogeneous material for which the state can only be reduced up to finite parts, the quantum particles, whose internal motion is undetermined.
-The state space for quantum particles is a complex inner product space $(\Psi, \langle \cdot , \cdot \rangle)$; deterministic and reversible evolution for such a state space is a unitary transformation.
-\end{itemize}
-
-While we are sure that some mathematical details can be improved, we feel that the overall ``conceptual castle" is solid.
-The starting physical assumptions are simple, the translation into math is often straightforward and the intuitive ideas that emerge help unveil connections within and across different branches of math and physics.
-It is the tight fitting of these different pieces into a broader picture that validates the merit of the approach.
-
-While the scope of this work is limited to refounding and reunderstanding physical theories already experimentally established, it would not be unexpected for a better grasp of old theories to lead to insights and ideas for new ones.
-
-%We also have a more unified picture of fundamental physics which, after all ideas are digested, helps the intuition to gain insight that span across the different field.
-It also illustrate a new approach in fundamental physics: one where we better clarify our starting physical assumption and definition, where physics comes first, and they are later captured with mathematical definitions.
-
-\begin{acknowledgements}
-We'd like to thank Erik Curiel, Stephen DiIorio, Jiahua Gu, Josh Hunt, Isaac Mooney and Pavel Okun for providing feedback on drafts of this manuscript, Mark Greenfield for further insights in the topological aspects and Alan Phillips for contributing figures.
-We additionally thank Paul Stankus for interest and support in the earlier stages of this project.
-Funding for this work was provided in part by the MCubed program of the University of Michigan.
-\end{acknowledgements}
-
-\appendix
-\section*{Appendix A. Mathematical proofs}
-\stepcounter{section}
-
-In this section we include mathematical demonstrations in support of the physical justifications of the paper.
-
-\begin{thrm}[Extended Riesz theorem]\label{extended_riesz_theorem}
-	Let $(S, \tau)$ be a locally compact Hausdorff space.
-Let $\Lambda = \{\Lambda_U : C(S) \rightarrow \mathbb{R}\}_{U \subseteq S}$ a family of positive linear functionals such that $\forall U \subseteq S \; \Lambda_U = \Lambda_{\mathrm{int}(U)}$ and $\Lambda_{U_1} + \Lambda_{U_2} = \Lambda_{U_1 \cup U_2} + \Lambda_{U_1 \cap U_2} \; \forall U_1, U_2 \subseteq S$.
-Then there exists a unique Borel measure $\mu$ such that $\Lambda_U (\rho) = \int_{U} \rho d\mu \; \forall \rho \in C(S)$.
-\end{thrm}
-
-\begin{proof}
-	We claim that at each $s \in S$ there exists a compact neighborhood $U \subseteq S$ endowed with a Borel measure $\mu_U$ such that $\Lambda_U (\rho) = \int_U \rho d \mu_U \; \forall \rho \in C(S)$.
-$S$ is locally compact.
-$\forall s \in S \; \exists U \subseteq S$ such that $U$ is compact.
-$U$ with the subspace topology is a compact Hausdorff topological space on which is defined a positive linear functional $\Lambda_U : C(S) \rightarrow \mathbb{R}$.
-For the Riesz representation theorem for linear functionals, there exists a unique Borel measure $\mu_U$ such that $\Lambda_U (\mathcal{c}) = \int_U \rho d \mu_U \; \forall \rho \in C(S)$.
-
-	We claim that $S$ is endowed with a unique Borel measure $\mu$ such that $\Lambda_U (\rho) = \int_U \rho d \mu$.
-Let $U \subseteq V \subseteq \mathcal{S}$ two compact subsets.
-Let $\rho \in C(S)$.
-$\Lambda_V (\rho) = \int_V \rho d \mu_V = \int_U \rho d \mu_V + \int_{V\backslash U} \rho d \mu_V$.
-Also $\Lambda_V (\rho) = \Lambda_U (\rho) + \Lambda_{V\backslash U} (\rho) = \int_U \rho d \mu_U + \int_{V \backslash U} \rho d \mu_{\overline{V \backslash U}} $.
-Therefore $\int_{U} \rho_{\mathcal{c}} d \mu_U = \int_{U} \rho_{\mathcal{c}} d \mu_V$.
-$d\mu$ is unique and does not depend on the choice of neighborhood.
-$S$ is locally compact.
-It admits a cover $\{U_\alpha\}_{\alpha \in A}$ where each $U_\alpha$ is compact.
-	\begin{align*}
-	\Lambda_{\mathcal{S}}(\mathcal{c}) &= \sum \limits_{\alpha \in A} \Lambda_{U_\alpha}(\mathcal{c}) - \sum \limits_{\alpha \in A} \sum \limits_{\beta \in A}^{\beta \neq\alpha} \Lambda_{U_\alpha \cap U_\beta}(\mathcal{c}) \\
-	&= \sum \limits_{\alpha \in A} \int_{U_\alpha} \rho_{\mathcal{c}} d\mu - \sum \limits_{\alpha \in A} \sum \limits_{\beta \in A}^{\beta \neq\alpha} \int_{U_\alpha \cap U_\beta}(\mathcal{c}) \rho_{\mathcal{c}} d\mu \\
-	&= \int_{\mathcal{S}} \rho_{\mathcal{c}} d\mu
-	\end{align*}
-	The measure is unique on the whole space.
-\end{proof}
-
-\begin{thrm}\label{everywhere_integrable_is_lebesgue_integrable}
-	Let $\mathcal{L}(S,\mu) \equiv \{ \rho : S \rightarrow \mathbb{R} \; | \;\; |\int_{U} \rho d\mu| < \infty \; \forall U \subseteq S\}$.
-$\mathcal{L}(S,\mu)=L^1(S,\mu)$ where $L^1(S,\mu) = \{ \rho : S \rightarrow \mathbb{R} \; | \;\; \int_{S} |\rho| d\mu < \infty \}$
-\end{thrm}
-
-\begin{proof}
-	We claim $\mathcal{L}(S,\mu) \supseteq L^1(S,\mu)$.
-Let $\rho \in L^1(S,\mu)$.
-Let $U \subseteq S$.
-$\lVert \rho \rVert_U \equiv \int_{U} |\rho| d\mu < \int_{\mathcal{S}} |\rho| d\mu < \infty$.
-$\lVert \rho \rVert_U = \lVert \rho^+ \rVert_U + \lVert \rho^- \rVert_U$.
-$\lVert \rho^+ \rVert_U < \infty$ and $\lVert \rho^- \rVert_U < \infty$.
-$|\int_{U} \rho d\mu| = |\int_{U} \rho^+ d\mu - \int_{U} \rho^- d\mu| = |\int_{U} \rho^+ d\mu| - |\int_{U} \rho^- d\mu| = \lVert \rho^+ \rVert_U - \lVert \rho^- \rVert_U < \infty$.
-$\rho \in\mathcal{L}(S,\mu)$.
-
-	We claim $L^1(S,\mu) \supseteq \mathcal{L}(S,\mu)$.
-Let $\rho \in \mathcal{L}(S,\mu)$.
-Let $S^+ \equiv  \{ s \in S \; | \; \rho(s) > 0\}$.
-Let $S^- \equiv  \{ s \in S \; | \; \rho(s) < 0\}$.
-$|\int_{S^+} \rho d\mu| = |\int_{S} \rho^+ d\mu| = \lVert \rho^+ \rVert < \infty$ and $|\int_{S^-} \rho d\mu| = |\int_{S} \rho^- d\mu| = \lVert \rho^- \rVert < \infty$.
-$\lVert \rho \rVert = \lVert \rho^+ \rVert + \lVert \rho^- \rVert < \infty$.
-$\rho \in L^1(S,\mu)$.
-
-	We claim $L^1(S,\mu) = \mathcal{L}(S,\mu)$. $\mathcal{L}(S,\mu) \supseteq L^1(S,\mu)$ and $L^1(S,\mu) \supseteq \mathcal{L}(S,\mu)$.	
-\end{proof}
-
-\begin{thrm}\label{symplectomorphism_generator}
-	Let $(M, \omega)$ a symplectic manifold.
-Let $f: (M, \omega) \rightarrow (M, \omega)$ an infinitesimal self-symplectomorphism.
-Let $S \in TM \; | \; f(\xi^a(m)) = \xi^a(m) + S^a(m)dt \; \forall m \in M$ be the infinitesimal displacement.
-There exists a function $H: M \rightarrow \mathbb{R}$ such that $S^{a} \omega_{ab} = \partial_{b}H$ and $H \in C^2(M, \mathbb{R})$.
-\end{thrm}
-
-\begin{proof}
-	We claim the vector field $S \in T\mathcal{Q}$ admits a potential $H$ such that $S^{a} \omega_{ab} = \partial_{b}H$.
-Let $v, w \in T_m M$ be two vectors defined at a point $m \in M$.
-Let $v^a, w^b$ be their components.
-Let $v'\equiv f v, w' \in T_{f(m)}M$ be the pushforward of $v, w$ by $f$.
-Since $f$ is a symplectomorphism we have $v^{a} \omega_{ab} w^{b} = v'^{a} \omega_{ab} w'^{b}$.
-The vector components change according to $v'^a = \partial_b \xi(t+dt)^a v^b = (\delta^a_b + \partial_b S^a dt) v^b$.
-We have:
-	\begin{align*}
-	v^{a} \omega_{ab} w^{b} &= v'^{a} \omega_{ab} w'^{b}  \\
-	&= (v^{a} + \partial_{c} S^{a} v^{c} dt) \omega_{ab} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
-	&= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} + v^{a} \omega_{ab} \partial_{d} S^{b} w^{d}) dt \\ &+ O(dt^2)
-	\end{align*}
-	$v^{c} w^{b} \partial_{c} S_{b} - v^{a} w^{d} \partial_{d} S_{a} = 0$ where $S_{b} \equiv S^{a} \omega_{ab}$.
-The relationship must be true for and pair of vector, therefore $\partial_{a} S_{b} - \partial_{b} S_{a} = curl(S_{a}) = 0$.
-$S$ is a curl free vector field and admits a potential $H$ such that $S_{a} = \partial_{a}H$.
-
-	We claim $H \in C^2(M, \mathbb{R})$.
-$S$ is differentiable as it is the displacement of a symplectomorphism.
-The derivatives of $H$ are differentiable as $S_{a} = \partial_{a}H$.
-$H$ is twice differentiable.
-\end{proof}
-
-\begin{thrm}\label{thrm:inner_product}
-	Let $V$ be a complex vector space.
-Let $| \cdot | ^2 : V \rightarrow \mathbb{R}$ such that $|x|^2 \ge 0 \; \forall x \in V$ and $|x|^2 = 0 \Leftrightarrow x = 0$.
-Let $x \in V$ and $\mathsf{P}_x : V \rightarrow V$ be a family of projections (i.e.
-$\mathsf{P}_x$ linear and $\mathsf{P}_x \circ \mathsf{P}_x = \mathsf{P}_x$) such that: $\mathsf{P}_x (y) = y \Leftrightarrow y = a x$ for some $a \in \mathbb{C}$; $|x|^2|\mathsf{P}_x (y)|^2=|y|^2|\mathsf{P}_y (x)|^2$; $\mathsf{P}_x (\mathsf{P}_y (x)) = r x$ for some $r \in \mathbb{R}$.
-	Then $V$ is an inner product space with the product $\langle \cdot , \cdot \rangle : V \times V \rightarrow \mathbb{C}$ defined by $\langle x , y \rangle x = |x|^2 \mathsf{P}_x (y)$.
-\end{thrm}
-
-\begin{proof}
-	We claim $\langle \cdot , \cdot \rangle$ is well defined.
-Let $x, y \in V$.
-$\mathsf{P}_x (\mathsf{P}_x (y)) =\mathsf{P}_x(y)$ as $\mathsf{P}_x$ is a projection.
-$\mathsf{P}_x(y) = a x $ for some $a \in \mathbb{C}$ by the first property of $\mathsf{P}_x$.
-$\langle x , y \rangle = |x|^2 a$ exists and is unique.
-
-	We claim $\langle \cdot , \cdot \rangle$ is positive definite.
-Let $x \in V$.
-$\langle x , x \rangle x = |x|^2 \mathsf{P}_x(x) = |x|^2 x $.
-Therefore $\langle x , x \rangle = |x|^2 \ge 0$ and $\langle x , x \rangle = |x|^2 = 0 \Leftrightarrow x = 0$.
-
-	We claim $\langle \cdot , \cdot \rangle$ is linear in the second argument.
-Let $x,y,z \in V$ and $a,b \in \mathbb{C}$.
-$\langle x , a y + b z \rangle x = |x|^2 \mathsf{P}_x(a y + b z)=a  |x|^2 \mathsf{P}_x(y) + b  |x|^2 \mathsf{P}_x(z) = a \langle x , y \rangle x + b \langle x, z \rangle x$ by linearity of $\mathsf{P}_x$.
-Therefore $\langle x , a y + b z \rangle = a \langle x , y \rangle + b \langle x, z \rangle$.
-
-	We claim $\langle \cdot , \cdot \rangle$ is conjugate symmetric.
-Let $x,y \in V$.
-$ | |x|^2\mathsf{P}_x(y)|^2 = |x|^4|\mathsf{P}_x(y)|^2 = |\langle x , y \rangle|^2 |x|^2$.
-Using the second property of the projection we have $|\langle x , y \rangle|^2 = |x|^2|\mathsf{P}_x(y)|^2 =  |y|^2|\mathsf{P}_y(x)|^2 = |\langle y , x \rangle|^2$.
-The modulus of the product is symmetric as a function of the arguments.
-Now consider $|x|^2|y|^2\mathsf{P}_x(\mathsf{P}_y(x)) = |x|^2 \langle y , x \rangle \mathsf{P}_x(y) = \langle y , x \rangle \langle x , y \rangle x$.
-The product $\langle x , y \rangle \langle y , x \rangle \in \mathbb{R}$ by the third property of the projection.
-$\arg(\langle x , y \rangle) + \arg(\langle y , x \rangle)$ is $0$ or $\pi$.
-$\pi$ is excluded since $\langle x , x \rangle = |x|^2 \ge 0$.
-$\langle y , x \rangle$ and $\langle x , y \rangle$ have equal modulus and opposite phase.
-$\langle x , y \rangle = \langle y , x \rangle^\dagger$
-\end{proof}
-
-\begin{thrm}\label{thrm:covariant_operator}
-	Let $C^1(\mathcal{Q}, \mathbb{C})$ be the space of differentiable complex functions defined over $\mathcal{Q}$.
-Let $q : \mathcal{Q} \rightarrow \mathbb{R}$ a differentiable function.
-Let $K_q : C^1(\mathcal{Q}, \mathbb{C}) \rightarrow C^1(\mathcal{Q}, \mathbb{C})$ be a covariant linear operator, that is given an invertible differentiable function $\hat{q} : \mathbb{R} \rightarrow \mathbb{R}$ $K_{\hat{q}} = \partial_{\hat{q}} q K_{q}$.
-Then $K_q = a \partial_q$ for some $a \in \mathbb{C}$.
-\end{thrm}
-
-\begin{proof}
-	We claim $K_q(f(q))= \partial_q f K_q(q)$.
-Let $I : \mathbb{R} \rightarrow \mathbb{R}$ be the identify function, that is $I \circ f = f \; \forall f : \mathbb{R} \rightarrow \mathbb{R}$.
-$K_q(q) = K_q(I(q)) = \partial_q I(q) K_{I}(I) = K_{I}(I)$ by covariance.
-	Let $f : \mathbb{R} \rightarrow \mathbb{R}$ be a differentiable invertible function.
-$K_q(f(q)) =K_q(I(f(q))) = \partial_q f K_{f}(I(f)) = \partial_q f \partial_f I(f) K_{I}(I) = \partial_q f K_q(q)$.
-Now let $f : \mathbb{R} \rightarrow \mathbb{R}$ be a differentiable function non necessarily invertible.
-It can be written as $f = f_1 + f_2$ with $f_1, f_2$ differentiable and invertible.
-By linearity $K_q(f)=K_q(f_1) + K_q(f_2) = \partial_q f_1 K_q(q) + \partial_q f_2 K_q(q)  = \partial_q f K_q(q)$.
-
-	We claim $K_q$ is a derivation.
-Let $f,g,h$ be differentiable functions such that $h(q)=f(q)g(q)$.
-$K_q(ln(h(q)))=\partial_q ln(h(q)) K_q(q) = \frac{1}{h} \partial_q h K_q(q) = \frac{1}{h} K_q(h(q)) = K_q(ln(f(q)g(q))) = K_q(ln(f(q))) + K_q(ln(g(q))) = \frac{1}{f} K_q(f(q)) + \frac{1}{g} K_q(g(q))$.
-Multiplying by $h=fg$ we have $K_q(h(q)) = g K_q(f(q)) + f K_q(g(q))$.
-$K_q$ is a linear operator that satisfies the product rule.
-$K_q$ is a derivation by definition.
-
-	We claim $K_q = a \partial_q$ for some $a \in \mathbb{C}$.
-$K_q$ is a derivation.
-As $K_q(f(q))= \partial_q f K_q(q)$ it is a derivation along $q$, which can be expressed as $K_q(q)=a \partial_q$ for some $a \in \mathbb{C}$.
-\end{proof}
-
-\begin{thrm}\label{thrm:antihermitian_derivative}
-	Let $G \equiv C^1(\mathcal{Q}, \mathbb{C}) \cap L^1(\mathcal{Q}, \mathbb{C})$ be the space of differentiable and integrable complex functions defined over $\mathcal{Q}$.
-Let $K$ be a linear adjoint operator of the form $a \partial_q$ with $a \in \mathbb{C}$.
-Let $K$ be self-adjoint with respect to the inner product $\langle f, g \rangle =
-	\int_{\mathcal{Q}} f^\dagger(q) g(q) dq$. Then $K= \imath r \partial_q$ with $r \in \mathbb{R}$.
-\end{thrm}
-
-\begin{proof}
-	We claim $K= \imath r \partial_q$ with $r \in \mathbb{R}$.
-Let $f,g \in G$.
-We have $\langle f, K g \rangle = \int_{\mathcal{Q}}f^\dagger a \partial_q g dq = f^\dagger g |_{\partial \mathcal{Q}}  - \int_{\mathcal{Q}} a \partial_qf^\dagger  g dq= \int_{\mathcal{Q}} (- a\partial_q f^\dagger) g dq = \int_{\mathcal{Q}} (- a^\dagger\partial_q f)^\dagger g dq = \langle - a^\dagger\partial_q f, g \rangle$ using integration by part and integrability of $f$ and $g$.
-$\langle - a^\dagger\partial_q f, g \rangle =\langle K f, g \rangle$ as $K$ is self-adjoint.
-$- a^\dagger\partial_q = a \partial_q$.
-$a = - a^\dagger$.
-$a = \imath r$ with $r \in \mathbb{R}$.
-\end{proof}
-
-\begin{thebibliography}{0}
-\bibitem{Roland}  
-Thomas, R. E., Rosa, A. J.: The analysis and design of linear circuits. Wiley (2001)
-
-\bibitem{Barron} Barron, R. F.: 8.2 Lumped-parameter analysis. In: Barron, R. F.: Industrial Noise Control and Acoustics. CRC Press (2002)
-%online at http://ssu.ac.ir/cms/fileadmin/user_upload/Daneshkadaha/dbehdasht/behdasht_herfei/e_book/_Randall_F._Barons_Industrial_Noise_Control_and_.org__1_.pdf
-
-\bibitem{Borutzky} Borutzky, W.: Bond graph methodology: development and analysis of multidisciplinary dynamic system models.
-Springer (2010)
-
-\bibitem{Freidin} Freidin, R., Lasnik, H.: Syntax: critical concepts in linguistics. Routledge (2006)
-
-\bibitem{Friedman} Friedman, D. P., Wand, M., Haynes, C.T.: Essentials of Programming Languages. MIT Press (2001)
-
-\bibitem{LeeTM} Lee, J. M.: Introduction to Topological Manifolds. Springer (2000)
-
-\bibitem{Halmos} Halmos, P: Measure Theory. D. van Nostrand and Co. (1950)
-
-\bibitem{Lee} Lee, J. M.: Introduction to Smooth Manifolds. Springer (2003)
-
-\bibitem{Lang} Lang, S. L.: Algebra. Springer (2002)
-\bibitem{Young} Young, N.: An Introduction to Hilbert Space. Cambridge Univ. Press, Cambridge, UK (1988)
-
-\bibitem{Grimmet} Grimmett, G. R., Stirzaker, D. R.: Probability and random processes. Oxford Univ. Press 3. ed. (2001)
-
-\bibitem{Pierce} Pierce, J. R.: An Introduction to Information Theory: Symbols, Signals and Noise. Dover Publications (1980)
-
-\bibitem{classical_dynamics} Jos\'{e}, J. V., Saletan, E. J.: Classical Dynamics. Cambridge University Press, (1998)
-
-\bibitem{Landau} Landau, L.D., Lifshitz, E.M.: Quantum Mechanics: Non-Relativistic Theory. Vol. 3 (3rd ed.). Pergamon Press (1977)
-
-\bibitem{Weinberg} Weinberg, S.: The Quantum Theory of Fields, vol I. Cambridge University Press (2002)
-
-\bibitem{Van Ness} Van Ness, H. C.: Understanding Thermodynamics. Dover Publications (1983)
-
-\bibitem{French} French, A. P.: Special Relativity. W. W. Norton \& Company (1968)
-
-\bibitem{Katok} Katok, A., Hasselblatt, B.: Introduction to the modern theory of dynamical systems. Cambridge (1996)
-
-\bibitem{Sontang} Sontag, E. D.: Mathematical Control Theory: Deterministic Finite Dimensional Systems (2nd ed.). Springer (1999)
-
-\bibitem{Shannon} Shannon, C. E.: A mathematical theory of communication. The Bell System Technical Journal, Vol. 27, pp. 379--423, 623--656, (1948)
-
-\bibitem{Jaynes} Jaynes, E. T.: Information theory and statistical mechanics. Statistical Physics 3, pp. 181--218, (1963)
-
-\bibitem{Ruse} Ruse, M.: Entry for ``Reductionism". The Oxford Companion to Philosophy (2nd ed.). Oxford University Press. p. 793 (2005)
-
-\bibitem{Linares} Galindo-Linares, E., Navarro-Morale, E., Silva-Ortigoza, G., Suárez-Xique, R., Marciano-Melchor, M., Silva-Ortigoza, R., Román-Hernández, E.: Any Hamiltonian System Is Locally Equivalent to a Free Particle.
-World Journal of Mechanics, 2012, 2, pp.
-246--252 (2012)
-
-\bibitem{BEC} Pitaevskii, L. P., Stringari, S.: Bose–-Einstein Condensation. Clarendon Press, Oxford (2003)
-
-\bibitem{Gromov} Gromov, M. L.: Pseudo holomorphic curves in symplectic manifolds. Inventiones Mathematicae 82, pp. 307--347, (1985)
-
-\bibitem{deGosson} de Gosson, M. A.: The symplectic camel and the uncertainty principle: the tip of an iceberg?. Foundations of Physics 39, pp. 194--214, (2009)
-
-\bibitem{Stewart} Stewart, I.: The symplectic camel, Nature 329, pp. 17--18, (1987)
-
-\bibitem{Grant} Grant, C. P.: Theory of Ordinary Differential Equations. CreateSpace Independent Publishing Platform (2014)
-
-%\bibitem{Treumann} Treumann, R. A., Nakamura, R., Baumjohann,  W.: Relativistic transformation of phase-space distributions. Annales Geophysicae (2011)
-
-\bibitem{Gourgoulhon} Gourgoulhon, E.: 3+1 Formalism and Bases of Numerical Relativity. arXiv. http://arxiv.org/abs/gr-qc/0703035 (2007)
-
-\bibitem{Synge} Synge, J. L.: Classical dynamics. Encyclopedia of Physics Vol 3/1, Springer (1960)
-
-\bibitem{Lanczos} Lanczos, C.: The variational principles of mechanics, University of Toronto Press, (1949)
-
-\bibitem{Struckmeier} Struckmeier, J.: Hamiltonian dynamics on the symplectic extended phase space for autonomous and non-autonomous systems. J. Phys. A: Math. Gen. 38, pp. 1257--1278, (2005)
-
-\bibitem{EPR} Einstein, A., Podolsky, B.,  Rosen, N.: Can quantum-mechanical description of physical reality be considered complete? Phys. Rev. 47 777 (1935)
-
-\bibitem{Von Neumann} Von Neumann, J.: Mathematical Foundations of Quantum Mechanics. Princeton University Press (1955)
-
-\bibitem{Bohr} Bohr, N.: The Quantum Postulate and the Recent Development of Atomic Theory. Nature 121, pp. 580-–590 (1928)
 \end{thebibliography}

### 2017-02-16 Notes/GalileanLorentzTransformationInPhaseSpace.tex
Adding notes on other issues

\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Notes for Galilean and Lorentz transformation in phase space}

\subsection{Standard phase-space}

Change of coordinates associated with Galilean boost.

\begin{align*}
\hat{q} &= q + vt \\
q &= \hat{q} - vt \\
d\hat{q} &= dq + v dt \\
\end{align*}

Note that $d\hat{q}$ cannot be properly expressed just in terms of $dq$ and $dp$ as $t$ is not a function of $q$ and $p$. We (improperly) treat $t$ as a constant, define $d\hat{q}$ at constant $t$ and have $d\hat{q} = dq$.

Momentum transforms like a covector under coordinate transformations.

\begin{align*}
\hat{p} &= \frac{\partial q}{\partial \hat{q}} p = p \\
d \hat{p} &= dp
\end{align*}

Already note that momentum does not transform as one would expect. We have $d\hat{q} \wedge d\hat{p} = dq \wedge dp$.

Given $\hat{H} : T^*\mathcal{Q} \rightarrow \mathbbm{R}$, the equations of motions transform in the following way:
\begin{align*}
\frac{d\hat{q}}{dt} &= \frac{dq}{dt} + v = \frac{\partial \hat{H}}{\partial \hat{p}}= \frac{\partial \hat{H}}{\partial q} \frac{\partial q}{\partial \hat{p}} + \frac{\partial \hat{H}}{\partial p} \frac{\partial p}{\partial \hat{p}} \\
&= \frac{\partial \hat{H}}{\partial p} \\
\frac{dq}{dt} &= \frac{\partial \hat{H}}{\partial p} - v = \frac{\partial (\hat{H} - vp)}{\partial p} \\
\frac{d\hat{p}}{dt} &= \frac{dp}{dt} = - \frac{\partial \hat{H}}{\partial \hat{q}} = - \frac{\partial \hat{H}}{\partial q} \frac{\partial q}{\partial \hat{q}} - \frac{\partial \hat{H}}{\partial p} \frac{\partial p}{\partial \hat{q}} \\
&=  - \frac{\partial \hat{H}}{\partial q} = \frac{\partial (\hat{H} - vp)}{\partial q}
\end{align*}

We have $H = \hat{H} - vp$: the Hamiltonian is not invariant. The fact that the Hamiltonian, a function of phase space only, has to change signals another problem. That is: we are not being consistent on how we are treating $t$. For the differential, we treated it as a constant. For the derivative, we treated it as an independent variable.

Suppose $\hat{H} = \frac{1}{2m} \hat{p}^2 + V(\hat{q})$. We have:
\begin{align*}
H &= \hat{H} - vp = \frac{1}{2m} \hat{p}^2 -v\hat{p} + V(\hat{q})\\
&= \frac{1}{2m} (\hat{p} - mv)^2 - \frac{1}{2} mv^2 + V(\hat{q}) \\
\end{align*}

We can define:
\begin{align*}
\bar{q} &= \hat{q} - vt \\
\bar{p} &= \hat{p} - mv \\
\bar{H} &= \frac{1}{2m} \bar{p}^2 - \frac{1}{2} mv^2 + V(\bar{q} + vt)
\end{align*}
and obtain the expected transformation. Note that the new Hamiltonian is no longer just a function of $T^*\mathcal{Q}$: it's a function of t as well.

In other words, to perform a Galilean boost in phase space we have to:
\begin{itemize}
	\item Ambivalently treat $t$ as a constant or an independent variable depending on the situation
	\item Adjust the Hamiltonian based on how the equations of motion change
	\item Transform conjugate momentum not simply as a covector, but based on how the form of the Hamiltonian and how it transforms
\end{itemize}

That is why one does not do this at all. Typically one starts with the Lagrangian to get the new expressions for the conjugate momentum and the Hamiltonian. But even in that case, the transformation rules applied to the velocity are not the ones for vectors: they are the ones for derivatives in $t$.

\subsection{Extended phase-space}

In extended phase space, the change of coordinates associated with Galilean boost is:

\begin{align*}
\hat{q} &= q + vt \\
\hat{t} &= t \\
q &= \hat{q} - vt \\
t &= \hat{t} \\
d\hat{q} &= dq + v dt \\
d\hat{t} &= dt
\end{align*}

We can now properly express the differentials. Momentum transforms like a covector under coordinate transformations.

TODO: Fix here

\begin{align*}
\hat{p} &= \frac{\partial q}{\partial \hat{q}} p + \frac{\partial t}{\partial \hat{q}} E = p \\
\hat{E} &= \frac{\partial q}{\partial \hat{t}} p + \frac{\partial t}{\partial \hat{t}} E = -vp + E \\
d \hat{p} &= dp \\
d \hat{E} &= -v dp + dE
\end{align*}

The symplectic form $\omega = d\hat{q} \wedge d\hat{p} - d\hat{t} \wedge d\hat{E}$ transforms like:

\begin{align*}
\omega &= d\hat{q} \wedge d\hat{p} - d\hat{t} \wedge d\hat{E} \\
&= (dq + v dt) \wedge dp - dt \wedge (-v dp + dE)\\
&= dq \wedge dp - dt \wedge dE + 2v dt \wedge dp
\end{align*}

and therefore is not preserved.

The change of coordinates associated with Lorentz boost is:

\begin{align*}
\hat{q} &= q cosh \zeta  + ct sinh \zeta  \\
\hat{t} &= \frac{1}{c} q sinh \zeta + t cosh \zeta \\
q &= \hat{q} cosh \zeta - c\hat{t} sinh \zeta \\
t &= - \frac{1}{c} \hat{q} sinh \zeta + \hat{t} cosh \zeta \\
d\hat{q} &= dq cosh \zeta  + cdt sinh \zeta \\
d\hat{t} &= \frac{1}{c} dq sinh \zeta + dt cosh \zeta
\end{align*}

Momentum transforms like a covector under coordinate transformations.

\begin{align*}
\hat{p} &= \frac{\partial q}{\partial \hat{q}} p + \frac{\partial t}{\partial \hat{q}} E = p cosh \zeta - \frac{1}{c} E sinh \\
\hat{E} &= \frac{\partial q}{\partial \hat{t}} p + \frac{\partial t}{\partial \hat{t}} E = - c p sinh + E cosh \zeta \\
d \hat{p} &= dp cosh \zeta - \frac{1}{c} dE sinh \\
d \hat{E} &= - c dp sinh + dE cosh \zeta
\end{align*}

The symplectic form $\omega = d\hat{q} \wedge d\hat{p} - d\hat{t} \wedge d\hat{E}$ transforms like:

\begin{align*}
\omega &= d\hat{q} \wedge d\hat{p} - d\hat{t} \wedge d\hat{E} \\
&= (dq cosh \zeta  + cdt sinh \zeta) \wedge (dp cosh \zeta - \frac{1}{c} dE sinh) \\
&- (\frac{1}{c} dq sinh \zeta + dt cosh \zeta) \wedge (- c dp sinh + dE cosh)\\
&= (cosh^2 ) dq \wedge dp - dt \wedge dE + 2v dt \wedge dp
\end{align*}

\end{document}

### 2017-02-16 Notes/UnderstandingLagrangian.tex
Adding notes on other issues

\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Notes for Lagrangian}

Define a Lagrangian $L : T\mathcal{Q} \rightarrow \mathbb{R}$. Define the action $S = \int_{t_0}^{t_1} L dt$. Minimize/maximize the action $\delta S = 0$. Obtain the Euler-Lagrange equations $\partial_{q^i}L-d_t \partial_{\dot{q}^i}L=0$.

When is the minimization/maximization problem well defined? Conjecture: it is only well defined when L is twice differentiable and concave in $\dot{q}^i$ (hyper-regular).

One d.o.f. case. Expand the total derivative:
\begin{align*}
0&=\partial_{q}L-d_t \partial_{\dot{q}}L \\
&=\partial_{q}L- \partial_{\dot{q}} (d_tL) \\
&=\partial_{q}L- \partial_{\dot{q}} (\partial_{q}L \; d_t q + \partial_{\dot{q}}L \; d_t \dot{q}) \\
&=\partial_{q}L- \partial_{\dot{q}} (\partial_{q}L \; \dot{q} + \partial_{\dot{q}}L \; \ddot{q}) \\
&=\partial_{q}L- \partial_{\dot{q}} \partial_{q}L \; \dot{q} + \partial_{\dot{q}}\partial_{\dot{q}}L \; \ddot{q} \\
\end{align*}

$L$ has to be twice differentiable for the equation to make sense. If $\partial_{\dot{q}}\partial_{\dot{q}}L \neq 0$ then it is an equation in $\ddot{q}$ with a unique solution. If not, it becomes the constrain $\partial_{q}L- \partial_{\dot{q}} \partial_{q}L \; \dot{q} = 0$ on $L$. What does this constrain corresponds to?

Suppose we are in a region such that $\partial_{\dot{q}}\partial_{\dot{q}}L = 0$ and $L$ satisfies the constrain $\partial_{q}L- \partial_{\dot{q}} \partial_{q}L \; \dot{q} = 0$. Let $M(q, \dot{q}) \equiv \partial_{q}L$. We have $\partial_{\dot{q}} M \; \dot{q} = M$. Then we have $M=N(q)\dot{q}$. Which means $\partial_{q}L = N(q)\dot{q}$, $L = R(q) \dot{q} + A(\dot{q})$ where $N = \partial_q R$. Since $\partial_{\dot{q}}\partial_{\dot{q}}L = 0$ on the region, $A(\dot{q})$ is a constant. As adding a constant to $L$ does not change the equations of motion, we choose $A(\dot{q})=0$ on the region.

Consider now the action $S = \int_{t_0}^{t_1} L dt = \int_{t_0}^{t_1} R(q) \dot{q} dt = \int_{t_0}^{t_1} R(q) \frac{dq}{dt} dt = \int_{q_0}^{q_1} R(q) dq$. Let $U$ be the anti-derivative of $R$, that is $\partial_q U = R$. $S = U(q_1) - U(q_0)$. That is: the action is path independent.

Therefore in a region where $\partial_{\dot{q}}\partial_{\dot{q}}L = 0$, the only solution of the equations corresponds to the case when the action is path independent. All path minimizes/maximize the action.

If the minimization/maximization of the action is well defined only if the Lagrangian is convex, then functions that do not satisfy this requirement are not to be considered valid Lagrangians. This means that all Lagrangians will have a corresponding Hamiltonian. But the reverse is not true: all twice differentiable functions are valid Hamiltonian and only the convex in $p$ will have a corresponding Lagrangian. That is: the space of valid Hamiltonian systems is bigger than that of valid Lagrangian systems.

\end{document}

### 2017-02-17 2017HamiltonianInformation/hamiltonianInformation.tex
Starting new paper on Hamiltonian mech and inf entropy
Adding marginal distributions proof sketches

@@
 
+$\mathcal{M} = T^*\mathcal{Q}$.
 $(T^*\mathcal{Q}, \omega)$, $(q^i, p_j)$ set of canonical coordinates, $F: T^*\mathcal{Q} \rightarrow T^*\mathcal{Q}$ such that $\omega = F^* \omega$ where $F^*$ is the pullback.
@@
 $\rho(q^i, p_j) : \mathbb{R}^{2n} \rightarrow \mathbb{R}$ integrable and differentiable.
-$\rho (q^i, p_j) \, dq^n \wedge dp_n = \rho (q^i, p_j) \, \omega^n = \rho (q^i(\hat{q}^i, \hat{p}_j), p_j(\hat{q}^i, \hat{p}_j)) \, d\hat{q}^n \wedge d\hat{p}_n$.
+$\rho (q^i, p_j) \, dq^n \wedge dp_n = \rho (q^i, p_j) \, \frac{\omega^n}{n!} = \rho (q^i(\hat{q}^i, \hat{p}_j), p_j(\hat{q}^i, \hat{p}_j)) \, d\hat{q}^n \wedge d\hat{p}_n$.
 $\rho$, defined on the symplectic structure, transforms like a function.
@@
 
+$\mathcal{Q} = \mathcal{Q}_1 \times \mathcal{Q}_2$.
+$(T^*\mathcal{Q}, \omega)= (T^*\mathcal{Q}_1, \omega_1) \times (T^*\mathcal{Q}_2, \omega_2)$, $(q^i, p_j)$ set of canonical coordinates divides into $(q_1^i, p_{1j})$ and $(q_2^i, p_{2j})$ canonical coordinates.
+
+$\rho : (T^*\mathcal{Q}, \omega) \rightarrow \mathbb{R}$.
+$\frac{(\omega)^n}{n!} = \frac{(\omega_1 + \omega_2)^n}{n!} = {n \choose n_1} \frac{(\omega_1)^{n_1} \wedge (\omega_2)^{n_2}}{n!} = \frac{n!}{n_1!n_2!}\frac{(\omega_1)^{n_1} \wedge (\omega_2)^{n_2}}{n!} = \frac{(\omega_1)^{n_1}}{n_1!} \wedge \frac{(\omega_2)^{n_2}}{n_2!} $.
+$\int_{T^*\mathcal{Q_1}}\rho \frac{(\omega)^n}{n!} = \int_{T^*\mathcal{Q_1}}\rho \frac{(\omega_1)^{n_1}}{n_1!} \wedge \frac{(\omega_2)^{n_2}}{n_2!} = \frac{(\omega_2)^{n_2}}{n_2!} \wedge \int_{T^*\mathcal{Q_1}}\rho \frac{(\omega_1)^{n_1}}{n_1!}$.
+Let $\rho_2 : (T^*\mathcal{Q}_2, \omega) \rightarrow \mathbb{R}$ such that $\rho_2 = \int_{T^*\mathcal{Q_1}}\rho \frac{(\omega_1)^{n_1}}{n_1!}$.
+$\int_{T^*\mathcal{Q_1}}\rho \frac{(\omega)^n}{n!} = \rho_2 \frac{(\omega_2)^{n_2}}{n_2!}$.
+Similarly, $\int_{T^*\mathcal{Q_2}}\rho \frac{(\omega)^n}{n!} = \rho_1 \frac{(\omega_1)^{n_1}}{n_1!}$ with $\rho_1 : (T^*\mathcal{Q}_1, \omega) \rightarrow \mathbb{R}$.
+
+$F(T^*\mathcal{Q}, \omega) = F((T^*\mathcal{Q}_1, \omega_1) \times (T^*\mathcal{Q}_2, \omega_2)) = F_1(T^*\mathcal{Q}_1, \omega_1) \times F_2(T^*\mathcal{Q}_2, \omega_2) = (\hat{\mathcal{M}}_1, \hat{\omega}_1) \times (\hat{\mathcal{M}}_2, \hat{\omega}_2)$.
+Let $\hat{\rho}_1 : (\hat{\mathcal{M}}_1, \hat{\omega}_1) \rightarrow \mathbb{R}$ and $\hat{\rho}_2 : (\hat{\mathcal{M}}_2, \hat{\omega}_2) \rightarrow \mathbb{R}$ such that $\hat{\rho}_1 = \rho_1 \circ F^{-1 }$ and $\hat{\rho}_2 = \rho_2 \circ F^{-1 }$.
+$\hat{\rho}_1(\hat{q}_1^i, \hat{p}_{1j})=\rho_1(q_1^i, p_{1j})$ and $\hat{\rho}_2(\hat{q}_2^i, \hat{p}_{2j})=\rho_2(q_2^i, p_{2j})$.
+
+Let $\rho=\rho_1\rho_2$.
+$\hat{\rho}(\hat{q}^i, \hat{p}_j) =\rho(q^i, p_j)=\rho_1(q_1^i, p_{1j})\rho_2(q_2^i, p_{2j})=\hat{\rho}_1(\hat{q}_1^i, \hat{p}_{1j})\hat{\rho}_2(\hat{q}_2^i, \hat{p}_{2j})$.
+$\hat{\rho}=\hat{\rho}_1\hat{\rho}_2$.
+
+Prove the opposite: from transport and independence of marginal distributions to canonical transformation.
+
+Conjecture: the manifold for the marginal distributions is not always a cotangent bundle.
+
 \begin{prop}

### 2017-02-21 2015CompleteDerivation/completeDerivation.tex
Adding Josh comments

@@
 The aim of this work is to show that particle mechanics, both classical and quantum, Hamiltonian and Lagrangian, can be derived from few simple physical assumptions.
-Assuming deterministic and reversible time evolution will give us a dynamical system whose set of states form a topological space and whose law of evolution is a self-homeomorphism.
-Assuming the system is infinitesimally reducible, that giving the state and the dynamics of the whole system is equivalent to giving the state and the dynamics of its infinitesimal parts, will give us a classical Hamiltonian system.
-Assuming the system is irreducible, that giving the state and the dynamics of the whole system tells us nothing about the state and the dynamics of its substructure, will give us a quantum Hamiltonian system.
+Assuming deterministic and reversible time evolution will give us a dynamical system whose set of states forms a topological space and whose law of evolution is a self-homeomorphism.
+Assuming the system is infinitesimally reducible---specifying the state and the dynamics of the whole system is equivalent to giving the state and the dynamics of its infinitesimal parts---will give us a classical Hamiltonian system.
+Assuming the system is irreducible---specifying the state and the dynamics of the whole system tells us nothing about the state and the dynamics of its substructure---will give us a quantum Hamiltonian system.
 Assuming kinematic equivalence, that studying trajectories is equivalent to studying state evolution, will give us Lagrangian mechanics and limit the form of the Hamiltonian/Lagrangian to the one with scalar and vector potential forces.
@@
 In many cases, mathematical ideas were what allowed progress in theoretical physics and ultimately led to experimental discoveries.
-While these ideas are indeed useful---anything that can advance our understanding is welcome---they cannot, however, be the whole story.
+While these ideas are indeed useful---anything that can advance our understanding is welcome---they cannot be the whole story.
 
@@
 
-The real trouble comes when the mathematical description essentially becomes the foundation of a physical theory.
+The real problem begins when the mathematical description essentially becomes the foundation of a physical theory.
 While some branches of physics, such as Newtonian mechanics, thermodynamics and special relativity, are founded on physical laws or assumptions, others, such as Hamiltonian, Lagrangian and quantum mechanics, start by setting their mathematical structure.
@@
 As we pushed further back toward the basic mathematical definitions, we ended up with physical concepts of a more general nature and found that, in some cases, a single physical idea could have more than one mathematical consequence.
-That is: by putting physics at the center, we could achieve a unification of ideas that was not evident in the math alone.
+That is: by putting physics at the center, we could achieve a unification of ideas that was not evident before.
 The unexpected success of the approach led us to think that, perhaps, the picture we were developing was a better framework to understand and link different areas of basic science and mathematics.
@@
 
-Rederiving classical, quantum and (to some extent) relativistic mechanics within the same work is therefore crucial to show that the concepts we develop are truly general, that they can be applied to many different cases.
+Rederiving classical, quantum and (to some extent) relativistic mechanics within the same work is therefore crucial to show that the concepts we develop are truly general, and they can be applied to many different cases.
 If we could derive classical Hamiltonian mechanics on assumptions that are inapplicable to the quantum case, what good would it do?
@@
 Interspersed between each step of the more formal treatment, there are numerous discussions that conceptually explain motivations and results in a more approachable way.
-These can use more concrete and intuitive examples which may, in some cases, make forward references.
+These often provide more concrete and intuitive examples which may, in some cases, foreshadow later material.
 As both the formal and more intuitive parts need to cover the same material, this will inevitably lead to some repetition.
@@
 \subsection{Infinitesimal reducibility}
+\label{subsec:infinitesimal_reducibility}
 
@@
 As they have the same net effect, if we apply all three we go back where we started.
-This gives us a very intuitive classical particle mechanics analogue for the $CPT$ theorem~\cite{Weinberg}.
+This gives us a very intuitive classical particle mechanics analogue for the $CPT$ theorem~\cite{Weinberg,ClCPT}.
 
@@
 As the material is homogeneous, all its particles carry the same amount of material and are described by the same state space.
-Each quantum particle is irreducible: its state does not tell us anything about the state of its parts.
+Each quantum particle is \textit{irreducible}: its state does not tell us anything about the state of its parts.
 While this idea should be familiar enough, it is best to iron out some details to avoid possible misconceptions.
 
-The first possible source of confusion is between divisibility, the idea that an object can be divided into parts, and reducibility, the idea that the state of the whole is equivalent to the state of the parts.
+The first possible source of confusion is between \textit{divisibility}, the idea that an object can be divided into parts, and \textit{reducibility}, the idea that the state of the whole is equivalent to the state of the parts.
 More precisely, a system is divisible if there is a time evolution $\mathcal{T}_{\Delta t}: \mathcal{S} \rightarrow \mathcal{S}_1 \times \mathcal{S}_2$ that takes the state of a system and returns the state of two separate isolated systems.
@@
 
-Even more subtle is the difference between divisibility and decomposability, the idea that the object has parts.
+Even more subtle is the difference between divisibility and \textit{decomposability}, the idea that the object has parts.
 That is, a system is decomposable if there is a law $+ : \mathcal{S} \times \mathcal{S} \rightarrow \mathcal{S}$ to combine two states to get a third.
-While irreducibility says nothing on whether a quantum particle can be divided, it implicitly assumes that it can be decomposed.
+While irreducibility says nothing about whether a quantum particle can be divided, it implicitly assumes that it can be decomposed.
 Irreducibility describes the relationship between the state of the whole and the state of the parts, therefore the existence of parts is implicit in the definition.
@@
 This is why one may not dwell on their difference and use the word ``part" interchangeably in the three cases.
-Now that we assume irreducibility, it is better to use more precise language to avoid confusion.
+In the quantum context, instead, we assume irreducibility, so it is better to use more precise language to avoid confusion.
 Therefore we will call \emph{fragments} the infinitesimal parts of the material that cannot be described independently.
@@
 This operator is linear in $\psi_1$: the component of the combined system that is prepared like $\psi_2$ is the sum of the subcomponents that are prepared like $\psi_2$.
-It also gives us the same result if applied once or twice: once we found the component that is prepared like $\psi_2$, it remains the same.
+It also gives us the same result if applied once or twice: once we find the component that is prepared like $\psi_2$, it remains the same.
 $\mathsf{P}_{\psi_2}$ is therefore a projection: a linear operator such that $\mathsf{P}_{\psi_2}(\mathsf{P}_{\psi_2} (\psi_1))= \mathsf{P}_{\psi_2} (\psi_1)$.
@@
 That is: $\mathsf{P}_{\psi_2} (\psi_1) = \psi_1 \Leftrightarrow \psi_1 = \alpha \psi_2$.
-Because of this, we can define a non degenerate complex product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbb{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{\langle \psi_2|\psi_1 \rangle}{|\psi_2|^2} \psi_2$.
+Because of this, we can define a non-degenerate complex product $\langle \cdot |\cdot \rangle : \Psi \times \Psi \rightarrow \mathbb{C}$ such that $\mathsf{P}_{\psi_2} (\psi_1) \equiv \frac{\langle \psi_2|\psi_1 \rangle}{|\psi_2|^2} \psi_2$.
 
@@
 The whole state $\phi$ is prepared as $\psi$.
-We can obtain $\phi$ from $\psi$ by a relabeling of the fragment configurations.
+We can obtain $\phi$ from $\psi$ by relabeling the fragment configurations.
 $\phi = c \psi$ for some $c \in \mathbb{C}$.
@@
 Let $\phi$ be the empty state.
-$|\phi|^2=0$ because it has no material and $|\mathsf{P}_{\psi} (\phi)|^2$ as no component of the empty state is prepared like $\psi$.
+$|\phi|^2=0$ because it has no material and $|\mathsf{P}_{\psi} (\phi)|^2 = 0$ as no component of the empty state is prepared like $\psi$.
 
@@
 Note that the inner product only has an indirect physical definition, based on the norm and the projections.
-These two concepts are therefore more useful to gain physical intuition as the first does not have a straightforward physical meaning.
+These latter two concepts therefore provide more physical intuition since the first does not have a straightforward physical meaning.
 
@@
 For mathematical convenience, one can take its completion with respect to the norm induced by the inner product and obtain a Hilbert space.
-Yet, such object no longer represents a physically meaningful set of states.
+Yet, such an object no longer represents a physically meaningful set of states.
 
 Therefore we need to make sure that the rest of our definitions and justifications do not require a complete metric space to be valid or they would not be physically meaningful.
-In particular, we don't want to use the usual link $\psi(q) = \langle q | \psi \rangle$ between the wave function and the inner product as the material cannot be prepared at a specific value when $q$ is a continuous variable.
+In particular, we don't want to use the usual link $\psi(q_0) = \langle q_0 | \psi \rangle$ between the wave function and the inner product as the material cannot be prepared at a specific value $q_0$ when $q$ is a continuous variable.
 Similarly, we cannot show that to each state variable is associated a self-adjoint operator $Q = \sum q |q\rangle \langle q |$ simply by expressing it in terms of eigenstates as they may not represent physical states.
@@
 As each state $\psi$ can be identified by a wave function on either space, there must be an operator $\Upsilon^q_{\hat{q}}(\psi(q)) = \psi(\hat{q})$ that converts the wave function on one space to the other.\footnote{Note that a neighborhood of $q$ is not necessarily a neighborhood of $\hat{q}$.
-They are only if $q$ and $\hat{q}$ can be expressed as a bijection of each other $\hat{q}^i(q^i)$: they are different coordinates of the same manifold.} As a linear combination of states corresponds to a linear combination of wave functions, $\Upsilon^q_{\hat{q}}$ is a linear operator.
+This relation holds only if $q$ and $\hat{q}$ can be expressed as a bijection of each other $\hat{q}^i(q^i)$, i.e.~if they are different coordinates of the same manifold.} As a linear combination of states corresponds to a linear combination of wave functions, $\Upsilon^q_{\hat{q}}$ is a linear operator.
 As converting the space again must give us the original wave function, $\Upsilon^{\hat{q}}_{q} \circ \Upsilon^q_{\hat{q}}= I$ is unitary.
@@
 \begin{prop}\label{prop:self_adjoint_operators}
-	For each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbb{R}$ and a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ that takes a state $\psi \in \Psi$ and returns a wave function $\psi(\hat{q}^i)$ from $\Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ the space of wave functions over $\hat{\mathcal{Q}}$.
+	For each fragment variable $f$ there exists a topological space $\hat{\mathcal{Q}}$ such that $f : \hat{\mathcal{Q}} \rightarrow \mathbb{R}$ and a vector space isomorphism $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \leftrightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ that takes a state $\psi \in \Psi$ and returns a wave function $\psi(\hat{q}^i)$ from $\Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$, the space of wave functions over $\hat{\mathcal{Q}}$.
 There also exists a corresponding self-adjoint linear operator $F_{[q^i]} : \Psi(\mathcal{Q}, dq^i) \rightarrow C(\mathcal{Q}, \mathbb{C})$ such that $\int \psi^\dagger(q^i) F_{[q^i]}( \psi(q^i))d^nq $ is the expectation of $f$ multiplied by the amount of material.
@@
 $\hat{q}^i$ provides a maximal set of fragment variables.
-There must exist a wave function $\psi(\hat{q}^i) : \mathbb{R}^n \rightarrow \mathbb{C}$ associated with $\psi$ such that it gives the maximal description for the state.
+There must exist a wave function $\psi(\hat{q}^i) : \mathbb{R}^n \rightarrow \mathbb{C}$ associated with $\psi$ such that it gives the maximal description for the state as seen in \ref{prop:wavefuntion}.
 Let $\Upsilon_{\hat{\mathcal{q}}^i} : \Psi \rightarrow \Psi(\hat{\mathcal{Q}}, d\hat{q}^i)$ be the map that associates a state with its wave function over $\hat{\mathcal{Q}}$.
@@
 The topology of $\mathcal{P}$, the space charted by conjugate momentum, depends globally on the topology of $\mathcal{Q}$.
-For example, if $\mathcal{Q}$ is an n-dimensional torus, there is no infinity and $\mathcal{P}$ has a discrete topology.
+For example, if $\mathcal{Q}$ is an n-dimensional torus, $\mathcal{P}$ has a discrete topology.
 That is why, in the end, we restricted ourselves to $\mathbb{R}^n$.
@@
 
-While the link between math and physics may not be as elegant and as general as one would like, we still managed the goal: we identified the states of an irreducible material and saw how these are related to the states in quantum mechanics.
+While the link between math and physics may not be as elegant and as general as one would like, we still managed to meet the goal: we identified the states of an irreducible material and saw how these are related to the states in quantum mechanics.
 
@@
 
-	As we saw in assumption \ref{ass:infinitesimal_reducibility}, infinitesimal reducibility had its problems. How about the irreducibility assumption? How do they compare?
-
-	The methodological problem assumes a different connotation.
+	As we saw in section \ref{subsec:infinitesimal_reducibility} during the rationale for assumption \ref{ass:infinitesimal_reducibility}, infinitesimal reducibility had its problems.
+How about the irreducibility assumption?
+How do they compare?
+
+	The methodological problem takes on a different character.
 In order for the assumption to hold, we need to show that there are no processes at our disposal that can probe the fragments of a quantum particle.
 For example: an electron as a whole interacts with a photon as a whole.
-Note that this does not require showing that no such process exists: conceptually specific parts of the photon may be interacting with specific parts of the electron.
+Note that this does not require showing that no such process exists: conceptually, specific parts of the photon may be interacting with specific parts of the electron.
 But if those processes are not available to us, say because we can only manipulate whole photons and they always interact with a whole electron, they cannot be used to create a more detailed picture.
@@
 Each degree of freedom comes with its own unstated part (i.e.
-the motion of the fragments within that d.o.f.), therefore there is no pretense of a perfect description under perfect isolation.
+the motion of the fragments within that d.o.f.); therefore there is no pretense of a perfect description under perfect isolation.
 Since we cannot describe the nature of the chaotic motion, we cannot say whether it is internally or externally driven.
@@
 
-	The inability to define time using a deterministic and reversible evolution is still a problem.
+	The inability to define time using a deterministic and reversible evolution remains a problem.
 As the system is homogeneous and irreducible, the unstated part cannot give us a measurable quantity that changes in time as that would provide a description of the unstated part.
@@
 In fact if the internal dynamics were out of equilibrium, the material would not be homogeneous which would contradict the assumption.
-Deterministic and reversible evolution (which we'll see corresponds to unitary evolution) is necessarily quasi static: at each moment a state must be well defined, at each moment the system is in equilibrium, the system changes slowly.
+Deterministic and reversible evolution (which we'll see corresponds to unitary evolution) is necessarily quasi-static: at each moment a state must be well defined, at each moment the system is in equilibrium, the system changes slowly.
 Therefore quantum states, as formulated here, cannot and should not be expected to provide a valid description during non-deterministic evolution.
@@
 
-	If we look more closely, we realize that the process that the S-matrix captures is a limit where initial states are at $t \rightarrow - \infty$ and final states are at $t \rightarrow + \infty$, one that takes an infinite amount of time.
-That is, to calculate predictions for a fast out of equilibrium process we are approximating it using slow quasi-static evolution.
+	If we look more closely, we realize that the process that the S-matrix captures is a limit where initial states are at $t \rightarrow - \infty$ and final states are at $t \rightarrow + \infty$, a process that takes an infinite amount of time.
+That is, to calculate predictions for a fast out-of-equilibrium process we approximate it using slow quasi-static evolution.
 This is actually not so uncommon: it is a standard first approximation in thermodynamics.
-The question is why would it work here?
+The question is why should it work here?
 As we'll see in more detail when talking about the projection postulate, it's because of the irreducibility assumption itself.
@@
 For example, for a system to reach its ground state it has to be able to radiate energy: it cannot be isolated.
-Therefore this dynamics cannot be described under the deterministic and reversible assumption.
+Therefore this dynamics cannot be described under the assumption of deterministic and reversible evolution.
 This already hints that deterministic and reversible evolution is not the only dynamics possible.
 Other considerations, outside of the scope of this work, can be used to recover this behavior (e.g.
-thermodynamics, decoherence, ...).
+thermodynamics, decoherence, ...)~\cite{Weiss}.
 
@@
 The next step would be to integrate the quantum description with time dependent evolution (i.e.
-relativistic mechanics) and the kinematic assumption (i.e.
+time dependent Hamiltonians and relativistic mechanics) and the kinematic assumption (i.e.
 Lagrangian mechanics).
@@
 They give more insight to all the pieces in a comprehensive way and help tear down the artificial walls between the different physical theories.
-This allows us to form a picture of fundamental physics that is more unified, something that has been sorely missing.
+This allows us to form a picture of fundamental physics that is more unified, that draws upon a set of common concepts.
 
@@
 It is clear that it has to be absorbed as it is within the aperture of the whole array.
-It is also clear that it must be absorbed whole: as it is an irreducible system, the detector cannot partially interact with it, it cannot half absorb it.
+It is also clear that it must be absorbed as a whole system: as it is an irreducible system, the detector cannot partially interact with it, it cannot half absorb it.
 So a single detector will capture it.
@@
 Note that while this process is of a non-deterministic nature, it is of a specific non-deterministic nature.
-It requires that no other state can influence the final distribution as the match between the physical material distribution and the statistical measurement distribution was recovered only under the assumption that the detector does not introduce any bias.
+It requires that no other state can influence the final distribution because the match between the physical material distribution and the statistical measurement distribution was recovered only under the assumption that the detector does not introduce any bias.
 That is: the non-deterministic nature cannot be understood as an averaging over some unknown external state.
@@
 So much so that we can substitute it with a deterministic one to compute the probability distribution.
-In particular, given only the initial state, we have no way of predicting which set of final outcomes the system will project to as that depends on the actual choice of detector setup, nor when that projection actually happens.
+In particular, given only the initial state, we have no way of predicting which set of final outcomes the system will project to, nor when that projection actually happens, as that depends on the actual choice of detector setup.
 
@@
 
-While we are sure that some mathematical details can be improved, we feel that the overall ``conceptual castle" is solid.
+While we are sure that the details of some justifications can be improved, we feel that the overall ``conceptual castle" is solid.
 The starting physical assumptions are simple, the translation into math is often straightforward and the intuitive ideas that emerge help unveil connections within and across different branches of math and physics.
@@
 
+\bibitem{ClCPT} Bell, J. S.: Time reversal in field theory. Proceedings of the Royal Society of London A, 231 (1187), 479-95, (1955)
+
 \bibitem{EPR} Einstein, A., Podolsky, B.,  Rosen, N.: Can quantum-mechanical description of physical reality be considered complete? Phys. Rev. 47 777 (1935)
@@
 \bibitem{Bohr} Bohr, N.: The Quantum Postulate and the Recent Development of Atomic Theory. Nature 121, pp. 580-–590 (1928)
+
+\bibitem{Weiss} Weiss, U.: Quantum Dissipative Systems. World Scientific (1992)
 \end{thebibliography}

### 2017-02-22 2015CompleteDerivation/completeDerivation.tex
Changing title

@@
 
-\title{From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics}
+\title{From physical assumptions to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics}
 \author{Gabriele Carcassi, Christine A. Aidala, David J. Baker, Lydia Bieri}
@@
 	% As the page margins are so large, the full title does not fit
-	\titlerunning{From physical principles to classical and quantum particle mechanics}
+	\titlerunning{From physical assumptions to classical and quantum particle mechanics}
 \fi

### 2017-03-01 2017HamiltonianInformation/hamiltonianInformation.tex
Extended to information theory

@@
 
+Suppose we can write $\rho=\prod\rho_i(q^i,p_i)$.
+Then $1 = \int_{\mathcal{M}} \rho dq^n\wedge dp_n$ but also $1 = \prod \int_{\mathcal{M}_i} \rho_i dq^i\wedge dp_i$.
+$\int_{\mathcal{M}} \prod\rho_i(q^i,p_i) dq^n\wedge dp_n = \prod \int_{\mathcal{M}_i} \rho_i dq^i\wedge dp_i$.
+$dq^n\wedge dp_n = \bigwedge dq^i\wedge dp_i = \frac{(\sum dq^i\wedge dp_i)^n}{n!}$.
+$\omega = \sum dq^i\wedge dp_i$.
+$(q^i,p_j)$ are canonical coordinates.
+Suppose $F$ is such that $\hat{\rho}(\hat{q}^i, \hat{p}_j) = \rho(q^i, p_j)$, $\hat{\rho}_1(\hat{q}_1^i, \hat{p}_{1j})=\rho_1(q_1^i, p_{1j})$ and $\hat{\rho}_2(\hat{q}_2^i, \hat{p}_{2j})=\rho_2(q_2^i, p_{2j})$.
+For each pair $(q^i, p_i)$ we have $\int_{\mathcal{M}_i} \rho_i dq^i \wedge dp_i = \int_{\hat{\mathcal{M}}_i} \hat{\rho}_i d\hat{q}^i \wedge d\hat{p}_i$ for all $\rho$ and $\hat{\rho}$: $dq^i \wedge dp_i = d\hat{q}^i \wedge d\hat{p}_i$.
+$\sum dq^i \wedge dp_i = \sum d\hat{q}^i \wedge d\hat{p}_i$.
+$(\hat{q}^i,\hat{p}_j)$ are canonical coordinates.
+$F$ is a canonical transformation.
+
 Conjecture: the manifold for the marginal distributions is not always a cotangent bundle.
@@
 
+If $F$ canonical transformation, $I_{ij} = \int_{T^*\mathcal{Q}} \rho \log (\frac{\rho}{\rho_i \rho_j}) dq^n dp_n = \int_{T^*\mathcal{Q}} \hat{\rho} \log (\frac{\hat{\rho}}{\hat{\rho}_i \hat{\rho}_j}) dq^n dp_n = \hat{I}_{ij}$.
+
+Suppose $I_{ij} = \int_{T^*\mathcal{Q}} \rho \log (\frac{\rho}{\rho_i \rho_j}) dq^n dp_n = \int_{T^*\mathcal{Q}} \hat{\rho} \log (\frac{\hat{\rho}}{\hat{\rho}_i \hat{\rho}_j}) dq^n dp_n = \hat{I}_{ij}$ for all $\rho$.
+Then $\rho(q^i, p_j)=\hat{\rho}(\hat{q}^i, \hat{p}_j)$ and $\rho_i(q^i, p_i)=\hat{\rho}_i(\hat{q}^i, \hat{p}_i)$.
+In particular, if $I_{ij} = \hat{I}_{ij} = 0$ if $i\neq j$ we can write $\rho=\prod\rho_i(q^i,p_i) = \hat{\rho} = \prod\rho_i(\hat{q}^i,\hat{p}_i)$.
+$F$ is a canonical transformation as it preserves independence of marginal distributions.
+
 \begin{thebibliography}{0}

### 2017-03-09 2017HamiltonianInformation/hamiltonianInformation.tex
Starting to add some text on definitions of distributions

@@
 
+\section{Distributions as observer dependent}
+
+The main concept at the heart of this work is the idea of a distribution or density.
+What we want to do is talk about its information entropy (i.e.
+how many bits are required to identify one of its elements) and its evolution in time.
+But before we can do that, we have to clarify some issues that appear when dealing with distributions: mainly that they are not invariant under change of variables.
+Partially to address this problem, differential geometry and probability theory formalize the concept in different ways.
+This creates further difficulties as we want to mix ideas from both camps.
+For example, we will want to talk about marginal distributions on manifolds.
+In this section we want to clarify what the problems are, how they are formally addressed in both disciplines so that we can build a bridge and highlight the connection.
+
+We start with our physical intuition.
+Distributions are widely used tools in all branches of physics: mass or charge densities, probability distributions over energy, cross sections and so on.
+All these objects are the limit of a ration between two quantities which can be integrated to get final values.
+For example, if $\mathbf{x}$ is the coordinate of a point, $V$ is a volume and $\rho$ is a mass distribution, the integral
+\begin{equation}
+m =\int_V \rho(\mathbf{x}) dV
+\end{equation}
+tells us the total mass within the given volume.
+
+Similarly, we can define the information entropy for the distribution as
+\begin{equation}
+I(\rho) =-\int \rho(\mathbf{x}) \ln (\rho(\mathbf{x})) dV
+\end{equation}
+which quantifies the amount of information needed to identify an element within the distribution.
+
+Densities, though, are not invariant.
+Intuitively, if we change units of the denominator the distribution changes value.
+For example, $1$ kg/m becomes $1000$ kg/km: the number changes but the units allow us to understand that the overall quantity remains the same.
+In general, if $\mathbf{y}=\mathbf{y}(\mathbf{x})$ we can recover how the density changes by requiring that the mass in each volume is the same no matter what coordinates are used.
+That is
+\begin{align*}
+m &=\int_V \rho_x(\mathbf{x}) dV_x = \int_V \rho_y(\mathbf{y}) dV_y \\
+&=\int_V\rho_y(\mathbf{y}) \left|\frac{\partial \mathbf{y}}{\partial \mathbf{x}}\right| dV_x
+\end{align*}
+\begin{equation}
+\rho_x(\mathbf{x}) = \rho_y(\mathbf{y}) \left|\frac{\partial \mathbf{y}}{\partial \mathbf{x}}\right|
+\end{equation}
+the density is multiplied by the Jacobian of the transformation.
+
+Similarly, we can see that information entropy is not invariant
+\begin{align*}
+I(\rho_x) &=-\int \rho_x(\mathbf{x}) \ln (\rho_x(\mathbf{x})) dV_x \\
+&=-\int \rho_y(\mathbf{y}) \left|\frac{\partial \mathbf{y}}{\partial \mathbf{x}}\right| \ln (\rho_y(\mathbf{y}) \left|\frac{\partial \mathbf{y}}{\partial \mathbf{x}}\right|) dV_x \\
+&=-\int \rho_y(\mathbf{y}) \ln (\rho_y(\mathbf{y}) \left|\frac{\partial \mathbf{y}}{\partial \mathbf{x}}\right|) dV_y \\
+&=-\int \rho_y(\mathbf{y}) \ln (\rho_y(\mathbf{y})) dV_y - \int \rho_y(\mathbf{y}) \ln (\left|\frac{\partial \mathbf{y}}{\partial \mathbf{x}}\right|) dV_y \\&=I(\rho_y) - \int \rho_y(\mathbf{y}) \ln (\left|\frac{\partial \mathbf{y}}{\partial \mathbf{x}}\right|) dV_y
+\end{align*}
+The reason is that, in the continuous case, the information entropy is not an absolute number of bits, but it's relative to uniform distribution over a unit size.
+That is: if the distribution is over a continuous variable, it would take an infinite amount of information to specify an element.
+But we can tell, for example, that to specify an element within a uniform distribution between $[0,2]$ we need a bit more than to specify an element within a uniform distribution between $[0,1]$.
+If we change coordinate what happens is that we change our reference therefore the information entropy appears to change.
+
+The punchline is the following: while physically we think as a distribution as a number associated to a point, mathematically it is not.
+That is, given the manifold $\mathcal{M}$ on which we define the distribution, the latter is \emph{not} $\rho : \mathcal{M} \rightarrow \mathbb{R}$.
+That would be invariant under coordinate transformation, as a point is a point in any coordinates.
+
+In differential geometry, a density is defined at each point as a function that takes a parallelepiped, identified by n vectors, and returns a number.
+That is $\mu : V \times ...
+\times V \rightarrow \mathbb{R}$.
+Physically, this includes both the distribution and the volume: it bundles $\rho(\mathbf{x}) dV$ because this is the invariant object.
+While this allows us to recast integration in a coordinate independent way, the object we have does not quite map to our physics intuition.
+It also does not allow to easily express quantities, like information entropy or expectations, that operate on $\rho(\mathbf{x})$ as a function at a point.
+
+In statistics, instead, the approach is different.
+The probability measure is a function that takes a set of events and assigns it a number between zero and one.
+That is $Pr : \sigma(\Omega) \rightarrow [0,1]$ does not on how the events are labeled.
+Random variables $X : \Omega \rightarrow \mathbb{R}$ are simply real valued functions of the events, not coordinates for some point on a manifold.
+For distributions, the starting point is the cumulative distribution function $F_X(x)=Pr(X<x)$ and the probability density function is its Radon–Nikodym derivative $\rho_X = \frac{dF_X}{dx}$.
+That is: the integral is the primal objects while the density is derived.\footnote{This actually makes more physical sense: what we measure is the amount of mass in a finite region and the density is the limit for a smaller and smaller one.} We therefore have densities defined at values, but they are not invariant objects.
+This is not an issue as probabilities are not objective entities.
+For example, there is no absolute probability for an electron to be in a state of spin up: it depends on the ensemble considered (e.g.
+an electron gas, electron pairs in a superconductor, ...).
+Therefore it does not matter if they depend on the choice of labels.
+
+Given the situation, though, a question arises spontaneously: what if we wanted to define probabilities at each point on a manifold stuch that they were invariant?
+What kind of mathematical structure would we get?
+We already know that such manifold would have to be differentiable, because only on a set of coordinates that are differentiable we can define the Jacobian and therefore we can properly transform the density and the information entropy.
+What other structure is needed?
+
+As we'll see, the answer is: we need, and only need, a symplectic structure.
+We need the type of structure we have in phase space for Hamiltonian mechanics.
+And if we require such distributions and their information entropy to be preserved over time, we get Hamiltonian motion.
+
 \section{Something}

### 2017-03-19 2017HamiltonianInformation/hamiltonianInformation.tex
More work on single degree of freedom

--- 2017-03-01 2017HamiltonianInformation/hamiltonianInformation.tex
@@
 
+\section{Distributions as observer dependent}
+
+The main subjects of this work are distributions and densities: their properties, like marginal distributions and information entropy, and their time evolution.
+Let's briefly review their basic features.
+
+Distributions may be used to represent different  physical entities: mass or charge densities, probability distributions over energy, cross sections and so on.
+All these objects are the limit of a ratio between two quantities, which can be integrated over an arbitrary region to get a finite value.
+For example, if $\mathbf{x}$ is the coordinate of a point, $V$ is a volume, $M$ is the total mass and $\rho$ is the normalized mass distribution, the integral
+\begin{equation}
+m_V =\int_V M \rho(\mathbf{x}) dV
+\end{equation}
+tells us the total mass within the given volume.
+
+Similarly, we can define the information entropy for the distribution as
+\begin{equation}
+I(\rho) =-\int \rho(\mathbf{x}) \ln (\rho(\mathbf{x})) dV
+\end{equation}
+which quantifies the amount of information needed to identify an element within the distribution.
+Technically, given the more convenient choice of the natural logarithm, the unit of information is the nat, which is equal to $\frac{1}{\ln 2}$ bits.
+However, in the discussion we will refer to it as measured in bits, as we believe that the majority of the readers is more conformable with the latter unit.
+
+Densities, though, are not invariant.
+Intuitively, if we change units of the denominator the distribution changes value.
+For example, $1$ kg/m becomes $1000$ kg/km: the number changes but the units allow us to understand that the overall quantity remains the same.
+In general, if $\mathbf{y}=\mathbf{y}(\mathbf{x})$ we can recover how the density changes by requiring that the mass in each volume is the same no matter what coordinates are used.
+That is
+\begin{align*}
+m &=\int_V \rho_x(\mathbf{x}) dV_x = \int_V \rho_y(\mathbf{y}) dV_y \\
+&=\int_V\rho_y(\mathbf{y}) \left|\frac{\partial \mathbf{y}}{\partial \mathbf{x}}\right| dV_x
+\end{align*}
+\begin{equation}\label{density_transformation}
+\rho_x(\mathbf{x}) = \rho_y(\mathbf{y}) \left|\frac{\partial \mathbf{y}}{\partial \mathbf{x}}\right|
+\end{equation}
+the density is multiplied by the Jacobian of the transformation.
+
+Similarly, we can also see that information entropy is not invariant
+\begin{align*}
+I(\rho_x) &=-\int \rho_x(\mathbf{x}) \ln (\rho_x(\mathbf{x})) dV_x \\
+&=-\int \rho_y(\mathbf{y}) \left|\frac{\partial \mathbf{y}}{\partial \mathbf{x}}\right| \ln (\rho_y(\mathbf{y}) \left|\frac{\partial \mathbf{y}}{\partial \mathbf{x}}\right|) dV_x \\
+&=-\int \rho_y(\mathbf{y}) \ln (\rho_y(\mathbf{y}) \left|\frac{\partial \mathbf{y}}{\partial \mathbf{x}}\right|) dV_y \\
+&=-\int \rho_y(\mathbf{y}) \ln (\rho_y(\mathbf{y})) dV_y - \int \rho_y(\mathbf{y}) \ln (\left|\frac{\partial \mathbf{y}}{\partial \mathbf{x}}\right|) dV_y
+\end{align*}
+\begin{equation}\label{entropy_transformation}
+I(\rho_x) =I(\rho_y) - \int \rho_y(\mathbf{y}) \ln (\left|\frac{\partial \mathbf{y}}{\partial \mathbf{x}}\right|) dV_y
+\end{equation}
+The reason is that, in the continuous case, the information entropy is not an absolute number of bits.
+In fact, if the distribution is over a continuous variable, it would take an infinite amount of information to specify an element.
+Instead it's the relative number of bits needed to identify an element compared to a uniform distribution over a unit size.
+For example, to specify an element within a uniform distribution between $[0,2]$ we need a bit more than to specify an element within a uniform distribution between $[0,1]$.
+If we change coordinates we also change our reference therefore the information entropy appears to change.
+
+The punchline is the following: while physically we may regard a distribution as a number associated to a point, mathematically it is not.
+Given the manifold $\mathcal{M}$ on which we define the distribution, the latter is \emph{not} $\rho : \mathcal{M} \rightarrow \mathbb{R}$.
+That would be invariant under coordinate transformation, as a point is a point in any coordinates.
+As a consequence, differential geometry and statistics have a different way to handle them.
+
+In differential geometry, a density is defined at each point as a function that takes a parallelepiped, identified by n vectors, and returns a number.
+That is $\mu : V \times ...
+\times V \rightarrow \mathbb{R}$.
+Physically, this includes both the distribution and the volume: it bundles $\rho(\mathbf{x}) dV$ because this is the invariant object.
+While this allows us to recast integration in a coordinate independent way, the object we have does not quite map to our physics intuition.
+It also does not allow to easily express quantities, like information entropy or expectations, that operate on $\rho(\mathbf{x})$ as a function at a point.
+
+In statistics, one starts with the probability measure: a function $Pr : \sigma(\Omega) \rightarrow [0,1]$ that takes a set of events and assigns it a number between zero and one.
+This function is independent on how the events are labeled.
+Random variables $X : \Omega \rightarrow \mathbb{R}$ are simply real valued functions of the events (i.e.
+they are not thought as coordinates for some point on a manifold).
+For each random variable we can define its cumulative distribution function $F_X(x)=Pr(X<x)$ and the probability density function is its Radon–Nikodym derivative $\rho_X = \frac{dF_X}{dx}$.
+That is: the integral is the primal object while the density is derived.\footnote{This actually makes more physical sense: what we measure is the amount of mass in a finite region and the density is the limit for a smaller and smaller one.} We therefore have densities defined at values, but they are not invariant objects.
+However, as probabilities depend on our choice of ensemble, it is not inconsistent that they also depend on our labeling.
+For example, there is no absolute probability for an electron to be in a state of spin up: it depends on the ensemble considered (e.g.
+an electron gas, electron pairs in a superconductor, ...).
+
+Given the situation, though, a question arises spontaneously: what if we wanted to define densities at each point on a manifold such that they \emph{were} invariant?
+This may be desirable for something like a mass distribution, which we would like to think like coordinate independent function of the point: the fact that some mass is in a particular state should be objective and therefore independent on the choice of coordinates.
+What kind of mathematical structure would we get?
+We already know that such manifold would have to be differentiable, because only on a set of coordinates that are differentiable we can define the Jacobian and therefore we can properly transform the density and the information entropy.
+What other structure is needed?
+
+As we'll see, the answer is: we need, and only need, a symplectic structure: the type of structure we have in phase space for Hamiltonian mechanics, with its conjugate quantities and Poisson brackets.
+And if we require such distributions and their information entropy to be preserved over time, we get Hamiltonian motion.
+
+\section{Single degree of freedom}
+
+We begin with studying the simplest case, one with a single degree of freedom.
+In this case, it is easy to show that canonical transformations preserve both the probability and the information entropy.
+
+Suppose we have a distribution in phase space $\rho(q, p)$ and perform a canonical change variables $Q=Q(q.p)$ $P=P(q,p)$.
+The Jacobian of the transformation $|J| = \frac{\partial Q}{\partial q}\frac{\partial P}{\partial p} - \frac{\partial P}{\partial q} \frac{\partial Q}{\partial p} = \{Q, P\}$ is equal to the Poisson bracket between $Q$ and $P$.
+But the definition of a canonical transformation is that it preserves the Poisson brackets: $\{q,p\} = \{Q, P\} = 1$.
+Therefore we have
+\begin{align}
+\rho(q, p) &= \rho(Q,P) \\
+I(\rho(q,p)) &= I(\rho(Q,P))
+\end{align}
+The density and the information entropy associated with it do not change.
+
+Conversely, suppose we have a transformation that preserves the density and the information entropy.
+Then the Jacobian determinant of such transformation has to be unitary everywhere.
+But, for one degree of freedom, the Jacobian coincides with the Poisson bracket: the transformation must be canonical.
+
+We recast this reasoning using symplectic geometry, as this will be useful later.
+For each distribution $\rho(q, p)$ we can associate a two form $\rho(q, p) \omega$ where $\omega=dq\wedge dp$ is the two form that returns the are of an infinitesimal surface.
+In other words, the distribution is the component of the two form associated with it.
+The integral of the two form over a region will give us the integral of the distribution over a region.
+
+Under a change of coordinates, we have $\rho(Q, P) dQ \wedge dP = \rho(q,p) dq \wedge dp$.
+If the transformation is canonical $\omega = dq \wedge dp = dQ \wedge dP$.
+Therefore $\rho(Q, P) \omega = \rho(q,p) \omega$ which gives $\rho(Q, P) = \rho(q,p)$.
+Conversely, if $\rho(Q, P) = \rho(q,p)$, we find $dQ \wedge dP = dq \wedge dp = \omega$: the transformation is canonical.
+In other words: a symplectic structure is exactly the structure needed to define coordinate invariant distributions on a 2 dimensional smooth manifold.
+
+If we require the distribution and its information entropy to be preserved through time evolution, then time evolution is a canonical transformation as well.
+As such, it allows a potential $H : \mathcal{M} \rightarrow \mathbb{R}$ such that
+\begin{align}
+S^q \equiv \frac{dq}{dt} &= \frac{\partial H}{\partial p} \\
+S^p \equiv \frac{dp}{dt} &= - \frac{\partial H}{\partial q}
+\end{align} 
+where the vector $S$ tells us the infinitesimal displacement of each state.
+As time passes, both the distribution value and the volume where it is defined are conserved.
+This means that $S$, as a vector field, is divergence free.
+As such, it admits a potential such that $S = \mathrm{curl}(H)$ which leads to the previous equations.
+
+We are now in a position to ``have our invariant distribution and integrate it too".
+On the symplectic structure, our density is a scalar function $\rho : \mathcal{M} \rightarrow \mathbb{R}$.
+It has the two form $\rho \omega$ associate with it, in case we want to speak the language of differential geometry, and it has the function $\rho(q,p)$ in local coordinate, in case we want to speak the language of statistics and define quantities like the information entropy.
+
+\section{Multiple degree of freedom}
+
+The generalization to multiple degrees of freedom is conceptually very simple: not only we will need to preserve the distribution and the entropy, but we will do it for each independent degree of freedom.
+One needs to be careful, though, as it may be misconstrued to mean something much stronger that it actually is.
+
+$\varrho = \{q,p\} \rho(q,p)$
+
+$\omega = \frac{dq\wedge dp}{\{q,p\}}$
+
+$\varrho \omega = \rho(q,p) dq\wedge dp$
+
 \section{Something}

### 2017-03-19 Notes/GalileanLorentzTransformationInPhaseSpace.tex
Adding change of time units

@@
 \bar{p} &= \hat{p} - mv \\
-\bar{H} &= \frac{1}{2m} \bar{p}^2 - \frac{1}{2} mv^2 + V(\bar{q} + vt)
+\bar{H} &= \frac{1}{2m} \bar{p}^2 mv^2 + V(\bar{q} + vt)
 \end{align*}
@@
 But even in that case, the transformation rules applied to the velocity are not the ones for vectors: they are the ones for derivatives in $t$.
+
+Let's now change the time unit.
+\begin{align*}
+\hat{q} &= q \\
+\hat{p} &= p \\
+\hat{t} &= a t \\
+\end{align*}
+It is canonical in the sense that $d\hat{q} \wedge d\hat{p} = dq \wedge dp$.
+But it does not preserve the density $\rho(\hat{q},\hat{p},\hat{t}) =\frac{1}{a} \rho(q,p,t)$.
+The equations of motion transform in the following way:
+\begin{align*}
+\frac{d\hat{q}}{d\hat{t}} &= \frac{1}{a} \frac{dq}{dt}= \frac{\partial \hat{H}}{\partial \hat{p}}= \frac{\partial \hat{H}}{\partial p} \\
+\frac{d\hat{p}}{dt} &= \frac{1}{a} \frac{dp}{dt} = - \frac{\partial \hat{H}}{\partial \hat{q}} = - \frac{\partial \hat{H}}{\partial q}
+\end{align*}
+
+We have $\hat{H}= \frac{1}{a}H$. The Hamiltonian is not invariant, but rather changes as a covariant time component.
 

### 2017-11-14 2017HamiltonianInformation/hamiltonianInformation.tex
More work on the information entropy article

@@
 
-\title{DRAFT \\ Hamiltonian mechanics is conservation of information}
+\title{DRAFT \\ Hamiltonian mechanics is conservation of information \\ (or why isolated systems conserve energy)}
 \author{Gabriele Carcassi, Christine A. Aidala}
@@
 
-We recast this reasoning using symplectic geometry, as this will be useful later.
-For each distribution $\rho(q, p)$ we can associate a two form $\rho(q, p) \omega$ where $\omega=dq\wedge dp$ is the two form that returns the are of an infinitesimal surface.
-In other words, the distribution is the component of the two form associated with it.
-The integral of the two form over a region will give us the integral of the distribution over a region.
+We can formalize this intuition using symplectic geometry.
+A distribution is a function $\varrho : \mathcal{M} \rightarrow \mathbb{R}$ that at each point gives the value of the density measured over a standardized invariant (canonical) unit.
+For example, for a mass distribution over phase space, it would be in $\frac{Kg}{m \times Kg \frac{m}{s}} = \frac{s}{m^2}$.
+Canonical units measure in a standard way the count of possible states.
 
-Under a change of coordinates, we have $\rho(Q, P) dQ \wedge dP = \rho(q,p) dq \wedge dp$.
-If the transformation is canonical $\omega = dq \wedge dp = dQ \wedge dP$.
-Therefore $\rho(Q, P) \omega = \rho(q,p) \omega$ which gives $\rho(Q, P) = \rho(q,p)$.
-Conversely, if $\rho(Q, P) = \rho(q,p)$, we find $dQ \wedge dP = dq \wedge dp = \omega$: the transformation is canonical.
-In other words: a symplectic structure is exactly the structure needed to define coordinate invariant distributions on a 2 dimensional smooth manifold.
+Given of two smooth variables $f,g \in C^\infty(\mathcal{M})$ of phase space, we have an operation $\{ f , g \} : C^\infty(\mathcal{M}) \times C^\infty(\mathcal{M}) \rightarrow C^\infty(\mathcal{M})$ that tells us, at each point, the scale factor between the area in units of $(f,g)$ and the area in canonical units.
+As in the example before, it would be $\frac{\textsf{units}(f) \times \textsf{units}(g)}{m \times Kg \frac{m}{s}}$.
+We have $\rho(f,g) = \frac{\varrho(q,p)}{\{ f , g \}}$.
+Note that the density cannot be expressed in terms of those variables if the bracket is zero.
+
+Given two vectors $v,w \in T\mathcal{M}$ we have a form $\omega(v,w) : T\mathcal{M} \times T\mathcal{M} \rightarrow \mathbb{R}$ that returns the area of the infinitesimal surface given by those two vectors in canonical units.
+The form is linear $\omega(u,\alpha v+ \beta w)=\alpha\omega(u, v)+\beta\omega(u,w)$ since we are just combining rectangular shapes.
+The form is skew-symmetric as the area does not change under a $90$ degree rotation $\omega(v,w) = \omega(w,-v) = -\omega(w,v)$.
+
+Two variables $q,p$ are canonical if they measure area in canonical units.
+For example, if $m \times Kg \frac{m}{s}$ is the canonical unit, angle in radians and angular momentum in $Kg \frac{m^2}{s}$ are canonical.
+By the previous definitions, if $q,p$ are canonical then $\{ q , p \} = 1$ and $\omega = dq \wedge dp$.
+
+The area given by vectors $e_q$ and $e_p$ is unitary in canonical units.
+If we express it in units of $(f,q)$ we have the scale factor $\{ f , g \}$.
+We can do so by using the differentials $(df\wedge dg) (e_q, e_p) = df(e_q)\wedge dg(e_p) - df(e_p)\wedge dg(e_q)$.
+Each element extracts the component of the vector along the variable.
+Therefore we obtain $\{ f , g \} = \frac{\partial f}{\partial q} \frac{\partial g}{\partial p}- \frac{\partial f}{\partial p} \frac{\partial g}{\partial q}$, the standard expression for the Poisson bracket.
+This can then be understood, instead of in terms of flow, in terms of ratio of areas, which is also confirmed by its units.
+
+To express the density in terms of two generic variables $(f, g)$ we need to know the area given by a square in the standardized units.
+This allows to express the density in terms of the units of $(f, g)$ instead of the standardized unit.
+This means that given the two variables $f,g \in C^\infty(\mathcal{M})$ on phase space, we must have an operation $\{ f , g \} : C^\infty(\mathcal{M}) \times C^\infty(\mathcal{M}) \rightarrow C^\infty(\mathcal{M})$ that tells us, at each point, the scale factor between the area in units of $(f,g)$ and the area in standardized units.
+As in the example before, it would be $\frac{\textsf{units}(f) \times \textsf{units}(g)}{m \times Kg \frac{m}{s}}$.
+
+We can integrate the density using $\int_{\Sigma}\varrho \, \omega(d\Sigma)$.
+$\varrho \, \omega(d\Sigma)$, in fact, will give us the amount of the distribution in the infinitesimal area.
+Note that $\int_{\Sigma}\varrho \, \omega(d\Sigma) = \int_{\Sigma}\frac{\varrho}{\{f,g\}} \, \{f,g\} \omega(d\Sigma) =\int_{\Sigma}\rho(f,g) \, df\wedge dg$.
+That is, we can express integration in any coordinates for which the Poisson bracket is non-zero.
+
+In other words, a symplectic manifold $(\mathcal{M}, \omega)$ is exactly the structure needed to define invariant distributions.
+In the same way the metric tensor $g$ defines lengths in a Riemannian manifold, the symplectic form $\omega$ defines areas over which we can define our densities.
 
@@
 \end{align} 
-where the vector $S$ tells us the infinitesimal displacement of each state.
+where the vector field $S$ tells us the infinitesimal displacement of each state.
+This can be understood in terms of simple vector calculus.
 As time passes, both the distribution value and the volume where it is defined are conserved.
@@
 We are now in a position to ``have our invariant distribution and integrate it too".
-On the symplectic structure, our density is a scalar function $\rho : \mathcal{M} \rightarrow \mathbb{R}$.
+On the symplectic structure, our density is a scalar function $\varrho : \mathcal{M} \rightarrow \mathbb{R}$.
 It has the two form $\rho \omega$ associate with it, in case we want to speak the language of differential geometry, and it has the function $\rho(q,p)$ in local coordinate, in case we want to speak the language of statistics and define quantities like the information entropy.
@@
 One needs to be careful, though, as it may be misconstrued to mean something much stronger that it actually is.
+
+Suppose we have system characterized by $n$ independent degrees of freedom.
+To each will correspond a set of canonical coordinates $(q^i, p_i)$ and together they will give us all the possible states for our system.
+Moreover, if we define a distribution $\rho_i(q^i,p_i)$ for each d.o.f., we must be able to define a joint distribution $\rho(q^i,p_j) = \prod \rho_i(q^i,p_i)$ on the whole system.
+
+That is: not only want to define the distribution as a whole in a way that is invariant.
+We want to make sure each marginal distribution on each independent d.o.f.
+is invariant as well.
+Not only the Jacobian of the whole transformation needs to be unitary, but also the two dimensional Jacobian on each independent d.o.f.
+and that is precisely what the preservation of the Poisson brackets do.
+
+We can again formalize this intuition using symplectic geometry.
+We have $\mathcal{M} = \mathcal{M}_1 \times \dots \times \mathcal{M}_n$.
+Each individual manifold has a symplectic form $\omega_1$.
+For each $\mathcal{M}_i$ we have an invariant marginal distribution $\varrho_i : \mathcal{M}_i \rightarrow \mathbb{R}$.
+On the whole space we want an invariant distribution $\varrho : \mathcal{M} \rightarrow \mathbb{R}$ such that $\prod \int_{\Sigma_i} \varrho_i \omega_i = \int_{\Sigma} \varrho \bigwedge \omega_i $
 

### 2018-02-23 2015CompleteDerivation/refereeResponse.tex
Sent to Christine for edits

% Using template from
% Title:    A LaTeX Template For Responses To a Referees' Reports
% Author:   Petr Zemek <s3rvac@gmail.com>
% Homepage: https://blog.petrzemek.net/2016/07/17/latex-template-for-responses-to-referees-reports/
% License:  CC BY 4.0 (https://creativecommons.org/licenses/by/4.0/)
\documentclass[10pt]{article}

% Allow Unicode input (alternatively, you can use XeLaTeX or LuaLaTeX)
\usepackage[utf8]{inputenc}

\usepackage{microtype,xparse,tcolorbox,amsfonts}
\newenvironment{reviewer-comment }{}{}
\tcbuselibrary{skins}
\tcolorboxenvironment{reviewer-comment }{empty,
  left = 1em, top = 1ex, bottom = 1ex,
  borderline west = {2pt} {0pt} {black!20},
}
\ExplSyntaxOn
\NewDocumentEnvironment {response} { +m O{black!20} } {
  \IfValueT {#1} {
    \begin{reviewer-comment~}
      \setlength\parindent{2em}
      \noindent
%      \ttfamily
       #1
    \end{reviewer-comment~}
  }
  \par\noindent\ignorespaces
} { \bigskip\par }

\NewDocumentCommand \Reviewer { m } {
  \section*{Comments~by~Reviewer~#1}
}
\ExplSyntaxOff
\AtBeginDocument{\maketitle\thispagestyle{empty}\noindent}

% You can get probably get rid of these definitions:
\newcommand\meta[1]{$\langle\hbox{#1}\rangle$}
\newcommand\PaperTitle[1]{``\textit{#1}''}

\title{Statement on the Revision of \meta{Paper ID} \\
  Based on the Referees' Report}
\author{Author1 \and Author2 \and Author3}
\date{\today}

\begin{document}
This statement concerns our revision of the \meta{Paper ID} paper,
entitled \PaperTitle{\meta{Paper Title}}, based on the referees'
report.

\Reviewer{\#1}

\begin{response}{Classical Mechanics (Lagrangian and Hamiltonian, and also Newtonian Mechanics), Quantumn Mechanics, and Relativity Theory all have their quite well understood, albeit different, mathematical formulations, where for each a sensible space of physical states are defined, equipped with meningful topology, and if possible by a differentiable structure, so that one can study time evolution operators that describe the dynamics in these spaces.

		This is hardly news! In this very long paper, I was unable to dig out,  however much I wished for it to be there, truly new unifying principles than went beyond eloquent rephrasings of what is already known.

		``Advancing Knowledge in the field" is a criterion for acceptance to this journal, and the authors really must work harder to convince this referee (and probably many others) that they have brought new knowledge, new insight, new perspectives, to the field. To me, they don't seem to be advancing anything.  }

As the first sentence in the abstract says ``the aim of this work is to show that particle mechanics, both classical and quantum, Hamiltonian and Lagrangian, can be derived from few simple physical assumptions." That is the claim of the paper, which we personally find new and unexpected. That we can justify the basic laws of physics, not by experiment, not by arbitrarily setting an arbitrary mathematical framework, but by logical reasoning.

If one opens a textbook about computer science theory, one finds it starts by defining what performing computation means and from that the fundamental results in decidability and complexity theory follow. If one opens a textbook about group theory, one finds it starts by defining what a group is and from that the fundamental results in representation theory follow. The point of the paper is to show that we can do that for physics as well. Once we fully specify what we are studying and how, we already get the laws of motion.

The reaction from a few colleagues to our work is along these lines: ``Well, you recover the same equations that we already have? So there are no new predictions. So there is no new physics. Why is this interesting?" This seems to be the reaction of the reviewer when he says he was unable to dig out anything new beyond the eloquent rephrasing. Indeed, if one is looking for ``new physics" or ``new mathematical results", one will be disappointed. However, we made no claims to the contrary. We state, over and over, that we derive only the known particle mechanics. Why would one expect something different?

The advancement, in this case, lies in renewed understanding. This is akin to what happened at the beginning of the nineteenth century when mathematicians started reworking the foundations of math, reformulating it on logic and set theory. At the beginning, that work did not immediately produce new concrete results as well and some were skeptical of its usefulness. Yet now axiomatic set theory is the basis of all mathematics. Unfortunately, the physics community never really engaged in a similar process. Most of the people who now work on foundations of physics are really looking for new physics, and not in trying to clarify what we already have, what are the most basic requirements to do physics and how much physics is already contained in those assumptions.

Now, it is perfectly fine that the referee is not interested in this topic or sees no value in this approach at this point. Maybe he will be more interested when we will expand it to the case where the assumption underpinning quantum mechanics (i.e. irreducibility) fails and, by using similar techniques, we try to derive what mathematical framework needs to follow, having first identified exactly what mathematical operation can no longer be justified. But before we can do that, we have to check we can at least get what we know. And we can guarantee the referee that there are others in the scientific community that believe this is a very reasonable way to ``Advance Knowledge in the field": we have met them at difference conferences.

\end{response}
\begin{response}{

		Niels Bohr famously made the distinction between precision and clarity, and it seems that here the authors have chosen precision to such a degree that all traces of a clear message are completely lost.

		It is possible that, hidden somewhere in the expanse of over 40 journal pages, there may be snippets of new insight, new approaches to constructing mathematical pictures. If so, the task is on the authors to destill these decisive differencs from conventional wisdom, and then present them more clearly (i.e., briefly), instead of hiding them in a capacious and overwhelming word gush.

		I recommend re-submission of a much shorter version. I suggest the authors provide examples where their method/approach shines, and the standard approach is not so good. Then we may see if they have advanced knowledge in the field of mathematical modelling. 
	}
We strongly agree that this work is far too detailed to be accessible by the largest audience. In fact, we have numerous other activities (e.g. talks, videos, target student lectures, ...) that take some of the insights and make them more accessible. However, this is not that work.

To boil it down to the essentials, the article consists of a \textbf{claim} (i.e. particle mechanics can be derived from physical principles) and a \textbf{demonstration} (not a formal mathematical proof, but still a series of arguments that follow from hypothesis to conclusion). The claim is essentially made in the abstract. It takes the rest of the work to show the claim is valid. While we can create a shorter version with just the highlights of the derivation, we still need to write somewhere the full demonstration. Otherwise we would just be giving a set of unsubstantiated claims. If the proof of a mathematical theorem takes 400 pages, it takes 400 pages and it's published in full. It's not rejected because ``it's too long" or ``it's too detailed". Those are not valid arguments against a demonstration. Yes, few people would actually bother to go through the whole 400 pages. Yes, most people will read the highlights. But it still needs to be written down.

We can ask the question: is the paper unreasonably long for what it sets itself up to do? Maybe the reviewer is right and this can be accomplished in 10 pages. We offer a back of the envelope calculation. We are deriving from scratch the underpinning of classical and quantum Hamiltonian/Lagrangian particle mechanics, including the relativistic case, under scalar/vector potential forces. These are taught at an undergraduate level in, let's say, 3 to 4 classes. We can say, for the sake of argument, that the derivation presented in this paper would replace the chapters of the books that introduce the equations and the basic concepts. Let's say a chapter each, of approximately 30-40 pages. We are at 90-160 pages. We are talking to a more advanced audience, so let's divide by 3. We get 30-50 pages, which is in the ballpark of the article. So, objectively, we would be skeptical that in 10 pages one is able to prove anything of this kind.

What we find particularly frustrating is that we are already in a world where detailed and thoughtful discussion is disparaged in favor of comments limited to 140 characters. We would hope that the scientific community as a whole still values that someone still takes the time to get into the nitty-gritty details to check whether something is correct or merely ``sounds" correct. We agree that not everybody can go into the excruciatingly boring details of everything, and we do not fault the reviewer for not being interested in those details. Yet, the idea that such type of work should not be conducted and published is what the referee is, maybe unintentionally, advocating.

If the referee is of the opinion that ``the basic laws of physics can be derived from few physical assumptions" is of no interest to the entire physics community, then fine. If the referee is of the opinion that the demonstration is wrong, then the referee should highlight the part where he believes it to be incorrect. But the referee doesn't do either. We hope that the referee recognizes that what we are doing is of value to the scientific community in general, though maybe not to him in particular, and maybe reconsiders his position.

In conclusion, we do not believe there is a practical way to address the referee's concerns. The first objection is that it is difficult to find what is the point of this paper. The abstract starts with ``The aim of this work is to show that particle mechanics, both classical and quantum, Hamiltonian and Lagrangian, can be derived from few simple physical assumptions." The introduction states: ``The main goal is to show that it is possible to derive classical/quantum Hamiltonian/Lagrangian particle mechanics starting from few physical assumptions." And the title is ``From physical assumptions to classical and quantum Hamiltonian and Lagrangian particle mechanics". We cannot find a way to make the claim more clear than this. The claim is novel (no such comprehensive work has been attempted) and is of interested (even to those that argue against us). The second objection is that the paper should be much shorter. As argued before, this would severely compromise the ability to demonstrate the claim, reducing it to a mere conjecture. We agree that another shorter submission with the highlights should be submitted, but not as a replacement to the comprehensive work.

\end{response}

\Reviewer{\#2}
\begin{response}{Let me start, contrary to the custom, with the conclusion: I recommend this article
		for publication in Journal of Physics Communication. The reason is not that I agree with
		authors or have similar point of view. The real reason is that in my opinion what is lacking
		in the contemporary theoretical physics is a deep considerations of its foundations. It seems
		to me that we, as a community of physicists, have arrived to the boundaries of our abilities
		to create meaningful physical theories. Some of us study, often with success, ``particular
		case of particular case" of a physical system while others create ``supersymmetric quasi-q-
		quantum" field theories in n dimensions and drift into pure mathematics not really having
		any physical example. Very rarely we go back into the foundations of even well established
		theories trying to understand their origin and maybe understand what prevents us from
		developing it further, or what might be a reason for arising difficulties. Authors made the
		effort of justifying the mathematical language used in physics by examining the very basic
		physical assumptions than one should make to proceed.}

We appreciate that the reviewer has grasped in full the aim and the spirit of our work.

\end{response}
\begin{response}{On page 7 there is Proposition IV.4 that says ``A map $f : \mathcal{S}_1 \to \mathcal{S}_2$ between two sets
		of physically distinguishable elements $\mathcal{S}_1$ and $\mathcal{S}_2$ is a continuous map." It was explained
		earlier in the text that $\mathcal{S}_1$ and $\mathcal{S}_2$ are topological spaces. Reading the above proposition
		literary we should conclude that the topology of $\mathcal{S}_1$ needs to be trivial in a sense that any
		subset is open. At least if we understand that `a map' means usually `any map' or `all
		the maps'. Later on it is said that this proposition explains why in physics we use `well
		behaved functions'. Does it mean that all are well behaved? I do not suppose so.}
 We thank the referee for catching the imprecision. We modified the proposition to read ``A map $f:\mathcal{S}_1 \rightarrow \mathcal{S}_2$ that represents a physical relationship between two sets of physically distinguishable elements $\mathcal{S}_1$ and $\mathcal{S}_2$ is a continuous map." This qualifies that, yes, mathematically we are free to choose whatever function we may want. But if the function has to describe a physical relationship then it has to map measurable outcomes to measurable outcomes.
\end{response}

\begin{response}{From the section about Hamiltonian mechanics I gained the impression that authors
consider Hamiltonian mechanics as fundamental while Lagrangian mechanics plays the
secondary role. I must admit that my opinion is quite the opposite. I think Lagrangian
mechanics is fundamental while Hamiltonian mechanics is in a sense derived and plays
the secondary role.}
Indeed there is ample debate in philosophy of science on whether Hamiltonian or Lagrangian mechanics plays the more fundamental role. See for example:
\begin{itemize}
	\item Jill North, \emph{The ``Structure" of physics}, 
	Journal of Philosophy 106 (2):57-88 (2009)
	\item  Erik Curiel, \emph{Classical Mechanics is Lagrangian; It Is Not Hamiltonian}, Brit. J. Phil. Sci. 65 (2014), 269–321
	\item Thomas William Barrett, \emph{On the Structure of Classical Mechanics},	
	The British Journal for the Philosophy of Science, Volume 66, Issue 4, 1 December 2015, Pages 801–828
\end{itemize}

This paper does not take a side per se. To make that more clear, we added a footnote right before the introduction of the Lagrangian that states: ``Given that we introduce Lagrangian Mechanics only now, the reader may get the impression that we believe that Hamiltonian mechanics is somehow more fundamental than Lagrangian mechanics. This is not the case. What happens is that under the assumption of deterministic and reversible evolution, as stated by V.I. Arnold, ``Lagrangian mechanics is contained in Hamiltonian mechanics as a special case". In this work deriving Hamiltonian mechanics first is therefore more natural as the argument proceeds from the more general to the more specific, and as we are interested in seeing how much can be derived from each individual assumption."

In other words, assuming deterministic and reversible evolution leads to assuming Lagrangians are regular and therefore admit a Hamiltonian. Hamiltonian systems need not be convex to be deterministic and reversible. For example, $H=c|p|$ is a perfectly valid Hamiltonian that given initial conditions $q_0$ and $p_0$ leads to the solution $q(t)=q_0 + c \, sign(p_0)$ and $p(t)=p_0$ and it does not admit a Lagrangian. In this setting, Lagrangian systems are a proper subset of Hamiltonian systems, so focusing on phase-space evolution yields the more general picture. But under different assumptions, the prevalent picture would be different.

\end{response}

\begin{response}{I could argue by showing, that the version of relativistic Hamiltoniam
		mechanics authors propose as a direct consequence of their assumptions, is in my opinion
		not correct. Let us think of the simplest case of a free particle in Special Relativity Theory.
		From Minkowski we have learned how to deal with problems of relativity, i.e. problem of
		working in coordinates relative to a reference frame and constant checking if our physical
		quantities transform properly by Lorenz transformation. I assume here that the space-time
		is a Minkowski space, i.e. four dimensional affine space modeled on four dimesional vector space $V$ with bilinear form of sygnature (1; 3) (or (3; 1) if you are more physicist than math-
		ematician) and with time orientation. Than the behaviour of a particle is fully described
		by its world line which is one dimensional oriented submanifold of $M$. We can parameterize
		this submanifold by its proper time, but in principle there is no need of it. Knowing the
		submanifold we can calculate the parametrisation, but it does not add any new knowledge
		about the particle itself. Phase trajectories in $T*M$ are also one dimensional oriented sub-
		manifolds in the mass shell i.e. satisfying the condition $\eta(p, p) = m^2$. An equation for such
		object as one dimensional submanifold is one dimensional distribution rather than vector
		field. More precisely, in case of oriented submanifolds, half-distribution to encode orientation. Such an object can be obtained in the realm of Hamiltonian mechanics, in case we
		think in terms of Lagrangian submanifolds rather than Hamiltonian vector fields and allow
		more general generating objects for Lagrangian submanifolds than just functions. For me
		the correct Hamiltonian description of a free relativistic particle in Minkowski space-time is
		the Lagrangian submanifold of $TT^*M$ generated by the following Morse family of functions
		$H : \mathbb{R}_+ \times T*M \to R,$ $H(r, x, p) = r(
		\sqrt{\eta(p, p)} - m)$ where $x \in M, p \in V*, \mathbb{R} \ni r > 0$.
		This can be derived form invariant but singular Lagrangian $L(x, v) = m \sqrt{\eta(v, v)}$ defined
		on $TM$. I suppose I have assumed $c = 1$ here. The details of this approach can be found
		in a paper W. M. Tulczyjew, P. Urbanski, Slow and careful Legendre transformation for
		singular Lagrangians, Acta Phys. Polon. B 30 (1999), 2909-2978. which is available also
		in arXiv.}
\end{response}

The version of relativistic Hamiltonian mechanics we derive is one of the standard formulations. To our knowledge, it was first developed by Synge, and was named Hamiltonian mechanics on the ``extended phase-space" by Lanczos (Synge preferred ``space of states and energy"). It was developed by other authors and is included in V.I.Arnold Mathematical Methods of Classical Mechanics. It returns the usual predictions, including for the case of a free particle.

We assume that by ``correct" the reviewer means ``the right way to look at it". The only claim we make is that it is a good way as long as the assumption we derived it from holds: if one is studying the deterministic motion of an infinitesimally reducible material, then it will work very well. Outside of that scope (e.g. a system described by a singular Lagrangian, a system under dissipative forces, etc.) we give no guarantees.

That said, there are numerous points of contact between Hamiltonian mechanics on the extended phase space and the formalism shown by the reviewer. The behavior of a particle is in general a line in $T^*M$ parameterized by $s$ (not necessarily proper time). If the system is also Lagrangian, then the projection of that line on $M$ will be unique. In the justification for Proposition V.7 we introduce the infinitesimal displacement $S \in T\mathcal{S}$. Since $\mathcal{S}$ is $T^*\mathcal{Q}$, $S \in TT^*\mathcal{Q}$, in line with the reviewer comment. The extended Hamiltonian for a free particle is given at VI.8 and is $\mathcal{H} = \frac{1}{2m} ( p^i p_i - (E/c)^2 + m^2c^2)$. Since $\mathcal{H}=0$, we can change the sign of $\mathcal{H}$, which will simply change the parametrization of the trajectories from $s$ to $-s$. We choose $\eta_{\alpha \beta}$ with signature $(+,-,-,-)$, so we have $(E/c)^2 - p^i p_i = p_\alpha \eta^{\alpha \beta} p_\beta$. We can rewrite $\hat{\mathcal{H}} = -\mathcal{H} = \frac{1}{2m} (\sqrt{p_\alpha \eta^{\alpha \beta} p_\beta} + mc) ( \sqrt{p_\alpha \eta^{\alpha \beta} p_\beta} - mc) = r ( \sqrt{p_\alpha \eta^{\alpha \beta} p_\beta} - mc)$. This fits the form set in the comment by setting $r$ appropriately.

The main difference is that in this formalism the extended Lagrangian is $\mathcal{L}=\frac{1}{2} m(u^\alpha \eta_{\alpha \beta} u^\beta - c^2)$ which is in terms of the four velocity. The Hessian is $\partial_{u^\alpha}\partial_{u^\beta} \mathcal{L} = \frac{1}{2}m \eta_{\alpha \beta}$ so the Lagrangian is not singular.

We are unsure, though, that these types of comparisons belong to this paper which is already quite long. We feel a critique to the different ``flavors" of mechanics would distract from the main objective of the paper, and would belong to a different work.

\begin{response}{Summarising, I really think we should discuss foundations of physics much more than
		we do. For this the paper From physical assumptions to classical and quantum Hamiltonian
		and Lagrangian particle mechanics might be an excellent starting point.}
\end{response}

We couldn't agree more.

\end{document}

### 2018-02-23 2015CompleteDerivation/refereeResponse.tex
Revised by Christine

@@
 
-\title{Statement on the Revision of \meta{Paper ID} \\
-  Based on the Referees' Report}
-\author{Author1 \and Author2 \and Author3}
+\title{Statement on the Revision of JPCO-100515 \\
+  Based on the Referees' Reports}
+\author{Gabriele Carcassi \and Christine A. Aidala \and David J. Baker \and Lydia Bieri}
 \date{\today}
@@
 \begin{document}
-This statement concerns our revision of the \meta{Paper ID} paper,
-entitled \PaperTitle{\meta{Paper Title}}, based on the referees'
-report.
+This statement concerns our revision of submission JPCO-100515,
+entitled \textit{From physical assumptions to classical and quantum Hamiltonian and Lagrangian particle mechanics}, based on the referees'
+reports.
 
@@
 That we can justify the basic laws of physics, not by experiment, not by arbitrarily setting an arbitrary mathematical framework, but by logical reasoning.
+The same point is reiterated in the introduction and is also present in the title.
+We cannot find a way to make the claim more clear than this.
 
@@
 The point of the paper is to show that we can do that for physics as well.
-Once we fully specify what we are studying and how, we already get the laws of motion.
+Once we fully specify what we are studying and how, we already get the laws of evolution.
 
@@
 So there is no new physics.
-Why is this interesting?" This seems to be the reaction of the reviewer when he says he was unable to dig out anything new beyond the eloquent rephrasing.
+Why is this interesting?" This seems to be the reaction of the reviewer when he says he was unable to dig out anything new beyond eloquent rephrasing.
 Indeed, if one is looking for ``new physics" or ``new mathematical results", one will be disappointed.
@@
 We state, over and over, that we derive only the known particle mechanics.
-Why would one expect something different?
 
@@
 This is akin to what happened at the beginning of the nineteenth century when mathematicians started reworking the foundations of math, reformulating it on logic and set theory.
-At the beginning, that work did not immediately produce new concrete results as well and some were skeptical of its usefulness.
+At the beginning, that work did not immediately produce new concrete results either and some were skeptical of its usefulness.
 Yet now axiomatic set theory is the basis of all mathematics.
 Unfortunately, the physics community never really engaged in a similar process.
-Most of the people who now work on foundations of physics are really looking for new physics, and not in trying to clarify what we already have, what are the most basic requirements to do physics and how much physics is already contained in those assumptions.
+Most of the people who now work on foundations of physics are really looking for new physics, and not trying to clarify what we already have, what are the most basic requirements to do physics and how much physics is already contained in those assumptions.
 
 Now, it is perfectly fine that the referee is not interested in this topic or sees no value in this approach at this point.
-Maybe he will be more interested when we will expand it to the case where the assumption underpinning quantum mechanics (i.e.
-irreducibility) fails and, by using similar techniques, we try to derive what mathematical framework needs to follow, having first identified exactly what mathematical operation can no longer be justified.
-But before we can do that, we have to check we can at least get what we know.
-And we can guarantee the referee that there are others in the scientific community that believe this is a very reasonable way to ``Advance Knowledge in the field": we have met them at difference conferences.
+Maybe he will be more interested when we expand it to the case where the assumption underpinning quantum mechanics (i.e.
+irreducibility) fails and, by using similar techniques, we try to derive what new mathematical framework needs to follow, having first identified exactly what mathematical operation can no longer be justified.
+But before we can do that, we have to check we can at least recover what we know.
+And indeed there are others in the scientific community that believe this is a very reasonable way to advance knowledge in the field: we have met them at different conferences.
 
@@
 We strongly agree that this work is far too detailed to be accessible by the largest audience.
-In fact, we have numerous other activities (e.g.
-talks, videos, target student lectures, ...) that take some of the insights and make them more accessible.
+In fact, we have numerous other activities (e.g.~talks, videos, target student lectures, ...) that take some of the insights and make them more accessible.
 However, this is not that work.
@@
 It takes the rest of the work to show the claim is valid.
-While we can create a shorter version with just the highlights of the derivation, we still need to write somewhere the full demonstration.
-Otherwise we would just be giving a set of unsubstantiated claims.
+While we could create a shorter version with just the highlights of the derivation, we would still need to write somewhere the full demonstration, otherwise we would just be giving a set of unsubstantiated claims.
 If the proof of a mathematical theorem takes 400 pages, it takes 400 pages and it's published in full.
@@
 Yes, few people would actually bother to go through the whole 400 pages.
-Yes, most people will read the highlights.
-But it still needs to be written down.
+Yes, most people will read only the highlights.
+But the full proof still needs to be written down.
 
-We can ask the question: is the paper unreasonably long for what it sets itself up to do?
-Maybe the reviewer is right and this can be accomplished in 10 pages.
+We can ask the question: is the paper unreasonably long for what it sets out to do?
 We offer a back of the envelope calculation.
@@
 We get 30-50 pages, which is in the ballpark of the article.
-So, objectively, we would be skeptical that in 10 pages one is able to prove anything of this kind.
+This means that a significantly shorter paper would severely compromise the ability to demonstrate the claim, reducing it to a mere conjecture.
 
 What we find particularly frustrating is that we are already in a world where detailed and thoughtful discussion is disparaged in favor of comments limited to 140 characters.
-We would hope that the scientific community as a whole still values that someone still takes the time to get into the nitty-gritty details to check whether something is correct or merely ``sounds" correct.
-We agree that not everybody can go into the excruciatingly boring details of everything, and we do not fault the reviewer for not being interested in those details.
+We would hope that the scientific community as a whole still values someone taking the time to get into the nitty-gritty details to check whether something is correct or merely ``sounds" correct.
+We agree that not everybody can go into the excruciating details of everything, and we do not fault the reviewer for not being interested in those details.
 Yet, the idea that such type of work should not be conducted and published is what the referee is, maybe unintentionally, advocating.
@@
 If the referee is of the opinion that ``the basic laws of physics can be derived from few physical assumptions" is of no interest to the entire physics community, then fine.
-If the referee is of the opinion that the demonstration is wrong, then the referee should highlight the part where he believes it to be incorrect.
-But the referee doesn't do either.
-We hope that the referee recognizes that what we are doing is of value to the scientific community in general, though maybe not to him in particular, and maybe reconsiders his position.
-
-In conclusion, we do not believe there is a practical way to address the referee's concerns.
-The first objection is that it is difficult to find what is the point of this paper.
-The abstract starts with ``The aim of this work is to show that particle mechanics, both classical and quantum, Hamiltonian and Lagrangian, can be derived from few simple physical assumptions." The introduction states: ``The main goal is to show that it is possible to derive classical/quantum Hamiltonian/Lagrangian particle mechanics starting from few physical assumptions." And the title is ``From physical assumptions to classical and quantum Hamiltonian and Lagrangian particle mechanics".
-We cannot find a way to make the claim more clear than this.
-The claim is novel (no such comprehensive work has been attempted) and is of interested (even to those that argue against us).
-The second objection is that the paper should be much shorter.
-As argued before, this would severely compromise the ability to demonstrate the claim, reducing it to a mere conjecture.
-We agree that another shorter submission with the highlights should be submitted, but not as a replacement to the comprehensive work.
+If the referee is of the opinion that the demonstration is wrong, then he should highlight the part where he believes it to be incorrect.
+But he doesn't do either.
+We hope that the referee recognizes the potential of what we are doing and how it can be of value to the scientific community in general, though maybe not to him in particular, and perhaps reconsiders his position.
 
@@
 
-We appreciate that the reviewer has grasped in full the aim and the spirit of our work.
+We thank the referee for the kind words and appreciate that the referee has grasped in full the aim and the spirit of our work.
 
@@
 This paper does not take a side per se.
-To make that more clear, we added a footnote right before the introduction of the Lagrangian that states: ``Given that we introduce Lagrangian Mechanics only now, the reader may get the impression that we believe that Hamiltonian mechanics is somehow more fundamental than Lagrangian mechanics.
+To make that more clear, we added a footnote right before the introduction of the Lagrangian that states: ``Given that we introduce Lagrangian mechanics only now, the reader may get the impression that we believe that Hamiltonian mechanics is somehow more fundamental than Lagrangian mechanics.
 This is not the case.
@@
 Hamiltonian systems need not be convex to be deterministic and reversible.
-For example, $H=c|p|$ is a perfectly valid Hamiltonian that given initial conditions $q_0$ and $p_0$ leads to the solution $q(t)=q_0 + c \, sign(p_0)$ and $p(t)=p_0$ and it does not admit a Lagrangian.
+For example, $H=c|p|$ is a perfectly valid Hamiltonian that given initial conditions $q_0$ and $p_0$ leads to the solution $q(t)=q_0 + c \, sign(p_0)$ and $p(t)=p_0$, and it does not admit a Lagrangian.
 In this setting, Lagrangian systems are a proper subset of Hamiltonian systems, so focusing on phase-space evolution yields the more general picture.
@@
 To our knowledge, it was first developed by Synge, and was named Hamiltonian mechanics on the ``extended phase-space" by Lanczos (Synge preferred ``space of states and energy").
-It was developed by other authors and is included in V.I.Arnold Mathematical Methods of Classical Mechanics.
+It was developed by other authors and is included in V.I.Arnold's Mathematical Methods of Classical Mechanics.
 It returns the usual predictions, including for the case of a free particle.
@@
 That said, there are numerous points of contact between Hamiltonian mechanics on the extended phase space and the formalism shown by the reviewer.
-The behavior of a particle is in general a line in $T^*M$ parameterized by $s$ (not necessarily proper time).
-If the system is also Lagrangian, then the projection of that line on $M$ will be unique.
+The behavior of a particle is in general a line in $T^*\mathcal{M}$ parameterized by $s$ (not necessarily proper time).
+If the system is also Lagrangian, then the projection of that line on $\mathcal{M}$ will be unique.
 In the justification for Proposition V.7 we introduce the infinitesimal displacement $S \in T\mathcal{S}$.
-Since $\mathcal{S}$ is $T^*\mathcal{Q}$, $S \in TT^*\mathcal{Q}$, in line with the reviewer comment.
+Since $\mathcal{S}$ is $T^*\mathcal{M}$, $S \in TT^*\mathcal{M}$, in line with the reviewer comment.
 The extended Hamiltonian for a free particle is given at VI.8 and is $\mathcal{H} = \frac{1}{2m} ( p^i p_i - (E/c)^2 + m^2c^2)$.
@@
 We can rewrite $\hat{\mathcal{H}} = -\mathcal{H} = \frac{1}{2m} (\sqrt{p_\alpha \eta^{\alpha \beta} p_\beta} + mc) ( \sqrt{p_\alpha \eta^{\alpha \beta} p_\beta} - mc) = r ( \sqrt{p_\alpha \eta^{\alpha \beta} p_\beta} - mc)$.
-This fits the form set in the comment by setting $r$ appropriately.
+This fits the form in the comment by setting $r$ appropriately.
 
@@
 
-We are unsure, though, that these types of comparisons belong to this paper which is already quite long.
-We feel a critique to the different ``flavors" of mechanics would distract from the main objective of the paper, and would belong to a different work.
+We are unsure, though, that these types of comparisons belong in this paper which is already quite long.
+We feel a critique of the different ``flavors" of mechanics would distract from the main objective of the paper, and would belong to a different work.
 
@@
 
-We couldn't agree more.
+We again thank the referee for the positive remarks. The type of issues raised and the discussions they prompt are exactly of the type we hope to have with more individuals in the community.
 

### 2018-03-03 2015CompleteDerivation/completeDerivation.tex
Prepared cover letter
Addressing referees comments

@@
 \begin{prop}\label{prop:continuity}
-	A map $f:\mathcal{S_1} \rightarrow \mathcal{S_2}$ between two sets of physically distinguishable elements $\mathcal{S_1}$ and $\mathcal{S_2}$ is a continuous map.
+	A map $f:\mathcal{S_1} \rightarrow \mathcal{S_2}$ that represents a physical relationship between two sets of physically distinguishable elements $\mathcal{S_1}$ and $\mathcal{S_2}$ is a continuous map.
 \end{prop}
@@
 Such transformation is the Lagrangian of the system and the motion is given by the Euler-Lagrange equation.
-For the time independent case, we define $L=v^i p_i - H$ where $v^i=d_t x^i$.
+For the time independent case, we define $L=v^i p_i - H$ where $v^i=d_t x^i$.\footnote{Given that we introduce Lagrangian mechanics only now, the reader may get the impression that we believe that Hamiltonian mechanics is somehow more fundamental than Lagrangian mechanics.
+This is not the case.
+What happens is that under the assumption of deterministic and reversible evolution, as stated by V.I.
+Arnold, ``Lagrangian mechanics is contained in Hamiltonian mechanics as a special case".\cite{Arnold} In this work deriving Hamiltonian mechanics first is therefore more natural as the argument proceeds from the more general to the more specific, and as we are interested in seeing how much can be derived from each individual assumption.}
 
@@
 
+\bibitem{Arnold} Arnold, V.I.: Mathematical Methods of Classical Mechanics, Springer, (1989)
+
 \bibitem{EPR} Einstein, A., Podolsky, B.,  Rosen, N.: Can quantum-mechanical description of physical reality be considered complete? Phys. Rev. 47 777 (1935)

### 2018-04-25 2015CompleteDerivation/changes.tex
Last changes before publication

% Using template from
% Title:    A LaTeX Template For Responses To a Referees' Reports
% Author:   Petr Zemek <s3rvac@gmail.com>
% Homepage: https://blog.petrzemek.net/2016/07/17/latex-template-for-responses-to-referees-reports/
% License:  CC BY 4.0 (https://creativecommons.org/licenses/by/4.0/)
\documentclass[10pt]{article}

% Allow Unicode input (alternatively, you can use XeLaTeX or LuaLaTeX)
\usepackage[utf8]{inputenc}

\usepackage{microtype,xparse,tcolorbox,amsfonts}
\newenvironment{reviewer-comment }{}{}
\tcbuselibrary{skins}
\tcolorboxenvironment{reviewer-comment }{empty,
  left = 1em, top = 1ex, bottom = 1ex,
  borderline west = {2pt} {0pt} {black!20},
}
\ExplSyntaxOn
\NewDocumentEnvironment {response} { +m O{black!20} } {
  \IfValueT {#1} {
    \begin{reviewer-comment~}
      \setlength\parindent{2em}
      \noindent
%      \ttfamily
       #1
    \end{reviewer-comment~}
  }
  \par\noindent\ignorespaces
} { \bigskip\par }

\NewDocumentCommand \Reviewer { m } {
  \section*{Comments~by~Reviewer~#1}
}
\ExplSyntaxOff
\AtBeginDocument{\maketitle\thispagestyle{empty}\noindent}

% You can get probably get rid of these definitions:
\newcommand\meta[1]{$\langle\hbox{#1}\rangle$}
\newcommand\PaperTitle[1]{``\textit{#1}''}

\title{Changes of JPCO-100515 \\
  Based on the Referees' Reports}
\author{Gabriele Carcassi \and Christine A. Aidala \and David J. Baker \and Lydia Bieri}
\date{\today}

\begin{document}
This statement concerns the changes of submission JPCO-100515,
entitled \textit{From physical assumptions to classical and quantum Hamiltonian and Lagrangian particle mechanics}, based on the referees'
reports.

~\newline

On page 7, we modified Proposition IV.4 from ``A map $f : \mathcal{S}_1 \to \mathcal{S}_2$ between two sets
 of physically distinguishable elements $\mathcal{S}_1$ and $\mathcal{S}_2$ is a continuous map." to  ``A map $f:\mathcal{S}_1 \rightarrow \mathcal{S}_2$ that represents a physical relationship between two sets of physically distinguishable elements $\mathcal{S}_1$ and $\mathcal{S}_2$ is a continuous map."

~\newline

On page 225, before Proposition VII.2, we added a footnote right before the introduction of the Lagrangian that states: ``Given that we introduce Lagrangian mechanics only now, the reader may get the impression that we believe that Hamiltonian mechanics is somehow more fundamental than Lagrangian mechanics. This is not the case. What happens is that under the assumption of deterministic and reversible evolution, as stated by V.I. Arnold, ``Lagrangian mechanics is contained in Hamiltonian mechanics as a special case". In this work deriving Hamiltonian mechanics first is therefore more natural as the argument proceeds from the more general to the more specific, and as we are interested in seeing how much can be derived from each individual assumption."

~\newline

In the bibliography, we added reference to Arnold, V.I.: Mathematical Methods of Classical Mechanics, Springer, (1989).

\end{document}
